### 2020-02-09 draft.Rmd
first commit

---
title: "A White Noise Approach to Evolutionary Ecology"
author: "Bob Week, Steve Krone, Scott Nuismer"
date: "8/19/2019"
abstract: "We derive the dynamics of the distribution of a quantitative character and the abundance of a biological population from a stochastic partial differential equation driven by space-time white noise. In the process we develop a useful set of heuristics to operationalize the powerful, but abstract theory of white noise. This approach allows us to compute the full implications of a stochastic process such as demographic stochasticity on phenotypic distributions and abundances of populations. We demonstrate the utility of our approach by deriving a model of diffuse coevolution mediated by exploitative competition for a continuum of resources. Other than trait and abundance distributions, this model predicts interaction networks parameterized by rates of interactions, competition coefficients, and selection gradients. We compute these quantities under the two scenarios of weak and moderate competition and investigate the relationship between selection gradients and competition coefficients. This illustrative investigation suggests magnitudes of selection gradients only weakly correspond to competition coefficients. Future work building on this formalization can contribute to the understanding of feedbacks between evolution and ecology and how random processes such as demographic stochasticity and random genetic drift modulate the strength of these feedbacks."
bibliography: ref.bib
output:
  pdf_document:
    toc: false
    number_sections: true
    fig_width: 4
    fig_height: 3
    fig_caption: true
header-includes:
  - \usepackage{float}
  - \usepackage{mathrsfs}
  - \usepackage{hyperref}
  - \usepackage[left]{lineno}
  - \usepackage{amsmath}
  - \linenumbers
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
#knitr::opts_chunk$set(fig.pos = 'H')
```

# Introduction

A primary aim of theoretical evolutionary ecology is the development of mathematical approaches to describe the evolution of biological populations and their interactions with both the biotic and abiotic environments in which they are embedded. Given this consideration a natural scope for such an approach centers on quantifying the abundance dynamics of populations and the evolution of traits mediating their interactions as functions of relevant abiotic factors. Although taking into account abundance, phenotype and environment provides the basis for a partial understanding of the complex nature of biological communities, a deeper understanding must account for the effects of contemporary dispersal and the phylogeographic history of interacting lineages [@Kraft2007; @Hickerson2010; @Manceau2016; @markmcpeek2017] along with the genetic basis of ecologically relevant traits [@jeffreyconner2004; @FUSSMANN2007] and feedbacks between populations and the biogeochemical cycles they ultimately depend on [@michelloreau2010; @9781107648258]. It is therefore ideal that the development of any such mathematical approach anticipates extensions to account for these important factors shaping ecological communities, especially as empirical and conceptual work in these directions continues to grow [@AbdalaRoberts2014; @Klzsch2015; @Crutsinger2015; @Fitzpatrick2015; @Fitzpatrick2017; @Marx2017; @Rudman2017; @Skovmand2018; @VanNuland2019; @Harmon2019]. Furthermore, the approach should have a stochastic component to capture the chance nature of biological reality [@Lande2003; @DeMeester2018; @Mubayi2019] and serve as a basis for the construction of statistical methods that measure evolutionary and ecological processes occuring in the wild. Such methods will tether theory to reality and allow for rigorous tests of hypotheses on the structure and behavior of ecological communities. In this paper we take one small step in this direction by establishing a formal connection between the continuous-time dynamics of abundance and quantitative traits in stochastically evolving populations and investigate the relationship between strengths of ecological interactions and coevolution.

Current theoretical approaches to synthesize evolution and ecology have capitalized on the fact that biological fitness plays a key role in determining both sets of dynamics. While correlation of fitness and genotype is the basis of evolution by natural selection, the mean fitness across all individuals in a population determines the growth, stasis or decline of abundance. In section \ref{deterministic} we review the mathematical formalization of this connection, which has been established in the contexts of population genetics [@9781932846126], evolutionary game theory [@9780674023383] and quantitative genetics [@Lion2018]. Although reviewing these accomplishments reveals a beautiful synthesis of evolution and population ecology, it also reveals a gap in theoretical approaches to incorporate the intrinsically random nature of populations. Specifically, in theoretical quantitative genetics the derivation of a populations' response to random genetic drift is derived in discrete time under the assumption of constant population (or effective population) size [@Lande1976]. Although this approach conveniently mimics the formalism provided by the Wright-Fisher model of population genetics, real population sizes fluctuate over time. Furthermore, since these fluctuations are themselves stochastic, it seems natural to derive expressions for the evolutionary response to demographic stochasticity and consider how the results relate to characterizations of random genetic drift. This can be done in continuous time for population genetic models without too much technical overhead, assuming a finite number of alleles. However, for populations with a continuum of types, such as a quantitative trait, this becomes a serious mathematical endeavor. Here we close this gap by applying some theory of stochastic partial differential equations (SPDE). Our goal is two-fold: 1) Communicate some useful results of SPDE to as wide of audience as possible. With this goal in mind we will not provide a rigorous treatment of SPDE. Instead, we introduce a set of heuristics that only require the basic concepts of Reimann integration, partial differentiation and some exposure to stochastic ordinary differential equations (SDE). For a concise introduction to SDE, we recommend the primer by Evans [-@9781470410544]. 2) Establish a novel synthetic approach to theoretical evolutionary ecology. To demonstrate the utility of our approach, we derive a model of coevolution in an ecological network and use it to investigate the relationship between competition coefficients and selection gradients.

We begin in section \ref{deterministic} with a brief review of progress in modelling the deterministic dynamics of populations. Section \ref{wnc} then introduces a useful set of mathematical tools based on the calculus of white noise. In section \ref{stochastic} we outline how to employ these tools to derive a system of SDE modelling abundance dynamics and the evolution of the mean and variance of a quantitative trait. In section \ref{particular} we derive particular results under the assumption of a normally distributed trait. Section \ref{inheritance} extends the previous results under the condition of imperfect inheritance. Section \ref{coev} demonstrates the utility of our framework through a derivation of a model of coevolution across a guild of $S$ species participating in exploitative competition along a common resource continuum. In section \ref{dynamics} we numerically compute distributions of traits and abundances along with interaction networks parameterized by selection gradients under the scenarios of weak and moderate competition, yielding theoretical possibilities for the distribution of pairwise coevolution in ecological communities. In section \ref{ecoevo} we investigate the relationship between the strengths of competition and coevolution. Finally, section \ref{conclusion} concludes with a few remarks on the limits of this approach along with future directions to incorporate more explicitly the genetic architecture of populations and feedbacks with ecosystem processes.

# The framework

At the core of our approach is a stochastic analog of the replicator equation with mutation in continuous time and phenotypic space [@Taylor1978; @Schuster1983]. From this stochastic replicator-mutator equation we derive a system of SDE for the dynamics of abundance, mean trait and additive genetic variance of a population. Hence, our approach develops a quantitative genetic theory of evolutionary ecology. A popular alternative to quantitative genetics is the theory of adaptive dynamics. As demonstrated by Page and Nowak [-@PAGE2002], the canonical equation of adaptive dynamics can be derived from the replicator-mutator equation. Thus, one could start from the atomic roots of our approach and pursue a stochastic adaptive dynamic theory instead. We choose the former in anticipation of an extension of our approach that explicitly models the genetic details of populations.

In this section we review the derivations of the replicator-mutator equation and trait dynamics from abundance dynamics and extend these formulae along with related results to the stochastic case. The results established in this section provide the framework from which larger scale ecological stuctures, such as species abundance distributions and interaction networks, can be derived.

## Deterministic dynamics \label{deterministic}

The connection between the deterministic dynamics of the abundance of a population and the frequencies of discrete phenotypes within it is well established via heterogeneity in the vital rates of differing phenotypes [ie, demographic heterogeneity, @Mubayi2019]. This is particularly evident in the derivation of the evolution of allele frequencies due to natural selection provided by Crow and Kimura [-@9781932846126, \S 5.3]. Specifically, denoting $N$ the abundance of the population, $m_i$ the growth rate [called the Malthusian parameter in @9781932846126] of allele $i$ and $p_i$ the frequency of this allele, we find (using their numbering scheme)

\begin{equation}
\dot N=N\sum_ip_im_i=\bar mN,\tag{5.3.3}
\end{equation}

\begin{equation}\label{popgen}
\dot p_i=(m_i-\bar m)p_i.\tag{5.3.5}
\end{equation}

Equation (\ref{popgen}) is known in the field of evolutionary game theory as the replicator equation [@9780674023383; @Lion2018]. Instead of being explicitly focused on alleles, the replicator equation describes the fluctuations of relative abundances of various _types_ in a population in terms of the vital rates of each type [@Taylor1978; @Schuster1983]. It is straight-forward to extend the replicator equation to include mutation, which is known as the replicator-mutator equation [@9780674023383]. To derive the replicator-mutator equation one can take the same steps to derive (\ref{popgen}) and inject a model of mutation via a matrix of transition rates [@9780674023383, eqn. 13.25]. Building on these insights, we derive an analog of the replicator-mutator equation for a continuum of types (that is, for a quantitative trait). In particular, for a continuously reproducing population at time $t$ with trait values $x\in\mathbb{R}$ and an abundance density $\nu(x,t)$, the demographic dynamics can be modelled by the general PDE

\begin{equation}\label{eq1}
\dot{\nu}(x,t)=F(\nu(x,t),x).
\end{equation}

More specifically, assuming selection is determined by the growth rate $m(\nu(x,t),x)$, mutation is captured by diffusion with coefficient $\frac{\mu}{2}$, and ignoring all other processes, we set

\begin{equation}\label{rxn_diff}
F(\nu(x,t),x)=m(\nu(x,t),x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t).
\end{equation}

In this case equation (\ref{eq1}) becomes a nonlinear reaction-diffusion equation. For brevity we often write the growth rate as just $m(\nu,x)$. As an example, logistic growth combined with quadratic stabilizing selection can be captured using the growth rate $m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_\mathbb{R}\nu(x,t)dx$ where $a\geq0$ is the strength of stabilizing selection around the phenotypic optimum $\theta\in\mathbb{R}$, $c\geq0$ captures the effect of intraspecific competition and $r\in\mathbb{R}$ is the intrinsic growth rate in the absence of abiotic selection. We will always assume $r>0$. In section \ref{coev} we will derive a model of coevolution in which the associated fitness function shares some important properties with this fitness function. In particular, we show in SM \S\ref{finite} that fitness functions $m(\nu,x)$ satisfying
\begin{equation}\label{m_cond}
r-\frac{a}{2}(\theta-x)^2-c\int_\mathbb{R}\nu(x,t)dx\leq m(\nu,x)\leq r-\frac{a}{2}(\theta-x)^2
\end{equation}
for some $r,a,c\geq0$ guarantee finite total abundance $N(t)$, trait mean $\bar x(t)$ and trait variance $\sigma^2(t)$ for all $t\geq0$, so long as these variables have finite initial conditions. Since this is the only form of fitness we consider, for the remainder of this paper we assume
\begin{subequations}
\begin{align}
N(t)=&\int_\mathbb{R}\nu(x,t)dx <\infty, \\
|\bar x(t)|=&\frac{1}{N(t)}\left|\int_\mathbb{R}x\nu(x,t)dx\right|<\infty, \\ 
\sigma^2(t)=&\frac{1}{N(t)}\int_\mathbb{R}(x-\bar x(t))^2\nu(x,t)dx <\infty, \ \forall \ t\geq0.
\end{align}
\end{subequations}

To derive a replicator-mutator equation from the combination of equations (\ref{eq1}) and (\ref{rxn_diff}), we employ the chain rule from elementary calculus. Defining $p(x,t)=\nu(x,t)/N(t)$ as the relative density of trait $x$ at time $t$, we have

\begin{multline}
\dot N(t)=\frac{d}{dt}\int_\mathbb{R}\nu(x,t)dx=\int_\mathbb{R}\dot{\nu}(x,t)dx \\
=\int_\mathbb{R}m(\nu,x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)dx \\
=N(t)\int_\mathbb{R}m(\nu,x)p(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t)dx=\bar m(t) N(t)
\end{multline}

where, for $f:\mathbb{R}\to\mathbb{R}$, we define, when it exists, $\bar f(t)=\int_\mathbb{R}f(x)p(x,t)dx$. In particular $\bar m(t)$ is the mean fitness across all trait values in the population. Since $\nu(x,t)$ is continuous in $x$ and $\int_\mathbb{R}\nu(x,t)dx<\infty$, the result $\int_\mathbb{R}\frac{\partial^2}{\partial x^2}\nu(x,t)dx=0$ can be derived from the fundamental theorem of calculus. Biological reasoning agrees with this result since mutation neither creates nor destroys individuals, but merely changes their type from their parental type. Taking the same approach, we derive the dynamics of $p(x,t)$ as

\begin{multline}
\dot p(x,t)=\frac{\partial}{\partial t}\frac{\nu(x,t)}{N(t)}=\frac{N(t)\dot{\nu}(x,t)-\nu(x,t)\dot N(t)}{N^2(t)} \\
=\frac{1}{N(t)}\Big(m(\nu,x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)-\bar m(t)\nu(x,t)\Big) \\
=(m(\nu,x)-\bar m(t)) p(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t).
\end{multline}

This result closely resembles Kimura's continuum-of-alleles model [@Kimura1965]. The primary difference being that our model utilizes diffusion instead of convolution with an arbitrary mutation kernel. Of course, our model of mutation can be derived as an approximation to Kimura's model, which has been referred to as the Gaussian allelic approximation [in reference to the distribution of mutational effects at a given locus of a genome on the values of traits, @9780471986539], the infinitesimal model [in reference to modelling continuous traits as being encoded by an infinite number of loci each having infinitesimal effect, @Barton2017] and the Gaussian descendants approximation [in reference to offspring trait values being normally distributed around their parental values, @Turelli2017]. Alternatively, since diffusion is the continuous-time equivalent to convolution against a Gaussian kernel (SM \S\ref{diffconvequiv}), our model can also be seen as a special case of the continuum-of-alleles model. Under our model of mutation, the dynamics of the mean trait $\bar x(t)$ are derived as
\begin{multline}\label{detmean}
\dot{\bar x}(t)=\frac{d}{dt}\int_\mathbb{R}x p(x,t)dx=\int_\mathbb{R}x\dot p(x,t)dx \\
=\int_\mathbb{R}x\big(m(\nu,x)-\bar m(t)\big) p(x,t)+x\frac{\mu}{2}\frac{\partial^2}{\partial x^2} p(x,t)dx \\
=\mathrm{Cov}_t\Big(x,m(\nu,x)\Big)+\frac{\mu}{2}\int_\mathbb{R} x\frac{\partial^2}{\partial x^2} p(x,t)dx,
\end{multline}

where, for $f,g:\mathbb{R}\to\mathbb{R}$, we define, when it exists,

\begin{equation}
\mathrm{Cov}_t\Big(f(x),g(x)\Big)=\int_\mathbb{R}\big(f(x)-\bar f(t)\big)\big(g(x)-\bar g(t)\big)p(x,t)dx.
\end{equation}

Equation (\ref{detmean}) is a continuous time analog of the well known Robertson-Price equation [@Robertson1966; @PRICE1970; @FRANK2012; @Queller2017; @Lion2018]. In the particular case developed here, transmission bias only occurs when $p(x,t)$ is not symmetric around $\bar x$. In particular, when $p(x,t)$ is a Gaussian curve $\int_\mathbb{R}x\frac{\partial^2}{\partial x^2}p(x,t) dx=0$. Following this approach, in SM \S\ref{var_deriv} we derive the dynamics of the variance $\sigma^2(t)=\mathrm{Cov}_t(x,x)$ as

\begin{equation}\label{vardyn_det}
\dot{\sigma}^2(t)=\mathrm{Cov}_t\Big((x-\bar x)^2,m(\nu,x)\Big)+\frac{\mu}{2}\int_\mathbb{R}(x-\bar x)^2\frac{\partial^2}{\partial x^2}p(x,t)dx.
\end{equation}

In the absence of mutation equation (\ref{vardyn_det}) agrees with the result derived by Lion [-@Lion2018] for discrete phenotypes. When $p$ is Gaussian the increase in variance due to mutation simplifies to $\mu$.

In the following sections we extend these results to the case when PDE (1) is driven by a space-time white noise process $\dot W(x,t)$. We begin by defining space-time white noise and illustrating its relevant properties. We then introduce a set of heuristics for performing calculations with respect to $\dot W(x,t)$. Since our aim is to present this material to as wide of audience as possible, our approach deviates from standard definitions to remove the need for a detailed technical treatment. For those not interested in the white noise calculus and would rather jump straight into more biologically relevant results, we recommend skipping to \S\ref{particular}.

## White noise calculus and superprocesses \label{wnc}

### Definition and basic properties \label{wnc_intro}

We define $L^2_C(\mathbb{R}\times[0,\infty))$ as the set of functions $f:\mathbb{R}\times[0,\infty)\to\mathbb R$ that are continuous in their second argument (time) and satisfy $\|f\|_2(t)=\sqrt{\int_\mathbb R f^2(x,t)dx}<\infty$. We define a generalized stochastic process $\mathbf W$ that maps functions $f\in L^2_C(\mathbb{R}\times[0,\infty))$ to real-valued stochastic processes indexed by $t\in[0,\infty)$. To evaluate this process for a function $f\in L^2_C(\mathbb{R}\times[0,\infty))$ and some time $t>0$ we write $\mathbf W_t(f)$. Specifically, for any $f,g\in L^2_C(\mathbb{R}\times[0,\infty))$, we define $\mathbf W(f)$ and $\mathbf W(g)$ to be Gaussian processes satisfying, for any $t,t_1,t_2\geq0$,

\begin{equation}\label{exp_WN}
\mathbb{E}\big(\mathbf W_t(f)\big)=\mathbb{E}\big(\mathbf W_t(g)\big)=0,
\end{equation}

\begin{equation}\label{cov_WN}
\mathbb{C}\big(\mathbf W_{t_1}(f),\mathbf W_{t_2}(f)\big)=\int_0^{t_1\wedge t_2}\int_\mathbb{R} f(x,s)g(x,s)dxds,
\end{equation}

where $t_1\wedge t_2=\min(t_1,t_2)$ and $\mathbb{E},\mathbb{C}$ denote expectation and covariance with respect to the underlying probability space. These operators are to be distinguished from $\bar f(t)$ and $\mathrm{Cov}_t(f,g)$ which denote expectation and covariance with respect to phenotypic diversity at time $t\geq0$.

Since Gaussian processes are characterized by their expectations and covariances, the processes $\mathbf W(f)$ and $\mathbf W(g)$ are well defined (under the convention $f=g$ whenever $\|f-g\|_2(t)=0$ for each $t>0$). In particular, if $f\in L^2_C(\mathbb{R}\times[0,\infty))$ is independent of time, then $\mathbf W(f)$ is a Brownian motion with variance $t\|f\|_2^2$. With the generalized process $\mathbf W$ defined, we define the space-time white noise $\dot W(x,t)$ implicitly via

\begin{equation}
``\int_0^t\int_\mathbb{R}f(x,s)\dot W(x,s)dxds"=\mathbf W_t(f), \ \forall \ f\in L^2_C(\mathbb{R}\times[0,\infty)), \ t>0.
\end{equation}

We place quotations around the left-hand side of the above expression to emphasize its informal nature. Since space-time white noise, represented by the symbol $\dot W(x,t)$, does not exist as a function in the classical sense it cannot be integrated as suggested above. However, similar to the approach taken here, space-time white noise can be made sense of in a more abstract setting as a _measure_-valued process [@Dawson1975; @Walsh]. Hence, our use of integral notation is merely symbolic to provide intuition and should not be confused with classical Riemann integrals. One can think of $\dot W(x,t)$ as the static seen on old television sets or infinitely detailed random dust spread across both time and space. Figure \ref{wn} illustrates sample paths of white noise in one (left) and two (right) dimensions.

```{r, include=F, eval=F}
require("ggplot2")
require("ggthemes")
require("plot3D")
require("gridExtra")

num1 = 1e4
num2 = 1e4

xi_1 = rnorm(num1)
xi_2 = rnorm(num2)

x1 = seq(0,1,length.out = num1)
x2 = seq(0,1,length.out = sqrt(num2))

M = expand.grid(x=x2,y=x2)
M$value = xi_2

df_1 = data.frame(xi=xi_1,x=x1)

wn1d = ggplot(df_1)+geom_point(aes(x=x,y=xi),size=0.05)+theme_bw()+xlab("Time")+ylab("Value")
wn2d = ggplot(M, aes(x, y, color = value))+scale_color_gradient(low="black", high="white")+geom_point(size=0.001, shape=22, show.legend = FALSE)+theme_bw()+xlab("Time")+ylab("Space")

wn = grid.arrange(wn1d,wn2d,nrow=1)

ggsave("~/Research/White Noise/wn.png",wn,width=5,height=2.5)
```

```{r, echo=F, fig.align='center', fig.cap='\\label{wn}Approximations of sample paths of temporal white noise (left) and space-time white noise (right) with brightness scaled to value.'}
knitr::include_graphics('wn.png')
```

Following our definition of white noise, when attempting to compute the value of this random field we sample it using $L^2_C(\mathbb{R}\times[0,\infty))$ functions. For example, integrating white noise over a region $D\times[0,t]$, with $t>0$ and $D$ a bounded subset of $\mathbb{R}$, is equivalent to evaluating $\mathbf W_t(I_{D\times [0,\infty)})$ where

\begin{equation}
I_{D\times [0,\infty)}(x,t)=\left\{\begin{matrix}
0, & x\notin D \\
1, & x\in D
\end{matrix}\right..
\end{equation}

Since $\int_\mathbb{R}I_{D\times [0,\infty)}^2(x,t)dx=\int_Ddx=|D|$, where $|D|$ denotes the length of $D$, $I_{D\times [0,\infty)}\in L^2_C(\mathbb{R}\times[0,\infty))$. Thus, by equations (\ref{exp_WN}) and (\ref{cov_WN}), we have

\begin{equation}
\mathbb{E}\left(\int_0^t\int_D\dot W(x,s)dxds\right)=0,
\end{equation}

\begin{equation}
\mathbb{V}\left(\int_0^t\int_D\dot W(x,s)dxds\right)=t|D|,
\end{equation}

where $\mathbb{V}$ denotes the variance operator with respect to the underlying probability space. Following this notation, equations (\ref{exp_WN}) and (\ref{cov_WN}) can be rewritten as

\begin{equation}\label{exp_WN_xi}
\mathbb{E}\left(\int_0^t\int_\mathbb{R}f(x,s)\dot W(x,s)dxds\right)=0,
\end{equation}

\begin{equation}\label{cov_WN_xi}
\mathbb{C}\left(\int_0^{t_1}\int_\mathbb{R}f(x,s)\dot W(x,s)dxds,\int_0^{t_2}\int_\mathbb{R}g(x,s)\dot W(x,s)dxds\right)
=\int_0^{t_1\wedge t_2}\int_\mathbb{R}f(x,s)g(x,s)dxds.
\end{equation}

We write $d\hat{\mathbf W}_t(f)=\frac{1}{\|f\|_2(t)}\left(\int_\mathbb{R}f(x,t)\dot W(x,t)dx\right)dt$ (and typically suppress the dependency on time) so that

\begin{equation}
\int_0^td\hat{\mathbf W}(f)=\int_0^t\int_\mathbb{R}\frac{f(x,s)}{\sqrt{\int_\mathbb{R}f^2(s,y)dy}}\dot W(x,s)dxds.
\end{equation}

It follows that $\mathbb{E}\big(\int_0^td\hat{\mathbf W}(f)\big)=0$ and $\mathbb{C}\big(\int_0^{t_1}d\hat{\mathbf W}(f),\int_0^{t_2}d\hat{\mathbf W}(f)\big)=t_1\wedge t_2$. This implies that, as a function of $t$, $\int_0^td\hat{\mathbf W}(f)$ is a standard Brownian motion for any $f\in L^2_C(\mathbb{R}\times[0,\infty))$. Hence, $d\hat{\mathbf W}(f)$ is analogous to the traditional shorthand used to denote stochastic differentials. Thus, equation (\ref{cov_WN_xi}) effectively extends It&ocirc;'s multiplication table to:

\begin{table}[H]
\centering\caption{An extension of It\^o's multiplication table.}\vspace{0.2cm}
\begin{tabular}{l|lll}
             & $d\hat{\mathbf W}(f)$               & $d\hat{\mathbf W}(g)$                & $dt$ \\ \hline
             &                            &                            &      \\
$d\hat{\mathbf W}(f)$ & $dt$                       & $\frac{\left(\int_\mathbb{R}f \ g \ dx\right)dt}{\|f\|_2\|g\|_2}$ & $0$  \\
             &                            &                            &      \\
$d\hat{\mathbf W}(g)$ & $\frac{\left(\int_\mathbb{R}f \ g \ dx\right)dt}{\|f\|_2\|g\|_2}$ & $dt$                       & $0$  \\
             &                            &                            &      \\
$dt$         & $0$                        & $0$                        & $0$
\end{tabular}\label{mult}
\end{table}

The extension of It&ocirc;'s multiplication table and properties of white noise outlined in this subsection provide a useful set of tools for working with SPDE. In SM \S\ref{SDEriv} we employ these tools to derive SDE that track the dynamics of abundance, mean trait and phenotypic variance of a population from a particular SPDE. In the following subsection, we review how this SPDE naturally arises from diffusion limits of branching processes.

## From branching processes to SPDE \label{stochastic}

In real populations individuals are born and potentially reproduce before they ultimately die. These three events provide the basic ingredients of a branching process. Mathematical investigations of such processes have a relatively deep history [@Kendall1966]. The most simple branching process, known as the Galton-Watson process, describes the number of individuals alive at a given time $t\geq0$ as a non-negative integer [@Kimmel2015]. Feller [-@feller1951] introduced a formal method to approximate branching processes with diffusion processes which are continuous in state (ie, population size is approximated as a continuous quantity). Since diffusion processes possess greater analytical tractability than branching processes, Feller's method, known as the diffusion limit, has acquired immense popularity, particularly in the field of mathematical population genetics [@Ewens2004]. For over the past half of a century a great deal of accomplishments have been achieved in formalizing the diffusion limits of branching processes that describe biological populations with some continuous structure [@Watanabe1968; @Dawson1975; @Mlard1993; @edwinarendperkins1995; @zeng1998absolute; @alisonetheridge2000; @Bertoin2003; @Etheridge2008; @Barton2019]. This structure can represent geographic space or, relevant to our context, phenotypic space. One particularly important branching process with continuous structure is known as the branching Brownian motion. Unlike other stochastic processes that take values in $d$-dimensional Euclidean space, branching Brownian motion takes values in the set of non-negative finite measures over $\mathbb{R}^d$. Intuitively, one can think of a measure as a function that maps bounded subsets of $\mathbb{R}^d$ to real numbers. As a technical aside, we only consider the Borel subsets of $\mathbb{R}^d$. In particular, denoting $X_t$ a branching Brownian motion, for any $D\subset\mathbb{R}^d$, $X_t(D)$ returns the number of individuals alive within the region $D$ at time $t\geq0$. The branching Brownian motion has three main components:

  1) __Spatial movement:__ Individuals are born at the current location of their parent and then move around space according to Brownian motion with diffusion parameter $\sqrt\mu$. In our context we interpret spatial movement as mutation so that the _location_ of an individual represents the value of its phenotype.

  2) __Branching rate:__ Lifetimes of these individuals are often assumed to be exponentially distributed. At the death of an individual, a branching event occurs where a random (possibly zero) number of offspring are left. However, in our case we will assume lifetimes are non-random and equal to a single unit of time for all individuals. From a mathematical standpoint this assumption makes little difference since many alternative branching processes lead to the same diffusion limit. However, we make this assumption because we are interpreting movement as mutation and are assuming the effects of mutations are independent and identically distributed for the trait values inherited by each offspring. Note that a Brownian motion with diffusion parameter $\sqrt\mu$ integrated over the unit interval is a zero-mean Gaussian random variable with variance $\mu$. This fact creates a vital link to the deterministic dynamics reviewed above. Indeed, in the absence of selection, the deterministic PDE combining equations (\ref{eq1}) and (\ref{rxn_diff}) reduces to the Fokker-Planck equation for a Brownian motion with diffusion parameter $\sqrt\mu$.

  3) __Branching mechanism:__ At a branching event, the parent leaves a random number of offspring. The distribution of offspring left is called the branching mechanism. Under the so-called critical condition the average number of offspring left is set equal to one. We denote the variance in reproductive output by $V$.

To obtain a SPDE from the branching Brownian motion we take a diffusion limit. For a population of size $n$ at time $t$ we rescale the mass of individuals and time by $1/n$, space by $1/\sqrt{n}$ and consider the limit as $n\to\infty$. That is, denoting $\sqrt n D=\{\sqrt n x:x\in D\subset\mathbb{R}^d\}$, we consider the rescaled process $X_t^{(n)}(D)=\frac{1}{n}X_{nt}(\sqrt n D)$. The limiting process $\mathscr{X}_t=\lim_{n\to\infty}X_t^{(n)}$ is called a super-Brownian motion and is also a non-negative finite measure-valued process [@Watanabe1968]. It turns out that for spatial dimension $d=1$, $\mathscr{X}_t$ is absolutely continuous for each $t\geq0$. This means that we can write $\mathscr{X}_t(D)=\int_D \nu(x,t)dx$ for some density process $\nu(x,t)$ [@Konno1988; @Reimers1989]. Under the critical condition this density process satisfies the SPDE

\begin{equation}\label{neutral_SPDE}
\dot\nu(x,t)=\frac{\mu}{2}\Delta\nu(x,t)+\sqrt{V\nu(x,t)}\dot W(x,t).
\end{equation}

The symbol $\Delta$, referred to as the Laplacian or Laplace operator, represents here the continuous extension of the second derivative $\frac{\partial^2}{\partial x^2}$ to the space of bounded continuous functions on $\mathbb{R}$ vanishing at $\pm\infty$. To rigorously interpret SPDE (\ref{neutral_SPDE}), we integrate the solution $\nu(x,t)$ against functions $f\in C_b^\infty(\mathbb{R})$, where $C_b^\infty(\mathbb{R})$ is the set of bounded and infinitely continuously differentiable functions on $\mathbb{R}$. That is, equation (\ref{neutral_SPDE}) is just an abbreviation for
\begin{multline}\label{neutral_int}
\int_\mathbb{R}\nu(x,t)f(x)dx-\int_\mathbb{R}\nu(x,0)f(x)dx \\
=\int_0^t\int_\mathbb{R}\nu(x,s)\frac{\mu}{2}\Delta f(x)dsdx+\int_0^t\int_\mathbb{R}f(x)\sqrt{V\nu(x,s)}\dot W(x,s)dxds.
\end{multline}

For more on the general theory of SPDE see Walsh [-@Walsh]. Note that since $\nu(x,t)$ is the density of a finite measure, it is integrable for each $t\geq0$. Thus, since $|f(x)|\leq M$ for every $x$ and some $M>0$, setting $\varphi(x,t)=f(x)\sqrt{V\nu(x,t)}$ implies $\varphi\in L^2_C(\mathbb{R}\times[0,\infty))$. Hence, the white noise integral on the right-hand side of equation (\ref{neutral_int}) can be understood using the heuristics introduced above. Evaluating equation (\ref{neutral_int}) in the particular case of $f(x)\equiv1$ returns the total mass process, which we refer to as the total abundance $N(t)$.

A convergence theorem for the diffusion limit of a more general version of the branching Brownian motion was established by Watanabe [-@Watanabe1968]. Dawson [-@Dawson1975] suggested that, for spatial dimension $d=1$, this diffusion limit should admit a density process that satisfies a SPDE. Konna and Shiga [-@Konno1988] and Reimers [-@Reimers1989] independently proved that this is the case. The diffusion limit of this more general branching process (in arbitrary spatial dimension) is referred to as the Dawson-Watanabe superprocess [@alisonetheridge2000]. Conditioning the Dawson-Watanabe superprocess to have constant mass returns the Fleming-Viot process [@Etheridge1991; @Perkins1991] which has been popular in studies of spatial population genetics. In particular, an extension of the Fleming-Viot process, known as the $\Lambda$-Fleming-Viot process, was introduced by Bertoin and Le Gall [-@Bertoin2003] and used (and coined) by Etheridge [-@Etheridge2008] to resolve some technical challenges in modelling isolation by distance [@Felsenstein1975; see also @Barton2013; and @Barton2019].

A more general version of the Dawson-Watanabe superprocess that includes interactions (which can represent, among other phenomena, density-dependent growth rates and frequency-dependent selection) was studied by Perkins [-@Perkins1992; -@edwinarendperkins1995]. When the interactions occur only at branching events, which in our case translates to independence of mutation and selection, the Dawson-Watanabe superprocess with interactions on one dimensional trait space has also been shown to satisfy a SPDE that admits function-valued solutions which are positive, integrable and continuous in time and space [@zeng1998absolute]. Hence, this process provides a very natural and mathematically convenient candidate to base our framework on.

Instead of formally developing the diffusion limit of a branching Brownian motion with interaction we skip directly to a special case of the SPDE found by Li [-@zeng1998absolute] mentioned above:
\begin{equation}\label{SPDE}
\dot{\nu}(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\Delta\nu(x,t)+\sqrt{V\nu(x,t)}\dot W(x,t).
\end{equation}

Note that, comparing equation (\ref{SPDE}) to equation (3.5) of Li [-@zeng1998absolute], our $m$ and $V$ correspond to Li's $b$ and $c$ respectively. Denoting $C_1(\mathbb{R})^+$ the set of non-negative, continuous, integrable functions on $\mathbb{R}$, it is important to note equation (\ref{SPDE}) is only formal when $m(h,x)$ is bounded across all $h\in C_1(\mathbb{R})^+$ and $x\in\mathbb{R}$. However, recalling condition (\ref{m_cond}), the $m(h,x)$ we consider are only bounded above. Replacing $m(h,x)$ with its upper-bound demonstrates that the associated superprocess should still take values in the set of finite measures. Hence, we should still be able to compute the total mass process $N(t)$. What remains is to show the spatial moments $\int_\mathbb{R}|x|^n\nu(x,t)dx$ are finite for each $t>0$ and $n\in\mathbb{N}$, assuming finite initial conditions. In SM \S\ref{numerical} we investigate this claim by comparing simulations of the Branching brownian motion with selection to analytical solutions of the PDE obtained by combining equations (\ref{eq1}) and (\ref{rxn_diff}). Our results suggest this assertion is sound, but it has yet to be rigorously justified. For now, we leave this as an open problem and assume all spatial moments are finite for every $t\geq0$.

### Derivation of SDE for $N$, $\bar x$ and $\sigma^2$ \label{SDEriv}

As noted above, the dynamics of abundance (referred to in the superprocess literature as the total mass process) is obtained by defining the process $N(t)=\int_{\mathbb{R}}\nu(x,t)dx$ and evaluating expression

\begin{multline}
\int_\mathbb{R}f(x)\nu(x,t)dx-\int_\mathbb{R}f(x)\nu(x,0)dx \\
=\int_0^t\int_\mathbb{R}\nu(x,s)\left(m(\nu,x)f(x)+\frac{\mu}{2}\Delta f(x)\right)dxds+\int_0^t\int_\mathbb{R}f(x)\sqrt{V\nu(x,s)}\dot W(x,s)dxds
\end{multline}

with $f(x)\equiv1$ to obtain

\begin{equation}
N(t)-N(0)=\bar m(t)N(t)+\int_0^t\int_\mathbb{R}\sqrt{V\nu(x,s)}\dot W(x,s)dxds.
\end{equation}

Equation (\ref{cov_WN_xi}) implies

\begin{equation}
\mathbb{V}\left(\int_0^t\int_\mathbb{R}\sqrt{V\nu(x,s)}\dot W(x,s)dxds\right)=\int_0^t\int_\mathbb{R}V\nu(x,s)dxds=\int_0^tV N(s)ds.
\end{equation}

Following the notation introduced in \S\ref{wnc_intro}, we have

\begin{equation}
\int_0^td\hat{\mathbf W}_s(\sqrt{V\nu(x,s)})=\int_0^t\int_\mathbb{R}\frac{\sqrt{V\nu(x,s)}}{\sqrt{V N(s)}}\dot W(x,s)dxds
\end{equation}

is, as a function of $t$, a standard Brownian motion. To clean up notation set $dW_1(t)=d\hat{\mathbf W}_t(\sqrt{V\nu(x,t)})$. Then we can write the SDE for $N$ as

\begin{equation}\label{N}
dN(t)=\bar m(t)N(t)dt+\sqrt{V N(t)}dW_1(t).
\end{equation}

Under the assmption that $\int_\mathbb{R}(|x|+x^2)\nu(x,t)dx<\infty$ for each $t\geq0$, we show in SM \S\ref{SDE_DERIV} SDE for $\bar x$ and $\sigma^2$ can be expressed as

\begin{equation}\label{xbar_gen}
d\bar x(t)=\mathrm{Cov}_t\Big(x,m(\nu,x)\Big)dt+\frac{\mu}{2}\int_\mathbb{R}x\Delta p(x,t)dx+\sqrt{V\frac{\sigma^2(t)}{N(t)}}dW_2(t),
\end{equation}

\begin{multline}\label{sig2_gen}
d\sigma^2(t)=\left(\mathrm{Cov}_t\Big((x-\bar x(t))^2,m(\nu,x)\Big)+\frac{\mu}{2}\int_\mathbb{R}(x-\bar x(t))^2\Delta p(x,t)dx-V\frac{\sigma^2(t)}{N(t)}\right)dt \\ +\sqrt{V\frac{\overline{(x-\bar x(t))^4}-\sigma^4(t)}{N(t)}}dW_3(t).
\end{multline}

where $p(x,t)=\nu(x,t)/N(t)$ and $W_2,W_3$ are standard Brownian motions. In SM \S\ref{SDE_DERIV} we show that in general $W_1$ is independent of both $W_2$ and $W_3$, but $W_2$ and $W_3$ covary. The expressions derived can be used to investigate the dynamics of the mean and variance for general $\nu$. However, in the next section we further simplify these expressions by approximating $\nu$ with a Gaussian curve. In SM \S\ref{SDE_DERIV} we show that under the Gaussian case $W_1,W_2$ and $W_3$ are independent.

Before we make the Gaussian assumption, there is still quite a bit we can learn from expressions (\ref{xbar_gen}) and (\ref{sig2_gen}). Firstly, setting $V=0$ recovers the deterministic dynamics derived in \S\ref{deterministic}. Relating the response to demographic stochasticity derived here to the effect of random genetic drift derived in classic quantitative genetic theory, if we pretend that $\sigma^2(t)$ and $N(t)$ are constant with respect to time, then integrating the stochastic term in equation (\ref{xbar_gen}) over a single unit of time returns a normally distributed random variable with mean zero and variance equal to $V\sigma^2/N$. In particular, assuming perfect inheritance, when $V=1$ this random variable coincides with the effect of random genetic drift on the change in mean trait over a single generation derived using sampling arguments [@Lande1976]. There is also an interesting connection with classical population genetics. A fundamental result from early population genetic theory is the expected reduction in diversity due to the chance loss of alleles in finite populations [@Fisher1923; @Wright97]. This expected reduction in diversity due to random genetic drift is captured by the third term in the deterministic component of expression (\ref{sig2_gen}), but is now parameterized by the variance in reproductive output $V$.

## Particular results assuming a Gaussian phenotypic distribution \label{particular}

Gaussian phenotypic distributions are often obtained through Gaussian, exponential or weak selection approximations together with a simplified model of inheritance and random mating [@pmid17248993; @Turelli1984; @Turelli1986; @Turelli2017; @9780471986539]. Alternatively, it has been shown that a Gaussian distribution can provide a reasonable approximation even when selection is strong and non-Gaussian [@pmid7851785]. However, our approach adds an additional layer of difficulty. Even with Gaussian selection, the resulting solution to SPDE (\ref{SPDE}) will only be Gaussian in expectation, assuming a Gaussian initial condition. Yet this difficulty is not as challenging as it may first appear. Indeed, since SPDE (\ref{SPDE}) can be derived as a diffusion limit we know that, under the appropriate assumptions on selection, genetic architecture and reproduction, the stochastic departure from a Gaussian curve is negligible when the ratio $V/N$ is small (ie, when the variance in reproductive output is much smaller than the population size). In SM \S\ref{numerical} we demonstrate this result using numerical methods. Mathematically, this requirement restricts model parameters to regions that maintain large population sizes. Biologically, this implies populations are not at risk of extinction. Hence, models developed in this framework are not suitable for studying colonization-extinction dynamics or evolutionary rescue. Allowing for these restrictions, we may safely assume that $\nu$ is approximately Gaussian and justify writing

\begin{equation}
\nu(x,t)=\frac{N(t)}{\sqrt{2\pi\sigma^2(t)}}\exp\left(-\frac{\big(x-\bar x(t)\big)^2}{2\sigma^2(t)}\right).
\end{equation}

Under this assumption, Lande [-@Lande1976] found (suppressing the dependency on $t$)

\begin{equation}
\mathrm{Cov}(x,m)=\sigma^2\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right).
\end{equation}

Under the same assumption, in SM \S\ref{var_deriv} we show

\begin{equation}
\mathrm{Cov}\Big((x-\bar x)^2,m\Big)=2\sigma^4\left(\frac{\partial\bar m}{\partial\sigma^2}-\overline{\frac{\partial m}{\partial\sigma^2}}\right)
\end{equation}

and $\overline{(x-\bar x)^4}=3\sigma^4$. In particular, this implies

\begin{equation}\label{xbar}
d\bar x=\sigma^2\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right)dt+\sqrt{V\frac{\sigma^2}{N}}dW_2,
\end{equation}

\begin{equation}\label{G}
d\sigma^2=2\sigma^4\left(\frac{\partial\bar m}{\partial\sigma^2}-\overline{\frac{\partial m}{\partial\sigma^2}}\right)dt +\left(\mu-V\frac{\sigma^2}{N}\right)dt+\sigma^2\sqrt{\frac{2V}{N}}dW_3.
\end{equation}

Note that in the above expressions, the partial derivatives of $\bar m$ represent frequency independent selection and the averaged partial derivatives of $m$ represent frequency dependent selection. This relationship has already been pointed out by Lande [-@Lande1976] for the evolution of trait means, but here we see this relationship holds also for the evolution of trait variance. In the next subsection we generalize this result to the case when traits are imperfectly inherited. In this case, the phenotypic variance $\sigma^2$ is replaced by a genetic variance $G$ that represents the proportion of $\sigma^2$ explained by additive interactions of alleles at the same locus, but on different chromosomes, as well as the additive interactions among loci encoding for the focal phenotype [@joanroughgarden1979; @bulmer1980; @michaellynch1998]. For completeness we mention that any non-additive effect of interacting alleles at the same locus, but on different chromosomes, is referred to as dominance. Similarly, any non-additive effects among loci encoding for a quantitative character is referred to as epistasis. The variance $G$ is therefore fittingly referred to as the additive genetic variance.

## The evolution of additive genetic variance \label{inheritance}

To model imperfect heritability we consider the relationship between the expressed phenotype $x$ and its genetic component $g\in\mathbb{R}$ known as the _breeding value_ [@michaellynch1998]. We assume the absence of dominance and epistasis. The additive genetic variance can be modelled as the variance of breeding values in a population [@bulmer1980]. For any given individual with breeding value $g$, we assume the expressed phenotype is normally distributed around $g$ with variance $\eta$. We denote the density of this distribution by $\psi$ so that

\begin{equation}
\psi(x|g)=\frac{1}{\sqrt{2\pi\eta}}\exp\left(-\frac{(x-g)^2}{2\eta}\right).
\end{equation}

In terms of classic quantitative genetics $\eta$ is the known as the _variance of environmental deviation_ [@michaellynch1998] and $\psi$ plays a similar role as the _segregation kernel_ [@joanroughgarden1979]. To include this relationship in our framework, we write $\rho(g,t)$ as the abundance density of breeding values at time $t$. Hence $\int_\mathbb{R}\rho(g,t)dg=\int_\mathbb{R}\nu(x,t)dx=N(t)$. We switch our focus from directly modelling the evolution of $\nu(x,t)$ to modelling the evolution of $\rho(g,t)$. Once $\rho(g,t)$ is defined, we can compute $\nu(x,t)$ via

\begin{equation}\label{nu_from_rho}
\nu(x,t)=\int_\mathbb{R}\rho(g,t)\psi(x|g)dg.
\end{equation}

However, since selection acts on expressed phenotypes, we have to use our assumed relationship between breeding values and expressed traits to calculate the fitness of breeding values. Using equation (\ref{nu_from_rho}) we can write the Malthusian fitness of trait value $x$ as

\begin{equation}
m\big(\nu(x,t),x\big)=m\left(\int_\mathbb{R}\rho(g,t)\psi(x|g)dg,x\right)=\tilde m(\rho,x).
\end{equation}

Hence, we can also calculate the net Malthusian fitness of a breeding value $g$ as

\begin{equation}
m^*(\rho,g)=\int_\mathbb{R}\tilde m(\rho,x)\psi(x|g)dx.
\end{equation}

This is similar to the approach taken by Kimura and Crow [-@Kimura1978] to calculate the overall effects of selection for expressed characters onto the changes in the distribution of alleles encoding those characters. However, instead of focusing on the frequencies of alleles at particular loci, we focus on the densities of breeding values. With the relationship between $m(\nu,x)$ and $m^*(\rho,g)$ established, we define the evolution of $\rho(g,t)$ by

\begin{equation}
\dot\rho(g,t)=m^*(\rho,g)\rho(g,t)+\frac{\mu}{2}\Delta\rho(g,t)+\sqrt{V\rho(g,t)}\dot W(g,t).
\end{equation}

If we assume $\rho(g,t)$ is Gaussian its mode coincides with $\bar x$ and, denoting $G$ the variance of breeding values (ie, the additive genetic variance), $\sigma^2=G+\eta$. Hence, narrow sense heritability can be computed as $h^2=G/(G+\eta)$. Futhermore, since

\begin{multline}
\bar m(t)= \int_\mathbb{R}m(\nu,x) p(x,t)dx =\frac{1}{N(t)}\int_\mathbb{R}\tilde m(\rho,x)\int_\mathbb{R}\rho(g,t)\psi(x|g)dgdx \\
=\frac{1}{N(t)}\int_\mathbb{R}\rho(g,t)\int_\mathbb{R}\tilde m(\rho,x)\psi(x|g)dxdg=\frac{1}{N(t)}\int_\mathbb{R}m^* (\rho,g)\rho(g,t)dg = \bar m^*(t),
\end{multline}

equations (\ref{xbar}) and (\ref{G}) become

\begin{equation}\label{xbarfinal}
d\bar x=G\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right)dt+\sqrt{V\frac{G}{N}}dW_2,
\end{equation}

\begin{equation}\label{Gfinal}
dG=2G^2\left(\frac{\partial\bar m}{\partial G}-\overline{\frac{\partial m}{\partial G}}\right)dt+\left(\mu-V\frac{G}{N}\right)dt+ G\sqrt{\frac{2V}{N}}dW_3.
\end{equation}

# A model of diffuse coevolution \label{coev}

## Formulation

In this section we demonstrate the use of our framework by formulating a model of diffuse coevolution across a guild of $S$ species whose interactions are mediated by resource competition along a single niche axis. Our approach mirrors closely the theory developed by MacArthur and Levins [-@Macarthur1967], Levins [-@9780691080628] and MacArthur [-@Arthur1969; -@MacArthur1970; -@9780691023823]. The most significant difference, aside from allowing evolution to occur, is the treatment of resource quality, which we replace with a model of abiotic stabilizing selection. A derivation is provided in SM \S\ref{diffuse}.

For species $i$ we inherit the above notation for trait value, distribution, average, variance, abundance, etc except with an $i$ in the subscript. Real world examples of niche axes include the body size of prey for lizard predators and the date of activity in a season for pollinators competing for floral resources. For mathematical convenience, we model the axis of resources by the real line $\mathbb{R}$. The value of a resouce along this axis is denoted by the symbol $\zeta$. For an individual in species $i$, we assume the resource utilization curve $u_i$ can be written as

\begin{equation}
u_i(\zeta,x_i)=\frac{U_i}{\sqrt{2\pi w_i}}\exp\left(-\frac{(x_i-\zeta)^2}{2w_i}\right).
\end{equation}

We further assume the niche center $x_i$ is normally distributed among individuals in species $i$, but the niche breadth $w_i$ and total niche utilization $U_i$ are constant across individuals in species $i$ and therefore cannot evolve. Suppose $\theta_i\in\mathbb R$ is the optimal location along the niche axis for species $i$ such that, in the absence of competition, individuals leave on average $Q_i$ offspring when concentrated at $\theta_i$. We capture the rate by which the fitness falls as niche location $\zeta$ leaves the optimum $\theta_i$ by the parameter $A_i\geq0$. Hence, abiotic stabilizing selection along the resource axis can be modelled by the curve

\begin{equation}
e_i(\zeta)=Q_i\exp\left(-\frac{A_i}{2}(\theta_i-\zeta)^2\right).
\end{equation}

The effect of abiotic stabilizing selection on the fitness for an individual of species $i$ with niche location $x_i$ is then given by

\begin{equation}
\int_\mathbb{R}e_i(\zeta)u_i(\zeta,x_i)d\zeta=\frac{Q_iU_i}{\sqrt{A_iw_i+1}}\exp\left(-\frac{A_i}{2(A_iw_i+1)}(\theta_i-x_i)^2\right).
\end{equation}

To determine the potential for competition between individuals with niche locations $x_i$ and $x_j$, belonging to species $i$ and $j$ respectively, we compute the niche overlap

\begin{equation}
\mathcal O_{ij}(x_i,x_j)=\int_\mathbb{R}u_i(\zeta,x_i)u_j(\zeta,x_j)d\zeta=\frac{U_iU_j}{\sqrt{2\pi(w_i+w_j)}}\exp\left(-\frac{(x_i-x_j)^2}{2(w_i+w_j)}\right).
\end{equation}

In SM \S\ref{diffuse} we combine this niche model with equations (\ref{N}), (\ref{xbarfinal}) and (\ref{Gfinal}) to find

\begin{subequations}\label{comm_dynamics}
\begin{multline}
dN_i = \left\{R_i-\frac{a_i}{2}\Big((\bar x_i-\theta_i)^2+G_i+\eta_i\Big) + c_i\left(\sqrt{\frac{b_{ii}}{2\pi}}-\sum_{j=1}^SN_j\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\right)\right\}N_idt \\ + \sqrt{V_iN_i}dW_1,
\end{multline}

\begin{multline}
d\bar x_i = \left\{a_iG_i(\theta_i-\bar x_i)-c_iG_i\left(\sum_{j=1}^SN_jb_{ij}(\bar x_j-\bar x_i)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\right)\right\}dt+\sqrt{V_i\frac{G_i}{N_i}}dW_2,
\end{multline}

\begin{multline}
dG_i = \left\{\mu_i-a_i{G_i}^2+c_i{G_i}^2\left(\sum_{j=1}^SN_jb_{ij}\left(1-b_{ij}(\bar x_i-\bar x_j)^2\right)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\right)-V_i\frac{G_i}{N_i}\right\}dt \\+G_i\sqrt{\frac{2V_i}{N_i}}dW_3,
\end{multline}
\end{subequations}

where

\begin{subequations}
\begin{align}
R_i = & \ \ln \left(\frac{Q_iU_i}{\sqrt{1+A_iw_i}}\right), \\
a_i = & \ \frac{A_i}{1+A_iw_i}, \\
b_{ij}(t) = b_{ji}(t) = & \ \big(w_i+w_j+\eta_i+\eta_j+G_i(t)+G_j(t)\big)^{-1}, \\
c_i \geq & \ 0.
\end{align}
\end{subequations}

## Community dynamics \label{dynamics}

For the sake of illustration we numerically integrated system (\ref{comm_dynamics}) for a richness of $S=50$ species under two scenarios. For the first scenario of weak competition we randomly chose $c_i\sim\mathrm{Exp}(\mathtt{0.01})$ for each $i=1,\dots,S$ and for the second scenario of moderately strong competition we randomly chose $c_i\sim\mathrm{Exp}(\mathtt{0.1})$ for each $i=1,\dots,S$. Background parameters are drawn independently for each species from a common distribution. Specifically, we set model parameters according to Table \ref{par_vals}.

\begin{table}[H]
\centering\caption{UNDER CONSTRUCTION Distributions of model parameters under the scenarios of weak and moderate competition. The index $i$ runs through all $1,\dots, S$ species.}\vspace{0.2cm}
\begin{tabular}{l|l|l|l}
Parameter     & Description & Weak Competition             & Moderate Competition                                         \\ \hline
$R_i$         & innate growth rate, see \S\ref{ecoevo}  & $\mathrm{Unif}(\mathtt{0.0},\ \mathtt{1.0})$ & \text{same}    \\
$\theta_i$    & abiotic optimum & $\mathrm{Unif}(\mathtt{-1.0},\ \mathtt{1.0})$            & \text{same}                  \\
$a_i$         & strength of abiotic selection & $\mathrm{Exp}(\mathtt{0.01})$              & \text{same}                  \\
$c_i$         & strength of competition & $\mathrm{Exp}(\mathtt{0.01})$                    & $\mathrm{Exp}(\mathtt{0.1})$ \\
$w_i$         & niche breadth & $\mathrm{Unif}(\mathtt{0.01},\ \mathtt{1.0})$              & \text{same}                  \\
$U_i$         & total niche use & $\mathrm{Unif}(\mathtt{0.01},\ \mathtt{1.0})$            & \text{same}                  \\
$\eta_i$      & segregation variance & $\mathrm{Unif}(\mathtt{1.0e-5}, \ \mathtt{1.0e-3})$ & \text{same}                  \\
$\mu_i$       & mutation rate & $\mathrm{Exp}(\mathtt{0.01})$                              & \text{same}                  \\
$V_i$ & variance of reproductive output & $\mathrm{Exp}(\mathtt{0.01})$            & \text{same}
\end{tabular}\label{par_vals}
\end{table}

With these two sets of model parameters, we simulated our model for \texttt{1.0e3} units of time. For both scenarios, we initialized the abundances to $N_i=$\texttt{5.0e3}, trait means to $\bar x_i=$\texttt{0.0} and heritability to $h^2_i=$\texttt{0.5} for each $i=1,\dots,S$. Temporal dynamics for each scenario are provided in Figure \ref{temporal}. Interaction networks and the distribution of coevolution at $t=$\texttt{1e3} for each scenario are displayed in Figure \ref{net}.

```{r, message=FALSE, echo=F, fig.align='center', fig.cap='\\label{temporal}Temporal dynamics of mean trait (top), heritability (middle) and abundance (bottom) for the scenario of weak competition (left) and strong competition (right). Red lines indicate average trend across species.'}
knitr::include_graphics('community_dynamics.png')
```

## The relation between the strength of ecological interactions and coevolution \label{ecoevo}

Relating our treatment of the niche to modern coexistence theory [@Chesson2000], the absolute competition coefficient $\alpha_{ij}$ becomes a dynamical quantity that can be written as

\begin{equation}
\alpha_{ij}(t)=\frac{c_i}{r_i(t)}\int_{\mathbb R}\int_{\mathbb R}p_i(x,t)p_j(y,t)\mathcal O_{ij}(x,y) dxdy =\frac{c_i}{r_i(t)}\sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right),
\end{equation}

where

\begin{equation}
r_i(t)=R_i-\frac{a_i}{2}\Big((\bar x_i(t)-\theta_i)^2+G_i(t)+\eta_i\Big) + c_i\sqrt{\frac{b_{ii}(t)}{2\pi}}.
\end{equation}

Hence, $dN_i(t)$ can be expressed as

\begin{equation}
dN_i(t)=r_i(t)\left(1-\sum_{j=1}^S\alpha_{ij}(t)N_j(t)\right)N_i(t)dt+\sqrt{V_iN_i(t)}dW_1(t).
\end{equation}

Note that although $r_i(t)$ is referred to in the coexistence literature as the intrinsic growth rate of the population, $R_i$ is a deeper intrinsic quantity. For now we refer to $R_i$ as the _innate_ growth rate. With this connection formally established, researchers may pursue a postmodern coexistence theory that naturally includes the evolutionary dynamics of populations and the effects of demographic stochasticity.

In SM \S\ref{diffuse} we show that the standardized directional selection gradient [sensu @Lande1983] induced by species $j$ on species $i$ can be computed as

\begin{equation}
\beta_{ij}(t)=c_iN_j(t)b_{ij}(t)\big(\bar x_i(t)-\bar x_j(t)\big) \sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right).
\end{equation}

Our notation differs from Lande and Arnold [-@Lande1983] in that subscripts here denote species instead of components of multivariate traits and we drop the prime that distinguishes between selection gradients and standardized selection gradients.

Below we compare interaction networks parameterized by $\alpha_{ij}$ with those parameterized by standardized selection gradients to investigate the correspondence of interaction intensity and coevolutionary change. However, we can already identify one major discrepancy; $\alpha_{ij}$ is maximized when $\bar x_i=\bar x_j$, but $\beta_{ij}=0$ under the same condition. We therefore include in our metric of selection the standardized stabilizing selection gradient $\gamma$ which measures the effect of stabilizing or disruptive selection on phenotypic variance [@Lande1983]. In SM \S\ref{diffuse} we show that the standardized stabilizing selection gradient induced by species $j$ on species $i$ can be computed as

\begin{equation}
\gamma_{ij}(t)=c_iN_j(t)b_{ij}(t)\left(1-b_{ij}(t)\big(\bar x_i(t)-\bar x_j(t)\big)^2\right) \sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right).
\end{equation}

Note that when $b_{ij}^{-1}>(\bar x_i-\bar x_j)^2$, $\gamma_{ij}>0$ and when $b_{ij}^{-1}<(\bar x_i-\bar x_j)^2$, $\gamma_{ij}<0$. To measure the total evolutionary change in species $i$ induced by species $j$, we form the metric $\Psi_{ij}=|\beta_{ij}|+|\gamma_{ij}|$. The top row of Figure \ref{net} displays interaction networks under weak and moderate competition where the edge width connecting species $i$ and $j$ is proportional to $\alpha_{ij}\alpha_{ji}$. The bottom row of Figure \ref{net} displays the distributions of pairwise coevolutionary change, which we measure via $\Psi_{ij}\Psi_{ji}$, under weak and moderate competition.

```{r, echo=F, fig.align='center', fig.cap='\\label{net}UNDER CONSTRUCTION Networks of interspecific interactions (top row) and distributions of pairwise coevolution (bottom row) under the scenarios of weak competition (left) and moderate competition (right) at time $t=$\\texttt{1.0e3}. Node sizes are proportional to population sizes. Edge widths and shade are monotonically increasing functions of pairwise coevolution.'}
knitr::include_graphics('interaction_networks.png')
```

We now make use of the expressions derived for competition coeffecients and selection gradients to investigate their relationship. As a first pass, let us assume all model parameters are equivalent across species and that each species has the same abundance and trait variance. Let us further assume that richness $S$ is large and the distribution of mean trait values is normal with mean $\bar{\bar x}$, variance $V_{\bar X}$ and density $f_{\bar X}$. Such assumptions are typical when deriving analytical results in the field of theoretical coevolutionary community ecology [@Nuismer2012; @Nuismer2018]. If $V_{\bar X}$ is not too large, then we may approximate $r_i$ with

\begin{equation}
\bar r=\int_\mathbb{R}\left(R-\frac{a}{2}\Big((\bar x-\theta)^2+G+\eta\Big)+c\sqrt{\frac{b}{2\pi}}\right)f_{\bar X}(\bar x)d\bar x
=R-\frac{a}{2}\Big((\bar{\bar x}-\theta)^2+V_{\bar X}+G+\eta\Big)+c\sqrt{\frac{b}{2\pi}}
\end{equation}

Write $\alpha(\bar x_i,\bar x_j)=\alpha_{ij}$, $\beta(\bar x_i,\bar x_j)=\beta_{ij}$ and $\gamma(\bar x_i,\bar x_j)=\gamma_{ij}$. In SM \S\ref{coef_grad_moms} we use these assumptions to calculate the first and second order moments describing the joint distribution of competition coefficients and selection gradients across the community. In particular, the correlations between these quantities are found to be:

UNDER CONSTRUCTION

\begin{subequations}

\begin{equation}
\mathrm{Corr}_{f_{\bar X}}(\alpha,\beta)=
\end{equation}

\begin{equation}
\mathrm{Corr}_{f_{\bar X}}(\alpha,\gamma)=
\end{equation}

\begin{equation}
\mathrm{Corr}_{f_{\bar X}}(\gamma,\beta)=
\end{equation}

\end{subequations}

# Conclusion \label{conclusion}

We have introduced an approach to derive models of evolutionary ecology using the calculus of white noise, demonstrated our approach by deriving a model of diffuse coevolution and investigated the relationship between coevolutionary change and competition coefficients, finding that these quantities are weakly correlated with each other. Although our approach has the merit of rigorously synthesizing the dynamics of abundance and phenotypic evolution of populations, there remains nuance details and important limitations that have yet to be confronted for gaining a more thorough and rigorous understanding of ecological communities. We touch on just five of them here.

__Limitations of diffusion limits__

As noted by Feller [-@feller1951], although diffusion limits provide reasonable approximations for large populations, relatively small populations require discrete models. Hence, as a diffusion limit, SPDE (\ref{SPDE}) restricts parameter values to regions that maintain large population sizes. This puts an important restriction on any model developed under this framework by implying populations are not at risk of extinction. Hence, studies of evolutionary rescue and colonization-extinction dynamics must be pursued via a different approach. A natural alternative can be developed utilizing the underlying branching Brownian motion process that SPDE (\ref{SPDE}) is a diffusion limit of (see section \ref{stochastic}). This process explicitly tracks individuals throughout their life-history and captures reproduction as branching events and therefore models population size $n$ as a non-negative integer instead of a continuously varying positive number. In particular, there is no restriction that prevents $n=0$. We avoided use of this process here because the resulting models are less analytically tractable and numerical implementations become computationally expensive for large $n$.

__The genetic architecture and distributions of traits__

Our model of coevolution rests on the assumption of normally distributed traits, which itself depends on a simplistic model of genetic architecture (see sections \ref{deterministic} and \ref{inheritance}). However, real traits are not encoded by an infinite number of loci each contributing an additive infinitesimal effect and mutations are not inherited as normally distributed deviations from parental breeding values. Departures from this model of genetic architecture can produce more complex distributions of breeding values and expressed traits. Such deviations can be reinforced via strong non-Gaussian selection surfaces, including the surface $m(\nu,x)$ we have derived from niche theory, along with non-random mating. Gaussian approximations are convenient since they are defined by their mean and variance. Hence, future work investigating the effects of non-normally distributed traits on the structure and dynamics of ecological communities will need to confront higher moments such as skew and kurtosis, ideally integrating previously established mathematical approaches to derive the dynamics of these higher moments [@Dbarre2015EvolutionOQ].

An alternative approach to breaking the assumption of normally distributed trait values is the development of explicit multilocus models. These models describe the contributions of alleles at particular loci in the genome to the development of quantitative traits. Tracking the fluctuations of allele frequencies then allows theoreticians to investigate the dynamics of phenotypic distributions that deviate from normality. While this approach has a long history in theoretical quantitative genetics [@bulmer1980; @pmid7851785; @Kirkpatrick1727] and coevolutionary theory [@Nuismer2005; @Kopp2006; @Nuismer2007], work to investigate the role of genetic architecture in mediating feedbacks between the dynamics of population abundances and the distributions of traits mediating ecological interactions has only just begun [@Patel2019].

__The distinction of niche overlap and competition__

Past work has demonstrated several limitations in measuring competition via niche overlap [@Abrams1980; @Holt1987]. In particular, Abrams [-@Abrams1980] argues the MacArthur-Levins framework utilized here is unnacceptable when nothing is known about the biological mechanism mediating competition. Hence, we must stress that our approach should not be blindly applied to arbitrary communities, but that a great deal of work interpreting our model with respect to the system is a prerequisite task. For example, in the case of a sub-alpine flowering plant community, the niche axis can be interpreted as day-of-year (under a suitable transformation). Our model then assumes within-season fluctuations in the ensemble of abiotic factors including temperature, precipitation, soil nutrients and pollinator activity induces an emergent abiotic optimal date that maximizes fitness in the absence of competition. This optimum likely differs across species with significantly different physiologies and potentially carries a phylogenetic signal. We therefore expect $\theta_i\neq\theta_j$ for our interpretation with respect to a sub-alpine flower community. Individual plants competing for their optimal date to flower then induce the exploitative competition we are modelling.

Another component of Abrams' critique is the distinction between niche overlap and measurements of competition. We also agree that this is an important distinction and is explicitly captured in our model by decomposing the effect of competition on fitness for individuals in species $i$ as the product of niche overlap and a number $c_i>0$ which we have dubbed the strength of competition. In our model, niche overlap merely determines whether or not an interaction occurs and to what frequency. The strength of competition then determines what effect this interaction has on fitness. A natural generalization is to assume the strength of competition between species $i$ and $j$ can be captured by a function of some other trait values. For example, if $y_i$ represents some other trait in species $i$, potentially morphological (flower shape) or physiological (water use efficiency), determining the outcome of indirect interspecific interactions between competitors (pollinator preference or nutrient availability) we can model the strength of competition between species $i$ and $j$ as a function $f$ of the distribution of this trait in each species. If this trait correlates with niche centers across individuals within species, or if resource utilization curves $u_i$ depend on $y_i$, or if $f$ depends on niche centers, the effect on fitness due to competition can no longer be decomposed multiplicatively into distinct components. Hence, our model also requires the necessary assumptions for this decomposition to hold, which vary from system to system.

In another critique by Holt [-@Holt1987], it is argued that utilizations of qualitatively distinct resources may have non-multiplicative effects on components of fitness. This consideration has also been incorporated into our model. In our derivation of fitness in SM \S\ref{diffuse}, we show that distinct components of the growth rate of populations can be decomposed additively under certain circumstances, though further work is needed to develop a general roadmap for when components of fitness can be decomposed additively or multiplicatively.

__The role of ecological stoichiometry__

Our treatment of both biotic and abiotic selection neglects important chemical constraints of biological reality. For instance, the resource we assume species are competing over is modelled as a static quantity. However, real resources can be dynamic quantities. Such theoretical quantities may reflect abiotic cycles of material and energy or whole trophic layers comprised of prey, hosts or mutualists. Although resource dynamics have been captured theoretically in consumer-resource models [@davidtilman1982], developing a more realistic model of resource competition must incorporate details on the ecophysiology of organisms [@michelloreau2010]. Doing so will help clarify the feedback between the evolution of populations and the ecosystem processes they are a part of.

Using plant-pollinator communities as an example, the role of nitrogen mediating interspecific interactions has been reviewed by David, Storkey and Stevens [-@David2019] and the evolutionary ecology of the nutritional content of nectar has been reviewed by Parachnowitsch, Manson and Sletvold [-@Parachnowitsch2018]. These studies demonstrate the need for further research to understand how soil nutrient availability and organismal ecophysiology structures communities of plants and pollinators. Theoretical pursuits in this directions may profit from interfacing the framework we have outlined here with population-ecosystem models such as that developed by Fridley [-@Fridley2017].

__Accounting for macroevolutionary history__

To understand patterns found in ecological communities, considerations must push beyond microevolutionary and contemporary ecological processes and consider the macroevolutionary dynamics of ancestral lineages leading to extant populations. Returning to our example of sub-alpine flower communities, one can observe a very strict ordering of phenology across broad geographic ranges. In particular, whether in the Colorado Rocky mountains (such as Gothic, Colorado) or on an outlier of the Idaho batholith (such as Kamiak butte near Palouse, Washington), one will almost surely observe a very conspicuous order of flowers emerging in early spring: at the very beginning of the season blooms _Claytonia lanceolata_ followed by _Erythronium grandiflorum_ which precedes _Delphinium nuttallianum_ (B. Week, personal observations). If contemporary phenological coevolution is rampant, why should this pattern be so well preserved across a thousand miles of rugged and diverse terrain? A more parsimoniuous hypothesis suggests the phenology and physiology of these species slowly evolved independently over macroevolutionary time scales to take advantage of the specific conditions available within each of these windows of the flowering season. However, this could not have carried out in the Rocky mountains since this terrain only became habitable just over ten thousand years ago as the glaciers of the Pleistocene began to retreat (Paul CaraDonna, personal communications). Hence, given these considerations, it appears that an understanding of early season phenology patterns should focus on how these communities are assembled as opposed to contemporary evolutionary dynamics. Indeed, recent work testing models of phylogeography ignores the potential for contemporary evolution and instead suggests alpine flower communities tend to follow neutral assembly where flowers merely compete for who can disperse to new habitat first, as opposed to a selective process where a regional species pool is filtered for those species adapted to the newly available habitat [@Marx2017].

Of course microevolutionary and ecological dynamics are not completely irrelevant for understanding patterns in communities that are primarily structured by deep evolutionary processes. In particular, macroevolutionary trait evolution is simply the aggregation of microevolutionary change occuring over large spans of time. This suggests a road forward to connect the theory we have introduced to models of macroevolutionary trait evolution that differs from previous approaches [@Manceau2016]. These previous approaches make the implicit assumption that trait evolution is scale free (ie, self-similar) so that the stochastic evolution of traits on macroevolutionary time scales has the same properties of trait evolution on microevolutionary time scales. In fact, many of these models simply repurpose microevolutionary models by blindly rescaling time from the units of generations to millions of years [@Nuismer2014]. Although some stochastic processes including Brownian motion do exhibit self-similarity, others do not. For example, consider a modification of the Ornstein-Uhlenbeck process defined by the SDE

\begin{equation}
dX_t=a(\theta_t-X_t)dt+bdW_t
\end{equation}

where $a,b>0$, $W_t$ is a standard Brownian motion and $\theta_t$ is itself a Markov process that takes normally distributed jumps centered on its current location at exponentially distributed time intervals. If we assume the rate $\lambda$ at which jumps occur is much smaller than $a$ and the variance in jumping is much larger than $b^2$, then, even though the sample paths of $X_t$ are actually continuous, over long intervals of time sample paths of $X_t$ will begin to appear to have periods of continuity interrupted by an occasional discontinuous jump and thus approach a qualitatively distinct process. These emergent properties can be formally characterized by L&eacute;vy processes and have been successfully employed in comparative phylogenetics to fit phenotypic data from extant populations and the fossil record [@Landis2017]. It would therefore be interesting to investigate whether a separation of time scales argument for the rate of environmental change ($\lambda$) versus the rate of evolutionary and ecological change ($a$) to microevolutionary models derived using our framework can be used to obtain macroevolutionary models that include not only mean trait evolution, but also the evolution of trait variance and abundance. The resulting macroevolutionary models can give rise to novel comparative phylogenetic methods and provide initial conditions for microevolutionary models that capture contemporary dynamics.

__Final remarks__

Although top-down approaches to community ecology, such as the Maximum Entropy Theory of Ecology [@Harte2011], have enjoyed some success in describing community-level patterns [@Harte2014; @Xiao2015], a mechanistic understanding of why these patterns emerge and how they will change remains a formidable task. Such an understanding must take both bottom-up and top-down approaches integrating considerations from the ecophysiology of individual organisms that reveal the economics of interspecific interactions [@Sterner2008], to the phylogeographic history of taxa that sets the stage for contemporary dynamics [@Hickerson2010]. Through connecting these dots we can increase the variance explained in observations of ecological communities by specific mechanisms and come closer to a predictive theory of evolutionary community ecology. Despite the long list of equations derived in this paper, this work takes just one small step towards capturing these many details. However, we hope the theoretical framework outlined here along with the demonstration of its use in modelling competitive communities provides some helpful tools to aid quantitative evolutionary ecologists in reaching such lofty goals.

\newpage

# Supplementary material (SM)

## Sufficient conditions for finite mean, variance and total abundance \label{finite}

We assume, for $m:C_1(\mathbb{R})^+\times\mathbb{R}\to\mathbb{R}$, there exists $\theta\in\mathbb{R}$ and $a,c,r>0$ such that, for each $(f,x)\in C_1(\mathbb{R})^+\times\mathbb{R}$,

\begin{equation}\label{m_ineq}
r-\frac{a}{2}(\theta-x)^2-c\int_\mathbb{R}f(x)dx \leq m(f,x)\leq r-\frac{a}{2}(\theta-x)^2.
\end{equation}

Assume that $\nu_0\in C_1(\mathbb{R})^+$ is twice continuously differentiable and satisfies

\begin{equation}
\int_\mathbb{R}|x|\nu_0(x)dx,\int_\mathbb{R}x^2\nu_0(x)dx<\infty.
\end{equation}

Then

\begin{multline}
\int_\mathbb{R}\big|m(\nu_0,x)\big|\nu_0(x)dx\leq\int_\mathbb{R}\left(r+\frac{a}{2}x^2+c\int_\mathbb{R}\nu_0(y)dy\right)\nu_0(x)dx \\ \leq \left(r+\frac{a}{2}\Big((\theta-\bar x_0)^2+\sigma^2_0\Big)+cN_0\right)N_0,
\end{multline}

where the initial phenotypic mean $\bar x_0$, phenotypic variance $\sigma^2_0$ and total abundance $N_0$ are finite by our assumptions on $\nu_0(x)$. To simplify analysis, we further assume

\begin{equation}
\nu_0(x)=\frac{N_0}{\sqrt{2\pi\sigma_0^2}}\left(-\frac{(x-\bar x_0)^2}{2\sigma_0^2}\right).
\end{equation}

If some function $f(x,t)$ that maps $\mathbb{R}\times[0,\infty)$ to $\mathbb{R}$ satisfies, for each fixed $t\geq0$, $f(\cdot,t)\in C_1(\mathbb{R})^+$, then by $m(f,x)$ we mean $m(f(\cdot,t),x)$ for each $t\geq0$. With this convention, consider the following initial value problem (IVP):

\begin{equation}\label{IVproblem}
\left\{\begin{matrix}
\dot\nu(x,t)=& m(\nu,x)\nu(x,t) + \frac{\mu}{2}\Delta\nu(x,t), & (x,t)\in\mathbb{R}\times(0,\infty)\\
\nu(x,t)=&\nu_0(x), & (x,t)\in\mathbb{R}\times\{0\}
\end{matrix}\right.
\end{equation}

To solve for the lower bound of $\nu(x,t)$, let us assume $m(f,x)=r-\frac{a}{2}(\theta-x)^2-c\int_\mathbb{R}f(x)dx$. Then

\begin{equation}
\bar m(t)=\frac{\int_\mathbb{R}m(\nu,x)\nu(x,t)dx}{\int_\mathbb{R}\nu(x,t)dx}=r-\frac{a}{2}\Big(\big(\theta-\bar x(t)\big)^2+\sigma^2(t)\Big)-cN(t).
\end{equation}

If the solution $\nu(x,t)$ to the IVP (\ref{IVproblem}) is Gaussian, then we can derive differential equations for $\bar x$, $\sigma^2$ and $N$ as in the main text to get

\begin{equation}
\dot{\bar x}=\sigma^2\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right)=a\sigma^2\big(\theta-\bar x\big)
\end{equation}
\begin{equation}
\dot\sigma^2=\mu+2\sigma^4\left(\frac{\partial\bar m}{\partial\sigma^2}-\overline{\frac{\partial m}{\partial\sigma^2}}\right)=\mu-a\sigma^4
\end{equation}
\begin{equation}
\dot N=\bar m N=\Big[r-\frac{a}{2}\Big(\big(\theta-\bar x\big)^2+\sigma^2\Big)-cN\Big]N.
\end{equation}

Under these assumptions, we can calculate the time derivative of $\nu(x,t)$ as

\begin{multline}
\dot\nu(x,t)=\left(\frac{\dot N}{N}+\frac{2\dot{\bar x}(x-\bar x)\sigma^2+\big((x-\bar x)^2-\sigma^2\big)\dot\sigma^2}{2\sigma^4}\right)\nu(x,t) \\
=\left(\bar m+a(x-\bar x)(\theta-\bar x)+\frac{\mu}{2}\frac{\big((x-\bar x)^2-\sigma^2\big)}{\sigma^4}-\frac{a}{2}\big((x-\bar x)^2-\sigma^2\big)\right)\nu(x,t) \\
=\left(r-\frac{a}{2}\Big[(\theta-\bar x)^2+(x-\bar x)^2\Big]+a(x-\bar x)(\theta-\bar x)-cN\right)\nu(x,t)+\frac{\mu}{2}\Delta\nu(x,t) \\
=\left(r-\frac{a}{2}(\theta-x)^2-cN\right)\nu(x,t)+\frac{\mu}{2}\Delta\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\Delta\nu(x,t).
\end{multline}

Thus, IVP (\ref{IVproblem}) admits a Gaussian solution. To get the upper bound on $\nu(x,t)$, one can perform the same series of calculations to find that IVP (\ref{IVproblem}) admits a Gaussian solution when $c=0$. Hence, for each $t\geq0$ $\nu(x,t)$ will have finite phenotypic mean, variance and total abundance for general $m$ satisfying inequality (\ref{m_ineq}).

## The relation between diffusion and convolution with a Gaussian kernel \label{diffconvequiv}

Let $g:\mathbb{R}^d\to\mathbb{R}$ be smooth. Consider the deterministic Cauchy problem

\begin{equation}\label{heateqn}
\left\{\begin{matrix}
\dot f(x,t)=&\Delta f(x,t), & (x,t)\in\mathbb{R}^d\times(0,\infty)\\
f(x,t)=&g(x), & (x,t)\in\mathbb{R}^d\times\{0\}.
\end{matrix}\right.  \tag{SM1.1}
\end{equation}

According to Evans [-@lawrenceevans2010], the fundamental solution of (\ref{heateqn}) is

\begin{equation}
\Phi(x,t)=\frac{1}{(4\pi t)^{d/2}}\exp\left(-\frac{|x|^2}{4t}\right), \ (x,t)\in(0,\infty)\times\mathbb{R}^d, \tag{SM1.2}
\end{equation}

where $|x|=\sqrt{\sum_ix_i^2}$. The solution $f(x,t)$ of PDE (\ref{heateqn}) is then given by the convolution

\begin{equation}
f(x,t)=\int_{\mathbb{R}^d}\Phi(x-y,t)g(y)dy, \ (x,t)\in(0,\infty)\times\mathbb{R}^d. \tag{SM1.3}
\end{equation}

Hence, by the fundamental theorem of calculus,

\begin{equation}
f(x,t)+\int_t^{t+1}\dot f(x,s)ds=f(x,t+1) \tag{SM1.4}
\end{equation}

$$=\int_{\mathbb{R}^d}\Phi(x-y,t+1)g(y)dy=\int_{\mathbb{R}^d}\int_{\mathbb{R}^d}\Phi(x-y,1)\Phi(y-z,t)g(z)dzdy$$

$$=\int_{\mathbb{R}^d}\Phi(x-y,1)f(t,y)dy.$$

In particular,

\begin{equation}
f(x,t)+\int_t^{t+1}\Delta f(x,s)ds=\int_{\mathbb{R}^d}\Phi(1,x-y)f(y,t)dy. \tag{SM1.5}
\end{equation}

## Deterministic dynamics of $\sigma^2(t)$ \label{var_deriv}

Picking up from the main text \S2.1,

\begin{equation}
\dot{\sigma}^2(t)=\frac{d}{dt}\int_\mathbb{R}(x-\bar x(t))^2p(x,t)dx
\end{equation}

$$=\int_\mathbb{R}2(x-\bar x(t))\dot{\bar x}(t)+(x-\bar x(t))^2\dot p(x,t)dx$$

$$=\int_\mathbb{R}(x-\bar x(t))^2\left((m(\nu,x)-\bar m(t))p(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t)\right)dx$$

$$=\int_\mathbb{R}\left((x-\bar x(t))^2-\sigma^2(t)+\sigma^2(t)\right)(m(\nu,x)-\bar m(t))p(x,t)+(x-\bar x(t))^2\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t)dx$$

$$=\mathrm{Cov}_t\Big((x-\bar x(t))^2,m(\nu,x)\Big)+\frac{\mu}{2}\int_\mathbb{R}(x-\bar x(t))^2\frac{\partial^2}{\partial x^2}p(x,t)dx.$$

In particular, when $p(x,t)$ is Gaussian,

\begin{equation}
\frac{\partial^2}{\partial x^2}p(x,t)=\frac{(x-\bar x(t))^2-\sigma^2(t)}{\sigma^4(t)}p(x,t)
\end{equation}

and hence, referring to the well-known moments of the Gaussian distribution, we find

\begin{equation}
\int_\mathbb{R}(x-\bar x(t))^2\frac{\partial^2}{\partial x^2}p(x,t)dx
\end{equation}

$$=\int_\mathbb{R}\frac{(x-\bar x(t))^4-(x-\bar x(t))^2\sigma^2(t)}{\sigma^4(t)}p(x,t)dx=2.$$

In the Gaussian case, we also have

\begin{equation}
2\sigma^4\left(\frac{\partial\bar m}{\partial\sigma^2}-\overline{\frac{\partial m}{\partial\sigma^2}}\right)=2\sigma^4\left(\frac{\partial}{\partial\sigma^2}\int_\mathbb{R}m(\nu,x)p(x,t)dx-\int_\mathbb{R}p(x,t)\frac{\partial}{\partial\sigma^2}m(\nu,x)dx\right)
\end{equation}

$$=2\sigma^4\int_\mathbb{R}\frac{(x-\bar x)^2-\sigma^2}{2\sigma^4}m(\nu,x)p(x,t)dx=\int_\mathbb{R}\left((x-\bar x)^2-\sigma^2\right)\big(m(\nu,x)-\bar m\big)p(x,t)dx$$

$$=\mathrm{Cov}_t\Big((x-\bar x)^2,m\Big).$$

## Numerical evidence of finite moments \label{numerical}

Here we use a numerical argument to suggest, for

\begin{equation}\label{m_ineq}
r-\frac{a}{2}(\theta-x)^2-c\int_\mathbb{R}f(x)dx \leq m(f,x)\leq r-\frac{a}{2}(\theta-x)^2, \ \forall (f,x)\in C_1(\mathbb{R})^+\times\mathbb{R},
\end{equation}

the density process $\nu(x,t)$ defined by SPDE (\ref{}) of the main text satisfies $\int_\mathbb{R}|x|^n\nu(x,t)dx<\infty$ for each $n\in\mathbb{N}$.

## Derivation of SDE for $\bar x$ and $\sigma^2$ \label{SDE_DERIV}

For $\nu(x,t)$ defined in the main text, $h\in C(\mathbb{R})$ and $t\geq0$ we make the following assumptions:

\begin{equation}\label{L1_nu}
\mathbb{E}\left(\int_\mathbb{R}|h(x)|\nu(x,t)dx\right)<\infty,
\end{equation}
\begin{equation}\label{L2_nu}
\mathbb{E}\left(\int_\mathbb{R}h^2(x)\nu(x,t)dx\right)<\infty,
\end{equation}
\begin{equation}\label{L1_deriv}
\mathbb{E}\left(\int_0^t\int_\mathbb{R}\nu(x,s)\Big|h(x)\Big|\left|m(\nu,x)f(x)+\frac{\mu}{2}\Delta f(x)\right|dxds\right)<\infty, \ \forall f\in C_b(\mathbb{R}).
\end{equation}

Put $H(t)=\int_\mathbb{R}h(x)\nu(x,t)dx$. Then, for non-random and non-negative $\nu_1(x,t)$ that is continuous in both arguments and integrable in $x$,
\begin{equation}
\star:=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{E}[H(t+\varepsilon)-H(t)|\nu(x,t)=\nu_1(x,t)]
\end{equation}
\begin{equation}
=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{E}\left[\int_\mathbb{R}h(x)\big(\nu(x,t+\varepsilon)-\nu(x,t)\big)dx\Big|\nu(x,t)=\nu_1(x,t)\right]
\end{equation}
\begin{equation}
=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{E}\left[\int_\mathbb{R}\int_t^{t+\varepsilon}f(x)\left(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)
+\sqrt{V\nu(x,s)}\dot W(x,s)\right)dsdx\Big|\nu(x,t)=\nu_1(x,t)\right].
\end{equation}

By assumption (\ref{L1_deriv}) we can use Fubini's theorem to write, with probability one,

\begin{equation}
\int_\mathbb{R}\int_t^{t+\varepsilon}h(x)\big(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)dsdx=\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\big(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)dxds.
\end{equation}

By assumption (\ref{L2_nu}) we have $\varphi(x,t)=h(x)\sqrt{V\nu(x,t)}$ implies $\varphi\in L^2_C(\mathbb{R}\times[0,\infty))$. Is it true by definition that

\begin{multline}
\mathbb{E}\left[\int_\mathbb{R}\int_t^{t+\varepsilon}h(x)\sqrt{V\nu(x,s)}\dot W(x,s)dsdx\Big|\nu(x,t)=\nu_1(x,t)\right] \\ =\mathbb{E}\left[\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\sqrt{V\nu(x,s)}\dot W(x,s)dxds\Big|\nu(x,t)=\nu_1(x,t)\right]=0?
\end{multline}

Hence,

\begin{equation}
\star=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{E}\left[\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\left(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)\right)dxds\Big|\nu(x,t)=\nu_1(x,t)\right].
\end{equation}

By assumption (\ref{L1_deriv}) we know that there exists a $\delta>0$ such that for each positive $\varepsilon<\delta$ and with probability one

\begin{multline}
\left|\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\left(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)\right)dxds\right| \\
\leq\int_t^{t+\delta}\int_\mathbb{R}\Big|h(x)\Big|\left|m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)\right|dxds<\infty
\end{multline}

almost surely. Thus, by Lebesgue's dominated convergence theorem, the drift component of the process $F(t)$ can be computed as

\begin{multline}
\star=\mathbb{E}\left[\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\left(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)\right)dxds\Big|\nu(x,t)=\nu_1(x,t)\right]\\
=\int_\mathbb{R}h(x)\big(m(\nu_1,x)\nu_1(x,t)+\frac{\mu}{2}\Delta\nu_1(x,t)\big)dx.
\end{multline}

To find an expression for the diffusion component of $H(t)$ set

\begin{multline}\label{star2}
\star\star:=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{V}[H(t+\varepsilon)-H(t)|\nu(x,t)=\nu_1(x,t)] \\
=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{V}\left[\int_\mathbb{R}h(x)\big(\nu(x,t+h)-\nu(x,t)\big)dx\Big|\nu(x,t)=\nu_1(x,t)\right].
\end{multline}

We can rewrite the integral inside expression (\ref{star2}) as

\begin{equation}
\int_\mathbb{R}\int_t^{t+\varepsilon}h(x)\big(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)+\sqrt{V\nu(x,s)}\dot W(x,s)\big)dsdx.
\end{equation}

By our assumption (\ref{L2_nu}) that $\int_\mathbb{R}h^2(x)\nu(x,t)dx<\infty$ for each $t\geq0$, we have $\varphi(x,t)=h(x)\sqrt{V\nu(x,t)}$ implies $\varphi\in L^2_C(\mathbb{R}\times[0,\infty))$. We have already found

\begin{multline}
\mathbb{E}\left[\int_\mathbb{R}\int_t^{t+\varepsilon}f(x)\big(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)+\sqrt{V\nu(x,s)}\dot W(x,s)\big)dsdx\Big|\nu(x,t)=\nu_1(x,t)\right] \\
=\mathbb{E}\left[\int_t^{t+\varepsilon}\int_\mathbb{R}f(x)\big(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)\big)dxds\Big|\nu(x,t)=\nu_1(x,t)\right].
\end{multline}

Then, since $h^2(x)\nu(x,t)$ is integrable by assumption (\ref{L2_nu}), we have

\begin{multline}
\star\star=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{E}\left[\left(\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\sqrt{V\nu(x,s)}\dot W(x,s)dsdx\right)^2\Big| \ \nu(x,t)=\nu_1(x,t)\right] \\
=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{E}\left[\int_t^{t+\varepsilon}\int_\mathbb{R}V h^2(x)\nu(x,s)dxds\Big|\nu(x,t)=\nu_1(x,t)\right].
\end{multline}

Thus, for any $\delta\geq\varepsilon\geq0$, we have, with probability one,

\begin{equation}
\int_t^{t+\varepsilon}\int_\mathbb{R}V h^2(x)\nu(x,s)dxds\leq\int_t^{t+\delta}\int_\mathbb{R}V h^2(x)\nu(x,s)dxds.
\end{equation}

We can therefore use Lebesgue's dominated convergence theorem to justify

\begin{equation}
\star\star=\mathbb{E}\left[\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\int_t^{t+\varepsilon}\int_\mathbb{R}V h^2(x)\nu(x,s)dxds\Big|\nu(x,t)=\nu_1(x,t)\right]=\int_\mathbb{R}V h^2(x)\nu_1(x,s)dxds.
\end{equation}

Thus, using standard SDE notation, we have

\begin{equation}\label{SDE_deriv}
dH(t)=\left(\int_\mathbb{R}f(x)\left(m(\nu,x)\nu(x,t)+\frac{\mu}{2}\Delta\nu(x,t)\right)dx\right)dt+\sqrt{V\int_\mathbb{R}h(x)\nu(x,t)dx} \ dW_H(t)
\end{equation}

where $W_H$ is a standard Brownian motion. In the following subsections we employ this formula under the cases $h(x)=x,x^2$ to obtain SDE for the phenotypic mean and variance.

<!--

### Preliminary results

We begin by reviewing a bit of calculus in Banach spaces. The standard reference for nonlinear functional analysis is Schwartz [-@jacobschwartz1969], but the particular material we need can be found in Kuo [-@h.kuo1975]. Note that the notation used in this subsection diverges from the notation used throughout the main text and rest of the supplementary material.

Let $\mathscr{H}$ denote a real separable Hilbert space with inner product $\langle\cdot,\cdot\rangle$ and norm $\|\cdot\|_\mathscr{H}=\sqrt{\langle\cdot,\cdot\rangle}$. Let $\|\cdot\|_\mathbf{B}$ be a measurable semi-norm on $\mathscr{H}$ and denote $\mathbf{B}$ the completion of $\mathscr{H}$ with respect to $\|\cdot\|_\mathbf{B}$. We denote $\mathbf{B}^*$ the restriction of the members of the dual of $\mathbf{B}$ to $\mathscr{H}$ and write the associated duality as $(\cdot,\cdot)$. We write $\iota:\mathscr{H}\hookrightarrow\mathbf{B}$ as the inclusion map from $\mathscr{H}$ into $\mathbf{B}$ and $\kappa:\mathbf{B}^*\hookrightarrow\mathscr{H}^*$ as the embedding by restriction of $\mathbf{B}^*$ into $\mathscr{H}^*$. We make the identification $\mathscr{H}^*\equiv\mathscr{H}$ and $\mathbf{B}^*\equiv\kappa\mathbf{B}^*$ so that 
\begin{equation}
\mathbf{B}^*\subset\mathscr{H}\subset\mathbf{B}.
\end{equation}
Then $(\iota,\mathscr{H},\mathbf{B})$ is referred to as an _abstract Wiener space_ [@h.kuo1975]. We define $W$ as a $\mathbf{B}$-valued Wiener process on $(\Omega,\mathscr{F},P)$. To save space, we refer to Kuo [-@kuo1972; -@h.kuo1975] and Dawson [-@Dawson1975] for precise definitions of this process. Let $F:\mathbf{B}\to\mathbb{R}$ be defined in a neighborhood $U$ of $\mathfrak{x}_0\in\mathbf{B}$. Then $F$ is said to be _Fr&eacute;chet differentiable_ in _$\mathscr{H}$-directions_ if there exists an element $b(\mathfrak{x})\in\mathscr{H}$ such that
\begin{equation}
\left|F(\mathfrak{x}+h)-F(\mathfrak{x})-\langle b(\mathfrak{x}),h\rangle\right|=o(\|h\|_\mathscr{H})
\end{equation}
for small $h$. When it exists, $b(\mathfrak{x})$ is unique and we denote it by $DF(\mathfrak{x})$. Second Fr&eacute;chet differentiability in $\mathscr{H}$-directions is defined in an analogous manner. We denote by $\mathscr{L}_1(\mathscr{H})$ the set of all trace class operators on $\mathscr{H}$. We now recite the infinite-dimensional It&ocirc; lemma [established by @kuo1972; but our notation is inspired by @Dawson1975]:

> __Lemma__ Let $F$ denote a real-valued function on $\mathbf{B}$. Suppose that $F$ is twice Fr&eacute;chet differentiable in $\mathscr{H}$-directions with $D^2F(x)\in\mathscr{L}_1(\mathscr{H})$ for each $x\in\mathbf{B}$, $DF$ is continuous from $\mathbf{B}$ into $\mathscr{H}$ and $D^2F$ is continuous from $\mathbf{B}$ into $\mathscr{L}_1(\mathscr{H})$. Then, if
\begin{equation}
\mathfrak{X}(t)=\mathfrak{x}_0+\int_0^t\zeta(s)dW_s+\int_0^t\sigma(s)ds
\end{equation}
> where $\zeta(t)$ is a bounded linear nonanticipating transformation on $\mathbf{B}$ and $\sigma(t)\in\mathscr{H}$ is nonanticipating,
\begin{multline}\label{ito_out}
F(\mathfrak{X}(t))=F(\mathfrak{x}_0)+\int_0^t(\zeta^*(s)DF(\mathfrak{X}(s)),dW_s) \\
+\int_0^t\left[\langle DF(\mathfrak{X}(s)),\sigma(s)\rangle
+\frac{1}{2}\mathrm{trace}\left\{\zeta^*(s)D^2F(\mathfrak{X}(s))\zeta(s)\right\}\right]ds,
\end{multline}
> where $^*$ denotes the adjoint operator when restricted to $\mathscr{H}$.

### Applying the infinite-dimensional It&ocirc; lemma

In this subsection we unpack the above result for the situation described in main text. As we progress, we show how each component involved in expressing equation (\ref{ito_out}) relates to a particular component in the main text. Once these connections have been made, we rewrite equation (\ref{ito_out}) using the simple heuristics presented in the main text. The following subsection then applies this simplification to derive SDE for the first and second moments of $\nu(x,t)$.

For the situation in the main text, we set $\mathscr{H}=L^2(\mathbb{R})$ and $\mathbf{B}^*=C_1(\mathbb{R})$, where $C_1(\mathbb{R})$ denotes the set of bounded, continuous, integrable functions on $\mathbb{R}$. This is sensible because, if $\varphi\in C_1(\mathbb{R})$, then, by H&ouml;lder's inequality,

\begin{equation}
\int_\mathbb{R}\varphi^2(x)dx\leq\left(\int_\mathbb{R}|\varphi(x)|dx\right)\left(\sup_{x\in\mathbb{R}}|\varphi(x)|\right)<\infty.
\end{equation}

Hence $C_1(\mathbb{R})\subset L^2(\mathbb{R})$ and, for $\varphi,\varphi_1,\varphi_2\in C_1(\mathbb{R})$,

\begin{subequations}
\begin{equation}
\langle\varphi_1,\varphi_2\rangle=(\varphi_1,\varphi_2)=\int_\mathbb{R}\varphi_1(x)\varphi_2(x)dx,
\end{equation}
\begin{equation}
\|\varphi\|_\mathscr{H}=\|\varphi\|_\mathbf{B}=\sqrt{\int_\mathbb{R}\varphi^2(x)dx}.
\end{equation}
\end{subequations}

As a $\mathbf{B}$-valued process, $W_t$ can be thought of as a random continuous linear functional over a space that includes $C_1(\mathbb{R})$ for each $t>0$. In particular, for each $\varphi_1,\varphi_2\in C_1(\mathbb{R})$,

\begin{subequations}\label{Wprops}
\begin{equation}
\mathbb{E}\Big(W_t(\varphi_1)\Big)=\mathbb{E}\Big(W_t(\varphi_2)\Big)=0,
\end{equation}
\begin{equation}
\mathbb{C}\Big(W_t(\varphi_1),W_t(\varphi_2)\Big)=t\int_\mathbb{R}\varphi_1(x)\varphi_2(x)dx.
\end{equation}
\end{subequations}

We also write $W_t$ as the extension of $W_t$ to $L^2(\mathbb{R})$, defined by equations (\ref{Wprops}), but for any $\varphi_1,\varphi_2\in L^2(\mathbb{R})$ and under the identity $\varphi_1\equiv\varphi_2$ whenever $\|\varphi_1-\varphi_2\|_\mathscr{H}=0$. Hence, referring to $L^2_C(\mathbb{R}\times[0,\infty))$ and $\mathbf{W}$ from the main text, if $\varphi\in L^2_C(\mathbb{R}\times[0,\infty))$ does not depend on time, then, setting $\varphi'(x)=\varphi(x,t)$ for each $x\in\mathbb{R}$ and $t\geq0$, $W_t(\varphi')=\mathbf{W}_t(\varphi)$. Hence, if $\Phi(t)\in L^2(\mathbb{R})$ for each $t\geq0$ such that $\Phi(t)(x)=\varphi(x,t)$ and is continous in $t$, then $\varphi\in L^2_C(\mathbb{R}\times[0,\infty))$ and, using the heuristics presented in the main text,

\begin{equation}
\int_0^t(\Phi(s),dW_s)=\int_0^t\int_\mathbb{R}\varphi(x,s)\dot W(x,s)dxds.
\end{equation}

Referring to $\nu(x,t)$ from the main text, we set $\mathfrak{X}(t)(x)=\nu(x,t)$. This implies

\begin{equation}
\zeta(t)\varphi(x)=\zeta^*(t)\varphi(x)=\varphi(x)\sqrt{V\nu(x,t)}
\end{equation}

and

\begin{equation}
\sigma(t)=m(n,x)\nu(x,t)+\frac{\mu}{2}\Delta\nu(x,t).
\end{equation}

Assume that, for some $f\in L^2(\mathbb{R})$, $F(\varphi)=\int_\mathbb{R}f(x)\varphi(x)dx$ for each $\varphi\in L^2(\mathbb{R})$. Then $DF(\varphi)=f$ and $D^2F(\varphi_1)\varphi_2=f$ for each $\varphi,\varphi_1,\varphi_2\in L^2(\mathbb{R})$. Hence, equation (\ref{ito_out}) can be written as

\begin{multline}
\int_\mathbb{R}f(x)\nu(x,t)dx=\int_\mathbb{R}f(x)\nu(x,0)dx+\int_0^t\int_\mathbb{R}f(x)\sqrt{V\nu(x,t)}\dot W(x,s)dxds\\+\int_0^t\left[\int_\mathbb{R}f(x)\left(m(n,x)\nu(x,t)+\frac{\mu}{2}\Delta\nu(x,t)\right)dx+\frac{1}{2}\mathrm{trace}\{\mathfrak{G}(s)\}\right]ds,
\end{multline}

where $\mathfrak{G}(t)\varphi(x)=V f(x)\nu(x,t)\varphi(x)$ for each $\varphi\in L^2(\mathbb{R})$. Note that $\mathfrak{G}(t)$ does not have eigenvalues (ie, its resolvent is empty) for $t\geq0$. Hence $\mathrm{trace}\{\mathfrak{G}(t)\}=0$ for each $t\geq0$. This brings us to the desired formula for $f\in L^2(\mathbb{R}))$. However, when $f(x)=x,x^2$, $f\notin L^2(\mathbb{R})$.

UNDER CONSTRUCTION

-->

### Derivation for trait mean

Setting $\tilde x(t)=\int_\mathbb{R}x\nu(x,t)$ and applying formula (\ref{SDE_deriv}) provides

\begin{equation}
d\tilde x=\left(\overline{xm}N+\frac{\mu}{2}\int_\mathbb{R}x\Delta\nu dx\right)dt+\sqrt{V\overline{x^2}N}d\tilde W_2,
\end{equation}

where,

\begin{equation}
d\tilde{ W}_2=d\hat{\mathbf W}_{\sqrt{V x^2\nu}}=\frac{1}{\sqrt{V\overline{x^2}N}}\int_\mathbb{R}x\sqrt{V\nu(x,t)}\dot W(x,t)dxdt.
\end{equation}

Using It&ocirc;'s (finite-dimensional) quotient rule on $\bar x=\tilde x/N$, we obtain

\begin{equation}
d\bar x = d\left(\frac{\tilde x}{N}\right)=\frac{\tilde x}{N}\left(\frac{d\tilde x}{\tilde x}-\frac{dN}{N}-\frac{d\tilde x}{\tilde x}\frac{dN}{N}+\Big(\frac{dN}{N}\Big)^2\right)
\end{equation}

\begin{equation*}
= \frac{d\tilde x}{N}-\bar x\frac{dN}{N}-\frac{d\tilde x}{N}\frac{dN}{N}+\bar x\Big(\frac{dN}{N}\Big)^2
\end{equation*}

\begin{equation*}
= \overline{xm}dt+\frac{\mu}{2}\int_\mathbb{R}x\Delta p(x,t)dx+\frac{V}{2}\overline{x^2} +\sqrt{V\frac{\overline{x^2}}{N}}d\tilde W_2-\bar x\bar mdt
-V\sqrt{\frac{\bar x}{N}}d W_1-V\frac{\bar x}{N}dt+V\frac{\bar x}{N}dt
\end{equation*}

\begin{equation*}
= \left(\mathrm{Cov}_t(x,m)+\frac{\mu}{2}\int_\mathbb{R}x\Delta p(x,t)dx\right)dt
+\sqrt{V\frac{\overline{x^2}}{N}}d\tilde W_2-\bar x\sqrt{\frac{V}{N}}d W_1.
\end{equation*}

Note that

\begin{equation}
\sqrt{V\frac{\overline{x^2}}{N}}d\tilde W_2-\bar x\sqrt{\frac{V}{N}}d W_1=\frac{1}{N}\int_\mathbb{R}x\sqrt{V\nu(x,t)}\dot W(x,t)dx-\frac{\bar x}{N}\int_\mathbb{R}\sqrt{V\nu(x,t)}\dot W(x,t)dx
\end{equation}

\begin{equation*}
=\int_\mathbb{R}\frac{x-\bar x}{N}\sqrt{V\nu(x,t)}\dot W(x,t)dx
\end{equation*}

and

\begin{equation}
\mathbb{V}\left(\int_\mathbb{R}\frac{x-\bar x}{N}\sqrt{V\nu(x,t)}\dot W(x,t)dx\right)=\frac{V}{N}\int_\mathbb{R}(x-\bar x)^2p(x,t)dx=V\frac{\sigma^2}{N}.
\end{equation}

Hence, by setting

\begin{equation}
d W_2=\frac{\int_\mathbb{R}\frac{(x-\bar x)}{N}\sqrt{V\nu(x,t)}\dot W(x,t)dx}{\sqrt{V\sigma^2/N}}
\end{equation}

we can write

\begin{equation}
d\bar x=\left(\mathrm{Cov}_t(x,m)+\frac{\mu}{2}\int_\mathbb{R}x\Delta p(x,t)dx\right)dt
+\sqrt{V\frac{\sigma^2}{N}}d W_2.
\end{equation}

### Derivation for trait variance

UNDER CONSTRUCTION

Setting $\tilde{\tilde\sigma}^2(t)=\int_{\mathbb R} x^2\nu(x,t)dx$ and applying formula (\ref{SDE_deriv}) provides

\begin{equation}
d\tilde{\tilde\sigma}^2=\left(\overline{x^2m}N+\frac{\mu}{2}\int_\mathbb{R}x^2\Delta\nu dx\right)dt+\sqrt{V\overline{x^4}N}d\tilde W_3
\end{equation}

where

\begin{equation}
d\tilde{W}_3=d\hat{\mathbf W}_{\sqrt{V x^4\nu}}
=\frac{1}{\sqrt{V\overline{x^4}N}}\int_\mathbb{R}x^2\sqrt{V\nu(x,t)}\dot W(x,t)dx.
\end{equation}

Use It&ocirc;'s quotient rule on $\tilde\sigma^2=\tilde{\tilde\sigma}^2/N$ to obtain:

\begin{equation}
d\tilde\sigma^2 = d\left(\frac{\tilde{\tilde\sigma}^2}{N}\right)=\frac{\tilde{\tilde\sigma}^2}{N}\left(\frac{d\tilde{\tilde\sigma}^2}{\tilde{\tilde\sigma}^2}-\frac{dN}{N}-\frac{d\tilde{\tilde\sigma}^2}{\tilde{\tilde\sigma}^2}\frac{dN}{N}+\Big(\frac{dN}{N}\Big)^2\right)
\end{equation}

\begin{equation}
=\frac{d\tilde{\tilde\sigma}^2}{N}-\tilde\sigma^2\frac{dN}{N}-\frac{d\tilde{\tilde\sigma}^2}{N}\frac{dN}{N}+\tilde{\tilde\sigma}^2\left(\frac{dN}{N}\right)^2.
\end{equation}

Table 1 of the main text implies

\begin{equation}
\left(\sqrt{V\frac{\overline{x^4}}{N}}d\tilde W_3\right)\left(\sqrt\frac{V}{N}d W_1\right)=\frac{\int_\mathbb{R}\sqrt{x^4V\nu(x,t)}\sqrt{V\nu(x,t)}dx}{N^2}dt=V\frac{\overline{x^2}}{N}dt.
\end{equation}

and hence

\begin{equation}
d\tilde\sigma^2=\left(\overline{x^2m} + \frac{\mu}{2}\int_\mathbb{R}x^2\Delta p(x,t)dx\right)dt+\sqrt{V\frac{\overline{x^4}}{N}}d\tilde W_3
-\tilde\sigma^2 \bar mdt+\tilde\sigma^2\sqrt{\frac{V}{N}}d W_1
-V\frac{\overline{x^2}}{N}dt-\tilde\sigma^2\frac{V}{N}dt.
\end{equation}

\begin{equation}
=\left(\mathrm{Cov}_t\Big(x^2,m\Big) + \frac{\mu}{2}\int_\mathbb{R}x^2\Delta pdx \right)dt+\sqrt{V\frac{\overline{x^4}}{N}}d\tilde W_3-\tilde\sigma^2\sqrt\frac{V}{N}d W_1.
\end{equation}

Setting $F(y,z)=y-z^2$, use It&ocirc;'s formula on $\sigma^2=F(\tilde\sigma^2,\bar x)=\tilde\sigma^2-\bar x^2$ to obtain:

\begin{equation}
d\sigma^2=d\tilde\sigma^2-2\bar xd\bar x-(d\bar x)^2
\end{equation}

$$=\left(\mathrm{Cov}_t\Big(x^2,m\Big) + \mu \right)dt+\sqrt\frac{V\overline{x^4}}{N}d\tilde W_3-\tilde\sigma^2\sqrt\frac{V}{N}d W_1$$

$$-2\bar x\left(\mathrm{Cov}_t(x,m)dt+\sqrt{\frac{V\sigma^2}{N}}d W_2\right)-\left(\mathrm{Cov}_t(x,m)dt+\sqrt{\frac{V\sigma^2}{N}}d W_2\right)^2$$

$$=\left(\mathrm{Cov}_t\Big(x^2-2\bar xx,m\Big) + \mu \right)dt+\sqrt\frac{V\overline{x^4}}{N}d\tilde W_3-\tilde\sigma^2\sqrt\frac{V}{N}d W_1-2\bar x\sqrt{\frac{V\sigma^2}{N}}d W_2-\left(\frac{V\sigma^2}{N}\right)dt$$

$$=\left(\mathrm{Cov}_t\Big(x-\bar x)^2,m\Big) + \mu - \frac{V\sigma^2}{N}\right)dt+\sqrt\frac{V\overline{x^4}}{N}d\tilde W_3-\tilde\sigma^2\sqrt\frac{V}{N}d W_1-2\bar xd W_2$$

In light of

\begin{equation}
\sqrt\frac{V\overline{x^4}}{N}d\tilde W_3-\tilde\sigma^2\sqrt\frac{V}{N}d W_1-2\bar xd W_2=\frac{1}{N}\int_\mathbb{R}\left(x^2-\tilde\sigma^2-2\bar x(x-\bar x)\right)\sqrt{V\nu(x,t)}\dot W(x,t)dx
\end{equation}

$$=\frac{1}{N}\int_\mathbb{R}\left((x-\bar x)^2-\sigma^2\right)\sqrt{V\nu(x,t)}\dot W(x,t)dx$$

we set

\begin{equation}
\mathscr{V}^2=\mathbb{V}\left(\frac{1}{N}\int_\mathbb{R}\left((x-\bar x)^2-\sigma^2\right)\sqrt{V\nu(x,s)}\dot W(x,s)dx\right)
\end{equation}

\begin{equation}
=\frac{V}{N}\left(\int_\mathbb{R}((x-\bar x)^4-2(x-\bar x)^2\sigma^2+\sigma^4)p(x,t)dx\right)
\end{equation}

\begin{equation}
=\frac{V}{N}\left(\overline{(x-\bar x)^4}-\sigma^4\right)
\end{equation}

and

\begin{equation}
d W_3=\frac{1}{N\mathscr{V}}\int_\mathbb{R}\left((x-\bar x)^2-\sigma^2\right)\sqrt{V\nu(x,t)}\dot W(x,t)dx
\end{equation}

so that

\begin{equation}
d\sigma^2=\left(\mathrm{Cov}_t\Big((x-\bar x)^2,m\Big)+\frac{\mu}{2}\int_\mathbb{R}(x-\bar x)^2\Delta\nu(x,t)dx-V\frac{\sigma^2}{N}\right)dt+\sqrt{V\frac{\overline{(x-\bar x)^4}-\sigma^4}{N}}d W_3.
\end{equation}

Table \ref{mult} of the main text implies

\begin{equation}
d W_1d W_2=\frac{\int_\mathbb{R}(x-\bar x)\nu(x,t)dx}{\sqrt{N\sigma^2}}=0,
\end{equation}

\begin{equation}
d W_1d W_3=\frac{\int_\mathbb{R}\big((x-\bar x)^2-\sigma^2\big)\nu(x,t)dx}{\sqrt{\overline{(x-\bar x)^4}-\sigma^4}}=0,
\end{equation}

\begin{equation}
d W_2d W_3=\frac{\int_\mathbb{R}(x-\bar x)\big((x-\bar x)^2-\sigma^2\big)\nu(x,t)dx}{\sqrt{\sigma^2\big(\overline{(x-\bar x)^4}-\sigma^4\big)}}
=\frac{N\overline{(x-\bar x)^3}}{\sqrt{\sigma^2\big(\overline{(x-\bar x)^4}-\sigma^4\big)}}.
\end{equation}

In particular, when $p$ is a Gaussian curve $d W_2d W_3=0$.

## Derivation of diffuse coevolution model \label{diffuse}

In this section we provide a derivation of our model of diffuse coevolution driven by competition. Since most of the work in this derivation has already been completed in Supplementary Material \S\ref{SDE_DERIV}, we focus here on deriving the Malthusian fitness $m$ as a function of trait value $x$. We begin with discrete populations of individuals. In particular, we begin by assuming population size $n_i$ is an integer for each species $i=1,\dots,S$ before passing to the large population size limit.

The reduction in fitness for an individual of species $i$ caused by competition is captured multiplicatively by $0<C_i\leq1$. Biologically this assumes all competitors affect individuals of a given species equally by consuming the same amount of resources. This is a mean-field interaction since any individual that consumes resources has an effect on the fitness of all other individuals competing for the same resources. Denote by $x_{ij}$ the trait value of the $j$-th individual belonging to species $i$. The set of trait values across all individuals in the community at time $t\geq0$ is written $X=\{x_{ij}\}$. We denote by $\mathscr{B}_{ij}$ a function that maps $X_t$ to the cumulative effect of all competitive interactions on the fitness of the $j$-th individual in species $i$. Since individuals do not compete with themselves the net multiplicative effects on fitness of both interspecific and intraspecific competition on the $j$-th individual in species $i$ is summarized by

\begin{equation}
\mathscr B_{ij}(X)=C_i^{\sum_{l\neq j}\mathcal{O}_{ii}(x_{ij},x_{il})+\sum_{k\neq i}\sum_{l=1}^{n_k}\mathcal{O}_{ik}(x_{ij},x_{kl})},
\end{equation}

where $\mathcal{O}_{ij}$, defined in the main text, measures the overlap in resource use between individuals of species $i$ and $j$ as a function of their niche-centers. Writing $\mathscr W_{ij}(X)$ as the average number of offspring left by the $j$-th individual of species $i$, we have

\begin{equation}
\mathscr W_{ij}(X)=\mathscr{A}_i(x_{ij})\mathscr{B}_{ij}(X),
\end{equation}

where $\mathscr{A}_i(x)=\int_\mathbb{R}e_i(\zeta)u_i(\zeta,x)d\zeta$ accounts for abiotic selection and $e_i$ has been defined in the main text. The Malthusian fitness $m$ is related to $\mathscr W$ by $m=\ln\mathscr{W}$ [@9781932846126]. Hence,

\begin{multline}
m_{ij}(X)=\ln\mathscr{A}_i(x_{ij})+\ln\mathscr{B}_{ij}(X) \\
=\ln\mathscr{A}_i(x_{ij})+\left(\sum_{l\neq j}\mathcal{O}_{ii}(x_{ij},x_{il})+\sum_{k\neq i}\sum_{l=1}^{n_k}\mathcal{O}_{ik}(x_{ij},x_{kl})\right)\ln C_i.
\end{multline}

In the diffusion limit we rescale individual mass of species $i$ by $1/n_i$. We set $\pmb{n}=(n_1,\dots,n_S)$ and write the rescaled limit of the Malthusian fitness as

\begin{equation}
m^{(\pmb{n})}_{ij}(X_t)=\ln\mathscr{A}_i(x_{ij})+\left(\frac{1}{n_i}\sum_{l\neq j}\mathcal{O}_{ii}(x_{ij},x_{il})+\sum_{k\neq i}\frac{1}{n_k}\sum_{l=1}^{n_k}\mathcal{O}_{ik}(x_{ij},x_{kl})\right)\ln C_i.
\end{equation}

Denoting the density process describing the distribution of trait values in species $i$ by $\nu_i(x,t)$ and writing $\pmb\nu=(\nu_1,\dots,\nu_S)$, the diffusion limit of Malthusian fitness for trait value $x$ in species $i$ at time $t\geq0$ becomes

\begin{multline}
m_i(\pmb\nu,x)=\ln\mathscr{A}_i(x_{ij}) \\ +\left(\int_\mathbb{R}\mathcal{O}_{ii}(x_{ij},x)\nu_i(x,t)dx-\frac{\int_\mathbb{R}\mathcal{O}_{ii}(x_{ij},x)\nu_i(x,t)dx}{\int_\mathbb{R}\nu_i(x,t)dx}+\sum_{k\neq i}\int_\mathbb{R}\mathcal{O}_{ik}(x_{ij},x)\nu_k(x,t)dx\right)\ln C_i \\
=\ln\mathscr{A}_i(x_{ij})+\left(-\bar{\mathcal{O}}_{ii}(x,t_{ij})+\sum_{k=1}^{S}N_k(t)\bar{\mathcal{O}}_{ik}(x,t_{ij})\right)\ln C_i.
\end{multline}

When $\nu_i(x,t)$ and $\nu_j(x,t)$ are Gaussian curves, the expected niche overlap between an individual at niche location $x$ with individuals in species $j$ is

\begin{equation}
\bar{\mathcal{O}}_{ij}(x,t)=\frac{\int_\mathbb{R}\mathcal O_{ij}(x,x_j) \nu_j(x_j,t)dx_j}{\int_\mathbb{R}\nu_j(x_j,t)dx_j}
=\frac{1}{\sqrt{2\pi(w_i+w_j+{\sigma_j}^2(t))}}\exp\left(-\frac{(x-\bar x_j(t))^2}{2(w_i+w_j+{\sigma_j}^2(t))}\right).
\end{equation}

Setting

\begin{subequations}
\begin{align}
r_i = & \ \ln \left(\frac{Q_iU_i}{\sqrt{1+A_iw_i}}\right), \\
a_i = & \ \frac{A_i}{1+A_iw_i}, \\
\tilde b_{ij}(t) = & \ \frac{1}{w_i+w_j+\sigma_j^2(t)}, \\
c_i = & -\ln C_i,
\end{align}
\end{subequations}

we get

\begin{equation}\label{malth_fit}
m_i(\pmb\nu,x)=r_i-\frac{a_i}{2}(x-\theta_i)^2+c_i\left(\sqrt{\frac{\tilde b_{ii}(t)}{2\pi}}e^{-\frac{\tilde b_{ii}(t)}{2}(x-\bar x_i(t))^2}-\sum_{j=1}^SN_j(t)\sqrt{\frac{\tilde b_{ij}(t)}{2\pi}}e^{-\frac{\tilde b_{ij}(t)}{2}(x-\bar x_j(t))^2}\right).
\end{equation}

For the remainder of the derivation we suppress notation indicating dependency on $t$ and $x$. From (\ref{malth_fit}) we calculate

\begin{equation}
\frac{\partial m_i}{\partial\bar x_i}=c_i\tilde b_{ii}(x-\bar x_i)\sqrt{\frac{\tilde b_{ii}}{2\pi}}e^{-\frac{\tilde b_{ii}}{2}(x-\bar x_i)^2}
\end{equation}

\begin{equation}
\frac{\partial m_i}{\partial G_i}=\frac{c_i\left((x-\bar x_i)^2-{\sigma_i}^2-2w_i\right)}{2({\sigma_i}^2+2w_i)^2}\sqrt{\frac{\tilde b_{ii}}{2\pi}}e^{-\frac{\tilde b_{ii}}{2}(x-\bar x_i)^2}
\end{equation}

and hence

\begin{equation}
\overline{\frac{\partial m_i}{\partial\bar x_i}}=0,
\end{equation}

\begin{multline}
\overline{\frac{\partial m_i}{\partial G_i}}=\frac{c_i}{2({\sigma_i}^2+2w_i)^2}\left(\frac{({\sigma_i}^2+2w_i){\sigma_i}^2}{2(w_i+{\sigma_i}^2)}-{\sigma_i}^2-2w_i\right)\sqrt{\frac{b_{ii}}{2\pi}} \\
=-\frac{c_i}{2({\sigma_i}^2+2w_i)^2}\frac{({\sigma_i}^2+2w_i)^2}{2({\sigma_i}^2+w_i)}\sqrt{\frac{b_{ii}}{2\pi}}=-\frac{c_ib_{ii}}{2}\sqrt{\frac{b_{ii}}{2\pi}},
\end{multline}

where

\begin{equation}
b_{ij}=\frac{1}{w_i+w_j+{\sigma_i}^2+{\sigma_j}^2}.
\end{equation}

The average fitness for species $i$ is

\begin{equation}
\bar m_i=r_i-\frac{a_i}{2}\Big((\bar x_i-\theta_i)^2+G_i+\eta_iN_i\Big) + c_i\left(\sqrt{\frac{b_{ii}}{2\pi}}-\sum_{j=1}^SN_j\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\right).
\end{equation}

Hence

\begin{equation}
\frac{\partial\bar m_i}{\partial\bar x_i}=a_i(\theta_i-\bar x_i)-c_i\sum_jN_jb_{ij}(\bar x_j-\bar x_i)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2},
\end{equation}

\begin{equation}
\frac{\partial\bar m_i}{\partial G_i}=-\frac{a_i}{2}-\frac{c_i}{2}\left(b_{ii}\sqrt\frac{b_{ii}}{2\pi}-\sum_{j=1}^SN_jb_{ij}\left(1-b_{ij}(\bar x_i-\bar x_j)^2\right)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\right).
\end{equation}

Applying equations (\ref{N}), (\ref{xbarfinal}) and (\ref{Gfinal}) recovers system (\ref{comm_dynamics}) of the main text.

## Moments of competition coefficients and selection gradients \label{coef_grad_moms}

The calculations are straightforward applications of Gaussian products. That is, if

\begin{equation}
f_1(x)=\frac{1}{\sqrt{2\pi\sigma_1^2}}\exp\left(-\frac{(\mu_1-x)^2}{2\sigma_1^2}\right), \ f_2(x)=\frac{1}{\sqrt{2\pi\sigma_2^2}}\exp\left(-\frac{(\mu_2-x)^2}{2\sigma_2^2}\right),
\end{equation}

then

\begin{equation}
f_1(x)f_2(x)=\frac{1}{\sqrt{2\pi(\sigma_1^2+\sigma_2^2)}}\exp\left(-\frac{(\mu_1-\mu_2)^2}{2(\sigma_1^2+\sigma_2^2)}\right)\frac{1}{\sqrt{2\pi\tilde\sigma^2}}\exp\left(-\frac{(\tilde\mu-x)^2}{2\tilde\sigma^2}\right),
\end{equation}

where

\begin{equation}
\tilde\mu=\frac{\sigma_2^2\mu_1+\sigma_1^2\mu_2}{\sigma_1^2+\sigma_2^2}, \ \tilde\sigma^2=\frac{\sigma_1^2\sigma_2^2}{\sigma_1^2+\sigma_2^2}.
\end{equation}

Hence:

\begin{multline}
\bar\alpha=
\int_\mathbb{R}\int_\mathbb{R}\alpha(\bar x_i,\bar x_j)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
=\frac{c}{\bar r}\int_\mathbb{R}\frac{1}{\sqrt{2\pi(b^{-1}+V_{\bar X})}}\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(b^{-1}+V_{\bar X})}\right)f_{\bar X}(\bar x_j)d\bar x_j
=\frac{c/\bar r}{\sqrt{2\pi(b^{-1}+2V_{\bar X})}},
\end{multline}

\begin{multline}
\bar\beta=
\int_\mathbb{R}\int_\mathbb{R}\beta(\bar x_i,\bar x_j)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
= cNb\int_\mathbb{R}(\bar{\bar x}-\bar x_j)\frac{1}{\sqrt{2\pi(b^{-1}+V_{\bar X})}}\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(b^{-1}+V_{\bar X})}\right)f_{\bar X}(\bar x_j)d\bar x_j=0,
\end{multline}

\begin{multline}
\bar\gamma=
\int_\mathbb{R}\int_\mathbb{R}\gamma(\bar x_i,\bar x_j)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
= cNb\int_\mathbb{R}\Big(1-b\big((\bar{\bar x}-\bar x_j)^2+V_{\bar X}\big)\Big)\frac{1}{\sqrt{2\pi(b^{-1}+V_{\bar X})}}\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(b^{-1}+V_{\bar X})}\right)f_{\bar X}(\bar x_j)d\bar x_j \\
=\frac{cNb(1-2bV_{\bar X})}{\sqrt{2\pi(b^{-1}+2V_{\bar X})}}
\end{multline}

\begin{multline}
\mathrm{Var}_{f_{\bar X}}(\alpha)=
\int_\mathbb{R}\int_\mathbb{R}\big(\bar\alpha-\alpha(\bar x_i,\bar x_j)\big)^2f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j\\
=\frac{c^2}{\bar r^2}\int_\mathbb{R}\int_\mathbb{R}\left(\frac{1}{\sqrt{2\pi(b^{-1}+2V_{\bar X})}}-\sqrt{\frac{b}{2\pi}}\exp\left(-\frac{b}{2}(\bar x_i-\bar x_j)^2\right)\right)^2f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
=\frac{c^2}{\bar r^2}\left(\frac{1}{2\pi(b^{-1}+2V_{\bar X})}-\frac{1}{\pi(b^{-1}+2V_{\bar X})}+\sqrt{\frac{b}{2\pi}}\frac{1}{\sqrt{2\pi(b^{-1}+2V_{\bar X})}}\right) \\
=\frac{c^2}{2\pi\bar r^2}\left(\frac{1+2bV_{\bar X}}{b^{-1}+2V_{\bar X}}-\frac{1}{b^{-1}+2V_{\bar X}}\right)=\frac{c^2}{2\pi\bar r^2}\frac{2bV_{\bar X}}{b^{-1}+2V_{\bar X}},
\end{multline}

\begin{multline}
\mathrm{Var}_{f_{\bar X}}(\beta)=
\int_\mathbb{R}\int_\mathbb{R}\big(\bar\beta-\beta(\bar x_i,\bar x_j)\big)^2f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
=\int_\mathbb{R}\int_\mathbb{R}c^2N^2b^2(\bar x_i-\bar x_j)^2\frac{b}{2\pi}\exp\left(-\frac{b}{2}(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
=\sqrt{\frac{b}{2\pi}}c^2N^2b^2\left(b^{-1}+2V_{\bar X}\right),
\end{multline}

\begin{multline}
\mathrm{Var}_{f_{\bar X}}(\gamma)=
\int_\mathbb{R}\int_\mathbb{R}\big(\bar\gamma-\gamma(\bar x_i,\bar x_j)\big)^2f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
=
\end{multline}

\begin{multline}
\mathrm{Cov}_{f_{\bar X}}(\alpha,\beta)=
\int_\mathbb{R}\int_\mathbb{R}\big(\bar\alpha-\alpha(\bar x_i,\bar x_j)\big)\big(\bar\beta-\beta(\bar x_i,\bar x_j)\big)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
=
\end{multline}

\begin{multline}
\mathrm{Cov}_{f_{\bar X}}(\alpha,\gamma)=
\int_\mathbb{R}\int_\mathbb{R}\big(\bar\alpha-\alpha(\bar x_i,\bar x_j)\big)\big(\bar\gamma-\gamma(\bar x_i,\bar x_j)\big)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
=
\end{multline}

\begin{multline}
\mathrm{Cov}_{f_{\bar X}}(\gamma,\beta)=
\int_\mathbb{R}\int_\mathbb{R}\big(\bar\gamma-\gamma(\bar x_i,\bar x_j)\big)\big(\bar\beta-\beta(\bar x_i,\bar x_j)\big)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
=
\end{multline}

\newpage

# References

### 2020-02-09 bbm.jl

@@ -1,0 +1,73 @@
+using Parameters, Statistics, Random, LinearAlgebra, Distributions,
+	DifferentialEquations, StatsBase, StatsPlots, Plots, DataFrames, CSV, Optim
+
+include("/home/bob/Research/White\ Noise/bbm_functions_structs.jl")
+
+#
+# a branching Brownian motion
+#
+
+# number of generations to halt at
+T = 1000
+
+# initial number of individuals
+N₀ = 100
+
+# initial trait values
+x₀=rand(Normal(0.0,1.0),N₀)
+
+# set up initial population
+X = population(x=x₀, x̄=mean(x₀), σ²=var(x₀), N=N₀,
+	r=1.0, a=0.01, θ=0.0, c=0.01, μ=2.0, V=exp(1.0))
+
+##
+## VERY IMPORTANT REQUIREMENT  -->   V >= exp(r)
+##
+
+# always a good idea to inspect a single iteration
+update(X)
+
+# set up history of population
+Xₕ = fill(X,T)
+
+# set up containers for paths of N, x\bar and σ²
+Nₕ = fill(X.N,T)
+x̄ₕ = fill(X.x̄,T)
+σ²ₕ = fill(X.σ²,T)
+
+# simulate
+for i in 2:T
+
+	if Xₕ[i-1].N > 0
+
+		Xₕ[i] = update(Xₕ[i-1])
+
+	else
+
+		Xₕ[i] = Xₕ[i-1]
+
+	end
+
+	Nₕ[i] = Xₕ[i].N
+	x̄ₕ[i] = Xₕ[i].x̄
+	σ²ₕ[i] = Xₕ[i].σ²
+
+end
+
+# plot some results
+
+plot(1:T,Nₕ)
+
+plot(1:T,x̄ₕ)
+
+plot(1:T,σ²ₕ)
+
+# rescale time
+Tₛ = fill(1/N₀,T)
+for i in 2:T
+	Tₛ[i] = Tₛ[i-1]+1/Nₕ[i]
+end
+
+# plotting the rescaled process
+plot(Tₛ,x̄ₕ./.√Nₕ,ylims=(-1.0,1.0))
+plot(Tₛ,σ²ₕ./.√Nₕ)

### 2020-02-09 bbm_functions_structs.jl

@@ -1,0 +1,69 @@
+
+# data type that holds population and parameters
+@with_kw mutable struct population
+    x::Vector{Float64}  # trait values
+    N::Int64            # poulation size
+    x̄::Float64          # mean trait
+    σ²::Float64         # phenotypic variance
+    r::Float64          # intrinsic rate of growth
+    a::Float64          # strength of abiotic selection
+    θ::Float64          # abiotic optimum
+    c::Float64          # strength of competition
+    μ::Float64          # rate of diffusion (mutation)
+    V::Float64          # variance in reproductive output
+end
+
+
+function update(X)
+
+    @unpack x, N, x̄, σ², r, a, θ, c, μ, V = X
+
+    W = fill(0,N)
+
+    for j in 1:N
+
+        # mean fitness of individual j
+        w = exp( r - a*(θ-x[j])^2/2 - c*N )
+
+        # parameterizing the NegativeBinomial
+        p = w/V
+        rr = w^2/(V-w)
+
+        # draw random number of offspring
+        W[j] = rand( NegativeBinomial( rr, p ), 1)[1]
+
+    end
+
+    # total number of offspring
+    Np = sum(W)
+
+    # container for locations of offspring
+    xp = fill(0.0,Np)
+
+    # keeps track of which individual is being born
+    ct = 0
+
+    # loop throug parents
+    for j in 1:N
+
+        # birth each offspring
+        for k in 1:W[j]
+
+            # consider next individual
+            ct += 1
+
+            # draw random trait for this individual
+            xp[ct] = rand( Normal( x[j], √μ ), 1)[1]
+
+        end
+
+    end
+
+    x̄p = mean(xp)
+    σ²p= var(xp)
+
+    Xp = population(x=xp,N=Np,x̄=x̄p,σ²=σ²p,r=r,a=a,θ=θ,c=c,μ=μ,V=V)
+
+    return Xp
+
+end

### 2020-03-20 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
>

---
title: "A White Noise Approach to Evolutionary Ecology"
author: "Bob Week, Steve Krone, Scott L. Nuismer, Luke J. Harmon"
date: "3/1/2020"
abstract: "We derive the dynamics of the distribution of a quantitative character and the abundance of a biological population from a stochastic partial differential equation driven by space-time white noise. In the process we develop a useful set of heuristics to operationalize the powerful, but abstract theory of white noise and measure-valued Markov processes. This approach allows us to compute the full implications of a stochastic process such as demographic stochasticity on phenotypic distributions and abundances of populations. We demonstrate the utility of our approach by deriving a model of diffuse coevolution mediated by exploitative competition for a continuum of resources. Other than trait and abundance distributions, this model predicts interaction networks parameterized by rates of interactions, competition coefficients, and selection gradients. We briefly investigate the relationship between selection gradients and competition coefficients. This illustrative investigation suggests selection gradients can be either positively or negatively correlated with competition coefficients depending on the ratio of interspecific trait variation to intraspecific trait variation. Hence, this approach can contribute to the development of a synthetic theory of evolutionary ecology by formalizing first principle derivations of dynamical equations describing populations and communities which can then be used for rigorous investigations of the relationship between feedbacks of ecological and evolutionary processes and the patterns of diversity they produce."
bibliography: ref.bib
output:
  pdf_document:
    toc: false
    number_sections: true
    fig_width: 4
    fig_height: 3
    fig_caption: true
header-includes:
  - \usepackage{float}
  - \usepackage{mathrsfs}
  - \usepackage{hyperref}
  - \usepackage[left]{lineno}
  - \usepackage{amsmath}
  - \linenumbers
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
#knitr::opts_chunk$set(fig.pos = 'H')
```

# Introduction

Our goal in this manuscript is to develop a rigorous, but accessible approach to synthesize the stochastic dynamics of abundance, mean trait and heritable variation in biological populations for the study of theoretical evolutionary ecology. A primary aim of theoretical evolutionary ecology is the development of mathematical approaches to describe the evolution of populations and their interactions with both the biotic and abiotic environments in which they are embedded. Given this consideration, a natural scope for such an approach centers on quantifying the abundance dynamics of populations and the evolution of traits mediating their interactions as functions of relevant abiotic factors. Although taking into account abundance, phenotype and environment provides the basis for a partial understanding of the complex nature of biological communities, a deeper understanding must account for the effects of contemporary dispersal and the phylogeographic history of interacting lineages [@Kraft2007; @Hickerson2010; @Manceau2016; @markmcpeek2017] along with the genetic basis of ecologically relevant traits [@jeffreyconner2004; @FUSSMANN2007] and feedbacks between populations and the biogeochemical cycles they ultimately depend on [@michelloreau2010; @9781107648258]. It is therefore ideal that the development of any such mathematical approach anticipates extensions to account for these important factors shaping ecological communities, especially as empirical and conceptual work in these directions continues to grow [@AbdalaRoberts2014; @Klzsch2015; @Crutsinger2015; @Fitzpatrick2015; @Fitzpatrick2017; @Marx2017; @Rudman2017; @Skovmand2018; @VanNuland2019; @Harmon2019]. Furthermore, the approach would benefit from a stochastic component to capture the chance nature of biological reality [@Lande2003; @DeMeester2018; @Mubayi2019] and serve as a basis for the construction of statistical methods that measure evolutionary and ecological processes occuring in the wild. Such methods will tether theory to reality and allow for rigorous tests of hypotheses on the structure and behavior of ecological communities. In this paper we introduce a framework that establishes a formal connection between the continuous-time dynamics of abundance and quantitative traits in stochastically evolving populations, demonstrate it's utility through the derivation and analysis of a model of diffuse coevolution and discuss how it can be extended to account for the details mentioned above.

Current theoretical approaches to synthesize evolution and ecology have capitalized on the fact that biological fitness plays a key role in determining both sets of dynamics. While correlation of fitness and genotype is the basis of evolution by natural selection, the mean fitness across all individuals in a population determines the growth, stasis or decline of abundance. In section \ref{deterministic} we review the mathematical formalization of this connection, which has been established in the contexts of population genetics [@9781932846126; @joanroughgarden1979], evolutionary game theory [@Hofbauer1998; @9780674023383; @Lion2018], quantitative genetics [@Lande1982; @Doebeli1996; @Lion2018] and a unifying framework for these three distinct approaches to evolutionary theory [@Champagnat2006] which is intimately related to the approach we take here.

Although reviewing these accomplishments reveals a beautiful synthesis of evolution and population ecology, it also reveals a gap in theoretical approaches to incorporate the intrinsically random nature of populations. Specifically, in theoretical quantitative genetics the derivation of a population's response to random genetic drift is derived in discrete time under the assumption of constant effective population size using arguments based on properties of random samples [@Lande1976]. Although this approach conveniently mimics the formalism provided by the Wright-Fisher model of population genetics, real population sizes fluctuate over time. Furthermore, since these fluctuations are themselves stochastic, it seems natural to derive expressions for the evolutionary response to demographic stochasticity and consider how the results relate to characterizations of random genetic drift. This has been done in continuous time for population genetic models without too much technical overhead, assuming a finite number of alleles. However, for populations with a continuum of types, such as a quantitative trait, this becomes a vexing mathematical challenge. Here we close this gap by combining the calculus of white noise with results on rescaled limits of branching Brownian motion processes (BBM) and stochastic partial differential equations (SPDE). Our goal has two components: 1) Establish a novel synthetic approach to theoretical evolutionary ecology that provides a formal connection between demographic stochasticity and random genetic drift in the context of quantitative traits. To demonstrate the utility of our approach, we derive a model of coevolution in an ecological network and use it to investigate the relationship between competition coefficients and selection gradients. 2) Communicate some useful properties of space-time white noise, BBM and SPDE to as wide of audience as possible. With this goal in mind we will not provide a rigorous treatment of any of these deep subjects. Instead, we introduce a set of heuristics that only require the basic concepts of Reimann integration, partial differentiation and some exposure to Brownian motion and stochastic ordinary differential equations (SDE). For a concise introduction to SDE and Brownian motion, we recommend the primer by Evans [-@9781470410544]. Rigorous treatments of SPDE and rescaled limits of BBM can be found in Walsh [-@Walsh] and Dawson [-@dawson1993measure] respectively.

To provide motivation for the stochastic equations developed later and background for our model of coevolution, we begin with \S\ref{deterministic} by briefly summarizing derivations of deterministic dynamics of biological populations. Starting with a partial differential equation (PDE), we arrive at a general set of ordinary differential equations modelling the dynamics of abundance, trait mean and trait variance. From this we observe that replacing the PDE with a SPDE provides a path to derive SDE describing the evolutionary response to demographic stochasticity. We accomplish this in \S\ref{wnc} by introducing a set of mathematical tools based on the calculus of white noise and discuss how a diffusion limit of a spatially structured branching process leads to the natural SPDE appropriate for our study. The diffusion limit in turn provides a rigorous method for constructing fitness functions used in models of evolutionary ecology. We employ these tools to derive a system of SDE generalizing our deterministic results to account for demographic stochasticity. However, although biologically insightful, these equations remain difficult to analyze and implement numerically. In \S\ref{equations} we use an assumption of normally distributed trait values to simplify these expressions into formula that are much easier to work with. We then account for the constraint of adaptive evolution on the amount of heritable variation in a population by extending these results via a model of imperfect inheritance. The resulting equations coincide with classical results in quantitative genetics as a special case. We then demonstrate the utility of our approach by combining the derived equations of population dynamics with classical niche theory in \S\ref{coev} to formulate of a model of coevolution across a guild of $S$ species participating in exploitative competition along a common resource continuum. In SM \S\ref{diffuse} we apply a classical theorem on rescaled limits of BBM that allow for ecological interactions to provide a rigorous derivation. To gain biological insight, in \S\ref{dynamics} we numerically integrate our model of coevolution for $S=100$ species, tracking the dynamics of traits and abundances, under scenarios of weak and strong competition. We include an account of the natural history of the simulated system and discuss the significance of demographic stochasticity for structuring ecological communities. In \S\ref{ecoevo} we provide formula for selection gradients and competition coefficients implied by our model and use these to investigate the relationship between the strengths of competition and coevolution. Finally, \S\ref{conclusion} concludes with a summary of accomplishments, a few remarks on the limits of this approach and future directions to incorporate more explicitly the genetic architecture of populations, feedbacks with ecosystem processes and the macroevolutionary history of interacting lineages.

# The framework

At the core of our approach is a stochastic analog of the replicator equation with mutation in continuous time and phenotypic space [@Taylor1978; @Schuster1983]. From this stochastic replicator-mutator equation we derive a system of SDE for the dynamics of abundance, mean trait and additive genetic variance of a population. Hence, our approach develops a quantitative genetic theory of evolutionary ecology. A popular alternative to quantitative genetics is the theory of adaptive dynamics. As demonstrated by Page and Nowak [-@PAGE2002], the canonical equation of adaptive dynamics can be derived from the replicator-mutator equation. Thus, one could start from the atomic roots of our approach and pursue a stochastic adaptive dynamic theory instead. We choose the former in anticipation of an extension of our approach that explicitly models the genetic details of populations.

In this section we review the derivations of the replicator-mutator equation and trait dynamics from abundance dynamics and extend these formulae along with related results to the stochastic case. The results established in this section provide the framework from which larger scale ecological stuctures, such as species abundance distributions and interaction networks, can be computed.

## Deterministic dynamics \label{deterministic}

Our review begins by considering the dynamics of an asexually reproducing population in a homogeneous environment. Individuals are assumed to be haploid and carry one of $K$ alleles each with a different fitness expressed as growth rate. Under these assumptions, the derivation of the evolution of allele frequencies due to natural selection can be derived from expressions of exponential growth. This, and a few related approaches, have been provided by Crow and Kimura [-@9781932846126, \S 5.3]. Specifically, denoting $\nu_i$ the abundance of individuals with allele $i$ and $m_i$ the growth rate of allele $i$ [called the Malthusian parameter in @9781932846126], we have
\begin{equation}\label{first}
\frac{d\nu_i}{dt}=m_i\nu_i.
\end{equation}
Starting from this model, we get the total abundance of the population as $N=\sum_{i=1}^K\nu_i$, the frequency of allele $i$ as $p_i=\nu_i/N$ and the mean fitness of the population as $\bar m=\sum_{i=1}^Kp_im_i$. Hence, we can employ some elementary calculus to derive the dynamics of abundance $dN/dt$ and the dynamics of allele frequencies $dp_1/dt,\dots,dp_K/dt$ as
\begin{equation}
\frac{dN}{dt}=\sum_{i=1}^K\nu_im_i=N\sum_{i=1}^Kp_im_i=\bar mN,
\end{equation}
\begin{equation}\label{popgen}
\frac{dp_i}{dt}=\frac{d}{dt}\frac{\nu_i}{N}=\frac{1}{N^2}\left(N\frac{d\nu_i}{dt}-\frac{dN}{dt}\nu_i\right)
=\frac{1}{N}\left(m_i\nu_i-\bar m Np_i\right)=(m_i-\bar m)p_i.
\end{equation}
Two important observations of these equations include 1) mean fitness $\bar m$ determines the abundance dynamics of the entire population and 2) allele $i$ will increase (decrease) in frequency if $m_i>\bar m$ ($<\bar m$). Equation (\ref{popgen}) is known in the field of evolutionary game theory as the replicator equation [@Hofbauer1998; @9780674023383; @Lion2018]. Instead of being explicitly focused on alleles, the replicator equation describes the fluctuations of relative abundances of various _types_ in a population in terms of the vital rates of each type [@Taylor1978; @Schuster1983]. Using a matrix of transition rates between differing types, it is straight-forward to extend the replicator equation to include mutation, which is known as the replicator-mutator equation [@9780674023383].

Inspired by equations (\ref{first})-(\ref{popgen}), we derive an analog of the replicator-mutator equation for a continuum of types (that is, for a quantitative trait). In particular, we model a continuously reproducing population with trait values $x\in\mathbb{R}$ and an abundance density $\nu(x,t)$ that represents the amount of individuals in the population with trait value $x$ at time $t$. Hence, the abundance density satisfies $N(t)=\int_{-\infty}^{+\infty}\nu(x,t)dx$ and $p(x,t)=\nu(x,t)/N(t)$ is the relative density of trait $x$ which we also refer to as the phenotypic distribution.

To stay within the realm of biological plausibility, we require a set technical assumptions. First, we assume the initial abundance density is continuous, non-negative, integrable and has finite trait mean and variance. That is, we assume $\nu(x,0)$ is continuous in $x$, satisfies $\nu(x,0)\geq0$ for all $x\in\mathbb{R}$ and
\begin{equation}
N(0)=\int_{-\infty}^{+\infty}\nu(x,0)dx<+\infty,
\end{equation}
\begin{equation}
-\infty<\bar x(0)=\int_{-\infty}^{+\infty}xp(x,0)dx<+\infty,
\end{equation}
\begin{equation}
\sigma^2(0)=\int_{-\infty}^{+\infty}(x-\bar x(0))^2p(x,0)dx<+\infty,
\end{equation}
where $\bar x(t)$ and $\sigma^2(t)$ are respectively the mean trait and phenotypic variance at time $t\geq0$. We denote by $C_{1,c}^+(\mathbb{R}\times[0,\infty))$ the set of all non-negative integrable functions that are continuous in $x$ and $t$. Second, we assume selection is determined by the growth rate $m(\nu,x)$ that is differentiable with respect to $x\in\mathbb{R}$, continuous with respect to $\nu\in C^+_{1,c}(\mathbb{R}\times[0,\infty))$ and satisfies $m(\nu,x)\leq r$ for some $r\in\mathbb{R}$ and all $x\in\mathbb{R}$ and $\nu\in C_{1,c}^+(\mathbb{R}\times[0,\infty))$. Our continuity assumption implies; if a sequence of abundance densities $\{\nu_n\}_{n=1}^\infty\subset C^+_{1,c}(\mathbb{R}\times[0,\infty))$ satisfies $\lim_{n\to\infty}\nu_n=\nu\in C^+_{1,c}(\mathbb{R}\times[0,\infty))$, then $\lim_{n\to\infty}m(\nu_n,x)=m(\nu,x)$ for all $x\in\mathbb{R}$. Third, we assume mutation is captured by diffusion with coefficient $\frac{\mu}{2}$. With these technicalities aside, the demographic dynamics can be modelled by the PDE
\begin{equation}\label{eq1}
\frac{\partial}{\partial t}\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)
\end{equation}
with the initial condition $\nu(x,0)$ described above. This PDE is semilinear due to the dependency of the growth rate $m(\nu,x)$ on the solution $\nu(x,t)$ and is referred to as a scalar reaction-diffusion equation [@lawrenceevans2010]. When $\mu=0$, equation (\ref{eq1}) can be seen as an analog of equation (\ref{first}) for a continuum of types. By assuming mutation acts as diffusion the effect of mutation causes $\nu(x,t)$ to flatten out over time. In fact, if the growth rate is constant across $x$, then this model of mutation will cause $\nu(x,t)$ to converge to a flat line as $t\to\infty$. Although clearly an idealized representation of biological reality, this model is sufficiently general to capture a large class of dynamics including density dependent growth and frequency dependent selection. As an example, logistic growth combined with quadratic stabilizing selection can be captured using the growth rate $m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\nu(x,t)$ where $a\geq0$ is the strength of stabilizing selection around the phenotypic optimum $\theta\in\mathbb{R}$, $c\geq0$ captures the effect of intraspecific competition and $r\in\mathbb{R}$ is the intrinsic growth rate in the absence of abiotic selection.

To derive a replicator-mutator equation from equation (\ref{eq1}), we employ the chain rule from calculus. Writing $\bar m(t)=\int_{-\infty}^{+\infty}m(\nu,x)p(x,t)dx$ for the mean fitness, we have

\begin{multline}\label{N_det}
\frac{d}{dt}N(t)=\frac{d}{dt}\int_{-\infty}^{+\infty}\nu(x,t)dx=\int_{-\infty}^{+\infty}\frac{\partial}{\partial t}\nu(x,t)dx \\
=\int_{-\infty}^{+\infty} m(\nu,x)\nu(x,t)dx+\int_{-\infty}^{+\infty}\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)dx \\
=N(t)\int_{-\infty}^{+\infty} m(\nu,x)p(x,t)dx=\bar m(t) N(t).
\end{multline}
By our assumptions on mutation and rate of growth, $\nu(x,t)$ is twice differentiable with respect to $x$ and $\int_{-\infty}^{+\infty}\nu(x,t)dx<\infty$ for all $t\geq0$ (SM \S\ref{}). This implies that we are justified in swapping the order of differentiation and integration and the result $\int_{-\infty}^{+\infty}\frac{\partial^2}{\partial x^2}\nu(x,t)dx=0$ can be derived from the fundamental theorem of calculus. Biological reasoning agrees with this latter result since mutation neither creates nor destroys individuals, but merely changes their type from their parental type. Taking the same approach, we derive the dynamics of the phenotypic distribution $p(x,t)$ in response to selection and muation as

\begin{multline}\label{p_PDE}
\frac{\partial}{\partial t}p(x,t)=\frac{\partial}{\partial t}\frac{\nu(x,t)}{N(t)}=\frac{1}{N^2(t)}\left(N(t)\frac{\partial}{\partial t}\nu(x,t)-\nu(x,t)\frac{d}{dt}N(t)\right) \\
=\frac{1}{N(t)}\Big(m(\nu,x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)-\bar m(t)\nu(x,t)\Big) \\
=(m(\nu,x)-\bar m(t)) p(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t).
\end{multline}

This result closely resembles Kimura's continuum-of-alleles model [@Kimura1965; @Brger1986]. The primary difference being that our model utilizes diffusion instead of convolution with an arbitrary mutation kernel. Of course, our model of mutation can be derived as an approximation to Kimura's model, which has been referred to as the Gaussian allelic approximation [in reference to the distribution of mutational effects at a given locus of a genome on the values of traits, @9780471986539], the infinitesimal model [in reference to modelling continuous traits as being encoded by an infinite number of loci each having infinitesimal effect, @Barton2017] and the Gaussian descendants approximation [in reference to offspring trait values being normally distributed around their parental values, @Turelli2017]. Alternatively, since diffusion is the continuous-time equivalent to convolution against a Gaussian kernel (SM \S\ref{diffconvequiv}), equation (\ref{p_PDE}) can also be seen as a special case of the continuum-of-alleles model.

The covariance of fitness and phenotype across the population is defined as
\begin{equation}
\mathrm{Cov}_t\Big(m(\nu,x),x\Big)=\int_{-\infty}^{+\infty}\Big(m(\nu,x)-\bar m(t)\Big)\Big(x-\bar x(t)\Big)p(x,t)dx.
\end{equation}
Hence, the dynamics of the mean trait $\bar x(t)$ can be derived as

\begin{multline}\label{detmean}
\frac{d}{dt}\bar x(t)=\frac{d}{dt}\int_{-\infty}^{+\infty} x p(x,t)dx=\int_{-\infty}^{+\infty} x\frac{\partial}{\partial t} p(x,t)dx \\
=\int_{-\infty}^{+\infty} x\big(m(\nu,x)-\bar m(t)\big) p(x,t)+x\frac{\mu}{2}\frac{\partial^2}{\partial x^2} p(x,t)dx \\
=\mathrm{Cov}_t\Big(m(\nu,x),x\Big)+\frac{\mu}{2}\int_{-\infty}^{+\infty} x\frac{\partial^2}{\partial x^2} p(x,t)dx.
\end{multline}

Equation (\ref{detmean}) is a continuous time analog of the well known Robertson-Price equation [@Robertson1966; @PRICE1970; @FRANK2012; @Queller2017; @Lion2018]. The covariance of fitness and phenotype creates change in $\bar x$ to maximize mean fitness $\bar m$. Since this change is driven by a covariance with respect to phenotypic diversity, the response in mean trait to selection is mediated by the phenotypic variance. In particular, when $\sigma^2=0$, $\bar x$ will not respond to selection. The second term, due to mutation, represents the so-called _transmission bias_ [@FRANK2012]. When $p(x,t)$ is symmetric around $\bar x$, this term disappears due to equal bias in opposite directions cancelling the effect of mutation on $\bar x$. Following the approach taken to calculate the evolution of $\bar x$, we find the response of phenotypic variation to this model of selection and mutation is
\begin{equation}\label{vardyn_det}
\frac{d}{dt}\sigma^2(t)=\mathrm{Cov}_t\Big(m(\nu,x),(x-\bar x)^2\Big)+\frac{\mu}{2}\int_{-\infty}^{+\infty}(x-\bar x)^2\frac{\partial^2}{\partial x^2}p(x,t)dx.
\end{equation}
For the sake of space we relegate the derivation of $d\sigma^2/dt$ to SM \S\ref{var_deriv}. In the absence of mutation equation (\ref{vardyn_det}) agrees with the result derived by Lion [-@Lion2018] for discrete phenotypes. From a statistical perspective, if we think of $(x-\bar x)^2$ as a square error, then in analogy to the dynamics of the mean trait, we see that the response in $\sigma^2$ to selection can be expressed as a covariance of fitness and square error, which is defined in analogy to $\mathrm{Cov}_t(m(\nu,x),x)$. This covariance also creates change in $\sigma^2$ that maximizes mean fitness $\bar m$. In SM \S\ref{var_deriv} we show when $p(x,t)$ is a Gaussian bell curve, the increase in variance due to mutation simplifies to $\mu$.

In SM \S\ref{SDE_DERIV}, we extend these results to include the effects of demographic stochasticity. Similar to the approach taken by Champagnat, Ferriére and Méléard [-@Champagnat2006], we begin with a BBM that models populations as discrete sets of reproducing individuals whose vital rates depend on their trait value as well as the state of the entire population. Taking a large population size limit and keeping our assumption of single dimensional traits, we employ a pair of classic results that show, under the appropriate rescaling in time, phenotypic space and population density, a sequence of rescaled BBM converges to a limiting process that can be characterized by a SPDE [@Mlard1993; @zeng1998absolute]. The limiting processes of rescaled BBM have been referred to as measure-valued Markov processes [@dawson1993measure] or superprocesses [@alisonetheridge2000]. Under the simplifying assumptions inherited from our treatment of deterministic dynamics and an additional assumption on demographic stochasticity, we obtain as a special case a relatively simple expression for an SPDE that generalizes equation (\ref{eq1}). The simplicity of our special case allows us to use properties of space-time white noise processes to derive a set of SDE that generalize equations (\ref{N_det}), (\ref{detmean}) and (\ref{vardyn_det}) to include the effects of demographic stochasticity. Classic expressions for the effects of random genetic drift on the evolution of mean traits are obtained as a special case.

In the following section we provide the necessary mathematical tools needed to derive SDE from SPDE. Since our aim is to present this material to as wide of audience as possible, our approach deviates from standard definitions to remove the need for a detailed technical treatment. In addition to the notions of Reimann integration and partial differentiation already employed, the reader will only need some elementary probability and an intuitive understanding of SDE. Because space-time white noise, denoted by $\dot W(x,t)$, appears in the SPDE characterizing diffusion limits of BBM, we begin by defining $\dot W(x,t)$ and illustrating its relevant properties including a set of heuristics for performing calculations. Treating only the simplist of cases, we then provide a brief review of BBM, their diffusion limits and the SPDE that characterize them. For those not interested in the white noise calculus or superprocesses and would rather jump straight into more biologically relevant results, we recommend skipping to \S\ref{particular}.

## White noise calculus and superprocesses \label{wnc}

### Definition and basic properties of white noise \label{wnc_intro}

One can think of white noise as the static seen on old television sets or infinitely detailed random dust spread across both time and space. From a more mathematical, yet still informal perspective, white noise can be thought of as a stochastic process. That is, we can picture white noise as a collection of random variables indexed by time and possibly space. In relation to Brownian motion, denoted by $W$, white noise can be interpreted of as the derivative of Brownian motion with respect to time, denoted $\dot W$. Since Brownian motion takes infinitesimally small Gaussian distributed jumps at each time point, this leads to the conceptualization of white noise as a collection of Gaussian distributed random variables. Figure \ref{wn} illustrates realizations of this conceptualized white noise in one (left) and two (right) dimensions.

```{r, include=F, eval=F}
require("ggplot2")
require("ggthemes")
require("plot3D")
require("gridExtra")

num1 = 1e4
num2 = 1e4

xi_1 = rnorm(num1)
xi_2 = rnorm(num2)

x1 = seq(0,1,length.out = num1)
x2 = seq(0,1,length.out = sqrt(num2))

M = expand.grid(x=x2,y=x2)
M$value = xi_2

df_1 = data.frame(xi=xi_1,x=x1)

wn1d = ggplot(df_1)+geom_point(aes(x=x,y=xi),size=0.05)+theme_minimal()+xlab("Time")+ylab("Value")
wn2d = ggplot(M, aes(x, y, color = value))+scale_color_gradient(low="black", high="white")+geom_point(size=0.001, shape=22, show.legend = FALSE)+theme_minimal()+xlab("Time")+ylab("Space")

wn = grid.arrange(wn1d,wn2d,nrow=1)

ggsave("~/Research/A White Noise Approach To Evolutionary Ecology/wn.png",wn,width=5,height=2.5)
```

```{r, echo=F, fig.align='center', fig.cap='\\label{wn}Approximations of sample paths of temporal white noise (left) and space-time white noise (right) with brightness scaled to value.'}
knitr::include_graphics('wn.png')
```

However, it turns out that realizations of white noise do not exist as functions in the classical sense. Indeed, since Brownian motion is nowhere differentiable with respect to time, white noise cannot be formally understood as its time derivative. Thus our notation $\dot W$ is only meant to aid intuition and not be taken as formal. A formal understanding is possible by considering white noise as a _measure_-valued process [@Dawson1975; @Walsh] or as a _generalized_ process that acts on classically defined functions to return random variables or stochastic processes [@Krylov1981; @DaPrato2014]. The latter provides the general idea implemented here, though we do not closely follow the treatments of Krylov and Rozovskii [-@Krylov1981] or Da Prato and Zabczyk [-@DaPrato2014].

Throughout this section, we minimize notation by writing $\int_{\mathbb{R}}f(x)dx=\int_{-\infty}^{+\infty}f(x)dx$ and similarly $\int_Df(x)dx$ for the integral of $f$ on $D\subset\mathbb{R}$. We define $L^2_c(\mathbb{R}\times[0,\infty))$ as the set of functions $f(x,t)$ mapping $\mathbb{R}\times[0,\infty)$ to $\mathbb R$ that are continuous in $t$ and satisfy $\|f\|_2(t)=\sqrt{\int_\mathbb R f^2(x,t)dx}<\infty$ for each $t\in[0,\infty)$. We define a generalized stochastic process $\mathbf W$ that maps functions $f\in L^2_c(\mathbb{R}\times[0,\infty))$ to real-valued stochastic processes indexed by $t\in[0,\infty)$. To evaluate this process for a function $f\in L^2_c(\mathbb{R}\times[0,\infty))$ and some time $t\in[0,\infty)$ we write $\mathbf W_t(f)$. Specifically, for any $f,g\in L^2_c(\mathbb{R}\times[0,\infty))$, we define $\mathbf W(f)$ and $\mathbf W(g)$ to be Gaussian processes satisfying, for any $t,t_1,t_2\geq0$,
\begin{equation}\label{exp_WN}
\mathbb{E}\big(\mathbf W_t(f)\big)=\mathbb{E}\big(\mathbf W_t(g)\big)=0,
\end{equation}
\begin{equation}\label{cov_WN}
\mathbb{C}\big(\mathbf W_{t_1}(f),\mathbf W_{t_2}(g)\big)=\int_0^{t_1\wedge t_2}\int_\mathbb{R} f(x,s)g(x,s)dxds,
\end{equation}
where $t_1\wedge t_2=\min(t_1,t_2)$ and $\mathbb{E},\mathbb{C}$ denote expectation and covariance with respect to the underlying probability space. These operators are to be distinguished from $\bar f(t)$ and $\mathrm{Cov}_t(f,g)$ which denote expectation and covariance with respect to phenotypic diversity at time $t\geq0$.

Since Gaussian processes are characterized by their expectations and covariances, the processes $\mathbf W(f)$ and $\mathbf W(g)$ are well defined (under the convention $f=g$ whenever $\|f-g\|_2(t)=0$ for each $t\geq0$). In particular, if $f\in L^2_c(\mathbb{R}\times[0,\infty))$ is independent of time, then $\mathbf W(f)$ is a Brownian motion with variance at time $t\geq0$ equal to $t\|f\|_2^2$. With the generalized process $\mathbf W$ defined, we define the space-time white noise $\dot W(x,t)$ implicitly via the stochastic integral
\begin{equation}
``\int_0^t\int_\mathbb{R}f(x,s)\dot W(x,s)dxds"=``\int_\mathbb{R}\int_0^tf(x,s)\dot W(x,s)dxds"=\mathbf W_t(f), \ \forall \ f\in L^2_c(\mathbb{R}\times[0,\infty)), \ t\geq0.
\end{equation}
We place quotations in the above expression to emphasize its informal nature and that it should not be confused with classical Riemann integration. Following this definition of white noise, to compute its value we sample it using $L^2_c(\mathbb{R}\times[0,\infty))$ functions. For example, integrating white noise over a region $D\times[0,t]$, with $t>0$ and $D$ a bounded subset of $\mathbb{R}$, is equivalent to evaluating $\mathbf W_t(I_{D\times [0,\infty)})$ where
\begin{equation}
I_{D\times [0,\infty)}(x,t)=\left\{\begin{matrix}
0, & x\notin D \\
1, & x\in D
\end{matrix}\right..
\end{equation}
Since $\int_\mathbb{R}I_{D\times [0,\infty)}^2(x,t)dx=\int_Ddx=|D|$, where $|D|$ denotes the length of $D$, $I_{D\times [0,\infty)}\in L^2_c(\mathbb{R}\times[0,\infty))$. Thus, by equations (\ref{exp_WN}) and (\ref{cov_WN}), we have 
\begin{equation}
\mathbb{E}\left(\int_0^t\int_D\dot W(x,s)dxds\right)=0,
\end{equation}
\begin{equation}
\mathbb{V}\left(\int_0^t\int_D\dot W(x,s)dxds\right)=t|D|,
\end{equation}
where $\mathbb{V}$ denotes the variance operator with respect to the underlying probability space. Following this notation, equations (\ref{exp_WN}) and (\ref{cov_WN}) can be rewritten as
\begin{equation}\label{exp_WN_xi}
\mathbb{E}\left(\int_0^t\int_\mathbb{R}f(x,s)\dot W(x,s)dxds\right)=0,
\end{equation}
\begin{equation}\label{cov_WN_xi}
\mathbb{C}\left(\int_0^{t_1}\int_\mathbb{R}f(x,s)\dot W(x,s)dxds,\int_0^{t_2}\int_\mathbb{R}g(x,s)\dot W(x,s)dxds\right)
=\int_0^{t_1\wedge t_2}\int_\mathbb{R}f(x,s)g(x,s)dxds.
\end{equation}
To relate this material to the common notation used for SDE, we write $d\hat{\mathbf W}_t(f)=\frac{1}{\|f\|_2(t)}\left(\int_\mathbb{R}f(x,t)\dot W(x,t)dx\right)dt$ so that
\begin{equation}
\int_0^td\hat{\mathbf W}_s(f)=\int_0^t\int_\mathbb{R}\frac{f(x,s)}{\sqrt{\int_\mathbb{R}f^2(s,y)dy}}\dot W(x,s)dxds.
\end{equation}
This implies $\mathbb{E}\big(\int_0^td\hat{\mathbf W}_s(f)\big)=0$, $\mathbb{C}\big(\int_0^{t_1}d\hat{\mathbf W}_s(f),\int_0^{t_2}d\hat{\mathbf W}_s(f)\big)=t_1\wedge t_2$ and, as a function of $t$, $\int_0^td\hat{\mathbf W}_s(f)$ is a standard Brownian motion for any $f\in L^2_c(\mathbb{R}\times[0,\infty))$. Hence, $d\hat{\mathbf W}_t(f)$ is analogous to the traditional shorthand used to denote stochastic differentials. Thus, equation (\ref{cov_WN_xi}) effectively extends It&ocirc;'s multiplication table to:
\begin{table}[H]
\centering\caption{An extension of It\^o's multiplication table.}\vspace{0.2cm}
\begin{tabular}{l|lll}
             & $d\hat{\mathbf W}_t(f)$               & $d\hat{\mathbf W}_t(g)$                & $dt$ \\ \hline
             &                            &                            &      \\
$d\hat{\mathbf W}_t(f)$ & $dt$                       & $\frac{\left(\int_\mathbb{R}f \ g \ dx\right)dt}{\|f\|_2\|g\|_2}$ & $0$  \\
             &                            &                            &      \\
$d\hat{\mathbf W}_t(g)$ & $\frac{\left(\int_\mathbb{R}f \ g \ dx\right)dt}{\|f\|_2\|g\|_2}$ & $dt$                       & $0$  \\
             &                            &                            &      \\
$dt$         & $0$                        & $0$                        & $0$
\end{tabular}\label{mult}
\end{table}
The extension of It&ocirc;'s multiplication table and properties of white noise outlined in this subsection provide a useful set of tools for working with SPDE. In SM \S\ref{SDE_DERIV} we employ these tools to derive SDE that track the dynamics of abundance, mean trait and phenotypic variance of a population from a particular SPDE. In the following subsection, we review how this particular SPDE naturally arises as the diffusion limit of a BBM.

### From branching processes to SPDE \label{stochastic}

In real populations individuals are born and potentially reproduce before they ultimately die. These three events provide the basic ingredients of a branching process. Mathematical investigations of such processes have a relatively deep history [@Kendall1966]. The most simple branching process, known as the Galton-Watson process, describes the number of individuals alive at a given time $t\geq0$ as a non-negative integer [@Kimmel2015]. Feller [-@feller1951] introduced a formal method to approximate branching processes with diffusion processes which are continuous in state (i.e., population size is approximated as a continuous quantity). Since diffusion processes possess greater analytical tractability than branching processes, Feller's method, known as the diffusion limit, has acquired immense popularity particularly in the field of mathematical population genetics [@Ewens2004]. For over the past half of a century a great deal of accomplishments have been achieved in formalizing the diffusion limits of branching processes that describe populations of individuals occuring in some continuous space [@Watanabe1968; @Dawson1975; @Perkins1992; @Mlard1993; @edwinarendperkins1995; @zeng1998absolute; @Bertoin2003; @Etheridge2008; @Barton2019]. This space can represent geographic space or, relevant to our context, phenotypic space. In the following subsection, we describe the BBM process, which is a particularly important branching process with spatial stucture. This process has been very useful in the study of SPDE due to its simplifying assumption that individuals do not interact. However, this assumption imposes an unfortunate restriction by precluding the modelling of ecological interactions. We therefore follow our discussion of BBM with a review of a few important results on spatially structured branching processes that account for interactions.

__Branching Brownian motion__

A BBM tracks individuals navigating $d$-dimensional Euclidean space whose lifetimes are continuous random variables. Unlike other stochastic processes that take values in $\mathbb{R}^d$, BBM takes values in the set of _non-negative finite measures_ over $\mathbb{R}^d$. Intuitively, one can think of a finite measure as a function that maps subsets of $\mathbb{R}^d$ to real numbers (to be formal, we only consider the Borel subsets of $\mathbb{R}^d$ corresponding to the Euclidean metric, but understanding this technicality is not crucial to our discussion). In particular, denoting $X_t$ a BBM, for a subset $D\subset\mathbb{R}^d$, $X_t(D)$ returns the (random) number of individuals alive within the region $D$ at time $t\geq0$. The BBM has three main components:

  1) __Branching rate:__ Lifetimes of individuals are assumed to be exponentially distributed with death rate $\lambda>0$. Reproduction is assumed to occur simultaneously with death.

  2) __Reproductive law:__ At the death of an individual, a branching event occurs where a random (possibly zero) number of offspring are left. The distribution of offspring left is called the reproductive law or branching mechanism. We denote the mean and variance in reproductive output by $\mathscr{W}$ and $V$ respectively. The case of $\mathscr{W}=1$ is referred to as the critical condition.

  3) __Spatial movement:__ Each offspring is born at the current location of their parent. Immediately after birth they move around space according to Brownian motion with diffusion parameter $\sqrt\mu$. In our context we interpret spatial movement as mutation so that the location of an individual at death represents the value of its phenotype. Then an individual born at location $x\in\mathbb{R}^d$ that lives for $\tau>0$ units of time will have a normally distributed trait centered on $x$ with covariance matrix equal to $\tau\mu$ times the $d\times d$ identity matrix. Hence, offspring inherit normally distributed traits centered on their parental trait. This fact creates a vital link to the deterministic dynamics reviewed above. Indeed, in the absence of selection, the deterministic PDE (\ref{eq1}) reduces to the $d=1$-dimensional Kolmogorov forward equation for a Brownian motion with diffusion parameter $\sqrt\mu$.

To obtain a SPDE from a BBM we take a diffusion limit. There are several ways to do this, but a simple approach is to rescale the mass of individuals and time by $1/n$, diffusion by $\mu\to\mu/n$, branching rate by $\lambda\to n\lambda$, fitness by $\mathscr{W}\to\mathscr{W}^{1/n}$ and consider the limit as $n\to\infty$. Denoting the rescaled process by $X_t^{(n)}(D)$, the limiting process $\mathscr{X}_t=\lim_{n\to\infty}X_t^{(n)}$ is called a super-Brownian motion and is also a non-negative finite measure-valued process [@Watanabe1968]. Instead of returning the number of individuals alive in a region of space, super-Brownian motion returns the _mass_ of the population concentrated in a region of space. In particular, the value of $\mathscr{X}_t(D)$ is generally a continuously varying non-negative random variable for any $t\geq0$ and $D\subset\mathbb{R}^d$. It turns out that for spatial dimension $d=1$, $\mathscr{X}_t$ is absolutely continuous with respect to the Lebesgue measure for each $t\geq0$ [@Konno1988; @Reimers1989]. This means that we can write $\mathscr{X}_t(D)=\int_D \nu(x,t)dx$ for some density process $\nu(x,t)$. Setting $\lambda=1$ and $m=\ln\mathscr{W}$ this density process satisfies the SPDE
\begin{equation}\label{neutral_SPDE}
\frac{\partial}{\partial t}\nu(x,t)=m\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)+\sqrt{V\nu(x,t)}\dot W(x,t).
\end{equation}
Since $\nu(x,t)$ is not generally differentiable in $x$ or $t$, the derivatives in expression (\ref{neutral_SPDE}) are taken in the weak sense. That is, to rigorously interpret SPDE (\ref{neutral_SPDE}), we integrate the solution $\nu(x,t)$ against functions $f\in C_b^2(\mathbb{R})$, where $C_b^2(\mathbb{R})$ is the set of bounded and twice continuously differentiable functions on $\mathbb{R}$. Hence, equation (\ref{neutral_SPDE}) is just an abbreviation for

\begin{multline}\label{neutral_int}
\int_\mathbb{R}\nu(x,t)f(x)dx-\int_\mathbb{R}\nu(x,0)f(x)dx \\
=\int_0^t\int_\mathbb{R}\nu(x,s)\left(mf(x)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}f(x)\right)dsdx+\int_0^t\int_\mathbb{R}f(x)\sqrt{V\nu(x,s)}\dot W(x,s)dxds.
\end{multline}
For more on the general theory of SPDE see Walsh [-@Walsh]. Note that since $\nu(x,t)$ is the density of a finite measure, it is integrable for each $t\geq0$. Thus, since for some $M>0$, $|f(x)|\leq M$ for every $x\in\mathbb{R}$, setting $\varphi(x,t)=f(x)\sqrt{V\nu(x,t)}$ implies $\varphi\in L^2_c(\mathbb{R}\times[0,\infty))$. Hence, the white noise integral on the right-hand side of equation (\ref{neutral_int}) can be understood using the heuristics introduced above. Evaluating equation (\ref{neutral_int}) in the particular case of $f(x)\equiv1$ returns the total mass process, which we refer to as the total abundance $N(t)$.

A convergence theorem for the diffusion limit of a generalization of BBM was established by Watanabe [-@Watanabe1968]. Dawson [-@Dawson1975] suggested that, for spatial dimension $d=1$, this diffusion limit should admit a density process that satisfies a SPDE. Konna and Shiga [-@Konno1988] and Reimers [-@Reimers1989] independently proved Dawson's suggestion was indeed correct. The diffusion limit of this more general branching process (in arbitrary spatial dimension) is referred to as the Dawson-Watanabe superprocess [@alisonetheridge2000]. Conditioning the Dawson-Watanabe superprocess to have constant mass returns the Fleming-Viot process [@Etheridge1991; @Perkins1991] which has been popular in studies of spatial population genetics. In particular, an extension of the Fleming-Viot process, known as the $\Lambda$-Fleming-Viot process, was introduced by Bertoin and Le Gall [-@Bertoin2003] and coined by Etheridge [-@Etheridge2008] where it was used to resolve some technical challenges in modelling isolation by distance [@Felsenstein1975; see also @Barton2013; and @Barton2019]. Although this provides an impressive list of accomplishments, the Dawson-Watanabe superprocess falls short of our needs. In particular this process assumes individuals do not interact and thus precludes its ability to model ecological interactions. However, this concern has been addressed, leading to constuctions of superprocesses that account for interactions among individuals. In the next subsection we summarize the main results in this area and introduce the SPDE that provides the basis for our approach to theoretical evolutionary ecology.

__Interacting superprocesses__

The existence of diffusion limits for a class of measure-valued branching processes involving interactions among individuals has been treated by Méléard and Roelly [-@meleard1992interacting; -@Mlard1993]. The interactions can manifest as dependencies of the spatial movement or reproductive law of individuals on their position $x$ and the state of the whole population described by $X_t$. An important result of Méléard and Roelly [-@meleard1992interacting; -@Mlard1993] is a theorem that provides sufficient conditions to construct a sequence of rescaled measure-valued branching processes that converge to a generalization of the Dawson-Watanabe superprocess that includes interactions. The rescaling is analogous to that described above for non-interacting Dawson-Watanabe superprocesses, but now the reproductive law described by $\mathscr{W}(X_t,x)$ and $V(X_t,x)$, branching rate $\lambda(X_t,x)$ and diffusion parameter $\sqrt{\mu(X_t,x)}$ are allowed to depend on the whole population $X_t$ and individual location $x$. In Figure \ref{rescaled} we demonstrate this rescaling in discrete time for a population experiencing stabilizing selection and logistic growth. Since time is discretized, the process we simulate is formally a branching random walk. For further details on our simulation see SM \S\ref{numerical}.

```{r, eval=F, echo=F}
require(ggplot2)
require(ggthemes)
require(gridExtra)

n_1 = read.csv("/home/bob/Research/Branching Brownian Motion/n_1.csv")
n_2 = read.csv("/home/bob/Research/Branching Brownian Motion/n_2.csv")
n_3 = read.csv("/home/bob/Research/Branching Brownian Motion/n_3.csv")

n1p = ggplot(n_1,aes(x=time,y=x))+geom_point(size=1.0,stroke=0,alpha=0.1)+theme_minimal()+ylab("Trait Value \n (n=1)")+xlab("")+xlim(0,50)+ylim(-5,5)+ggtitle("Rescalings of a branching random walk")+theme(plot.title = element_text(hjust = 0.5))
n2p = ggplot(n_2,aes(x=time,y=x))+geom_point(size=0.5,stroke=0,alpha=0.1)+theme_minimal()+ylab("Trait Value \n (n=5)")+xlab("")+xlim(0,50)+ylim(-5,5)
n3p = ggplot(n_3,aes(x=time,y=x))+geom_point(size=0.25,stroke=0,alpha=0.1)+theme_minimal()+ylab("Trait Value \n (n=20)")+xlab("Time")+xlim(0,50)+ylim(-5,5)

rescaled_plots = grid.arrange(n1p,n2p,n3p,nrow=3)

ggsave("~/Research/A White Noise Approach To Evolutionary Ecology/rescaled_plots.png",rescaled_plots,device="png",width=6,height=6,units="in",dpi=350)
```

```{r, message=FALSE, echo=F, fig.align='center', fig.cap='\\label{rescaled}Rescaled sample paths of a branching random walk under stabilizing selection and logistic growth. The top plot displays a sample path without scaling $(n=1)$, the middle plot shows a sample path rescaled by $n=5$ and the bottom plot shows a sample path rescaled by $n=20$.'}
knitr::include_graphics('rescaled_plots.png')
```

Interactions that manifest in the spatial movement can be used to model mutation bias and those manisfesting in the reproductive law can model density-dependent growth rates and frequency-dependent selection. Perkins [-@Perkins1992; -@edwinarendperkins1995] developed a theory of stochastic integration with respect to the so-called _Brownian trees_ to characterize interacting superprocesses and establish properties of existence and uniqueness. Li [-@zeng1998absolute] built directly off of the construction of Méléard and Roelly [-@meleard1992interacting; -@Mlard1993] to study properties of density processes associated with interacting superprocesses, arriving at a SPDE that forms the foundation of our approach.

Recall, we use $\nu(x,t)$ to denote the density of a superprocess, given it exists. Assuming the interactions manifest only in the reproductive law and that spatial movement follows Brownian motion with diffusion parameter $\sqrt\mu$ independent of both $X_t$ and $x$, Li [-@zeng1998absolute] proved a result that implies the interacting superprocess on one dimensional trait space has a density $\nu(x,t)$ which is non-negative, integrable, continuous in time and space and satisfies the SPDE
\begin{equation}\label{SPDE}
\frac{\partial}{\partial t}\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)+\sqrt{V\nu(x,t)}\dot W(x,t).
\end{equation}
Note that, comparing equation (\ref{SPDE}) to equation (3.5) of Li [-@zeng1998absolute], our $m$ and $V$ correspond to Li's $b$ and $c$ respectively. Denoting $C_1^+(\mathbb{R})$ the set of non-negative, continuous, integrable functions on $\mathbb{R}$, it is important to note that, under the assumptions made in Méléard and Roelly [-@meleard1992interacting; -@Mlard1993] and Li [-@zeng1998absolute], equation (\ref{SPDE}) is only formal when $m(\nu,x)$ is bounded across all combinations of $\nu\in C_1^+(\mathbb{R})$ and $x\in\mathbb{R}$. However, recalling our condition $m(\nu,x)\leq r\in\mathbb{R}$, the growth rates we consider are only bounded above. Replacing $m(\nu,x)$ with its upper-bound demonstrates that the associated superprocess should still admit an integrable density. That is, we should still be able to compute the total mass process $N(t)$. In fact, in the proof of the construction of the interacting superprocess as the limit of rescaled branching diffusions, Méléard and Roelly [-@meleard1992interacting; -@Mlard1993] assumed $m(\nu,x)$ to be bounded to guarantee the total mass process will have finite mean and variance. This allowed them to employ a tightness criterion for sequences of measures and show the rescaled processes converge to a superprocess with finite total mass. Li's [-@zeng1998absolute] result builds directly off of Méléard and Roelly's construction, inheriting the assumption of boundedness for $m(\nu,x)$. However, in Li [-@zeng1998absolute], the sufficiency of $m(\nu,x)$ being bounded above is even more clear since Li works explicitly with a common upperbound for both $m(\nu,x)$ and $V$. Hence, one can repeat the necessary proofs replacing the assumption that $m(\nu,x)$ is bounded with the assumption that $m(\nu,x)$ is merely bounded above to derive the same results.

What remains is to show the spatial mean and variance are finite, assuming finite initial conditions. That is, setting $M_n(t)=\int_\mathbb{R}|x|^n\nu(x,t)dx$, we want $M_1(0),M_2(0),M_4(0)<\infty$ to imply $M_1(t),M_2(t),M_4(t)<\infty$ for each $t>0$ (it turns out we need $M_4(t)<\infty$ for the SDE describing the evolution of phenotypic variance). By condition (\ref{m_cond}), we see that $m(\nu,x)$ drops quadratically as $x$ leaves $\theta$. In SM \S\ref{finite} we show that, when $V=0$, this form of fitness generates a restoring force that ensures a finite mean and variance. Since mutation causes finite perturbations we do not expect it to overcome this restoring force in the stochastic case. If anything, we suspect a large mutation variance to merely increase the genetic load of the population and drive it to extinction. Indeed, this suggestion is supported by equilibrium of the corresponding deterministic system found in SM \S\ref{numerical}. However, to establish further support, in SM \S\ref{numerical} we investigate these claims by comparing simulations of the BBM with selection to analytical solutions of the PDE obtained from equation (\ref{SPDE}) by setting $V=0$. The results suggest our assertions are sound, but they have yet to be rigorously justified. For now, we leave this as an open problem and assume spatial means and variances are finite for every $t\in[0,\infty)$. In the next subsection we derive an SDE for the abundance of a population and report the SDE for the trait mean $\bar x$ and trait variance $\sigma^2$. For the sake of space, derivations of $d\bar x$ and $d\sigma^2$ can be found in SM \S\ref{SDE_DERIV}.

### Deriving SDE from a SPDE \label{SDEriv}

The dynamics of abundance (referred to in the superprocess literature as the total mass process) are obtained by defining the process $N(t)=\int_{\mathbb{R}}\nu(x,t)dx$ and evaluating expression

\begin{multline}\label{neat_trick}
\int_\mathbb{R}f(x)\nu(x,t)dx-\int_\mathbb{R}f(x)\nu(x,0)dx \\
=\int_0^t\int_\mathbb{R}\nu(x,s)\left(m(\nu,x)f(x)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2} f(x)\right)dxds+\int_0^t\int_\mathbb{R}f(x)\sqrt{V\nu(x,s)}\dot W(x,s)dxds
\end{multline}
with $f(x)\equiv1$ to obtain
\begin{equation}
N(t)-N(0)=\bar m(t)N(t)+\int_0^t\int_\mathbb{R}\sqrt{V\nu(x,s)}\dot W(x,s)dxds.
\end{equation}
Equation (\ref{cov_WN_xi}) implies
\begin{equation}
\mathbb{V}\left(\int_0^t\int_\mathbb{R}\sqrt{V\nu(x,s)}\dot W(x,s)dxds\right)=\int_0^t\int_\mathbb{R}V\nu(x,s)dxds=\int_0^tV N(s)ds.
\end{equation}
Following the notation introduced in \S\ref{wnc_intro}, we have
\begin{equation}
\int_0^td\hat{\mathbf W}_s(\sqrt{V\nu(x,s)})=\int_0^t\int_\mathbb{R}\frac{\sqrt{V\nu(x,s)}}{\sqrt{V N(s)}}\dot W(x,s)dxds
\end{equation}
is, as a function of $t$, a standard Brownian motion. To clean up notation set $dW_1(t)=d\hat{\mathbf W}_t(\sqrt{V\nu(x,t)})$. Then we can write an SDE for $N$ which we provide in the following section.

Unfortunately, the same trick does not work for calculating SDE for the mean and variance of traits. Indeed, setting $f(x)=x,x^2$ and evaluating expression (\ref{neat_trick}) will lead to SDE that miscalculate the effect of mutation and hence are clearly incorrect results. Obtaining the correct results turns out to require some very involved calculations which we perform in SM \S\ref{SDE_DERIV}.

## Equations of evolutionary and demographic dynamics \label{equations}

In SM \S\ref{SDE_DERIV} we use our assumptions that trait mean and variance are finite and that variance in reproductive output $V\geq0$ does not depend on $x$ or $\nu(x,t)$ to show SDE for $N(t)$, $\bar x(t)$ and $\sigma^2(t)$ can be expressed as
\begin{subequations}\label{gen_eqns}
\begin{equation}\label{N}
dN(t)=\bar m(t)N(t)dt+\sqrt{V N(t)}dW_1(t),
\end{equation}
\begin{equation}\label{xbar_gen}
d\bar x(t)=\mathrm{Cov}_t\Big(x,m(\nu,x)\Big)dt+\frac{\mu}{2}\int_{-\infty}^{+\infty}x\frac{\partial^2}{\partial x^2} p(x,t)dx+\sqrt{V\frac{\sigma^2(t)}{N(t)}}dW_2(t),
\end{equation}
\begin{multline}\label{sig2_gen}
d\sigma^2(t)=\left(\mathrm{Cov}_t\Big((x-\bar x(t))^2,m(\nu,x)\Big)+\frac{\mu}{2}\int_{-\infty}^{+\infty}(x-\bar x(t))^2\frac{\partial^2}{\partial x^2} p(x,t)dx-V\frac{\sigma^2(t)}{N(t)}\right)dt \\ +\sqrt{V\frac{\overline{(x-\bar x(t))^4}-\sigma^4(t)}{N(t)}}dW_3(t).
\end{multline}
\end{subequations}
where $W_1$, $W_2$ and $W_3$ are standard Brownian motions. Dividing by $dt$ one can interpret equations (\ref{gen_eqns}) as if they are ordinary differential equations, but this not technically rigorous since Brownian motion is nowhere differentiable with respect to time. In SM \S\ref{SDE_DERIV} we show that in general $W_1$ is independent of both $W_2$ and $W_3$, but $W_2$ and $W_3$ covary.

There is quite a bit we can learn from expressions (\ref{gen_eqns}). Firstly, setting $V=0$ recovers the deterministic dynamics derived in \S\ref{deterministic}. Alternatively, one can take $N(t)\to\infty$ to recover the deterministic dynamics for $\bar x(t)$ and $\sigma^2(t)$. Characteristically, we note the effect of demographic stochasticity on abundance grows with $\sqrt{N(t)}$. Hence, dividing by $N$, we find the effects of demographic stochasticity on the per-capita growth rate diminish with increased abundance. Relating the response to demographic stochasticity derived here to the effect of random genetic drift derived in classic quantitative genetic theory, if we set $\sigma^2(t)=\sigma^2$ and $N(t)=N$ constant with respect to time, then integrating the stochastic term in equation (\ref{xbar_gen}) over a single unit of time returns a normally distributed random variable with mean zero and variance equal to $V\sigma^2/N$. In particular, assuming perfect inheritance, when reproductive variance is unity ($V=1$) this random variable coincides with the effect of random genetic drift on the change in mean trait over a single generation derived using sampling arguments [@Lande1976]. There is also an interesting connection with classical population genetics. A fundamental result from early population genetic theory is the expected reduction in diversity due to the chance loss of alleles in finite populations [@Fisher1923; @Wright97]. This expected reduction in diversity due to random genetic drift is captured by the third term in the deterministic component of expression (\ref{sig2_gen}), particularly $-V\sigma^2(t)/N(t)$. The component of SDE (\ref{sig2_gen}) describing random fluctuations in $\sigma^2(t)$ is more complicated and is proportional to the root of the difference between the centralized fourth moment of $p(x,t)$ and $\sigma^4(t)$. These expressions can be used to investigate the dynamics of the mean and variance for general $\nu(x,t)$. However, in the next subsection we simplify these expressions by approximating $\nu(x,t)$ with a Gaussian curve. In SM \S\ref{SDE_DERIV} we show that under the Gaussian case $W_1,W_2$ and $W_3$ are independent.

### Particular results assuming a Gaussian phenotypic distribution \label{particular}

By assuming $\nu(x,t)$ can be approximated by a Gaussian curve for each $t\geq0$, expressions (\ref{N}), (\ref{xbar_gen}) and (\ref{xbar_gen}) transform into efficient tools for deriving the dynamics of populations given a fitness function $m(\nu,x)$. Gaussian phenotypic distributions are often obtained through Gaussian, exponential or weak selection approximations together with a simplified model of inheritance and random mating [@pmid17248993; @Turelli1984; @Turelli1986; @Turelli2017; @9780471986539]. Alternatively, it has been shown that a Gaussian distribution can provide a reasonable approximation even when selection is strong and non-Gaussian [@pmid7851785]. However, our approach adds an additional layer of difficulty. Even with Gaussian selection, the resulting solution to SPDE (\ref{SPDE}) will only be a Gaussian curve in expectation, assuming a Gaussian initial condition. Yet this difficulty is not as challenging as it may first appear. Indeed, since SPDE (\ref{SPDE}) can be derived as a diffusion limit we know that, under the appropriate assumptions on selection, genetic architecture and reproduction, the stochastic departure from a Gaussian curve is negligible when the ratio $V/N$ is small (i.e., when the variance in reproductive output is much smaller than the population size). In SM \S\ref{numerical} we demonstrate this result using numerical methods. Mathematically, this requirement restricts model parameters to regions that maintain large population sizes. Biologically, this implies populations are not at risk of extinction. Hence, models developed in this framework are not suitable for studying colonization-extinction dynamics or evolutionary rescue. Allowing for these restrictions, we may safely assume that $\nu$ is approximately Gaussian and justify writing
\begin{equation}
\nu(x,t)=\frac{N(t)}{\sqrt{2\pi\sigma^2(t)}}\exp\left(-\frac{\big(x-\bar x(t)\big)^2}{2\sigma^2(t)}\right).
\end{equation}
Under this assumption, Lande [-@Lande1976] found (suppressing the dependency on $t$)
\begin{equation}
\mathrm{Cov}(x,m)=\sigma^2\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right).
\end{equation}
Under the same assumption, in SM \S\ref{var_deriv} we show
\begin{equation}
\mathrm{Cov}\Big((x-\bar x)^2,m\Big)=2\sigma^4\left(\frac{\partial\bar m}{\partial\sigma^2}-\overline{\frac{\partial m}{\partial\sigma^2}}\right)
\end{equation}
and $\overline{(x-\bar x)^4}=3\sigma^4$. In particular, this implies
\begin{subequations}\label{no_inher}
\begin{equation}\label{xbar}
d\bar x=\sigma^2\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right)dt+\sqrt{V\frac{\sigma^2}{N}}dW_2,
\end{equation}
\begin{equation}\label{G}
d\sigma^2=2\sigma^4\left(\frac{\partial\bar m}{\partial\sigma^2}-\overline{\frac{\partial m}{\partial\sigma^2}}\right)dt +\left(\mu-V\frac{\sigma^2}{N}\right)dt+\sigma^2\sqrt{\frac{2V}{N}}dW_3.
\end{equation}
\end{subequations}
These equations allow us to derive the response in trait mean and variance by taking derivatives of fitness, a much more straightforward operation than calculating a covariance for general phenotypic distributions. Note that in the above expressions, the partial derivatives of $\bar m$ represent frequency independent selection and the averaged partial derivatives of $m$ represent frequency dependent selection. This relationship has already been pointed out by Lande [-@Lande1976] for the evolution of the mean trait, but here we see an analogous relationship holds also for the evolution of trait variance.

In the next subsection we generalize this result to the case when traits are imperfectly inherited. In this case, the phenotypic variance $\sigma^2$ is replaced by a genetic variance $G$. This genetic variance represents the component of the variance in expressed traits $\sigma^2$ explained by additive effects of different alleles among genetic loci encoding for the focal phenotype [@joanroughgarden1979; @bulmer1980; @michaellynch1998]. It is therefore fitting that $G$ is referred to as the additive genetic variance.

### The evolution of additive genetic variance \label{inheritance}

To model imperfect heritability we consider the relationship between expressed phenotypes $x\in\mathbb{R}$ and associated genetic values $g\in\mathbb{R}$ known as _breeding values_. The breeding value of an individual is the sum of additive effects of the alleles carried by the individual on its expressed trait. Since our derivations of evolutionary equations are based on branching processes that assume asexually reproducing populations (\S\ref{stochastic}), the additive genetic variance $G$ is just the variance of breeding values in a population. For a detailed treatment of breeding values and additive genetic variances, see Bulmer [-@bulmer1980] and Lynch and Walsh [-@michaellynch1998]. Following standard assumptions of classical quantitative genetics we assume that the expressed trait for any given individual is normally distributed around their breeding value with variance $\eta$. Hence, $\sigma^2=G+\eta$. This coincides with assuming breeding values can be predicted from expressed traits using ordinary least squares. In the case that all of the effects of alleles on an expressed trait are additive, $\eta$ is known as the _variance of environmental deviation_ [@michaellynch1998]. For a given breeding value, we denote the probability density of a randomly drawn expressed trait by $\psi(x,g)$. Hence,
\begin{equation}
\psi(x,g)=\frac{1}{\sqrt{2\pi\eta}}\exp\left(-\frac{(x-g)^2}{2\eta}\right).
\end{equation}
To include this relationship in our framework, we write $\rho(g,t)$ as the abundance density of breeding values at time $t$ so that $\int_{-\infty}^{+\infty}\rho(g,t)dg=\int_{-\infty}^{+\infty}\nu(x,t)dx=N(t)$. We switch our focus from directly modelling the evolution of $\nu(x,t)$ to modelling the evolution of $\rho(g,t)$. Once $\rho(g,t)$ is determined, we can compute $\nu(x,t)$ via
\begin{equation}\label{nu_from_rho}
\nu(x,t)=\int_{-\infty}^{+\infty}\rho(g,t)\psi(x,g)dg.
\end{equation}
However, since selection acts on expressed phenotypes, we use our assumed relationship between breeding values and expressed traits to calculate the fitness of breeding values. To motivate our approach, consider the problem of inferring the breeding value of an individual given its expressed trait $x$. Denote $\mathfrak{g}$ a random variable representing the unkown breeding value. Under our model of inheritance we know $x$ is a random sample from a normal distribution with mean $\mathfrak{g}$ and variance $\eta$. Maximizing likelihood suggests $x$ is our best guess for $\mathfrak{g}$, but the actual value of $\mathfrak{g}$ is normally distributed around $x$ with the variance $\eta$. Hence, for fixed $x$, we obtain $\psi(x,g)$ as the probability density of $\mathfrak{g}$. Thus, the mean fitness of a breeding value $g$ across all individuals carrying $g$ can be written as
\begin{equation}\label{m_rel}
m^*(\rho,g)=\int_{-\infty}^{+\infty}m(\nu,x)\psi(x,g)dx.
\end{equation}
This is similar to the approach taken by Kimura and Crow [-@Kimura1978] to calculate the overall effects of selection for expressed characters onto the changes in the distribution of alleles encoding those characters. However, instead of focusing on the frequencies of alleles at particular loci, we focus on the densities of breeding values. With the relationship between $m(\nu,x)$ and $m^*(\rho,g)$ established, we define the evolution of $\rho(g,t)$ by the SPDE
\begin{equation}
\dot\rho(g,t)=m^*(\rho,g)\rho(g,t)+\frac{\mu}{2}\frac{\partial^2}{\partial^2 g}\rho(g,t)+\sqrt{V\rho(g,t)}\dot W(g,t).
\end{equation}
We assume $\rho(g,t)$ is Gaussian which implies its mode coincides with $\bar x$. Furthermore, since $\sigma^2=G+\eta$, we can use equation (\ref{m_rel}) and the chain rule from calculus to find
\begin{subequations}
\begin{equation}
\frac{\partial\bar m}{\partial G}=\frac{\partial\bar m}{\partial\sigma^2}\frac{\partial\sigma^2}{\partial G}=\frac{\partial\bar m}{\partial\sigma^2},
\end{equation}
\begin{equation}
\overline{\frac{\partial m}{\partial G}}=\overline{\frac{\partial m}{\partial\sigma^2}\frac{\partial\sigma^2}{\partial G}}=\overline{\frac{\partial m}{\partial\sigma^2}}.
\end{equation}
\end{subequations}
Thus, equations (\ref{no_inher}) become
\begin{subequations}\label{inher}
\begin{equation}\label{xbarfinal}
d\bar x=G\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right)dt+\sqrt{V\frac{G}{N}}dW_2,
\end{equation}
\begin{equation}\label{Gfinal}
dG=2G^2\left(\frac{\partial\bar m}{\partial G}-\overline{\frac{\partial m}{\partial G}}\right)dt+\left(\mu-V\frac{G}{N}\right)dt+ G\sqrt{\frac{2V}{N}}dW_3.
\end{equation}
\end{subequations}

From expressions (\ref{inher}) we see that, under our model of inheritance, focusing on additive genetic variance $G$ instead the variance in expressed traits $\sigma^2$ makes no structural changes to the basic equations describing the dynamics of populations.

# A model of diffuse coevolution \label{coev}

## Formulation \label{form_coev}

In this section we demonstrate the use of our framework by developing a model of diffuse coevolution across a guild of $S$ species whose interactions are mediated by resource competition along a single niche axis. Because our approach treats abundance dynamics and evolutionary dynamics simultaneously, this model allows us to investigate the relationship between selection gradients and competition coefficients.

The dynamics of phenotypic distributions and abundances have been derived above and so the only task remaining is the formulation of a fitness function. Our approach mirrors closely the theory developed by MacArthur and Levins [-@Macarthur1967], Levins [-@9780691080628] and MacArthur [-@Arthur1969; -@MacArthur1970; -@9780691023823]. The most significant difference, aside from allowing evolution to occur, is the treatment of resource quality, which we replace with a model of abiotic stabilizing selection. A derivation is provided in SM \S\ref{diffuse}.

For species $i$ we inherit the above notation for trait value, distribution, average, variance, abundance, etc except with an $i$ in the subscript. Real world examples of niche axes include the body size of prey for lizard predators and the date of activity in a season for pollinators competing for floral resources. For mathematical convenience, we model the axis of resources by the real line $\mathbb{R}$. The value of a resouce along this axis is denoted by the symbol $\zeta$. For an individual in species $i$, we assume the resource utilization curve $u_i$ can be written as

\begin{equation}
u_i(\zeta,x_i)=\frac{U_i}{\sqrt{2\pi w_i}}\exp\left(-\frac{(x_i-\zeta)^2}{2w_i}\right).
\end{equation}

We further assume the niche center $x_i$ is normally distributed among individuals in species $i$, but the niche breadth $w_i$ and total niche utilization $U_i$ are constant across individuals in species $i$ and therefore cannot evolve. Suppose $\theta_i\in\mathbb R$ is the optimal location along the niche axis for species $i$ such that, in the absence of competition, individuals leave on average $Q_i$ offspring when concentrated at $\theta_i$. We capture the rate by which the fitness falls as niche location $\zeta$ leaves the optimum $\theta_i$ by the parameter $A_i\geq0$. Hence, abiotic stabilizing selection along the resource axis can be modelled by the curve

\begin{equation}
e_i(\zeta)=Q_i\exp\left(-\frac{A_i}{2}(\theta_i-\zeta)^2\right).
\end{equation}

The effect of abiotic stabilizing selection on the fitness for an individual of species $i$ with niche location $x_i$ is then given by

\begin{equation}
\int_{-\infty}^{+\infty}e_i(\zeta)u_i(\zeta,x_i)d\zeta=\frac{Q_iU_i}{\sqrt{A_iw_i+1}}\exp\left(-\frac{A_i}{2(A_iw_i+1)}(\theta_i-x_i)^2\right).
\end{equation}

To determine the potential for competition between individuals with niche locations $x_i$ and $x_j$, belonging to species $i$ and $j$ respectively, we compute the niche overlap

\begin{equation}
\mathcal O_{ij}(x_i,x_j)=\int_{-\infty}^{+\infty}u_i(\zeta,x_i)u_j(\zeta,x_j)d\zeta=\frac{U_iU_j}{\sqrt{2\pi(w_i+w_j)}}\exp\left(-\frac{(x_i-x_j)^2}{2(w_i+w_j)}\right).
\end{equation}

A notable criticism of using niche overlap to measure the intensity of competition points to cases where populations competiting on multiple niche axes exhibit overlap on at least one of the axes, but no overall niche overlap [@Holt1987]. Thus niche overlap on lower-dimensional projections of some multivariate niche space does not imply the populations compete. To illustrate with a simple example, consider two populations competing for space on the plane $\mathbb{R}^2$. If the spatial distributions of the two populations overlap, then they will overlap on both spatial axes. However, if the populations do not overlap on at least one of the spatial axes, they will have no overall spatial overlap. Furthermore, even if the species overlap on both spatial axes, they need not have any overall spatial overlap. This final result corresponds to the fact that components of niche space do not necessarily interact multiplicatively to determine the consequences for the intensity of competition. In another component of Holt's [-@Holt1987] critique, an argument is made for the potential of competition occuring without any overlap in niche space. However, this argument is based on the practical difficulty of identifying every resource axis populations are competing on and how these axes interact to determine fitness consequences. Our model avoids these caveats by assuming competition only occurs along a single dimensional resource gradient. To map the degree of niche overlap to fitness, we assume competition between individuals with niche locations $x_i$ and $x_j$ additively decreases the Malthusian fitness for the individual in species $i$ by $c_i\mathcal{O}_{ij}(x_i,x_j)$ for some $c_i\geq0$. In SM \S\ref{diffuse} we combine this niche model with equations (\ref{N}), (\ref{xbarfinal}) and (\ref{Gfinal}) to find

\begin{subequations}\label{comm_dynamics}
\begin{multline}\label{N_comm}
dN_i = \Bigg\{R_i-\frac{a_i}{2}\Big((\bar x_i-\theta_i)^2+G_i+\eta_i\Big) - c_i\sum_{j=1}^SN_jU_iU_j\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\Bigg\}N_idt + \sqrt{V_iN_i}dW_1,
\end{multline}

\begin{multline}\label{x_comm}
d\bar x_i = \Bigg\{a_iG_i(\theta_i-\bar x_i)-c_iG_i\bigg(\sum_{j=1}^SN_jU_iU_jb_{ij}(\bar x_j-\bar x_i)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\bigg)\Bigg\}dt+\sqrt{V_i\frac{G_i}{N_i}}dW_2,
\end{multline}

\begin{multline}\label{G_comm}
dG_i =  \Bigg\{c_i{G_i}^2\bigg(N_iU_i^2b_{ii}\sqrt\frac{b_{ii}}{2\pi}+\sum_{j=1}^SN_jU_iU_jb_{ij}\left(1-b_{ij}(\bar x_i-\bar x_j)^2\right)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\bigg) \\
+\mu_i-a_i{G_i}^2-V_i\frac{G_i}{N_i}\Bigg\}dt+G_i\sqrt{\frac{2V_i}{N_i}}dW_3,
\end{multline}
\end{subequations}

where

\begin{subequations}
\begin{align}
R_i = & \ \ln \left(\frac{Q_iU_i}{\sqrt{1+A_iw_i}}\right), \\
a_i = & \ \frac{A_i}{1+A_iw_i}, \\
b_{ij}(t) = b_{ji}(t) = & \ \big(w_i+w_j+\eta_i+\eta_j+G_i(t)+G_j(t)\big)^{-1}, \\
c_i \geq & \ 0.
\end{align}
\end{subequations}

Despite the convoluted appearance of system (\ref{comm_dynamics}), there are some nice features that reflect biological reasoning. For example, the dynamics of abundance are just a generalization of Lotka-Volterra dynamics. In particular, the effect of competition with species $j$ on the fitness of species $i$ grows linearly with $N_j$. However, as biotic selection pushes $\bar x_i$ away from $\bar x_j$, the effect of competition with species $j$ on the fitness of species $i$ rapidly diminishes, reflecting a reduction in niche overlap. The divergence of $\bar x_i$ and $\bar x_j$ due to competition is referred to in the community ecology literature as character displacement. We also see that the fitness of species $i$ drops quadratically with the difference between $\bar x_i$ and the abiotic optimum $\theta_i$. Hence, abiotic selection acts to pull $\bar x_i$ towards $\theta_i$. The response in mean trait $\bar x_i$ to natural selection is proportional to the amount of heritable variation in the population, represented by the additive genetic variance $G_i$. However, we have that $G_i$ is itself a dynamic quantity. Under our model, abiotic stabilizing selection erodes away heritable variation at a rate that is independent of both $N_i$ and $\bar x_i$. The effect of competition on $G_i$ is a bit more complicated. When $b_{ij}(\bar x_i-\bar x_j)^2<1$, competition with species $j$ acts as diversifying selection which tends to increase the amount of heritable variation. However, when $b_{ij}(\bar x_i-\bar x_j)^2>1$, competition with species $j$ acts as directional selection and reduces $G_i$. In the following subsections we demonstrate the behavior of system (\ref{comm_dynamics}) by plotting numerical solutions and investigate implications for the relationship between the strength of ecological interactions and selection.

## Community dynamics \label{dynamics}

For the sake of illustration we numerically integrated system (\ref{comm_dynamics}) for a richness of $S=100$ species. We assumed homogeneous model parameters across species in the community as summarized by Table \ref{par_vals}. We repeated numerical integration under the two scenarios of weak and strong competition. For the first scenario of weak competition we set $c=1.0\times10^{-7}$ and for the second scenario of strong competition we set $c=5.0\times10^{-6}$. With these two sets of model parameters, we simulated our model for $1000.0$ units of time. For both scenarios, we initialized the trait means to $\bar x_i=0.0$, additive genetic variances to $G_i=10.0$ and abundances to $N_i=1000.0$ for each $i=1,\dots,S$.

Temporal dynamics for each scenario are provided in Figure \ref{temporal}. This figure suggests weaker competition leads to smoother dynamics and a higher degree of organization within the community. Considering expression (\ref{N_comm}) we note that, all else equal, relaxed competition allows for larger growth rates which promote greater abundances. From (\ref{N_comm}) we also note that the per-capita effects on demographic stochasticity diminish with abundance. To see this, divide both sides by $N_i$. Inspecting expressions (\ref{x_comm}) and (\ref{G_comm}), we see that larger abundances also erode the effects of demographic stochasticity on the evolution of mean trait and additive genetic variance. These effects were already noted in \S\ref{equations}, and thus are not a consequence of our model of coevolution per-se, but we revisit them here since Figure \ref{temporal} demonstrates the importance of demographic stochasticity in structuring ecological communities even when populations are very large. Hence, contrary to the common assumption that stochastic effects can be ignored for large populations, we find that minute asymmetries by demographic stochasticity remain significant drivers of community structure. In particular, although we initialized the species with identical state variables and model parameters, we find an enourmous amount of asymmetry and even some potential phase changes. In the following two paragraphs we describe the natural history of the community as illustrated in Figure \ref{temporal}.

After about $125.0$ units of time, the community appears to have shaken off the initial conditions and entered into a qualitatively distinct phase of dynamics. Aside from a few outliers, most of the species remain clustered together in their state variables. This lasts for approximately $375.0$ units of time until, at around time $500.0$, a drastic change occurs. At this moment the tightly packed cluster of species begins to fan out in all three state variables. Simultaneously, we observe large a shift in mean traits for higher values and in additive genetic variances for lower values. Upon inspecting our calculations, we diagnose the reason for this shift. The outlier species that were initially pushed away from the common abiotic optimum ($0.0$ in this case) evolved a significant reduction in the quantity of heritable variation ($\approx60\%$) due to directional selection induced by competition. This reduction in heritable variation slowed adaptation, causing these species to linger on the outskirts of niche space, some longer than others. In the meantime the rest of the community, being tightly packed, experienced greater competition which led to diminished abundances for these species and caused some members of the core group to veer away from the abiotic optimum. The reduced abundances of the core group led to reduced competition overall. As a result, the outlier populations were given a slight increase in growth rate, enough to allow them to increase their abundances orders of magnitude higher than the species in the core group and giving them more weight in driving the evolution of other species. Many of these heavy-hitting outlier species had already been maintaining negative mean traits, but around time $500.0$ the high abundance species with positive mean traits began to experience enough intraspecific competition to override interspecific competition and begin evolving towards the abiotic optimum. The sudden imbalance of these high abundance species effectively induced a single large competitive exclusion event pushing the majority of the community far away from the abiotic optimum. After this shift the cluster began to slowly bloom in all three state variables as species took advantage of novel asymmetries in their competitive abilities mediated by a new distribution of mean trait values across the community. About $125.0$ units of time later, the community reached a qualitatively new phase of dynamics. If we kept running the numerical integrator, we would continue to see similar drama unfolding over and over again as minute stochastic changes contribute to asymmetries which slowly build into drastic shifts.

The strong competition scenario is not quite as showy. Although the dynamics of trait means and variances tend to be far more stochastic than in the weak competition scenario, the community overall appears to quickly reach some statistical equilibrium and remain there. However, the abundances across all species in the community are very low due to strength of competition being an order of magnitude higher than in the weak competition case. Although most of the species maintain abundances greater than $1000.0$, we found one species that dropped to an abundance of about $50.0$. If we let the numerical integrator run long enough in this case, we will likely see many of the species go extinct.

Although finding ways to interpret simulated dynamics provides a useful arena to exercise biological reasoning, it does not fulfill our desire to quantify the patterns and processes present in competing communities. In the next subsection we take a step in this direction by using our model to derive formula for selection gradients and competition coefficients. To investigate their relationship, we calculate their covariances using simplifying assumptions on species abundances and intraspecific trait variances. We then investigate how these covariances change with the ratio of variance of interspecific mean traits to variance of intraspecific individual traits and use a numerical approach to investigate correlations between the strength of pairwise coevolution and competition coefficients.

\begin{table}
\centering\caption{Values of model parameters used for numerical integration.}\vspace{0.2cm}
\begin{tabular}{l|l|l}
Parameter     & Description & Value                                           \\ \hline
$R$         & innate growth rate, see \S\ref{ecoevo}  & $1.0$      \\
$\theta$    & abiotic optimum & $0.0$                              \\
$a$         & strength of abiotic selection & $0.01$               \\
$c$         & strength of competition & $\{1.0\times10^{-7},5.0\times10^{-6}\}$ \\
$w$         & niche breadth & $0.1$                               \\
$U$         & total niche use & $1.0$                              \\
$\eta$      & segregation variance & $1.0$                         \\
$\mu$       & mutation rate & $1.0\times10^{-7}$                             \\
$V$ & variance of reproductive output & $5.0$
\end{tabular}\label{par_vals}
\end{table}

```{r, eval=F, echo=F}
require("ggplot2")
require("gridExtra")
require("reshape")

####################
#                  #
# weak competition #
#                  #
####################

smpl_wc = read.csv("/home/bob/Research/Branching Brownian Motion/sample_path_wc.csv")

x_wc = ggplot(smpl_wc,aes(x=time,y=x,group=spp))+theme_minimal()+xlab("")+ylab("Mean Trait")+
  ylim(-12,12)+geom_line(show.legend=F, lwd=0.1)+
  ggtitle("Weak Competition")+theme(plot.title = element_text(hjust = 0.5))

G_wc = ggplot(smpl_wc,aes(x=time,y=G,group=spp))+theme_minimal()+xlab("")+ylab("Additive Genetic \n Variance")+
  ylim(0,45)+geom_line(show.legend=F, lwd=0.1)

N_wc = ggplot(smpl_wc,aes(x=time,y=N,group=spp))+theme_minimal()+xlab("time")+ylab("Abundance")+
  scale_y_log10(limits=c(50,1e8))+geom_line(show.legend=F, lwd=0.1)

smpl_plt_wc = grid.arrange(x_wc,G_wc,N_wc,nrow=3)

#
# calculating strength of coevolution
#

tmax = max(smpl_wc$time)

final_comm = subset(smpl_wc,time==tmax)

x = final_comm$x
G = final_comm$G
N = final_comm$N

w = 0.1
c = 1e-7
eta = 1.0
R = 1.0
a = 1e-2
th = 0
S = length(x)

b = matrix(0,nc=S,nr=S)
C = matrix(0,nc=S,nr=S)
alpha = matrix(0,nc=S,nr=S)
beta = matrix(0,nc=S,nr=S)
gamma = matrix(0,nc=S,nr=S)
for( i in 1:S){
  for( j in 1:S ){

    r = R - a*( (x[i]-th)^2 + G[i] + eta )/2

    b[i,j] = 1/( G[i]+G[j]+2*(w+eta) )

    alpha[i,j] = c*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2) / r

    beta[i,j] = c*N[j]*b[i,j]*(x[i]-x[j])*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2)

    gamma[i,j] = c*N[j]*b[i,j]*(1 - b[i,j]*(x[i]-x[j])^2)*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2)

    C[i,j] = abs(beta[i,j]) + abs(gamma[i,j])

  }
}

Csp = C # defines a matrix that ignores intraspecific interactions
for( i in 1:S ){
  Csp[i,i] = 0
}

Cvec = NULL
for(i in 1:S){
  for(j in i:S)
    if( i!=j ) Cvec = c(Cvec,C[i,j])
}

#
# distribution of coevolution
#

Cdf = data.frame(val=Cvec)

WC_co <- ggplot(Cdf, aes(x=val)) + geom_histogram(aes(y=..count../sum(..count..)), fill='black', alpha=1.0, bins=50) + 
  theme_minimal() + xlab("Pairwise Coevolution") + ylab("Relative Frequency") + 
  scale_x_log10(limits=c(1e-7,1e-1))

#
# building a network
#
require(igraph)
require(plotwidgets)
require(network)
require(GGally)

asp = alpha # defines a matrix that ignores intraspecific interactions
for( i in 1:S ){
  asp[i,i] = 0
}

avec = NULL
for(i in 1:S){
  for(j in i:S)
    if( i!=j ) avec = c(avec,alpha[i,j])
}

# bivariate plot
Cadf_WC = data.frame(Coev=Cvec,Comp=avec, Which="WC")
Ca_WC_pl = ggplot(Cadf_WC, aes(Coev,Comp)) + geom_bin2d(bins=20, show.legend = FALSE) + 
  geom_density_2d(col='white') + scale_fill_gradient(low="black",high="grey",trans="log") +
  scale_x_log10(limits=c(5e-5,1e-2)) + scale_y_log10(limits=c(2e-9,2e-8)) + 
  theme_minimal() + xlab("Pairwise Coevolution") + ylab("Competition Coefficient") + 
  ggtitle("Weak \n Competition") + theme(plot.title = element_text(hjust = 0.5))

network <- graph_from_adjacency_matrix(ceiling(asp/max(asp)))

CDF <- function(x,X){
  L <- length(X)
  k <- 0
  for(i in 1:L){
    if(x >= X[i]) k <- k+1
  }
  k/L
}

resc_wts <- NULL
for(i in 1:length(avec)){
  resc_wts <- c(resc_wts, CDF(avec[i],avec))
}

colors <- colorRamp(c("white","black"))

edge_colors <- NULL
for(i in 1:length(avec)){
  edge_colors <- c(edge_colors,rgb2col(t(colors(resc_wts[i]))))
}

E(network)$width <- resc_wts
E(network)$weights <- resc_wts
E(network)$arrow.size <- 0.01
E(network)$color <- edge_colors
V(network)$size <- N/sum(N)
V(network)$frame.color <- "white"
V(network)$color <- "black"
V(network)$label <- ""
# l <- layout_in_circle(network)
l <- layout_with_fr(network)

WC_net <- network(asp, directed = F)

netWC <- ggnet2(WC_net, mode = "circle", edge.alpha=.1, edge.color = edge_colors, edge.size = 2*min(1/avec)*avec, size = N/sum(N), max_size = 3, node.color = 'black')+
  ggtitle("Weak \n Competition")+theme(plot.title = element_text(hjust = 0.5), legend.position = "none")

######################
#                    #
# strong competition #
#                    #
######################

smpl_mc = read.csv("/home/bob/Research/Branching Brownian Motion/sample_path_mc.csv")

x_mc = ggplot(smpl_mc,aes(x=time,y=x,group=spp))+theme_bw()+theme_minimal()+xlab("")+ylab("")+
  ylim(-12,12)+geom_line(show.legend=F, lwd=0.1)+ggtitle("Strong Competition")+theme(plot.title = element_text(hjust = 0.5))

G_mc = ggplot(smpl_mc,aes(x=time,y=G,group=spp))+theme_bw()+theme_minimal()+xlab("")+ylab("")+
  ylim(0,45)+geom_line(show.legend=F, lwd=0.1)

N_mc = ggplot(smpl_mc,aes(x=time,y=N,group=spp))+theme_bw()+theme_minimal()+xlab("time")+ylab("")+
  scale_y_log10(limits=c(50,1e8))+geom_line(show.legend=F, lwd=0.1)

smpl_plt_mc = grid.arrange(x_mc,G_mc,N_mc,nrow=3)

community_plot = grid.arrange(smpl_plt_wc,smpl_plt_mc,nrow=1)

ggsave("~/Research/A White Noise Approach To Evolutionary Ecology/community_dynamics.png",community_plot,width=6.5,height=5,dpi=320)

#
# calculating strength of coevolution
#

tmax = max(smpl_mc$time)

final_comm = subset(smpl_mc,time==tmax)

x = final_comm$x
G = final_comm$G
N = final_comm$N

w = 0.1
c = 1e-7
eta = 1.0
R = 1.0
a = 1e-2
th = 0
S = length(x)

b = matrix(0,nc=S,nr=S)
C = matrix(0,nc=S,nr=S)
alpha = matrix(0,nc=S,nr=S)
beta = matrix(0,nc=S,nr=S)
gamma = matrix(0,nc=S,nr=S)
for( i in 1:S){
  for( j in 1:S ){

    r = R - a*( (x[i]-th)^2 + G[i] + eta )/2

    b[i,j] = 1/( G[i]+G[j]+2*(w+eta) )

    alpha[i,j] = c*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2) / r

    beta[i,j] = c*N[j]*b[i,j]*(x[i]-x[j])*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2)

    gamma[i,j] = c*N[j]*b[i,j]*(1 - b[i,j]*(x[i]-x[j])^2)*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2)

    C[i,j] = abs(beta[i,j]) + abs(gamma[i,j])

  }
}

Csp = C # defines a matrix that ignores intraspecific interactions
for( i in 1:S ){
  Csp[i,i] = 0
}

Cvec = NULL
for(i in 1:S){
  for(j in i:S)
    if( i!=j ) Cvec = c(Cvec,C[i,j])
}

#
# distribution of coevolution
#

Cdf = data.frame(val=Cvec)

MC_co <- ggplot(Cdf, aes(x=val)) + geom_histogram(aes(y=..count../sum(..count..)), fill='black', alpha=1.0, bins=50) + 
  theme_minimal() + xlab("Pairwise Coevolution") + ylab("") + 
  scale_x_log10(limits=c(1e-7,1e-1))

#
# building a network
#
require(igraph)
require(plotwidgets)
require(network)
require(GGally)

asp = alpha # defines a matrix that ignores intraspecific interactions
for( i in 1:S ){
  asp[i,i] = 0
}

avec = NULL
for(i in 1:S){
  for(j in i:S)
    if( i!=j ) avec = c(avec,alpha[i,j])
}

# bivariate plot
Cadf_SC = data.frame(Coev=Cvec,Comp=avec, Which="SC")
Ca_SC_pl = ggplot(Cadf_SC, aes(Coev,Comp)) + geom_bin2d(bins=20, show.legend = FALSE) + 
  geom_density_2d(col='white') + scale_fill_gradient(low="black",high="grey",trans="log") +
  scale_x_log10(limits=c(5e-7,2e-4)) + scale_y_log10(limits=c(3e-9,1e-8)) + 
  theme_minimal() + xlab("Pairwise Coevolution") + ylab("") + ggtitle("Strong \n Competition") +
  theme(plot.title = element_text(hjust = 0.5))

bivplot = grid.arrange(Ca_WC_pl,Ca_SC_pl,nrow=1)

Cadf = rbind(Cadf_WC,Cadf_SC)
head(Cadf)

on_pl = ggplot(Cadf_WC, aes(Coev,Comp)) + geom_density_2d(col="black") +
  geom_density_2d(data=Cadf_SC,col="red") +
  scale_x_log10(limits=c(5e-7,1e-2)) + scale_y_log10(limits=c(2.5e-9,1.5e-8)) + 
  theme_minimal() + xlab("Pairwise Coevolution") + ylab("Competition Coefficient")

ggsave("~/Research/A White Noise Approach To Evolutionary Ecology/on_pl.png",on_pl,device="png",width=3.5,height=3.5,units="in",dpi=350)

network <- graph_from_adjacency_matrix(ceiling(asp/max(asp)))

CDF <- function(x,X){
  L <- length(X)
  k <- 0
  for(i in 1:L){
    if(x >= X[i]) k <- k+1
  }
  k/L
}

resc_wts <- NULL
for(i in 1:length(avec)){
  resc_wts <- c(resc_wts, CDF(avec[i],avec))
}

colors <- colorRamp(c("white","black"))

edge_colors <- NULL
for(i in 1:length(avec)){
  edge_colors <- c(edge_colors,rgb2col(t(colors(resc_wts[i]))))
}

E(network)$width <- resc_wts
E(network)$weights <- resc_wts
E(network)$arrow.size <- 0.01
E(network)$color <- edge_colors
V(network)$size <- N/sum(N)
V(network)$frame.color <- "white"
V(network)$color <- "black"
V(network)$label <- ""
# l <- layout_in_circle(network)
l <- layout_with_fr(network)

MC_net <- network(asp, directed = F)

netMC <- ggnet2(MC_net, mode = "circle", edge.alpha=.1, edge.color = edge_colors, edge.size = 2*min(1/avec)*avec, size = N/sum(N), max_size = 3, node.color = 'black')+
  ggtitle("Strong \n Competition")+theme(plot.title = element_text(hjust = 0.5), legend.position = "none")

coevs <- grid.arrange(netWC, netMC,
                      WC_co, MC_co, nrow=2)

ggsave("~/Research/A White Noise Approach To Evolutionary Ecology/interaction_networks.png",coevs,device="png",width=6,height=6,units="in",dpi=350)

```

```{r, message=FALSE, echo=F, fig.align='center', fig.cap='\\label{temporal}Temporal dynamics of mean trait (top), additive genetic variance (middle) and abundance (bottom) for the scenario of weak competition (left) and strong competition (right). Red lines indicate average trend across species.'}
knitr::include_graphics('community_dynamics.png')
```

## The relation between the strength of ecological interactions and coevolution \label{ecoevo}

Relating our treatment of the niche to modern coexistence theory [@Chesson2000], the absolute competition coefficient $\alpha_{ij}$ becomes a dynamical quantity that can be written as

\begin{equation}
\alpha_{ij}(t)=\frac{c_i}{r_i(t)}\int_{-\infty}^{+\infty}\int_{-\infty}^{+\infty}p_i(x,t)p_j(y,t)\mathcal O_{ij}(x,y) dxdy =\frac{c_iU_iU_j}{r_i(t)}\sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right),
\end{equation}

where

\begin{equation}
r_i(t)=R_i-\frac{a_i}{2}\Big((\bar x_i(t)-\theta_i)^2+G_i(t)+\eta_i\Big).
\end{equation}

Hence, $dN_i(t)$ can be expressed as

\begin{equation}
dN_i(t)=r_i(t)\left(1-\sum_{j=1}^S\alpha_{ij}(t)N_j(t)\right)N_i(t)dt+\sqrt{V_iN_i(t)}dW_1(t).
\end{equation}

Note that although $r_i(t)$ is referred to in the coexistence literature as the intrinsic growth rate of the population, $R_i$ is a deeper intrinsic quantity. For now we refer to $R_i$ as the _innate_ growth rate. With this connection formally established, researchers may pursue a postmodern coexistence theory that naturally includes the evolutionary dynamics of populations and the effects of demographic stochasticity.

In SM \S\ref{diffuse} we show that the standardized directional selection gradient [sensu @Lande1983] induced by species $j$ on species $i$ can be computed as

\begin{equation}
\beta_{ij}(t)=c_iU_iU_jN_j(t)b_{ij}(t)\big(\bar x_i(t)-\bar x_j(t)\big) \sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right).
\end{equation}

Our notation differs from Lande and Arnold [-@Lande1983] in that subscripts here denote species instead of components of multivariate traits and we drop the prime that distinguishes between selection gradients and standardized selection gradients.

Below we investigate the correspondence of interaction intensity and coevolutionary change. However, we can already identify one major discrepancy; $\alpha_{ij}$ is maximized when $\bar x_i=\bar x_j$, but $\beta_{ij}=0$ under the same condition. We therefore include in our metric of selection the standardized stabilizing selection gradient $\gamma$ which measures the effect of stabilizing or disruptive selection on phenotypic variance [@Lande1983]. In SM \S\ref{diffuse} we show that the standardized stabilizing selection gradient induced by species $j$ on species $i$ can be computed as

\begin{equation}
\gamma_{ij}(t)=c_iU_iU_jN_j(t)b_{ij}(t)\left(1-b_{ij}(t)\big(\bar x_i(t)-\bar x_j(t)\big)^2\right) \sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right).
\end{equation}

To measure the total evolutionary change in species $i$ induced by species $j$, we form the metric $\Psi_{ij}=|\beta_{ij}|+|\gamma_{ij}|$. The top row of Figure \ref{net} displays interaction networks under weak and strong competition where the edge width connecting species $i$ and $j$ is proportional to $\alpha_{ij}\alpha_{ji}$. The bottom row of Figure \ref{net} displays the distributions of pairwise coevolutionary change, which we measure for species $i$ and $j$ via $\mathfrak{C}_{ij}=\Psi_{ij}\Psi_{ji}$, under weak and strong competition.

```{r, echo=F, fig.align='center', fig.cap='\\label{net}Networks of interspecific interactions parameterized by competition coffecients (top row) and distributions of pairwise coevolution (bottom row) under the scenarios of weak competition (left) and strong competition (right) at time $t=1.0\\times 10^3$ Node sizes are proportional to population sizes. Edge widths and shade are monotonically increasing functions of pairwise coevolution.'}
knitr::include_graphics('on_pl.png')
```

We now make use of the expressions derived for competition coeffecients and selection gradients to investigate their relationship. As a first pass, let us assume all model parameters are equivalent across species and that each species has the same abundance and trait variance. Let us further assume that richness $S$ is large and the distribution of mean trait values is normal with mean $\bar{\bar x}$, variance $V_{\bar X}$ and density $f_{\bar X}$. Such assumptions are typical when deriving analytical results in the field of theoretical coevolutionary community ecology [@GuimaresJr2011; @Guimares2017; @Nuismer2012; @Nuismer2018]. If $\bar{\bar x}$ is near $\theta$ and $V_{\bar X}$ is much smaller than $|2R/a-G-\eta|$, then we may approximate $r_i$ with

\begin{equation}
\bar r=\int_{-\infty}^{+\infty}\left(R-\frac{a}{2}\Big((\bar x-\theta)^2+G+\eta\Big)\right)f_{\bar X}(\bar x)d\bar x
=R-\frac{a}{2}\Big((\bar{\bar x}-\theta)^2+V_{\bar X}+G+\eta\Big).
\end{equation}

In SM \S\ref{coef_grad_moms} we use these assumptions to calculate the first and second order moments describing the joint distribution of competition coefficients and selection gradients across the community. We find that the covariance between linear selection gradients and competition coeffecients are zero due to the symmetry implied by our assumptions. However, setting $\alpha(\bar x_i,\bar x_j)=\alpha_{ij}$, $\beta(\bar x_i,\bar x_j)=\beta_{ij}$ and $\gamma(\bar x_i,\bar x_j)=\gamma_{ij}$, the covariances between the magnitude of linear selection gradients and competition coefficients and between stabilizing selection gradients and competition coefficients can be written as

\begin{subequations}

\begin{equation}\label{cov_alpha_beta}
\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)=\frac{2c^2b^2U^4N}{\pi\bar r}\sqrt\frac{V_{\bar X}}{2\pi}\left(\frac{1}{(1+8bV_{\bar X})^{3/4}}-\frac{1}{(1+4bV_{\bar X})^{3/4}(1+2bV_{\bar X})^{1/2}}\right),
\end{equation}

\begin{equation}\label{cov_alpha_gamma}
\mathrm{Cov}_{f_{\bar X}}(\alpha,\gamma)=\frac{c^2b^2U^4N}{2\pi\bar r}(1-2bV_{\bar X})\left(\frac{1}{\sqrt{1+4bV_{\bar X}}}-\frac{1}{1+2bV_{\bar X}}\right).
\end{equation}

\end{subequations}

```{r, eval=F, echo=F}
c=1e-4
b=1e-1
r=0.1
N=1e10
Cov_ab <- function(x){
  # x <- 2*b*V_{\bar X}
  val = 2*c^2*b^2*N / (pi*r)
  val = val * sqrt(x/(2*b*pi))
  val2 = (1+8*x/2)^(-0.75)
  val3 = (1+4*x/2)^(-0.75) / sqrt(1+2*x/2)
  val = val * (val2 - val3)
  return(val)  
}
Cov_ag <- function(x){
  # x <- 2*b*V_{\bar X}
  val = 2*c^2*b^2*N / (2*pi*r)
  val = val * (1-2*x/2)
  val2 = 1 / sqrt(1+4*x/2)
  val3 = 1 / (1+2*x/2)
  val = val * (val2 - val3)
  return(val)  
}
require(ggplot2)
require(ggthemes)
require(gridExtra)
dummy <- ggplot(data = data.frame(x = 0), mapping = aes(x = x))
cov_ab_p <- dummy + stat_function(fun = Cov_ab) + ylim(-0.4,0.4) + xlim(0,2) + theme_minimal() + xlab("Approximate ratio of variance among \n species to variance within species") + ylab("Covariance of \u03B1 and |\u03B2|")
cov_ag_p <- dummy + stat_function(fun = Cov_ag) + ylim(-0.4,0.4) + xlim(0,2) + theme_minimal() + xlab("Approximate ratio of variance among \n species to variance within species") + ylab("Covariance of \u03B1 and \u03B3")
cov_plot <- grid.arrange(cov_ab_p,cov_ag_p,nrow=1) 
ggsave("~/Research/A White Noise Approach To Evolutionary Ecology/cov.png",cov_plot,width=6.5,height=2.5)
```

```{r, echo=F, fig.align='center', fig.cap='\\label{cov_fig}Curves representing the covariance between the magnitude of linear selection gradients and competition coefficients (left) and between stabilizing selection gradients and competition coefficients (right) as a function of $2bV_{\\bar X}$ which is approximately equal to the ratio of variance in mean traits among species to the intraspecific trait variance. In both plots we set $c=1.0\\times10^{-4}$, $b=0.1$, $\\bar r=0.1$ and $N=1.0\\times10^{10}$ and let $V_{\\bar X}$ vary.'}
knitr::include_graphics('/home/bob/Research/A White Noise Approach To Evolutionary Ecology/cov.png')
```

For fixed $c,b,\bar r$ and $N$, we visualize these relationships in Figure \ref{cov_fig}. To gain insight into the relationship between selection gradients and competition coefficients, note that our assumptions in this section imply $b^{-1}=2(\sigma^2+w)$. If we further assume $\sigma^2+w\approx\sigma^2$, then $2bV_{\bar X}\approx V_{\bar X}/\sigma^2$. That is, when populations are generalists and are comprised of specialist individuals, the value $2bV_{\bar X}$ is approximately equal to the ratio of interspecific mean trait variation to intraspecific individual trait variation. Hence, for both covariances we see that there is no relationship between selection gradients and competition coefficients when this ratio is zero. From equation (\ref{cov_alpha_beta}) we can use numerical optimization to find that when $V_{\bar X}/\sigma^2\approx1.25$ the relationship between the magnitudes of linear selection gradients and competition coefficients disappears, but when (approximately) $V_{\bar X}/\sigma^2<1.25$ ($>1.25$), this covariance becomes negative (positive). Equation (\ref{cov_alpha_gamma}) states that when $V_{\bar X}/\sigma^2$ is approximately equal to one (or slightly larger), there is no expected relationship between competition coefficients and quadratic selection gradients. However, when $V_{\bar X}/\sigma^2<1.0$ ($>1.0$), then we expect a positive (negative) relationship between $\alpha$ and $\gamma$. These results are true regardless of the chosen parameter values. In SM \S\ref{coef_grad_moms} we use simulations of system (\ref{comm_dynamics}) to show that these results do not qualitatively differ when allowing for heterogeneous population sizes and additive genetic variances across species.

From a biological perspective, if the ratio $V_{\bar X}/\sigma^2$ is small, then species are packed tightly in phenotypic space. In our model this occurs when abiotic stabilizing selection is much stronger than competition ($a\gg c$). This causes species to overlap more in niche space (i.e., large $\alpha$) and creates disruptive selection for greater intraspecific variance (i.e., positive $\gamma$), which explains the positive region of covariance between $\alpha$ and $\gamma$. However, as species begin to overlap in niche space, directional selection begins to vanish (i.e., small $|\beta|$), leading to a negative covariance between $\alpha$ and $|\beta|$. In the limiting case that two species have perfectly overlapping niches, they will exhibit zero directional selection since a shift in either direction will yield the same fitness advantages.

In the opposite scenario where competition is much stronger than abiotic stabilizing selection ($c\gg a$), species will not evolve to be as tightly packed and instead their niche-centers will be spread out with little ovelap in their resouce utilization curves (i.e., small $\alpha$). In this case biotic directional selection will be strong (i.e., large $|\beta|$), particularly for species towards the outer regions of niche space due to asymmetric fitness advantages conferred by shifts in niche-centers. This leads to a positive covariance between $\alpha$ and $|\beta|$. However, as noted above, this directional selection will also erode away at standing heritable variation (i.e., negative $\gamma$), reducing the rate at which adaptation can occur and creating a negative covariance between $\alpha$ and $\gamma$.

In summary, we see the relation between competition coefficients and selection is highly non-trivial and depends on the relative magnitudes of different ecological processes shaping the community. However, this does not address the relation between competition coefficients and coevolution per se. In SM \S\ref{coef_grad_moms} we show that calculating a formula for the covariance between competition coefficients and the metric of coevolution $\mathfrak{C}$ introduced above provides a difficult analytical challenge. Instead of confronting this challenge we build on our numerical approach used to justify analytical approximations of $\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)$ and $\mathrm{Cov}_{f_{\bar X}}(\alpha,\gamma)$ to approximate the correlation of $\alpha$ and $\mathfrak{C}$. This numerical approach inherits the assumptions of homogeneous background parameters such as the mutation rate $\mu$ and abiotic optima $\theta$, but allows us to relax the assumption that $N$ and $G$ are constant across species and time.

In particular, we numerically integrated system (\ref{comm_dynamics}) for $T_1=1000.0$ units of time and then continued to integrate for $T_2=1000.0$ units of time. We then calculate the covariance between the quantities $\alpha$ and $\mathfrak{C}$ across $S=100$ species for each of the last $T_2$ time steps. We assume the temporal average of these covariances across the last $T_2$ units of time approximates the expectation at equilibrium. We repeated this approach for randomly drawn $a$ and $c$ until our sample size reached $1000$. In Figure \ref{cor_coev_fig} we plot the temporally averaged values of $\mathrm{Cov}_{f_{\bar X}}(\alpha,\mathfrak{C})$ against the strength of competition $c$. Using a cubic regression, we see the correlation of coevolutionary selection gradients and competition coefficeints is negative at variance ratios below $0.5$, zero at variance ratios between $0.5$ and $1.0$, and drops below zero again at variance ratios above $1.0$.

```{r, eval=F, echo=F}
require(ggplot2)
require(gridExtra)
df = read.csv("~/Research/Branching Brownian Motion/corrs.csv")
Cαβp = ggplot(df,aes(x=df$V/df$σ,y=df$Cαβ))+geom_point(size=0.1)+stat_smooth(color="black", method = "lm", formula = y ~ x + I(x^2) + I(x^3))+theme_minimal() + ylab("Correlation of \u03B1 and \u03B2")+xlab("Ratio of variance among \n species to variance within species")
Cαabsβp = ggplot(df,aes(x=df$V/df$σ,y=df$Cαabsβ))+geom_point(size=0.1)+stat_smooth(color="black", method = "lm", formula = y ~ x + I(x^2) + I(x^3))+theme_minimal()+ylab("Correlation of \u03B1 and |\u03B2|")+xlab("Ratio of variance among \n species to variance within species")
Cαγp = ggplot(df,aes(x=df$V/df$σ,y=df$Cαγ))+geom_point(size=0.1)+stat_smooth(color="black", method = "lm", formula = y ~ x + I(x^2) + I(x^3))+theme_minimal()+ylab("Correlation of \u03B1 and \u03B3")+xlab("Ratio of variance among \n species to variance within species")
Cαℭp = ggplot(df,aes(x=df$V/df$σ,y=df$Cαℭ))+geom_point(size=0.1)+stat_smooth(color="black", method = "lm", formula = y ~ x + I(x^2) + I(x^3))+theme_minimal()+ylab("Correlation of \u03B1 and ℭ")+xlab("Ratio of variance among \n species to variance within species")
corrsp = grid.arrange(Cαabsβp,Cαγp,nrow=1)
ggsave("/home/bob/Research/A White Noise Approach To Evolutionary Ecology/corrs.png",  corrsp,  width=6.5,height=2.5)
ggsave("/home/bob/Research/A White Noise Approach To Evolutionary Ecology/Cαℭp.png",   Cαℭp,  width=4,height=2.5)
```

```{r, echo=F, fig.align='center', fig.cap='\\label{cor_coev_fig}Numerical estimates for the correlation of the strength of coevolution measured by $\\mathfrak{C}$ and competition coefficeints $\\alpha$ plotted against the variance of mean traits among species divided by the mean variance of traits within species. Each dot represents the result from a single simulation. The red line is a cubic regression.'}
knitr::include_graphics('/home/bob/Research/A White Noise Approach To Evolutionary Ecology/Cαℭp.png')
```

# Conclusion \label{conclusion}

We have introduced an approach to derive models of evolutionary ecology using the calculus of white noise, demonstrated our approach by deriving a model of diffuse coevolution and investigated the relationship between selection gradients and competition coefficients, finding that these quantities exhibit interesting relationships which shed light on the feedback between the structure and dynamics of ecological communities. However, although our approach has the merit of rigorously synthesizing the dynamics of abundance and phenotypic evolution for populations experiencing demographic stochasticity, there remains biological details and their associated technical challenges that have yet to be confronted for gaining a more thorough and rigorous understanding of ecological communities. We touch on just four of them here and provide suggestions for future research to approach these challenges.

__Limitations of diffusion limits__

As noted by Feller [-@feller1951], although diffusion limits provide reasonable approximations for large populations, relatively small populations require discrete models. Hence, as a diffusion limit, SPDE (\ref{SPDE}) restricts parameter values to regions that maintain large population sizes. Although this suggests an important restriction on any model developed under this framework by implying populations are not at risk of extinction, the SDE describing abundance dynamics technically permits extinction. However, for small abundances, pathology emerges in the evolution of trait means and variances. In particular, stochastic components of the SDE describing the evolution of $\bar x$ and $\sigma^2$ diverge towards infinite values as $N\to0$. Hence, studies of evolutionary rescue and colonization-extinction dynamics that incorporate phenotypic evolution should be pursued via a different approach. A natural alternative can be developed utilizing the underlying BBM that SPDE (\ref{SPDE}) is a diffusion limit of (see section \ref{stochastic}). This process explicitly tracks individuals throughout their life-history and captures reproduction as branching events. Hence, BBM processes model population size as a non-negative integer instead of a continuously varying number. In particular, the pathological behavior described for the diffusion limit does not occur for BBM as population size goes to zero.

__The genetic architecture and distributions of traits and the role of sexual reproduction__

Our treatment of inheritance and our approach to model coevolution rest on the assumptions of normally distributed breeding values and expressed phenotypes along with asexual reproduction. However, real traits are not encoded by an infinite number of loci each contributing an additive infinitesimal effect (as required by the infinitesimal model), mutations are not inherited as normally distributed deviations from parental breeding values (as required by the Gaussian descendents model) and many populations of interest exhibit non-random sexual reproduction. Departures from this model of genetic architecture can produce more complex distributions of breeding values and expressed traits. Such deviations can be reinforced via strong non-Gaussian selection surfaces, including the surface $m(\nu,x)$ we have derived from niche theory, along with non-random mating in sexually reproducing populations. However, Gaussian approximations are convenient since they are defined by their mean and variance. Future work investigating the effects of non-normally distributed traits on the structure and dynamics of ecological communities will therefore need to confront higher moments such as skew and kurtosis, ideally integrating previously established mathematical approaches to derive the dynamics of these higher moments [@Dbarre2015EvolutionOQ].

An alternative approach to breaking the assumption of normally distributed trait values is the development of explicit multilocus models. These models describe the contributions of alleles at particular loci in the genome to the development of quantitative traits. Tracking the fluctuations of allele frequencies then allows theoreticians to investigate the dynamics of phenotypic distributions that deviate from normality. While this approach has a long history in theoretical quantitative genetics [@bulmer1980; @pmid7851785; @Kirkpatrick1727] and coevolutionary theory [@Nuismer2005; @Kopp2006; @Nuismer2007], work to investigate the role of genetic architecture in mediating feedbacks between the dynamics of population abundances and the distributions of traits mediating ecological interactions has apparently only just begun [@Schreiber2018; @Patel2019].

__The role of ecological stoichiometry__

Our treatment of both biotic and abiotic selection neglects important chemical constraints of biological reality. For instance, the resource we assume species are competing over is modelled as a static quantity. However, real resources can be dynamic quantities. Such theoretical quantities may reflect abiotic cycles of material and energy or whole trophic layers comprised of prey, hosts or mutualists. Although resource dynamics have been captured theoretically in consumer-resource models [@davidtilman1982], developing a more realistic model of resource competition must incorporate details on the ecophysiology of organisms [@michelloreau2010]. Doing so will help clarify the feedback between the evolution of populations and the ecosystem processes they are a part of.

Using plant-pollinator communities as an example, the role of nitrogen mediating interspecific interactions has been reviewed by David, Storkey and Stevens [-@David2019] and the evolutionary ecology of the nutritional content of nectar has been reviewed by Parachnowitsch, Manson and Sletvold [-@Parachnowitsch2018]. These studies demonstrate the need for further research to understand how soil nutrient availability and organismal ecophysiology structures communities of plants and pollinators. Theoretical pursuits in this directions may profit from interfacing the framework we have outlined here with population-ecosystem models such as that developed by Fridley [-@Fridley2017].

__Accounting for macroevolutionary history__

To understand patterns found in ecological communities, considerations must push beyond microevolutionary and contemporary ecological processes and consider the macroevolutionary dynamics of ancestral lineages leading to extant populations. Using sub-alpine flower communities as an example, one can observe a very strict ordering of phenology across broad geographic ranges. In particular, whether in the Colorado Rocky mountains (such as Gothic, Colorado) or on an outlier of the Idaho batholith (such as Kamiak butte near Palouse, Washington), one will almost surely observe a very conspicuous order of flowers emerging in early spring: at the very beginning of the season blooms _Claytonia lanceolata_ followed by _Erythronium grandiflorum_ which precedes _Delphinium nuttallianum_ (B. Week, personal observations). If contemporary phenological coevolution is rampant, why should this pattern be so well preserved across a thousand miles of rugged and diverse terrain? Although it would be exciting to find that these species repeatedly coevolved this pattern in each location, a more parsimoniuous hypothesis suggests the phenology and physiology of these species slowly evolved independently over macroevolutionary time scales to take advantage of the specific conditions available within each of these windows of the flowering season. However, this could not have carried out in the Rocky mountains since this terrain only became habitable just over ten thousand years ago as the glaciers of the Pleistocene began to retreat (Paul CaraDonna, personal communications). Hence, given these considerations, it appears that an understanding of early season phenology patterns should focus on how these communities are assembled as opposed to contemporary evolutionary dynamics. Indeed, recent work testing models of phylogeography ignores the potential for contemporary evolution and instead suggests alpine flower communities tend to follow neutral assembly where flowers merely compete for who can disperse to new habitat first, as opposed to a selective process where a regional species pool is filtered for those species adapted to the newly available habitat [@Marx2017].

Of course microevolutionary and ecological dynamics are not completely irrelevant for understanding patterns in communities that are primarily structured by deep evolutionary processes. In particular, macroevolutionary trait evolution is simply the aggregation of microevolutionary change occuring over large spans of time. This suggests a road forward to connect the theory we have introduced to models of macroevolutionary trait evolution.

Some approaches to modelling macroevolutionary trait change simply repurpose microevolutionary models by blindly rescaling time from the units of generations to millions of years [@Nuismer2014; Luke, can you think of others?]. Such an approach makes the implicit assumption that trait evolution is statistically self-similar [sensu @kennethfalconer2014] so that the stochastic evolution of traits on macroevolutionary time scales has the same properties of trait evolution on microevolutionary time scales. Although some stochastic processes, including Brownian motion, do exhibit self-similarity, others do not. For example, consider a modification of the Ornstein-Uhlenbeck process defined by the SDE

\begin{equation}
dX_t=a(\theta_t-X_t)dt+bdW_t
\end{equation}

where $a,b>0$, $W_t$ is a standard Brownian motion and $\theta_t$ is itself a Markov process that takes normally distributed jumps centered on its current location at exponentially distributed time intervals. If we assume the rate $\lambda$ at which jumps occur is much smaller than $a$ and the variance in jumping is much larger than $b^2$, then, even though the sample paths of $X_t$ are actually continuous (if we zoom in close enough, they look like Brownian motion), over long intervals of time sample paths of $X_t$ will begin to appear to have periods of continuity interrupted by an occasional discontinuous jump and thus approach a qualitatively distinct process. These emergent properties can be formally characterized by L&eacute;vy processes and have been successfully employed in comparative phylogenetics to fit phenotypic data from extant populations and the fossil record [@Landis2017]. It would therefore be interesting to investigate whether an application of a separation of time scales argument for the rate of environmental change ($\lambda$) versus the rate of evolutionary and ecological change ($a$) to microevolutionary models derived using our framework can be used to obtain macroevolutionary models that include not only mean trait evolution, but also the evolution of trait variance and abundance. The resulting macroevolutionary models can give rise to novel comparative phylogenetic methods and provide initial conditions for microevolutionary models that capture contemporary dynamics.

__Final remarks__

Although top-down approaches to community ecology, such as the Maximum Entropy Theory of Ecology [@Harte2011], have enjoyed some success in describing community-level patterns [@Harte2014; @Xiao2015], a mechanistic understanding of why these patterns emerge and how they will change remains a formidable task. Such an understanding must take both bottom-up and top-down approaches integrating considerations from the ecophysiology of individual organisms that reveal the economics of interspecific interactions [@Sterner2008], to the phylogeographic history of taxa that sets the stage for contemporary dynamics [@Hickerson2010]. Through connecting these dots we can increase the variance explained in observations of ecological communities by specific mechanisms and come closer to a predictive theory of evolutionary community ecology. Despite the long list of equations derived in this paper, this work takes just one small step towards capturing these many details. However, we hope the theoretical framework outlined here along with the demonstration of its use in modelling competitive communities provides some helpful tools to aid quantitative evolutionary ecologists in reaching such lofty goals.

\newpage

# Supplementary material (SM)

Throughout this supplement, we set use dot notation for time derivatives so that $\dot f(x,t)=\frac{\partial }{\partial t}f(x,t)$ and set $\Delta=\frac{\partial^2}{\partial x^2}$, except in \S\ref{caC} where $\Delta$ represents a random variable.

## Sufficient conditions for finite mean, variance and total abundance in the deterministic case \label{finite}

Recall $C_{1,c}^+(\mathbb{R}\times[0,\infty))$ denotes the set of non-negative functions $\nu(x,t)$ continuous in both $x$ and $t$ such that $\int_{\mathbb{R}}\nu(x,t)dx<+\infty$. We assume, for $m:C_{1,c}^+(\mathbb{R}\times[0,\infty))\times\mathbb{R}\to\mathbb{R}$, there exists $r\in\mathbb{R}$ such that $m(\nu,x)\leq r$ for each $\nu\in C_{1,c}^+(\mathbb{R}\times[0,\infty))$ and $x\in\mathbb{R}$. As in the main text, we also assume the initial condition $\nu(x,0)$ satisfies
\begin{equation}
N(0)=\int_{\mathbb{R}}\nu(x,0)dx<+\infty,
\end{equation}
\begin{equation}
-\infty<\bar x(0)=\int_{\mathbb{R}}xp(x,0)dx<+\infty,
\end{equation}
\begin{equation}
\sigma^2(0)=\int_{\mathbb{R}}xp(x,0)dx<+\infty,
\end{equation}
where $p(x,0)=\nu(x,0)/N(0)$, and we consider the PDE
\begin{equation}\label{PDE_SM}
\dot\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\Delta\nu(x,t).
\end{equation}
Replacing $m$ with it's upper bound $r\in\mathbb{R}$, equation (\ref{PDE_SM}) reduces to a simple parabolic equation that can be solved using standard techniques [@stanleyfarlow1993]. In particular, when $m(\nu,x)\equiv0$ denote the solution to (\ref{PDE_SM}) by $\nu_0(x,t)$. Then, denoting
\begin{equation}
\Phi(x,t)=\frac{\exp\left(-x^2/2\mu t\right)}{\sqrt{2\pi\mu t}},
\end{equation}
we have
\begin{equation}
\nu_0(x,t)=\int_{\mathbb{R}}\Phi(x-y,t)\nu(y,0)dy.
\end{equation}
In the more general case, when $m(\nu,x)\equiv r\in\mathbb{R}$, equation (\ref{PDE_SM}) has the solution $\nu_r(x,t)=e^{rt}\nu_0(x,t)$. Hence, $\nu_r(x,t)\geq0$ for all $x\in\mathbb{R}$ and $\int_\mathbb{R}\nu_r(x,t)dx=e^{rt}N(0)<+\infty$ for all $t\geq0$. Furthermore, denoting $N_r(t)=\int_\mathbb{R}\nu_r(x,t)dx$, $p_r(x,t)=\nu_r(x,t)/N_r(t)$, $\bar x_r(t)=\int_\mathbb{R}xp_r(x,t)dx$ and $\sigma^2_r(t)=\int_\mathbb{R}(x-\bar x_r(t))^2p_r(x,t)dx$, we have
\begin{equation}
\bar x_r(t)=\int_\mathbb{R}x\int_\mathbb{R}\Phi(x-y,t)p(y,0)dydx=\int_\mathbb{R}yp(y,0)dy=\bar x(0),
\end{equation}
\begin{equation}
\sigma_r^2(t)=\int_\mathbb{R}(x-\bar x_r(t))^2\int_\mathbb{R}\Phi(x-y,t)p(y,0)dydx=\int_\mathbb{R}\Big((y-\bar x(0))^2+\mu t\Big)p(y,0)dy=\sigma^2(0)+\mu t.
\end{equation}

Hence, $|\bar x_r(t)|,\sigma_r^2(t)<+\infty$ for all $t\geq0$. For the sake of contradiction, suppose there exists $x\in\mathbb{R}$ and $t\geq0$ such that $\nu(x,t)>\nu_r(x,t)$. Then
\begin{equation}
\nu(x,t)-\nu(x,0)=\int_0^tm(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)ds>\int_0^tr\nu_r(x,s)+\frac{\mu}{2}\Delta\nu_r(x,s)ds=\nu_r(x,t)-\nu(x,0)
\end{equation}
which implies there exists $\nu\in C^+_{1,c}(\mathbb{R}\times[0,\infty))$ and $x\in\mathbb{R}$ such that $m(\nu,x)>r$. But this contradicts our assumption $m(\nu,x)\leq r$ for all $\nu\in C^+_{1,c}(\mathbb{R}\times[0,\infty))$ and $x\in\mathbb{R}$. So we have $\nu(x,t)\leq\nu_r(x,t)$ for each $x\in\mathbb{R}$ and $t\geq0$. This implies that $N(t)=\int_\mathbb{R}\nu(x,t)dx<+\infty$,
\begin{equation}
0<\int_\mathbb{R}x^2\nu(x,t)dx\leq\int_\mathbb{R}x^2\nu_r(x,t)dx<+\infty
\end{equation}
and in particular
\begin{equation}
0<\sigma^2(t)+\bar x^2(t)=\frac{1}{N(t)}\int_\mathbb{R}x^2\nu(x,t)dx<+\infty
\end{equation}
for each $t\geq0$.

## The relation between diffusion and convolution with a Gaussian kernel \label{diffconvequiv}

Let $g:\mathbb{R}^d\to\mathbb{R}$ be smooth. Consider the deterministic Cauchy problem

\begin{equation}\label{heateqn}
\left\{\begin{matrix}
\dot f(x,t)=&\Delta f(x,t), & (x,t)\in\mathbb{R}^d\times(0,\infty)\\
f(x,t)=&g(x), & (x,t)\in\mathbb{R}^d\times\{0\}.
\end{matrix}\right.  \tag{SM1.1}
\end{equation}

According to Evans [-@lawrenceevans2010], the fundamental solution of (\ref{heateqn}) is

\begin{equation}
\Phi(x,t)=\frac{1}{(4\pi t)^{d/2}}\exp\left(-\frac{|x|^2}{4t}\right), \ (x,t)\in(0,\infty)\times\mathbb{R}^d, \tag{SM1.2}
\end{equation}

where $|x|=\sqrt{\sum_ix_i^2}$. The solution $f(x,t)$ of PDE (\ref{heateqn}) is then given by the convolution

\begin{equation}
f(x,t)=\int_{\mathbb{R}^d}\Phi(x-y,t)g(y)dy, \ (x,t)\in(0,\infty)\times\mathbb{R}^d. \tag{SM1.3}
\end{equation}

Hence, by the fundamental theorem of calculus,

\begin{equation}
f(x,t)+\int_t^{t+1}\dot f(x,s)ds=f(x,t+1) \tag{SM1.4}
\end{equation}

$$=\int_{\mathbb{R}^d}\Phi(x-y,t+1)g(y)dy=\int_{\mathbb{R}^d}\int_{\mathbb{R}^d}\Phi(x-y,1)\Phi(y-z,t)g(z)dzdy$$

$$=\int_{\mathbb{R}^d}\Phi(x-y,1)f(t,y)dy.$$

In particular,

\begin{equation}
f(x,t)+\int_t^{t+1}\Delta f(x,s)ds=\int_{\mathbb{R}^d}\Phi(1,x-y)f(y,t)dy. \tag{SM1.5}
\end{equation}

## Deterministic dynamics of $\sigma^2(t)$ \label{var_deriv}

Picking up from the main text \S2.1,

\begin{equation}
\dot{\sigma}^2(t)=\frac{d}{dt}\int_\mathbb{R}(x-\bar x(t))^2p(x,t)dx
\end{equation}

$$=\int_\mathbb{R}2(x-\bar x(t))\dot{\bar x}(t)+(x-\bar x(t))^2\dot p(x,t)dx$$

$$=\int_\mathbb{R}(x-\bar x(t))^2\left((m(\nu,x)-\bar m(t))p(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t)\right)dx$$

$$=\int_\mathbb{R}\left((x-\bar x(t))^2-\sigma^2(t)+\sigma^2(t)\right)(m(\nu,x)-\bar m(t))p(x,t)+(x-\bar x(t))^2\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t)dx$$

$$=\mathrm{Cov}_t\Big((x-\bar x(t))^2,m(\nu,x)\Big)+\frac{\mu}{2}\int_\mathbb{R}(x-\bar x(t))^2\frac{\partial^2}{\partial x^2}p(x,t)dx.$$

In particular, when $p(x,t)$ is Gaussian,

\begin{equation}
\frac{\partial^2}{\partial x^2}p(x,t)=\frac{(x-\bar x(t))^2-\sigma^2(t)}{\sigma^4(t)}p(x,t)
\end{equation}

and hence, referring to the well-known moments of the Gaussian distribution, we find

\begin{equation}
\int_\mathbb{R}(x-\bar x(t))^2\frac{\partial^2}{\partial x^2}p(x,t)dx
\end{equation}

$$=\int_\mathbb{R}\frac{(x-\bar x(t))^4-(x-\bar x(t))^2\sigma^2(t)}{\sigma^4(t)}p(x,t)dx=2.$$

In the Gaussian case, we also have

\begin{equation}
2\sigma^4\left(\frac{\partial\bar m}{\partial\sigma^2}-\overline{\frac{\partial m}{\partial\sigma^2}}\right)=2\sigma^4\left(\frac{\partial}{\partial\sigma^2}\int_\mathbb{R}m(\nu,x)p(x,t)dx-\int_\mathbb{R}p(x,t)\frac{\partial}{\partial\sigma^2}m(\nu,x)dx\right)
\end{equation}

$$=2\sigma^4\int_\mathbb{R}\frac{(x-\bar x)^2-\sigma^2}{2\sigma^4}m(\nu,x)p(x,t)dx=\int_\mathbb{R}\left((x-\bar x)^2-\sigma^2\right)\big(m(\nu,x)-\bar m\big)p(x,t)dx$$

$$=\mathrm{Cov}_t\Big((x-\bar x)^2,m\Big).$$

## Numerical evidence of finite moments and approximate normality in the stochastic case \label{numerical}

Here we use a numerical argument to suggest, for

\begin{equation}
r-\frac{a}{2}(\theta-x)^2-c\int_\mathbb{R}\nu(x,t)dx \leq m(\nu,x)\leq r-\frac{a}{2}(\theta-x)^2, \ \forall (\nu,x)\in C_{1,c}^+(\mathbb{R})\times\mathbb{R},
\end{equation}

the density process $\nu(x,t)$ defined by SPDE (\ref{SPDE}) of the main text satisfies $\int_\mathbb{R}(|x|+x^2)\nu(x,t)dx<\infty$. From SM \S\ref{finite}, under the assumption $m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_\mathbb{R}\nu(x,t)dx$, we can derive the differential equations

\begin{subequations}
\begin{equation}
\dot{\bar x}=aG(\theta-\bar x)
\end{equation}
\begin{equation}
\dot G=\mu-aG^2
\end{equation}
\begin{equation}
\dot{N}=\left(r-\frac{a}{2}\left((\theta-\bar x)^2+G+\eta\right)-cN\right)N.
\end{equation}
\end{subequations}

Ignoring the trivial case of $N=0$, the equilibrium

\begin{subequations}
\begin{equation}
\hat{\bar x}=\theta,
\end{equation}
\begin{equation}
\hat G=\sqrt\frac{\mu}{a},
\end{equation}
\begin{equation}
\hat N=\frac{1}{c}\left(r-\frac{1}{2}(\eta a+\sqrt{\mu a})\right),
\end{equation}
\end{subequations}

is unique and globally stable for $a,c,\mu>0$. Setting $2r>\eta a+\sqrt{\mu a}$ ensures a positive equilibrium abundance and setting $c<r-(\eta a+\sqrt{\mu a})/2$ ensures $\hat N>1$, which is important for numerical simulations when $N$ is an integer. We use these results to help guide our choice of parameter values for simulations of the branching random walk. In the following section we provide a detailed description of the branching random walk and how we have chosen to rescale it. We then use the rescaled branching random walk to investigate finiteness of moments and normality.

### Description of simulation \label{description}

We begin by describing the branching random walk before introducing our scheme to rescale it. Our branching random walk follows closely the description of branching Brownian motion in the main text. However, we replace exponentially distributed lifetimes with deterministic unit time steps for easier implementation. Hence, we restrict time to $t=0,1,2,\dots$. Furthermore, we allow individual fitness to depend on both trait value and the state of the entire population. For time $t$ we write $\{\xi_1(t),\dots,\xi_{N(t)}(t)\}$ as the set of trait values across all $N(t)$ individuals alive in the population. Since our simulation treats individuals instead of continuous distributions of trait values, we can write

\begin{equation}
\nu(x,t)=\sum_{i=1}^{N(t)}\delta(x-\xi_i(t)),
\end{equation}

where $\delta(x-\xi_i(t))$ denotes the point mass located at $\xi_i(t)$. To allow for imperfect heritability, we also track the set of breeding values which, at time $t$, is denoted by $\{\gamma_1(t),\dots,\gamma_{N(t)}(t)\}$ and should not be confused with the quadratic selection gradients discussed in \S\ref{} of the main text. Then the distribution of breeding values can be written as

\begin{equation}
\rho(g,t)=\sum_{i=1}^{N(t)}\delta(g-\gamma_i(t)).
\end{equation}

Following our model of heritiability, the trait value $\xi_i(t)$ is drawn from a normal distribution centered on $\gamma_i(t)$ with variance $\eta$. At each iteration we draw, for each individual, a random number of offspring from a Negative-Binomial distribution. Recall the Negative-Binomial distribution models the number of failed Bernoulli trials that occur before a given number of successful trials. Denoting $q$ the probability of success for each trial and $s$ the number of successes, the mean and variance is given respectively by

\begin{equation}
\frac{s(1-q)}{q}, \ \frac{s(1-q)}{q^2}.
\end{equation}

Then if we require the $i$th individual to have mean number offspring $\mathscr W(\nu,\xi_i)$ and variance equal to $V$, the parameters of the associated Negative-Binomial distribution become

\begin{equation}
q(\nu,\xi_i)=\frac{\mathscr{W}(\nu,\xi_i)}{V}, \ s(\nu,\xi_i)=\frac{\mathscr{W}^2(\nu,\xi_i)}{V-\mathscr{W}(\nu,\xi_i)}.
\end{equation}

The imposes the restriction $V>\mathscr{W}(\nu,\xi_i)$. For each offpsring produced by the individual with breeding value $\gamma_i(t)$, we assign indepently drawn breeding values normally distributed around $\gamma_i(t)$ with variance $\mu$. After breeding values have been assigned, we randomly draw trait values for each offpsring as described above. For an overview of our model of inheritance, see \S\ref{} of the main text. This summarizes the basic structure of our simulation. To impose selection and density dependent growth rates, we set

\begin{equation}
\mathscr{W}(\nu,\xi_i)=\exp\left(r-\frac{a}{2}(\theta-\xi_i)^2-c\int_\mathbb{R}\nu(x,t)dx\right),
\end{equation}

where the above integral becomes $\int_\mathbb{R}\nu(x,t)dx=\sum_{i=1}^{N(t)}1=N(t)$.

__Rescaling__

To rescale the branching random walk by a positive integer $n$, we rescale segregation and mutational variance according to $\eta\to \eta$ and $\mu\to \mu/n$, time by $t\to t/n$ and the reproductive law by $V\to V$ and

\begin{equation}
\mathscr W(\nu,\xi_i)\to\mathscr W^{(n)}(\nu,\xi_i)=\exp\left(\frac{r}{n}-\frac{a}{2n}(\theta-\xi_i)^2-\frac{c}{n^2}N(t)\right)=\exp\left(\frac{r}{n}-\frac{a}{2n}(\theta-\xi_i)^2-\frac{c}{n}N^{(n)}(t)\right).
\end{equation}

We also replace individual mass with $\frac{1}{n}$ and write rescaled abundance as $N^{(n)}(t)=\frac{1}{n}N(nt)$. Under this rescaling the deterministic equilibrium of the raw numerical abundance becomes

\begin{equation}
\hat N=\frac{n^2}{c}\left(\frac{r}{n}-\frac{1}{2n}(\eta a+\sqrt{\mu a})\right)=\frac{n}{c}\left(r-\frac{1}{2}(\eta a+\sqrt{\mu a})\right).
\end{equation}

The deterministic equilibrium of the rescaled abundance is then

\begin{equation}
\hat N^{(n)}=\frac{1}{c}\left(r-\frac{1}{2}(\eta a+\sqrt{\mu a})\right).
\end{equation}

When it exists, we denote by $N^{(\infty)}(t)$ the limiting process of $N^{(n)}(t)$. Then

\begin{equation}
\lim_{n\to\infty}n\left(\mathscr W^{(n)}(\nu,\xi_i)-1\right)=r-\frac{a}{2}(\theta-\xi_i)^2-cN^{(\infty)}(t).
\end{equation}

Note that, following the notation of Theorem 1 in Méléard and Roelly [-@meleard1992interacting], setting $\lambda_n=n$, $m_n(\nu)=\mathscr W^{(n)}(\nu,\cdot)$ and $\varepsilon_n=1/n$ satisfies their hypotheses $(\mathscr H_0)$-$(\mathscr H_3)$ when $c=0$. We have implemented this simulation in the programming language Julia. A copy can be found at the url:

\texttt{\url{https://github.com/bobweek/branching.brownian.motion.and.spde}}

For the sake of illustration, we simulated the unscaled process ($n=1$) and the rescaled process with $n=5$ and $n=20$ for $50$ units of time. Results are shown in Figure \ref{rescaled}. In the following section we use a statistical test to show, for the lower bound on $m(\nu,x)$, the rescaled process converges to a Gaussian density as $n\to\infty$ and $V/N\to0$.

### Evidence of normality \label{ev_normal}

To demonstrate approximate normality of the phenotypic distribution when $V/N$ is small we utilized the one-sided Kolmogorov-Smirnov test. This test compares an empirical cumulative distribution function (i.e., a cumulative distribution function generated from simulated data) to a hypothetical cumulative distribution function by providing a distribution for the maximum distance between these curves. More precisely if $F_n(x)$ is the empricial distribution function for a sample of size $n$ and $F(x)$ is the hypothetical distribution function, Kolmogorov-Smirnov statistic is $D_n=\sup_x|F_n(x)-F(x)|$.

## Derivation of SDE for $\bar x$ and $\sigma^2$ \label{SDE_DERIV}

For $\nu(x,t)$ defined in the main text, $h\in C(\mathbb{R})$ and $t\geq0$ we make the following assumptions:

\begin{equation}\label{L1_nu}
\mathbb{E}\left(\int_\mathbb{R}|h(x)|\nu(x,t)dx\right)<\infty,
\end{equation}
\begin{equation}\label{L2_nu}
\mathbb{E}\left(\int_\mathbb{R}h^2(x)\nu(x,t)dx\right)<\infty,
\end{equation}
\begin{equation}\label{L1_deriv}
\mathbb{E}\left(\int_0^t\int_\mathbb{R}\nu(x,s)\Big|h(x)\Big|\left|m(\nu,x)f(x)+\frac{\mu}{2}\Delta f(x)\right|dxds\right)<\infty, \ \forall f\in C^2_b(\mathbb{R}).
\end{equation}

Put $H(t)=\int_\mathbb{R}h(x)\nu(x,t)dx$. Then, for non-random and non-negative $\nu_1(x,t)$ that is continuous in both arguments and integrable in $x$,
\begin{equation}
\star:=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{E}[H(t+\varepsilon)-H(t)|\nu(x,t)=\nu_1(x,t)]
\end{equation}
\begin{equation}
=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{E}\left[\int_\mathbb{R}h(x)\big(\nu(x,t+\varepsilon)-\nu(x,t)\big)dx\Big|\nu(x,t)=\nu_1(x,t)\right]
\end{equation}
\begin{equation}
=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{E}\left[\int_\mathbb{R}\int_t^{t+\varepsilon}f(x)\left(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)
+\sqrt{V\nu(x,s)}\dot W(x,s)\right)dsdx\Big|\nu(x,t)=\nu_1(x,t)\right].
\end{equation}

By assumption (\ref{L1_deriv}) we can use Fubini's theorem to write, with probability one,

\begin{equation}
\int_\mathbb{R}\int_t^{t+\varepsilon}h(x)\big(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)dsdx=\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\big(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)dxds.
\end{equation}

By assumption (\ref{L2_nu}) we have $\varphi(x,t)=h(x)\sqrt{V\nu(x,t)}$ implies $\varphi\in L^2_c(\mathbb{R}\times[0,\infty))$. Hence, the following is true by definition;

\begin{equation}
\int_\mathbb{R}\int_t^{t+\varepsilon}h(x)\sqrt{V\nu(x,s)}\dot W(x,s)dsdx = 
\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\sqrt{V\nu(x,s)}\dot W(x,s)dxds.
\end{equation}

Hence,

\begin{multline}
\mathbb{E}\left[\int_\mathbb{R}\int_t^{t+\varepsilon}h(x)\sqrt{V\nu(x,s)}\dot W(x,s)dsdx\Big|\nu(x,t)=\nu_1(x,t)\right] \\ =\mathbb{E}\left[\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\sqrt{V\nu(x,s)}\dot W(x,s)dxds\Big|\nu(x,t)=\nu_1(x,t)\right]=0
\end{multline}

and

\begin{equation}
\star=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{E}\left[\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\left(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)\right)dxds\Big|\nu(x,t)=\nu_1(x,t)\right].
\end{equation}

By assumption (\ref{L1_deriv}) we know that there exists a $\delta>0$ such that for each positive $\varepsilon<\delta$ the following holds almost surely:

\begin{multline}
\left|\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\left(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)\right)dxds\right| \\
\leq\int_t^{t+\delta}\int_\mathbb{R}\Big|h(x)\Big|\left|m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)\right|dxds<\infty.
\end{multline}

Thus, by Lebesgue's dominated convergence theorem, the drift component of the process $H(t)$ can be computed as

\begin{multline}
\star=\mathbb{E}\left[\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\left(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)\right)dxds\Big|\nu(x,t)=\nu_1(x,t)\right]\\
=\int_\mathbb{R}h(x)\big(m(\nu_1,x)\nu_1(x,t)+\frac{\mu}{2}\Delta\nu_1(x,t)\big)dx.
\end{multline}

To find an expression for the diffusion component of $H(t)$ set

\begin{multline}\label{star2}
\star\star:=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{V}[H(t+\varepsilon)-H(t)|\nu(x,t)=\nu_1(x,t)] \\
=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{V}\left[\int_\mathbb{R}h(x)\big(\nu(x,t+h)-\nu(x,t)\big)dx\Big|\nu(x,t)=\nu_1(x,t)\right].
\end{multline}

We can rewrite the integral inside expression (\ref{star2}) as

\begin{equation}
\int_\mathbb{R}\int_t^{t+\varepsilon}h(x)\big(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)+\sqrt{V\nu(x,s)}\dot W(x,s)\big)dsdx.
\end{equation}

We have already found

\begin{multline}
\mathbb{E}\left[\int_\mathbb{R}\int_t^{t+\varepsilon}h(x)\big(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)+\sqrt{V\nu(x,s)}\dot W(x,s)\big)dsdx\Big|\nu(x,t)=\nu_1(x,t)\right] \\
=\mathbb{E}\left[\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\big(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)\big)dxds\Big|\nu(x,t)=\nu_1(x,t)\right].
\end{multline}

Then, since $h^2(x)\nu(x,t)$ is integrable by assumption (\ref{L2_nu}), we have

\begin{multline}
\star\star=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{E}\left[\left(\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\sqrt{V\nu(x,s)}\dot W(x,s)dsdx\right)^2\Big| \ \nu(x,t)=\nu_1(x,t)\right] \\
=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{E}\left[\int_t^{t+\varepsilon}\int_\mathbb{R}V h^2(x)\nu(x,s)dxds\Big|\nu(x,t)=\nu_1(x,t)\right].
\end{multline}

Thus, for any $\delta\geq\varepsilon\geq0$, we have, with probability one,

\begin{equation}
\int_t^{t+\varepsilon}\int_\mathbb{R}V h^2(x)\nu(x,s)dxds\leq\int_t^{t+\delta}\int_\mathbb{R}V h^2(x)\nu(x,s)dxds.
\end{equation}

We can therefore use Lebesgue's dominated convergence theorem to justify

\begin{equation}
\star\star=\mathbb{E}\left[\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\int_t^{t+\varepsilon}\int_\mathbb{R}V h^2(x)\nu(x,s)dxds\Big|\nu(x,t)=\nu_1(x,t)\right]=\int_\mathbb{R}V h^2(x)\nu_1(x,t)dx.
\end{equation}

Then, using the notation of stochastic differentials, we have

\begin{equation}\label{SDE_deriv}
dH(t)=\left(\int_\mathbb{R}h(x)\left(m(\nu,x)\nu(x,t)+\frac{\mu}{2}\Delta\nu(x,t)\right)dx\right)dt+\sqrt{V\int_\mathbb{R}h^2(x)\nu(x,t)dx} \ dW(t)
\end{equation}

where $W$ is a standard Brownian motion. In the following subsections we employ this formula under the cases $h(x)=x,x^2$ to obtain SDE for the phenotypic mean and variance.

### Derivation for trait mean

We set $\tilde x(t)=\int_\mathbb{R}x\nu(x,t)$ and make use of the notation

\begin{equation}
\begin{matrix}
\|N\|_2 = \sqrt{V\int_\mathbb{R}\nu(x,t)dx}= \sqrt{VN} \\ \\
\|\tilde x\|_2 = \sqrt{V\int_\mathbb{R}x^2\nu(x,t)dx} \\ \\
\langle\tilde x,N\rangle = V\int_\mathbb{R}x\nu(x,t)dx=\bar x VN.
\end{matrix}
\end{equation}

Applying formula (\ref{SDE_deriv}) provides

\begin{equation}
d\tilde x=\left(\overline{xm}N+\frac{\mu}{2}\int_\mathbb{R}x\Delta\nu(x,t)dx\right)dt+\|\tilde x\|_2d\tilde W_2,
\end{equation}

where,

\begin{equation}
d\tilde{ W}_2=d\hat{\mathbf W}_{\sqrt{V x^2\nu}}=\frac{1}{\|\tilde x\|_2}\int_\mathbb{R}x\sqrt{V\nu(x,t)}\dot W(x,t)dxdt.
\end{equation}

Using It&ocirc;'s quotient rule on $\bar x=\tilde x/N$, we obtain

\begin{equation}
d\bar x = d\left(\frac{\tilde x}{N}\right)=\frac{\tilde x}{N}\left(\frac{d\tilde x}{\tilde x}-\frac{dN}{N}-\frac{d\tilde x}{\tilde x}\frac{dN}{N}+\Big(\frac{dN}{N}\Big)^2\right)
= \frac{d\tilde x}{N}-\bar x\frac{dN}{N}-\frac{d\tilde x}{N}\frac{dN}{N}+\bar x\Big(\frac{dN}{N}\Big)^2.
\end{equation}

From Table \ref{mult} of the main text $d\tilde xdN=\langle\tilde x,N\rangle$ and $dN^2=\|N\|_2^2$. Hence,

\begin{equation}
d\bar x= \overline{xm}dt+\frac{\mu}{2}\int_\mathbb{R}x\Delta p(x,t)dxdt+\frac{\|\tilde x\|_2}{N}d\tilde W_2-\bar x\left(\bar m dt+\sqrt\frac{V}{N}dW_1\right)-\frac{\langle\tilde x,N\rangle}{N^2}dt+\bar x\frac{\|N\|_2^2}{N^2}dt
\end{equation}

\begin{equation*}
= (\overline{xm}-\bar x\bar m)dt+\frac{\mu}{2}\int_\mathbb{R}x\Delta p(x,t)dxdt+\frac{\|\tilde x\|_2}{N}d\tilde W_2
-\bar x\sqrt{\frac{V}{N}}d W_1-V\frac{\bar x}{N}dt+V\frac{\bar x}{N}dt
\end{equation*}

\begin{equation*}
= \left(\mathrm{Cov}_t(x,m)+\frac{\mu}{2}\int_\mathbb{R}x\Delta p(x,t)dx\right)dt
+\frac{\|\tilde x\|_2}{N}d\tilde W_2-\bar x\sqrt{\frac{V}{N}}d W_1.
\end{equation*}

Note that

\begin{equation}
\frac{\|\tilde x\|_2}{N}d\tilde W_2-\bar x\sqrt{\frac{V}{N}}d W_1=\frac{1}{N}\int_\mathbb{R}x\sqrt{V\nu(x,t)}\dot W(x,t)dx-\frac{\bar x}{N}\int_\mathbb{R}\sqrt{V\nu(x,t)}\dot W(x,t)dx
\end{equation}

\begin{equation*}
=\int_\mathbb{R}\frac{x-\bar x}{N}\sqrt{V\nu(x,t)}\dot W(x,t)dx
\end{equation*}

and

\begin{equation}
\mathbb{V}\left(\int_\mathbb{R}\frac{x-\bar x}{N}\sqrt{V\nu(x,t)}\dot W(x,t)dx\right)=\frac{V}{N}\int_\mathbb{R}(x-\bar x)^2p(x,t)dx=V\frac{\sigma^2}{N}.
\end{equation}

Hence, by setting

\begin{equation}
d W_2=\frac{\int_\mathbb{R}\frac{(x-\bar x)}{N}\sqrt{V\nu(x,t)}\dot W(x,t)dx}{\sqrt{V\sigma^2/N}}
\end{equation}

we can write

\begin{equation}
d\bar x=\left(\mathrm{Cov}_t(x,m)+\frac{\mu}{2}\int_\mathbb{R}x\Delta p(x,t)dx\right)dt
+\sqrt{V\frac{\sigma^2}{N}}d W_2.
\end{equation}

### Derivation for trait variance

We set $\tilde{\tilde\sigma}^2(t)=\int_{\mathbb R} x^2\nu(x,t)dx$ and make use of the notation

\begin{equation}
\begin{matrix}
\|\tilde{\tilde\sigma}^2\|_2 = \sqrt{V\int_\mathbb{R}x^4\nu(x,t)dx} \\ \\
\langle\tilde{\tilde\sigma}^2,N\rangle=V\int_\mathbb{R}x^2\nu(x,t)dx=\overline{x^2}VN.
\end{matrix}
\end{equation}

Applying formula (\ref{SDE_deriv}) provides

\begin{equation}
d\tilde{\tilde\sigma}^2=\left(\overline{x^2m}N+\frac{\mu}{2}\int_\mathbb{R}x^2\Delta\nu(x,t) dx\right)dt+\|\tilde{\tilde\sigma}^2\|_2d\tilde W_3
\end{equation}

where

\begin{equation}
d\tilde{W}_3=d\hat{\mathbf W}_{\sqrt{V x^4\nu}}
=\frac{1}{\|\tilde{\tilde\sigma}^2\|_2}\int_\mathbb{R}x^2\sqrt{V\nu(x,t)}\dot W(x,t)dx.
\end{equation}

Using It&ocirc;'s quotient rule on $\overline{x^2}=\tilde{\tilde\sigma}^2/N$, we obtain

\begin{equation}
d\overline{x^2} = d\left(\frac{\tilde{\tilde\sigma}^2}{N}\right)=\frac{\tilde{\tilde\sigma}^2}{N}\left(\frac{d\tilde{\tilde\sigma}^2}{\tilde{\tilde\sigma}^2}-\frac{dN}{N}-\frac{d\tilde{\tilde\sigma}^2}{\tilde{\tilde\sigma}^2}\frac{dN}{N}+\Big(\frac{dN}{N}\Big)^2\right)
=\frac{d\tilde{\tilde\sigma}^2}{N}-\overline{x^2}\frac{dN}{N}-\frac{d\tilde{\tilde\sigma}^2}{N}\frac{dN}{N}+\overline{x^2}\left(\frac{dN}{N}\right)^2.
\end{equation}

Table \ref{mult} of the main text implies $d\tilde W_3d W_1=\langle\tilde{\tilde\sigma}^2,N\rangle$ and hence

\begin{multline}
d\overline{x^2}=\left(\overline{x^2m} + \frac{\mu}{2}\int_\mathbb{R}x^2\Delta p(x,t)dx\right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3
-\overline{x^2}\left(\bar m dt+\sqrt\frac{V}{N}dW_1\right)-\frac{\langle\tilde{\tilde\sigma}^2,N\rangle}{N^2}dt+\overline{x^2}\frac{\|N\|_2^2}{N^2}dt \\ \\
=\left(\overline{x^2m}-\overline{x^2}\bar mdt + \frac{\mu}{2}\int_\mathbb{R}x^2\Delta p(x,t)dx\right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3
-\overline{x^2}\sqrt{\frac{V}{N}}d W_1
-\overline{x^2}\frac{V}{N}dt+\overline{x^2}\frac{V}{N}dt \\ \\
=\left(\mathrm{Cov}_t\Big(x^2,m\Big) + \frac{\mu}{2}\int_\mathbb{R}x^2\Delta pdx \right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3-\overline{x^2}\sqrt\frac{V}{N}d W_1.
\end{multline}

Setting $F(y,z)=y-z^2$, use It&ocirc;'s formula on $\sigma^2=F(\overline{x^2},\bar x)=\overline{x^2}-\bar x^2$ to obtain:

\begin{multline}
d\sigma^2=d\overline{x^2}-2\bar xd\bar x-(d\bar x)^2
=\left(\mathrm{Cov}_t\Big(x^2,m\Big) + \frac{\mu}{2}\int_\mathbb{R}x^2\Delta p(x,t)dx \right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3-\overline{x^2}\sqrt\frac{V}{N}d W_1 \\ \\
-2\bar x\left(\mathrm{Cov}_t(x,m)dt+\frac{\mu}{2}\int_\mathbb{R}x\Delta p(x,t)dxdt+\sqrt{\frac{V\sigma^2}{N}}d W_2\right)-\left(\mathrm{Cov}_t(x,m)dt+\frac{\mu}{2}\int_\mathbb{R}x\Delta p(x,t)dxdt+\sqrt{\frac{V\sigma^2}{N}}d W_2\right)^2 \\ \\
=\left(\mathrm{Cov}_t\Big(x^2-2\bar xx,m\Big) + \frac{\mu}{2}\int_\mathbb{R}(x^2-x\bar x)\Delta p(x,t)dx \right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3-\overline{x^2}\sqrt\frac{V}{N}d W_1-2\bar x\sqrt{\frac{V\sigma^2}{N}}d W_2-\left(\frac{V\sigma^2}{N}\right)dt \\ \\
=\left(\mathrm{Cov}_t\Big(x-\bar x)^2,m\Big) + \frac{\mu}{2}\int_\mathbb{R}(x-\bar x)^2\Delta p(x,t)dx - \frac{V\sigma^2}{N}\right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3-\overline{x^2}\sqrt\frac{V}{N}d W_1-2\bar x\sqrt{\frac{V\sigma^2}{N}}d W_2.
\end{multline}

In light of

\begin{multline}
\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3-\overline{x^2}\sqrt\frac{V}{N}d W_1-2\bar x\sqrt{\frac{V\sigma^2}{N}}d W_2=\frac{1}{N}\int_\mathbb{R}\left(x^2-\tilde\sigma^2-2\bar x(x-\bar x)\right)\sqrt{V\nu(x,t)}\dot W(x,t)dx \\
=\frac{1}{N}\int_\mathbb{R}\left((x-\bar x)^2-\sigma^2\right)\sqrt{V\nu(x,t)}\dot W(x,t)dx
\end{multline}

and

\begin{multline}
\frac{1}{N}\int_\mathbb{R}\Big(\left((x-\bar x)^2-\sigma^2\right)\sqrt{V\nu(x,s)}\Big)^2dx
=\frac{V}{N}\left(\int_\mathbb{R}((x-\bar x)^4-2(x-\bar x)^2\sigma^2+\sigma^4)p(x,t)dx\right) \\
=\frac{V}{N}\left(\overline{(x-\bar x)^4}-\sigma^4\right)
\end{multline}

we set

\begin{equation}
d W_3=\frac{\int_\mathbb{R}\left((x-\bar x)^2-\sigma^2\right)\sqrt{V\nu(x,t)}\dot W(x,t)dx}{V\left(\overline{(x-\bar x)^4}-\sigma^4\right)}
\end{equation}

so that

\begin{equation}
d\sigma^2=\left(\mathrm{Cov}_t\Big((x-\bar x)^2,m\Big)+\frac{\mu}{2}\int_\mathbb{R}(x-\bar x)^2\Delta p(x,t)dx-V\frac{\sigma^2}{N}\right)dt+\sqrt{V\frac{\overline{(x-\bar x)^4}-\sigma^4}{N}}d W_3.
\end{equation}

Table \ref{mult} of the main text implies

\begin{equation}
d W_1d W_2=\frac{\int_\mathbb{R}(x-\bar x)\nu(x,t)dx}{\sqrt{N\sigma^2}}dt=0,
\end{equation}

\begin{equation}
d W_1d W_3=\frac{\int_\mathbb{R}\big((x-\bar x)^2-\sigma^2\big)\nu(x,t)dx}{\sqrt{\overline{(x-\bar x)^4}-\sigma^4}}dt=0,
\end{equation}

\begin{equation}
d W_2d W_3=\frac{\int_\mathbb{R}(x-\bar x)\big((x-\bar x)^2-\sigma^2\big)p(x,t)dx}{\sqrt{\sigma^2\big(\overline{(x-\bar x)^4}-\sigma^4\big)}}dt
=\frac{N\overline{(x-\bar x)^3}}{\sqrt{\sigma^2\big(\overline{(x-\bar x)^4}-\sigma^4\big)}}dt.
\end{equation}

In particular, when $p$ is a Gaussian curve $d W_2d W_3=0$.

## Relating fitness of expressed traits to fitness of breeding values

$$m^*(\rho,g)=\int_\mathbb{R}m(\nu,x)\psi(x,g)dx$$

$$\overline{\frac{\partial m^*}{\partial\bar x}}=\int_{\mathbb{R}}\frac{\rho(g,t)}{N(t)}\frac{\partial}{\partial\bar x}\int_\mathbb{R}m(\nu,x)\psi(x,g)dxdg=\int_{\mathbb{R}}\int_\mathbb{R}\frac{\rho(g,t)}{N(t)}\psi(x,g)dg\frac{\partial}{\partial\bar x}m(\nu,x)dx=\int_\mathbb{R}p(x,t)\frac{\partial}{\partial\bar x}m(\nu,x)dx=\overline{\frac{\partial m}{\partial\bar x}}$$

$$\overline{\frac{\partial m^*}{\partial G}}=\int_{\mathbb{R}}\frac{\rho(g,t)}{N(t)}\frac{\partial}{\partial G}\int_\mathbb{R}m(\nu,x)\psi(x,g)dxdg=\int_{\mathbb{R}}\int_\mathbb{R}\frac{\rho(g,t)}{N(t)}\psi(x,g)dg\frac{\partial m}{\partial G}dx=\int_\mathbb{R}p(x,t)\frac{\partial m}{\partial\sigma^2}\frac{\partial\sigma^2}{\partial G}dx=\overline{\frac{\partial m}{\partial\sigma^2}}$$

## Derivation of diffuse coevolution model \label{diffuse}

In this section we provide a derivation of our model of diffuse coevolution driven by competition. Since most of the work in this derivation has already been completed in Supplementary Material \S\ref{SDE_DERIV}, we focus here on deriving the Malthusian fitness $m$ as a function of trait value $x$. We begin with discrete populations of individuals. In particular, we begin by assuming population size $n_i$ is an integer for each species $i=1,\dots,S$ before passing to the large population size limit.

The reduction in fitness for an individual of species $i$ caused by competition is captured multiplicatively by $0<C_i\leq1$. Biologically this assumes all competitors affect individuals of a given species equally by consuming the same amount of resources. This is a mean-field interaction since any individual that consumes resources has an effect on the fitness of all other individuals competing for the same resources. Denote by $x_{ij}$ the trait value of the $j$-th individual belonging to species $i$. The set of trait values across all individuals in the community at time $t\geq0$ is written $X=\{x_{ij}\}$. We denote by $\mathscr{B}_{ij}$ a function that maps $X$ to the cumulative effect of all competitive interactions on the fitness of the $j$-th individual in species $i$. Since individuals do not compete with themselves the net multiplicative effects on fitness of both interspecific and intraspecific competition on the $j$-th individual in species $i$ can be summarized by

\begin{equation}
\mathscr B_{ij}(X)=C_i^{\sum_{l\neq j}\mathcal{O}_{ii}(x_{ij},x_{il})+\sum_{k\neq i}\sum_{l=1}^{n_k}\mathcal{O}_{ik}(x_{ij},x_{kl})},
\end{equation}

where $\mathcal{O}_{ij}$, defined in the main text, measures the overlap in resource use between individuals of species $i$ and $j$ as a function of their niche-centers. Writing $\mathscr W_{ij}(X)$ as the average number of offspring left by the $j$-th individual of species $i$, we have

\begin{equation}
\mathscr W_{ij}(X)=\mathscr{A}_i(x_{ij})\mathscr{B}_{ij}(X),
\end{equation}

where $\mathscr{A}_i(x)=\int_\mathbb{R}e_i(\zeta)u_i(\zeta,x)d\zeta$ accounts for abiotic selection and $e_i$ has been defined in the main text.

We now turn to a diffusion limit. Since we have more than one population, we take the diffusion limit for each population one at a time starting with population $1$. We write $\pmb n=(n_1,\dots,n_S)$. Following Méléard and Roelly [-@Mlard1993; -@meleard1992interacting] we rescale generation time and individual mass to $\frac{1}{n_1}$ and mean of the reproductive law to

\begin{equation}
\mathscr W_{1j}^{(\pmb n)}(X)=\mathscr{A}_1(x_{1j})^{1/n_1}\exp\left(\frac{\ln C_1}{n_1^2}\sum_{l\neq j}\mathcal{O}_{11}(x_{1j},x_{1l})+\frac{\ln C_1}{n_1}\sum_{k\neq i}\frac{1}{n_k}\sum_{l=1}^{n_k}\mathcal{O}_{1k}(x_{1j},x_{kl})\right).
\end{equation}

For large $n_1$, we have the approximation

\begin{equation}
\mathscr W_{1j}^{(\pmb n)}(X)\approx\mathscr{A}_1(x_{1j})^{1/n_1}\left(1+\frac{\ln C_1}{n_1^2}\sum_{l\neq j}\mathcal{O}_{11}(x_{1j},x_{1l})+\frac{\ln C_1}{n_1}\sum_{k\neq 1}\frac{1}{n_k}\sum_{l=1}^{n_k}\mathcal{O}_{1k}(x_{1j},x_{kl})\right).
\end{equation}
Hence
\begin{equation}
\lim_{n_1\to\infty}n_1\left(\mathscr W_{1j}^{(\pmb n)}(X)-1\right)=\ln\mathscr{A}_1(x_{1j})+\left(\int_\mathbb{R}\mathcal{O}_{11}(x_{1j},y)\nu_1(y,t)dy+\sum_{k\neq 1}\frac{1}{n_k}\sum_{l=1}^{n_k}\mathcal{O}_{1k}(x_{1j},x_{kl})\right)\ln C_1.
\end{equation}

We write $\lim_{\pmb n\to\pmb\infty}$ for the iterated limit $\lim_{n_S\to\infty}\cdots\lim_{n_1\to\infty}$ and, assuming  $\nu_i(\cdot,t)\in C_1^+(\mathbb{R})$ for $i=1,\dots,S$ and $t\in[0,\infty)$, we set $\pmb\nu=(\nu_1,\dots,\nu_S)$. Then, for any $\pmb\nu$, the the diffusion limits for the remaining populations provides the Malthusian parameter for individuals in species $i$ with trait value $x_{1j}$ as

\begin{equation}
m_1(\pmb\nu,x_{1j}):=\lim_{\pmb n\to\pmb\infty}n_1\left(\mathscr W_{1j}^{(\pmb n)}(X)-1\right)=\ln\mathscr{A}_1(x)+\left(\sum_{k=1}^S\int_\mathbb{R}\mathcal{O}_{1k}(x_{1j},y)\nu_k(y,t)dy\right)\ln C_1.
\end{equation}

We compute the average niche overlap of an individual in species $i$ with nich location $x$ across all individuals in species $j$ as

\begin{equation}
\bar{\mathcal{O}}_{ij}(x,t)=\frac{\int_\mathbb{R}\mathcal{O}_{ij}(x,y)\nu_j(y,t)dy}{\int_\mathbb{R}\nu_j(y,t)dy}.
\end{equation}

We now assume the resource utilization curves $u_i(\zeta)$ and phenotypic densities $\nu_i(x,t)$ are Gaussian curves for $i=1,\dots,S$. In this case $\bar{\mathcal{O}}_{ij}(x,t)$ simplifies to

\begin{equation}
\bar{\mathcal{O}}_{ij}(x,t)=\frac{\int_\mathbb{R}\mathcal O_{ij}(x,y)\nu_j(y,t)dy}{\int_\mathbb{R}\nu_j(y,t)dy}
=\frac{U_iU_j}{\sqrt{2\pi(w_i+w_j+{\sigma_j}^2(t))}}\exp\left(-\frac{(x-\bar x_j(t))^2}{2(w_i+w_j+{\sigma_j}^2(t))}\right).
\end{equation}

Setting

\begin{subequations}
\begin{align}
\sigma_i^2(t) = & \ G_i(t)+\eta_i, \\
R_i = & \ \ln \left(\frac{Q_iU_i}{\sqrt{1+A_iw_i}}\right), \\
a_i = & \ \frac{A_i}{1+A_iw_i}, \\
\tilde b_{ij}(t) = & \ \frac{1}{w_i+w_j+\sigma_j^2(t)}, \\
c_i = & -\ln C_i,
\end{align}
\end{subequations}

we get

\begin{equation}\label{malth_fit}
m_i(\pmb\nu,x)=R_i-\frac{a_i}{2}(x-\theta_i)^2-c_i\sum_{j=1}^SN_j(t)U_iU_j\sqrt{\frac{\tilde b_{ij}(t)}{2\pi}}e^{-\frac{\tilde b_{ij}(t)}{2}(x-\bar x_j(t))^2}.
\end{equation}

Hence, our fitness function satisfies condition (\ref{m_cond}) of the main text.

For the remainder of the derivation we suppress notation indicating dependency on $\pmb\nu$, $x$ and $t$. From (\ref{malth_fit}) we calculate

\begin{equation}
\frac{\partial m_i}{\partial\bar x_i}=c_iN_iU_i^2\tilde b_{ii}(x-\bar x_i)\sqrt{\frac{\tilde b_{ii}}{2\pi}}e^{-\frac{\tilde b_{ii}}{2}(x-\bar x_i)^2}
\end{equation}

\begin{multline}
\frac{\partial m_i}{\partial G_i}=\frac{c_iN_iU_i^2}{2}\left(\frac{(x-\bar x_i)^2-G_i-\eta_i-2w_i}{(G_i+\eta_i+2w_i)^2}\right)\sqrt{\frac{\tilde b_{ii}}{2\pi}}e^{-\frac{\tilde b_{ii}}{2}(x-\bar x_i)^2} \\
=\frac{c_iN_iU_i^2{\tilde b_{ii}}^2}{2}\left((x-\bar x_i)^2-{\sigma_i}^2-2w_i\right)\sqrt{\frac{\tilde b_{ii}}{2\pi}}e^{-\frac{\tilde b_{ii}}{2}(x-\bar x_i)^2}.
\end{multline}

Note that

\begin{multline}
\sqrt\frac{\tilde b_{ii}}{2\pi}\exp\left(-\frac{\tilde b_{ii}}{2}(x-\bar x_i)^2\right)\sqrt\frac{1}{2\pi{\sigma_i}^2}\exp\left(-\frac{(x-\bar x_i)^2}{2{\sigma_i}^2}\right) \\
=\sqrt\frac{1}{2\pi({\sigma_i}^2+1/\tilde b_{ii})}\sqrt\frac{{\sigma_i}^2+1/\tilde b_{ii}}{2\pi{\sigma_i}^2/\tilde b_{ii}}\exp\left(-\frac{{\sigma_i}^2+1/\tilde b_{ii}}{2{\sigma_i}^2/\tilde b_{ii}}(x-\bar x_i)^2\right) \\
=\sqrt{\frac{1}{4\pi({\sigma_i}^2+w_i)}}\sqrt\frac{2({\sigma_i}^2+w_i)}{2\pi{\sigma_i}^2({\sigma_i}^2+2w_i)}\exp\left(-\frac{{\sigma_i}^2({\sigma_i}^2+2w_i)}{4({\sigma_i}^2+w_i)}(x-\bar x_i)^2\right).
\end{multline}

Hence,

\begin{equation}
\overline{\frac{\partial m_i}{\partial\bar x_i}}=0,
\end{equation}

\begin{multline}
\overline{\frac{\partial m_i}{\partial G_i}}=\frac{c_iN_iU_i^2}{2({\sigma_i}^2+2w_i)^2}\left(\frac{({\sigma_i}^2+2w_i){\sigma_i}^2}{2(w_i+{\sigma_i}^2)}-{\sigma_i}^2-2w_i\right)\sqrt{\frac{b_{ii}}{2\pi}} \\
=\frac{c_iN_iU_i^2}{2({\sigma_i}^2+2w_i)}\left(\frac{{\sigma_i}^2}{2({\sigma_i}^2+w_i)}-1\right)\sqrt{\frac{b_{ii}}{2\pi}}=-\frac{c_iN_iU_i^2b_{ii}}{2}\sqrt{\frac{b_{ii}}{2\pi}},
\end{multline}

where

\begin{equation}
b_{ij}=\frac{1}{w_i+w_j+{\sigma_i}^2+{\sigma_j}^2}.
\end{equation}

The average fitness for species $i$ is

\begin{equation}
\bar m_i=R_i-\frac{a_i}{2}\Big((\bar x_i-\theta_i)^2+G_i+\eta_i\Big) - c_i\sum_{j=1}^SN_jU_iU_j\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}.
\end{equation}

Thus,

\begin{equation}
\frac{\partial\bar m_i}{\partial\bar x_i}=a_i(\theta_i-\bar x_i)-c_i\sum_jN_jU_iU_jb_{ij}(\bar x_j-\bar x_i)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2},
\end{equation}

\begin{equation}
\frac{\partial\bar m_i}{\partial G_i}=-\frac{a_i}{2}+\frac{c_i}{2}\sum_{j=1}^SN_jU_iU_jb_{ij}\left(1-b_{ij}(\bar x_i-\bar x_j)^2\right)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}.
\end{equation}

In particular

\begin{equation}
\frac{\partial\bar m_i}{\partial G_i}-\overline{\frac{\partial m_i}{\partial G_i}}=-\frac{a_i}{2}+\frac{c_i}{2}\left(N_iU_i^2b_{ii}\sqrt\frac{b_{ii}}{2\pi}+\sum_{j=1}^SN_jU_iU_jb_{ij}\left(1-b_{ij}(\bar x_i-\bar x_j)^2\right)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\right).
\end{equation}

Applying equations (\ref{N}), (\ref{xbarfinal}) and (\ref{Gfinal}) of the main text recovers system (\ref{comm_dynamics}) of the main text.

## The relation between competition coefficients and selection \label{coef_grad_moms}

### Derivation of analytical approximations

Just as with most caclulations in this work, the derivations are straightforward applications of Gaussian products. That is, if

\begin{equation}
f_1(x)=\frac{1}{\sqrt{2\pi\sigma_1^2}}\exp\left(-\frac{(\mu_1-x)^2}{2\sigma_1^2}\right), \ f_2(x)=\frac{1}{\sqrt{2\pi\sigma_2^2}}\exp\left(-\frac{(\mu_2-x)^2}{2\sigma_2^2}\right),
\end{equation}

then

\begin{equation}
f_1(x)f_2(x)=\frac{1}{\sqrt{2\pi(\sigma_1^2+\sigma_2^2)}}\exp\left(-\frac{(\mu_1-\mu_2)^2}{2(\sigma_1^2+\sigma_2^2)}\right)\frac{1}{\sqrt{2\pi\tilde\sigma^2}}\exp\left(-\frac{(\tilde\mu-x)^2}{2\tilde\sigma^2}\right),
\end{equation}

where

\begin{equation}
\tilde\mu=\frac{\sigma_2^2\mu_1+\sigma_1^2\mu_2}{\sigma_1^2+\sigma_2^2}, \ \tilde\sigma^2=\frac{\sigma_1^2\sigma_2^2}{\sigma_1^2+\sigma_2^2}.
\end{equation}

#### Caclulating $\mathrm{Cov}_{f_{\bar X}}(\alpha,\gamma)$

Recalling

\begin{equation}
\alpha(\bar x_i,\bar x_j)=\frac{c}{\bar r}\sqrt\frac{b}{2\pi}\exp\left(-\frac{b}{2}\big(\bar x_i-\bar x_j\big)^2\right),
\end{equation}

\begin{equation}
\gamma(\bar x_i,\bar x_j)=cNb\left(1-b\big(\bar x_i-\bar x_j\big)^2\right) \sqrt\frac{b}{2\pi}\exp\left(-\frac{b}{2}\big(\bar x_i-\bar x_j\big)^2\right),
\end{equation}

we have

\begin{multline}
\bar\alpha=
\int_\mathbb{R}\int_\mathbb{R}\alpha(\bar x_i,\bar x_j)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
=\frac{c}{\bar r}\int_\mathbb{R}\frac{1}{\sqrt{2\pi(b^{-1}+V_{\bar X})}}\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(b^{-1}+V_{\bar X})}\right)f_{\bar X}(\bar x_j)d\bar x_j
=\frac{c/\bar r}{\sqrt{2\pi(b^{-1}+2V_{\bar X})}},
\end{multline}

\begin{multline}
\bar\gamma=
\int_\mathbb{R}\int_\mathbb{R}\gamma(\bar x_i,\bar x_j)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
= cNb\int_\mathbb{R}\left\{1-\left[\left(\frac{\bar{\bar x}+bV_{\bar X}\bar x_j}{1+bV_{\bar X}}-\bar x_j\right)^2+\frac{V_{\bar X}}{1+bV_{\bar X}}\right]\right\}\frac{1}{\sqrt{2\pi(b^{-1}+V_{\bar X})}}\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(b^{-1}+V_{\bar X})}\right)f_{\bar X}(\bar x_j)d\bar x_j \\
= cNb\int_\mathbb{R}\left\{1-\left[\left(\frac{\bar{\bar x}-\bar x_j}{1+bV_{\bar X}}\right)^2+\frac{V_{\bar X}}{1+bV_{\bar X}}\right]\right\}\frac{1}{\sqrt{2\pi(b^{-1}+V_{\bar X})}}\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(b^{-1}+V_{\bar X})}\right)f_{\bar X}(\bar x_j)d\bar x_j \\
= cNb\left(1-\frac{(1+bV_{\bar X})V_{\bar X}}{1+2bV_{\bar X}}\frac{1}{(1+bV_{\bar X})^2}-\frac{V_{\bar X}}{1+bV_{\bar X}}\right)\frac{1}{\sqrt{2\pi(b^{-1}+2V_{\bar X})}} \\
= cNb\left[1-\left(\frac{1}{1+2bV_{\bar X}}+1\right)\frac{V_{\bar X}}{1+bV_{\bar X}}\right]\frac{1}{\sqrt{2\pi(b^{-1}+2V_{\bar X})}} \\
= cNb\left(1-\frac{2V_{\bar X}}{1+2bV_{\bar X}}\right)\sqrt\frac{b}{2\pi(1+2bV_{\bar X})},
\end{multline}

\begin{multline}
\mathrm{Var}_{f_{\bar X}}(\alpha)=
\int_\mathbb{R}\int_\mathbb{R}\big(\bar\alpha-\alpha(\bar x_i,\bar x_j)\big)^2f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
=\frac{c^2}{\bar r^2}\left(\sqrt\frac{b}{4\pi}\int_\mathbb{R}\int_\mathbb{R}\sqrt\frac{b}{\pi}\exp\left(-b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j-\frac{1}{2\pi(b^{-1}+2V_{\bar X})}\right) \\
=\frac{c^2}{\bar r^2}\left(\sqrt\frac{b}{4\pi}\int_\mathbb{R}\sqrt\frac{1}{2\pi(\frac{1}{2b}+V_{\bar X})}\exp\left(-b(\bar{\bar x}-\bar x_j)^2\right)f_{\bar X}(\bar x_j)d\bar x_j-\frac{1}{2\pi(b^{-1}+2V_{\bar X})}\right) \\
=\frac{c^2}{\bar r^2}\left(\sqrt\frac{b}{4\pi}\sqrt\frac{1}{2\pi(\frac{1}{2b}+2V_{\bar X})}-\frac{1}{2\pi(b^{-1}+2V_{\bar X})}\right)=\frac{c^2b}{2\pi\bar r^2}\left(\frac{1}{\sqrt{1+4bV_{\bar X}}}-\frac{1}{1+2bV_{\bar X}}\right),
\end{multline}

\begin{multline}
\mathrm{Cov}_{f_{\bar X}}(\alpha,\gamma)=
\int_\mathbb{R}\int_\mathbb{R}\big(\bar\alpha-\alpha(\bar x_i,\bar x_j)\big)\big(\bar\gamma-\gamma(\bar x_i,\bar x_j)\big)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
=\frac{c^2Nb}{2\bar r}\sqrt\frac{b}{\pi}\int_\mathbb{R}\int_\mathbb{R}\big(1-b(\bar x_i-\bar x_j)^2\big)\sqrt\frac{b}{\pi}\exp\left(-b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j-\bar\alpha\bar\gamma \\
=\frac{c^2Nb}{2\bar r}\sqrt\frac{b}{\pi}\frac{1-2bV_{\bar X}}{\sqrt{2\pi((2b)^{-1}+2V_{\bar X})}}-\frac{c^2Nb}{\bar r}\frac{1-2bV_{\bar X}}{2\pi(b^{-1}+2V_{\bar X})} \\
=\frac{c^2b^2N}{2\pi\bar r}(1-2bV_{\bar X})\left(\frac{1}{\sqrt{1+4bV_{\bar X}}}-\frac{1}{1+2bV_{\bar X}}\right).
\end{multline}

#### Caclulating $\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)$

To calculate moments of $|\beta|$ we note that, as a random variable, $|\beta|$ takes a folded normal distribution. Setting $\Phi(x)$ equal to the cumulative density function of the standard normal distribution and using the properties of the folded normal distribution, we find

\begin{equation}
\overline{|\beta|}=\sqrt\frac{2\mathrm{Var}_{f_{\bar X}}(\beta)}{\pi}\exp\left(-\frac{\bar\beta^2}{2\mathrm{Var}_{f_{\bar X}}(\beta)}\right)-\bar\beta\left[1-2\Phi\left(\frac{\bar\beta}{\mathrm{Var}_{f_{\bar X}}(\beta)}\right)\right]
\end{equation}

\begin{equation}
\mathrm{Var}_{f_{\bar X}}(|\beta|)=\bar\beta^2+\mathrm{Var}_{f_{\bar X}}(\beta)-\overline{|\beta|}^2.
\end{equation}

Recall that

\begin{equation}
\beta(\bar x_i,\bar x_j)=cNb\big(\bar x_i-\bar x_j\big)\sqrt\frac{b}{2\pi}\exp\left(-\frac{b}{2}\big(\bar x_i-\bar x_j\big)^2\right)
\end{equation}

and hence

\begin{multline}
\bar\beta=
\int_\mathbb{R}\int_\mathbb{R}\beta(\bar x_i,\bar x_j)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
= cNb\int_\mathbb{R}(\bar{\bar x}-\bar x_j)\frac{1}{\sqrt{2\pi(b^{-1}+V_{\bar X})}}\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(b^{-1}+V_{\bar X})}\right)f_{\bar X}(\bar x_j)d\bar x_j=0,
\end{multline}

\begin{multline}
\mathrm{Var}_{f_{\bar X}}(\beta)=
\int_\mathbb{R}\int_\mathbb{R}\big(\bar\beta-\beta(\bar x_i,\bar x_j)\big)^2f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
=\int_\mathbb{R}\int_\mathbb{R}c^2N^2b^2(\bar x_i-\bar x_j)^2\frac{b}{2\pi}\exp\left(-b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
=\sqrt{\frac{b}{4\pi}}c^2N^2b^2\int_\mathbb{R}\left[\left(\frac{\bar{\bar x}+2bV_{\bar X}\bar x_j}{1+2bV_{\bar X}}-\bar x_j\right)^2+\frac{V_{\bar X}}{1+2bV_{\bar X}}\right]\frac{\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(\frac{1}{2b}+V_{\bar X})}\right)}{\sqrt{2\pi(\frac{1}{2b}+V_{\bar X})}}f_{\bar X}(\bar x_j)d\bar x_j \\
=\sqrt{\frac{b}{4\pi}}c^2N^2b^2\int_\mathbb{R}\left[\frac{(\bar{\bar x}-\bar x_j)^2}{(1+2bV_{\bar X})^2}+\frac{V_{\bar X}}{1+2bV_{\bar X}}\right]\frac{\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(\frac{1}{2b}+V_{\bar X})}\right)}{\sqrt{2\pi(\frac{1}{2b}+V_{\bar X})}}f_{\bar X}(\bar x_j)d\bar x_j \\
=\sqrt{\frac{b}{4\pi}}c^2N^2b^2\left[\frac{(1+2bV_{\bar X})V_{\bar X}}{1+4bV_{\bar X}}\frac{1}{(1+2bV_{\bar X})^2}+\frac{V_{\bar X}}{1+2bV_{\bar X}}\right]\frac{1}{\sqrt{2\pi(\frac{1}{2b}+2V_{\bar X})}} \\
=\frac{b}{\pi}\frac{c^2N^2b^2}{\sqrt{1+4bV_{\bar X}}}\frac{V_{\bar X}}{1+2bV_{\bar X}}\left(\frac{1}{1+4bV_{\bar X}}+1\right)=\frac{2c^2N^2b^3V_{\bar X}}{\pi(1+4bV_{\bar X})^{3/2}}.
\end{multline}

Thus, using properties of the folded normal distribution, we find

\begin{equation}
\overline{|\beta|}=\sqrt{\frac{2}{\pi}}\frac{cNb^{3/2}}{(1+4bV_{\bar X})^{3/4}}\sqrt\frac{2V_{\bar X}}{\pi}=\frac{2}{\pi}\frac{cNb^{3/2}}{(1+4bV_{\bar X})^{3/4}}\sqrt{V_{\bar X}},
\end{equation}

\begin{equation}
\mathrm{Var}_{f_{\bar X}}(|\beta|)=\frac{c^2N^2b^3}{(1+4bV_{\bar X})^{3/2}}\frac{2V_{\bar X}}{\pi}\left(1-\frac{2}{\pi}\right).
\end{equation}

We also calculate

\begin{multline}
\mathrm{Cov}_{f_{\bar X}}(\alpha,\beta)=
\int_\mathbb{R}\int_\mathbb{R}\big(\bar\alpha-\alpha(\bar x_i,\bar x_j)\big)\big(\bar\beta-\beta(\bar x_i,\bar x_j)\big)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
=\frac{c^2Nb}{2\bar r}\sqrt\frac{b}{\pi}\int_\mathbb{R}\int_\mathbb{R}(\bar x_i-\bar x_j)\sqrt\frac{b}{\pi}\exp\left(-b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j=0.
\end{multline}

In attempt to calculate $\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)$ we find

\begin{multline}
\mathrm{Cov}f_{\bar X}(\alpha,|\beta|)=\int_\mathbb{R}\int_\mathbb{R}\alpha(\bar x_i,\bar x_j)|\beta(\bar x_i,\bar x_j)|f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j-\bar\alpha\overline{|\beta|} \\
=\int_\mathbb{R}\int_\mathbb{R}\frac{c}{\bar r}\sqrt\frac{b}{2\pi}\exp\left(-\frac{b}{2}(\bar x_i-\bar x_j)^2\right)cNb|\bar x_i-\bar x_j|\sqrt\frac{b}{2\pi}\exp\left(-\frac{b}{2}(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j-\bar\alpha\overline{|\beta|} \\
=\frac{c^2Nb}{\bar r}\sqrt\frac{b}{4\pi}\int_\mathbb{R}\int_\mathbb{R}|\bar x_i-\bar x_j|\sqrt\frac{b}{\pi}\exp\left(-b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j-\bar\alpha\overline{|\beta|}.
\end{multline}

Just as we used the folded normal to find $\overline{|\beta|}$ and $\mathrm{Var}_{f_{\bar X}}(|\beta|)$, we can calculate $\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)$ by considering

\begin{equation}
\int_\mathbb{R}\int_\mathbb{R}(\bar x_i-\bar x_j)\sqrt\frac{b}{\pi}\exp\left(-b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j=0
\end{equation}

and

\begin{multline}
\int_\mathbb{R}\int_\mathbb{R}(\bar x_i-\bar x_j)^2\frac{b}{\pi}\exp\left(-2b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
=\sqrt\frac{2b}{\pi}\int_\mathbb{R}\int_\mathbb{R}(\bar x_i-\bar x_j)^2\frac{1}{\sqrt{2\pi\frac{1}{4b}}}\exp\left(-\frac{(\bar x_i-\bar x_j)^2}{2\frac{1}{4b}}\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
=\sqrt\frac{2b}{\pi}\int_\mathbb{R}\left[\left(\frac{\bar{\bar x}+4bV_{\bar X}\bar x_j}{1+4bV_{\bar X}}-\bar x_j\right)^2+\frac{V_{\bar X}}{1+4bV_{\bar X}}\right]\frac{1}{\sqrt{2\pi(\frac{1}{4b}+V_{\bar X})}}\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(\frac{1}{4b}+V_{\bar X})}\right)f_{\bar X}(\bar x_j)d\bar x_j \\
=\sqrt\frac{2b}{\pi}\int_\mathbb{R}\left[\left(\frac{\bar{\bar x}-\bar x_j}{1+4bV_{\bar X}}\right)^2+\frac{V_{\bar X}}{1+4bV_{\bar X}}\right]\frac{1}{\sqrt{2\pi(\frac{1}{4b}+V_{\bar X})}}\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(\frac{1}{4b}+V_{\bar X})}\right)f_{\bar X}(\bar x_j)d\bar x_j \\
=\sqrt\frac{2b}{\pi}\left[\frac{(1+4bV_{\bar X})V_{\bar X}}{1+8bV_{\bar X}}\frac{1}{(1+4bV_{\bar X})^2}+\frac{V_{\bar X}}{1+4bV_{\bar X}}\right]\frac{1}{\sqrt{2\pi(\frac{1}{4b}+2V_{\bar X})}} \\
=\sqrt\frac{2b}{\pi}\frac{2V_{\bar X}}{1+8bV_{\bar X}}\sqrt{\frac{4b}{2\pi(1+8bV_{\bar X})}}=\frac{b}{\pi}\frac{4V_{\bar X}}{(1+8bV_{\bar X})^{3/2}}.
\end{multline}

Hence

\begin{equation}
\int_\mathbb{R}\int_\mathbb{R}|\bar x_i-\bar x_j|\sqrt\frac{b}{\pi}\exp\left(-b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j=\sqrt\frac{2}{\pi}\sqrt{\frac{b}{\pi}\frac{4V_{\bar X}}{(1+8bV_{\bar X})^{3/2}}}=\frac{2}{\pi}\frac{\sqrt{2bV_{\bar X}}}{(1+8bV_{\bar X})^{3/4}}
\end{equation}

and

\begin{multline}
\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)=\frac{c^2Nb}{\bar r}\sqrt\frac{b}{4\pi}\frac{2}{\pi}\frac{\sqrt{2bV_{\bar X}}}{(1+8bV_{\bar X})^{3/4}}-\bar\alpha\overline{|\beta|} \\
=\frac{2c^2Nb^2}{\pi\bar r(1+8bV_{\bar X})^{3/4}}\sqrt\frac{V_{\bar X}}{2\pi}-\frac{c}{\bar r}\sqrt\frac{b}{2\pi(1+2bV_{\bar X})}\frac{2}{\pi}\frac{cNb^{3/2}}{(1+4bV_{\bar X})^{3/4}}\sqrt{V_{\bar X}} \\
=\frac{2c^2Nb^2}{\pi\bar r(1+8bV_{\bar X})^{3/4}}\sqrt\frac{V_{\bar X}}{2\pi}-\frac{2c^2Nb^2}{\pi\bar r(1+4bV_{\bar X})^{3/4}}\sqrt\frac{V_{\bar X}}{2\pi(1+2bV_{\bar X})} \\
=\frac{2c^2Nb^2}{\pi\bar r}\sqrt\frac{V_{\bar X}}{2\pi}\left(\frac{1}{(1+8bV_{\bar X})^{3/4}}-\frac{1}{(1+4bV_{\bar X})^{3/4}(1+2bV_{\bar X})^{1/2}}\right).
\end{multline}

#### Starting the caclulation of $\mathrm{Cov}_{f_{\bar X}}(\alpha,\mathfrak{C})$ \label{caC}

We have

\begin{equation}
\mathfrak{C}(\bar x_i,\bar x_j)=c^2N^2b^2\Big(|\bar x_i-\bar x_j|+\big|1-b(\bar x_i-\bar x_j)^2\big|\Big)^2\exp\left(-\frac{b}{2}(\bar x_i-\bar x_j)^2\right).
\end{equation}

Note that the random variable $\delta=\bar x_i-\bar x_j$ is a mean zero Gaussian random variable with variance $2V_{\bar X}$. We write the probability density function of $\delta$ as $f_{\Delta}(\delta)$. Substituting in $\delta$, we can write

\begin{multline}
\mathfrak{C}(\delta,0)=c^2N^2b^2\Big(|\delta|+\big|1-b\delta^2\big|\Big)^2\exp\left(-\frac{b}{2}\delta^2\right) \\
=c^2N^2b^2\Big(\delta^2+2\big||\delta|-b|\delta|^3\big|+\big(1-b\delta^2\big)^2\Big)\exp\left(-\frac{b}{2}\delta^2\right).
\end{multline}

From this expression, we see properties of the folded normal distribution can be used to calculate several components of the intregal $\mathrm{Cov}_{f_{\bar X}}(\alpha,\mathfrak{C})$, but a major technical challenge lies in calculating

\begin{equation}
\int_{\mathbb{R}}\big||\delta|-b|\delta|^3\big|\exp\left(-\frac{b}{2}\delta^2\right)f_{\Delta}(\delta)d\delta.
\end{equation}

Instead of overcoming this challenge to find an analytical form of $\mathrm{Cov}_{f_{\bar X}}(\alpha,\mathfrak{C})$ we turn to a numerical approach outlined in the following section.

### Numerical estimates for heterogeneous $N$ and $G$

```{r, echo=F, fig.align='center', fig.cap='Numerical estimate for the correlations of selection gradients and competition coefficients.'}
knitr::include_graphics('/home/bob/Research/A White Noise Approach To Evolutionary Ecology/corrs.png')
```

Details on simulations, table of parameters, distributions of $a$ and $c$.

\newpage

# References

### 2020-03-20 .A White Noise Approach To Evolutionary Ecology/draft_01_24_20.Rmd
>

---
title: "A White Noise Approach to Evolutionary Ecology"
author: "Bob Week, Steve Krone, Scott Nuismer"
date: "8/19/2019"
abstract: "We derive the dynamics of the distribution of a quantitative character and the abundance of a biological population from a stochastic partial differential equation driven by space-time white noise. In the process we develop a useful set of heuristics to operationalize the powerful, but abstract theory of white noise. This approach allows us to compute the full implications of a stochastic process such as demographic stochasticity on phenotypic distributions and abundances of populations. We demonstrate the utility of our approach by deriving a model of diffuse coevolution mediated by resource competition. Alongside trait distributions, this model predicts species abundance distributions and interaction networks. We compute these quantities under the two scenarios of weak and moderate competition. A method for the statistical inference of model parameters based on these quantities is then outlined. Results derived from this formalization can contribute to the resolution of long-standing debates on the feedbacks between evolution and ecology by providing statistical methods for the quantification of processes shaping ecological communities from data on phenotypes, abundances and interaction networks."
bibliography: ref.bib
output:
  pdf_document:
    toc: false
    number_sections: true
    fig_width: 4
    fig_height: 3
    fig_caption: true
header-includes:
  - \usepackage{float}
  - \usepackage{mathrsfs}
  - \usepackage{hyperref}
  - \usepackage[left]{lineno}
  - \linenumbers
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.pos = 'H')
```

# Introduction

A primary aim of theoretical evolutionary ecology is the development of mathematical approaches to describe the evolution of biological populations and their interactions with both the biotic and abiotic environments in which they are embedded. Given this consideration a natural scope for such an approach centers on quantifying the abundance dynamics of populations and the evolution of traits mediating their interactions as functions of relevant abiotic factors. Although taking into account abundance, phenotype and environment provides the basis for a partial understanding of the complex nature of biological communities, for a deeper understanding one must account for the effects of contemporary dispersal and the phylogeographic history of interacting lineages [@Kraft2007; @Hickerson2010; @Manceau2016; @markmcpeek2017] along with the genetic basis of ecologically relevant traits [@jeffreyconner2004; @FUSSMANN2007] and feedbacks between populations and the biogeochemical cycles they ultimately depend on [@michelloreau2010; @9781107648258]. It is therefore ideal that the development of any such mathematical approach anticipates extensions to account for these important factors shaping ecological communities, especially as empirical and conceptual work in these directions continues to grow [@AbdalaRoberts2014; @Klzsch2015; @Crutsinger2015; @Fitzpatrick2015; @Fitzpatrick2017; @Marx2017; @Rudman2017; @Skovmand2018; @VanNuland2019; @Harmon2019]. Furthermore, the approach should have a stochastic component to capture the chance nature of biological reality [@Lande2003; @DeMeester2018; @Mubayi2019] and serve as a basis for the construction of statistical methods that measure evolutionary and ecological processes occuring in the wild. Such methods will tether theory to reality and allow for rigorous tests of hypotheses on the structure and behavior of ecological communities. In this paper we take one small step in this direction by establishing a formal connection between the continuous-time dynamics of abundance and quantitative traits in stochastically evolving populations and introduce some statistical methods based on this connection.

Current theoretical approaches to synthesize evolution and ecology have capitalized on the fact that biological fitness plays a key role in determining both sets of dynamics. While correlation of fitness and genotype is the basis of evolution by natural selection, the mean fitness across all individuals in a population determines the growth, stasis or decline of abundance. In section \ref{deterministic} we review the mathematical formalization of this connection, which has been established in the contexts of population genetics [@9781932846126], evolutionary game theory [@9780674023383] and quantitative genetics [@Lion2018]. Although reviewing these accomplishments reveals a beautiful synthesis of evolution and population ecology, it also reveals a gap in theoretical approaches to incorporate the intrinsically random nature of populations. Specifically, in theoretical quantitative genetics the derivation of a populations' response to random genetic drift is derived in discrete time under the assumption of constant population (or effective population) size [@Lande1976]. Although this approach conveniently mimics the formalism provided by the Wright-Fisher model of population genetics, real population sizes fluctuate over time. Furthermore, since these fluctuations are themselves stochastic, it seems natural to derive expressions for the evolutionary response to demographic stochasticity and consider how the results relate to characterizations of random genetic drift. This can be done in continuous time for population genetic models without too much technical overhead, assuming a finite number of alleles. However, for populations with a continuum of types, such as a quantitative trait, this becomes a serious mathematical endeavor. Here we close this gap by applying some theory of stochastic partial differential equations (SPDE). Our goal is two-fold: 1) Communicate some useful results of SPDE to as wide of audience as possible. With this goal in mind we will not provide a rigorous treatment of SPDE. Instead, we introduce a set of heuristics that only require the basic concepts of Reimann integration, partial differentiation and some exposure to stochastic ordinary differential equations (SDE). For a concise introduction to SDE, we recommend the primer by Evans [-@9781470410544]. 2) Establish a novel synthetic approach to theoretical evolutionary ecology. To demonstrate that our approach can yield important statistical tools for the inference of evolutionary and ecological processes, we derive a model of coevolution in an ecological network and outline some methods of inference based on this model.

We begin in section \ref{deterministic} with a brief review of progress in modelling the deterministic dynamics of populations. Section \ref{wnc} then introduces a useful set of mathematical tools based on the calculus of white noise. In section \ref{stochastic} we outline how to employ these tools to derive a system of SDE modelling abundance dynamics and the evolution of the mean and variance of a quantitative trait. In section \ref{particular} we derive particular results under the assumption of a normally distributed trait. Section \ref{inheritance} extends the previous results under the condition of imperfect inheritance. Section \ref{coev} demonstrates the utility of our framework through a derivation of a model of coevolution across a guild of $S$ species competing along a common resource axis. We numerically compute distributions of traits and abundances along with interaction networks under the scenarios of weak and moderate competition. In section \ref{stats} we discuss an approach to develop statistical methods based on our framework for the inference of evolutionary and ecological processes using trait, abundance and network data. Finally, section \ref{conclusion} concludes with a few remarks on the limits of this approach along with future directions to incorporate more explicitly the genetic architecture of populations and feedbacks with ecosystem processes.

# The framework

At the core of our approach is a stochastic analog of the replicator equation with mutation in continuous time and phenotypic space [@Taylor1978; @Schuster1983]. From this stochastic replicator-mutator equation we derive a system of SDE for the dynamics of abundance, mean trait and additive genetic variance of a population. Hence, our approach develops a quantitative genetic theory of evolutionary ecology. A popular alternative to quantitative genetics is the theory of adaptive dynamics. As demonstrated by Page and Nowak [-@PAGE2002], the canonical equation of adaptive dynamics can be derived from the replicator-mutator equation. Thus, one could start from the atomic roots of our approach and pursue a stochastic adaptive dynamic theory instead. We choose the former in anticipation of an extension of our approach that explicitly models the genetic details of populations.

In this section we review the derivations of the replicator-mutator equation and trait dynamics from abundance dynamics and extend these formulae along with related results to the stochastic case. The results established in this section provide the framework from which larger scale ecological stuctures, such as species abundance distributions and interaction networks, can be derived.

## Deterministic dynamics \label{deterministic}

The connection between the deterministic dynamics of the abundance of a population and the frequencies of discrete phenotypes within it is well established via heterogeneity in the vital rates of differing phenotypes [ie, demographic heterogeneity, @Mubayi2019]. This is particularly evident in the derivation of the evolution of allele frequencies due to natural selection provided by Crow and Kimura [-@9781932846126, \S 5.3]. Specifically, denoting $N$ the abundance of the population, $m_i$ the growth rate [called the Malthusian parameter in @9781932846126] of allele $i$ and $p_i$ the frequency of this allele, we find (using their numbering scheme)

\begin{equation}
\dot N=N\sum_ip_im_i=\bar mN,\tag{5.3.3}
\end{equation}

\begin{equation}\label{popgen}
\dot p_i=(m_i-\bar m)p_i.\tag{5.3.5}
\end{equation}

Equation (\ref{popgen}) is known in the field of evolutionary game theory as the replicator equation [@9780674023383; @Lion2018]. Instead of being explicitly focused on alleles, the replicator equation describes the fluctuations of relative abundances of various _types_ in a population in terms of the vital rates of each type [@Taylor1978; @Schuster1983]. It is straight-forward to extend the replicator equation to include mutation, which is known as the replicator-mutator equation [@9780674023383]. To derive the replicator-mutator equation one can take the same steps to derive (\ref{popgen}) and inject a model of mutation via a matrix of transition rates [@9780674023383, eqn. 13.25]. Building on these insights, we derive an analog of the replicator-mutator equation for a continuum of types (that is, for a quantitative trait). In particular, for a continuously reproducing population at time $t$ with trait values $x\in\mathbb{R}$ and an abundance density $n(t,x)$, the demographic dynamics can be modelled by the general PDE

\begin{equation}\label{eq1}
\dot n(t,x)=F(n,x).
\end{equation}

More specifically, assuming selection is determined by the growth rate $m(n,x)$ (which we assume to be continuous in $n$ and smooth in $x$), mutation is captured by diffusion with coefficient $\frac{\mu}{2}$, and ignoring all other processes, we set

\begin{equation}
F(n,x)=m(n,x)n(t,x)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}n(t,x).
\end{equation}

In this case equation (\ref{eq1}) becomes a reaction-diffusion equation. As an example, logistic growth combined with stabilizing selection can be captured using the growth rate $m(n,x)=r-\frac{a}{2}(\theta-x)^2-n(t,x)$ where $r>0$ determines the carrying capacity and $a\geq0$ determines the strength of stabilizing selection around the phenotypic optimum $\theta\in\mathbb{R}$.

To derive a replicator-mutator equation from (\ref{eq1}), we employ the chain rule from elementary calculus. Defining $N(t)=\int_\mathbb{R}n(t,x)dx$ as the total abundance of the population and $p(t,x)=n(t,x)/N(t)$ as the relative density of trait $x$ at time $t$, we have

\begin{equation}
\dot N(t)=\frac{d}{dt}\int_\mathbb{R}n(t,x)dx=\int_\mathbb{R}\dot n(t,x)dx
\end{equation}

$$=\int_\mathbb{R}m(n,x)n(t,x)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}n(t,x)dx$$

$$=N(t)\int_\mathbb{R}m(n,x)p(t,x)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(t,x)dx=\bar m(t) N(t)$$

where, for $f:\mathbb{R}\to\mathbb{R}$, we define (when it exists) $\bar f(t)=\int_\mathbb{R}f(x)p(t,x)dx$. In particular $\bar m(t)$ is the mean fitness across all trait values in the population. Assuming $n(t,x)>0$  for all $(t,x)\in(0,\infty)\times\mathbb{R}$ (which, given $n(0,x)\geq0$ for all $x$ and $n(t,x)>0$ for some $x$, we expect due to the infinite-propagation property of diffusion [@lawrenceevans2010]), and assuming $\int_\mathbb{R}n(t,x)dx<\infty$ (which requires $m(n,x)$ to not grow too fast in either of its arguments), the result $\int_\mathbb{R}\frac{\partial^2}{\partial x^2}n(t,x)dx=0$ can be derived from the fundamental theorem of calculus. Biological intuition agrees with this result since mutation neither creates nor destroys individuals, but merely changes their type from their parental type. Taking the same appraoch, we derive the dynamics of $p(t,x)$ as

\begin{equation}
\dot p(t,x)=\frac{\partial}{\partial t}\frac{n(t,x)}{N(t)}=\frac{N(t)\dot n(t,x)-n(t,x)\dot N(t)}{N^2(t)}
\end{equation}

$$=\frac{1}{N(t)}\Big(m(n,x)n(t,x)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}n(t,x)-\bar m(t)n(t,x)\Big)$$

$$=(m(n,x)-\bar m(t)) p(t,x)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(t,x).$$

This result closely resembles Kimura's continuum-of-alleles model [@Kimura1965]. The primary difference being that our model utilizes diffusion instead of convolution with an arbitrary mutation kernel. Of course, our model can be derived as an approximation to Kimura's model, which has been referred to as the Gaussian allelic approximation [@9780471986539], the infinitesimal model [@Barton2017] and the Gaussian descendants approximation [@Turelli2017]. Alternatively, since diffusion is the continuous-time equivalent to convolution against a Gaussian kernel (SM \S\ref{diffconvequiv}), our model can also be seen as a special case of the continuum-of-alleles model. The dynamics of the mean trait $\bar x(t)$ are derived as

\begin{equation}\label{detmean}
\dot{\bar x}(t)=\frac{d}{dt}\int_\mathbb{R}x p(t,x)dx=\int_\mathbb{R}x\dot p(t,x)dx
\end{equation}

$$=\int_\mathbb{R}x\big(m(n,x)-\bar m(t)\big) p(t,x)+x\frac{\mu}{2}\frac{\partial^2}{\partial x^2} p(t,x)dx,$$

$$=\mathrm{Cov}_ p\Big(x,m(n,x)\Big)+\frac{\mu}{2}\int_\mathbb{R} x\frac{\partial^2}{\partial x^2} p(t,x)dx,$$

where, for $f,g:\mathbb{R}\to\mathbb{R}$, we define (when it exists)

\begin{equation}
\mathrm{Cov}_ p\Big(f(x),g(x)\Big)=\int_\mathbb{R}\big(f(x)-\bar f(t)\big)\big(g(x)-\bar g(t)\big)p(t,x)dx.
\end{equation}

Equation (\ref{detmean}) is a special case of the well known Robertson-Price equation [@Robertson1966; @PRICE1970; @FRANK2012; @Queller2017; @Lion2018]. Conveniently, when $p$ is Gaussian $\int_\mathbb{R}x\frac{\partial^2}{\partial x^2}p(t,x) dx=0$. Following this approach, in SM \S\ref{var_deriv} we derive the dynamics of the variance $\sigma^2(t)=\mathrm{Cov}_p(x,x)$ as

\begin{equation}\label{vardyn_det}
\dot{\sigma}^2(t)=\mathrm{Cov}_p\big((x-\bar x)^2,m\big)+\frac{\mu}{2}\int_\mathbb{R}(x-\bar x)^2\frac{\partial^2}{\partial x^2}p(t,x)dx.
\end{equation}

In the absence of mutation equation (\ref{vardyn_det}) agrees with the result derived by Lion [-@Lion2018] for discrete phenotypes. When $p$ is Gaussian the increase in variance due to mutation simplifies to $\mu$.

In the following sections we extend these results to the case when PDE (1) is driven by a space-time white noise process $\dot\xi(t,x)$. We begin by defining space-time white noise and illustrating its relevant properties. We then introduce a set of heuristics for performing calculations with respect to $\dot\xi(t,x)$. Since our aim is to present this material to as wide of audience as possible, our approach deviates from standard definitions to remove the need for a detailed technical treatment. For those not interested in the white noise calculus and would rather jump straight into more biologically relevant results, we recommend skipping to \S\ref{particular}.

## White noise calculus \label{wnc}

### Definition and basic properties \label{wnc_intro}

We define $L^2_C$ to be the set of functions $f:[0,\infty)\times\mathbb{R}\to\mathbb R$ that are continuous in their first argument (time) and satisfy $\|f\|_2(t)=\sqrt{\int_\mathbb R f^2(t,x)dx}<\infty$. We define a generalized stochastic process $\Xi$ that maps functions $f\in L^2_C$ to real-valued stochastic processes. More specifically, for any $f,g\in L^2_C$, we define $\Xi_f$ and $\Xi_g$ as Gaussian processes satisfying, for any $t,t_1,t_2\geq0$,

\begin{equation}\label{exp_WN}
\mathbb{E}\big(\Xi_f(t)\big)=\mathbb{E}\big(\Xi_g(t)\big)=0,
\end{equation}

\begin{equation}\label{cov_WN}
\mathbb{C}\big(\Xi_f(t_1),\Xi_g(t_2)\big)=\int_0^{t_1\wedge t_2}\int_\mathbb{R} f(s,x)g(s,x)dxds,
\end{equation}

where $t_1\wedge t_2=\min(t_1,t_2)$ and $\mathbb{E},\mathbb{C}$ denote expectation and covariance with respect to the underlying probability space. These operators are to be distinguished from $\bar f$ and $\mathrm{Cov}_ p(f,g)$ which denote expectation and covariance with respect to phenotypic diversity.

Since Gaussian processes are characterized by their expectations and covariances, the processes $\Xi_f$ and $\Xi_g$ are well defined (under the convention $f=g$ whenever $\|f-g\|_2(t)=0$ for each $t>0$). In particular, if $f\in L^2_C$ is independent of time, then $\Xi_f$ is a Brownian motion with variance $t\|f\|_2^2$. With the generalized process $\Xi$ defined, we define the space-time white noise $\dot\xi(t,x)$ implicitly via

\begin{equation}
``\int_0^t\int_\mathbb{R}f(s,x)\dot\xi(s,x)dxds"=\Xi_f(t), \ \forall \ f\in L^2_C.
\end{equation}

We place quotations around the left-hand side of the above expression to emphasize its informal nature. Since space-time white noise, represented by the symbol $\dot\xi(t,x)$, does not exist as a function in the classical sense it cannot be integrated as suggested above. However, similar to the approach taken here, space-time white noise can be made sense of in a more abstract setting as a _measure_-valued process [@Dawson1975; @vladimirbogachev1998]. Hence, our use of integral notation is merely symbolic to provide intuition and should not be confused with classical Riemann integrals. One can think of $\dot\xi(t,x)$ as the static seen on old television sets or infinitely detailed random dust spread across both time and space. Figure \ref{wn} illustrates sample paths of white noise in one (left) and two (right) dimensions.

```{r, include=F, eval=F}
require("ggplot2")
require("ggthemes")
require("plot3D")
require("gridExtra")

num1 = 1e4
num2 = 1e4

xi_1 = rnorm(num1)
xi_2 = rnorm(num2)

x1 = seq(0,1,length.out = num1)
x2 = seq(0,1,length.out = sqrt(num2))

M = expand.grid(x=x2,y=x2)
M$value = xi_2

df_1 = data.frame(xi=xi_1,x=x1)

wn1d = ggplot(df_1)+geom_point(aes(x=x,y=xi),size=0.05)+theme_bw()+xlab("Time")+ylab("Value")
wn2d = ggplot(M, aes(x, y, color = value))+scale_color_gradient(low="black", high="white")+geom_point(size=0.001, shape=22, show.legend = FALSE)+theme_bw()+xlab("Time")+ylab("Space")

wn = grid.arrange(wn1d,wn2d,nrow=1)

ggsave("~/Research/WNEE/wn.png",wn,width=5,height=2.5)
```

```{r, echo=F, fig.align='center', fig.cap='\\label{wn}Approximations of sample paths of temporal white noise (left) and space-time white noise (right) with brightness scaled to value.'}
knitr::include_graphics('wn.png')
```

Following our definition of white noise, when attempting to measure the value of this random field we sample it using $L^2_C$ functions. For example, integrating white noise over a region $[0,t]\times D$, with $t>0$ and $D$ a bounded subset of $\mathbb{R}$, is equivalent to evaluating $\Xi_{I_{[0,\infty)\times D}}(t)$ where

\begin{equation}
I_{[0,\infty)\times D}(t,x)=\left\{\begin{matrix}
0, & x\notin D \\
1, & x\in D
\end{matrix}\right..
\end{equation}

Since $\int_\mathbb{R}{I_{[0,\infty)\times D}}^2(t,x)dx=\int_Ddx=|D|$, where $|D|$ denotes the length of $D$, $I_{[0,\infty)\times D}\in L^2_C$. Thus, by equations (\ref{exp_WN}) and (\ref{cov_WN}), we have

\begin{equation}
\mathbb{E}\left(\int_0^t\int_D\dot\xi(s,x)dxds\right)=0,
\end{equation}

\begin{equation}
\mathbb{V}\left(\int_0^t\int_D\dot\xi(s,x)dxds\right)=t|D|,
\end{equation}

where $\mathbb{V}$ denotes the variance operator with respect to the underlying probability space. Following this notation, equations (\ref{exp_WN}) and (\ref{cov_WN}) can be rewritten as

\begin{equation}\label{exp_WN_xi}
\mathbb{E}\left(\int_0^t\int_\mathbb{R}f(s,x)\dot\xi(s,x)dxds\right)=0,
\end{equation}

\begin{equation}\label{cov_WN_xi}
\mathbb{C}\left(\int_0^{t_1}\int_\mathbb{R}f(s,x)\dot\xi(s,x)dxds,\int_0^{t_2}\int_\mathbb{R}g(s,x)\dot\xi(s,x)dxds\right)
=\int_0^{t_1\wedge t_2}\int_\mathbb{R}f(s,x)g(s,x)dxds.
\end{equation}

We write $d\hat\Xi_f(t)=\frac{1}{\|f\|_2(t)}\left(\int_\mathbb{R}f(t,x)\dot\xi(t,x)dx\right)dt$ (and typically suppress the dependency on $t$) so that

\begin{equation}
\int_0^td\hat\Xi_f(s)=\int_0^t\int_\mathbb{R}\frac{f(s,x)}{\sqrt{\int_\mathbb{R}f^2(s,y)dy}}\dot\xi(s,x)dxds.
\end{equation}

It follows that $\mathbb{E}\big(\int_0^td\hat\Xi_f\big)=0$ and $\mathbb{C}\big(\int_0^{t_1}d\hat\Xi_f,\int_0^{t_2}d\hat\Xi_f\big)=t_1\wedge t_2$. This implies that, as a function of $t$, $\int_0^td\hat\Xi_f(s)$ is a standard Brownian motion for any $f\in L^2_C$. Hence, $d\hat\Xi_f$ is analogous to the traditional shorthand used to denote stochastic differentials. Thus, equation (\ref{cov_WN_xi}) effectively extends It&ocirc;'s multiplication table to:

\begin{table}[H]
\centering\caption{An extension of It\^o's multiplication table.}\vspace{0.2cm}
\begin{tabular}{l|lll}
             & $d\hat\Xi_f$                 & $d\hat\Xi_g$                 & $dt$ \\ \hline
             &                            &                            &      \\
$d\hat\Xi_f$ & $dt$                       & $\frac{\left(\int_\mathbb{R}f \ g \ dx\right)dt}{\|f\|_2\|g\|_2}$ & $0$  \\
             &                            &                            &      \\
$d\hat\Xi_g$ & $\frac{\left(\int_\mathbb{R}f \ g \ dx\right)dt}{\|f\|_2\|g\|_2}$ & $dt$                       & $0$  \\
             &                            &                            &      \\
$dt$         & $0$                        & $0$                        & $0$
\end{tabular}\label{mult}
\end{table}

### Interpretation of SPDE

Since $\dot\xi(t,x)$ is only well defined with respect to the generalized stochastic process $\Xi$, SPDE such as

\begin{equation}\label{SPDE}
\dot n(t,x)=m(n,x)n(t,x)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}n(t,x)+\sqrt{V(n,x)n(t,x)}\dot\xi(t,x)
\end{equation}

should be understood more precisely as a shorthand for defining the generalized stochastic process $n$. However, although $\Xi$ maps $L^2_C$ functions to real-valued processes, the generalized process $n$ maps bounded countinuous functions to real-valued processes [@zeng1998absolute]. For more on solutions of SPDE as generalized processes see Walsh [-@Walsh]  and Dawson [-@dawson1993measure]. When $m(n,x)$ and $V(n,x)$ are bounded across all $n$ and $x$, the generalized solution $n$ can be represented by a legitimate function, which we denote by $n(t,x)$, that is non-negative, continous in both $t$ and $x$ and integrable in $x$ for any $t>0$ [@zeng1998absolute]. More specifically, for any bounded continuous function $f$,

\begin{equation}\label{SOLUTION}
\int_\mathbb{R}\big(n(t,x)-n(0,x)\big)f(x)dx=\int_0^t\int_\mathbb{R}n(s,x)\left(m(n,x)f(x)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}f(x)\right)dxds
\end{equation}

\begin{equation*}
+\int_0^t\int_\mathbb{R}\sqrt{V(n,x)n(s,x)}f(x)\dot\xi(s,x)dxds.
\end{equation*}

In particular, setting $f(x)=1$ allows us to derive to total mass process, which we refer to as the

This property of solutions turns out to be quite useful in our situation. In particular, setting $h(x)=1,x,$ or $x^2$ provides the first steps to derive a system of SDE to track the total abundance, mean trait or phenotypic variance respectively. We demonstate the utility of this definition in the next section.

## Stochastic dynamics of abundance, mean trait and phenotypic variance \label{stochastic}

We base our formulation on the SPDE (\ref{SPDE}). A special case of this SPDE with $m(n,x)\equiv0$ was suggested by Dawson [-@Dawson1975] to be satisfied by the diffusion limit of a branching Brownian motion, which itself was initially studied by Watanabe [-@Watanabe1968]. The solution $n$ is referred to as a Dawson-Watanabe superprocess [@alisonetheridge2000]. This connection was made rigorous for spatial dimension $d=1$ by Konna and Shiga [-@Konno1988] and Reimers [-@Reimers1989]. Furthermore, the Dawson-Watanabe superprocess is closely related to the $\Lambda$-Fleming-Viot process introduced by Bertoin and Le Gall [-@Bertoin2003] and used by Etheridge [-@Etheridge2008] to resolve some technical challenges in modelling isolation by distance [@Felsenstein1975; see also @Barton2013; and @Barton2019].

The branching Brownian motion process is, essentially, a birth-death process with spatial structure. It has three main components: spatial movement, branching rate and branching mechanism. Individuals are born at the current location of their mothers and then move around geographic space according to Brownian motion. Lifetimes of these individuals are typically assumed to be exponentially distributed. The branching mechanism describes the distribution of offspring left by the parent. For more information on branching Brownian motion and an introduction to superprocesses see Etheridge [-@alisonetheridge2000]. In our context we replace geographic space with phenotypic space where movement is reinterpreted as mutation. To make this intepretation more rigorous, we require the duration of movement to be non-random $1/N$ units of time so that mutation rates remain equal across individuals. We take $N\to\infty$ in the diffusion limit. In this case the only difference between life-histories of organisms is the number of offspring left. Selection can be imposed by allowing the mean number of offspring produced to vary with phenotype.

A more general version of the Dawson-Watanabe superprocess that includes interactions, which can represent, among other phenomena, density-dependent growth rates and frequency-dependent selection, was studied by Perkins [-@Perkins1992; -@edwinarendperkins1995]. When the interactions occur only at branching events, which in our case translates to independence of mutation and selection, the superprocess with interactions on one dimensional trait space corresponds to a generalization of SPDE (\ref{SPDE}) [@zeng1998absolute]. As noted above, equation (\ref{SPDE}) has been shown to admit function-valued solutions that are positive, integrable and continuous in time and space. Hence, SPDE (\ref{SPDE}) provides a very natural and mathematically convenient candidate to base our framework on.

### Derivation of SDE for $N$, $\bar x$ and $\sigma^2$

The dynamics of abundance (referred to in the superprocess literature as the total mass process) is obtained by defining the process $N(t)=\int_{\mathbb{R}}n(t,x)dx$ and evaluating expression (\ref{SOLUTION}) with $h(x)=1$ to obtain

\begin{equation}
N(t)-N(0)=\bar m(t)N(t)+\int_0^t\int_\mathbb{R}\sqrt{V(n,x)n(s,x)}\dot\xi(s,x)dxds.
\end{equation}

By equation (\ref{cov_WN_xi}) we have

\begin{equation}
\mathbb{V}\left(\int_0^t\int_\mathbb{R}\sqrt{V(n,x)n(s,x)}\dot\xi(s,x)dxds\right)=\int_0^t\int_\mathbb{R}V(n,x)n(s,x)dxds=\int_0^t\bar V(s)N(s)ds,
\end{equation}

where $\bar V(t)=\int_\mathbb{R}V(n,x)p(t,x)dx$. Following the notation introduced in \S\ref{wnc_intro}, we have

\begin{equation}
\int_0^td\hat\Xi_{\sqrt{Vn}}=\int_0^t\int_\mathbb{R}\frac{\sqrt{V(n,x)n(s,x)}}{\sqrt{\bar V(s)N(s)}}\dot\xi(s,x)dxds
\end{equation}

is, as a function of $t$, a standard Brownian motion. To clean up notation set $d\xi_1=d\hat\Xi_{\sqrt{Vn}}$. Then we can write the SDE for $N(t)$ as

\begin{equation}
dN=\bar mNdt+\sqrt{\bar VN}d\xi_1.
\end{equation}

Defining the process $p(t,x)=n(t,x)/N(t)$, in SM \S\ref{SDE_DERIV} we extend this approach to derive SDE for $\bar x(t)$ and $\sigma^2(t)$ which provides

\begin{equation}
d\bar x=\left(\mathrm{Cov}_ p(x,m)+\frac{\mu}{2}\int_\mathbb{R}x\frac{\partial^2}{\partial x^2}p(t,x)dx-\mathrm{Cov}_ p\Big(x,\frac{V}{N}\Big)\right)dt+\sqrt{\frac{\sigma^2\bar V+\mathrm{Cov}_p\left((x-\bar x)^2,V\right)}{N}}d\xi_2,
\end{equation}

\begin{equation}
d\sigma^2=\left(\mathrm{Cov}_ p\Big((x-\bar x)^2,m\Big)+\frac{\mu}{2}\int_\mathbb{R}(x-\bar x)^2\frac{\partial^2}{\partial x^2}p(t,x)dx-\frac{\sigma^2\bar V}{N}\right)dt
\end{equation}

\begin{equation}
+\sqrt\frac{\left(\mathrm{Cov}_p\big((x-\bar x)^4,V\big)+\overline{(x-\bar x)^4}\bar V-2\sigma^2\mathrm{Cov}_p\big((x-\bar x)^2,V\big)-\sigma^4\bar V\right)}{N}d\xi_3
\end{equation}

where $\xi_2$ and $\xi_3$ are standard Brownian motions and derivatives with respect to $x$ are taken in the weak sense [for more on weak derivatives see @lawrenceevans2010]. In SM \S\ref{SDE_DERIV} we show that in general $\xi_1$ is independent of both $\xi_2$ and $\xi_3$, but $\xi_2$ and $\xi_3$ covary. We simpify these results by assuming $\sqrt{V(n,x)}\equiv\nu>0$ so that

\begin{equation}\label{N}
dN=\bar mNdt+\nu\sqrt{N}d\xi_1,
\end{equation}

\begin{equation}\label{xbar_gen}
d\bar x=\mathrm{Cov}_ p(x,m)dt+\frac{\mu}{2}\int_\mathbb{R}x\frac{\partial^2}{\partial x^2}p(t,x)dx+\frac{\nu\sigma}{\sqrt N}d\xi_2,
\end{equation}

\begin{equation}\label{sig2_gen}
d\sigma^2=\left(\mathrm{Cov}_ p\Big((x-\bar x)^2,m\Big)+\frac{\mu}{2}\int_\mathbb{R}(x-\bar x)^2\frac{\partial^2}{\partial x^2}p(t,x)dx-\frac{\nu^2\sigma^2}{N}\right)dt+\nu\sqrt{\frac{\overline{(x-\bar x)^4}-\sigma^4}{N}}d\xi_3.
\end{equation}

The expressions derived can be used to investigate the dynamics of the mean and variance for general $n$. However, in the next section we further simplify these expressions by approximating $n$ with a Gaussian curve. In SM \S\ref{SDE_DERIV} we show that under the Gaussian case $\xi_1,\xi_2$ and $\xi_3$ are independent.

## Particular results assuming a Gaussian phenotypic distribution \label{particular}

Gaussian phenotypic distributions are often obtained through Gaussian, exponential or weak selection approximations together with a simplified model of inheritance and random mating [@pmid17248993; @Turelli1984; @Turelli1986; @Turelli2017; @9780471986539]. Alternatively, it has been shown that a Gaussian distribution can provide a reasonable approximation even when selection is strong and non-Gaussian [@pmid7851785]. However, our approach adds an additional layer of difficulty. Even with Gaussian selection, the resulting solution to SPDE (\ref{SPDE}) will only be Gaussian in expectation, assuming a Gaussian initial condition. Yet this difficulty is not as challenging as it may first appear. Indeed, since SPDE (\ref{SPDE}) can be derived as a diffusion limit we know that, under the appropriate assumptions on selection, genetic architecture and reproduction, the stochastic departure from a Gaussian curve is negligible when the ratio $\nu^2/N$ is small (ie, when the variance in reproductive output is much smaller than the population size). Mathematically, this requirement restricts model parameters to regions that maintain large population sizes. Biologically, this implies populations are not at risk of extinction. Hence, models developed in this framework are not suitable for studying colonization-extinction dynamics or evolutionary rescue. Allowing for these restrictions, we may safely assume that $n$ is approximately Gaussian and justify writing

\begin{equation}
n(t,x)=\frac{N(t)}{\sqrt{2\pi\sigma^2(t)}}\exp\left(-\frac{\big(x-\bar x(t)\big)^2}{2\sigma^2(t)}\right).
\end{equation}

Under this assumption we find (suppressing the dependency on $t$)

\begin{equation}
\mathrm{Cov}_ p(x,m)=\sigma^2\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right),
\end{equation}

\begin{equation}
\mathrm{Cov}_ p\Big((x-\bar x)^2,m\Big)=2\sigma^4\left(\frac{\partial\bar m}{\partial\sigma^2}-\overline{\frac{\partial m}{\partial\sigma^2}}\right)
\end{equation}

and $\overline{(x-\bar x)^4}=3\sigma^4$. In particular, this implies

\begin{equation}\label{xbar}
d\bar x=\sigma^2\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right)dt+\frac{\nu\sigma}{\sqrt N}d\xi_2,
\end{equation}

\begin{equation}\label{G}
d\sigma^2=2\sigma^4\left(\frac{\partial\bar m}{\partial\sigma^2}-\overline{\frac{\partial m}{\partial\sigma^2}}\right)dt +\left(\mu-\frac{\nu^2\sigma^2}{N}\right)dt +\nu\sigma^2\sqrt{\frac{2}{N}}d\xi_3.
\end{equation}

## The evolution of additive genetic variance \label{inheritance}

To model imperfect heritability we consider the relationship between the expressed phenotype $x$ and its genetic component $g\in\mathbb{R}$ known as the _breeding value_ [@michaellynch1998]. For any given individual with breeding value $g$, we assume the expressed phenotype is normally distributed around $g$ with variance $\eta$. We denote the density of this distribution by $\psi$ so that

\begin{equation}
\psi(x|g)=\frac{1}{\sqrt{2\pi\eta}}\exp\left(-\frac{(x-g)^2}{2\eta}\right).
\end{equation}

In terms of classical quantitative genetics $\eta$ is the _variance of environmental deviation_ and $\psi$ plays a similar role as the _segregation kernel_ [@michaellynch1998;@jonathanroughgarden1979]. To include this relationship in our framework, we write $\gamma(t,g)$ as the abundance density of breeding values at time $t$. In an actual population with a finite number of individuals $\gamma(t,g)$ can be written as a sum of point masses. In particular, denoting $\delta$ the Dirac distribution having unit mass centered at zero, for a population with breeding values $g_1,\dots,g_{N(t)}$ at time $t$ we have

\begin{equation}
\gamma(t,g)=\sum_{i=1}^{N(t)}\delta(g-g_i).
\end{equation}

Denoting by $X_i$ the normally distributed expressed phenotype for individual $i$, $n(t,x)$ can be written as

\begin{equation}
n(t,x)=\sum_{i=1}^{N(t)}\delta(x-X_i).
\end{equation}

The random variables $X_1-g_1,\dots,X_{N(t)}-g_{N(t)}$ are identically and independently distributed with mean zero and variance $\eta$. Hence, as $N(t)\to\infty$

\begin{equation}
n(t,x)\to\sum_{i=1}^{N(t)}\int_\mathbb{R}\delta(g-g_i)\psi(x|g)dg=\sum_{i=1}^{N(t)}\frac{1}{\sqrt{2\pi\eta}}\exp\left(-\frac{(x-g_i)^2}{2\eta}\right)
\end{equation}

$$
=\int_\mathbb{R}\gamma(t,g)\frac{1}{\sqrt{2\pi\eta}}\exp\left(-\frac{(x-g)^2}{2\eta}\right)dg.
$$

Note that when $\gamma$ converges to a Gaussian distribution in the diffusion limit its mode coincides with $\bar x$ and, denoting $G$ the variance of breeding values (ie, the additive genetic variance), $\sigma^2=G+\eta$. Hence, narrow sense heritability can be computed as $h^2=G/(G+\eta)$. Using the assumed relationship between expressed phenotypes and breeding values, we can compute the net Malthusian fitness of a breeding value $g$ via

\begin{equation}
m^*(\gamma,g)=\int_\mathbb{R}m(n,x)\frac{1}{\sqrt{2\pi\eta}}\exp\left(-\frac{(x-g)^2}{2\eta}\right)dg.
\end{equation}

This is similar to the approach taken by Kimura and Crow [-@Kimura1978] to calculate the overall effects of selection for expressed characters onto the changes in the distribution of alleles encoding those characters. With $m^*$ defined, the evolution of $\gamma$ is then captured by

\begin{equation}
\dot\gamma(t,g)=m^*(\gamma,g)\gamma(t,g)+\frac{\mu}{2}\frac{\partial^2}{\partial g^2}\gamma(t,g)+\nu\sqrt{\gamma(t,g)}\dot\xi(t,g).
\end{equation}

Futhermore, since

\begin{equation}
\bar m(t)= \int_\mathbb{R}m(n,x) p(t,x)dx
\end{equation}

$$=\int_\mathbb{R}\int_\mathbb{R}m^*(\gamma,g) p(t,x)\frac{1}{\sqrt{2\pi\eta}}\exp\left(-\frac{(x-g)^2}{2\eta}\right)dgdx$$

$$=\frac{1}{N(t)}\int_\mathbb{R}m^*(\gamma,g)\gamma(t,g)dg = \bar m^*(t),$$

equations (\ref{xbar}) and (\ref{G}) become

\begin{equation}\label{xbarfinal}
\dot{\bar x}=G\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right)+\nu\sqrt{\frac{G}{N}}\dot\xi_2,
\end{equation}

\begin{equation}\label{Gfinal}
\dot G=2G^2\left(\frac{\partial\bar m}{\partial G}-\overline{\frac{\partial m}{\partial G}}\right)+\mu-\frac{\nu^2G}{N}+\nu G\sqrt{\frac{2}{N}}\dot\xi_3.
\end{equation}

# A model of diffuse coevolution \label{coev}

## Formulation

In this section we demonstrate the use of our framework by formulating a model of diffuse coevolution across a guild of $S$ species whose interactions are mediated by resource competition along a single niche axis. Our approach mirrors closely the theory developed by MacArthur and Levins [-@Macarthur1967], Levins [-@9780691080628] and MacArthur [-@Arthur1969; -@MacArthur1970; -@9780691023823]. The most significant difference, aside from allowing evolution to occur, is the treatment of resource quality, which we replace with a model of abiotic stabilizing selection. A derivation is provided in SM \S\ref{diffuse}.

For species $i$ we inherit the above notation for trait value, distribution, average, variance, abundance, etc except with an $i$ in the subscript. Real world examples of niche axes include the body size of prey for lizard predators and the date of activity in a season for pollinators competing for floral resources. For mathematical convenience, we model the axis of resources by the real line $\mathbb{R}$. The value of a resouce along this axis is denoted by the symbol $\zeta$. For an individual in species $i$, we assume the resource utilization curve $u_i$ can be written as

\begin{equation}
u_i(\zeta,x_i)=\frac{U_i}{\sqrt{2\pi w_i}}\exp\left(-\frac{(x_i-\zeta)^2}{2w_i}\right).
\end{equation}

We further assume the niche center $x_i$ is normally distributed among individuals in species $i$, but the niche breadth $w_i$ and total niche utilization $U_i$ are constant across individuals in species $i$ and therefore cannot evolve. Suppose $\theta_i\in\mathbb R$ is the optimal location along the niche axis for species $i$ such that, in the absence of competition, individuals leave on average $W_i$ offspring when concentrated at $\theta_i$. We capture the rate by which the fitness falls as niche location $\zeta$ leaves the optimum $\theta_i$ by the parameter $A_i\geq0$. Hence, abiotic stabilizing selection along the resource axis can be modelled by the curve

\begin{equation}
e_i(\zeta)=W_i\exp\left(-\frac{A_i}{2}(\theta_i-\zeta)^2\right).
\end{equation}

The effect of abiotic stabilizing selection on the fitness for an individual of species $i$ with niche location $x_i$ is then given by

\begin{equation}
\int_\mathbb{R}e_i(\zeta)u_i(\zeta,x_i)d\zeta=\frac{W_iU_i}{\sqrt{A_iw_i+1}}\exp\left(-\frac{A_i}{2(A_iw_i+1)}(\theta_i-x_i)^2\right).
\end{equation}

To determine the potential for competition between individuals with niche locations $x_i$ and $x_j$, belonging to species $i$ and $j$ respectively, we compute the niche overlap

\begin{equation}
\mathcal O_{ij}(x_i,x_j)=\int_\mathbb{R}u_i(\zeta,x_i)u_j(\zeta,x_j)d\zeta=\frac{U_iU_j}{\sqrt{2\pi(w_i+w_j)}}\exp\left(-\frac{(x_i-x_j)^2}{2(w_i+w_j)}\right).
\end{equation}

Relating our treatment of the niche to classic competition theory [@michelloreau2010], the competition coefficient $\alpha_{ij}$ becomes a dynamical quantity that can be written as

\begin{equation}
\alpha_{ij}(t)=\frac{\int_{\mathbb R}\int_{\mathbb R}p_i(t,x_i)p_j(t,x_j)\mathcal O_{ij}(x_i,x_j) dx_idx_j}{\int_{\mathbb R}\int_{\mathbb R}p_i(t,x_i)p_i(t,x_i')\mathcal O_{ii}(x_i,x_i') dx_idx_i'}
\end{equation}

$$=\frac{U_j}{U_i}\sqrt{\frac{2(w_i+{\sigma_i}^2(t))}{w_i+w_j+{\sigma_i}^2(t)+{\sigma_j}^2(t)}}\exp\left(-\frac{(\bar x_i(t)-\bar x_j(t))^2}{2(w_i+w_j+{\sigma_i}^2(t)+{\sigma_j}^2(t))}\right).$$

In SM \S\ref{diffuse} we combine this niche model with equations (\ref{N}), (\ref{xbarfinal}) and (\ref{Gfinal}) to find

\begin{subequations}\label{comm_dynamics}
\begin{align}
\dot N_i = & \left\{r_i-\frac{a_i}{2}\Big((\bar x_i-\theta_i)^2+G_i+\eta_iN_i\Big) + c_i\left(\sqrt{\frac{b_{ii}}{2\pi}}-\sum_{j=1}^SN_j\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\right)\right\}N_i+\nu_i\sqrt{N_i}\dot\xi_1, \\
\dot{\bar x}_i = & \ a_iG_i(\theta_i-\bar x_i)-c_iG_i\left(\sum_{j=1}^SN_jb_{ij}(\bar x_j-\bar x_i)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\right)+\nu_i\sqrt{\frac{G_i}{N_i}}\dot\xi_2, \\
\dot G_i = & \ \mu_i-a_i{G_i}^2+c_i{G_i}^2\left(\sum_{j=1}^SN_jb_{ij}\left(1-b_{ij}(\bar x_i-\bar x_j)^2\right)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\right)-\frac{\nu_i^2G_i}{N_i}+\nu_iG_i\sqrt{\frac{2}{N_i}}\dot\xi_3,
\end{align}
\end{subequations}

where

\begin{subequations}
\begin{align}
r_i = & \ \ln \left(\frac{W_iU_i}{\sqrt{1+A_iw_i}}\right), \\
a_i = & \ \frac{A_i}{1+A_iw_i}, \\
b_{ij}(t) = b_{ji}(t) = & \ \frac{1}{w_i+w_j+\sigma_i^2(t)+\sigma_j^2(t)}, \\
c_i \geq & \ 0.
\end{align}
\end{subequations}

### The Strength of Pairwise Coevolution

In Week and Nuismer [-@Week2019], the strength of pairwise coevolution was defined in terms of the offset-matching model, given by the difference equation

\begin{equation}
\begin{matrix}
\Delta\bar x_1(t) = G_1A_1(\theta_1-\bar x_1(t))+G_1B_1(\bar x_2(t) +\delta_1-\bar x_1(t)) + \frac{G_1}{N_1}\xi_1(t) \\
\Delta\bar x_2(t) = G_2A_2(\theta_2-\bar x_2(t))+G_2B_2(\bar x_1(t) +\delta_2-\bar x_2(t)) + \frac{G_2}{N_2}\xi_2(t)
\end{matrix}
\end{equation}

where $t=0,1,2,\dots$, $A_i$ denotes the strength of abiotic selection, $\theta_i$ is the abiotic phenotypic optimum, $B_i$ is the strength of biotic selection, $\delta_i$ is the optimal offset and $\xi_i$ is a sequence of independent unit normal random variables for species $i$. With respect to the offset-matching model, the strength of pairwise coevolution was defined as $\sqrt{B_iB_j}$. Motivated by this definition, we define the strength of pairwise coevolution between species $i$ and $j$ with respect to our model of diffuse coevolution by

\begin{equation}
\mathfrak{C}_{ij}=\sqrt{\frac{c_ic_jG_iG_jN_iN_jb_{ij}b_{ji}\sqrt{b_{ij}b_{ji}}}{2\pi}}\exp\left(-\frac{b_{ij}+b_{ji}}{4}(\bar x_j-\bar x_i)^2\right)
\end{equation}

$$=b_{ij}\sqrt{\frac{c_ic_jG_iG_jN_iN_jb_{ij}}{2\pi}}\exp\left(-\frac{b_{ij}}{2}(\bar x_j-\bar x_i)^2\right).$$

This fundamentally differs from the definition given in Week and Nuismer [-@Week2019] by including the additive genetic variance of both species, which both must be positive numbers for coevolution to occur.

## Community dynamics

For the sake of illustration we numerically integrated system (\ref{comm_dynamics}) for a richness of $S=50$ species under two scenarios. For the first scenario of weak competition we randomly chose $c_i\approx$\texttt{0.0} for each $i=1,\dots,S$ and for the second scenario of moderately strong competition we randomly chose $c_i\ll$\texttt{1.0} for each $i=1,\dots,S$. Background parameters are drawn independently for each species from a common distribution. Specifically, we set model parameters according to Table \ref{par_vals}.

\begin{table}[H]
\centering\caption{Distributions of model parameters under the scenarios of weak and moderate competition. The index $i$ runs through all $1,\dots, S$ species.}\vspace{0.2cm}
\begin{tabular}{l|l|l}
Parameter   & Weak Competition             & Moderate Competition     \\ \hline
$r_i$       & $\mathrm{Unif}(\mathtt{0.0},\ \mathtt{1.0})$   & \text{same}        \\
$\theta_i$  & $\mathrm{Unif}(\mathtt{-1.0},\ \mathtt{1.0})$  & \text{same}        \\
$a_i$       & $\mathrm{Exp}(\mathtt{0.01})$         & \text{same}        \\
$c_i$       & $\mathrm{Exp}(\mathtt{0.01})$         & $\mathrm{Exp}(\mathtt{0.1})$ \\
$w_i$       & $\mathrm{Unif}(\mathtt{0.01},\ \mathtt{1.0})$  & \text{same}        \\
$U_i$       & $\mathrm{Unif}(\mathtt{0.01},\ \mathtt{1.0})$  & \text{same}        \\
$\eta_i$    & $\mathrm{Unif}(\mathtt{1.0e-5}, \ \mathtt{1.0e-3})$         & \text{same}        \\
$\mu_i$     & $\mathrm{Exp}(\mathtt{0.01})$         & \text{same}        \\
$\nu_i$     & $\mathrm{Exp}(\mathtt{0.01})$         & \text{same}
\end{tabular}\label{par_vals}
\end{table}

With these two sets of model parameters, we simulated our model for \texttt{1.0e3} units of time. Whether or not the units of time reflect years, months or even hours depends on the life cycles of the species involved. For both scenarios, we initialized the abundances to $N_i=$\texttt{5.0e3}, trait means to $\bar x_i=$\texttt{0.0} and heritability to $h^2_i=$\texttt{0.5} for each $i=1,\dots,S$. Temporal dynamics for each scenario are provided in Figure \ref{temporal}. Interaction networks and the distribution of coevolution at $t=$\texttt{1e3} for each scenario are displayed in Figure \ref{net}.

```{r, message=FALSE, echo=F, fig.align='center', fig.cap='\\label{temporal}Temporal dynamics of mean trait (top), heritability (middle) and abundance (bottom) for the scenario of weak competition (left) and strong competition (right). Red lines indicate average trend across species.'}
knitr::include_graphics('community_dynamics.png')
```

```{r, echo=F, fig.align='center', fig.cap='\\label{net}Networks of interspecific interactions (top row) and distributions of pairwise coevolution (bottom row) under the scenarios of weak competition (left) and moderate competition (right) at time $t=$\\texttt{1.0e3}. Node sizes are proportional to population sizes. Edge widths and shade are monotonically increasing functions of pairwise coevolution.'}
knitr::include_graphics('interaction_networks.png')
```

# Tying theory to data \label{stats}

The focus of this paper thusfar has been centered on the introduction of white noise calculus and its use in deriving stochastic models of evolutionary ecology. However, this stochasticity is meant to reflect the intrinsically random nature of populations and does not account for the additional layer of sampling effects that blur empirical observations. In this section we introduce a simple model that mimics the sampling of ecological communities conducted by field ecologists. This sampling model provides the crucial mapping that connects theoretical results to real world data and hence is key to the creation of statistical methods. For the sake of clarity, we develop this model with respect to our model of diffuse coevolution. We show that the cross-sectional data of trait distributions, abundance distributions and interaction networks provide only partial information for directly inferring model parameters from equations relating these data. The remaining parameters must be inferred indirectly by numerically integrating the model. We therefore split this section into two subsections. The first covers parameters that can be inferred directly from cross-sectional data, which happen to be niche breadth and relative niche use. The second outlines a numerical approach to approximating the posterior distribution of the remaining model parameters.

## Inference of niche parameters

We treat estimates of trait distributions, abundances, and interaction counts independently. To model the collection of abundance data in the field, we assume individuals are sampled at random from the entire community so that a randomly drawn individual will belong to species $i$ with probability $s_i=N_i/\sum_jN_j$. Traits are assumed to be drawn independently. For each species $i$, a sample of size $K_i$ is drawn from a normal distribution with mean $\bar x_i$ and variance ${\sigma_i}^2=G_i+\eta_i$. Examples of trait and abundance distributions sampled at $t=$\texttt{1.0e3} from the community depicted in Figure \ref{temporal} are presented in Figure \ref{smpl_dists}.

```{r, echo=F, fig.align='center', fig.cap='\\label{smpl_dists}Histograms of mean trait (top row) and abundance distributions (bottom row) drawn from the communities in Figure \\ref{temporal} at time $t=$\\texttt{1.0e3} under weak competition (left column) and moderate competition (right column) for a sample size of \\texttt{1e3}.'}
knitr::include_graphics('distributions.png')
```

We assume the rate of interactions between species $i$ and $j$ is given by their total niche overlap

\begin{equation}
R_{ij}=\int_\mathbb{R}\int_\mathbb{R}\mathcal{O}_{ij}(x_i,x_j)n_i(x_i)n_j(x_j)dx_idx_j=N_iN_jU_iU_j\sqrt{\frac{b_{ij}}{2\pi}}\exp\left(-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2\right).
\end{equation}

Ignoring intraspecific interactions, the relative rate of interaction between species $i$ and $j$ is defined as

\begin{equation}
\rho_{ij}=\frac{R_{ij}}{\sum_{k=1}^S\sum_{l>k}R_{kl}}.
\end{equation}

We assume $\rho_{ij}$ is also the probability that, given an interaction is observed, the species involved are $i$ and $j$. Hence, observed interactions are modelled as being drawn independently from the same distribution defined by $\rho$. In turn, a randomly drawn sample of interactions provides estimates of the relative rates of interactions $\hat\rho_{ij}$.

<!--

Examples of empirical networks predicted by our model are displayed in Figure \ref{sample_net}.

```{r, echo=F, fig.align='center', fig.cap='\\label{sample_net}Sample networks of interspecific interactions drawn from the networks found in Figure \\ref{net} for a sample size of \\texttt{1e3}.'}
knitr::include_graphics('sample_networks.png')
```

-->

With the estimates $\hat\rho_{ij}$, $\hat s_i$, $\hat{\bar x}_i$ and ${\hat\sigma_i}^2$, we can directly calculate estimates for the niche breadth $w_i$ and the relative niche use $v_i=U_i/\sum_lU_l$ for each species $i$ under consideration. In particular, we define the function

\begin{equation}\label{rho_est}
\tilde\rho_{ij}(w_1,\dots,w_S,v_1,\dots,v_S)=\frac{\tilde{R}_{ij}}{\sum_{k=1}^S\sum_{l>k}\tilde{R}_{kl}},
\end{equation}

where

\begin{equation}
\tilde{R}_{ij}=\hat s_i\hat s_jv_iv_j\sqrt{\frac{\hat b_{ij}}{2\pi}}\exp\left(-\frac{\hat b_{ij}}{2}(\hat {\bar x}_j-\hat {\bar x}_i)^2\right),
\end{equation}

\begin{equation}
\hat b_{ij}=\hat b_{ji}=\frac{1}{w_i+w_j+{\hat \sigma_i}^2+{\hat \sigma_j}^2}.
\end{equation}

Note that the absolute total niche use $U_i$ cannot be inferred from interactions rates since the sum $\Omega=\sum_iU_i$ disappears from the quotient defining $\rho_{ij}$. However, the parameter $\Omega$ scales the amount of niche overlap between species which then dictates the strength of biotic selection. Hence, $\Omega$ will need to be inferred later when we develop an approach to inferring remaining model parameters.

The estimates $\hat w_i$ and $\hat v_i$ can then be found by minimizing the Kullback-Liebler divergence from $\hat\rho$ to $\tilde\rho$, given by

\begin{equation}
\mathcal{D}_{\mathrm{KL}}(\hat\rho\|\tilde\rho)=\sum_{i=1}^S\sum_{j>i}\hat\rho_{ij}\ln\left(\frac{\hat\rho_{ij}}{\tilde\rho_{ij}}\right),
\end{equation}

across all $w_1,\dots,w_S,v_1,\dots,v_S>0$ with $\sum_iv_i=1$. Since $\tilde\rho_{ij}>0$ for all $w_1,\dots,w_S,v_1,\dots,v_S>0$ with $\sum_iv_i=1$ and $i,j=1,\dots,S$, $i\neq j$, the quantity $\mathcal{D}_{\mathrm{KL}}(\hat\rho\|\tilde\rho)$ is well-defined under the convention $0\ln(0)=0$. However, since we may observe $\hat\rho_{ij}=0$ for some $i,j$, the quantity $\mathcal{D}_{\mathrm{KL}}(\tilde\rho\|\hat\rho)$ is not well-defined and thus we exclude it from our objective function.

<!--

either a metric, such as an $l^p$ metric defined by

\begin{equation}
\|\hat\rho-\tilde\rho\|_p=\left(\sum_{i=1}^S\sum_{j>i}|\hat\rho_{ij}-\tilde\rho_{ij}|^p\right)^{1/p}
\end{equation}

with $p\geq1$, or other measures such as

-->

From a purely algebraic perspective, we note that inference of the $w_i$ and $v_i$ using the general approach outlined above constrains the minimal number of species included in the community. In particular, with $S$ species, the above method attempts to infer $2S$ parameters from $S(S-1)/2$ relationships (which is just the number of interspecific interactions possible between $S$ species) given by minimizing <!-- either $\|\hat\rho-\tilde\rho\|_p$ or --> $\mathcal{D}_{KL}(\hat\rho\|\tilde\rho)$. Hence, for inferring all the $w_i$ and $v_i$, we require $S\geq5$. However, by assuming ${\sigma_i}^2\gg w_i$ so that ${\sigma_i}^2+w_i\approx{\sigma_i}^2$, we can form the approximation

\begin{equation}
b_{ij}\approx\frac{1}{{\sigma_i}^2+{\sigma_j}^2}
\end{equation}

to relax the constraint on species richness to $S\geq3$. Biologically, this assumes that individuals are far more specialized in their niche-use than the species they belong to. For example, conspecifics may focus solely on prey body sizes within very narrow ranges that have little overlap as a result of intraspecific competition. While certainly not the rule, support for this assumption has been produced from previous studies [@Bolnick2003 and references therein].

Mathematically, this approximation reduces the number of parameters inferred to $S$ and thereby provides an enormous simplification by halving the dimension of the inference problem. Given the ubiquitous curse of dimensionality among statistical methods and their numerical implementations [@Wainwright2019; @Vershynin2018; @1701.02434], any opportunity to reduce the dimensionality of a problem without making a fatal sacrifice to its empirical relevance is worth taking, at least as a first step. Furthermore, by omitting the $w_i$'s one can transform the system defined by equation (\ref{rho_est}) into a linear system with respect to the $v_i$'s. Because of these benefits, for the remainder of this paper, we adopt the assumption ${\sigma_i}^2\gg w_i$.

### Implementation

As a proof of concept, we implemented this approach in the programming language _Julia_ [Version 1.3.0, @bezanson2017julia]. Code for this solution is available at \texttt{\url{https://github.com/bobweek/WNEE}}. Setting $S=20$, we integrated the model for $\tau=$\texttt{1e6} units of time to compute the simulated interaction distribution $\rho$. We drew \texttt{1e4} interactions at random weighted by $\rho$ to calculate $\hat\rho$, \texttt{100} traits for each species from normal distributions defined by $\bar x_i$ and ${\sigma_i}^2$ to calculate $\hat{\bar x}_i$ and ${\hat\sigma_i}^2$, and \texttt{1e3} individuals from across the community, weighted by $s$, to estimate the relative abundances $\hat s$. Scatter plots of estimated relative niche-use parameters against specified parameters for the two scenarios of weak and moderate competition are shown in Figure \ref{sctr}.

```{r, include=F, eval=F}
require("ggplot2")
require("gridExtra")
ecol_reg1 = read.csv("/home/bob/Research/WNEE/ecol_reg1.csv")
ecol_reg2 = read.csv("/home/bob/Research/WNEE/ecol_reg2.csv")
ecol_plt1 = ggplot(ecol_reg1)+geom_point(aes(x=actual,y=inferred))+xlim(0,.11)+ylim(0,.11)+
  geom_abline(intercept=0,slope=1,linetype="dashed")+theme_minimal()+xlab("Specified v")+ylab("Estimated v")+
  ggtitle("Weak Competition")+theme(plot.title = element_text(hjust = 0.5))
ecol_plt2 = ggplot(ecol_reg2)+geom_point(aes(x=actual,y=inferred))+xlim(0,.11)+ylim(0,.11)+
  geom_abline(intercept=0,slope=1,linetype="dashed")+theme_minimal()+xlab("Specified v")+ylab("")+
  ggtitle("Moderate Competition")+theme(plot.title = element_text(hjust = 0.5))
ecol_plt = grid.arrange(ecol_plt1,ecol_plt2,nrow=1)
ggsave("~/Research/White Noise/ecol_plt.png",ecol_plt,width=5,height=2.5)
```

```{r, echo=F, fig.align='center', fig.cap='\\label{sctr} Two examples of inferring species specific relative niche use parameters. Dashed lines with zero-intercept and unit-slope are provided for reference.'}
knitr::include_graphics('ecol_plt.png')
```

The results found in Figure \ref{sctr} demonstrate that our method can accurately infer species specific niche-use parameters. However, although this method provides important information about ecological parameters, it is insufficient to measure the distribution of coevolution. In the following section we outline a numerical method to calculate the likelihood of our model and a Markov Chain Monte Carlo (MCMC) based on this method to estimate posterior distributions of the remaining model parameters. Remaining parameters include absolute niche use scaling $\Omega$, intrinsic rates of growth $r_i$, phenotypic optima $\theta_i$, strengths of abiotic selection $a_i$, strengths of competition $c_i$, segregation variances $\eta_i$, mutation rates $\mu_i$ and magnitudes of demographic stochasticity $\nu_i$ for $i=1,\dots,S$.

## Inference of remaining parameters

The direct inference of niche use for each species depends on the explicit relationships between these parameter values and the collected data predicted by our model. However, assuming the community has reached approximate statistical equilibrium with respect to our model, we demonstrate that remaining model parameters can be inferred from the distributions of mean traits, trait variances and relative abundances. In principle one could derive the equilibrium distribution of system (\ref{comm_dynamics}) as the likelihood of the remaining model parameters given the data in order to harness the latent statistical power of our model. However, given the non-linearity and high dimensionality of system (\ref{comm_dynamics}) even for modestly sized communities, deriving an analytical expression for the likelihood surface is at best difficult, but likely impossiple. Instead, we take an alternative approach to approximate this likelihood surface. This approach makes use of the property that the temporal fluctuations of system (\ref{comm_dynamics}) will converge in distribution to the equilibrium distribution across states. Viewing the equilibrium distribution of (\ref{comm_dynamics}) as a function of model parameters conditioned on the state of the system, we see this distribution coincides with the likelihood surface mentioned above. Hence, this approach provides a numerical solution for approximating the likelihood surface.

To implement this approach, for each set of model parameters, we numerically integrated the deterministic component of system (\ref{comm_dynamics}) for $\tau=$\texttt{1.0e6} units of time initialized with $\bar x_i(0)=\theta_i$, $\ln{\sigma_i}^2(0)\sim\mathrm{Unif}(\mathtt{0.0},\mathtt{2.0})$ and $\ln N_i(0)\sim\mathrm{Unif}(\mathtt{3.0},\mathtt{10.0})$. We denote the solution to the deterministic system by the $\mathbb{R}^{3S}$-valued function $X(t)$, $0\leq t\leq \tau$ with components given by

\begin{equation}
X_i(t) = \left\{\begin{matrix}
\bar x_i(t) , & 1\leq i\leq S \\
{\sigma_i}^2(t), & (S+1)\leq i \leq 2S \\
N_i(t), & (2S+1)\leq i\leq 3S.
\end{matrix}\right.
\end{equation}

We checked for equilibrium by calculating the greatest difference

\begin{equation}
\delta=\max_{i=1,\dots,3S}\Big(X_i(\tau)-X_i(\tau-\Delta\tau)\Big)
\end{equation}

among state variables between $\Delta\tau=$\texttt{100.0} units of time. If $\delta<\varepsilon=$\texttt{1.0e-3} we assumed $X(\tau)$ approximates the deterministic equilibrium, which in turn we assumed to also approximate the expected state at equilibrium of the stochastic system (\ref{comm_dynamics}). Keeping the same set of parameters, we numerically integrated system (\ref{comm_dynamics}) using $X(\tau)$ as the initial condition for $\tau^*=$\texttt{1.0e3} units of time. Denoting this numerical solution by $X^*(t)$, $\tau\leq t\leq\tau+\tau^*$, we assume the distribution of the random variable defined by $X^*(T)$, $T\sim\mathrm{Unif}(\tau,\tau+\tau^*)$, approximates the equilibrium distribution of system (\ref{comm_dynamics}) and therefore approximates the likelihood of our model. However, using the sample path of $X^*$ to compute likelihood requires a method to aggregate the simulated time-series into either a smooth density function using kernel density estimation or binned discretely into a histogram. Either of these objects become costly to compute in high dimensions. We therefore utlize a pseudo-likelihood approximation. In particular, denoting $x$ the $3S$-dimensional vector of state variables and $\Theta$ the $(7S+1)$-dimensional vector of remaining model parameters, we compute (normalized) histograms $H_i(x|\Theta)$ of the marginal equilibrium distributions for each component $X^*_i$ and approximate likelihood as $L(\Theta|x)\approx\prod_iH_i(x|\Theta)$.

Since our model assumes patterns of interaction are completely described by niche use and breadth parameters along with phenotype and abundance data, the interaction data provides no further information to infer evolutionary parameters. This leaves $7S+1$ parameters to infer from $3S$ data points. Unless the parameter values exhibit significant interdependence, we cannot expect the $3S$ data points to uniquely define the values of the $7S+1$ parameters. We overcome this dilemma by assuming these evolutionary parameters are equivalent across species. That is, we assume $\mu=\mu_1=\mu_2=\dots$, $\theta=\theta_1=\theta_2=\dots,$ and so on. This simplifying assumption leaves just eight parameters to infer. With an approximation of likelihood in hand we develop a Bayesian approach to infer posterior distributions of the eight remaining model parameters. In particular, we implement an MCMC approach to sample values from the posteriors.

Our MCMC approach utilizes the Metropolis-Hastings algorithm to exlore parameter space. For parameters taking positive values, we set the jump distributions to log-normal with mean zero and variance \texttt{1.0}. For $\theta$, the only parameter that can be any real number, we set the jump distribution to a standard normal distribution. We assumed each of these jump distributions to be independent of the others.

### Implementation

To demonstrate our approach to inference we implemented a solution using the programming language _Julia_ [Version 1.3.0, @bezanson2017julia]. Code for this solution is available along with code for inferring niche-use parameters at \texttt{\url{https://github.com/bobweek/WNEE}}.

# Conclusion \label{conclusion}

We have introduced an approach to derive models of evolutionary ecology using the calculus of white noise, demonstrated our approach by deriving a model of diffuse coevolution and outlined methods of statistical inference based on this model. Although our approach has the merit of rigorously synthesizing the dynamics of abundance and phenotypic evolution of populations, there remains nuance details and important limitations that have yet to be confronted for gaining a more thorough and rigorous understanding of ecological communities. We touch on just five of them here.

__Limitations of diffusion limits__

As noted by Feller [-@feller1951], although diffusion limits provide reasonable approximations for large populations, relatively small populations require discrete models. Hence, as a diffusion limit, SPDE (\ref{SPDE}) restricts parameter values to regions that maintain large population sizes. This puts an important restriction on any model developed under this framework by implying populations are not at risk of extinction. Hence, studies of evolutionary rescue and colonization-extinction dynamics must be pursued via a different approach. A natural alternative can be developed utilizing the underlying branching Brownian motion process that SPDE (\ref{SPDE}) is a diffusion limit of [@Perkins1992; @Mlard1993; @zhao1997interacting; @xuelei1997absolute; @zeng1998absolute]. This process explicitly tracks individuals throughout their life-history and captures reproduction as branching events and therefore models population size $N$ as a non-negative integer instead of a continuously varying positive number. In particular, there is no restriction that prevents $N=0$. We avoided use of this process here because the resulting models are less analytically tractable and numerical implementations become computationally expensive for large $N$.

__The genetic architecture and distributions of traits__

Our model of coevolution rests on the assumption of normally distributed traits, which is not independent of our simplistic model of the genetic architecture of traits. Real traits are not encoded by an infinite number of loci each contributing an additive infinitesimal effect and mutations are not inherited as normally distributed deviations from parental breeding values. Departures from this model of genetic architecture can produce more complex distributions of breeding values and expressed traits. Such deviations can be reinforced via strong non-Gaussian selection surfaces, including the surface $m(n,x)$ we have derived from niche theory, along with non-random mating. Gaussian approximations are convenient since they are defined by their mean and variance and for their tractable algebraic structure (infinite divisibility). Hence, future work investigating the effects of non-normally distributed traits on the structure and dynamics of ecological communities will need to confront higher moments such as skew and kurtosis, ideally integrating previously established mathematical approaches to derive the dynamics of higher moments [@Dbarre2015EvolutionOQ].

__The relation between niche overlap and competition__

Past work has demonstrated several limitations in measuring competition via niche overlap [@Abrams1980; @Holt1987]. In particular, Abrams [-@Abrams1980] argues the MacArthur-Levins framework utilized here is unnacceptable when nothing is known about the biological mechanism mediating competition. Hence, we must stress that our approach should not be blindly applied to arbitrary communities, but that a great deal of work interpreting our model with respect to the system is a prerequisite task. For example, in the case of a sub-alpine flowering plant community, the niche axis can be interpreted as day-of-year (under a suitable transformation). Our model then assumes within-season fluctuations in the ensemble of abiotic factors including temperature, precipitation, soil nutrients and pollinator activity induces an emergent abiotic optimal date that maximizes fitness in the absence of competition. This optimum likely differs across species and potentially carries a phylogenetic signal, but for the sake of parsimony we have assumed it to be constant across these sub-alpine flowers. Given estimates of these parameters through, say, experimental measurements, the model generalizes in an obvious way ($\theta_i\neq\theta_j$) to account for such details.

Another component of Abrams' critique is the distinction between niche overlap and measurements of competition. We also agree that this is an important distinction and is explicitly captured in our model by decomposing the effect of competition on fitness as the product of niche overlap and a number $c>0$ which we have dubbed the strength of competition. According to our model the niche overlap between each pair of species in the community differs (with probability one). However, biological intuition suggests the strength of competition $c$ also differs across species pairs. Hence, guild-wise constancy of $c$ is also a parsimonious assumption. The natural generalization is to assume the strength of competition between species $i$ and $j$ can be captured by the pairwise specific parameter $c_{ij}>0$. Allowing $c_{ij}\neq c_{ji}$ then captures asymmetries in the sensitivities of species' fitness to competition along this resource gradient. Pushing this idea further, if $y_i$ represents some other trait in species $i$, potentially morphological (flower shape) or physiological (water use efficiency), determining the outcome of indirect interspecific interactions between competitors (pollinator preference or nutrient availability) we can model the strength of competition between species $i$ and $j$ as a function $f$ of the distribution of this trait in each species. If this trait correlates with niche centers across individuals within species, or if resource utilization curves $u_i$ depend on $y_i$, or if $f$ depends on niche centers, the effect on fitness due to competition can no longer be decomposed multiplicatively into distinct components. Hence, our model also requires the necessary assumptions for this decomposition to hold, which vary from system to system.

In another critique by Holt [-@Holt1987], it is argued that utilizations of qualitatively distinct resources may have non-multiplicative effects on components of fitness. This consideration has also been incorporated into our model. In our derivation of fitness in SM \S\ref{diffuse}, we show that distinct components of the growth rate of populations can be decomposed additively under certain circumstances, though further work is needed to develop a general roadmap for when components of fitness can be decomposed additively or multiplicatively.

__The role of ecological stoichiometry__

Our treatment of both biotic and abiotic selection neglects important chemical constraints of biological reality. For instance, the resource we assume species are competing over is modelled as a static quantity. However, real resources can be dynamic quantities. Such theoretical quantities may reflect abiotic cycles of material and energy or whole trophic layers comprised of prey, hosts or mutualists. Although resource dynamics have been captured theoretically in consumer-resource models [@davidtilman1982], developing a more realistic model of resource competition must incorporate details on the ecophysiology of organisms [@michelloreau2010]. Doing so will help clarify the feedback between the evolution of populations and the ecosystem processes they are a part of.

For example, in the context of plant-pollinator communities, the role of nitrogen mediating interspecific interactions has been reviewed by David, Storkey and Stevens [-@David2019] and the evolutionary ecology of the nutritional content of nectar has been reviewed by Parachnowitsch, Manson and Sletvold [-@Parachnowitsch2018]. These studies demonstrate the need for further research to understand how soil nutrient availability and organismal ecophysiology structures communities of plants and pollinators. Theoretical pursuits in this directions may profit from interfacing the framework we have outlined here with population-ecosystem models such as that developed by Fridley [-@Fridley2017].

__Challenges in statistical inference__

The statistical methods developed above depend on a long list of assumptions, both biological and technical. For example, we significantly reduced the dimensionality of the inference problem by assuming generalist populations of specialist individuals. While there is evidence for this assumption in some populations [@Bolnick2003], it may not be a general feature. Furthermore, our symmetry assumptions in mutation rates, abiotic optima and so on were made purely for the sake of statistical tractability. Relaxing these assumptions would allow for a more realistic treatment of populations at the cost of increased dimensionality of parameter space. However, even with these assumptions in place, precise calculations of likelihood become difficult for species rich communities due to increased dimensionality of state space. There are two distinct repurcussions of richness. Firstly, our method to calculate likelihood is based on simulating the equilibrium distribution of state variables. Since the length of the calculation that updates state variables increases quadratically with species richness and since high dimensional processes take longer to converge, this approach becomes costly for communities of many species. Secondly, computing likelihood from the simulated time-series of state variables requires approximation of the probability density function for the system using either kernel density estimation or a histogram. However, high dimensional density estimates and histograms are computationally prohibitive. Thus, although our approach to parameter estimation is theoretically possible, future work will need to confront the challenges involved with high-dimensional statistics [@Wainwright2019].

__Final remarks__

The reality of ecological communities is immensely complex. Although top-down approaches to community ecology, such as the Maximum Entropy Theory of Ecology [@Harte2011], have enjoyed some success in describing community-level patterns [@Harte2014; @Xiao2015], a mechanistic understanding of why these patterns emerge and how they will change remains a formidable task. Such an understanding must take both bottom-up and top-down approaches integrating considerations from the ecophysiology of individual organisms that reveal the economics of interspecific interactions [@Sterner2008], to the phylogeographic history of taxa that sets the stage for contemporary dynamics [@Hickerson2010]. Through connecting these dots we can increase the variance explained in observations of ecological communities by specific mechanisms and come closer to a much needed predictive theory of evolutionary ecology. Despite the long list of equations derived in this paper, this work takes just one small step towards capturing these many details. However, we hope the theoretical framework outlined here, along with the demonstration of its use in modelling competitive communities and inferring model parameters, provides some helpful tools to aid quantitative evolutionary ecologists in reaching such lofty goals.

\newpage

# Supplementary material (SM)

## The relation between diffusion and convolution with a Gaussian kernel \label{diffconvequiv}

Let $g:\mathbb{R}^d\to\mathbb{R}$ be smooth and consider the deterministic Cauchy Problem

\begin{equation}\label{heateqn}
\left\{\begin{matrix}
\dot f(t,x)=&\nabla^2 f(t,x), & (t,x)\in(0,\infty)\times\mathbb{R}^d\\
f(t,x)=&g(x), & (t,x)\in\{0\}\times\mathbb{R}^d.
\end{matrix}\right.  \tag{SM1.1}
\end{equation}

According to Evans [-@lawrenceevans2010], the fundamental solution of (\ref{heateqn}) is

\begin{equation}
\Phi(t,x)=\frac{1}{(4\pi t)^{d/2}}\exp\left(-\frac{|x|^2}{4t}\right), \ (t,x)\in(0,\infty)\times\mathbb{R}^d, \tag{SM1.2}
\end{equation}

where $|x|=\sqrt{\sum_ix_i^2}$. The solution $f(t,x)$ of PDE (\ref{heateqn}) is then given by the convolution

\begin{equation}
f(t,x)=\int_{\mathbb{R}^d}\Phi(t,x-y)g(y)dy, \ (t,x)\in(0,\infty)\times\mathbb{R}^d. \tag{SM1.3}
\end{equation}

Hence, by the fundamental theorem of calculus,

\begin{equation}
f(t,x)+\int_t^{t+1}\dot f(s,x)ds=f(t+1,x) \tag{SM1.4}
\end{equation}

$$=\int_{\mathbb{R}^d}\Phi(t+1,x-y)g(y)dy=\int_{\mathbb{R}^d}\int_{\mathbb{R}^d}\Phi(1,x-y)\Phi(t,y-z)g(z)dzdy$$

$$=\int_{\mathbb{R}^d}\Phi(1,x-y)f(t,y)dy.$$

In particular,

\begin{equation}
f(t,x)+\int_t^{t+1}\nabla^2f(s,x)ds=\int_{\mathbb{R}^d}\Phi(1,x-y)f(t,y)dy. \tag{SM1.5}
\end{equation}

## Deterministic dynamics of $\sigma^2(t)$ \label{var_deriv}

Picking up from the main text \S2.1,

\begin{equation}
\dot{\sigma}^2(t)=\frac{d}{dt}\int_\mathbb{R}(x-\bar x(t))^2p(t,x)dx
\end{equation}

$$=\int_\mathbb{R}2(x-\bar x(t))\dot{\bar x}(t)+(x-\bar x(t))^2\dot p(t,x)dx$$

$$=\int_\mathbb{R}(x-\bar x(t))^2\left((m(n,x)-\bar m(t))p(t,x)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(t,x)\right)dx$$

$$=\int_\mathbb{R}\left((x-\bar x(t))^2-\sigma^2(t)+\sigma^2(t)\right)(m(n,x)-\bar m(t))p(t,x)+(x-\bar x(t))^2\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(t,x)dx$$

$$=\mathrm{Cov}_p\Big((x-\bar x(t))^2,m(n,x)\Big)+\frac{\mu}{2}\int_\mathbb{R}(x-\bar x(t))^2\frac{\partial^2}{\partial x^2}p(t,x)dx.$$

In particular, when $p(t,x)$ is Gaussian,

\begin{equation}
\frac{\partial^2}{\partial x^2}p(t,x)=\frac{(x-\bar x(t))^2-\sigma^2(t)}{\sigma^4(t)}p(t,x)
\end{equation}

and hence, referring to the well-known moments of the Gaussian distribution, we find

\begin{equation}
\int_\mathbb{R}(x-\bar x(t))^2\frac{\partial^2}{\partial x^2}p(t,x)dx
\end{equation}

$$=\int_\mathbb{R}\frac{(x-\bar x(t))^4-(x-\bar x(t))^2\sigma^2(t)}{\sigma^4(t)}p(t,x)dx=2.$$

In the Gaussian case, we also have

\begin{equation}
2\sigma^4\left(\frac{\partial\bar m}{\partial\sigma^2}-\overline{\frac{\partial m}{\partial\sigma^2}}\right)=2\sigma^4\left(\frac{\partial}{\partial\sigma^2}\int_\mathbb{R}m(n,x)p(t,x)dx-\int_\mathbb{R}p(t,x)\frac{\partial}{\partial\sigma^2}m(n,x)dx\right)
\end{equation}

$$=2\sigma^4\int_\mathbb{R}\frac{(x-\bar x)^2-\sigma^2}{2\sigma^4}m(n,x)p(t,x)dx=\int_\mathbb{R}\left((x-\bar x)^2-\sigma^2\right)\big(m(n,x)-\bar m\big)p(t,x)dx$$

$$=\mathrm{Cov}_p\Big((x-\bar x)^2,m\Big).$$

## Solutions to SPDE (\ref{SPDE}) have finite moments \label{rapid_decay}

In Li [-@zeng1998absolute] solutions to a generalization of SPDE (\ref{SPDE}) were defined by

\begin{equation}
\int_{\mathbb{R}}(n(t,x)-n(0,x))f(x)dx=\int_0^t\int_\mathbb{R}n(s,x)\left(m(n,x)f(x)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}f(x)\right)dxds
\end{equation}

$$+\int_0^t\int_\mathbb{R}\sqrt{V(n,x)n(s,x)}f(x)\dot\xi(s,x)dxds$$

for any bounded and continuous $f$. This conflicts with our application since polynomials are not bounded. However, one can use bounded functions to approximate polynomials. In particular, for any positive integers $k,l$, we can approximate $x^k$ via

\begin{equation}
\alpha_l(x)=\left\{\begin{matrix}
x^k , & |x| <l \\
l^k , & x > l \\
(-l)^k , & x<-l .
\end{matrix}\right.
\end{equation}

Hence, $\alpha_l(x)$ is continous and bounded and $\lim_{l\to\infty}\alpha_l(x)=x^k$. In addition, we have $|\alpha_l(x)|\leq|\alpha_{l+1}(x)|$. Then, by monotone convergence, $\int_0^\infty x^kn(t,x)dx<\infty$ and, when $k$ is even, $\int_{-\infty}^0x^kn(t,x)dx<\infty$ for any $t\geq0$. When $k$ is odd we apply the same argument to $-\alpha_l(x)$ to obtain $-\int_{-\infty}^0x^kn(t,x)dx<\infty$ for any $t\geq0$. Thus,

\begin{equation}
\left|\int_\mathbb{R}x^kn(t,x)dx\right|=\left|\int_0^\infty x^kn(t,x)dx+\int_{-\infty}^0x^kn(t,x)dx\right|<\infty
\end{equation}

for any positive integer $k$ and $t\geq0$. Linearity of integration then provides $|\int_\mathbb{R}h(x)n(t,x)dx|<\infty$ for any polynomial $h(x)$ and $t\geq0$. Since $\frac{\partial^2}{\partial x^2}h(x)$ and $h^2(x)$ are polynomials and assuming $m(n,x)$ and $V(n,x)$ are bounded and continuous in both $n$ and $x$, we can use the same logic to justify

\begin{equation}
\left|\int_\mathbb{R}n(t,x)m(n,x)h(x)dx\right|,\left|\int n(t,x)\frac{\mu}{2}\frac{\partial^2}{\partial x^2}h(x)dx\right|,\left|\int_\mathbb{R}V(n,x)n(t,x)h^2(x)\right|<\infty
\end{equation}

for any $t\geq0$. Hence,

\begin{equation}\label{spde_ext}
\int_{\mathbb{R}}(n(t,x)-n(0,x))h(x)dx=\int_0^t\int_\mathbb{R}n(s,x)\left(m(n,x)h(x)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}h(x)\right)dxds
\end{equation}

$$+\int_0^t\int_\mathbb{R}\sqrt{V(n,x)n(s,x)}h(x)\dot\xi(s,x)dxds$$

for any polynomial $h(x)$ and $t\geq0$. By the same logic we see equation (\ref{spde_ext}) holds for $m(n,x)$ that grows no faster than polynomial in $x$ and $N=\int n(t,x)dx$. This is of particular relevance since our derivation of fitness implies $m(n,x)$ is a linear combination polynomials and bounded functions in $N$ and $x$.

## Derivation of SDE for $\bar x$ and $\sigma^2$ \label{SDE_DERIV}

NEW NEW:

Denote by $\varphi_1(t,N|N_0)$, $\varphi_2(t,\bar x|\bar x_0)$ and $\varphi_3(t,\sigma^2|\sigma^2_0)$ the probability density functions of the processes $N(t)$, $\bar x(t)$ and $\sigma^2(t)$ given $N(0)=N_0$, $\bar x(0)=\bar x_0$ and $\sigma^2(0)=\sigma_0^2$. We will from hereon simply write these densities as $\varphi_1(t,N)$, $\varphi_2(t,\bar x)$ and $\varphi_3(t,\sigma^2)$.

NEW:

Before taking the diffusion limit, we can write $n(t,x)$ as

\begin{equation}
n(t,x)=\sum_{i=1}^{N(t)}\delta(x-x_i).
\end{equation}

Hence, the phenotypic mean and variance can be computed as

\begin{equation}
\bar x(t)=\frac{1}{N(t)}\int_\mathbb{R}xn(t,x)dx=\frac{1}{N(t)}\sum_{i=1}^{N(t)}x_i\delta(x-x_i),
\end{equation}

\begin{equation}
\sigma^2(t)=\left(\frac{1}{N(t)}\int_\mathbb{R}x^2n(t,x)dx\right)-\bar x^2(t)=\left(\frac{1}{N(t)}\sum_{i=1}^{N(t)}x_i^2\delta(x-x_i)\right)-\bar x^2(t).
\end{equation}

Set $\tilde x(t)=N(t)\bar x(t)$ and $\tilde{\tilde\sigma}^2(t)=N(t)\left(\sigma^2(t)+\bar x^2(t)\right)$. Then

\begin{equation}
N(t)\left(\tilde x\left(t+\frac{1}{N(t)}\right)-\tilde x(t)\right)=N(t)\int_\mathbb{R}x\left[n\left(t+\frac{1}{N(t)},x\right)-n(t,x)\right]dx
\end{equation}

\begin{equation}
N(t)\left(\tilde{\tilde\sigma}^2\left(t+\frac{1}{N(t)}\right)-\tilde{\tilde\sigma}^2(t)\right)=N(t)\int_\mathbb{R}x^2\left[n\left(t+\frac{1}{N(t)},x\right)-n(t,x)\right]dx.
\end{equation}

If the diffusion limits of these expressions converge, then by construction of $n(t,x)$ they converge to

\begin{equation}
\dot{\tilde x}(t)=\int_\mathbb{R} x\dot n(t,x)dx=\int_\mathbb{R}x\left(m(n,x)n(t,x)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}n(t,x)+\sqrt{V(n,x)n(t,x)}\dot\xi(t,x)\right)dx
\end{equation}

\begin{equation}
\dot{\tilde{\tilde\sigma}}^2(t)=\int_\mathbb{R} x^2\dot n(t,x)dx=\int_\mathbb{R}x^2\left(m(n,x)n(t,x)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}n(t,x)+\sqrt{V(n,x)n(t,x)}\dot\xi(t,x)\right)dx.
\end{equation}

Note that these expressions are merely shorthand for

\begin{equation}
\tilde x(t)-\tilde x(0)=\int_\mathbb{R}x(n(t,x)-n(0,x))dx=\int_0^t\int_\mathbb{R}xn(s,x)\left(m(n,x)f(x)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}f(x)\right)dxds
\end{equation}

$$+\int_0^t\int_\mathbb{R}x\sqrt{V(n,x)n(s,x)}f(x)\dot\xi(s,x)dxds$$

and

\begin{equation}
\tilde{\tilde\sigma}^2(t)-\tilde{\tilde\sigma}^2(0)=\int_\mathbb{R}x^2(n(t,x)-n(0,x))dx=\int_0^t\int_\mathbb{R}xn(s,x)\left(m(n,x)f(x)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}f(x)\right)dxds
\end{equation}

$$+\int_0^t\int_\mathbb{R}x^2\sqrt{V(n,x)n(s,x)}f(x)\dot\xi(s,x)dxds$$

which are satisfied for any bounded continous function $f$ when $\frac{\partial^2}{\partial x^2}$ is taken as the extension of the Laplacian to bounded continous functions. We assume $\dot{\tilde x}(t)$ and $\dot{\tilde{\tilde\sigma}}(t)$ are given by the above expressions which can be simplified to

\begin{equation}
\dot{\tilde x}=\overline{xm}N+\frac{\mu}{2}\int_\mathbb{R} x\frac{\partial^2}{\partial x^2} n(t,x)dx+\sqrt{\overline{x^2V}(t)N(t)}\dot{\tilde\xi}_2(t),
\end{equation}

\begin{equation}\label{vardyn}
\dot{\tilde{\tilde\sigma}}^2(t)=\overline{x^2m}N+\frac{\mu}{2}\int_\mathbb{R}x^2\frac{\partial^2}{\partial x^2}n(t,x)dx.
\end{equation}

OLD:

The derivation of mean trait dynamics follows two steps:

  1. Define the process $\tilde x(t)=\int_{\mathbb R} xn(t,x)dx$ and evaluate expression (\ref{SOLUTION}) of the main text with $h(x)=x$ to obtain:

\begin{equation}
\tilde x(t)-\tilde x(0)=\int_0^t\sqrt{\overline{x^2V}(s)N(s)}d\tilde{\xi}_2(s) + \overline{xm}(t)N(t),
\end{equation}

>where

\begin{equation}
d\tilde{\xi}_2=d\hat\Xi_{\sqrt{x^2Vn}}=\frac{1}{\sqrt{\overline{x^2V}N}}\int_\mathbb{R}x\sqrt{V(n,x)n(t,x)}\dot\xi(t,x)dxdt.
\end{equation}

> Hence

\begin{equation}
d\tilde x = \overline{xm}Ndt + \sqrt{\overline{x^2V}N}d\tilde{\xi}_2.
\end{equation}

  2. Use It&ocirc;'s quotient rule on $\bar x=\tilde x/N$ to obtain:

\begin{equation}
d\bar x = d\left(\frac{\tilde x}{N}\right)=\frac{\tilde x}{N}\left(\frac{d\tilde x}{\tilde x}-\frac{dN}{N}-\frac{d\tilde x}{\tilde x}\frac{dN}{N}+\Big(\frac{dN}{N}\Big)^2\right)
\end{equation}

\begin{equation*}
= \frac{d\tilde x}{N}-\bar x\frac{dN}{N}-\frac{d\tilde x}{N}\frac{dN}{N}+\bar x\Big(\frac{dN}{N}\Big)^2= \overline{xm}dt+\sqrt{\frac{\overline{x^2V}}{N}}d\tilde\xi_2-\bar x\bar mdt-\bar x\sqrt{\frac{\bar V}{N}}d\xi_1-\frac{\overline{xV}}{N}dt+\frac{\bar x\bar V}{N}dt
\end{equation*}

\begin{equation*}
= \left(\mathrm{Cov}_ p(x,m)-\mathrm{Cov}_ p\Big(x,\frac{V}{N}\Big)\right)dt+\sqrt{\frac{\overline{x^2V}}{N}}d\tilde\xi_2-\bar x\sqrt{\frac{\bar V}{N}}d\xi_1.
\end{equation*}

Note that

\begin{equation}
\sqrt{\frac{\overline{x^2V}}{N}}d\tilde\xi_2-\bar x\sqrt{\frac{\bar V}{N}}d\xi_1=\frac{1}{N}\int_\mathbb{R}x\sqrt{V(n,x)n(t,x)}\dot\xi(t,x)dx-\frac{\bar x}{N}\int_\mathbb{R}\sqrt{V(n,x)n(t,x)}\dot\xi(t,x)dx
\end{equation}

\begin{equation*}
=\int_\mathbb{R}\frac{x-\bar x}{N}\sqrt{V(n,x)n(t,x)}\dot\xi(t,x)dx
\end{equation*}

and

\begin{equation}
\mathbb{V}\left(\int_\mathbb{R}\frac{x-\bar x(t)}{N(t)}\sqrt{V(n,x)n(t,x)}\dot\xi(t,x)dx\right)=\frac{1}{N(t)}\int_\mathbb{R}(x-\bar x(t))^2V(n,x)p(t,x)dx
\end{equation}

\begin{equation*}
=\frac{\sigma^2(t)\bar V(t)+\mathrm{Cov}_p\Big((x-\bar x(t))^2,V(n,x)\Big)}{N(t)}.
\end{equation*}

Hence, by setting

\begin{equation}
d\xi_2=\frac{\int_\mathbb{R}\frac{(x-\bar x(t))}{N(t)}\sqrt{V(n,x)n(t,x)}\dot\xi(t,x)dx}{\sqrt{\frac{1}{N(t)}\left(\sigma^2(t)\bar V(t)+\mathrm{Cov}_p\Big((x-\bar x(t))^2,V(n,x)\Big)\right)}}
\end{equation}

we can write

\begin{equation}
d\bar x=\left(\mathrm{Cov}_ p(x,m)-\mathrm{Cov}_ p\Big(x,\frac{V}{N}\Big)\right)dt+\sqrt{\frac{\sigma^2\bar V+\mathrm{Cov}_p\left((x-\bar x)^2,V\right)}{N}}d\xi_2.
\end{equation}

This approach is easily adapted to derive the evolution of the phenotypic variance $\sigma^2$ using the steps:

  1. Define $\tilde{\tilde\sigma}^2(t)=\int_{\mathbb R} x^2n(t,x)dx$ and evaluate expression (\ref{SOLUTION}) of the main text with $h(x)=x^2$ to obtain:

\begin{equation}
\tilde{\tilde\sigma}^2(t)-\tilde{\tilde\sigma}^2(0)=\int_0^t\sqrt{\overline{x^4V}(s)N(s)}d\tilde\xi_3(s)+\overline{x^2m}(t)N(t)+\mu N(t)
\end{equation}

> where

\begin{equation}
d\tilde{\xi}_3=d\hat\Xi_{\sqrt{x^4Vn}}=\frac{1}{\sqrt{\overline{x^4V}N}}\int_\mathbb{R}x^2\sqrt{V(n,x)n(t,x)}\dot\xi(t,x)dx.
\end{equation}

> Hence

\begin{equation}
d\tilde{\tilde\sigma}^2 = (\overline{x^2m}N + \mu N)dt + \sqrt{\overline{x^4V}N}d\tilde{\xi}_3.
\end{equation}

  2. Use It&ocirc;'s quotient rule on $\tilde\sigma^2=\tilde{\tilde\sigma}^2/N$ to obtain:

\begin{equation}
d\tilde\sigma^2 = d\left(\frac{\tilde{\tilde\sigma}^2}{N}\right)=\frac{\tilde{\tilde\sigma}^2}{N}\left(\frac{d\tilde{\tilde\sigma}^2}{\tilde{\tilde\sigma}^2}-\frac{dN}{N}-\frac{d\tilde{\tilde\sigma}^2}{\tilde{\tilde\sigma}^2}\frac{dN}{N}+\Big(\frac{dN}{N}\Big)^2\right)
\end{equation}

\begin{equation}
=\frac{d\tilde{\tilde\sigma}^2}{N}-\tilde\sigma^2\frac{dN}{N}-\frac{d\tilde{\tilde\sigma}^2}{N}\frac{dN}{N}+\tilde{\tilde\sigma}^2\left(\frac{dN}{N}\right)^2.
\end{equation}

Note that

\begin{equation}
\left(\sqrt\frac{\overline{x^4 V}}{N}d\tilde\xi_3\right)\left(\sqrt\frac{\bar V}{N}d\xi_1\right)=\frac{\int_\mathbb{R}\sqrt{x^4V(n,x)n(t,x)}\sqrt{V(n,x)n(t,x)}dx}{N^2}dt=\frac{\overline{x^2V}}{N}dt.
\end{equation}

and hence

\begin{equation}
d\tilde\sigma^2=\left(\overline{x^2m} + \mu\right)dt+\sqrt{\frac{\overline{x^4V}}{N}}d\tilde\xi_3-\tilde\sigma^2 \bar mdt+\tilde\sigma^2\sqrt{\frac{\bar V}{N}}d\xi_1-\frac{\overline{x^2V}}{N}dt-\tilde\sigma^2\frac{\bar V}{N}dt.
\end{equation}

\begin{equation}
=\left(\mathrm{Cov}_p\Big(x^2,m\Big) - \mathrm{Cov}_p\Big(x^2,\frac{V}{N}\Big) + \mu \right)dt+\sqrt\frac{\overline{x^4V}}{N}d\tilde\xi_3-\tilde\sigma^2\sqrt\frac{\bar V}{N}d\xi_1.
\end{equation}

  3. Setting $F(y,z)=y-z^2$, use It&ocirc;'s formula on $\sigma^2=F(\tilde\sigma^2,\bar x)=\tilde\sigma^2-\bar x^2$ to obtain:

\begin{equation}
d\sigma^2=d\tilde\sigma^2-2\bar xd\bar x-(d\bar x)^2
\end{equation}

$$=\left(\mathrm{Cov}_p\Big(x^2,m\Big)-\mathrm{Cov}_p\Big(x^2,\frac{V}{N}\Big) + \mu \right)dt+\sqrt\frac{\overline{x^4V}}{N}d\tilde\xi_3-\tilde\sigma^2\sqrt\frac{\bar V}{N}d\xi_1$$

$$-2\bar x\left(\left(\mathrm{Cov}_ p(x,m)-\mathrm{Cov}_ p\Big(x,\frac{V}{N}\Big)\right)dt+\sqrt{\frac{\sigma^2\bar V+\mathrm{Cov}_p\left((x-\bar x)^2,V\right)}{N}}d\xi_2\right)$$

$$-\left(\left(\mathrm{Cov}_ p(x,m)-\mathrm{Cov}_ p\Big(x,\frac{V}{N}\Big)\right)dt+\sqrt{\frac{\sigma^2\bar V+\mathrm{Cov}_p\left((x-\bar x)^2,V\right)}{N}}d\xi_2\right)^2$$

$$=\left(\mathrm{Cov}_p\Big(x^2-2\bar xx,m\Big)-\mathrm{Cov}_p\Big(x^2-2\bar xx,\frac{V}{N}\Big) + \mu \right)dt+\sqrt\frac{\overline{x^4V}}{N}d\tilde\xi_3-\tilde\sigma^2\sqrt\frac{\bar V}{N}d\xi_1$$

$$-2\bar x\sqrt{\frac{\sigma^2\bar V+\mathrm{Cov}_p\left((x-\bar x)^2,V\right)}{N}}d\xi_2-\left(\frac{\sigma^2\bar V}{N}-\mathrm{Cov}_p\left((x-\bar x)^2,\frac{V}{N}\right)\right)dt$$

$$=\left(\mathrm{Cov}_p\Big(x-\bar x)^2,m\Big) + \mu - \frac{\sigma^2\bar V}{N}\right)dt+\sqrt\frac{\overline{x^4V}}{N}d\tilde\xi_3-\tilde\sigma^2\sqrt\frac{\bar V}{N}d\xi_1-2\bar xd\xi_2$$

In light of

\begin{equation}
\sqrt\frac{\overline{x^4V}}{N}d\tilde\xi_3-\tilde\sigma^2\sqrt\frac{\bar V}{N}d\xi_1-2\bar xd\xi_2=\frac{1}{N}\int_\mathbb{R}\left(x^2-\tilde\sigma^2-2\bar x(x-\bar x)\right)\sqrt{V(n,x)n(t,x)}\dot\xi(t,x)dx
\end{equation}

$$=\frac{1}{N}\int_\mathbb{R}\left((x-\bar x)^2-\sigma^2\right)\sqrt{V(n,x)n(t,x)}\dot\xi(t,x)dx$$

we set

\begin{equation}
\mathscr{V}^2=\mathbb{V}\left(\frac{1}{N}\int_\mathbb{R}\left((x-\bar x)^2-\sigma^2\right)\sqrt{V(n,x)n(s,x)}\dot\xi(s,x)dx\right)
\end{equation}

\begin{equation}
=\frac{1}{N}\left(\int_\mathbb{R}((x-\bar x)^4-2(x-\bar x)^2\sigma^2+\sigma^4)V(n,x)p(t,x)dx\right)
\end{equation}

\begin{equation}
=\frac{1}{N}\left(\overline{(x-\bar x)^4V}-2\sigma^2\overline{(x-\bar x)^2V}+\sigma^4\bar V\right)
\end{equation}

\begin{equation}
=\frac{1}{N}\left(\mathrm{Cov}_p\big((x-\bar x)^4,V\big)+\overline{(x-\bar x)^4}\bar V-2\sigma^2\Big(\mathrm{Cov}_p\big((x-\bar x)^2,V\big)+\sigma^2\bar V\Big)+\sigma^4\bar V\right)
\end{equation}

and

\begin{equation}
d\xi_3=\frac{1}{N\mathscr{V}}\int_\mathbb{R}\left((x-\bar x)^2-\sigma^2\right)\sqrt{V(n,x)n(t,x)}\dot\xi(t,x)dx
\end{equation}

so that

\begin{equation}
d\sigma^2=\left(\mathrm{Cov}_ p\Big((x-\bar x)^2,m\Big)+\mu-\frac{\sigma^2\bar V}{N}\right)dt+\mathscr{V}d\xi_3.
\end{equation}

From here we assume $V(n,x)\equiv\nu>0$. Table \ref{mult} of the main text implies

\begin{equation}
d\xi_1d\xi_2=\frac{\int_\mathbb{R}(x-\bar x)n(t,x)dx}{\sqrt{N\sigma^2}}=0,
\end{equation}

\begin{equation}
d\xi_1d\xi_3=\frac{\int_\mathbb{R}\big((x-\bar x)^2-\sigma^2\big)n(t,x)dx}{\sqrt{\overline{(x-\bar x)^4}-\sigma^4}}=0,
\end{equation}

\begin{equation}
d\xi_2d\xi_3=\frac{\int_\mathbb{R}(x-\bar x)\big((x-\bar x)^2-\sigma^2\big)n(t,x)dx}{\sqrt{\sigma^2\big(\overline{(x-\bar x)^4}-\sigma^4\big)}}
=\frac{N\overline{(x-\bar x)^3}}{\sqrt{\sigma^2\big(\overline{(x-\bar x)^4}-\sigma^4\big)}}.
\end{equation}

In particular, when $p$ is Gaussian $d\xi_2d\xi_3=0$.

## Derivation of diffuse coevolution model \label{diffuse}

In this section we provide a derivation of our model of diffuse coevolution driven by competition. Since most of the work in this derivation has already been completed in Supplementary Material \S\ref{SDE_DERIV}, we focus here on deriving the Malthusian fitness $m$ as a function of trait value $x$. We begin with discrete populations of individuals. In particular, we begin by assuming $N_i(t)$ is an integer for each species $i=1,\dots,S$ and time $t\geq0$ before passing to the large population size limit.

The reduction in fitness for an individual of species $i$ caused by competition is captured multiplicatively by $0<C_i\leq1$. Biologically this assumes all competitors affect individuals of a given species equally by consuming the same amount of resources. This is a mean-field interaction since any individual that consumes resources has an effect on the fitness of all other individuals competing for the same resources. Denote by $x_{ij}$ the trait value of the $j$-th individual belonging to species $i$. The set of trait values across all individuals in the community at time $t\geq0$ is written $X_t=\{x_{ij}\}$. We denote by $\mathscr{B}_{ij}$ a function that maps $X_t$ to the cumulative effect of all competitive interactions on the fitness of the $j$-th individual in species $i$. Since individuals do not compete with themselves the net multiplicative effects on fitness of both interspecific and intraspecific competition on the $j$-th individual in species $i$ is summarized by

\begin{equation}
\mathscr B_{ij}(X_t)=C_i^{\sum_{l\neq j}\mathcal{O}_{ii}(x_{ij},x_{il})+\sum_{k\neq i}\sum_{l=1}^{N_k(t)}\mathcal{O}_{ik}(x_{ij},x_{kl})},
\end{equation}

where $\mathcal{O}_{ij}$, defined in the main text, measures the overlap in resource use between individuals of species $i$ and $j$ as a function of their niche-centers. Writing $\mathscr W_{ij}(X_t)$ as the number of offspring left by the $j$-th individual of species $i$, we have

\begin{equation}
\mathscr W_{ij}(X_t)=\mathscr{A}_i(x_{ij})\mathscr{B}_{ij}(X_t),
\end{equation}

where $\mathscr{A}_i(x)=\int_\mathbb{R}e_i(\zeta)u_i(\zeta,x)d\zeta$ and $e_i$ has been defined in the main text. The Malthusian fitness $m$ is related to $\mathscr W$ by $m=\ln\mathscr{W}$ [@9781932846126]. Hence,

\begin{equation}
m_{ij}(X_t)=\ln\mathscr{A}_i(x_{ij})+\ln\mathscr{B}_{ij}(X_t)
\end{equation}

$$=\ln\mathscr{A}_i(x_{ij})+\left(\sum_{l\neq j}\mathcal{O}_{ii}(x_{ij},x_{il})+\sum_{k\neq i}\sum_{l=1}^{N_k(t)}\mathcal{O}_{ik}(x_{ij},x_{kl})\right)\ln C_i.$$

The phenotypic distribution in species $i$ is given by $\tilde p_i(t,x)=\frac{1}{N_i(t)}\sum_j\delta(x-x_{ij})$ with $\delta$ the Dirac distribution having unit mass centered at zero. Taking $N_1(t),\dots,N_S(t)\to\infty$ and assuming the limit $p_i(t,x)$ of the distribution $\tilde p_i(t,x)$ is a probability density function, $m_{ij}(X_t)$ approaches

\begin{equation}
m(t,x_{ij})=\ln\mathscr{A}_i(x_{ij})+\left((N_i(t)-1)\int_\mathbb{R}\mathcal{O}_{ii}(x_{ij},x)p_i(t,x)dx+\sum_{k\neq i}N_k(t)\int_\mathbb{R}\mathcal{O}_{ik}(x_{ij},x)p_k(t,x)dx\right)\ln C_i
\end{equation}

$$=\ln\mathscr{A}_i(x_{ij})+\left(-\bar{\mathcal{O}}_{ii}(t,x_{ij})+\sum_{k=1}^{S}N_k(t)\bar{\mathcal{O}}_{ik}(t,x_{ij})\right)\ln C_i.$$

When $p_i(t,x)$ and $p_j(t,x)$ are normal, the expected niche overlap between an individual at niche location $x_i$ with individuals in species $j$ is

\begin{equation}
\bar{\mathcal{O}}_{ij}(t,x_i)=\int_\mathbb{R}\mathcal O(x_i,x_j) p_j(t,x_j)dx_j=\frac{1}{\sqrt{2\pi(w_i+w_j+{\sigma_j}^2(t))}}\exp\left(-\frac{(x_i-\bar x_j)^2}{2(w_i+w_j+{\sigma_j}^2(t))}\right).
\end{equation}

Setting

\begin{subequations}
\begin{align}
r_i = & \ \ln \left(\frac{W_iU_i}{\sqrt{1+A_iw_i}}\right), \\
a_i = & \ \frac{A_i}{1+A_iw_i}, \\
\tilde b_{ij}(t) = & \ \frac{1}{w_i+w_j+\sigma_j^2(t)}, \\
c_i = & -\ln C_i,
\end{align}
\end{subequations}

we get

\begin{equation}\label{malth_fit}
m_i(t,x)=r_i-\frac{a_i}{2}(x-\theta_i)^2+c_i\left(\sqrt{\frac{\tilde b_{ii}(t)}{2\pi}}e^{-\frac{\tilde b_{ii}(t)}{2}(x-\bar x_i(t))^2}-\sum_{j=1}^SN_j(t)\sqrt{\frac{\tilde b_{ij}(t)}{2\pi}}e^{-\frac{\tilde b_{ij}(t)}{2}(x-\bar x_j(t))^2}\right).
\end{equation}

For the remainder of the derivation we suppress notation indicating dependency on $t$ and $x$. From (\ref{malth_fit}) we calclate

\begin{equation}
\frac{\partial m_i}{\partial\bar x_i}=c_i\tilde b_{ii}(x-\bar x_i)\sqrt{\frac{\tilde b_{ii}}{2\pi}}e^{-\frac{\tilde b_{ii}}{2}(x-\bar x_i)^2}
\end{equation}

\begin{equation}
\frac{\partial m_i}{\partial G_i}=\frac{c_i\left((x-\bar x_i)^2-{\sigma_i}^2-2w_i\right)}{2({\sigma_i}^2+2w_i)^2}\sqrt{\frac{\tilde b_{ii}}{2\pi}}e^{-\frac{\tilde b_{ii}}{2}(x-\bar x_i)^2}
\end{equation}

and hence

\begin{equation}
\overline{\frac{\partial m_i}{\partial\bar x_i}}=0,
\end{equation}

\begin{equation}
\overline{\frac{\partial m_i}{\partial G_i}}=\frac{c_i}{2({\sigma_i}^2+2w_i)^2}\left(\frac{({\sigma_i}^2+2w_i){\sigma_i}^2}{2(w_i+{\sigma_i}^2)}-{\sigma_i}^2-2w_i\right)\sqrt{\frac{b_{ii}}{2\pi}}
\end{equation}

$$=-\frac{c_i}{2({\sigma_i}^2+2w_i)^2}\frac{({\sigma_i}^2+2w_i)^2}{2({\sigma_i}^2+w_i)}\sqrt{\frac{b_{ii}}{2\pi}}=-\frac{c_ib_{ii}}{2}\sqrt{\frac{b_{ii}}{2\pi}},$$

where

\begin{equation}
b_{ij}=\frac{1}{w_i+w_j+{\sigma_i}^2+{\sigma_j}^2}.
\end{equation}

The average fitness for species $i$ is

\begin{equation}
\bar m_i=r_i-\frac{a_i}{2}\Big((\bar x_i-\theta_i)^2+G_i+\eta_iN_i\Big) + c_i\left(\sqrt{\frac{b_{ii}}{2\pi}}-\sum_{j=1}^SN_j\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\right).
\end{equation}

Hence

\begin{equation}
\frac{\partial\bar m_i}{\partial\bar x_i}=a_i(\theta_i-\bar x_i)-c_i\sum_jN_jb_{ij}(\bar x_j-\bar x_i)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2},
\end{equation}

\begin{equation}
\frac{\partial\bar m_i}{\partial G_i}=-\frac{a_i}{2}+\frac{c_i}{2}\left(-b_{ii}\sqrt\frac{b_{ii}}{2\pi}+\sum_{j=1}^SN_jb_{ij}\left(1-b_{ij}(\bar x_i-\bar x_j)^2\right)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\right).
\end{equation}

Applying equations (\ref{N}), (\ref{xbarfinal}) and (\ref{Gfinal}) of the main text provides

\begin{subequations}
\begin{align}
\dot N_i = & \left\{r_i-\frac{a_i}{2}\Big((\bar x_i-\theta_i)^2+G_i+\eta_iN_i\Big) + c_i\left(\sqrt{\frac{b_{ii}}{2\pi}}-\sum_{j=1}^SN_j\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\right)\right\}N_i+\nu_i\sqrt{N_i}\dot\xi_1, \\
\dot{\bar x}_i = & \ a_iG_i(\theta_i-\bar x_i)-c_iG_i\sum_{j=1}^SN_jb_{ij}(\bar x_j-\bar x_i)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}+\nu_i\sqrt{\frac{G_i}{N_i}}\dot\xi_2, \\
\dot G_i = & \ \mu_i-a_i{G_i}^2+c_i{G_i}^2\left(\sum_{j=1}^SN_jb_{ij}\left(1-b_{ij}(\bar x_i-\bar x_j)^2\right)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\right)-\frac{\nu_i^2G_i}{N_i}+\nu_iG_i\sqrt{\frac{2}{N_i}}\dot\xi_3.
\end{align}
\end{subequations}

\newpage

# References

### 2020-03-20 bbm.jl

@@ -1,62 +1,56 @@
 using Parameters, Statistics, Random, LinearAlgebra, Distributions,
-	StatsBase, StatsPlots, Plots, DataFrames, CSV, Optim
+	StatsBase, StatsPlots, Plots, DataFrames, CSV, Optim, KernelDensity,
+	QuadGK, HypothesisTests
 
 include("/home/bob/Research/Branching Brownian Motion/bbm_functions_structs.jl")
 
-#
-# a branching Brownian motion
-#
+#######################################################
+#                                                     #
+# a branching Brownian motion for a single population #
+#                                                     #
+#######################################################
 
 # parameter values
-S = 1
-w = fill(0.1, S)  # niche breadths
-U = fill(1.0, S)  # total niche use
-c = fill(0.0003,S)      # strengths of competition
-Ω = sum(U) # niche use scaling
-η = fill(1.0, S)  # segregation variances
-μ = fill(1e-7, S) # mutation rates
-V = fill(5.0, S)  # magnitudes of drift
-R = fill(1.0, S) # innate rate of growth
-a = fill(1e-2,S)       # strengths of abiotic selection
-θ = fill(0.0, S)  # phenotypic optima
-
-# equilibrium abundance an the absence of interspecific interactions
-# we use this as the initial abundance
-
-C = c.*.√( U.^2 ./ .√(4*π.*w) )
-
-N₀ = Int64.(floor.( (R.-0.5*.√(μ.*a))./C ) )
-
-# initial breeding values
-g₀ = rand.(Normal(0.0,1.0),N₀)
-
-# initial trait values
-x₀ = fill(zeros(0),S)
-for i in 1:S
-	ηₘ = √η[i]*Matrix(I, N₀[i], N₀[i])
-	x₀[i] = vec(rand(MvNormal(g₀[i],ηₘ),1))
-end
+w = 0.1  # niche breadths
+U = 1.0  # total niche use
+c = 2e-3 # strengths of competition
[…]

### 2020-03-20 bbm_functions_structs.jl

@@ -317,24 +317,7 @@ function rescaled_lower(X)
             # this follows exactly from SM §5.6
             #
 
-            # container for aggregating effects of competition
-            B = 0.0
-
-            # collect effects of competition with other individuals
-            # within the same population
-            for k in filter(x -> x≠j, 1:N[i])
-                B += U[i]^2 / √(4*π*w[i])
-            end
-
-            # collect effects of competition with other individuals
-            # in other populations
-            for k in filter(x -> x≠i, 1:S)
-                for l in 1:N[k]
-                    B += U[i]*U[k] / √(2*π*(w[i]+w[k]))
-                end
-            end
-
-            𝒲 = exp( (R[i] - a[i]*(θ[i]-x[i][j])^2/2.0 - c[i]*B/n[i]) / n[i] )
+            𝒲 = exp( ( R[i] - (a[i]*(θ[i]-x[i][j])^2/2.0) - c[i]*N[i]/n[i] ) / n[i] )
 
             # parameterizing the NegativeBinomial
             q = 𝒲/V[i]
@@ -345,6 +328,9 @@ function rescaled_lower(X)
 
         end
 
+        # tracks the current offspring
+        count = Int64(1)
+
         # loop through parents
         for j in 1:N[i]
 
@@ -352,10 +338,12 @@ function rescaled_lower(X)
             for k in 1:W[j]
 
                 # draw random breeding value for this individual
-                append!(gₚ[i], rand( Normal( g[i][j], √μ[i] ), 1)[1])
+                append!( gₚ[i], rand( Normal( g[i][j], √(μ[i]/n[i]) ), 1)[1] )
 
                 # draw random trait value for this individual
-                append!(xₚ[i], rand( Normal( gₚ[i][k], √η[i] ), 1)[1])
+                append!( xₚ[i], rand( Normal( gₚ[i][count], √η[i] ), 1)[1] )
+
+                count += 1
 
             end
 
@@ -364,7 +352,7 @@ function rescaled_lower(X)
         x̄ₚ[i] = mean(xₚ[i])
         σₚ²[i]= var(xₚ[i])
         Gₚ[i] = var(gₚ[i])
-        Nₚ[i] = length(gₚ[i])
+        Nₚ[i] = sum(W)
 
     end
 

### 2020-03-20 bbm_gaussian.jl

@@ -4,62 +4,63 @@ using Parameters, Statistics, Random, LinearAlgebra, Distributions,
 
 include("/home/bob/Research/Branching Brownian Motion/bbm_functions_structs.jl")
 
-#
-# demonstrating approximate Gaussian distribution for small V/N
-#
-# corresponds
-#
-
-# number of generations to halt at
-T = 200
-
-# initial trait values
-x₀ = rand(Normal(0.0,1.0),n)
+#######################################################
+#                                                     #
+# a branching Brownian motion for a single population #
+#                                                     #
+#######################################################
 
 # parameter values
-r = 2.0
-a = 0.01
-θ = 0.0
-c = 0.0001
-μ = 1.0
-V = 20.0
-
-# equilibrium abundance
-# we use this as the initial abundance
-N̂ = floor((r-.5*√(μ*a))/c)
+w = 0.1  # niche breadths
+U = 1.0  # total niche use
+c = 2e-3 # strengths of competition
+η = 1e-5 # segregation variances
+μ = 1e-2 # mutation rates
+V = 2.0  # magnitudes of drift
+r = 1.0  # innate rate of growth
+a = 1e-2 # strengths of abiotic selection
+θ = 0.0  # phenotypic optima
+n = 20.0 # scaling parameter
 
 ##
 ## VERY IMPORTANT REQUIREMENT   -->  V >= exp(r)
 ##
 ## this inequality must be satisfied to use negative binomial sampling
 ##
-##
-## TWO MORE IMPORTANT REQUIREMENTS --> 2*r > √(μ*a) && c > r - √(μ*a)/2
-##
-## these inequalities must be satisfied for positive equilibrium abundance
-##
+
+# equilibrium abundance an the absence of interspecific interactions
+# we use this as the initial abundance
+
[…]

### 2020-03-20 sde_functions.jl

@@ -118,7 +118,11 @@ function f(u,p,t)
 
   end
 
-  return( cat(dx,dG,dN,dims=1) )
+  du = dx
+  append!(du,dG)
+  append!(du,dN)
+
+  return(du)
 
 end
 
@@ -161,7 +165,11 @@ function g(u,p,t)
 
   end
 
-  return( cat(dx,dG,dN,dims=1) )
+  du = dx
+  append!(du,dG)
+  append!(du,dN)
+
+  return(du)
 
 end
 

### 2020-03-20 sde_num_sol.jl

@@ -6,113 +6,18 @@ include("/home/bob/Research/Branching Brownian Motion/sde_functions.jl")
 j=1
 k=2
 # strengths of competition
-cc = [1e-5,1e-3]
+cc = [1e-7,5e-6]
 
 # durations
 TT = [1e3,5e3]
 
-dumm = true
-
-while dumm
-
-	try for j in 1:2
-
-		for k in 1:2
-
-			tspan = (0.0,TT[k])
-
-			# background parameters
-			S = 100
-			w = fill(0.1, S)  # niche breadths
-			U = fill(1.0, S)  # total niche use
-			c = fill(cc[i],S)/S      # strengths of competition
-			Ω = sum(U) # niche use scaling
-			η = fill(1.0, S)  # segregation variances
-			μ = fill(1e-7, S) # mutation rates
-			V = fill(5.0, S)  # magnitudes of drift
-			R = fill(1.0, S) # innate rate of growth
-			a = fill(1e-2,S)       # strengths of abiotic selection
-			θ = fill(0.0, S)  # phenotypic optima
-
-			pars = ModelParameters(S=S, w=w, U=U, η=η, c=c, a=a, μ=μ, V=V, R=R, θ=θ, Ω=Ω)
-
-			#
-			# find deterministic equilibrium
-			#
-
-			# approximation for equilibrium abundance
-			C = c.*.√( U.^2 ./ .√(4*π.*w) )
-			N₀ = Int64.(floor.( (R.-0.5*.√(μ.*a))./C ) )
-
-			# initial condition
-			u₀ = cat(θ,fill(10.0, S),fill(1000.0, S),dims=1)
-
-			# numerically solve SDE
-			prob = SDEProblem(f,g,u₀,tspan,pars)
-			sol = solve(prob,maxiters=1e8)
-
-			plot(sol)
-
-			# state at time T
-			endT = maximum(sol.t)
-			x̃ = sol(endT)
-
-			# extract state variables
-			x̄_sol = copy(sol[(0*S+1):(1*S),:])
-			G_sol = copy(sol[(1*S+1):(2*S),:])
-			N_sol = copy(sol[(2*S+1):(3*S),:])
-
-			# build dataframes
-			spp = string("Species", 1)
-			global df = DataFrame(spp = spp, x = x̄_sol[1,:], G = G_sol[1,:],
-			N = N_sol[1,:], time = sol.t)
-			for i in 2:S
-				spp = string("Species", i)
[…]

### 2020-03-29 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
+

@@
 Such methods will tether theory to reality and allow for rigorous tests of hypotheses on the structure and behavior of ecological communities.
-In this paper we introduce a framework that establishes a formal connection between the continuous-time dynamics of abundance and quantitative traits in stochastically evolving populations, demonstrate it's utility through the derivation and analysis of a model of diffuse coevolution and discuss how it can be extended to account for the details mentioned above.
+In this paper we introduce a framework that establishes a formal connection between the continuous-time dynamics of abundance and quantitative traits in stochastically evolving populations.
+We then demonstrate the utility of our framework through the derivation and analysis of a model of diffuse coevolution and discuss how it can be extended to account for the details mentioned above.
 
@@
 
-Although reviewing these accomplishments reveals a beautiful synthesis of evolution and population ecology, it also reveals a gap in theoretical approaches to incorporate the intrinsically random nature of populations.
+Reviewing these accomplishments reveals a beautiful synthesis of evolution and population ecology.
+However, it also reveals a gap in theoretical approaches to incorporate the intrinsically random nature of populations.
 Specifically, in theoretical quantitative genetics the derivation of a population's response to random genetic drift is derived in discrete time under the assumption of constant effective population size using arguments based on properties of random samples [@Lande1976].
-Although this approach conveniently mimics the formalism provided by the Wright-Fisher model of population genetics, real population sizes fluctuate over time.
+Though this approach conveniently mimics the formalism provided by the Wright-Fisher model of population genetics, real population sizes fluctuate over time.
 Furthermore, since these fluctuations are themselves stochastic, it seems natural to derive expressions for the evolutionary response to demographic stochasticity and consider how the results relate to characterizations of random genetic drift.
@@
 Our goal has two components: 1) Establish a novel synthetic approach to theoretical evolutionary ecology that provides a formal connection between demographic stochasticity and random genetic drift in the context of quantitative traits.
-To demonstrate the utility of our approach, we derive a model of coevolution in an ecological network and use it to investigate the relationship between competition coefficients and selection gradients.
+To show that our approach can be used to develop useful biological insights we derive a model of coevolution in an ecological network and use it to investigate the relationship between competition coefficients and selection gradients.
 2) Communicate some useful properties of space-time white noise, BBM and SPDE to as wide of audience as possible.
@@
 However, although biologically insightful, these equations remain difficult to analyze and implement numerically.
-In \S\ref{equations} we use an assumption of normally distributed trait values to simplify these expressions into formula that are much easier to work with.
+In \S\ref{equations} we use an assumption of normally distributed trait values to simplify these expressions into formulae that are much easier to work with.
 We then account for the constraint of adaptive evolution on the amount of heritable variation in a population by extending these results via a model of imperfect inheritance.
 The resulting equations coincide with classical results in quantitative genetics as a special case.
-We then demonstrate the utility of our approach by combining the derived equations of population dynamics with classical niche theory in \S\ref{coev} to formulate of a model of coevolution across a guild of $S$ species participating in exploitative competition along a common resource continuum.
+In \S\ref{coev} we combine the derived equations of population dynamics with classical niche theory to formulate a model of coevolution across a guild of $S$ species participating in exploitative competition along a common resource continuum.
 In SM \S\ref{diffuse} we apply a classical theorem on rescaled limits of BBM that allow for ecological interactions to provide a rigorous derivation.
@@
 Two important observations of these equations include 1) mean fitness $\bar m$ determines the abundance dynamics of the entire population and 2) allele $i$ will increase (decrease) in frequency if $m_i>\bar m$ ($<\bar m$).
-Equation (\ref{popgen}) is known in the field of evolutionary game theory as the replicator equation [@Hofbauer1998; @9780674023383; @Lion2018].
-Instead of being explicitly focused on alleles, the replicator equation describes the fluctuations of relative abundances of various _types_ in a population in terms of the vital rates of each type [@Taylor1978; @Schuster1983].
+Equation (\ref{popgen}) is known in the field of evolutionary game theory as the replicator equation [@Hofbauer1998; @9780674023383; @Lion2018; @Taylor1978; @Schuster1983].
+Instead of being explicitly focused on alleles, the replicator equation describes the fluctuations of relative abundances of various _types_ in a population in terms of the vital rates of each type.
 Using a matrix of transition rates between differing types, it is straight-forward to extend the replicator equation to include mutation, which is known as the replicator-mutator equation [@9780674023383].
@@
 
-To stay within the realm of biological plausibility, we require a set technical assumptions.
+To stay within the realm of biological plausibility we require a set technical assumptions.
 First, we assume the initial abundance density is continuous, non-negative, integrable and has finite trait mean and variance.
@@
 From a statistical perspective, if we think of $(x-\bar x)^2$ as a square error, then in analogy to the dynamics of the mean trait, we see that the response in $\sigma^2$ to selection can be expressed as a covariance of fitness and square error, which is defined in analogy to $\mathrm{Cov}_t(m(\nu,x),x)$.
-This covariance also creates change in $\sigma^2$ that maximizes mean fitness $\bar m$.
+This covariance also creates change in $\sigma^2$ that maximizes mean fitness $\bar m$, which can be positive or negative depending on whether selection is stabilizing or disruptive.
 In SM \S\ref{var_deriv} we show when $p(x,t)$ is a Gaussian bell curve, the increase in variance due to mutation simplifies to $\mu$.
@@
 Similar to the approach taken by Champagnat, Ferriére and Méléard [-@Champagnat2006], we begin with a BBM that models populations as discrete sets of reproducing individuals whose vital rates depend on their trait value as well as the state of the entire population.
-Taking a large population size limit and keeping our assumption of single dimensional traits, we employ a pair of classic results that show, under the appropriate rescaling in time, phenotypic space and population density, a sequence of rescaled BBM converges to a limiting process that can be characterized by a SPDE [@Mlard1993; @zeng1998absolute].
+Taking a large population size limit and keeping our assumption of single dimensional traits, we employ a pair of classical results that show, under the appropriate rescaling in time, phenotypic space and population density, a sequence of rescaled BBM converges to a limiting process that can be characterized by a SPDE [@Mlard1993; @zeng1998absolute].
 The limiting processes of rescaled BBM have been referred to as measure-valued Markov processes [@dawson1993measure] or superprocesses [@alisonetheridge2000].
@@
 The simplicity of our special case allows us to use properties of space-time white noise processes to derive a set of SDE that generalize equations (\ref{N_det}), (\ref{detmean}) and (\ref{vardyn_det}) to include the effects of demographic stochasticity.
-Classic expressions for the effects of random genetic drift on the evolution of mean traits are obtained as a special case.
+Classical expressions for the effects of random genetic drift on the evolution of mean traits are obtained as a special case.
 
@@
 Since our aim is to present this material to as wide of audience as possible, our approach deviates from standard definitions to remove the need for a detailed technical treatment.
-In addition to the notions of Reimann integration and partial differentiation already employed, the reader will only need some elementary probability and an intuitive understanding of SDE.
+In addition to the notions of Reimann integration and partial differentiation already employed, the reader will only need some elementary probability and an intuitive understanding of SDE, including Brownian motion.
 Because space-time white noise, denoted by $\dot W(x,t)$, appears in the SPDE characterizing diffusion limits of BBM, we begin by defining $\dot W(x,t)$ and illustrating its relevant properties including a set of heuristics for performing calculations.
@@
 In relation to Brownian motion, denoted by $W$, white noise can be interpreted of as the derivative of Brownian motion with respect to time, denoted $\dot W$.
-Since Brownian motion takes infinitesimally small Gaussian distributed jumps at each time point, this leads to the conceptualization of white noise as a collection of Gaussian distributed random variables.
+Since Brownian motion can be thought to take infinitesimally small Gaussian distributed jumps at each time point, this leads to the conceptualization of white noise as a collection of Gaussian distributed random variables.
 Figure \ref{wn} illustrates realizations of this conceptualized white noise in one (left) and two (right) dimensions.
@@
 wn1d = ggplot(df_1)+geom_point(aes(x=x,y=xi),size=0.05)+theme_minimal()+xlab("Time")+ylab("Value")
-wn2d = ggplot(M, aes(x, y, color = value))+scale_color_gradient(low="black", high="white")+geom_point(size=0.001, shape=22, show.legend = FALSE)+theme_minimal()+xlab("Time")+ylab("Space")
+wn2d = ggplot(M, aes(x, y, color = value))+scale_color_gradient2(low="black",mid="grey50", high="white")+geom_point(size=0.001, shape=22, show.legend = FALSE)+theme_minimal()+xlab("Time")+ylab("Space")
 
@@
 
-ggsave("~/Research/A White Noise Approach To Evolutionary Ecology/wn.png",wn,width=5,height=2.5)
+ggsave("~/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/wn.png",wn,width=5,height=2.5)
 ```
@@
 Thus our notation $\dot W$ is only meant to aid intuition and not be taken as formal.
-A formal understanding is possible by considering white noise as a _measure_-valued process [@Dawson1975; @Walsh] or as a _generalized_ process that acts on classically defined functions to return random variables or stochastic processes [@Krylov1981; @DaPrato2014].
-The latter provides the general idea implemented here, though we do not closely follow the treatments of Krylov and Rozovskii [-@Krylov1981] or Da Prato and Zabczyk [-@DaPrato2014].
-
-Throughout this section, we minimize notation by writing $\int_{\mathbb{R}}f(x)dx=\int_{-\infty}^{+\infty}f(x)dx$ and similarly $\int_Df(x)dx$ for the integral of $f$ on $D\subset\mathbb{R}$.
-We define $L^2_c(\mathbb{R}\times[0,\infty))$ as the set of functions $f(x,t)$ mapping $\mathbb{R}\times[0,\infty)$ to $\mathbb R$ that are continuous in $t$ and satisfy $\|f\|_2(t)=\sqrt{\int_\mathbb R f^2(x,t)dx}<\infty$ for each $t\in[0,\infty)$.
-We define a generalized stochastic process $\mathbf W$ that maps functions $f\in L^2_c(\mathbb{R}\times[0,\infty))$ to real-valued stochastic processes indexed by $t\in[0,\infty)$.
-To evaluate this process for a function $f\in L^2_c(\mathbb{R}\times[0,\infty))$ and some time $t\in[0,\infty)$ we write $\mathbf W_t(f)$.
-Specifically, for any $f,g\in L^2_c(\mathbb{R}\times[0,\infty))$, we define $\mathbf W(f)$ and $\mathbf W(g)$ to be Gaussian processes satisfying, for any $t,t_1,t_2\geq0$,
+A formal understanding is possible by considering white noise as a _measure_-valued process [@Dawson1975; @Walsh] or as a _generalized_ process that acts on classically defined functions or stochastic processes to return random variables or stochastic processes [@Krylov1981; @DaPrato2014].
+Since a measure-valued process can be defined from a generalized process and a generalized process can be defined from a measure-valued process, the distinction between the two is more or less a matter of perspective.
+However, we find the perspective of white-noise as a generalized process to be a more efficient route for developing heuristics to help with some routine calculations.
+Hence, this latter approach provides the general idea implemented here.
+Although the treatments provided by Krylov and Rozovskii [-@Krylov1981] and Da Prato and Zabczyk [-@DaPrato2014] extend the theory of SDE to formally treat SPDE in a general and elegant fashion, they require the navigation of an enourmous amount of technical definitions and detailed proofs.
+To extract some particularly useful results from this theory relevant to our goal of synthesizing the stochastic dynamics of biological populations, we provide a streamlined approach by capitolizing on the solid ground these authors have established.
+For instance, instead of rigorously proving properties of white-noise, we simply define them to be so, taking solice in the fact that the technical details have been worked out elsewhere.
+
+To shed a bit of light on the idea of a generalized process, we note that they are the stochastic analog of a generalized function, such as the Dirac delta function $\delta$.
+Just as a generalized function operates on classical functions to return a value (e.g., $\delta(f)=f(0)$), a generalized process acts on a set of functions to return a classically defined stochastic process.
+For a breif overview of the theory of generalized functions, see the addendum to chapter 3 of Kolmogorov and Fomin [-@kolmogorov1999elements].
+
+Throughout this section, we minimize notation by writing $\int_{\mathbb{R}}f(x)dx=\int_{-\infty}^{+\infty}f(x)dx$ and similarly $\int_Df(x)dx$ for the integral of $f$ over $D\subset\mathbb{R}$.
+We define $\mathscr{N}_2$ as the set of stochastic processes $f(x,t)$ that are continuous in both $x$ and $t$ and satisfy $\mathbb{E}\left(\int_0^t\int_\mathbb{R}f^2(x,s)dxds\right)<+\infty$ for each $t\geq0$.
+The operator $\mathbb{E}$ denotes expectation with respect to the underlying probability space.
+For each $t\geq0$ we set
+
+\begin{equation}
+\|f\|_t=\sqrt{\mathbb{E}\left(\int_0^t\int_\mathbb{R}f^2(x,s)dxds\right)}.
+\end{equation}
+
+We define a generalized stochastic process $\mathbf W$ that maps processes $f\in\mathscr{N}_2$ to real-valued stochastic processes indexed by time $t\geq0$, but not by space.
+To evaluate $\mathbf W$ for a process $f\in\mathscr{N}_2$ and some time $t\geq0$ we write $\mathbf W_t(f)$.
+Specifically, for any $f,g\in\mathscr{N}_2$, we define $\mathbf W(f)$ and $\mathbf W(g)$ to be Gaussian processes satisfying, for any $t,t_1,t_2\geq0$,
 \begin{equation}\label{exp_WN}
@@
 \begin{equation}\label{cov_WN}
-\mathbb{C}\big(\mathbf W_{t_1}(f),\mathbf W_{t_2}(g)\big)=\int_0^{t_1\wedge t_2}\int_\mathbb{R} f(x,s)g(x,s)dxds,
-\end{equation}
-where $t_1\wedge t_2=\min(t_1,t_2)$ and $\mathbb{E},\mathbb{C}$ denote expectation and covariance with respect to the underlying probability space.
-These operators are to be distinguished from $\bar f(t)$ and $\mathrm{Cov}_t(f,g)$ which denote expectation and covariance with respect to phenotypic diversity at time $t\geq0$.
-
-Since Gaussian processes are characterized by their expectations and covariances, the processes $\mathbf W(f)$ and $\mathbf W(g)$ are well defined (under the convention $f=g$ whenever $\|f-g\|_2(t)=0$ for each $t\geq0$).
-In particular, if $f\in L^2_c(\mathbb{R}\times[0,\infty))$ is independent of time, then $\mathbf W(f)$ is a Brownian motion with variance at time $t\geq0$ equal to $t\|f\|_2^2$.
+\mathbb{C}\big(\mathbf W_{t_1}(f),\mathbf W_{t_2}(g)\big)=\mathbb{E}\left(\int_0^{t_1\wedge t_2}\int_\mathbb{R} f(x,s)g(x,s)dxds\right),
+\end{equation}
+where $t_1\wedge t_2=\min(t_1,t_2)$ and $\mathbb{C}$ denotes covariance with respect to the underlying probability space.
+The operators $\mathbb{E}$ and $\mathbb{C}$ are to be distinguished from $\bar f(t)$ and $\mathrm{Cov}_t(f,g)$ which denote expectation and covariance with respect to phenotypic diversity at time $t\geq0$.
+
+Since Gaussian processes are characterized by their expectations and covariances and since we assume the $\mathscr{N}_2$ processes are continuous in both time and space, the processes $\mathbf W(f)$ and $\mathbf W(g)$ are well defined.
+In particular, if $f\in\mathscr{N}_2$ is independent of time, then $\mathbf W(f)$ is a Brownian motion with variance at time $t\geq0$ equal to $t \ \mathbb{E}(\int_\mathbb{R}f^2(x,0)dx)$.
 With the generalized process $\mathbf W$ defined, we define the space-time white noise $\dot W(x,t)$ implicitly via the stochastic integral
 \begin{equation}
-``\int_0^t\int_\mathbb{R}f(x,s)\dot W(x,s)dxds"=``\int_\mathbb{R}\int_0^tf(x,s)\dot W(x,s)dxds"=\mathbf W_t(f), \ \forall \ f\in L^2_c(\mathbb{R}\times[0,\infty)), \ t\geq0.
+``\int_0^t\int_\mathbb{R}f(x,s)\dot W(x,s)dxds"=``\int_\mathbb{R}\int_0^tf(x,s)\dot W(x,s)dxds"=\mathbf W_t(f), \ \forall \ f\in\mathscr{N}_2, \ t\geq0.
 \end{equation}
 We place quotations in the above expression to emphasize its informal nature and that it should not be confused with classical Riemann integration.
-Following this definition of white noise, to compute its value we sample it using $L^2_c(\mathbb{R}\times[0,\infty))$ functions.
+Following this definition of white noise, we compute its value by sampling it using $\mathscr{N}_2$ processes.
 For example, integrating white noise over a region $D\times[0,t]$, with $t>0$ and $D$ a bounded subset of $\mathbb{R}$, is equivalent to evaluating $\mathbf W_t(I_{D\times [0,\infty)})$ where
@@
 \end{equation}
-where $\mathbb{V}$ denotes the variance operator with respect to the underlying probability space. Following this notation, equations (\ref{exp_WN}) and (\ref{cov_WN}) can be rewritten as
+where $\mathbb{V}$ denotes the variance operator with respect to the underlying probability space. Using this notation equations (\ref{exp_WN}) and (\ref{cov_WN}) can be rewritten as
 \begin{equation}\label{exp_WN_xi}
@@
 \end{equation}
-To relate this material to the common notation used for SDE, we write $d\hat{\mathbf W}_t(f)=\frac{1}{\|f\|_2(t)}\left(\int_\mathbb{R}f(x,t)\dot W(x,t)dx\right)dt$ so that
+To relate these formulae to the common notation used for SDE, we write $d\hat{\mathbf W}_t(f)=\frac{1}{\|f\|_2(t)}\left(\int_\mathbb{R}f(x,t)\dot W(x,t)dx\right)dt$ so that
 \begin{equation}
@@
 
-A BBM tracks individuals navigating $d$-dimensional Euclidean space whose lifetimes are continuous random variables.
+A BBM tracks individuals navigating $d$-dimensional Euclidean space that reproduce and senesce between exponentially distributed intervals.
 Unlike other stochastic processes that take values in $\mathbb{R}^d$, BBM takes values in the set of _non-negative finite measures_ over $\mathbb{R}^d$.
-Intuitively, one can think of a finite measure as a function that maps subsets of $\mathbb{R}^d$ to real numbers (to be formal, we only consider the Borel subsets of $\mathbb{R}^d$ corresponding to the Euclidean metric, but understanding this technicality is not crucial to our discussion).
+Intuitively, one can think of a finite measure as a function that maps subsets of $\mathbb{R}^d$ to real numbers.
+To be formal, we only consider the Borel subsets of $\mathbb{R}^d$ corresponding to the Euclidean metric, but understanding this technicality is not crucial to our discussion.
 In particular, denoting $X_t$ a BBM, for a subset $D\subset\mathbb{R}^d$, $X_t(D)$ returns the (random) number of individuals alive within the region $D$ at time $t\geq0$.
@@
 
-  1) __Branching rate:__ Lifetimes of individuals are assumed to be exponentially distributed with death rate $\lambda>0$. Reproduction is assumed to occur simultaneously with death.
-
-  2) __Reproductive law:__ At the death of an individual, a branching event occurs where a random (possibly zero) number of offspring are left.
+  1) __Branching rate:__ In our formulation of BBM we assume Lifetimes of individuals are exponentially distributed with death rate $\lambda>0$ and reproduction occurs simultaneously with death.
+Biologically, this implies individuals are semelparous An alternative formulation treats birth and death events separately to model iteroparity.
+However, under the appropriate rescaling, both approaches converge to the same diffusion limit.
+We therefore choose the former approach for the sake of simplicity.
+
+  2) __Reproductive law:__ When a birth event occurs a random (possibly zero) number of offspring are left.
 The distribution of offspring left is called the reproductive law or branching mechanism.
@@
 The case of $\mathscr{W}=1$ is referred to as the critical condition.
+Under the critical condition the probability that extinction occurs in finite time is equal to one.
 
   3) __Spatial movement:__ Each offspring is born at the current location of their parent.
-Immediately after birth they move around space according to Brownian motion with diffusion parameter $\sqrt\mu$.
+Immediately after birth they move around space according to $d$-dimensional Brownian motion with diffusion parameter $\sqrt\mu$.
 In our context we interpret spatial movement as mutation so that the location of an individual at death represents the value of its phenotype.
@@
 Denoting the rescaled process by $X_t^{(n)}(D)$, the limiting process $\mathscr{X}_t=\lim_{n\to\infty}X_t^{(n)}$ is called a super-Brownian motion and is also a non-negative finite measure-valued process [@Watanabe1968].
-Instead of returning the number of individuals alive in a region of space, super-Brownian motion returns the _mass_ of the population concentrated in a region of space.
-In particular, the value of $\mathscr{X}_t(D)$ is generally a continuously varying non-negative random variable for any $t\geq0$ and $D\subset\mathbb{R}^d$.
+However, instead of returning the number of individuals alive in a region of space, super-Brownian motion returns the _mass_ of the population concentrated in a region of space.
+Since we have rescaled individual mass by $1/n$ and took the limit $n\to\infty$, individuals are no longer discrete units.
+The particle view of the population gets replaced by a continuous blob spread across $d$-dimensional space.
+In particular, the value of $\mathscr{X}_t(D)$ is a continuously varying non-negative random variable for any $t\geq0$ and $D\subset\mathbb{R}^d$.
 It turns out that for spatial dimension $d=1$, $\mathscr{X}_t$ is absolutely continuous with respect to the Lebesgue measure for each $t\geq0$ [@Konno1988; @Reimers1989].
@@
 \end{equation}
-Since $\nu(x,t)$ is not generally differentiable in $x$ or $t$, the derivatives in expression (\ref{neutral_SPDE}) are taken in the weak sense.
+Since $\nu(x,t)$ is not generally differentiable in $x$ or $t$, the derivatives in expression (\ref{neutral_SPDE}) are taken in the _weak_ sense.
 That is, to rigorously interpret SPDE (\ref{neutral_SPDE}), we integrate the solution $\nu(x,t)$ against functions $f\in C_b^2(\mathbb{R})$, where $C_b^2(\mathbb{R})$ is the set of bounded and twice continuously differentiable functions on $\mathbb{R}$.
@@
 \begin{multline}\label{neutral_int}
-\int_\mathbb{R}\nu(x,t)f(x)dx-\int_\mathbb{R}\nu(x,0)f(x)dx \\
-=\int_0^t\int_\mathbb{R}\nu(x,s)\left(mf(x)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}f(x)\right)dsdx+\int_0^t\int_\mathbb{R}f(x)\sqrt{V\nu(x,s)}\dot W(x,s)dxds.
-\end{multline}
+\int_\mathbb{R}\nu(x,t)f(x)dx-\int_\mathbb{R}\nu(x,0)f(x)dx=\int_0^t\int_\mathbb{R}\nu(x,s)\left(mf(x)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}f(x)\right)dsdx \\
++\int_0^t\int_\mathbb{R}f(x)\sqrt{V\nu(x,s)}\dot W(x,s)dxds, \ \forall \ f\in C_b^2(\mathbb{R}).
+\end{multline}
+Evans [-@lawrenceevans2010] provides a nice introduction to weak derivatives.
 For more on the general theory of SPDE see Walsh [-@Walsh].
@@
 
-n_1 = read.csv("/home/bob/Research/Branching Brownian Motion/n_1.csv")
-n_2 = read.csv("/home/bob/Research/Branching Brownian Motion/n_2.csv")
-n_3 = read.csv("/home/bob/Research/Branching Brownian Motion/n_3.csv")
+n_1 = read.csv("/home/bb/Gits/branching.brownian.motion.and.spde/n_1.csv")
+n_2 = read.csv("/home/bb/Gits/branching.brownian.motion.and.spde/n_2.csv")
+n_3 = read.csv("/home/bb/Gits/branching.brownian.motion.and.spde/n_3.csv")
 
@@
 
-ggsave("~/Research/A White Noise Approach To Evolutionary Ecology/rescaled_plots.png",rescaled_plots,device="png",width=6,height=6,units="in",dpi=350)
+ggsave("~/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/rescaled_plots.png",rescaled_plots,device="png",width=6,height=6,units="in",dpi=350)
 ```
@@
 Hence, contrary to the common assumption that stochastic effects can be ignored for large populations, we find that minute asymmetries by demographic stochasticity remain significant drivers of community structure.
-In particular, although we initialized the species with identical state variables and model parameters, we find an enourmous amount of asymmetry and even some potential phase changes.
+In particular, we initialized the species with identical state variables and model parameters, but found an enourmous amount of asymmetry and even some potential phase changes.
 In the following two paragraphs we describe the natural history of the community as illustrated in Figure \ref{temporal}.
@@
 However, the abundances across all species in the community are very low due to strength of competition being an order of magnitude higher than in the weak competition case.
-Although most of the species maintain abundances greater than $1000.0$, we found one species that dropped to an abundance of about $50.0$.
+Most of the species maintain abundances greater than $1000.0$, but we found one species that dropped to an abundance of about $50.0$.
 If we let the numerical integrator run long enough in this case, we will likely see many of the species go extinct.
 
-Although finding ways to interpret simulated dynamics provides a useful arena to exercise biological reasoning, it does not fulfill our desire to quantify the patterns and processes present in competing communities.
+Finding ways to interpret simulated dynamics provides a useful arena to exercise biological reasoning.
+However, it does not fulfill our desire to quantify the patterns and processes present in competing communities.
 In the next subsection we take a step in this direction by using our model to derive formula for selection gradients and competition coefficients.
@@
 
-smpl_wc = read.csv("/home/bob/Research/Branching Brownian Motion/sample_path_wc.csv")
+smpl_wc = read.csv("/home/bb/Gits/branching.brownian.motion.and.spde/sample_path_wc.csv")
 
@@
 
-smpl_mc = read.csv("/home/bob/Research/Branching Brownian Motion/sample_path_mc.csv")
+smpl_mc = read.csv("/home/bb/Gits/branching.brownian.motion.and.spde/sample_path_mc.csv")
 
@@
 
-ggsave("~/Research/A White Noise Approach To Evolutionary Ecology/community_dynamics.png",community_plot,width=6.5,height=5,dpi=320)
+ggsave("~/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/community_dynamics.png",community_plot,width=6.5,height=5,dpi=320)
 
@@
 
-ggsave("~/Research/A White Noise Approach To Evolutionary Ecology/on_pl.png",on_pl,device="png",width=3.5,height=3.5,units="in",dpi=350)
+ggsave("~/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/on_pl.png",on_pl,device="png",width=3.5,height=3.5,units="in",dpi=350)
 
@@
 
-ggsave("~/Research/A White Noise Approach To Evolutionary Ecology/interaction_networks.png",coevs,device="png",width=6,height=6,units="in",dpi=350)
+ggsave("~/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/interaction_networks.png",coevs,device="png",width=6,height=6,units="in",dpi=350)
 
@@
 cov_plot <- grid.arrange(cov_ab_p,cov_ag_p,nrow=1) 
-ggsave("~/Research/A White Noise Approach To Evolutionary Ecology/cov.png",cov_plot,width=6.5,height=2.5)
+ggsave("~/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/cov.png",cov_plot,width=6.5,height=2.5)
 ```
@@
 In both plots we set $c=1.0\\times10^{-4}$, $b=0.1$, $\\bar r=0.1$ and $N=1.0\\times10^{10}$ and let $V_{\\bar X}$ vary.'}
-knitr::include_graphics('/home/bob/Research/A White Noise Approach To Evolutionary Ecology/cov.png')
+knitr::include_graphics('/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/cov.png')
 ```
@@
 corrsp = grid.arrange(Cαabsβp,Cαγp,nrow=1)
-ggsave("/home/bob/Research/A White Noise Approach To Evolutionary Ecology/corrs.png",  corrsp,  width=6.5,height=2.5)
-ggsave("/home/bob/Research/A White Noise Approach To Evolutionary Ecology/Cαℭp.png",   Cαℭp,  width=4,height=2.5)
+ggsave("/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/corrs.png",  corrsp,  width=6.5,height=2.5)
+ggsave("/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/Cαℭp.png",   Cαℭp,  width=4,height=2.5)
 ```
@@
 The red line is a cubic regression.'}
-knitr::include_graphics('/home/bob/Research/A White Noise Approach To Evolutionary Ecology/Cαℭp.png')
+knitr::include_graphics('/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/Cαℭp.png')
 ```
@@
 We have introduced an approach to derive models of evolutionary ecology using the calculus of white noise, demonstrated our approach by deriving a model of diffuse coevolution and investigated the relationship between selection gradients and competition coefficients, finding that these quantities exhibit interesting relationships which shed light on the feedback between the structure and dynamics of ecological communities.
-However, although our approach has the merit of rigorously synthesizing the dynamics of abundance and phenotypic evolution for populations experiencing demographic stochasticity, there remains biological details and their associated technical challenges that have yet to be confronted for gaining a more thorough and rigorous understanding of ecological communities.
+
+Our approach has the merit of rigorously synthesizing the dynamics of abundance and phenotypic evolution for populations experiencing demographic stochasticity.
+Yet, there remains biological details and their associated technical challenges that need to be confronted for gaining a more thorough and rigorous understanding of ecological communities.
 We touch on just four of them here and provide suggestions for future research to approach these challenges.
@@
 
-As noted by Feller [-@feller1951], although diffusion limits provide reasonable approximations for large populations, relatively small populations require discrete models.
+As noted by Feller [-@feller1951], diffusion limits provide reasonable approximations for large populations, but relatively small populations require discrete models.
 Hence, as a diffusion limit, SPDE (\ref{SPDE}) restricts parameter values to regions that maintain large population sizes.
@@
 Tracking the fluctuations of allele frequencies then allows theoreticians to investigate the dynamics of phenotypic distributions that deviate from normality.
-While this approach has a long history in theoretical quantitative genetics [@bulmer1980; @pmid7851785; @Kirkpatrick1727] and coevolutionary theory [@Nuismer2005; @Kopp2006; @Nuismer2007], work to investigate the role of genetic architecture in mediating feedbacks between the dynamics of population abundances and the distributions of traits mediating ecological interactions has apparently only just begun [@Schreiber2018; @Patel2019].
+This approach has a long history in theoretical quantitative genetics [@bulmer1980; @pmid7851785; @Kirkpatrick1727] and coevolutionary theory [@Nuismer2005; @Kopp2006; @Nuismer2007].
+However, work to investigate the role of genetic architecture in mediating feedbacks between the dynamics of population abundances and the distributions of traits mediating ecological interactions has apparently only just begun [@Schreiber2018; @Patel2019].
 
@@
 \begin{equation}
-\sigma^2(0)=\int_{\mathbb{R}}xp(x,0)dx<+\infty,
+\sigma^2(0)=\int_{\mathbb{R}}(x-\bar x(0))^2p(x,0)dx<+\infty,
 \end{equation}
@@
 ```{r, echo=F, fig.align='center', fig.cap='Numerical estimate for the correlations of selection gradients and competition coefficients.'}
-knitr::include_graphics('/home/bob/Research/A White Noise Approach To Evolutionary Ecology/corrs.png')
+knitr::include_graphics('/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/corrs.png')
 ```

### 2020-03-29 sde_num_sol.jl

@@ -1,7 +1,7 @@
 using Parameters, Statistics, Random, LinearAlgebra, Distributions,
 	DifferentialEquations, StatsBase, StatsPlots, Plots, DataFrames, CSV, Optim
 
-include("/home/bob/Research/Branching Brownian Motion/sde_functions.jl")
+include("/home/bb/Gits/branching.brownian.motion.and.spde/sde_functions.jl")
 
 j=1
 k=2

### 2020-03-30 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
intXprts

@@
 We include an account of the natural history of the simulated system and discuss the significance of demographic stochasticity for structuring ecological communities.
-In \S\ref{ecoevo} we provide formula for selection gradients and competition coefficients implied by our model and use these to investigate the relationship between the strengths of competition and coevolution.
+In \S\ref{ecoevo} we provide expressions for selection gradients and competition coefficients implied by our model and use these expressions to investigate the relationship between the strengths of competition and coevolution.
 Finally, \S\ref{conclusion} concludes with a summary of accomplishments, a few remarks on the limits of this approach and future directions to incorporate more explicitly the genetic architecture of populations, feedbacks with ecosystem processes and the macroevolutionary history of interacting lineages.
@@
 where $\bar x(t)$ and $\sigma^2(t)$ are respectively the mean trait and phenotypic variance at time $t\geq0$.
-We denote by $C_{1,c}^+(\mathbb{R}\times[0,\infty))$ the set of all non-negative integrable functions that are continuous in $x$ and $t$.
-Second, we assume selection is determined by the growth rate $m(\nu,x)$ that is differentiable with respect to $x\in\mathbb{R}$, continuous with respect to $\nu\in C^+_{1,c}(\mathbb{R}\times[0,\infty))$ and satisfies $m(\nu,x)\leq r$ for some $r\in\mathbb{R}$ and all $x\in\mathbb{R}$ and $\nu\in C_{1,c}^+(\mathbb{R}\times[0,\infty))$.
-Our continuity assumption implies; if a sequence of abundance densities $\{\nu_n\}_{n=1}^\infty\subset C^+_{1,c}(\mathbb{R}\times[0,\infty))$ satisfies $\lim_{n\to\infty}\nu_n=\nu\in C^+_{1,c}(\mathbb{R}\times[0,\infty))$, then $\lim_{n\to\infty}m(\nu_n,x)=m(\nu,x)$ for all $x\in\mathbb{R}$.
+Second, we assume selection is determined by the growth rate $m(\nu(x,t),x)$ that is differentiable with respect to both arguments and satisfies $m(y,x)\leq r$ for some $r\in\mathbb{R}$ and for all $x\in\mathbb{R}$, $y\geq0$.
+From here on we abbreviate $m(\nu(x,t),x)$ to $m(\nu,x)$.
 Third, we assume mutation is captured by diffusion with coefficient $\frac{\mu}{2}$.
@@
 \end{multline}
-By our assumptions on mutation and rate of growth, $\nu(x,t)$ is twice differentiable with respect to $x$ and $\int_{-\infty}^{+\infty}\nu(x,t)dx<\infty$ for all $t\geq0$ (SM \S\ref{}).
+Using our assumptions on mutation and rate of growth, we show in SM \S\ref{} $\nu(x,t)$ is twice differentiable with respect to $x$ and $\int_{-\infty}^{+\infty}\nu(x,t)dx<\infty$ for all $t\geq0$.
 This implies that we are justified in swapping the order of differentiation and integration and the result $\int_{-\infty}^{+\infty}\frac{\partial^2}{\partial x^2}\nu(x,t)dx=0$ can be derived from the fundamental theorem of calculus.
@@
 \frac{d}{dt}\bar x(t)=\frac{d}{dt}\int_{-\infty}^{+\infty} x p(x,t)dx=\int_{-\infty}^{+\infty} x\frac{\partial}{\partial t} p(x,t)dx \\
-=\int_{-\infty}^{+\infty} x\big(m(\nu,x)-\bar m(t)\big) p(x,t)+x\frac{\mu}{2}\frac{\partial^2}{\partial x^2} p(x,t)dx \\
-=\mathrm{Cov}_t\Big(m(\nu,x),x\Big)+\frac{\mu}{2}\int_{-\infty}^{+\infty} x\frac{\partial^2}{\partial x^2} p(x,t)dx.
-\end{multline}
-
-Equation (\ref{detmean}) is a continuous time analog of the well known Robertson-Price equation [@Robertson1966; @PRICE1970; @FRANK2012; @Queller2017; @Lion2018].
+=\int_{-\infty}^{+\infty} x\big(m(\nu,x)-\bar m(t)\big) p(x,t)+x\frac{\mu}{2}\frac{\partial^2}{\partial x^2} p(x,t)dx
+=\mathrm{Cov}_t\Big(m(\nu,x),x\Big).
+\end{multline}
+
+Equation (\ref{detmean}) is a continuous time analog of the well known Robertson-Price equation without transmission bias [@Robertson1966; @PRICE1970; @FRANK2012; @Queller2017; @Lion2018].
 The covariance of fitness and phenotype creates change in $\bar x$ to maximize mean fitness $\bar m$.
@@
 In particular, when $\sigma^2=0$, $\bar x$ will not respond to selection.
-The second term, due to mutation, represents the so-called _transmission bias_ [@FRANK2012].
-When $p(x,t)$ is symmetric around $\bar x$, this term disappears due to equal bias in opposite directions cancelling the effect of mutation on $\bar x$.
+The result $\int_{-\infty}^{+\infty} x\frac{\partial^2}{\partial x^2} p(x,t)dx=0$ can be found by applying integration by parts.
 Following the approach taken to calculate the evolution of $\bar x$, we find the response of phenotypic variation to this model of selection and mutation is
 \begin{equation}\label{vardyn_det}
-\frac{d}{dt}\sigma^2(t)=\mathrm{Cov}_t\Big(m(\nu,x),(x-\bar x)^2\Big)+\frac{\mu}{2}\int_{-\infty}^{+\infty}(x-\bar x)^2\frac{\partial^2}{\partial x^2}p(x,t)dx.
+\frac{d}{dt}\sigma^2(t)=\mathrm{Cov}_t\Big(m(\nu,x),(x-\bar x)^2\Big)+\mu.
 \end{equation}
@@
 This covariance also creates change in $\sigma^2$ that maximizes mean fitness $\bar m$, which can be positive or negative depending on whether selection is stabilizing or disruptive.
-In SM \S\ref{var_deriv} we show when $p(x,t)$ is a Gaussian bell curve, the increase in variance due to mutation simplifies to $\mu$.
 
@@
 Thus our notation $\dot W$ is only meant to aid intuition and not be taken as formal.
-A formal understanding is possible by considering white noise as a _measure_-valued process [@Dawson1975; @Walsh] or as a _generalized_ process that acts on classically defined functions or stochastic processes to return random variables or stochastic processes [@Krylov1981; @DaPrato2014].
+A formal understanding is possible by considering white noise as a _measure_-valued process [@Dawson1975; @Walsh] or as a _generalized_ process that acts on classically defined functions or stochastic processes to return either random variables or stochastic processes [@Krylov1981; @DaPrato2014].
 Since a measure-valued process can be defined from a generalized process and a generalized process can be defined from a measure-valued process, the distinction between the two is more or less a matter of perspective.
-However, we find the perspective of white-noise as a generalized process to be a more efficient route for developing heuristics to help with some routine calculations.
-Hence, this latter approach provides the general idea implemented here.
-Although the treatments provided by Krylov and Rozovskii [-@Krylov1981] and Da Prato and Zabczyk [-@DaPrato2014] extend the theory of SDE to formally treat SPDE in a general and elegant fashion, they require the navigation of an enourmous amount of technical definitions and detailed proofs.
+However, we find the perspective of white-noise as a generalized process to be a more efficient route for developing heuristics to help with some routine calculations involved with deriving SDE from SPDE.
+Hence, the notion of a generalized process provides the general idea implemented here.
+Although the treatments of SPDE provided by Krylov and Rozovskii [-@Krylov1981] and Da Prato and Zabczyk [-@DaPrato2014] extend the theory of SDE to formally treat SPDE in a general and elegant fashion, they require the navigation of an enourmous amount of technical definitions and detailed proofs.
 To extract some particularly useful results from this theory relevant to our goal of synthesizing the stochastic dynamics of biological populations, we provide a streamlined approach by capitolizing on the solid ground these authors have established.
@@
 
-To shed a bit of light on the idea of a generalized process, we note that they are the stochastic analog of a generalized function, such as the Dirac delta function $\delta$.
+Before diving in, we shed a bit of light on the idea of a generalized process.
+A generalized process is the stochastic analog of a generalized function, such as the Dirac delta function $\delta$.
 Just as a generalized function operates on classical functions to return a value (e.g., $\delta(f)=f(0)$), a generalized process acts on a set of functions to return a classically defined stochastic process.
-For a breif overview of the theory of generalized functions, see the addendum to chapter 3 of Kolmogorov and Fomin [-@kolmogorov1999elements].
+For a breif primer on the theory of generalized functions, see the addendum to chapter 3 of Kolmogorov and Fomin [-@kolmogorov1999elements].
 
 Throughout this section, we minimize notation by writing $\int_{\mathbb{R}}f(x)dx=\int_{-\infty}^{+\infty}f(x)dx$ and similarly $\int_Df(x)dx$ for the integral of $f$ over $D\subset\mathbb{R}$.
-We define $\mathscr{N}_2$ as the set of stochastic processes $f(x,t)$ that are continuous in both $x$ and $t$ and satisfy $\mathbb{E}\left(\int_0^t\int_\mathbb{R}f^2(x,s)dxds\right)<+\infty$ for each $t\geq0$.
+We define $\mathscr{N}_2$ as the set of stochastic processes $f(x,t)$ that are continuous in $t$ and satisfy $\mathbb{E}\left(\int_0^t\int_\mathbb{R}f^2(x,s)dxds\right)<+\infty$ for each $t\geq0$.
 The operator $\mathbb{E}$ denotes expectation with respect to the underlying probability space.
@@
 \begin{equation}
-\|f\|_t=\sqrt{\mathbb{E}\left(\int_0^t\int_\mathbb{R}f^2(x,s)dxds\right)}.
-\end{equation}
+\|f\|_t=\sqrt{\mathbb{E}\left(\int_0^t\int_\mathbb{R}f^2(x,s)dxds\right)},
+\end{equation}
+
+and make use of the convention $f=g$ if $\|f-g\|_t=0$ for all $t\geq0$.
 
@@
 where $t_1\wedge t_2=\min(t_1,t_2)$ and $\mathbb{C}$ denotes covariance with respect to the underlying probability space.
+In particular, denoting $\mathbb{V}$ the variance operator with respect to the underlying probability space, we have $\mathbb{V}\big(\mathbf W_t(f)\big)=\|f\|_t^2$ for all $t\geq0$ and $f\in\mathscr{N}_2$.
 The operators $\mathbb{E}$ and $\mathbb{C}$ are to be distinguished from $\bar f(t)$ and $\mathrm{Cov}_t(f,g)$ which denote expectation and covariance with respect to phenotypic diversity at time $t\geq0$.
 
-Since Gaussian processes are characterized by their expectations and covariances and since we assume the $\mathscr{N}_2$ processes are continuous in both time and space, the processes $\mathbf W(f)$ and $\mathbf W(g)$ are well defined.
-In particular, if $f\in\mathscr{N}_2$ is independent of time, then $\mathbf W(f)$ is a Brownian motion with variance at time $t\geq0$ equal to $t \ \mathbb{E}(\int_\mathbb{R}f^2(x,0)dx)$.
+Since Gaussian processes are characterized by their expectations and covariances and since we assume the $\mathscr{N}_2$ processes are continuous in time, the processes $\mathbf W(f)$ and $\mathbf W(g)$ are well defined.
+As an example, if $f\in\mathscr{N}_2$ is independent of time, then $\mathbf W(f)$ is a Brownian motion with variance at time $t\geq0$ equal to $\|f\|_t^2=t \ \mathbb{E}(\int_\mathbb{R}f^2(x,0)dx)$.
 With the generalized process $\mathbf W$ defined, we define the space-time white noise $\dot W(x,t)$ implicitly via the stochastic integral
-\begin{equation}
+\begin{equation}\label{informal}
 ``\int_0^t\int_\mathbb{R}f(x,s)\dot W(x,s)dxds"=``\int_\mathbb{R}\int_0^tf(x,s)\dot W(x,s)dxds"=\mathbf W_t(f), \ \forall \ f\in\mathscr{N}_2, \ t\geq0.
@@
 Following this definition of white noise, we compute its value by sampling it using $\mathscr{N}_2$ processes.
-For example, integrating white noise over a region $D\times[0,t]$, with $t>0$ and $D$ a bounded subset of $\mathbb{R}$, is equivalent to evaluating $\mathbf W_t(I_{D\times [0,\infty)})$ where
-\begin{equation}
-I_{D\times [0,\infty)}(x,t)=\left\{\begin{matrix}
+For example, integrating white noise over a region $D\times[0,t]$, with $t>0$ and $D$ a bounded subset of $\mathbb{R}$, is equivalent to evaluating $\mathbf W_t(I_{D\times [0,+\infty)})$ for the deterministic process
+\begin{equation}
+I_{D\times [0,+\infty)}(x,t)=\left\{\begin{matrix}
 0, & x\notin D \\
@@
 \end{equation}
-Since $\int_\mathbb{R}I_{D\times [0,\infty)}^2(x,t)dx=\int_Ddx=|D|$, where $|D|$ denotes the length of $D$, $I_{D\times [0,\infty)}\in L^2_c(\mathbb{R}\times[0,\infty))$.
-Thus, by equations (\ref{exp_WN}) and (\ref{cov_WN}), we have
+Since 
+\begin{equation}
+\|I_{D\times[0,+\infty))}\|_t^2=\mathbb{E}\left(\int_0^t\int_\mathbb{R}I_{D\times [0,+\infty)}^2(x,s)dxds\right)=t\int_Ddx=t|D|<+\infty, 
+\end{equation}
+where $|D|$ denotes the length of $D$, we have $I_{D\times [0,+\infty)}\in\mathscr{N}_2$.
+Thus, using equations (\ref{exp_WN}) and (\ref{cov_WN}) and adopting the informal notation introduced in equation (\ref{informal}), we can write the following
 \begin{equation}
@@
 \begin{equation}
-\mathbb{V}\left(\int_0^t\int_D\dot W(x,s)dxds\right)=t|D|,
-\end{equation}
-where $\mathbb{V}$ denotes the variance operator with respect to the underlying probability space. Using this notation equations (\ref{exp_WN}) and (\ref{cov_WN}) can be rewritten as
+\mathbb{V}\left(\int_0^t\int_D\dot W(x,s)dxds\right)=t|D|.
+\end{equation}
+Using this informal notation, equations (\ref{exp_WN}) and (\ref{cov_WN}) can be rewritten as
 \begin{equation}\label{exp_WN_xi}
@@
 \end{equation}
-This implies $\mathbb{E}\big(\int_0^td\hat{\mathbf W}_s(f)\big)=0$, $\mathbb{C}\big(\int_0^{t_1}d\hat{\mathbf W}_s(f),\int_0^{t_2}d\hat{\mathbf W}_s(f)\big)=t_1\wedge t_2$ and, as a function of $t$, $\int_0^td\hat{\mathbf W}_s(f)$ is a standard Brownian motion for any $f\in L^2_c(\mathbb{R}\times[0,\infty))$.
+This implies $\mathbb{E}\big(\int_0^td\hat{\mathbf W}_s(f)\big)=0$, $\mathbb{C}\big(\int_0^{t_1}d\hat{\mathbf W}_s(f),\int_0^{t_2}d\hat{\mathbf W}_s(f)\big)=t_1\wedge t_2$ and, as a function of $t$, $\int_0^td\hat{\mathbf W}_s(f)$ is a standard Brownian motion for any $f\in\mathscr{N}_2$.
 Hence, $d\hat{\mathbf W}_t(f)$ is analogous to the traditional shorthand used to denote stochastic differentials.
@@
 Note that since $\nu(x,t)$ is the density of a finite measure, it is integrable for each $t\geq0$.
-Thus, since for some $M>0$, $|f(x)|\leq M$ for every $x\in\mathbb{R}$, setting $\varphi(x,t)=f(x)\sqrt{V\nu(x,t)}$ implies $\varphi\in L^2_c(\mathbb{R}\times[0,\infty))$.
+Thus, since for some $M>0$, $|f(x)|\leq M$ for every $x\in\mathbb{R}$, setting $\varphi(x,t)=f(x)\sqrt{V\nu(x,t)}$ implies $\varphi\in\mathscr{N}_2$.
 Hence, the white noise integral on the right-hand side of equation (\ref{neutral_int}) can be understood using the heuristics introduced above.
@@
 Note that, comparing equation (\ref{SPDE}) to equation (3.5) of Li [-@zeng1998absolute], our $m$ and $V$ correspond to Li's $b$ and $c$ respectively.
-Denoting $C_1^+(\mathbb{R})$ the set of non-negative, continuous, integrable functions on $\mathbb{R}$, it is important to note that, under the assumptions made in Méléard and Roelly [-@meleard1992interacting; -@Mlard1993] and Li [-@zeng1998absolute], equation (\ref{SPDE}) is only formal when $m(\nu,x)$ is bounded across all combinations of $\nu\in C_1^+(\mathbb{R})$ and $x\in\mathbb{R}$.
-However, recalling our condition $m(\nu,x)\leq r\in\mathbb{R}$, the growth rates we consider are only bounded above.
-Replacing $m(\nu,x)$ with its upper-bound demonstrates that the associated superprocess should still admit an integrable density.
+Denoting $C_1^+(\mathbb{R})$ the set of non-negative, continuous, integrable functions on $\mathbb{R}$, it is important to note that, under the assumptions made in Méléard and Roelly [-@meleard1992interacting; -@Mlard1993] and Li [-@zeng1998absolute], equation (\ref{SPDE}) is only formal when $m(y,x)$ is bounded across all combinations of $y\geq0$ and $x\in\mathbb{R}$.
+However, recalling our condition $m(y,x)\leq r\in\mathbb{R}$, the growth rates we consider are only bounded above.
+Replacing $m(y,x)$ with its upper-bound demonstrates that the associated superprocess should still admit an integrable density.
 That is, we should still be able to compute the total mass process $N(t)$.
@@
 \begin{equation}\label{xbar_gen}
-d\bar x(t)=\mathrm{Cov}_t\Big(x,m(\nu,x)\Big)dt+\frac{\mu}{2}\int_{-\infty}^{+\infty}x\frac{\partial^2}{\partial x^2} p(x,t)dx+\sqrt{V\frac{\sigma^2(t)}{N(t)}}dW_2(t),
-\end{equation}
-\begin{multline}\label{sig2_gen}
-d\sigma^2(t)=\left(\mathrm{Cov}_t\Big((x-\bar x(t))^2,m(\nu,x)\Big)+\frac{\mu}{2}\int_{-\infty}^{+\infty}(x-\bar x(t))^2\frac{\partial^2}{\partial x^2} p(x,t)dx-V\frac{\sigma^2(t)}{N(t)}\right)dt \\ +\sqrt{V\frac{\overline{(x-\bar x(t))^4}-\sigma^4(t)}{N(t)}}dW_3(t).
-\end{multline}
+d\bar x(t)=\mathrm{Cov}_t\Big(x,m(\nu,x)\Big)dt+\sqrt{V\frac{\sigma^2(t)}{N(t)}}dW_2(t),
+\end{equation}
+\begin{equation}\label{sig2_gen}
+d\sigma^2(t)=\left(\mathrm{Cov}_t\Big((x-\bar x(t))^2,m(\nu,x)\Big)+\mu-V\frac{\sigma^2(t)}{N(t)}\right)dt+\sqrt{V\frac{\overline{(x-\bar x(t))^4}-\sigma^4(t)}{N(t)}}dW_3(t),
+\end{equation}
 \end{subequations}
@@
 \end{equation}
-Under this assumption, Lande [-@Lande1976] found (suppressing the dependency on $t$)
-\begin{equation}
-\mathrm{Cov}(x,m)=\sigma^2\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right).
-\end{equation}
-Under the same assumption, in SM \S\ref{var_deriv} we show
+Under this assumption we find in SM \S\ref{cov2deriv} the results (suppressing the dependency on $t$)
+\begin{equation}\label{covxm}
+\mathrm{Cov}(x,m)=\sigma^2\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right),
+\end{equation}
 \begin{equation}
@@
 \end{equation}
-and $\overline{(x-\bar x)^4}=3\sigma^4$. In particular, this implies
+and $\overline{(x-\bar x)^4}=3\sigma^4$. Equation (\ref{covxm}) is the continuous time equivalent to equation (9) in Lande [-@Lande1976]. In particular, these results imply
 \begin{subequations}\label{no_inher}
@@
 
-Recall $C_{1,c}^+(\mathbb{R}\times[0,\infty))$ denotes the set of non-negative functions $\nu(x,t)$ continuous in both $x$ and $t$ such that $\int_{\mathbb{R}}\nu(x,t)dx<+\infty$.
-We assume, for $m:C_{1,c}^+(\mathbb{R}\times[0,\infty))\times\mathbb{R}\to\mathbb{R}$, there exists $r\in\mathbb{R}$ such that $m(\nu,x)\leq r$ for each $\nu\in C_{1,c}^+(\mathbb{R}\times[0,\infty))$ and $x\in\mathbb{R}$.
+Recall that $m(\nu,x)$ is shorthand for $m(\nu(x,t),x)$.
+That is, $m:[0,+\infty)\times\mathbb{R}\to\mathbb{R}$.
+Following our assumptions of the main text, we have that $m(y,x)$ is differentiable with respect to both $x$ and $y$ and there exists $r\in\mathbb{R}$ such that $m(y,x)\leq r$ for each $y\geq0$ and $x\in\mathbb{R}$.
+Hence, we can apply Proposition 1.21 of Cantrell and Cosner [-@Cantrell2004] to show that $\nu(x,t)$ remains differentiable with respect to $t$ and twice differentiable with respect to $x$ for all $t\geq0$.
+
 As in the main text, we also assume the initial condition $\nu(x,0)$ satisfies
@@
 \end{equation}
+
 Replacing $m$ with it's upper bound $r\in\mathbb{R}$, equation (\ref{PDE_SM}) reduces to a simple parabolic equation that can be solved using standard techniques [@stanleyfarlow1993].
@@
 \end{equation}
-which implies there exists $\nu\in C^+_{1,c}(\mathbb{R}\times[0,\infty))$ and $x\in\mathbb{R}$ such that $m(\nu,x)>r$.
-But this contradicts our assumption $m(\nu,x)\leq r$ for all $\nu\in C^+_{1,c}(\mathbb{R}\times[0,\infty))$ and $x\in\mathbb{R}$.
+which implies there exists $y\geq0$ and $x\in\mathbb{R}$ such that $m(y,x)>r$.
+But this contradicts our assumption $m(y,x)\leq r$ for all $y\geq0$ and $x\in\mathbb{R}$.
 So we have $\nu(x,t)\leq\nu_r(x,t)$ for each $x\in\mathbb{R}$ and $t\geq0$.
@@
 \begin{equation}
-0<\int_\mathbb{R}x^2\nu(x,t)dx\leq\int_\mathbb{R}x^2\nu_r(x,t)dx<+\infty
+0\leq\int_\mathbb{R}x^2\nu(x,t)dx\leq\int_\mathbb{R}x^2\nu_r(x,t)dx<+\infty
 \end{equation}
@@
 \begin{equation}
-0<\sigma^2(t)+\bar x^2(t)=\frac{1}{N(t)}\int_\mathbb{R}x^2\nu(x,t)dx<+\infty
+0\leq\sigma^2(t)+\bar x^2(t)=\frac{1}{N(t)}\int_\mathbb{R}x^2\nu(x,t)dx<+\infty
 \end{equation}
@@
 
-Picking up from the main text \S2.1,
-
-\begin{equation}
-\dot{\sigma}^2(t)=\frac{d}{dt}\int_\mathbb{R}(x-\bar x(t))^2p(x,t)dx
-\end{equation}
-
-$$=\int_\mathbb{R}2(x-\bar x(t))\dot{\bar x}(t)+(x-\bar x(t))^2\dot p(x,t)dx$$
-
-$$=\int_\mathbb{R}(x-\bar x(t))^2\left((m(\nu,x)-\bar m(t))p(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t)\right)dx$$
-
-$$=\int_\mathbb{R}\left((x-\bar x(t))^2-\sigma^2(t)+\sigma^2(t)\right)(m(\nu,x)-\bar m(t))p(x,t)+(x-\bar x(t))^2\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t)dx$$
-
-$$=\mathrm{Cov}_t\Big((x-\bar x(t))^2,m(\nu,x)\Big)+\frac{\mu}{2}\int_\mathbb{R}(x-\bar x(t))^2\frac{\partial^2}{\partial x^2}p(x,t)dx.$$
-
-In particular, when $p(x,t)$ is Gaussian,
-
-\begin{equation}
-\frac{\partial^2}{\partial x^2}p(x,t)=\frac{(x-\bar x(t))^2-\sigma^2(t)}{\sigma^4(t)}p(x,t)
-\end{equation}
-
-and hence, referring to the well-known moments of the Gaussian distribution, we find
-
-\begin{equation}
-\int_\mathbb{R}(x-\bar x(t))^2\frac{\partial^2}{\partial x^2}p(x,t)dx
-\end{equation}
-
-$$=\int_\mathbb{R}\frac{(x-\bar x(t))^4-(x-\bar x(t))^2\sigma^2(t)}{\sigma^4(t)}p(x,t)dx=2.$$
-
-In the Gaussian case, we also have
-
-\begin{equation}
-2\sigma^4\left(\frac{\partial\bar m}{\partial\sigma^2}-\overline{\frac{\partial m}{\partial\sigma^2}}\right)=2\sigma^4\left(\frac{\partial}{\partial\sigma^2}\int_\mathbb{R}m(\nu,x)p(x,t)dx-\int_\mathbb{R}p(x,t)\frac{\partial}{\partial\sigma^2}m(\nu,x)dx\right)
-\end{equation}
-
-$$=2\sigma^4\int_\mathbb{R}\frac{(x-\bar x)^2-\sigma^2}{2\sigma^4}m(\nu,x)p(x,t)dx=\int_\mathbb{R}\left((x-\bar x)^2-\sigma^2\right)\big(m(\nu,x)-\bar m\big)p(x,t)dx$$
-
-$$=\mathrm{Cov}_t\Big((x-\bar x)^2,m\Big).$$
+Picking up from the main text \S\ref{deterministic},
+
+\begin{multline}
+\dot{\sigma}^2(t)=\frac{d}{dt}\int_\mathbb{R}(x-\bar x(t))^2p(x,t)dx=\int_\mathbb{R}2(x-\bar x(t))\dot{\bar x}(t)+(x-\bar x(t))^2\dot p(x,t)dx\\=\int_\mathbb{R}(x-\bar x(t))^2\left((m(\nu,x)-\bar m(t))p(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t)\right)dx\\=\int_\mathbb{R}\left((x-\bar x(t))^2-\sigma^2(t)+\sigma^2(t)\right)(m(\nu,x)-\bar m(t))p(x,t)+(x-\bar x(t))^2\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t)dx\\=\mathrm{Cov}_t\Big((x-\bar x(t))^2,m(\nu,x)\Big)+\frac{\mu}{2}\int_\mathbb{R}(x-\bar x(t))^2\frac{\partial^2}{\partial x^2}p(x,t)dx.
+\end{multline}
+
+Applying integration by parts twice yields
+
+\begin{equation}
+\int_{-\infty}^{+\infty}(x-\bar x(t))^2\frac{\partial^2}{\partial x^2}p(x,t)dx=2.
+\end{equation}
+
+## Simplifying covariances with fitness under the assumption of a Gaussian density \label{cov2deriv}
+
+By assuming
+
+\begin{equation}
+p(x,t)=\frac{\exp\left(-\frac{(x-\bar x(t))^2}{2\sigma^2(t)}\right)}{\sqrt{2\pi\sigma^2(t)}}
+\end{equation}
+
+we have
+
+\begin{multline}
+\sigma^2\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right)=\sigma^2\left(\frac{\partial}{\partial\bar x}\int_\mathbb{R}m(\nu,x)p(x,t)dx-\int_\mathbb{R}p(x,t)\frac{\partial}{\bar x}m(\nu,x)dx\right)\\=\sigma^2\int_\mathbb{R}m(\nu,x)\frac{\partial}{\partial\bar x}p(x,t)dx=\sigma^2\int_\mathbb{R}\frac{x-\bar x(t)}{\sigma^2}m(\nu,x)p(x,t)dx\\=\int_\mathbb{R}(x-\bar x)(m(\nu,x)-\bar m)p(x,t)dx=\mathrm{Cov}_t(m,x),
+\end{multline}
+
+and
+
+\begin{multline}
+2\sigma^4\left(\frac{\partial\bar m}{\partial\sigma^2}-\overline{\frac{\partial m}{\partial\sigma^2}}\right)=2\sigma^4\left(\frac{\partial}{\partial\sigma^2}\int_\mathbb{R}m(\nu,x)p(x,t)dx-\int_\mathbb{R}p(x,t)\frac{\partial}{\partial\sigma^2}m(\nu,x)dx\right)\\=2\sigma^4\int_\mathbb{R}\frac{(x-\bar x)^2-\sigma^2}{2\sigma^4}m(\nu,x)p(x,t)dx=\int_\mathbb{R}\left((x-\bar x)^2-\sigma^2\right)\big(m(\nu,x)-\bar m\big)p(x,t)dx\\=\mathrm{Cov}_t\Big((x-\bar x)^2,m\Big).
+\end{multline}
 

### 2020-04-02 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
_

@@
 \end{equation}
-To relate these formulae to the common notation used for SDE, we write $d\hat{\mathbf W}_t(f)=\frac{1}{\|f\|_2(t)}\left(\int_\mathbb{R}f(x,t)\dot W(x,t)dx\right)dt$ so that
+To relate these formulae to the common notation used for SDE, we write 
+\begin{equation}
+d\hat{\mathbf W}_t(f)=\frac{1}{\|f\|_t}\left(\int_\mathbb{R}f(x,t)\dot W(x,t)dx\right)dt,
+\end{equation}
+so that
 \begin{equation}
@@
              &                            &                            &      \\
-$d\hat{\mathbf W}_t(f)$ & $dt$                       & $\frac{\left(\int_\mathbb{R}f \ g \ dx\right)dt}{\|f\|_2\|g\|_2}$ & $0$  \\
+$d\hat{\mathbf W}_t(f)$ & $dt$                       & $\frac{\left(\int_\mathbb{R}f \ g \ dx\right)dt}{\|f\|_t\|g\|_t}$ & $0$  \\
              &                            &                            &      \\
-$d\hat{\mathbf W}_t(g)$ & $\frac{\left(\int_\mathbb{R}f \ g \ dx\right)dt}{\|f\|_2\|g\|_2}$ & $dt$                       & $0$  \\
+$d\hat{\mathbf W}_t(g)$ & $\frac{\left(\int_\mathbb{R}f \ g \ dx\right)dt}{\|f\|_t\|g\|_t}$ & $dt$                       & $0$  \\
              &                            &                            &      \\
@@
 \end{multline}
+This expression is referred to as the _mild_ solution of (\ref{neutral_SPDE}).
 Evans [-@lawrenceevans2010] provides a nice introduction to weak derivatives.
-For more on the general theory of SPDE see Walsh [-@Walsh].
+For more on the general theory of SPDE see Walsh [-@Walsh] and Da Prato and Zabczyk [-@DaPrato2014].
 Note that since $\nu(x,t)$ is the density of a finite measure, it is integrable for each $t\geq0$.
@@
 Konna and Shiga [-@Konno1988] and Reimers [-@Reimers1989] independently proved Dawson's suggestion was indeed correct.
-The diffusion limit of this more general branching process (in arbitrary spatial dimension) is referred to as the Dawson-Watanabe superprocess [@alisonetheridge2000].
-Conditioning the Dawson-Watanabe superprocess to have constant mass returns the Fleming-Viot process [@Etheridge1991; @Perkins1991] which has been popular in studies of spatial population genetics.
-In particular, an extension of the Fleming-Viot process, known as the $\Lambda$-Fleming-Viot process, was introduced by Bertoin and Le Gall [-@Bertoin2003] and coined by Etheridge [-@Etheridge2008] where it was used to resolve some technical challenges in modelling isolation by distance [@Felsenstein1975; see also @Barton2013; and @Barton2019].
+The diffusion limit of this more general branching process (in arbitrary spatial dimension) is referred to as a Dawson-Watanabe superprocess [@alisonetheridge2000].
+Conditioning a Dawson-Watanabe superprocess to have constant mass returns a Fleming-Viot process [@Etheridge1991; @Perkins1991] which has been popular in studies of spatial population genetics.
+In particular, an extension of Fleming-Viot processes, known as $\Lambda$-Fleming-Viot processes, were introduced by Bertoin and Le Gall [-@Bertoin2003] and coined by Etheridge [-@Etheridge2008] where it was used to resolve some technical challenges in modelling isolation by distance [@Felsenstein1975; see also @Barton2013; and @Barton2019].
 Although this provides an impressive list of accomplishments, the Dawson-Watanabe superprocess falls short of our needs.
@@
 \end{equation}
-Note that, comparing equation (\ref{SPDE}) to equation (3.5) of Li [-@zeng1998absolute], our $m$ and $V$ correspond to Li's $b$ and $c$ respectively.
-Denoting $C_1^+(\mathbb{R})$ the set of non-negative, continuous, integrable functions on $\mathbb{R}$, it is important to note that, under the assumptions made in Méléard and Roelly [-@meleard1992interacting; -@Mlard1993] and Li [-@zeng1998absolute], equation (\ref{SPDE}) is only formal when $m(y,x)$ is bounded across all combinations of $y\geq0$ and $x\in\mathbb{R}$.
+Comparing equation (\ref{SPDE}) to equation (3.5) of Li [-@zeng1998absolute], our $m$ and $V$ correspond to Li's $b$ and $c$ respectively.
+It is important to note that, under the assumptions made in Méléard and Roelly [-@meleard1992interacting; -@Mlard1993] and Li [-@zeng1998absolute], equation (\ref{SPDE}) is only formal when $m(y,x)$ is bounded across all combinations of $y\geq0$ and $x\in\mathbb{R}$.
 However, recalling our condition $m(y,x)\leq r\in\mathbb{R}$, the growth rates we consider are only bounded above.
-Replacing $m(y,x)$ with its upper-bound demonstrates that the associated superprocess should still admit an integrable density.
-That is, we should still be able to compute the total mass process $N(t)$.
-In fact, in the proof of the construction of the interacting superprocess as the limit of rescaled branching diffusions, Méléard and Roelly [-@meleard1992interacting; -@Mlard1993] assumed $m(\nu,x)$ to be bounded to guarantee the total mass process will have finite mean and variance.
-This allowed them to employ a tightness criterion for sequences of measures and show the rescaled processes converge to a superprocess with finite total mass.
-Li's [-@zeng1998absolute] result builds directly off of Méléard and Roelly's construction, inheriting the assumption of boundedness for $m(\nu,x)$.
-However, in Li [-@zeng1998absolute], the sufficiency of $m(\nu,x)$ being bounded above is even more clear since Li works explicitly with a common upperbound for both $m(\nu,x)$ and $V$.
-Hence, one can repeat the necessary proofs replacing the assumption that $m(\nu,x)$ is bounded with the assumption that $m(\nu,x)$ is merely bounded above to derive the same results.
-
-What remains is to show the spatial mean and variance are finite, assuming finite initial conditions.
-That is, setting $M_n(t)=\int_\mathbb{R}|x|^n\nu(x,t)dx$, we want $M_1(0),M_2(0),M_4(0)<\infty$ to imply $M_1(t),M_2(t),M_4(t)<\infty$ for each $t>0$ (it turns out we need $M_4(t)<\infty$ for the SDE describing the evolution of phenotypic variance).
-By condition (\ref{m_cond}), we see that $m(\nu,x)$ drops quadratically as $x$ leaves $\theta$.
-In SM \S\ref{finite} we show that, when $V=0$, this form of fitness generates a restoring force that ensures a finite mean and variance.
-Since mutation causes finite perturbations we do not expect it to overcome this restoring force in the stochastic case.
-If anything, we suspect a large mutation variance to merely increase the genetic load of the population and drive it to extinction.
-Indeed, this suggestion is supported by equilibrium of the corresponding deterministic system found in SM \S\ref{numerical}.
-However, to establish further support, in SM \S\ref{numerical} we investigate these claims by comparing simulations of the BBM with selection to analytical solutions of the PDE obtained from equation (\ref{SPDE}) by setting $V=0$.
-The results suggest our assertions are sound, but they have yet to be rigorously justified.
-For now, we leave this as an open problem and assume spatial means and variances are finite for every $t\in[0,\infty)$.
-In the next subsection we derive an SDE for the abundance of a population and report the SDE for the trait mean $\bar x$ and trait variance $\sigma^2$.
-For the sake of space, derivations of $d\bar x$ and $d\sigma^2$ can be found in SM \S\ref{SDE_DERIV}.
-
-### Deriving SDE from a SPDE \label{SDEriv}
-
-The dynamics of abundance (referred to in the superprocess literature as the total mass process) are obtained by defining the process $N(t)=\int_{\mathbb{R}}\nu(x,t)dx$ and evaluating expression
-
-\begin{multline}\label{neat_trick}
-\int_\mathbb{R}f(x)\nu(x,t)dx-\int_\mathbb{R}f(x)\nu(x,0)dx \\
-=\int_0^t\int_\mathbb{R}\nu(x,s)\left(m(\nu,x)f(x)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2} f(x)\right)dxds+\int_0^t\int_\mathbb{R}f(x)\sqrt{V\nu(x,s)}\dot W(x,s)dxds
-\end{multline}
-with $f(x)\equiv1$ to obtain
-\begin{equation}
-N(t)-N(0)=\bar m(t)N(t)+\int_0^t\int_\mathbb{R}\sqrt{V\nu(x,s)}\dot W(x,s)dxds.
-\end{equation}
-Equation (\ref{cov_WN_xi}) implies
-\begin{equation}
-\mathbb{V}\left(\int_0^t\int_\mathbb{R}\sqrt{V\nu(x,s)}\dot W(x,s)dxds\right)=\int_0^t\int_\mathbb{R}V\nu(x,s)dxds=\int_0^tV N(s)ds.
-\end{equation}
-Following the notation introduced in \S\ref{wnc_intro}, we have
-\begin{equation}
-\int_0^td\hat{\mathbf W}_s(\sqrt{V\nu(x,s)})=\int_0^t\int_\mathbb{R}\frac{\sqrt{V\nu(x,s)}}{\sqrt{V N(s)}}\dot W(x,s)dxds
-\end{equation}
-is, as a function of $t$, a standard Brownian motion.
-To clean up notation set $dW_1(t)=d\hat{\mathbf W}_t(\sqrt{V\nu(x,t)})$.
-Then we can write an SDE for $N$ which we provide in the following section.
-
-Unfortunately, the same trick does not work for calculating SDE for the mean and variance of traits.
-Indeed, setting $f(x)=x,x^2$ and evaluating expression (\ref{neat_trick}) will lead to SDE that miscalculate the effect of mutation and hence are clearly incorrect results.
-Obtaining the correct results turns out to require some very involved calculations which we perform in SM \S\ref{SDE_DERIV}.
+Yet, in the proof of the construction of the interacting superprocess as the limit of rescaled branching diffusions, Méléard and Roelly [-@meleard1992interacting; -@Mlard1993] assumed $m(y,x)$ to be bounded to guarantee the total mass process will have finite mean and variance, for finite $t\geq0$.
+This allowed the authors to employ a tightness criterion for sequences of measures and show the rescaled processes converge to a superprocess with finite total mass.
+Li's [-@zeng1998absolute] result builds directly off of Méléard and Roelly's construction, inheriting the assumption of boundedness for $m(y,x)$.
+However, in Li [-@zeng1998absolute], the sufficiency of $m(y,x)$ being bounded above is even more clear since Li works explicitly with a common upperbound for both $m(y,x)$ and $V$.
+Hence, one can repeat the necessary proofs replacing the assumption that $m(y,x)$ is bounded with the assumption that $m(y,x)$ is merely bounded above to derive the same results.
+
+With solutions to SPDE (\ref{SPDE}) well defined for any growth rate bounded above, we can calculate the total mass process $N(t)$ using the mild solution of (\ref{SPDE}) with $f(x)\equiv1\in C_b^2(\mathbb{R})$ (the symbol "$\equiv$" means equal to for every $x$).
+That is,
+\begin{multline}
+N(t)=N(0)+\int_0^t\int_\mathbb{R}\nu(x,s)\left(m(\nu,x)\cdot1+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}1\right)+1\sqrt{V\nu(x,s)}\dot W(x,s)dsdx \\
+=N(0)+\int_0^t\bar m(s)N(s)dt+\int_0^t\sqrt{VN(s)}d\hat{\mathbf{W}}_s(\sqrt{\nu(x,s)}),
+\end{multline}
+
+where
+\begin{equation}
+\bar m(t)=\frac{1}{N(t)}\int_\mathbb{R}m(\nu,x)\nu(x,t)dx,
+\end{equation}
+and
+\begin{equation}
+\int_0^td\hat{\mathbf{W}}_s(\sqrt{\nu(x,s)})=\int_0^t\int_\mathbb{R}\frac{\sqrt{\nu(x,s)}}{\sqrt{\int_\mathbb{R}\nu(x,s)dx}}\dot W(x,s)dxds.
+\end{equation}
+Setting $W_1(t)=\hat{\mathbf{W}}_t(\sqrt{\nu(x,t)})$, we can use traditional stochastic differential notation to write
+\begin{equation}
+dN=\bar mNdt+\sqrt{VN}dW_1.
+\end{equation}
+To find the associated SDE for $\bar x(t)$ and $\sigma^2(t)$, we want to repeat the same approach for $f(x)=x,x^2$ and apply It&ocirc;'s lemma.
+However, for these cases $f\notin C^2_b(\mathbb{R})$ since $f$ will not be bounded.
+Hence, the goal is to show that $\int_\mathbb{R}(|x|+x^2)\nu(x,t)dx<+\infty$ for all $t>0$ given this condition is satisfied by $\nu(x,0)$.
+Let us begin by supposing we can apply the mild solution to $f(x)=x,x^2$.
+Setting $\tilde x(t)=\int_\mathbb{R}x\nu(x,t)dx$, we have
+\begin{equation}
+\tilde x(t)=\tilde x(0)+\int_0^t\int_\mathbb{R}\nu(x,s)m(\nu,x)x+x\sqrt{V\nu(x,s)}\dot W(x,s)dxds.
+\end{equation}
+Similarly, setting $\tilde{\tilde\sigma}^2(t)=\int_\mathbb{R}x^2\nu(x,t)dx$, we have
+\begin{equation}
+\tilde{\tilde\sigma}^2(t)=\tilde{\tilde\sigma}^2(0)+\int_0^t\int_\mathbb{R}\nu(x,s)\left(m(\nu,x)x^2+\mu\right)+x^2\sqrt{V\nu(x,s)}\dot W(x,s)dxds.
+\end{equation}
+Since $\bar x(t)=\tilde x(t)/N(t)$ and $\sigma^2(t)=\tilde{\tilde\sigma}^2(t)/N(t)-\bar x^2(t)$, we can use It&ocirc;'s lemma to derive SDE for $\bar x(t)$ and $\sigma^2(t)$, which we perform in SM \S\ref{SDE_DERIV}.
+We do not make any general assertions about the existence or uniqueness of $\bar x(t)$ or $\sigma^2(t)$.
+
+<!--
+
+__Superprocesses without densities__
+
+The superprocesses we consider here allow function-valued representations (i.e., densities such as $\nu(x,t)$) only in the case of spatial dimension $d=1$.
+However, if we are willing to let go of function-valued representations, we are free to study superprocesses for greater spatial dimensions.
+Dawson [-@dawson1993measure] provides an excellent resource for this subject.
+Futhermore, the literature on these more general processes includes models of ecological interactions including competition.
+We summarize a few results important to our model of diffuse coevolution here.
+
+Denoting $C^k(\mathbb{R}^d)$ the set of functions on $\mathbb{R}^d$ with continuous derivatives up to order $k$ and $\Delta=\sum_{i=1}^d\frac{\partial^2}{\partial x_i^2}$ the Laplacian operator on $C^2(\mathbb{R}^d)$, we define the operator $A:C^2(\mathbb{R}^d)\to C^0(\mathbb{R}^d)$ by
+\begin{equation}
+A=\frac{\mu}{2}\Delta+r-\frac{a}{2}(\theta-x)^2,
+\end{equation}
+for fixed $r,\theta\in\mathbb{R}$ and $\mu,a>0$.
+
+-->
 
@@
 
-In SM \S\ref{SDE_DERIV} we use our assumptions that trait mean and variance are finite and that variance in reproductive output $V\geq0$ does not depend on $x$ or $\nu(x,t)$ to show SDE for $N(t)$, $\bar x(t)$ and $\sigma^2(t)$ can be expressed as
+In SM \S\ref{SDE_DERIV} we show SDE for $N(t)$, $\bar x(t)$ and $\sigma^2(t)$ can be expressed as
 \begin{subequations}\label{gen_eqns}
@@
 where $W_1$, $W_2$ and $W_3$ are standard Brownian motions.
+We note that conditions on the growth rate $m$ to guarantee existence and uniqueness of solutions to (\ref{xbar_gen}) and (\ref{sig2_gen}) have yet to be investigated.
+However, our results on the deterministic PDE suggest that $m(y,x)$ bounded above and differentiable in both arguments is sufficient.
 Dividing by $dt$ one can interpret equations (\ref{gen_eqns}) as if they are ordinary differential equations, but this not technically rigorous since Brownian motion is nowhere differentiable with respect to time.

### 2020-04-02 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
easy_fig3

@@
 \begin{equation}\label{informal}
-``\int_0^t\int_\mathbb{R}f(x,s)\dot W(x,s)dxds"=``\int_\mathbb{R}\int_0^tf(x,s)\dot W(x,s)dxds"=\mathbf W_t(f), \ \forall \ f\in\mathscr{N}_2, \ t\geq0.
+``\int_0^t\int_\mathbb{R}f(x,s)\dot W(x,s)dxds"=\mathbf W_t(f), \ \forall \ f\in\mathscr{N}_2, \ t\geq0.
 \end{equation}
@@
 Since we have rescaled individual mass by $1/n$ and took the limit $n\to\infty$, individuals are no longer discrete units.
-The particle view of the population gets replaced by a continuous blob spread across $d$-dimensional space.
+Instead, the particle view of the population gets replaced by a blob spread across $d$-dimensional space.
 In particular, the value of $\mathscr{X}_t(D)$ is a continuously varying non-negative random variable for any $t\geq0$ and $D\subset\mathbb{R}^d$.
-It turns out that for spatial dimension $d=1$, $\mathscr{X}_t$ is absolutely continuous with respect to the Lebesgue measure for each $t\geq0$ [@Konno1988; @Reimers1989].
+
+Unfortunately, just as with cream cheese spread across too much toast, the blob perspective of the population may exhibit some complicated spatial discontinuities.
+However, for spatial dimension $d=1$, it turns out that $\mathscr{X}_t$ is absolutely continuous with respect to the Lebesgue measure for each $t\geq0$ [@Konno1988; @Reimers1989].
 This means that we can write $\mathscr{X}_t(D)=\int_D \nu(x,t)dx$ for some density process $\nu(x,t)$.
@@
 \end{equation}
-Since $\nu(x,t)$ is not generally differentiable in $x$ or $t$, the derivatives in expression (\ref{neutral_SPDE}) are taken in the _weak_ sense.
+Since $\nu(x,t)$ is not generally differentiable in $x$ or $t$, the derivatives in expression (\ref{neutral_SPDE}) are taken in the _weak_ sense [sensu @lawrenceevans2010].
 That is, to rigorously interpret SPDE (\ref{neutral_SPDE}), we integrate the solution $\nu(x,t)$ against functions $f\in C_b^2(\mathbb{R})$, where $C_b^2(\mathbb{R})$ is the set of bounded and twice continuously differentiable functions on $\mathbb{R}$.
@@
 This expression is referred to as the _mild_ solution of (\ref{neutral_SPDE}).
-Evans [-@lawrenceevans2010] provides a nice introduction to weak derivatives.
 For more on the general theory of SPDE see Walsh [-@Walsh] and Da Prato and Zabczyk [-@DaPrato2014].
@@
 However, for these cases $f\notin C^2_b(\mathbb{R})$ since $f$ will not be bounded.
-Hence, the goal is to show that $\int_\mathbb{R}(|x|+x^2)\nu(x,t)dx<+\infty$ for all $t>0$ given this condition is satisfied by $\nu(x,0)$.
-Let us begin by supposing we can apply the mild solution to $f(x)=x,x^2$.
+However, if we can show that $\int_\mathbb{R}(|x|+x^2+x^4)\nu(x,t)dx<+\infty$ for all $t>0$ given this condition is satisfied by $\nu(x,0)$, then we can apply the mild solution of (\ref{SPDE}) to derive SDE for $\bar x(t)$ and $\sigma^2(t)$.
+To illustrate, let us suppose this is the case.
 Setting $\tilde x(t)=\int_\mathbb{R}x\nu(x,t)dx$, we have
@@
 The component of SDE (\ref{sig2_gen}) describing random fluctuations in $\sigma^2(t)$ is more complicated and is proportional to the root of the difference between the centralized fourth moment of $p(x,t)$ and $\sigma^4(t)$.
+
 These expressions can be used to investigate the dynamics of the mean and variance for general $\nu(x,t)$.
 However, in the next subsection we simplify these expressions by approximating $\nu(x,t)$ with a Gaussian curve.
-In SM \S\ref{SDE_DERIV} we show that under the Gaussian case $W_1,W_2$ and $W_3$ are independent.
+By assuming $\nu(x,t)$ is Gaussian for $t\geq0$, we guarantee the existence of $\bar x(t)$ and $\sigma^2(t)$ for all $t\geq0$.
+Furthermore, in SM \S\ref{SDE_DERIV} we show that under the Gaussian case $W_1,W_2$ and $W_3$ are independent.
 
@@
 Red lines indicate average trend across species.'}
-knitr::include_graphics('community_dynamics.png')
+knitr::include_graphics('community_dynamics.jpg')
 ```
@@
 
-## Numerical evidence of finite moments and approximate normality in the stochastic case \label{numerical}
+## Simulating the rescaled process and numerical evidence of approximate normality \label{numerical}
 
@@
 
+Picking up from \S\ref{stochastic} of the main text, we have
+\begin{equation}
+\tilde x(t)=\int_\mathbb{R}x\nu(x,t)dx, \ \ \tilde{\tilde\sigma}^2(t)=\int_\mathbb{R}x^2\nu(x,t)dx
+\end{equation}
+and
+\begin{equation}\label{xtilde}
+\tilde x(t)=\tilde x(0)+\int_0^t\int_\mathbb{R}\nu(x,s)m(\nu,x)x+x\sqrt{V\nu(x,s)}\dot W(x,s)dxds,
+\end{equation}
+\begin{equation}\label{stilde}
+\tilde{\tilde\sigma}^2(t)=\tilde{\tilde\sigma}^2(0)+\int_0^t\int_\mathbb{R}\nu(x,s)\left(m(\nu,x)x^2+\mu\right)+x^2\sqrt{V\nu(x,s)}\dot W(x,s)dxds.
+\end{equation}
+
+<!--
+
 For $\nu(x,t)$ defined in the main text, $h\in C(\mathbb{R})$ and $t\geq0$ we make the following assumptions:
@@
 
+-->
+
 ### Derivation for trait mean
 
-We set $\tilde x(t)=\int_\mathbb{R}x\nu(x,t)$ and make use of the notation
-
-\begin{equation}
-\begin{matrix}
+We make use of the notation
+
+\begin{equation}
+\left\{\begin{matrix}
 \|N\|_2 = \sqrt{V\int_\mathbb{R}\nu(x,t)dx}= \sqrt{VN} \\ \\
@@
 \langle\tilde x,N\rangle = V\int_\mathbb{R}x\nu(x,t)dx=\bar x VN.
-\end{matrix}
-\end{equation}
-
-Applying formula (\ref{SDE_deriv}) provides
+\end{matrix}\right.
+\end{equation}
+
+Rewriting formula (\ref{xtilde}) as an SDE provides
 
@@
 \begin{equation}
-d\bar x= \overline{xm}dt+\frac{\mu}{2}\int_\mathbb{R}x\Delta p(x,t)dxdt+\frac{\|\tilde x\|_2}{N}d\tilde W_2-\bar x\left(\bar m dt+\sqrt\frac{V}{N}dW_1\right)-\frac{\langle\tilde x,N\rangle}{N^2}dt+\bar x\frac{\|N\|_2^2}{N^2}dt
+d\bar x= \overline{xm}dt+\frac{\|\tilde x\|_2}{N}d\tilde W_2-\bar x\left(\bar m dt+\sqrt\frac{V}{N}dW_1\right)-\frac{\langle\tilde x,N\rangle}{N^2}dt+\bar x\frac{\|N\|_2^2}{N^2}dt
 \end{equation}
@@
 \begin{equation*}
-= (\overline{xm}-\bar x\bar m)dt+\frac{\mu}{2}\int_\mathbb{R}x\Delta p(x,t)dxdt+\frac{\|\tilde x\|_2}{N}d\tilde W_2
+= (\overline{xm}-\bar x\bar m)dt+\frac{\|\tilde x\|_2}{N}d\tilde W_2
 -\bar x\sqrt{\frac{V}{N}}d W_1-V\frac{\bar x}{N}dt+V\frac{\bar x}{N}dt
@@
 \begin{equation*}
-= \left(\mathrm{Cov}_t(x,m)+\frac{\mu}{2}\int_\mathbb{R}x\Delta p(x,t)dx\right)dt
+=\mathrm{Cov}_t(x,m)
 +\frac{\|\tilde x\|_2}{N}d\tilde W_2-\bar x\sqrt{\frac{V}{N}}d W_1.
@@
 \begin{equation}
-d\bar x=\left(\mathrm{Cov}_t(x,m)+\frac{\mu}{2}\int_\mathbb{R}x\Delta p(x,t)dx\right)dt
+d\bar x=\mathrm{Cov}_t(x,m)dt
 +\sqrt{V\frac{\sigma^2}{N}}d W_2.
@@
 
-We set $\tilde{\tilde\sigma}^2(t)=\int_{\mathbb R} x^2\nu(x,t)dx$ and make use of the notation
-
-\begin{equation}
-\begin{matrix}
+We make use of the notation
+
+\begin{equation}
+\left\{\begin{matrix}
 \|\tilde{\tilde\sigma}^2\|_2 = \sqrt{V\int_\mathbb{R}x^4\nu(x,t)dx} \\ \\
 \langle\tilde{\tilde\sigma}^2,N\rangle=V\int_\mathbb{R}x^2\nu(x,t)dx=\overline{x^2}VN.
-\end{matrix}
-\end{equation}
-
-Applying formula (\ref{SDE_deriv}) provides
-
-\begin{equation}
-d\tilde{\tilde\sigma}^2=\left(\overline{x^2m}N+\frac{\mu}{2}\int_\mathbb{R}x^2\Delta\nu(x,t) dx\right)dt+\|\tilde{\tilde\sigma}^2\|_2d\tilde W_3
+\end{matrix}\right.
+\end{equation}
+
+Applying formula (\ref{stilde}) provides
+
+\begin{equation}
+d\tilde{\tilde\sigma}^2=\left(\overline{x^2m}N+\mu N\right)dt+\|\tilde{\tilde\sigma}^2\|_2d\tilde W_3
 \end{equation}
@@
 \begin{multline}
-d\overline{x^2}=\left(\overline{x^2m} + \frac{\mu}{2}\int_\mathbb{R}x^2\Delta p(x,t)dx\right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3
+d\overline{x^2}=\left(\overline{x^2m} + \mu \right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3
 -\overline{x^2}\left(\bar m dt+\sqrt\frac{V}{N}dW_1\right)-\frac{\langle\tilde{\tilde\sigma}^2,N\rangle}{N^2}dt+\overline{x^2}\frac{\|N\|_2^2}{N^2}dt \\ \\
-=\left(\overline{x^2m}-\overline{x^2}\bar mdt + \frac{\mu}{2}\int_\mathbb{R}x^2\Delta p(x,t)dx\right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3
+=\left(\overline{x^2m}-\overline{x^2}\bar mdt + \mu\right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3
 -\overline{x^2}\sqrt{\frac{V}{N}}d W_1
 -\overline{x^2}\frac{V}{N}dt+\overline{x^2}\frac{V}{N}dt \\ \\
-=\left(\mathrm{Cov}_t\Big(x^2,m\Big) + \frac{\mu}{2}\int_\mathbb{R}x^2\Delta pdx \right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3-\overline{x^2}\sqrt\frac{V}{N}d W_1.
+=\left(\mathrm{Cov}_t\Big(x^2,m\Big) + \mu \right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3-\overline{x^2}\sqrt\frac{V}{N}d W_1.
 \end{multline}
@@
 d\sigma^2=d\overline{x^2}-2\bar xd\bar x-(d\bar x)^2
-=\left(\mathrm{Cov}_t\Big(x^2,m\Big) + \frac{\mu}{2}\int_\mathbb{R}x^2\Delta p(x,t)dx \right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3-\overline{x^2}\sqrt\frac{V}{N}d W_1 \\ \\
--2\bar x\left(\mathrm{Cov}_t(x,m)dt+\frac{\mu}{2}\int_\mathbb{R}x\Delta p(x,t)dxdt+\sqrt{\frac{V\sigma^2}{N}}d W_2\right)-\left(\mathrm{Cov}_t(x,m)dt+\frac{\mu}{2}\int_\mathbb{R}x\Delta p(x,t)dxdt+\sqrt{\frac{V\sigma^2}{N}}d W_2\right)^2 \\ \\
+=\left(\mathrm{Cov}_t\Big(x^2,m\Big) + \mu \right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3-\overline{x^2}\sqrt\frac{V}{N}d W_1 \\ \\
+-2\bar x\left(\mathrm{Cov}_t(x,m)+\mu dt+\sqrt{\frac{V\sigma^2}{N}}d W_2\right)-\left(\mathrm{Cov}_t(x,m)dt+\mu dt+\sqrt{\frac{V\sigma^2}{N}}d W_2\right)^2 \\ \\
 =\left(\mathrm{Cov}_t\Big(x^2-2\bar xx,m\Big) + \frac{\mu}{2}\int_\mathbb{R}(x^2-x\bar x)\Delta p(x,t)dx \right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3-\overline{x^2}\sqrt\frac{V}{N}d W_1-2\bar x\sqrt{\frac{V\sigma^2}{N}}d W_2-\left(\frac{V\sigma^2}{N}\right)dt \\ \\
@@
 \begin{equation}
-d\sigma^2=\left(\mathrm{Cov}_t\Big((x-\bar x)^2,m\Big)+\frac{\mu}{2}\int_\mathbb{R}(x-\bar x)^2\Delta p(x,t)dx-V\frac{\sigma^2}{N}\right)dt+\sqrt{V\frac{\overline{(x-\bar x)^4}-\sigma^4}{N}}d W_3.
+d\sigma^2=\left(\mathrm{Cov}_t\Big((x-\bar x)^2,m\Big)+\mu-V\frac{\sigma^2}{N}\right)dt+\sqrt{V\frac{\overline{(x-\bar x)^4}-\sigma^4}{N}}d W_3.
 \end{equation}

### 2020-04-02 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
fig4Updated!

@@
 \end{multline}
-Using our assumptions on mutation and rate of growth, we show in SM \S\ref{} $\nu(x,t)$ is twice differentiable with respect to $x$ and $\int_{-\infty}^{+\infty}\nu(x,t)dx<\infty$ for all $t\geq0$.
+Using our assumptions on mutation and rate of growth, we show in SM \S\ref{finite} $\nu(x,t)$ is twice differentiable with respect to $x$ and $\int_{-\infty}^{+\infty}\nu(x,t)dx<\infty$ for all $t\geq0$.
 This implies that we are justified in swapping the order of differentiation and integration and the result $\int_{-\infty}^{+\infty}\frac{\partial^2}{\partial x^2}\nu(x,t)dx=0$ can be derived from the fundamental theorem of calculus.
@@
 
-```{r, echo=F, fig.align='center', fig.cap='\\label{net}Networks of interspecific interactions parameterized by competition coffecients (top row) and distributions of pairwise coevolution (bottom row) under the scenarios of weak competition (left) and strong competition (right) at time $t=1.0\\times 10^3$ Node sizes are proportional to population sizes.
-Edge widths and shade are monotonically increasing functions of pairwise coevolution.'}
+```{r, eval=F, echo=F}
+require("ggplot2")
+require("gridExtra")
+require("reshape")
+
+####################
+#                  #
+# weak competition #
+#                  #
+####################
+
+smpl_wc1e3 = read.csv("/home/bb/Gits/branching.brownian.motion.and.spde/sample_path_wc1e3.csv")
+
+#
+# calculating strength of coevolution
+#
+
+x = smpl_wc1e3$x
+G = smpl_wc1e3$G
+N = smpl_wc1e3$N
+
+w = 0.1
+c = 1e-8
+eta = 1.0
+R = 1.0
+a = 1e-2
+th = 0
+S = length(x)
+
+b = matrix(0,nc=S,nr=S)
+C = matrix(0,nc=S,nr=S)
+alpha = matrix(0,nc=S,nr=S)
+beta = matrix(0,nc=S,nr=S)
+gamma = matrix(0,nc=S,nr=S)
+for( i in 1:S){
+  for( j in 1:S ){
+
+    r = R - a*( (x[i]-th)^2 + G[i] + eta )/2
+
+    b[i,j] = 1/( G[i]+G[j]+2*(w+eta) )
+
+    alpha[i,j] = c*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2) / r
+
+    beta[i,j] = c*N[j]*b[i,j]*(x[i]-x[j])*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2)
+
+    gamma[i,j] = c*N[j]*b[i,j]*(1 - b[i,j]*(x[i]-x[j])^2)*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2)
+
+    C[i,j] = abs(beta[i,j]) + abs(gamma[i,j])
+
+  }
+}
+
+Cvec = NULL
+for(i in 1:S){
+  for(j in i:S)
+    if( i!=j ) Cvec = c(Cvec,C[i,j])
+}
+
+avec = NULL
+for(i in 1:S){
+  for(j in i:S)
+    if( i!=j ) avec = c(avec,alpha[i,j])
+}
+
+Cadf_WC = data.frame(Coev=Cvec,Comp=avec, Which="WC")
+
+######################
+#                    #
+# strong competition #
+#                    #
+######################
+
+smpl_mc1e3 = read.csv("/home/bb/Gits/branching.brownian.motion.and.spde/sample_path_mc1e3.csv")
+
+#
+# calculating strength of coevolution
+#
+
+x = smpl_mc1e3$x
+G = smpl_mc1e3$G
+N = smpl_mc1e3$N
+
+w = 0.1
+c = 5e-7
+eta = 1.0
+R = 1.0
+a = 1e-2
+th = 0
+S = length(x)
+
+b = matrix(0,nc=S,nr=S)
+C = matrix(0,nc=S,nr=S)
+alpha = matrix(0,nc=S,nr=S)
+beta = matrix(0,nc=S,nr=S)
+gamma = matrix(0,nc=S,nr=S)
+for( i in 1:S){
+  for( j in 1:S ){
+
+    r = R - a*( (x[i]-th)^2 + G[i] + eta )/2
+
+    b[i,j] = 1/( G[i]+G[j]+2*(w+eta) )
+
+    alpha[i,j] = c*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2) / r
+
+    beta[i,j] = c*N[j]*b[i,j]*(x[i]-x[j])*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2)
+
+    gamma[i,j] = c*N[j]*b[i,j]*(1 - b[i,j]*(x[i]-x[j])^2)*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2)
+
+    C[i,j] = abs(beta[i,j]) + abs(gamma[i,j])
+
+  }
+}
+
+Cvec = NULL
+for(i in 1:S){
+  for(j in i:S)
+    if( i!=j ) Cvec = c(Cvec,C[i,j])
+}
+
+avec = NULL
+for(i in 1:S){
+  for(j in i:S)
+    if( i!=j ) avec = c(avec,alpha[i,j])
+}
+
+Cadf_SC = data.frame(Coev=Cvec,Comp=avec, Which="SC")
+
+# bivariate plot
+Cadf = rbind(Cadf_WC,Cadf_SC)
+
+on_pl = ggplot(Cadf_SC, aes(Coev,Comp)) + 
+  scale_x_log10(limits=c(1e-5,2e-4)) + 
+  geom_density_2d(col="black",n=200) +
+  geom_density_2d(data=Cadf_WC, linetype="dashed", col="black") +
+  theme_minimal() + xlab("Pairwise Coevolution") + ylab("")
+
+on_pl2 = on_pl + facet_grid(rows=vars(Which), scales="free_y") + 
+  ylab("Competition Coefficient") + theme(strip.background = element_blank(), strip.text.y = element_blank())
+
+ggsave("~/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/on_pl.png",on_pl2,device="png",width=3.5,height=3.5,units="in")
+```
+
+```{r, echo=F, fig.align='center', fig.cap='\\label{net}Bivariate distributions of competition coffecients (y-axis) and pairwise coevolution (x-axis) under the scenarios of weak competition (dashed line) and strong competition (solid line) after simulating for $t=1000.0$ units of time.
+Simulations were ran for $S=1000$ species.
+Parameters are the same as in Table 2, except we used $c=1\\times 10^{-8}$ for weak competition and $c=5\\times 10^{-7}$ for strong competition in order to account for increased species richness.'}
 knitr::include_graphics('on_pl.png')
@@
 
-__The genetic architecture and distributions of traits and the role of sexual reproduction__
+__The genetic architecture and distributions of traits__
 
@@
 
-__The role of ecological stoichiometry__
+__Ecological stoichiometry__
 
@@
 where $\delta(x-\xi_i(t))$ denotes the point mass located at $\xi_i(t)$.
-To allow for imperfect heritability, we also track the set of breeding values which, at time $t$, is denoted by $\{\gamma_1(t),\dots,\gamma_{N(t)}(t)\}$ and should not be confused with the quadratic selection gradients discussed in \S\ref{} of the main text.
+To allow for imperfect heritability, we also track the set of breeding values which, at time $t$, is denoted by $\{\gamma_1(t),\dots,\gamma_{N(t)}(t)\}$ and should not be confused with the quadratic selection gradients discussed in \S\ref{diffuse} of the main text.
 Then the distribution of breeding values can be written as
@@
 After breeding values have been assigned, we randomly draw trait values for each offpsring as described above.
-For an overview of our model of inheritance, see \S\ref{} of the main text.
+For an overview of our model of inheritance, see \S\ref{inheritance} of the main text.
 This summarizes the basic structure of our simulation.
@@
 -2\bar x\left(\mathrm{Cov}_t(x,m)+\mu dt+\sqrt{\frac{V\sigma^2}{N}}d W_2\right)-\left(\mathrm{Cov}_t(x,m)dt+\mu dt+\sqrt{\frac{V\sigma^2}{N}}d W_2\right)^2 \\ \\
-=\left(\mathrm{Cov}_t\Big(x^2-2\bar xx,m\Big) + \frac{\mu}{2}\int_\mathbb{R}(x^2-x\bar x)\Delta p(x,t)dx \right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3-\overline{x^2}\sqrt\frac{V}{N}d W_1-2\bar x\sqrt{\frac{V\sigma^2}{N}}d W_2-\left(\frac{V\sigma^2}{N}\right)dt \\ \\
-=\left(\mathrm{Cov}_t\Big(x-\bar x)^2,m\Big) + \frac{\mu}{2}\int_\mathbb{R}(x-\bar x)^2\Delta p(x,t)dx - \frac{V\sigma^2}{N}\right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3-\overline{x^2}\sqrt\frac{V}{N}d W_1-2\bar x\sqrt{\frac{V\sigma^2}{N}}d W_2.
+=\left(\mathrm{Cov}_t\Big(x^2-2\bar xx,m\Big) + \mu \right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3-\overline{x^2}\sqrt\frac{V}{N}d W_1-2\bar x\sqrt{\frac{V\sigma^2}{N}}d W_2-\left(\frac{V\sigma^2}{N}\right)dt \\ \\
+=\left(\mathrm{Cov}_t\Big(x-\bar x)^2,m\Big) + \mu - \frac{V\sigma^2}{N}\right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3-\overline{x^2}\sqrt\frac{V}{N}d W_1-2\bar x\sqrt{\frac{V\sigma^2}{N}}d W_2.
 \end{multline}

### 2020-04-02 sde_num_sol.jl

@@ -4,9 +4,13 @@ using Parameters, Statistics, Random, LinearAlgebra, Distributions,
 include("/home/bb/Gits/branching.brownian.motion.and.spde/sde_functions.jl")
 
 j=1
-k=2
+k=1
 # strengths of competition
 cc = [1e-7,5e-6]
+
+# for S = 1000
+cc = [1e-8,5e-7]
+
 
 # durations
 TT = [1e3,5e3]
@@ -14,7 +18,7 @@ TT = [1e3,5e3]
 tspan = (0.0,TT[j])
 
 # background parameters
-S = 100
+S = 1000
 w = fill(0.1, S)  # niche breadths
 U = fill(1.0, S)  # total niche use
 c = fill(cc[k],S)      # strengths of competition
@@ -43,8 +47,6 @@ u₀ = cat(θ,fill(10.0, S),fill(1000.0, S),dims=1)
 prob = SDEProblem(f,g,u₀,tspan,pars)
 sol = solve(prob)
 
-plot(sol,ylim=(-10,10))
-
 # extract state variables
 x̄_sol = copy(.-sol[(0*S+1):(1*S),:])
 G_sol = copy(sol[(1*S+1):(2*S),:])
@@ -64,4 +66,30 @@ end
 #CSV.write("/home/bob/Research/Branching Brownian Motion/sample_path_wc.csv", df)
 
 # moderate competition
-CSV.write("/home/bob/Research/Branching Brownian Motion/sample_path_mc.csv", df)
+#CSV.write("/home/bob/Research/Branching Brownian Motion/sample_path_mc.csv", df)
+
+# for S = 1000
+
+LT = length(sol)
+
+# extract state variables
+x̄_sol = copy(.-sol[(0*S+1):(1*S),LT])
+G_sol = copy(sol[(1*S+1):(2*S),LT])
+N_sol = copy(sol[(2*S+1):(3*S),LT])
+
+# build dataframes
+spp = string("Species", 1)
+df = DataFrame(spp = spp, x = x̄_sol[1], G = G_sol[1], N = N_sol[1])
+for i in 2:S
+	spp = string("Species", i)
+	df_dummy = DataFrame(spp = spp, x = x̄_sol[i], G = G_sol[i],
+	N = N_sol[i])
+	global df = append!(df,df_dummy)
+end
+
+
+# weak competition
+CSV.write("/home/bb/Gits/branching.brownian.motion.and.spde/sample_path_wc1e3.csv", df)
[…]

### 2020-04-02 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
fig4Updated!

@@
   pdf_document:
+    extra_dependencies: ["float","flafter"]
     toc: false
@@
 header-includes:
-  - \usepackage{float}
   - \usepackage{mathrsfs}
@@
 ```{r global_options, include=FALSE}
-#knitr::opts_chunk$set(fig.pos = 'H')
+#knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
 ```
@@
 
+__Metric of pairwise coevolution__
+
 Below we investigate the correspondence of interaction intensity and coevolutionary change.
@@
 To measure the total evolutionary change in species $i$ induced by species $j$, we form the metric $\Psi_{ij}=|\beta_{ij}|+|\gamma_{ij}|$.
-The top row of Figure \ref{net} displays interaction networks under weak and strong competition where the edge width connecting species $i$ and $j$ is proportional to $\alpha_{ij}\alpha_{ji}$.
-The bottom row of Figure \ref{net} displays the distributions of pairwise coevolutionary change, which we measure for species $i$ and $j$ via $\mathfrak{C}_{ij}=\Psi_{ij}\Psi_{ji}$, under weak and strong competition.
+Figure \ref{net} displays the joint distribution of $\mathfrak{C}_{ij}=\Psi_{ij}\Psi_{ji}$, our metric of pairwise coevolution, and the product of competition coefficients $\alpha_{ij}\alpha_{ji}$ for two simulated communities, both with richness $S=1000$.
+The solid contour represents the case of strong selection ($c=5\times10^{-7}$) and the dashed contour represents the case of weak sensitivity to competition ($c=1\times10^{-8}$).
+In both cases we see that $\mathfrak{C}_{ij}$ and $\alpha_{ij}\alpha_{ji}$ are, essentially, unrelated.
+However, this depiction is only representative of a specific set of parameters.
+Next, we provide analytical approximations of the covariance between selection gradients and competition coefficients and a numerical estimate for the relationship between pairwise coevolution and competition coefficients for a range of parameters.
 
@@
 ```
+
+__Covariance of selection and competition as a function of diversity__
 
@@
 
-An alternative approach to breaking the assumption of normally distributed trait values is the development of explicit multilocus models.
+Another approach to breaking the assumption of normally distributed trait values is the development of explicit multilocus models.
 These models describe the contributions of alleles at particular loci in the genome to the development of quantitative traits.
@@
 This approach has a long history in theoretical quantitative genetics [@bulmer1980; @pmid7851785; @Kirkpatrick1727] and coevolutionary theory [@Nuismer2005; @Kopp2006; @Nuismer2007].
-However, work to investigate the role of genetic architecture in mediating feedbacks between the dynamics of population abundances and the distributions of traits mediating ecological interactions has apparently only just begun [@Schreiber2018; @Patel2019].
+However, work to investigate the role of complex genetic architectures in mediating feedbacks between the dynamics of population abundances and the distributions of traits mediating ecological interactions has apparently only just begun [@Patel2019].
 
@@
 These studies demonstrate the need for further research to understand how soil nutrient availability and organismal ecophysiology structures communities of plants and pollinators.
-Theoretical pursuits in this directions may profit from interfacing the framework we have outlined here with population-ecosystem models such as that developed by Fridley [-@Fridley2017].
+Theoretical pursuits in this direction may profit from interfacing the framework we have outlined here with population-ecosystem models such as that developed by Fridley [-@Fridley2017].
 
@@
 To understand patterns found in ecological communities, considerations must push beyond microevolutionary and contemporary ecological processes and consider the macroevolutionary dynamics of ancestral lineages leading to extant populations.
-Using sub-alpine flower communities as an example, one can observe a very strict ordering of phenology across broad geographic ranges.
+Using sub-alpine flower communities as an example, one can observe a very strict ordering of phenology across a broad geographic range.
 In particular, whether in the Colorado Rocky mountains (such as Gothic, Colorado) or on an outlier of the Idaho batholith (such as Kamiak butte near Palouse, Washington), one will almost surely observe a very conspicuous order of flowers emerging in early spring: at the very beginning of the season blooms _Claytonia lanceolata_ followed by _Erythronium grandiflorum_ which precedes _Delphinium nuttallianum_ (B.

### 2020-04-02 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
fig4Updated!

@@
 Since diffusion processes possess greater analytical tractability than branching processes, Feller's method, known as the diffusion limit, has acquired immense popularity particularly in the field of mathematical population genetics [@Ewens2004].
-For over the past half of a century a great deal of accomplishments have been achieved in formalizing the diffusion limits of branching processes that describe populations of individuals occuring in some continuous space [@Watanabe1968; @Dawson1975; @Perkins1992; @Mlard1993; @edwinarendperkins1995; @zeng1998absolute; @Bertoin2003; @Etheridge2008; @Barton2019].
+For over the past half of a century a great deal of accomplishments have been achieved in formalizing the diffusion limits of branching processes that describe populations of individuals occuring in some continuous space [@Watanabe1968; @Dawson1975; @Dawson1978; @Perkins1992; @Mlard1993; @edwinarendperkins1995; @zeng1998absolute; @Bertoin2003; @Etheridge2008; @Barton2019].
 This space can represent geographic space or, relevant to our context, phenotypic space.
@@
 Conditioning a Dawson-Watanabe superprocess to have constant mass returns a Fleming-Viot process [@Etheridge1991; @Perkins1991] which has been popular in studies of spatial population genetics.
-In particular, an extension of Fleming-Viot processes, known as $\Lambda$-Fleming-Viot processes, were introduced by Bertoin and Le Gall [-@Bertoin2003] and coined by Etheridge [-@Etheridge2008] where it was used to resolve some technical challenges in modelling isolation by distance [@Felsenstein1975; see also @Barton2013; and @Barton2019].
+In particular, an extension of the Fleming-Viot process, known as the $\Lambda$-Fleming-Viot process, was introduced by Bertoin and Le Gall [-@Bertoin2003] and coined by Etheridge [-@Etheridge2008] where it was used to resolve some technical challenges in modelling isolation by distance [@Felsenstein1975; see also @Barton2013; and @Barton2019].
 Although this provides an impressive list of accomplishments, the Dawson-Watanabe superprocess falls short of our needs.
@@
 That is,
+
 \begin{multline}
@@
 \end{multline}
-
 where

### 2020-04-02 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
DaPrato&Zabczyk

@@
 header-includes:
+  - \usepackage{mathtools}
   - \usepackage{mathrsfs}
@@
 Furthermore, since these fluctuations are themselves stochastic, it seems natural to derive expressions for the evolutionary response to demographic stochasticity and consider how the results relate to characterizations of random genetic drift.
-This has been done in continuous time for population genetic models without too much technical overhead, assuming a finite number of alleles.
+This can be done in continuous time for population genetic models without too much technical overhead, assuming a finite number of alleles.
 However, for populations with a continuum of types, such as a quantitative trait, this becomes a vexing mathematical challenge.
@@
 We include an account of the natural history of the simulated system and discuss the significance of demographic stochasticity for structuring ecological communities.
-In \S\ref{ecoevo} we provide expressions for selection gradients and competition coefficients implied by our model and use these expressions to investigate the relationship between the strengths of competition and coevolution.
+In \S\ref{ecoevo} we provide expressions for selection gradients and competition coefficients implied by our model and use these expressions to investigate the relationship between the degree of competition and coevolution.
 Finally, \S\ref{conclusion} concludes with a summary of accomplishments, a few remarks on the limits of this approach and future directions to incorporate more explicitly the genetic architecture of populations, feedbacks with ecosystem processes and the macroevolutionary history of interacting lineages.
@@
 The primary difference being that our model utilizes diffusion instead of convolution with an arbitrary mutation kernel.
-Of course, our model of mutation can be derived as an approximation to Kimura's model, which has been referred to as the Gaussian allelic approximation [in reference to the distribution of mutational effects at a given locus of a genome on the values of traits, @9780471986539], the infinitesimal model [in reference to modelling continuous traits as being encoded by an infinite number of loci each having infinitesimal effect, @Barton2017] and the Gaussian descendants approximation [in reference to offspring trait values being normally distributed around their parental values, @Turelli2017].
+Of course, our model of mutation can be derived as an approximation to Kimura's model, which has been referred to as the Gaussian allelic approximation [in reference to the distribution of mutational effects at a given locus of a genome on the values of traits, @Lande1975; but see also @9780471986539], the infinitesimal model [in reference to modelling continuous traits as being encoded by an infinite number of loci each having infinitesimal effect, @Barton2017] and the Gaussian descendants approximation [in reference to offspring trait values being normally distributed around their parental values, @Turelli2017].
 Alternatively, since diffusion is the continuous-time equivalent to convolution against a Gaussian kernel (SM \S\ref{diffconvequiv}), equation (\ref{p_PDE}) can also be seen as a special case of the continuum-of-alleles model.
@@
 Equation (\ref{detmean}) is a continuous time analog of the well known Robertson-Price equation without transmission bias [@Robertson1966; @PRICE1970; @FRANK2012; @Queller2017; @Lion2018].
-The covariance of fitness and phenotype creates change in $\bar x$ to maximize mean fitness $\bar m$.
+Whether or not the covariance of fitness and phenotype creates change in $\bar x$ to maximize mean fitness $\bar m$ depends on the degree to which selection is frequency dependent [@Lande1976].
 Since this change is driven by a covariance with respect to phenotypic diversity, the response in mean trait to selection is mediated by the phenotypic variance.
@@
 From a statistical perspective, if we think of $(x-\bar x)^2$ as a square error, then in analogy to the dynamics of the mean trait, we see that the response in $\sigma^2$ to selection can be expressed as a covariance of fitness and square error, which is defined in analogy to $\mathrm{Cov}_t(m(\nu,x),x)$.
-This covariance also creates change in $\sigma^2$ that maximizes mean fitness $\bar m$, which can be positive or negative depending on whether selection is stabilizing or disruptive.
+Just as for the evolution of $\bar x(t)$, this covariance also creates change in $\sigma^2$ that can either increase or decrease mean fitness $\bar m$, depending on whether or not individual fitness depends on the distribution of traits across the entire population.
+The effect of selection on phenotypic variance can be positive or negative depending on whether selection is stabilizing or disruptive.
 
@@
 To map the degree of niche overlap to fitness, we assume competition between individuals with niche locations $x_i$ and $x_j$ additively decreases the Malthusian fitness for the individual in species $i$ by $c_i\mathcal{O}_{ij}(x_i,x_j)$ for some $c_i\geq0$.
+We refer to $c_i$ as the sensitivity to competition for species $i$.
 In SM \S\ref{diffuse} we combine this niche model with equations (\ref{N}), (\ref{xbarfinal}) and (\ref{Gfinal}) to find
@@
 Although the dynamics of trait means and variances tend to be far more stochastic than in the weak competition scenario, the community overall appears to quickly reach some statistical equilibrium and remain there.
-However, the abundances across all species in the community are very low due to strength of competition being an order of magnitude higher than in the weak competition case.
+However, the abundances across all species in the community are very low due to competition sensitivies being an order of magnitude higher than in the weak competition case.
 Most of the species maintain abundances greater than $1000.0$, but we found one species that dropped to an abundance of about $50.0$.
@@
 $a$         & strength of abiotic selection & $0.01$               \\
-$c$         & strength of competition & $\{1.0\times10^{-7},5.0\times10^{-6}\}$ \\
+$c$         & sensitivity to competition & $\{1.0\times10^{-7},5.0\times10^{-6}\}$ \\
 $w$         & niche breadth & $0.1$                               \\
@@
 For now we refer to $R_i$ as the _innate_ growth rate.
-With this connection formally established, researchers may pursue a postmodern coexistence theory that naturally includes the evolutionary dynamics of populations and the effects of demographic stochasticity.
+Previous work has shown the importance of demographic stochasticity [@Schreiber2017] and evolutionary dynamics [@Schreiber2018] in determining ceoxistence of competing species.
+However, these studies consider evolution and stochasticity separately.
+With the connection outlined in this section formally established, researchers may pursue a postmodern coexistence theory that naturally includes both the evolutionary dynamics of populations and the effects of demographic stochasticity in a simple synthetic framework.
 
@@
 We repeated this approach for randomly drawn $a$ and $c$ until our sample size reached $1000$.
-In Figure \ref{cor_coev_fig} we plot the temporally averaged values of $\mathrm{Cov}_{f_{\bar X}}(\alpha,\mathfrak{C})$ against the strength of competition $c$.
+In Figure \ref{cor_coev_fig} we plot the temporally averaged values of $\mathrm{Cov}_{f_{\bar X}}(\alpha,\mathfrak{C})$ against the sensitivity to competition $c$.
 Using a cubic regression, we see the correlation of coevolutionary selection gradients and competition coefficeints is negative at variance ratios below $0.5$, zero at variance ratios between $0.5$ and $1.0$, and drops below zero again at variance ratios above $1.0$.
@@
 That is, $m:[0,+\infty)\times\mathbb{R}\to\mathbb{R}$.
-Following our assumptions of the main text, we have that $m(y,x)$ is differentiable with respect to both $x$ and $y$ and there exists $r\in\mathbb{R}$ such that $m(y,x)\leq r$ for each $y\geq0$ and $x\in\mathbb{R}$.
+Following our assumptions of the main text, we have that $m(y,x)$ is differentiable with respect to both $x$ and $y$ and there exists $r\in\mathbb{R}$ such that $m(y,x)\leq r$ across all $y\geq0$ and $x\in\mathbb{R}$.
 Hence, we can apply Proposition 1.21 of Cantrell and Cosner [-@Cantrell2004] to show that $\nu(x,t)$ remains differentiable with respect to $t$ and twice differentiable with respect to $x$ for all $t\geq0$.
@@
 Hence, $\nu_r(x,t)\geq0$ for all $x\in\mathbb{R}$ and $\int_\mathbb{R}\nu_r(x,t)dx=e^{rt}N(0)<+\infty$ for all $t\geq0$.
-Furthermore, denoting $N_r(t)=\int_\mathbb{R}\nu_r(x,t)dx$, $p_r(x,t)=\nu_r(x,t)/N_r(t)$, $\bar x_r(t)=\int_\mathbb{R}xp_r(x,t)dx$ and $\sigma^2_r(t)=\int_\mathbb{R}(x-\bar x_r(t))^2p_r(x,t)dx$, we have
+Furthermore, denoting
+\begin{equation}
+\begin{matrix}
+N_r(t)&=&\int_\mathbb{R}\nu_r(x,t)dx, \\
+p_r(x,t)&=&\nu_r(x,t)/N_r(t), \\
+\bar x_r(t)&=&\int_\mathbb{R}xp_r(x,t)dx, \\
+\sigma^2_r(t)&=&\int_\mathbb{R}(x-\bar x_r(t))^2p_r(x,t)dx, 
+\end{matrix}
+\end{equation}
+we have
 \begin{equation}
@@
 \end{equation}
-
 Hence, $|\bar x_r(t)|,\sigma_r^2(t)<+\infty$ for all $t\geq0$. For the sake of contradiction, suppose there exists $x\in\mathbb{R}$ and $t\geq0$ such that $\nu(x,t)>\nu_r(x,t)$. Then
@@
 
-Let $g:\mathbb{R}^d\to\mathbb{R}$ be smooth. Consider the deterministic Cauchy problem
-
+For continuous $g:\mathbb{R}^d\to\mathbb{R}$, consider the deterministic Cauchy problem
 \begin{equation}\label{heateqn}
@@
 f(x,t)=&g(x), & (x,t)\in\mathbb{R}^d\times\{0\}.
-\end{matrix}\right.  \tag{SM1.1}
-\end{equation}
-
+\end{matrix}\right.
+\end{equation}
 According to Evans [-@lawrenceevans2010], the fundamental solution of (\ref{heateqn}) is
-
-\begin{equation}
-\Phi(x,t)=\frac{1}{(4\pi t)^{d/2}}\exp\left(-\frac{|x|^2}{4t}\right), \ (x,t)\in(0,\infty)\times\mathbb{R}^d, \tag{SM1.2}
-\end{equation}
-
+\begin{equation}
+\Phi(x,t)=\frac{1}{(4\pi t)^{d/2}}\exp\left(-\frac{|x|^2}{4t}\right), \ (x,t)\in(0,\infty)\times\mathbb{R}^d,
+\end{equation}
 where $|x|=\sqrt{\sum_ix_i^2}$. The solution $f(x,t)$ of PDE (\ref{heateqn}) is then given by the convolution
-
-\begin{equation}
-f(x,t)=\int_{\mathbb{R}^d}\Phi(x-y,t)g(y)dy, \ (x,t)\in(0,\infty)\times\mathbb{R}^d. \tag{SM1.3}
-\end{equation}
-
+\begin{equation}
+f(x,t)=\int_{\mathbb{R}^d}\Phi(x-y,t)g(y)dy, \ (x,t)\in(0,\infty)\times\mathbb{R}^d.
+\end{equation}
 Hence, by the fundamental theorem of calculus,
 
-\begin{equation}
-f(x,t)+\int_t^{t+1}\dot f(x,s)ds=f(x,t+1) \tag{SM1.4}
-\end{equation}
-
-$$=\int_{\mathbb{R}^d}\Phi(x-y,t+1)g(y)dy=\int_{\mathbb{R}^d}\int_{\mathbb{R}^d}\Phi(x-y,1)\Phi(y-z,t)g(z)dzdy$$
-
-$$=\int_{\mathbb{R}^d}\Phi(x-y,1)f(t,y)dy.$$
-
+\begin{multline}
+f(x,t)+\int_t^{t+1}\dot f(x,s)ds=f(x,t+1) \\
+=\int_{\mathbb{R}^d}\Phi(x-y,t+1)g(y)dy=\int_{\mathbb{R}^d}\int_{\mathbb{R}^d}\Phi(x-y,1)\Phi(y-z,t)g(z)dzdy \\
+=\int_{\mathbb{R}^d}\Phi(x-y,1)f(t,y)dy.
+\end{multline}
 In particular,
-
-\begin{equation}
-f(x,t)+\int_t^{t+1}\Delta f(x,s)ds=\int_{\mathbb{R}^d}\Phi(1,x-y)f(y,t)dy. \tag{SM1.5}
+\begin{equation}
+f(x,t)+\int_t^{t+1}\Delta f(x,s)ds=\int_{\mathbb{R}^d}\Phi(1,x-y)f(y,t)dy.
 \end{equation}
@@
 \end{multline}
-
 Applying integration by parts twice yields
-
 \begin{equation}
@@
 \end{multline}
+
+## Comparing our treatment of white noise to Da Prato and Zabczyk (2014)
+
+Our approach in the main text is inspired by the treatment provided in \S4.2 of Da Prato and Zabczyk [-@DaPrato2014].
+Here the authors develop a stochastic integral of operator-valued processes.
+In particular, they consider processes indexed by time $t\geq0$ valued as Hilbert-Schmidt operators $\Phi(t)$ and define the norm
+\begin{equation}
+\|\Phi\|_t=\sqrt{\mathbb{E}\left(\int_0^t\mathrm{Tr}[\Phi(s)\Phi^*(s)]ds\right)}, \ t\geq0.
+\end{equation}
+In our case, we only consider the so-called multiplication operators.
+That is, processes that consist of operators $\Phi(t)$ having the form $\Phi(t)g(x)=\varphi(x,t)g(x)$ such that $\varphi(\cdot,t)\in L^2(\mathbb{R})$ a.s.
+for each $t\geq0$.
+In this case $\Phi(t)=\Phi^*(t)$ and
+\begin{equation}
+\|\Phi\|_t=\|\varphi\|_t=\sqrt{\mathbb{E}\left(\int_0^t\int_\mathbb{R}\varphi^2(x,s)dxds\right)}, \ t\geq0.
+\end{equation}
+Da Prato and Zabczyk [-@DaPrato2014] form the space $\mathscr{N}^2_W(0,T)$ of Hilbert-Schmidt operator-valued predictable processes $\Phi(t)$ that satisfy $\|\Phi\|_T<+\infty$ for some $T>0$.
+This corresponds to our more specialized space $\mathscr{N}_2$ that consists of $L^2(\mathbb{R})$ valued processes $\varphi(x,t)$ such that $\|\varphi\|_t<+\infty$ for all $t\geq0$.
+In their treatment, space-time white noise is denoted by $W(t)$, where the spatial dependency is suppressed for the sake of rigor.
+For $\Phi\in\mathscr{N}^2_W(0,T)$, they denote the stochastic integral for $t\in[0,T]$ by $\Phi\cdot W(t)$.
+The authors then prove the following:
+
+  > __Proposition 4.28__ _Assume that $\Phi_1,\Phi_2\in\mathscr{N}^2_W(0,T)$. Then_
+  >
+  > $$\mathbb{E}(\Phi_i\cdot W(t))=0, \ \ \mathbb{E}(\|\Phi_i\cdot W(t)\|^2)<+\infty,$$
+  >
+  > __Corollory 4.29__ _Under the same assumptions as Proposition 4.28,_
+  >
+  > $$\mathbb{C}(\Phi_1\cdot W(t),\Phi_2\cdot W(s))=\mathbb{E}\left(\int_0^{t\wedge s}\mathrm{Tr}[\Phi_2(r)\Phi_1^*(r)]dr\right).$$
+
+Simplifying these expressions for the multiplication operators described above returns equations (\ref{exp_WN_xi}) and (\ref{cov_WN_xi}) of the main text.
 
@@
 \begin{equation}
-d\tilde{ W}_2=d\hat{\mathbf W}_{\sqrt{V x^2\nu}}=\frac{1}{\|\tilde x\|_2}\int_\mathbb{R}x\sqrt{V\nu(x,t)}\dot W(x,t)dxdt.
+d\tilde{ W}_2=d\hat{\mathbf W}_t(\sqrt{V x^2\nu})=\frac{1}{\|\tilde x\|_2}\int_\mathbb{R}x\sqrt{V\nu(x,t)}\dot W(x,t)dxdt.
 \end{equation}
@@
 \begin{equation}
-d\tilde{W}_3=d\hat{\mathbf W}_{\sqrt{V x^4\nu}}
+d\tilde{W}_3=d\hat{\mathbf W}_t(\sqrt{V x^4\nu})
 =\frac{1}{\|\tilde{\tilde\sigma}^2\|_2}\int_\mathbb{R}x^2\sqrt{V\nu(x,t)}\dot W(x,t)dx.

### 2020-04-02 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
GenProcDiscussion

@@
 For instance, instead of rigorously proving properties of white-noise, we simply define them to be so, taking solice in the fact that the technical details have been worked out elsewhere.
+In SM \S\ref{DPZ14} we show how our informal treatment is related to the rigorous treatment provided by Da Prato and Zabczyk [-@DaPrato2014].
 
@@
 A generalized process is the stochastic analog of a generalized function, such as the Dirac delta function $\delta$.
-Just as a generalized function operates on classical functions to return a value (e.g., $\delta(f)=f(0)$), a generalized process acts on a set of functions to return a classically defined stochastic process.
-For a breif primer on the theory of generalized functions, see the addendum to chapter 3 of Kolmogorov and Fomin [-@kolmogorov1999elements].
+Often one sees $\delta$ defined as a function satisfying the properties $\delta(x)=0$ for $x\neq0$ and $\int_{-\infty}^{+\infty}\delta(x)dx=1$.
+However, since there is no function that satisfies these properties, we refer to $\delta$ as a generalized function.
+An alternative definition of $\delta$ considers its action on classically defined functions $f$.
+In particular, $\delta(f)=f(0)$, which can be heuristically written as $\int_{-\infty}^{+\infty}f(x)\delta(x)dx=f(0)$.
+Similarly, other generalized functions can be defined by their action on classically defined functions.
+Then, just as a generalized function operates on classical functions to return a value, a generalized process acts on a set of functions (or processes) to return a random variable (or a classically defined stochastic process).
+For a brief primer on the theory of generalized functions, see the addendum to chapter 3 of Kolmogorov and Fomin [-@kolmogorov1999elements].
 
@@
 \begin{equation}
-d\hat{\mathbf W}_t(f)=\frac{1}{\|f\|_t}\left(\int_\mathbb{R}f(x,t)\dot W(x,t)dx\right)dt,
+\hat f(x,t)=\frac{f(x,t)}{\sqrt{\int_\mathbb{R}f^2(y,t)dy}} \ \text{ and } \ 
+d\hat{\mathbf W}_t(f)=\left(\int_\mathbb{R}\hat f(x,t)\dot W(x,t)dx\right)dt
 \end{equation}
@@
 \end{equation}
-This implies $\mathbb{E}\big(\int_0^td\hat{\mathbf W}_s(f)\big)=0$, $\mathbb{C}\big(\int_0^{t_1}d\hat{\mathbf W}_s(f),\int_0^{t_2}d\hat{\mathbf W}_s(f)\big)=t_1\wedge t_2$ and, as a function of $t$, $\int_0^td\hat{\mathbf W}_s(f)$ is a standard Brownian motion for any $f\in\mathscr{N}_2$.
+This implies 
+\begin{equation}
+\mathbb{E}\left(\int_0^td\hat{\mathbf W}_s(f)\right)=0, \ \mathbb{C}\left(\int_0^{t_1}d\hat{\mathbf W}_s(f),\int_0^{t_2}d\hat{\mathbf W}_s(f)\right)=t_1\wedge t_2
+\end{equation}
+and in particular, as a function of $t$, $\int_0^td\hat{\mathbf W}_s(f)$ is a standard Brownian motion for any $f\in\mathscr{N}_2$.
 Hence, $d\hat{\mathbf W}_t(f)$ is analogous to the traditional shorthand used to denote stochastic differentials.
@@
              &                            &                            &      \\
-$d\hat{\mathbf W}_t(f)$ & $dt$                       & $\frac{\left(\int_\mathbb{R}f \ g \ dx\right)dt}{\|f\|_t\|g\|_t}$ & $0$  \\
+$d\hat{\mathbf W}_t(f)$ & $dt$                       & $\left(\int_\mathbb{R}\hat f(x,t)\hat g(x,t)dx\right)dt$ & $0$  \\
              &                            &                            &      \\
-$d\hat{\mathbf W}_t(g)$ & $\frac{\left(\int_\mathbb{R}f \ g \ dx\right)dt}{\|f\|_t\|g\|_t}$ & $dt$                       & $0$  \\
+$d\hat{\mathbf W}_t(g)$ & $\left(\int_\mathbb{R}\hat f(x,t)\hat g(x,t)dx\right)dt$ & $dt$                       & $0$  \\
              &                            &                            &      \\
@@
 \end{equation}
-We assume $\rho(g,t)$ is Gaussian which implies its mode coincides with $\bar x$. Furthermore, since $\sigma^2=G+\eta$, we can use equation (\ref{m_rel}) and the chain rule from calculus to find
+We assume $\rho(g,t)$ is Gaussian which implies its mode coincides with $\bar x$.
+Furthermore, since $\sigma^2=G+\eta$, we can use equation (\ref{m_rel}) and the chain rule from calculus to find (see SM \S\ref{fit2fit})
 \begin{subequations}
@@
 
-## Comparing our treatment of white noise to Da Prato and Zabczyk (2014)
+## Comparing our treatment of white noise to Da Prato and Zabczyk (2014) \label{DPZ14}
 
@@
 
-## Relating fitness of expressed traits to fitness of breeding values
-
-$$m^*(\rho,g)=\int_\mathbb{R}m(\nu,x)\psi(x,g)dx$$
-
-$$\overline{\frac{\partial m^*}{\partial\bar x}}=\int_{\mathbb{R}}\frac{\rho(g,t)}{N(t)}\frac{\partial}{\partial\bar x}\int_\mathbb{R}m(\nu,x)\psi(x,g)dxdg=\int_{\mathbb{R}}\int_\mathbb{R}\frac{\rho(g,t)}{N(t)}\psi(x,g)dg\frac{\partial}{\partial\bar x}m(\nu,x)dx=\int_\mathbb{R}p(x,t)\frac{\partial}{\partial\bar x}m(\nu,x)dx=\overline{\frac{\partial m}{\partial\bar x}}$$
-
-$$\overline{\frac{\partial m^*}{\partial G}}=\int_{\mathbb{R}}\frac{\rho(g,t)}{N(t)}\frac{\partial}{\partial G}\int_\mathbb{R}m(\nu,x)\psi(x,g)dxdg=\int_{\mathbb{R}}\int_\mathbb{R}\frac{\rho(g,t)}{N(t)}\psi(x,g)dg\frac{\partial m}{\partial G}dx=\int_\mathbb{R}p(x,t)\frac{\partial m}{\partial\sigma^2}\frac{\partial\sigma^2}{\partial G}dx=\overline{\frac{\partial m}{\partial\sigma^2}}$$
+## Relating fitness of expressed traits to fitness of breeding values \label{fit2fit}
+
+Following \S\ref{inheritance} of the main text, we have $\sigma^2=G+\eta$ and
+
+\begin{equation}
+m^*(\rho,g)=\int_\mathbb{R}m(\nu,x)\psi(x,g)dx.
+\end{equation}
+
+Hence,
+
+\begin{multline}
+\overline{\frac{\partial m^*}{\partial\bar x}}=\int_{\mathbb{R}}\frac{\rho(g,t)}{N(t)}\frac{\partial}{\partial\bar x}\int_\mathbb{R}m(\nu,x)\psi(x,g)dxdg= \\
+\int_{\mathbb{R}}\int_\mathbb{R}\frac{\rho(g,t)}{N(t)}\psi(x,g)dg\frac{\partial}{\partial\bar x}m(\nu,x)dx=\int_\mathbb{R}p(x,t)\frac{\partial}{\partial\bar x}m(\nu,x)dx=\overline{\frac{\partial m}{\partial\bar x}}
+\end{multline}
+and
+
+\begin{multline}
+\overline{\frac{\partial m^*}{\partial G}}=\int_{\mathbb{R}}\frac{\rho(g,t)}{N(t)}\frac{\partial}{\partial G}\int_\mathbb{R}m(\nu,x)\psi(x,g)dxdg=\int_{\mathbb{R}}\int_\mathbb{R}\frac{\rho(g,t)}{N(t)}\psi(x,g)dg\frac{\partial m}{\partial G}dx= \\
+\int_\mathbb{R}p(x,t)\frac{\partial m}{\partial\sigma^2}\frac{\partial\sigma^2}{\partial G}dx=\overline{\frac{\partial m}{\partial\sigma^2}}.
+\end{multline}
 

### 2020-04-03 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
mention lande 1975

@@
 This PDE is semilinear due to the dependency of the growth rate $m(\nu,x)$ on the solution $\nu(x,t)$ and is referred to as a scalar reaction-diffusion equation [@lawrenceevans2010].
+Interpreting the trait value $x$ as spatial location, this model of population dynamics has been intensely studied [for a review, see @Cantrell2004].
 When $\mu=0$, equation (\ref{eq1}) can be seen as an analog of equation (\ref{first}) for a continuum of types.
@@
 Although clearly an idealized representation of biological reality, this model is sufficiently general to capture a large class of dynamics including density dependent growth and frequency dependent selection.
-As an example, logistic growth combined with quadratic stabilizing selection can be captured using the growth rate $m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\nu(x,t)$ where $a\geq0$ is the strength of stabilizing selection around the phenotypic optimum $\theta\in\mathbb{R}$, $c\geq0$ captures the effect of intraspecific competition and $r\in\mathbb{R}$ is the intrinsic growth rate in the absence of abiotic selection.
+As an example, logistic growth combined with quadratic stabilizing selection can be captured using the growth rate
+\begin{equation}
+m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\nu(x,t)
+\end{equation}
+where $a\geq0$ is the strength of stabilizing selection around the phenotypic optimum $\theta\in\mathbb{R}$, $c\geq0$ captures the sensitivity of fitness to intraspecific competition and $r\in\mathbb{R}$ is the intrinsic growth rate in the absence of abiotic selection.
+This exemplary fitness function has a few convenient properties that we breifly take note of.
+First, the effect of competition induces a carrying capacity on the population, leading to a finite equilibrium abundance over bounded subsets of $\mathbb{R}$.
+Second, abiotic selection prevents the abundance density from diffusing too far from the abiotic optimum, leading to a finite equilibrium abundance across all of $\mathbb{R}$.
+A stochastic generalization of this model, closely related to our model of diffuse coevolution presented later, has been studied by Dawson [-@Dawson1978].
 
@@
 Taking the same approach, we derive the dynamics of the phenotypic distribution $p(x,t)$ in response to selection and muation as
-
 \begin{multline}\label{p_PDE}
@@
 The primary difference being that our model utilizes diffusion instead of convolution with an arbitrary mutation kernel.
-Of course, our model of mutation can be derived as an approximation to Kimura's model, which has been referred to as the Gaussian allelic approximation [in reference to the distribution of mutational effects at a given locus of a genome on the values of traits, @Lande1975; but see also @9780471986539], the infinitesimal model [in reference to modelling continuous traits as being encoded by an infinite number of loci each having infinitesimal effect, @Barton2017] and the Gaussian descendants approximation [in reference to offspring trait values being normally distributed around their parental values, @Turelli2017].
-Alternatively, since diffusion is the continuous-time equivalent to convolution against a Gaussian kernel (SM \S\ref{diffconvequiv}), equation (\ref{p_PDE}) can also be seen as a special case of the continuum-of-alleles model.
+Of course, our model of mutation can be derived as an approximation to Kimura's model, which has been referred to as the Gaussian allelic approximation in reference to the distribution of mutational effects at loci in a genome on the values of traits [@Lande1975; but see also @9780471986539], the infinitesimal model in reference to modelling continuous traits as being encoded by an infinite number of loci each having infinitesimal effect [@Barton2017] and the Gaussian descendants approximation in reference to offspring trait values being normally distributed around their parental values [@Turelli2017].
+Alternatively, since diffusion is the continuous-time equivalent to convolution against a Gaussian kernel (SM \S\ref{diffconvequiv}), equation (\ref{p_PDE}) can also be seen as a special case of Kimura's continuum-of-alleles model.
+We return to these concepts in \S\ref{inheritance} where we review a simple model of inheritance that leads us to an expression for the stochastic evolution of additive genetic variance.
 
@@
 
-### Definition and basic properties of white noise \label{wnc_intro}
+### Introduction to white noise
 
@@
 However, it turns out that realizations of white noise do not exist as functions in the classical sense.
-Indeed, since Brownian motion is nowhere differentiable with respect to time, white noise cannot be formally understood as its time derivative.
+Indeed, since Brownian motion is nowhere differentiable with respect to time, white noise cannot be formally understood as a time derivative.
 Thus our notation $\dot W$ is only meant to aid intuition and not be taken as formal.
@@
 For a brief primer on the theory of generalized functions, see the addendum to chapter 3 of Kolmogorov and Fomin [-@kolmogorov1999elements].
+
+### Definition and basic properties of white noise \label{wnc_intro}
 
@@
 For a detailed treatment of breeding values and additive genetic variances, see Bulmer [-@bulmer1980] and Lynch and Walsh [-@michaellynch1998].
-Following standard assumptions of classical quantitative genetics we assume that the expressed trait for any given individual is normally distributed around their breeding value with variance $\eta$.
+
+Our treatment of the relationship between breeding values and expressed traits follows classical quantitative genetic assumptions such as those used in the seminal paper by Lande [-@Lande1975] to investigate the maintenance of genetic variation.
+In particular, we assume that the expressed trait for any given individual is independent of environmental conditions and normally distributed around their breeding value with variance $\eta$.
 Hence, $\sigma^2=G+\eta$.
-This coincides with assuming breeding values can be predicted from expressed traits using ordinary least squares.
-In the case that all of the effects of alleles on an expressed trait are additive, $\eta$ is known as the _variance of environmental deviation_ [@michaellynch1998].
+In the case that all of the effects of alleles on an expressed trait are additive, $\eta$ is known as the _variance of environmental deviation_ [@Lande1975; @michaellynch1998].
 For a given breeding value, we denote the probability density of a randomly drawn expressed trait by $\psi(x,g)$.
@@
 \end{equation}
-However, since selection acts on expressed phenotypes, we use our assumed relationship between breeding values and expressed traits to calculate the fitness of breeding values.
-To motivate our approach, consider the problem of inferring the breeding value of an individual given its expressed trait $x$.
+However, since selection acts on expressed phenotypes, we use the assumed relationship between breeding values and expressed traits to calculate the fitness of breeding values.
+To motivate the approach taken, consider the problem of inferring the breeding value of an individual given its expressed trait $x$.
 Denote $\mathfrak{g}$ a random variable representing the unkown breeding value.
-Under our model of inheritance we know $x$ is a random sample from a normal distribution with mean $\mathfrak{g}$ and variance $\eta$.
+Under this model of inheritance we know $x$ is a random sample from a normal distribution with mean $\mathfrak{g}$ and variance $\eta$.
 Maximizing likelihood suggests $x$ is our best guess for $\mathfrak{g}$, but the actual value of $\mathfrak{g}$ is normally distributed around $x$ with the variance $\eta$.
@@
 This is similar to the approach taken by Kimura and Crow [-@Kimura1978] to calculate the overall effects of selection for expressed characters onto the changes in the distribution of alleles encoding those characters.
-However, instead of focusing on the frequencies of alleles at particular loci, we focus on the densities of breeding values.
+However, instead of focusing on the frequencies of alleles at particular loci, our results focus on the densities of breeding values.
 With the relationship between $m(\nu,x)$ and $m^*(\rho,g)$ established, we define the evolution of $\rho(g,t)$ by the SPDE
-\begin{equation}
+\begin{equation}\label{rho_SPDE}
 \dot\rho(g,t)=m^*(\rho,g)\rho(g,t)+\frac{\mu}{2}\frac{\partial^2}{\partial^2 g}\rho(g,t)+\sqrt{V\rho(g,t)}\dot W(g,t).
 \end{equation}
+Equation (\ref{rho_SPDE}) is a stochastic generalization of the deterministic PDE (\ref{eq1}) from \S\ref{deterministic}, but describes the evolution of the distribution of breeding values instead of expressed characters.
+In \S\ref{wnc} we review the origins of this equation and provide some theory to help make sense of it, particularly the term $\dot W$.
+
 We assume $\rho(g,t)$ is Gaussian which implies its mode coincides with $\bar x$.
-Furthermore, since $\sigma^2=G+\eta$, we can use equation (\ref{m_rel}) and the chain rule from calculus to find (see SM \S\ref{fit2fit})
+Furthermore, since $\sigma^2=G+\eta$, we can use equation (\ref{m_rel}) and the chain rule from calculus (see SM \S\ref{fit2fit}) to find
 \begin{subequations}
@@
 
-From expressions (\ref{inher}) we see that, under our model of inheritance, focusing on additive genetic variance $G$ instead the variance in expressed traits $\sigma^2$ makes no structural changes to the basic equations describing the dynamics of populations.
+From expressions (\ref{inher}) we see that, under this model of inheritance, focusing on additive genetic variance $G$ instead the variance in expressed traits $\sigma^2$ makes no structural changes to the basic equations describing the dynamics of populations.
+In the next section, we make use of these expressions to develop a model of diffuse coevolution in a guild of $S$ species competing along a resource continuum.
 
@@
 Our model avoids these caveats by assuming competition only occurs along a single dimensional resource gradient.
+
 To map the degree of niche overlap to fitness, we assume competition between individuals with niche locations $x_i$ and $x_j$ additively decreases the Malthusian fitness for the individual in species $i$ by $c_i\mathcal{O}_{ij}(x_i,x_j)$ for some $c_i\geq0$.
 We refer to $c_i$ as the sensitivity to competition for species $i$.
-In SM \S\ref{diffuse} we combine this niche model with equations (\ref{N}), (\ref{xbarfinal}) and (\ref{Gfinal}) to find
+The term $c_i\mathcal{O}_{ij}(x_i,x_j)$ coincides with a special case of a term used to capture competition in Dawson's geostochastic logistic model, an SPDE model developed to study the combined effects of demographic stochasticity, spatial dispersion and locally finite carrying capacity [@Dawson1978].
+In SM \S\ref{diffuse} we combine our treatment of resource competition with equations (\ref{N}), (\ref{xbarfinal}) and (\ref{Gfinal}) to find
 
@@
 These effects were already noted in \S\ref{equations}, and thus are not a consequence of our model of coevolution per-se, but we revisit them here since Figure \ref{temporal} demonstrates the importance of demographic stochasticity in structuring ecological communities even when populations are very large.
-Hence, contrary to the common assumption that stochastic effects can be ignored for large populations, we find that minute asymmetries by demographic stochasticity remain significant drivers of community structure.
+Hence, contrary to the common assumption that stochastic effects can be ignored for large populations, we find that minute asymmetries generated by demographic stochasticity remain significant drivers of community structure.
 In particular, we initialized the species with identical state variables and model parameters, but found an enourmous amount of asymmetry and even some potential phase changes.
@@
 
+We begin by describing the weak competition scenario.
 After about $125.0$ units of time, the community appears to have shaken off the initial conditions and entered into a qualitatively distinct phase of dynamics.
@@
 Upon inspecting our calculations, we diagnose the reason for this shift.
-The outlier species that were initially pushed away from the common abiotic optimum ($0.0$ in this case) evolved a significant reduction in the quantity of heritable variation ($\approx60\%$) due to directional selection induced by competition.
+The outlier species that were initially pushed away from the common abiotic optimum ($0.0$ in this case) evolved a significant reduction in the amount of heritable variation ($\approx60\%$) due to directional selection induced by competition.
 This reduction in heritable variation slowed adaptation, causing these species to linger on the outskirts of niche space, some longer than others.
@@
 As a result, the outlier populations were given a slight increase in growth rate, enough to allow them to increase their abundances orders of magnitude higher than the species in the core group and giving them more weight in driving the evolution of other species.
-Many of these heavy-hitting outlier species had already been maintaining negative mean traits, but around time $500.0$ the high abundance species with positive mean traits began to experience enough intraspecific competition to override interspecific competition and begin evolving towards the abiotic optimum.
+Many of these heavy-hitting outlier species had already been maintaining negative mean traits, but around time $500.0$ the high abundance species with positive mean traits began to experience enough intraspecific competition to override interspecific competition.
+This generated a net selection gradient and associated evolutionary response towards the abiotic optimum.
 The sudden imbalance of these high abundance species effectively induced a single large competitive exclusion event pushing the majority of the community far away from the abiotic optimum.
@@
 
-Relating our treatment of the niche to modern coexistence theory [@Chesson2000], the absolute competition coefficient $\alpha_{ij}$ becomes a dynamical quantity that can be written as
+Relating our treatment of resource competition to modern coexistence theory [@Chesson2000], the absolute competition coefficient $\alpha_{ij}$ becomes a dynamical quantity that can be written as
 
@@
 This corresponds to our more specialized space $\mathscr{N}_2$ that consists of $L^2(\mathbb{R})$ valued processes $\varphi(x,t)$ such that $\|\varphi\|_t<+\infty$ for all $t\geq0$.
-In their treatment, space-time white noise is denoted by $W(t)$, where the spatial dependency is suppressed for the sake of rigor.
+In their treatment, $W(t)$ plays a similar role to our generalized process $\mathbf{W}_t$.
 For $\Phi\in\mathscr{N}^2_W(0,T)$, they denote the stochastic integral for $t\in[0,T]$ by $\Phi\cdot W(t)$.
+Hence, for $\Phi(t) g(x)=\varphi(x,t)g(x)$ as above, $\mathbf{W}_t(\varphi)=\Phi\cdot W(t)$.
 The authors then prove the following:
@@
   >
-  > $$\mathbb{E}(\Phi_i\cdot W(t))=0, \ \ \mathbb{E}(\|\Phi_i\cdot W(t)\|^2)<+\infty,$$
+  > $$\mathbb{E}(\Phi_i\cdot W(t))=0, \ \ \mathbb{E}(\|\Phi_i\cdot W(t)\|^2)<+\infty, \ \ \forall \ t\in[0,T].$$
   >
@@
   >
-  > $$\mathbb{C}(\Phi_1\cdot W(t),\Phi_2\cdot W(s))=\mathbb{E}\left(\int_0^{t\wedge s}\mathrm{Tr}[\Phi_2(r)\Phi_1^*(r)]dr\right).$$
+  > $$\mathbb{C}(\Phi_1\cdot W(t),\Phi_2\cdot W(s))=\mathbb{E}\left(\int_0^{t\wedge s}\mathrm{Tr}[\Phi_2(r)\Phi_1^*(r)]dr\right), \ \ \forall \ t,s\in[0,T].$$
 
@@
 
-Hence, our fitness function satisfies condition (\ref{m_cond}) of the main text.
+Hence, our fitness function is bounded above, as required in the main text.
 

### 2020-04-06 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
ALL of the rigorous proofs are finally complete

@@
 where $\bar x(t)$ and $\sigma^2(t)$ are respectively the mean trait and phenotypic variance at time $t\geq0$.
-Second, we assume selection is determined by the growth rate $m(\nu(x,t),x)$ that is differentiable with respect to both arguments and satisfies $m(y,x)\leq r$ for some $r\in\mathbb{R}$ and for all $x\in\mathbb{R}$, $y\geq0$.
-From here on we abbreviate $m(\nu(x,t),x)$ to $m(\nu,x)$.
+Second, we assume selection is determined by the growth rate $m(h,x)$ that is differentiable with respect to both arguments and is bounded above by some $r\in\mathbb{R}$ for all $x\in\mathbb{R}$ and $h\geq0$.
+We denote by $K$ an operator that takes functions such as $\nu(x,t)$ as its argument and returns other functions that may still depend on the spatial variable $x$ and the temporal variable $t$.
+For our application we will be focused on operators $K$ that can be written as $K(\nu)(x,t)=\int_{-\infty}^{+\infty}\kappa(x,y)\nu(y,t)dy$ for some non-negative and bounded bivariate function $\kappa$.
+We set $h=K(\nu)(x,t)$ to account for nonlocal effects of the abundance density $\nu(x,t)$ on the fitness of individuals with trait value $x$ [@Volpert2014].
+An example of a nonlocal effect is the competition for a common resource among all individuals regardless of their trait value so that individuals with differing trait values effect each others fitness.
+The term nonlocal comes from interpreting the trait value as spatial location.
+From here on we abbreviate $m(h,x)=m(K(\nu)(x,t),x)$ to just $m(\nu,x)$.
 Third, we assume mutation is captured by diffusion with coefficient $\frac{\mu}{2}$.
-With these technicalities aside, the demographic dynamics can be modelled by the PDE
+Now that these technicalities have been treated we can model the demographic dynamics of a population and the dynamics of a quantitative character simultaneously by the PDE
 \begin{equation}\label{eq1}
@@
 This PDE is semilinear due to the dependency of the growth rate $m(\nu,x)$ on the solution $\nu(x,t)$ and is referred to as a scalar reaction-diffusion equation [@lawrenceevans2010].
-Interpreting the trait value $x$ as spatial location, this model of population dynamics has been intensely studied [for a review, see @Cantrell2004].
+Interpreting the trait value $x$ as spatial location and ignoring nonlocal effects, which can be satisfied by setting $K(\nu)=\nu$, this model of spatially distributed population dynamics has been intensely studied [for a review, see @Cantrell2004].
 When $\mu=0$, equation (\ref{eq1}) can be seen as an analog of equation (\ref{first}) for a continuum of types.
-By assuming mutation acts as diffusion the effect of mutation causes $\nu(x,t)$ to flatten out over time.
+By assuming mutation acts as diffusion, the effect of mutation causes $\nu(x,t)$ to flatten out over time.
 In fact, if the growth rate is constant across $x$, then this model of mutation will cause $\nu(x,t)$ to converge to a flat line as $t\to\infty$.
@@
 \begin{equation}
-m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\nu(x,t)
-\end{equation}
-where $a\geq0$ is the strength of stabilizing selection around the phenotypic optimum $\theta\in\mathbb{R}$, $c\geq0$ captures the sensitivity of fitness to intraspecific competition and $r\in\mathbb{R}$ is the intrinsic growth rate in the absence of abiotic selection.
+m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_{-\infty}^{+\infty}\nu(y,t)dy
+\end{equation}
+where $r\in\mathbb{R}$ is the intrinsic growth rate in the absence of abiotic selection, $a\geq0$ is the strength of stabilizing selection around the phenotypic optimum $\theta\in\mathbb{R}$, $c\geq0$ captures the sensitivity of fitness to intraspecific competition and we have set $\kappa(x,y)=1$.
 This exemplary fitness function has a few convenient properties that we breifly take note of.
-First, the effect of competition induces a carrying capacity on the population, leading to a finite equilibrium abundance over bounded subsets of $\mathbb{R}$.
-Second, abiotic selection prevents the abundance density from diffusing too far from the abiotic optimum, leading to a finite equilibrium abundance across all of $\mathbb{R}$.
-A stochastic generalization of this model, closely related to our model of diffuse coevolution presented later, has been studied by Dawson [-@Dawson1978].
+First, the effect of competition induces a local carrying capacity on the population, leading to a finite equilibrium abundance over bounded subsets of $\mathbb{R}$.
+Second, abiotic selection prevents the abundance density from diffusing too far from the abiotic optimum, leading to a finite equilibrial abundance across all of $\mathbb{R}$, $\lim_{t\to\infty}N(t)=\tfrac{1}{c}(r-\tfrac{1}{2}\sqrt{a\mu})$, finite equilibrial mean trait $\lim_{t\to\infty}\bar x(t)=\theta$ and finite equilibrial phenotypic variance $\lim_{t\to\infty}\sigma^2(t)=\sqrt{\mu/a}$, so long as $|\bar x(0)|<+\infty$, $\sigma^2(0),N(0)>0$ and $r\geq\sqrt{a\mu}/2$.
+We prove these statements in SM \S\ref{equilib}.
+By replacing $\kappa(x,y)=1$ with a particular Gaussian function, we will find in \S\ref{coev} and SM \S\ref{diffuse} this fitness function becomes equivalent to a fitness function derived from niche theory that we use to model coevolution driven by resource competition.
 
@@
 The limiting processes of rescaled BBM have been referred to as measure-valued Markov processes [@dawson1993measure] or superprocesses [@alisonetheridge2000].
-Under the simplifying assumptions inherited from our treatment of deterministic dynamics and an additional assumption on demographic stochasticity, we obtain as a special case a relatively simple expression for an SPDE that generalizes equation (\ref{eq1}).
+Under the simplifying assumptions inherited from our treatment of deterministic dynamics and the additional assumption that the magnitude of demographic stochasticity is independent of trait values, we obtain as a special case a relatively simple expression for an SPDE that generalizes equation (\ref{eq1}).
 The simplicity of our special case allows us to use properties of space-time white noise processes to derive a set of SDE that generalize equations (\ref{N_det}), (\ref{detmean}) and (\ref{vardyn_det}) to include the effects of demographic stochasticity.
@@
 Treating only the simplist of cases, we then provide a brief review of BBM, their diffusion limits and the SPDE that characterize them.
-For those not interested in the white noise calculus or superprocesses and would rather jump straight into more biologically relevant results, we recommend skipping to \S\ref{particular}.
+For those not interested in the white noise calculus or superprocesses and would rather jump straight into more biologically relevant results, we recommend skipping to \S\ref{equations}.
 
@@
 
-Throughout this supplement, we set use dot notation for time derivatives so that $\dot f(x,t)=\frac{\partial }{\partial t}f(x,t)$ and set $\Delta=\frac{\partial^2}{\partial x^2}$, except in \S\ref{caC} where $\Delta$ represents a random variable.
+Throughout this supplement, we set use dot notation for time derivatives so that $\dot f(x,t)=\frac{\partial }{\partial t}f(x,t)$ and set $\Delta=\sum_{i=1}^d\frac{\partial^2}{\partial x_i^2}$ the Laplace operator on $\mathbb{R}^d$, except in \S\ref{caC} where $\Delta$ represents a random variable.
 
@@
 Following our assumptions of the main text, we have that $m(y,x)$ is differentiable with respect to both $x$ and $y$ and there exists $r\in\mathbb{R}$ such that $m(y,x)\leq r$ across all $y\geq0$ and $x\in\mathbb{R}$.
-Hence, we can apply Proposition 1.21 of Cantrell and Cosner [-@Cantrell2004] to show that $\nu(x,t)$ remains differentiable with respect to $t$ and twice differentiable with respect to $x$ for all $t\geq0$.
-
 As in the main text, we also assume the initial condition $\nu(x,0)$ satisfies
@@
 
+## Regularity of deterministic solution \label{regular}
+
+Here we show that solutions to the PDE described in the main text are continuous in time and space, twice continuously differentiable with respect to space and differentiable with respect to time.
+Our approach is an adaptation of that taken by Volpert [-@Volpert2014].
+
+Recall that the PDE we consider is of the form
+\begin{equation}\label{PDE_supp}
+\dot\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\Delta\nu(x,t),
+\end{equation}
+where $m(\nu,x)$ bounded above by $r\in\mathbb{R}$ is shorthand for $m(K(\nu)(x,t),x)$ and $K(\nu)(x,t)=\int_\mathbb{R}\kappa(x,y)\nu(y,t)dy$ for some non-negative, bounded $\kappa$.
+Denoting $L^1(\mathbb{R})$ the Lebesgue space of integrable functions on $\mathbb{R}$ (ie, the set of functions $f:\mathbb{R}\to\mathbb{R}$ that satisfy $\|f\|_1=\int_\mathbb{R}|f(x)|dx<+\infty$ under the equivalence relation $f=g$ iff $\|f-g\|_1=0$), we assume $\nu(\cdot,0)\in L^1(\mathbb{R})$.
+Under these assumptions, we have already proved in SM \S\ref{finite} that $\nu(\cdot,t)\in L^1(\mathbb{R})$ for all $t\geq0$.
+Before proving the regularity of $\nu(x,t)$, we review the approach taken by Volpert [-@Volpert2014].
+
+__Volperts approach__
+
+In chapter 9 \S2.2 of Volpert [-@Volpert2014] complimentary conditions are assumed where $\kappa(x,y)=\varphi(x-y)$ for some integrable function $\varphi$ and initial condition $\nu(\cdot,0)\in L^\infty(\mathbb{R})$, where $L^\infty(\mathbb{R})$ is the Lebesgue space of bounded functions on $\mathbb{R}$ (ie, the set of functions $f:\mathbb{R}\to\mathbb{R}$ satisfying $\|f\|_\infty=\sup_{x\in\mathbb{R}}|f(x)|<+\infty$ under the equivalence relation $f=g$ iff $\|f-g\|_\infty=0$).
+Instead of the growth term $m(\nu,x)\nu(x,t)$ of PDE (\ref{PDE_supp}), Volpert worked with $f(\nu(x,t),K(\nu)(x,t))$ for some function $f:\mathbb{R}^2\to\mathbb{R}$ differentiable with respect to both arguments and satisfying, for some $q\geq0$,
+\begin{equation}
+f(x,y)\leq q(1+|x|+|y|), \ \forall \ x,y\in\mathbb{R}.
+\end{equation}
+Under these conditions, Volpert used semigroup theory to prove the solution $\nu(x,t)$ takes values in $L^\infty(\mathbb{R})$ for all $t\geq0$ and is differentiable with respect to time and twice continuously differentiable with respect to space.
+In particular, defining $\nu(t)$ as the function $x\to\nu(x,t)$ and keeping $\Phi(x,t)$ as defined in SM \S\ref{finite}, Volpert states the semigroup $\{S(t)\}_{t\geq0}$ given by
+\begin{equation}
+S(t)\nu(0)=\int_0^t\frac{\mu}{2}\Delta\nu(x,s)ds=\int_\mathbb{R}\Phi(x-y,t)\nu(y,0)dy, \ \ \forall \ \ t\geq0, \ \nu(0)\in L^\infty(\mathbb{R})
+\end{equation}
+satisfies $\|S(t)\nu(0)\|_\infty\leq\|\nu(0)\|_\infty$ for all $t\geq0$ and $\nu(0)\in L^\infty(\mathbb{R})$. Then, by setting
+\begin{equation}
+F(v)=f\left(v(x),\int_\mathbb{R}\varphi(x-y)v(y)dy\right), \ \ \forall \ \ v\in L^\infty(\mathbb{R}),
+\end{equation}
+Volpert notes that $F$ is locally Lipschitz continuous on $L^\infty(\mathbb{R})$ and in particular $\|F(v)\|_\infty\leq q\|v\|_\infty$.
+Hence, under Volperts conditions, solutions to PDE (\ref{PDE_supp}) can be written as
+\begin{equation}\label{volp_sol}
+\nu(t)=S(t)\nu(0)+\int_0^tS(t-s)F(\nu(s))ds\in L^\infty(\mathbb{R}) \ \ \forall \ \ t\geq0, \ \nu(0)\in L^\infty(\mathbb{R}).
+\end{equation}
+Thanks to properties of the semigroup $\{S(t)\}_{t\geq0}$, solutions $\nu(t)$ given by (\ref{volp_sol}) satisfy the claimed regularity conditions.
+
+__Adapting Volperts approach to our context__
+
+To adapt Volperts approach to our situation, note that $\|S(t)\nu(0)\|_1=\|\nu(0)\|_1$ for all $\nu(0)\in L^1(\mathbb{R})$ and, setting
+\begin{equation}
+F(v)=m\left(\int_\mathbb{R}\kappa(x,y)v(y)dy,x\right)v(x), \ \ \forall \ \ v\in L^1(\mathbb{R}),
+\end{equation}
+we have $\|F(v)\|_1\leq r\|v\|_1$ for all $v\in L^1(\mathbb{R})$.
+Hence, using the same arguments employed by Volpert [-@Volpert2014], we have that solutions to PDE (\ref{PDE_supp}) under our assumptions can be written as
+\begin{equation}
+\nu(t)=S(t)\nu(0)+\int_0^tS(t-s)F(\nu(s))ds\in L^1(\mathbb{R}) \ \ \forall \ \ t\geq0, \ \nu(0)\in L^1(\mathbb{R})
+\end{equation}
+which therefore satisfy the claimed regularity conditions.
+
 ## The relation between diffusion and convolution with a Gaussian kernel \label{diffconvequiv}
@@
 
+## Equilibrium moments for a deterministic population experiencing logistic growth and stabilizing selection \label{equilib}
+
+Here we set out to show, given appropriate initial conditions, $N_\infty=\lim_{t\to\infty}N(t)=\tfrac{1}{c}(r-\tfrac{1}{2}\sqrt{a\mu})$, $\bar x_\infty=\lim_{t\to\infty}\bar x(t)=\theta$ and $\sigma^2_\infty=\lim_{t\to\infty}\sigma^2(t)=\sqrt{\tfrac{\mu}{a}}$ when
+\begin{equation}\label{m_log_stab}
+m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_\mathbb{R}\nu(y,t)dy
+\end{equation}
+for $\theta\in\mathbb{R}$, $\mu\geq0$, $a,c>0$ and $r\geq\tfrac{1}{2}\sqrt{\mu a}$.
+Set $\nu_\infty(x)=\lim_{t\to\infty}\nu(x,t)$ and $p_\infty(x)=\nu_\infty(x)/N_\infty$, when these quantities exist.
+The mean fitness can be calculated as
+\begin{equation}
+\bar m(t)=r-\frac{a}{2}((\theta-\bar x(t))^2-\sigma^2(t))-cN(t), \ \ \forall \ \ t\geq0.
+\end{equation}
+Solving for $N_\infty$ amounts to setting $\dot N(t)=0$ and solving for $N(t)$. This reduces to solving $\bar m(t)=0$ for $N(t)$, which returns
+\begin{equation}
+N_\infty=\frac{1}{c}\left(r-\frac{a}{2}\left((\theta-\bar x_\infty)^2-\sigma^2_\infty\right)\right).
+\end{equation}
+To find $\nu_\infty(x)$ we solve $\tfrac{\partial}{\partial t}\nu(x,t)=0$ for $\nu(x,t)$. This implies the following ordinary differential equation
+\begin{equation}
+\frac{d^2}{dx^2}\nu_\infty(x)=\left(\frac{2c}{\mu}N_\infty+\frac{a}{\mu}(\theta-x)^2-\frac{2r}{\mu}\right)\nu_\infty(x)
+\end{equation}
+which has the solution
+\begin{equation}
+\nu_\infty(x)=\frac{N_\infty}{\sqrt{2\pi}}\left(\frac{a}{\mu}\right)^{\frac{1}{4}}\exp\left(-\frac{(\theta-x)^2}{2\sqrt{\mu/a}}\right).
+\end{equation}
+From this expression we infer $\bar x_\infty=\theta$ and $\sigma^2_\infty=\sqrt{\mu/a}$.
+Hence $N_\infty=\frac{1}{c}\left(r-\frac{1}{2}\sqrt{a\mu}\right)$.
+This holds for any initial $\nu(\cdot,0)\in L^1(\mathbb{R})$ such that $\bar x(0)\in\mathbb{R}, \ \sigma^2(0),N(0)>0$.
+
 ## Deterministic dynamics of $\sigma^2(t)$ \label{var_deriv}
@@
 \begin{multline}
-\dot{\sigma}^2(t)=\frac{d}{dt}\int_\mathbb{R}(x-\bar x(t))^2p(x,t)dx=\int_\mathbb{R}2(x-\bar x(t))\dot{\bar x}(t)+(x-\bar x(t))^2\dot p(x,t)dx\\=\int_\mathbb{R}(x-\bar x(t))^2\left((m(\nu,x)-\bar m(t))p(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t)\right)dx\\=\int_\mathbb{R}\left((x-\bar x(t))^2-\sigma^2(t)+\sigma^2(t)\right)(m(\nu,x)-\bar m(t))p(x,t)+(x-\bar x(t))^2\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t)dx\\=\mathrm{Cov}_t\Big((x-\bar x(t))^2,m(\nu,x)\Big)+\frac{\mu}{2}\int_\mathbb{R}(x-\bar x(t))^2\frac{\partial^2}{\partial x^2}p(x,t)dx.
+\frac{d}{dt}\sigma^2(t)=\frac{d}{dt}\int_\mathbb{R}(x-\bar x(t))^2p(x,t)dx=\int_\mathbb{R}2(x-\bar x(t))\frac{d}{dt}{\bar x}(t)+(x-\bar x(t))^2\frac{\partial}{\partial t} p(x,t)dx\\=\int_\mathbb{R}(x-\bar x(t))^2\left((m(\nu,x)-\bar m(t))p(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t)\right)dx\\=\int_\mathbb{R}\left((x-\bar x(t))^2-\sigma^2(t)+\sigma^2(t)\right)(m(\nu,x)-\bar m(t))p(x,t)+(x-\bar x(t))^2\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t)dx\\=\mathrm{Cov}_t\Big((x-\bar x(t))^2,m(\nu,x)\Big)+\frac{\mu}{2}\int_\mathbb{R}(x-\bar x(t))^2\frac{\partial^2}{\partial x^2}p(x,t)dx.
 \end{multline}
@@
 \begin{multline}
-\sigma^2\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right)=\sigma^2\left(\frac{\partial}{\partial\bar x}\int_\mathbb{R}m(\nu,x)p(x,t)dx-\int_\mathbb{R}p(x,t)\frac{\partial}{\bar x}m(\nu,x)dx\right)\\=\sigma^2\int_\mathbb{R}m(\nu,x)\frac{\partial}{\partial\bar x}p(x,t)dx=\sigma^2\int_\mathbb{R}\frac{x-\bar x(t)}{\sigma^2}m(\nu,x)p(x,t)dx\\=\int_\mathbb{R}(x-\bar x)(m(\nu,x)-\bar m)p(x,t)dx=\mathrm{Cov}_t(m,x),
+\sigma^2\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right)=\sigma^2\left(\frac{\partial}{\partial\bar x}\int_\mathbb{R}m(\nu,x)p(x,t)dx-\int_\mathbb{R}p(x,t)\frac{\partial}{\partial \bar x}m(\nu,x)dx\right)\\=\sigma^2\int_\mathbb{R}m(\nu,x)\frac{\partial}{\partial\bar x}p(x,t)dx=\sigma^2\int_\mathbb{R}\frac{x-\bar x(t)}{\sigma^2}m(\nu,x)p(x,t)dx\\=\int_\mathbb{R}(x-\bar x)(m(\nu,x)-\bar m)p(x,t)dx=\mathrm{Cov}_t(m,x),
 \end{multline}
@@
 the density process $\nu(x,t)$ defined by SPDE (\ref{SPDE}) of the main text satisfies $\int_\mathbb{R}(|x|+x^2)\nu(x,t)dx<\infty$.
-From SM \S\ref{finite}, under the assumption $m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_\mathbb{R}\nu(x,t)dx$, we can derive the differential equations
-
-\begin{subequations}
-\begin{equation}
-\dot{\bar x}=aG(\theta-\bar x)
-\end{equation}
-\begin{equation}
-\dot G=\mu-aG^2
-\end{equation}
-\begin{equation}
-\dot{N}=\left(r-\frac{a}{2}\left((\theta-\bar x)^2+G+\eta\right)-cN\right)N.
-\end{equation}
-\end{subequations}
-
-Ignoring the trivial case of $N=0$, the equilibrium
-
-\begin{subequations}
-\begin{equation}
-\hat{\bar x}=\theta,
-\end{equation}
-\begin{equation}
-\hat G=\sqrt\frac{\mu}{a},
-\end{equation}
-\begin{equation}
-\hat N=\frac{1}{c}\left(r-\frac{1}{2}(\eta a+\sqrt{\mu a})\right),
-\end{equation}
-\end{subequations}
-
-is unique and globally stable for $a,c,\mu>0$.
-Setting $2r>\eta a+\sqrt{\mu a}$ ensures a positive equilibrium abundance and setting $c<r-(\eta a+\sqrt{\mu a})/2$ ensures $\hat N>1$, which is important for numerical simulations when $N$ is an integer.
+
+and setting $c<r-(\eta a+\sqrt{\mu a})/2$ ensures $\hat N>1$, which is important for numerical simulations when $N$ is an integer.
 We use these results to help guide our choice of parameter values for simulations of the branching random walk.

### 2020-04-06 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
ALL of the rigorous proofs are finally complete

@@
 \end{equation}
-where $r\in\mathbb{R}$ is the intrinsic growth rate in the absence of abiotic selection, $a\geq0$ is the strength of stabilizing selection around the phenotypic optimum $\theta\in\mathbb{R}$, $c\geq0$ captures the sensitivity of fitness to intraspecific competition and we have set $\kappa(x,y)=1$.
+where $r\in\mathbb{R}$ is the intrinsic growth rate in the absence of abiotic selection, $a>0$ is the strength of stabilizing selection around the phenotypic optimum $\theta\in\mathbb{R}$, $c>0$ captures the sensitivity of fitness to intraspecific competition and we have set $\kappa(x,y)=1$.
 This exemplary fitness function has a few convenient properties that we breifly take note of.

### 2020-04-06 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
ALL of the rigorous proofs are finally complete

@@
 Taking the same approach, we derive the dynamics of the phenotypic distribution $p(x,t)$ in response to selection and muation as
+
 \begin{multline}\label{p_PDE}
@@
 \end{multline}
-
 This result closely resembles Kimura's continuum-of-alleles model [@Kimura1965; @Brger1986].
@@
 The term $c_i\mathcal{O}_{ij}(x_i,x_j)$ coincides with a special case of a term used to capture competition in Dawson's geostochastic logistic model, an SPDE model developed to study the combined effects of demographic stochasticity, spatial dispersion and locally finite carrying capacity [@Dawson1978].
+In relation to the example fitness function discussed in \S\ref{deterministic}, consider $\kappa(x_i,x_j)=\mathcal{O}_{ij}(x_i,x_j)$.
 In SM \S\ref{diffuse} we combine our treatment of resource competition with equations (\ref{N}), (\ref{xbarfinal}) and (\ref{Gfinal}) to find
@@
 Previous work has shown the importance of demographic stochasticity [@Schreiber2017] and evolutionary dynamics [@Schreiber2018] in determining ceoxistence of competing species.
-However, these studies consider evolution and stochasticity separately.
+However, these studies consider evolution and demographic stochasticity separately.
 With the connection outlined in this section formally established, researchers may pursue a postmodern coexistence theory that naturally includes both the evolutionary dynamics of populations and the effects of demographic stochasticity in a simple synthetic framework.

### 2020-04-06 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
fixed a few typos

@@
 First, the effect of competition induces a local carrying capacity on the population, leading to a finite equilibrium abundance over bounded subsets of $\mathbb{R}$.
-Second, abiotic selection prevents the abundance density from diffusing too far from the abiotic optimum, leading to a finite equilibrial abundance across all of $\mathbb{R}$, $\lim_{t\to\infty}N(t)=\tfrac{1}{c}(r-\tfrac{1}{2}\sqrt{a\mu})$, finite equilibrial mean trait $\lim_{t\to\infty}\bar x(t)=\theta$ and finite equilibrial phenotypic variance $\lim_{t\to\infty}\sigma^2(t)=\sqrt{\mu/a}$, so long as $|\bar x(0)|<+\infty$, $\sigma^2(0),N(0)>0$ and $r\geq\sqrt{a\mu}/2$.
+Second, abiotic selection prevents the abundance density from diffusing too far from the abiotic optimum, leading to a finite equilibrial abundance across all of $\mathbb{R}$, $\lim_{t\to\infty}N(t)=\tfrac{1}{c}(r-\tfrac{1}{2}\sqrt{a\mu})$, finite equilibrial mean trait $\lim_{t\to\infty}\bar x(t)=\theta$ and finite equilibrial phenotypic variance $\lim_{t\to\infty}\sigma^2(t)=\sqrt{\mu/a}$, so long as $\bar x(0)\in\mathbb{R}$, $\sigma^2(0),N(0)\in(0,+\infty)$ and $r>\sqrt{a\mu}/2$.
 We prove these statements in SM \S\ref{equilib}.
@@
 
-Recall that $m(\nu,x)$ is shorthand for $m(\nu(x,t),x)$.
+Recall that $m(\nu,x)$ is shorthand for $m(K(\nu)(x,t),x)$.
 That is, $m:[0,+\infty)\times\mathbb{R}\to\mathbb{R}$.
-Following our assumptions of the main text, we have that $m(y,x)$ is differentiable with respect to both $x$ and $y$ and there exists $r\in\mathbb{R}$ such that $m(y,x)\leq r$ across all $y\geq0$ and $x\in\mathbb{R}$.
+Following our assumptions of the main text, we have that $m(h,x)$ is differentiable with respect to both $x$ and $h$ and there exists $r\in\mathbb{R}$ such that $m(h,x)\leq r$ across all $h\geq0$ and $x\in\mathbb{R}$.
 As in the main text, we also assume the initial condition $\nu(x,0)$ satisfies
@@
 \end{equation}
-which implies there exists $y\geq0$ and $x\in\mathbb{R}$ such that $m(y,x)>r$.
-But this contradicts our assumption $m(y,x)\leq r$ for all $y\geq0$ and $x\in\mathbb{R}$.
+which implies there exists $h\geq0$ and $x\in\mathbb{R}$ such that $m(h,x)>r$.
+But this contradicts our assumption $m(h,x)\leq r$ for all $h\geq0$ and $x\in\mathbb{R}$.
 So we have $\nu(x,t)\leq\nu_r(x,t)$ for each $x\in\mathbb{R}$ and $t\geq0$.
@@
 \end{equation}
-for $\theta\in\mathbb{R}$, $\mu\geq0$, $a,c>0$ and $r\geq\tfrac{1}{2}\sqrt{\mu a}$.
-Set $\nu_\infty(x)=\lim_{t\to\infty}\nu(x,t)$ and $p_\infty(x)=\nu_\infty(x)/N_\infty$, when these quantities exist.
-The mean fitness can be calculated as
+for $\theta\in\mathbb{R}$, $\mu\geq0$, $a,c>0$ and $r>\tfrac{1}{2}\sqrt{\mu a}$. When it exists, set $\nu_\infty(x)=\lim_{t\to\infty}\nu(x,t)$. The mean fitness can be calculated as
 \begin{equation}
@@
 Hence $N_\infty=\frac{1}{c}\left(r-\frac{1}{2}\sqrt{a\mu}\right)$.
-This holds for any initial $\nu(\cdot,0)\in L^1(\mathbb{R})$ such that $\bar x(0)\in\mathbb{R}, \ \sigma^2(0),N(0)>0$.
+This holds for any initial $\nu(\cdot,0)\in L^1(\mathbb{R})$ such that $\bar x(0)\in\mathbb{R}, \ \sigma^2(0),N(0)\in(0,+\infty)$.
 

### 2020-04-07 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
expanded volperts argument

@@
 We denote by $K$ an operator that takes functions such as $\nu(x,t)$ as its argument and returns other functions that may still depend on the spatial variable $x$ and the temporal variable $t$.
-For our application we will be focused on operators $K$ that can be written as $K(\nu)(x,t)=\int_{-\infty}^{+\infty}\kappa(x,y)\nu(y,t)dy$ for some non-negative and bounded bivariate function $\kappa$.
-We set $h=K(\nu)(x,t)$ to account for nonlocal effects of the abundance density $\nu(x,t)$ on the fitness of individuals with trait value $x$ [@Volpert2014].
+For our application we will be focused on operators $K$ that can be written as $(K\nu)(x,t)=\int_{-\infty}^{+\infty}\kappa(x,y)\nu(y,t)dy$ for some non-negative and bounded bivariate function $\kappa$.
+We set $h=(K\nu)(x,t)$ to account for nonlocal effects of the abundance density $\nu(x,t)$ on the fitness of individuals with trait value $x$ [@Volpert2014].
 An example of a nonlocal effect is the competition for a common resource among all individuals regardless of their trait value so that individuals with differing trait values effect each others fitness.
 The term nonlocal comes from interpreting the trait value as spatial location.
-From here on we abbreviate $m(h,x)=m(K(\nu)(x,t),x)$ to just $m(\nu,x)$.
+From here on we abbreviate $m(h,x)=m\big((K\nu)(x,t),x\big)$ to just $m(\nu,x)$.
 Third, we assume mutation is captured by diffusion with coefficient $\frac{\mu}{2}$.
@@
 This PDE is semilinear due to the dependency of the growth rate $m(\nu,x)$ on the solution $\nu(x,t)$ and is referred to as a scalar reaction-diffusion equation [@lawrenceevans2010].
-Interpreting the trait value $x$ as spatial location and ignoring nonlocal effects, which can be satisfied by setting $K(\nu)=\nu$, this model of spatially distributed population dynamics has been intensely studied [for a review, see @Cantrell2004].
+Interpreting the trait value $x$ as spatial location and ignoring nonlocal effects, which can be satisfied by setting $K\nu=\nu$, this model of spatially distributed population dynamics has been intensely studied [for a review, see @Cantrell2004].
 When $\mu=0$, equation (\ref{eq1}) can be seen as an analog of equation (\ref{first}) for a continuum of types.
@@
 
-Recall that $m(\nu,x)$ is shorthand for $m(K(\nu)(x,t),x)$.
+Recall that $m(\nu,x)$ is shorthand for $m\big((K\nu)(x,t),x\big)$.
 That is, $m:[0,+\infty)\times\mathbb{R}\to\mathbb{R}$.
@@
 
-Here we show that solutions to the PDE described in the main text are continuous in time and space, twice continuously differentiable with respect to space and differentiable with respect to time.
+Here we show that solutions to the PDE described in the main text are twice continuously differentiable with respect to space and differentiable with respect to time.
 Our approach is an adaptation of that taken by Volpert [-@Volpert2014].
@@
 \end{equation}
-where $m(\nu,x)$ bounded above by $r\in\mathbb{R}$ is shorthand for $m(K(\nu)(x,t),x)$ and $K(\nu)(x,t)=\int_\mathbb{R}\kappa(x,y)\nu(y,t)dy$ for some non-negative, bounded $\kappa$.
+where $m(\nu,x)$ bounded above by $r\in\mathbb{R}$ is shorthand for $m((K\nu)(x,t),x)$ and $(K\nu)(x,t)=\int_\mathbb{R}\kappa(x,y)\nu(y,t)dy$ for some non-negative, bounded $\kappa$.
 Denoting $L^1(\mathbb{R})$ the Lebesgue space of integrable functions on $\mathbb{R}$ (ie, the set of functions $f:\mathbb{R}\to\mathbb{R}$ that satisfy $\|f\|_1=\int_\mathbb{R}|f(x)|dx<+\infty$ under the equivalence relation $f=g$ iff $\|f-g\|_1=0$), we assume $\nu(\cdot,0)\in L^1(\mathbb{R})$.
@@
 In chapter 9 \S2.2 of Volpert [-@Volpert2014] complimentary conditions are assumed where $\kappa(x,y)=\varphi(x-y)$ for some integrable function $\varphi$ and initial condition $\nu(\cdot,0)\in L^\infty(\mathbb{R})$, where $L^\infty(\mathbb{R})$ is the Lebesgue space of bounded functions on $\mathbb{R}$ (ie, the set of functions $f:\mathbb{R}\to\mathbb{R}$ satisfying $\|f\|_\infty=\sup_{x\in\mathbb{R}}|f(x)|<+\infty$ under the equivalence relation $f=g$ iff $\|f-g\|_\infty=0$).
-Instead of the growth term $m(\nu,x)\nu(x,t)$ of PDE (\ref{PDE_supp}), Volpert worked with $f(\nu(x,t),K(\nu)(x,t))$ for some function $f:\mathbb{R}^2\to\mathbb{R}$ differentiable with respect to both arguments and satisfying, for some $q\geq0$,
+Instead of the growth term $m(\nu,x)\nu(x,t)$ of PDE (\ref{PDE_supp}), Volpert worked with $f(\nu(x,t),(K\nu)(x,t))$ for some function $f:\mathbb{R}^2\to\mathbb{R}$ differentiable with respect to both arguments and satisfying, for some $q\geq0$,
 \begin{equation}
@@
 \end{equation}
-satisfies $\|S(t)\nu(0)\|_\infty\leq\|\nu(0)\|_\infty$ for all $t\geq0$ and $\nu(0)\in L^\infty(\mathbb{R})$. Then, by setting
+satisfies $\|S(t)\nu(0)\|_\infty\leq\|\nu(0)\|_\infty$ for all $t\geq0$ and $\nu(0)\in L^\infty(\mathbb{R})$. By setting
 \begin{equation}
@@
 Volpert notes that $F$ is locally Lipschitz continuous on $L^\infty(\mathbb{R})$ and in particular $\|F(v)\|_\infty\leq q\|v\|_\infty$.
-Hence, under Volperts conditions, solutions to PDE (\ref{PDE_supp}) can be written as
+Hence, under Volperts conditions, solutions to PDE (\ref{PDE_supp}), given they exist, can be written as
 \begin{equation}\label{volp_sol}
-\nu(t)=S(t)\nu(0)+\int_0^tS(t-s)F(\nu(s))ds\in L^\infty(\mathbb{R}) \ \ \forall \ \ t\geq0, \ \nu(0)\in L^\infty(\mathbb{R}).
-\end{equation}
-Thanks to properties of the semigroup $\{S(t)\}_{t\geq0}$, solutions $\nu(t)$ given by (\ref{volp_sol}) satisfy the claimed regularity conditions.
+\nu(t)=S(t)\nu(0)+\int_0^tS(t-s)F(\nu(s))ds \ \ \forall \ \ t\geq0, \ \nu(0)\in L^\infty(\mathbb{R}).
+\end{equation}
+Volpert interprets the right-hand-side of equation (\ref{volp_sol}) as an operator $T$ acting on functions $v\in L^\infty(\mathbb{R}\times[0,\tau])$ with $\tau>0$.
+The space $L^\infty(\mathbb{R}\times[0,\tau])$ is given the norm $\|v\|_\infty^\infty=\max_{t\in[0,\tau]}\|v(t)\|_\infty$.
+The operator $T$ is claimed to have a single fixed point for sufficiently small $\tau$, which means that there exists exactly one function $v$ such that $Tv=v$.
+That is, there exists $v\in L^\infty(\mathbb{R}\times[0,\tau])$ such that
+\begin{equation}
+v(t)=(Tv)(t)=S(t)v(0)+\int_0^tS(t-s)(Fv)(s)ds\in L^\infty(\mathbb{R}) \ \ \forall \ \ t\in[0,\tau].
+\end{equation}
+Volpert merely states this is easy to conclude and the regularity conditions follow from it. We expand on why this is so.
+
+Combining the results $\|S(t)v(s)\|_\infty\leq\|v(s)\|_\infty$, $\|Fv(s)\|_\infty\leq q\|v(s)\|_\infty$ for all $t\geq0$, $s\in[0,\tau]$ and $v\in L^\infty(\mathbb{R}\times[0,\tau])$ with Minkowski's inequality, we have
+\begin{equation}
+\|Tv\|_\infty^\infty\leq\|v\|_\infty^\infty+q\tau\|v\|_\infty^\infty=(1+q\tau)\|v\|_\infty^\infty.
+\end{equation}
+Hence, for each $\tau>0$, $T$ is a contraction mapping on $L^\infty(\mathbb{R}\times[0,\tau])$.
+Banach's fixed point theorem therefore implies $T$ has exactly one fixed point.
+This holds for all $\tau>0$ which implies the existence and uniqueness of solutions to PDE (\ref{PDE_supp}) under Volpert's conditions.
+Hence, the solutions are differentiable with respect to time.
+Since $S(t)v$ is twice differentiable with respect to space for all $t\geq0$ and since $f$ is differentiable with respect to both arguments, $\int_0^tS(t-s)(Fv)(s)ds$ is twice differentiable with respect to space.
+Hence, solutions to PDE (\ref{PDE_supp}) are twice differentiable with respect to space.
 
@@
 \begin{equation}
-F(v)=m\left(\int_\mathbb{R}\kappa(x,y)v(y)dy,x\right)v(x), \ \ \forall \ \ v\in L^1(\mathbb{R}),
-\end{equation}
-we have $\|F(v)\|_1\leq r\|v\|_1$ for all $v\in L^1(\mathbb{R})$.
-Hence, using the same arguments employed by Volpert [-@Volpert2014], we have that solutions to PDE (\ref{PDE_supp}) under our assumptions can be written as
-\begin{equation}
-\nu(t)=S(t)\nu(0)+\int_0^tS(t-s)F(\nu(s))ds\in L^1(\mathbb{R}) \ \ \forall \ \ t\geq0, \ \nu(0)\in L^1(\mathbb{R})
-\end{equation}
-which therefore satisfy the claimed regularity conditions.
+(Fv)(x)=m\left(\int_\mathbb{R}\kappa(x,y)v(y)dy,x\right)v(x), \ \ \forall \ \ v\in L^1(\mathbb{R}),
+\end{equation}
+we have $\|Fv\|_1\leq r\|v\|_1$ for all $v\in L^1(\mathbb{R})$. Hence, when they exist, solutions to PDE (\ref{PDE_supp}) under our assumptions can be written as
+\begin{equation}
+\nu(t)=S(t)\nu(0)+\int_0^tS(t-s)(F\nu(s))ds \ \ \forall \ \ t\geq0, \ \nu(0)\in L^1(\mathbb{R}).
+\end{equation}
+Then, using the same arguments employed by Volpert [-@Volpert2014], we conclude solutions to PDE (\ref{PDE_supp}) under our conditions exist, are unique and exhibit the desired regularity conditions.
 

### 2020-04-09 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
improved proof of existence and uniqueness for the PDE

--- 2020-04-06 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
@@
 
+## Existence, uniqueness and regularity of solution to deterministic PDE (\ref{eq1}) of the main text \label{regular}
+
+Here we show that solutions to the PDE described in the main text exist, are unique and are twice continuously differentiable with respect to space.
+Our approach is an adaptation of that taken in chapter 9 \S2.2 of Volpert [-@Volpert2014].
+
+Recall that the PDE we consider is of the form
+\begin{equation}\label{PDE_supp}
+\begin{matrix}
+\dot\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\Delta\nu(x,t), \\
+\\
+\nu(x,0)=u(x),
+\end{matrix}
+\end{equation}
+where $m(\nu,x)\leq r\in\mathbb{R}$ is shorthand for $m((K\nu)(x,t),x)$ and $(K\nu)(x,t)=\int_\mathbb{R}\kappa(x,y)\nu(y,t)dy$ for some non-negative, bounded $\kappa$.
+Denoting $C(\mathbb{R})$ the set of continuous functions on $\mathbb{R}$ and $L^1(\mathbb{R})$ the Lebesgue space of integrable functions on $\mathbb{R}$ (ie, the set of functions $f:\mathbb{R}\to\mathbb{R}$ that satisfy $\|f\|_1=\int_\mathbb{R}|f(x)|dx<+\infty$ under the equivalence relation $f=g$ iff $\|f-g\|_1=0$), we assume a continuous and integrable initial condition $\nu(\cdot,0)=u\in L^1(\mathbb{R})\cap C(\mathbb{R})$.
+We now prove the existence and uniqueness of solutions to this Cauchy problem.
+
+Set 
+\begin{equation}
+\Phi(x,t)=\frac{\exp\left(-x^2/2\mu t\right)}{\sqrt{2\pi\mu t}}.
+\end{equation}
+Then the semigroup $\{S(t)\}_{t\geq0}$ given by 
+\begin{equation}
+S(t)v=\left\{\begin{matrix}
+\int_0^t\frac{\mu}{2}\Delta\nu(x,s)ds=\int_\mathbb{R}\Phi(x-y,t)v(y)dy, & t>0 \\
+ & \\
+v, &t=0
+\end{matrix}\right.
+\end{equation}
+for all $v\in L^\infty(\mathbb{R})$ satisfies $\|S(t)v\|_1=\|v\|_1$ for all $t\geq0$ and $v\in L^1(\mathbb{R})$. We also define the operator
+\begin{equation}
+(Fv)(x)=m\left(\int_\mathbb{R}\kappa(x,y)v(y)dy,x\right)v(x), \ \ \forall \ \ v\in L^1(\mathbb{R}).
+\end{equation}
+Thus $\|Fv\|_1\leq |r| \ \|v\|_1$ for all $v\in L^1(\mathbb{R})$ which implies $F$ is Lipschitz continuous.
+For $T>0$ set $L^{\infty,1}([0,T]\times\mathbb{R})$ as the set of functions $\nu:[0,T]\to L^1(\mathbb{R})$ that satisfy $\|\nu\|_{1,T}=\sup_{t\in[0,T]}\|\nu(t)\|_1<+\infty$ and consider the operator $G$ defined on $L^{\infty,1}([0,T]\times\mathbb{R})$ for fixed $u\in L^1(\mathbb{R})$ by
+\begin{equation}
+(Gv)(t)=S(t)u+\int_0^tS(t-s)(Fv(s))ds \ \ \forall \ \ t\in[0,T].
+\end{equation}
+Then for any $t\in[0,T]$ and $\nu_1,\nu_2\in L^{\infty,1}([0,T]\times\mathbb{R})$,
+
+\begin{multline}
+\|G\nu_1-G\nu_2\|_{1,T}=\sup_{t\in[0,T]}\left\|\int_0^tS(t-s)(F\nu_1(s)-F\nu_2(s))ds\right\|_1 \\ \leq \sup_{t\in[0,T]}t|r|\|\nu_1(t)-\nu_2(t)\|_1\leq T|r|\|\nu_1-\nu_2\|_{1,T}.
+\end{multline}
+Hence, when $T<\frac{1}{|r|}$ the operator $G$ is a contraction operator.
+In this case we apply Banach's fixed point theorem to demonstrate $G$ has a unique fixed point.
+That is, there exists a unique $\nu\in L^{\infty,1}([0,T]\times\mathbb{R})$ such that $G\nu=\nu$.
+Then, for $\nu=\nu(x,t)$ and considering $\nu(t)$ the function $x\to\nu(x,t)$, we have by definition $\nu$ is the mild solution of PDE (\ref{PDE_supp}) with initial condition $\nu(0)=u$.
+To extend this to arbitrarily large $T$ we take a similar approach as Evans [-@lawrenceevans2010] and choose $T_1<\frac{1}{|r|}$ and apply the above argument to the interval $[T_1,2T_1]$.
+Repeating this task a finite number of times shows $\nu$ is the mild solution of PDE (\ref{PDE_supp}) with initial condition $\nu(0)=u$ for arbitrarily large $T>0$.
+Applying lemma 3.3.2 of Henry [-@Henry1981] we have that the mild solution $\nu$ is in fact a strong solution to the Cauchy problem defined by PDE (\ref{PDE_supp}) with initial condition $\nu(0)=u$ for $t\in[0,T]$.
+In particular, following example 3.6 of Henry [-@Henry1981], this implies that $\nu(t)\in D(\Delta)$ for each $t\in(0,T)$ where $D(\Delta)=W^{2,2}(\mathbb{R})$ is the Sobolov space of square integrable functions with weak derivatives at least of order $2$.
+Then, according to a Sobolov embedding theorem [@Henry1981], we have $D(\Delta)\subset C^2(\mathbb{R})$ and hence solutions to the Cauchy problem (\ref{PDE_supp}) are twice continuously differentiable in their spatial parameter.
+
 ## Sufficient conditions for finite mean, variance and total abundance in the deterministic case \label{finite}
 
-Recall that $m(\nu,x)$ is shorthand for $m(K(\nu)(x,t),x)$.
+Recall that $m(\nu,x)$ is shorthand for $m\big((K\nu)(x,t),x\big)$.
 That is, $m:[0,+\infty)\times\mathbb{R}\to\mathbb{R}$.
 Following our assumptions of the main text, we have that $m(h,x)$ is differentiable with respect to both $x$ and $h$ and there exists $r\in\mathbb{R}$ such that $m(h,x)\leq r$ across all $h\geq0$ and $x\in\mathbb{R}$.
-As in the main text, we also assume the initial condition $\nu(x,0)$ satisfies
+As in the main text, we also assume the initial condition $\nu(x,0)$ is continuous in $x$ and satisfies
 \begin{equation}
@@
 
+In SM \S\ref{regular} we have shown PDE (\ref{PDE_SM}) with an integrable initial condition has a classical solution.
 Replacing $m$ with it's upper bound $r\in\mathbb{R}$, equation (\ref{PDE_SM}) reduces to a simple parabolic equation that can be solved using standard techniques [@stanleyfarlow1993].
@@
 for each $t\geq0$.
-
-## Regularity of deterministic solution \label{regular}
-
-Here we show that solutions to the PDE described in the main text are continuous in time and space, twice continuously differentiable with respect to space and differentiable with respect to time.
-Our approach is an adaptation of that taken by Volpert [-@Volpert2014].
-
-Recall that the PDE we consider is of the form
-\begin{equation}\label{PDE_supp}
-\dot\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\Delta\nu(x,t),
-\end{equation}
-where $m(\nu,x)$ bounded above by $r\in\mathbb{R}$ is shorthand for $m(K(\nu)(x,t),x)$ and $K(\nu)(x,t)=\int_\mathbb{R}\kappa(x,y)\nu(y,t)dy$ for some non-negative, bounded $\kappa$.
-Denoting $L^1(\mathbb{R})$ the Lebesgue space of integrable functions on $\mathbb{R}$ (ie, the set of functions $f:\mathbb{R}\to\mathbb{R}$ that satisfy $\|f\|_1=\int_\mathbb{R}|f(x)|dx<+\infty$ under the equivalence relation $f=g$ iff $\|f-g\|_1=0$), we assume $\nu(\cdot,0)\in L^1(\mathbb{R})$.
-Under these assumptions, we have already proved in SM \S\ref{finite} that $\nu(\cdot,t)\in L^1(\mathbb{R})$ for all $t\geq0$.
-Before proving the regularity of $\nu(x,t)$, we review the approach taken by Volpert [-@Volpert2014].
-
-__Volperts approach__
-
-In chapter 9 \S2.2 of Volpert [-@Volpert2014] complimentary conditions are assumed where $\kappa(x,y)=\varphi(x-y)$ for some integrable function $\varphi$ and initial condition $\nu(\cdot,0)\in L^\infty(\mathbb{R})$, where $L^\infty(\mathbb{R})$ is the Lebesgue space of bounded functions on $\mathbb{R}$ (ie, the set of functions $f:\mathbb{R}\to\mathbb{R}$ satisfying $\|f\|_\infty=\sup_{x\in\mathbb{R}}|f(x)|<+\infty$ under the equivalence relation $f=g$ iff $\|f-g\|_\infty=0$).
-Instead of the growth term $m(\nu,x)\nu(x,t)$ of PDE (\ref{PDE_supp}), Volpert worked with $f(\nu(x,t),K(\nu)(x,t))$ for some function $f:\mathbb{R}^2\to\mathbb{R}$ differentiable with respect to both arguments and satisfying, for some $q\geq0$,
-\begin{equation}
-f(x,y)\leq q(1+|x|+|y|), \ \forall \ x,y\in\mathbb{R}.
-\end{equation}
-Under these conditions, Volpert used semigroup theory to prove the solution $\nu(x,t)$ takes values in $L^\infty(\mathbb{R})$ for all $t\geq0$ and is differentiable with respect to time and twice continuously differentiable with respect to space.
-In particular, defining $\nu(t)$ as the function $x\to\nu(x,t)$ and keeping $\Phi(x,t)$ as defined in SM \S\ref{finite}, Volpert states the semigroup $\{S(t)\}_{t\geq0}$ given by
-\begin{equation}
-S(t)\nu(0)=\int_0^t\frac{\mu}{2}\Delta\nu(x,s)ds=\int_\mathbb{R}\Phi(x-y,t)\nu(y,0)dy, \ \ \forall \ \ t\geq0, \ \nu(0)\in L^\infty(\mathbb{R})
-\end{equation}
-satisfies $\|S(t)\nu(0)\|_\infty\leq\|\nu(0)\|_\infty$ for all $t\geq0$ and $\nu(0)\in L^\infty(\mathbb{R})$. Then, by setting
-\begin{equation}
-F(v)=f\left(v(x),\int_\mathbb{R}\varphi(x-y)v(y)dy\right), \ \ \forall \ \ v\in L^\infty(\mathbb{R}),
-\end{equation}
-Volpert notes that $F$ is locally Lipschitz continuous on $L^\infty(\mathbb{R})$ and in particular $\|F(v)\|_\infty\leq q\|v\|_\infty$.
-Hence, under Volperts conditions, solutions to PDE (\ref{PDE_supp}) can be written as
-\begin{equation}\label{volp_sol}
-\nu(t)=S(t)\nu(0)+\int_0^tS(t-s)F(\nu(s))ds\in L^\infty(\mathbb{R}) \ \ \forall \ \ t\geq0, \ \nu(0)\in L^\infty(\mathbb{R}).
-\end{equation}
-Thanks to properties of the semigroup $\{S(t)\}_{t\geq0}$, solutions $\nu(t)$ given by (\ref{volp_sol}) satisfy the claimed regularity conditions.
-
-__Adapting Volperts approach to our context__
-
-To adapt Volperts approach to our situation, note that $\|S(t)\nu(0)\|_1=\|\nu(0)\|_1$ for all $\nu(0)\in L^1(\mathbb{R})$ and, setting
-\begin{equation}
-F(v)=m\left(\int_\mathbb{R}\kappa(x,y)v(y)dy,x\right)v(x), \ \ \forall \ \ v\in L^1(\mathbb{R}),
-\end{equation}
-we have $\|F(v)\|_1\leq r\|v\|_1$ for all $v\in L^1(\mathbb{R})$.
-Hence, using the same arguments employed by Volpert [-@Volpert2014], we have that solutions to PDE (\ref{PDE_supp}) under our assumptions can be written as
-\begin{equation}
-\nu(t)=S(t)\nu(0)+\int_0^tS(t-s)F(\nu(s))ds\in L^1(\mathbb{R}) \ \ \forall \ \ t\geq0, \ \nu(0)\in L^1(\mathbb{R})
-\end{equation}
-which therefore satisfy the claimed regularity conditions.
 

### 2020-04-10 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
improved formatting

@@
 We set $h=(K\nu)(x,t)$ to account for nonlocal effects of the abundance density $\nu(x,t)$ on the fitness of individuals with trait value $x$ [@Volpert2014].
-An example of a nonlocal effect is the competition for a common resource among all individuals regardless of their trait value so that individuals with differing trait values effect each others fitness.
+An example of a nonlocal effect comes from competition for a common resource among all individuals in a population regardless of their trait value.
+In this case all individuals in the population effect each others fitness' even though they may exhibit diverse trait values.
 The term nonlocal comes from interpreting the trait value as spatial location.
@@
 and make use of the convention $f=g$ if $\|f-g\|_t=0$ for all $t\geq0$.
+Later, when we investigate stochastic dynamics of the abundance density $\nu(x,t)$, we will find $\sqrt\nu\in\mathscr{N}_2$.
+This enables us to utilize the heuristics developed in this section for the derivation of SDE describing the stochastic dynamics of $N(t)$, $\bar x(t)$ and $\sigma^2(t)$.
 
@@
 \end{equation}
-Since $\nu(x,t)$ is not generally differentiable in $x$ or $t$, the derivatives in expression (\ref{neutral_SPDE}) are taken in the _weak_ sense [sensu @lawrenceevans2010].
+Since $\nu(x,t)$ is not generally differentiable in $x$ or $t$, the derivatives in expression (\ref{neutral_SPDE}) are taken in the _weak_ sense [for examples, see @lawrenceevans2010].
 That is, to rigorously interpret SPDE (\ref{neutral_SPDE}), we integrate the solution $\nu(x,t)$ against functions $f\in C_b^2(\mathbb{R})$, where $C_b^2(\mathbb{R})$ is the set of bounded and twice continuously differentiable functions on $\mathbb{R}$.
@@
 Comparing equation (\ref{SPDE}) to equation (3.5) of Li [-@zeng1998absolute], our $m$ and $V$ correspond to Li's $b$ and $c$ respectively.
-It is important to note that, under the assumptions made in Méléard and Roelly [-@meleard1992interacting; -@Mlard1993] and Li [-@zeng1998absolute], equation (\ref{SPDE}) is only formal when $m(y,x)$ is bounded across all combinations of $y\geq0$ and $x\in\mathbb{R}$.
-However, recalling our condition $m(y,x)\leq r\in\mathbb{R}$, the growth rates we consider are only bounded above.
-Yet, in the proof of the construction of the interacting superprocess as the limit of rescaled branching diffusions, Méléard and Roelly [-@meleard1992interacting; -@Mlard1993] assumed $m(y,x)$ to be bounded to guarantee the total mass process will have finite mean and variance, for finite $t\geq0$.
+It is important to note that, under the assumptions made in Méléard and Roelly [-@meleard1992interacting; -@Mlard1993] and Li [-@zeng1998absolute], equation (\ref{SPDE}) is only formal when $m(h,x)$ is bounded across all combinations of $h\geq0$ and $x\in\mathbb{R}$.
+However, recalling our condition $m(h,x)\leq r\in\mathbb{R}$, the growth rates we consider are only bounded above.
+Yet, in the proof of the construction of the interacting superprocess as the limit of rescaled branching diffusions, Méléard and Roelly [-@meleard1992interacting; -@Mlard1993] assumed $m(h,x)$ to be bounded to guarantee the total mass process will have finite mean and variance, for finite $t\geq0$.
 This allowed the authors to employ a tightness criterion for sequences of measures and show the rescaled processes converge to a superprocess with finite total mass.
-Li's [-@zeng1998absolute] result builds directly off of Méléard and Roelly's construction, inheriting the assumption of boundedness for $m(y,x)$.
-However, in Li [-@zeng1998absolute], the sufficiency of $m(y,x)$ being bounded above is even more clear since Li works explicitly with a common upperbound for both $m(y,x)$ and $V$.
-Hence, one can repeat the necessary proofs replacing the assumption that $m(y,x)$ is bounded with the assumption that $m(y,x)$ is merely bounded above to derive the same results.
+Li's [-@zeng1998absolute] result builds directly off of Méléard and Roelly's construction, inheriting the assumption of boundedness for $m(h,x)$.
+However, in Li [-@zeng1998absolute], the sufficiency of $m(h,x)$ being bounded above is even more clear since Li works explicitly with a common upperbound for both $m(h,x)$ and $V$.
+Hence, one can repeat the necessary proofs replacing the assumption that $m(h,x)$ is bounded with the assumption that $m(h,x)$ is merely bounded above to derive the same results.
 
@@
 \begin{equation}\label{sig2_gen}
-d\sigma^2(t)=\left(\mathrm{Cov}_t\Big((x-\bar x(t))^2,m(\nu,x)\Big)+\mu-V\frac{\sigma^2(t)}{N(t)}\right)dt+\sqrt{V\frac{\overline{(x-\bar x(t))^4}-\sigma^4(t)}{N(t)}}dW_3(t),
+d\sigma^2(t)=\mathrm{Cov}_t\Big((x-\bar x(t))^2,m(\nu,x)\Big)dt+\left(\mu-V\frac{\sigma^2(t)}{N(t)}\right)dt+\sqrt{V\frac{\overline{(x-\bar x(t))^4}-\sigma^4(t)}{N(t)}}dW_3(t),
 \end{equation}
@@
 
-By assuming $\nu(x,t)$ can be approximated by a Gaussian curve for each $t\geq0$, expressions (\ref{N}), (\ref{xbar_gen}) and (\ref{xbar_gen}) transform into efficient tools for deriving the dynamics of populations given a fitness function $m(\nu,x)$.
+By assuming $\nu(x,t)$ can be approximated by a Gaussian curve for each $t\geq0$, expressions (\ref{N}), (\ref{xbar_gen}) and (\ref{sig2_gen}) transform into efficient tools for deriving the dynamics of populations given a fitness function $m(\nu,x)$.
 Gaussian phenotypic distributions are often obtained through Gaussian, exponential or weak selection approximations together with a simplified model of inheritance and random mating [@pmid17248993; @Turelli1984; @Turelli1986; @Turelli2017; @9780471986539].
@@
 However, our approach adds an additional layer of difficulty.
-Even with Gaussian selection, the resulting solution to SPDE (\ref{SPDE}) will only be a Gaussian curve in expectation, assuming a Gaussian initial condition.
+Even with Gaussian selection, the resulting solution to SPDE (\ref{SPDE}) will only be a Gaussian curve in expectation, assuming a Gaussian curve as the initial condition.
 Yet this difficulty is not as challenging as it may first appear.
@@
 In this section we demonstrate the use of our framework by developing a model of diffuse coevolution across a guild of $S$ species whose interactions are mediated by resource competition along a single niche axis.
-Because our approach treats abundance dynamics and evolutionary dynamics simultaneously, this model allows us to investigate the relationship between selection gradients and competition coefficients.
+Because our approach treats abundance dynamics and evolutionary dynamics simultaneously, this model allows us to investigate the relationship between selection gradients and competition coefficients, which we carry out in what follows.
 
@@
 
-```{r, message=FALSE, echo=F, fig.align='center', fig.cap='\\label{temporal}Temporal dynamics of mean trait (top), additive genetic variance (middle) and abundance (bottom) for the scenario of weak competition (left) and strong competition (right).
+```{r, message=FALSE, echo=F, fig.align='center', out.width='100%', fig.cap='\\label{temporal}Temporal dynamics of mean trait (top), additive genetic variance (middle) and abundance (bottom) for the scenario of weak competition (left) and strong competition (right).
 Red lines indicate average trend across species.'}

### 2020-04-10 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
improved formatting

@@
 This illustrative investigation suggests selection gradients can be either positively or negatively correlated with competition coefficients depending on the ratio of interspecific trait variation to intraspecific trait variation.
-Hence, this approach can contribute to the development of a synthetic theory of evolutionary ecology by formalizing first principle derivations of dynamical equations describing populations and communities which can then be used for rigorous investigations of the relationship between feedbacks of ecological and evolutionary processes and the patterns of diversity they produce."
+Hence, this approach contributes to the development of a synthetic theory of evolutionary ecology by formalizing first principle derivations of stochastic models that underlie rigorous investigations of the relationship between feedbacks of biological processes and the patterns of diversity they produce."
 bibliography: ref.bib
@@
 \end{equation}
-where $r\in\mathbb{R}$ is the intrinsic growth rate in the absence of abiotic selection, $a>0$ is the strength of stabilizing selection around the phenotypic optimum $\theta\in\mathbb{R}$, $c>0$ captures the sensitivity of fitness to intraspecific competition and we have set $\kappa(x,y)=1$.
+where $r\in\mathbb{R}$ is the intrinsic growth rate in the absence of abiotic stabilizing selection, $a>0$ is the strength of abiotic stabilizing selection around the phenotypic optimum $\theta\in\mathbb{R}$, $c>0$ captures the sensitivity of fitness to intraspecific competition and we have set $\kappa(x,y)=1$.
 This exemplary fitness function has a few convenient properties that we breifly take note of.
@@
 \end{equation}
-In our case, we only consider the so-called multiplication operators.
+In our case we only consider the so-called multiplication operators.
 That is, processes that consist of operators $\Phi(t)$ having the form $\Phi(t)g(x)=\varphi(x,t)g(x)$ such that $\varphi(\cdot,t)\in L^2(\mathbb{R})$ a.s.
@@
 \begin{equation}
-d\sigma^2=\left(\mathrm{Cov}_t\Big((x-\bar x)^2,m\Big)+\mu-V\frac{\sigma^2}{N}\right)dt+\sqrt{V\frac{\overline{(x-\bar x)^4}-\sigma^4}{N}}d W_3.
+d\sigma^2=\mathrm{Cov}_t\Big((x-\bar x)^2,m\Big)dt+\left(\mu-V\frac{\sigma^2}{N}\right)dt+\sqrt{V\frac{\overline{(x-\bar x)^4}-\sigma^4}{N}}d W_3.
 \end{equation}

### 2020-04-10 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
improved proof of stable equilibrium for logistic growth and stabilizing selection

@@
 Third, we assume mutation is captured by diffusion with coefficient $\frac{\mu}{2}$.
-Now that these technicalities have been treated we can model the demographic dynamics of a population and the dynamics of a quantitative character simultaneously by the PDE
+With these technicalities addressed we model the demographic dynamics of a population and the dynamics of a quantitative character simultaneously by the PDE
 \begin{equation}\label{eq1}
@@
 with the initial condition $\nu(x,0)$ described above.
-This PDE is semilinear due to the dependency of the growth rate $m(\nu,x)$ on the solution $\nu(x,t)$ and is referred to as a scalar reaction-diffusion equation [@lawrenceevans2010].
+This PDE is semilinear due to the dependency of the growth rate $m(\nu,x)$ on the solution $\nu(x,t)$ and is referred to as a scalar reaction-diffusion equation [@Henry1981; @lawrenceevans2010].
 Interpreting the trait value $x$ as spatial location and ignoring nonlocal effects, which can be satisfied by setting $K\nu=\nu$, this model of spatially distributed population dynamics has been intensely studied [for a review, see @Cantrell2004].
@@
 As an example, logistic growth combined with quadratic stabilizing selection can be captured using the growth rate
-\begin{equation}
+\begin{equation}\label{special_m}
 m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_{-\infty}^{+\infty}\nu(y,t)dy
 \end{equation}
-where $r\in\mathbb{R}$ is the intrinsic growth rate in the absence of abiotic stabilizing selection, $a>0$ is the strength of abiotic stabilizing selection around the phenotypic optimum $\theta\in\mathbb{R}$, $c>0$ captures the sensitivity of fitness to intraspecific competition and we have set $\kappa(x,y)=1$.
+where $r\in\mathbb{R}$ is the intrinsic growth rate in the absence of abiotic stabilizing selection, $a>0$ is the strength of abiotic stabilizing selection around the phenotypic optimum $\theta\in\mathbb{R}$, $c>0$ captures the sensitivity of fitness to intraspecific competition and we have set $\kappa(x,y)=1$ so that competitive interactions cause the same reduction in fitness regardless of trait value.
 This exemplary fitness function has a few convenient properties that we breifly take note of.
 First, the effect of competition induces a local carrying capacity on the population, leading to a finite equilibrium abundance over bounded subsets of $\mathbb{R}$.
-Second, abiotic selection prevents the abundance density from diffusing too far from the abiotic optimum, leading to a finite equilibrial abundance across all of $\mathbb{R}$, $\lim_{t\to\infty}N(t)=\tfrac{1}{c}(r-\tfrac{1}{2}\sqrt{a\mu})$, finite equilibrial mean trait $\lim_{t\to\infty}\bar x(t)=\theta$ and finite equilibrial phenotypic variance $\lim_{t\to\infty}\sigma^2(t)=\sqrt{\mu/a}$, so long as $\bar x(0)\in\mathbb{R}$, $\sigma^2(0),N(0)\in(0,+\infty)$ and $r>\sqrt{a\mu}/2$.
+Second, abiotic selection prevents the abundance density from diffusing too far from the abiotic optimum, leading to a finite equilibrial abundance across all of $\mathbb{R}$, $\lim_{t\to\infty}N(t)=\tfrac{1}{c}(r-\tfrac{1}{2}\sqrt{a\mu})$, finite equilibrial mean trait $\lim_{t\to\infty}\bar x(t)=\theta$ and finite equilibrial phenotypic variance $\lim_{t\to\infty}\sigma^2(t)=\sqrt{\frac{\mu}{a}}$, so long as $\bar x(0)\in\mathbb{R}$, $\sigma^2(0)\in[0,+\infty)$, $N(0)\in(0,+\infty)$ and $r>\frac{1}{2}\sqrt{a\mu}>0$.
 We prove these statements in SM \S\ref{equilib}.
-By replacing $\kappa(x,y)=1$ with a particular Gaussian function, we will find in \S\ref{coev} and SM \S\ref{diffuse} this fitness function becomes equivalent to a fitness function derived from niche theory that we use to model coevolution driven by resource competition.
+By replacing $\kappa(x,y)=1$ in expression (\ref{special_m}) with a particular Gaussian function, we will find in \S\ref{coev} and SM \S\ref{diffuse} this fitness function becomes equivalent to a fitness function derived from niche theory that we use to model coevolution driven by resource competition.
 
@@
 The primary difference being that our model utilizes diffusion instead of convolution with an arbitrary mutation kernel.
-Of course, our model of mutation can be derived as an approximation to Kimura's model, which has been referred to as the Gaussian allelic approximation in reference to the distribution of mutational effects at loci in a genome on the values of traits [@Lande1975; but see also @9780471986539], the infinitesimal model in reference to modelling continuous traits as being encoded by an infinite number of loci each having infinitesimal effect [@Barton2017] and the Gaussian descendants approximation in reference to offspring trait values being normally distributed around their parental values [@Turelli2017].
+Of course, our model of mutation can be derived as an approximation to Kimura's model, which, for sexually reproducing populations, has been referred to as the Gaussian allelic approximation in reference to the distribution of mutational effects at loci in a genome on the values of traits [@Lande1975; but see also @9780471986539], the infinitesimal model in reference to modelling continuous traits as being encoded by an infinite number of loci each having infinitesimal effect [@Barton2017] and the Gaussian descendants approximation in reference to offspring trait values being normally distributed around their parental values [@Turelli2017].
 Alternatively, since diffusion is the continuous-time equivalent to convolution against a Gaussian kernel (SM \S\ref{diffconvequiv}), equation (\ref{p_PDE}) can also be seen as a special case of Kimura's continuum-of-alleles model.
 We return to these concepts in \S\ref{inheritance} where we review a simple model of inheritance that leads us to an expression for the stochastic evolution of additive genetic variance.
+To clarify the distinction between this model and previous models of phenotypic evolution we refer to it as the Deterministic Asexual Gaussian allelic model with Abundance dynamics (abbreviated DAGA).
+Later, we will extend this model to include the effects of demographic stochasticity, which we refer to as the Stochastic Asexual Gaussian allelic model with Abundance dynamics (abbreviated SAGA).
 
@@
 \end{equation}
+As mentioned above, we refer to SPDE (\ref{SPDE}) as the Stochastic Asexual Gaussian allelic model with Abundance dynamics (abbreviated SAGA).
 Comparing equation (\ref{SPDE}) to equation (3.5) of Li [-@zeng1998absolute], our $m$ and $V$ correspond to Li's $b$ and $c$ respectively.
@@
 Equation (\ref{rho_SPDE}) is a stochastic generalization of the deterministic PDE (\ref{eq1}) from \S\ref{deterministic}, but describes the evolution of the distribution of breeding values instead of expressed characters.
+However, whether modelling expressed characters or breeding values, we refer to SPDE of the form (\ref{SPDE}) as Stochastic Asexual Gaussian allelic models with Abundance dynamics (abbreviated SAGA).
+By replacing $\psi(x,g)$ with different probability densities produces different developmental models.
+This includes densities that may depend on environmental variables and hence model phenotypic plasticy.
+To leave room for the naming of other models where $\psi(x,g)$ is Gaussian, we refer to the model of development we have chosen as the _centered_ Gaussian developmental model.
 In \S\ref{wnc} we review the origins of this equation and provide some theory to help make sense of it, particularly the term $\dot W$.
@@
 
-```{r, message=FALSE, echo=F, fig.align='center', out.width='100%', fig.cap='\\label{temporal}Temporal dynamics of mean trait (top), additive genetic variance (middle) and abundance (bottom) for the scenario of weak competition (left) and strong competition (right).
-Red lines indicate average trend across species.'}
+```{r, message=FALSE, echo=F, fig.align='center', out.width='100%', fig.cap='\\label{temporal}Temporal dynamics of mean trait (top), additive genetic variance (middle) and abundance (bottom) for the scenario of weak competition (left) and strong competition (right).'}
 knitr::include_graphics('community_dynamics.jpg')
@@
 Let us further assume that richness $S$ is large and the distribution of mean trait values is normal with mean $\bar{\bar x}$, variance $V_{\bar X}$ and density $f_{\bar X}$.
-Such assumptions are typical when deriving analytical results in the field of theoretical coevolutionary community ecology [@GuimaresJr2011; @Guimares2017; @Nuismer2012; @Nuismer2018].
+Such assumptions are typical when deriving analytical results in the field of theoretical coevolutionary community ecology [@Nuismer2012; @Nuismer2018].
 If $\bar{\bar x}$ is near $\theta$ and $V_{\bar X}$ is much smaller than $|2R/a-G-\eta|$, then we may approximate $r_i$ with
@@
 
-We have introduced an approach to derive models of evolutionary ecology using the calculus of white noise, demonstrated our approach by deriving a model of diffuse coevolution and investigated the relationship between selection gradients and competition coefficients, finding that these quantities exhibit interesting relationships which shed light on the feedback between the structure and dynamics of ecological communities.
-
-Our approach has the merit of rigorously synthesizing the dynamics of abundance and phenotypic evolution for populations experiencing demographic stochasticity.
-Yet, there remains biological details and their associated technical challenges that need to be confronted for gaining a more thorough and rigorous understanding of ecological communities.
+We have introduced an approach to derive models of evolutionary ecology using the calculus of white noise, developed SAGA, a novel model of phenotypic evolution that accounts for demographic stochasticity, demonstrated our approach by deriving a model of diffuse coevolution from classical competition theory, and used this model to investigate the relationship between selection gradients and competition coefficients.
+This work demonstrates that connecting contemporary theoretical approaches to evolutionary ecology with some fundamental results in the theory of measure-valued Markov processes allows for the development of a rigorous, yet flexible approach to synthesizing the dynamics of abundance and distributions of quantitative characters.
+Although this approach enjoys these important merits, there remains biological details and their associated technical challenges that need to be confronted for gaining a more thorough and accurate understanding of ecological communities.
 We touch on just four of them here and provide suggestions for future research to approach these challenges.
 
-__Limitations of diffusion limits__
+__Limitations of diffusion limits and validating coexistence theory__
 
@@
 In particular, the pathological behavior described for the diffusion limit does not occur for BBM as population size goes to zero.
+Relating this concern to our comment in \S\ref{ecoevo} on using our model of diffuse coevolution to extend modern coexistence theory [sensu @Chesson2000] in a way that accounts for phenotypic evolution and abundance dynamics, it seems necessary to determine whether or not results pertaining to the coexistence of competing populations based on deterministic models that ignore phenotypic evolution hold using stochastic individual-based simulations.
+To stay within the framework outlined in this paper and thereby produce comparable results, these simulations should be extensions of BBM that account for the effects of competition.
 
@@
 
-Here we set out to show, given appropriate initial conditions, $N_\infty=\lim_{t\to\infty}N(t)=\tfrac{1}{c}(r-\tfrac{1}{2}\sqrt{a\mu})$, $\bar x_\infty=\lim_{t\to\infty}\bar x(t)=\theta$ and $\sigma^2_\infty=\lim_{t\to\infty}\sigma^2(t)=\sqrt{\tfrac{\mu}{a}}$ when
+Here we set out to show, given the initial conditions $\nu(\cdot,0)\in L^1(\mathbb{R})$ such that $\bar x(0)\in\mathbb{R}$, $\sigma^2(0)\in[0,+\infty)$ and $N(0)\in(0,+\infty)$, and given the growth rate
 \begin{equation}\label{m_log_stab}
@@
 \end{equation}
-for $\theta\in\mathbb{R}$, $\mu\geq0$, $a,c>0$ and $r>\tfrac{1}{2}\sqrt{\mu a}$. When it exists, set $\nu_\infty(x)=\lim_{t\to\infty}\nu(x,t)$. The mean fitness can be calculated as
-\begin{equation}
-\bar m(t)=r-\frac{a}{2}((\theta-\bar x(t))^2-\sigma^2(t))-cN(t), \ \ \forall \ \ t\geq0.
-\end{equation}
-Solving for $N_\infty$ amounts to setting $\dot N(t)=0$ and solving for $N(t)$. This reduces to solving $\bar m(t)=0$ for $N(t)$, which returns
-\begin{equation}
-N_\infty=\frac{1}{c}\left(r-\frac{a}{2}\left((\theta-\bar x_\infty)^2-\sigma^2_\infty\right)\right).
-\end{equation}
-To find $\nu_\infty(x)$ we solve $\tfrac{\partial}{\partial t}\nu(x,t)=0$ for $\nu(x,t)$. This implies the following ordinary differential equation
-\begin{equation}
-\frac{d^2}{dx^2}\nu_\infty(x)=\left(\frac{2c}{\mu}N_\infty+\frac{a}{\mu}(\theta-x)^2-\frac{2r}{\mu}\right)\nu_\infty(x)
+such that $\theta\in\mathbb{R}$, $a,c,\mu>0$ and $r>\tfrac{1}{2}\sqrt{\mu a}$, we have the following stable equilibrium $N_\infty=\lim_{t\to\infty}N(t)=\tfrac{1}{c}(r-\tfrac{1}{2}\sqrt{a\mu})$, $\bar x_\infty=\lim_{t\to\infty}\bar x(t)=\theta$ and $\sigma^2_\infty=\lim_{t\to\infty}\sigma^2(t)=\sqrt{\tfrac{\mu}{a}}$.
+
+The mean fitness can be calculated as
+\begin{equation}
+\bar m(t)=r-\frac{a}{2}((\theta-\bar x(t))^2+\sigma^2(t))-cN(t).
+\end{equation}
+Recalling the ODE derived for $N(t)$,
+\begin{equation}
+\frac{d}{dt}N(t)=\bar m(t)N(t)=(r-\frac{a}{2}((\theta-\bar x(t))^2+\sigma^2(t))-cN(t))N(t),
+\end{equation}
+solving for equilibrium total abundance $\hat N$ amounts to setting $\frac{d}{dt}N(t)=0$ and solving for $N(t)$.
+Ignoring the equilibrium $N(t)=0$, this reduces to solving $\bar m(t)=0$ for $N(t)$, which returns
+\begin{equation}
+\hat N=\frac{1}{c}\left(r-\frac{a}{2}\left((\theta-\hat{\bar x})^2-\hat\sigma^2\right)\right).
+\end{equation}
+Unfortunately, deriving ODE for $\bar x(t)$ and $\sigma^2(t)$ leads to expressions involving higher moments and finding ODE for these higher moments will lead to expressions involving yet even higher moments.
+To avoid this infinite regression, we find the equilibrium abundance density $\hat\nu(x)$ by solving $\frac{\partial}{\partial t}\nu(x,t)=0$ for $\nu(x,t)$.
+This implies the following ordinary differential equation
+\begin{equation}
+\frac{d^2}{dx^2}\hat\nu(x)=\left(\frac{2c}{\mu}\hat N+\frac{a}{\mu}(\theta-x)^2-\frac{2r}{\mu}\right)\hat\nu(x)
 \end{equation}
@@
 \begin{equation}
-\nu_\infty(x)=\frac{N_\infty}{\sqrt{2\pi}}\left(\frac{a}{\mu}\right)^{\frac{1}{4}}\exp\left(-\frac{(\theta-x)^2}{2\sqrt{\mu/a}}\right).
-\end{equation}
-From this expression we infer $\bar x_\infty=\theta$ and $\sigma^2_\infty=\sqrt{\mu/a}$.
-Hence $N_\infty=\frac{1}{c}\left(r-\frac{1}{2}\sqrt{a\mu}\right)$.
-This holds for any initial $\nu(\cdot,0)\in L^1(\mathbb{R})$ such that $\bar x(0)\in\mathbb{R}, \ \sigma^2(0),N(0)\in(0,+\infty)$.
+\hat\nu(x)=\frac{\hat N}{\sqrt{2\pi}}\left(\frac{a}{\mu}\right)^{\frac{1}{4}}\exp\left(-\sqrt\frac{a}{\mu}\frac{(\theta-x)^2}{2}\right).
+\end{equation}
+From this expression we infer $\hat{\bar x}=\theta$ and $\hat\sigma^2=\sqrt{\frac{\mu}{a}}$.
+Hence $\hat N=\frac{1}{c}\left(r-\frac{1}{2}\sqrt{a\mu}\right)$.
+To show that $N_\infty=\hat N$, $\bar x_\infty=\hat{\bar x}$ and $\sigma^2_\infty=\hat\sigma^2$, we linearize ODE for $N(t)$, $\bar x(t)$ and $\sigma^2(t)$ using the equilibrium $\hat\nu(x)$.
+In particular, since $\hat\nu(x)$ is Gaussian, we do not run into the same issue with higher moments as above.
+Assuming the equilibrium $\hat\nu(x)$ does not change the ODE for $N(t)$, but the ODE for $\bar x(t)$ and $\sigma^2(t)$ can now be expressed as
+\begin{equation}
+\frac{d}{dt}\bar x(t)=\sigma^2(t)\left(\frac{\partial\bar m(t)}{\partial\bar x(t)}-\overline{\frac{\partial m(t)}{\partial\bar x(t)}}\right)=a\sigma^2(t)(\theta-\bar x(t)),
+\end{equation}
+\begin{equation}
+\frac{d}{dt}\sigma^2(t)=2\sigma^4(t)\left(\frac{\partial\bar m(t)}{\partial\sigma^2(t)}-\overline{\frac{\partial m(t)}{\partial\sigma^2(t)}}\right)+\mu=\mu-a\sigma^4(t),
+\end{equation}
+where $\hat p(x)=\hat\nu(x)/\hat N$. These expressions confirm our findings that $\hat{\bar x}=\theta$ and $\hat\sigma^2=\sqrt{\frac{\mu}{a}}$. Furthermore, calculating
+\begin{equation}
+\frac{\partial}{\partial\sigma^2(t)}\frac{d}{dt}\sigma^2(t)=-2a\sigma^2(t)
+\end{equation}
+and evaluating at $\sigma^2(t)=\hat\sigma^2$ demonstrates the equilibrium phenotypic variance is stable when $a,\mu>0$. Hence, calculating
+\begin{equation}
+\frac{\partial}{\partial\bar x(t)}\frac{d}{dt}\bar x(t)=-a\sigma^2(t)
+\end{equation}
+and evaluating at $\sigma^2(t)=\hat\sigma^2$ and $\bar x(t)=\hat{\bar x}$ demonstrates the equilibrium phenotypic mean is stable when $a,\mu>0$. Finally, calculating
+\begin{equation}
+\frac{\partial}{\partial N(t)}\frac{d}{dt}N(t)=r-\frac{a}{2}((\theta-\bar x(t))^2+\sigma^2(t))-2cN(t)
+\end{equation}
+and evaluating at $\sigma^2(t)=\hat\sigma^2$, $\bar x(t)=\hat{\bar x}$ and $N(t)=\hat N$ demonstrates the equilibrium total abundance is stable when $a,c,\mu>0$, and $r>\frac{1}{2}\sqrt{a\mu}$.
 

### 2020-04-11 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
the models name is SAGA

@@
 However, for populations with a continuum of types, such as a quantitative trait, this becomes a vexing mathematical challenge.
-Here we close this gap by combining the calculus of white noise with results on rescaled limits of branching Brownian motion processes (BBM) and stochastic partial differential equations (SPDE).
+Here we close this gap by combining the calculus of white noise with results on rescaled limits of branching Brownian motion processes (BBM) and stochastic partial differential equations (SPDE), which are stochastic analogs of deterministic partial differential equations (PDE).
 Our goal has two components: 1) Establish a novel synthetic approach to theoretical evolutionary ecology that provides a formal connection between demographic stochasticity and random genetic drift in the context of quantitative traits.
@@
 To provide motivation for the stochastic equations developed later and background for our model of coevolution, we begin with \S\ref{deterministic} by briefly summarizing derivations of deterministic dynamics of biological populations.
-Starting with a partial differential equation (PDE), we arrive at a general set of ordinary differential equations modelling the dynamics of abundance, trait mean and trait variance.
-From this we observe that replacing the PDE with a SPDE provides a path to derive SDE describing the evolutionary response to demographic stochasticity.
+Starting with a PDE (which we refer to as the Deterministic Asexual Gaussian allelic model with Abundance dynamics, abbreviated DAGA), we arrive at a general set of ordinary differential equations modelling the dynamics of abundance, trait mean and trait variance.
+We discuss the relationship between DAGA and previous models of phenotypic evolution that have been employed in theoretic quantitative genetics.
+Observing that generalizing the PDE to a SPDE, we find a path to derive SDE describing the evolutionary response to demographic stochasticity.
 We accomplish this in \S\ref{wnc} by introducing a set of mathematical tools based on the calculus of white noise and discuss how a diffusion limit of a spatially structured branching process leads to the natural SPDE appropriate for our study.
+We refer to this SPDE as the Stochastic Asexual Gaussian allelic model with Abundance dynamics (abbreviated SAGA).
 The diffusion limit in turn provides a rigorous method for constructing fitness functions used in models of evolutionary ecology.
 We employ these tools to derive a system of SDE generalizing our deterministic results to account for demographic stochasticity.
+We find that the effect of random genetic drift on the evolution of mean trait and phenotypic variance characterizes the process of random genetic drift in continuous time.
 However, although biologically insightful, these equations remain difficult to analyze and implement numerically.
@@
 In SM \S\ref{diffuse} we apply a classical theorem on rescaled limits of BBM that allow for ecological interactions to provide a rigorous derivation.
-To gain biological insight, in \S\ref{dynamics} we numerically integrate our model of coevolution for $S=100$ species, tracking the dynamics of traits and abundances, under scenarios of weak and strong competition.
+To gain biological insight, in \S\ref{dynamics} we numerically integrate our model of coevolution for $S=100$ species, tracking the dynamics of traits and abundances under scenarios of weak and strong competition.
 We include an account of the natural history of the simulated system and discuss the significance of demographic stochasticity for structuring ecological communities.
 In \S\ref{ecoevo} we provide expressions for selection gradients and competition coefficients implied by our model and use these expressions to investigate the relationship between the degree of competition and coevolution.
-Finally, \S\ref{conclusion} concludes with a summary of accomplishments, a few remarks on the limits of this approach and future directions to incorporate more explicitly the genetic architecture of populations, feedbacks with ecosystem processes and the macroevolutionary history of interacting lineages.
+Finally, \S\ref{conclusion} concludes with a summary of accomplishments, a few remarks on the limits of this approach and future directions to incorporate more explicitly the effects of small populations, the genetic architecture of quantitative traits, feedbacks with ecosystem processes and the macroevolutionary history of interacting lineages.
 
@@
 We return to these concepts in \S\ref{inheritance} where we review a simple model of inheritance that leads us to an expression for the stochastic evolution of additive genetic variance.
-To clarify the distinction between this model and previous models of phenotypic evolution we refer to it as the Deterministic Asexual Gaussian allelic model with Abundance dynamics (abbreviated DAGA).
+To clarify the distinction between this model and previous models of phenotypic evolution we refer to the PDE (\ref{eq1}) from which (\ref{p_PDE}) was derived as the Deterministic Asexual Gaussian allelic model with Abundance dynamics (abbreviated DAGA).
 Later, we will extend this model to include the effects of demographic stochasticity, which we refer to as the Stochastic Asexual Gaussian allelic model with Abundance dynamics (abbreviated SAGA).
@@
 In SM \S\ref{SDE_DERIV}, we extend these results to include the effects of demographic stochasticity.
+The key realization in this direction is that a rigorous derivation of a SPDE that naturally generalizes (\ref{eq1}) in the context of biological populations that accounts for ecological interactions has been provided by Li [-@zeng1998absolute].
+To understand how this SPDE was derived, we provide an informal discussion of measure-valued branching processes and their diffusion limits.
 Similar to the approach taken by Champagnat, Ferriére and Méléard [-@Champagnat2006], we begin with a BBM that models populations as discrete sets of reproducing individuals whose vital rates depend on their trait value as well as the state of the entire population.
@@
 Although this provides an impressive list of accomplishments, the Dawson-Watanabe superprocess falls short of our needs.
-In particular this process assumes individuals do not interact and thus precludes its ability to model ecological interactions.
+In particular this process assumes individuals do not interact and thus precludes its ability to model nonlocal effects on the fitness of individuals, such as those produced via ecological interactions.
 However, this concern has been addressed, leading to constuctions of superprocesses that account for interactions among individuals.
@@
 
-The existence of diffusion limits for a class of measure-valued branching processes involving interactions among individuals has been treated by Méléard and Roelly [-@meleard1992interacting; -@Mlard1993].
+In the above subsection we reviewed the diffusion limit of an especially tractable measure-valued branching process.
+However, we found the simplicity of this process restricts us from modelling nonlocal effects on the fitness of individuals.
+Here, we discuss superprocesses that account for such effects.
+The existence of diffusion limits for a class of measure-valued branching processes involving nonlocal effects has been treated by Méléard and Roelly [-@meleard1992interacting; -@Mlard1993].
 The interactions can manifest as dependencies of the spatial movement or reproductive law of individuals on their position $x$ and the state of the whole population described by $X_t$.
@@
 Equation (\ref{rho_SPDE}) is a stochastic generalization of the deterministic PDE (\ref{eq1}) from \S\ref{deterministic}, but describes the evolution of the distribution of breeding values instead of expressed characters.
-However, whether modelling expressed characters or breeding values, we refer to SPDE of the form (\ref{SPDE}) as Stochastic Asexual Gaussian allelic models with Abundance dynamics (abbreviated SAGA).
+However, whether modelling expressed characters or breeding values, we refer to SPDE of the form (\ref{rho_SPDE}) as Stochastic Asexual Gaussian allelic models with Abundance dynamics (abbreviated SAGA).
 By replacing $\psi(x,g)$ with different probability densities produces different developmental models.
@@
 To leave room for the naming of other models where $\psi(x,g)$ is Gaussian, we refer to the model of development we have chosen as the _centered_ Gaussian developmental model.
-In \S\ref{wnc} we review the origins of this equation and provide some theory to help make sense of it, particularly the term $\dot W$.
-
-We assume $\rho(g,t)$ is Gaussian which implies its mode coincides with $\bar x$.
+In \S\ref{wnc} we review the origins of equation (\ref{rho_SPDE}) and provide some theory to help make sense of it, particularly the term $\dot W$.
+
+Assuming $\rho(g,t)$ is Gaussian implies its mode coincides with $\bar x$.
 Furthermore, since $\sigma^2=G+\eta$, we can use equation (\ref{m_rel}) and the chain rule from calculus (see SM \S\ref{fit2fit}) to find
@@
 
-We have introduced an approach to derive models of evolutionary ecology using the calculus of white noise, developed SAGA, a novel model of phenotypic evolution that accounts for demographic stochasticity, demonstrated our approach by deriving a model of diffuse coevolution from classical competition theory, and used this model to investigate the relationship between selection gradients and competition coefficients.
-This work demonstrates that connecting contemporary theoretical approaches to evolutionary ecology with some fundamental results in the theory of measure-valued Markov processes allows for the development of a rigorous, yet flexible approach to synthesizing the dynamics of abundance and distributions of quantitative characters.
+We have introduced an approach to derive models of evolutionary ecology using the calculus of white noise and developed SAGA, a novel model of phenotypic evolution that accounts for demographic stochasticity.
+From SAGA we derived SDE that track the evolution of abundance, mean trait and additive genetic variance.
+Observing the expressions of these SDE we find the effects of demographic stochasticity on the evolution of mean trait and additive genetic variance characterize the effects of random genetic drift.
+By considering the statistics of random samples, this has previously been in models of fixed population size in discrete time [@Lande1976].
+However, we have shown random genetic drift is a result of demographic stochasticity for populations with fluctuating abundances in continuous time.
+To demonstrate the relevance of our approach to studies of evolutionary ecology, we comined our SDE with classical competition theory to derive a model of diffuse coevolution and used this model to investigate the relationship between selection gradients and competition coefficients.
+This work demonstrates that connecting contemporary theoretical approaches to evolutionary ecology with some fundamental results in the theory of measure-valued Markov processes allows for the development of a rigorous, yet flexible approach to synthesizing the dynamics of abundance and distributions of quantitative characters, in agreement with Champagnat et.
+al.
+[-@Champagnat2006].
 Although this approach enjoys these important merits, there remains biological details and their associated technical challenges that need to be confronted for gaining a more thorough and accurate understanding of ecological communities.
 We touch on just four of them here and provide suggestions for future research to approach these challenges.
+For the first three, we recommend the use of individual-based models such as BBM to confront these challenges.
+To summarize why, we state here that analytical results become difficult to derive for small populations and complicated models of mutation and selection.
 
@@
 Relating this concern to our comment in \S\ref{ecoevo} on using our model of diffuse coevolution to extend modern coexistence theory [sensu @Chesson2000] in a way that accounts for phenotypic evolution and abundance dynamics, it seems necessary to determine whether or not results pertaining to the coexistence of competing populations based on deterministic models that ignore phenotypic evolution hold using stochastic individual-based simulations.
-To stay within the framework outlined in this paper and thereby produce comparable results, these simulations should be extensions of BBM that account for the effects of competition.
+To stay within the framework outlined in this paper and thereby produce comparable results, these simulations should be extensions of branching processes such as BBM that account for the effects of competition.
+An extension of BBM that accounts for competition and stabilizing selection was used to produce Figure \ref{rescaled} in \S\ref{stochastic}.
+Code for this simulation can be found in the associated github repository or by contacting the lead author of this paper.
 
@@
 However, work to investigate the role of complex genetic architectures in mediating feedbacks between the dynamics of population abundances and the distributions of traits mediating ecological interactions has apparently only just begun [@Patel2019].
+However, since analytical results are difficult to obtain for general models of mutation, development and selection  [but see @9780471986539], we again suggest the use of individual-based simulations.
+For example, by assuming the mutational effects in the BBM model is Gaussian with variance an order of magnitude greater than the equilibrium phenotypic (or genetic) variance, we obtain the so-called _House-of-cards_ model [@Turelli1984].
+However, perhaps considering heavy-tailed distributions, such as the Laplace-distributions considered in Park et.
+al.
+[-@Park2017], will provide more realistic models of mutational effects.
+This has important implications in conservation genetics where small populations may evolve primarily in response to mutation and random genetic drift [@Lande1995].
 
@@
 Theoretical pursuits in this direction may profit from interfacing the framework we have outlined here with population-ecosystem models such as that developed by Fridley [-@Fridley2017].
+However, these population-ecosystem models will likely yield complicated fitness functions that are difficult to analyze using rigorous mathematical techniques.
+Hence, we again recommend the use of individual-based simulation approaches that generalize BBM to understand when analytical approximations hold.
 

### 2020-04-11 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
the models name is SAGA
expanding on coexistence

@@
 However, for these cases $f\notin C^2_b(\mathbb{R})$ since $f$ will not be bounded.
-However, if we can show that $\int_\mathbb{R}(|x|+x^2+x^4)\nu(x,t)dx<+\infty$ for all $t>0$ given this condition is satisfied by $\nu(x,0)$, then we can apply the mild solution of (\ref{SPDE}) to derive SDE for $\bar x(t)$ and $\sigma^2(t)$.
+But if we can show $\int_\mathbb{R}(|x|+x^2+x^4)\nu(x,t)dx<+\infty$ for all $t>0$ given this condition is satisfied by $\nu(x,0)$, then we can apply the mild solution of (\ref{SPDE}) to derive SDE for $\bar x(t)$ and $\sigma^2(t)$.
 To illustrate, let us suppose this is the case.
@@
 Since $\bar x(t)=\tilde x(t)/N(t)$ and $\sigma^2(t)=\tilde{\tilde\sigma}^2(t)/N(t)-\bar x^2(t)$, we can use It&ocirc;'s lemma to derive SDE for $\bar x(t)$ and $\sigma^2(t)$, which we perform in SM \S\ref{SDE_DERIV}.
-We do not make any general assertions about the existence or uniqueness of $\bar x(t)$ or $\sigma^2(t)$.
+We make no attempt in finding sufficient conditions for $\int_\mathbb{R}(|x|+x^2+x^4)\nu(x,t)dx<+\infty$ to hold and hence make no general assertions about the existence or uniqueness of $\bar x(t)$ or $\sigma^2(t)$.
+Regardless, we will later assume $\nu(x,t)$ can be approximated by a Gaussian curve in $x$ for all $t\geq0$.
+This assumption implies $\int_\mathbb{R}|x|^n\nu(x,t)dx<+\infty$ for all $n\in\mathbb{N}$ and for all $t\geq0$.
 
@@
 We note that conditions on the growth rate $m$ to guarantee existence and uniqueness of solutions to (\ref{xbar_gen}) and (\ref{sig2_gen}) have yet to be investigated.
-However, our results on the deterministic PDE suggest that $m(y,x)$ bounded above and differentiable in both arguments is sufficient.
+However, our results on the deterministic PDE suggest that $m(h,x)$ bounded above and differentiable in both arguments is sufficient.
 Dividing by $dt$ one can interpret equations (\ref{gen_eqns}) as if they are ordinary differential equations, but this not technically rigorous since Brownian motion is nowhere differentiable with respect to time.
@@
 For now we refer to $R_i$ as the _innate_ growth rate.
-Previous work has shown the importance of demographic stochasticity [@Schreiber2017] and evolutionary dynamics [@Schreiber2018] in determining ceoxistence of competing species.
+Previous work, to list just a few citations in this literature, has shown the importance of demographic stochasticity [@Turelli1980; @Schreiber2017] and evolutionary dynamics [@Case1986; @Schreiber2018] in determining coexistence of competing species.
 However, these studies consider evolution and demographic stochasticity separately.
+To our knowledge, there are no rigorous theoretical investigations of coexistence that account for evolutionary dynamics of quantitative characters and demographic stochasticity simultaneously.
+In fact, it appears that theoretical investigations combining these processes in a single species model to understand persistence have only just begun [@Gomulkiewicz2017].
 With the connection outlined in this section formally established, researchers may pursue a postmodern coexistence theory that naturally includes both the evolutionary dynamics of populations and the effects of demographic stochasticity in a simple synthetic framework.
@@
 In particular, the pathological behavior described for the diffusion limit does not occur for BBM as population size goes to zero.
-Relating this concern to our comment in \S\ref{ecoevo} on using our model of diffuse coevolution to extend modern coexistence theory [sensu @Chesson2000] in a way that accounts for phenotypic evolution and abundance dynamics, it seems necessary to determine whether or not results pertaining to the coexistence of competing populations based on deterministic models that ignore phenotypic evolution hold using stochastic individual-based simulations.
+This concern relates to our comment in \S\ref{ecoevo} on using our model of diffuse coevolution to extend modern coexistence theory [sensu @Chesson2000] in a way that accounts for phenotypic evolution and demographic stochasticity.
+Just as previous work has sought to understand the effects of demographic stochasticity and phenotypic evolution separately, it seems necessary to determine whether or not results pertaining to the coexistence of competing populations based on deterministic models that ignore phenotypic evolution hold using stochastic individual-based simulations.
 To stay within the framework outlined in this paper and thereby produce comparable results, these simulations should be extensions of branching processes such as BBM that account for the effects of competition.

### 2020-04-12 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
including champagnat's unifying approach

@@
 abstract: "We derive the dynamics of the distribution of a quantitative character and the abundance of a biological population from a stochastic partial differential equation driven by space-time white noise.
-In the process we develop a useful set of heuristics to operationalize the powerful, but abstract theory of white noise and measure-valued Markov processes.
-This approach allows us to compute the full implications of a stochastic process such as demographic stochasticity on phenotypic distributions and abundances of populations.
-We demonstrate the utility of our approach by deriving a model of diffuse coevolution mediated by exploitative competition for a continuum of resources.
+In the process we develop a useful set of heuristics to operationalize the powerful, but abstract theory of white noise and measure-valued stochastic processes.
+This approach allows us to compute the full implications of demographic stochasticity on phenotypic distributions and abundances of populations.
+We demonstrate the utility of our approach by deriving a quantitative genetic model of diffuse coevolution mediated by exploitative competition for a continuum of resources.
 Other than trait and abundance distributions, this model predicts interaction networks parameterized by rates of interactions, competition coefficients, and selection gradients.
@@
 Given this consideration, a natural scope for such an approach centers on quantifying the abundance dynamics of populations and the evolution of traits mediating their interactions as functions of relevant abiotic factors.
-Although taking into account abundance, phenotype and environment provides the basis for a partial understanding of the complex nature of biological communities, a deeper understanding must account for the effects of contemporary dispersal and the phylogeographic history of interacting lineages [@Kraft2007; @Hickerson2010; @Manceau2016; @markmcpeek2017] along with the genetic basis of ecologically relevant traits [@jeffreyconner2004; @FUSSMANN2007] and feedbacks between populations and the biogeochemical cycles they ultimately depend on [@michelloreau2010; @9781107648258].
+Although taking into account abundance, phenotype and environment provides the basis for a partial understanding of the complex nature of biological communities, a deeper understanding must account for the effects of dispersal and the phylogeographic history of interacting lineages [@Kraft2007; @Hickerson2010; @Manceau2016; @markmcpeek2017] along with the genetic basis of ecologically relevant traits [@jeffreyconner2004; @FUSSMANN2007] and feedbacks between populations and the biogeochemical cycles they ultimately depend on [@michelloreau2010; @9781107648258].
 It is therefore ideal that the development of any such mathematical approach anticipates extensions to account for these important factors shaping ecological communities, especially as empirical and conceptual work in these directions continues to grow [@AbdalaRoberts2014; @Klzsch2015; @Crutsinger2015; @Fitzpatrick2015; @Fitzpatrick2017; @Marx2017; @Rudman2017; @Skovmand2018; @VanNuland2019; @Harmon2019].
@@
 While correlation of fitness and genotype is the basis of evolution by natural selection, the mean fitness across all individuals in a population determines the growth, stasis or decline of abundance.
-In section \ref{deterministic} we review the mathematical formalization of this connection, which has been established in the contexts of population genetics [@9781932846126; @joanroughgarden1979], evolutionary game theory [@Hofbauer1998; @9780674023383; @Lion2018], quantitative genetics [@Lande1982; @Doebeli1996; @Lion2018] and a unifying framework for these three distinct approaches to evolutionary theory [@Champagnat2006] which is intimately related to the approach we take here.
-
-Reviewing these accomplishments reveals a beautiful synthesis of evolution and population ecology.
+In section \ref{deterministic} we review the mathematical formalization of this connection, which has been established in the contexts of population genetics [@9781932846126; @joanroughgarden1979], evolutionary game theory [@Hofbauer1998; @9780674023383; @Lion2018], quantitative genetics [@Lande1982; @Doebeli1996; @Lion2018] and a unifying framework for these three distinct approaches to evolutionary theory [@Champagnat2006] which includes our approach as a special case.
+We note here there are two primary differences between our aims and those of Champagnat, Ferrière, and Méléard [-@Champagnat2006].
+First, instead of developing a unifying approach to evolutionary theory, we focus on developing a stochastic synthesis of population dynamics and quantitative genetics including specific expressions for the dynamics of abundance, mean trait and additive genetic variance.
+Second, although Champagnat, Ferrière, and Méléard [-@Champagnat2006] provide rigorous proofs of results relevant to our approach, they do not communicate the nessecary tools we make use of for deriving the dynamics of abundance, mean trait and additive genetic variance.
+We aim to communicate these tools to a wide audience by providing an intuitive, yet accurate treatment of white noise and associated heuristics useful for deriving new stochastic models of evolutionary ecology.
+
+Reviewing the above accomplishments reveals a beautiful synthesis of evolution and population ecology.
 However, it also reveals a gap in theoretical approaches to incorporate the intrinsically random nature of populations.
@@
 Furthermore, since these fluctuations are themselves stochastic, it seems natural to derive expressions for the evolutionary response to demographic stochasticity and consider how the results relate to characterizations of random genetic drift.
-This can be done in continuous time for population genetic models without too much technical overhead, assuming a finite number of alleles.
-However, for populations with a continuum of types, such as a quantitative trait, this becomes a vexing mathematical challenge.
-Here we close this gap by combining the calculus of white noise with results on rescaled limits of branching Brownian motion processes (BBM) and stochastic partial differential equations (SPDE), which are stochastic analogs of deterministic partial differential equations (PDE).
+This can be done in continuous time for population genetic models without too much technical overhead, assuming a finite number of alleles [@Lande2009; @Parsons2010; @Gomulkiewicz2017].
+However, for populations with a continuum of types, such as a quantitative trait, finding a formal approach to derive the evolutionary response to demographic stochasticity has remained a vexing mathematical challenge.
+Previous work integrating stochastic processes with quantitative genetic models have profited from the simplifying assumption that phenotypic and genetic variances remain constant over time, yeilding analytical approximations for the joint dynamics of abundance and the mean of a quantitative trait [@Engen2013].
+These results not only ignore the evolution genetic variance, but they focus on environmental stochasticity and assume sufficiently large populations so that demographic stochasticity is negligible.
+Here we account for both demographic stochasticity and the evolution of additive genetic variance by combining the calculus of white noise with results on rescaled limits of branching Brownian motion processes (BBM) and stochastic partial differential equations (SPDE), which are stochastic analogs of deterministic partial differential equations (PDE).
 Our goal has two components: 1) Establish a novel synthetic approach to theoretical evolutionary ecology that provides a formal connection between demographic stochasticity and random genetic drift in the context of quantitative traits.
-To show that our approach can be used to develop useful biological insights we derive a model of coevolution in an ecological network and use it to investigate the relationship between competition coefficients and selection gradients.
 2) Communicate some useful properties of space-time white noise, BBM and SPDE to as wide of audience as possible.
@@
 
+Through accomplishing these goals this work yields five main results.
+Firstly, we develop a concise, but accurate approach for understanding SPDE and provide heuristics to empower mathematical evolutionary ecologists to derive novel models that relax key assumptions commonly made in combining models of evolutionary and abundance dynamics.
+Second, we formally derive general SDE that express the joint dynamics of abundance, mean trait and additive genetic variance for populations exhibiting a wide range of potential phenotypic distributions.
+Third, we simplify these SDE by approximating the abundance density of phenotypes with a Gaussian curve and integrate the results with a model of imperfect inheritance to find formula that generalize classical quantitative genetic expressions.
+This includes an expression for additive genetic variance analogous to the expression for the response in mean phenotype found by Lande [-@Lande1976].
+Fourth, following classical niche theory we provide a rigorous mechanistic derivation of fitness for populations experiencing resource competition and abiotic stabilizing selection using diffusion limits of measure-valued branching processes.
+This calculation provides a demonstration for performing such diffusion limits that can be applied in other contexts.
+Fifth, we explore implications of the resulting model of diffuse coevolution by investigating the relationship between selection gradients and competition coefficients.
+
+We now summarize the basic structure of the paper.
 To provide motivation for the stochastic equations developed later and background for our model of coevolution, we begin with \S\ref{deterministic} by briefly summarizing derivations of deterministic dynamics of biological populations.
 Starting with a PDE (which we refer to as the Deterministic Asexual Gaussian allelic model with Abundance dynamics, abbreviated DAGA), we arrive at a general set of ordinary differential equations modelling the dynamics of abundance, trait mean and trait variance.
-We discuss the relationship between DAGA and previous models of phenotypic evolution that have been employed in theoretic quantitative genetics.
-Observing that generalizing the PDE to a SPDE, we find a path to derive SDE describing the evolutionary response to demographic stochasticity.
-We accomplish this in \S\ref{wnc} by introducing a set of mathematical tools based on the calculus of white noise and discuss how a diffusion limit of a spatially structured branching process leads to the natural SPDE appropriate for our study.
+We discuss the relationship between DAGA and previous models of phenotypic evolution that have been employed in theoretical quantitative genetics.
+Noting that a natural stochastic generalization of DAGA has been previously derived, we find a path to derive SDE describing the evolutionary response to demographic stochasticity.
+We accomplish this in \S\ref{wnc} by discussing how a diffusion limit of a spatially structured branching process (i.e., a measure-valued branching process) leads to the natural SPDE appropriate for our study.
 We refer to this SPDE as the Stochastic Asexual Gaussian allelic model with Abundance dynamics (abbreviated SAGA).
 The diffusion limit in turn provides a rigorous method for constructing fitness functions used in models of evolutionary ecology.
-We employ these tools to derive a system of SDE generalizing our deterministic results to account for demographic stochasticity.
-We find that the effect of random genetic drift on the evolution of mean trait and phenotypic variance characterizes the process of random genetic drift in continuous time.
+However, before we discuss measure-valued braching processes and their diffusion limits, we find it necessary to review some fundamental results on the calculus of white noise and introducing a set of mathematical tools based on these results.
+In \S\ref{wnc} and SM \S\ref{SDE_DERIV} we employ these tools to derive a system of SDE generalizing our deterministic results to account for demographic stochasticity.
+We find that the effect of demographic stochasticity on the evolution of mean trait and phenotypic variance characterizes the process of random genetic drift in continuous time.
 However, although biologically insightful, these equations remain difficult to analyze and implement numerically.
@@
 We include an account of the natural history of the simulated system and discuss the significance of demographic stochasticity for structuring ecological communities.
-In \S\ref{ecoevo} we provide expressions for selection gradients and competition coefficients implied by our model and use these expressions to investigate the relationship between the degree of competition and coevolution.
+In \S\ref{ecoevo} we provide expressions for selection gradients and competition coefficients implied by our model and use these expressions to investigate the relationship between the degree of competition and strength of coevolution.
 Finally, \S\ref{conclusion} concludes with a summary of accomplishments, a few remarks on the limits of this approach and future directions to incorporate more explicitly the effects of small populations, the genetic architecture of quantitative traits, feedbacks with ecosystem processes and the macroevolutionary history of interacting lineages.
@@
 A popular alternative to quantitative genetics is the theory of adaptive dynamics.
-As demonstrated by Page and Nowak [-@PAGE2002], the canonical equation of adaptive dynamics can be derived from the replicator-mutator equation.
+As demonstrated by Page and Nowak [-@PAGE2002] and Champagnat, Ferrière, and Méléard [-@Champagnat2006], the canonical equation of adaptive dynamics can be derived from the replicator-mutator equation.
 Thus, one could start from the atomic roots of our approach and pursue a stochastic adaptive dynamic theory instead.
@@
 Our review begins by considering the dynamics of an asexually reproducing population in a homogeneous environment.
-Individuals are assumed to be haploid and carry one of $K$ alleles each with a different fitness expressed as growth rate.
+For simplicity, we first assume individuals are haploid and carry one of $K$ alleles each with a different fitness expressed as growth rate before introducing a model involving a quantitative trait.
 Under these assumptions, the derivation of the evolution of allele frequencies due to natural selection can be derived from expressions of exponential growth.
@@
 We prove these statements in SM \S\ref{equilib}.
+Note in particular the coincidence of equilibrial phenotypic variance predicted by this model and that predicted by the classic Gaussian allelic model of quantitative genetics [@Lande1975; @9780471986539; @Johnson2005].
 By replacing $\kappa(x,y)=1$ in expression (\ref{special_m}) with a particular Gaussian function, we will find in \S\ref{coev} and SM \S\ref{diffuse} this fitness function becomes equivalent to a fitness function derived from niche theory that we use to model coevolution driven by resource competition.
@@
 The primary difference being that our model utilizes diffusion instead of convolution with an arbitrary mutation kernel.
-Of course, our model of mutation can be derived as an approximation to Kimura's model, which, for sexually reproducing populations, has been referred to as the Gaussian allelic approximation in reference to the distribution of mutational effects at loci in a genome on the values of traits [@Lande1975; but see also @9780471986539], the infinitesimal model in reference to modelling continuous traits as being encoded by an infinite number of loci each having infinitesimal effect [@Barton2017] and the Gaussian descendants approximation in reference to offspring trait values being normally distributed around their parental values [@Turelli2017].
+Of course, our model of mutation can be derived as an approximation to Kimura's model, which, for sexually reproducing populations, has been referred to as the Gaussian allelic approximation in reference to the distribution of mutational effects at loci in a genome on the values of traits [@Lande1975; but see also @9780471986539; and @Johnson2005], the infinitesimal model in reference to modelling continuous traits as being encoded by an infinite number of loci each having infinitesimal effect [@Barton2017] and the Gaussian descendants approximation in reference to offspring trait values being normally distributed around their parental values [@Turelli2017].
 Alternatively, since diffusion is the continuous-time equivalent to convolution against a Gaussian kernel (SM \S\ref{diffconvequiv}), equation (\ref{p_PDE}) can also be seen as a special case of Kimura's continuum-of-alleles model.
@@
 In SM \S\ref{SDE_DERIV}, we extend these results to include the effects of demographic stochasticity.
-The key realization in this direction is that a rigorous derivation of a SPDE that naturally generalizes (\ref{eq1}) in the context of biological populations that accounts for ecological interactions has been provided by Li [-@zeng1998absolute].
-To understand how this SPDE was derived, we provide an informal discussion of measure-valued branching processes and their diffusion limits.
+The key realization in this direction is that a rigorous derivation of a SPDE that naturally generalizes DAGA and accounts for ecological interactions has been provided by Li [-@zeng1998absolute].
+To understand how this SPDE was derived, we provide an informal discussion of measure-valued branching processes (which treat populations as sets of discrete individuals) and their diffusion limits, the so-called superprocesses (which treat population size as a continuously varying number).
 Similar to the approach taken by Champagnat, Ferriére and Méléard [-@Champagnat2006], we begin with a BBM that models populations as discrete sets of reproducing individuals whose vital rates depend on their trait value as well as the state of the entire population.
@@
 The limiting processes of rescaled BBM have been referred to as measure-valued Markov processes [@dawson1993measure] or superprocesses [@alisonetheridge2000].
-Under the simplifying assumptions inherited from our treatment of deterministic dynamics and the additional assumption that the magnitude of demographic stochasticity is independent of trait values, we obtain as a special case a relatively simple expression for an SPDE that generalizes equation (\ref{eq1}).
+Under the simplifying assumptions inherited from our treatment of deterministic dynamics and the additional assumption that the magnitude of demographic stochasticity is independent of trait values, we obtain as a special case a relatively simple expression for an SPDE that generalizes DAGA.
 The simplicity of our special case allows us to use properties of space-time white noise processes to derive a set of SDE that generalize equations (\ref{N_det}), (\ref{detmean}) and (\ref{vardyn_det}) to include the effects of demographic stochasticity.
@@
 Hence, one can repeat the necessary proofs replacing the assumption that $m(h,x)$ is bounded with the assumption that $m(h,x)$ is merely bounded above to derive the same results.
+An alternative approach to establishing this result that allows for our relaxed assumption on the growth rate has carried out by Champagnat, Ferrière, and Méléard [-@Champagnat2006] using techniques developed by Evans and Perkins [-@Evans1994].
 
@@
 However, these studies consider evolution and demographic stochasticity separately.
-To our knowledge, there are no rigorous theoretical investigations of coexistence that account for evolutionary dynamics of quantitative characters and demographic stochasticity simultaneously.
-In fact, it appears that theoretical investigations combining these processes in a single species model to understand persistence have only just begun [@Gomulkiewicz2017].
-With the connection outlined in this section formally established, researchers may pursue a postmodern coexistence theory that naturally includes both the evolutionary dynamics of populations and the effects of demographic stochasticity in a simple synthetic framework.
+To our knowledge, there are no rigorous theoretical investigations of coexistence that account for evolutionary dynamics of quantitative characters and demographic stochasticity simultaneously and, in particular, no studies combining these processes that also include the evolution of additive genetic variance.
+In fact, only recently have rigorous theoretical investigations successfully combined these processes to understand persistence of a single population [@Gomulkiewicz2017].
+Even so, these authors focused on a single species model with a finite number of types (i.e., a population genetic model).
+With the connection outlined in this section formally established, researchers may pursue a postmodern coexistence theory that naturally includes both the evolutionary dynamics of quantitative characters and the effects of demographic stochasticity in a simple synthetic framework.
 
@@
 Figure \ref{net} displays the joint distribution of $\mathfrak{C}_{ij}=\Psi_{ij}\Psi_{ji}$, our metric of pairwise coevolution, and the product of competition coefficients $\alpha_{ij}\alpha_{ji}$ for two simulated communities, both with richness $S=1000$.
-The solid contour represents the case of strong selection ($c=5\times10^{-7}$) and the dashed contour represents the case of weak sensitivity to competition ($c=1\times10^{-8}$).
+The solid contour represents the case of relatively high sensitivity to competition ($c=5\times10^{-7}$) and the dashed contour represents the case of relatively weak sensitivity to competition ($c=1\times10^{-8}$).
 In both cases we see that $\mathfrak{C}_{ij}$ and $\alpha_{ij}\alpha_{ji}$ are, essentially, unrelated.
@@
 
-We have introduced an approach to derive models of evolutionary ecology using the calculus of white noise and developed SAGA, a novel model of phenotypic evolution that accounts for demographic stochasticity.
+We have introduced an approach to derive models of evolutionary ecology using the calculus of white noise and coined SAGA, a SPDE model of phenotypic evolution that accounts for demographic stochasticity.
 From SAGA we derived SDE that track the evolution of abundance, mean trait and additive genetic variance.
@@
 However, we have shown random genetic drift is a result of demographic stochasticity for populations with fluctuating abundances in continuous time.
-To demonstrate the relevance of our approach to studies of evolutionary ecology, we comined our SDE with classical competition theory to derive a model of diffuse coevolution and used this model to investigate the relationship between selection gradients and competition coefficients.
-This work demonstrates that connecting contemporary theoretical approaches to evolutionary ecology with some fundamental results in the theory of measure-valued Markov processes allows for the development of a rigorous, yet flexible approach to synthesizing the dynamics of abundance and distributions of quantitative characters, in agreement with Champagnat et.
-al.
-[-@Champagnat2006].
+To illustrate the relevance of our approach to studies of evolutionary ecology, we comined our SDE with classical competition theory to derive a model of diffuse coevolution and used this model to investigate the relationship between selection gradients and competition coefficients.
+This work demonstrates that connecting contemporary theoretical approaches to evolutionary ecology with some fundamental results in the theory of measure-valued stochastic processes allows for the development of a rigorous, yet flexible approach to synthesizing the dynamics of abundance and distributions of quantitative characters, in agreement with Champagnat, Ferrière, and Méléard [-@Champagnat2006].
 Although this approach enjoys these important merits, there remains biological details and their associated technical challenges that need to be confronted for gaining a more thorough and accurate understanding of ecological communities.
 We touch on just four of them here and provide suggestions for future research to approach these challenges.
-For the first three, we recommend the use of individual-based models such as BBM to confront these challenges.
+For the first three, we recommend the use of individual-based simulations of measure-valued processes to confront these challenges.
 To summarize why, we state here that analytical results become difficult to derive for small populations and complicated models of mutation and selection.
 
-__Limitations of diffusion limits and validating coexistence theory__
+__Limitations of diffusion limits and generalizing coexistence theory__
 
@@
 In particular, stochastic components of the SDE describing the evolution of $\bar x$ and $\sigma^2$ diverge towards infinite values as $N\to0$.
-Hence, studies of evolutionary rescue and colonization-extinction dynamics that incorporate phenotypic evolution should be pursued via a different approach.
-A natural alternative can be developed utilizing the underlying BBM that SPDE (\ref{SPDE}) is a diffusion limit of (see section \ref{stochastic}).
+Hence, studies involving extinction and phenotypic evolution should be pursued through a different approach.
+A natural alternative can be developed utilizing the underlying measure-valued branching process that SAGA is a diffusion limit of (see section \ref{stochastic}).
 This process explicitly tracks individuals throughout their life-history and captures reproduction as branching events.
-Hence, BBM processes model population size as a non-negative integer instead of a continuously varying number.
-In particular, the pathological behavior described for the diffusion limit does not occur for BBM as population size goes to zero.
+Hence, branching processes model population size as a non-negative integer instead of a continuously varying number.
+In particular, the pathological behavior described for the SDE of $\bar x$ and $\sigma^2$ does not occur for measure-valued branching processes.
 This concern relates to our comment in \S\ref{ecoevo} on using our model of diffuse coevolution to extend modern coexistence theory [sensu @Chesson2000] in a way that accounts for phenotypic evolution and demographic stochasticity.
-Just as previous work has sought to understand the effects of demographic stochasticity and phenotypic evolution separately, it seems necessary to determine whether or not results pertaining to the coexistence of competing populations based on deterministic models that ignore phenotypic evolution hold using stochastic individual-based simulations.
-To stay within the framework outlined in this paper and thereby produce comparable results, these simulations should be extensions of branching processes such as BBM that account for the effects of competition.
-An extension of BBM that accounts for competition and stabilizing selection was used to produce Figure \ref{rescaled} in \S\ref{stochastic}.
-Code for this simulation can be found in the associated github repository or by contacting the lead author of this paper.
+Just as previous work has sought to understand the implications of processes for coexistence separately, it seems necessary to determine whether or not results pertaining to the coexistence of competing populations hold using models that account for phenotypic evolution and demographic stochasticity simultaneously.
+Because such investigations consider both extinction and phenotypic evolution we argue the models used should be either simulations of, or whenever possible, analytical results based on measure-valued branching processes.
 
@@
 
-Our treatment of inheritance and our approach to model coevolution rest on the assumptions of normally distributed breeding values and expressed phenotypes along with asexual reproduction.
-However, real traits are not encoded by an infinite number of loci each contributing an additive infinitesimal effect (as required by the infinitesimal model), mutations are not inherited as normally distributed deviations from parental breeding values (as required by the Gaussian descendents model) and many populations of interest exhibit non-random sexual reproduction.
+Our treatment of inheritance and our approach to model coevolution rests on the assumptions of normally distributed breeding values and expressed phenotypes along with asexual reproduction.
+However, real traits are not encoded by an infinite number of loci each contributing an additive infinitesimal effect (as required by the infinitesimal model), mutations are not inherited as normally distributed deviations from parental breeding values (as required by the Gaussian allelic model) and many populations of interest exhibit non-random sexual reproduction.
 Departures from this model of genetic architecture can produce more complex distributions of breeding values and expressed traits.
@@
 However, work to investigate the role of complex genetic architectures in mediating feedbacks between the dynamics of population abundances and the distributions of traits mediating ecological interactions has apparently only just begun [@Patel2019].
-However, since analytical results are difficult to obtain for general models of mutation, development and selection  [but see @9780471986539], we again suggest the use of individual-based simulations.
+Since analytical results are difficult to obtain for general models of mutation, development and selection [c.f., @9780471986539], we again suggest the use of measure-valued branching processes and their individual-based simulations.
 For example, by assuming the mutational effects in the BBM model is Gaussian with variance an order of magnitude greater than the equilibrium phenotypic (or genetic) variance, we obtain the so-called _House-of-cards_ model [@Turelli1984].
@@
 al.
-[-@Park2017], will provide more realistic models of mutational effects.
-This has important implications in conservation genetics where small populations may evolve primarily in response to mutation and random genetic drift [@Lande1995].
+[-@Park2017], will provide more realistic, yet less analytically tractable models of mutational effects.
+Such models have important implications in conservation genetics where small populations may evolve primarily in response to mutation and random genetic drift [@Lande1995].
 
@@
 However, these population-ecosystem models will likely yield complicated fitness functions that are difficult to analyze using rigorous mathematical techniques.
-Hence, we again recommend the use of individual-based simulation approaches that generalize BBM to understand when analytical approximations hold.
+Hence, we again recommend the use of simulations of measure-valued branching processes that generalize BBM to understand when analytical approximations hold.
 

### 2020-04-12 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
classical solutions

@@
 We denote by $K$ an operator that takes functions such as $\nu(x,t)$ as its argument and returns other functions that may still depend on the spatial variable $x$ and the temporal variable $t$.
-For our application we will be focused on operators $K$ that can be written as $(K\nu)(x,t)=\int_{-\infty}^{+\infty}\kappa(x,y)\nu(y,t)dy$ for some non-negative and bounded bivariate function $\kappa$.
+For our application we will be focused on operators $K$ that can be written as $(K\nu)(x,t)=\int_{-\infty}^{+\infty}\kappa(x-y)\nu(y,t)dy$ for some non-negative and bounded function $\kappa$.
 We set $h=(K\nu)(x,t)$ to account for nonlocal effects of the abundance density $\nu(x,t)$ on the fitness of individuals with trait value $x$ [@Volpert2014].
@@
 \end{equation}
-where $r\in\mathbb{R}$ is the intrinsic growth rate in the absence of abiotic stabilizing selection, $a>0$ is the strength of abiotic stabilizing selection around the phenotypic optimum $\theta\in\mathbb{R}$, $c>0$ captures the sensitivity of fitness to intraspecific competition and we have set $\kappa(x,y)=1$ so that competitive interactions cause the same reduction in fitness regardless of trait value.
+where $r\in\mathbb{R}$ is the intrinsic growth rate in the absence of abiotic stabilizing selection, $a>0$ is the strength of abiotic stabilizing selection around the phenotypic optimum $\theta\in\mathbb{R}$, $c>0$ captures the sensitivity of fitness to intraspecific competition and we have set $\kappa(x-y)=1$ so that competitive interactions cause the same reduction in fitness regardless of trait value.
 This exemplary fitness function has a few convenient properties that we breifly take note of.
@@
 Note in particular the coincidence of equilibrial phenotypic variance predicted by this model and that predicted by the classic Gaussian allelic model of quantitative genetics [@Lande1975; @9780471986539; @Johnson2005].
-By replacing $\kappa(x,y)=1$ in expression (\ref{special_m}) with a particular Gaussian function, we will find in \S\ref{coev} and SM \S\ref{diffuse} this fitness function becomes equivalent to a fitness function derived from niche theory that we use to model coevolution driven by resource competition.
+By replacing $\kappa(x-y)=1$ in expression (\ref{special_m}) with a particular Gaussian function, we will find in \S\ref{coev} and SM \S\ref{diffuse} this fitness function becomes equivalent to a fitness function derived from niche theory that we use to model coevolution driven by resource competition.
 
@@
 \begin{equation}
-\mathcal O_{ij}(x_i,x_j)=\int_{-\infty}^{+\infty}u_i(\zeta,x_i)u_j(\zeta,x_j)d\zeta=\frac{U_iU_j}{\sqrt{2\pi(w_i+w_j)}}\exp\left(-\frac{(x_i-x_j)^2}{2(w_i+w_j)}\right).
+\mathcal O_{ij}(x_i-x_j)=\int_{-\infty}^{+\infty}u_i(\zeta,x_i)u_j(\zeta,x_j)d\zeta=\frac{U_iU_j}{\sqrt{2\pi(w_i+w_j)}}\exp\left(-\frac{(x_i-x_j)^2}{2(w_i+w_j)}\right).
 \end{equation}
@@
 
-To map the degree of niche overlap to fitness, we assume competition between individuals with niche locations $x_i$ and $x_j$ additively decreases the Malthusian fitness for the individual in species $i$ by $c_i\mathcal{O}_{ij}(x_i,x_j)$ for some $c_i\geq0$.
+To map the degree of niche overlap to fitness, we assume competition between individuals with niche locations $x_i$ and $x_j$ additively decreases the Malthusian fitness for the individual in species $i$ by $c_i\mathcal{O}_{ij}(x_i-x_j)$ for some $c_i\geq0$.
 We refer to $c_i$ as the sensitivity to competition for species $i$.
-The term $c_i\mathcal{O}_{ij}(x_i,x_j)$ coincides with a special case of a term used to capture competition in Dawson's geostochastic logistic model, an SPDE model developed to study the combined effects of demographic stochasticity, spatial dispersion and locally finite carrying capacity [@Dawson1978].
-In relation to the example fitness function discussed in \S\ref{deterministic}, consider $\kappa(x_i,x_j)=\mathcal{O}_{ij}(x_i,x_j)$.
+The term $c_i\mathcal{O}_{ij}(x_i-x_j)$ coincides with a special case of a term used to capture competition in Dawson's geostochastic logistic model, an SPDE model developed to study the combined effects of demographic stochasticity, spatial dispersion and locally finite carrying capacity [@Dawson1978].
+In relation to the example fitness function discussed in \S\ref{deterministic}, consider $\kappa(x_i-x_j)=\mathcal{O}_{ij}(x_i-x_j)$.
 In SM \S\ref{diffuse} we combine our treatment of resource competition with equations (\ref{N}), (\ref{xbarfinal}) and (\ref{Gfinal}) to find
@@
 \end{equation}
-where $m(\nu,x)\leq r\in\mathbb{R}$ is shorthand for $m((K\nu)(x,t),x)$ and $(K\nu)(x,t)=\int_\mathbb{R}\kappa(x,y)\nu(y,t)dy$ for some non-negative, bounded $\kappa$.
+where $m(\nu,x)\leq r\in\mathbb{R}$ is shorthand for $m((K\nu)(x,t),x)$ and $(K\nu)(x,t)=\int_\mathbb{R}\kappa(x-y)\nu(y,t)dy$ for some non-negative, bounded $\kappa$.
 Denoting $C(\mathbb{R})$ the set of continuous functions on $\mathbb{R}$ and $L^1(\mathbb{R})$ the Lebesgue space of integrable functions on $\mathbb{R}$ (ie, the set of functions $f:\mathbb{R}\to\mathbb{R}$ that satisfy $\|f\|_1=\int_\mathbb{R}|f(x)|dx<+\infty$ under the equivalence relation $f=g$ iff $\|f-g\|_1=0$), we assume a continuous and integrable initial condition $\nu(\cdot,0)=u\in L^1(\mathbb{R})\cap C(\mathbb{R})$.
@@
 \begin{equation}
-(Fv)(x)=m\left(\int_\mathbb{R}\kappa(x,y)v(y)dy,x\right)v(x), \ \ \forall \ \ v\in L^1(\mathbb{R}).
+(Fv)(x)=m\left(\int_\mathbb{R}\kappa(x-y)v(y)dy,x\right)v(x), \ \ \forall \ \ v\in L^1(\mathbb{R}).
 \end{equation}
@@
 Repeating this task a finite number of times shows $\nu$ is the mild solution of PDE (\ref{PDE_supp}) with initial condition $\nu(0)=u$ for arbitrarily large $T>0$.
-Applying lemma 3.3.2 of Henry [-@Henry1981] we have that the mild solution $\nu$ is in fact a strong solution to the Cauchy problem defined by PDE (\ref{PDE_supp}) with initial condition $\nu(0)=u$ for $t\in[0,T]$.
-In particular, following example 3.6 of Henry [-@Henry1981], this implies that $\nu(t)\in D(\Delta)$ for each $t\in(0,T)$ where $D(\Delta)=W^{2,2}(\mathbb{R})$ is the Sobolov space of square integrable functions with weak derivatives at least of order $2$.
-Then, according to a Sobolov embedding theorem [@Henry1981], we have $D(\Delta)\subset C^2(\mathbb{R})$ and hence solutions to the Cauchy problem (\ref{PDE_supp}) are twice continuously differentiable in their spatial parameter.
+
+We shift our focus to showing mild solutions $\nu$ to PDE (\ref{PDE_supp}) are twice differentiable in $x$ and hence are classical solutions.
+Applying lemma 3.3.2 of Henry [-@Henry1981] we have that the mild solution $\nu$ is in fact a strong (but necessarily classical) solution to the Cauchy problem defined by PDE (\ref{PDE_supp}) with initial condition $\nu(0)=u$ for $t\in[0,T]$.
+In particular, this implies that $\nu(t)\in D(\Delta)$ for each $t\in(0,T)$ where $D(\Delta)=W^{2,1}(\mathbb{R})$ is the Sobolov space of integrable functions with weak derivatives at least of order $2$.
+Since $\frac{1}{p}\geq\frac{1}{p+1}\geq\frac{1}{p}-2$ for all $p\geq1$, the Sobolov embedding theorem [@Henry1981] implies $D(\Delta)\subset L^p(\mathbb{R})$, the Lebesgue space of $p$-integrable functions on $\mathbb{R}$, for all $p\geq1$.
+Hence $D(\Delta)\subset W^{2,p}(\mathbb{R})$, the Sobolov space of $p$-integrable functions with weak derivatives at least of order $2$, for all $p\geq1$.
+
+If $v\in C^{2+\delta}(\mathbb{R})$ for some $\delta>0$, then 
+\begin{equation}
+(Kv)(\cdot)=\int_\mathbb{R}\kappa(\cdot-y)v(y)dy=\int_\mathbb{R}\kappa(y)v(\cdot-y)dy\in C^{2+\delta}(\mathbb{R}),
+\end{equation}
+and, since we assume $m(h,x)$ is continuous in both arguments, $Fv=m(Kv,\cdot)v\in C^\delta$.
+
+Why is $F\nu\in C^\delta(\mathbb{R})$ for some $\delta>0$?
+Why does this imply $\nu(\cdot,t)\in C^{2+\delta}(\mathbb{R})$ for all $t\in(0,T)$?
+This argument holds for all $T>0$.
+Hence, solutions to the Cauchy problem (\ref{PDE_supp}) are twice continuously differentiable in their spatial parameter for all $t>0$.
 
@@
 \begin{equation}
-\mathscr B_{ij}(X)=C_i^{\sum_{l\neq j}\mathcal{O}_{ii}(x_{ij},x_{il})+\sum_{k\neq i}\sum_{l=1}^{n_k}\mathcal{O}_{ik}(x_{ij},x_{kl})},
+\mathscr B_{ij}(X)=C_i^{\sum_{l\neq j}\mathcal{O}_{ii}(x_{ij}-x_{il})+\sum_{k\neq i}\sum_{l=1}^{n_k}\mathcal{O}_{ik}(x_{ij}-x_{kl})},
 \end{equation}
@@
 \begin{equation}
-\mathscr W_{1j}^{(\pmb n)}(X)=\mathscr{A}_1(x_{1j})^{1/n_1}\exp\left(\frac{\ln C_1}{n_1^2}\sum_{l\neq j}\mathcal{O}_{11}(x_{1j},x_{1l})+\frac{\ln C_1}{n_1}\sum_{k\neq i}\frac{1}{n_k}\sum_{l=1}^{n_k}\mathcal{O}_{1k}(x_{1j},x_{kl})\right).
+\mathscr W_{1j}^{(\pmb n)}(X)=\mathscr{A}_1(x_{1j})^{1/n_1}\exp\left(\frac{\ln C_1}{n_1^2}\sum_{l\neq j}\mathcal{O}_{11}(x_{1j}-x_{1l})+\frac{\ln C_1}{n_1}\sum_{k\neq i}\frac{1}{n_k}\sum_{l=1}^{n_k}\mathcal{O}_{1k}(x_{1j}-x_{kl})\right).
 \end{equation}
@@
 \begin{equation}
-\mathscr W_{1j}^{(\pmb n)}(X)\approx\mathscr{A}_1(x_{1j})^{1/n_1}\left(1+\frac{\ln C_1}{n_1^2}\sum_{l\neq j}\mathcal{O}_{11}(x_{1j},x_{1l})+\frac{\ln C_1}{n_1}\sum_{k\neq 1}\frac{1}{n_k}\sum_{l=1}^{n_k}\mathcal{O}_{1k}(x_{1j},x_{kl})\right).
+\mathscr W_{1j}^{(\pmb n)}(X)\approx\mathscr{A}_1(x_{1j})^{1/n_1}\left(1+\frac{\ln C_1}{n_1^2}\sum_{l\neq j}\mathcal{O}_{11}(x_{1j}-x_{1l})+\frac{\ln C_1}{n_1}\sum_{k\neq 1}\frac{1}{n_k}\sum_{l=1}^{n_k}\mathcal{O}_{1k}(x_{1j}-x_{kl})\right).
 \end{equation}
@@
 \begin{equation}
-\lim_{n_1\to\infty}n_1\left(\mathscr W_{1j}^{(\pmb n)}(X)-1\right)=\ln\mathscr{A}_1(x_{1j})+\left(\int_\mathbb{R}\mathcal{O}_{11}(x_{1j},y)\nu_1(y,t)dy+\sum_{k\neq 1}\frac{1}{n_k}\sum_{l=1}^{n_k}\mathcal{O}_{1k}(x_{1j},x_{kl})\right)\ln C_1.
+\lim_{n_1\to\infty}n_1\left(\mathscr W_{1j}^{(\pmb n)}(X)-1\right)=\ln\mathscr{A}_1(x_{1j})+\left(\int_\mathbb{R}\mathcal{O}_{11}(x_{1j}-y)\nu_1(y,t)dy+\sum_{k\neq 1}\frac{1}{n_k}\sum_{l=1}^{n_k}\mathcal{O}_{1k}(x_{1j}-x_{kl})\right)\ln C_1.
 \end{equation}
@@
 \begin{equation}
-m_1(\pmb\nu,x_{1j}):=\lim_{\pmb n\to\pmb\infty}n_1\left(\mathscr W_{1j}^{(\pmb n)}(X)-1\right)=\ln\mathscr{A}_1(x)+\left(\sum_{k=1}^S\int_\mathbb{R}\mathcal{O}_{1k}(x_{1j},y)\nu_k(y,t)dy\right)\ln C_1.
+m_1(\pmb\nu,x_{1j}):=\lim_{\pmb n\to\pmb\infty}n_1\left(\mathscr W_{1j}^{(\pmb n)}(X)-1\right)=\ln\mathscr{A}_1(x)+\left(\sum_{k=1}^S\int_\mathbb{R}\mathcal{O}_{1k}(x_{1j}-y)\nu_k(y,t)dy\right)\ln C_1.
 \end{equation}
@@
 \begin{equation}
-\bar{\mathcal{O}}_{ij}(x,t)=\frac{\int_\mathbb{R}\mathcal{O}_{ij}(x,y)\nu_j(y,t)dy}{\int_\mathbb{R}\nu_j(y,t)dy}.
+\bar{\mathcal{O}}_{ij}(x,t)=\frac{\int_\mathbb{R}\mathcal{O}_{ij}(x-y)\nu_j(y,t)dy}{\int_\mathbb{R}\nu_j(y,t)dy}.
 \end{equation}
@@
 \begin{equation}
-\bar{\mathcal{O}}_{ij}(x,t)=\frac{\int_\mathbb{R}\mathcal O_{ij}(x,y)\nu_j(y,t)dy}{\int_\mathbb{R}\nu_j(y,t)dy}
+\bar{\mathcal{O}}_{ij}(x,t)=\frac{\int_\mathbb{R}\mathcal O_{ij}(x-y)\nu_j(y,t)dy}{\int_\mathbb{R}\nu_j(y,t)dy}
 =\frac{U_iU_j}{\sqrt{2\pi(w_i+w_j+{\sigma_j}^2(t))}}\exp\left(-\frac{(x-\bar x_j(t))^2}{2(w_i+w_j+{\sigma_j}^2(t))}\right).

### 2020-04-12 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
classical solutions

@@
 \end{equation}
-Thus $\|Fv\|_1\leq |r| \ \|v\|_1$ for all $v\in L^1(\mathbb{R})$ which implies $F$ is Lipschitz continuous.
+Thus $\|Fv\|_1\leq |r| \ \|v\|_1$ for all $v\in L^1(\mathbb{R})$. Then, for any $v,w\in L^1(\mathbb{R})$,
+\begin{multline}
+Fv-Fw=F(v+w)-F(v+w)+Fv-Fw \\
+=m(v+w,x)(v+w)-m(w,x)w-m(v+w,x)(v+w)+m(v,x)v \\
+=\Big(m(v+w,x)-m(w,x)\Big)w-\Big(m(v+w,x)-m(v,x)\Big)v+m(v+w,x)(v-w).
+\end{multline}
+Hence $\|Fv-Fw\|_1\leq2|r|(\|w\|_1+\|v\|_2)+|r|\|v-w\|_1\leq\|r\|\|v-w\|_1$ which implies $F$ is Lipschitz continuous on $L^1(\mathbb{R})$.
 For $T>0$ set $L^{\infty,1}([0,T]\times\mathbb{R})$ as the set of functions $\nu:[0,T]\to L^1(\mathbb{R})$ that satisfy $\|\nu\|_{1,T}=\sup_{t\in[0,T]}\|\nu(t)\|_1<+\infty$ and consider the operator $G$ defined on $L^{\infty,1}([0,T]\times\mathbb{R})$ for fixed $u\in L^1(\mathbb{R})$ by
@@
 We shift our focus to showing mild solutions $\nu$ to PDE (\ref{PDE_supp}) are twice differentiable in $x$ and hence are classical solutions.
-Applying lemma 3.3.2 of Henry [-@Henry1981] we have that the mild solution $\nu$ is in fact a strong (but necessarily classical) solution to the Cauchy problem defined by PDE (\ref{PDE_supp}) with initial condition $\nu(0)=u$ for $t\in[0,T]$.
+Applying lemma 3.3.2 of Henry [-@Henry1981] we have that the mild solution $\nu$ is in fact a strong solution (i.e., differentiable with respect to $t\in(0,T)$) to the abstract Cauchy problem defined by PDE (\ref{PDE_supp}) with initial condition $\nu(0)=u$ for $t\in[0,T]$.
 In particular, this implies that $\nu(t)\in D(\Delta)$ for each $t\in(0,T)$ where $D(\Delta)=W^{2,1}(\mathbb{R})$ is the Sobolov space of integrable functions with weak derivatives at least of order $2$.

### 2020-04-13 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
improved proof of classical solutions

@@
 
+<!--
+
 ## Existence, uniqueness and regularity of solution to deterministic PDE (\ref{eq1}) of the main text \label{regular}
@@
 Applying lemma 3.3.2 of Henry [-@Henry1981] we have that the mild solution $\nu$ is in fact a strong solution (i.e., differentiable with respect to $t\in(0,T)$) to the abstract Cauchy problem defined by PDE (\ref{PDE_supp}) with initial condition $\nu(0)=u$ for $t\in[0,T]$.
-In particular, this implies that $\nu(t)\in D(\Delta)$ for each $t\in(0,T)$ where $D(\Delta)=W^{2,1}(\mathbb{R})$ is the Sobolov space of integrable functions with weak derivatives at least of order $2$.
+In particular, this implies that $\nu(t)\in D(\Delta)$ for each $t\in(0,T)$ where $D(\Delta)=W^{2,1}(\mathbb{R})$ is the Sobolov space of integrable functions with integrable weak derivatives at least of order $2$.
 Since $\frac{1}{p}\geq\frac{1}{p+1}\geq\frac{1}{p}-2$ for all $p\geq1$, the Sobolov embedding theorem [@Henry1981] implies $D(\Delta)\subset L^p(\mathbb{R})$, the Lebesgue space of $p$-integrable functions on $\mathbb{R}$, for all $p\geq1$.
@@
 
+-->
+
 ## Sufficient conditions for finite mean, variance and total abundance in the deterministic case \label{finite}
@@
 Following our assumptions of the main text, we have that $m(h,x)$ is differentiable with respect to both $x$ and $h$ and there exists $r\in\mathbb{R}$ such that $m(h,x)\leq r$ across all $h\geq0$ and $x\in\mathbb{R}$.
-As in the main text, we also assume the initial condition $\nu(x,0)$ is continuous in $x$ and satisfies
-\begin{equation}
-N(0)=\int_{\mathbb{R}}\nu(x,0)dx<+\infty,
-\end{equation}
-\begin{equation}
--\infty<\bar x(0)=\int_{\mathbb{R}}xp(x,0)dx<+\infty,
-\end{equation}
-\begin{equation}
-\sigma^2(0)=\int_{\mathbb{R}}(x-\bar x(0))^2p(x,0)dx<+\infty,
-\end{equation}
-where $p(x,0)=\nu(x,0)/N(0)$, and we consider the PDE
+As in the main text, we also assume the initial condition $u(x)$ is continuous and integrable in $x$ and satisfies
+\begin{equation}
+0\leq\int_{\mathbb{R}}(|x|+x^2)u(x)dx<+\infty
+\end{equation}
+and consider the Cauchy problem
 \begin{equation}\label{PDE_SM}
-\dot\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\Delta\nu(x,t).
-\end{equation}
-
-In SM \S\ref{regular} we have shown PDE (\ref{PDE_SM}) with an integrable initial condition has a classical solution.
-Replacing $m$ with it's upper bound $r\in\mathbb{R}$, equation (\ref{PDE_SM}) reduces to a simple parabolic equation that can be solved using standard techniques [@stanleyfarlow1993].
+\left\{\begin{matrix}
+\dot\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\Delta\nu(x,t) & t>0\\
+\nu(x,0)=u(x) & t=0.
+\end{matrix}\right.
+\end{equation}
+
+As mentioned in the main text, theorem 2.5.6 of Zheng [-@zheng2004nonlinear] states that if $m(\nu,x)$ is locally Lipschitz in both arguments then for some maximal $T>0$ problem (\ref{PDE_SM}) admits a unique local classical solution $\nu(x,t)$ for $t\in[0,T)$.
+Furthermore, either $T=+\infty$ or $T<+\infty$ and $\lim_{t\uparrow T}\int_\mathbb{R}|\nu(t,x)|dx=+\infty$.
+In this section we show that our assumption $m(h,x)\leq r$ for all $h\geq0$ and $x\in\mathbb{R}$ implies $T=+\infty$.
+Replacing $m$ with it's upper bound $r\in\mathbb{R}$, equation (\ref{PDE_SM}) reduces to a simple parabolic equation that can be solved using elementary techniques [@stanleyfarlow1993].
 In particular, when $m(\nu,x)\equiv0$ denote the solution to (\ref{PDE_SM}) by $\nu_0(x,t)$.

### 2020-04-13 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
local lipschitz condition in main text

@@
   pdf_document: 
-    extra_dependencies: ["float","flafter"]
-    toc: false
-    number_sections: true
+    extra_dependencies:
+      - float
+      - flafter
+    fig_caption: yes
+    fig_height: 3
     fig_width: 4
-    fig_height: 3
-    fig_caption: true
+    keep_tex: yes
+    number_sections: yes
 header-includes:
@@
 We set $h=(K\nu)(x,t)$ to account for nonlocal effects of the abundance density $\nu(x,t)$ on the fitness of individuals with trait value $x$ [@Volpert2014].
-An example of a nonlocal effect comes from competition for a common resource among all individuals in a population regardless of their trait value.
-In this case all individuals in the population effect each others fitness' even though they may exhibit diverse trait values.
+As an example of a nonlocal effect, consider a population experiencing intraspecific competition for a common resource such that the competitive interactions are not mediated by any set of traits.
+In this case all individuals in the population effect the fitness of one another even though they may exhibit a diverse range of trait values.
 The term nonlocal comes from interpreting the trait value as spatial location.
@@
 with the initial condition $\nu(x,0)$ described above.
-This PDE is semilinear due to the dependency of the growth rate $m(\nu,x)$ on the solution $\nu(x,t)$ and is referred to as a scalar reaction-diffusion equation [@Henry1981; @lawrenceevans2010].
+This PDE is semilinear due to the dependency of the growth rate $m(\nu,x)$ on the solution $\nu(x,t)$ and is referred to as a scalar reaction-diffusion equation [@zheng2004nonlinear; @lawrenceevans2010].
+To ensure existence and uniqueness of solutions to (\ref{eq1}), we further assume that for each $M>0$ there exists a constant $L_M>0$ depending on $M$ such that $\int_\mathbb{R}\nu_1(x)dx,\int_\mathbb{R}\nu_2(x)dx\leq M$ implies
+\begin{equation}\label{local_lipschitz}
+\int_\mathbb{R}\left|m(\nu_1,x)\nu_1(x)-m(\nu_2,x)\nu_2(x)\right|dx\leq L_M\int_\mathbb{R}\nu_1(x)-\nu_2(x)dx.
+\end{equation}
+In SM \S\ref{finite} we combine the above assumptions on growth rate, mutation and initial condition to prove that solutions to (\ref{eq1}) satisfy $N(t),|\bar x(t)|,\sigma^2(t)<+\infty$ for all $t\geq0$.
+
 Interpreting the trait value $x$ as spatial location and ignoring nonlocal effects, which can be satisfied by setting $K\nu=\nu$, this model of spatially distributed population dynamics has been intensely studied [for a review, see @Cantrell2004].
@@
 where $r\in\mathbb{R}$ is the intrinsic growth rate in the absence of abiotic stabilizing selection, $a>0$ is the strength of abiotic stabilizing selection around the phenotypic optimum $\theta\in\mathbb{R}$, $c>0$ captures the sensitivity of fitness to intraspecific competition and we have set $\kappa(x-y)=1$ so that competitive interactions cause the same reduction in fitness regardless of trait value.
-This exemplary fitness function has a few convenient properties that we breifly take note of.
+This exemplary fitness function has a few convenient properties.
 First, the effect of competition induces a local carrying capacity on the population, leading to a finite equilibrium abundance over bounded subsets of $\mathbb{R}$.
@@
 
-As mentioned in the main text, theorem 2.5.6 of Zheng [-@zheng2004nonlinear] states that if $m(\nu,x)$ is locally Lipschitz in both arguments then for some maximal $T>0$ problem (\ref{PDE_SM}) admits a unique local classical solution $\nu(x,t)$ for $t\in[0,T)$.
-Furthermore, either $T=+\infty$ or $T<+\infty$ and $\lim_{t\uparrow T}\int_\mathbb{R}|\nu(t,x)|dx=+\infty$.
+According to Theorem 2.5.6 of Zheng [-@zheng2004nonlinear], if the operator $F$ defined by $\nu(x,t)\to m(\nu,x)\nu(x,t)$ is locally Lipschitz, corresponding to equation (\ref{local_lipschitz}) of the main text, then for some maximal $T>0$ problem (\ref{PDE_SM}) admits a unique local classical solution $\nu(x,t)$ for $t\in[0,T)$.
+Furthermore, either $T=+\infty$ or $T<+\infty$ and $\lim_{t\uparrow T}\int_\mathbb{R}|\nu(x,t)|dx=+\infty$.
 In this section we show that our assumption $m(h,x)\leq r$ for all $h\geq0$ and $x\in\mathbb{R}$ implies $T=+\infty$.
@@
 
+## Equilibrium moments for a deterministic population experiencing logistic growth and stabilizing selection \label{equilib}
+
+Here we set out to show, given the initial conditions $\nu(\cdot,0)\in L^1(\mathbb{R})$ such that $\bar x(0)\in\mathbb{R}$, $\sigma^2(0)\in[0,+\infty)$ and $N(0)\in(0,+\infty)$, and given the growth rate
+\begin{equation}\label{m_log_stab}
+m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_\mathbb{R}\nu(y,t)dy
+\end{equation}
+such that $\theta\in\mathbb{R}$, $a,c,\mu>0$ and $r>\tfrac{1}{2}\sqrt{\mu a}$, we have the following stable equilibrium $N_\infty=\lim_{t\to\infty}N(t)=\tfrac{1}{c}(r-\tfrac{1}{2}\sqrt{a\mu})$, $\bar x_\infty=\lim_{t\to\infty}\bar x(t)=\theta$ and $\sigma^2_\infty=\lim_{t\to\infty}\sigma^2(t)=\sqrt{\tfrac{\mu}{a}}$.
+
+The mean fitness can be calculated as
+\begin{equation}
+\bar m(t)=r-\frac{a}{2}((\theta-\bar x(t))^2+\sigma^2(t))-cN(t).
+\end{equation}
+Recalling the ODE derived for $N(t)$,
+\begin{equation}
+\frac{d}{dt}N(t)=\bar m(t)N(t)=(r-\frac{a}{2}((\theta-\bar x(t))^2+\sigma^2(t))-cN(t))N(t),
+\end{equation}
+solving for equilibrium total abundance $\hat N$ amounts to setting $\frac{d}{dt}N(t)=0$ and solving for $N(t)$.
+Ignoring the equilibrium $N(t)=0$, this reduces to solving $\bar m(t)=0$ for $N(t)$, which returns
+\begin{equation}
+\hat N=\frac{1}{c}\left(r-\frac{a}{2}\left((\theta-\hat{\bar x})^2-\hat\sigma^2\right)\right).
+\end{equation}
+Unfortunately, deriving ODE for $\bar x(t)$ and $\sigma^2(t)$ leads to expressions involving higher moments and finding ODE for these higher moments will lead to expressions involving yet even higher moments.
+To avoid this infinite regression, we find the equilibrium abundance density $\hat\nu(x)$ by solving $\frac{\partial}{\partial t}\nu(x,t)=0$ for $\nu(x,t)$.
+This implies the following ordinary differential equation
+\begin{equation}
+\frac{d^2}{dx^2}\hat\nu(x)=\left(\frac{2c}{\mu}\hat N+\frac{a}{\mu}(\theta-x)^2-\frac{2r}{\mu}\right)\hat\nu(x)
+\end{equation}
+which has the solution
+\begin{equation}
+\hat\nu(x)=\frac{\hat N}{\sqrt{2\pi}}\left(\frac{a}{\mu}\right)^{\frac{1}{4}}\exp\left(-\sqrt\frac{a}{\mu}\frac{(\theta-x)^2}{2}\right).
+\end{equation}
+From this expression we infer $\hat{\bar x}=\theta$ and $\hat\sigma^2=\sqrt{\frac{\mu}{a}}$.
+Hence $\hat N=\frac{1}{c}\left(r-\frac{1}{2}\sqrt{a\mu}\right)$.
+To show that $N_\infty=\hat N$, $\bar x_\infty=\hat{\bar x}$ and $\sigma^2_\infty=\hat\sigma^2$, we linearize ODE for $N(t)$, $\bar x(t)$ and $\sigma^2(t)$ using the equilibrium $\hat\nu(x)$.
+In particular, since $\hat\nu(x)$ is Gaussian, we do not run into the same issue with higher moments as above.
+Assuming the equilibrium $\hat\nu(x)$ does not change the ODE for $N(t)$, but the ODE for $\bar x(t)$ and $\sigma^2(t)$ can now be expressed as
+\begin{equation}
+\frac{d}{dt}\bar x(t)=\sigma^2(t)\left(\frac{\partial\bar m(t)}{\partial\bar x(t)}-\overline{\frac{\partial m(t)}{\partial\bar x(t)}}\right)=a\sigma^2(t)(\theta-\bar x(t)),
+\end{equation}
+\begin{equation}
+\frac{d}{dt}\sigma^2(t)=2\sigma^4(t)\left(\frac{\partial\bar m(t)}{\partial\sigma^2(t)}-\overline{\frac{\partial m(t)}{\partial\sigma^2(t)}}\right)+\mu=\mu-a\sigma^4(t),
+\end{equation}
+where $\hat p(x)=\hat\nu(x)/\hat N$. These expressions confirm our findings that $\hat{\bar x}=\theta$ and $\hat\sigma^2=\sqrt{\frac{\mu}{a}}$. Furthermore, calculating
+\begin{equation}
+\frac{\partial}{\partial\sigma^2(t)}\frac{d}{dt}\sigma^2(t)=-2a\sigma^2(t)
+\end{equation}
+and evaluating at $\sigma^2(t)=\hat\sigma^2$ demonstrates the equilibrium phenotypic variance is stable when $a,\mu>0$. Hence, calculating
+\begin{equation}
+\frac{\partial}{\partial\bar x(t)}\frac{d}{dt}\bar x(t)=-a\sigma^2(t)
+\end{equation}
+and evaluating at $\sigma^2(t)=\hat\sigma^2$ and $\bar x(t)=\hat{\bar x}$ demonstrates the equilibrium phenotypic mean is stable when $a,\mu>0$. Finally, calculating
+\begin{equation}
+\frac{\partial}{\partial N(t)}\frac{d}{dt}N(t)=r-\frac{a}{2}((\theta-\bar x(t))^2+\sigma^2(t))-2cN(t)
+\end{equation}
+and evaluating at $\sigma^2(t)=\hat\sigma^2$, $\bar x(t)=\hat{\bar x}$ and $N(t)=\hat N$ demonstrates the equilibrium total abundance is stable when $a,c,\mu>0$, and $r>\frac{1}{2}\sqrt{a\mu}$.
+
 ## The relation between diffusion and convolution with a Gaussian kernel \label{diffconvequiv}
@@
 \end{equation}
-
-## Equilibrium moments for a deterministic population experiencing logistic growth and stabilizing selection \label{equilib}
-
-Here we set out to show, given the initial conditions $\nu(\cdot,0)\in L^1(\mathbb{R})$ such that $\bar x(0)\in\mathbb{R}$, $\sigma^2(0)\in[0,+\infty)$ and $N(0)\in(0,+\infty)$, and given the growth rate
-\begin{equation}\label{m_log_stab}
-m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_\mathbb{R}\nu(y,t)dy
-\end{equation}
-such that $\theta\in\mathbb{R}$, $a,c,\mu>0$ and $r>\tfrac{1}{2}\sqrt{\mu a}$, we have the following stable equilibrium $N_\infty=\lim_{t\to\infty}N(t)=\tfrac{1}{c}(r-\tfrac{1}{2}\sqrt{a\mu})$, $\bar x_\infty=\lim_{t\to\infty}\bar x(t)=\theta$ and $\sigma^2_\infty=\lim_{t\to\infty}\sigma^2(t)=\sqrt{\tfrac{\mu}{a}}$.
-
-The mean fitness can be calculated as
-\begin{equation}
-\bar m(t)=r-\frac{a}{2}((\theta-\bar x(t))^2+\sigma^2(t))-cN(t).
-\end{equation}
-Recalling the ODE derived for $N(t)$,
-\begin{equation}
-\frac{d}{dt}N(t)=\bar m(t)N(t)=(r-\frac{a}{2}((\theta-\bar x(t))^2+\sigma^2(t))-cN(t))N(t),
-\end{equation}
-solving for equilibrium total abundance $\hat N$ amounts to setting $\frac{d}{dt}N(t)=0$ and solving for $N(t)$.
-Ignoring the equilibrium $N(t)=0$, this reduces to solving $\bar m(t)=0$ for $N(t)$, which returns
-\begin{equation}
-\hat N=\frac{1}{c}\left(r-\frac{a}{2}\left((\theta-\hat{\bar x})^2-\hat\sigma^2\right)\right).
-\end{equation}
-Unfortunately, deriving ODE for $\bar x(t)$ and $\sigma^2(t)$ leads to expressions involving higher moments and finding ODE for these higher moments will lead to expressions involving yet even higher moments.
-To avoid this infinite regression, we find the equilibrium abundance density $\hat\nu(x)$ by solving $\frac{\partial}{\partial t}\nu(x,t)=0$ for $\nu(x,t)$.
-This implies the following ordinary differential equation
-\begin{equation}
-\frac{d^2}{dx^2}\hat\nu(x)=\left(\frac{2c}{\mu}\hat N+\frac{a}{\mu}(\theta-x)^2-\frac{2r}{\mu}\right)\hat\nu(x)
-\end{equation}
-which has the solution
-\begin{equation}
-\hat\nu(x)=\frac{\hat N}{\sqrt{2\pi}}\left(\frac{a}{\mu}\right)^{\frac{1}{4}}\exp\left(-\sqrt\frac{a}{\mu}\frac{(\theta-x)^2}{2}\right).
-\end{equation}
-From this expression we infer $\hat{\bar x}=\theta$ and $\hat\sigma^2=\sqrt{\frac{\mu}{a}}$.
-Hence $\hat N=\frac{1}{c}\left(r-\frac{1}{2}\sqrt{a\mu}\right)$.
-To show that $N_\infty=\hat N$, $\bar x_\infty=\hat{\bar x}$ and $\sigma^2_\infty=\hat\sigma^2$, we linearize ODE for $N(t)$, $\bar x(t)$ and $\sigma^2(t)$ using the equilibrium $\hat\nu(x)$.
-In particular, since $\hat\nu(x)$ is Gaussian, we do not run into the same issue with higher moments as above.
-Assuming the equilibrium $\hat\nu(x)$ does not change the ODE for $N(t)$, but the ODE for $\bar x(t)$ and $\sigma^2(t)$ can now be expressed as
-\begin{equation}
-\frac{d}{dt}\bar x(t)=\sigma^2(t)\left(\frac{\partial\bar m(t)}{\partial\bar x(t)}-\overline{\frac{\partial m(t)}{\partial\bar x(t)}}\right)=a\sigma^2(t)(\theta-\bar x(t)),
-\end{equation}
-\begin{equation}
-\frac{d}{dt}\sigma^2(t)=2\sigma^4(t)\left(\frac{\partial\bar m(t)}{\partial\sigma^2(t)}-\overline{\frac{\partial m(t)}{\partial\sigma^2(t)}}\right)+\mu=\mu-a\sigma^4(t),
-\end{equation}
-where $\hat p(x)=\hat\nu(x)/\hat N$. These expressions confirm our findings that $\hat{\bar x}=\theta$ and $\hat\sigma^2=\sqrt{\frac{\mu}{a}}$. Furthermore, calculating
-\begin{equation}
-\frac{\partial}{\partial\sigma^2(t)}\frac{d}{dt}\sigma^2(t)=-2a\sigma^2(t)
-\end{equation}
-and evaluating at $\sigma^2(t)=\hat\sigma^2$ demonstrates the equilibrium phenotypic variance is stable when $a,\mu>0$. Hence, calculating
-\begin{equation}
-\frac{\partial}{\partial\bar x(t)}\frac{d}{dt}\bar x(t)=-a\sigma^2(t)
-\end{equation}
-and evaluating at $\sigma^2(t)=\hat\sigma^2$ and $\bar x(t)=\hat{\bar x}$ demonstrates the equilibrium phenotypic mean is stable when $a,\mu>0$. Finally, calculating
-\begin{equation}
-\frac{\partial}{\partial N(t)}\frac{d}{dt}N(t)=r-\frac{a}{2}((\theta-\bar x(t))^2+\sigma^2(t))-2cN(t)
-\end{equation}
-and evaluating at $\sigma^2(t)=\hat\sigma^2$, $\bar x(t)=\hat{\bar x}$ and $N(t)=\hat N$ demonstrates the equilibrium total abundance is stable when $a,c,\mu>0$, and $r>\frac{1}{2}\sqrt{a\mu}$.
 

### 2020-04-13 .A White Noise Approach To Evolutionary Ecology/draft.tex
local lipschitz condition in main text

--- 2020-04-13 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
@@
----
-title: "A White Noise Approach to Evolutionary Ecology"
-author: "Bob Week, Steve Krone, Scott L. Nuismer, Luke J. Harmon"
-date: "3/1/2020"
-abstract: "We derive the dynamics of the distribution of a quantitative character and the abundance of a biological population from a stochastic partial differential equation driven by space-time white noise.
-In the process we develop a useful set of heuristics to operationalize the powerful, but abstract theory of white noise and measure-valued stochastic processes.
-This approach allows us to compute the full implications of demographic stochasticity on phenotypic distributions and abundances of populations.
-We demonstrate the utility of our approach by deriving a quantitative genetic model of diffuse coevolution mediated by exploitative competition for a continuum of resources.
-Other than trait and abundance distributions, this model predicts interaction networks parameterized by rates of interactions, competition coefficients, and selection gradients.
-We briefly investigate the relationship between selection gradients and competition coefficients.
-This illustrative investigation suggests selection gradients can be either positively or negatively correlated with competition coefficients depending on the ratio of interspecific trait variation to intraspecific trait variation.
-Hence, this approach contributes to the development of a synthetic theory of evolutionary ecology by formalizing first principle derivations of stochastic models that underlie rigorous investigations of the relationship between feedbacks of biological processes and the patterns of diversity they produce."
-bibliography: ref.bib
-output:
-  pdf_document: 
-    extra_dependencies:
-      - float
-      - flafter
-    fig_caption: yes
-    fig_height: 3
-    fig_width: 4
-    keep_tex: yes
-    number_sections: yes
-header-includes:
-  - \usepackage{mathtools}
-  - \usepackage{mathrsfs}
-  - \usepackage{hyperref}
-  - \usepackage[left]{lineno}
-  - \usepackage{amsmath}
-  - \linenumbers
-editor_options: 
-  chunk_output_type: console
----
-
-```{r global_options, include=FALSE}
-#knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
-```
-
-# Introduction
-
-Our goal in this manuscript is to develop a rigorous, but accessible approach to synthesize the stochastic dynamics of abundance, mean trait and heritable variation in biological populations for the study of theoretical evolutionary ecology.
-A primary aim of theoretical evolutionary ecology is the development of mathematical approaches to describe the evolution of populations and their interactions with both the biotic and abiotic environments in which they are embedded.
-Given this consideration, a natural scope for such an approach centers on quantifying the abundance dynamics of populations and the evolution of traits mediating their interactions as functions of relevant abiotic factors.
-Although taking into account abundance, phenotype and environment provides the basis for a partial understanding of the complex nature of biological communities, a deeper understanding must account for the effects of dispersal and the phylogeographic history of interacting lineages [@Kraft2007; @Hickerson2010; @Manceau2016; @markmcpeek2017] along with the genetic basis of ecologically relevant traits [@jeffreyconner2004; @FUSSMANN2007] and feedbacks between populations and the biogeochemical cycles they ultimately depend on [@michelloreau2010; @9781107648258].
-It is therefore ideal that the development of any such mathematical approach anticipates extensions to account for these important factors shaping ecological communities, especially as empirical and conceptual work in these directions continues to grow [@AbdalaRoberts2014; @Klzsch2015; @Crutsinger2015; @Fitzpatrick2015; @Fitzpatrick2017; @Marx2017; @Rudman2017; @Skovmand2018; @VanNuland2019; @Harmon2019].
-Furthermore, the approach would benefit from a stochastic component to capture the chance nature of biological reality [@Lande2003; @DeMeester2018; @Mubayi2019] and serve as a basis for the construction of statistical methods that measure evolutionary and ecological processes occuring in the wild.
-Such methods will tether theory to reality and allow for rigorous tests of hypotheses on the structure and behavior of ecological communities.
-In this paper we introduce a framework that establishes a formal connection between the continuous-time dynamics of abundance and quantitative traits in stochastically evolving populations.
-We then demonstrate the utility of our framework through the derivation and analysis of a model of diffuse coevolution and discuss how it can be extended to account for the details mentioned above.
-
-Current theoretical approaches to synthesize evolution and ecology have capitalized on the fact that biological fitness plays a key role in determining both sets of dynamics.
-While correlation of fitness and genotype is the basis of evolution by natural selection, the mean fitness across all individuals in a population determines the growth, stasis or decline of abundance.
-In section \ref{deterministic} we review the mathematical formalization of this connection, which has been established in the contexts of population genetics [@9781932846126; @joanroughgarden1979], evolutionary game theory [@Hofbauer1998; @9780674023383; @Lion2018], quantitative genetics [@Lande1982; @Doebeli1996; @Lion2018] and a unifying framework for these three distinct approaches to evolutionary theory [@Champagnat2006] which includes our approach as a special case.
-We note here there are two primary differences between our aims and those of Champagnat, Ferrière, and Méléard [-@Champagnat2006].
-First, instead of developing a unifying approach to evolutionary theory, we focus on developing a stochastic synthesis of population dynamics and quantitative genetics including specific expressions for the dynamics of abundance, mean trait and additive genetic variance.
-Second, although Champagnat, Ferrière, and Méléard [-@Champagnat2006] provide rigorous proofs of results relevant to our approach, they do not communicate the nessecary tools we make use of for deriving the dynamics of abundance, mean trait and additive genetic variance.
-We aim to communicate these tools to a wide audience by providing an intuitive, yet accurate treatment of white noise and associated heuristics useful for deriving new stochastic models of evolutionary ecology.
-
-Reviewing the above accomplishments reveals a beautiful synthesis of evolution and population ecology.
-However, it also reveals a gap in theoretical approaches to incorporate the intrinsically random nature of populations.
-Specifically, in theoretical quantitative genetics the derivation of a population's response to random genetic drift is derived in discrete time under the assumption of constant effective population size using arguments based on properties of random samples [@Lande1976].
-Though this approach conveniently mimics the formalism provided by the Wright-Fisher model of population genetics, real population sizes fluctuate over time.
-Furthermore, since these fluctuations are themselves stochastic, it seems natural to derive expressions for the evolutionary response to demographic stochasticity and consider how the results relate to characterizations of random genetic drift.
-This can be done in continuous time for population genetic models without too much technical overhead, assuming a finite number of alleles [@Lande2009; @Parsons2010; @Gomulkiewicz2017].
-However, for populations with a continuum of types, such as a quantitative trait, finding a formal approach to derive the evolutionary response to demographic stochasticity has remained a vexing mathematical challenge.
-Previous work integrating stochastic processes with quantitative genetic models have profited from the simplifying assumption that phenotypic and genetic variances remain constant over time, yeilding analytical approximations for the joint dynamics of abundance and the mean of a quantitative trait [@Engen2013].
-These results not only ignore the evolution genetic variance, but they focus on environmental stochasticity and assume sufficiently large populations so that demographic stochasticity is negligible.
-Here we account for both demographic stochasticity and the evolution of additive genetic variance by combining the calculus of white noise with results on rescaled limits of branching Brownian motion processes (BBM) and stochastic partial differential equations (SPDE), which are stochastic analogs of deterministic partial differential equations (PDE).
-Our goal has two components: 1) Establish a novel synthetic approach to theoretical evolutionary ecology that provides a formal connection between demographic stochasticity and random genetic drift in the context of quantitative traits.
-2) Communicate some useful properties of space-time white noise, BBM and SPDE to as wide of audience as possible.
-With this goal in mind we will not provide a rigorous treatment of any of these deep subjects.
-Instead, we introduce a set of heuristics that only require the basic concepts of Reimann integration, partial differentiation and some exposure to Brownian motion and stochastic ordinary differential equations (SDE).
-For a concise introduction to SDE and Brownian motion, we recommend the primer by Evans [-@9781470410544].
-Rigorous treatments of SPDE and rescaled limits of BBM can be found in Walsh [-@Walsh] and Dawson [-@dawson1993measure] respectively.
+\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[]{article}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provides euro and other symbols
+\else % if luatex or xelatex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage[]{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
+\usepackage{hyperref}
+\hypersetup{
+            pdftitle={A White Noise Approach to Evolutionary Ecology},
+            pdfauthor={Bob Week, Steve Krone, Scott L. Nuismer, Luke J. Harmon},
+            pdfborder={0 0 0},
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
+\usepackage[margin=1in]{geometry}
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{5}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
+
+% set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+
+\usepackage{mathtools}
+\usepackage{mathrsfs}
+\usepackage{hyperref}
+\usepackage[left]{lineno}
+\usepackage{amsmath}
+\linenumbers
+\usepackage{float}
+\usepackage{flafter}
+
+\title{A White Noise Approach to Evolutionary Ecology}
+\author{Bob Week, Steve Krone, Scott L. Nuismer, Luke J. Harmon}
+\date{3/1/2020}
+
+\begin{document}
+\maketitle
+\begin{abstract}
+We derive the dynamics of the distribution of a quantitative character
+and the abundance of a biological population from a stochastic partial
+differential equation driven by space-time white noise. In the process
+we develop a useful set of heuristics to operationalize the powerful,
+but abstract theory of white noise and measure-valued stochastic
+processes. This approach allows us to compute the full implications of
+demographic stochasticity on phenotypic distributions and abundances of
+populations. We demonstrate the utility of our approach by deriving a
+quantitative genetic model of diffuse coevolution mediated by
+exploitative competition for a continuum of resources. Other than trait
+and abundance distributions, this model predicts interaction networks
+parameterized by rates of interactions, competition coefficients, and
+selection gradients. We briefly investigate the relationship between
+selection gradients and competition coefficients. This illustrative
+investigation suggests selection gradients can be either positively or
+negatively correlated with competition coefficients depending on the
+ratio of interspecific trait variation to intraspecific trait variation.
+Hence, this approach contributes to the development of a synthetic
+theory of evolutionary ecology by formalizing first principle
+derivations of stochastic models that underlie rigorous investigations
+of the relationship between feedbacks of biological processes and the
+patterns of diversity they produce.
+\end{abstract}
+
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
+
+Our goal in this manuscript is to develop a rigorous, but accessible
+approach to synthesize the stochastic dynamics of abundance, mean trait
+and heritable variation in biological populations for the study of
+theoretical evolutionary ecology. A primary aim of theoretical
+evolutionary ecology is the development of mathematical approaches to
+describe the evolution of populations and their interactions with both
+the biotic and abiotic environments in which they are embedded. Given
+this consideration, a natural scope for such an approach centers on
+quantifying the abundance dynamics of populations and the evolution of
+traits mediating their interactions as functions of relevant abiotic
+factors. Although taking into account abundance, phenotype and
+environment provides the basis for a partial understanding of the
+complex nature of biological communities, a deeper understanding must
+account for the effects of dispersal and the phylogeographic history of
+interacting lineages (Kraft et al. 2007; Hickerson et al. 2010; Manceau,
+Lambert, and Morlon 2016; McPeek 2017) along with the genetic basis of
+ecologically relevant traits (Conner 2004; Fussman, Loreau, and Abrams
+2007) and feedbacks between populations and the biogeochemical cycles
+they ultimately depend on (Loreau 2010; Ågren and Andersson 2012). It is
+therefore ideal that the development of any such mathematical approach
+anticipates extensions to account for these important factors shaping
+ecological communities, especially as empirical and conceptual work in
+these directions continues to grow (Abdala-Roberts and Mooney 2014;
+Kölzsch et al. 2015; Crutsinger 2015; Fitzpatrick et al. 2015, 2017;
+Marx et al. 2017; Rudman et al. 2017; Skovmand et al. 2018; Nuland et
+al. 2019; Harmon et al. 2019). Furthermore, the approach would benefit
+from a stochastic component to capture the chance nature of biological
+reality (Lande, Engen, and SÆther 2003; Meester et al. 2018; Mubayi et
+al. 2019) and serve as a basis for the construction of statistical
+methods that measure evolutionary and ecological processes occuring in
+the wild. Such methods will tether theory to reality and allow for
+rigorous tests of hypotheses on the structure and behavior of ecological
+communities. In this paper we introduce a framework that establishes a
+formal connection between the continuous-time dynamics of abundance and
+quantitative traits in stochastically evolving populations. We then
+demonstrate the utility of our framework through the derivation and
+analysis of a model of diffuse coevolution and discuss how it can be
+extended to account for the details mentioned above.
+
+Current theoretical approaches to synthesize evolution and ecology have
+capitalized on the fact that biological fitness plays a key role in
+determining both sets of dynamics. While correlation of fitness and
+genotype is the basis of evolution by natural selection, the mean
+fitness across all individuals in a population determines the growth,
+stasis or decline of abundance. In section \ref{deterministic} we review
+the mathematical formalization of this connection, which has been
+established in the contexts of population genetics (Crow and Kimura
+1970; Roughgarden 1979), evolutionary game theory (Hofbauer and Sigmund
+1998; Nowak 2006; Lion 2018), quantitative genetics (Lande 1982; Doebeli
+1996; Lion 2018) and a unifying framework for these three distinct
+approaches to evolutionary theory (Champagnat, Ferrière, and Méléard
+2006) which includes our approach as a special case. We note here there
+are two primary differences between our aims and those of Champagnat,
+Ferrière, and Méléard (2006). First, instead of developing a unifying
+approach to evolutionary theory, we focus on developing a stochastic
+synthesis of population dynamics and quantitative genetics including
+specific expressions for the dynamics of abundance, mean trait and
+additive genetic variance. Second, although Champagnat, Ferrière, and
+Méléard (2006) provide rigorous proofs of results relevant to our
+approach, they do not communicate the nessecary tools we make use of for
+deriving the dynamics of abundance, mean trait and additive genetic
+variance. We aim to communicate these tools to a wide audience by
+providing an intuitive, yet accurate treatment of white noise and
+associated heuristics useful for deriving new stochastic models of
+evolutionary ecology.
+
+Reviewing the above accomplishments reveals a beautiful synthesis of
+evolution and population ecology. However, it also reveals a gap in
+theoretical approaches to incorporate the intrinsically random nature of
+populations. Specifically, in theoretical quantitative genetics the
+derivation of a population's response to random genetic drift is derived
+in discrete time under the assumption of constant effective population
+size using arguments based on properties of random samples (Lande 1976).
+Though this approach conveniently mimics the formalism provided by the
+Wright-Fisher model of population genetics, real population sizes
+fluctuate over time. Furthermore, since these fluctuations are
+themselves stochastic, it seems natural to derive expressions for the
+evolutionary response to demographic stochasticity and consider how the
+results relate to characterizations of random genetic drift. This can be
+done in continuous time for population genetic models without too much
+technical overhead, assuming a finite number of alleles (Lande, Engen,
+and Sæther 2009; Parsons, Quince, and Plotkin 2010; Gomulkiewicz, Krone,
+and Remien 2017). However, for populations with a continuum of types,
+such as a quantitative trait, finding a formal approach to derive the
+evolutionary response to demographic stochasticity has remained a vexing
+mathematical challenge. Previous work integrating stochastic processes
+with quantitative genetic models have profited from the simplifying
+assumption that phenotypic and genetic variances remain constant over
+time, yeilding analytical approximations for the joint dynamics of
+abundance and the mean of a quantitative trait (Engen, Lande, and Sæther
+2013). These results not only ignore the evolution genetic variance, but
+they focus on environmental stochasticity and assume sufficiently large
+populations so that demographic stochasticity is negligible. Here we
+account for both demographic stochasticity and the evolution of additive
+genetic variance by combining the calculus of white noise with results
+on rescaled limits of branching Brownian motion processes (BBM) and
+stochastic partial differential equations (SPDE), which are stochastic
+analogs of deterministic partial differential equations (PDE). Our goal
+has two components: 1) Establish a novel synthetic approach to
+theoretical evolutionary ecology that provides a formal connection
+between demographic stochasticity and random genetic drift in the
+context of quantitative traits. 2) Communicate some useful properties of
+space-time white noise, BBM and SPDE to as wide of audience as possible.
+With this goal in mind we will not provide a rigorous treatment of any
+of these deep subjects. Instead, we introduce a set of heuristics that
+only require the basic concepts of Reimann integration, partial
+differentiation and some exposure to Brownian motion and stochastic
+ordinary differential equations (SDE). For a concise introduction to SDE
+and Brownian motion, we recommend the primer by Evans (2014). Rigorous
+treatments of SPDE and rescaled limits of BBM can be found in Walsh
+(1986) and Dawson (1993) respectively.
 
 Through accomplishing these goals this work yields five main results.
-Firstly, we develop a concise, but accurate approach for understanding SPDE and provide heuristics to empower mathematical evolutionary ecologists to derive novel models that relax key assumptions commonly made in combining models of evolutionary and abundance dynamics.
-Second, we formally derive general SDE that express the joint dynamics of abundance, mean trait and additive genetic variance for populations exhibiting a wide range of potential phenotypic distributions.
-Third, we simplify these SDE by approximating the abundance density of phenotypes with a Gaussian curve and integrate the results with a model of imperfect inheritance to find formula that generalize classical quantitative genetic expressions.
-This includes an expression for additive genetic variance analogous to the expression for the response in mean phenotype found by Lande [-@Lande1976].
-Fourth, following classical niche theory we provide a rigorous mechanistic derivation of fitness for populations experiencing resource competition and abiotic stabilizing selection using diffusion limits of measure-valued branching processes.
-This calculation provides a demonstration for performing such diffusion limits that can be applied in other contexts.
-Fifth, we explore implications of the resulting model of diffuse coevolution by investigating the relationship between selection gradients and competition coefficients.
-
-We now summarize the basic structure of the paper.
-To provide motivation for the stochastic equations developed later and background for our model of coevolution, we begin with \S\ref{deterministic} by briefly summarizing derivations of deterministic dynamics of biological populations.
-Starting with a PDE (which we refer to as the Deterministic Asexual Gaussian allelic model with Abundance dynamics, abbreviated DAGA), we arrive at a general set of ordinary differential equations modelling the dynamics of abundance, trait mean and trait variance.
-We discuss the relationship between DAGA and previous models of phenotypic evolution that have been employed in theoretical quantitative genetics.
-Noting that a natural stochastic generalization of DAGA has been previously derived, we find a path to derive SDE describing the evolutionary response to demographic stochasticity.
-We accomplish this in \S\ref{wnc} by discussing how a diffusion limit of a spatially structured branching process (i.e., a measure-valued branching process) leads to the natural SPDE appropriate for our study.
-We refer to this SPDE as the Stochastic Asexual Gaussian allelic model with Abundance dynamics (abbreviated SAGA).
-The diffusion limit in turn provides a rigorous method for constructing fitness functions used in models of evolutionary ecology.
-However, before we discuss measure-valued braching processes and their diffusion limits, we find it necessary to review some fundamental results on the calculus of white noise and introducing a set of mathematical tools based on these results.
-In \S\ref{wnc} and SM \S\ref{SDE_DERIV} we employ these tools to derive a system of SDE generalizing our deterministic results to account for demographic stochasticity.
-We find that the effect of demographic stochasticity on the evolution of mean trait and phenotypic variance characterizes the process of random genetic drift in continuous time.
-However, although biologically insightful, these equations remain difficult to analyze and implement numerically.
-In \S\ref{equations} we use an assumption of normally distributed trait values to simplify these expressions into formulae that are much easier to work with.
-We then account for the constraint of adaptive evolution on the amount of heritable variation in a population by extending these results via a model of imperfect inheritance.
-The resulting equations coincide with classical results in quantitative genetics as a special case.
-In \S\ref{coev} we combine the derived equations of population dynamics with classical niche theory to formulate a model of coevolution across a guild of $S$ species participating in exploitative competition along a common resource continuum.
-In SM \S\ref{diffuse} we apply a classical theorem on rescaled limits of BBM that allow for ecological interactions to provide a rigorous derivation.
-To gain biological insight, in \S\ref{dynamics} we numerically integrate our model of coevolution for $S=100$ species, tracking the dynamics of traits and abundances under scenarios of weak and strong competition.
-We include an account of the natural history of the simulated system and discuss the significance of demographic stochasticity for structuring ecological communities.
-In \S\ref{ecoevo} we provide expressions for selection gradients and competition coefficients implied by our model and use these expressions to investigate the relationship between the degree of competition and strength of coevolution.
-Finally, \S\ref{conclusion} concludes with a summary of accomplishments, a few remarks on the limits of this approach and future directions to incorporate more explicitly the effects of small populations, the genetic architecture of quantitative traits, feedbacks with ecosystem processes and the macroevolutionary history of interacting lineages.
-
-# The framework
-
-At the core of our approach is a stochastic analog of the replicator equation with mutation in continuous time and phenotypic space [@Taylor1978; @Schuster1983].
-From this stochastic replicator-mutator equation we derive a system of SDE for the dynamics of abundance, mean trait and additive genetic variance of a population.
-Hence, our approach develops a quantitative genetic theory of evolutionary ecology.
-A popular alternative to quantitative genetics is the theory of adaptive dynamics.
-As demonstrated by Page and Nowak [-@PAGE2002] and Champagnat, Ferrière, and Méléard [-@Champagnat2006], the canonical equation of adaptive dynamics can be derived from the replicator-mutator equation.
-Thus, one could start from the atomic roots of our approach and pursue a stochastic adaptive dynamic theory instead.
-We choose the former in anticipation of an extension of our approach that explicitly models the genetic details of populations.
-
-In this section we review the derivations of the replicator-mutator equation and trait dynamics from abundance dynamics and extend these formulae along with related results to the stochastic case.
-The results established in this section provide the framework from which larger scale ecological stuctures, such as species abundance distributions and interaction networks, can be computed.
-
-## Deterministic dynamics \label{deterministic}
-
-Our review begins by considering the dynamics of an asexually reproducing population in a homogeneous environment.
-For simplicity, we first assume individuals are haploid and carry one of $K$ alleles each with a different fitness expressed as growth rate before introducing a model involving a quantitative trait.
-Under these assumptions, the derivation of the evolution of allele frequencies due to natural selection can be derived from expressions of exponential growth.
-This, and a few related approaches, have been provided by Crow and Kimura [-@9781932846126, \S 5.3].
-Specifically, denoting $\nu_i$ the abundance of individuals with allele $i$ and $m_i$ the growth rate of allele $i$ [called the Malthusian parameter in @9781932846126], we have
+Firstly, we develop a concise, but accurate approach for understanding
+SPDE and provide heuristics to empower mathematical evolutionary
+ecologists to derive novel models that relax key assumptions commonly
+made in combining models of evolutionary and abundance dynamics. Second,
+we formally derive general SDE that express the joint dynamics of
+abundance, mean trait and additive genetic variance for populations
+exhibiting a wide range of potential phenotypic distributions. Third, we
+simplify these SDE by approximating the abundance density of phenotypes
+with a Gaussian curve and integrate the results with a model of
+imperfect inheritance to find formula that generalize classical
+quantitative genetic expressions. This includes an expression for
+additive genetic variance analogous to the expression for the response
+in mean phenotype found by Lande (1976). Fourth, following classical
+niche theory we provide a rigorous mechanistic derivation of fitness for
+populations experiencing resource competition and abiotic stabilizing
+selection using diffusion limits of measure-valued branching processes.
+This calculation provides a demonstration for performing such diffusion
+limits that can be applied in other contexts. Fifth, we explore
+implications of the resulting model of diffuse coevolution by
+investigating the relationship between selection gradients and
+competition coefficients.
+
+We now summarize the basic structure of the paper. To provide motivation
+for the stochastic equations developed later and background for our
+model of coevolution, we begin with \S\ref{deterministic} by briefly
+summarizing derivations of deterministic dynamics of biological
+populations. Starting with a PDE (which we refer to as the Deterministic
+Asexual Gaussian allelic model with Abundance dynamics, abbreviated
+DAGA), we arrive at a general set of ordinary differential equations
+modelling the dynamics of abundance, trait mean and trait variance. We
+discuss the relationship between DAGA and previous models of phenotypic
+evolution that have been employed in theoretical quantitative genetics.
+Noting that a natural stochastic generalization of DAGA has been
+previously derived, we find a path to derive SDE describing the
+evolutionary response to demographic stochasticity. We accomplish this
+in \S\ref{wnc} by discussing how a diffusion limit of a spatially
+structured branching process (i.e., a measure-valued branching process)
+leads to the natural SPDE appropriate for our study. We refer to this
+SPDE as the Stochastic Asexual Gaussian allelic model with Abundance
+dynamics (abbreviated SAGA). The diffusion limit in turn provides a
+rigorous method for constructing fitness functions used in models of
+evolutionary ecology. However, before we discuss measure-valued braching
+processes and their diffusion limits, we find it necessary to review
+some fundamental results on the calculus of white noise and introducing
+a set of mathematical tools based on these results. In \S\ref{wnc} and
+SM \S\ref{SDE_DERIV} we employ these tools to derive a system of SDE
+generalizing our deterministic results to account for demographic
+stochasticity. We find that the effect of demographic stochasticity on
+the evolution of mean trait and phenotypic variance characterizes the
+process of random genetic drift in continuous time. However, although
+biologically insightful, these equations remain difficult to analyze and
+implement numerically. In \S\ref{equations} we use an assumption of
+normally distributed trait values to simplify these expressions into
+formulae that are much easier to work with. We then account for the
+constraint of adaptive evolution on the amount of heritable variation in
+a population by extending these results via a model of imperfect
+inheritance. The resulting equations coincide with classical results in
+quantitative genetics as a special case. In \S\ref{coev} we combine the
+derived equations of population dynamics with classical niche theory to
+formulate a model of coevolution across a guild of \(S\) species
+participating in exploitative competition along a common resource
+continuum. In SM \S\ref{diffuse} we apply a classical theorem on
+rescaled limits of BBM that allow for ecological interactions to provide
+a rigorous derivation. To gain biological insight, in \S\ref{dynamics}
+we numerically integrate our model of coevolution for \(S=100\) species,
+tracking the dynamics of traits and abundances under scenarios of weak
+and strong competition. We include an account of the natural history of
+the simulated system and discuss the significance of demographic
+stochasticity for structuring ecological communities. In \S\ref{ecoevo}
+we provide expressions for selection gradients and competition
+coefficients implied by our model and use these expressions to
+investigate the relationship between the degree of competition and
+strength of coevolution. Finally, \S\ref{conclusion} concludes with a
+summary of accomplishments, a few remarks on the limits of this approach
+and future directions to incorporate more explicitly the effects of
+small populations, the genetic architecture of quantitative traits,
+feedbacks with ecosystem processes and the macroevolutionary history of
+interacting lineages.
+
+\hypertarget{the-framework}{%
+\section{The framework}\label{the-framework}}
+
+At the core of our approach is a stochastic analog of the replicator
+equation with mutation in continuous time and phenotypic space (Taylor
+and Jonker 1978; Schuster and Sigmund 1983). From this stochastic
+replicator-mutator equation we derive a system of SDE for the dynamics
+of abundance, mean trait and additive genetic variance of a population.
+Hence, our approach develops a quantitative genetic theory of
+evolutionary ecology. A popular alternative to quantitative genetics is
+the theory of adaptive dynamics. As demonstrated by Page and Nowak
+(2002) and Champagnat, Ferrière, and Méléard (2006), the canonical
+equation of adaptive dynamics can be derived from the replicator-mutator
+equation. Thus, one could start from the atomic roots of our approach
+and pursue a stochastic adaptive dynamic theory instead. We choose the
+former in anticipation of an extension of our approach that explicitly
+models the genetic details of populations.
+
+In this section we review the derivations of the replicator-mutator
+equation and trait dynamics from abundance dynamics and extend these
+formulae along with related results to the stochastic case. The results
+established in this section provide the framework from which larger
+scale ecological stuctures, such as species abundance distributions and
+interaction networks, can be computed.
+
+\hypertarget{deterministic-dynamics}{%
+\subsection{\texorpdfstring{Deterministic dynamics
+\label{deterministic}}{Deterministic dynamics }}\label{deterministic-dynamics}}
+
+Our review begins by considering the dynamics of an asexually
+reproducing population in a homogeneous environment. For simplicity, we
+first assume individuals are haploid and carry one of \(K\) alleles each
+with a different fitness expressed as growth rate before introducing a
+model involving a quantitative trait. Under these assumptions, the
+derivation of the evolution of allele frequencies due to natural
+selection can be derived from expressions of exponential growth. This,
+and a few related approaches, have been provided by Crow and Kimura
+(1970, \S 5.3). Specifically, denoting \(\nu_i\) the abundance of
+individuals with allele \(i\) and \(m_i\) the growth rate of allele
+\(i\) (called the Malthusian parameter in Crow and Kimura 1970), we have
 \begin{equation}\label{first}
 \frac{d\nu_i}{dt}=m_i\nu_i.
-\end{equation}
-Starting from this model, we get the total abundance of the population as $N=\sum_{i=1}^K\nu_i$, the frequency of allele $i$ as $p_i=\nu_i/N$ and the mean fitness of the population as $\bar m=\sum_{i=1}^Kp_im_i$.
-Hence, we can employ some elementary calculus to derive the dynamics of abundance $dN/dt$ and the dynamics of allele frequencies $dp_1/dt,\dots,dp_K/dt$ as
-\begin{equation}
+\end{equation} Starting from this model, we get the total abundance of
+the population as \(N=\sum_{i=1}^K\nu_i\), the frequency of allele \(i\)
+as \(p_i=\nu_i/N\) and the mean fitness of the population as
+\(\bar m=\sum_{i=1}^Kp_im_i\). Hence, we can employ some elementary
+calculus to derive the dynamics of abundance \(dN/dt\) and the dynamics
+of allele frequencies \(dp_1/dt,\dots,dp_K/dt\) as \begin{equation}
 \frac{dN}{dt}=\sum_{i=1}^K\nu_im_i=N\sum_{i=1}^Kp_im_i=\bar mN,
-\end{equation}
-\begin{equation}\label{popgen}
+\end{equation} \begin{equation}\label{popgen}
 \frac{dp_i}{dt}=\frac{d}{dt}\frac{\nu_i}{N}=\frac{1}{N^2}\left(N\frac{d\nu_i}{dt}-\frac{dN}{dt}\nu_i\right)
 =\frac{1}{N}\left(m_i\nu_i-\bar m Np_i\right)=(m_i-\bar m)p_i.
-\end{equation}
-Two important observations of these equations include 1) mean fitness $\bar m$ determines the abundance dynamics of the entire population and 2) allele $i$ will increase (decrease) in frequency if $m_i>\bar m$ ($<\bar m$).
-Equation (\ref{popgen}) is known in the field of evolutionary game theory as the replicator equation [@Hofbauer1998; @9780674023383; @Lion2018; @Taylor1978; @Schuster1983].
-Instead of being explicitly focused on alleles, the replicator equation describes the fluctuations of relative abundances of various _types_ in a population in terms of the vital rates of each type.
-Using a matrix of transition rates between differing types, it is straight-forward to extend the replicator equation to include mutation, which is known as the replicator-mutator equation [@9780674023383].
-
-Inspired by equations (\ref{first})-(\ref{popgen}), we derive an analog of the replicator-mutator equation for a continuum of types (that is, for a quantitative trait).
-In particular, we model a continuously reproducing population with trait values $x\in\mathbb{R}$ and an abundance density $\nu(x,t)$ that represents the amount of individuals in the population with trait value $x$ at time $t$.
-Hence, the abundance density satisfies $N(t)=\int_{-\infty}^{+\infty}\nu(x,t)dx$ and $p(x,t)=\nu(x,t)/N(t)$ is the relative density of trait $x$ which we also refer to as the phenotypic distribution.
-
-To stay within the realm of biological plausibility we require a set technical assumptions.
-First, we assume the initial abundance density is continuous, non-negative, integrable and has finite trait mean and variance.
-That is, we assume $\nu(x,0)$ is continuous in $x$, satisfies $\nu(x,0)\geq0$ for all $x\in\mathbb{R}$ and
+\end{equation} Two important observations of these equations include 1)
+mean fitness \(\bar m\) determines the abundance dynamics of the entire
+population and 2) allele \(i\) will increase (decrease) in frequency if
+\(m_i>\bar m\) (\(<\bar m\)). Equation (\ref{popgen}) is known in the
+field of evolutionary game theory as the replicator equation (Hofbauer
+and Sigmund 1998; Nowak 2006; Lion 2018; Taylor and Jonker 1978;
+Schuster and Sigmund 1983). Instead of being explicitly focused on
+alleles, the replicator equation describes the fluctuations of relative
+abundances of various \emph{types} in a population in terms of the vital
+rates of each type. Using a matrix of transition rates between differing
+types, it is straight-forward to extend the replicator equation to
+include mutation, which is known as the replicator-mutator equation
+(Nowak 2006).
+
+Inspired by equations (\ref{first})-(\ref{popgen}), we derive an analog
+of the replicator-mutator equation for a continuum of types (that is,
+for a quantitative trait). In particular, we model a continuously
+reproducing population with trait values \(x\in\mathbb{R}\) and an
+abundance density \(\nu(x,t)\) that represents the amount of individuals
+in the population with trait value \(x\) at time \(t\). Hence, the
+abundance density satisfies \(N(t)=\int_{-\infty}^{+\infty}\nu(x,t)dx\)
+and \(p(x,t)=\nu(x,t)/N(t)\) is the relative density of trait \(x\)
+which we also refer to as the phenotypic distribution.
+
+To stay within the realm of biological plausibility we require a set
+technical assumptions. First, we assume the initial abundance density is
+continuous, non-negative, integrable and has finite trait mean and
+variance. That is, we assume \(\nu(x,0)\) is continuous in \(x\),
+satisfies \(\nu(x,0)\geq0\) for all \(x\in\mathbb{R}\) and
 \begin{equation}
 N(0)=\int_{-\infty}^{+\infty}\nu(x,0)dx<+\infty,
-\end{equation}
-\begin{equation}
+\end{equation} \begin{equation}
 -\infty<\bar x(0)=\int_{-\infty}^{+\infty}xp(x,0)dx<+\infty,
-\end{equation}
-\begin{equation}
+\end{equation} \begin{equation}
 \sigma^2(0)=\int_{-\infty}^{+\infty}(x-\bar x(0))^2p(x,0)dx<+\infty,
-\end{equation}
-where $\bar x(t)$ and $\sigma^2(t)$ are respectively the mean trait and phenotypic variance at time $t\geq0$.
-Second, we assume selection is determined by the growth rate $m(h,x)$ that is differentiable with respect to both arguments and is bounded above by some $r\in\mathbb{R}$ for all $x\in\mathbb{R}$ and $h\geq0$.
-We denote by $K$ an operator that takes functions such as $\nu(x,t)$ as its argument and returns other functions that may still depend on the spatial variable $x$ and the temporal variable $t$.
-For our application we will be focused on operators $K$ that can be written as $(K\nu)(x,t)=\int_{-\infty}^{+\infty}\kappa(x-y)\nu(y,t)dy$ for some non-negative and bounded function $\kappa$.
-We set $h=(K\nu)(x,t)$ to account for nonlocal effects of the abundance density $\nu(x,t)$ on the fitness of individuals with trait value $x$ [@Volpert2014].
-As an example of a nonlocal effect, consider a population experiencing intraspecific competition for a common resource such that the competitive interactions are not mediated by any set of traits.
-In this case all individuals in the population effect the fitness of one another even though they may exhibit a diverse range of trait values.
-The term nonlocal comes from interpreting the trait value as spatial location.
-From here on we abbreviate $m(h,x)=m\big((K\nu)(x,t),x\big)$ to just $m(\nu,x)$.
-Third, we assume mutation is captured by diffusion with coefficient $\frac{\mu}{2}$.
-With these technicalities addressed we model the demographic dynamics of a population and the dynamics of a quantitative character simultaneously by the PDE
+\end{equation} where \(\bar x(t)\) and \(\sigma^2(t)\) are respectively
+the mean trait and phenotypic variance at time \(t\geq0\). Second, we
+assume selection is determined by the growth rate \(m(h,x)\) that is
+differentiable with respect to both arguments and is bounded above by
+some \(r\in\mathbb{R}\) for all \(x\in\mathbb{R}\) and \(h\geq0\). We
+denote by \(K\) an operator that takes functions such as \(\nu(x,t)\) as
+its argument and returns other functions that may still depend on the
+spatial variable \(x\) and the temporal variable \(t\). For our
+application we will be focused on operators \(K\) that can be written as
+\((K\nu)(x,t)=\int_{-\infty}^{+\infty}\kappa(x-y)\nu(y,t)dy\) for some
+non-negative and bounded function \(\kappa\). We set \(h=(K\nu)(x,t)\)
+to account for nonlocal effects of the abundance density \(\nu(x,t)\) on
+the fitness of individuals with trait value \(x\) (Volpert 2014). As an
+example of a nonlocal effect, consider a population experiencing
+intraspecific competition for a common resource such that the
+competitive interactions are not mediated by any set of traits. In this
+case all individuals in the population effect the fitness of one another
+even though they may exhibit a diverse range of trait values. The term
+nonlocal comes from interpreting the trait value as spatial location.
+From here on we abbreviate \(m(h,x)=m\big((K\nu)(x,t),x\big)\) to just
+\(m(\nu,x)\). Third, we assume mutation is captured by diffusion with
+coefficient \(\frac{\mu}{2}\). With these technicalities addressed we
+model the demographic dynamics of a population and the dynamics of a
+quantitative character simultaneously by the PDE
 \begin{equation}\label{eq1}
 \frac{\partial}{\partial t}\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)
-\end{equation}
-with the initial condition $\nu(x,0)$ described above.
-This PDE is semilinear due to the dependency of the growth rate $m(\nu,x)$ on the solution $\nu(x,t)$ and is referred to as a scalar reaction-diffusion equation [@zheng2004nonlinear; @lawrenceevans2010].
-To ensure existence and uniqueness of solutions to (\ref{eq1}), we further assume that for each $M>0$ there exists a constant $L_M>0$ depending on $M$ such that $\int_\mathbb{R}\nu_1(x)dx,\int_\mathbb{R}\nu_2(x)dx\leq M$ implies
+\end{equation} with the initial condition \(\nu(x,0)\) described above.
+This PDE is semilinear due to the dependency of the growth rate
+\(m(\nu,x)\) on the solution \(\nu(x,t)\) and is referred to as a scalar
+reaction-diffusion equation (Zheng 2004; Evans 2010). To ensure
+existence and uniqueness of solutions to (\ref{eq1}), we further assume
+that for each \(M>0\) there exists a constant \(L_M>0\) depending on
+\(M\) such that
+\(\int_\mathbb{R}\nu_1(x)dx,\int_\mathbb{R}\nu_2(x)dx\leq M\) implies
 \begin{equation}\label{local_lipschitz}
 \int_\mathbb{R}\left|m(\nu_1,x)\nu_1(x)-m(\nu_2,x)\nu_2(x)\right|dx\leq L_M\int_\mathbb{R}\nu_1(x)-\nu_2(x)dx.
-\end{equation}
-In SM \S\ref{finite} we combine the above assumptions on growth rate, mutation and initial condition to prove that solutions to (\ref{eq1}) satisfy $N(t),|\bar x(t)|,\sigma^2(t)<+\infty$ for all $t\geq0$.
-
-Interpreting the trait value $x$ as spatial location and ignoring nonlocal effects, which can be satisfied by setting $K\nu=\nu$, this model of spatially distributed population dynamics has been intensely studied [for a review, see @Cantrell2004].
-When $\mu=0$, equation (\ref{eq1}) can be seen as an analog of equation (\ref{first}) for a continuum of types.
-By assuming mutation acts as diffusion, the effect of mutation causes $\nu(x,t)$ to flatten out over time.
-In fact, if the growth rate is constant across $x$, then this model of mutation will cause $\nu(x,t)$ to converge to a flat line as $t\to\infty$.
-Although clearly an idealized representation of biological reality, this model is sufficiently general to capture a large class of dynamics including density dependent growth and frequency dependent selection.
-As an example, logistic growth combined with quadratic stabilizing selection can be captured using the growth rate
+\end{equation} In SM \S\ref{finite} we combine the above assumptions on
+growth rate, mutation and initial condition to prove that solutions to
+(\ref{eq1}) satisfy \(N(t),|\bar x(t)|,\sigma^2(t)<+\infty\) for all
+\(t\geq0\).
+
+Interpreting the trait value \(x\) as spatial location and ignoring
+nonlocal effects, which can be satisfied by setting \(K\nu=\nu\), this
+model of spatially distributed population dynamics has been intensely
+studied (for a review, see Cantrell and Cosner 2004). When \(\mu=0\),
+equation (\ref{eq1}) can be seen as an analog of equation (\ref{first})
+for a continuum of types. By assuming mutation acts as diffusion, the
+effect of mutation causes \(\nu(x,t)\) to flatten out over time. In
+fact, if the growth rate is constant across \(x\), then this model of
+mutation will cause \(\nu(x,t)\) to converge to a flat line as
+\(t\to\infty\). Although clearly an idealized representation of
+biological reality, this model is sufficiently general to capture a
+large class of dynamics including density dependent growth and frequency
+dependent selection. As an example, logistic growth combined with
+quadratic stabilizing selection can be captured using the growth rate
 \begin{equation}\label{special_m}
 m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_{-\infty}^{+\infty}\nu(y,t)dy
-\end{equation}
-where $r\in\mathbb{R}$ is the intrinsic growth rate in the absence of abiotic stabilizing selection, $a>0$ is the strength of abiotic stabilizing selection around the phenotypic optimum $\theta\in\mathbb{R}$, $c>0$ captures the sensitivity of fitness to intraspecific competition and we have set $\kappa(x-y)=1$ so that competitive interactions cause the same reduction in fitness regardless of trait value.
-This exemplary fitness function has a few convenient properties.
-First, the effect of competition induces a local carrying capacity on the population, leading to a finite equilibrium abundance over bounded subsets of $\mathbb{R}$.
-Second, abiotic selection prevents the abundance density from diffusing too far from the abiotic optimum, leading to a finite equilibrial abundance across all of $\mathbb{R}$, $\lim_{t\to\infty}N(t)=\tfrac{1}{c}(r-\tfrac{1}{2}\sqrt{a\mu})$, finite equilibrial mean trait $\lim_{t\to\infty}\bar x(t)=\theta$ and finite equilibrial phenotypic variance $\lim_{t\to\infty}\sigma^2(t)=\sqrt{\frac{\mu}{a}}$, so long as $\bar x(0)\in\mathbb{R}$, $\sigma^2(0)\in[0,+\infty)$, $N(0)\in(0,+\infty)$ and $r>\frac{1}{2}\sqrt{a\mu}>0$.
-We prove these statements in SM \S\ref{equilib}.
-Note in particular the coincidence of equilibrial phenotypic variance predicted by this model and that predicted by the classic Gaussian allelic model of quantitative genetics [@Lande1975; @9780471986539; @Johnson2005].
-By replacing $\kappa(x-y)=1$ in expression (\ref{special_m}) with a particular Gaussian function, we will find in \S\ref{coev} and SM \S\ref{diffuse} this fitness function becomes equivalent to a fitness function derived from niche theory that we use to model coevolution driven by resource competition.
-
-To derive a replicator-mutator equation from equation (\ref{eq1}), we employ the chain rule from calculus. Writing $\bar m(t)=\int_{-\infty}^{+\infty}m(\nu,x)p(x,t)dx$ for the mean fitness, we have
+\end{equation} where \(r\in\mathbb{R}\) is the intrinsic growth rate in
+the absence of abiotic stabilizing selection, \(a>0\) is the strength of
+abiotic stabilizing selection around the phenotypic optimum
+\(\theta\in\mathbb{R}\), \(c>0\) captures the sensitivity of fitness to
+intraspecific competition and we have set \(\kappa(x-y)=1\) so that
+competitive interactions cause the same reduction in fitness regardless
+of trait value. This exemplary fitness function has a few convenient
+properties. First, the effect of competition induces a local carrying
+capacity on the population, leading to a finite equilibrium abundance
+over bounded subsets of \(\mathbb{R}\). Second, abiotic selection
+prevents the abundance density from diffusing too far from the abiotic
+optimum, leading to a finite equilibrial abundance across all of
+\(\mathbb{R}\),
+\(\lim_{t\to\infty}N(t)=\tfrac{1}{c}(r-\tfrac{1}{2}\sqrt{a\mu})\),
+finite equilibrial mean trait \(\lim_{t\to\infty}\bar x(t)=\theta\) and
+finite equilibrial phenotypic variance
+\(\lim_{t\to\infty}\sigma^2(t)=\sqrt{\frac{\mu}{a}}\), so long as
+\(\bar x(0)\in\mathbb{R}\), \(\sigma^2(0)\in[0,+\infty)\),
+\(N(0)\in(0,+\infty)\) and \(r>\frac{1}{2}\sqrt{a\mu}>0\). We prove
+these statements in SM \S\ref{equilib}. Note in particular the
+coincidence of equilibrial phenotypic variance predicted by this model
+and that predicted by the classic Gaussian allelic model of quantitative
+genetics (Lande 1975; Bürger 2000; Johnson and Barton 2005). By
+replacing \(\kappa(x-y)=1\) in expression (\ref{special_m}) with a
+particular Gaussian function, we will find in \S\ref{coev} and SM
+\S\ref{diffuse} this fitness function becomes equivalent to a fitness
+function derived from niche theory that we use to model coevolution
+driven by resource competition.
+
+To derive a replicator-mutator equation from equation (\ref{eq1}), we
+employ the chain rule from calculus. Writing
+\(\bar m(t)=\int_{-\infty}^{+\infty}m(\nu,x)p(x,t)dx\) for the mean
+fitness, we have
 
@@
 =N(t)\int_{-\infty}^{+\infty} m(\nu,x)p(x,t)dx=\bar m(t) N(t).
-\end{multline}
-Using our assumptions on mutation and rate of growth, we show in SM \S\ref{finite} $\nu(x,t)$ is twice differentiable with respect to $x$ and $\int_{-\infty}^{+\infty}\nu(x,t)dx<\infty$ for all $t\geq0$.
-This implies that we are justified in swapping the order of differentiation and integration and the result $\int_{-\infty}^{+\infty}\frac{\partial^2}{\partial x^2}\nu(x,t)dx=0$ can be derived from the fundamental theorem of calculus.
-Biological reasoning agrees with this latter result since mutation neither creates nor destroys individuals, but merely changes their type from their parental type.
-Taking the same approach, we derive the dynamics of the phenotypic distribution $p(x,t)$ in response to selection and muation as
+\end{multline} Using our assumptions on mutation and rate of growth, we
+show in SM \S\ref{finite} \(\nu(x,t)\) is twice differentiable with
+respect to \(x\) and \(\int_{-\infty}^{+\infty}\nu(x,t)dx<\infty\) for
+all \(t\geq0\). This implies that we are justified in swapping the order
+of differentiation and integration and the result
+\(\int_{-\infty}^{+\infty}\frac{\partial^2}{\partial x^2}\nu(x,t)dx=0\)
+can be derived from the fundamental theorem of calculus. Biological
+reasoning agrees with this latter result since mutation neither creates
+nor destroys individuals, but merely changes their type from their
+parental type. Taking the same approach, we derive the dynamics of the
+phenotypic distribution \(p(x,t)\) in response to selection and muation
+as
 
@@
 =(m(\nu,x)-\bar m(t)) p(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t).
-\end{multline}
-This result closely resembles Kimura's continuum-of-alleles model [@Kimura1965; @Brger1986].
-The primary difference being that our model utilizes diffusion instead of convolution with an arbitrary mutation kernel.
-Of course, our model of mutation can be derived as an approximation to Kimura's model, which, for sexually reproducing populations, has been referred to as the Gaussian allelic approximation in reference to the distribution of mutational effects at loci in a genome on the values of traits [@Lande1975; but see also @9780471986539; and @Johnson2005], the infinitesimal model in reference to modelling continuous traits as being encoded by an infinite number of loci each having infinitesimal effect [@Barton2017] and the Gaussian descendants approximation in reference to offspring trait values being normally distributed around their parental values [@Turelli2017].
-Alternatively, since diffusion is the continuous-time equivalent to convolution against a Gaussian kernel (SM \S\ref{diffconvequiv}), equation (\ref{p_PDE}) can also be seen as a special case of Kimura's continuum-of-alleles model.
-We return to these concepts in \S\ref{inheritance} where we review a simple model of inheritance that leads us to an expression for the stochastic evolution of additive genetic variance.
-To clarify the distinction between this model and previous models of phenotypic evolution we refer to the PDE (\ref{eq1}) from which (\ref{p_PDE}) was derived as the Deterministic Asexual Gaussian allelic model with Abundance dynamics (abbreviated DAGA).
-Later, we will extend this model to include the effects of demographic stochasticity, which we refer to as the Stochastic Asexual Gaussian allelic model with Abundance dynamics (abbreviated SAGA).
-
-The covariance of fitness and phenotype across the population is defined as
-\begin{equation}
+\end{multline} This result closely resembles Kimura's
+continuum-of-alleles model (Kimura 1965; Bürger 1986). The primary
+difference being that our model utilizes diffusion instead of
+convolution with an arbitrary mutation kernel. Of course, our model of
+mutation can be derived as an approximation to Kimura's model, which,
+for sexually reproducing populations, has been referred to as the
+Gaussian allelic approximation in reference to the distribution of
+mutational effects at loci in a genome on the values of traits (Lande
+1975; but see also Bürger 2000; and Johnson and Barton 2005), the
+infinitesimal model in reference to modelling continuous traits as being
+encoded by an infinite number of loci each having infinitesimal effect
+(Barton, Etheridge, and Véber 2017) and the Gaussian descendants
+approximation in reference to offspring trait values being normally
+distributed around their parental values (Turelli 2017). Alternatively,
+since diffusion is the continuous-time equivalent to convolution against
+a Gaussian kernel (SM \S\ref{diffconvequiv}), equation (\ref{p_PDE}) can
+also be seen as a special case of Kimura's continuum-of-alleles model.
+We return to these concepts in \S\ref{inheritance} where we review a
+simple model of inheritance that leads us to an expression for the
+stochastic evolution of additive genetic variance. To clarify the
+distinction between this model and previous models of phenotypic
+evolution we refer to the PDE (\ref{eq1}) from which (\ref{p_PDE}) was
+derived as the Deterministic Asexual Gaussian allelic model with
+Abundance dynamics (abbreviated DAGA). Later, we will extend this model
+to include the effects of demographic stochasticity, which we refer to
+as the Stochastic Asexual Gaussian allelic model with Abundance dynamics
+(abbreviated SAGA).
+
+The covariance of fitness and phenotype across the population is defined
+as \begin{equation}
 \mathrm{Cov}_t\Big(m(\nu,x),x\Big)=\int_{-\infty}^{+\infty}\Big(m(\nu,x)-\bar m(t)\Big)\Big(x-\bar x(t)\Big)p(x,t)dx.
-\end{equation}
-Hence, the dynamics of the mean trait $\bar x(t)$ can be derived as
+\end{equation} Hence, the dynamics of the mean trait \(\bar x(t)\) can
+be derived as
 
@@
 
-Equation (\ref{detmean}) is a continuous time analog of the well known Robertson-Price equation without transmission bias [@Robertson1966; @PRICE1970; @FRANK2012; @Queller2017; @Lion2018].
-Whether or not the covariance of fitness and phenotype creates change in $\bar x$ to maximize mean fitness $\bar m$ depends on the degree to which selection is frequency dependent [@Lande1976].
-Since this change is driven by a covariance with respect to phenotypic diversity, the response in mean trait to selection is mediated by the phenotypic variance.
-In particular, when $\sigma^2=0$, $\bar x$ will not respond to selection.
-The result $\int_{-\infty}^{+\infty} x\frac{\partial^2}{\partial x^2} p(x,t)dx=0$ can be found by applying integration by parts.
-Following the approach taken to calculate the evolution of $\bar x$, we find the response of phenotypic variation to this model of selection and mutation is
+Equation (\ref{detmean}) is a continuous time analog of the well known
+Robertson-Price equation without transmission bias (Robertson 1966;
+Price 1970; Frank 2012; Queller 2017; Lion 2018). Whether or not the
+covariance of fitness and phenotype creates change in \(\bar x\) to
+maximize mean fitness \(\bar m\) depends on the degree to which
+selection is frequency dependent (Lande 1976). Since this change is
+driven by a covariance with respect to phenotypic diversity, the
+response in mean trait to selection is mediated by the phenotypic
+variance. In particular, when \(\sigma^2=0\), \(\bar x\) will not
+respond to selection. The result
+\(\int_{-\infty}^{+\infty} x\frac{\partial^2}{\partial x^2} p(x,t)dx=0\)
+can be found by applying integration by parts. Following the approach
+taken to calculate the evolution of \(\bar x\), we find the response of
+phenotypic variation to this model of selection and mutation is
 \begin{equation}\label{vardyn_det}
 \frac{d}{dt}\sigma^2(t)=\mathrm{Cov}_t\Big(m(\nu,x),(x-\bar x)^2\Big)+\mu.
-\end{equation}
-For the sake of space we relegate the derivation of $d\sigma^2/dt$ to SM \S\ref{var_deriv}.
-In the absence of mutation equation (\ref{vardyn_det}) agrees with the result derived by Lion [-@Lion2018] for discrete phenotypes.
-From a statistical perspective, if we think of $(x-\bar x)^2$ as a square error, then in analogy to the dynamics of the mean trait, we see that the response in $\sigma^2$ to selection can be expressed as a covariance of fitness and square error, which is defined in analogy to $\mathrm{Cov}_t(m(\nu,x),x)$.
-Just as for the evolution of $\bar x(t)$, this covariance also creates change in $\sigma^2$ that can either increase or decrease mean fitness $\bar m$, depending on whether or not individual fitness depends on the distribution of traits across the entire population.
-The effect of selection on phenotypic variance can be positive or negative depending on whether selection is stabilizing or disruptive.
-
-In SM \S\ref{SDE_DERIV}, we extend these results to include the effects of demographic stochasticity.
-The key realization in this direction is that a rigorous derivation of a SPDE that naturally generalizes DAGA and accounts for ecological interactions has been provided by Li [-@zeng1998absolute].
-To understand how this SPDE was derived, we provide an informal discussion of measure-valued branching processes (which treat populations as sets of discrete individuals) and their diffusion limits, the so-called superprocesses (which treat population size as a continuously varying number).
-Similar to the approach taken by Champagnat, Ferriére and Méléard [-@Champagnat2006], we begin with a BBM that models populations as discrete sets of reproducing individuals whose vital rates depend on their trait value as well as the state of the entire population.
-Taking a large population size limit and keeping our assumption of single dimensional traits, we employ a pair of classical results that show, under the appropriate rescaling in time, phenotypic space and population density, a sequence of rescaled BBM converges to a limiting process that can be characterized by a SPDE [@Mlard1993; @zeng1998absolute].
-The limiting processes of rescaled BBM have been referred to as measure-valued Markov processes [@dawson1993measure] or superprocesses [@alisonetheridge2000].
-Under the simplifying assumptions inherited from our treatment of deterministic dynamics and the additional assumption that the magnitude of demographic stochasticity is independent of trait values, we obtain as a special case a relatively simple expression for an SPDE that generalizes DAGA.
-The simplicity of our special case allows us to use properties of space-time white noise processes to derive a set of SDE that generalize equations (\ref{N_det}), (\ref{detmean}) and (\ref{vardyn_det}) to include the effects of demographic stochasticity.
-Classical expressions for the effects of random genetic drift on the evolution of mean traits are obtained as a special case.
-
-In the following section we provide the necessary mathematical tools needed to derive SDE from SPDE.
-Since our aim is to present this material to as wide of audience as possible, our approach deviates from standard definitions to remove the need for a detailed technical treatment.
-In addition to the notions of Reimann integration and partial differentiation already employed, the reader will only need some elementary probability and an intuitive understanding of SDE, including Brownian motion.
-Because space-time white noise, denoted by $\dot W(x,t)$, appears in the SPDE characterizing diffusion limits of BBM, we begin by defining $\dot W(x,t)$ and illustrating its relevant properties including a set of heuristics for performing calculations.
-Treating only the simplist of cases, we then provide a brief review of BBM, their diffusion limits and the SPDE that characterize them.
-For those not interested in the white noise calculus or superprocesses and would rather jump straight into more biologically relevant results, we recommend skipping to \S\ref{equations}.
-
-## White noise calculus and superprocesses \label{wnc}
-
-### Introduction to white noise
-
-One can think of white noise as the static seen on old television sets or infinitely detailed random dust spread across both time and space.
-From a more mathematical, yet still informal perspective, white noise can be thought of as a stochastic process.
-That is, we can picture white noise as a collection of random variables indexed by time and possibly space.
-In relation to Brownian motion, denoted by $W$, white noise can be interpreted of as the derivative of Brownian motion with respect to time, denoted $\dot W$.
-Since Brownian motion can be thought to take infinitesimally small Gaussian distributed jumps at each time point, this leads to the conceptualization of white noise as a collection of Gaussian distributed random variables.
-Figure \ref{wn} illustrates realizations of this conceptualized white noise in one (left) and two (right) dimensions.
-
-```{r, include=F, eval=F}
-require("ggplot2")
-require("ggthemes")
-require("plot3D")
-require("gridExtra")
-
-num1 = 1e4
-num2 = 1e4
-
-xi_1 = rnorm(num1)
-xi_2 = rnorm(num2)
-
-x1 = seq(0,1,length.out = num1)
-x2 = seq(0,1,length.out = sqrt(num2))
-
-M = expand.grid(x=x2,y=x2)
-M$value = xi_2
-
-df_1 = data.frame(xi=xi_1,x=x1)
-
-wn1d = ggplot(df_1)+geom_point(aes(x=x,y=xi),size=0.05)+theme_minimal()+xlab("Time")+ylab("Value")
-wn2d = ggplot(M, aes(x, y, color = value))+scale_color_gradient2(low="black",mid="grey50", high="white")+geom_point(size=0.001, shape=22, show.legend = FALSE)+theme_minimal()+xlab("Time")+ylab("Space")
-
-wn = grid.arrange(wn1d,wn2d,nrow=1)
-
-ggsave("~/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/wn.png",wn,width=5,height=2.5)
-```
-
-```{r, echo=F, fig.align='center', fig.cap='\\label{wn}Approximations of sample paths of temporal white noise (left) and space-time white noise (right) with brightness scaled to value.'}
-knitr::include_graphics('wn.png')
-```
-
-However, it turns out that realizations of white noise do not exist as functions in the classical sense.
-Indeed, since Brownian motion is nowhere differentiable with respect to time, white noise cannot be formally understood as a time derivative.
-Thus our notation $\dot W$ is only meant to aid intuition and not be taken as formal.
-A formal understanding is possible by considering white noise as a _measure_-valued process [@Dawson1975; @Walsh] or as a _generalized_ process that acts on classically defined functions or stochastic processes to return either random variables or stochastic processes [@Krylov1981; @DaPrato2014].
-Since a measure-valued process can be defined from a generalized process and a generalized process can be defined from a measure-valued process, the distinction between the two is more or less a matter of perspective.
-However, we find the perspective of white-noise as a generalized process to be a more efficient route for developing heuristics to help with some routine calculations involved with deriving SDE from SPDE.
-Hence, the notion of a generalized process provides the general idea implemented here.
-Although the treatments of SPDE provided by Krylov and Rozovskii [-@Krylov1981] and Da Prato and Zabczyk [-@DaPrato2014] extend the theory of SDE to formally treat SPDE in a general and elegant fashion, they require the navigation of an enourmous amount of technical definitions and detailed proofs.
-To extract some particularly useful results from this theory relevant to our goal of synthesizing the stochastic dynamics of biological populations, we provide a streamlined approach by capitolizing on the solid ground these authors have established.
-For instance, instead of rigorously proving properties of white-noise, we simply define them to be so, taking solice in the fact that the technical details have been worked out elsewhere.
-In SM \S\ref{DPZ14} we show how our informal treatment is related to the rigorous treatment provided by Da Prato and Zabczyk [-@DaPrato2014].
-
-Before diving in, we shed a bit of light on the idea of a generalized process.
-A generalized process is the stochastic analog of a generalized function, such as the Dirac delta function $\delta$.
-Often one sees $\delta$ defined as a function satisfying the properties $\delta(x)=0$ for $x\neq0$ and $\int_{-\infty}^{+\infty}\delta(x)dx=1$.
-However, since there is no function that satisfies these properties, we refer to $\delta$ as a generalized function.
-An alternative definition of $\delta$ considers its action on classically defined functions $f$.
-In particular, $\delta(f)=f(0)$, which can be heuristically written as $\int_{-\infty}^{+\infty}f(x)\delta(x)dx=f(0)$.
-Similarly, other generalized functions can be defined by their action on classically defined functions.
-Then, just as a generalized function operates on classical functions to return a value, a generalized process acts on a set of functions (or processes) to return a random variable (or a classically defined stochastic process).
-For a brief primer on the theory of generalized functions, see the addendum to chapter 3 of Kolmogorov and Fomin [-@kolmogorov1999elements].
-
-### Definition and basic properties of white noise \label{wnc_intro}
-
-Throughout this section, we minimize notation by writing $\int_{\mathbb{R}}f(x)dx=\int_{-\infty}^{+\infty}f(x)dx$ and similarly $\int_Df(x)dx$ for the integral of $f$ over $D\subset\mathbb{R}$.
-We define $\mathscr{N}_2$ as the set of stochastic processes $f(x,t)$ that are continuous in $t$ and satisfy $\mathbb{E}\left(\int_0^t\int_\mathbb{R}f^2(x,s)dxds\right)<+\infty$ for each $t\geq0$.
-The operator $\mathbb{E}$ denotes expectation with respect to the underlying probability space.
-For each $t\geq0$ we set
+\end{equation} For the sake of space we relegate the derivation of
+\(d\sigma^2/dt\) to SM \S\ref{var_deriv}. In the absence of mutation
+equation (\ref{vardyn_det}) agrees with the result derived by Lion
+(2018) for discrete phenotypes. From a statistical perspective, if we
+think of \((x-\bar x)^2\) as a square error, then in analogy to the
+dynamics of the mean trait, we see that the response in \(\sigma^2\) to
+selection can be expressed as a covariance of fitness and square error,
+which is defined in analogy to \(\mathrm{Cov}_t(m(\nu,x),x)\). Just as
+for the evolution of \(\bar x(t)\), this covariance also creates change
+in \(\sigma^2\) that can either increase or decrease mean fitness
+\(\bar m\), depending on whether or not individual fitness depends on
+the distribution of traits across the entire population. The effect of
+selection on phenotypic variance can be positive or negative depending
+on whether selection is stabilizing or disruptive.
+
+In SM \S\ref{SDE_DERIV}, we extend these results to include the effects
+of demographic stochasticity. The key realization in this direction is
+that a rigorous derivation of a SPDE that naturally generalizes DAGA and
+accounts for ecological interactions has been provided by Li (1998). To
+understand how this SPDE was derived, we provide an informal discussion
+of measure-valued branching processes (which treat populations as sets
+of discrete individuals) and their diffusion limits, the so-called
+superprocesses (which treat population size as a continuously varying
+number). Similar to the approach taken by Champagnat, Ferriére and
+Méléard (2006), we begin with a BBM that models populations as discrete
+sets of reproducing individuals whose vital rates depend on their trait
+value as well as the state of the entire population. Taking a large
+population size limit and keeping our assumption of single dimensional
+traits, we employ a pair of classical results that show, under the
+appropriate rescaling in time, phenotypic space and population density,
+a sequence of rescaled BBM converges to a limiting process that can be
+characterized by a SPDE (Méléard and Roelly 1993; Li 1998). The limiting
+processes of rescaled BBM have been referred to as measure-valued Markov
+processes (Dawson 1993) or superprocesses (Etheridge 2000). Under the
+simplifying assumptions inherited from our treatment of deterministic
+dynamics and the additional assumption that the magnitude of demographic
+stochasticity is independent of trait values, we obtain as a special
+case a relatively simple expression for an SPDE that generalizes DAGA.
+The simplicity of our special case allows us to use properties of
+space-time white noise processes to derive a set of SDE that generalize
+equations (\ref{N_det}), (\ref{detmean}) and (\ref{vardyn_det}) to
+include the effects of demographic stochasticity. Classical expressions
+for the effects of random genetic drift on the evolution of mean traits
+are obtained as a special case.
+
+In the following section we provide the necessary mathematical tools
+needed to derive SDE from SPDE. Since our aim is to present this
+material to as wide of audience as possible, our approach deviates from
+standard definitions to remove the need for a detailed technical
+treatment. In addition to the notions of Reimann integration and partial
+differentiation already employed, the reader will only need some
+elementary probability and an intuitive understanding of SDE, including
+Brownian motion. Because space-time white noise, denoted by
+\(\dot W(x,t)\), appears in the SPDE characterizing diffusion limits of
+BBM, we begin by defining \(\dot W(x,t)\) and illustrating its relevant
+properties including a set of heuristics for performing calculations.
+Treating only the simplist of cases, we then provide a brief review of
+BBM, their diffusion limits and the SPDE that characterize them. For
+those not interested in the white noise calculus or superprocesses and
+would rather jump straight into more biologically relevant results, we
+recommend skipping to \S\ref{equations}.
+
+\hypertarget{white-noise-calculus-and-superprocesses}{%
+\subsection{\texorpdfstring{White noise calculus and superprocesses
+\label{wnc}}{White noise calculus and superprocesses }}\label{white-noise-calculus-and-superprocesses}}
+
+\hypertarget{introduction-to-white-noise}{%
+\subsubsection{Introduction to white
+noise}\label{introduction-to-white-noise}}
+
+One can think of white noise as the static seen on old television sets
+or infinitely detailed random dust spread across both time and space.
+From a more mathematical, yet still informal perspective, white noise
+can be thought of as a stochastic process. That is, we can picture white
+noise as a collection of random variables indexed by time and possibly
+space. In relation to Brownian motion, denoted by \(W\), white noise can
+be interpreted of as the derivative of Brownian motion with respect to
+time, denoted \(\dot W\). Since Brownian motion can be thought to take
+infinitesimally small Gaussian distributed jumps at each time point,
+this leads to the conceptualization of white noise as a collection of
+Gaussian distributed random variables. Figure \ref{wn} illustrates
+realizations of this conceptualized white noise in one (left) and two
+(right) dimensions.
+
+\begin{figure}
+
+{\centering \includegraphics{wn}
+
+}
+
+\caption{\label{wn}Approximations of sample paths of temporal white noise (left) and space-time white noise (right) with brightness scaled to value.}\label{fig:unnamed-chunk-2}
+\end{figure}
+
+However, it turns out that realizations of white noise do not exist as
+functions in the classical sense. Indeed, since Brownian motion is
+nowhere differentiable with respect to time, white noise cannot be
+formally understood as a time derivative. Thus our notation \(\dot W\)
+is only meant to aid intuition and not be taken as formal. A formal
+understanding is possible by considering white noise as a
+\emph{measure}-valued process (Dawson 1975; Walsh 1986) or as a
+\emph{generalized} process that acts on classically defined functions or
+stochastic processes to return either random variables or stochastic
+processes (Krylov and Rozovskii 1981; Da Prato and Zabczyk 2014). Since
+a measure-valued process can be defined from a generalized process and a
+generalized process can be defined from a measure-valued process, the
+distinction between the two is more or less a matter of perspective.
+However, we find the perspective of white-noise as a generalized process
+to be a more efficient route for developing heuristics to help with some
+routine calculations involved with deriving SDE from SPDE. Hence, the
+notion of a generalized process provides the general idea implemented
+here. Although the treatments of SPDE provided by Krylov and Rozovskii
+(1981) and Da Prato and Zabczyk (2014) extend the theory of SDE to
+formally treat SPDE in a general and elegant fashion, they require the
+navigation of an enourmous amount of technical definitions and detailed
+proofs. To extract some particularly useful results from this theory
+relevant to our goal of synthesizing the stochastic dynamics of
+biological populations, we provide a streamlined approach by
+capitolizing on the solid ground these authors have established. For
+instance, instead of rigorously proving properties of white-noise, we
+simply define them to be so, taking solice in the fact that the
+technical details have been worked out elsewhere. In SM \S\ref{DPZ14} we
+show how our informal treatment is related to the rigorous treatment
+provided by Da Prato and Zabczyk (2014).
+
+Before diving in, we shed a bit of light on the idea of a generalized
+process. A generalized process is the stochastic analog of a generalized
+function, such as the Dirac delta function \(\delta\). Often one sees
+\(\delta\) defined as a function satisfying the properties
+\(\delta(x)=0\) for \(x\neq0\) and
+\(\int_{-\infty}^{+\infty}\delta(x)dx=1\). However, since there is no
+function that satisfies these properties, we refer to \(\delta\) as a
+generalized function. An alternative definition of \(\delta\) considers
+its action on classically defined functions \(f\). In particular,
+\(\delta(f)=f(0)\), which can be heuristically written as
+\(\int_{-\infty}^{+\infty}f(x)\delta(x)dx=f(0)\). Similarly, other
+generalized functions can be defined by their action on classically
+defined functions. Then, just as a generalized function operates on
+classical functions to return a value, a generalized process acts on a
+set of functions (or processes) to return a random variable (or a
+classically defined stochastic process). For a brief primer on the
+theory of generalized functions, see the addendum to chapter 3 of
+Kolmogorov and Fomin (1999).
+
+\hypertarget{definition-and-basic-properties-of-white-noise}{%
+\subsubsection{\texorpdfstring{Definition and basic properties of white
+noise
+\label{wnc_intro}}{Definition and basic properties of white noise }}\label{definition-and-basic-properties-of-white-noise}}
+
+Throughout this section, we minimize notation by writing
+\(\int_{\mathbb{R}}f(x)dx=\int_{-\infty}^{+\infty}f(x)dx\) and similarly
+\(\int_Df(x)dx\) for the integral of \(f\) over \(D\subset\mathbb{R}\).
+We define \(\mathscr{N}_2\) as the set of stochastic processes
+\(f(x,t)\) that are continuous in \(t\) and satisfy
+\(\mathbb{E}\left(\int_0^t\int_\mathbb{R}f^2(x,s)dxds\right)<+\infty\)
+for each \(t\geq0\). The operator \(\mathbb{E}\) denotes expectation
+with respect to the underlying probability space. For each \(t\geq0\) we
+set
 
@@
 
-and make use of the convention $f=g$ if $\|f-g\|_t=0$ for all $t\geq0$.
-Later, when we investigate stochastic dynamics of the abundance density $\nu(x,t)$, we will find $\sqrt\nu\in\mathscr{N}_2$.
-This enables us to utilize the heuristics developed in this section for the derivation of SDE describing the stochastic dynamics of $N(t)$, $\bar x(t)$ and $\sigma^2(t)$.
-
-We define a generalized stochastic process $\mathbf W$ that maps processes $f\in\mathscr{N}_2$ to real-valued stochastic processes indexed by time $t\geq0$, but not by space.
-To evaluate $\mathbf W$ for a process $f\in\mathscr{N}_2$ and some time $t\geq0$ we write $\mathbf W_t(f)$.
-Specifically, for any $f,g\in\mathscr{N}_2$, we define $\mathbf W(f)$ and $\mathbf W(g)$ to be Gaussian processes satisfying, for any $t,t_1,t_2\geq0$,
-\begin{equation}\label{exp_WN}
+and make use of the convention \(f=g\) if \(\|f-g\|_t=0\) for all
+\(t\geq0\). Later, when we investigate stochastic dynamics of the
+abundance density \(\nu(x,t)\), we will find
+\(\sqrt\nu\in\mathscr{N}_2\). This enables us to utilize the heuristics
+developed in this section for the derivation of SDE describing the
+stochastic dynamics of \(N(t)\), \(\bar x(t)\) and \(\sigma^2(t)\).
+
+We define a generalized stochastic process \(\mathbf W\) that maps
+processes \(f\in\mathscr{N}_2\) to real-valued stochastic processes
+indexed by time \(t\geq0\), but not by space. To evaluate \(\mathbf W\)
+for a process \(f\in\mathscr{N}_2\) and some time \(t\geq0\) we write
+\(\mathbf W_t(f)\). Specifically, for any \(f,g\in\mathscr{N}_2\), we
+define \(\mathbf W(f)\) and \(\mathbf W(g)\) to be Gaussian processes
+satisfying, for any \(t,t_1,t_2\geq0\), \begin{equation}\label{exp_WN}
 \mathbb{E}\big(\mathbf W_t(f)\big)=\mathbb{E}\big(\mathbf W_t(g)\big)=0,
-\end{equation}
-\begin{equation}\label{cov_WN}
+\end{equation} \begin{equation}\label{cov_WN}
 \mathbb{C}\big(\mathbf W_{t_1}(f),\mathbf W_{t_2}(g)\big)=\mathbb{E}\left(\int_0^{t_1\wedge t_2}\int_\mathbb{R} f(x,s)g(x,s)dxds\right),
-\end{equation}
-where $t_1\wedge t_2=\min(t_1,t_2)$ and $\mathbb{C}$ denotes covariance with respect to the underlying probability space.
-In particular, denoting $\mathbb{V}$ the variance operator with respect to the underlying probability space, we have $\mathbb{V}\big(\mathbf W_t(f)\big)=\|f\|_t^2$ for all $t\geq0$ and $f\in\mathscr{N}_2$.
-The operators $\mathbb{E}$ and $\mathbb{C}$ are to be distinguished from $\bar f(t)$ and $\mathrm{Cov}_t(f,g)$ which denote expectation and covariance with respect to phenotypic diversity at time $t\geq0$.
-
-Since Gaussian processes are characterized by their expectations and covariances and since we assume the $\mathscr{N}_2$ processes are continuous in time, the processes $\mathbf W(f)$ and $\mathbf W(g)$ are well defined.
-As an example, if $f\in\mathscr{N}_2$ is independent of time, then $\mathbf W(f)$ is a Brownian motion with variance at time $t\geq0$ equal to $\|f\|_t^2=t \ \mathbb{E}(\int_\mathbb{R}f^2(x,0)dx)$.
-With the generalized process $\mathbf W$ defined, we define the space-time white noise $\dot W(x,t)$ implicitly via the stochastic integral
+\end{equation} where \(t_1\wedge t_2=\min(t_1,t_2)\) and \(\mathbb{C}\)
+denotes covariance with respect to the underlying probability space. In
+particular, denoting \(\mathbb{V}\) the variance operator with respect
+to the underlying probability space, we have
+\(\mathbb{V}\big(\mathbf W_t(f)\big)=\|f\|_t^2\) for all \(t\geq0\) and
+\(f\in\mathscr{N}_2\). The operators \(\mathbb{E}\) and \(\mathbb{C}\)
+are to be distinguished from \(\bar f(t)\) and \(\mathrm{Cov}_t(f,g)\)
+which denote expectation and covariance with respect to phenotypic
+diversity at time \(t\geq0\).
+
+Since Gaussian processes are characterized by their expectations and
+covariances and since we assume the \(\mathscr{N}_2\) processes are
+continuous in time, the processes \(\mathbf W(f)\) and \(\mathbf W(g)\)
+are well defined. As an example, if \(f\in\mathscr{N}_2\) is independent
+of time, then \(\mathbf W(f)\) is a Brownian motion with variance at
+time \(t\geq0\) equal to
+\(\|f\|_t^2=t \ \mathbb{E}(\int_\mathbb{R}f^2(x,0)dx)\). With the
+generalized process \(\mathbf W\) defined, we define the space-time
+white noise \(\dot W(x,t)\) implicitly via the stochastic integral
 \begin{equation}\label{informal}
 ``\int_0^t\int_\mathbb{R}f(x,s)\dot W(x,s)dxds"=\mathbf W_t(f), \ \forall \ f\in\mathscr{N}_2, \ t\geq0.
-\end{equation}
-We place quotations in the above expression to emphasize its informal nature and that it should not be confused with classical Riemann integration.
-Following this definition of white noise, we compute its value by sampling it using $\mathscr{N}_2$ processes.
-For example, integrating white noise over a region $D\times[0,t]$, with $t>0$ and $D$ a bounded subset of $\mathbb{R}$, is equivalent to evaluating $\mathbf W_t(I_{D\times [0,+\infty)})$ for the deterministic process
-\begin{equation}
+\end{equation} We place quotations in the above expression to emphasize
+its informal nature and that it should not be confused with classical
+Riemann integration. Following this definition of white noise, we
+compute its value by sampling it using \(\mathscr{N}_2\) processes. For
+example, integrating white noise over a region \(D\times[0,t]\), with
+\(t>0\) and \(D\) a bounded subset of \(\mathbb{R}\), is equivalent to
+evaluating \(\mathbf W_t(I_{D\times [0,+\infty)})\) for the
+deterministic process \begin{equation}
 I_{D\times [0,+\infty)}(x,t)=\left\{\begin{matrix}
@@
 \end{matrix}\right..
-\end{equation}
-Since 
-\begin{equation}
+\end{equation} Since \begin{equation}
 \|I_{D\times[0,+\infty))}\|_t^2=\mathbb{E}\left(\int_0^t\int_\mathbb{R}I_{D\times [0,+\infty)}^2(x,s)dxds\right)=t\int_Ddx=t|D|<+\infty, 
-\end{equation}
-where $|D|$ denotes the length of $D$, we have $I_{D\times [0,+\infty)}\in\mathscr{N}_2$.
-Thus, using equations (\ref{exp_WN}) and (\ref{cov_WN}) and adopting the informal notation introduced in equation (\ref{informal}), we can write the following
+\end{equation} where \(|D|\) denotes the length of \(D\), we have
+\(I_{D\times [0,+\infty)}\in\mathscr{N}_2\). Thus, using equations
+(\ref{exp_WN}) and (\ref{cov_WN}) and adopting the informal notation
+introduced in equation (\ref{informal}), we can write the following
 \begin{equation}
 \mathbb{E}\left(\int_0^t\int_D\dot W(x,s)dxds\right)=0,
-\end{equation}
-\begin{equation}
+\end{equation} \begin{equation}
 \mathbb{V}\left(\int_0^t\int_D\dot W(x,s)dxds\right)=t|D|.
-\end{equation}
-Using this informal notation, equations (\ref{exp_WN}) and (\ref{cov_WN}) can be rewritten as
-\begin{equation}\label{exp_WN_xi}
+\end{equation} Using this informal notation, equations (\ref{exp_WN})
+and (\ref{cov_WN}) can be rewritten as \begin{equation}\label{exp_WN_xi}
 \mathbb{E}\left(\int_0^t\int_\mathbb{R}f(x,s)\dot W(x,s)dxds\right)=0,
-\end{equation}
-\begin{equation}\label{cov_WN_xi}
+\end{equation} \begin{equation}\label{cov_WN_xi}
 \mathbb{C}\left(\int_0^{t_1}\int_\mathbb{R}f(x,s)\dot W(x,s)dxds,\int_0^{t_2}\int_\mathbb{R}g(x,s)\dot W(x,s)dxds\right)
 =\int_0^{t_1\wedge t_2}\int_\mathbb{R}f(x,s)g(x,s)dxds.
-\end{equation}
-To relate these formulae to the common notation used for SDE, we write
-\begin{equation}
+\end{equation} To relate these formulae to the common notation used for
+SDE, we write \begin{equation}
 \hat f(x,t)=\frac{f(x,t)}{\sqrt{\int_\mathbb{R}f^2(y,t)dy}} \ \text{ and } \ 
 d\hat{\mathbf W}_t(f)=\left(\int_\mathbb{R}\hat f(x,t)\dot W(x,t)dx\right)dt
-\end{equation}
-so that
-\begin{equation}
+\end{equation} so that \begin{equation}
 \int_0^td\hat{\mathbf W}_s(f)=\int_0^t\int_\mathbb{R}\frac{f(x,s)}{\sqrt{\int_\mathbb{R}f^2(s,y)dy}}\dot W(x,s)dxds.
-\end{equation}
-This implies 
-\begin{equation}
+\end{equation} This implies \begin{equation}
 \mathbb{E}\left(\int_0^td\hat{\mathbf W}_s(f)\right)=0, \ \mathbb{C}\left(\int_0^{t_1}d\hat{\mathbf W}_s(f),\int_0^{t_2}d\hat{\mathbf W}_s(f)\right)=t_1\wedge t_2
-\end{equation}
-and in particular, as a function of $t$, $\int_0^td\hat{\mathbf W}_s(f)$ is a standard Brownian motion for any $f\in\mathscr{N}_2$.
-Hence, $d\hat{\mathbf W}_t(f)$ is analogous to the traditional shorthand used to denote stochastic differentials.
-Thus, equation (\ref{cov_WN_xi}) effectively extends It&ocirc;'s multiplication table to:
+\end{equation} and in particular, as a function of \(t\),
+\(\int_0^td\hat{\mathbf W}_s(f)\) is a standard Brownian motion for any
+\(f\in\mathscr{N}_2\). Hence, \(d\hat{\mathbf W}_t(f)\) is analogous to
+the traditional shorthand used to denote stochastic differentials. Thus,
+equation (\ref{cov_WN_xi}) effectively extends Itô's multiplication
+table to:
+
 \begin{table}[H]
@@
 \end{table}
-The extension of It&ocirc;'s multiplication table and properties of white noise outlined in this subsection provide a useful set of tools for working with SPDE.
-In SM \S\ref{SDE_DERIV} we employ these tools to derive SDE that track the dynamics of abundance, mean trait and phenotypic variance of a population from a particular SPDE.
-In the following subsection, we review how this particular SPDE naturally arises as the diffusion limit of a BBM.
-
-### From branching processes to SPDE \label{stochastic}
-
-In real populations individuals are born and potentially reproduce before they ultimately die.
-These three events provide the basic ingredients of a branching process.
-Mathematical investigations of such processes have a relatively deep history [@Kendall1966].
-The most simple branching process, known as the Galton-Watson process, describes the number of individuals alive at a given time $t\geq0$ as a non-negative integer [@Kimmel2015].
-Feller [-@feller1951] introduced a formal method to approximate branching processes with diffusion processes which are continuous in state (i.e., population size is approximated as a continuous quantity).
-Since diffusion processes possess greater analytical tractability than branching processes, Feller's method, known as the diffusion limit, has acquired immense popularity particularly in the field of mathematical population genetics [@Ewens2004].
-For over the past half of a century a great deal of accomplishments have been achieved in formalizing the diffusion limits of branching processes that describe populations of individuals occuring in some continuous space [@Watanabe1968; @Dawson1975; @Dawson1978; @Perkins1992; @Mlard1993; @edwinarendperkins1995; @zeng1998absolute; @Bertoin2003; @Etheridge2008; @Barton2019].
-This space can represent geographic space or, relevant to our context, phenotypic space.
-In the following subsection, we describe the BBM process, which is a particularly important branching process with spatial stucture.
-This process has been very useful in the study of SPDE due to its simplifying assumption that individuals do not interact.
-However, this assumption imposes an unfortunate restriction by precluding the modelling of ecological interactions.
-We therefore follow our discussion of BBM with a review of a few important results on spatially structured branching processes that account for interactions.
-
-__Branching Brownian motion__
-
-A BBM tracks individuals navigating $d$-dimensional Euclidean space that reproduce and senesce between exponentially distributed intervals.
-Unlike other stochastic processes that take values in $\mathbb{R}^d$, BBM takes values in the set of _non-negative finite measures_ over $\mathbb{R}^d$.
-Intuitively, one can think of a finite measure as a function that maps subsets of $\mathbb{R}^d$ to real numbers.
-To be formal, we only consider the Borel subsets of $\mathbb{R}^d$ corresponding to the Euclidean metric, but understanding this technicality is not crucial to our discussion.
-In particular, denoting $X_t$ a BBM, for a subset $D\subset\mathbb{R}^d$, $X_t(D)$ returns the (random) number of individuals alive within the region $D$ at time $t\geq0$.
-The BBM has three main components:
-
-  1) __Branching rate:__ In our formulation of BBM we assume Lifetimes of individuals are exponentially distributed with death rate $\lambda>0$ and reproduction occurs simultaneously with death.
-Biologically, this implies individuals are semelparous An alternative formulation treats birth and death events separately to model iteroparity.
-However, under the appropriate rescaling, both approaches converge to the same diffusion limit.
-We therefore choose the former approach for the sake of simplicity.
-
-  2) __Reproductive law:__ When a birth event occurs a random (possibly zero) number of offspring are left.
-The distribution of offspring left is called the reproductive law or branching mechanism.
-We denote the mean and variance in reproductive output by $\mathscr{W}$ and $V$ respectively.
-The case of $\mathscr{W}=1$ is referred to as the critical condition.
-Under the critical condition the probability that extinction occurs in finite time is equal to one.
-
-  3) __Spatial movement:__ Each offspring is born at the current location of their parent.
-Immediately after birth they move around space according to $d$-dimensional Brownian motion with diffusion parameter $\sqrt\mu$.
-In our context we interpret spatial movement as mutation so that the location of an individual at death represents the value of its phenotype.
-Then an individual born at location $x\in\mathbb{R}^d$ that lives for $\tau>0$ units of time will have a normally distributed trait centered on $x$ with covariance matrix equal to $\tau\mu$ times the $d\times d$ identity matrix.
-Hence, offspring inherit normally distributed traits centered on their parental trait.
-This fact creates a vital link to the deterministic dynamics reviewed above.
-Indeed, in the absence of selection, the deterministic PDE (\ref{eq1}) reduces to the $d=1$-dimensional Kolmogorov forward equation for a Brownian motion with diffusion parameter $\sqrt\mu$.
-
-To obtain a SPDE from a BBM we take a diffusion limit.
-There are several ways to do this, but a simple approach is to rescale the mass of individuals and time by $1/n$, diffusion by $\mu\to\mu/n$, branching rate by $\lambda\to n\lambda$, fitness by $\mathscr{W}\to\mathscr{W}^{1/n}$ and consider the limit as $n\to\infty$.
-Denoting the rescaled process by $X_t^{(n)}(D)$, the limiting process $\mathscr{X}_t=\lim_{n\to\infty}X_t^{(n)}$ is called a super-Brownian motion and is also a non-negative finite measure-valued process [@Watanabe1968].
-However, instead of returning the number of individuals alive in a region of space, super-Brownian motion returns the _mass_ of the population concentrated in a region of space.
-Since we have rescaled individual mass by $1/n$ and took the limit $n\to\infty$, individuals are no longer discrete units.
-Instead, the particle view of the population gets replaced by a blob spread across $d$-dimensional space.
-In particular, the value of $\mathscr{X}_t(D)$ is a continuously varying non-negative random variable for any $t\geq0$ and $D\subset\mathbb{R}^d$.
-
-Unfortunately, just as with cream cheese spread across too much toast, the blob perspective of the population may exhibit some complicated spatial discontinuities.
-However, for spatial dimension $d=1$, it turns out that $\mathscr{X}_t$ is absolutely continuous with respect to the Lebesgue measure for each $t\geq0$ [@Konno1988; @Reimers1989].
-This means that we can write $\mathscr{X}_t(D)=\int_D \nu(x,t)dx$ for some density process $\nu(x,t)$.
-Setting $\lambda=1$ and $m=\ln\mathscr{W}$ this density process satisfies the SPDE
-\begin{equation}\label{neutral_SPDE}
+
+The extension of Itô's multiplication table and properties of white
+noise outlined in this subsection provide a useful set of tools for
+working with SPDE. In SM \S\ref{SDE_DERIV} we employ these tools to
+derive SDE that track the dynamics of abundance, mean trait and
+phenotypic variance of a population from a particular SPDE. In the
+following subsection, we review how this particular SPDE naturally
+arises as the diffusion limit of a BBM.
+
+\hypertarget{from-branching-processes-to-spde}{%
+\subsubsection{\texorpdfstring{From branching processes to SPDE
+\label{stochastic}}{From branching processes to SPDE }}\label{from-branching-processes-to-spde}}
+
+In real populations individuals are born and potentially reproduce
+before they ultimately die. These three events provide the basic
+ingredients of a branching process. Mathematical investigations of such
+processes have a relatively deep history (Kendall 1966). The most simple
+branching process, known as the Galton-Watson process, describes the
+number of individuals alive at a given time \(t\geq0\) as a non-negative
+integer (Kimmel and Axelrod 2015). Feller (1951) introduced a formal
+method to approximate branching processes with diffusion processes which
+are continuous in state (i.e., population size is approximated as a
+continuous quantity). Since diffusion processes possess greater
+analytical tractability than branching processes, Feller's method, known
+as the diffusion limit, has acquired immense popularity particularly in
+the field of mathematical population genetics (Ewens 2004). For over the
+past half of a century a great deal of accomplishments have been
+achieved in formalizing the diffusion limits of branching processes that
+describe populations of individuals occuring in some continuous space
+(Watanabe 1968; Dawson 1975, 1978; Perkins 1992, 1995; Méléard and
+Roelly 1993; Li 1998; Bertoin and Le Gall 2003; Etheridge 2008; Barton
+and Etheridge 2019). This space can represent geographic space or,
+relevant to our context, phenotypic space. In the following subsection,
+we describe the BBM process, which is a particularly important branching
+process with spatial stucture. This process has been very useful in the
+study of SPDE due to its simplifying assumption that individuals do not
+interact. However, this assumption imposes an unfortunate restriction by
+precluding the modelling of ecological interactions. We therefore follow
+our discussion of BBM with a review of a few important results on
+spatially structured branching processes that account for interactions.
+
+\textbf{Branching Brownian motion}
+
+A BBM tracks individuals navigating \(d\)-dimensional Euclidean space
+that reproduce and senesce between exponentially distributed intervals.
+Unlike other stochastic processes that take values in \(\mathbb{R}^d\),
+BBM takes values in the set of \emph{non-negative finite measures} over
+\(\mathbb{R}^d\). Intuitively, one can think of a finite measure as a
+function that maps subsets of \(\mathbb{R}^d\) to real numbers. To be
+formal, we only consider the Borel subsets of \(\mathbb{R}^d\)
+corresponding to the Euclidean metric, but understanding this
+technicality is not crucial to our discussion. In particular, denoting
+\(X_t\) a BBM, for a subset \(D\subset\mathbb{R}^d\), \(X_t(D)\) returns
+the (random) number of individuals alive within the region \(D\) at time
+\(t\geq0\). The BBM has three main components:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi})}
+\item
+  \textbf{Branching rate:} In our formulation of BBM we assume Lifetimes
+  of individuals are exponentially distributed with death rate
+  \(\lambda>0\) and reproduction occurs simultaneously with death.
+  Biologically, this implies individuals are semelparous An alternative
+  formulation treats birth and death events separately to model
+  iteroparity. However, under the appropriate rescaling, both approaches
+  converge to the same diffusion limit. We therefore choose the former
+  approach for the sake of simplicity.
+\item
+  \textbf{Reproductive law:} When a birth event occurs a random
+  (possibly zero) number of offspring are left. The distribution of
+  offspring left is called the reproductive law or branching mechanism.
+  We denote the mean and variance in reproductive output by
+  \(\mathscr{W}\) and \(V\) respectively. The case of \(\mathscr{W}=1\)
+  is referred to as the critical condition. Under the critical condition
+  the probability that extinction occurs in finite time is equal to one.
+\item
+  \textbf{Spatial movement:} Each offspring is born at the current
+  location of their parent. Immediately after birth they move around
+  space according to \(d\)-dimensional Brownian motion with diffusion
+  parameter \(\sqrt\mu\). In our context we interpret spatial movement
+  as mutation so that the location of an individual at death represents
+  the value of its phenotype. Then an individual born at location
+  \(x\in\mathbb{R}^d\) that lives for \(\tau>0\) units of time will have
+  a normally distributed trait centered on \(x\) with covariance matrix
+  equal to \(\tau\mu\) times the \(d\times d\) identity matrix. Hence,
+  offspring inherit normally distributed traits centered on their
+  parental trait. This fact creates a vital link to the deterministic
+  dynamics reviewed above. Indeed, in the absence of selection, the
+  deterministic PDE (\ref{eq1}) reduces to the \(d=1\)-dimensional
+  Kolmogorov forward equation for a Brownian motion with diffusion
+  parameter \(\sqrt\mu\).
+\end{enumerate}
+
+To obtain a SPDE from a BBM we take a diffusion limit. There are several
+ways to do this, but a simple approach is to rescale the mass of
+individuals and time by \(1/n\), diffusion by \(\mu\to\mu/n\), branching
+rate by \(\lambda\to n\lambda\), fitness by
+\(\mathscr{W}\to\mathscr{W}^{1/n}\) and consider the limit as
+\(n\to\infty\). Denoting the rescaled process by \(X_t^{(n)}(D)\), the
+limiting process \(\mathscr{X}_t=\lim_{n\to\infty}X_t^{(n)}\) is called
+a super-Brownian motion and is also a non-negative finite measure-valued
+process (Watanabe 1968). However, instead of returning the number of
+individuals alive in a region of space, super-Brownian motion returns
+the \emph{mass} of the population concentrated in a region of space.
+Since we have rescaled individual mass by \(1/n\) and took the limit
+\(n\to\infty\), individuals are no longer discrete units. Instead, the
+particle view of the population gets replaced by a blob spread across
+\(d\)-dimensional space. In particular, the value of
+\(\mathscr{X}_t(D)\) is a continuously varying non-negative random
+variable for any \(t\geq0\) and \(D\subset\mathbb{R}^d\).
+
+Unfortunately, just as with cream cheese spread across too much toast,
+the blob perspective of the population may exhibit some complicated
+spatial discontinuities. However, for spatial dimension \(d=1\), it
+turns out that \(\mathscr{X}_t\) is absolutely continuous with respect
+to the Lebesgue measure for each \(t\geq0\) (Konno and Shiga 1988;
+Reimers 1989). This means that we can write
+\(\mathscr{X}_t(D)=\int_D \nu(x,t)dx\) for some density process
+\(\nu(x,t)\). Setting \(\lambda=1\) and \(m=\ln\mathscr{W}\) this
+density process satisfies the SPDE \begin{equation}\label{neutral_SPDE}
 \frac{\partial}{\partial t}\nu(x,t)=m\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)+\sqrt{V\nu(x,t)}\dot W(x,t).
-\end{equation}
-Since $\nu(x,t)$ is not generally differentiable in $x$ or $t$, the derivatives in expression (\ref{neutral_SPDE}) are taken in the _weak_ sense [for examples, see @lawrenceevans2010].
-That is, to rigorously interpret SPDE (\ref{neutral_SPDE}), we integrate the solution $\nu(x,t)$ against functions $f\in C_b^2(\mathbb{R})$, where $C_b^2(\mathbb{R})$ is the set of bounded and twice continuously differentiable functions on $\mathbb{R}$.
-Hence, equation (\ref{neutral_SPDE}) is just an abbreviation for
+\end{equation} Since \(\nu(x,t)\) is not generally differentiable in
+\(x\) or \(t\), the derivatives in expression (\ref{neutral_SPDE}) are
+taken in the \emph{weak} sense (for examples, see Evans 2010). That is,
+to rigorously interpret SPDE (\ref{neutral_SPDE}), we integrate the
+solution \(\nu(x,t)\) against functions \(f\in C_b^2(\mathbb{R})\),
+where \(C_b^2(\mathbb{R})\) is the set of bounded and twice continuously
+differentiable functions on \(\mathbb{R}\). Hence, equation
+(\ref{neutral_SPDE}) is just an abbreviation for
 
@@
 +\int_0^t\int_\mathbb{R}f(x)\sqrt{V\nu(x,s)}\dot W(x,s)dxds, \ \forall \ f\in C_b^2(\mathbb{R}).
-\end{multline}
-This expression is referred to as the _mild_ solution of (\ref{neutral_SPDE}).
-For more on the general theory of SPDE see Walsh [-@Walsh] and Da Prato and Zabczyk [-@DaPrato2014].
-Note that since $\nu(x,t)$ is the density of a finite measure, it is integrable for each $t\geq0$.
-Thus, since for some $M>0$, $|f(x)|\leq M$ for every $x\in\mathbb{R}$, setting $\varphi(x,t)=f(x)\sqrt{V\nu(x,t)}$ implies $\varphi\in\mathscr{N}_2$.
-Hence, the white noise integral on the right-hand side of equation (\ref{neutral_int}) can be understood using the heuristics introduced above.
-Evaluating equation (\ref{neutral_int}) in the particular case of $f(x)\equiv1$ returns the total mass process, which we refer to as the total abundance $N(t)$.
-
-A convergence theorem for the diffusion limit of a generalization of BBM was established by Watanabe [-@Watanabe1968].
-Dawson [-@Dawson1975] suggested that, for spatial dimension $d=1$, this diffusion limit should admit a density process that satisfies a SPDE.
-Konna and Shiga [-@Konno1988] and Reimers [-@Reimers1989] independently proved Dawson's suggestion was indeed correct.
-The diffusion limit of this more general branching process (in arbitrary spatial dimension) is referred to as a Dawson-Watanabe superprocess [@alisonetheridge2000].
-Conditioning a Dawson-Watanabe superprocess to have constant mass returns a Fleming-Viot process [@Etheridge1991; @Perkins1991] which has been popular in studies of spatial population genetics.
-In particular, an extension of the Fleming-Viot process, known as the $\Lambda$-Fleming-Viot process, was introduced by Bertoin and Le Gall [-@Bertoin2003] and coined by Etheridge [-@Etheridge2008] where it was used to resolve some technical challenges in modelling isolation by distance [@Felsenstein1975; see also @Barton2013; and @Barton2019].
-Although this provides an impressive list of accomplishments, the Dawson-Watanabe superprocess falls short of our needs.
-In particular this process assumes individuals do not interact and thus precludes its ability to model nonlocal effects on the fitness of individuals, such as those produced via ecological interactions.
-However, this concern has been addressed, leading to constuctions of superprocesses that account for interactions among individuals.
-In the next subsection we summarize the main results in this area and introduce the SPDE that provides the basis for our approach to theoretical evolutionary ecology.
-
-__Interacting superprocesses__
-
-In the above subsection we reviewed the diffusion limit of an especially tractable measure-valued branching process.
-However, we found the simplicity of this process restricts us from modelling nonlocal effects on the fitness of individuals.
-Here, we discuss superprocesses that account for such effects.
-The existence of diffusion limits for a class of measure-valued branching processes involving nonlocal effects has been treated by Méléard and Roelly [-@meleard1992interacting; -@Mlard1993].
-The interactions can manifest as dependencies of the spatial movement or reproductive law of individuals on their position $x$ and the state of the whole population described by $X_t$.
-An important result of Méléard and Roelly [-@meleard1992interacting; -@Mlard1993] is a theorem that provides sufficient conditions to construct a sequence of rescaled measure-valued branching processes that converge to a generalization of the Dawson-Watanabe superprocess that includes interactions.
-The rescaling is analogous to that described above for non-interacting Dawson-Watanabe superprocesses, but now the reproductive law described by $\mathscr{W}(X_t,x)$ and $V(X_t,x)$, branching rate $\lambda(X_t,x)$ and diffusion parameter $\sqrt{\mu(X_t,x)}$ are allowed to depend on the whole population $X_t$ and individual location $x$.
-In Figure \ref{rescaled} we demonstrate this rescaling in discrete time for a population experiencing stabilizing selection and logistic growth.
-Since time is discretized, the process we simulate is formally a branching random walk.
-For further details on our simulation see SM \S\ref{numerical}.
-
-```{r, eval=F, echo=F}
-require(ggplot2)
-require(ggthemes)
-require(gridExtra)
-
-n_1 = read.csv("/home/bb/Gits/branching.brownian.motion.and.spde/n_1.csv")
-n_2 = read.csv("/home/bb/Gits/branching.brownian.motion.and.spde/n_2.csv")
-n_3 = read.csv("/home/bb/Gits/branching.brownian.motion.and.spde/n_3.csv")
-
-n1p = ggplot(n_1,aes(x=time,y=x))+geom_point(size=1.0,stroke=0,alpha=0.1)+theme_minimal()+ylab("Trait Value \n (n=1)")+xlab("")+xlim(0,50)+ylim(-5,5)+ggtitle("Rescalings of a branching random walk")+theme(plot.title = element_text(hjust = 0.5))
-n2p = ggplot(n_2,aes(x=time,y=x))+geom_point(size=0.5,stroke=0,alpha=0.1)+theme_minimal()+ylab("Trait Value \n (n=5)")+xlab("")+xlim(0,50)+ylim(-5,5)
-n3p = ggplot(n_3,aes(x=time,y=x))+geom_point(size=0.25,stroke=0,alpha=0.1)+theme_minimal()+ylab("Trait Value \n (n=20)")+xlab("Time")+xlim(0,50)+ylim(-5,5)
-
-rescaled_plots = grid.arrange(n1p,n2p,n3p,nrow=3)
-
-ggsave("~/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/rescaled_plots.png",rescaled_plots,device="png",width=6,height=6,units="in",dpi=350)
-```
-
-```{r, message=FALSE, echo=F, fig.align='center', fig.cap='\\label{rescaled}Rescaled sample paths of a branching random walk under stabilizing selection and logistic growth.
-The top plot displays a sample path without scaling $(n=1)$, the middle plot shows a sample path rescaled by $n=5$ and the bottom plot shows a sample path rescaled by $n=20$.'}
-knitr::include_graphics('rescaled_plots.png')
-```
-
-Interactions that manifest in the spatial movement can be used to model mutation bias and those manisfesting in the reproductive law can model density-dependent growth rates and frequency-dependent selection.
-Perkins [-@Perkins1992; -@edwinarendperkins1995] developed a theory of stochastic integration with respect to the so-called _Brownian trees_ to characterize interacting superprocesses and establish properties of existence and uniqueness.
-Li [-@zeng1998absolute] built directly off of the construction of Méléard and Roelly [-@meleard1992interacting; -@Mlard1993] to study properties of density processes associated with interacting superprocesses, arriving at a SPDE that forms the foundation of our approach.
-
-Recall, we use $\nu(x,t)$ to denote the density of a superprocess, given it exists.
-Assuming the interactions manifest only in the reproductive law and that spatial movement follows Brownian motion with diffusion parameter $\sqrt\mu$ independent of both $X_t$ and $x$, Li [-@zeng1998absolute] proved a result that implies the interacting superprocess on one dimensional trait space has a density $\nu(x,t)$ which is non-negative, integrable, continuous in time and space and satisfies the SPDE
-\begin{equation}\label{SPDE}
+\end{multline} This expression is referred to as the \emph{mild}
+solution of (\ref{neutral_SPDE}). For more on the general theory of SPDE
+see Walsh (1986) and Da Prato and Zabczyk (2014). Note that since
+\(\nu(x,t)\) is the density of a finite measure, it is integrable for
+each \(t\geq0\). Thus, since for some \(M>0\), \(|f(x)|\leq M\) for
+every \(x\in\mathbb{R}\), setting \(\varphi(x,t)=f(x)\sqrt{V\nu(x,t)}\)
+implies \(\varphi\in\mathscr{N}_2\). Hence, the white noise integral on
+the right-hand side of equation (\ref{neutral_int}) can be understood
+using the heuristics introduced above. Evaluating equation
+(\ref{neutral_int}) in the particular case of \(f(x)\equiv1\) returns
+the total mass process, which we refer to as the total abundance
+\(N(t)\).
+
+A convergence theorem for the diffusion limit of a generalization of BBM
+was established by Watanabe (1968). Dawson (1975) suggested that, for
+spatial dimension \(d=1\), this diffusion limit should admit a density
+process that satisfies a SPDE. Konna and Shiga (1988) and Reimers (1989)
+independently proved Dawson's suggestion was indeed correct. The
+diffusion limit of this more general branching process (in arbitrary
+spatial dimension) is referred to as a Dawson-Watanabe superprocess
+(Etheridge 2000). Conditioning a Dawson-Watanabe superprocess to have
+constant mass returns a Fleming-Viot process (Etheridge and March 1991;
+Perkins 1991) which has been popular in studies of spatial population
+genetics. In particular, an extension of the Fleming-Viot process, known
+as the \(\Lambda\)-Fleming-Viot process, was introduced by Bertoin and
+Le Gall (2003) and coined by Etheridge (2008) where it was used to
+resolve some technical challenges in modelling isolation by distance
+(Felsenstein 1975; see also Barton, Etheridge, and Véber 2013; and
+Barton and Etheridge 2019). Although this provides an impressive list of
+accomplishments, the Dawson-Watanabe superprocess falls short of our
+needs. In particular this process assumes individuals do not interact
+and thus precludes its ability to model nonlocal effects on the fitness
+of individuals, such as those produced via ecological interactions.
+However, this concern has been addressed, leading to constuctions of
+superprocesses that account for interactions among individuals. In the
+next subsection we summarize the main results in this area and introduce
+the SPDE that provides the basis for our approach to theoretical
+evolutionary ecology.
+
+\textbf{Interacting superprocesses}
+
+In the above subsection we reviewed the diffusion limit of an especially
+tractable measure-valued branching process. However, we found the
+simplicity of this process restricts us from modelling nonlocal effects
+on the fitness of individuals. Here, we discuss superprocesses that
+account for such effects. The existence of diffusion limits for a class
+of measure-valued branching processes involving nonlocal effects has
+been treated by Méléard and Roelly (1992, 1993). The interactions can
+manifest as dependencies of the spatial movement or reproductive law of
+individuals on their position \(x\) and the state of the whole
+population described by \(X_t\). An important result of Méléard and
+Roelly (1992, 1993) is a theorem that provides sufficient conditions to
+construct a sequence of rescaled measure-valued branching processes that
+converge to a generalization of the Dawson-Watanabe superprocess that
+includes interactions. The rescaling is analogous to that described
+above for non-interacting Dawson-Watanabe superprocesses, but now the
+reproductive law described by \(\mathscr{W}(X_t,x)\) and \(V(X_t,x)\),
+branching rate \(\lambda(X_t,x)\) and diffusion parameter
+\(\sqrt{\mu(X_t,x)}\) are allowed to depend on the whole population
+\(X_t\) and individual location \(x\). In Figure \ref{rescaled} we
+demonstrate this rescaling in discrete time for a population
+experiencing stabilizing selection and logistic growth. Since time is
+discretized, the process we simulate is formally a branching random
+walk. For further details on our simulation see SM \S\ref{numerical}.
+
+\begin{figure}
+
+{\centering \includegraphics{rescaled_plots}
+
+}
+
+\caption{\label{rescaled}Rescaled sample paths of a branching random walk under stabilizing selection and logistic growth.
+The top plot displays a sample path without scaling $(n=1)$, the middle plot shows a sample path rescaled by $n=5$ and the bottom plot shows a sample path rescaled by $n=20$.}\label{fig:unnamed-chunk-4}
+\end{figure}
+
+Interactions that manifest in the spatial movement can be used to model
+mutation bias and those manisfesting in the reproductive law can model
+density-dependent growth rates and frequency-dependent selection.
+Perkins (1992, 1995) developed a theory of stochastic integration with
+respect to the so-called \emph{Brownian trees} to characterize
+interacting superprocesses and establish properties of existence and
+uniqueness. Li (1998) built directly off of the construction of Méléard
+and Roelly (1992, 1993) to study properties of density processes
+associated with interacting superprocesses, arriving at a SPDE that
+forms the foundation of our approach.
+
+Recall, we use \(\nu(x,t)\) to denote the density of a superprocess,
+given it exists. Assuming the interactions manifest only in the
+reproductive law and that spatial movement follows Brownian motion with
+diffusion parameter \(\sqrt\mu\) independent of both \(X_t\) and \(x\),
+Li (1998) proved a result that implies the interacting superprocess on
+one dimensional trait space has a density \(\nu(x,t)\) which is
+non-negative, integrable, continuous in time and space and satisfies the
+SPDE \begin{equation}\label{SPDE}
 \frac{\partial}{\partial t}\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)+\sqrt{V\nu(x,t)}\dot W(x,t).
-\end{equation}
-As mentioned above, we refer to SPDE (\ref{SPDE}) as the Stochastic Asexual Gaussian allelic model with Abundance dynamics (abbreviated SAGA).
-Comparing equation (\ref{SPDE}) to equation (3.5) of Li [-@zeng1998absolute], our $m$ and $V$ correspond to Li's $b$ and $c$ respectively.
-It is important to note that, under the assumptions made in Méléard and Roelly [-@meleard1992interacting; -@Mlard1993] and Li [-@zeng1998absolute], equation (\ref{SPDE}) is only formal when $m(h,x)$ is bounded across all combinations of $h\geq0$ and $x\in\mathbb{R}$.
-However, recalling our condition $m(h,x)\leq r\in\mathbb{R}$, the growth rates we consider are only bounded above.
-Yet, in the proof of the construction of the interacting superprocess as the limit of rescaled branching diffusions, Méléard and Roelly [-@meleard1992interacting; -@Mlard1993] assumed $m(h,x)$ to be bounded to guarantee the total mass process will have finite mean and variance, for finite $t\geq0$.
-This allowed the authors to employ a tightness criterion for sequences of measures and show the rescaled processes converge to a superprocess with finite total mass.
-Li's [-@zeng1998absolute] result builds directly off of Méléard and Roelly's construction, inheriting the assumption of boundedness for $m(h,x)$.
-However, in Li [-@zeng1998absolute], the sufficiency of $m(h,x)$ being bounded above is even more clear since Li works explicitly with a common upperbound for both $m(h,x)$ and $V$.
-Hence, one can repeat the necessary proofs replacing the assumption that $m(h,x)$ is bounded with the assumption that $m(h,x)$ is merely bounded above to derive the same results.
-An alternative approach to establishing this result that allows for our relaxed assumption on the growth rate has carried out by Champagnat, Ferrière, and Méléard [-@Champagnat2006] using techniques developed by Evans and Perkins [-@Evans1994].
-
-With solutions to SPDE (\ref{SPDE}) well defined for any growth rate bounded above, we can calculate the total mass process $N(t)$ using the mild solution of (\ref{SPDE}) with $f(x)\equiv1\in C_b^2(\mathbb{R})$ (the symbol "$\equiv$" means equal to for every $x$).
-That is,
+\end{equation} As mentioned above, we refer to SPDE (\ref{SPDE}) as the
+Stochastic Asexual Gaussian allelic model with Abundance dynamics
+(abbreviated SAGA). Comparing equation (\ref{SPDE}) to equation (3.5) of
+Li (1998), our \(m\) and \(V\) correspond to Li's \(b\) and \(c\)
+respectively. It is important to note that, under the assumptions made
+in Méléard and Roelly (1992, 1993) and Li (1998), equation (\ref{SPDE})
+is only formal when \(m(h,x)\) is bounded across all combinations of
+\(h\geq0\) and \(x\in\mathbb{R}\). However, recalling our condition
+\(m(h,x)\leq r\in\mathbb{R}\), the growth rates we consider are only
+bounded above. Yet, in the proof of the construction of the interacting
+superprocess as the limit of rescaled branching diffusions, Méléard and
+Roelly (1992, 1993) assumed \(m(h,x)\) to be bounded to guarantee the
+total mass process will have finite mean and variance, for finite
+\(t\geq0\). This allowed the authors to employ a tightness criterion for
+sequences of measures and show the rescaled processes converge to a
+superprocess with finite total mass. Li's (1998) result builds directly
+off of Méléard and Roelly's construction, inheriting the assumption of
+boundedness for \(m(h,x)\). However, in Li (1998), the sufficiency of
+\(m(h,x)\) being bounded above is even more clear since Li works
+explicitly with a common upperbound for both \(m(h,x)\) and \(V\).
+Hence, one can repeat the necessary proofs replacing the assumption that
+\(m(h,x)\) is bounded with the assumption that \(m(h,x)\) is merely
+bounded above to derive the same results. An alternative approach to
+establishing this result that allows for our relaxed assumption on the
+growth rate has carried out by Champagnat, Ferrière, and Méléard (2006)
+using techniques developed by Evans and Perkins (1994).
+
+With solutions to SPDE (\ref{SPDE}) well defined for any growth rate
+bounded above, we can calculate the total mass process \(N(t)\) using
+the mild solution of (\ref{SPDE}) with
+\(f(x)\equiv1\in C_b^2(\mathbb{R})\) (the symbol ``\(\equiv\)'' means
+equal to for every \(x\)). That is,
 
@@
 =N(0)+\int_0^t\bar m(s)N(s)dt+\int_0^t\sqrt{VN(s)}d\hat{\mathbf{W}}_s(\sqrt{\nu(x,s)}),
-\end{multline}
-where
-\begin{equation}
+\end{multline} where \begin{equation}
 \bar m(t)=\frac{1}{N(t)}\int_\mathbb{R}m(\nu,x)\nu(x,t)dx,
-\end{equation}
-and
-\begin{equation}
+\end{equation} and \begin{equation}
 \int_0^td\hat{\mathbf{W}}_s(\sqrt{\nu(x,s)})=\int_0^t\int_\mathbb{R}\frac{\sqrt{\nu(x,s)}}{\sqrt{\int_\mathbb{R}\nu(x,s)dx}}\dot W(x,s)dxds.
-\end{equation}
-Setting $W_1(t)=\hat{\mathbf{W}}_t(\sqrt{\nu(x,t)})$, we can use traditional stochastic differential notation to write
+\end{equation} Setting \(W_1(t)=\hat{\mathbf{W}}_t(\sqrt{\nu(x,t)})\),
+we can use traditional stochastic differential notation to write
 \begin{equation}
 dN=\bar mNdt+\sqrt{VN}dW_1.
-\end{equation}
-To find the associated SDE for $\bar x(t)$ and $\sigma^2(t)$, we want to repeat the same approach for $f(x)=x,x^2$ and apply It&ocirc;'s lemma.
-However, for these cases $f\notin C^2_b(\mathbb{R})$ since $f$ will not be bounded.
-But if we can show $\int_\mathbb{R}(|x|+x^2+x^4)\nu(x,t)dx<+\infty$ for all $t>0$ given this condition is satisfied by $\nu(x,0)$, then we can apply the mild solution of (\ref{SPDE}) to derive SDE for $\bar x(t)$ and $\sigma^2(t)$.
-To illustrate, let us suppose this is the case.
-Setting $\tilde x(t)=\int_\mathbb{R}x\nu(x,t)dx$, we have
+\end{equation} To find the associated SDE for \(\bar x(t)\) and
+\(\sigma^2(t)\), we want to repeat the same approach for \(f(x)=x,x^2\)
+and apply Itô's lemma. However, for these cases
+\(f\notin C^2_b(\mathbb{R})\) since \(f\) will not be bounded. But if we
+can show \(\int_\mathbb{R}(|x|+x^2+x^4)\nu(x,t)dx<+\infty\) for all
+\(t>0\) given this condition is satisfied by \(\nu(x,0)\), then we can
+apply the mild solution of (\ref{SPDE}) to derive SDE for \(\bar x(t)\)
+and \(\sigma^2(t)\). To illustrate, let us suppose this is the case.
+Setting \(\tilde x(t)=\int_\mathbb{R}x\nu(x,t)dx\), we have
 \begin{equation}
 \tilde x(t)=\tilde x(0)+\int_0^t\int_\mathbb{R}\nu(x,s)m(\nu,x)x+x\sqrt{V\nu(x,s)}\dot W(x,s)dxds.
-\end{equation}
-Similarly, setting $\tilde{\tilde\sigma}^2(t)=\int_\mathbb{R}x^2\nu(x,t)dx$, we have
+\end{equation} Similarly, setting
+\(\tilde{\tilde\sigma}^2(t)=\int_\mathbb{R}x^2\nu(x,t)dx\), we have
 \begin{equation}
 \tilde{\tilde\sigma}^2(t)=\tilde{\tilde\sigma}^2(0)+\int_0^t\int_\mathbb{R}\nu(x,s)\left(m(\nu,x)x^2+\mu\right)+x^2\sqrt{V\nu(x,s)}\dot W(x,s)dxds.
-\end{equation}
-Since $\bar x(t)=\tilde x(t)/N(t)$ and $\sigma^2(t)=\tilde{\tilde\sigma}^2(t)/N(t)-\bar x^2(t)$, we can use It&ocirc;'s lemma to derive SDE for $\bar x(t)$ and $\sigma^2(t)$, which we perform in SM \S\ref{SDE_DERIV}.
-We make no attempt in finding sufficient conditions for $\int_\mathbb{R}(|x|+x^2+x^4)\nu(x,t)dx<+\infty$ to hold and hence make no general assertions about the existence or uniqueness of $\bar x(t)$ or $\sigma^2(t)$.
-Regardless, we will later assume $\nu(x,t)$ can be approximated by a Gaussian curve in $x$ for all $t\geq0$.
-This assumption implies $\int_\mathbb{R}|x|^n\nu(x,t)dx<+\infty$ for all $n\in\mathbb{N}$ and for all $t\geq0$.
-
-<!--
-
-__Superprocesses without densities__
-
-The superprocesses we consider here allow function-valued representations (i.e., densities such as $\nu(x,t)$) only in the case of spatial dimension $d=1$.
-However, if we are willing to let go of function-valued representations, we are free to study superprocesses for greater spatial dimensions.
-Dawson [-@dawson1993measure] provides an excellent resource for this subject.
-Futhermore, the literature on these more general processes includes models of ecological interactions including competition.
-We summarize a few results important to our model of diffuse coevolution here.
-
-Denoting $C^k(\mathbb{R}^d)$ the set of functions on $\mathbb{R}^d$ with continuous derivatives up to order $k$ and $\Delta=\sum_{i=1}^d\frac{\partial^2}{\partial x_i^2}$ the Laplacian operator on $C^2(\mathbb{R}^d)$, we define the operator $A:C^2(\mathbb{R}^d)\to C^0(\mathbb{R}^d)$ by
-\begin{equation}
-A=\frac{\mu}{2}\Delta+r-\frac{a}{2}(\theta-x)^2,
-\end{equation}
-for fixed $r,\theta\in\mathbb{R}$ and $\mu,a>0$.
-
--->
-
-## Equations of evolutionary and demographic dynamics \label{equations}
-
-In SM \S\ref{SDE_DERIV} we show SDE for $N(t)$, $\bar x(t)$ and $\sigma^2(t)$ can be expressed as
+\end{equation} Since \(\bar x(t)=\tilde x(t)/N(t)\) and
+\(\sigma^2(t)=\tilde{\tilde\sigma}^2(t)/N(t)-\bar x^2(t)\), we can use
+Itô's lemma to derive SDE for \(\bar x(t)\) and \(\sigma^2(t)\), which
+we perform in SM \S\ref{SDE_DERIV}. We make no attempt in finding
+sufficient conditions for
+\(\int_\mathbb{R}(|x|+x^2+x^4)\nu(x,t)dx<+\infty\) to hold and hence
+make no general assertions about the existence or uniqueness of
+\(\bar x(t)\) or \(\sigma^2(t)\). Regardless, we will later assume
+\(\nu(x,t)\) can be approximated by a Gaussian curve in \(x\) for all
+\(t\geq0\). This assumption implies
+\(\int_\mathbb{R}|x|^n\nu(x,t)dx<+\infty\) for all \(n\in\mathbb{N}\)
+and for all \(t\geq0\).
+
+\hypertarget{equations-of-evolutionary-and-demographic-dynamics}{%
+\subsection{\texorpdfstring{Equations of evolutionary and demographic
+dynamics
+\label{equations}}{Equations of evolutionary and demographic dynamics }}\label{equations-of-evolutionary-and-demographic-dynamics}}
+
+In SM \S\ref{SDE_DERIV} we show SDE for \(N(t)\), \(\bar x(t)\) and
+\(\sigma^2(t)\) can be expressed as
+
 \begin{subequations}\label{gen_eqns}
@@
 \end{subequations}
-where $W_1$, $W_2$ and $W_3$ are standard Brownian motions.
-We note that conditions on the growth rate $m$ to guarantee existence and uniqueness of solutions to (\ref{xbar_gen}) and (\ref{sig2_gen}) have yet to be investigated.
-However, our results on the deterministic PDE suggest that $m(h,x)$ bounded above and differentiable in both arguments is sufficient.
-Dividing by $dt$ one can interpret equations (\ref{gen_eqns}) as if they are ordinary differential equations, but this not technically rigorous since Brownian motion is nowhere differentiable with respect to time.
-In SM \S\ref{SDE_DERIV} we show that in general $W_1$ is independent of both $W_2$ and $W_3$, but $W_2$ and $W_3$ covary.
+
+where \(W_1\), \(W_2\) and \(W_3\) are standard Brownian motions. We
+note that conditions on the growth rate \(m\) to guarantee existence and
+uniqueness of solutions to (\ref{xbar_gen}) and (\ref{sig2_gen}) have
+yet to be investigated. However, our results on the deterministic PDE
+suggest that \(m(h,x)\) bounded above and differentiable in both
+arguments is sufficient. Dividing by \(dt\) one can interpret equations
+(\ref{gen_eqns}) as if they are ordinary differential equations, but
+this not technically rigorous since Brownian motion is nowhere
+differentiable with respect to time. In SM \S\ref{SDE_DERIV} we show
+that in general \(W_1\) is independent of both \(W_2\) and \(W_3\), but
+\(W_2\) and \(W_3\) covary.
 
 There is quite a bit we can learn from expressions (\ref{gen_eqns}).
-Firstly, setting $V=0$ recovers the deterministic dynamics derived in \S\ref{deterministic}.
-Alternatively, one can take $N(t)\to\infty$ to recover the deterministic dynamics for $\bar x(t)$ and $\sigma^2(t)$.
-Characteristically, we note the effect of demographic stochasticity on abundance grows with $\sqrt{N(t)}$.
-Hence, dividing by $N$, we find the effects of demographic stochasticity on the per-capita growth rate diminish with increased abundance.
-Relating the response to demographic stochasticity derived here to the effect of random genetic drift derived in classic quantitative genetic theory, if we set $\sigma^2(t)=\sigma^2$ and $N(t)=N$ constant with respect to time, then integrating the stochastic term in equation (\ref{xbar_gen}) over a single unit of time returns a normally distributed random variable with mean zero and variance equal to $V\sigma^2/N$.
-In particular, assuming perfect inheritance, when reproductive variance is unity ($V=1$) this random variable coincides with the effect of random genetic drift on the change in mean trait over a single generation derived using sampling arguments [@Lande1976].
-There is also an interesting connection with classical population genetics.
-A fundamental result from early population genetic theory is the expected reduction in diversity due to the chance loss of alleles in finite populations [@Fisher1923; @Wright97].
-This expected reduction in diversity due to random genetic drift is captured by the third term in the deterministic component of expression (\ref{sig2_gen}), particularly $-V\sigma^2(t)/N(t)$.
-The component of SDE (\ref{sig2_gen}) describing random fluctuations in $\sigma^2(t)$ is more complicated and is proportional to the root of the difference between the centralized fourth moment of $p(x,t)$ and $\sigma^4(t)$.
-
-These expressions can be used to investigate the dynamics of the mean and variance for general $\nu(x,t)$.
-However, in the next subsection we simplify these expressions by approximating $\nu(x,t)$ with a Gaussian curve.
-By assuming $\nu(x,t)$ is Gaussian for $t\geq0$, we guarantee the existence of $\bar x(t)$ and $\sigma^2(t)$ for all $t\geq0$.
-Furthermore, in SM \S\ref{SDE_DERIV} we show that under the Gaussian case $W_1,W_2$ and $W_3$ are independent.
-
-### Particular results assuming a Gaussian phenotypic distribution \label{particular}
-
-By assuming $\nu(x,t)$ can be approximated by a Gaussian curve for each $t\geq0$, expressions (\ref{N}), (\ref{xbar_gen}) and (\ref{sig2_gen}) transform into efficient tools for deriving the dynamics of populations given a fitness function $m(\nu,x)$.
-Gaussian phenotypic distributions are often obtained through Gaussian, exponential or weak selection approximations together with a simplified model of inheritance and random mating [@pmid17248993; @Turelli1984; @Turelli1986; @Turelli2017; @9780471986539].
-Alternatively, it has been shown that a Gaussian distribution can provide a reasonable approximation even when selection is strong and non-Gaussian [@pmid7851785].
-However, our approach adds an additional layer of difficulty.
-Even with Gaussian selection, the resulting solution to SPDE (\ref{SPDE}) will only be a Gaussian curve in expectation, assuming a Gaussian curve as the initial condition.
-Yet this difficulty is not as challenging as it may first appear.
-Indeed, since SPDE (\ref{SPDE}) can be derived as a diffusion limit we know that, under the appropriate assumptions on selection, genetic architecture and reproduction, the stochastic departure from a Gaussian curve is negligible when the ratio $V/N$ is small (i.e., when the variance in reproductive output is much smaller than the population size).
-In SM \S\ref{numerical} we demonstrate this result using numerical methods.
-Mathematically, this requirement restricts model parameters to regions that maintain large population sizes.
+Firstly, setting \(V=0\) recovers the deterministic dynamics derived in
+\S\ref{deterministic}. Alternatively, one can take \(N(t)\to\infty\) to
+recover the deterministic dynamics for \(\bar x(t)\) and
+\(\sigma^2(t)\). Characteristically, we note the effect of demographic
+stochasticity on abundance grows with \(\sqrt{N(t)}\). Hence, dividing
+by \(N\), we find the effects of demographic stochasticity on the
+per-capita growth rate diminish with increased abundance. Relating the
+response to demographic stochasticity derived here to the effect of
+random genetic drift derived in classic quantitative genetic theory, if
+we set \(\sigma^2(t)=\sigma^2\) and \(N(t)=N\) constant with respect to
+time, then integrating the stochastic term in equation (\ref{xbar_gen})
+over a single unit of time returns a normally distributed random
+variable with mean zero and variance equal to \(V\sigma^2/N\). In
+particular, assuming perfect inheritance, when reproductive variance is
+unity (\(V=1\)) this random variable coincides with the effect of random
+genetic drift on the change in mean trait over a single generation
+derived using sampling arguments (Lande 1976). There is also an
+interesting connection with classical population genetics. A fundamental
+result from early population genetic theory is the expected reduction in
+diversity due to the chance loss of alleles in finite populations
+(Fisher 1923; Wright 1931). This expected reduction in diversity due to
+random genetic drift is captured by the third term in the deterministic
+component of expression (\ref{sig2_gen}), particularly
+\(-V\sigma^2(t)/N(t)\). The component of SDE (\ref{sig2_gen}) describing
+random fluctuations in \(\sigma^2(t)\) is more complicated and is
+proportional to the root of the difference between the centralized
+fourth moment of \(p(x,t)\) and \(\sigma^4(t)\).
+
+These expressions can be used to investigate the dynamics of the mean
+and variance for general \(\nu(x,t)\). However, in the next subsection
+we simplify these expressions by approximating \(\nu(x,t)\) with a
+Gaussian curve. By assuming \(\nu(x,t)\) is Gaussian for \(t\geq0\), we
+guarantee the existence of \(\bar x(t)\) and \(\sigma^2(t)\) for all
+\(t\geq0\). Furthermore, in SM \S\ref{SDE_DERIV} we show that under the
+Gaussian case \(W_1,W_2\) and \(W_3\) are independent.
+
+\hypertarget{particular-results-assuming-a-gaussian-phenotypic-distribution}{%
+\subsubsection{\texorpdfstring{Particular results assuming a Gaussian
+phenotypic distribution
+\label{particular}}{Particular results assuming a Gaussian phenotypic distribution }}\label{particular-results-assuming-a-gaussian-phenotypic-distribution}}
+
+By assuming \(\nu(x,t)\) can be approximated by a Gaussian curve for
+each \(t\geq0\), expressions (\ref{N}), (\ref{xbar_gen}) and
+(\ref{sig2_gen}) transform into efficient tools for deriving the
+dynamics of populations given a fitness function \(m(\nu,x)\). Gaussian
+phenotypic distributions are often obtained through Gaussian,
+exponential or weak selection approximations together with a simplified
+model of inheritance and random mating (Lande 1980; Turelli 1984, 1986,
+2017; Bürger 2000). Alternatively, it has been shown that a Gaussian
+distribution can provide a reasonable approximation even when selection
+is strong and non-Gaussian (Turelli and Barton 1994). However, our
+approach adds an additional layer of difficulty. Even with Gaussian
+selection, the resulting solution to SPDE (\ref{SPDE}) will only be a
+Gaussian curve in expectation, assuming a Gaussian curve as the initial
+condition. Yet this difficulty is not as challenging as it may first
+appear. Indeed, since SPDE (\ref{SPDE}) can be derived as a diffusion
+limit we know that, under the appropriate assumptions on selection,
+genetic architecture and reproduction, the stochastic departure from a
+Gaussian curve is negligible when the ratio \(V/N\) is small (i.e., when
+the variance in reproductive output is much smaller than the population
+size). In SM \S\ref{numerical} we demonstrate this result using
+numerical methods. Mathematically, this requirement restricts model
+parameters to regions that maintain large population sizes.
 Biologically, this implies populations are not at risk of extinction.
-Hence, models developed in this framework are not suitable for studying colonization-extinction dynamics or evolutionary rescue.
-Allowing for these restrictions, we may safely assume that $\nu$ is approximately Gaussian and justify writing
-\begin{equation}
+Hence, models developed in this framework are not suitable for studying
+colonization-extinction dynamics or evolutionary rescue. Allowing for
+these restrictions, we may safely assume that \(\nu\) is approximately
+Gaussian and justify writing \begin{equation}
 \nu(x,t)=\frac{N(t)}{\sqrt{2\pi\sigma^2(t)}}\exp\left(-\frac{\big(x-\bar x(t)\big)^2}{2\sigma^2(t)}\right).
-\end{equation}
-Under this assumption we find in SM \S\ref{cov2deriv} the results (suppressing the dependency on $t$)
+\end{equation} Under this assumption we find in SM \S\ref{cov2deriv} the
+results (suppressing the dependency on \(t\))
 \begin{equation}\label{covxm}
 \mathrm{Cov}(x,m)=\sigma^2\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right),
-\end{equation}
-\begin{equation}
+\end{equation} \begin{equation}
 \mathrm{Cov}\Big((x-\bar x)^2,m\Big)=2\sigma^4\left(\frac{\partial\bar m}{\partial\sigma^2}-\overline{\frac{\partial m}{\partial\sigma^2}}\right)
-\end{equation}
-and $\overline{(x-\bar x)^4}=3\sigma^4$. Equation (\ref{covxm}) is the continuous time equivalent to equation (9) in Lande [-@Lande1976]. In particular, these results imply
+\end{equation} and \(\overline{(x-\bar x)^4}=3\sigma^4\). Equation
+(\ref{covxm}) is the continuous time equivalent to equation (9) in Lande
+(1976). In particular, these results imply
+
 \begin{subequations}\label{no_inher}
@@
 \end{subequations}
-These equations allow us to derive the response in trait mean and variance by taking derivatives of fitness, a much more straightforward operation than calculating a covariance for general phenotypic distributions.
-Note that in the above expressions, the partial derivatives of $\bar m$ represent frequency independent selection and the averaged partial derivatives of $m$ represent frequency dependent selection.
-This relationship has already been pointed out by Lande [-@Lande1976] for the evolution of the mean trait, but here we see an analogous relationship holds also for the evolution of trait variance.
-
-In the next subsection we generalize this result to the case when traits are imperfectly inherited.
-In this case, the phenotypic variance $\sigma^2$ is replaced by a genetic variance $G$.
-This genetic variance represents the component of the variance in expressed traits $\sigma^2$ explained by additive effects of different alleles among genetic loci encoding for the focal phenotype [@joanroughgarden1979; @bulmer1980; @michaellynch1998].
-It is therefore fitting that $G$ is referred to as the additive genetic variance.
-
-### The evolution of additive genetic variance \label{inheritance}
-
-To model imperfect heritability we consider the relationship between expressed phenotypes $x\in\mathbb{R}$ and associated genetic values $g\in\mathbb{R}$ known as _breeding values_.
-The breeding value of an individual is the sum of additive effects of the alleles carried by the individual on its expressed trait.
-Since our derivations of evolutionary equations are based on branching processes that assume asexually reproducing populations (\S\ref{stochastic}), the additive genetic variance $G$ is just the variance of breeding values in a population.
-For a detailed treatment of breeding values and additive genetic variances, see Bulmer [-@bulmer1980] and Lynch and Walsh [-@michaellynch1998].
-
-Our treatment of the relationship between breeding values and expressed traits follows classical quantitative genetic assumptions such as those used in the seminal paper by Lande [-@Lande1975] to investigate the maintenance of genetic variation.
-In particular, we assume that the expressed trait for any given individual is independent of environmental conditions and normally distributed around their breeding value with variance $\eta$.
-Hence, $\sigma^2=G+\eta$.
-In the case that all of the effects of alleles on an expressed trait are additive, $\eta$ is known as the _variance of environmental deviation_ [@Lande1975; @michaellynch1998].
-For a given breeding value, we denote the probability density of a randomly drawn expressed trait by $\psi(x,g)$.
-Hence,
+
+These equations allow us to derive the response in trait mean and
+variance by taking derivatives of fitness, a much more straightforward
+operation than calculating a covariance for general phenotypic
+distributions. Note that in the above expressions, the partial
+derivatives of \(\bar m\) represent frequency independent selection and
+the averaged partial derivatives of \(m\) represent frequency dependent
+selection. This relationship has already been pointed out by Lande
+(1976) for the evolution of the mean trait, but here we see an analogous
+relationship holds also for the evolution of trait variance.
+
+In the next subsection we generalize this result to the case when traits
+are imperfectly inherited. In this case, the phenotypic variance
+\(\sigma^2\) is replaced by a genetic variance \(G\). This genetic
+variance represents the component of the variance in expressed traits
+\(\sigma^2\) explained by additive effects of different alleles among
+genetic loci encoding for the focal phenotype (Roughgarden 1979; Bulmer
+1980; Lynch and Walsh 1998). It is therefore fitting that \(G\) is
+referred to as the additive genetic variance.
+
+\hypertarget{the-evolution-of-additive-genetic-variance}{%
+\subsubsection{\texorpdfstring{The evolution of additive genetic
+variance
+\label{inheritance}}{The evolution of additive genetic variance }}\label{the-evolution-of-additive-genetic-variance}}
+
+To model imperfect heritability we consider the relationship between
+expressed phenotypes \(x\in\mathbb{R}\) and associated genetic values
+\(g\in\mathbb{R}\) known as \emph{breeding values}. The breeding value
+of an individual is the sum of additive effects of the alleles carried
+by the individual on its expressed trait. Since our derivations of
+evolutionary equations are based on branching processes that assume
+asexually reproducing populations (\S\ref{stochastic}), the additive
+genetic variance \(G\) is just the variance of breeding values in a
+population. For a detailed treatment of breeding values and additive
+genetic variances, see Bulmer (1980) and Lynch and Walsh (1998).
+
+Our treatment of the relationship between breeding values and expressed
+traits follows classical quantitative genetic assumptions such as those
+used in the seminal paper by Lande (1975) to investigate the maintenance
+of genetic variation. In particular, we assume that the expressed trait
+for any given individual is independent of environmental conditions and
+normally distributed around their breeding value with variance \(\eta\).
+Hence, \(\sigma^2=G+\eta\). In the case that all of the effects of
+alleles on an expressed trait are additive, \(\eta\) is known as the
+\emph{variance of environmental deviation} (Lande 1975; Lynch and Walsh
+1998). For a given breeding value, we denote the probability density of
+a randomly drawn expressed trait by \(\psi(x,g)\). Hence,
 \begin{equation}
 \psi(x,g)=\frac{1}{\sqrt{2\pi\eta}}\exp\left(-\frac{(x-g)^2}{2\eta}\right).
-\end{equation}
-To include this relationship in our framework, we write $\rho(g,t)$ as the abundance density of breeding values at time $t$ so that $\int_{-\infty}^{+\infty}\rho(g,t)dg=\int_{-\infty}^{+\infty}\nu(x,t)dx=N(t)$.
-We switch our focus from directly modelling the evolution of $\nu(x,t)$ to modelling the evolution of $\rho(g,t)$.
-Once $\rho(g,t)$ is determined, we can compute $\nu(x,t)$ via
+\end{equation} To include this relationship in our framework, we write
+\(\rho(g,t)\) as the abundance density of breeding values at time \(t\)
+so that
+\(\int_{-\infty}^{+\infty}\rho(g,t)dg=\int_{-\infty}^{+\infty}\nu(x,t)dx=N(t)\).
+We switch our focus from directly modelling the evolution of
+\(\nu(x,t)\) to modelling the evolution of \(\rho(g,t)\). Once
+\(\rho(g,t)\) is determined, we can compute \(\nu(x,t)\) via
 \begin{equation}\label{nu_from_rho}
 \nu(x,t)=\int_{-\infty}^{+\infty}\rho(g,t)\psi(x,g)dg.
-\end{equation}
-However, since selection acts on expressed phenotypes, we use the assumed relationship between breeding values and expressed traits to calculate the fitness of breeding values.
-To motivate the approach taken, consider the problem of inferring the breeding value of an individual given its expressed trait $x$.
-Denote $\mathfrak{g}$ a random variable representing the unkown breeding value.
-Under this model of inheritance we know $x$ is a random sample from a normal distribution with mean $\mathfrak{g}$ and variance $\eta$.
-Maximizing likelihood suggests $x$ is our best guess for $\mathfrak{g}$, but the actual value of $\mathfrak{g}$ is normally distributed around $x$ with the variance $\eta$.
-Hence, for fixed $x$, we obtain $\psi(x,g)$ as the probability density of $\mathfrak{g}$.
-Thus, the mean fitness of a breeding value $g$ across all individuals carrying $g$ can be written as
+\end{equation} However, since selection acts on expressed phenotypes, we
+use the assumed relationship between breeding values and expressed
+traits to calculate the fitness of breeding values. To motivate the
+approach taken, consider the problem of inferring the breeding value of
+an individual given its expressed trait \(x\). Denote \(\mathfrak{g}\) a
+random variable representing the unkown breeding value. Under this model
+of inheritance we know \(x\) is a random sample from a normal
+distribution with mean \(\mathfrak{g}\) and variance \(\eta\).
+Maximizing likelihood suggests \(x\) is our best guess for
+\(\mathfrak{g}\), but the actual value of \(\mathfrak{g}\) is normally
+distributed around \(x\) with the variance \(\eta\). Hence, for fixed
+\(x\), we obtain \(\psi(x,g)\) as the probability density of
+\(\mathfrak{g}\). Thus, the mean fitness of a breeding value \(g\)
+across all individuals carrying \(g\) can be written as
 \begin{equation}\label{m_rel}
 m^*(\rho,g)=\int_{-\infty}^{+\infty}m(\nu,x)\psi(x,g)dx.
-\end{equation}
-This is similar to the approach taken by Kimura and Crow [-@Kimura1978] to calculate the overall effects of selection for expressed characters onto the changes in the distribution of alleles encoding those characters.
-However, instead of focusing on the frequencies of alleles at particular loci, our results focus on the densities of breeding values.
-With the relationship between $m(\nu,x)$ and $m^*(\rho,g)$ established, we define the evolution of $\rho(g,t)$ by the SPDE
-\begin{equation}\label{rho_SPDE}
+\end{equation} This is similar to the approach taken by Kimura and Crow
+(1978) to calculate the overall effects of selection for expressed
+characters onto the changes in the distribution of alleles encoding
+those characters. However, instead of focusing on the frequencies of
+alleles at particular loci, our results focus on the densities of
+breeding values. With the relationship between \(m(\nu,x)\) and
+\(m^*(\rho,g)\) established, we define the evolution of \(\rho(g,t)\) by
+the SPDE \begin{equation}\label{rho_SPDE}
 \dot\rho(g,t)=m^*(\rho,g)\rho(g,t)+\frac{\mu}{2}\frac{\partial^2}{\partial^2 g}\rho(g,t)+\sqrt{V\rho(g,t)}\dot W(g,t).
-\end{equation}
-Equation (\ref{rho_SPDE}) is a stochastic generalization of the deterministic PDE (\ref{eq1}) from \S\ref{deterministic}, but describes the evolution of the distribution of breeding values instead of expressed characters.
-However, whether modelling expressed characters or breeding values, we refer to SPDE of the form (\ref{rho_SPDE}) as Stochastic Asexual Gaussian allelic models with Abundance dynamics (abbreviated SAGA).
-By replacing $\psi(x,g)$ with different probability densities produces different developmental models.
-This includes densities that may depend on environmental variables and hence model phenotypic plasticy.
-To leave room for the naming of other models where $\psi(x,g)$ is Gaussian, we refer to the model of development we have chosen as the _centered_ Gaussian developmental model.
-In \S\ref{wnc} we review the origins of equation (\ref{rho_SPDE}) and provide some theory to help make sense of it, particularly the term $\dot W$.
-
-Assuming $\rho(g,t)$ is Gaussian implies its mode coincides with $\bar x$.
-Furthermore, since $\sigma^2=G+\eta$, we can use equation (\ref{m_rel}) and the chain rule from calculus (see SM \S\ref{fit2fit}) to find
+\end{equation} Equation (\ref{rho_SPDE}) is a stochastic generalization
+of the deterministic PDE (\ref{eq1}) from \S\ref{deterministic}, but
+describes the evolution of the distribution of breeding values instead
+of expressed characters. However, whether modelling expressed characters
+or breeding values, we refer to SPDE of the form (\ref{rho_SPDE}) as
+Stochastic Asexual Gaussian allelic models with Abundance dynamics
+(abbreviated SAGA). By replacing \(\psi(x,g)\) with different
+probability densities produces different developmental models. This
+includes densities that may depend on environmental variables and hence
+model phenotypic plasticy. To leave room for the naming of other models
+where \(\psi(x,g)\) is Gaussian, we refer to the model of development we
+have chosen as the \emph{centered} Gaussian developmental model. In
+\S\ref{wnc} we review the origins of equation (\ref{rho_SPDE}) and
+provide some theory to help make sense of it, particularly the term
+\(\dot W\).
+
+Assuming \(\rho(g,t)\) is Gaussian implies its mode coincides with
+\(\bar x\). Furthermore, since \(\sigma^2=G+\eta\), we can use equation
+(\ref{m_rel}) and the chain rule from calculus (see SM \S\ref{fit2fit})
+to find
+
 \begin{subequations}
@@
 \end{subequations}
+
 Thus, equations (\ref{no_inher}) become
+
 \begin{subequations}\label{inher}
@@
 
-From expressions (\ref{inher}) we see that, under this model of inheritance, focusing on additive genetic variance $G$ instead the variance in expressed traits $\sigma^2$ makes no structural changes to the basic equations describing the dynamics of populations.
-In the next section, we make use of these expressions to develop a model of diffuse coevolution in a guild of $S$ species competing along a resource continuum.
-
-# A model of diffuse coevolution \label{coev}
-
-## Formulation \label{form_coev}
-
-In this section we demonstrate the use of our framework by developing a model of diffuse coevolution across a guild of $S$ species whose interactions are mediated by resource competition along a single niche axis.
-Because our approach treats abundance dynamics and evolutionary dynamics simultaneously, this model allows us to investigate the relationship between selection gradients and competition coefficients, which we carry out in what follows.
-
-The dynamics of phenotypic distributions and abundances have been derived above and so the only task remaining is the formulation of a fitness function.
-Our approach mirrors closely the theory developed by MacArthur and Levins [-@Macarthur1967], Levins [-@9780691080628] and MacArthur [-@Arthur1969; -@MacArthur1970; -@9780691023823].
-The most significant difference, aside from allowing evolution to occur, is the treatment of resource quality, which we replace with a model of abiotic stabilizing selection.
-A derivation is provided in SM \S\ref{diffuse}.
-
-For species $i$ we inherit the above notation for trait value, distribution, average, variance, abundance, etc except with an $i$ in the subscript.
-Real world examples of niche axes include the body size of prey for lizard predators and the date of activity in a season for pollinators competing for floral resources.
-For mathematical convenience, we model the axis of resources by the real line $\mathbb{R}$.
-The value of a resouce along this axis is denoted by the symbol $\zeta$.
-For an individual in species $i$, we assume the resource utilization curve $u_i$ can be written as
+From expressions (\ref{inher}) we see that, under this model of
+inheritance, focusing on additive genetic variance \(G\) instead the
+variance in expressed traits \(\sigma^2\) makes no structural changes to
+the basic equations describing the dynamics of populations. In the next
+section, we make use of these expressions to develop a model of diffuse
+coevolution in a guild of \(S\) species competing along a resource
+continuum.
+
+\hypertarget{a-model-of-diffuse-coevolution}{%
+\section{\texorpdfstring{A model of diffuse coevolution
+\label{coev}}{A model of diffuse coevolution }}\label{a-model-of-diffuse-coevolution}}
+
+\hypertarget{formulation}{%
+\subsection{\texorpdfstring{Formulation
+\label{form_coev}}{Formulation }}\label{formulation}}
+
+In this section we demonstrate the use of our framework by developing a
+model of diffuse coevolution across a guild of \(S\) species whose
+interactions are mediated by resource competition along a single niche
+axis. Because our approach treats abundance dynamics and evolutionary
+dynamics simultaneously, this model allows us to investigate the
+relationship between selection gradients and competition coefficients,
+which we carry out in what follows.
+
+The dynamics of phenotypic distributions and abundances have been
+derived above and so the only task remaining is the formulation of a
+fitness function. Our approach mirrors closely the theory developed by
+MacArthur and Levins (1967), Levins (1968) and MacArthur (1969, 1970,
+1972). The most significant difference, aside from allowing evolution to
+occur, is the treatment of resource quality, which we replace with a
+model of abiotic stabilizing selection. A derivation is provided in SM
+\S\ref{diffuse}.
+
+For species \(i\) we inherit the above notation for trait value,
+distribution, average, variance, abundance, etc except with an \(i\) in
+the subscript. Real world examples of niche axes include the body size
+of prey for lizard predators and the date of activity in a season for
+pollinators competing for floral resources. For mathematical
+convenience, we model the axis of resources by the real line
+\(\mathbb{R}\). The value of a resouce along this axis is denoted by the
+symbol \(\zeta\). For an individual in species \(i\), we assume the
+resource utilization curve \(u_i\) can be written as
 
@@
 
-We further assume the niche center $x_i$ is normally distributed among individuals in species $i$, but the niche breadth $w_i$ and total niche utilization $U_i$ are constant across individuals in species $i$ and therefore cannot evolve.
-Suppose $\theta_i\in\mathbb R$ is the optimal location along the niche axis for species $i$ such that, in the absence of competition, individuals leave on average $Q_i$ offspring when concentrated at $\theta_i$.
-We capture the rate by which the fitness falls as niche location $\zeta$ leaves the optimum $\theta_i$ by the parameter $A_i\geq0$.
-Hence, abiotic stabilizing selection along the resource axis can be modelled by the curve
+We further assume the niche center \(x_i\) is normally distributed among
+individuals in species \(i\), but the niche breadth \(w_i\) and total
+niche utilization \(U_i\) are constant across individuals in species
+\(i\) and therefore cannot evolve. Suppose \(\theta_i\in\mathbb R\) is
+the optimal location along the niche axis for species \(i\) such that,
+in the absence of competition, individuals leave on average \(Q_i\)
+offspring when concentrated at \(\theta_i\). We capture the rate by
+which the fitness falls as niche location \(\zeta\) leaves the optimum
+\(\theta_i\) by the parameter \(A_i\geq0\). Hence, abiotic stabilizing
+selection along the resource axis can be modelled by the curve
 
@@
 
-The effect of abiotic stabilizing selection on the fitness for an individual of species $i$ with niche location $x_i$ is then given by
+The effect of abiotic stabilizing selection on the fitness for an
+individual of species \(i\) with niche location \(x_i\) is then given by
 
@@
 
-To determine the potential for competition between individuals with niche locations $x_i$ and $x_j$, belonging to species $i$ and $j$ respectively, we compute the niche overlap
+To determine the potential for competition between individuals with
+niche locations \(x_i\) and \(x_j\), belonging to species \(i\) and
+\(j\) respectively, we compute the niche overlap
 
@@
 
-A notable criticism of using niche overlap to measure the intensity of competition points to cases where populations competiting on multiple niche axes exhibit overlap on at least one of the axes, but no overall niche overlap [@Holt1987].
-Thus niche overlap on lower-dimensional projections of some multivariate niche space does not imply the populations compete.
-To illustrate with a simple example, consider two populations competing for space on the plane $\mathbb{R}^2$.
-If the spatial distributions of the two populations overlap, then they will overlap on both spatial axes.
-However, if the populations do not overlap on at least one of the spatial axes, they will have no overall spatial overlap.
-Furthermore, even if the species overlap on both spatial axes, they need not have any overall spatial overlap.
-This final result corresponds to the fact that components of niche space do not necessarily interact multiplicatively to determine the consequences for the intensity of competition.
-In another component of Holt's [-@Holt1987] critique, an argument is made for the potential of competition occuring without any overlap in niche space.
-However, this argument is based on the practical difficulty of identifying every resource axis populations are competing on and how these axes interact to determine fitness consequences.
-Our model avoids these caveats by assuming competition only occurs along a single dimensional resource gradient.
-
-To map the degree of niche overlap to fitness, we assume competition between individuals with niche locations $x_i$ and $x_j$ additively decreases the Malthusian fitness for the individual in species $i$ by $c_i\mathcal{O}_{ij}(x_i-x_j)$ for some $c_i\geq0$.
-We refer to $c_i$ as the sensitivity to competition for species $i$.
-The term $c_i\mathcal{O}_{ij}(x_i-x_j)$ coincides with a special case of a term used to capture competition in Dawson's geostochastic logistic model, an SPDE model developed to study the combined effects of demographic stochasticity, spatial dispersion and locally finite carrying capacity [@Dawson1978].
-In relation to the example fitness function discussed in \S\ref{deterministic}, consider $\kappa(x_i-x_j)=\mathcal{O}_{ij}(x_i-x_j)$.
-In SM \S\ref{diffuse} we combine our treatment of resource competition with equations (\ref{N}), (\ref{xbarfinal}) and (\ref{Gfinal}) to find
+A notable criticism of using niche overlap to measure the intensity of
+competition points to cases where populations competiting on multiple
+niche axes exhibit overlap on at least one of the axes, but no overall
+niche overlap (Holt 1987). Thus niche overlap on lower-dimensional
+projections of some multivariate niche space does not imply the
+populations compete. To illustrate with a simple example, consider two
+populations competing for space on the plane \(\mathbb{R}^2\). If the
+spatial distributions of the two populations overlap, then they will
+overlap on both spatial axes. However, if the populations do not overlap
+on at least one of the spatial axes, they will have no overall spatial
+overlap. Furthermore, even if the species overlap on both spatial axes,
+they need not have any overall spatial overlap. This final result
+corresponds to the fact that components of niche space do not
+necessarily interact multiplicatively to determine the consequences for
+the intensity of competition. In another component of Holt's (1987)
+critique, an argument is made for the potential of competition occuring
+without any overlap in niche space. However, this argument is based on
+the practical difficulty of identifying every resource axis populations
+are competing on and how these axes interact to determine fitness
+consequences. Our model avoids these caveats by assuming competition
+only occurs along a single dimensional resource gradient.
+
+To map the degree of niche overlap to fitness, we assume competition
+between individuals with niche locations \(x_i\) and \(x_j\) additively
+decreases the Malthusian fitness for the individual in species \(i\) by
+\(c_i\mathcal{O}_{ij}(x_i-x_j)\) for some \(c_i\geq0\). We refer to
+\(c_i\) as the sensitivity to competition for species \(i\). The term
+\(c_i\mathcal{O}_{ij}(x_i-x_j)\) coincides with a special case of a term
+used to capture competition in Dawson's geostochastic logistic model, an
+SPDE model developed to study the combined effects of demographic
+stochasticity, spatial dispersion and locally finite carrying capacity
+(Dawson 1978). In relation to the example fitness function discussed in
+\S\ref{deterministic}, consider
+\(\kappa(x_i-x_j)=\mathcal{O}_{ij}(x_i-x_j)\). In SM \S\ref{diffuse} we
+combine our treatment of resource competition with equations (\ref{N}),
+(\ref{xbarfinal}) and (\ref{Gfinal}) to find
 
@@
 
-Despite the convoluted appearance of system (\ref{comm_dynamics}), there are some nice features that reflect biological reasoning.
-For example, the dynamics of abundance are just a generalization of Lotka-Volterra dynamics.
-In particular, the effect of competition with species $j$ on the fitness of species $i$ grows linearly with $N_j$.
-However, as biotic selection pushes $\bar x_i$ away from $\bar x_j$, the effect of competition with species $j$ on the fitness of species $i$ rapidly diminishes, reflecting a reduction in niche overlap.
-The divergence of $\bar x_i$ and $\bar x_j$ due to competition is referred to in the community ecology literature as character displacement.
-We also see that the fitness of species $i$ drops quadratically with the difference between $\bar x_i$ and the abiotic optimum $\theta_i$.
-Hence, abiotic selection acts to pull $\bar x_i$ towards $\theta_i$.
-The response in mean trait $\bar x_i$ to natural selection is proportional to the amount of heritable variation in the population, represented by the additive genetic variance $G_i$.
-However, we have that $G_i$ is itself a dynamic quantity.
-Under our model, abiotic stabilizing selection erodes away heritable variation at a rate that is independent of both $N_i$ and $\bar x_i$.
-The effect of competition on $G_i$ is a bit more complicated.
-When $b_{ij}(\bar x_i-\bar x_j)^2<1$, competition with species $j$ acts as diversifying selection which tends to increase the amount of heritable variation.
-However, when $b_{ij}(\bar x_i-\bar x_j)^2>1$, competition with species $j$ acts as directional selection and reduces $G_i$.
-In the following subsections we demonstrate the behavior of system (\ref{comm_dynamics}) by plotting numerical solutions and investigate implications for the relationship between the strength of ecological interactions and selection.
-
-## Community dynamics \label{dynamics}
-
-For the sake of illustration we numerically integrated system (\ref{comm_dynamics}) for a richness of $S=100$ species.
-We assumed homogeneous model parameters across species in the community as summarized by Table \ref{par_vals}.
-We repeated numerical integration under the two scenarios of weak and strong competition.
-For the first scenario of weak competition we set $c=1.0\times10^{-7}$ and for the second scenario of strong competition we set $c=5.0\times10^{-6}$.
-With these two sets of model parameters, we simulated our model for $1000.0$ units of time.
-For both scenarios, we initialized the trait means to $\bar x_i=0.0$, additive genetic variances to $G_i=10.0$ and abundances to $N_i=1000.0$ for each $i=1,\dots,S$.
-
-Temporal dynamics for each scenario are provided in Figure \ref{temporal}.
-This figure suggests weaker competition leads to smoother dynamics and a higher degree of organization within the community.
-Considering expression (\ref{N_comm}) we note that, all else equal, relaxed competition allows for larger growth rates which promote greater abundances.
-From (\ref{N_comm}) we also note that the per-capita effects on demographic stochasticity diminish with abundance.
-To see this, divide both sides by $N_i$.
-Inspecting expressions (\ref{x_comm}) and (\ref{G_comm}), we see that larger abundances also erode the effects of demographic stochasticity on the evolution of mean trait and additive genetic variance.
-These effects were already noted in \S\ref{equations}, and thus are not a consequence of our model of coevolution per-se, but we revisit them here since Figure \ref{temporal} demonstrates the importance of demographic stochasticity in structuring ecological communities even when populations are very large.
-Hence, contrary to the common assumption that stochastic effects can be ignored for large populations, we find that minute asymmetries generated by demographic stochasticity remain significant drivers of community structure.
-In particular, we initialized the species with identical state variables and model parameters, but found an enourmous amount of asymmetry and even some potential phase changes.
-In the following two paragraphs we describe the natural history of the community as illustrated in Figure \ref{temporal}.
-
-We begin by describing the weak competition scenario.
-After about $125.0$ units of time, the community appears to have shaken off the initial conditions and entered into a qualitatively distinct phase of dynamics.
-Aside from a few outliers, most of the species remain clustered together in their state variables.
-This lasts for approximately $375.0$ units of time until, at around time $500.0$, a drastic change occurs.
-At this moment the tightly packed cluster of species begins to fan out in all three state variables.
-Simultaneously, we observe large a shift in mean traits for higher values and in additive genetic variances for lower values.
-Upon inspecting our calculations, we diagnose the reason for this shift.
-The outlier species that were initially pushed away from the common abiotic optimum ($0.0$ in this case) evolved a significant reduction in the amount of heritable variation ($\approx60\%$) due to directional selection induced by competition.
-This reduction in heritable variation slowed adaptation, causing these species to linger on the outskirts of niche space, some longer than others.
-In the meantime the rest of the community, being tightly packed, experienced greater competition which led to diminished abundances for these species and caused some members of the core group to veer away from the abiotic optimum.
-The reduced abundances of the core group led to reduced competition overall.
-As a result, the outlier populations were given a slight increase in growth rate, enough to allow them to increase their abundances orders of magnitude higher than the species in the core group and giving them more weight in driving the evolution of other species.
-Many of these heavy-hitting outlier species had already been maintaining negative mean traits, but around time $500.0$ the high abundance species with positive mean traits began to experience enough intraspecific competition to override interspecific competition.
-This generated a net selection gradient and associated evolutionary response towards the abiotic optimum.
-The sudden imbalance of these high abundance species effectively induced a single large competitive exclusion event pushing the majority of the community far away from the abiotic optimum.
-After this shift the cluster began to slowly bloom in all three state variables as species took advantage of novel asymmetries in their competitive abilities mediated by a new distribution of mean trait values across the community.
-About $125.0$ units of time later, the community reached a qualitatively new phase of dynamics.
-If we kept running the numerical integrator, we would continue to see similar drama unfolding over and over again as minute stochastic changes contribute to asymmetries which slowly build into drastic shifts.
-
-The strong competition scenario is not quite as showy.
-Although the dynamics of trait means and variances tend to be far more stochastic than in the weak competition scenario, the community overall appears to quickly reach some statistical equilibrium and remain there.
-However, the abundances across all species in the community are very low due to competition sensitivies being an order of magnitude higher than in the weak competition case.
-Most of the species maintain abundances greater than $1000.0$, but we found one species that dropped to an abundance of about $50.0$.
-If we let the numerical integrator run long enough in this case, we will likely see many of the species go extinct.
-
-Finding ways to interpret simulated dynamics provides a useful arena to exercise biological reasoning.
-However, it does not fulfill our desire to quantify the patterns and processes present in competing communities.
-In the next subsection we take a step in this direction by using our model to derive formula for selection gradients and competition coefficients.
-To investigate their relationship, we calculate their covariances using simplifying assumptions on species abundances and intraspecific trait variances.
-We then investigate how these covariances change with the ratio of variance of interspecific mean traits to variance of intraspecific individual traits and use a numerical approach to investigate correlations between the strength of pairwise coevolution and competition coefficients.
+Despite the convoluted appearance of system (\ref{comm_dynamics}), there
+are some nice features that reflect biological reasoning. For example,
+the dynamics of abundance are just a generalization of Lotka-Volterra
+dynamics. In particular, the effect of competition with species \(j\) on
+the fitness of species \(i\) grows linearly with \(N_j\). However, as
+biotic selection pushes \(\bar x_i\) away from \(\bar x_j\), the effect
+of competition with species \(j\) on the fitness of species \(i\)
+rapidly diminishes, reflecting a reduction in niche overlap. The
+divergence of \(\bar x_i\) and \(\bar x_j\) due to competition is
+referred to in the community ecology literature as character
+displacement. We also see that the fitness of species \(i\) drops
+quadratically with the difference between \(\bar x_i\) and the abiotic
+optimum \(\theta_i\). Hence, abiotic selection acts to pull \(\bar x_i\)
+towards \(\theta_i\). The response in mean trait \(\bar x_i\) to natural
+selection is proportional to the amount of heritable variation in the
+population, represented by the additive genetic variance \(G_i\).
+However, we have that \(G_i\) is itself a dynamic quantity. Under our
+model, abiotic stabilizing selection erodes away heritable variation at
+a rate that is independent of both \(N_i\) and \(\bar x_i\). The effect
+of competition on \(G_i\) is a bit more complicated. When
+\(b_{ij}(\bar x_i-\bar x_j)^2<1\), competition with species \(j\) acts
+as diversifying selection which tends to increase the amount of
+heritable variation. However, when \(b_{ij}(\bar x_i-\bar x_j)^2>1\),
+competition with species \(j\) acts as directional selection and reduces
+\(G_i\). In the following subsections we demonstrate the behavior of
+system (\ref{comm_dynamics}) by plotting numerical solutions and
+investigate implications for the relationship between the strength of
+ecological interactions and selection.
+
+\hypertarget{community-dynamics}{%
+\subsection{\texorpdfstring{Community dynamics
+\label{dynamics}}{Community dynamics }}\label{community-dynamics}}
+
+For the sake of illustration we numerically integrated system
+(\ref{comm_dynamics}) for a richness of \(S=100\) species. We assumed
+homogeneous model parameters across species in the community as
+summarized by Table \ref{par_vals}. We repeated numerical integration
+under the two scenarios of weak and strong competition. For the first
+scenario of weak competition we set \(c=1.0\times10^{-7}\) and for the
+second scenario of strong competition we set \(c=5.0\times10^{-6}\).
+With these two sets of model parameters, we simulated our model for
+\(1000.0\) units of time. For both scenarios, we initialized the trait
+means to \(\bar x_i=0.0\), additive genetic variances to \(G_i=10.0\)
+and abundances to \(N_i=1000.0\) for each \(i=1,\dots,S\).
+
+Temporal dynamics for each scenario are provided in Figure
+\ref{temporal}. This figure suggests weaker competition leads to
+smoother dynamics and a higher degree of organization within the
+community. Considering expression (\ref{N_comm}) we note that, all else
+equal, relaxed competition allows for larger growth rates which promote
+greater abundances. From (\ref{N_comm}) we also note that the per-capita
+effects on demographic stochasticity diminish with abundance. To see
+this, divide both sides by \(N_i\). Inspecting expressions
+(\ref{x_comm}) and (\ref{G_comm}), we see that larger abundances also
+erode the effects of demographic stochasticity on the evolution of mean
+trait and additive genetic variance. These effects were already noted in
+\S\ref{equations}, and thus are not a consequence of our model of
+coevolution per-se, but we revisit them here since Figure \ref{temporal}
+demonstrates the importance of demographic stochasticity in structuring
+ecological communities even when populations are very large. Hence,
+contrary to the common assumption that stochastic effects can be ignored
+for large populations, we find that minute asymmetries generated by
+demographic stochasticity remain significant drivers of community
+structure. In particular, we initialized the species with identical
+state variables and model parameters, but found an enourmous amount of
+asymmetry and even some potential phase changes. In the following two
+paragraphs we describe the natural history of the community as
+illustrated in Figure \ref{temporal}.
+
+We begin by describing the weak competition scenario. After about
+\(125.0\) units of time, the community appears to have shaken off the
+initial conditions and entered into a qualitatively distinct phase of
+dynamics. Aside from a few outliers, most of the species remain
+clustered together in their state variables. This lasts for
+approximately \(375.0\) units of time until, at around time \(500.0\), a
+drastic change occurs. At this moment the tightly packed cluster of
+species begins to fan out in all three state variables. Simultaneously,
+we observe large a shift in mean traits for higher values and in
+additive genetic variances for lower values. Upon inspecting our
+calculations, we diagnose the reason for this shift. The outlier species
+that were initially pushed away from the common abiotic optimum (\(0.0\)
+in this case) evolved a significant reduction in the amount of heritable
+variation (\(\approx60\%\)) due to directional selection induced by
+competition. This reduction in heritable variation slowed adaptation,
+causing these species to linger on the outskirts of niche space, some
+longer than others. In the meantime the rest of the community, being
+tightly packed, experienced greater competition which led to diminished
+abundances for these species and caused some members of the core group
+to veer away from the abiotic optimum. The reduced abundances of the
+core group led to reduced competition overall. As a result, the outlier
+populations were given a slight increase in growth rate, enough to allow
+them to increase their abundances orders of magnitude higher than the
+species in the core group and giving them more weight in driving the
+evolution of other species. Many of these heavy-hitting outlier species
+had already been maintaining negative mean traits, but around time
+\(500.0\) the high abundance species with positive mean traits began to
+experience enough intraspecific competition to override interspecific
+competition. This generated a net selection gradient and associated
+evolutionary response towards the abiotic optimum. The sudden imbalance
+of these high abundance species effectively induced a single large
+competitive exclusion event pushing the majority of the community far
+away from the abiotic optimum. After this shift the cluster began to
+slowly bloom in all three state variables as species took advantage of
+novel asymmetries in their competitive abilities mediated by a new
+distribution of mean trait values across the community. About \(125.0\)
+units of time later, the community reached a qualitatively new phase of
+dynamics. If we kept running the numerical integrator, we would continue
+to see similar drama unfolding over and over again as minute stochastic
+changes contribute to asymmetries which slowly build into drastic
+shifts.
+
+The strong competition scenario is not quite as showy. Although the
+dynamics of trait means and variances tend to be far more stochastic
+than in the weak competition scenario, the community overall appears to
+quickly reach some statistical equilibrium and remain there. However,
+the abundances across all species in the community are very low due to
+competition sensitivies being an order of magnitude higher than in the
+weak competition case. Most of the species maintain abundances greater
+than \(1000.0\), but we found one species that dropped to an abundance
+of about \(50.0\). If we let the numerical integrator run long enough in
+this case, we will likely see many of the species go extinct.
+
+Finding ways to interpret simulated dynamics provides a useful arena to
+exercise biological reasoning. However, it does not fulfill our desire
+to quantify the patterns and processes present in competing communities.
+In the next subsection we take a step in this direction by using our
+model to derive formula for selection gradients and competition
+coefficients. To investigate their relationship, we calculate their
+covariances using simplifying assumptions on species abundances and
+intraspecific trait variances. We then investigate how these covariances
+change with the ratio of variance of interspecific mean traits to
+variance of intraspecific individual traits and use a numerical approach
+to investigate correlations between the strength of pairwise coevolution
+and competition coefficients.
 
@@
 
-```{r, eval=F, echo=F}
-require("ggplot2")
-require("gridExtra")
-require("reshape")
-
-####################
-#                  #
-# weak competition #
-#                  #
-####################
-
-smpl_wc = read.csv("/home/bb/Gits/branching.brownian.motion.and.spde/sample_path_wc.csv")
-
-x_wc = ggplot(smpl_wc,aes(x=time,y=x,group=spp))+theme_minimal()+xlab("")+ylab("Mean Trait")+
-  ylim(-12,12)+geom_line(show.legend=F, lwd=0.1)+
-  ggtitle("Weak Competition")+theme(plot.title = element_text(hjust = 0.5))
-
-G_wc = ggplot(smpl_wc,aes(x=time,y=G,group=spp))+theme_minimal()+xlab("")+ylab("Additive Genetic \n Variance")+
-  ylim(0,45)+geom_line(show.legend=F, lwd=0.1)
-
-N_wc = ggplot(smpl_wc,aes(x=time,y=N,group=spp))+theme_minimal()+xlab("time")+ylab("Abundance")+
-  scale_y_log10(limits=c(50,1e8))+geom_line(show.legend=F, lwd=0.1)
-
-smpl_plt_wc = grid.arrange(x_wc,G_wc,N_wc,nrow=3)
-
-#
-# calculating strength of coevolution
-#
-
-tmax = max(smpl_wc$time)
-
-final_comm = subset(smpl_wc,time==tmax)
-
-x = final_comm$x
-G = final_comm$G
-N = final_comm$N
-
-w = 0.1
-c = 1e-7
-eta = 1.0
-R = 1.0
-a = 1e-2
-th = 0
-S = length(x)
-
-b = matrix(0,nc=S,nr=S)
-C = matrix(0,nc=S,nr=S)
-alpha = matrix(0,nc=S,nr=S)
-beta = matrix(0,nc=S,nr=S)
-gamma = matrix(0,nc=S,nr=S)
-for( i in 1:S){
-  for( j in 1:S ){
-
-    r = R - a*( (x[i]-th)^2 + G[i] + eta )/2
-
-    b[i,j] = 1/( G[i]+G[j]+2*(w+eta) )
-
-    alpha[i,j] = c*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2) / r
-
-    beta[i,j] = c*N[j]*b[i,j]*(x[i]-x[j])*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2)
-
-    gamma[i,j] = c*N[j]*b[i,j]*(1 - b[i,j]*(x[i]-x[j])^2)*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2)
-
-    C[i,j] = abs(beta[i,j]) + abs(gamma[i,j])
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{community_dynamics}
 
 }
+
+\caption{\label{temporal}Temporal dynamics of mean trait (top), additive genetic variance (middle) and abundance (bottom) for the scenario of weak competition (left) and strong competition (right).}\label{fig:unnamed-chunk-6}
+\end{figure}
+
+\hypertarget{the-relation-between-the-strength-of-ecological-interactions-and-coevolution}{%
+\subsection{\texorpdfstring{The relation between the strength of
+ecological interactions and coevolution
+\label{ecoevo}}{The relation between the strength of ecological interactions and coevolution }}\label{the-relation-between-the-strength-of-ecological-interactions-and-coevolution}}
+
+Relating our treatment of resource competition to modern coexistence
+theory (Chesson 2000), the absolute competition coefficient
+\(\alpha_{ij}\) becomes a dynamical quantity that can be written as
+
+\begin{equation}
+\alpha_{ij}(t)=\frac{c_i}{r_i(t)}\int_{-\infty}^{+\infty}\int_{-\infty}^{+\infty}p_i(x,t)p_j(y,t)\mathcal O_{ij}(x,y) dxdy =\frac{c_iU_iU_j}{r_i(t)}\sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right),
+\end{equation}
+
+where
+
+\begin{equation}
+r_i(t)=R_i-\frac{a_i}{2}\Big((\bar x_i(t)-\theta_i)^2+G_i(t)+\eta_i\Big).
+\end{equation}
+
+Hence, \(dN_i(t)\) can be expressed as
+
+\begin{equation}
+dN_i(t)=r_i(t)\left(1-\sum_{j=1}^S\alpha_{ij}(t)N_j(t)\right)N_i(t)dt+\sqrt{V_iN_i(t)}dW_1(t).
+\end{equation}
+
+Note that although \(r_i(t)\) is referred to in the coexistence
+literature as the intrinsic growth rate of the population, \(R_i\) is a
+deeper intrinsic quantity. For now we refer to \(R_i\) as the
+\emph{innate} growth rate. Previous work, to list just a few citations
+in this literature, has shown the importance of demographic
+stochasticity (Turelli 1980; Schreiber 2017) and evolutionary dynamics
+(Case and Taper 1986; Schreiber, Patel, and terHorst 2018) in
+determining coexistence of competing species. However, these studies
+consider evolution and demographic stochasticity separately. To our
+knowledge, there are no rigorous theoretical investigations of
+coexistence that account for evolutionary dynamics of quantitative
+characters and demographic stochasticity simultaneously and, in
+particular, no studies combining these processes that also include the
+evolution of additive genetic variance. In fact, only recently have
+rigorous theoretical investigations successfully combined these
+processes to understand persistence of a single population
+(Gomulkiewicz, Krone, and Remien 2017). Even so, these authors focused
+on a single species model with a finite number of types (i.e., a
+population genetic model). With the connection outlined in this section
+formally established, researchers may pursue a postmodern coexistence
+theory that naturally includes both the evolutionary dynamics of
+quantitative characters and the effects of demographic stochasticity in
+a simple synthetic framework.
+
+In SM \S\ref{diffuse} we show that the standardized directional
+selection gradient (sensu Lande and Arnold 1983) induced by species
+\(j\) on species \(i\) can be computed as
+
+\begin{equation}
+\beta_{ij}(t)=c_iU_iU_jN_j(t)b_{ij}(t)\big(\bar x_i(t)-\bar x_j(t)\big) \sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right).
+\end{equation}
+
+Our notation differs from Lande and Arnold (1983) in that subscripts
+here denote species instead of components of multivariate traits and we
+drop the prime that distinguishes between selection gradients and
+standardized selection gradients.
+
+\textbf{Metric of pairwise coevolution}
+
+Below we investigate the correspondence of interaction intensity and
+coevolutionary change. However, we can already identify one major
+discrepancy; \(\alpha_{ij}\) is maximized when \(\bar x_i=\bar x_j\),
+but \(\beta_{ij}=0\) under the same condition. We therefore include in
+our metric of selection the standardized stabilizing selection gradient
+\(\gamma\) which measures the effect of stabilizing or disruptive
+selection on phenotypic variance (Lande and Arnold 1983). In SM
+\S\ref{diffuse} we show that the standardized stabilizing selection
+gradient induced by species \(j\) on species \(i\) can be computed as
+
+\begin{equation}
+\gamma_{ij}(t)=c_iU_iU_jN_j(t)b_{ij}(t)\left(1-b_{ij}(t)\big(\bar x_i(t)-\bar x_j(t)\big)^2\right) \sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right).
+\end{equation}
+
+To measure the total evolutionary change in species \(i\) induced by
+species \(j\), we form the metric
+\(\Psi_{ij}=|\beta_{ij}|+|\gamma_{ij}|\). Figure \ref{net} displays the
+joint distribution of \(\mathfrak{C}_{ij}=\Psi_{ij}\Psi_{ji}\), our
+metric of pairwise coevolution, and the product of competition
+coefficients \(\alpha_{ij}\alpha_{ji}\) for two simulated communities,
+both with richness \(S=1000\). The solid contour represents the case of
+relatively high sensitivity to competition (\(c=5\times10^{-7}\)) and
+the dashed contour represents the case of relatively weak sensitivity to
+competition (\(c=1\times10^{-8}\)). In both cases we see that
+\(\mathfrak{C}_{ij}\) and \(\alpha_{ij}\alpha_{ji}\) are, essentially,
+unrelated. However, this depiction is only representative of a specific
+set of parameters. Next, we provide analytical approximations of the
+covariance between selection gradients and competition coefficients and
+a numerical estimate for the relationship between pairwise coevolution
+and competition coefficients for a range of parameters.
+
+\begin{figure}
+
+{\centering \includegraphics{on_pl}
+
 }
 
-Csp = C # defines a matrix that ignores intraspecific interactions
-for( i in 1:S ){
-  Csp[i,i] = 0
-}
-
-Cvec = NULL
-for(i in 1:S){
-  for(j in i:S)
-    if( i!=j ) Cvec = c(Cvec,C[i,j])
-}
-
-#
-# distribution of coevolution
-#
-
-Cdf = data.frame(val=Cvec)
-
-WC_co <- ggplot(Cdf, aes(x=val)) + geom_histogram(aes(y=..count../sum(..count..)), fill='black', alpha=1.0, bins=50) + 
-  theme_minimal() + xlab("Pairwise Coevolution") + ylab("Relative Frequency") + 
-  scale_x_log10(limits=c(1e-7,1e-1))
-
-#
-# building a network
-#
-require(igraph)
-require(plotwidgets)
-require(network)
-require(GGally)
-
-asp = alpha # defines a matrix that ignores intraspecific interactions
-for( i in 1:S ){
-  asp[i,i] = 0
-}
-
-avec = NULL
-for(i in 1:S){
-  for(j in i:S)
-    if( i!=j ) avec = c(avec,alpha[i,j])
-}
-
-# bivariate plot
-Cadf_WC = data.frame(Coev=Cvec,Comp=avec, Which="WC")
-Ca_WC_pl = ggplot(Cadf_WC, aes(Coev,Comp)) + geom_bin2d(bins=20, show.legend = FALSE) + 
-  geom_density_2d(col='white') + scale_fill_gradient(low="black",high="grey",trans="log") +
-  scale_x_log10(limits=c(5e-5,1e-2)) + scale_y_log10(limits=c(2e-9,2e-8)) + 
-  theme_minimal() + xlab("Pairwise Coevolution") + ylab("Competition Coefficient") + 
-  ggtitle("Weak \n Competition") + theme(plot.title = element_text(hjust = 0.5))
-
-network <- graph_from_adjacency_matrix(ceiling(asp/max(asp)))
-
-CDF <- function(x,X){
-  L <- length(X)
-  k <- 0
-  for(i in 1:L){
-    if(x >= X[i]) k <- k+1
-  }
-  k/L
-}
-
-resc_wts <- NULL
-for(i in 1:length(avec)){
-  resc_wts <- c(resc_wts, CDF(avec[i],avec))
-}
-
-colors <- colorRamp(c("white","black"))
-
-edge_colors <- NULL
-for(i in 1:length(avec)){
-  edge_colors <- c(edge_colors,rgb2col(t(colors(resc_wts[i]))))
-}
-
-E(network)$width <- resc_wts
-E(network)$weights <- resc_wts
-E(network)$arrow.size <- 0.01
-E(network)$color <- edge_colors
-V(network)$size <- N/sum(N)
-V(network)$frame.color <- "white"
-V(network)$color <- "black"
-V(network)$label <- ""
-# l <- layout_in_circle(network)
-l <- layout_with_fr(network)
-
-WC_net <- network(asp, directed = F)
-
-netWC <- ggnet2(WC_net, mode = "circle", edge.alpha=.1, edge.color = edge_colors, edge.size = 2*min(1/avec)*avec, size = N/sum(N), max_size = 3, node.color = 'black')+
-  ggtitle("Weak \n Competition")+theme(plot.title = element_text(hjust = 0.5), legend.position = "none")
-
-######################
-#                    #
-# strong competition #
-#                    #
-######################
-
-smpl_mc = read.csv("/home/bb/Gits/branching.brownian.motion.and.spde/sample_path_mc.csv")
-
-x_mc = ggplot(smpl_mc,aes(x=time,y=x,group=spp))+theme_bw()+theme_minimal()+xlab("")+ylab("")+
-  ylim(-12,12)+geom_line(show.legend=F, lwd=0.1)+ggtitle("Strong Competition")+theme(plot.title = element_text(hjust = 0.5))
-
-G_mc = ggplot(smpl_mc,aes(x=time,y=G,group=spp))+theme_bw()+theme_minimal()+xlab("")+ylab("")+
-  ylim(0,45)+geom_line(show.legend=F, lwd=0.1)
-
-N_mc = ggplot(smpl_mc,aes(x=time,y=N,group=spp))+theme_bw()+theme_minimal()+xlab("time")+ylab("")+
-  scale_y_log10(limits=c(50,1e8))+geom_line(show.legend=F, lwd=0.1)
-
-smpl_plt_mc = grid.arrange(x_mc,G_mc,N_mc,nrow=3)
-
-community_plot = grid.arrange(smpl_plt_wc,smpl_plt_mc,nrow=1)
-
-ggsave("~/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/community_dynamics.png",community_plot,width=6.5,height=5,dpi=320)
-
-#
-# calculating strength of coevolution
-#
-
-tmax = max(smpl_mc$time)
-
-final_comm = subset(smpl_mc,time==tmax)
-
-x = final_comm$x
-G = final_comm$G
-N = final_comm$N
-
-w = 0.1
-c = 1e-7
-eta = 1.0
-R = 1.0
-a = 1e-2
-th = 0
-S = length(x)
-
-b = matrix(0,nc=S,nr=S)
-C = matrix(0,nc=S,nr=S)
-alpha = matrix(0,nc=S,nr=S)
-beta = matrix(0,nc=S,nr=S)
-gamma = matrix(0,nc=S,nr=S)
-for( i in 1:S){
-  for( j in 1:S ){
-
-    r = R - a*( (x[i]-th)^2 + G[i] + eta )/2
-
-    b[i,j] = 1/( G[i]+G[j]+2*(w+eta) )
-
-    alpha[i,j] = c*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2) / r
-
-    beta[i,j] = c*N[j]*b[i,j]*(x[i]-x[j])*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2)
-
-    gamma[i,j] = c*N[j]*b[i,j]*(1 - b[i,j]*(x[i]-x[j])^2)*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2)
-
-    C[i,j] = abs(beta[i,j]) + abs(gamma[i,j])
-
-  }
-}
-
-Csp = C # defines a matrix that ignores intraspecific interactions
-for( i in 1:S ){
-  Csp[i,i] = 0
-}
-
-Cvec = NULL
-for(i in 1:S){
-  for(j in i:S)
-    if( i!=j ) Cvec = c(Cvec,C[i,j])
-}
-
-#
-# distribution of coevolution
-#
-
-Cdf = data.frame(val=Cvec)
-
-MC_co <- ggplot(Cdf, aes(x=val)) + geom_histogram(aes(y=..count../sum(..count..)), fill='black', alpha=1.0, bins=50) + 
-  theme_minimal() + xlab("Pairwise Coevolution") + ylab("") + 
-  scale_x_log10(limits=c(1e-7,1e-1))
-
-#
-# building a network
-#
-require(igraph)
-require(plotwidgets)
-require(network)
-require(GGally)
-
-asp = alpha # defines a matrix that ignores intraspecific interactions
-for( i in 1:S ){
-  asp[i,i] = 0
-}
-
-avec = NULL
-for(i in 1:S){
-  for(j in i:S)
-    if( i!=j ) avec = c(avec,alpha[i,j])
-}
-
-# bivariate plot
-Cadf_SC = data.frame(Coev=Cvec,Comp=avec, Which="SC")
-Ca_SC_pl = ggplot(Cadf_SC, aes(Coev,Comp)) + geom_bin2d(bins=20, show.legend = FALSE) + 
-  geom_density_2d(col='white') + scale_fill_gradient(low="black",high="grey",trans="log") +
-  scale_x_log10(limits=c(5e-7,2e-4)) + scale_y_log10(limits=c(3e-9,1e-8)) + 
-  theme_minimal() + xlab("Pairwise Coevolution") + ylab("") + ggtitle("Strong \n Competition") +
-  theme(plot.title = element_text(hjust = 0.5))
-
-bivplot = grid.arrange(Ca_WC_pl,Ca_SC_pl,nrow=1)
-
-Cadf = rbind(Cadf_WC,Cadf_SC)
-head(Cadf)
-
-on_pl = ggplot(Cadf_WC, aes(Coev,Comp)) + geom_density_2d(col="black") +
-  geom_density_2d(data=Cadf_SC,col="red") +
-  scale_x_log10(limits=c(5e-7,1e-2)) + scale_y_log10(limits=c(2.5e-9,1.5e-8)) + 
-  theme_minimal() + xlab("Pairwise Coevolution") + ylab("Competition Coefficient")
-
-ggsave("~/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/on_pl.png",on_pl,device="png",width=3.5,height=3.5,units="in",dpi=350)
-
-network <- graph_from_adjacency_matrix(ceiling(asp/max(asp)))
-
-CDF <- function(x,X){
-  L <- length(X)
-  k <- 0
-  for(i in 1:L){
-    if(x >= X[i]) k <- k+1
-  }
-  k/L
-}
-
-resc_wts <- NULL
-for(i in 1:length(avec)){
-  resc_wts <- c(resc_wts, CDF(avec[i],avec))
-}
-
-colors <- colorRamp(c("white","black"))
-
-edge_colors <- NULL
-for(i in 1:length(avec)){
-  edge_colors <- c(edge_colors,rgb2col(t(colors(resc_wts[i]))))
-}
-
-E(network)$width <- resc_wts
-E(network)$weights <- resc_wts
-E(network)$arrow.size <- 0.01
-E(network)$color <- edge_colors
-V(network)$size <- N/sum(N)
-V(network)$frame.color <- "white"
-V(network)$color <- "black"
-V(network)$label <- ""
-# l <- layout_in_circle(network)
-l <- layout_with_fr(network)
-
-MC_net <- network(asp, directed = F)
-
-netMC <- ggnet2(MC_net, mode = "circle", edge.alpha=.1, edge.color = edge_colors, edge.size = 2*min(1/avec)*avec, size = N/sum(N), max_size = 3, node.color = 'black')+
-  ggtitle("Strong \n Competition")+theme(plot.title = element_text(hjust = 0.5), legend.position = "none")
-
-coevs <- grid.arrange(netWC, netMC,
-                      WC_co, MC_co, nrow=2)
-
-ggsave("~/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/interaction_networks.png",coevs,device="png",width=6,height=6,units="in",dpi=350)
-
-```
-
-```{r, message=FALSE, echo=F, fig.align='center', out.width='100%', fig.cap='\\label{temporal}Temporal dynamics of mean trait (top), additive genetic variance (middle) and abundance (bottom) for the scenario of weak competition (left) and strong competition (right).'}
-knitr::include_graphics('community_dynamics.jpg')
-```
-
-## The relation between the strength of ecological interactions and coevolution \label{ecoevo}
-
-Relating our treatment of resource competition to modern coexistence theory [@Chesson2000], the absolute competition coefficient $\alpha_{ij}$ becomes a dynamical quantity that can be written as
-
-\begin{equation}
-\alpha_{ij}(t)=\frac{c_i}{r_i(t)}\int_{-\infty}^{+\infty}\int_{-\infty}^{+\infty}p_i(x,t)p_j(y,t)\mathcal O_{ij}(x,y) dxdy =\frac{c_iU_iU_j}{r_i(t)}\sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right),
-\end{equation}
-
-where
-
-\begin{equation}
-r_i(t)=R_i-\frac{a_i}{2}\Big((\bar x_i(t)-\theta_i)^2+G_i(t)+\eta_i\Big).
-\end{equation}
-
-Hence, $dN_i(t)$ can be expressed as
-
-\begin{equation}
-dN_i(t)=r_i(t)\left(1-\sum_{j=1}^S\alpha_{ij}(t)N_j(t)\right)N_i(t)dt+\sqrt{V_iN_i(t)}dW_1(t).
-\end{equation}
-
-Note that although $r_i(t)$ is referred to in the coexistence literature as the intrinsic growth rate of the population, $R_i$ is a deeper intrinsic quantity.
-For now we refer to $R_i$ as the _innate_ growth rate.
-Previous work, to list just a few citations in this literature, has shown the importance of demographic stochasticity [@Turelli1980; @Schreiber2017] and evolutionary dynamics [@Case1986; @Schreiber2018] in determining coexistence of competing species.
-However, these studies consider evolution and demographic stochasticity separately.
-To our knowledge, there are no rigorous theoretical investigations of coexistence that account for evolutionary dynamics of quantitative characters and demographic stochasticity simultaneously and, in particular, no studies combining these processes that also include the evolution of additive genetic variance.
-In fact, only recently have rigorous theoretical investigations successfully combined these processes to understand persistence of a single population [@Gomulkiewicz2017].
-Even so, these authors focused on a single species model with a finite number of types (i.e., a population genetic model).
-With the connection outlined in this section formally established, researchers may pursue a postmodern coexistence theory that naturally includes both the evolutionary dynamics of quantitative characters and the effects of demographic stochasticity in a simple synthetic framework.
-
-In SM \S\ref{diffuse} we show that the standardized directional selection gradient [sensu @Lande1983] induced by species $j$ on species $i$ can be computed as
-
-\begin{equation}
-\beta_{ij}(t)=c_iU_iU_jN_j(t)b_{ij}(t)\big(\bar x_i(t)-\bar x_j(t)\big) \sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right).
-\end{equation}
-
-Our notation differs from Lande and Arnold [-@Lande1983] in that subscripts here denote species instead of components of multivariate traits and we drop the prime that distinguishes between selection gradients and standardized selection gradients.
-
-__Metric of pairwise coevolution__
-
-Below we investigate the correspondence of interaction intensity and coevolutionary change.
-However, we can already identify one major discrepancy; $\alpha_{ij}$ is maximized when $\bar x_i=\bar x_j$, but $\beta_{ij}=0$ under the same condition.
-We therefore include in our metric of selection the standardized stabilizing selection gradient $\gamma$ which measures the effect of stabilizing or disruptive selection on phenotypic variance [@Lande1983].
-In SM \S\ref{diffuse} we show that the standardized stabilizing selection gradient induced by species $j$ on species $i$ can be computed as
-
-\begin{equation}
-\gamma_{ij}(t)=c_iU_iU_jN_j(t)b_{ij}(t)\left(1-b_{ij}(t)\big(\bar x_i(t)-\bar x_j(t)\big)^2\right) \sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right).
-\end{equation}
-
-To measure the total evolutionary change in species $i$ induced by species $j$, we form the metric $\Psi_{ij}=|\beta_{ij}|+|\gamma_{ij}|$.
-Figure \ref{net} displays the joint distribution of $\mathfrak{C}_{ij}=\Psi_{ij}\Psi_{ji}$, our metric of pairwise coevolution, and the product of competition coefficients $\alpha_{ij}\alpha_{ji}$ for two simulated communities, both with richness $S=1000$.
-The solid contour represents the case of relatively high sensitivity to competition ($c=5\times10^{-7}$) and the dashed contour represents the case of relatively weak sensitivity to competition ($c=1\times10^{-8}$).
-In both cases we see that $\mathfrak{C}_{ij}$ and $\alpha_{ij}\alpha_{ji}$ are, essentially, unrelated.
-However, this depiction is only representative of a specific set of parameters.
-Next, we provide analytical approximations of the covariance between selection gradients and competition coefficients and a numerical estimate for the relationship between pairwise coevolution and competition coefficients for a range of parameters.
-
-```{r, eval=F, echo=F}
-require("ggplot2")
-require("gridExtra")
-require("reshape")
-
-####################
-#                  #
-# weak competition #
-#                  #
-####################
-
-smpl_wc1e3 = read.csv("/home/bb/Gits/branching.brownian.motion.and.spde/sample_path_wc1e3.csv")
-
-#
-# calculating strength of coevolution
-#
-
-x = smpl_wc1e3$x
-G = smpl_wc1e3$G
-N = smpl_wc1e3$N
-
-w = 0.1
-c = 1e-8
-eta = 1.0
-R = 1.0
-a = 1e-2
-th = 0
-S = length(x)
-
-b = matrix(0,nc=S,nr=S)
-C = matrix(0,nc=S,nr=S)
-alpha = matrix(0,nc=S,nr=S)
-beta = matrix(0,nc=S,nr=S)
-gamma = matrix(0,nc=S,nr=S)
-for( i in 1:S){
-  for( j in 1:S ){
-
-    r = R - a*( (x[i]-th)^2 + G[i] + eta )/2
-
-    b[i,j] = 1/( G[i]+G[j]+2*(w+eta) )
-
-    alpha[i,j] = c*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2) / r
-
-    beta[i,j] = c*N[j]*b[i,j]*(x[i]-x[j])*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2)
-
-    gamma[i,j] = c*N[j]*b[i,j]*(1 - b[i,j]*(x[i]-x[j])^2)*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2)
-
-    C[i,j] = abs(beta[i,j]) + abs(gamma[i,j])
-
-  }
-}
-
-Cvec = NULL
-for(i in 1:S){
-  for(j in i:S)
-    if( i!=j ) Cvec = c(Cvec,C[i,j])
-}
-
-avec = NULL
-for(i in 1:S){
-  for(j in i:S)
-    if( i!=j ) avec = c(avec,alpha[i,j])
-}
-
-Cadf_WC = data.frame(Coev=Cvec,Comp=avec, Which="WC")
-
-######################
-#                    #
-# strong competition #
-#                    #
-######################
-
-smpl_mc1e3 = read.csv("/home/bb/Gits/branching.brownian.motion.and.spde/sample_path_mc1e3.csv")
-
-#
-# calculating strength of coevolution
-#
-
-x = smpl_mc1e3$x
-G = smpl_mc1e3$G
-N = smpl_mc1e3$N
-
-w = 0.1
-c = 5e-7
-eta = 1.0
-R = 1.0
-a = 1e-2
-th = 0
-S = length(x)
-
-b = matrix(0,nc=S,nr=S)
-C = matrix(0,nc=S,nr=S)
-alpha = matrix(0,nc=S,nr=S)
-beta = matrix(0,nc=S,nr=S)
-gamma = matrix(0,nc=S,nr=S)
-for( i in 1:S){
-  for( j in 1:S ){
-
-    r = R - a*( (x[i]-th)^2 + G[i] + eta )/2
-
-    b[i,j] = 1/( G[i]+G[j]+2*(w+eta) )
-
-    alpha[i,j] = c*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2) / r
-
-    beta[i,j] = c*N[j]*b[i,j]*(x[i]-x[j])*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2)
-
-    gamma[i,j] = c*N[j]*b[i,j]*(1 - b[i,j]*(x[i]-x[j])^2)*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2)
-
-    C[i,j] = abs(beta[i,j]) + abs(gamma[i,j])
-
-  }
-}
-
-Cvec = NULL
-for(i in 1:S){
-  for(j in i:S)
-    if( i!=j ) Cvec = c(Cvec,C[i,j])
-}
-
-avec = NULL
-for(i in 1:S){
-  for(j in i:S)
-    if( i!=j ) avec = c(avec,alpha[i,j])
-}
-
-Cadf_SC = data.frame(Coev=Cvec,Comp=avec, Which="SC")
-
-# bivariate plot
-Cadf = rbind(Cadf_WC,Cadf_SC)
-
-on_pl = ggplot(Cadf_SC, aes(Coev,Comp)) + 
-  scale_x_log10(limits=c(1e-5,2e-4)) + 
-  geom_density_2d(col="black",n=200) +
-  geom_density_2d(data=Cadf_WC, linetype="dashed", col="black") +
-  theme_minimal() + xlab("Pairwise Coevolution") + ylab("")
-
-on_pl2 = on_pl + facet_grid(rows=vars(Which), scales="free_y") + 
-  ylab("Competition Coefficient") + theme(strip.background = element_blank(), strip.text.y = element_blank())
-
-ggsave("~/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/on_pl.png",on_pl2,device="png",width=3.5,height=3.5,units="in")
-```
-
-```{r, echo=F, fig.align='center', fig.cap='\\label{net}Bivariate distributions of competition coffecients (y-axis) and pairwise coevolution (x-axis) under the scenarios of weak competition (dashed line) and strong competition (solid line) after simulating for $t=1000.0$ units of time.
+\caption{\label{net}Bivariate distributions of competition coffecients (y-axis) and pairwise coevolution (x-axis) under the scenarios of weak competition (dashed line) and strong competition (solid line) after simulating for $t=1000.0$ units of time.
 Simulations were ran for $S=1000$ species.
-Parameters are the same as in Table 2, except we used $c=1\\times 10^{-8}$ for weak competition and $c=5\\times 10^{-7}$ for strong competition in order to account for increased species richness.'}
-knitr::include_graphics('on_pl.png')
-```
-
-__Covariance of selection and competition as a function of diversity__
-
-We now make use of the expressions derived for competition coeffecients and selection gradients to investigate their relationship.
-As a first pass, let us assume all model parameters are equivalent across species and that each species has the same abundance and trait variance.
-Let us further assume that richness $S$ is large and the distribution of mean trait values is normal with mean $\bar{\bar x}$, variance $V_{\bar X}$ and density $f_{\bar X}$.
-Such assumptions are typical when deriving analytical results in the field of theoretical coevolutionary community ecology [@Nuismer2012; @Nuismer2018].
-If $\bar{\bar x}$ is near $\theta$ and $V_{\bar X}$ is much smaller than $|2R/a-G-\eta|$, then we may approximate $r_i$ with
+Parameters are the same as in Table 2, except we used $c=1\times 10^{-8}$ for weak competition and $c=5\times 10^{-7}$ for strong competition in order to account for increased species richness.}\label{fig:unnamed-chunk-8}
+\end{figure}
+
+\textbf{Covariance of selection and competition as a function of
+diversity}
+
+We now make use of the expressions derived for competition coeffecients
+and selection gradients to investigate their relationship. As a first
+pass, let us assume all model parameters are equivalent across species
+and that each species has the same abundance and trait variance. Let us
+further assume that richness \(S\) is large and the distribution of mean
+trait values is normal with mean \(\bar{\bar x}\), variance
+\(V_{\bar X}\) and density \(f_{\bar X}\). Such assumptions are typical
+when deriving analytical results in the field of theoretical
+coevolutionary community ecology (Nuismer, Jordano, and Bascompte 2012;
+Nuismer, Week, and Aizen 2018). If \(\bar{\bar x}\) is near \(\theta\)
+and \(V_{\bar X}\) is much smaller than \(|2R/a-G-\eta|\), then we may
+approximate \(r_i\) with
 
@@
 
-In SM \S\ref{coef_grad_moms} we use these assumptions to calculate the first and second order moments describing the joint distribution of competition coefficients and selection gradients across the community.
-We find that the covariance between linear selection gradients and competition coeffecients are zero due to the symmetry implied by our assumptions.
-However, setting $\alpha(\bar x_i,\bar x_j)=\alpha_{ij}$, $\beta(\bar x_i,\bar x_j)=\beta_{ij}$ and $\gamma(\bar x_i,\bar x_j)=\gamma_{ij}$, the covariances between the magnitude of linear selection gradients and competition coefficients and between stabilizing selection gradients and competition coefficients can be written as
+In SM \S\ref{coef_grad_moms} we use these assumptions to calculate the
+first and second order moments describing the joint distribution of
+competition coefficients and selection gradients across the community.
+We find that the covariance between linear selection gradients and
+competition coeffecients are zero due to the symmetry implied by our
+assumptions. However, setting \(\alpha(\bar x_i,\bar x_j)=\alpha_{ij}\),
+\(\beta(\bar x_i,\bar x_j)=\beta_{ij}\) and
+\(\gamma(\bar x_i,\bar x_j)=\gamma_{ij}\), the covariances between the
+magnitude of linear selection gradients and competition coefficients and
+between stabilizing selection gradients and competition coefficients can
+be written as
 
@@
 
-```{r, eval=F, echo=F}
-c=1e-4
-b=1e-1
-r=0.1
-N=1e10
-Cov_ab <- function(x){
-  # x <- 2*b*V_{\bar X}
-  val = 2*c^2*b^2*N / (pi*r)
-  val = val * sqrt(x/(2*b*pi))
-  val2 = (1+8*x/2)^(-0.75)
-  val3 = (1+4*x/2)^(-0.75) / sqrt(1+2*x/2)
-  val = val * (val2 - val3)
-  return(val)  
+\begin{figure}
+
+{\centering \includegraphics{/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/cov}
+
 }
-Cov_ag <- function(x){
-  # x <- 2*b*V_{\bar X}
-  val = 2*c^2*b^2*N / (2*pi*r)
-  val = val * (1-2*x/2)
-  val2 = 1 / sqrt(1+4*x/2)
-  val3 = 1 / (1+2*x/2)
-  val = val * (val2 - val3)
-  return(val)  
+
+\caption{\label{cov_fig}Curves representing the covariance between the magnitude of linear selection gradients and competition coefficients (left) and between stabilizing selection gradients and competition coefficients (right) as a function of $2bV_{\bar X}$ which is approximately equal to the ratio of variance in mean traits among species to the intraspecific trait variance.
+In both plots we set $c=1.0\times10^{-4}$, $b=0.1$, $\bar r=0.1$ and $N=1.0\times10^{10}$ and let $V_{\bar X}$ vary.}\label{fig:unnamed-chunk-10}
+\end{figure}
+
+For fixed \(c,b,\bar r\) and \(N\), we visualize these relationships in
+Figure \ref{cov_fig}. To gain insight into the relationship between
+selection gradients and competition coefficients, note that our
+assumptions in this section imply \(b^{-1}=2(\sigma^2+w)\). If we
+further assume \(\sigma^2+w\approx\sigma^2\), then
+\(2bV_{\bar X}\approx V_{\bar X}/\sigma^2\). That is, when populations
+are generalists and are comprised of specialist individuals, the value
+\(2bV_{\bar X}\) is approximately equal to the ratio of interspecific
+mean trait variation to intraspecific individual trait variation. Hence,
+for both covariances we see that there is no relationship between
+selection gradients and competition coefficients when this ratio is
+zero. From equation (\ref{cov_alpha_beta}) we can use numerical
+optimization to find that when \(V_{\bar X}/\sigma^2\approx1.25\) the
+relationship between the magnitudes of linear selection gradients and
+competition coefficients disappears, but when (approximately)
+\(V_{\bar X}/\sigma^2<1.25\) (\(>1.25\)), this covariance becomes
+negative (positive). Equation (\ref{cov_alpha_gamma}) states that when
+\(V_{\bar X}/\sigma^2\) is approximately equal to one (or slightly
+larger), there is no expected relationship between competition
+coefficients and quadratic selection gradients. However, when
+\(V_{\bar X}/\sigma^2<1.0\) (\(>1.0\)), then we expect a positive
+(negative) relationship between \(\alpha\) and \(\gamma\). These results
+are true regardless of the chosen parameter values. In SM
+\S\ref{coef_grad_moms} we use simulations of system
+(\ref{comm_dynamics}) to show that these results do not qualitatively
+differ when allowing for heterogeneous population sizes and additive
+genetic variances across species.
+
+From a biological perspective, if the ratio \(V_{\bar X}/\sigma^2\) is
+small, then species are packed tightly in phenotypic space. In our model
+this occurs when abiotic stabilizing selection is much stronger than
+competition (\(a\gg c\)). This causes species to overlap more in niche
+space (i.e., large \(\alpha\)) and creates disruptive selection for
+greater intraspecific variance (i.e., positive \(\gamma\)), which
+explains the positive region of covariance between \(\alpha\) and
+\(\gamma\). However, as species begin to overlap in niche space,
+directional selection begins to vanish (i.e., small \(|\beta|\)),
+leading to a negative covariance between \(\alpha\) and \(|\beta|\). In
+the limiting case that two species have perfectly overlapping niches,
+they will exhibit zero directional selection since a shift in either
+direction will yield the same fitness advantages.
+
+In the opposite scenario where competition is much stronger than abiotic
+stabilizing selection (\(c\gg a\)), species will not evolve to be as
+tightly packed and instead their niche-centers will be spread out with
+little ovelap in their resouce utilization curves (i.e., small
+\(\alpha\)). In this case biotic directional selection will be strong
+(i.e., large \(|\beta|\)), particularly for species towards the outer
+regions of niche space due to asymmetric fitness advantages conferred by
+shifts in niche-centers. This leads to a positive covariance between
+\(\alpha\) and \(|\beta|\). However, as noted above, this directional
+selection will also erode away at standing heritable variation (i.e.,
+negative \(\gamma\)), reducing the rate at which adaptation can occur
+and creating a negative covariance between \(\alpha\) and \(\gamma\).
+
+In summary, we see the relation between competition coefficients and
+selection is highly non-trivial and depends on the relative magnitudes
+of different ecological processes shaping the community. However, this
+does not address the relation between competition coefficients and
+coevolution per se. In SM \S\ref{coef_grad_moms} we show that
+calculating a formula for the covariance between competition
+coefficients and the metric of coevolution \(\mathfrak{C}\) introduced
+above provides a difficult analytical challenge. Instead of confronting
+this challenge we build on our numerical approach used to justify
+analytical approximations of
+\(\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)\) and
+\(\mathrm{Cov}_{f_{\bar X}}(\alpha,\gamma)\) to approximate the
+correlation of \(\alpha\) and \(\mathfrak{C}\). This numerical approach
+inherits the assumptions of homogeneous background parameters such as
+the mutation rate \(\mu\) and abiotic optima \(\theta\), but allows us
+to relax the assumption that \(N\) and \(G\) are constant across species
+and time.
+
+In particular, we numerically integrated system (\ref{comm_dynamics})
+for \(T_1=1000.0\) units of time and then continued to integrate for
+\(T_2=1000.0\) units of time. We then calculate the covariance between
+the quantities \(\alpha\) and \(\mathfrak{C}\) across \(S=100\) species
+for each of the last \(T_2\) time steps. We assume the temporal average
+of these covariances across the last \(T_2\) units of time approximates
+the expectation at equilibrium. We repeated this approach for randomly
+drawn \(a\) and \(c\) until our sample size reached \(1000\). In Figure
+\ref{cor_coev_fig} we plot the temporally averaged values of
+\(\mathrm{Cov}_{f_{\bar X}}(\alpha,\mathfrak{C})\) against the
+sensitivity to competition \(c\). Using a cubic regression, we see the
+correlation of coevolutionary selection gradients and competition
+coefficeints is negative at variance ratios below \(0.5\), zero at
+variance ratios between \(0.5\) and \(1.0\), and drops below zero again
+at variance ratios above \(1.0\).
+
+\begin{figure}
+
+{\centering \includegraphics{/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/Cαℭp}
+
 }
-require(ggplot2)
-require(ggthemes)
-require(gridExtra)
-dummy <- ggplot(data = data.frame(x = 0), mapping = aes(x = x))
-cov_ab_p <- dummy + stat_function(fun = Cov_ab) + ylim(-0.4,0.4) + xlim(0,2) + theme_minimal() + xlab("Approximate ratio of variance among \n species to variance within species") + ylab("Covariance of \u03B1 and |\u03B2|")
-cov_ag_p <- dummy + stat_function(fun = Cov_ag) + ylim(-0.4,0.4) + xlim(0,2) + theme_minimal() + xlab("Approximate ratio of variance among \n species to variance within species") + ylab("Covariance of \u03B1 and \u03B3")
-cov_plot <- grid.arrange(cov_ab_p,cov_ag_p,nrow=1) 
-ggsave("~/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/cov.png",cov_plot,width=6.5,height=2.5)
-```
-
-```{r, echo=F, fig.align='center', fig.cap='\\label{cov_fig}Curves representing the covariance between the magnitude of linear selection gradients and competition coefficients (left) and between stabilizing selection gradients and competition coefficients (right) as a function of $2bV_{\\bar X}$ which is approximately equal to the ratio of variance in mean traits among species to the intraspecific trait variance.
-In both plots we set $c=1.0\\times10^{-4}$, $b=0.1$, $\\bar r=0.1$ and $N=1.0\\times10^{10}$ and let $V_{\\bar X}$ vary.'}
-knitr::include_graphics('/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/cov.png')
-```
-
-For fixed $c,b,\bar r$ and $N$, we visualize these relationships in Figure \ref{cov_fig}.
-To gain insight into the relationship between selection gradients and competition coefficients, note that our assumptions in this section imply $b^{-1}=2(\sigma^2+w)$.
-If we further assume $\sigma^2+w\approx\sigma^2$, then $2bV_{\bar X}\approx V_{\bar X}/\sigma^2$.
-That is, when populations are generalists and are comprised of specialist individuals, the value $2bV_{\bar X}$ is approximately equal to the ratio of interspecific mean trait variation to intraspecific individual trait variation.
-Hence, for both covariances we see that there is no relationship between selection gradients and competition coefficients when this ratio is zero.
-From equation (\ref{cov_alpha_beta}) we can use numerical optimization to find that when $V_{\bar X}/\sigma^2\approx1.25$ the relationship between the magnitudes of linear selection gradients and competition coefficients disappears, but when (approximately) $V_{\bar X}/\sigma^2<1.25$ ($>1.25$), this covariance becomes negative (positive).
-Equation (\ref{cov_alpha_gamma}) states that when $V_{\bar X}/\sigma^2$ is approximately equal to one (or slightly larger), there is no expected relationship between competition coefficients and quadratic selection gradients.
-However, when $V_{\bar X}/\sigma^2<1.0$ ($>1.0$), then we expect a positive (negative) relationship between $\alpha$ and $\gamma$.
-These results are true regardless of the chosen parameter values.
-In SM \S\ref{coef_grad_moms} we use simulations of system (\ref{comm_dynamics}) to show that these results do not qualitatively differ when allowing for heterogeneous population sizes and additive genetic variances across species.
-
-From a biological perspective, if the ratio $V_{\bar X}/\sigma^2$ is small, then species are packed tightly in phenotypic space.
-In our model this occurs when abiotic stabilizing selection is much stronger than competition ($a\gg c$).
-This causes species to overlap more in niche space (i.e., large $\alpha$) and creates disruptive selection for greater intraspecific variance (i.e., positive $\gamma$), which explains the positive region of covariance between $\alpha$ and $\gamma$.
-However, as species begin to overlap in niche space, directional selection begins to vanish (i.e., small $|\beta|$), leading to a negative covariance between $\alpha$ and $|\beta|$.
-In the limiting case that two species have perfectly overlapping niches, they will exhibit zero directional selection since a shift in either direction will yield the same fitness advantages.
-
-In the opposite scenario where competition is much stronger than abiotic stabilizing selection ($c\gg a$), species will not evolve to be as tightly packed and instead their niche-centers will be spread out with little ovelap in their resouce utilization curves (i.e., small $\alpha$).
-In this case biotic directional selection will be strong (i.e., large $|\beta|$), particularly for species towards the outer regions of niche space due to asymmetric fitness advantages conferred by shifts in niche-centers.
-This leads to a positive covariance between $\alpha$ and $|\beta|$.
-However, as noted above, this directional selection will also erode away at standing heritable variation (i.e., negative $\gamma$), reducing the rate at which adaptation can occur and creating a negative covariance between $\alpha$ and $\gamma$.
-
-In summary, we see the relation between competition coefficients and selection is highly non-trivial and depends on the relative magnitudes of different ecological processes shaping the community.
-However, this does not address the relation between competition coefficients and coevolution per se.
-In SM \S\ref{coef_grad_moms} we show that calculating a formula for the covariance between competition coefficients and the metric of coevolution $\mathfrak{C}$ introduced above provides a difficult analytical challenge.
-Instead of confronting this challenge we build on our numerical approach used to justify analytical approximations of $\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)$ and $\mathrm{Cov}_{f_{\bar X}}(\alpha,\gamma)$ to approximate the correlation of $\alpha$ and $\mathfrak{C}$.
-This numerical approach inherits the assumptions of homogeneous background parameters such as the mutation rate $\mu$ and abiotic optima $\theta$, but allows us to relax the assumption that $N$ and $G$ are constant across species and time.
-
-In particular, we numerically integrated system (\ref{comm_dynamics}) for $T_1=1000.0$ units of time and then continued to integrate for $T_2=1000.0$ units of time.
-We then calculate the covariance between the quantities $\alpha$ and $\mathfrak{C}$ across $S=100$ species for each of the last $T_2$ time steps.
-We assume the temporal average of these covariances across the last $T_2$ units of time approximates the expectation at equilibrium.
-We repeated this approach for randomly drawn $a$ and $c$ until our sample size reached $1000$.
-In Figure \ref{cor_coev_fig} we plot the temporally averaged values of $\mathrm{Cov}_{f_{\bar X}}(\alpha,\mathfrak{C})$ against the sensitivity to competition $c$.
-Using a cubic regression, we see the correlation of coevolutionary selection gradients and competition coefficeints is negative at variance ratios below $0.5$, zero at variance ratios between $0.5$ and $1.0$, and drops below zero again at variance ratios above $1.0$.
-
-```{r, eval=F, echo=F}
-require(ggplot2)
-require(gridExtra)
-df = read.csv("~/Research/Branching Brownian Motion/corrs.csv")
-Cαβp = ggplot(df,aes(x=df$V/df$σ,y=df$Cαβ))+geom_point(size=0.1)+stat_smooth(color="black", method = "lm", formula = y ~ x + I(x^2) + I(x^3))+theme_minimal() + ylab("Correlation of \u03B1 and \u03B2")+xlab("Ratio of variance among \n species to variance within species")
-Cαabsβp = ggplot(df,aes(x=df$V/df$σ,y=df$Cαabsβ))+geom_point(size=0.1)+stat_smooth(color="black", method = "lm", formula = y ~ x + I(x^2) + I(x^3))+theme_minimal()+ylab("Correlation of \u03B1 and |\u03B2|")+xlab("Ratio of variance among \n species to variance within species")
-Cαγp = ggplot(df,aes(x=df$V/df$σ,y=df$Cαγ))+geom_point(size=0.1)+stat_smooth(color="black", method = "lm", formula = y ~ x + I(x^2) + I(x^3))+theme_minimal()+ylab("Correlation of \u03B1 and \u03B3")+xlab("Ratio of variance among \n species to variance within species")
-Cαℭp = ggplot(df,aes(x=df$V/df$σ,y=df$Cαℭ))+geom_point(size=0.1)+stat_smooth(color="black", method = "lm", formula = y ~ x + I(x^2) + I(x^3))+theme_minimal()+ylab("Correlation of \u03B1 and ℭ")+xlab("Ratio of variance among \n species to variance within species")
-corrsp = grid.arrange(Cαabsβp,Cαγp,nrow=1)
-ggsave("/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/corrs.png",  corrsp,  width=6.5,height=2.5)
-ggsave("/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/Cαℭp.png",   Cαℭp,  width=4,height=2.5)
-```
-
-```{r, echo=F, fig.align='center', fig.cap='\\label{cor_coev_fig}Numerical estimates for the correlation of the strength of coevolution measured by $\\mathfrak{C}$ and competition coefficeints $\\alpha$ plotted against the variance of mean traits among species divided by the mean variance of traits within species.
+
+\caption{\label{cor_coev_fig}Numerical estimates for the correlation of the strength of coevolution measured by $\mathfrak{C}$ and competition coefficeints $\alpha$ plotted against the variance of mean traits among species divided by the mean variance of traits within species.
 Each dot represents the result from a single simulation.
-The red line is a cubic regression.'}
-knitr::include_graphics('/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/Cαℭp.png')
-```
-
-# Conclusion \label{conclusion}
-
-We have introduced an approach to derive models of evolutionary ecology using the calculus of white noise and coined SAGA, a SPDE model of phenotypic evolution that accounts for demographic stochasticity.
-From SAGA we derived SDE that track the evolution of abundance, mean trait and additive genetic variance.
-Observing the expressions of these SDE we find the effects of demographic stochasticity on the evolution of mean trait and additive genetic variance characterize the effects of random genetic drift.
-By considering the statistics of random samples, this has previously been in models of fixed population size in discrete time [@Lande1976].
-However, we have shown random genetic drift is a result of demographic stochasticity for populations with fluctuating abundances in continuous time.
-To illustrate the relevance of our approach to studies of evolutionary ecology, we comined our SDE with classical competition theory to derive a model of diffuse coevolution and used this model to investigate the relationship between selection gradients and competition coefficients.
-This work demonstrates that connecting contemporary theoretical approaches to evolutionary ecology with some fundamental results in the theory of measure-valued stochastic processes allows for the development of a rigorous, yet flexible approach to synthesizing the dynamics of abundance and distributions of quantitative characters, in agreement with Champagnat, Ferrière, and Méléard [-@Champagnat2006].
-Although this approach enjoys these important merits, there remains biological details and their associated technical challenges that need to be confronted for gaining a more thorough and accurate understanding of ecological communities.
-We touch on just four of them here and provide suggestions for future research to approach these challenges.
-For the first three, we recommend the use of individual-based simulations of measure-valued processes to confront these challenges.
-To summarize why, we state here that analytical results become difficult to derive for small populations and complicated models of mutation and selection.
-
-__Limitations of diffusion limits and generalizing coexistence theory__
-
-As noted by Feller [-@feller1951], diffusion limits provide reasonable approximations for large populations, but relatively small populations require discrete models.
-Hence, as a diffusion limit, SPDE (\ref{SPDE}) restricts parameter values to regions that maintain large population sizes.
-Although this suggests an important restriction on any model developed under this framework by implying populations are not at risk of extinction, the SDE describing abundance dynamics technically permits extinction.
-However, for small abundances, pathology emerges in the evolution of trait means and variances.
-In particular, stochastic components of the SDE describing the evolution of $\bar x$ and $\sigma^2$ diverge towards infinite values as $N\to0$.
-Hence, studies involving extinction and phenotypic evolution should be pursued through a different approach.
-A natural alternative can be developed utilizing the underlying measure-valued branching process that SAGA is a diffusion limit of (see section \ref{stochastic}).
-This process explicitly tracks individuals throughout their life-history and captures reproduction as branching events.
-Hence, branching processes model population size as a non-negative integer instead of a continuously varying number.
-In particular, the pathological behavior described for the SDE of $\bar x$ and $\sigma^2$ does not occur for measure-valued branching processes.
-This concern relates to our comment in \S\ref{ecoevo} on using our model of diffuse coevolution to extend modern coexistence theory [sensu @Chesson2000] in a way that accounts for phenotypic evolution and demographic stochasticity.
-Just as previous work has sought to understand the implications of processes for coexistence separately, it seems necessary to determine whether or not results pertaining to the coexistence of competing populations hold using models that account for phenotypic evolution and demographic stochasticity simultaneously.
-Because such investigations consider both extinction and phenotypic evolution we argue the models used should be either simulations of, or whenever possible, analytical results based on measure-valued branching processes.
-
-__The genetic architecture and distributions of traits__
-
-Our treatment of inheritance and our approach to model coevolution rests on the assumptions of normally distributed breeding values and expressed phenotypes along with asexual reproduction.
-However, real traits are not encoded by an infinite number of loci each contributing an additive infinitesimal effect (as required by the infinitesimal model), mutations are not inherited as normally distributed deviations from parental breeding values (as required by the Gaussian allelic model) and many populations of interest exhibit non-random sexual reproduction.
-Departures from this model of genetic architecture can produce more complex distributions of breeding values and expressed traits.
-Such deviations can be reinforced via strong non-Gaussian selection surfaces, including the surface $m(\nu,x)$ we have derived from niche theory, along with non-random mating in sexually reproducing populations.
-However, Gaussian approximations are convenient since they are defined by their mean and variance.
-Future work investigating the effects of non-normally distributed traits on the structure and dynamics of ecological communities will therefore need to confront higher moments such as skew and kurtosis, ideally integrating previously established mathematical approaches to derive the dynamics of these higher moments [@Dbarre2015EvolutionOQ].
-
-Another approach to breaking the assumption of normally distributed trait values is the development of explicit multilocus models.
-These models describe the contributions of alleles at particular loci in the genome to the development of quantitative traits.
-Tracking the fluctuations of allele frequencies then allows theoreticians to investigate the dynamics of phenotypic distributions that deviate from normality.
-This approach has a long history in theoretical quantitative genetics [@bulmer1980; @pmid7851785; @Kirkpatrick1727] and coevolutionary theory [@Nuismer2005; @Kopp2006; @Nuismer2007].
-However, work to investigate the role of complex genetic architectures in mediating feedbacks between the dynamics of population abundances and the distributions of traits mediating ecological interactions has apparently only just begun [@Patel2019].
-Since analytical results are difficult to obtain for general models of mutation, development and selection [c.f., @9780471986539], we again suggest the use of measure-valued branching processes and their individual-based simulations.
-For example, by assuming the mutational effects in the BBM model is Gaussian with variance an order of magnitude greater than the equilibrium phenotypic (or genetic) variance, we obtain the so-called _House-of-cards_ model [@Turelli1984].
-However, perhaps considering heavy-tailed distributions, such as the Laplace-distributions considered in Park et.
-al.
-[-@Park2017], will provide more realistic, yet less analytically tractable models of mutational effects.
-Such models have important implications in conservation genetics where small populations may evolve primarily in response to mutation and random genetic drift [@Lande1995].
-
-__Ecological stoichiometry__
-
-Our treatment of both biotic and abiotic selection neglects important chemical constraints of biological reality.
-For instance, the resource we assume species are competing over is modelled as a static quantity.
-However, real resources can be dynamic quantities.
-Such theoretical quantities may reflect abiotic cycles of material and energy or whole trophic layers comprised of prey, hosts or mutualists.
-Although resource dynamics have been captured theoretically in consumer-resource models [@davidtilman1982], developing a more realistic model of resource competition must incorporate details on the ecophysiology of organisms [@michelloreau2010].
-Doing so will help clarify the feedback between the evolution of populations and the ecosystem processes they are a part of.
-
-Using plant-pollinator communities as an example, the role of nitrogen mediating interspecific interactions has been reviewed by David, Storkey and Stevens [-@David2019] and the evolutionary ecology of the nutritional content of nectar has been reviewed by Parachnowitsch, Manson and Sletvold [-@Parachnowitsch2018].
-These studies demonstrate the need for further research to understand how soil nutrient availability and organismal ecophysiology structures communities of plants and pollinators.
-Theoretical pursuits in this direction may profit from interfacing the framework we have outlined here with population-ecosystem models such as that developed by Fridley [-@Fridley2017].
-However, these population-ecosystem models will likely yield complicated fitness functions that are difficult to analyze using rigorous mathematical techniques.
-Hence, we again recommend the use of simulations of measure-valued branching processes that generalize BBM to understand when analytical approximations hold.
-
-__Accounting for macroevolutionary history__
-
-To understand patterns found in ecological communities, considerations must push beyond microevolutionary and contemporary ecological processes and consider the macroevolutionary dynamics of ancestral lineages leading to extant populations.
-Using sub-alpine flower communities as an example, one can observe a very strict ordering of phenology across a broad geographic range.
-In particular, whether in the Colorado Rocky mountains (such as Gothic, Colorado) or on an outlier of the Idaho batholith (such as Kamiak butte near Palouse, Washington), one will almost surely observe a very conspicuous order of flowers emerging in early spring: at the very beginning of the season blooms _Claytonia lanceolata_ followed by _Erythronium grandiflorum_ which precedes _Delphinium nuttallianum_ (B.
-Week, personal observations).
-If contemporary phenological coevolution is rampant, why should this pattern be so well preserved across a thousand miles of rugged and diverse terrain?
-Although it would be exciting to find that these species repeatedly coevolved this pattern in each location, a more parsimoniuous hypothesis suggests the phenology and physiology of these species slowly evolved independently over macroevolutionary time scales to take advantage of the specific conditions available within each of these windows of the flowering season.
-However, this could not have carried out in the Rocky mountains since this terrain only became habitable just over ten thousand years ago as the glaciers of the Pleistocene began to retreat (Paul CaraDonna, personal communications).
-Hence, given these considerations, it appears that an understanding of early season phenology patterns should focus on how these communities are assembled as opposed to contemporary evolutionary dynamics.
-Indeed, recent work testing models of phylogeography ignores the potential for contemporary evolution and instead suggests alpine flower communities tend to follow neutral assembly where flowers merely compete for who can disperse to new habitat first, as opposed to a selective process where a regional species pool is filtered for those species adapted to the newly available habitat [@Marx2017].
-
-Of course microevolutionary and ecological dynamics are not completely irrelevant for understanding patterns in communities that are primarily structured by deep evolutionary processes.
-In particular, macroevolutionary trait evolution is simply the aggregation of microevolutionary change occuring over large spans of time.
-This suggests a road forward to connect the theory we have introduced to models of macroevolutionary trait evolution.
-
-Some approaches to modelling macroevolutionary trait change simply repurpose microevolutionary models by blindly rescaling time from the units of generations to millions of years [@Nuismer2014; Luke, can you think of others?].
-Such an approach makes the implicit assumption that trait evolution is statistically self-similar [sensu @kennethfalconer2014] so that the stochastic evolution of traits on macroevolutionary time scales has the same properties of trait evolution on microevolutionary time scales.
-Although some stochastic processes, including Brownian motion, do exhibit self-similarity, others do not.
-For example, consider a modification of the Ornstein-Uhlenbeck process defined by the SDE
+The red line is a cubic regression.}\label{fig:unnamed-chunk-12}
+\end{figure}
+
+\hypertarget{conclusion}{%
+\section{\texorpdfstring{Conclusion
+\label{conclusion}}{Conclusion }}\label{conclusion}}
+
+We have introduced an approach to derive models of evolutionary ecology
+using the calculus of white noise and coined SAGA, a SPDE model of
+phenotypic evolution that accounts for demographic stochasticity. From
+SAGA we derived SDE that track the evolution of abundance, mean trait
+and additive genetic variance. Observing the expressions of these SDE we
+find the effects of demographic stochasticity on the evolution of mean
+trait and additive genetic variance characterize the effects of random
+genetic drift. By considering the statistics of random samples, this has
+previously been in models of fixed population size in discrete time
+(Lande 1976). However, we have shown random genetic drift is a result of
+demographic stochasticity for populations with fluctuating abundances in
+continuous time. To illustrate the relevance of our approach to studies
+of evolutionary ecology, we comined our SDE with classical competition
+theory to derive a model of diffuse coevolution and used this model to
+investigate the relationship between selection gradients and competition
+coefficients. This work demonstrates that connecting contemporary
+theoretical approaches to evolutionary ecology with some fundamental
+results in the theory of measure-valued stochastic processes allows for
+the development of a rigorous, yet flexible approach to synthesizing the
+dynamics of abundance and distributions of quantitative characters, in
+agreement with Champagnat, Ferrière, and Méléard (2006). Although this
+approach enjoys these important merits, there remains biological details
+and their associated technical challenges that need to be confronted for
+gaining a more thorough and accurate understanding of ecological
+communities. We touch on just four of them here and provide suggestions
+for future research to approach these challenges. For the first three,
+we recommend the use of individual-based simulations of measure-valued
+processes to confront these challenges. To summarize why, we state here
+that analytical results become difficult to derive for small populations
+and complicated models of mutation and selection.
+
+\textbf{Limitations of diffusion limits and generalizing coexistence
+theory}
+
+As noted by Feller (1951), diffusion limits provide reasonable
+approximations for large populations, but relatively small populations
+require discrete models. Hence, as a diffusion limit, SPDE (\ref{SPDE})
+restricts parameter values to regions that maintain large population
+sizes. Although this suggests an important restriction on any model
+developed under this framework by implying populations are not at risk
+of extinction, the SDE describing abundance dynamics technically permits
+extinction. However, for small abundances, pathology emerges in the
+evolution of trait means and variances. In particular, stochastic
+components of the SDE describing the evolution of \(\bar x\) and
+\(\sigma^2\) diverge towards infinite values as \(N\to0\). Hence,
+studies involving extinction and phenotypic evolution should be pursued
+through a different approach. A natural alternative can be developed
+utilizing the underlying measure-valued branching process that SAGA is a
+diffusion limit of (see section \ref{stochastic}). This process
+explicitly tracks individuals throughout their life-history and captures
+reproduction as branching events. Hence, branching processes model
+population size as a non-negative integer instead of a continuously
+varying number. In particular, the pathological behavior described for
+the SDE of \(\bar x\) and \(\sigma^2\) does not occur for measure-valued
+branching processes. This concern relates to our comment in
+\S\ref{ecoevo} on using our model of diffuse coevolution to extend
+modern coexistence theory (sensu Chesson 2000) in a way that accounts
+for phenotypic evolution and demographic stochasticity. Just as previous
+work has sought to understand the implications of processes for
+coexistence separately, it seems necessary to determine whether or not
+results pertaining to the coexistence of competing populations hold
+using models that account for phenotypic evolution and demographic
+stochasticity simultaneously. Because such investigations consider both
+extinction and phenotypic evolution we argue the models used should be
+either simulations of, or whenever possible, analytical results based on
+measure-valued branching processes.
+
+\textbf{The genetic architecture and distributions of traits}
+
+Our treatment of inheritance and our approach to model coevolution rests
+on the assumptions of normally distributed breeding values and expressed
+phenotypes along with asexual reproduction. However, real traits are not
+encoded by an infinite number of loci each contributing an additive
+infinitesimal effect (as required by the infinitesimal model), mutations
+are not inherited as normally distributed deviations from parental
+breeding values (as required by the Gaussian allelic model) and many
+populations of interest exhibit non-random sexual reproduction.
+Departures from this model of genetic architecture can produce more
+complex distributions of breeding values and expressed traits. Such
+deviations can be reinforced via strong non-Gaussian selection surfaces,
+including the surface \(m(\nu,x)\) we have derived from niche theory,
+along with non-random mating in sexually reproducing populations.
+However, Gaussian approximations are convenient since they are defined
+by their mean and variance. Future work investigating the effects of
+non-normally distributed traits on the structure and dynamics of
+ecological communities will therefore need to confront higher moments
+such as skew and kurtosis, ideally integrating previously established
+mathematical approaches to derive the dynamics of these higher moments
+(Débarre, Yeaman, and Guillaume 2015).
+
+Another approach to breaking the assumption of normally distributed
+trait values is the development of explicit multilocus models. These
+models describe the contributions of alleles at particular loci in the
+genome to the development of quantitative traits. Tracking the
+fluctuations of allele frequencies then allows theoreticians to
+investigate the dynamics of phenotypic distributions that deviate from
+normality. This approach has a long history in theoretical quantitative
+genetics (Bulmer 1980; Turelli and Barton 1994; Kirkpatrick, Johnson,
+and Barton 2002) and coevolutionary theory (Nuismer, Doebeli, and
+Browning 2005; Kopp and Gavrilets 2006; Nuismer, Ridenhour, and Oswald
+2007). However, work to investigate the role of complex genetic
+architectures in mediating feedbacks between the dynamics of population
+abundances and the distributions of traits mediating ecological
+interactions has apparently only just begun (Patel and Bürger 2019).
+Since analytical results are difficult to obtain for general models of
+mutation, development and selection (c.f., Bürger 2000), we again
+suggest the use of measure-valued branching processes and their
+individual-based simulations. For example, by assuming the mutational
+effects in the BBM model is Gaussian with variance an order of magnitude
+greater than the equilibrium phenotypic (or genetic) variance, we obtain
+the so-called \emph{House-of-cards} model (Turelli 1984). However,
+perhaps considering heavy-tailed distributions, such as the
+Laplace-distributions considered in Park et. al. (2017), will provide
+more realistic, yet less analytically tractable models of mutational
+effects. Such models have important implications in conservation
+genetics where small populations may evolve primarily in response to
+mutation and random genetic drift (Lande 1995).
+
+\textbf{Ecological stoichiometry}
+
+Our treatment of both biotic and abiotic selection neglects important
+chemical constraints of biological reality. For instance, the resource
+we assume species are competing over is modelled as a static quantity.
+However, real resources can be dynamic quantities. Such theoretical
+quantities may reflect abiotic cycles of material and energy or whole
+trophic layers comprised of prey, hosts or mutualists. Although resource
+dynamics have been captured theoretically in consumer-resource models
+(Tilman 1982), developing a more realistic model of resource competition
+must incorporate details on the ecophysiology of organisms (Loreau
+2010). Doing so will help clarify the feedback between the evolution of
+populations and the ecosystem processes they are a part of.
+
+Using plant-pollinator communities as an example, the role of nitrogen
+mediating interspecific interactions has been reviewed by David, Storkey
+and Stevens (2019) and the evolutionary ecology of the nutritional
+content of nectar has been reviewed by Parachnowitsch, Manson and
+Sletvold (2018). These studies demonstrate the need for further research
+to understand how soil nutrient availability and organismal
+ecophysiology structures communities of plants and pollinators.
+Theoretical pursuits in this direction may profit from interfacing the
+framework we have outlined here with population-ecosystem models such as
+that developed by Fridley (2017). However, these population-ecosystem
+models will likely yield complicated fitness functions that are
+difficult to analyze using rigorous mathematical techniques. Hence, we
+again recommend the use of simulations of measure-valued branching
+processes that generalize BBM to understand when analytical
+approximations hold.
+
+\textbf{Accounting for macroevolutionary history}
+
+To understand patterns found in ecological communities, considerations
+must push beyond microevolutionary and contemporary ecological processes
+and consider the macroevolutionary dynamics of ancestral lineages
+leading to extant populations. Using sub-alpine flower communities as an
+example, one can observe a very strict ordering of phenology across a
+broad geographic range. In particular, whether in the Colorado Rocky
+mountains (such as Gothic, Colorado) or on an outlier of the Idaho
+batholith (such as Kamiak butte near Palouse, Washington), one will
+almost surely observe a very conspicuous order of flowers emerging in
+early spring: at the very beginning of the season blooms \emph{Claytonia
+lanceolata} followed by \emph{Erythronium grandiflorum} which precedes
+\emph{Delphinium nuttallianum} (B. Week, personal observations). If
+contemporary phenological coevolution is rampant, why should this
+pattern be so well preserved across a thousand miles of rugged and
+diverse terrain? Although it would be exciting to find that these
+species repeatedly coevolved this pattern in each location, a more
+parsimoniuous hypothesis suggests the phenology and physiology of these
+species slowly evolved independently over macroevolutionary time scales
+to take advantage of the specific conditions available within each of
+these windows of the flowering season. However, this could not have
+carried out in the Rocky mountains since this terrain only became
+habitable just over ten thousand years ago as the glaciers of the
+Pleistocene began to retreat (Paul CaraDonna, personal communications).
+Hence, given these considerations, it appears that an understanding of
+early season phenology patterns should focus on how these communities
+are assembled as opposed to contemporary evolutionary dynamics. Indeed,
+recent work testing models of phylogeography ignores the potential for
+contemporary evolution and instead suggests alpine flower communities
+tend to follow neutral assembly where flowers merely compete for who can
+disperse to new habitat first, as opposed to a selective process where a
+regional species pool is filtered for those species adapted to the newly
+available habitat (Marx et al. 2017).
+
+Of course microevolutionary and ecological dynamics are not completely
+irrelevant for understanding patterns in communities that are primarily
+structured by deep evolutionary processes. In particular,
+macroevolutionary trait evolution is simply the aggregation of
+microevolutionary change occuring over large spans of time. This
+suggests a road forward to connect the theory we have introduced to
+models of macroevolutionary trait evolution.
+
+Some approaches to modelling macroevolutionary trait change simply
+repurpose microevolutionary models by blindly rescaling time from the
+units of generations to millions of years {[}Nuismer and Harmon (2014);
+Luke, can you think of others?{]}. Such an approach makes the implicit
+assumption that trait evolution is statistically self-similar (sensu
+Falconer 2014) so that the stochastic evolution of traits on
+macroevolutionary time scales has the same properties of trait evolution
+on microevolutionary time scales. Although some stochastic processes,
+including Brownian motion, do exhibit self-similarity, others do not.
+For example, consider a modification of the Ornstein-Uhlenbeck process
+defined by the SDE
 
@@
 
-where $a,b>0$, $W_t$ is a standard Brownian motion and $\theta_t$ is itself a Markov process that takes normally distributed jumps centered on its current location at exponentially distributed time intervals.
-If we assume the rate $\lambda$ at which jumps occur is much smaller than $a$ and the variance in jumping is much larger than $b^2$, then, even though the sample paths of $X_t$ are actually continuous (if we zoom in close enough, they look like Brownian motion), over long intervals of time sample paths of $X_t$ will begin to appear to have periods of continuity interrupted by an occasional discontinuous jump and thus approach a qualitatively distinct process.
-These emergent properties can be formally characterized by L&eacute;vy processes and have been successfully employed in comparative phylogenetics to fit phenotypic data from extant populations and the fossil record [@Landis2017].
-It would therefore be interesting to investigate whether an application of a separation of time scales argument for the rate of environmental change ($\lambda$) versus the rate of evolutionary and ecological change ($a$) to microevolutionary models derived using our framework can be used to obtain macroevolutionary models that include not only mean trait evolution, but also the evolution of trait variance and abundance.
-The resulting macroevolutionary models can give rise to novel comparative phylogenetic methods and provide initial conditions for microevolutionary models that capture contemporary dynamics.
-
-__Final remarks__
-
-Although top-down approaches to community ecology, such as the Maximum Entropy Theory of Ecology [@Harte2011], have enjoyed some success in describing community-level patterns [@Harte2014; @Xiao2015], a mechanistic understanding of why these patterns emerge and how they will change remains a formidable task.
-Such an understanding must take both bottom-up and top-down approaches integrating considerations from the ecophysiology of individual organisms that reveal the economics of interspecific interactions [@Sterner2008], to the phylogeographic history of taxa that sets the stage for contemporary dynamics [@Hickerson2010].
-Through connecting these dots we can increase the variance explained in observations of ecological communities by specific mechanisms and come closer to a predictive theory of evolutionary community ecology.
-Despite the long list of equations derived in this paper, this work takes just one small step towards capturing these many details.
-However, we hope the theoretical framework outlined here along with the demonstration of its use in modelling competitive communities provides some helpful tools to aid quantitative evolutionary ecologists in reaching such lofty goals.
+where \(a,b>0\), \(W_t\) is a standard Brownian motion and \(\theta_t\)
+is itself a Markov process that takes normally distributed jumps
+centered on its current location at exponentially distributed time
+intervals. If we assume the rate \(\lambda\) at which jumps occur is
+much smaller than \(a\) and the variance in jumping is much larger than
+\(b^2\), then, even though the sample paths of \(X_t\) are actually
+continuous (if we zoom in close enough, they look like Brownian motion),
+over long intervals of time sample paths of \(X_t\) will begin to appear
+to have periods of continuity interrupted by an occasional discontinuous
+jump and thus approach a qualitatively distinct process. These emergent
+properties can be formally characterized by Lévy processes and have been
+successfully employed in comparative phylogenetics to fit phenotypic
+data from extant populations and the fossil record (Landis and Schraiber
+2017). It would therefore be interesting to investigate whether an
+application of a separation of time scales argument for the rate of
+environmental change (\(\lambda\)) versus the rate of evolutionary and
+ecological change (\(a\)) to microevolutionary models derived using our
+framework can be used to obtain macroevolutionary models that include
+not only mean trait evolution, but also the evolution of trait variance
+and abundance. The resulting macroevolutionary models can give rise to
+novel comparative phylogenetic methods and provide initial conditions
+for microevolutionary models that capture contemporary dynamics.
+
+\textbf{Final remarks}
+
+Although top-down approaches to community ecology, such as the Maximum
+Entropy Theory of Ecology (Harte 2011), have enjoyed some success in
+describing community-level patterns (Harte and Newman 2014; Xiao,
+McGlinn, and White 2015), a mechanistic understanding of why these
+patterns emerge and how they will change remains a formidable task. Such
+an understanding must take both bottom-up and top-down approaches
+integrating considerations from the ecophysiology of individual
+organisms that reveal the economics of interspecific interactions
+(Sterner and Elser 2008), to the phylogeographic history of taxa that
+sets the stage for contemporary dynamics (Hickerson et al. 2010).
+Through connecting these dots we can increase the variance explained in
+observations of ecological communities by specific mechanisms and come
+closer to a predictive theory of evolutionary community ecology. Despite
+the long list of equations derived in this paper, this work takes just
+one small step towards capturing these many details. However, we hope
+the theoretical framework outlined here along with the demonstration of
+its use in modelling competitive communities provides some helpful tools
+to aid quantitative evolutionary ecologists in reaching such lofty
+goals.
 
@@
 
-# Supplementary material (SM)
-
-Throughout this supplement, we set use dot notation for time derivatives so that $\dot f(x,t)=\frac{\partial }{\partial t}f(x,t)$ and set $\Delta=\sum_{i=1}^d\frac{\partial^2}{\partial x_i^2}$ the Laplace operator on $\mathbb{R}^d$, except in \S\ref{caC} where $\Delta$ represents a random variable.
-
-<!--
-
-## Existence, uniqueness and regularity of solution to deterministic PDE (\ref{eq1}) of the main text \label{regular}
-
-Here we show that solutions to the PDE described in the main text exist, are unique and are twice continuously differentiable with respect to space.
-Our approach is an adaptation of that taken in chapter 9 \S2.2 of Volpert [-@Volpert2014].
-
-Recall that the PDE we consider is of the form
-\begin{equation}\label{PDE_supp}
-\begin{matrix}
-\dot\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\Delta\nu(x,t), \\
-\\
-\nu(x,0)=u(x),
-\end{matrix}
-\end{equation}
-where $m(\nu,x)\leq r\in\mathbb{R}$ is shorthand for $m((K\nu)(x,t),x)$ and $(K\nu)(x,t)=\int_\mathbb{R}\kappa(x-y)\nu(y,t)dy$ for some non-negative, bounded $\kappa$.
-Denoting $C(\mathbb{R})$ the set of continuous functions on $\mathbb{R}$ and $L^1(\mathbb{R})$ the Lebesgue space of integrable functions on $\mathbb{R}$ (ie, the set of functions $f:\mathbb{R}\to\mathbb{R}$ that satisfy $\|f\|_1=\int_\mathbb{R}|f(x)|dx<+\infty$ under the equivalence relation $f=g$ iff $\|f-g\|_1=0$), we assume a continuous and integrable initial condition $\nu(\cdot,0)=u\in L^1(\mathbb{R})\cap C(\mathbb{R})$.
-We now prove the existence and uniqueness of solutions to this Cauchy problem.
-
-Set 
-\begin{equation}
-\Phi(x,t)=\frac{\exp\left(-x^2/2\mu t\right)}{\sqrt{2\pi\mu t}}.
-\end{equation}
-Then the semigroup $\{S(t)\}_{t\geq0}$ given by 
-\begin{equation}
-S(t)v=\left\{\begin{matrix}
-\int_0^t\frac{\mu}{2}\Delta\nu(x,s)ds=\int_\mathbb{R}\Phi(x-y,t)v(y)dy, & t>0 \\
- & \\
-v, &t=0
-\end{matrix}\right.
-\end{equation}
-for all $v\in L^\infty(\mathbb{R})$ satisfies $\|S(t)v\|_1=\|v\|_1$ for all $t\geq0$ and $v\in L^1(\mathbb{R})$. We also define the operator
-\begin{equation}
-(Fv)(x)=m\left(\int_\mathbb{R}\kappa(x-y)v(y)dy,x\right)v(x), \ \ \forall \ \ v\in L^1(\mathbb{R}).
-\end{equation}
-Thus $\|Fv\|_1\leq |r| \ \|v\|_1$ for all $v\in L^1(\mathbb{R})$. Then, for any $v,w\in L^1(\mathbb{R})$,
-\begin{multline}
-Fv-Fw=F(v+w)-F(v+w)+Fv-Fw \\
-=m(v+w,x)(v+w)-m(w,x)w-m(v+w,x)(v+w)+m(v,x)v \\
-=\Big(m(v+w,x)-m(w,x)\Big)w-\Big(m(v+w,x)-m(v,x)\Big)v+m(v+w,x)(v-w).
-\end{multline}
-Hence $\|Fv-Fw\|_1\leq2|r|(\|w\|_1+\|v\|_2)+|r|\|v-w\|_1\leq\|r\|\|v-w\|_1$ which implies $F$ is Lipschitz continuous on $L^1(\mathbb{R})$.
-For $T>0$ set $L^{\infty,1}([0,T]\times\mathbb{R})$ as the set of functions $\nu:[0,T]\to L^1(\mathbb{R})$ that satisfy $\|\nu\|_{1,T}=\sup_{t\in[0,T]}\|\nu(t)\|_1<+\infty$ and consider the operator $G$ defined on $L^{\infty,1}([0,T]\times\mathbb{R})$ for fixed $u\in L^1(\mathbb{R})$ by
-\begin{equation}
-(Gv)(t)=S(t)u+\int_0^tS(t-s)(Fv(s))ds \ \ \forall \ \ t\in[0,T].
-\end{equation}
-Then for any $t\in[0,T]$ and $\nu_1,\nu_2\in L^{\infty,1}([0,T]\times\mathbb{R})$,
-
-\begin{multline}
-\|G\nu_1-G\nu_2\|_{1,T}=\sup_{t\in[0,T]}\left\|\int_0^tS(t-s)(F\nu_1(s)-F\nu_2(s))ds\right\|_1 \\ \leq \sup_{t\in[0,T]}t|r|\|\nu_1(t)-\nu_2(t)\|_1\leq T|r|\|\nu_1-\nu_2\|_{1,T}.
-\end{multline}
-Hence, when $T<\frac{1}{|r|}$ the operator $G$ is a contraction operator.
-In this case we apply Banach's fixed point theorem to demonstrate $G$ has a unique fixed point.
-That is, there exists a unique $\nu\in L^{\infty,1}([0,T]\times\mathbb{R})$ such that $G\nu=\nu$.
-Then, for $\nu=\nu(x,t)$ and considering $\nu(t)$ the function $x\to\nu(x,t)$, we have by definition $\nu$ is the mild solution of PDE (\ref{PDE_supp}) with initial condition $\nu(0)=u$.
-To extend this to arbitrarily large $T$ we take a similar approach as Evans [-@lawrenceevans2010] and choose $T_1<\frac{1}{|r|}$ and apply the above argument to the interval $[T_1,2T_1]$.
-Repeating this task a finite number of times shows $\nu$ is the mild solution of PDE (\ref{PDE_supp}) with initial condition $\nu(0)=u$ for arbitrarily large $T>0$.
-
-We shift our focus to showing mild solutions $\nu$ to PDE (\ref{PDE_supp}) are twice differentiable in $x$ and hence are classical solutions.
-Applying lemma 3.3.2 of Henry [-@Henry1981] we have that the mild solution $\nu$ is in fact a strong solution (i.e., differentiable with respect to $t\in(0,T)$) to the abstract Cauchy problem defined by PDE (\ref{PDE_supp}) with initial condition $\nu(0)=u$ for $t\in[0,T]$.
-In particular, this implies that $\nu(t)\in D(\Delta)$ for each $t\in(0,T)$ where $D(\Delta)=W^{2,1}(\mathbb{R})$ is the Sobolov space of integrable functions with integrable weak derivatives at least of order $2$.
-Since $\frac{1}{p}\geq\frac{1}{p+1}\geq\frac{1}{p}-2$ for all $p\geq1$, the Sobolov embedding theorem [@Henry1981] implies $D(\Delta)\subset L^p(\mathbb{R})$, the Lebesgue space of $p$-integrable functions on $\mathbb{R}$, for all $p\geq1$.
-Hence $D(\Delta)\subset W^{2,p}(\mathbb{R})$, the Sobolov space of $p$-integrable functions with weak derivatives at least of order $2$, for all $p\geq1$.
-
-If $v\in C^{2+\delta}(\mathbb{R})$ for some $\delta>0$, then 
-\begin{equation}
-(Kv)(\cdot)=\int_\mathbb{R}\kappa(\cdot-y)v(y)dy=\int_\mathbb{R}\kappa(y)v(\cdot-y)dy\in C^{2+\delta}(\mathbb{R}),
-\end{equation}
-and, since we assume $m(h,x)$ is continuous in both arguments, $Fv=m(Kv,\cdot)v\in C^\delta$.
-
-Why is $F\nu\in C^\delta(\mathbb{R})$ for some $\delta>0$?
-Why does this imply $\nu(\cdot,t)\in C^{2+\delta}(\mathbb{R})$ for all $t\in(0,T)$?
-This argument holds for all $T>0$.
-Hence, solutions to the Cauchy problem (\ref{PDE_supp}) are twice continuously differentiable in their spatial parameter for all $t>0$.
-
--->
-
-## Sufficient conditions for finite mean, variance and total abundance in the deterministic case \label{finite}
-
-Recall that $m(\nu,x)$ is shorthand for $m\big((K\nu)(x,t),x\big)$.
-That is, $m:[0,+\infty)\times\mathbb{R}\to\mathbb{R}$.
-Following our assumptions of the main text, we have that $m(h,x)$ is differentiable with respect to both $x$ and $h$ and there exists $r\in\mathbb{R}$ such that $m(h,x)\leq r$ across all $h\geq0$ and $x\in\mathbb{R}$.
-As in the main text, we also assume the initial condition $u(x)$ is continuous and integrable in $x$ and satisfies
-\begin{equation}
+\hypertarget{supplementary-material-sm}{%
+\section{Supplementary material (SM)}\label{supplementary-material-sm}}
+
+Throughout this supplement, we set use dot notation for time derivatives
+so that \(\dot f(x,t)=\frac{\partial }{\partial t}f(x,t)\) and set
+\(\Delta=\sum_{i=1}^d\frac{\partial^2}{\partial x_i^2}\) the Laplace
+operator on \(\mathbb{R}^d\), except in \S\ref{caC} where \(\Delta\)
+represents a random variable.
+
+\hypertarget{sufficient-conditions-for-finite-mean-variance-and-total-abundance-in-the-deterministic-case}{%
+\subsection{\texorpdfstring{Sufficient conditions for finite mean,
+variance and total abundance in the deterministic case
+\label{finite}}{Sufficient conditions for finite mean, variance and total abundance in the deterministic case }}\label{sufficient-conditions-for-finite-mean-variance-and-total-abundance-in-the-deterministic-case}}
+
+Recall that \(m(\nu,x)\) is shorthand for \(m\big((K\nu)(x,t),x\big)\).
+That is, \(m:[0,+\infty)\times\mathbb{R}\to\mathbb{R}\). Following our
+assumptions of the main text, we have that \(m(h,x)\) is differentiable
+with respect to both \(x\) and \(h\) and there exists \(r\in\mathbb{R}\)
+such that \(m(h,x)\leq r\) across all \(h\geq0\) and \(x\in\mathbb{R}\).
+As in the main text, we also assume the initial condition \(u(x)\) is
+continuous and integrable in \(x\) and satisfies \begin{equation}
 0\leq\int_{\mathbb{R}}(|x|+x^2)u(x)dx<+\infty
-\end{equation}
-and consider the Cauchy problem
+\end{equation} and consider the Cauchy problem
 \begin{equation}\label{PDE_SM}
@@
 
-According to Theorem 2.5.6 of Zheng [-@zheng2004nonlinear], if the operator $F$ defined by $\nu(x,t)\to m(\nu,x)\nu(x,t)$ is locally Lipschitz, corresponding to equation (\ref{local_lipschitz}) of the main text, then for some maximal $T>0$ problem (\ref{PDE_SM}) admits a unique local classical solution $\nu(x,t)$ for $t\in[0,T)$.
-Furthermore, either $T=+\infty$ or $T<+\infty$ and $\lim_{t\uparrow T}\int_\mathbb{R}|\nu(x,t)|dx=+\infty$.
-In this section we show that our assumption $m(h,x)\leq r$ for all $h\geq0$ and $x\in\mathbb{R}$ implies $T=+\infty$.
-Replacing $m$ with it's upper bound $r\in\mathbb{R}$, equation (\ref{PDE_SM}) reduces to a simple parabolic equation that can be solved using elementary techniques [@stanleyfarlow1993].
-In particular, when $m(\nu,x)\equiv0$ denote the solution to (\ref{PDE_SM}) by $\nu_0(x,t)$.
-Then, denoting
+According to Theorem 2.5.6 of Zheng (2004), if the operator \(F\)
+defined by \(\nu(x,t)\to m(\nu,x)\nu(x,t)\) is locally Lipschitz,
+corresponding to equation (\ref{local_lipschitz}) of the main text, then
+for some maximal \(T>0\) problem (\ref{PDE_SM}) admits a unique local
+classical solution \(\nu(x,t)\) for \(t\in[0,T)\). Furthermore, either
+\(T=+\infty\) or \(T<+\infty\) and
+\(\lim_{t\uparrow T}\int_\mathbb{R}|\nu(x,t)|dx=+\infty\). In this
+section we show that our assumption \(m(h,x)\leq r\) for all \(h\geq0\)
+and \(x\in\mathbb{R}\) implies \(T=+\infty\). Replacing \(m\) with it's
+upper bound \(r\in\mathbb{R}\), equation (\ref{PDE_SM}) reduces to a
+simple parabolic equation that can be solved using elementary techniques
+(Farlow 1993). In particular, when \(m(\nu,x)\equiv0\) denote the
+solution to (\ref{PDE_SM}) by \(\nu_0(x,t)\). Then, denoting
 \begin{equation}
 \Phi(x,t)=\frac{\exp\left(-x^2/2\mu t\right)}{\sqrt{2\pi\mu t}},
-\end{equation}
-we have
-\begin{equation}
+\end{equation} we have \begin{equation}
 \nu_0(x,t)=\int_{\mathbb{R}}\Phi(x-y,t)\nu(y,0)dy.
-\end{equation}
-In the more general case, when $m(\nu,x)\equiv r\in\mathbb{R}$, equation (\ref{PDE_SM}) has the solution $\nu_r(x,t)=e^{rt}\nu_0(x,t)$.
-Hence, $\nu_r(x,t)\geq0$ for all $x\in\mathbb{R}$ and $\int_\mathbb{R}\nu_r(x,t)dx=e^{rt}N(0)<+\infty$ for all $t\geq0$.
-Furthermore, denoting
-\begin{equation}
+\end{equation} In the more general case, when
+\(m(\nu,x)\equiv r\in\mathbb{R}\), equation (\ref{PDE_SM}) has the
+solution \(\nu_r(x,t)=e^{rt}\nu_0(x,t)\). Hence, \(\nu_r(x,t)\geq0\) for
+all \(x\in\mathbb{R}\) and
+\(\int_\mathbb{R}\nu_r(x,t)dx=e^{rt}N(0)<+\infty\) for all \(t\geq0\).
+Furthermore, denoting \begin{equation}
 \begin{matrix}
@@
 \end{matrix}
-\end{equation}
-we have
-\begin{equation}
+\end{equation} we have \begin{equation}
 \bar x_r(t)=\int_\mathbb{R}x\int_\mathbb{R}\Phi(x-y,t)p(y,0)dydx=\int_\mathbb{R}yp(y,0)dy=\bar x(0),
-\end{equation}
-\begin{equation}
+\end{equation} \begin{equation}
 \sigma_r^2(t)=\int_\mathbb{R}(x-\bar x_r(t))^2\int_\mathbb{R}\Phi(x-y,t)p(y,0)dydx=\int_\mathbb{R}\Big((y-\bar x(0))^2+\mu t\Big)p(y,0)dy=\sigma^2(0)+\mu t.
-\end{equation}
-Hence, $|\bar x_r(t)|,\sigma_r^2(t)<+\infty$ for all $t\geq0$. For the sake of contradiction, suppose there exists $x\in\mathbb{R}$ and $t\geq0$ such that $\nu(x,t)>\nu_r(x,t)$. Then
-\begin{equation}
+\end{equation} Hence, \(|\bar x_r(t)|,\sigma_r^2(t)<+\infty\) for all
+\(t\geq0\). For the sake of contradiction, suppose there exists
+\(x\in\mathbb{R}\) and \(t\geq0\) such that \(\nu(x,t)>\nu_r(x,t)\).
+Then \begin{equation}
 \nu(x,t)-\nu(x,0)=\int_0^tm(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)ds>\int_0^tr\nu_r(x,s)+\frac{\mu}{2}\Delta\nu_r(x,s)ds=\nu_r(x,t)-\nu(x,0)
-\end{equation}
-which implies there exists $h\geq0$ and $x\in\mathbb{R}$ such that $m(h,x)>r$.
-But this contradicts our assumption $m(h,x)\leq r$ for all $h\geq0$ and $x\in\mathbb{R}$.
-So we have $\nu(x,t)\leq\nu_r(x,t)$ for each $x\in\mathbb{R}$ and $t\geq0$.
-This implies that $N(t)=\int_\mathbb{R}\nu(x,t)dx<+\infty$,
-\begin{equation}
+\end{equation} which implies there exists \(h\geq0\) and
+\(x\in\mathbb{R}\) such that \(m(h,x)>r\). But this contradicts our
+assumption \(m(h,x)\leq r\) for all \(h\geq0\) and \(x\in\mathbb{R}\).
+So we have \(\nu(x,t)\leq\nu_r(x,t)\) for each \(x\in\mathbb{R}\) and
+\(t\geq0\). This implies that
+\(N(t)=\int_\mathbb{R}\nu(x,t)dx<+\infty\), \begin{equation}
 0\leq\int_\mathbb{R}x^2\nu(x,t)dx\leq\int_\mathbb{R}x^2\nu_r(x,t)dx<+\infty
-\end{equation}
-and in particular
-\begin{equation}
+\end{equation} and in particular \begin{equation}
 0\leq\sigma^2(t)+\bar x^2(t)=\frac{1}{N(t)}\int_\mathbb{R}x^2\nu(x,t)dx<+\infty
-\end{equation}
-for each $t\geq0$.
-
-## Equilibrium moments for a deterministic population experiencing logistic growth and stabilizing selection \label{equilib}
-
-Here we set out to show, given the initial conditions $\nu(\cdot,0)\in L^1(\mathbb{R})$ such that $\bar x(0)\in\mathbb{R}$, $\sigma^2(0)\in[0,+\infty)$ and $N(0)\in(0,+\infty)$, and given the growth rate
+\end{equation} for each \(t\geq0\).
+
+\hypertarget{equilibrium-moments-for-a-deterministic-population-experiencing-logistic-growth-and-stabilizing-selection}{%
+\subsection{\texorpdfstring{Equilibrium moments for a deterministic
+population experiencing logistic growth and stabilizing selection
+\label{equilib}}{Equilibrium moments for a deterministic population experiencing logistic growth and stabilizing selection }}\label{equilibrium-moments-for-a-deterministic-population-experiencing-logistic-growth-and-stabilizing-selection}}
+
+Here we set out to show, given the initial conditions
+\(\nu(\cdot,0)\in L^1(\mathbb{R})\) such that
+\(\bar x(0)\in\mathbb{R}\), \(\sigma^2(0)\in[0,+\infty)\) and
+\(N(0)\in(0,+\infty)\), and given the growth rate
 \begin{equation}\label{m_log_stab}
 m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_\mathbb{R}\nu(y,t)dy
-\end{equation}
-such that $\theta\in\mathbb{R}$, $a,c,\mu>0$ and $r>\tfrac{1}{2}\sqrt{\mu a}$, we have the following stable equilibrium $N_\infty=\lim_{t\to\infty}N(t)=\tfrac{1}{c}(r-\tfrac{1}{2}\sqrt{a\mu})$, $\bar x_\infty=\lim_{t\to\infty}\bar x(t)=\theta$ and $\sigma^2_\infty=\lim_{t\to\infty}\sigma^2(t)=\sqrt{\tfrac{\mu}{a}}$.
-
-The mean fitness can be calculated as
-\begin{equation}
+\end{equation} such that \(\theta\in\mathbb{R}\), \(a,c,\mu>0\) and
+\(r>\tfrac{1}{2}\sqrt{\mu a}\), we have the following stable equilibrium
+\(N_\infty=\lim_{t\to\infty}N(t)=\tfrac{1}{c}(r-\tfrac{1}{2}\sqrt{a\mu})\),
+\(\bar x_\infty=\lim_{t\to\infty}\bar x(t)=\theta\) and
+\(\sigma^2_\infty=\lim_{t\to\infty}\sigma^2(t)=\sqrt{\tfrac{\mu}{a}}\).
+
+The mean fitness can be calculated as \begin{equation}
 \bar m(t)=r-\frac{a}{2}((\theta-\bar x(t))^2+\sigma^2(t))-cN(t).
-\end{equation}
-Recalling the ODE derived for $N(t)$,
-\begin{equation}
+\end{equation} Recalling the ODE derived for \(N(t)\), \begin{equation}
 \frac{d}{dt}N(t)=\bar m(t)N(t)=(r-\frac{a}{2}((\theta-\bar x(t))^2+\sigma^2(t))-cN(t))N(t),
-\end{equation}
-solving for equilibrium total abundance $\hat N$ amounts to setting $\frac{d}{dt}N(t)=0$ and solving for $N(t)$.
-Ignoring the equilibrium $N(t)=0$, this reduces to solving $\bar m(t)=0$ for $N(t)$, which returns
-\begin{equation}
+\end{equation} solving for equilibrium total abundance \(\hat N\)
+amounts to setting \(\frac{d}{dt}N(t)=0\) and solving for \(N(t)\).
+Ignoring the equilibrium \(N(t)=0\), this reduces to solving
+\(\bar m(t)=0\) for \(N(t)\), which returns \begin{equation}
 \hat N=\frac{1}{c}\left(r-\frac{a}{2}\left((\theta-\hat{\bar x})^2-\hat\sigma^2\right)\right).
-\end{equation}
-Unfortunately, deriving ODE for $\bar x(t)$ and $\sigma^2(t)$ leads to expressions involving higher moments and finding ODE for these higher moments will lead to expressions involving yet even higher moments.
-To avoid this infinite regression, we find the equilibrium abundance density $\hat\nu(x)$ by solving $\frac{\partial}{\partial t}\nu(x,t)=0$ for $\nu(x,t)$.
-This implies the following ordinary differential equation
-\begin{equation}
+\end{equation} Unfortunately, deriving ODE for \(\bar x(t)\) and
+\(\sigma^2(t)\) leads to expressions involving higher moments and
+finding ODE for these higher moments will lead to expressions involving
+yet even higher moments. To avoid this infinite regression, we find the
+equilibrium abundance density \(\hat\nu(x)\) by solving
+\(\frac{\partial}{\partial t}\nu(x,t)=0\) for \(\nu(x,t)\). This implies
+the following ordinary differential equation \begin{equation}
 \frac{d^2}{dx^2}\hat\nu(x)=\left(\frac{2c}{\mu}\hat N+\frac{a}{\mu}(\theta-x)^2-\frac{2r}{\mu}\right)\hat\nu(x)
-\end{equation}
-which has the solution
-\begin{equation}
+\end{equation} which has the solution \begin{equation}
 \hat\nu(x)=\frac{\hat N}{\sqrt{2\pi}}\left(\frac{a}{\mu}\right)^{\frac{1}{4}}\exp\left(-\sqrt\frac{a}{\mu}\frac{(\theta-x)^2}{2}\right).
-\end{equation}
-From this expression we infer $\hat{\bar x}=\theta$ and $\hat\sigma^2=\sqrt{\frac{\mu}{a}}$.
-Hence $\hat N=\frac{1}{c}\left(r-\frac{1}{2}\sqrt{a\mu}\right)$.
-To show that $N_\infty=\hat N$, $\bar x_\infty=\hat{\bar x}$ and $\sigma^2_\infty=\hat\sigma^2$, we linearize ODE for $N(t)$, $\bar x(t)$ and $\sigma^2(t)$ using the equilibrium $\hat\nu(x)$.
-In particular, since $\hat\nu(x)$ is Gaussian, we do not run into the same issue with higher moments as above.
-Assuming the equilibrium $\hat\nu(x)$ does not change the ODE for $N(t)$, but the ODE for $\bar x(t)$ and $\sigma^2(t)$ can now be expressed as
+\end{equation} From this expression we infer \(\hat{\bar x}=\theta\) and
+\(\hat\sigma^2=\sqrt{\frac{\mu}{a}}\). Hence
+\(\hat N=\frac{1}{c}\left(r-\frac{1}{2}\sqrt{a\mu}\right)\). To show
+that \(N_\infty=\hat N\), \(\bar x_\infty=\hat{\bar x}\) and
+\(\sigma^2_\infty=\hat\sigma^2\), we linearize ODE for \(N(t)\),
+\(\bar x(t)\) and \(\sigma^2(t)\) using the equilibrium \(\hat\nu(x)\).
+In particular, since \(\hat\nu(x)\) is Gaussian, we do not run into the
+same issue with higher moments as above. Assuming the equilibrium
+\(\hat\nu(x)\) does not change the ODE for \(N(t)\), but the ODE for
+\(\bar x(t)\) and \(\sigma^2(t)\) can now be expressed as
 \begin{equation}
 \frac{d}{dt}\bar x(t)=\sigma^2(t)\left(\frac{\partial\bar m(t)}{\partial\bar x(t)}-\overline{\frac{\partial m(t)}{\partial\bar x(t)}}\right)=a\sigma^2(t)(\theta-\bar x(t)),
-\end{equation}
-\begin{equation}
+\end{equation} \begin{equation}
 \frac{d}{dt}\sigma^2(t)=2\sigma^4(t)\left(\frac{\partial\bar m(t)}{\partial\sigma^2(t)}-\overline{\frac{\partial m(t)}{\partial\sigma^2(t)}}\right)+\mu=\mu-a\sigma^4(t),
-\end{equation}
-where $\hat p(x)=\hat\nu(x)/\hat N$. These expressions confirm our findings that $\hat{\bar x}=\theta$ and $\hat\sigma^2=\sqrt{\frac{\mu}{a}}$. Furthermore, calculating
+\end{equation} where \(\hat p(x)=\hat\nu(x)/\hat N\). These expressions
+confirm our findings that \(\hat{\bar x}=\theta\) and
+\(\hat\sigma^2=\sqrt{\frac{\mu}{a}}\). Furthermore, calculating
 \begin{equation}
 \frac{\partial}{\partial\sigma^2(t)}\frac{d}{dt}\sigma^2(t)=-2a\sigma^2(t)
-\end{equation}
-and evaluating at $\sigma^2(t)=\hat\sigma^2$ demonstrates the equilibrium phenotypic variance is stable when $a,\mu>0$. Hence, calculating
-\begin{equation}
+\end{equation} and evaluating at \(\sigma^2(t)=\hat\sigma^2\)
+demonstrates the equilibrium phenotypic variance is stable when
+\(a,\mu>0\). Hence, calculating \begin{equation}
 \frac{\partial}{\partial\bar x(t)}\frac{d}{dt}\bar x(t)=-a\sigma^2(t)
-\end{equation}
-and evaluating at $\sigma^2(t)=\hat\sigma^2$ and $\bar x(t)=\hat{\bar x}$ demonstrates the equilibrium phenotypic mean is stable when $a,\mu>0$. Finally, calculating
-\begin{equation}
+\end{equation} and evaluating at \(\sigma^2(t)=\hat\sigma^2\) and
+\(\bar x(t)=\hat{\bar x}\) demonstrates the equilibrium phenotypic mean
+is stable when \(a,\mu>0\). Finally, calculating \begin{equation}
 \frac{\partial}{\partial N(t)}\frac{d}{dt}N(t)=r-\frac{a}{2}((\theta-\bar x(t))^2+\sigma^2(t))-2cN(t)
-\end{equation}
-and evaluating at $\sigma^2(t)=\hat\sigma^2$, $\bar x(t)=\hat{\bar x}$ and $N(t)=\hat N$ demonstrates the equilibrium total abundance is stable when $a,c,\mu>0$, and $r>\frac{1}{2}\sqrt{a\mu}$.
-
-## The relation between diffusion and convolution with a Gaussian kernel \label{diffconvequiv}
-
-For continuous $g:\mathbb{R}^d\to\mathbb{R}$, consider the deterministic Cauchy problem
-\begin{equation}\label{heateqn}
+\end{equation} and evaluating at \(\sigma^2(t)=\hat\sigma^2\),
+\(\bar x(t)=\hat{\bar x}\) and \(N(t)=\hat N\) demonstrates the
+equilibrium total abundance is stable when \(a,c,\mu>0\), and
+\(r>\frac{1}{2}\sqrt{a\mu}\).
+
+\hypertarget{the-relation-between-diffusion-and-convolution-with-a-gaussian-kernel}{%
+\subsection{\texorpdfstring{The relation between diffusion and
+convolution with a Gaussian kernel
+\label{diffconvequiv}}{The relation between diffusion and convolution with a Gaussian kernel }}\label{the-relation-between-diffusion-and-convolution-with-a-gaussian-kernel}}
+
+For continuous \(g:\mathbb{R}^d\to\mathbb{R}\), consider the
+deterministic Cauchy problem \begin{equation}\label{heateqn}
 \left\{\begin{matrix}
@@
 \end{matrix}\right.
-\end{equation}
-According to Evans [-@lawrenceevans2010], the fundamental solution of (\ref{heateqn}) is
-\begin{equation}
+\end{equation} According to Evans (2010), the fundamental solution of
+(\ref{heateqn}) is \begin{equation}
 \Phi(x,t)=\frac{1}{(4\pi t)^{d/2}}\exp\left(-\frac{|x|^2}{4t}\right), \ (x,t)\in(0,\infty)\times\mathbb{R}^d,
-\end{equation}
-where $|x|=\sqrt{\sum_ix_i^2}$. The solution $f(x,t)$ of PDE (\ref{heateqn}) is then given by the convolution
-\begin{equation}
+\end{equation} where \(|x|=\sqrt{\sum_ix_i^2}\). The solution \(f(x,t)\)
+of PDE (\ref{heateqn}) is then given by the convolution \begin{equation}
 f(x,t)=\int_{\mathbb{R}^d}\Phi(x-y,t)g(y)dy, \ (x,t)\in(0,\infty)\times\mathbb{R}^d.
-\end{equation}
-Hence, by the fundamental theorem of calculus,
+\end{equation} Hence, by the fundamental theorem of calculus,
 
@@
 =\int_{\mathbb{R}^d}\Phi(x-y,1)f(t,y)dy.
-\end{multline}
-In particular,
-\begin{equation}
+\end{multline} In particular, \begin{equation}
 f(x,t)+\int_t^{t+1}\Delta f(x,s)ds=\int_{\mathbb{R}^d}\Phi(1,x-y)f(y,t)dy.
@@
 
-## Deterministic dynamics of $\sigma^2(t)$ \label{var_deriv}
+\hypertarget{deterministic-dynamics-of-sigma2t}{%
+\subsection{\texorpdfstring{Deterministic dynamics of \(\sigma^2(t)\)
+\label{var_deriv}}{Deterministic dynamics of \textbackslash{}sigma\^{}2(t) }}\label{deterministic-dynamics-of-sigma2t}}
 
@@
 \frac{d}{dt}\sigma^2(t)=\frac{d}{dt}\int_\mathbb{R}(x-\bar x(t))^2p(x,t)dx=\int_\mathbb{R}2(x-\bar x(t))\frac{d}{dt}{\bar x}(t)+(x-\bar x(t))^2\frac{\partial}{\partial t} p(x,t)dx\\=\int_\mathbb{R}(x-\bar x(t))^2\left((m(\nu,x)-\bar m(t))p(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t)\right)dx\\=\int_\mathbb{R}\left((x-\bar x(t))^2-\sigma^2(t)+\sigma^2(t)\right)(m(\nu,x)-\bar m(t))p(x,t)+(x-\bar x(t))^2\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t)dx\\=\mathrm{Cov}_t\Big((x-\bar x(t))^2,m(\nu,x)\Big)+\frac{\mu}{2}\int_\mathbb{R}(x-\bar x(t))^2\frac{\partial^2}{\partial x^2}p(x,t)dx.
-\end{multline}
-Applying integration by parts twice yields
+\end{multline} Applying integration by parts twice yields
 \begin{equation}
@@
 
-## Simplifying covariances with fitness under the assumption of a Gaussian density \label{cov2deriv}
+\hypertarget{simplifying-covariances-with-fitness-under-the-assumption-of-a-gaussian-density}{%
+\subsection{\texorpdfstring{Simplifying covariances with fitness under
+the assumption of a Gaussian density
+\label{cov2deriv}}{Simplifying covariances with fitness under the assumption of a Gaussian density }}\label{simplifying-covariances-with-fitness-under-the-assumption-of-a-gaussian-density}}
 
@@
 
-## Comparing our treatment of white noise to Da Prato and Zabczyk (2014) \label{DPZ14}
-
-Our approach in the main text is inspired by the treatment provided in \S4.2 of Da Prato and Zabczyk [-@DaPrato2014].
-Here the authors develop a stochastic integral of operator-valued processes.
-In particular, they consider processes indexed by time $t\geq0$ valued as Hilbert-Schmidt operators $\Phi(t)$ and define the norm
-\begin{equation}
+\hypertarget{comparing-our-treatment-of-white-noise-to-da-prato-and-zabczyk-2014}{%
+\subsection{\texorpdfstring{Comparing our treatment of white noise to Da
+Prato and Zabczyk (2014)
+\label{DPZ14}}{Comparing our treatment of white noise to Da Prato and Zabczyk (2014) }}\label{comparing-our-treatment-of-white-noise-to-da-prato-and-zabczyk-2014}}
+
+Our approach in the main text is inspired by the treatment provided in
+\S4.2 of Da Prato and Zabczyk (2014). Here the authors develop a
+stochastic integral of operator-valued processes. In particular, they
+consider processes indexed by time \(t\geq0\) valued as Hilbert-Schmidt
+operators \(\Phi(t)\) and define the norm \begin{equation}
 \|\Phi\|_t=\sqrt{\mathbb{E}\left(\int_0^t\mathrm{Tr}[\Phi(s)\Phi^*(s)]ds\right)}, \ t\geq0.
-\end{equation}
-In our case we only consider the so-called multiplication operators.
-That is, processes that consist of operators $\Phi(t)$ having the form $\Phi(t)g(x)=\varphi(x,t)g(x)$ such that $\varphi(\cdot,t)\in L^2(\mathbb{R})$ a.s.
-for each $t\geq0$.
-In this case $\Phi(t)=\Phi^*(t)$ and
-\begin{equation}
+\end{equation} In our case we only consider the so-called multiplication
+operators. That is, processes that consist of operators \(\Phi(t)\)
+having the form \(\Phi(t)g(x)=\varphi(x,t)g(x)\) such that
+\(\varphi(\cdot,t)\in L^2(\mathbb{R})\) a.s. for each \(t\geq0\). In
+this case \(\Phi(t)=\Phi^*(t)\) and \begin{equation}
 \|\Phi\|_t=\|\varphi\|_t=\sqrt{\mathbb{E}\left(\int_0^t\int_\mathbb{R}\varphi^2(x,s)dxds\right)}, \ t\geq0.
-\end{equation}
-Da Prato and Zabczyk [-@DaPrato2014] form the space $\mathscr{N}^2_W(0,T)$ of Hilbert-Schmidt operator-valued predictable processes $\Phi(t)$ that satisfy $\|\Phi\|_T<+\infty$ for some $T>0$.
-This corresponds to our more specialized space $\mathscr{N}_2$ that consists of $L^2(\mathbb{R})$ valued processes $\varphi(x,t)$ such that $\|\varphi\|_t<+\infty$ for all $t\geq0$.
-In their treatment, $W(t)$ plays a similar role to our generalized process $\mathbf{W}_t$.
-For $\Phi\in\mathscr{N}^2_W(0,T)$, they denote the stochastic integral for $t\in[0,T]$ by $\Phi\cdot W(t)$.
-Hence, for $\Phi(t) g(x)=\varphi(x,t)g(x)$ as above, $\mathbf{W}_t(\varphi)=\Phi\cdot W(t)$.
-The authors then prove the following:
-
-  > __Proposition 4.28__ _Assume that $\Phi_1,\Phi_2\in\mathscr{N}^2_W(0,T)$. Then_
-  >
-  > $$\mathbb{E}(\Phi_i\cdot W(t))=0, \ \ \mathbb{E}(\|\Phi_i\cdot W(t)\|^2)<+\infty, \ \ \forall \ t\in[0,T].$$
-  >
-  > __Corollory 4.29__ _Under the same assumptions as Proposition 4.28,_
-  >
-  > $$\mathbb{C}(\Phi_1\cdot W(t),\Phi_2\cdot W(s))=\mathbb{E}\left(\int_0^{t\wedge s}\mathrm{Tr}[\Phi_2(r)\Phi_1^*(r)]dr\right), \ \ \forall \ t,s\in[0,T].$$
-
-Simplifying these expressions for the multiplication operators described above returns equations (\ref{exp_WN_xi}) and (\ref{cov_WN_xi}) of the main text.
-
-## Simulating the rescaled process and numerical evidence of approximate normality \label{numerical}
+\end{equation} Da Prato and Zabczyk (2014) form the space
+\(\mathscr{N}^2_W(0,T)\) of Hilbert-Schmidt operator-valued predictable
+processes \(\Phi(t)\) that satisfy \(\|\Phi\|_T<+\infty\) for some
+\(T>0\). This corresponds to our more specialized space
+\(\mathscr{N}_2\) that consists of \(L^2(\mathbb{R})\) valued processes
+\(\varphi(x,t)\) such that \(\|\varphi\|_t<+\infty\) for all \(t\geq0\).
+In their treatment, \(W(t)\) plays a similar role to our generalized
+process \(\mathbf{W}_t\). For \(\Phi\in\mathscr{N}^2_W(0,T)\), they
+denote the stochastic integral for \(t\in[0,T]\) by \(\Phi\cdot W(t)\).
+Hence, for \(\Phi(t) g(x)=\varphi(x,t)g(x)\) as above,
+\(\mathbf{W}_t(\varphi)=\Phi\cdot W(t)\). The authors then prove the
+following:
+
+\begin{quote}
+\textbf{Proposition 4.28} \emph{Assume that
+\(\Phi_1,\Phi_2\in\mathscr{N}^2_W(0,T)\). Then}
+
+\[\mathbb{E}(\Phi_i\cdot W(t))=0, \ \ \mathbb{E}(\|\Phi_i\cdot W(t)\|^2)<+\infty, \ \ \forall \ t\in[0,T].\]
+
+\textbf{Corollory 4.29} \emph{Under the same assumptions as Proposition
+4.28,}
+
+\[\mathbb{C}(\Phi_1\cdot W(t),\Phi_2\cdot W(s))=\mathbb{E}\left(\int_0^{t\wedge s}\mathrm{Tr}[\Phi_2(r)\Phi_1^*(r)]dr\right), \ \ \forall \ t,s\in[0,T].\]
+\end{quote}
+
+Simplifying these expressions for the multiplication operators described
+above returns equations (\ref{exp_WN_xi}) and (\ref{cov_WN_xi}) of the
+main text.
+
+\hypertarget{simulating-the-rescaled-process-and-numerical-evidence-of-approximate-normality}{%
+\subsection{\texorpdfstring{Simulating the rescaled process and
+numerical evidence of approximate normality
+\label{numerical}}{Simulating the rescaled process and numerical evidence of approximate normality }}\label{simulating-the-rescaled-process-and-numerical-evidence-of-approximate-normality}}
 
@@
 
-the density process $\nu(x,t)$ defined by SPDE (\ref{SPDE}) of the main text satisfies $\int_\mathbb{R}(|x|+x^2)\nu(x,t)dx<\infty$.
-
-and setting $c<r-(\eta a+\sqrt{\mu a})/2$ ensures $\hat N>1$, which is important for numerical simulations when $N$ is an integer.
-We use these results to help guide our choice of parameter values for simulations of the branching random walk.
-In the following section we provide a detailed description of the branching random walk and how we have chosen to rescale it.
-We then use the rescaled branching random walk to investigate finiteness of moments and normality.
-
-### Description of simulation \label{description}
-
-We begin by describing the branching random walk before introducing our scheme to rescale it.
-Our branching random walk follows closely the description of branching Brownian motion in the main text.
-However, we replace exponentially distributed lifetimes with deterministic unit time steps for easier implementation.
-Hence, we restrict time to $t=0,1,2,\dots$.
-Furthermore, we allow individual fitness to depend on both trait value and the state of the entire population.
-For time $t$ we write $\{\xi_1(t),\dots,\xi_{N(t)}(t)\}$ as the set of trait values across all $N(t)$ individuals alive in the population.
-Since our simulation treats individuals instead of continuous distributions of trait values, we can write
+the density process \(\nu(x,t)\) defined by SPDE (\ref{SPDE}) of the
+main text satisfies \(\int_\mathbb{R}(|x|+x^2)\nu(x,t)dx<\infty\).
+
+and setting \(c<r-(\eta a+\sqrt{\mu a})/2\) ensures \(\hat N>1\), which
+is important for numerical simulations when \(N\) is an integer. We use
+these results to help guide our choice of parameter values for
+simulations of the branching random walk. In the following section we
+provide a detailed description of the branching random walk and how we
+have chosen to rescale it. We then use the rescaled branching random
+walk to investigate finiteness of moments and normality.
+
+\hypertarget{description-of-simulation}{%
+\subsubsection{\texorpdfstring{Description of simulation
+\label{description}}{Description of simulation }}\label{description-of-simulation}}
+
+We begin by describing the branching random walk before introducing our
+scheme to rescale it. Our branching random walk follows closely the
+description of branching Brownian motion in the main text. However, we
+replace exponentially distributed lifetimes with deterministic unit time
+steps for easier implementation. Hence, we restrict time to
+\(t=0,1,2,\dots\). Furthermore, we allow individual fitness to depend on
+both trait value and the state of the entire population. For time \(t\)
+we write \(\{\xi_1(t),\dots,\xi_{N(t)}(t)\}\) as the set of trait values
+across all \(N(t)\) individuals alive in the population. Since our
+simulation treats individuals instead of continuous distributions of
+trait values, we can write
 
@@
 
-where $\delta(x-\xi_i(t))$ denotes the point mass located at $\xi_i(t)$.
-To allow for imperfect heritability, we also track the set of breeding values which, at time $t$, is denoted by $\{\gamma_1(t),\dots,\gamma_{N(t)}(t)\}$ and should not be confused with the quadratic selection gradients discussed in \S\ref{diffuse} of the main text.
-Then the distribution of breeding values can be written as
+where \(\delta(x-\xi_i(t))\) denotes the point mass located at
+\(\xi_i(t)\). To allow for imperfect heritability, we also track the set
+of breeding values which, at time \(t\), is denoted by
+\(\{\gamma_1(t),\dots,\gamma_{N(t)}(t)\}\) and should not be confused
+with the quadratic selection gradients discussed in \S\ref{diffuse} of
+the main text. Then the distribution of breeding values can be written
+as
 
@@
 
-Following our model of heritiability, the trait value $\xi_i(t)$ is drawn from a normal distribution centered on $\gamma_i(t)$ with variance $\eta$.
-At each iteration we draw, for each individual, a random number of offspring from a Negative-Binomial distribution.
-Recall the Negative-Binomial distribution models the number of failed Bernoulli trials that occur before a given number of successful trials.
-Denoting $q$ the probability of success for each trial and $s$ the number of successes, the mean and variance is given respectively by
+Following our model of heritiability, the trait value \(\xi_i(t)\) is
+drawn from a normal distribution centered on \(\gamma_i(t)\) with
+variance \(\eta\). At each iteration we draw, for each individual, a
+random number of offspring from a Negative-Binomial distribution. Recall
+the Negative-Binomial distribution models the number of failed Bernoulli
+trials that occur before a given number of successful trials. Denoting
+\(q\) the probability of success for each trial and \(s\) the number of
+successes, the mean and variance is given respectively by
 
@@
 
-Then if we require the $i$th individual to have mean number offspring $\mathscr W(\nu,\xi_i)$ and variance equal to $V$, the parameters of the associated Negative-Binomial distribution become
+Then if we require the \(i\)th individual to have mean number offspring
+\(\mathscr W(\nu,\xi_i)\) and variance equal to \(V\), the parameters of
+the associated Negative-Binomial distribution become
 
@@
 
-The imposes the restriction $V>\mathscr{W}(\nu,\xi_i)$.
-For each offpsring produced by the individual with breeding value $\gamma_i(t)$, we assign indepently drawn breeding values normally distributed around $\gamma_i(t)$ with variance $\mu$.
-After breeding values have been assigned, we randomly draw trait values for each offpsring as described above.
-For an overview of our model of inheritance, see \S\ref{inheritance} of the main text.
-This summarizes the basic structure of our simulation.
-To impose selection and density dependent growth rates, we set
+The imposes the restriction \(V>\mathscr{W}(\nu,\xi_i)\). For each
+offpsring produced by the individual with breeding value
+\(\gamma_i(t)\), we assign indepently drawn breeding values normally
+distributed around \(\gamma_i(t)\) with variance \(\mu\). After breeding
+values have been assigned, we randomly draw trait values for each
+offpsring as described above. For an overview of our model of
+inheritance, see \S\ref{inheritance} of the main text. This summarizes
+the basic structure of our simulation. To impose selection and density
+dependent growth rates, we set
 
@@
 
-where the above integral becomes $\int_\mathbb{R}\nu(x,t)dx=\sum_{i=1}^{N(t)}1=N(t)$.
-
-__Rescaling__
-
-To rescale the branching random walk by a positive integer $n$, we rescale segregation and mutational variance according to $\eta\to \eta$ and $\mu\to \mu/n$, time by $t\to t/n$ and the reproductive law by $V\to V$ and
+where the above integral becomes
+\(\int_\mathbb{R}\nu(x,t)dx=\sum_{i=1}^{N(t)}1=N(t)\).
+
+\textbf{Rescaling}
+
+To rescale the branching random walk by a positive integer \(n\), we
+rescale segregation and mutational variance according to
+\(\eta\to \eta\) and \(\mu\to \mu/n\), time by \(t\to t/n\) and the
+reproductive law by \(V\to V\) and
 
@@
 
-We also replace individual mass with $\frac{1}{n}$ and write rescaled abundance as $N^{(n)}(t)=\frac{1}{n}N(nt)$.
-Under this rescaling the deterministic equilibrium of the raw numerical abundance becomes
+We also replace individual mass with \(\frac{1}{n}\) and write rescaled
+abundance as \(N^{(n)}(t)=\frac{1}{n}N(nt)\). Under this rescaling the
+deterministic equilibrium of the raw numerical abundance becomes
 
@@
 
-When it exists, we denote by $N^{(\infty)}(t)$ the limiting process of $N^{(n)}(t)$. Then
+When it exists, we denote by \(N^{(\infty)}(t)\) the limiting process of
+\(N^{(n)}(t)\). Then
 
@@
 
-Note that, following the notation of Theorem 1 in Méléard and Roelly [-@meleard1992interacting], setting $\lambda_n=n$, $m_n(\nu)=\mathscr W^{(n)}(\nu,\cdot)$ and $\varepsilon_n=1/n$ satisfies their hypotheses $(\mathscr H_0)$-$(\mathscr H_3)$ when $c=0$.
-We have implemented this simulation in the programming language Julia.
-A copy can be found at the url:
+Note that, following the notation of Theorem 1 in Méléard and Roelly
+(1992), setting \(\lambda_n=n\),
+\(m_n(\nu)=\mathscr W^{(n)}(\nu,\cdot)\) and \(\varepsilon_n=1/n\)
+satisfies their hypotheses \((\mathscr H_0)\)-\((\mathscr H_3)\) when
+\(c=0\). We have implemented this simulation in the programming language
+Julia. A copy can be found at the url:
 
@@
 
-For the sake of illustration, we simulated the unscaled process ($n=1$) and the rescaled process with $n=5$ and $n=20$ for $50$ units of time.
-Results are shown in Figure \ref{rescaled}.
-In the following section we use a statistical test to show, for the lower bound on $m(\nu,x)$, the rescaled process converges to a Gaussian density as $n\to\infty$ and $V/N\to0$.
-
-### Evidence of normality \label{ev_normal}
-
-To demonstrate approximate normality of the phenotypic distribution when $V/N$ is small we utilized the one-sided Kolmogorov-Smirnov test.
-This test compares an empirical cumulative distribution function (i.e., a cumulative distribution function generated from simulated data) to a hypothetical cumulative distribution function by providing a distribution for the maximum distance between these curves.
-More precisely if $F_n(x)$ is the empricial distribution function for a sample of size $n$ and $F(x)$ is the hypothetical distribution function, Kolmogorov-Smirnov statistic is $D_n=\sup_x|F_n(x)-F(x)|$.
-
-## Derivation of SDE for $\bar x$ and $\sigma^2$ \label{SDE_DERIV}
+For the sake of illustration, we simulated the unscaled process
+(\(n=1\)) and the rescaled process with \(n=5\) and \(n=20\) for \(50\)
+units of time. Results are shown in Figure \ref{rescaled}. In the
+following section we use a statistical test to show, for the lower bound
+on \(m(\nu,x)\), the rescaled process converges to a Gaussian density as
+\(n\to\infty\) and \(V/N\to0\).
+
+\hypertarget{evidence-of-normality}{%
+\subsubsection{\texorpdfstring{Evidence of normality
+\label{ev_normal}}{Evidence of normality }}\label{evidence-of-normality}}
+
+To demonstrate approximate normality of the phenotypic distribution when
+\(V/N\) is small we utilized the one-sided Kolmogorov-Smirnov test. This
+test compares an empirical cumulative distribution function (i.e., a
+cumulative distribution function generated from simulated data) to a
+hypothetical cumulative distribution function by providing a
+distribution for the maximum distance between these curves. More
+precisely if \(F_n(x)\) is the empricial distribution function for a
+sample of size \(n\) and \(F(x)\) is the hypothetical distribution
+function, Kolmogorov-Smirnov statistic is \(D_n=\sup_x|F_n(x)-F(x)|\).
+
+\hypertarget{derivation-of-sde-for-bar-x-and-sigma2}{%
+\subsection{\texorpdfstring{Derivation of SDE for \(\bar x\) and
+\(\sigma^2\)
+\label{SDE_DERIV}}{Derivation of SDE for \textbackslash{}bar x and \textbackslash{}sigma\^{}2 }}\label{derivation-of-sde-for-bar-x-and-sigma2}}
 
@@
 \tilde x(t)=\int_\mathbb{R}x\nu(x,t)dx, \ \ \tilde{\tilde\sigma}^2(t)=\int_\mathbb{R}x^2\nu(x,t)dx
-\end{equation}
-and
-\begin{equation}\label{xtilde}
+\end{equation} and \begin{equation}\label{xtilde}
 \tilde x(t)=\tilde x(0)+\int_0^t\int_\mathbb{R}\nu(x,s)m(\nu,x)x+x\sqrt{V\nu(x,s)}\dot W(x,s)dxds,
-\end{equation}
-\begin{equation}\label{stilde}
+\end{equation} \begin{equation}\label{stilde}
 \tilde{\tilde\sigma}^2(t)=\tilde{\tilde\sigma}^2(0)+\int_0^t\int_\mathbb{R}\nu(x,s)\left(m(\nu,x)x^2+\mu\right)+x^2\sqrt{V\nu(x,s)}\dot W(x,s)dxds.
@@
 
-<!--
-
-For $\nu(x,t)$ defined in the main text, $h\in C(\mathbb{R})$ and $t\geq0$ we make the following assumptions:
-
-\begin{equation}\label{L1_nu}
-\mathbb{E}\left(\int_\mathbb{R}|h(x)|\nu(x,t)dx\right)<\infty,
-\end{equation}
-\begin{equation}\label{L2_nu}
-\mathbb{E}\left(\int_\mathbb{R}h^2(x)\nu(x,t)dx\right)<\infty,
-\end{equation}
-\begin{equation}\label{L1_deriv}
-\mathbb{E}\left(\int_0^t\int_\mathbb{R}\nu(x,s)\Big|h(x)\Big|\left|m(\nu,x)f(x)+\frac{\mu}{2}\Delta f(x)\right|dxds\right)<\infty, \ \forall f\in C^2_b(\mathbb{R}).
-\end{equation}
-
-Put $H(t)=\int_\mathbb{R}h(x)\nu(x,t)dx$. Then, for non-random and non-negative $\nu_1(x,t)$ that is continuous in both arguments and integrable in $x$,
-\begin{equation}
-\star:=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{E}[H(t+\varepsilon)-H(t)|\nu(x,t)=\nu_1(x,t)]
-\end{equation}
-\begin{equation}
-=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{E}\left[\int_\mathbb{R}h(x)\big(\nu(x,t+\varepsilon)-\nu(x,t)\big)dx\Big|\nu(x,t)=\nu_1(x,t)\right]
-\end{equation}
-\begin{equation}
-=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{E}\left[\int_\mathbb{R}\int_t^{t+\varepsilon}f(x)\left(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)
-+\sqrt{V\nu(x,s)}\dot W(x,s)\right)dsdx\Big|\nu(x,t)=\nu_1(x,t)\right].
-\end{equation}
-
-By assumption (\ref{L1_deriv}) we can use Fubini's theorem to write, with probability one,
-
-\begin{equation}
-\int_\mathbb{R}\int_t^{t+\varepsilon}h(x)\big(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)dsdx=\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\big(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)dxds.
-\end{equation}
-
-By assumption (\ref{L2_nu}) we have $\varphi(x,t)=h(x)\sqrt{V\nu(x,t)}$ implies $\varphi\in L^2_c(\mathbb{R}\times[0,\infty))$. Hence, the following is true by definition;
-
-\begin{equation}
-\int_\mathbb{R}\int_t^{t+\varepsilon}h(x)\sqrt{V\nu(x,s)}\dot W(x,s)dsdx = 
-\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\sqrt{V\nu(x,s)}\dot W(x,s)dxds.
-\end{equation}
-
-Hence,
-
-\begin{multline}
-\mathbb{E}\left[\int_\mathbb{R}\int_t^{t+\varepsilon}h(x)\sqrt{V\nu(x,s)}\dot W(x,s)dsdx\Big|\nu(x,t)=\nu_1(x,t)\right] \\ =\mathbb{E}\left[\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\sqrt{V\nu(x,s)}\dot W(x,s)dxds\Big|\nu(x,t)=\nu_1(x,t)\right]=0
-\end{multline}
-
-and
-
-\begin{equation}
-\star=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{E}\left[\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\left(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)\right)dxds\Big|\nu(x,t)=\nu_1(x,t)\right].
-\end{equation}
-
-By assumption (\ref{L1_deriv}) we know that there exists a $\delta>0$ such that for each positive $\varepsilon<\delta$ the following holds almost surely:
-
-\begin{multline}
-\left|\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\left(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)\right)dxds\right| \\
-\leq\int_t^{t+\delta}\int_\mathbb{R}\Big|h(x)\Big|\left|m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)\right|dxds<\infty.
-\end{multline}
-
-Thus, by Lebesgue's dominated convergence theorem, the drift component of the process $H(t)$ can be computed as
-
-\begin{multline}
-\star=\mathbb{E}\left[\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\left(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)\right)dxds\Big|\nu(x,t)=\nu_1(x,t)\right]\\
-=\int_\mathbb{R}h(x)\big(m(\nu_1,x)\nu_1(x,t)+\frac{\mu}{2}\Delta\nu_1(x,t)\big)dx.
-\end{multline}
-
-To find an expression for the diffusion component of $H(t)$ set
-
-\begin{multline}\label{star2}
-\star\star:=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{V}[H(t+\varepsilon)-H(t)|\nu(x,t)=\nu_1(x,t)] \\
-=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{V}\left[\int_\mathbb{R}h(x)\big(\nu(x,t+h)-\nu(x,t)\big)dx\Big|\nu(x,t)=\nu_1(x,t)\right].
-\end{multline}
-
-We can rewrite the integral inside expression (\ref{star2}) as
-
-\begin{equation}
-\int_\mathbb{R}\int_t^{t+\varepsilon}h(x)\big(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)+\sqrt{V\nu(x,s)}\dot W(x,s)\big)dsdx.
-\end{equation}
-
-We have already found
-
-\begin{multline}
-\mathbb{E}\left[\int_\mathbb{R}\int_t^{t+\varepsilon}h(x)\big(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)+\sqrt{V\nu(x,s)}\dot W(x,s)\big)dsdx\Big|\nu(x,t)=\nu_1(x,t)\right] \\
-=\mathbb{E}\left[\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\big(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)\big)dxds\Big|\nu(x,t)=\nu_1(x,t)\right].
-\end{multline}
-
-Then, since $h^2(x)\nu(x,t)$ is integrable by assumption (\ref{L2_nu}), we have
-
-\begin{multline}
-\star\star=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{E}\left[\left(\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\sqrt{V\nu(x,s)}\dot W(x,s)dsdx\right)^2\Big| \ \nu(x,t)=\nu_1(x,t)\right] \\
-=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{E}\left[\int_t^{t+\varepsilon}\int_\mathbb{R}V h^2(x)\nu(x,s)dxds\Big|\nu(x,t)=\nu_1(x,t)\right].
-\end{multline}
-
-Thus, for any $\delta\geq\varepsilon\geq0$, we have, with probability one,
-
-\begin{equation}
-\int_t^{t+\varepsilon}\int_\mathbb{R}V h^2(x)\nu(x,s)dxds\leq\int_t^{t+\delta}\int_\mathbb{R}V h^2(x)\nu(x,s)dxds.
-\end{equation}
-
-We can therefore use Lebesgue's dominated convergence theorem to justify
-
-\begin{equation}
-\star\star=\mathbb{E}\left[\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\int_t^{t+\varepsilon}\int_\mathbb{R}V h^2(x)\nu(x,s)dxds\Big|\nu(x,t)=\nu_1(x,t)\right]=\int_\mathbb{R}V h^2(x)\nu_1(x,t)dx.
-\end{equation}
-
-Then, using the notation of stochastic differentials, we have
-
-\begin{equation}\label{SDE_deriv}
-dH(t)=\left(\int_\mathbb{R}h(x)\left(m(\nu,x)\nu(x,t)+\frac{\mu}{2}\Delta\nu(x,t)\right)dx\right)dt+\sqrt{V\int_\mathbb{R}h^2(x)\nu(x,t)dx} \ dW(t)
-\end{equation}
-
-where $W$ is a standard Brownian motion. In the following subsections we employ this formula under the cases $h(x)=x,x^2$ to obtain SDE for the phenotypic mean and variance.
-
--->
-
-### Derivation for trait mean
+\hypertarget{derivation-for-trait-mean}{%
+\subsubsection{Derivation for trait
+mean}\label{derivation-for-trait-mean}}
 
@@
 
-Using It&ocirc;'s quotient rule on $\bar x=\tilde x/N$, we obtain
+Using Itô's quotient rule on \(\bar x=\tilde x/N\), we obtain
 
@@
 
-From Table \ref{mult} of the main text $d\tilde xdN=\langle\tilde x,N\rangle$ and $dN^2=\|N\|_2^2$. Hence,
+From Table \ref{mult} of the main text
+\(d\tilde xdN=\langle\tilde x,N\rangle\) and \(dN^2=\|N\|_2^2\). Hence,
 
@@
 
-### Derivation for trait variance
+\hypertarget{derivation-for-trait-variance}{%
+\subsubsection{Derivation for trait
+variance}\label{derivation-for-trait-variance}}
 
@@
 
-Using It&ocirc;'s quotient rule on $\overline{x^2}=\tilde{\tilde\sigma}^2/N$, we obtain
+Using Itô's quotient rule on
+\(\overline{x^2}=\tilde{\tilde\sigma}^2/N\), we obtain
 
@@
 
-Table \ref{mult} of the main text implies $d\tilde W_3d W_1=\langle\tilde{\tilde\sigma}^2,N\rangle$ and hence
+Table \ref{mult} of the main text implies
+\(d\tilde W_3d W_1=\langle\tilde{\tilde\sigma}^2,N\rangle\) and hence
 
@@
 
-Setting $F(y,z)=y-z^2$, use It&ocirc;'s formula on $\sigma^2=F(\overline{x^2},\bar x)=\overline{x^2}-\bar x^2$ to obtain:
+Setting \(F(y,z)=y-z^2\), use Itô's formula on
+\(\sigma^2=F(\overline{x^2},\bar x)=\overline{x^2}-\bar x^2\) to obtain:
 
@@
 
-In particular, when $p$ is a Gaussian curve $d W_2d W_3=0$.
-
-## Relating fitness of expressed traits to fitness of breeding values \label{fit2fit}
-
-Following \S\ref{inheritance} of the main text, we have $\sigma^2=G+\eta$ and
+In particular, when \(p\) is a Gaussian curve \(d W_2d W_3=0\).
+
+\hypertarget{relating-fitness-of-expressed-traits-to-fitness-of-breeding-values}{%
+\subsection{\texorpdfstring{Relating fitness of expressed traits to
+fitness of breeding values
+\label{fit2fit}}{Relating fitness of expressed traits to fitness of breeding values }}\label{relating-fitness-of-expressed-traits-to-fitness-of-breeding-values}}
+
+Following \S\ref{inheritance} of the main text, we have
+\(\sigma^2=G+\eta\) and
 
@@
 \int_{\mathbb{R}}\int_\mathbb{R}\frac{\rho(g,t)}{N(t)}\psi(x,g)dg\frac{\partial}{\partial\bar x}m(\nu,x)dx=\int_\mathbb{R}p(x,t)\frac{\partial}{\partial\bar x}m(\nu,x)dx=\overline{\frac{\partial m}{\partial\bar x}}
-\end{multline}
-and
+\end{multline} and
 
@@
 
-## Derivation of diffuse coevolution model \label{diffuse}
-
-In this section we provide a derivation of our model of diffuse coevolution driven by competition.
-Since most of the work in this derivation has already been completed in Supplementary Material \S\ref{SDE_DERIV}, we focus here on deriving the Malthusian fitness $m$ as a function of trait value $x$.
-We begin with discrete populations of individuals.
-In particular, we begin by assuming population size $n_i$ is an integer for each species $i=1,\dots,S$ before passing to the large population size limit.
-
-The reduction in fitness for an individual of species $i$ caused by competition is captured multiplicatively by $0<C_i\leq1$.
-Biologically this assumes all competitors affect individuals of a given species equally by consuming the same amount of resources.
-This is a mean-field interaction since any individual that consumes resources has an effect on the fitness of all other individuals competing for the same resources.
-Denote by $x_{ij}$ the trait value of the $j$-th individual belonging to species $i$.
-The set of trait values across all individuals in the community at time $t\geq0$ is written $X=\{x_{ij}\}$.
-We denote by $\mathscr{B}_{ij}$ a function that maps $X$ to the cumulative effect of all competitive interactions on the fitness of the $j$-th individual in species $i$.
-Since individuals do not compete with themselves the net multiplicative effects on fitness of both interspecific and intraspecific competition on the $j$-th individual in species $i$ can be summarized by
+\hypertarget{derivation-of-diffuse-coevolution-model}{%
+\subsection{\texorpdfstring{Derivation of diffuse coevolution model
+\label{diffuse}}{Derivation of diffuse coevolution model }}\label{derivation-of-diffuse-coevolution-model}}
+
+In this section we provide a derivation of our model of diffuse
+coevolution driven by competition. Since most of the work in this
+derivation has already been completed in Supplementary Material
+\S\ref{SDE_DERIV}, we focus here on deriving the Malthusian fitness
+\(m\) as a function of trait value \(x\). We begin with discrete
+populations of individuals. In particular, we begin by assuming
+population size \(n_i\) is an integer for each species \(i=1,\dots,S\)
+before passing to the large population size limit.
+
+The reduction in fitness for an individual of species \(i\) caused by
+competition is captured multiplicatively by \(0<C_i\leq1\). Biologically
+this assumes all competitors affect individuals of a given species
+equally by consuming the same amount of resources. This is a mean-field
+interaction since any individual that consumes resources has an effect
+on the fitness of all other individuals competing for the same
+resources. Denote by \(x_{ij}\) the trait value of the \(j\)-th
+individual belonging to species \(i\). The set of trait values across
+all individuals in the community at time \(t\geq0\) is written
+\(X=\{x_{ij}\}\). We denote by \(\mathscr{B}_{ij}\) a function that maps
+\(X\) to the cumulative effect of all competitive interactions on the
+fitness of the \(j\)-th individual in species \(i\). Since individuals
+do not compete with themselves the net multiplicative effects on fitness
+of both interspecific and intraspecific competition on the \(j\)-th
+individual in species \(i\) can be summarized by
 
@@
 
-where $\mathcal{O}_{ij}$, defined in the main text, measures the overlap in resource use between individuals of species $i$ and $j$ as a function of their niche-centers.
-Writing $\mathscr W_{ij}(X)$ as the average number of offspring left by the $j$-th individual of species $i$, we have
+where \(\mathcal{O}_{ij}\), defined in the main text, measures the
+overlap in resource use between individuals of species \(i\) and \(j\)
+as a function of their niche-centers. Writing \(\mathscr W_{ij}(X)\) as
+the average number of offspring left by the \(j\)-th individual of
+species \(i\), we have
 
@@
 
-where $\mathscr{A}_i(x)=\int_\mathbb{R}e_i(\zeta)u_i(\zeta,x)d\zeta$ accounts for abiotic selection and $e_i$ has been defined in the main text.
-
-We now turn to a diffusion limit.
-Since we have more than one population, we take the diffusion limit for each population one at a time starting with population $1$.
-We write $\pmb n=(n_1,\dots,n_S)$.
-Following Méléard and Roelly [-@Mlard1993; -@meleard1992interacting] we rescale generation time and individual mass to $\frac{1}{n_1}$ and mean of the reproductive law to
+where \(\mathscr{A}_i(x)=\int_\mathbb{R}e_i(\zeta)u_i(\zeta,x)d\zeta\)
+accounts for abiotic selection and \(e_i\) has been defined in the main
+text.
+
+We now turn to a diffusion limit. Since we have more than one
+population, we take the diffusion limit for each population one at a
+time starting with population \(1\). We write
+\(\pmb n=(n_1,\dots,n_S)\). Following Méléard and Roelly (1993, 1992) we
+rescale generation time and individual mass to \(\frac{1}{n_1}\) and
+mean of the reproductive law to
 
@@
 
-For large $n_1$, we have the approximation
+For large \(n_1\), we have the approximation
 
@@
 \mathscr W_{1j}^{(\pmb n)}(X)\approx\mathscr{A}_1(x_{1j})^{1/n_1}\left(1+\frac{\ln C_1}{n_1^2}\sum_{l\neq j}\mathcal{O}_{11}(x_{1j}-x_{1l})+\frac{\ln C_1}{n_1}\sum_{k\neq 1}\frac{1}{n_k}\sum_{l=1}^{n_k}\mathcal{O}_{1k}(x_{1j}-x_{kl})\right).
-\end{equation}
-Hence
-\begin{equation}
+\end{equation} Hence \begin{equation}
 \lim_{n_1\to\infty}n_1\left(\mathscr W_{1j}^{(\pmb n)}(X)-1\right)=\ln\mathscr{A}_1(x_{1j})+\left(\int_\mathbb{R}\mathcal{O}_{11}(x_{1j}-y)\nu_1(y,t)dy+\sum_{k\neq 1}\frac{1}{n_k}\sum_{l=1}^{n_k}\mathcal{O}_{1k}(x_{1j}-x_{kl})\right)\ln C_1.
@@
 
-We write $\lim_{\pmb n\to\pmb\infty}$ for the iterated limit $\lim_{n_S\to\infty}\cdots\lim_{n_1\to\infty}$ and, assuming  $\nu_i(\cdot,t)\in C_1^+(\mathbb{R})$ for $i=1,\dots,S$ and $t\in[0,\infty)$, we set $\pmb\nu=(\nu_1,\dots,\nu_S)$.
-Then, for any $\pmb\nu$, the the diffusion limits for the remaining populations provides the Malthusian parameter for individuals in species $i$ with trait value $x_{1j}$ as
+We write \(\lim_{\pmb n\to\pmb\infty}\) for the iterated limit
+\(\lim_{n_S\to\infty}\cdots\lim_{n_1\to\infty}\) and, assuming
+\(\nu_i(\cdot,t)\in C_1^+(\mathbb{R})\) for \(i=1,\dots,S\) and
+\(t\in[0,\infty)\), we set \(\pmb\nu=(\nu_1,\dots,\nu_S)\). Then, for
+any \(\pmb\nu\), the the diffusion limits for the remaining populations
+provides the Malthusian parameter for individuals in species \(i\) with
+trait value \(x_{1j}\) as
 
@@
 
-We compute the average niche overlap of an individual in species $i$ with nich location $x$ across all individuals in species $j$ as
+We compute the average niche overlap of an individual in species \(i\)
+with nich location \(x\) across all individuals in species \(j\) as
 
@@
 
-We now assume the resource utilization curves $u_i(\zeta)$ and phenotypic densities $\nu_i(x,t)$ are Gaussian curves for $i=1,\dots,S$. In this case $\bar{\mathcal{O}}_{ij}(x,t)$ simplifies to
+We now assume the resource utilization curves \(u_i(\zeta)\) and
+phenotypic densities \(\nu_i(x,t)\) are Gaussian curves for
+\(i=1,\dots,S\). In this case \(\bar{\mathcal{O}}_{ij}(x,t)\) simplifies
+to
 
@@
 
-Hence, our fitness function is bounded above, as required in the main text.
-
-For the remainder of the derivation we suppress notation indicating dependency on $\pmb\nu$, $x$ and $t$. From (\ref{malth_fit}) we calculate
+Hence, our fitness function is bounded above, as required in the main
+text.
+
+For the remainder of the derivation we suppress notation indicating
+dependency on \(\pmb\nu\), \(x\) and \(t\). From (\ref{malth_fit}) we
+calculate
 
@@
 
-The average fitness for species $i$ is
+The average fitness for species \(i\) is
 
@@
 
-Applying equations (\ref{N}), (\ref{xbarfinal}) and (\ref{Gfinal}) of the main text recovers system (\ref{comm_dynamics}) of the main text.
-
-## The relation between competition coefficients and selection \label{coef_grad_moms}
-
-### Derivation of analytical approximations
-
-Just as with most caclulations in this work, the derivations are straightforward applications of Gaussian products. That is, if
+Applying equations (\ref{N}), (\ref{xbarfinal}) and (\ref{Gfinal}) of
+the main text recovers system (\ref{comm_dynamics}) of the main text.
+
+\hypertarget{the-relation-between-competition-coefficients-and-selection}{%
+\subsection{\texorpdfstring{The relation between competition
+coefficients and selection
+\label{coef_grad_moms}}{The relation between competition coefficients and selection }}\label{the-relation-between-competition-coefficients-and-selection}}
+
+\hypertarget{derivation-of-analytical-approximations}{%
+\subsubsection{Derivation of analytical
+approximations}\label{derivation-of-analytical-approximations}}
+
+Just as with most caclulations in this work, the derivations are
+straightforward applications of Gaussian products. That is, if
 
@@
 
-#### Caclulating $\mathrm{Cov}_{f_{\bar X}}(\alpha,\gamma)$
+\hypertarget{caclulating-mathrmcov_f_bar-xalphagamma}{%
+\paragraph{\texorpdfstring{Caclulating
+\(\mathrm{Cov}_{f_{\bar X}}(\alpha,\gamma)\)}{Caclulating \textbackslash{}mathrm\{Cov\}\_\{f\_\{\textbackslash{}bar X\}\}(\textbackslash{}alpha,\textbackslash{}gamma)}}\label{caclulating-mathrmcov_f_bar-xalphagamma}}
 
@@
 
-#### Caclulating $\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)$
-
-To calculate moments of $|\beta|$ we note that, as a random variable, $|\beta|$ takes a folded normal distribution.
-Setting $\Phi(x)$ equal to the cumulative density function of the standard normal distribution and using the properties of the folded normal distribution, we find
+\hypertarget{caclulating-mathrmcov_f_bar-xalphabeta}{%
+\paragraph{\texorpdfstring{Caclulating
+\(\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)\)}{Caclulating \textbackslash{}mathrm\{Cov\}\_\{f\_\{\textbackslash{}bar X\}\}(\textbackslash{}alpha,\textbar{}\textbackslash{}beta\textbar{})}}\label{caclulating-mathrmcov_f_bar-xalphabeta}}
+
+To calculate moments of \(|\beta|\) we note that, as a random variable,
+\(|\beta|\) takes a folded normal distribution. Setting \(\Phi(x)\)
+equal to the cumulative density function of the standard normal
+distribution and using the properties of the folded normal distribution,
+we find
 
@@
 
-In attempt to calculate $\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)$ we find
+In attempt to calculate \(\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)\) we
+find
 
@@
 
-Just as we used the folded normal to find $\overline{|\beta|}$ and $\mathrm{Var}_{f_{\bar X}}(|\beta|)$, we can calculate $\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)$ by considering
+Just as we used the folded normal to find \(\overline{|\beta|}\) and
+\(\mathrm{Var}_{f_{\bar X}}(|\beta|)\), we can calculate
+\(\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)\) by considering
 
@@
 
-#### Starting the caclulation of $\mathrm{Cov}_{f_{\bar X}}(\alpha,\mathfrak{C})$ \label{caC}
+\hypertarget{starting-the-caclulation-of-mathrmcov_f_bar-xalphamathfrakc}{%
+\paragraph{\texorpdfstring{Starting the caclulation of
+\(\mathrm{Cov}_{f_{\bar X}}(\alpha,\mathfrak{C})\)
+\label{caC}}{Starting the caclulation of \textbackslash{}mathrm\{Cov\}\_\{f\_\{\textbackslash{}bar X\}\}(\textbackslash{}alpha,\textbackslash{}mathfrak\{C\}) }}\label{starting-the-caclulation-of-mathrmcov_f_bar-xalphamathfrakc}}
 
@@
 
-Note that the random variable $\delta=\bar x_i-\bar x_j$ is a mean zero Gaussian random variable with variance $2V_{\bar X}$.
-We write the probability density function of $\delta$ as $f_{\Delta}(\delta)$.
-Substituting in $\delta$, we can write
+Note that the random variable \(\delta=\bar x_i-\bar x_j\) is a mean
+zero Gaussian random variable with variance \(2V_{\bar X}\). We write
+the probability density function of \(\delta\) as
+\(f_{\Delta}(\delta)\). Substituting in \(\delta\), we can write
 
@@
 
-From this expression, we see properties of the folded normal distribution can be used to calculate several components of the intregal $\mathrm{Cov}_{f_{\bar X}}(\alpha,\mathfrak{C})$, but a major technical challenge lies in calculating
+From this expression, we see properties of the folded normal
+distribution can be used to calculate several components of the intregal
+\(\mathrm{Cov}_{f_{\bar X}}(\alpha,\mathfrak{C})\), but a major
+technical challenge lies in calculating
 
@@
 
-Instead of overcoming this challenge to find an analytical form of $\mathrm{Cov}_{f_{\bar X}}(\alpha,\mathfrak{C})$ we turn to a numerical approach outlined in the following section.
-
-### Numerical estimates for heterogeneous $N$ and $G$
-
-```{r, echo=F, fig.align='center', fig.cap='Numerical estimate for the correlations of selection gradients and competition coefficients.'}
-knitr::include_graphics('/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/corrs.png')
-```
-
-Details on simulations, table of parameters, distributions of $a$ and $c$.
+Instead of overcoming this challenge to find an analytical form of
+\(\mathrm{Cov}_{f_{\bar X}}(\alpha,\mathfrak{C})\) we turn to a
+numerical approach outlined in the following section.
+
+\hypertarget{numerical-estimates-for-heterogeneous-n-and-g}{%
+\subsubsection{\texorpdfstring{Numerical estimates for heterogeneous
+\(N\) and
+\(G\)}{Numerical estimates for heterogeneous N and G}}\label{numerical-estimates-for-heterogeneous-n-and-g}}
+
+\begin{figure}
+
+{\centering \includegraphics{/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/corrs}
+
+}
+
+\caption{Numerical estimate for the correlations of selection gradients and competition coefficients.}\label{fig:unnamed-chunk-13}
+\end{figure}
+
+Details on simulations, table of parameters, distributions of \(a\) and
+\(c\).
 
@@
 
-# References
+\hypertarget{references}{%
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
+
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-AbdalaRoberts2014}{}%
+Abdala-Roberts, Luis, and Kailen A. Mooney. 2014. ``Ecological and
+Evolutionary Consequences of Plant Genotype Diversity in a Tri-Trophic
+System.'' \emph{Ecology} 95 (10). Wiley: 2879--93.
+
+\leavevmode\hypertarget{ref-9781107648258}{}%
+Ågren, Göran I., and Folke O. Andersson. 2012. \emph{Terrestrial
+Ecosystem Ecology: Principles and Applications}. Cambridge University
+Press.
+
+\leavevmode\hypertarget{ref-Barton2017}{}%
+Barton, N.H., A.M. Etheridge, and A. Véber. 2017. ``The Infinitesimal
+Model: Definition, Derivation, and Implications.'' \emph{Theoretical
+Population Biology} 118 (December). Elsevier BV: 50--73.
+
+\leavevmode\hypertarget{ref-Barton2019}{}%
+Barton, Nick, and Alison Etheridge. 2019. ``Mathematical Models in
+Population Genetics.'' Wiley.
+
+\leavevmode\hypertarget{ref-Barton2013}{}%
+Barton, Nick, Alison Etheridge, and Amandine Véber. 2013. ``Modelling
+Evolution in a Spatial Continuum.'' \emph{Journal of Statistical
+Mechanics: Theory and Experiment} 2013 (01). IOP Publishing: P01002.
+
+\leavevmode\hypertarget{ref-Bertoin2003}{}%
+Bertoin, Jean, and Jean-François Le Gall. 2003. ``Stochastic Flows
+Associated to Coalescent Processes.'' \emph{Probability Theory and
+Related Fields} 126 (2). Springer Science; Business Media: 261--88.
+
+\leavevmode\hypertarget{ref-bulmer1980}{}%
+Bulmer. 1980. \emph{The Mathematical Theory of Quantitative Genetics}.
+Oxford University Press.
+
+\leavevmode\hypertarget{ref-Brger1986}{}%
+Bürger, Reinhard. 1986. ``On the Maintenance of Genetic Variation:
+Global Analysis of Kimuras Continuum-of-Alleles Model.'' \emph{Journal
+of Mathematical Biology} 24 (3). Springer Nature: 341--51.
+
+\leavevmode\hypertarget{ref-9780471986539}{}%
+---------. 2000. \emph{The Mathematical Theory of Selection,
+Recombination, and Mutation}. Wiley.
+
+\leavevmode\hypertarget{ref-Cantrell2004}{}%
+Cantrell, Robert Stephen, and Chris Cosner. 2004. \emph{Spatial Ecology
+via Reaction-Diffusion Equations}. Wiley.
+
+\leavevmode\hypertarget{ref-Case1986}{}%
+Case, Ted J., and Mark L. Taper. 1986. ``ON THE COEXISTENCE AND
+COEVOLUTION OF ASEXUAL AND SEXUAL COMPETITORS.'' \emph{Evolution} 40
+(2). Wiley: 366--87.
+
+\leavevmode\hypertarget{ref-Champagnat2006}{}%
+Champagnat, Nicolas, Régis Ferrière, and Sylvie Méléard. 2006.
+``Unifying Evolutionary Dynamics: From Individual Stochastic Processes
+to Macroscopic Models.'' \emph{Theoretical Population Biology} 69 (3).
+Elsevier BV: 297--321.
+
+\leavevmode\hypertarget{ref-Chesson2000}{}%
+Chesson, Peter. 2000. ``Mechanisms of Maintenance of Species
+Diversity.'' \emph{Annual Review of Ecology and Systematics} 31 (1).
+Annual Reviews: 343--66.
+
+\leavevmode\hypertarget{ref-jeffreyconner2004}{}%
+Conner, Jeffrey K. 2004. \emph{A Primer of Ecological Genetics}. Sinauer
+Associates.
+
+\leavevmode\hypertarget{ref-9781932846126}{}%
+Crow, James F., and Motoo Kimura. 1970. \emph{An Introduction to
+Population Genetics Theory}. The Blackburn Press.
+
+\leavevmode\hypertarget{ref-Crutsinger2015}{}%
+Crutsinger, Gregory M. 2015. ``A Community Genetics Perspective:
+Opportunities for the Coming Decade.'' \emph{New Phytologist} 210 (1).
+Wiley: 65--70.
+
+\leavevmode\hypertarget{ref-DaPrato2014}{}%
+Da Prato, Giuseppe, and Jerzy Zabczyk. 2014. \emph{Stochastic Equations
+in Infinite Dimensions}. Cambridge University Press.
+
+\leavevmode\hypertarget{ref-David2019}{}%
+David, Thomas I., Jonathan Storkey, and Carly J. Stevens. 2019.
+``Understanding How Changing Soil Nitrogen Affects Plantpollinator
+Interactions.'' \emph{Arthropod-Plant Interactions} 13 (5). Springer
+Science; Business Media LLC: 671--84.
+
+\leavevmode\hypertarget{ref-Dawson1975}{}%
+Dawson, Donald A. 1975. ``Stochastic Evolution Equations and Related
+Measure Processes.'' \emph{Journal of Multivariate Analysis} 5 (1).
+Elsevier BV: 1--52.
+
+\leavevmode\hypertarget{ref-Dawson1978}{}%
+---------. 1978. ``Geostochastic Calculus.'' \emph{Canadian Journal of
+Statistics} 6 (2). Wiley: 143--68.
+
+\leavevmode\hypertarget{ref-dawson1993measure}{}%
+---------. 1993. ``Measure-Valued Markov Processes.'' In \emph{École
+d'été de Probabilités de Saint-Flour Xxi-1991}, 1--260. Springer.
+
+\leavevmode\hypertarget{ref-Dbarre2015EvolutionOQ}{}%
+Débarre, Florence, Sam Yeaman, and Frédéric Guillaume. 2015. ``Evolution
+of Quantitative Traits Under a Migration-Selection Balance: When Does
+Skew Matter?'' \emph{The American Naturalist} 186 Suppl 1: S37--47.
+
+\leavevmode\hypertarget{ref-Doebeli1996}{}%
+Doebeli, Michael. 1996. ``Quantitative Genetics and Population
+Dynamics.'' \emph{Evolution} 50 (2). Wiley: 532--46.
+
+\leavevmode\hypertarget{ref-Engen2013}{}%
+Engen, Steinar, Russell Lande, and Bernt-Erik Sæther. 2013. ``A
+Quantitative Genetic Model of R- and K-Selection in a Fluctuating
+Population.'' \emph{The American Naturalist} 181 (6). University of
+Chicago Press: 725--36.
+
+\leavevmode\hypertarget{ref-alisonetheridge2000}{}%
+Etheridge, Alison M. 2000. \emph{An Introduction to Superprocesses}.
+American Mathematical Society.
+
+\leavevmode\hypertarget{ref-Etheridge2008}{}%
+---------. 2008. ``Drift, Draft and Structure: Some Mathematical Models
+of Evolution.'' In \emph{Stochastic Models in Biological Sciences}.
+Institute of Mathematics Polish Academy of Sciences.
+
+\leavevmode\hypertarget{ref-Etheridge1991}{}%
+Etheridge, Alison, and Peter March. 1991. ``A Note on Superprocesses.''
+\emph{Probability Theory and Related Fields} 89 (2). Springer: 141--47.
+
+\leavevmode\hypertarget{ref-lawrenceevans2010}{}%
+Evans, Lawrence C. 2010. \emph{Partial Differential Equations: Second
+Edition}. American Mathematical Society.
+
+\leavevmode\hypertarget{ref-9781470410544}{}%
+---------. 2014. \emph{An Introduction to Stochastic Differential
+Equations}. American Mathematical Society.
+
+\leavevmode\hypertarget{ref-Evans1994}{}%
+Evans, Steven N., and Edwin A. Perkins. 1994. ``Measure-Valued Branching
+Diffusions with Singular Interactions.'' \emph{Canadian Journal of
+Mathematics} 46 (1). Canadian Mathematical Society: 120--68.
+
+\leavevmode\hypertarget{ref-Ewens2004}{}%
+Ewens, Warren J. 2004. \emph{Mathematical Population Genetics}. Springer
+New York.
+
+\leavevmode\hypertarget{ref-kennethfalconer2014}{}%
+Falconer, Kenneth. 2014. \emph{Fractal Geometry: Mathematical
+Foundations and Applications}. Wiley.
+
+\leavevmode\hypertarget{ref-stanleyfarlow1993}{}%
+Farlow, Stanley J. 1993. \emph{Partial Differential Equations for
+Scientists and Engineers}. Dover.
+
+\leavevmode\hypertarget{ref-feller1951}{}%
+Feller, William. 1951. ``Diffusion Processes in Genetics.'' In
+\emph{Proceedings of the Second Berkeley Symposium on Mathematical
+Statistics and Probability}, 227--46. University of California Press.
+
+\leavevmode\hypertarget{ref-Felsenstein1975}{}%
+Felsenstein, Joseph. 1975. ``A Pain in the Torus: Some Difficulties with
+Models of Isolation by Distance.'' \emph{The American Naturalist} 109
+(967). University of Chicago Press: 359--68.
+
+\leavevmode\hypertarget{ref-Fisher1923}{}%
+Fisher, R. A. 1923. ``XXI.-on the Dominance Ratio.'' \emph{Proceedings
+of the Royal Society of Edinburgh} 42. Cambridge University Press:
+321--41.
+
+\leavevmode\hypertarget{ref-Fitzpatrick2015}{}%
+Fitzpatrick, Connor R., Anurag A. Agrawal, Nathan Basiliko, Amy P.
+Hastings, Marney E. Isaac, Michael Preston, and Marc T. J. Johnson.
+2015. ``The Importance of Plant Genotype and Contemporary Evolution for
+Terrestrial Ecosystem Processes.'' \emph{Ecology} 96 (10). Wiley:
+2632--42.
+
+\leavevmode\hypertarget{ref-Fitzpatrick2017}{}%
+Fitzpatrick, Connor R., Anna V. Mikhailitchenko, Daniel N. Anstett, and
+Marc T. J. Johnson. 2017. ``The Influence of Range-Wide Plant Genetic
+Variation on Soil Invertebrate Communities.'' \emph{Ecography} 41 (7).
+Wiley: 1135--46.
+
+\leavevmode\hypertarget{ref-FRANK2012}{}%
+Frank, S. A. 2012. ``Natural Selection. IV. The Price Equation.''
+\emph{Journal of Evolutionary Biology} 25 (6). Wiley: 1002--19.
+
+\leavevmode\hypertarget{ref-Fridley2017}{}%
+Fridley, Jason D. 2017. ``Plant Energetics and the Synthesis of
+Population and Ecosystem Ecology.'' Edited by David Gibson.
+\emph{Journal of Ecology} 105 (1). Wiley: 95--110.
+
+\leavevmode\hypertarget{ref-FUSSMANN2007}{}%
+Fussman, G. F., M. Loreau, and P. A. Abrams. 2007. ``Eco-Evolutionary
+Dynamics of Communities and Ecosystems.'' \emph{Functional Ecology} 21
+(3). Wiley: 465--77.
+
+\leavevmode\hypertarget{ref-Gomulkiewicz2017}{}%
+Gomulkiewicz, Richard, Stephen M. Krone, and Christopher H. Remien.
+2017. ``Evolution and the Duration of a Doomed Population.''
+\emph{Evolutionary Applications} 10 (5). Wiley: 471--84.
+
+\leavevmode\hypertarget{ref-Harmon2019}{}%
+Harmon, Luke J., Cecilia S. Andreazzi, Florence Débarre, Jonathan Drury,
+Emma E. Goldberg, Ayana B. Martins, Carlos J. Melián, et al. 2019.
+``Detecting the Macroevolutionary Signal of Species Interactions.''
+\emph{Journal of Evolutionary Biology} 32 (8). Wiley: 769--82.
+
+\leavevmode\hypertarget{ref-Harte2011}{}%
+Harte, John. 2011. \emph{Maximum Entropy and Ecology}. Oxford University
+Press.
+
+\leavevmode\hypertarget{ref-Harte2014}{}%
+Harte, John, and Erica A. Newman. 2014. ``Maximum Information Entropy: A
+Foundation for Ecological Theory.'' \emph{Trends in Ecology \&
+Evolution} 29 (7). Elsevier BV: 384--89.
+
+\leavevmode\hypertarget{ref-Hickerson2010}{}%
+Hickerson, M.J., B.C. Carstens, J. Cavender-Bares, K.A. Crandall, C.H.
+Graham, J.B. Johnson, L. Rissler, P.F. Victoriano, and A.D. Yoder. 2010.
+``Phylogeography's Past, Present, and Future: 10 Years After Avise,
+2000.'' \emph{Molecular Phylogenetics and Evolution} 54 (1). Elsevier
+BV: 291--301.
+
+\leavevmode\hypertarget{ref-Hofbauer1998}{}%
+Hofbauer, Josef, and Karl Sigmund. 1998. \emph{Evolutionary Games and
+Population Dynamics}. Cambridge University Press.
+
+\leavevmode\hypertarget{ref-Holt1987}{}%
+Holt, Robert D. 1987. ``On the Relation Between Niche Overlap and
+Competition: The Effect of Incommensurable Niche Dimensions.''
+\emph{Oikos} 48 (1). JSTOR: 110.
+
+\leavevmode\hypertarget{ref-Johnson2005}{}%
+Johnson, Toby, and Nick Barton. 2005. ``Theoretical Models of Selection
+and Mutation on Quantitative Traits.'' \emph{Philosophical Transactions
+of the Royal Society B: Biological Sciences} 360 (1459). The Royal
+Society: 1411--25.
+
+\leavevmode\hypertarget{ref-Kendall1966}{}%
+Kendall, David G. 1966. ``Branching Processes Since 1873.''
+\emph{Journal of the London Mathematical Society} s1-41 (1). Wiley:
+385--406.
+
+\leavevmode\hypertarget{ref-Kimmel2015}{}%
+Kimmel, Marek, and David E. Axelrod. 2015. \emph{Branching Processes in
+Biology}. Springer New York.
+
+\leavevmode\hypertarget{ref-Kimura1965}{}%
+Kimura, M. 1965. ``A Stochastic Model Concerning the Maintenance of
+Genetic Variability in Quantitative Characters.'' \emph{Proceedings of
+the National Academy of Sciences} 54 (3). Proceedings of the National
+Academy of Sciences: 731--36.
+
+\leavevmode\hypertarget{ref-Kimura1978}{}%
+Kimura, M., and J. F. Crow. 1978. ``Effect of Overall Phenotypic
+Selection on Genetic Change at Individual Loci.'' \emph{Proceedings of
+the National Academy of Sciences} 75 (12). Proceedings of the National
+Academy of Sciences: 6168--71.
+
+\leavevmode\hypertarget{ref-Kirkpatrick1727}{}%
+Kirkpatrick, Mark, Toby Johnson, and Nick Barton. 2002. ``General Models
+of Multilocus Evolution.'' \emph{Genetics} 161 (4). Genetics: 1727--50.
+
+\leavevmode\hypertarget{ref-kolmogorov1999elements}{}%
+Kolmogorov, A.N., and S.V. Fomin. 1999. \emph{Elements of the Theory of
+Functions and Functional Analysis}. v. 1. Dover.
+
+\leavevmode\hypertarget{ref-Konno1988}{}%
+Konno, N., and T. Shiga. 1988. ``Stochastic Partial Differential
+Equations for Some Measure-Valued Diffusions.'' \emph{Probability Theory
+and Related Fields} 79 (2). Springer Nature: 201--25.
+
+\leavevmode\hypertarget{ref-Kopp2006}{}%
+Kopp, Michael, and Sergey Gavrilets. 2006. ``Multilocus Gentics and the
+Coevolution of Quantitative Traits.'' \emph{Evolution} 60 (7). Wiley:
+1321--36.
+
+\leavevmode\hypertarget{ref-Klzsch2015}{}%
+Kölzsch, Andrea, Adriana Alzate, Frederic Bartumeus, Monique de Jager,
+Ellen J. Weerman, Geerten M. Hengeveld, Marc Naguib, Bart A. Nolet, and
+Johan van de Koppel. 2015. ``Experimental Evidence for Inherent Lévy
+Search Behaviour in Foraging Animals.'' \emph{Proceedings of the Royal
+Society B: Biological Sciences} 282 (1807). The Royal Society: 20150424.
+
+\leavevmode\hypertarget{ref-Kraft2007}{}%
+Kraft, Nathan J. B., William K. Cornwell, Campbell O. Webb, and David D.
+Ackerly. 2007. ``Trait Evolution, Community Assembly, and the
+Phylogenetic Structure of Ecological Communities.'' \emph{The American
+Naturalist} 170 (2). University of Chicago Press: 271--83.
+
+\leavevmode\hypertarget{ref-Krylov1981}{}%
+Krylov, N. V., and B. L. Rozovskii. 1981. ``Stochastic Evolution
+Equations.'' \emph{Journal of Soviet Mathematics} 16 (4). Springer
+Science; Business Media LLC: 1233--77.
+
+\leavevmode\hypertarget{ref-Lande1975}{}%
+Lande, Russell. 1975. ``The Maintenance of Genetic Variability by
+Mutation in a Polygenic Character with Linked Loci.'' \emph{Genetical
+Research} 26 (3). Cambridge University Press (CUP): 221--35.
+
+\leavevmode\hypertarget{ref-Lande1976}{}%
+---------. 1976. ``Natural Selection and Random Genetic Drift in
+Phenotypic Evolution.'' \emph{Evolution} 30 (2). Wiley: 314--34.
+
+\leavevmode\hypertarget{ref-pmid17248993}{}%
+---------. 1980. ``The Genetic Covariance between Characters Maintained
+by Pleiotropic Mutations.'' \emph{Genetics} 94 (1): 203--15.
+
+\leavevmode\hypertarget{ref-Lande1982}{}%
+---------. 1982. ``A Quantitative Genetic Theory of Life History
+Evolution.'' \emph{Ecology} 63 (3). Wiley: 607--15.
+
+\leavevmode\hypertarget{ref-Lande1995}{}%
+---------. 1995. ``Mutation and Conservation.'' \emph{Conservation
+Biology} 9 (4). Wiley: 782--91.
+
+\leavevmode\hypertarget{ref-Lande1983}{}%
+Lande, Russell, and Stevan J. Arnold. 1983. ``The Measurement of
+Selection on Correlated Characters.'' \emph{Evolution} 37 (6). JSTOR:
+1210.
+
+\leavevmode\hypertarget{ref-Lande2003}{}%
+Lande, Russell, Steinar Engen, and Bernt-Erik SÆther. 2003.
+``Demographic and Environmental Stochasticity.'' In \emph{Stochastic
+Population Dynamics in Ecology and Conservation}, 1--24. Oxford
+University Press.
+
+\leavevmode\hypertarget{ref-Lande2009}{}%
+Lande, Russell, Steinar Engen, and Bernt-Erik Sæther. 2009. ``An
+Evolutionary Maximum Principle for Density-Dependent Population Dynamics
+in a Fluctuating Environment.'' \emph{Philosophical Transactions of the
+Royal Society B: Biological Sciences} 364 (1523). The Royal Society:
+1511--8.
+
+\leavevmode\hypertarget{ref-Landis2017}{}%
+Landis, Michael J., and Joshua G. Schraiber. 2017. ``Pulsed Evolution
+Shaped Modern Vertebrate Body Sizes.'' \emph{Proceedings of the National
+Academy of Sciences} 114 (50). Proceedings of the National Academy of
+Sciences: 13224--9.
+
+\leavevmode\hypertarget{ref-9780691080628}{}%
+Levins, Richard. 1968. \emph{Evolution in Changing Environments: Some
+Theoretical Explorations. (MPB-2) (Monographs in Population Biology)}.
+Princeton University Press.
+
+\leavevmode\hypertarget{ref-zeng1998absolute}{}%
+Li, Zeng-Hu. 1998. ``Absolute Continuity of Measure Branching Processes
+with Interaction.'' \emph{Chinese Journal of Applied Probability and
+Statistics} 14. Citeseer: 231--42.
+
+\leavevmode\hypertarget{ref-Lion2018}{}%
+Lion, Sébastien. 2018. ``Theoretical Approaches in Evolutionary Ecology:
+Environmental Feedback as a Unifying Perspective.'' \emph{The American
+Naturalist} 191 (1). University of Chicago Press: 21--44.
+
+\leavevmode\hypertarget{ref-michelloreau2010}{}%
+Loreau, Michel. 2010. \emph{From Populations to Ecosystems: Theoretical
+Foundations for a New Ecological Synthesis}. Princeton University Press.
+
+\leavevmode\hypertarget{ref-michaellynch1998}{}%
+Lynch, Michael, and Bruce Walsh. 1998. \emph{Genetics and Analysis of
+Quantitative Traits}. Sinauer Associates is an imprint of Oxford
+University Press.
+
+\leavevmode\hypertarget{ref-Arthur1969}{}%
+MacArthur, Robert H. 1969. ``Species Packing, and what Competition
+Minimizes.'' \emph{Proceedings of the National Academy of Sciences} 64
+(4). Proceedings of the National Academy of Sciences: 1369--71.
+
+\leavevmode\hypertarget{ref-MacArthur1970}{}%
+---------. 1970. ``Species Packing and Competitive Equilibrium for Many
+Species.'' \emph{Theoretical Population Biology} 1 (1). Elsevier BV:
+1--11.
+
+\leavevmode\hypertarget{ref-9780691023823}{}%
+---------. 1972. \emph{Geographical Ecology}. Princeton University
+Press.
+
+\leavevmode\hypertarget{ref-Macarthur1967}{}%
+MacArthur, Robert H., and Richard Levins. 1967. ``The Limiting
+Similarity, Convergence, and Divergence of Coexisting Species.''
+\emph{The American Naturalist} 101 (921). University of Chicago Press:
+377--85.
+
+\leavevmode\hypertarget{ref-Manceau2016}{}%
+Manceau, Marc, Amaury Lambert, and Hélène Morlon. 2016. ``A Unifying
+Comparative Phylogenetic Framework Including Traits Coevolving Across
+Interacting Lineages.'' \emph{Systematic Biology}, December. Oxford
+University Press (OUP), syw115.
+
+\leavevmode\hypertarget{ref-Marx2017}{}%
+Marx, Hannah E., Cédric Dentant, Julien Renaud, Romain Delunel, David C.
+Tank, and Sébastien Lavergne. 2017. ``Riders in the Sky (Islands): Using
+a Mega-Phylogenetic Approach to Understand Plant Species Distribution
+and Coexistence at the Altitudinal Limits of Angiosperm Plant Life.''
+\emph{Journal of Biogeography} 44 (11). Wiley: 2618--30.
+
+\leavevmode\hypertarget{ref-markmcpeek2017}{}%
+McPeek, Mark A. 2017. \emph{Evolutionary Community Ecology}. Princeton
+University Press.
+
+\leavevmode\hypertarget{ref-DeMeester2018}{}%
+Meester, Luc De, Kristien I. Brans, Lynn Govaert, Caroline Souffreau,
+Shinjini Mukherjee, Héléne Vanvelk, Konrad Korzeniowski, et al. 2018.
+``Analyzing Eco-Evolutionary Dynamics - the Challenging Complexity of
+the Real World.'' \emph{Functional Ecology}. Wiley.
+
+\leavevmode\hypertarget{ref-meleard1992interacting}{}%
+Méléard, M, and S Roelly. 1992. ``Interacting Branching Measure
+Processes.'' \emph{Stochastic Partial Differential Equations and
+Applications (G. Da Prato and L. Tubaro, Eds.)}, 246--56.
+
+\leavevmode\hypertarget{ref-Mlard1993}{}%
+---------. 1993. ``Interacting Measure Branching Processes. Some Bounds
+for the Support.'' \emph{Stochastics and Stochastic Reports} 44 (1-2).
+Informa UK Limited: 103--21.
+
+\leavevmode\hypertarget{ref-Mubayi2019}{}%
+Mubayi, Anuj, Christopher Kribs, Viswanathan Arunachalam, and Carlos
+Castillo-Chavez. 2019. ``Studying Complexity and Risk Through Stochastic
+Population Dynamics: Persistence, Resonance, and Extinction in
+Ecosystems.'' In \emph{Handbook of Statistics}, 157--93. Elsevier.
+
+\leavevmode\hypertarget{ref-9780674023383}{}%
+Nowak, Martin A. 2006. \emph{Evolutionary Dynamics: Exploring the
+Equations of Life}. Belknap Press.
+
+\leavevmode\hypertarget{ref-Nuismer2005}{}%
+Nuismer, Scott L., Michael Doebeli, and Danny Browning. 2005. ``The
+Coevolutionary Dynamics of Antagonistic Interactions Mediated by
+Quantitative Traits with Evolving Variances.'' \emph{Evolution} 59 (10).
+The Society for the Study of Evolution: 2073.
+
+\leavevmode\hypertarget{ref-Nuismer2014}{}%
+Nuismer, Scott L., and Luke J. Harmon. 2014. ``Predicting Rates of
+Interspecific Interaction from Phylogenetic Trees.'' Edited by Jerome
+Chave. \emph{Ecology Letters} 18 (1). Wiley: 17--27.
+
+\leavevmode\hypertarget{ref-Nuismer2012}{}%
+Nuismer, Scott L., Pedro Jordano, and Jordi Bascompte. 2012.
+``Coevolution and the Architecture of Mutualistic Networks.''
+\emph{Evolution} 67 (2). Wiley: 338--54.
+
+\leavevmode\hypertarget{ref-Nuismer2007}{}%
+Nuismer, Scott L., Benjamin J. Ridenhour, and Benjamin P. Oswald. 2007.
+``Antagonistic Coevolution Mediated by Phenotypic Differenes Between
+Quantitative Traits.'' \emph{Evolution} 61 (8). Wiley: 1823--34.
+
+\leavevmode\hypertarget{ref-Nuismer2018}{}%
+Nuismer, Scott L., Bob Week, and Marcelo A. Aizen. 2018. ``Coevolution
+Slows the Disassembly of Mutualistic Networks.'' \emph{The American
+Naturalist} 192 (4). University of Chicago Press: 490--502.
+
+\leavevmode\hypertarget{ref-VanNuland2019}{}%
+Nuland, Michael E. Van, Ian M. Ware, Joseph K. Bailey, and Jennifer A.
+Schweitzer. 2019. ``Ecosystem Feedbacks Contribute to Geographic
+Variation in Plant-Soil Eco-Evolutionary Dynamics Across a Fertility
+Gradient.'' Edited by Franziska Brunner. \emph{Functional Ecology} 33
+(1). Wiley: 95--106.
+
+\leavevmode\hypertarget{ref-PAGE2002}{}%
+Page, Karen M., and Martin A. Nowak. 2002. ``Unifying Evolutionary
+Dynamics.'' \emph{Journal of Theoretical Biology} 219 (1). Elsevier BV:
+93--98.
+
+\leavevmode\hypertarget{ref-Parachnowitsch2018}{}%
+Parachnowitsch, Amy L, Jessamyn S Manson, and Nina Sletvold. 2018.
+``Evolutionary Ecology of Nectar.'' \emph{Annals of Botany} 123 (2).
+Oxford University Press (OUP): 247--61.
+
+\leavevmode\hypertarget{ref-Park2017}{}%
+Park, Briton, Matthew T. Rutter, Charles B. Fenster, V. Vaughan Symonds,
+Mark C. Ungerer, and Jeffrey P. Townsend. 2017. ``Distributions of
+Mutational Effects and the Estimation of Directional Selection in
+Divergent Lineages ofArabidopsis Thaliana.'' \emph{Genetics} 206 (4).
+Genetics Society of America: 2105--17.
+
+\leavevmode\hypertarget{ref-Parsons2010}{}%
+Parsons, Todd L., Christopher Quince, and Joshua B. Plotkin. 2010.
+``Some Consequences of Demographic Stochasticity in Population
+Genetics.'' \emph{Genetics} 185 (4). Genetics Society of America:
+1345--54.
+
+\leavevmode\hypertarget{ref-Patel2019}{}%
+Patel, Swati, and Reinhard Bürger. 2019. ``Eco-Evolutionary Feedbacks
+Between Prey Densities and Linkage Disequilibrium in the Predator
+Maintain Diversity.'' \emph{Evolution} 73 (8). Wiley: 1533--48.
+
+\leavevmode\hypertarget{ref-Perkins1991}{}%
+Perkins, Edwin A. 1991. ``Conditional Dawson-Watanabe Processes and
+Fleming-Viot Processes.'' In \emph{Seminar on Stochastic Processes,
+1991}, 143--56. Birkhäuser Boston.
+
+\leavevmode\hypertarget{ref-Perkins1992}{}%
+---------. 1992. ``Measure-Valued Branching Diffusions with Spatial
+Interactions.'' \emph{Probability Theory and Related Fields} 94 (2).
+Springer Science; Business Media LLC: 189--245.
+
+\leavevmode\hypertarget{ref-edwinarendperkins1995}{}%
+---------. 1995. \emph{On the Martingale Problem for Interactive
+Measure-Valued Branching Diffusions}. Amer Mathematical Society.
+
+\leavevmode\hypertarget{ref-PRICE1970}{}%
+Price, George R. 1970. ``Selection and Covariance.'' \emph{Nature} 227
+(5257). Springer Nature: 520--21.
+
+\leavevmode\hypertarget{ref-Queller2017}{}%
+Queller, David C. 2017. ``Fundamental Theorems of Evolution.'' \emph{The
+American Naturalist} 189 (4). University of Chicago Press: 345--53.
+
+\leavevmode\hypertarget{ref-Reimers1989}{}%
+Reimers, Mark. 1989. ``One Dimensional Stochastic Partial Differential
+Equations and the Branching Measure Diffusion.'' \emph{Probability
+Theory and Related Fields} 81 (3). Springer Nature: 319--40.
+
+\leavevmode\hypertarget{ref-Robertson1966}{}%
+Robertson, Alan. 1966. ``A Mathematical Model of the Culling Process in
+Dairy Cattle.'' \emph{Animal Science} 8 (1). Cambridge University Press:
+95--108.
+
+\leavevmode\hypertarget{ref-joanroughgarden1979}{}%
+Roughgarden, Joan. 1979. \emph{Theory of Population Genetics and
+Evolutionary Ecology: An Introduction}. Macmillan.
+
+\leavevmode\hypertarget{ref-Rudman2017}{}%
+Rudman, Seth M., Matthew A. Barbour, Katalin Csilléry, Phillip Gienapp,
+Frederic Guillaume, Nelson G. Hairston Jr, Andrew P. Hendry, et al.
+2017. ``What Genomic Data Can Reveal About Eco-Evolutionary Dynamics.''
+\emph{Nature Ecology \& Evolution} 2 (1). Springer Nature: 9--15.
+
+\leavevmode\hypertarget{ref-Schreiber2017}{}%
+Schreiber, Sebastian J. 2017. ``Coexistence in the Face of
+Uncertainty.'' In \emph{Recent Progress and Modern Challenges in Applied
+Mathematics, Modeling and Computational Science}, 349--84. Springer New
+York.
+
+\leavevmode\hypertarget{ref-Schreiber2018}{}%
+Schreiber, Sebastian J., Swati Patel, and Casey terHorst. 2018.
+``Evolution as a Coexistence Mechanism: Does Genetic Architecture
+Matter?'' \emph{The American Naturalist} 191 (3). University of Chicago
+Press: 407--20.
+
+\leavevmode\hypertarget{ref-Schuster1983}{}%
+Schuster, Peter, and Karl Sigmund. 1983. ``Replicator Dynamics.''
+\emph{Journal of Theoretical Biology} 100 (3). Elsevier BV: 533--38.
+
+\leavevmode\hypertarget{ref-Skovmand2018}{}%
+Skovmand, Lotte H., Charles C.Y. Xu, Maria R. Servedio, Patrik Nosil,
+Rowan D.H. Barrett, and Andrew P. Hendry. 2018. ``Keystone Genes.''
+\emph{Trends in Ecology \& Evolution} 33 (9). Elsevier BV: 689--700.
+
+\leavevmode\hypertarget{ref-Sterner2008}{}%
+Sterner, R.W., and J.J. Elser. 2008. ``Ecological Stoichiometry:
+Overview.'' In \emph{Encyclopedia of Ecology}, 1101--16. Elsevier.
+
+\leavevmode\hypertarget{ref-Taylor1978}{}%
+Taylor, Peter D., and Leo B. Jonker. 1978. ``Evolutionary Stable
+Strategies and Game Dynamics.'' \emph{Mathematical Biosciences} 40
+(1-2). Elsevier BV: 145--56.
+
+\leavevmode\hypertarget{ref-davidtilman1982}{}%
+Tilman, David. 1982. \emph{Resource Competition and Community
+Structure}. Princeton University Press.
+
+\leavevmode\hypertarget{ref-Turelli1980}{}%
+Turelli, Michael. 1980. ``Niche Overlap and Invasion of Competitors in
+Random Environments II. The Effects of Demographic Stochasticity.'' In
+\emph{Biological Growth and Spread}, 119--29. Springer Berlin
+Heidelberg.
+
+\leavevmode\hypertarget{ref-Turelli1984}{}%
+---------. 1984. ``Heritable Genetic Variation via Mutation-Selection
+Balance: Lerchs Zeta Meets the Abdominal Bristle.'' \emph{Theoretical
+Population Biology} 25 (2). Elsevier: 138--93.
+
+\leavevmode\hypertarget{ref-Turelli1986}{}%
+---------. 1986. ``Gaussian Versus Non-Gaussian Genetic Analyses of
+Polygenic Mutation-Selection Balance.'' In \emph{Evolutionary Processes
+and Theory}, 607--28. Academic Press.
+
+\leavevmode\hypertarget{ref-Turelli2017}{}%
+---------. 2017. ``Commentary: Fisher's Infinitesimal Model: A Story for
+the Ages.'' \emph{Theoretical Population Biology} 118 (December).
+Elsevier BV: 46--49.
+
+\leavevmode\hypertarget{ref-pmid7851785}{}%
+Turelli, Michael, and Nick Barton. 1994. ``Genetic and statistical
+analyses of strong selection on polygenic traits: what, me normal?''
+\emph{Genetics} 138 (3): 913--41.
+
+\leavevmode\hypertarget{ref-Volpert2014}{}%
+Volpert, Vitaly. 2014. \emph{Elliptic Partial Differential Equations:
+Volume 2: Reaction-Diffusion Equations}. Springer Basel.
+
+\leavevmode\hypertarget{ref-Walsh}{}%
+Walsh, John B. 1986. ``An Introduction to Stochastic Partial
+Differential Equations.'' In \emph{Lecture Notes in Mathematics},
+265--439. Springer Berlin Heidelberg.
+
+\leavevmode\hypertarget{ref-Watanabe1968}{}%
+Watanabe, Shinzo. 1968. ``A Limit Theorem of Branching Processes and
+Continuous State Branching Processes.'' \emph{Journal of Mathematics of
+Kyoto University} 8 (1). Duke University Press: 141--67.
+
+\leavevmode\hypertarget{ref-Wright97}{}%
+Wright, Sewall. 1931. ``Evolution in Mendelian Populations.''
+\emph{Genetics} 16 (2). Genetics: 97--159.
+
+\leavevmode\hypertarget{ref-Xiao2015}{}%
+Xiao, Xiao, Daniel J. McGlinn, and Ethan P. White. 2015. ``A Strong Test
+of the Maximum Entropy Theory of Ecology.'' \emph{The American
+Naturalist} 185 (3). University of Chicago Press: E70--E80.
+
+\leavevmode\hypertarget{ref-zheng2004nonlinear}{}%
+Zheng, Songmu. 2004. \emph{Nonlinear Evolution Equations}. Boca Raton,
+Fla: Chapman \& Hall/CRC Press.
+
+\end{document}

### 2020-04-18 .A White Noise Approach To Evolutionary Ecology/draft.tex
final draft

@@
 selection gradients and competition coefficients. This illustrative
-investigation suggests selection gradients can be either positively or
-negatively correlated with competition coefficients depending on the
-ratio of interspecific trait variation to intraspecific trait variation.
+investigation suggests absolute values of linear selection gradients are
+negatively correlated with competition coefficients when the ratio of
+interspecific trait variation to intraspecific trait variation is
+approximately less than \(1.25\) and positively correlated otherwise.
+The inverse relationship holds for quadratic selection gradients except
+with threshold diversity ratio equal to unity instead of \(1.25\).
 Hence, this approach contributes to the development of a synthetic
@@
 2006) which includes our approach as a special case. We note here there
-are two primary differences between our aims and those of Champagnat,
-Ferrière, and Méléard (2006). First, instead of developing a unifying
-approach to evolutionary theory, we focus on developing a stochastic
-synthesis of population dynamics and quantitative genetics including
-specific expressions for the dynamics of abundance, mean trait and
-additive genetic variance. Second, although Champagnat, Ferrière, and
-Méléard (2006) provide rigorous proofs of results relevant to our
-approach, they do not communicate the nessecary tools we make use of for
-deriving the dynamics of abundance, mean trait and additive genetic
-variance. We aim to communicate these tools to a wide audience by
-providing an intuitive, yet accurate treatment of white noise and
-associated heuristics useful for deriving new stochastic models of
-evolutionary ecology.
-
-Reviewing the above accomplishments reveals a beautiful synthesis of
-evolution and population ecology. However, it also reveals a gap in
-theoretical approaches to incorporate the intrinsically random nature of
-populations. Specifically, in theoretical quantitative genetics the
-derivation of a population's response to random genetic drift is derived
-in discrete time under the assumption of constant effective population
-size using arguments based on properties of random samples (Lande 1976).
-Though this approach conveniently mimics the formalism provided by the
-Wright-Fisher model of population genetics, real population sizes
-fluctuate over time. Furthermore, since these fluctuations are
-themselves stochastic, it seems natural to derive expressions for the
-evolutionary response to demographic stochasticity and consider how the
-results relate to characterizations of random genetic drift. This can be
-done in continuous time for population genetic models without too much
-technical overhead, assuming a finite number of alleles (Lande, Engen,
-and Sæther 2009; Parsons, Quince, and Plotkin 2010; Gomulkiewicz, Krone,
-and Remien 2017). However, for populations with a continuum of types,
-such as a quantitative trait, finding a formal approach to derive the
-evolutionary response to demographic stochasticity has remained a vexing
-mathematical challenge. Previous work integrating stochastic processes
-with quantitative genetic models have profited from the simplifying
-assumption that phenotypic and genetic variances remain constant over
-time, yeilding analytical approximations for the joint dynamics of
-abundance and the mean of a quantitative trait (Engen, Lande, and Sæther
-2013). These results not only ignore the evolution genetic variance, but
-they focus on environmental stochasticity and assume sufficiently large
-populations so that demographic stochasticity is negligible. Here we
-account for both demographic stochasticity and the evolution of additive
-genetic variance by combining the calculus of white noise with results
-on rescaled limits of branching Brownian motion processes (BBM) and
-stochastic partial differential equations (SPDE), which are stochastic
-analogs of deterministic partial differential equations (PDE). Our goal
-has two components: 1) Establish a novel synthetic approach to
-theoretical evolutionary ecology that provides a formal connection
-between demographic stochasticity and random genetic drift in the
-context of quantitative traits. 2) Communicate some useful properties of
-space-time white noise, BBM and SPDE to as wide of audience as possible.
-With this goal in mind we will not provide a rigorous treatment of any
-of these deep subjects. Instead, we introduce a set of heuristics that
-only require the basic concepts of Reimann integration, partial
-differentiation and some exposure to Brownian motion and stochastic
-ordinary differential equations (SDE). For a concise introduction to SDE
-and Brownian motion, we recommend the primer by Evans (2014). Rigorous
-treatments of SPDE and rescaled limits of BBM can be found in Walsh
-(1986) and Dawson (1993) respectively.
+are two primary differences between the results presented in this paper
+and those of Champagnat, Ferrière and Méléard (2006). First, instead of
+developing a unifying approach to evolutionary theory, we focus on
+developing a stochastic synthesis of population dynamics and
+quantitative genetics including specific expressions for the dynamics of
+abundance, mean trait and additive genetic variance. Second, although
+Champagnat, Ferrière and Méléard (2006) provide rigorous proofs of
+results relevant to our approach, they do not communicate the nessecary
+tools we make use of for deriving the dynamics of abundance, mean trait
+and additive genetic variance. We aim to communicate these tools to a
+wide audience by striking a balance between intuitive appeal and
+mathematical rigor.
+
+Although reviewing previously established theoretical results reveals a
+beautiful synthesis of evolution and population ecology, it also reveals
+a gap in theoretical approaches to incorporate the intrinsically random
+nature of populations. Specifically, in theoretical quantitative
+genetics the derivation of a population's response to random genetic
+drift is derived in discrete time under the assumption of constant
+effective population size using arguments based on properties of random
+samples (Lande 1976). Though this approach conveniently mimics the
+formalism provided by the Wright-Fisher model of population genetics,
+real population sizes fluctuate over time. Furthermore, since these
+fluctuations are themselves stochastic, it seems natural to derive
+expressions for the evolutionary response to demographic stochasticity
+and consider how the results relate to characterizations of random
+genetic drift. This can be done in continuous time for population
+genetic models without too much technical overhead, assuming a finite
+number of alleles (Lande, Engen, and Sæther 2009; Parsons, Quince, and
+Plotkin 2010; Gomulkiewicz, Krone, and Remien 2017). However, for
+populations with a continuum of types, such as a quantitative trait,
+finding a formal approach to derive the evolutionary response to
+demographic stochasticity has remained a vexing mathematical challenge.
+In this paper we close this gap by combining the calculus of white noise
+with results on rescaled limits of branching Brownian motion processes
+(BBM) and stochastic partial differential equations (SPDE), which are
+stochastic analogs of deterministic partial differential equations
+(PDE). Our goal has two components: 1) Establish a novel synthetic
+approach to theoretical evolutionary ecology that provides a formal
+connection between demographic stochasticity and random genetic drift in
+the context of quantitative traits. 2) Communicate some useful
+properties of space-time white noise, BBM and SPDE to as wide of
+audience as possible. With this goal in mind we will not provide a
+rigorous treatment of any of these deep subjects. Instead, we introduce
+a set of heuristics that only require the basic concepts of Reimann
+integration, partial differentiation and some exposure to Brownian
+motion and stochastic ordinary differential equations (SDE). For a
+concise introduction to SDE and Brownian motion, we recommend the primer
+by Evans (2014). Rigorous treatments of SPDE and rescaled limits of BBM
+can be found in Walsh (1986) and Dawson (1993) respectively.
 
@@
 made in combining models of evolutionary and abundance dynamics. Second,
-we formally derive general SDE that express the joint dynamics of
-abundance, mean trait and additive genetic variance for populations
-exhibiting a wide range of potential phenotypic distributions. Third, we
-simplify these SDE by approximating the abundance density of phenotypes
-with a Gaussian curve and integrate the results with a model of
-imperfect inheritance to find formula that generalize classical
-quantitative genetic expressions. This includes an expression for
-additive genetic variance analogous to the expression for the response
+we formally derive general SDE driven by demographic stochasticity that
+express the joint dynamics of abundance, mean trait and additive genetic
+variance for populations exhibiting a wide range of potential phenotypic
+distributions. Although these dynamics have been considered before while
+accounting for demographic stochasticity (Polechová and Barton 2015), to
+our knowledge this is the first time they have been rigorously derived
+from first principles. Third, we simplify these SDE by approximating the
+abundance density of phenotypes with a Gaussian curve and integrate the
+results with a model of imperfect inheritance to find formula that
+generalize classical quantitative genetic expressions. This includes an
+expression for the response of additive genetic variance to selection in
+terms of fitness gradients analogous to the expression for the response
 in mean phenotype found by Lande (1976). Fourth, following classical
@@
 investigating the relationship between selection gradients and
-competition coefficients.
+competition coefficients, discovering the conditions under which these
+quantities positively or negatively covary.
 
@@
 model of coevolution, we begin with \S\ref{deterministic} by briefly
-summarizing derivations of deterministic dynamics of biological
-populations. Starting with a PDE (which we refer to as the Deterministic
+summarizing derivations of deterministic evolutionary and demographic
+dynamics. Starting with a PDE (which we refer to as the Deterministic
 Asexual Gaussian allelic model with Abundance dynamics, abbreviated
@@
 evolution that have been employed in theoretical quantitative genetics.
-Noting that a natural stochastic generalization of DAGA has been
-previously derived, we find a path to derive SDE describing the
-evolutionary response to demographic stochasticity. We accomplish this
-in \S\ref{wnc} by discussing how a diffusion limit of a spatially
+In \S\ref{wnc} we discuss how a diffusion limit of a spatially
 structured branching process (i.e., a measure-valued branching process)
-leads to the natural SPDE appropriate for our study. We refer to this
-SPDE as the Stochastic Asexual Gaussian allelic model with Abundance
-dynamics (abbreviated SAGA). The diffusion limit in turn provides a
-rigorous method for constructing fitness functions used in models of
-evolutionary ecology. However, before we discuss measure-valued braching
-processes and their diffusion limits, we find it necessary to review
-some fundamental results on the calculus of white noise and introducing
-a set of mathematical tools based on these results. In \S\ref{wnc} and
-SM \S\ref{SDE_DERIV} we employ these tools to derive a system of SDE
+leads to a SPDE that provides a natural stochastic generalization of
+DAGA. We refer to this SPDE as the Stochastic Asexual Gaussian allelic
+model with Abundance dynamics (abbreviated SAGA). The diffusion limit in
+turn provides a rigorous method for constructing fitness functions used
+in models of evolutionary ecology. However, before we discuss
+measure-valued braching processes and their diffusion limits, we find it
+necessary to review some fundamental results on the calculus of white
+noise and introduce a set of mathematical tools based on these results.
+In SM \S\ref{SDE_DERIV} we employ these tools to derive a system of SDE
 generalizing our deterministic results to account for demographic
-stochasticity. We find that the effect of demographic stochasticity on
-the evolution of mean trait and phenotypic variance characterizes the
+stochasticity. We find the effect of demographic stochasticity on the
+evolution of mean trait and phenotypic variance characterizes the
 process of random genetic drift in continuous time. However, although
@@
 normally distributed trait values to simplify these expressions into
-formulae that are much easier to work with. We then account for the
-constraint of adaptive evolution on the amount of heritable variation in
-a population by extending these results via a model of imperfect
+tractable formula that express the evolutionary response to selection in
+terms of fitness gradients. We then account for the constraint of
+adaptive evolution on the amount of heritable variation available in a
+population by extending these results via a model of imperfect
 inheritance. The resulting equations coincide with classical results in
@@
 participating in exploitative competition along a common resource
-continuum. In SM \S\ref{diffuse} we apply a classical theorem on
-rescaled limits of BBM that allow for ecological interactions to provide
-a rigorous derivation. To gain biological insight, in \S\ref{dynamics}
-we numerically integrate our model of coevolution for \(S=100\) species,
-tracking the dynamics of traits and abundances under scenarios of weak
-and strong competition. We include an account of the natural history of
-the simulated system and discuss the significance of demographic
-stochasticity for structuring ecological communities. In \S\ref{ecoevo}
-we provide expressions for selection gradients and competition
-coefficients implied by our model and use these expressions to
-investigate the relationship between the degree of competition and
+continuum. In SM \S\ref{diffuse} we apply a result on rescaled limits of
+BBM to provide a rigorous derivation. In \S\ref{dynamics} we illustrate
+our model of coevolution through numerical integration for \(S=100\)
+species, tracking the dynamics of traits and abundances under scenarios
+of weak and strong competition. We include an account of the natural
+history of the simulated system and discuss the significance of
+demographic stochasticity for structuring ecological communities. In
+\S\ref{ecoevo} we provide expressions for selection gradients and
+competition coefficients implied by our model and use these expressions
+to investigate the relationship between the degree of competition and
 strength of coevolution. Finally, \S\ref{conclusion} concludes with a
@@
 the theory of adaptive dynamics. As demonstrated by Page and Nowak
-(2002) and Champagnat, Ferrière, and Méléard (2006), the canonical
+(2002) and Champagnat, Ferrière and Méléard (2006), the canonical
 equation of adaptive dynamics can be derived from the replicator-mutator
@@
 equation and trait dynamics from abundance dynamics and extend these
-formulae along with related results to the stochastic case. The results
+formula along with related results to the stochastic case. The results
 established in this section provide the framework from which larger
@@
 =\frac{1}{N}\left(m_i\nu_i-\bar m Np_i\right)=(m_i-\bar m)p_i.
-\end{equation} Two important observations of these equations include 1)
+\end{equation} Two important observations of these equations include (i)
 mean fitness \(\bar m\) determines the abundance dynamics of the entire
-population and 2) allele \(i\) will increase (decrease) in frequency if
-\(m_i>\bar m\) (\(<\bar m\)). Equation (\ref{popgen}) is known in the
+population and (ii) allele \(i\) will increase (decrease) in frequency
+if \(m_i>\bar m\) (\(<\bar m\)). Equation (\ref{popgen}) is known in the
 field of evolutionary game theory as the replicator equation (Hofbauer
@@
 some \(r\in\mathbb{R}\) for all \(x\in\mathbb{R}\) and \(h\geq0\). We
-denote by \(K\) an operator that takes functions such as \(\nu(x,t)\) as
-its argument and returns other functions that may still depend on the
-spatial variable \(x\) and the temporal variable \(t\). For our
-application we will be focused on operators \(K\) that can be written as
+also assume \((1+h)C\leq m(h,x)\) for some constant \(C>0\) so that
+\(m(h,x)\) is linear or more than linear in \(h\). We denote by \(K\) an
+operator that takes functions such as \(\nu(x,t)\) as its argument and
+returns other functions that may still depend on the spatial variable
+\(x\) and the temporal variable \(t\). For our application we will be
+focused on operators \(K\) that can be written as
 \((K\nu)(x,t)=\int_{-\infty}^{+\infty}\kappa(x-y)\nu(y,t)dy\) for some
@@
 \end{equation} with the initial condition \(\nu(x,0)\) described above.
-This PDE is semilinear due to the dependency of the growth rate
-\(m(\nu,x)\) on the solution \(\nu(x,t)\) and is referred to as a scalar
-reaction-diffusion equation (Zheng 2004; Evans 2010). To ensure
+Equation (\ref{eq1}) is a semilinear parabolic PDE and is referred to as
+a scalar reaction-diffusion equation (Zheng 2004; Evans 2010). To ensure
 existence and uniqueness of solutions to (\ref{eq1}), we further assume
@@
 \begin{equation}\label{local_lipschitz}
-\int_\mathbb{R}\left|m(\nu_1,x)\nu_1(x)-m(\nu_2,x)\nu_2(x)\right|dx\leq L_M\int_\mathbb{R}\nu_1(x)-\nu_2(x)dx.
+\int_\mathbb{R}\left|m(\nu_1,x)\nu_1(x)-m(\nu_2,x)\nu_2(x)\right|dx\leq L_M\int_\mathbb{R}|\nu_1(x)-\nu_2(x)|dx.
 \end{equation} In SM \S\ref{finite} we combine the above assumptions on
@@
 mutation will cause \(\nu(x,t)\) to converge to a flat line as
-\(t\to\infty\). Although clearly an idealized representation of
-biological reality, this model is sufficiently general to capture a
-large class of dynamics including density dependent growth and frequency
-dependent selection. As an example, logistic growth combined with
-quadratic stabilizing selection can be captured using the growth rate
+\(t\to\infty\).
+
+Although clearly an idealized representation of biological reality, this
+model is sufficiently general to capture a large class of dynamics
+including density dependent growth and frequency dependent selection. As
+an example, logistic growth combined with quadratic stabilizing
+selection can be captured using the growth rate
 \begin{equation}\label{special_m}
 m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_{-\infty}^{+\infty}\nu(y,t)dy
-\end{equation} where \(r\in\mathbb{R}\) is the intrinsic growth rate in
-the absence of abiotic stabilizing selection, \(a>0\) is the strength of
-abiotic stabilizing selection around the phenotypic optimum
+\end{equation} where \(r\in\mathbb{R}\) is the intrinsic rate of growth
+in the absence of abiotic stabilizing selection, \(a>0\) is the strength
+of abiotic stabilizing selection around the phenotypic optimum
 \(\theta\in\mathbb{R}\), \(c>0\) captures the sensitivity of fitness to
-intraspecific competition and we have set \(\kappa(x-y)=1\) so that
-competitive interactions cause the same reduction in fitness regardless
-of trait value. This exemplary fitness function has a few convenient
-properties. First, the effect of competition induces a local carrying
-capacity on the population, leading to a finite equilibrium abundance
-over bounded subsets of \(\mathbb{R}\). Second, abiotic selection
-prevents the abundance density from diffusing too far from the abiotic
-optimum, leading to a finite equilibrial abundance across all of
-\(\mathbb{R}\),
+intraspecific competition and \(\kappa(x-y)=1\) so that competitive
+interactions cause the same reduction in fitness regardless of trait
+value. This exemplary fitness function has a few convenient properties.
+First, the effect of competition induces a local carrying capacity on
+the population, leading to a finite equilibrium abundance over bounded
+subsets of \(\mathbb{R}\). Second, abiotic selection prevents the
+abundance density from diffusing too far from the abiotic optimum,
+leading to a finite equilibrial abundance across all of \(\mathbb{R}\),
 \(\lim_{t\to\infty}N(t)=\tfrac{1}{c}(r-\tfrac{1}{2}\sqrt{a\mu})\),
@@
 and that predicted by the classic Gaussian allelic model of quantitative
-genetics (Lande 1975; Bürger 2000; Johnson and Barton 2005). By
-replacing \(\kappa(x-y)=1\) in expression (\ref{special_m}) with a
-particular Gaussian function, we will find in \S\ref{coev} and SM
-\S\ref{diffuse} this fitness function becomes equivalent to a fitness
-function derived from niche theory that we use to model coevolution
-driven by resource competition.
+genetics (Lande 1975; Bürger 2000; Johnson and Barton 2005). In
+\S\ref{coev} and SM \S\ref{diffuse} we find this fitness function is
+closely related to a fitness function derived from niche theory that we
+use to model coevolution driven by resource competition.
 
@@
 =\frac{1}{N(t)}\Big(m(\nu,x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)-\bar m(t)\nu(x,t)\Big) \\
-=(m(\nu,x)-\bar m(t)) p(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t).
+=\Big(m(\nu,x)-\bar m(t)\Big) p(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t).
 \end{multline} This result closely resembles Kimura's
@@
 mutational effects at loci in a genome on the values of traits (Lande
-1975; but see also Bürger 2000; and Johnson and Barton 2005), the
+1975; but see also Bürger 2000; along with Johnson and Barton 2005), the
 infinitesimal model in reference to modelling continuous traits as being
@@
 simple model of inheritance that leads us to an expression for the
-stochastic evolution of additive genetic variance. To clarify the
-distinction between this model and previous models of phenotypic
-evolution we refer to the PDE (\ref{eq1}) from which (\ref{p_PDE}) was
-derived as the Deterministic Asexual Gaussian allelic model with
-Abundance dynamics (abbreviated DAGA). Later, we will extend this model
-to include the effects of demographic stochasticity, which we refer to
-as the Stochastic Asexual Gaussian allelic model with Abundance dynamics
-(abbreviated SAGA).
+stochastic evolution of additive genetic variance. To distinguish this
+model from previous models of phenotypic evolution we refer to PDE
+(\ref{eq1}) from which (\ref{p_PDE}) was derived as the Deterministic
+Asexual Gaussian allelic model with Abundance dynamics (abbreviated
+DAGA). Later, we will extend this model to include the effects of
+demographic stochasticity, which we refer to as the Stochastic Asexual
+Gaussian allelic model with Abundance dynamics (abbreviated SAGA).
 
@@
 that a rigorous derivation of a SPDE that naturally generalizes DAGA and
-accounts for ecological interactions has been provided by Li (1998). To
+accounts for ecological interactions has been provided by Li (1998) and
+slightly generalized by Champagnat, Ferriére and Méléard (2006). To
 understand how this SPDE was derived, we provide an informal discussion
@@
 a sequence of rescaled BBM converges to a limiting process that can be
-characterized by a SPDE (Méléard and Roelly 1993; Li 1998). The limiting
-processes of rescaled BBM have been referred to as measure-valued Markov
-processes (Dawson 1993) or superprocesses (Etheridge 2000). Under the
-simplifying assumptions inherited from our treatment of deterministic
-dynamics and the additional assumption that the magnitude of demographic
+characterized by a SPDE (Méléard and Roelly 1992, 1993; Li 1998;
+Champagnat, Ferrière, and Méléard 2006). The limiting processes of
+rescaled BBM have been referred to as measure-valued Markov processes
+(Dawson 1993) or superprocesses (Etheridge 2000). Under the simplifying
+assumptions inherited from our treatment of deterministic dynamics and
+the additional assumption that the magnitude of demographic
 stochasticity is independent of trait values, we obtain as a special
@@
 for the effects of random genetic drift on the evolution of mean traits
-are obtained as a special case.
+are obtained as a further special case.
 
@@
 this leads to the conceptualization of white noise as a collection of
-Gaussian distributed random variables. Figure \ref{wn} illustrates
-realizations of this conceptualized white noise in one (left) and two
-(right) dimensions.
+independent Gaussian distributed random variables. Figure \ref{wn}
+illustrates realizations of this conceptualized white noise in one
+(left) and two (right) dimensions.
 
@@
 capitolizing on the solid ground these authors have established. For
-instance, instead of rigorously proving properties of white-noise, we
+instance, instead of rigorously proving properties of white noise, we
 simply define them to be so, taking solice in the fact that the
@@
 \(t>0\) and \(D\) a bounded subset of \(\mathbb{R}\), is equivalent to
-evaluating \(\mathbf W_t(I_{D\times [0,+\infty)})\) for the
-deterministic process \begin{equation}
-I_{D\times [0,+\infty)}(x,t)=\left\{\begin{matrix}
+evaluating \(\mathbf W_t(I_D)\) for the deterministic process
+\begin{equation}
+I_D(x,t)=\left\{\begin{matrix}
 0, & x\notin D \\
@@
 \end{equation} Since \begin{equation}
-\|I_{D\times[0,+\infty))}\|_t^2=\mathbb{E}\left(\int_0^t\int_\mathbb{R}I_{D\times [0,+\infty)}^2(x,s)dxds\right)=t\int_Ddx=t|D|<+\infty, 
+\|I_D\|_t^2=\mathbb{E}\left(\int_0^t\int_\mathbb{R}I_D^2(x,s)dxds\right)=t\int_Ddx=t|D|<+\infty, 
 \end{equation} where \(|D|\) denotes the length of \(D\), we have
-\(I_{D\times [0,+\infty)}\in\mathscr{N}_2\). Thus, using equations
-(\ref{exp_WN}) and (\ref{cov_WN}) and adopting the informal notation
-introduced in equation (\ref{informal}), we can write the following
-\begin{equation}
+\(I_D\in\mathscr{N}_2\). Thus, using equations (\ref{exp_WN}) and
+(\ref{cov_WN}) and adopting the informal notation introduced in equation
+(\ref{informal}), we can write the following \begin{equation}
 \mathbb{E}\left(\int_0^t\int_D\dot W(x,s)dxds\right)=0,
@@
 =\int_0^{t_1\wedge t_2}\int_\mathbb{R}f(x,s)g(x,s)dxds.
-\end{equation} To relate these formulae to the common notation used for
+\end{equation} To relate these formula to the common notation used for
 SDE, we write \begin{equation}
@@
 we describe the BBM process, which is a particularly important branching
-process with spatial stucture. This process has been very useful in the
+process with spatial structure. This process has been very useful in the
 study of SPDE due to its simplifying assumption that individuals do not
@@
   iteroparity. However, under the appropriate rescaling, both approaches
-  converge to the same diffusion limit. We therefore choose the former
-  approach for the sake of simplicity.
+  converge to the same diffusion limit (Champagnat, Ferrière, and
+  Méléard 2006). We therefore choose the former approach for the sake of
+  simplicity.
 \item
-  \textbf{Reproductive law:} When a birth event occurs a random
-  (possibly zero) number of offspring are left. The distribution of
-  offspring left is called the reproductive law or branching mechanism.
-  We denote the mean and variance in reproductive output by
-  \(\mathscr{W}\) and \(V\) respectively. The case of \(\mathscr{W}=1\)
-  is referred to as the critical condition. Under the critical condition
-  the probability that extinction occurs in finite time is equal to one.
+  \textbf{Reproductive law:} When a birth event occurs we assume a
+  random (possibly zero) number of offspring are produced. The
+  distribution of offspring left is called the reproductive law or
+  branching mechanism. We denote the mean and variance in reproductive
+  output by \(\mathscr{W}\) and \(V\) respectively. The case of
+  \(\mathscr{W}=1\) is referred to as the critical condition. Since
+  \(X_t(\mathbb{R}^d)\) is a Galton-Watson branching process, it is well
+  known the critical condition implies extinction in finite time with
+  probability one, given finite \(X_0(\mathbb{R}^d)\) (Athreya and Ney
+  1972).
 \item
@@
 \(x\) or \(t\), the derivatives in expression (\ref{neutral_SPDE}) are
-taken in the \emph{weak} sense (for examples, see Evans 2010). That is,
-to rigorously interpret SPDE (\ref{neutral_SPDE}), we integrate the
-solution \(\nu(x,t)\) against functions \(f\in C_b^2(\mathbb{R})\),
-where \(C_b^2(\mathbb{R})\) is the set of bounded and twice continuously
+taken in the \emph{weak} sense (for a definition see Evans 2010). That
+is, to rigorously interpret SPDE (\ref{neutral_SPDE}), we integrate the
+solution \(\nu(x,t)\) against functions \(f\in C_b^2(\mathbb{R})\) where
+\(C_b^2(\mathbb{R})\) is the set of bounded and twice continuously
 differentiable functions on \(\mathbb{R}\). Hence, equation
@@
 and thus precludes its ability to model nonlocal effects on the fitness
-of individuals, such as those produced via ecological interactions.
+of individuals, such as those produced via competitive interactions.
 However, this concern has been addressed, leading to constuctions of
@@
 uniqueness. Li (1998) built directly off of the construction of Méléard
-and Roelly (1992, 1993) to study properties of density processes
-associated with interacting superprocesses, arriving at a SPDE that
-forms the foundation of our approach.
-
-Recall, we use \(\nu(x,t)\) to denote the density of a superprocess,
-given it exists. Assuming the interactions manifest only in the
-reproductive law and that spatial movement follows Brownian motion with
-diffusion parameter \(\sqrt\mu\) independent of both \(X_t\) and \(x\),
-Li (1998) proved a result that implies the interacting superprocess on
-one dimensional trait space has a density \(\nu(x,t)\) which is
-non-negative, integrable, continuous in time and space and satisfies the
-SPDE \begin{equation}\label{SPDE}
+and Roelly (1992, 1993) to study properties of interacting
+superprocesses and, by assuming individual spatial movement occurs
+independently of location \(x\) and the entire population \(X_t\),
+showed the evolution of associated density processes can be described by
+SPDE.
+
+Recall \(\nu(x,t)\) denotes the density of a superprocess, given it
+exists. Assuming the interactions manifest only in the reproductive law
+and that spatial movement follows Brownian motion with diffusion
+parameter \(\sqrt\mu\) independent of both \(X_t\) and \(x\) and the
+growth rate is bounded both above and below, Li's (1998) result implies
+the interacting superprocess on one dimensional trait space has a
+density \(\nu(x,t)\) which is non-negative, integrable, continuous in
+time and space and satisfies the SPDE \begin{equation}\label{SPDE}
 \frac{\partial}{\partial t}\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)+\sqrt{V\nu(x,t)}\dot W(x,t).
@@
 Li (1998), our \(m\) and \(V\) correspond to Li's \(b\) and \(c\)
-respectively. It is important to note that, under the assumptions made
-in Méléard and Roelly (1992, 1993) and Li (1998), equation (\ref{SPDE})
-is only formal when \(m(h,x)\) is bounded across all combinations of
-\(h\geq0\) and \(x\in\mathbb{R}\). However, recalling our condition
-\(m(h,x)\leq r\in\mathbb{R}\), the growth rates we consider are only
-bounded above. Yet, in the proof of the construction of the interacting
-superprocess as the limit of rescaled branching diffusions, Méléard and
-Roelly (1992, 1993) assumed \(m(h,x)\) to be bounded to guarantee the
-total mass process will have finite mean and variance, for finite
-\(t\geq0\). This allowed the authors to employ a tightness criterion for
-sequences of measures and show the rescaled processes converge to a
-superprocess with finite total mass. Li's (1998) result builds directly
-off of Méléard and Roelly's construction, inheriting the assumption of
-boundedness for \(m(h,x)\). However, in Li (1998), the sufficiency of
-\(m(h,x)\) being bounded above is even more clear since Li works
-explicitly with a common upperbound for both \(m(h,x)\) and \(V\).
-Hence, one can repeat the necessary proofs replacing the assumption that
-\(m(h,x)\) is bounded with the assumption that \(m(h,x)\) is merely
-bounded above to derive the same results. An alternative approach to
-establishing this result that allows for our relaxed assumption on the
-growth rate has carried out by Champagnat, Ferrière, and Méléard (2006)
-using techniques developed by Evans and Perkins (1994).
-
-With solutions to SPDE (\ref{SPDE}) well defined for any growth rate
-bounded above, we can calculate the total mass process \(N(t)\) using
-the mild solution of (\ref{SPDE}) with
+respectively. Using techniques developed by Evans and Perkins (1994),
+Champagnat, Ferrière and Méléard (2006) generalized this result to the
+case when \(m(h,x)\) is bounded above and is linear or more than linear
+in \(h\). In particular, SAGA corresponds to a special case of equation
+(4.6) in Champagnat, Ferrière and Méléard (2006) and is valid for the
+assumptions on \(m\) made in \S\ref{deterministic}.
+
+With solutions to SAGA well defined, we can calculate the total mass
+process \(N(t)\) using the mild solution of (\ref{SPDE}) with
 \(f(x)\equiv1\in C_b^2(\mathbb{R})\) (the symbol ``\(\equiv\)'' means
@@
 we perform in SM \S\ref{SDE_DERIV}. We make no attempt in finding
-sufficient conditions for
-\(\int_\mathbb{R}(|x|+x^2+x^4)\nu(x,t)dx<+\infty\) to hold and hence
-make no general assertions about the existence or uniqueness of
-\(\bar x(t)\) or \(\sigma^2(t)\). Regardless, we will later assume
-\(\nu(x,t)\) can be approximated by a Gaussian curve in \(x\) for all
-\(t\geq0\). This assumption implies
-\(\int_\mathbb{R}|x|^n\nu(x,t)dx<+\infty\) for all \(n\in\mathbb{N}\)
-and for all \(t\geq0\).
+sufficient conditions to ensure
+\(\int_\mathbb{R}(|x|+x^2+x^4)\nu(x,t)dx<+\infty\) and hence make no
+general assertions about the existence or uniqueness of \(\bar x(t)\) or
+\(\sigma^2(t)\). Regardless, we will later assume \(\nu(x,t)\) can be
+approximated by a Gaussian curve in \(x\) for all \(t\geq0\). This
+assumption implies \(\int_\mathbb{R}|x|^n\nu(x,t)dx<+\infty\) for all
+\(n\in\mathbb{N}\) and for all \(t\geq0\).
 
@@
 stochasticity on abundance grows with \(\sqrt{N(t)}\). Hence, dividing
-by \(N\), we find the effects of demographic stochasticity on the
+by \(N(t)\), we find the effects of demographic stochasticity on the
 per-capita growth rate diminish with increased abundance. Relating the
@@
 These expressions can be used to investigate the dynamics of the mean
-and variance for general \(\nu(x,t)\). However, in the next subsection
-we simplify these expressions by approximating \(\nu(x,t)\) with a
-Gaussian curve. By assuming \(\nu(x,t)\) is Gaussian for \(t\geq0\), we
-guarantee the existence of \(\bar x(t)\) and \(\sigma^2(t)\) for all
-\(t\geq0\). Furthermore, in SM \S\ref{SDE_DERIV} we show that under the
-Gaussian case \(W_1,W_2\) and \(W_3\) are independent.
+and variance for general \(\nu(x,t)\), assuming
+\(\int_\mathbb{R}(|x|+x^2+x^4)\nu(x,t)dx<+\infty\) for all \(t\geq0\).
+However, in the next subsection we simplify these expressions by
+approximating \(\nu(x,t)\) with a Gaussian curve. By assuming
+\(\nu(x,t)\) is Gaussian for \(t\geq0\), we guarantee the existence of
+\(\bar x(t)\) and \(\sigma^2(t)\) for all \(t\geq0\). Furthermore, in SM
+\S\ref{SDE_DERIV} we show that under the Gaussian case \(W_1,W_2\) and
+\(W_3\) are independent.
 
@@
 
-By assuming \(\nu(x,t)\) can be approximated by a Gaussian curve for
-each \(t\geq0\), expressions (\ref{N}), (\ref{xbar_gen}) and
-(\ref{sig2_gen}) transform into efficient tools for deriving the
-dynamics of populations given a fitness function \(m(\nu,x)\). Gaussian
-phenotypic distributions are often obtained through Gaussian,
-exponential or weak selection approximations together with a simplified
-model of inheritance and random mating (Lande 1980; Turelli 1984, 1986,
-2017; Bürger 2000). Alternatively, it has been shown that a Gaussian
-distribution can provide a reasonable approximation even when selection
-is strong and non-Gaussian (Turelli and Barton 1994). However, our
-approach adds an additional layer of difficulty. Even with Gaussian
-selection, the resulting solution to SPDE (\ref{SPDE}) will only be a
-Gaussian curve in expectation, assuming a Gaussian curve as the initial
-condition. Yet this difficulty is not as challenging as it may first
-appear. Indeed, since SPDE (\ref{SPDE}) can be derived as a diffusion
-limit we know that, under the appropriate assumptions on selection,
-genetic architecture and reproduction, the stochastic departure from a
-Gaussian curve is negligible when the ratio \(V/N\) is small (i.e., when
-the variance in reproductive output is much smaller than the population
-size). In SM \S\ref{numerical} we demonstrate this result using
-numerical methods. Mathematically, this requirement restricts model
-parameters to regions that maintain large population sizes.
-Biologically, this implies populations are not at risk of extinction.
-Hence, models developed in this framework are not suitable for studying
-colonization-extinction dynamics or evolutionary rescue. Allowing for
-these restrictions, we may safely assume that \(\nu\) is approximately
-Gaussian and justify writing \begin{equation}
+By approximating \(\nu(x,t)\) with a Gaussian curve in \(x\) for each
+\(t\geq0\), expressions (\ref{gen_eqns}) transform into efficient tools
+for deriving the dynamics of populations given a fitness function
+\(m(\nu,x)\). Gaussian phenotypic distributions are often obtained
+through Gaussian, exponential or weak selection approximations together
+with a simplified model of mutation and asexual reproduction or random
+mating (Lande 1980; Turelli 1984, 1986, 2017; Bürger 2000).
+Alternatively, it has been shown that a Gaussian distribution can
+provide a reasonable approximation even when selection is strong and
+non-Gaussian (Turelli and Barton 1994). Hence, given appropriate
+assumptions on selection, mutation and reproduction, the abundance
+density \(\nu(x,t)\) can be approximated as a Gaussian curve in \(x\)
+when the ratio \(V/N\) is small (i.e., when the variance in reproductive
+output is much smaller than the population size). As with any diffusion
+approximation, this requires a sufficiently large abundance to
+accurately reflect the dynamics of populations. Hence, models developed
+in this framework are not suitable for studies involving very small
+population sizes. Allowing for these restrictions, we assume
+\begin{equation}
 \nu(x,t)=\frac{N(t)}{\sqrt{2\pi\sigma^2(t)}}\exp\left(-\frac{\big(x-\bar x(t)\big)^2}{2\sigma^2(t)}\right).
@@
 \mathrm{Cov}\Big((x-\bar x)^2,m\Big)=2\sigma^4\left(\frac{\partial\bar m}{\partial\sigma^2}-\overline{\frac{\partial m}{\partial\sigma^2}}\right)
-\end{equation} and \(\overline{(x-\bar x)^4}=3\sigma^4\). Equation
-(\ref{covxm}) is the continuous time equivalent to equation (9) in Lande
-(1976). In particular, these results imply
+\end{equation} and \(\overline{(x-\bar x)^4}=3\sigma^4\). In particular,
+these results imply
 
@@
 selection. This relationship has already been pointed out by Lande
-(1976) for the evolution of the mean trait, but here we see an analogous
-relationship holds also for the evolution of trait variance.
+(1976) for the evolution of the mean trait in discrete time, but here we
+see an analogous relationship holds in continuous time and also for the
+evolution of trait variance.
 
@@
 Stochastic Asexual Gaussian allelic models with Abundance dynamics
-(abbreviated SAGA). By replacing \(\psi(x,g)\) with different
-probability densities produces different developmental models. This
-includes densities that may depend on environmental variables and hence
-model phenotypic plasticy. To leave room for the naming of other models
-where \(\psi(x,g)\) is Gaussian, we refer to the model of development we
-have chosen as the \emph{centered} Gaussian developmental model. In
+(abbreviated SAGA). Replacing \(\psi(x,g)\) with different probability
+densities produces different developmental models. This includes
+densities that may depend on environmental variables and hence model
+phenotypic plasticy. To leave room for the naming of other models where
+\(\psi(x,g)\) is Gaussian, we refer to the model of development we have
+chosen as the \emph{centered} Gaussian developmental model. In
 \S\ref{wnc} we review the origins of equation (\ref{rho_SPDE}) and
@@
 For species \(i\) we inherit the above notation for trait value,
-distribution, average, variance, abundance, etc except with an \(i\) in
-the subscript. Real world examples of niche axes include the body size
-of prey for lizard predators and the date of activity in a season for
-pollinators competing for floral resources. For mathematical
-convenience, we model the axis of resources by the real line
-\(\mathbb{R}\). The value of a resouce along this axis is denoted by the
-symbol \(\zeta\). For an individual in species \(i\), we assume the
-resource utilization curve \(u_i\) can be written as
+distribution, average, variance, abundance, et cetera, except with an
+\(i\) in the subscript. Real world examples of niche axes include the
+size of seeds consumed competing finch species and the date of activity
+in a season for pollinators competing for floral resources. For
+mathematical convenience, we model the axis of resources by the real
+line \(\mathbb{R}\). The value of a resouce along this axis is denoted
+by the symbol \(\zeta\). For an individual in species \(i\), we assume
+the resource utilization curve \(u_i\) can be written as
 
@@
 in the absence of competition, individuals leave on average \(Q_i\)
-offspring when concentrated at \(\theta_i\). We capture the rate by
-which the fitness falls as niche location \(\zeta\) leaves the optimum
-\(\theta_i\) by the parameter \(A_i\geq0\). Hence, abiotic stabilizing
-selection along the resource axis can be modelled by the curve
+offspring when concentrated at \(\theta_i\). We assume the expected
+reproductive output of an individual in species \(i\) decreases as niche
+location \(\zeta\) leaves the optimum \(\theta_i\) at the rate
+\(A_i>0\). Hence, abiotic stabilizing selection along the resource axis
+can be modelled by the curve
 
@@
 To map the degree of niche overlap to fitness, we assume competition
-between individuals with niche locations \(x_i\) and \(x_j\) additively
-decreases the Malthusian fitness for the individual in species \(i\) by
-\(c_i\mathcal{O}_{ij}(x_i-x_j)\) for some \(c_i\geq0\). We refer to
-\(c_i\) as the sensitivity to competition for species \(i\). The term
+between individuals with niche locations \(x_i\) and \(x_j\) decreases
+the expected reproductive output for the individual in species \(i\) at
+the rate \(c_i\mathcal{O}_{ij}(x_i-x_j)\) for some \(c_i>0\). We refer
+to \(c_i\) as the sensitivity to competition for species \(i\). The term
 \(c_i\mathcal{O}_{ij}(x_i-x_j)\) coincides with a special case of a term
@@
 \end{multline}
-
 \begin{multline}\label{x_comm}
@@
 \end{multline}
-
 \begin{multline}\label{G_comm}
@@
 referred to in the community ecology literature as character
-displacement. We also see that the fitness of species \(i\) drops
-quadratically with the difference between \(\bar x_i\) and the abiotic
-optimum \(\theta_i\). Hence, abiotic selection acts to pull \(\bar x_i\)
-towards \(\theta_i\). The response in mean trait \(\bar x_i\) to natural
-selection is proportional to the amount of heritable variation in the
-population, represented by the additive genetic variance \(G_i\).
-However, we have that \(G_i\) is itself a dynamic quantity. Under our
-model, abiotic stabilizing selection erodes away heritable variation at
-a rate that is independent of both \(N_i\) and \(\bar x_i\). The effect
-of competition on \(G_i\) is a bit more complicated. When
-\(b_{ij}(\bar x_i-\bar x_j)^2<1\), competition with species \(j\) acts
-as diversifying selection which tends to increase the amount of
-heritable variation. However, when \(b_{ij}(\bar x_i-\bar x_j)^2>1\),
-competition with species \(j\) acts as directional selection and reduces
-\(G_i\). In the following subsections we demonstrate the behavior of
-system (\ref{comm_dynamics}) by plotting numerical solutions and
-investigate implications for the relationship between the strength of
-ecological interactions and selection.
+displacement (Brown and Wilson 1956). We also see that the fitness of
+species \(i\) drops quadratically with the difference between
+\(\bar x_i\) and the abiotic optimum \(\theta_i\). Hence, abiotic
+selection acts to pull \(\bar x_i\) towards \(\theta_i\). The response
+in mean trait \(\bar x_i\) to natural selection is proportional to the
+amount of heritable variation in the population, represented by the
+additive genetic variance \(G_i\). However, we have that \(G_i\) is
+itself a dynamic quantity. Under our model, abiotic stabilizing
+selection erodes away heritable variation at a rate that is independent
+of both \(N_i\) and \(\bar x_i\). The effect of competition on \(G_i\)
+is a bit more complicated. When \(b_{ij}(\bar x_i-\bar x_j)^2<1\),
+competition with species \(j\) acts as diversifying selection which
+tends to increase the amount of heritable variation. However, when
+\(b_{ij}(\bar x_i-\bar x_j)^2>1\), competition with species \(j\) acts
+as directional selection and reduces \(G_i\). In the following
+subsections we demonstrate the behavior of system (\ref{comm_dynamics})
+by plotting numerical solutions and investigate implications for the
+relationship between the strength of ecological interactions and
+selection.
 
@@
 \(\alpha_{ij}\) becomes a dynamical quantity that can be written as
-
 \begin{equation}
 \alpha_{ij}(t)=\frac{c_i}{r_i(t)}\int_{-\infty}^{+\infty}\int_{-\infty}^{+\infty}p_i(x,t)p_j(y,t)\mathcal O_{ij}(x,y) dxdy =\frac{c_iU_iU_j}{r_i(t)}\sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right),
-\end{equation}
-
-where
-
-\begin{equation}
+\end{equation} where \begin{equation}
 r_i(t)=R_i-\frac{a_i}{2}\Big((\bar x_i(t)-\theta_i)^2+G_i(t)+\eta_i\Big).
-\end{equation}
-
-Hence, \(dN_i(t)\) can be expressed as
-
-\begin{equation}
+\end{equation} Hence, \(dN_i(t)\) can be expressed as \begin{equation}
 dN_i(t)=r_i(t)\left(1-\sum_{j=1}^S\alpha_{ij}(t)N_j(t)\right)N_i(t)dt+\sqrt{V_iN_i(t)}dW_1(t).
-\end{equation}
-
-Note that although \(r_i(t)\) is referred to in the coexistence
-literature as the intrinsic growth rate of the population, \(R_i\) is a
-deeper intrinsic quantity. For now we refer to \(R_i\) as the
-\emph{innate} growth rate. Previous work, to list just a few citations
-in this literature, has shown the importance of demographic
-stochasticity (Turelli 1980; Schreiber 2017) and evolutionary dynamics
-(Case and Taper 1986; Schreiber, Patel, and terHorst 2018) in
-determining coexistence of competing species. However, these studies
-consider evolution and demographic stochasticity separately. To our
-knowledge, there are no rigorous theoretical investigations of
-coexistence that account for evolutionary dynamics of quantitative
-characters and demographic stochasticity simultaneously and, in
-particular, no studies combining these processes that also include the
-evolution of additive genetic variance. In fact, only recently have
-rigorous theoretical investigations successfully combined these
-processes to understand persistence of a single population
-(Gomulkiewicz, Krone, and Remien 2017). Even so, these authors focused
-on a single species model with a finite number of types (i.e., a
-population genetic model). With the connection outlined in this section
-formally established, researchers may pursue a postmodern coexistence
-theory that naturally includes both the evolutionary dynamics of
-quantitative characters and the effects of demographic stochasticity in
-a simple synthetic framework.
+\end{equation} Note that although \(r_i(t)\) is referred to in the
+coexistence literature as the intrinsic growth rate of the population,
+\(R_i\) is a deeper intrinsic quantity. For now we refer to \(R_i\) as
+the \emph{innate} growth rate. Previous work has shown the importance of
+demographic stochasticity (Turelli 1980; Schreiber 2017) and
+evolutionary dynamics (Case and Taper 1986; Schreiber, Patel, and
+terHorst 2018) in determining coexistence of competing species. However,
+these studies consider evolution and demographic stochasticity
+separately. To our knowledge, there are no rigorous theoretical
+investigations of coexistence that account for evolutionary dynamics of
+quantitative characters and demographic stochasticity simultaneously
+and, in particular, no studies combining these processes that also
+include the evolution of additive genetic variance. With the connection
+outlined in this section formally established, researchers may pursue a
+postmodern coexistence theory that naturally includes both the
+evolutionary dynamics of quantitative characters and the effects of
+demographic stochasticity in a simple synthetic framework.
 
@@
 selection gradient (sensu Lande and Arnold 1983) induced by species
-\(j\) on species \(i\) can be computed as
-
-\begin{equation}
+\(j\) on species \(i\) can be computed as \begin{equation}
 \beta_{ij}(t)=c_iU_iU_jN_j(t)b_{ij}(t)\big(\bar x_i(t)-\bar x_j(t)\big) \sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right).
-\end{equation}
-
-Our notation differs from Lande and Arnold (1983) in that subscripts
-here denote species instead of components of multivariate traits and we
-drop the prime that distinguishes between selection gradients and
-standardized selection gradients.
+\end{equation} Our notation differs from Lande and Arnold (1983) in that
+subscripts here denote species instead of components of multivariate
+traits and we drop the prime that distinguishes between selection
+gradients and standardized selection gradients.
 
@@
 gradient induced by species \(j\) on species \(i\) can be computed as
-
 \begin{equation}
 \gamma_{ij}(t)=c_iU_iU_jN_j(t)b_{ij}(t)\left(1-b_{ij}(t)\big(\bar x_i(t)-\bar x_j(t)\big)^2\right) \sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right).
-\end{equation}
-
-To measure the total evolutionary change in species \(i\) induced by
-species \(j\), we form the metric
-\(\Psi_{ij}=|\beta_{ij}|+|\gamma_{ij}|\). Figure \ref{net} displays the
-joint distribution of \(\mathfrak{C}_{ij}=\Psi_{ij}\Psi_{ji}\), our
-metric of pairwise coevolution, and the product of competition
-coefficients \(\alpha_{ij}\alpha_{ji}\) for two simulated communities,
-both with richness \(S=1000\). The solid contour represents the case of
-relatively high sensitivity to competition (\(c=5\times10^{-7}\)) and
-the dashed contour represents the case of relatively weak sensitivity to
-competition (\(c=1\times10^{-8}\)). In both cases we see that
-\(\mathfrak{C}_{ij}\) and \(\alpha_{ij}\alpha_{ji}\) are, essentially,
-unrelated. However, this depiction is only representative of a specific
-set of parameters. Next, we provide analytical approximations of the
-covariance between selection gradients and competition coefficients and
-a numerical estimate for the relationship between pairwise coevolution
-and competition coefficients for a range of parameters.
+\end{equation} To measure evolutionary change in species \(i\) induced
+by species \(j\), we form the metric
+\(\Psi_{ij}=|\beta_{ij}|+|\gamma_{ij}|\) and define the strength of
+coevolution as \(\mathfrak{C}_{ij}=\Psi_{ij}\Psi_{ji}\). Figure
+\ref{net} displays the joint distribution of \(\mathfrak{C}_{ij}\) and
+the product of competition coefficients \(\alpha_{ij}\alpha_{ji}\) for
+two simulated communities, both with richness \(S=1000\). The solid
+contour represents the case of relatively high sensitivity to
+competition (\(c=5\times10^{-7}\)) and the dashed contour represents the
+case of relatively weak sensitivity to competition
+(\(c=1\times10^{-8}\)). In both cases we see that \(\mathfrak{C}_{ij}\)
+and \(\alpha_{ij}\alpha_{ji}\) are, essentially, unrelated. However,
+this depiction is only representative of a specific set of parameters.
+Next, we provide analytical approximations of the covariance between
+selection gradients and competition coefficients and a numerical
+estimate for the relationship between pairwise coevolution and
+competition coefficients for a range of parameters.
 
@@
 Nuismer, Week, and Aizen 2018). If \(\bar{\bar x}\) is near \(\theta\)
-and \(V_{\bar X}\) is much smaller than \(|2R/a-G-\eta|\), then we may
-approximate \(r_i\) with
-
-\begin{equation}
+and \(V_{\bar X}\) is much smaller than \(|2\frac{R}{a}-G-\eta|\), then
+we may approximate \(r_i\) with \begin{equation}
 \bar r=\int_{-\infty}^{+\infty}\left(R-\frac{a}{2}\Big((\bar x-\theta)^2+G+\eta\Big)\right)f_{\bar X}(\bar x)d\bar x
 =R-\frac{a}{2}\Big((\bar{\bar x}-\theta)^2+V_{\bar X}+G+\eta\Big).
-\end{equation}
-
-In SM \S\ref{coef_grad_moms} we use these assumptions to calculate the
-first and second order moments describing the joint distribution of
-competition coefficients and selection gradients across the community.
-We find that the covariance between linear selection gradients and
-competition coeffecients are zero due to the symmetry implied by our
-assumptions. However, setting \(\alpha(\bar x_i,\bar x_j)=\alpha_{ij}\),
+\end{equation} In SM \S\ref{coef_grad_moms} we use these assumptions to
+calculate the first and second order moments describing the joint
+distribution of competition coefficients and selection gradients across
+the community. We find that the covariance between linear selection
+gradients and competition coeffecients are zero due to the symmetry
+implied by our assumptions. However, setting
+\(\alpha(\bar x_i,\bar x_j)=\alpha_{ij}\),
 \(\beta(\bar x_i,\bar x_j)=\beta_{ij}\) and
@@
 
-\begin{subequations}
-
+\begin{subequations}\label{cov_eqns}
 \begin{equation}\label{cov_alpha_beta}
@@
 \end{equation}
-
 \begin{equation}\label{cov_alpha_gamma}
@@
 \end{equation}
-
 \end{subequations}
@@
 
-For fixed \(c,b,\bar r\) and \(N\), we visualize these relationships in
-Figure \ref{cov_fig}. To gain insight into the relationship between
-selection gradients and competition coefficients, note that our
-assumptions in this section imply \(b^{-1}=2(\sigma^2+w)\). If we
+We visualize these relationships in Figure \ref{cov_fig} for fixed
+\(c,b,\bar r,U\) and \(N\). In SM \S\ref{coef_grad_moms} we verify the
+qualitative nature of these relationships by repeating numerical
+integrations of system (\ref{comm_dynamics}) across a set of randomly
+drawn model parameters. We designed our numerical approach to inherit
+the assumptions of symmetric model parameters such as mutation rates
+\(\mu_1=\dots=\mu_S\) and abiotic optima \(\theta_1=\dots=\theta_S\) and
+to relax the key assumptions of normally distributed mean traits along
+with constant additive genetic variances and abundances across species.
+
+Our assumptions in this section imply \(b^{-1}=2(\sigma^2+w)\). If we
 further assume \(\sigma^2+w\approx\sigma^2\), then
@@
 \(2bV_{\bar X}\) is approximately equal to the ratio of interspecific
-mean trait variation to intraspecific individual trait variation. Hence,
-for both covariances we see that there is no relationship between
-selection gradients and competition coefficients when this ratio is
-zero. From equation (\ref{cov_alpha_beta}) we can use numerical
+mean trait variation to intraspecific individual trait variation.
+Glancing at Figure (\ref{cov_fig}) we see there is no relationship
+between selection gradients and competition coefficients when this ratio
+is zero. From equation (\ref{cov_alpha_beta}) we can use numerical
 optimization to find that when \(V_{\bar X}/\sigma^2\approx1.25\) the
 relationship between the magnitudes of linear selection gradients and
-competition coefficients disappears, but when (approximately)
-\(V_{\bar X}/\sigma^2<1.25\) (\(>1.25\)), this covariance becomes
-negative (positive). Equation (\ref{cov_alpha_gamma}) states that when
-\(V_{\bar X}/\sigma^2\) is approximately equal to one (or slightly
-larger), there is no expected relationship between competition
-coefficients and quadratic selection gradients. However, when
-\(V_{\bar X}/\sigma^2<1.0\) (\(>1.0\)), then we expect a positive
+competition coefficients disappears, but when
+\(V_{\bar X}/\sigma^2\lessapprox1.25\) (\(\gtrapprox1.25\)), this
+covariance becomes negative (positive). Equation (\ref{cov_alpha_gamma})
+states that when \(V_{\bar X}/\sigma^2\) is approximately equal to one
+(or slightly larger), there is no expected relationship between
+competition coefficients and quadratic selection gradients. However,
+when \(V_{\bar X}/\sigma^2<1.0\) (\(>1.0\)), then we expect a positive
 (negative) relationship between \(\alpha\) and \(\gamma\). These results
@@
 small, then species are packed tightly in phenotypic space. In our model
-this occurs when abiotic stabilizing selection is much stronger than
-competition (\(a\gg c\)). This causes species to overlap more in niche
-space (i.e., large \(\alpha\)) and creates disruptive selection for
-greater intraspecific variance (i.e., positive \(\gamma\)), which
-explains the positive region of covariance between \(\alpha\) and
-\(\gamma\). However, as species begin to overlap in niche space,
-directional selection begins to vanish (i.e., small \(|\beta|\)),
+this occurs when fitness is more sensitive to abiotic stabilizing
+selection than competition (\(a\gg c\)). This causes species to overlap
+more in niche space (i.e., large \(\alpha\)) and creates disruptive
+selection for greater intraspecific variance (i.e., positive
+\(\gamma\)), which explains the positive region of covariance between
+\(\alpha\) and \(\gamma\). However, as species begin to overlap in niche
+space, directional selection begins to vanish (i.e., small \(|\beta|\)),
 leading to a negative covariance between \(\alpha\) and \(|\beta|\). In
@@
 they will exhibit zero directional selection since a shift in either
-direction will yield the same fitness advantages.
-
-In the opposite scenario where competition is much stronger than abiotic
-stabilizing selection (\(c\gg a\)), species will not evolve to be as
-tightly packed and instead their niche-centers will be spread out with
-little ovelap in their resouce utilization curves (i.e., small
-\(\alpha\)). In this case biotic directional selection will be strong
-(i.e., large \(|\beta|\)), particularly for species towards the outer
-regions of niche space due to asymmetric fitness advantages conferred by
-shifts in niche-centers. This leads to a positive covariance between
-\(\alpha\) and \(|\beta|\). However, as noted above, this directional
-selection will also erode away at standing heritable variation (i.e.,
-negative \(\gamma\)), reducing the rate at which adaptation can occur
-and creating a negative covariance between \(\alpha\) and \(\gamma\).
-
-In summary, we see the relation between competition coefficients and
-selection is highly non-trivial and depends on the relative magnitudes
-of different ecological processes shaping the community. However, this
-does not address the relation between competition coefficients and
-coevolution per se. In SM \S\ref{coef_grad_moms} we show that
-calculating a formula for the covariance between competition
+direction will yield identical fitness advantages.
+
+In the opposite scenario where fitness is much more sensitive to
+competition than abiotic stabilizing selection (\(c\gg a\)), species
+will not evolve to be as tightly packed and instead their niche-centers
+will be spread out with little ovelap in their resouce utilization
+curves (i.e., small \(\alpha\)). In this case biotic directional
+selection will be strong (i.e., large \(|\beta|\)), particularly for
+species towards the outer regions of niche space due to asymmetric
+fitness advantages conferred by shifts in niche-centers. This leads to a
+positive covariance between \(\alpha\) and \(|\beta|\). However, as
+noted above, this directional selection will also erode away at standing
+heritable variation (i.e., negative \(\gamma\)), reducing the rate at
+which adaptation can occur and creating a negative covariance between
+\(\alpha\) and \(\gamma\).
+
+In summary, we see the relations between competition coefficients and
+selection gradients are non-trivial and depend on the relative
+magnitudes of different ecological processes shaping the community.
+However, this does not address the relation between competition
+coefficients and coevolution per se. In SM \S\ref{coef_grad_moms} we
+show that calculating a formula for the covariance between competition
 coefficients and the metric of coevolution \(\mathfrak{C}\) introduced
@@
 \(\mathrm{Cov}_{f_{\bar X}}(\alpha,\gamma)\) to approximate the
-correlation of \(\alpha\) and \(\mathfrak{C}\). This numerical approach
-inherits the assumptions of homogeneous background parameters such as
-the mutation rate \(\mu\) and abiotic optima \(\theta\), but allows us
-to relax the assumption that \(N\) and \(G\) are constant across species
-and time.
-
-In particular, we numerically integrated system (\ref{comm_dynamics})
-for \(T_1=1000.0\) units of time and then continued to integrate for
-\(T_2=1000.0\) units of time. We then calculate the covariance between
-the quantities \(\alpha\) and \(\mathfrak{C}\) across \(S=100\) species
-for each of the last \(T_2\) time steps. We assume the temporal average
-of these covariances across the last \(T_2\) units of time approximates
-the expectation at equilibrium. We repeated this approach for randomly
-drawn \(a\) and \(c\) until our sample size reached \(1000\). In Figure
-\ref{cor_coev_fig} we plot the temporally averaged values of
-\(\mathrm{Cov}_{f_{\bar X}}(\alpha,\mathfrak{C})\) against the
-sensitivity to competition \(c\). Using a cubic regression, we see the
-correlation of coevolutionary selection gradients and competition
-coefficeints is negative at variance ratios below \(0.5\), zero at
-variance ratios between \(0.5\) and \(1.0\), and drops below zero again
-at variance ratios above \(1.0\).
+correlation of \(\alpha\) and \(\mathfrak{C}\). In Figure
+\ref{cor_coev_fig} we plot values of
+\(\mathrm{Corr}_{f_{\bar X}}(\alpha,\mathfrak{C})\) against the exact
+ratio of interspecific trait variance to intraspecific trait variance.
+Using a cubic regression, we see the correlation of coevolutionary
+selection gradients and competition coefficeints is negative at variance
+ratios below \(0.5\), zero at variance ratios between \(0.5\) and
+\(1.0\), and drops below zero again at variance ratios above \(1.0\).
 
@@
 phenotypic evolution that accounts for demographic stochasticity. From
-SAGA we derived SDE that track the evolution of abundance, mean trait
-and additive genetic variance. Observing the expressions of these SDE we
+SAGA we derived SDE that track the dynamics of abundance, mean trait and
+additive genetic variance. Observing the expressions of these SDE we
 find the effects of demographic stochasticity on the evolution of mean
 trait and additive genetic variance characterize the effects of random
-genetic drift. By considering the statistics of random samples, this has
-previously been in models of fixed population size in discrete time
-(Lande 1976). However, we have shown random genetic drift is a result of
-demographic stochasticity for populations with fluctuating abundances in
-continuous time. To illustrate the relevance of our approach to studies
-of evolutionary ecology, we comined our SDE with classical competition
-theory to derive a model of diffuse coevolution and used this model to
-investigate the relationship between selection gradients and competition
-coefficients. This work demonstrates that connecting contemporary
-theoretical approaches to evolutionary ecology with some fundamental
-results in the theory of measure-valued stochastic processes allows for
-the development of a rigorous, yet flexible approach to synthesizing the
-dynamics of abundance and distributions of quantitative characters, in
-agreement with Champagnat, Ferrière, and Méléard (2006). Although this
-approach enjoys these important merits, there remains biological details
-and their associated technical challenges that need to be confronted for
+genetic drift. Although Lande (1976) has previously characterized the
+effects of random genetic drift on mean trait evolution in quantitative
+genetic models, the approach taken assumed constant effective population
+size and discrete non-overlapping generations. In contrast, our approach
+shows random genetic drift is a result of demographic stochasticity for
+continuously reproducing populations with fluctuating abundances. To
+illustrate the relevance of our approach to studies of evolutionary
+ecology, we comined our SDE with classical competition theory to derive
+a model of diffuse coevolution and used this model to investigate the
+relationship between standardized selection gradients and competition
+coefficients. We found the absolute value of linear selection gradients
+are negatively related with competition coefficients when the ratio of
+interspecific phenotypic variation to intraspecific phenotypic variation
+is approximately less than \(1.25\) and positive when this ratio is
+approximately greater than \(1.25\). We found the opposite relationship
+between quadratic selection gradients and competition coefficients with
+a threshold diversity ratio equal to unity instead of \(1.25\). This
+work demonstrates that connecting contemporary theoretical approaches to
+evolutionary ecology with some fundamental results in the theory of
+measure-valued stochastic processes allows for the development of a
+rigorous, yet flexible approach to synthesizing the dynamics of
+abundance and distributions of quantitative characters, in agreement
+with Champagnat, Ferrière and Méléard (2006). Although this approach
+enjoys these important merits, there remains biological details and
+their associated technical challenges that need to be confronted for
 gaining a more thorough and accurate understanding of ecological
@@
 
-\textbf{The genetic architecture and distributions of traits}
+\textbf{Genetic architecture and intraspecific trait distributions}
 
@@
 effects in the BBM model is Gaussian with variance an order of magnitude
-greater than the equilibrium phenotypic (or genetic) variance, we obtain
-the so-called \emph{House-of-cards} model (Turelli 1984). However,
-perhaps considering heavy-tailed distributions, such as the
-Laplace-distributions considered in Park et. al. (2017), will provide
-more realistic, yet less analytically tractable models of mutational
-effects. Such models have important implications in conservation
-genetics where small populations may evolve primarily in response to
-mutation and random genetic drift (Lande 1995).
+greater than the equilibrium phenotypic (or additive genetic) variance,
+we obtain the \emph{House-of-cards} model (Turelli 1984). However,
+perhaps considering heavy-tailed distributions such as the
+Laplace-distributions considered by Park et. al. (2017) will provide
+more realistic yet less analytically tractable models of mutational
+effects. Such mutation models have important implications in
+conservation genetics where small populations evolve primarily in
+response to mutation and random genetic drift (Lande 1995).
 
@@
 
-To understand patterns found in ecological communities, considerations
+To understand patterns found in ecological communities considerations
 must push beyond microevolutionary and contemporary ecological processes
@@
 repurpose microevolutionary models by blindly rescaling time from the
-units of generations to millions of years {[}Nuismer and Harmon (2014);
-Luke, can you think of others?{]}. Such an approach makes the implicit
-assumption that trait evolution is statistically self-similar (sensu
-Falconer 2014) so that the stochastic evolution of traits on
-macroevolutionary time scales has the same properties of trait evolution
-on microevolutionary time scales. Although some stochastic processes,
-including Brownian motion, do exhibit self-similarity, others do not.
-For example, consider a modification of the Ornstein-Uhlenbeck process
-defined by the SDE
+units of generations to millions of years (e.g., Nuismer and Harmon
+2014). Such an approach makes the implicit assumption that trait
+evolution is statistically self-similar (sensu Falconer 2014) so that
+the stochastic evolution of traits on macroevolutionary time scales has
+the same properties of trait evolution on microevolutionary time scales.
+Although some stochastic processes, including Brownian motion, do
+exhibit self-similarity, others do not. For example, consider a
+modification of the Ornstein-Uhlenbeck process defined by the SDE
 
@@
 environmental change (\(\lambda\)) versus the rate of evolutionary and
-ecological change (\(a\)) to microevolutionary models derived using our
+ecological change (\(a\)) to microevolutionary models derived via our
 framework can be used to obtain macroevolutionary models that include
@@
 
-\hypertarget{simulating-the-rescaled-process-and-numerical-evidence-of-approximate-normality}{%
-\subsection{\texorpdfstring{Simulating the rescaled process and
-numerical evidence of approximate normality
-\label{numerical}}{Simulating the rescaled process and numerical evidence of approximate normality }}\label{simulating-the-rescaled-process-and-numerical-evidence-of-approximate-normality}}
-
-Here we use a numerical argument to suggest, for
-
-\begin{equation}
-r-\frac{a}{2}(\theta-x)^2-c\int_\mathbb{R}\nu(x,t)dx \leq m(\nu,x)\leq r-\frac{a}{2}(\theta-x)^2, \ \forall (\nu,x)\in C_{1,c}^+(\mathbb{R})\times\mathbb{R},
-\end{equation}
-
-the density process \(\nu(x,t)\) defined by SPDE (\ref{SPDE}) of the
-main text satisfies \(\int_\mathbb{R}(|x|+x^2)\nu(x,t)dx<\infty\).
-
-and setting \(c<r-(\eta a+\sqrt{\mu a})/2\) ensures \(\hat N>1\), which
-is important for numerical simulations when \(N\) is an integer. We use
-these results to help guide our choice of parameter values for
-simulations of the branching random walk. In the following section we
-provide a detailed description of the branching random walk and how we
-have chosen to rescale it. We then use the rescaled branching random
-walk to investigate finiteness of moments and normality.
+\hypertarget{simulating-the-rescaled-process}{%
+\subsection{\texorpdfstring{Simulating the rescaled process
+\label{numerical}}{Simulating the rescaled process }}\label{simulating-the-rescaled-process}}
+
+Here we provide a detailed description of the branching random walk and
+how we have chosen to rescale it. We focus on the case \begin{equation}
+m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_\mathbb{R}\nu(x,t)dx.
+\end{equation}
 
@@
 steps for easier implementation. Hence, we restrict time to
-\(t=0,1,2,\dots\). Furthermore, we allow individual fitness to depend on
-both trait value and the state of the entire population. For time \(t\)
-we write \(\{\xi_1(t),\dots,\xi_{N(t)}(t)\}\) as the set of trait values
-across all \(N(t)\) individuals alive in the population. Since our
-simulation treats individuals instead of continuous distributions of
-trait values, we can write
-
-\begin{equation}
+\(t=0,1,2,\dots,\) and so on. Furthermore, we allow individual fitness
+to depend on both trait value and the state of the entire population.
+For time \(t\) we write \(\{\xi_1(t),\dots,\xi_{N(t)}(t)\}\) as the set
+of trait values across all \(N(t)\) individuals alive in the population.
+Since our simulation follows discrete individuals instead of continuous
+distributions of trait values, we can write \begin{equation}
 \nu(x,t)=\sum_{i=1}^{N(t)}\delta(x-\xi_i(t)),
-\end{equation}
-
-where \(\delta(x-\xi_i(t))\) denotes the point mass located at
-\(\xi_i(t)\). To allow for imperfect heritability, we also track the set
-of breeding values which, at time \(t\), is denoted by
-\(\{\gamma_1(t),\dots,\gamma_{N(t)}(t)\}\) and should not be confused
-with the quadratic selection gradients discussed in \S\ref{diffuse} of
-the main text. Then the distribution of breeding values can be written
-as
-
-\begin{equation}
-\rho(g,t)=\sum_{i=1}^{N(t)}\delta(g-\gamma_i(t)).
-\end{equation}
-
-Following our model of heritiability, the trait value \(\xi_i(t)\) is
-drawn from a normal distribution centered on \(\gamma_i(t)\) with
-variance \(\eta\). At each iteration we draw, for each individual, a
-random number of offspring from a Negative-Binomial distribution. Recall
-the Negative-Binomial distribution models the number of failed Bernoulli
+\end{equation} where \(\delta(x-\xi_i(t))\) denotes the point mass
+located at \(\xi_i(t)\). For simplicity we assume perfect heritability.
+At each iteration we draw, for each individual, a random number of
+offspring from a Negative-Binomial distribution. Recall the
+Negative-Binomial distribution models the number of failed Bernoulli
 trials that occur before a given number of successful trials. Denoting
@@
 successes, the mean and variance is given respectively by
-
 \begin{equation}
 \frac{s(1-q)}{q}, \ \frac{s(1-q)}{q^2}.
-\end{equation}
-
-Then if we require the \(i\)th individual to have mean number offspring
-\(\mathscr W(\nu,\xi_i)\) and variance equal to \(V\), the parameters of
-the associated Negative-Binomial distribution become
-
+\end{equation} Then if we require the \(i\)th individual to have mean
+number offspring \(\mathscr W(\nu,\xi_i)\) and variance equal to \(V\),
+the parameters of the associated Negative-Binomial distribution become
 \begin{equation}
 q(\nu,\xi_i)=\frac{\mathscr{W}(\nu,\xi_i)}{V}, \ s(\nu,\xi_i)=\frac{\mathscr{W}^2(\nu,\xi_i)}{V-\mathscr{W}(\nu,\xi_i)}.
-\end{equation}
-
-The imposes the restriction \(V>\mathscr{W}(\nu,\xi_i)\). For each
-offpsring produced by the individual with breeding value
-\(\gamma_i(t)\), we assign indepently drawn breeding values normally
-distributed around \(\gamma_i(t)\) with variance \(\mu\). After breeding
-values have been assigned, we randomly draw trait values for each
-offpsring as described above. For an overview of our model of
-inheritance, see \S\ref{inheritance} of the main text. This summarizes
-the basic structure of our simulation. To impose selection and density
-dependent growth rates, we set
-
+\end{equation} This imposes the restriction
+\(V>\mathscr{W}(\nu,\xi_i)\). For each offpsring produced by the
+individual with trait value \(\xi_i(t)\), we assign indepently drawn
+trait values normally distributed around \(\xi_i(t)\) with variance
+\(\mu\). This summarizes the basic structure of our simulation. To
+impose selection and density dependent growth rates, we set
 \begin{equation}
 \mathscr{W}(\nu,\xi_i)=\exp\left(r-\frac{a}{2}(\theta-\xi_i)^2-c\int_\mathbb{R}\nu(x,t)dx\right),
-\end{equation}
-
-where the above integral becomes
+\end{equation} where the above integral becomes
 \(\int_\mathbb{R}\nu(x,t)dx=\sum_{i=1}^{N(t)}1=N(t)\).
@@
 To rescale the branching random walk by a positive integer \(n\), we
-rescale segregation and mutational variance according to
-\(\eta\to \eta\) and \(\mu\to \mu/n\), time by \(t\to t/n\) and the
-reproductive law by \(V\to V\) and
-
-\begin{equation}
+rescale mutational variance by \(\mu\to \mu/n\), time by \(t\to t/n\)
+and the reproductive law by \(V\to V\) and \begin{equation}
 \mathscr W(\nu,\xi_i)\to\mathscr W^{(n)}(\nu,\xi_i)=\exp\left(\frac{r}{n}-\frac{a}{2n}(\theta-\xi_i)^2-\frac{c}{n^2}N(t)\right)=\exp\left(\frac{r}{n}-\frac{a}{2n}(\theta-\xi_i)^2-\frac{c}{n}N^{(n)}(t)\right).
-\end{equation}
-
-We also replace individual mass with \(\frac{1}{n}\) and write rescaled
-abundance as \(N^{(n)}(t)=\frac{1}{n}N(nt)\). Under this rescaling the
-deterministic equilibrium of the raw numerical abundance becomes
-
-\begin{equation}
-\hat N=\frac{n^2}{c}\left(\frac{r}{n}-\frac{1}{2n}(\eta a+\sqrt{\mu a})\right)=\frac{n}{c}\left(r-\frac{1}{2}(\eta a+\sqrt{\mu a})\right).
-\end{equation}
-
-The deterministic equilibrium of the rescaled abundance is then
-
-\begin{equation}
-\hat N^{(n)}=\frac{1}{c}\left(r-\frac{1}{2}(\eta a+\sqrt{\mu a})\right).
-\end{equation}
-
-When it exists, we denote by \(N^{(\infty)}(t)\) the limiting process of
-\(N^{(n)}(t)\). Then
-
-\begin{equation}
+\end{equation} We also replace individual mass with \(\frac{1}{n}\) and
+write rescaled abundance as \(N^{(n)}(t)=\frac{1}{n}N(nt)\). When it
+exists, we denote by \(N^{(\infty)}(t)\) the limiting process of
+\(N^{(n)}(t)\). Then \begin{equation}
 \lim_{n\to\infty}n\left(\mathscr W^{(n)}(\nu,\xi_i)-1\right)=r-\frac{a}{2}(\theta-\xi_i)^2-cN^{(\infty)}(t).
-\end{equation}
-
-Note that, following the notation of Theorem 1 in Méléard and Roelly
-(1992), setting \(\lambda_n=n\),
-\(m_n(\nu)=\mathscr W^{(n)}(\nu,\cdot)\) and \(\varepsilon_n=1/n\)
-satisfies their hypotheses \((\mathscr H_0)\)-\((\mathscr H_3)\) when
-\(c=0\). We have implemented this simulation in the programming language
-Julia. A copy can be found at the url:
-
-\texttt{\url{https://github.com/bobweek/branching.brownian.motion.and.spde}}
-
+\end{equation} Note that, since the limiting fitness function is bounded
+above and decreases linearly with respect to \(N^{(\infty)}(t)\), it
+satisfies the hypotheses of Champagnat, Ferrière and Méléard (2006). We
+have implemented this simulation in the programming language Julia. A
+copy can be found at the url:
+\[\texttt{\url{https://github.com/bobweek/branching.brownian.motion.and.spde}}\]
 For the sake of illustration, we simulated the unscaled process
 (\(n=1\)) and the rescaled process with \(n=5\) and \(n=20\) for \(50\)
-units of time. Results are shown in Figure \ref{rescaled}. In the
-following section we use a statistical test to show, for the lower bound
-on \(m(\nu,x)\), the rescaled process converges to a Gaussian density as
-\(n\to\infty\) and \(V/N\to0\).
-
-\hypertarget{evidence-of-normality}{%
-\subsubsection{\texorpdfstring{Evidence of normality
-\label{ev_normal}}{Evidence of normality }}\label{evidence-of-normality}}
-
-To demonstrate approximate normality of the phenotypic distribution when
-\(V/N\) is small we utilized the one-sided Kolmogorov-Smirnov test. This
-test compares an empirical cumulative distribution function (i.e., a
-cumulative distribution function generated from simulated data) to a
-hypothetical cumulative distribution function by providing a
-distribution for the maximum distance between these curves. More
-precisely if \(F_n(x)\) is the empricial distribution function for a
-sample of size \(n\) and \(F(x)\) is the hypothetical distribution
-function, Kolmogorov-Smirnov statistic is \(D_n=\sup_x|F_n(x)-F(x)|\).
+units of time. Results are shown in Figure \ref{rescaled} of the main
+text.
 
@@
 
-Hence, our fitness function is bounded above, as required in the main
-text.
-
-For the remainder of the derivation we suppress notation indicating
-dependency on \(\pmb\nu\), \(x\) and \(t\). From (\ref{malth_fit}) we
-calculate
+Hence, our fitness function is bounded above and decreases linearly with
+total abundance, as required by the assumptions of Champagnat, Ferrière
+and Méléard (2006). For the remainder of the derivation we suppress
+notation indicating dependency on \(\pmb\nu\), \(x\) and \(t\). From
+(\ref{malth_fit}) we calculate
 
@@
 
+To verify our analytical estimates for the covariance of selection
+gradients and competition coefficients, equations (\ref{cov_eqns}) of
+the main text, and for approximating the correlation of competition
+coefficients and a metric of pairwise coevolution, Figure \ref{cov_fig}
+of the main text, we developed a numerical approach by repeatedly
+simulating our system (\ref{comm_dynamics}) of the main text across a
+set of randomly drawn model parameters. In particular, we ran each
+numerical integration with \(S=100\) species for \(T_1=1000.0\) units of
+time. We then continued to integrate for \(T_2=1000.0\) units of time
+and calculated the covariances between competition coefficients and the
+quantities \(|\beta|\), \(\gamma\) and \(\mathfrak{C}\) for each of the
+last \(T_2\) time steps. We assume the temporal average of these
+covariances across the last \(T_2\) units of time approximates the
+expectation at equilibrium. We repeated this approach for randomly drawn
+\(a\) and \(c\) until our sample size reached \(1000\). We drew values
+for \(a\) and \(c\) from condition log-normal distributions. For \(a\),
+we set the log-normal parameters equal to \(\mu=-10\) and \(\sigma=6\)
+and conditioned on \(a<0.01\) to prevent extinction. For \(c\) we set
+the log-normal parameters equal to \(\mu=-4\) and \(\sigma=4\) and
+conditioned on \(c<1\times 10^{-4}\) to prevent very large interspecific
+phenotypic variation and extinction. When extinction occurs, we aborted
+the simulation and restarted with newly drawn parameters. Results for
+\(\mathrm{Corr}_{f_{\bar X}}(\alpha,|\beta|)\) and
+\(\mathrm{Corr}_{f_{\bar X}}(\alpha,\gamma)\) are displayed in SM Figure
+\ref{num_corr}.
+
 \begin{figure}
@@
 
-\caption{Numerical estimate for the correlations of selection gradients and competition coefficients.}\label{fig:unnamed-chunk-13}
+\caption{\label{num_corr} Numerical estimate for the correlations of selection gradients and competition coefficients.}\label{fig:unnamed-chunk-13}
 \end{figure}
-
-Details on simulations, table of parameters, distributions of \(a\) and
-\(c\).
 
@@
 
+\leavevmode\hypertarget{ref-Athreya1972}{}%
+Athreya, Krishna B., and Peter E. Ney. 1972. \emph{Branching Processes}.
+Springer Berlin Heidelberg.
+
 \leavevmode\hypertarget{ref-9781107648258}{}%
@@
 
+\leavevmode\hypertarget{ref-Brown1956}{}%
+Brown, W. L., and E. O. Wilson. 1956. ``Character Displacement.''
+\emph{Systematic Zoology} 5 (2). Oxford University Press (OUP): 49.
+
 \leavevmode\hypertarget{ref-bulmer1980}{}%
@@
 \leavevmode\hypertarget{ref-Case1986}{}%
-Case, Ted J., and Mark L. Taper. 1986. ``ON THE COEXISTENCE AND
-COEVOLUTION OF ASEXUAL AND SEXUAL COMPETITORS.'' \emph{Evolution} 40
+Case, Ted J., and Mark L. Taper. 1986. ``On the Coexistence and
+Coevolution of Asexual and Sexual Competitors.'' \emph{Evolution} 40
 (2). Wiley: 366--87.
@@
 Dynamics.'' \emph{Evolution} 50 (2). Wiley: 532--46.
-
-\leavevmode\hypertarget{ref-Engen2013}{}%
-Engen, Steinar, Russell Lande, and Bernt-Erik Sæther. 2013. ``A
-Quantitative Genetic Model of R- and K-Selection in a Fluctuating
-Population.'' \emph{The American Naturalist} 181 (6). University of
-Chicago Press: 725--36.
 
@@
 
+\leavevmode\hypertarget{ref-Polechov2015}{}%
+Polechová, Jitka, and Nicholas H. Barton. 2015. ``Limits to Adaptation
+Along Environmental Gradients.'' \emph{Proceedings of the National
+Academy of Sciences} 112 (20). Proceedings of the National Academy of
+Sciences: 6401--6.
+
 \leavevmode\hypertarget{ref-PRICE1970}{}%

### 2020-04-18 .A White Noise Approach To Evolutionary Ecology/draft_04_18_20.Rmd
final draft

--- 2020-04-13 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
@@
 We briefly investigate the relationship between selection gradients and competition coefficients.
-This illustrative investigation suggests selection gradients can be either positively or negatively correlated with competition coefficients depending on the ratio of interspecific trait variation to intraspecific trait variation.
+This illustrative investigation suggests absolute values of linear selection gradients are negatively correlated with competition coefficients when the ratio of interspecific trait variation to intraspecific trait variation is approximately less than $1.25$ and positively correlated otherwise.
+The inverse relationship holds for quadratic selection gradients except with threshold diversity ratio equal to unity instead of $1.25$.
 Hence, this approach contributes to the development of a synthetic theory of evolutionary ecology by formalizing first principle derivations of stochastic models that underlie rigorous investigations of the relationship between feedbacks of biological processes and the patterns of diversity they produce."
@@
 In section \ref{deterministic} we review the mathematical formalization of this connection, which has been established in the contexts of population genetics [@9781932846126; @joanroughgarden1979], evolutionary game theory [@Hofbauer1998; @9780674023383; @Lion2018], quantitative genetics [@Lande1982; @Doebeli1996; @Lion2018] and a unifying framework for these three distinct approaches to evolutionary theory [@Champagnat2006] which includes our approach as a special case.
-We note here there are two primary differences between our aims and those of Champagnat, Ferrière, and Méléard [-@Champagnat2006].
+We note here there are two primary differences between the results presented in this paper and those of Champagnat, Ferrière and Méléard [-@Champagnat2006].
 First, instead of developing a unifying approach to evolutionary theory, we focus on developing a stochastic synthesis of population dynamics and quantitative genetics including specific expressions for the dynamics of abundance, mean trait and additive genetic variance.
-Second, although Champagnat, Ferrière, and Méléard [-@Champagnat2006] provide rigorous proofs of results relevant to our approach, they do not communicate the nessecary tools we make use of for deriving the dynamics of abundance, mean trait and additive genetic variance.
-We aim to communicate these tools to a wide audience by providing an intuitive, yet accurate treatment of white noise and associated heuristics useful for deriving new stochastic models of evolutionary ecology.
-
-Reviewing the above accomplishments reveals a beautiful synthesis of evolution and population ecology.
-However, it also reveals a gap in theoretical approaches to incorporate the intrinsically random nature of populations.
+Second, although Champagnat, Ferrière and Méléard [-@Champagnat2006] provide rigorous proofs of results relevant to our approach, they do not communicate the nessecary tools we make use of for deriving the dynamics of abundance, mean trait and additive genetic variance.
+We aim to communicate these tools to a wide audience by striking a balance between intuitive appeal and mathematical rigor.
+
+Although reviewing previously established theoretical results reveals a beautiful synthesis of evolution and population ecology, it also reveals a gap in theoretical approaches to incorporate the intrinsically random nature of populations.
 Specifically, in theoretical quantitative genetics the derivation of a population's response to random genetic drift is derived in discrete time under the assumption of constant effective population size using arguments based on properties of random samples [@Lande1976].
@@
 However, for populations with a continuum of types, such as a quantitative trait, finding a formal approach to derive the evolutionary response to demographic stochasticity has remained a vexing mathematical challenge.
-Previous work integrating stochastic processes with quantitative genetic models have profited from the simplifying assumption that phenotypic and genetic variances remain constant over time, yeilding analytical approximations for the joint dynamics of abundance and the mean of a quantitative trait [@Engen2013].
-These results not only ignore the evolution genetic variance, but they focus on environmental stochasticity and assume sufficiently large populations so that demographic stochasticity is negligible.
-Here we account for both demographic stochasticity and the evolution of additive genetic variance by combining the calculus of white noise with results on rescaled limits of branching Brownian motion processes (BBM) and stochastic partial differential equations (SPDE), which are stochastic analogs of deterministic partial differential equations (PDE).
+In this paper we close this gap by combining the calculus of white noise with results on rescaled limits of branching Brownian motion processes (BBM) and stochastic partial differential equations (SPDE), which are stochastic analogs of deterministic partial differential equations (PDE).
 Our goal has two components: 1) Establish a novel synthetic approach to theoretical evolutionary ecology that provides a formal connection between demographic stochasticity and random genetic drift in the context of quantitative traits.
@@
 Firstly, we develop a concise, but accurate approach for understanding SPDE and provide heuristics to empower mathematical evolutionary ecologists to derive novel models that relax key assumptions commonly made in combining models of evolutionary and abundance dynamics.
-Second, we formally derive general SDE that express the joint dynamics of abundance, mean trait and additive genetic variance for populations exhibiting a wide range of potential phenotypic distributions.
+Second, we formally derive general SDE driven by demographic stochasticity that express the joint dynamics of abundance, mean trait and additive genetic variance for populations exhibiting a wide range of potential phenotypic distributions.
+Although these dynamics have been considered before while accounting for demographic stochasticity [@Polechov2015], to our knowledge this is the first time they have been rigorously derived from first principles.
 Third, we simplify these SDE by approximating the abundance density of phenotypes with a Gaussian curve and integrate the results with a model of imperfect inheritance to find formula that generalize classical quantitative genetic expressions.
-This includes an expression for additive genetic variance analogous to the expression for the response in mean phenotype found by Lande [-@Lande1976].
+This includes an expression for the response of additive genetic variance to selection in terms of fitness gradients analogous to the expression for the response in mean phenotype found by Lande [-@Lande1976].
 Fourth, following classical niche theory we provide a rigorous mechanistic derivation of fitness for populations experiencing resource competition and abiotic stabilizing selection using diffusion limits of measure-valued branching processes.
 This calculation provides a demonstration for performing such diffusion limits that can be applied in other contexts.
-Fifth, we explore implications of the resulting model of diffuse coevolution by investigating the relationship between selection gradients and competition coefficients.
+Fifth, we explore implications of the resulting model of diffuse coevolution by investigating the relationship between selection gradients and competition coefficients, discovering the conditions under which these quantities positively or negatively covary.
 
 We now summarize the basic structure of the paper.
-To provide motivation for the stochastic equations developed later and background for our model of coevolution, we begin with \S\ref{deterministic} by briefly summarizing derivations of deterministic dynamics of biological populations.
+To provide motivation for the stochastic equations developed later and background for our model of coevolution, we begin with \S\ref{deterministic} by briefly summarizing derivations of deterministic evolutionary and demographic dynamics.
 Starting with a PDE (which we refer to as the Deterministic Asexual Gaussian allelic model with Abundance dynamics, abbreviated DAGA), we arrive at a general set of ordinary differential equations modelling the dynamics of abundance, trait mean and trait variance.
 We discuss the relationship between DAGA and previous models of phenotypic evolution that have been employed in theoretical quantitative genetics.
-Noting that a natural stochastic generalization of DAGA has been previously derived, we find a path to derive SDE describing the evolutionary response to demographic stochasticity.
-We accomplish this in \S\ref{wnc} by discussing how a diffusion limit of a spatially structured branching process (i.e., a measure-valued branching process) leads to the natural SPDE appropriate for our study.
+In \S\ref{wnc} we discuss how a diffusion limit of a spatially structured branching process (i.e., a measure-valued branching process) leads to a SPDE that provides a natural stochastic generalization of DAGA.
 We refer to this SPDE as the Stochastic Asexual Gaussian allelic model with Abundance dynamics (abbreviated SAGA).
 The diffusion limit in turn provides a rigorous method for constructing fitness functions used in models of evolutionary ecology.
-However, before we discuss measure-valued braching processes and their diffusion limits, we find it necessary to review some fundamental results on the calculus of white noise and introducing a set of mathematical tools based on these results.
-In \S\ref{wnc} and SM \S\ref{SDE_DERIV} we employ these tools to derive a system of SDE generalizing our deterministic results to account for demographic stochasticity.
-We find that the effect of demographic stochasticity on the evolution of mean trait and phenotypic variance characterizes the process of random genetic drift in continuous time.
+However, before we discuss measure-valued braching processes and their diffusion limits, we find it necessary to review some fundamental results on the calculus of white noise and introduce a set of mathematical tools based on these results.
+In SM \S\ref{SDE_DERIV} we employ these tools to derive a system of SDE generalizing our deterministic results to account for demographic stochasticity.
+We find the effect of demographic stochasticity on the evolution of mean trait and phenotypic variance characterizes the process of random genetic drift in continuous time.
 However, although biologically insightful, these equations remain difficult to analyze and implement numerically.
-In \S\ref{equations} we use an assumption of normally distributed trait values to simplify these expressions into formulae that are much easier to work with.
-We then account for the constraint of adaptive evolution on the amount of heritable variation in a population by extending these results via a model of imperfect inheritance.
+In \S\ref{equations} we use an assumption of normally distributed trait values to simplify these expressions into tractable formula that express the evolutionary response to selection in terms of fitness gradients.
+We then account for the constraint of adaptive evolution on the amount of heritable variation available in a population by extending these results via a model of imperfect inheritance.
 The resulting equations coincide with classical results in quantitative genetics as a special case.
 In \S\ref{coev} we combine the derived equations of population dynamics with classical niche theory to formulate a model of coevolution across a guild of $S$ species participating in exploitative competition along a common resource continuum.
-In SM \S\ref{diffuse} we apply a classical theorem on rescaled limits of BBM that allow for ecological interactions to provide a rigorous derivation.
-To gain biological insight, in \S\ref{dynamics} we numerically integrate our model of coevolution for $S=100$ species, tracking the dynamics of traits and abundances under scenarios of weak and strong competition.
+In SM \S\ref{diffuse} we apply a result on rescaled limits of BBM to provide a rigorous derivation.
+In \S\ref{dynamics} we illustrate our model of coevolution through numerical integration for $S=100$ species, tracking the dynamics of traits and abundances under scenarios of weak and strong competition.
 We include an account of the natural history of the simulated system and discuss the significance of demographic stochasticity for structuring ecological communities.
@@
 A popular alternative to quantitative genetics is the theory of adaptive dynamics.
-As demonstrated by Page and Nowak [-@PAGE2002] and Champagnat, Ferrière, and Méléard [-@Champagnat2006], the canonical equation of adaptive dynamics can be derived from the replicator-mutator equation.
+As demonstrated by Page and Nowak [-@PAGE2002] and Champagnat, Ferrière and Méléard [-@Champagnat2006], the canonical equation of adaptive dynamics can be derived from the replicator-mutator equation.
 Thus, one could start from the atomic roots of our approach and pursue a stochastic adaptive dynamic theory instead.
@@
 
-In this section we review the derivations of the replicator-mutator equation and trait dynamics from abundance dynamics and extend these formulae along with related results to the stochastic case.
+In this section we review the derivations of the replicator-mutator equation and trait dynamics from abundance dynamics and extend these formula along with related results to the stochastic case.
 The results established in this section provide the framework from which larger scale ecological stuctures, such as species abundance distributions and interaction networks, can be computed.
@@
 \end{equation}
-Two important observations of these equations include 1) mean fitness $\bar m$ determines the abundance dynamics of the entire population and 2) allele $i$ will increase (decrease) in frequency if $m_i>\bar m$ ($<\bar m$).
+Two important observations of these equations include (i) mean fitness $\bar m$ determines the abundance dynamics of the entire population and (ii) allele $i$ will increase (decrease) in frequency if $m_i>\bar m$ ($<\bar m$).
 Equation (\ref{popgen}) is known in the field of evolutionary game theory as the replicator equation [@Hofbauer1998; @9780674023383; @Lion2018; @Taylor1978; @Schuster1983].
@@
 Second, we assume selection is determined by the growth rate $m(h,x)$ that is differentiable with respect to both arguments and is bounded above by some $r\in\mathbb{R}$ for all $x\in\mathbb{R}$ and $h\geq0$.
+We also assume $(1+h)C\leq m(h,x)$ for some constant $C>0$ so that $m(h,x)$ is linear or more than linear in $h$.
 We denote by $K$ an operator that takes functions such as $\nu(x,t)$ as its argument and returns other functions that may still depend on the spatial variable $x$ and the temporal variable $t$.
@@
 with the initial condition $\nu(x,0)$ described above.
-This PDE is semilinear due to the dependency of the growth rate $m(\nu,x)$ on the solution $\nu(x,t)$ and is referred to as a scalar reaction-diffusion equation [@zheng2004nonlinear; @lawrenceevans2010].
+Equation (\ref{eq1}) is a semilinear parabolic PDE and is referred to as a scalar reaction-diffusion equation [@zheng2004nonlinear; @lawrenceevans2010].
 To ensure existence and uniqueness of solutions to (\ref{eq1}), we further assume that for each $M>0$ there exists a constant $L_M>0$ depending on $M$ such that $\int_\mathbb{R}\nu_1(x)dx,\int_\mathbb{R}\nu_2(x)dx\leq M$ implies
 \begin{equation}\label{local_lipschitz}
-\int_\mathbb{R}\left|m(\nu_1,x)\nu_1(x)-m(\nu_2,x)\nu_2(x)\right|dx\leq L_M\int_\mathbb{R}\nu_1(x)-\nu_2(x)dx.
+\int_\mathbb{R}\left|m(\nu_1,x)\nu_1(x)-m(\nu_2,x)\nu_2(x)\right|dx\leq L_M\int_\mathbb{R}|\nu_1(x)-\nu_2(x)|dx.
 \end{equation}
@@
 In fact, if the growth rate is constant across $x$, then this model of mutation will cause $\nu(x,t)$ to converge to a flat line as $t\to\infty$.
+
 Although clearly an idealized representation of biological reality, this model is sufficiently general to capture a large class of dynamics including density dependent growth and frequency dependent selection.
@@
 \end{equation}
-where $r\in\mathbb{R}$ is the intrinsic growth rate in the absence of abiotic stabilizing selection, $a>0$ is the strength of abiotic stabilizing selection around the phenotypic optimum $\theta\in\mathbb{R}$, $c>0$ captures the sensitivity of fitness to intraspecific competition and we have set $\kappa(x-y)=1$ so that competitive interactions cause the same reduction in fitness regardless of trait value.
+where $r\in\mathbb{R}$ is the intrinsic rate of growth in the absence of abiotic stabilizing selection, $a>0$ is the strength of abiotic stabilizing selection around the phenotypic optimum $\theta\in\mathbb{R}$, $c>0$ captures the sensitivity of fitness to intraspecific competition and $\kappa(x-y)=1$ so that competitive interactions cause the same reduction in fitness regardless of trait value.
 This exemplary fitness function has a few convenient properties.
@@
 Note in particular the coincidence of equilibrial phenotypic variance predicted by this model and that predicted by the classic Gaussian allelic model of quantitative genetics [@Lande1975; @9780471986539; @Johnson2005].
-By replacing $\kappa(x-y)=1$ in expression (\ref{special_m}) with a particular Gaussian function, we will find in \S\ref{coev} and SM \S\ref{diffuse} this fitness function becomes equivalent to a fitness function derived from niche theory that we use to model coevolution driven by resource competition.
+In \S\ref{coev} and SM \S\ref{diffuse} we find this fitness function is closely related to a fitness function derived from niche theory that we use to model coevolution driven by resource competition.
 
@@
 =\frac{1}{N(t)}\Big(m(\nu,x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)-\bar m(t)\nu(x,t)\Big) \\
-=(m(\nu,x)-\bar m(t)) p(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t).
+=\Big(m(\nu,x)-\bar m(t)\Big) p(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t).
 \end{multline}
@@
 The primary difference being that our model utilizes diffusion instead of convolution with an arbitrary mutation kernel.
-Of course, our model of mutation can be derived as an approximation to Kimura's model, which, for sexually reproducing populations, has been referred to as the Gaussian allelic approximation in reference to the distribution of mutational effects at loci in a genome on the values of traits [@Lande1975; but see also @9780471986539; and @Johnson2005], the infinitesimal model in reference to modelling continuous traits as being encoded by an infinite number of loci each having infinitesimal effect [@Barton2017] and the Gaussian descendants approximation in reference to offspring trait values being normally distributed around their parental values [@Turelli2017].
+Of course, our model of mutation can be derived as an approximation to Kimura's model, which, for sexually reproducing populations, has been referred to as the Gaussian allelic approximation in reference to the distribution of mutational effects at loci in a genome on the values of traits [@Lande1975; but see also @9780471986539; along with @Johnson2005], the infinitesimal model in reference to modelling continuous traits as being encoded by an infinite number of loci each having infinitesimal effect [@Barton2017] and the Gaussian descendants approximation in reference to offspring trait values being normally distributed around their parental values [@Turelli2017].
 Alternatively, since diffusion is the continuous-time equivalent to convolution against a Gaussian kernel (SM \S\ref{diffconvequiv}), equation (\ref{p_PDE}) can also be seen as a special case of Kimura's continuum-of-alleles model.
 We return to these concepts in \S\ref{inheritance} where we review a simple model of inheritance that leads us to an expression for the stochastic evolution of additive genetic variance.
-To clarify the distinction between this model and previous models of phenotypic evolution we refer to the PDE (\ref{eq1}) from which (\ref{p_PDE}) was derived as the Deterministic Asexual Gaussian allelic model with Abundance dynamics (abbreviated DAGA).
+To distinguish this model from previous models of phenotypic evolution we refer to PDE (\ref{eq1}) from which (\ref{p_PDE}) was derived as the Deterministic Asexual Gaussian allelic model with Abundance dynamics (abbreviated DAGA).
 Later, we will extend this model to include the effects of demographic stochasticity, which we refer to as the Stochastic Asexual Gaussian allelic model with Abundance dynamics (abbreviated SAGA).
@@
 In SM \S\ref{SDE_DERIV}, we extend these results to include the effects of demographic stochasticity.
-The key realization in this direction is that a rigorous derivation of a SPDE that naturally generalizes DAGA and accounts for ecological interactions has been provided by Li [-@zeng1998absolute].
+The key realization in this direction is that a rigorous derivation of a SPDE that naturally generalizes DAGA and accounts for ecological interactions has been provided by Li [-@zeng1998absolute] and slightly generalized by Champagnat, Ferriére and Méléard [-@Champagnat2006].
 To understand how this SPDE was derived, we provide an informal discussion of measure-valued branching processes (which treat populations as sets of discrete individuals) and their diffusion limits, the so-called superprocesses (which treat population size as a continuously varying number).
 Similar to the approach taken by Champagnat, Ferriére and Méléard [-@Champagnat2006], we begin with a BBM that models populations as discrete sets of reproducing individuals whose vital rates depend on their trait value as well as the state of the entire population.
-Taking a large population size limit and keeping our assumption of single dimensional traits, we employ a pair of classical results that show, under the appropriate rescaling in time, phenotypic space and population density, a sequence of rescaled BBM converges to a limiting process that can be characterized by a SPDE [@Mlard1993; @zeng1998absolute].
+Taking a large population size limit and keeping our assumption of single dimensional traits, we employ a pair of classical results that show, under the appropriate rescaling in time, phenotypic space and population density, a sequence of rescaled BBM converges to a limiting process that can be characterized by a SPDE [@meleard1992interacting; @Mlard1993; @zeng1998absolute; @Champagnat2006].
 The limiting processes of rescaled BBM have been referred to as measure-valued Markov processes [@dawson1993measure] or superprocesses [@alisonetheridge2000].
@@
 The simplicity of our special case allows us to use properties of space-time white noise processes to derive a set of SDE that generalize equations (\ref{N_det}), (\ref{detmean}) and (\ref{vardyn_det}) to include the effects of demographic stochasticity.
-Classical expressions for the effects of random genetic drift on the evolution of mean traits are obtained as a special case.
+Classical expressions for the effects of random genetic drift on the evolution of mean traits are obtained as a further special case.
 
@@
 In relation to Brownian motion, denoted by $W$, white noise can be interpreted of as the derivative of Brownian motion with respect to time, denoted $\dot W$.
-Since Brownian motion can be thought to take infinitesimally small Gaussian distributed jumps at each time point, this leads to the conceptualization of white noise as a collection of Gaussian distributed random variables.
+Since Brownian motion can be thought to take infinitesimally small Gaussian distributed jumps at each time point, this leads to the conceptualization of white noise as a collection of independent Gaussian distributed random variables.
 Figure \ref{wn} illustrates realizations of this conceptualized white noise in one (left) and two (right) dimensions.
@@
 To extract some particularly useful results from this theory relevant to our goal of synthesizing the stochastic dynamics of biological populations, we provide a streamlined approach by capitolizing on the solid ground these authors have established.
-For instance, instead of rigorously proving properties of white-noise, we simply define them to be so, taking solice in the fact that the technical details have been worked out elsewhere.
+For instance, instead of rigorously proving properties of white noise, we simply define them to be so, taking solice in the fact that the technical details have been worked out elsewhere.
 In SM \S\ref{DPZ14} we show how our informal treatment is related to the rigorous treatment provided by Da Prato and Zabczyk [-@DaPrato2014].
@@
 Following this definition of white noise, we compute its value by sampling it using $\mathscr{N}_2$ processes.
-For example, integrating white noise over a region $D\times[0,t]$, with $t>0$ and $D$ a bounded subset of $\mathbb{R}$, is equivalent to evaluating $\mathbf W_t(I_{D\times [0,+\infty)})$ for the deterministic process
-\begin{equation}
-I_{D\times [0,+\infty)}(x,t)=\left\{\begin{matrix}
+For example, integrating white noise over a region $D\times[0,t]$, with $t>0$ and $D$ a bounded subset of $\mathbb{R}$, is equivalent to evaluating $\mathbf W_t(I_D)$ for the deterministic process
+\begin{equation}
+I_D(x,t)=\left\{\begin{matrix}
 0, & x\notin D \\
@@
 \begin{equation}
-\|I_{D\times[0,+\infty))}\|_t^2=\mathbb{E}\left(\int_0^t\int_\mathbb{R}I_{D\times [0,+\infty)}^2(x,s)dxds\right)=t\int_Ddx=t|D|<+\infty, 
-\end{equation}
-where $|D|$ denotes the length of $D$, we have $I_{D\times [0,+\infty)}\in\mathscr{N}_2$.
+\|I_D\|_t^2=\mathbb{E}\left(\int_0^t\int_\mathbb{R}I_D^2(x,s)dxds\right)=t\int_Ddx=t|D|<+\infty, 
+\end{equation}
+where $|D|$ denotes the length of $D$, we have $I_D\in\mathscr{N}_2$.
 Thus, using equations (\ref{exp_WN}) and (\ref{cov_WN}) and adopting the informal notation introduced in equation (\ref{informal}), we can write the following
@@
 \end{equation}
-To relate these formulae to the common notation used for SDE, we write
+To relate these formula to the common notation used for SDE, we write
 \begin{equation}
@@
 This space can represent geographic space or, relevant to our context, phenotypic space.
-In the following subsection, we describe the BBM process, which is a particularly important branching process with spatial stucture.
+In the following subsection, we describe the BBM process, which is a particularly important branching process with spatial structure.
 This process has been very useful in the study of SPDE due to its simplifying assumption that individuals do not interact.
@@
 Biologically, this implies individuals are semelparous An alternative formulation treats birth and death events separately to model iteroparity.
-However, under the appropriate rescaling, both approaches converge to the same diffusion limit.
+However, under the appropriate rescaling, both approaches converge to the same diffusion limit [@Champagnat2006].
 We therefore choose the former approach for the sake of simplicity.
 
-  2) __Reproductive law:__ When a birth event occurs a random (possibly zero) number of offspring are left.
+  2) __Reproductive law:__ When a birth event occurs we assume a random (possibly zero) number of offspring are produced.
 The distribution of offspring left is called the reproductive law or branching mechanism.
@@
 The case of $\mathscr{W}=1$ is referred to as the critical condition.
-Under the critical condition the probability that extinction occurs in finite time is equal to one.
+Since $X_t(\mathbb{R}^d)$ is a Galton-Watson branching process, it is well known the critical condition implies extinction in finite time with probability one, given finite $X_0(\mathbb{R}^d)$ [@Athreya1972].
 
@@
 \end{equation}
-Since $\nu(x,t)$ is not generally differentiable in $x$ or $t$, the derivatives in expression (\ref{neutral_SPDE}) are taken in the _weak_ sense [for examples, see @lawrenceevans2010].
-That is, to rigorously interpret SPDE (\ref{neutral_SPDE}), we integrate the solution $\nu(x,t)$ against functions $f\in C_b^2(\mathbb{R})$, where $C_b^2(\mathbb{R})$ is the set of bounded and twice continuously differentiable functions on $\mathbb{R}$.
+Since $\nu(x,t)$ is not generally differentiable in $x$ or $t$, the derivatives in expression (\ref{neutral_SPDE}) are taken in the _weak_ sense [for a definition see @lawrenceevans2010].
+That is, to rigorously interpret SPDE (\ref{neutral_SPDE}), we integrate the solution $\nu(x,t)$ against functions $f\in C_b^2(\mathbb{R})$ where $C_b^2(\mathbb{R})$ is the set of bounded and twice continuously differentiable functions on $\mathbb{R}$.
 Hence, equation (\ref{neutral_SPDE}) is just an abbreviation for
@@
 Although this provides an impressive list of accomplishments, the Dawson-Watanabe superprocess falls short of our needs.
-In particular this process assumes individuals do not interact and thus precludes its ability to model nonlocal effects on the fitness of individuals, such as those produced via ecological interactions.
+In particular this process assumes individuals do not interact and thus precludes its ability to model nonlocal effects on the fitness of individuals, such as those produced via competitive interactions.
 However, this concern has been addressed, leading to constuctions of superprocesses that account for interactions among individuals.
@@
 Perkins [-@Perkins1992; -@edwinarendperkins1995] developed a theory of stochastic integration with respect to the so-called _Brownian trees_ to characterize interacting superprocesses and establish properties of existence and uniqueness.
-Li [-@zeng1998absolute] built directly off of the construction of Méléard and Roelly [-@meleard1992interacting; -@Mlard1993] to study properties of density processes associated with interacting superprocesses, arriving at a SPDE that forms the foundation of our approach.
-
-Recall, we use $\nu(x,t)$ to denote the density of a superprocess, given it exists.
-Assuming the interactions manifest only in the reproductive law and that spatial movement follows Brownian motion with diffusion parameter $\sqrt\mu$ independent of both $X_t$ and $x$, Li [-@zeng1998absolute] proved a result that implies the interacting superprocess on one dimensional trait space has a density $\nu(x,t)$ which is non-negative, integrable, continuous in time and space and satisfies the SPDE
+Li [-@zeng1998absolute] built directly off of the construction of Méléard and Roelly [-@meleard1992interacting; -@Mlard1993] to study properties of interacting superprocesses and, by assuming individual spatial movement occurs independently of location $x$ and the entire population $X_t$, showed the evolution of associated density processes can be described by SPDE.
+
+Recall $\nu(x,t)$ denotes the density of a superprocess, given it exists.
+Assuming the interactions manifest only in the reproductive law and that spatial movement follows Brownian motion with diffusion parameter $\sqrt\mu$ independent of both $X_t$ and $x$ and the growth rate is bounded both above and below, Li's [-@zeng1998absolute] result implies the interacting superprocess on one dimensional trait space has a density $\nu(x,t)$ which is non-negative, integrable, continuous in time and space and satisfies the SPDE
 \begin{equation}\label{SPDE}
@@
 Comparing equation (\ref{SPDE}) to equation (3.5) of Li [-@zeng1998absolute], our $m$ and $V$ correspond to Li's $b$ and $c$ respectively.
-It is important to note that, under the assumptions made in Méléard and Roelly [-@meleard1992interacting; -@Mlard1993] and Li [-@zeng1998absolute], equation (\ref{SPDE}) is only formal when $m(h,x)$ is bounded across all combinations of $h\geq0$ and $x\in\mathbb{R}$.
-However, recalling our condition $m(h,x)\leq r\in\mathbb{R}$, the growth rates we consider are only bounded above.
-Yet, in the proof of the construction of the interacting superprocess as the limit of rescaled branching diffusions, Méléard and Roelly [-@meleard1992interacting; -@Mlard1993] assumed $m(h,x)$ to be bounded to guarantee the total mass process will have finite mean and variance, for finite $t\geq0$.
-This allowed the authors to employ a tightness criterion for sequences of measures and show the rescaled processes converge to a superprocess with finite total mass.
-Li's [-@zeng1998absolute] result builds directly off of Méléard and Roelly's construction, inheriting the assumption of boundedness for $m(h,x)$.
-However, in Li [-@zeng1998absolute], the sufficiency of $m(h,x)$ being bounded above is even more clear since Li works explicitly with a common upperbound for both $m(h,x)$ and $V$.
-Hence, one can repeat the necessary proofs replacing the assumption that $m(h,x)$ is bounded with the assumption that $m(h,x)$ is merely bounded above to derive the same results.
-An alternative approach to establishing this result that allows for our relaxed assumption on the growth rate has carried out by Champagnat, Ferrière, and Méléard [-@Champagnat2006] using techniques developed by Evans and Perkins [-@Evans1994].
-
-With solutions to SPDE (\ref{SPDE}) well defined for any growth rate bounded above, we can calculate the total mass process $N(t)$ using the mild solution of (\ref{SPDE}) with $f(x)\equiv1\in C_b^2(\mathbb{R})$ (the symbol "$\equiv$" means equal to for every $x$).
+Using techniques developed by Evans and Perkins [-@Evans1994], Champagnat, Ferrière and Méléard [-@Champagnat2006] generalized this result to the case when $m(h,x)$ is bounded above and is linear or more than linear in $h$.
+In particular, SAGA corresponds to a special case of equation (4.6) in Champagnat, Ferrière and Méléard [-@Champagnat2006] and is valid for the assumptions on $m$ made in \S\ref{deterministic}.
+
+With solutions to SAGA well defined, we can calculate the total mass process $N(t)$ using the mild solution of (\ref{SPDE}) with $f(x)\equiv1\in C_b^2(\mathbb{R})$ (the symbol "$\equiv$" means equal to for every $x$).
 That is,
@@
 Since $\bar x(t)=\tilde x(t)/N(t)$ and $\sigma^2(t)=\tilde{\tilde\sigma}^2(t)/N(t)-\bar x^2(t)$, we can use It&ocirc;'s lemma to derive SDE for $\bar x(t)$ and $\sigma^2(t)$, which we perform in SM \S\ref{SDE_DERIV}.
-We make no attempt in finding sufficient conditions for $\int_\mathbb{R}(|x|+x^2+x^4)\nu(x,t)dx<+\infty$ to hold and hence make no general assertions about the existence or uniqueness of $\bar x(t)$ or $\sigma^2(t)$.
+We make no attempt in finding sufficient conditions to ensure $\int_\mathbb{R}(|x|+x^2+x^4)\nu(x,t)dx<+\infty$ and hence make no general assertions about the existence or uniqueness of $\bar x(t)$ or $\sigma^2(t)$.
 Regardless, we will later assume $\nu(x,t)$ can be approximated by a Gaussian curve in $x$ for all $t\geq0$.
@@
 Characteristically, we note the effect of demographic stochasticity on abundance grows with $\sqrt{N(t)}$.
-Hence, dividing by $N$, we find the effects of demographic stochasticity on the per-capita growth rate diminish with increased abundance.
+Hence, dividing by $N(t)$, we find the effects of demographic stochasticity on the per-capita growth rate diminish with increased abundance.
 Relating the response to demographic stochasticity derived here to the effect of random genetic drift derived in classic quantitative genetic theory, if we set $\sigma^2(t)=\sigma^2$ and $N(t)=N$ constant with respect to time, then integrating the stochastic term in equation (\ref{xbar_gen}) over a single unit of time returns a normally distributed random variable with mean zero and variance equal to $V\sigma^2/N$.
@@
 
-These expressions can be used to investigate the dynamics of the mean and variance for general $\nu(x,t)$.
+These expressions can be used to investigate the dynamics of the mean and variance for general $\nu(x,t)$, assuming $\int_\mathbb{R}(|x|+x^2+x^4)\nu(x,t)dx<+\infty$ for all $t\geq0$.
 However, in the next subsection we simplify these expressions by approximating $\nu(x,t)$ with a Gaussian curve.
@@
 
-By assuming $\nu(x,t)$ can be approximated by a Gaussian curve for each $t\geq0$, expressions (\ref{N}), (\ref{xbar_gen}) and (\ref{sig2_gen}) transform into efficient tools for deriving the dynamics of populations given a fitness function $m(\nu,x)$.
-Gaussian phenotypic distributions are often obtained through Gaussian, exponential or weak selection approximations together with a simplified model of inheritance and random mating [@pmid17248993; @Turelli1984; @Turelli1986; @Turelli2017; @9780471986539].
+By approximating $\nu(x,t)$ with a Gaussian curve in $x$ for each $t\geq0$, expressions (\ref{gen_eqns}) transform into efficient tools for deriving the dynamics of populations given a fitness function $m(\nu,x)$.
+Gaussian phenotypic distributions are often obtained through Gaussian, exponential or weak selection approximations together with a simplified model of mutation and asexual reproduction or random mating [@pmid17248993; @Turelli1984; @Turelli1986; @Turelli2017; @9780471986539].
 Alternatively, it has been shown that a Gaussian distribution can provide a reasonable approximation even when selection is strong and non-Gaussian [@pmid7851785].
-However, our approach adds an additional layer of difficulty.
-Even with Gaussian selection, the resulting solution to SPDE (\ref{SPDE}) will only be a Gaussian curve in expectation, assuming a Gaussian curve as the initial condition.
-Yet this difficulty is not as challenging as it may first appear.
-Indeed, since SPDE (\ref{SPDE}) can be derived as a diffusion limit we know that, under the appropriate assumptions on selection, genetic architecture and reproduction, the stochastic departure from a Gaussian curve is negligible when the ratio $V/N$ is small (i.e., when the variance in reproductive output is much smaller than the population size).
-In SM \S\ref{numerical} we demonstrate this result using numerical methods.
-Mathematically, this requirement restricts model parameters to regions that maintain large population sizes.
-Biologically, this implies populations are not at risk of extinction.
-Hence, models developed in this framework are not suitable for studying colonization-extinction dynamics or evolutionary rescue.
-Allowing for these restrictions, we may safely assume that $\nu$ is approximately Gaussian and justify writing
+Hence, given appropriate assumptions on selection, mutation and reproduction, the abundance density $\nu(x,t)$ can be approximated as a Gaussian curve in $x$ when the ratio $V/N$ is small (i.e., when the variance in reproductive output is much smaller than the population size).
+As with any diffusion approximation, this requires a sufficiently large abundance to accurately reflect the dynamics of populations.
+Hence, models developed in this framework are not suitable for studies involving very small population sizes.
+Allowing for these restrictions, we assume
 \begin{equation}
@@
 \end{equation}
-and $\overline{(x-\bar x)^4}=3\sigma^4$. Equation (\ref{covxm}) is the continuous time equivalent to equation (9) in Lande [-@Lande1976]. In particular, these results imply
+and $\overline{(x-\bar x)^4}=3\sigma^4$. In particular, these results imply
 \begin{subequations}\label{no_inher}
@@
 Note that in the above expressions, the partial derivatives of $\bar m$ represent frequency independent selection and the averaged partial derivatives of $m$ represent frequency dependent selection.
-This relationship has already been pointed out by Lande [-@Lande1976] for the evolution of the mean trait, but here we see an analogous relationship holds also for the evolution of trait variance.
+This relationship has already been pointed out by Lande [-@Lande1976] for the evolution of the mean trait in discrete time, but here we see an analogous relationship holds in continuous time and also for the evolution of trait variance.
 
@@
 However, whether modelling expressed characters or breeding values, we refer to SPDE of the form (\ref{rho_SPDE}) as Stochastic Asexual Gaussian allelic models with Abundance dynamics (abbreviated SAGA).
-By replacing $\psi(x,g)$ with different probability densities produces different developmental models.
+Replacing $\psi(x,g)$ with different probability densities produces different developmental models.
 This includes densities that may depend on environmental variables and hence model phenotypic plasticy.
@@
 
-For species $i$ we inherit the above notation for trait value, distribution, average, variance, abundance, etc except with an $i$ in the subscript.
-Real world examples of niche axes include the body size of prey for lizard predators and the date of activity in a season for pollinators competing for floral resources.
+For species $i$ we inherit the above notation for trait value, distribution, average, variance, abundance, et cetera, except with an $i$ in the subscript.
+Real world examples of niche axes include the size of seeds consumed competing finch species and the date of activity in a season for pollinators competing for floral resources.
 For mathematical convenience, we model the axis of resources by the real line $\mathbb{R}$.
@@
 Suppose $\theta_i\in\mathbb R$ is the optimal location along the niche axis for species $i$ such that, in the absence of competition, individuals leave on average $Q_i$ offspring when concentrated at $\theta_i$.
-We capture the rate by which the fitness falls as niche location $\zeta$ leaves the optimum $\theta_i$ by the parameter $A_i\geq0$.
+We assume the expected reproductive output of an individual in species $i$ decreases as niche location $\zeta$ leaves the optimum $\theta_i$ at the rate $A_i>0$.
 Hence, abiotic stabilizing selection along the resource axis can be modelled by the curve
@@
 
-To map the degree of niche overlap to fitness, we assume competition between individuals with niche locations $x_i$ and $x_j$ additively decreases the Malthusian fitness for the individual in species $i$ by $c_i\mathcal{O}_{ij}(x_i-x_j)$ for some $c_i\geq0$.
+To map the degree of niche overlap to fitness, we assume competition between individuals with niche locations $x_i$ and $x_j$ decreases the expected reproductive output for the individual in species $i$ at the rate $c_i\mathcal{O}_{ij}(x_i-x_j)$ for some $c_i>0$.
 We refer to $c_i$ as the sensitivity to competition for species $i$.
@@
 In SM \S\ref{diffuse} we combine our treatment of resource competition with equations (\ref{N}), (\ref{xbarfinal}) and (\ref{Gfinal}) to find
-
 \begin{subequations}\label{comm_dynamics}
@@
 \end{subequations}
-
 where
-
 \begin{subequations}
@@
 \end{subequations}
-
 Despite the convoluted appearance of system (\ref{comm_dynamics}), there are some nice features that reflect biological reasoning.
@@
 However, as biotic selection pushes $\bar x_i$ away from $\bar x_j$, the effect of competition with species $j$ on the fitness of species $i$ rapidly diminishes, reflecting a reduction in niche overlap.
-The divergence of $\bar x_i$ and $\bar x_j$ due to competition is referred to in the community ecology literature as character displacement.
+The divergence of $\bar x_i$ and $\bar x_j$ due to competition is referred to in the community ecology literature as character displacement [@Brown1956].
 We also see that the fitness of species $i$ drops quadratically with the difference between $\bar x_i$ and the abiotic optimum $\theta_i$.
@@
 We then investigate how these covariances change with the ratio of variance of interspecific mean traits to variance of intraspecific individual traits and use a numerical approach to investigate correlations between the strength of pairwise coevolution and competition coefficients.
-
 \begin{table}
@@
 Relating our treatment of resource competition to modern coexistence theory [@Chesson2000], the absolute competition coefficient $\alpha_{ij}$ becomes a dynamical quantity that can be written as
-
 \begin{equation}
@@
 \end{equation}
-
 where
-
 \begin{equation}
@@
 \end{equation}
-
 Hence, $dN_i(t)$ can be expressed as
-
 \begin{equation}
@@
 \end{equation}
-
 Note that although $r_i(t)$ is referred to in the coexistence literature as the intrinsic growth rate of the population, $R_i$ is a deeper intrinsic quantity.
 For now we refer to $R_i$ as the _innate_ growth rate.
-Previous work, to list just a few citations in this literature, has shown the importance of demographic stochasticity [@Turelli1980; @Schreiber2017] and evolutionary dynamics [@Case1986; @Schreiber2018] in determining coexistence of competing species.
+Previous work has shown the importance of demographic stochasticity [@Turelli1980; @Schreiber2017] and evolutionary dynamics [@Case1986; @Schreiber2018] in determining coexistence of competing species.
 However, these studies consider evolution and demographic stochasticity separately.
 To our knowledge, there are no rigorous theoretical investigations of coexistence that account for evolutionary dynamics of quantitative characters and demographic stochasticity simultaneously and, in particular, no studies combining these processes that also include the evolution of additive genetic variance.
-In fact, only recently have rigorous theoretical investigations successfully combined these processes to understand persistence of a single population [@Gomulkiewicz2017].
-Even so, these authors focused on a single species model with a finite number of types (i.e., a population genetic model).
 With the connection outlined in this section formally established, researchers may pursue a postmodern coexistence theory that naturally includes both the evolutionary dynamics of quantitative characters and the effects of demographic stochasticity in a simple synthetic framework.
@@
 In SM \S\ref{diffuse} we show that the standardized directional selection gradient [sensu @Lande1983] induced by species $j$ on species $i$ can be computed as
-
 \begin{equation}
@@
 \end{equation}
-
 Our notation differs from Lande and Arnold [-@Lande1983] in that subscripts here denote species instead of components of multivariate traits and we drop the prime that distinguishes between selection gradients and standardized selection gradients.
@@
 In SM \S\ref{diffuse} we show that the standardized stabilizing selection gradient induced by species $j$ on species $i$ can be computed as
-
 \begin{equation}
@@
 \end{equation}
-
-To measure the total evolutionary change in species $i$ induced by species $j$, we form the metric $\Psi_{ij}=|\beta_{ij}|+|\gamma_{ij}|$.
-Figure \ref{net} displays the joint distribution of $\mathfrak{C}_{ij}=\Psi_{ij}\Psi_{ji}$, our metric of pairwise coevolution, and the product of competition coefficients $\alpha_{ij}\alpha_{ji}$ for two simulated communities, both with richness $S=1000$.
+To measure evolutionary change in species $i$ induced by species $j$, we form the metric $\Psi_{ij}=|\beta_{ij}|+|\gamma_{ij}|$ and define the strength of coevolution as $\mathfrak{C}_{ij}=\Psi_{ij}\Psi_{ji}$.
+Figure \ref{net} displays the joint distribution of $\mathfrak{C}_{ij}$ and the product of competition coefficients $\alpha_{ij}\alpha_{ji}$ for two simulated communities, both with richness $S=1000$.
 The solid contour represents the case of relatively high sensitivity to competition ($c=5\times10^{-7}$) and the dashed contour represents the case of relatively weak sensitivity to competition ($c=1\times10^{-8}$).
@@
 Such assumptions are typical when deriving analytical results in the field of theoretical coevolutionary community ecology [@Nuismer2012; @Nuismer2018].
-If $\bar{\bar x}$ is near $\theta$ and $V_{\bar X}$ is much smaller than $|2R/a-G-\eta|$, then we may approximate $r_i$ with
-
+If $\bar{\bar x}$ is near $\theta$ and $V_{\bar X}$ is much smaller than $|2\frac{R}{a}-G-\eta|$, then we may approximate $r_i$ with
 \begin{equation}
@@
 \end{equation}
-
 In SM \S\ref{coef_grad_moms} we use these assumptions to calculate the first and second order moments describing the joint distribution of competition coefficients and selection gradients across the community.
@@
 However, setting $\alpha(\bar x_i,\bar x_j)=\alpha_{ij}$, $\beta(\bar x_i,\bar x_j)=\beta_{ij}$ and $\gamma(\bar x_i,\bar x_j)=\gamma_{ij}$, the covariances between the magnitude of linear selection gradients and competition coefficients and between stabilizing selection gradients and competition coefficients can be written as
-
-\begin{subequations}
-
+\begin{subequations}\label{cov_eqns}
 \begin{equation}\label{cov_alpha_beta}
@@
 \end{equation}
-
 \end{subequations}
-
 ```{r, eval=F, echo=F}
@@
 ```
-
-For fixed $c,b,\bar r$ and $N$, we visualize these relationships in Figure \ref{cov_fig}.
-To gain insight into the relationship between selection gradients and competition coefficients, note that our assumptions in this section imply $b^{-1}=2(\sigma^2+w)$.
+We visualize these relationships in Figure \ref{cov_fig} for fixed $c,b,\bar r,U$ and $N$.
+In SM \S\ref{coef_grad_moms} we verify the qualitative nature of these relationships by repeating numerical integrations of system (\ref{comm_dynamics}) across a set of randomly drawn model parameters.
+We designed our numerical approach to inherit the assumptions of symmetric model parameters such as mutation rates $\mu_1=\dots=\mu_S$ and abiotic optima $\theta_1=\dots=\theta_S$ and to relax the key assumptions of normally distributed mean traits along with constant additive genetic variances and abundances across species.
+
+Our assumptions in this section imply $b^{-1}=2(\sigma^2+w)$.
 If we further assume $\sigma^2+w\approx\sigma^2$, then $2bV_{\bar X}\approx V_{\bar X}/\sigma^2$.
 That is, when populations are generalists and are comprised of specialist individuals, the value $2bV_{\bar X}$ is approximately equal to the ratio of interspecific mean trait variation to intraspecific individual trait variation.
-Hence, for both covariances we see that there is no relationship between selection gradients and competition coefficients when this ratio is zero.
-From equation (\ref{cov_alpha_beta}) we can use numerical optimization to find that when $V_{\bar X}/\sigma^2\approx1.25$ the relationship between the magnitudes of linear selection gradients and competition coefficients disappears, but when (approximately) $V_{\bar X}/\sigma^2<1.25$ ($>1.25$), this covariance becomes negative (positive).
+Glancing at Figure (\ref{cov_fig}) we see there is no relationship between selection gradients and competition coefficients when this ratio is zero.
+From equation (\ref{cov_alpha_beta}) we can use numerical optimization to find that when $V_{\bar X}/\sigma^2\approx1.25$ the relationship between the magnitudes of linear selection gradients and competition coefficients disappears, but when $V_{\bar X}/\sigma^2\lessapprox1.25$ ($\gtrapprox1.25$), this covariance becomes negative (positive).
 Equation (\ref{cov_alpha_gamma}) states that when $V_{\bar X}/\sigma^2$ is approximately equal to one (or slightly larger), there is no expected relationship between competition coefficients and quadratic selection gradients.
@@
 From a biological perspective, if the ratio $V_{\bar X}/\sigma^2$ is small, then species are packed tightly in phenotypic space.
-In our model this occurs when abiotic stabilizing selection is much stronger than competition ($a\gg c$).
+In our model this occurs when fitness is more sensitive to abiotic stabilizing selection than competition ($a\gg c$).
 This causes species to overlap more in niche space (i.e., large $\alpha$) and creates disruptive selection for greater intraspecific variance (i.e., positive $\gamma$), which explains the positive region of covariance between $\alpha$ and $\gamma$.
 However, as species begin to overlap in niche space, directional selection begins to vanish (i.e., small $|\beta|$), leading to a negative covariance between $\alpha$ and $|\beta|$.
-In the limiting case that two species have perfectly overlapping niches, they will exhibit zero directional selection since a shift in either direction will yield the same fitness advantages.
-
-In the opposite scenario where competition is much stronger than abiotic stabilizing selection ($c\gg a$), species will not evolve to be as tightly packed and instead their niche-centers will be spread out with little ovelap in their resouce utilization curves (i.e., small $\alpha$).
+In the limiting case that two species have perfectly overlapping niches, they will exhibit zero directional selection since a shift in either direction will yield identical fitness advantages.
+
+In the opposite scenario where fitness is much more sensitive to competition than abiotic stabilizing selection ($c\gg a$), species will not evolve to be as tightly packed and instead their niche-centers will be spread out with little ovelap in their resouce utilization curves (i.e., small $\alpha$).
 In this case biotic directional selection will be strong (i.e., large $|\beta|$), particularly for species towards the outer regions of niche space due to asymmetric fitness advantages conferred by shifts in niche-centers.
@@
 
-In summary, we see the relation between competition coefficients and selection is highly non-trivial and depends on the relative magnitudes of different ecological processes shaping the community.
+In summary, we see the relations between competition coefficients and selection gradients are non-trivial and depend on the relative magnitudes of different ecological processes shaping the community.
 However, this does not address the relation between competition coefficients and coevolution per se.
@@
 Instead of confronting this challenge we build on our numerical approach used to justify analytical approximations of $\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)$ and $\mathrm{Cov}_{f_{\bar X}}(\alpha,\gamma)$ to approximate the correlation of $\alpha$ and $\mathfrak{C}$.
-This numerical approach inherits the assumptions of homogeneous background parameters such as the mutation rate $\mu$ and abiotic optima $\theta$, but allows us to relax the assumption that $N$ and $G$ are constant across species and time.
-
-In particular, we numerically integrated system (\ref{comm_dynamics}) for $T_1=1000.0$ units of time and then continued to integrate for $T_2=1000.0$ units of time.
-We then calculate the covariance between the quantities $\alpha$ and $\mathfrak{C}$ across $S=100$ species for each of the last $T_2$ time steps.
-We assume the temporal average of these covariances across the last $T_2$ units of time approximates the expectation at equilibrium.
-We repeated this approach for randomly drawn $a$ and $c$ until our sample size reached $1000$.
-In Figure \ref{cor_coev_fig} we plot the temporally averaged values of $\mathrm{Cov}_{f_{\bar X}}(\alpha,\mathfrak{C})$ against the sensitivity to competition $c$.
+In Figure \ref{cor_coev_fig} we plot values of $\mathrm{Corr}_{f_{\bar X}}(\alpha,\mathfrak{C})$ against the exact ratio of interspecific trait variance to intraspecific trait variance.
 Using a cubic regression, we see the correlation of coevolutionary selection gradients and competition coefficeints is negative at variance ratios below $0.5$, zero at variance ratios between $0.5$ and $1.0$, and drops below zero again at variance ratios above $1.0$.
@@
 We have introduced an approach to derive models of evolutionary ecology using the calculus of white noise and coined SAGA, a SPDE model of phenotypic evolution that accounts for demographic stochasticity.
-From SAGA we derived SDE that track the evolution of abundance, mean trait and additive genetic variance.
+From SAGA we derived SDE that track the dynamics of abundance, mean trait and additive genetic variance.
 Observing the expressions of these SDE we find the effects of demographic stochasticity on the evolution of mean trait and additive genetic variance characterize the effects of random genetic drift.
-By considering the statistics of random samples, this has previously been in models of fixed population size in discrete time [@Lande1976].
-However, we have shown random genetic drift is a result of demographic stochasticity for populations with fluctuating abundances in continuous time.
-To illustrate the relevance of our approach to studies of evolutionary ecology, we comined our SDE with classical competition theory to derive a model of diffuse coevolution and used this model to investigate the relationship between selection gradients and competition coefficients.
-This work demonstrates that connecting contemporary theoretical approaches to evolutionary ecology with some fundamental results in the theory of measure-valued stochastic processes allows for the development of a rigorous, yet flexible approach to synthesizing the dynamics of abundance and distributions of quantitative characters, in agreement with Champagnat, Ferrière, and Méléard [-@Champagnat2006].
+Although Lande [-@Lande1976] has previously characterized the effects of random genetic drift on mean trait evolution in quantitative genetic models, the approach taken assumed constant effective population size and discrete non-overlapping generations.
+In contrast, our approach shows random genetic drift is a result of demographic stochasticity for continuously reproducing populations with fluctuating abundances.
+To illustrate the relevance of our approach to studies of evolutionary ecology, we comined our SDE with classical competition theory to derive a model of diffuse coevolution and used this model to investigate the relationship between standardized selection gradients and competition coefficients.
+We found the absolute value of linear selection gradients are negatively related with competition coefficients when the ratio of interspecific phenotypic variation to intraspecific phenotypic variation is approximately less than $1.25$ and positive when this ratio is approximately greater than $1.25$.
+We found the opposite relationship between quadratic selection gradients and competition coefficients with a threshold diversity ratio equal to unity instead of $1.25$.
+This work demonstrates that connecting contemporary theoretical approaches to evolutionary ecology with some fundamental results in the theory of measure-valued stochastic processes allows for the development of a rigorous, yet flexible approach to synthesizing the dynamics of abundance and distributions of quantitative characters, in agreement with Champagnat, Ferrière and Méléard [-@Champagnat2006].
 Although this approach enjoys these important merits, there remains biological details and their associated technical challenges that need to be confronted for gaining a more thorough and accurate understanding of ecological communities.
@@
 
-__The genetic architecture and distributions of traits__
+__Genetic architecture and intraspecific trait distributions__
 
@@
 Since analytical results are difficult to obtain for general models of mutation, development and selection [c.f., @9780471986539], we again suggest the use of measure-valued branching processes and their individual-based simulations.
-For example, by assuming the mutational effects in the BBM model is Gaussian with variance an order of magnitude greater than the equilibrium phenotypic (or genetic) variance, we obtain the so-called _House-of-cards_ model [@Turelli1984].
-However, perhaps considering heavy-tailed distributions, such as the Laplace-distributions considered in Park et.
+For example, by assuming the mutational effects in the BBM model is Gaussian with variance an order of magnitude greater than the equilibrium phenotypic (or additive genetic) variance, we obtain the _House-of-cards_ model [@Turelli1984].
+However, perhaps considering heavy-tailed distributions such as the Laplace-distributions considered by Park et.
 al.
-[-@Park2017], will provide more realistic, yet less analytically tractable models of mutational effects.
-Such models have important implications in conservation genetics where small populations may evolve primarily in response to mutation and random genetic drift [@Lande1995].
+[-@Park2017] will provide more realistic yet less analytically tractable models of mutational effects.
+Such mutation models have important implications in conservation genetics where small populations evolve primarily in response to mutation and random genetic drift [@Lande1995].
 
@@
 
-To understand patterns found in ecological communities, considerations must push beyond microevolutionary and contemporary ecological processes and consider the macroevolutionary dynamics of ancestral lineages leading to extant populations.
+To understand patterns found in ecological communities considerations must push beyond microevolutionary and contemporary ecological processes and consider the macroevolutionary dynamics of ancestral lineages leading to extant populations.
 Using sub-alpine flower communities as an example, one can observe a very strict ordering of phenology across a broad geographic range.
@@
 
-Some approaches to modelling macroevolutionary trait change simply repurpose microevolutionary models by blindly rescaling time from the units of generations to millions of years [@Nuismer2014; Luke, can you think of others?].
+Some approaches to modelling macroevolutionary trait change simply repurpose microevolutionary models by blindly rescaling time from the units of generations to millions of years [e.g., @Nuismer2014].
 Such an approach makes the implicit assumption that trait evolution is statistically self-similar [sensu @kennethfalconer2014] so that the stochastic evolution of traits on macroevolutionary time scales has the same properties of trait evolution on microevolutionary time scales.
@@
 These emergent properties can be formally characterized by L&eacute;vy processes and have been successfully employed in comparative phylogenetics to fit phenotypic data from extant populations and the fossil record [@Landis2017].
-It would therefore be interesting to investigate whether an application of a separation of time scales argument for the rate of environmental change ($\lambda$) versus the rate of evolutionary and ecological change ($a$) to microevolutionary models derived using our framework can be used to obtain macroevolutionary models that include not only mean trait evolution, but also the evolution of trait variance and abundance.
+It would therefore be interesting to investigate whether an application of a separation of time scales argument for the rate of environmental change ($\lambda$) versus the rate of evolutionary and ecological change ($a$) to microevolutionary models derived via our framework can be used to obtain macroevolutionary models that include not only mean trait evolution, but also the evolution of trait variance and abundance.
 The resulting macroevolutionary models can give rise to novel comparative phylogenetic methods and provide initial conditions for microevolutionary models that capture contemporary dynamics.
@@
 
-## Simulating the rescaled process and numerical evidence of approximate normality \label{numerical}
-
-Here we use a numerical argument to suggest, for
-
-\begin{equation}
-r-\frac{a}{2}(\theta-x)^2-c\int_\mathbb{R}\nu(x,t)dx \leq m(\nu,x)\leq r-\frac{a}{2}(\theta-x)^2, \ \forall (\nu,x)\in C_{1,c}^+(\mathbb{R})\times\mathbb{R},
-\end{equation}
-
-the density process $\nu(x,t)$ defined by SPDE (\ref{SPDE}) of the main text satisfies $\int_\mathbb{R}(|x|+x^2)\nu(x,t)dx<\infty$.
-
-and setting $c<r-(\eta a+\sqrt{\mu a})/2$ ensures $\hat N>1$, which is important for numerical simulations when $N$ is an integer.
-We use these results to help guide our choice of parameter values for simulations of the branching random walk.
-In the following section we provide a detailed description of the branching random walk and how we have chosen to rescale it.
-We then use the rescaled branching random walk to investigate finiteness of moments and normality.
+## Simulating the rescaled process \label{numerical}
+
+Here we provide a detailed description of the branching random walk and how we have chosen to rescale it. We focus on the case
+\begin{equation}
+m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_\mathbb{R}\nu(x,t)dx.
+\end{equation}
 
@@
 However, we replace exponentially distributed lifetimes with deterministic unit time steps for easier implementation.
-Hence, we restrict time to $t=0,1,2,\dots$.
+Hence, we restrict time to $t=0,1,2,\dots,$ and so on.
 Furthermore, we allow individual fitness to depend on both trait value and the state of the entire population.
 For time $t$ we write $\{\xi_1(t),\dots,\xi_{N(t)}(t)\}$ as the set of trait values across all $N(t)$ individuals alive in the population.
-Since our simulation treats individuals instead of continuous distributions of trait values, we can write
-
+Since our simulation follows discrete individuals instead of continuous distributions of trait values, we can write
 \begin{equation}
@@
 \end{equation}
-
 where $\delta(x-\xi_i(t))$ denotes the point mass located at $\xi_i(t)$.
-To allow for imperfect heritability, we also track the set of breeding values which, at time $t$, is denoted by $\{\gamma_1(t),\dots,\gamma_{N(t)}(t)\}$ and should not be confused with the quadratic selection gradients discussed in \S\ref{diffuse} of the main text.
-Then the distribution of breeding values can be written as
-
-\begin{equation}
-\rho(g,t)=\sum_{i=1}^{N(t)}\delta(g-\gamma_i(t)).
-\end{equation}
-
-Following our model of heritiability, the trait value $\xi_i(t)$ is drawn from a normal distribution centered on $\gamma_i(t)$ with variance $\eta$.
+For simplicity we assume perfect heritability.
 At each iteration we draw, for each individual, a random number of offspring from a Negative-Binomial distribution.
@@
 Denoting $q$ the probability of success for each trial and $s$ the number of successes, the mean and variance is given respectively by
-
 \begin{equation}
@@
 \end{equation}
-
 Then if we require the $i$th individual to have mean number offspring $\mathscr W(\nu,\xi_i)$ and variance equal to $V$, the parameters of the associated Negative-Binomial distribution become
-
 \begin{equation}
@@
 \end{equation}
-
-The imposes the restriction $V>\mathscr{W}(\nu,\xi_i)$.
-For each offpsring produced by the individual with breeding value $\gamma_i(t)$, we assign indepently drawn breeding values normally distributed around $\gamma_i(t)$ with variance $\mu$.
-After breeding values have been assigned, we randomly draw trait values for each offpsring as described above.
-For an overview of our model of inheritance, see \S\ref{inheritance} of the main text.
+This imposes the restriction $V>\mathscr{W}(\nu,\xi_i)$.
+For each offpsring produced by the individual with trait value $\xi_i(t)$, we assign indepently drawn trait values normally distributed around $\xi_i(t)$ with variance $\mu$.
 This summarizes the basic structure of our simulation.
 To impose selection and density dependent growth rates, we set
-
 \begin{equation}
@@
 \end{equation}
-
 where the above integral becomes $\int_\mathbb{R}\nu(x,t)dx=\sum_{i=1}^{N(t)}1=N(t)$.
@@
 
-To rescale the branching random walk by a positive integer $n$, we rescale segregation and mutational variance according to $\eta\to \eta$ and $\mu\to \mu/n$, time by $t\to t/n$ and the reproductive law by $V\to V$ and
-
+To rescale the branching random walk by a positive integer $n$, we rescale mutational variance by $\mu\to \mu/n$, time by $t\to t/n$ and the reproductive law by $V\to V$ and
 \begin{equation}
@@
 \end{equation}
-
 We also replace individual mass with $\frac{1}{n}$ and write rescaled abundance as $N^{(n)}(t)=\frac{1}{n}N(nt)$.
-Under this rescaling the deterministic equilibrium of the raw numerical abundance becomes
-
-\begin{equation}
-\hat N=\frac{n^2}{c}\left(\frac{r}{n}-\frac{1}{2n}(\eta a+\sqrt{\mu a})\right)=\frac{n}{c}\left(r-\frac{1}{2}(\eta a+\sqrt{\mu a})\right).
-\end{equation}
-
-The deterministic equilibrium of the rescaled abundance is then
-
-\begin{equation}
-\hat N^{(n)}=\frac{1}{c}\left(r-\frac{1}{2}(\eta a+\sqrt{\mu a})\right).
-\end{equation}
-
-When it exists, we denote by $N^{(\infty)}(t)$ the limiting process of $N^{(n)}(t)$. Then
-
+When it exists, we denote by $N^{(\infty)}(t)$ the limiting process of $N^{(n)}(t)$.
+Then
 \begin{equation}
@@
 \end{equation}
-
-Note that, following the notation of Theorem 1 in Méléard and Roelly [-@meleard1992interacting], setting $\lambda_n=n$, $m_n(\nu)=\mathscr W^{(n)}(\nu,\cdot)$ and $\varepsilon_n=1/n$ satisfies their hypotheses $(\mathscr H_0)$-$(\mathscr H_3)$ when $c=0$.
+Note that, since the limiting fitness function is bounded above and decreases linearly with respect to $N^{(\infty)}(t)$, it satisfies the hypotheses of Champagnat, Ferrière and Méléard [-@Champagnat2006].
 We have implemented this simulation in the programming language Julia.
 A copy can be found at the url:
-
-\texttt{\url{https://github.com/bobweek/branching.brownian.motion.and.spde}}
-
+$$\texttt{\url{https://github.com/bobweek/branching.brownian.motion.and.spde}}$$
 For the sake of illustration, we simulated the unscaled process ($n=1$) and the rescaled process with $n=5$ and $n=20$ for $50$ units of time.
-Results are shown in Figure \ref{rescaled}.
-In the following section we use a statistical test to show, for the lower bound on $m(\nu,x)$, the rescaled process converges to a Gaussian density as $n\to\infty$ and $V/N\to0$.
-
-### Evidence of normality \label{ev_normal}
-
-To demonstrate approximate normality of the phenotypic distribution when $V/N$ is small we utilized the one-sided Kolmogorov-Smirnov test.
-This test compares an empirical cumulative distribution function (i.e., a cumulative distribution function generated from simulated data) to a hypothetical cumulative distribution function by providing a distribution for the maximum distance between these curves.
-More precisely if $F_n(x)$ is the empricial distribution function for a sample of size $n$ and $F(x)$ is the hypothetical distribution function, Kolmogorov-Smirnov statistic is $D_n=\sup_x|F_n(x)-F(x)|$.
+Results are shown in Figure \ref{rescaled} of the main text.
 
@@
 
-Hence, our fitness function is bounded above, as required in the main text.
-
-For the remainder of the derivation we suppress notation indicating dependency on $\pmb\nu$, $x$ and $t$. From (\ref{malth_fit}) we calculate
+Hence, our fitness function is bounded above and decreases linearly with total abundance, as required by the assumptions of Champagnat, Ferrière and Méléard [-@Champagnat2006].
+For the remainder of the derivation we suppress notation indicating dependency on $\pmb\nu$, $x$ and $t$.
+From (\ref{malth_fit}) we calculate
 
@@
 
-```{r, echo=F, fig.align='center', fig.cap='Numerical estimate for the correlations of selection gradients and competition coefficients.'}
+To verify our analytical estimates for the covariance of selection gradients and competition coefficients, equations (\ref{cov_eqns}) of the main text, and for approximating the correlation of competition coefficients and a metric of pairwise coevolution, Figure \ref{cov_fig} of the main text, we developed a numerical approach by repeatedly simulating our system (\ref{comm_dynamics}) of the main text across a set of randomly drawn model parameters.
+In particular, we ran each numerical integration with $S=100$ species for $T_1=1000.0$ units of time.
+We then continued to integrate for $T_2=1000.0$ units of time and calculated the covariances between competition coefficients and the quantities $|\beta|$, $\gamma$ and $\mathfrak{C}$ for each of the last $T_2$ time steps.
+We assume the temporal average of these covariances across the last $T_2$ units of time approximates the expectation at equilibrium.
+We repeated this approach for randomly drawn $a$ and $c$ until our sample size reached $1000$.
+We drew values for $a$ and $c$ from condition log-normal distributions.
+For $a$, we set the log-normal parameters equal to $\mu=-10$ and $\sigma=6$ and conditioned on $a<0.01$ to prevent extinction.
+For $c$ we set the log-normal parameters equal to $\mu=-4$ and $\sigma=4$ and conditioned on $c<1\times 10^{-4}$ to prevent very large interspecific phenotypic variation and extinction.
+When extinction occurs, we aborted the simulation and restarted with newly drawn parameters.
+Results for $\mathrm{Corr}_{f_{\bar X}}(\alpha,|\beta|)$ and $\mathrm{Corr}_{f_{\bar X}}(\alpha,\gamma)$ are displayed in SM Figure \ref{num_corr}.
+
+```{r, echo=F, fig.align='center', fig.cap='\\label{num_corr} Numerical estimate for the correlations of selection gradients and competition coefficients.'}
 knitr::include_graphics('/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/corrs.png')
@@
 
-Details on simulations, table of parameters, distributions of $a$ and $c$.
-
 \newpage

### 2020-04-18 sde_comp_coev.jl

@@ -69,7 +69,7 @@ while num < 1000
         adraw = exp.(rand(Normal(-10,6),1))[1]
     end
 
-    # continue to draw c until it is less than 1e-3
+    # continue to draw c until it is less than 1e-4
     # this prevents astronomical Vₓ and small N
     while maximum(cdraw)>1e-4
         cdraw = exp.(rand(Normal(-20,4),1))[1]

### 2020-04-18 .A White Noise Approach To Evolutionary Ecology/SM.Rmd
split main text and SM

--- 2020-04-18 .A White Noise Approach To Evolutionary Ecology/draft_04_18_20.Rmd
@@
 ---
-title: "A White Noise Approach to Evolutionary Ecology"
-author: "Bob Week, Steve Krone, Scott L. Nuismer, Luke J. Harmon"
-date: "3/1/2020"
-abstract: "We derive the dynamics of the distribution of a quantitative character and the abundance of a biological population from a stochastic partial differential equation driven by space-time white noise.
-In the process we develop a useful set of heuristics to operationalize the powerful, but abstract theory of white noise and measure-valued stochastic processes.
-This approach allows us to compute the full implications of demographic stochasticity on phenotypic distributions and abundances of populations.
-We demonstrate the utility of our approach by deriving a quantitative genetic model of diffuse coevolution mediated by exploitative competition for a continuum of resources.
-Other than trait and abundance distributions, this model predicts interaction networks parameterized by rates of interactions, competition coefficients, and selection gradients.
-We briefly investigate the relationship between selection gradients and competition coefficients.
-This illustrative investigation suggests absolute values of linear selection gradients are negatively correlated with competition coefficients when the ratio of interspecific trait variation to intraspecific trait variation is approximately less than $1.25$ and positively correlated otherwise.
-The inverse relationship holds for quadratic selection gradients except with threshold diversity ratio equal to unity instead of $1.25$.
-Hence, this approach contributes to the development of a synthetic theory of evolutionary ecology by formalizing first principle derivations of stochastic models that underlie rigorous investigations of the relationship between feedbacks of biological processes and the patterns of diversity they produce."
+title: "Supplementary Material (SM)"
 bibliography: ref.bib
@@
 header-includes:
+  - \usepackage{xr}
   - \usepackage{mathtools}
@@
   - \linenumbers
+  - \externaldocument{draft}  
 editor_options: 
@@
 ```
-
-# Introduction
-
-Our goal in this manuscript is to develop a rigorous, but accessible approach to synthesize the stochastic dynamics of abundance, mean trait and heritable variation in biological populations for the study of theoretical evolutionary ecology.
-A primary aim of theoretical evolutionary ecology is the development of mathematical approaches to describe the evolution of populations and their interactions with both the biotic and abiotic environments in which they are embedded.
-Given this consideration, a natural scope for such an approach centers on quantifying the abundance dynamics of populations and the evolution of traits mediating their interactions as functions of relevant abiotic factors.
-Although taking into account abundance, phenotype and environment provides the basis for a partial understanding of the complex nature of biological communities, a deeper understanding must account for the effects of dispersal and the phylogeographic history of interacting lineages [@Kraft2007; @Hickerson2010; @Manceau2016; @markmcpeek2017] along with the genetic basis of ecologically relevant traits [@jeffreyconner2004; @FUSSMANN2007] and feedbacks between populations and the biogeochemical cycles they ultimately depend on [@michelloreau2010; @9781107648258].
-It is therefore ideal that the development of any such mathematical approach anticipates extensions to account for these important factors shaping ecological communities, especially as empirical and conceptual work in these directions continues to grow [@AbdalaRoberts2014; @Klzsch2015; @Crutsinger2015; @Fitzpatrick2015; @Fitzpatrick2017; @Marx2017; @Rudman2017; @Skovmand2018; @VanNuland2019; @Harmon2019].
-Furthermore, the approach would benefit from a stochastic component to capture the chance nature of biological reality [@Lande2003; @DeMeester2018; @Mubayi2019] and serve as a basis for the construction of statistical methods that measure evolutionary and ecological processes occuring in the wild.
-Such methods will tether theory to reality and allow for rigorous tests of hypotheses on the structure and behavior of ecological communities.
-In this paper we introduce a framework that establishes a formal connection between the continuous-time dynamics of abundance and quantitative traits in stochastically evolving populations.
-We then demonstrate the utility of our framework through the derivation and analysis of a model of diffuse coevolution and discuss how it can be extended to account for the details mentioned above.
-
-Current theoretical approaches to synthesize evolution and ecology have capitalized on the fact that biological fitness plays a key role in determining both sets of dynamics.
-While correlation of fitness and genotype is the basis of evolution by natural selection, the mean fitness across all individuals in a population determines the growth, stasis or decline of abundance.
-In section \ref{deterministic} we review the mathematical formalization of this connection, which has been established in the contexts of population genetics [@9781932846126; @joanroughgarden1979], evolutionary game theory [@Hofbauer1998; @9780674023383; @Lion2018], quantitative genetics [@Lande1982; @Doebeli1996; @Lion2018] and a unifying framework for these three distinct approaches to evolutionary theory [@Champagnat2006] which includes our approach as a special case.
-We note here there are two primary differences between the results presented in this paper and those of Champagnat, Ferrière and Méléard [-@Champagnat2006].
-First, instead of developing a unifying approach to evolutionary theory, we focus on developing a stochastic synthesis of population dynamics and quantitative genetics including specific expressions for the dynamics of abundance, mean trait and additive genetic variance.
-Second, although Champagnat, Ferrière and Méléard [-@Champagnat2006] provide rigorous proofs of results relevant to our approach, they do not communicate the nessecary tools we make use of for deriving the dynamics of abundance, mean trait and additive genetic variance.
-We aim to communicate these tools to a wide audience by striking a balance between intuitive appeal and mathematical rigor.
-
-Although reviewing previously established theoretical results reveals a beautiful synthesis of evolution and population ecology, it also reveals a gap in theoretical approaches to incorporate the intrinsically random nature of populations.
-Specifically, in theoretical quantitative genetics the derivation of a population's response to random genetic drift is derived in discrete time under the assumption of constant effective population size using arguments based on properties of random samples [@Lande1976].
-Though this approach conveniently mimics the formalism provided by the Wright-Fisher model of population genetics, real population sizes fluctuate over time.
-Furthermore, since these fluctuations are themselves stochastic, it seems natural to derive expressions for the evolutionary response to demographic stochasticity and consider how the results relate to characterizations of random genetic drift.
-This can be done in continuous time for population genetic models without too much technical overhead, assuming a finite number of alleles [@Lande2009; @Parsons2010; @Gomulkiewicz2017].
-However, for populations with a continuum of types, such as a quantitative trait, finding a formal approach to derive the evolutionary response to demographic stochasticity has remained a vexing mathematical challenge.
-In this paper we close this gap by combining the calculus of white noise with results on rescaled limits of branching Brownian motion processes (BBM) and stochastic partial differential equations (SPDE), which are stochastic analogs of deterministic partial differential equations (PDE).
-Our goal has two components: 1) Establish a novel synthetic approach to theoretical evolutionary ecology that provides a formal connection between demographic stochasticity and random genetic drift in the context of quantitative traits.
-2) Communicate some useful properties of space-time white noise, BBM and SPDE to as wide of audience as possible.
-With this goal in mind we will not provide a rigorous treatment of any of these deep subjects.
-Instead, we introduce a set of heuristics that only require the basic concepts of Reimann integration, partial differentiation and some exposure to Brownian motion and stochastic ordinary differential equations (SDE).
-For a concise introduction to SDE and Brownian motion, we recommend the primer by Evans [-@9781470410544].
-Rigorous treatments of SPDE and rescaled limits of BBM can be found in Walsh [-@Walsh] and Dawson [-@dawson1993measure] respectively.
-
-Through accomplishing these goals this work yields five main results.
-Firstly, we develop a concise, but accurate approach for understanding SPDE and provide heuristics to empower mathematical evolutionary ecologists to derive novel models that relax key assumptions commonly made in combining models of evolutionary and abundance dynamics.
-Second, we formally derive general SDE driven by demographic stochasticity that express the joint dynamics of abundance, mean trait and additive genetic variance for populations exhibiting a wide range of potential phenotypic distributions.
-Although these dynamics have been considered before while accounting for demographic stochasticity [@Polechov2015], to our knowledge this is the first time they have been rigorously derived from first principles.
-Third, we simplify these SDE by approximating the abundance density of phenotypes with a Gaussian curve and integrate the results with a model of imperfect inheritance to find formula that generalize classical quantitative genetic expressions.
-This includes an expression for the response of additive genetic variance to selection in terms of fitness gradients analogous to the expression for the response in mean phenotype found by Lande [-@Lande1976].
-Fourth, following classical niche theory we provide a rigorous mechanistic derivation of fitness for populations experiencing resource competition and abiotic stabilizing selection using diffusion limits of measure-valued branching processes.
-This calculation provides a demonstration for performing such diffusion limits that can be applied in other contexts.
-Fifth, we explore implications of the resulting model of diffuse coevolution by investigating the relationship between selection gradients and competition coefficients, discovering the conditions under which these quantities positively or negatively covary.
-
-We now summarize the basic structure of the paper.
-To provide motivation for the stochastic equations developed later and background for our model of coevolution, we begin with \S\ref{deterministic} by briefly summarizing derivations of deterministic evolutionary and demographic dynamics.
-Starting with a PDE (which we refer to as the Deterministic Asexual Gaussian allelic model with Abundance dynamics, abbreviated DAGA), we arrive at a general set of ordinary differential equations modelling the dynamics of abundance, trait mean and trait variance.
-We discuss the relationship between DAGA and previous models of phenotypic evolution that have been employed in theoretical quantitative genetics.
-In \S\ref{wnc} we discuss how a diffusion limit of a spatially structured branching process (i.e., a measure-valued branching process) leads to a SPDE that provides a natural stochastic generalization of DAGA.
-We refer to this SPDE as the Stochastic Asexual Gaussian allelic model with Abundance dynamics (abbreviated SAGA).
-The diffusion limit in turn provides a rigorous method for constructing fitness functions used in models of evolutionary ecology.
-However, before we discuss measure-valued braching processes and their diffusion limits, we find it necessary to review some fundamental results on the calculus of white noise and introduce a set of mathematical tools based on these results.
-In SM \S\ref{SDE_DERIV} we employ these tools to derive a system of SDE generalizing our deterministic results to account for demographic stochasticity.
-We find the effect of demographic stochasticity on the evolution of mean trait and phenotypic variance characterizes the process of random genetic drift in continuous time.
-However, although biologically insightful, these equations remain difficult to analyze and implement numerically.
-In \S\ref{equations} we use an assumption of normally distributed trait values to simplify these expressions into tractable formula that express the evolutionary response to selection in terms of fitness gradients.
-We then account for the constraint of adaptive evolution on the amount of heritable variation available in a population by extending these results via a model of imperfect inheritance.
-The resulting equations coincide with classical results in quantitative genetics as a special case.
-In \S\ref{coev} we combine the derived equations of population dynamics with classical niche theory to formulate a model of coevolution across a guild of $S$ species participating in exploitative competition along a common resource continuum.
-In SM \S\ref{diffuse} we apply a result on rescaled limits of BBM to provide a rigorous derivation.
-In \S\ref{dynamics} we illustrate our model of coevolution through numerical integration for $S=100$ species, tracking the dynamics of traits and abundances under scenarios of weak and strong competition.
-We include an account of the natural history of the simulated system and discuss the significance of demographic stochasticity for structuring ecological communities.
-In \S\ref{ecoevo} we provide expressions for selection gradients and competition coefficients implied by our model and use these expressions to investigate the relationship between the degree of competition and strength of coevolution.
-Finally, \S\ref{conclusion} concludes with a summary of accomplishments, a few remarks on the limits of this approach and future directions to incorporate more explicitly the effects of small populations, the genetic architecture of quantitative traits, feedbacks with ecosystem processes and the macroevolutionary history of interacting lineages.
-
-# The framework
-
-At the core of our approach is a stochastic analog of the replicator equation with mutation in continuous time and phenotypic space [@Taylor1978; @Schuster1983].
-From this stochastic replicator-mutator equation we derive a system of SDE for the dynamics of abundance, mean trait and additive genetic variance of a population.
-Hence, our approach develops a quantitative genetic theory of evolutionary ecology.
-A popular alternative to quantitative genetics is the theory of adaptive dynamics.
-As demonstrated by Page and Nowak [-@PAGE2002] and Champagnat, Ferrière and Méléard [-@Champagnat2006], the canonical equation of adaptive dynamics can be derived from the replicator-mutator equation.
-Thus, one could start from the atomic roots of our approach and pursue a stochastic adaptive dynamic theory instead.
-We choose the former in anticipation of an extension of our approach that explicitly models the genetic details of populations.
-
-In this section we review the derivations of the replicator-mutator equation and trait dynamics from abundance dynamics and extend these formula along with related results to the stochastic case.
-The results established in this section provide the framework from which larger scale ecological stuctures, such as species abundance distributions and interaction networks, can be computed.
-
-## Deterministic dynamics \label{deterministic}
-
-Our review begins by considering the dynamics of an asexually reproducing population in a homogeneous environment.
-For simplicity, we first assume individuals are haploid and carry one of $K$ alleles each with a different fitness expressed as growth rate before introducing a model involving a quantitative trait.
-Under these assumptions, the derivation of the evolution of allele frequencies due to natural selection can be derived from expressions of exponential growth.
-This, and a few related approaches, have been provided by Crow and Kimura [-@9781932846126, \S 5.3].
-Specifically, denoting $\nu_i$ the abundance of individuals with allele $i$ and $m_i$ the growth rate of allele $i$ [called the Malthusian parameter in @9781932846126], we have
-\begin{equation}\label{first}
-\frac{d\nu_i}{dt}=m_i\nu_i.
-\end{equation}
-Starting from this model, we get the total abundance of the population as $N=\sum_{i=1}^K\nu_i$, the frequency of allele $i$ as $p_i=\nu_i/N$ and the mean fitness of the population as $\bar m=\sum_{i=1}^Kp_im_i$.
-Hence, we can employ some elementary calculus to derive the dynamics of abundance $dN/dt$ and the dynamics of allele frequencies $dp_1/dt,\dots,dp_K/dt$ as
-\begin{equation}
-\frac{dN}{dt}=\sum_{i=1}^K\nu_im_i=N\sum_{i=1}^Kp_im_i=\bar mN,
-\end{equation}
-\begin{equation}\label{popgen}
-\frac{dp_i}{dt}=\frac{d}{dt}\frac{\nu_i}{N}=\frac{1}{N^2}\left(N\frac{d\nu_i}{dt}-\frac{dN}{dt}\nu_i\right)
-=\frac{1}{N}\left(m_i\nu_i-\bar m Np_i\right)=(m_i-\bar m)p_i.
-\end{equation}
-Two important observations of these equations include (i) mean fitness $\bar m$ determines the abundance dynamics of the entire population and (ii) allele $i$ will increase (decrease) in frequency if $m_i>\bar m$ ($<\bar m$).
-Equation (\ref{popgen}) is known in the field of evolutionary game theory as the replicator equation [@Hofbauer1998; @9780674023383; @Lion2018; @Taylor1978; @Schuster1983].
-Instead of being explicitly focused on alleles, the replicator equation describes the fluctuations of relative abundances of various _types_ in a population in terms of the vital rates of each type.
-Using a matrix of transition rates between differing types, it is straight-forward to extend the replicator equation to include mutation, which is known as the replicator-mutator equation [@9780674023383].
-
-Inspired by equations (\ref{first})-(\ref{popgen}), we derive an analog of the replicator-mutator equation for a continuum of types (that is, for a quantitative trait).
-In particular, we model a continuously reproducing population with trait values $x\in\mathbb{R}$ and an abundance density $\nu(x,t)$ that represents the amount of individuals in the population with trait value $x$ at time $t$.
-Hence, the abundance density satisfies $N(t)=\int_{-\infty}^{+\infty}\nu(x,t)dx$ and $p(x,t)=\nu(x,t)/N(t)$ is the relative density of trait $x$ which we also refer to as the phenotypic distribution.
-
-To stay within the realm of biological plausibility we require a set technical assumptions.
-First, we assume the initial abundance density is continuous, non-negative, integrable and has finite trait mean and variance.
-That is, we assume $\nu(x,0)$ is continuous in $x$, satisfies $\nu(x,0)\geq0$ for all $x\in\mathbb{R}$ and
-\begin{equation}
-N(0)=\int_{-\infty}^{+\infty}\nu(x,0)dx<+\infty,
-\end{equation}
-\begin{equation}
--\infty<\bar x(0)=\int_{-\infty}^{+\infty}xp(x,0)dx<+\infty,
-\end{equation}
-\begin{equation}
-\sigma^2(0)=\int_{-\infty}^{+\infty}(x-\bar x(0))^2p(x,0)dx<+\infty,
-\end{equation}
-where $\bar x(t)$ and $\sigma^2(t)$ are respectively the mean trait and phenotypic variance at time $t\geq0$.
-Second, we assume selection is determined by the growth rate $m(h,x)$ that is differentiable with respect to both arguments and is bounded above by some $r\in\mathbb{R}$ for all $x\in\mathbb{R}$ and $h\geq0$.
-We also assume $(1+h)C\leq m(h,x)$ for some constant $C>0$ so that $m(h,x)$ is linear or more than linear in $h$.
-We denote by $K$ an operator that takes functions such as $\nu(x,t)$ as its argument and returns other functions that may still depend on the spatial variable $x$ and the temporal variable $t$.
-For our application we will be focused on operators $K$ that can be written as $(K\nu)(x,t)=\int_{-\infty}^{+\infty}\kappa(x-y)\nu(y,t)dy$ for some non-negative and bounded function $\kappa$.
-We set $h=(K\nu)(x,t)$ to account for nonlocal effects of the abundance density $\nu(x,t)$ on the fitness of individuals with trait value $x$ [@Volpert2014].
-As an example of a nonlocal effect, consider a population experiencing intraspecific competition for a common resource such that the competitive interactions are not mediated by any set of traits.
-In this case all individuals in the population effect the fitness of one another even though they may exhibit a diverse range of trait values.
-The term nonlocal comes from interpreting the trait value as spatial location.
-From here on we abbreviate $m(h,x)=m\big((K\nu)(x,t),x\big)$ to just $m(\nu,x)$.
-Third, we assume mutation is captured by diffusion with coefficient $\frac{\mu}{2}$.
-With these technicalities addressed we model the demographic dynamics of a population and the dynamics of a quantitative character simultaneously by the PDE
-\begin{equation}\label{eq1}
-\frac{\partial}{\partial t}\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)
-\end{equation}
-with the initial condition $\nu(x,0)$ described above.
-Equation (\ref{eq1}) is a semilinear parabolic PDE and is referred to as a scalar reaction-diffusion equation [@zheng2004nonlinear; @lawrenceevans2010].
-To ensure existence and uniqueness of solutions to (\ref{eq1}), we further assume that for each $M>0$ there exists a constant $L_M>0$ depending on $M$ such that $\int_\mathbb{R}\nu_1(x)dx,\int_\mathbb{R}\nu_2(x)dx\leq M$ implies
-\begin{equation}\label{local_lipschitz}
-\int_\mathbb{R}\left|m(\nu_1,x)\nu_1(x)-m(\nu_2,x)\nu_2(x)\right|dx\leq L_M\int_\mathbb{R}|\nu_1(x)-\nu_2(x)|dx.
-\end{equation}
-In SM \S\ref{finite} we combine the above assumptions on growth rate, mutation and initial condition to prove that solutions to (\ref{eq1}) satisfy $N(t),|\bar x(t)|,\sigma^2(t)<+\infty$ for all $t\geq0$.
-
-Interpreting the trait value $x$ as spatial location and ignoring nonlocal effects, which can be satisfied by setting $K\nu=\nu$, this model of spatially distributed population dynamics has been intensely studied [for a review, see @Cantrell2004].
-When $\mu=0$, equation (\ref{eq1}) can be seen as an analog of equation (\ref{first}) for a continuum of types.
-By assuming mutation acts as diffusion, the effect of mutation causes $\nu(x,t)$ to flatten out over time.
-In fact, if the growth rate is constant across $x$, then this model of mutation will cause $\nu(x,t)$ to converge to a flat line as $t\to\infty$.
-
-Although clearly an idealized representation of biological reality, this model is sufficiently general to capture a large class of dynamics including density dependent growth and frequency dependent selection.
-As an example, logistic growth combined with quadratic stabilizing selection can be captured using the growth rate
-\begin{equation}\label{special_m}
-m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_{-\infty}^{+\infty}\nu(y,t)dy
-\end{equation}
-where $r\in\mathbb{R}$ is the intrinsic rate of growth in the absence of abiotic stabilizing selection, $a>0$ is the strength of abiotic stabilizing selection around the phenotypic optimum $\theta\in\mathbb{R}$, $c>0$ captures the sensitivity of fitness to intraspecific competition and $\kappa(x-y)=1$ so that competitive interactions cause the same reduction in fitness regardless of trait value.
-This exemplary fitness function has a few convenient properties.
-First, the effect of competition induces a local carrying capacity on the population, leading to a finite equilibrium abundance over bounded subsets of $\mathbb{R}$.
-Second, abiotic selection prevents the abundance density from diffusing too far from the abiotic optimum, leading to a finite equilibrial abundance across all of $\mathbb{R}$, $\lim_{t\to\infty}N(t)=\tfrac{1}{c}(r-\tfrac{1}{2}\sqrt{a\mu})$, finite equilibrial mean trait $\lim_{t\to\infty}\bar x(t)=\theta$ and finite equilibrial phenotypic variance $\lim_{t\to\infty}\sigma^2(t)=\sqrt{\frac{\mu}{a}}$, so long as $\bar x(0)\in\mathbb{R}$, $\sigma^2(0)\in[0,+\infty)$, $N(0)\in(0,+\infty)$ and $r>\frac{1}{2}\sqrt{a\mu}>0$.
-We prove these statements in SM \S\ref{equilib}.
-Note in particular the coincidence of equilibrial phenotypic variance predicted by this model and that predicted by the classic Gaussian allelic model of quantitative genetics [@Lande1975; @9780471986539; @Johnson2005].
-In \S\ref{coev} and SM \S\ref{diffuse} we find this fitness function is closely related to a fitness function derived from niche theory that we use to model coevolution driven by resource competition.
-
-To derive a replicator-mutator equation from equation (\ref{eq1}), we employ the chain rule from calculus. Writing $\bar m(t)=\int_{-\infty}^{+\infty}m(\nu,x)p(x,t)dx$ for the mean fitness, we have
-
-\begin{multline}\label{N_det}
-\frac{d}{dt}N(t)=\frac{d}{dt}\int_{-\infty}^{+\infty}\nu(x,t)dx=\int_{-\infty}^{+\infty}\frac{\partial}{\partial t}\nu(x,t)dx \\
-=\int_{-\infty}^{+\infty} m(\nu,x)\nu(x,t)dx+\int_{-\infty}^{+\infty}\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)dx \\
-=N(t)\int_{-\infty}^{+\infty} m(\nu,x)p(x,t)dx=\bar m(t) N(t).
-\end{multline}
-Using our assumptions on mutation and rate of growth, we show in SM \S\ref{finite} $\nu(x,t)$ is twice differentiable with respect to $x$ and $\int_{-\infty}^{+\infty}\nu(x,t)dx<\infty$ for all $t\geq0$.
-This implies that we are justified in swapping the order of differentiation and integration and the result $\int_{-\infty}^{+\infty}\frac{\partial^2}{\partial x^2}\nu(x,t)dx=0$ can be derived from the fundamental theorem of calculus.
-Biological reasoning agrees with this latter result since mutation neither creates nor destroys individuals, but merely changes their type from their parental type.
-Taking the same approach, we derive the dynamics of the phenotypic distribution $p(x,t)$ in response to selection and muation as
-
-\begin{multline}\label{p_PDE}
-\frac{\partial}{\partial t}p(x,t)=\frac{\partial}{\partial t}\frac{\nu(x,t)}{N(t)}=\frac{1}{N^2(t)}\left(N(t)\frac{\partial}{\partial t}\nu(x,t)-\nu(x,t)\frac{d}{dt}N(t)\right) \\
-=\frac{1}{N(t)}\Big(m(\nu,x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)-\bar m(t)\nu(x,t)\Big) \\
-=\Big(m(\nu,x)-\bar m(t)\Big) p(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t).
-\end{multline}
-This result closely resembles Kimura's continuum-of-alleles model [@Kimura1965; @Brger1986].
-The primary difference being that our model utilizes diffusion instead of convolution with an arbitrary mutation kernel.
-Of course, our model of mutation can be derived as an approximation to Kimura's model, which, for sexually reproducing populations, has been referred to as the Gaussian allelic approximation in reference to the distribution of mutational effects at loci in a genome on the values of traits [@Lande1975; but see also @9780471986539; along with @Johnson2005], the infinitesimal model in reference to modelling continuous traits as being encoded by an infinite number of loci each having infinitesimal effect [@Barton2017] and the Gaussian descendants approximation in reference to offspring trait values being normally distributed around their parental values [@Turelli2017].
-Alternatively, since diffusion is the continuous-time equivalent to convolution against a Gaussian kernel (SM \S\ref{diffconvequiv}), equation (\ref{p_PDE}) can also be seen as a special case of Kimura's continuum-of-alleles model.
-We return to these concepts in \S\ref{inheritance} where we review a simple model of inheritance that leads us to an expression for the stochastic evolution of additive genetic variance.
-To distinguish this model from previous models of phenotypic evolution we refer to PDE (\ref{eq1}) from which (\ref{p_PDE}) was derived as the Deterministic Asexual Gaussian allelic model with Abundance dynamics (abbreviated DAGA).
-Later, we will extend this model to include the effects of demographic stochasticity, which we refer to as the Stochastic Asexual Gaussian allelic model with Abundance dynamics (abbreviated SAGA).
-
-The covariance of fitness and phenotype across the population is defined as
-\begin{equation}
-\mathrm{Cov}_t\Big(m(\nu,x),x\Big)=\int_{-\infty}^{+\infty}\Big(m(\nu,x)-\bar m(t)\Big)\Big(x-\bar x(t)\Big)p(x,t)dx.
-\end{equation}
-Hence, the dynamics of the mean trait $\bar x(t)$ can be derived as
-
-\begin{multline}\label{detmean}
-\frac{d}{dt}\bar x(t)=\frac{d}{dt}\int_{-\infty}^{+\infty} x p(x,t)dx=\int_{-\infty}^{+\infty} x\frac{\partial}{\partial t} p(x,t)dx \\
-=\int_{-\infty}^{+\infty} x\big(m(\nu,x)-\bar m(t)\big) p(x,t)+x\frac{\mu}{2}\frac{\partial^2}{\partial x^2} p(x,t)dx
-=\mathrm{Cov}_t\Big(m(\nu,x),x\Big).
-\end{multline}
-
-Equation (\ref{detmean}) is a continuous time analog of the well known Robertson-Price equation without transmission bias [@Robertson1966; @PRICE1970; @FRANK2012; @Queller2017; @Lion2018].
-Whether or not the covariance of fitness and phenotype creates change in $\bar x$ to maximize mean fitness $\bar m$ depends on the degree to which selection is frequency dependent [@Lande1976].
-Since this change is driven by a covariance with respect to phenotypic diversity, the response in mean trait to selection is mediated by the phenotypic variance.
-In particular, when $\sigma^2=0$, $\bar x$ will not respond to selection.
-The result $\int_{-\infty}^{+\infty} x\frac{\partial^2}{\partial x^2} p(x,t)dx=0$ can be found by applying integration by parts.
-Following the approach taken to calculate the evolution of $\bar x$, we find the response of phenotypic variation to this model of selection and mutation is
-\begin{equation}\label{vardyn_det}
-\frac{d}{dt}\sigma^2(t)=\mathrm{Cov}_t\Big(m(\nu,x),(x-\bar x)^2\Big)+\mu.
-\end{equation}
-For the sake of space we relegate the derivation of $d\sigma^2/dt$ to SM \S\ref{var_deriv}.
-In the absence of mutation equation (\ref{vardyn_det}) agrees with the result derived by Lion [-@Lion2018] for discrete phenotypes.
-From a statistical perspective, if we think of $(x-\bar x)^2$ as a square error, then in analogy to the dynamics of the mean trait, we see that the response in $\sigma^2$ to selection can be expressed as a covariance of fitness and square error, which is defined in analogy to $\mathrm{Cov}_t(m(\nu,x),x)$.
-Just as for the evolution of $\bar x(t)$, this covariance also creates change in $\sigma^2$ that can either increase or decrease mean fitness $\bar m$, depending on whether or not individual fitness depends on the distribution of traits across the entire population.
-The effect of selection on phenotypic variance can be positive or negative depending on whether selection is stabilizing or disruptive.
-
-In SM \S\ref{SDE_DERIV}, we extend these results to include the effects of demographic stochasticity.
-The key realization in this direction is that a rigorous derivation of a SPDE that naturally generalizes DAGA and accounts for ecological interactions has been provided by Li [-@zeng1998absolute] and slightly generalized by Champagnat, Ferriére and Méléard [-@Champagnat2006].
-To understand how this SPDE was derived, we provide an informal discussion of measure-valued branching processes (which treat populations as sets of discrete individuals) and their diffusion limits, the so-called superprocesses (which treat population size as a continuously varying number).
-Similar to the approach taken by Champagnat, Ferriére and Méléard [-@Champagnat2006], we begin with a BBM that models populations as discrete sets of reproducing individuals whose vital rates depend on their trait value as well as the state of the entire population.
-Taking a large population size limit and keeping our assumption of single dimensional traits, we employ a pair of classical results that show, under the appropriate rescaling in time, phenotypic space and population density, a sequence of rescaled BBM converges to a limiting process that can be characterized by a SPDE [@meleard1992interacting; @Mlard1993; @zeng1998absolute; @Champagnat2006].
-The limiting processes of rescaled BBM have been referred to as measure-valued Markov processes [@dawson1993measure] or superprocesses [@alisonetheridge2000].
-Under the simplifying assumptions inherited from our treatment of deterministic dynamics and the additional assumption that the magnitude of demographic stochasticity is independent of trait values, we obtain as a special case a relatively simple expression for an SPDE that generalizes DAGA.
-The simplicity of our special case allows us to use properties of space-time white noise processes to derive a set of SDE that generalize equations (\ref{N_det}), (\ref{detmean}) and (\ref{vardyn_det}) to include the effects of demographic stochasticity.
-Classical expressions for the effects of random genetic drift on the evolution of mean traits are obtained as a further special case.
-
-In the following section we provide the necessary mathematical tools needed to derive SDE from SPDE.
-Since our aim is to present this material to as wide of audience as possible, our approach deviates from standard definitions to remove the need for a detailed technical treatment.
-In addition to the notions of Reimann integration and partial differentiation already employed, the reader will only need some elementary probability and an intuitive understanding of SDE, including Brownian motion.
-Because space-time white noise, denoted by $\dot W(x,t)$, appears in the SPDE characterizing diffusion limits of BBM, we begin by defining $\dot W(x,t)$ and illustrating its relevant properties including a set of heuristics for performing calculations.
-Treating only the simplist of cases, we then provide a brief review of BBM, their diffusion limits and the SPDE that characterize them.
-For those not interested in the white noise calculus or superprocesses and would rather jump straight into more biologically relevant results, we recommend skipping to \S\ref{equations}.
-
-## White noise calculus and superprocesses \label{wnc}
-
-### Introduction to white noise
-
-One can think of white noise as the static seen on old television sets or infinitely detailed random dust spread across both time and space.
-From a more mathematical, yet still informal perspective, white noise can be thought of as a stochastic process.
-That is, we can picture white noise as a collection of random variables indexed by time and possibly space.
-In relation to Brownian motion, denoted by $W$, white noise can be interpreted of as the derivative of Brownian motion with respect to time, denoted $\dot W$.
-Since Brownian motion can be thought to take infinitesimally small Gaussian distributed jumps at each time point, this leads to the conceptualization of white noise as a collection of independent Gaussian distributed random variables.
-Figure \ref{wn} illustrates realizations of this conceptualized white noise in one (left) and two (right) dimensions.
-
-```{r, include=F, eval=F}
-require("ggplot2")
-require("ggthemes")
-require("plot3D")
-require("gridExtra")
-
-num1 = 1e4
-num2 = 1e4
-
-xi_1 = rnorm(num1)
-xi_2 = rnorm(num2)
-
-x1 = seq(0,1,length.out = num1)
-x2 = seq(0,1,length.out = sqrt(num2))
-
-M = expand.grid(x=x2,y=x2)
-M$value = xi_2
-
-df_1 = data.frame(xi=xi_1,x=x1)
-
-wn1d = ggplot(df_1)+geom_point(aes(x=x,y=xi),size=0.05)+theme_minimal()+xlab("Time")+ylab("Value")
-wn2d = ggplot(M, aes(x, y, color = value))+scale_color_gradient2(low="black",mid="grey50", high="white")+geom_point(size=0.001, shape=22, show.legend = FALSE)+theme_minimal()+xlab("Time")+ylab("Space")
-
-wn = grid.arrange(wn1d,wn2d,nrow=1)
-
-ggsave("~/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/wn.png",wn,width=5,height=2.5)
-```
-
-```{r, echo=F, fig.align='center', fig.cap='\\label{wn}Approximations of sample paths of temporal white noise (left) and space-time white noise (right) with brightness scaled to value.'}
-knitr::include_graphics('wn.png')
-```
-
-However, it turns out that realizations of white noise do not exist as functions in the classical sense.
-Indeed, since Brownian motion is nowhere differentiable with respect to time, white noise cannot be formally understood as a time derivative.
-Thus our notation $\dot W$ is only meant to aid intuition and not be taken as formal.
-A formal understanding is possible by considering white noise as a _measure_-valued process [@Dawson1975; @Walsh] or as a _generalized_ process that acts on classically defined functions or stochastic processes to return either random variables or stochastic processes [@Krylov1981; @DaPrato2014].
-Since a measure-valued process can be defined from a generalized process and a generalized process can be defined from a measure-valued process, the distinction between the two is more or less a matter of perspective.
-However, we find the perspective of white-noise as a generalized process to be a more efficient route for developing heuristics to help with some routine calculations involved with deriving SDE from SPDE.
-Hence, the notion of a generalized process provides the general idea implemented here.
-Although the treatments of SPDE provided by Krylov and Rozovskii [-@Krylov1981] and Da Prato and Zabczyk [-@DaPrato2014] extend the theory of SDE to formally treat SPDE in a general and elegant fashion, they require the navigation of an enourmous amount of technical definitions and detailed proofs.
-To extract some particularly useful results from this theory relevant to our goal of synthesizing the stochastic dynamics of biological populations, we provide a streamlined approach by capitolizing on the solid ground these authors have established.
-For instance, instead of rigorously proving properties of white noise, we simply define them to be so, taking solice in the fact that the technical details have been worked out elsewhere.
-In SM \S\ref{DPZ14} we show how our informal treatment is related to the rigorous treatment provided by Da Prato and Zabczyk [-@DaPrato2014].
-
-Before diving in, we shed a bit of light on the idea of a generalized process.
-A generalized process is the stochastic analog of a generalized function, such as the Dirac delta function $\delta$.
-Often one sees $\delta$ defined as a function satisfying the properties $\delta(x)=0$ for $x\neq0$ and $\int_{-\infty}^{+\infty}\delta(x)dx=1$.
-However, since there is no function that satisfies these properties, we refer to $\delta$ as a generalized function.
-An alternative definition of $\delta$ considers its action on classically defined functions $f$.
-In particular, $\delta(f)=f(0)$, which can be heuristically written as $\int_{-\infty}^{+\infty}f(x)\delta(x)dx=f(0)$.
-Similarly, other generalized functions can be defined by their action on classically defined functions.
-Then, just as a generalized function operates on classical functions to return a value, a generalized process acts on a set of functions (or processes) to return a random variable (or a classically defined stochastic process).
-For a brief primer on the theory of generalized functions, see the addendum to chapter 3 of Kolmogorov and Fomin [-@kolmogorov1999elements].
-
-### Definition and basic properties of white noise \label{wnc_intro}
-
-Throughout this section, we minimize notation by writing $\int_{\mathbb{R}}f(x)dx=\int_{-\infty}^{+\infty}f(x)dx$ and similarly $\int_Df(x)dx$ for the integral of $f$ over $D\subset\mathbb{R}$.
-We define $\mathscr{N}_2$ as the set of stochastic processes $f(x,t)$ that are continuous in $t$ and satisfy $\mathbb{E}\left(\int_0^t\int_\mathbb{R}f^2(x,s)dxds\right)<+\infty$ for each $t\geq0$.
-The operator $\mathbb{E}$ denotes expectation with respect to the underlying probability space.
-For each $t\geq0$ we set
-
-\begin{equation}
-\|f\|_t=\sqrt{\mathbb{E}\left(\int_0^t\int_\mathbb{R}f^2(x,s)dxds\right)},
-\end{equation}
-
-and make use of the convention $f=g$ if $\|f-g\|_t=0$ for all $t\geq0$.
-Later, when we investigate stochastic dynamics of the abundance density $\nu(x,t)$, we will find $\sqrt\nu\in\mathscr{N}_2$.
-This enables us to utilize the heuristics developed in this section for the derivation of SDE describing the stochastic dynamics of $N(t)$, $\bar x(t)$ and $\sigma^2(t)$.
-
-We define a generalized stochastic process $\mathbf W$ that maps processes $f\in\mathscr{N}_2$ to real-valued stochastic processes indexed by time $t\geq0$, but not by space.
-To evaluate $\mathbf W$ for a process $f\in\mathscr{N}_2$ and some time $t\geq0$ we write $\mathbf W_t(f)$.
-Specifically, for any $f,g\in\mathscr{N}_2$, we define $\mathbf W(f)$ and $\mathbf W(g)$ to be Gaussian processes satisfying, for any $t,t_1,t_2\geq0$,
-\begin{equation}\label{exp_WN}
-\mathbb{E}\big(\mathbf W_t(f)\big)=\mathbb{E}\big(\mathbf W_t(g)\big)=0,
-\end{equation}
-\begin{equation}\label{cov_WN}
-\mathbb{C}\big(\mathbf W_{t_1}(f),\mathbf W_{t_2}(g)\big)=\mathbb{E}\left(\int_0^{t_1\wedge t_2}\int_\mathbb{R} f(x,s)g(x,s)dxds\right),
-\end{equation}
-where $t_1\wedge t_2=\min(t_1,t_2)$ and $\mathbb{C}$ denotes covariance with respect to the underlying probability space.
-In particular, denoting $\mathbb{V}$ the variance operator with respect to the underlying probability space, we have $\mathbb{V}\big(\mathbf W_t(f)\big)=\|f\|_t^2$ for all $t\geq0$ and $f\in\mathscr{N}_2$.
-The operators $\mathbb{E}$ and $\mathbb{C}$ are to be distinguished from $\bar f(t)$ and $\mathrm{Cov}_t(f,g)$ which denote expectation and covariance with respect to phenotypic diversity at time $t\geq0$.
-
-Since Gaussian processes are characterized by their expectations and covariances and since we assume the $\mathscr{N}_2$ processes are continuous in time, the processes $\mathbf W(f)$ and $\mathbf W(g)$ are well defined.
-As an example, if $f\in\mathscr{N}_2$ is independent of time, then $\mathbf W(f)$ is a Brownian motion with variance at time $t\geq0$ equal to $\|f\|_t^2=t \ \mathbb{E}(\int_\mathbb{R}f^2(x,0)dx)$.
-With the generalized process $\mathbf W$ defined, we define the space-time white noise $\dot W(x,t)$ implicitly via the stochastic integral
-\begin{equation}\label{informal}
-``\int_0^t\int_\mathbb{R}f(x,s)\dot W(x,s)dxds"=\mathbf W_t(f), \ \forall \ f\in\mathscr{N}_2, \ t\geq0.
-\end{equation}
-We place quotations in the above expression to emphasize its informal nature and that it should not be confused with classical Riemann integration.
-Following this definition of white noise, we compute its value by sampling it using $\mathscr{N}_2$ processes.
-For example, integrating white noise over a region $D\times[0,t]$, with $t>0$ and $D$ a bounded subset of $\mathbb{R}$, is equivalent to evaluating $\mathbf W_t(I_D)$ for the deterministic process
-\begin{equation}
-I_D(x,t)=\left\{\begin{matrix}
-0, & x\notin D \\
-1, & x\in D
-\end{matrix}\right..
-\end{equation}
-Since 
-\begin{equation}
-\|I_D\|_t^2=\mathbb{E}\left(\int_0^t\int_\mathbb{R}I_D^2(x,s)dxds\right)=t\int_Ddx=t|D|<+\infty, 
-\end{equation}
-where $|D|$ denotes the length of $D$, we have $I_D\in\mathscr{N}_2$.
-Thus, using equations (\ref{exp_WN}) and (\ref{cov_WN}) and adopting the informal notation introduced in equation (\ref{informal}), we can write the following
-\begin{equation}
-\mathbb{E}\left(\int_0^t\int_D\dot W(x,s)dxds\right)=0,
-\end{equation}
-\begin{equation}
-\mathbb{V}\left(\int_0^t\int_D\dot W(x,s)dxds\right)=t|D|.
-\end{equation}
-Using this informal notation, equations (\ref{exp_WN}) and (\ref{cov_WN}) can be rewritten as
-\begin{equation}\label{exp_WN_xi}
-\mathbb{E}\left(\int_0^t\int_\mathbb{R}f(x,s)\dot W(x,s)dxds\right)=0,
-\end{equation}
-\begin{equation}\label{cov_WN_xi}
-\mathbb{C}\left(\int_0^{t_1}\int_\mathbb{R}f(x,s)\dot W(x,s)dxds,\int_0^{t_2}\int_\mathbb{R}g(x,s)\dot W(x,s)dxds\right)
-=\int_0^{t_1\wedge t_2}\int_\mathbb{R}f(x,s)g(x,s)dxds.
-\end{equation}
-To relate these formula to the common notation used for SDE, we write
-\begin{equation}
-\hat f(x,t)=\frac{f(x,t)}{\sqrt{\int_\mathbb{R}f^2(y,t)dy}} \ \text{ and } \ 
-d\hat{\mathbf W}_t(f)=\left(\int_\mathbb{R}\hat f(x,t)\dot W(x,t)dx\right)dt
-\end{equation}
-so that
-\begin{equation}
-\int_0^td\hat{\mathbf W}_s(f)=\int_0^t\int_\mathbb{R}\frac{f(x,s)}{\sqrt{\int_\mathbb{R}f^2(s,y)dy}}\dot W(x,s)dxds.
-\end{equation}
-This implies 
-\begin{equation}
-\mathbb{E}\left(\int_0^td\hat{\mathbf W}_s(f)\right)=0, \ \mathbb{C}\left(\int_0^{t_1}d\hat{\mathbf W}_s(f),\int_0^{t_2}d\hat{\mathbf W}_s(f)\right)=t_1\wedge t_2
-\end{equation}
-and in particular, as a function of $t$, $\int_0^td\hat{\mathbf W}_s(f)$ is a standard Brownian motion for any $f\in\mathscr{N}_2$.
-Hence, $d\hat{\mathbf W}_t(f)$ is analogous to the traditional shorthand used to denote stochastic differentials.
-Thus, equation (\ref{cov_WN_xi}) effectively extends It&ocirc;'s multiplication table to:
-\begin{table}[H]
-\centering\caption{An extension of It\^o's multiplication table.}\vspace{0.2cm}
-\begin{tabular}{l|lll}
-             & $d\hat{\mathbf W}_t(f)$               & $d\hat{\mathbf W}_t(g)$                & $dt$ \\ \hline
-             &                            &                            &      \\
-$d\hat{\mathbf W}_t(f)$ & $dt$                       & $\left(\int_\mathbb{R}\hat f(x,t)\hat g(x,t)dx\right)dt$ & $0$  \\
-             &                            &                            &      \\
-$d\hat{\mathbf W}_t(g)$ & $\left(\int_\mathbb{R}\hat f(x,t)\hat g(x,t)dx\right)dt$ & $dt$                       & $0$  \\
-             &                            &                            &      \\
-$dt$         & $0$                        & $0$                        & $0$
-\end{tabular}\label{mult}
-\end{table}
-The extension of It&ocirc;'s multiplication table and properties of white noise outlined in this subsection provide a useful set of tools for working with SPDE.
-In SM \S\ref{SDE_DERIV} we employ these tools to derive SDE that track the dynamics of abundance, mean trait and phenotypic variance of a population from a particular SPDE.
-In the following subsection, we review how this particular SPDE naturally arises as the diffusion limit of a BBM.
-
-### From branching processes to SPDE \label{stochastic}
-
-In real populations individuals are born and potentially reproduce before they ultimately die.
-These three events provide the basic ingredients of a branching process.
-Mathematical investigations of such processes have a relatively deep history [@Kendall1966].
-The most simple branching process, known as the Galton-Watson process, describes the number of individuals alive at a given time $t\geq0$ as a non-negative integer [@Kimmel2015].
-Feller [-@feller1951] introduced a formal method to approximate branching processes with diffusion processes which are continuous in state (i.e., population size is approximated as a continuous quantity).
-Since diffusion processes possess greater analytical tractability than branching processes, Feller's method, known as the diffusion limit, has acquired immense popularity particularly in the field of mathematical population genetics [@Ewens2004].
-For over the past half of a century a great deal of accomplishments have been achieved in formalizing the diffusion limits of branching processes that describe populations of individuals occuring in some continuous space [@Watanabe1968; @Dawson1975; @Dawson1978; @Perkins1992; @Mlard1993; @edwinarendperkins1995; @zeng1998absolute; @Bertoin2003; @Etheridge2008; @Barton2019].
-This space can represent geographic space or, relevant to our context, phenotypic space.
-In the following subsection, we describe the BBM process, which is a particularly important branching process with spatial structure.
-This process has been very useful in the study of SPDE due to its simplifying assumption that individuals do not interact.
-However, this assumption imposes an unfortunate restriction by precluding the modelling of ecological interactions.
-We therefore follow our discussion of BBM with a review of a few important results on spatially structured branching processes that account for interactions.
-
-__Branching Brownian motion__
-
-A BBM tracks individuals navigating $d$-dimensional Euclidean space that reproduce and senesce between exponentially distributed intervals.
-Unlike other stochastic processes that take values in $\mathbb{R}^d$, BBM takes values in the set of _non-negative finite measures_ over $\mathbb{R}^d$.
-Intuitively, one can think of a finite measure as a function that maps subsets of $\mathbb{R}^d$ to real numbers.
-To be formal, we only consider the Borel subsets of $\mathbb{R}^d$ corresponding to the Euclidean metric, but understanding this technicality is not crucial to our discussion.
-In particular, denoting $X_t$ a BBM, for a subset $D\subset\mathbb{R}^d$, $X_t(D)$ returns the (random) number of individuals alive within the region $D$ at time $t\geq0$.
-The BBM has three main components:
-
-  1) __Branching rate:__ In our formulation of BBM we assume Lifetimes of individuals are exponentially distributed with death rate $\lambda>0$ and reproduction occurs simultaneously with death.
-Biologically, this implies individuals are semelparous An alternative formulation treats birth and death events separately to model iteroparity.
-However, under the appropriate rescaling, both approaches converge to the same diffusion limit [@Champagnat2006].
-We therefore choose the former approach for the sake of simplicity.
-
-  2) __Reproductive law:__ When a birth event occurs we assume a random (possibly zero) number of offspring are produced.
-The distribution of offspring left is called the reproductive law or branching mechanism.
-We denote the mean and variance in reproductive output by $\mathscr{W}$ and $V$ respectively.
-The case of $\mathscr{W}=1$ is referred to as the critical condition.
-Since $X_t(\mathbb{R}^d)$ is a Galton-Watson branching process, it is well known the critical condition implies extinction in finite time with probability one, given finite $X_0(\mathbb{R}^d)$ [@Athreya1972].
-
-  3) __Spatial movement:__ Each offspring is born at the current location of their parent.
-Immediately after birth they move around space according to $d$-dimensional Brownian motion with diffusion parameter $\sqrt\mu$.
-In our context we interpret spatial movement as mutation so that the location of an individual at death represents the value of its phenotype.
-Then an individual born at location $x\in\mathbb{R}^d$ that lives for $\tau>0$ units of time will have a normally distributed trait centered on $x$ with covariance matrix equal to $\tau\mu$ times the $d\times d$ identity matrix.
-Hence, offspring inherit normally distributed traits centered on their parental trait.
-This fact creates a vital link to the deterministic dynamics reviewed above.
-Indeed, in the absence of selection, the deterministic PDE (\ref{eq1}) reduces to the $d=1$-dimensional Kolmogorov forward equation for a Brownian motion with diffusion parameter $\sqrt\mu$.
-
-To obtain a SPDE from a BBM we take a diffusion limit.
-There are several ways to do this, but a simple approach is to rescale the mass of individuals and time by $1/n$, diffusion by $\mu\to\mu/n$, branching rate by $\lambda\to n\lambda$, fitness by $\mathscr{W}\to\mathscr{W}^{1/n}$ and consider the limit as $n\to\infty$.
-Denoting the rescaled process by $X_t^{(n)}(D)$, the limiting process $\mathscr{X}_t=\lim_{n\to\infty}X_t^{(n)}$ is called a super-Brownian motion and is also a non-negative finite measure-valued process [@Watanabe1968].
-However, instead of returning the number of individuals alive in a region of space, super-Brownian motion returns the _mass_ of the population concentrated in a region of space.
-Since we have rescaled individual mass by $1/n$ and took the limit $n\to\infty$, individuals are no longer discrete units.
-Instead, the particle view of the population gets replaced by a blob spread across $d$-dimensional space.
-In particular, the value of $\mathscr{X}_t(D)$ is a continuously varying non-negative random variable for any $t\geq0$ and $D\subset\mathbb{R}^d$.
-
-Unfortunately, just as with cream cheese spread across too much toast, the blob perspective of the population may exhibit some complicated spatial discontinuities.
-However, for spatial dimension $d=1$, it turns out that $\mathscr{X}_t$ is absolutely continuous with respect to the Lebesgue measure for each $t\geq0$ [@Konno1988; @Reimers1989].
-This means that we can write $\mathscr{X}_t(D)=\int_D \nu(x,t)dx$ for some density process $\nu(x,t)$.
-Setting $\lambda=1$ and $m=\ln\mathscr{W}$ this density process satisfies the SPDE
-\begin{equation}\label{neutral_SPDE}
-\frac{\partial}{\partial t}\nu(x,t)=m\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)+\sqrt{V\nu(x,t)}\dot W(x,t).
-\end{equation}
-Since $\nu(x,t)$ is not generally differentiable in $x$ or $t$, the derivatives in expression (\ref{neutral_SPDE}) are taken in the _weak_ sense [for a definition see @lawrenceevans2010].
-That is, to rigorously interpret SPDE (\ref{neutral_SPDE}), we integrate the solution $\nu(x,t)$ against functions $f\in C_b^2(\mathbb{R})$ where $C_b^2(\mathbb{R})$ is the set of bounded and twice continuously differentiable functions on $\mathbb{R}$.
-Hence, equation (\ref{neutral_SPDE}) is just an abbreviation for
-
-\begin{multline}\label{neutral_int}
-\int_\mathbb{R}\nu(x,t)f(x)dx-\int_\mathbb{R}\nu(x,0)f(x)dx=\int_0^t\int_\mathbb{R}\nu(x,s)\left(mf(x)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}f(x)\right)dsdx \\
-+\int_0^t\int_\mathbb{R}f(x)\sqrt{V\nu(x,s)}\dot W(x,s)dxds, \ \forall \ f\in C_b^2(\mathbb{R}).
-\end{multline}
-This expression is referred to as the _mild_ solution of (\ref{neutral_SPDE}).
-For more on the general theory of SPDE see Walsh [-@Walsh] and Da Prato and Zabczyk [-@DaPrato2014].
-Note that since $\nu(x,t)$ is the density of a finite measure, it is integrable for each $t\geq0$.
-Thus, since for some $M>0$, $|f(x)|\leq M$ for every $x\in\mathbb{R}$, setting $\varphi(x,t)=f(x)\sqrt{V\nu(x,t)}$ implies $\varphi\in\mathscr{N}_2$.
-Hence, the white noise integral on the right-hand side of equation (\ref{neutral_int}) can be understood using the heuristics introduced above.
-Evaluating equation (\ref{neutral_int}) in the particular case of $f(x)\equiv1$ returns the total mass process, which we refer to as the total abundance $N(t)$.
-
-A convergence theorem for the diffusion limit of a generalization of BBM was established by Watanabe [-@Watanabe1968].
-Dawson [-@Dawson1975] suggested that, for spatial dimension $d=1$, this diffusion limit should admit a density process that satisfies a SPDE.
-Konna and Shiga [-@Konno1988] and Reimers [-@Reimers1989] independently proved Dawson's suggestion was indeed correct.
-The diffusion limit of this more general branching process (in arbitrary spatial dimension) is referred to as a Dawson-Watanabe superprocess [@alisonetheridge2000].
-Conditioning a Dawson-Watanabe superprocess to have constant mass returns a Fleming-Viot process [@Etheridge1991; @Perkins1991] which has been popular in studies of spatial population genetics.
-In particular, an extension of the Fleming-Viot process, known as the $\Lambda$-Fleming-Viot process, was introduced by Bertoin and Le Gall [-@Bertoin2003] and coined by Etheridge [-@Etheridge2008] where it was used to resolve some technical challenges in modelling isolation by distance [@Felsenstein1975; see also @Barton2013; and @Barton2019].
-Although this provides an impressive list of accomplishments, the Dawson-Watanabe superprocess falls short of our needs.
-In particular this process assumes individuals do not interact and thus precludes its ability to model nonlocal effects on the fitness of individuals, such as those produced via competitive interactions.
-However, this concern has been addressed, leading to constuctions of superprocesses that account for interactions among individuals.
-In the next subsection we summarize the main results in this area and introduce the SPDE that provides the basis for our approach to theoretical evolutionary ecology.
-
-__Interacting superprocesses__
-
-In the above subsection we reviewed the diffusion limit of an especially tractable measure-valued branching process.
-However, we found the simplicity of this process restricts us from modelling nonlocal effects on the fitness of individuals.
-Here, we discuss superprocesses that account for such effects.
-The existence of diffusion limits for a class of measure-valued branching processes involving nonlocal effects has been treated by Méléard and Roelly [-@meleard1992interacting; -@Mlard1993].
-The interactions can manifest as dependencies of the spatial movement or reproductive law of individuals on their position $x$ and the state of the whole population described by $X_t$.
-An important result of Méléard and Roelly [-@meleard1992interacting; -@Mlard1993] is a theorem that provides sufficient conditions to construct a sequence of rescaled measure-valued branching processes that converge to a generalization of the Dawson-Watanabe superprocess that includes interactions.
-The rescaling is analogous to that described above for non-interacting Dawson-Watanabe superprocesses, but now the reproductive law described by $\mathscr{W}(X_t,x)$ and $V(X_t,x)$, branching rate $\lambda(X_t,x)$ and diffusion parameter $\sqrt{\mu(X_t,x)}$ are allowed to depend on the whole population $X_t$ and individual location $x$.
-In Figure \ref{rescaled} we demonstrate this rescaling in discrete time for a population experiencing stabilizing selection and logistic growth.
-Since time is discretized, the process we simulate is formally a branching random walk.
-For further details on our simulation see SM \S\ref{numerical}.
-
-```{r, eval=F, echo=F}
-require(ggplot2)
-require(ggthemes)
-require(gridExtra)
-
-n_1 = read.csv("/home/bb/Gits/branching.brownian.motion.and.spde/n_1.csv")
-n_2 = read.csv("/home/bb/Gits/branching.brownian.motion.and.spde/n_2.csv")
-n_3 = read.csv("/home/bb/Gits/branching.brownian.motion.and.spde/n_3.csv")
-
-n1p = ggplot(n_1,aes(x=time,y=x))+geom_point(size=1.0,stroke=0,alpha=0.1)+theme_minimal()+ylab("Trait Value \n (n=1)")+xlab("")+xlim(0,50)+ylim(-5,5)+ggtitle("Rescalings of a branching random walk")+theme(plot.title = element_text(hjust = 0.5))
-n2p = ggplot(n_2,aes(x=time,y=x))+geom_point(size=0.5,stroke=0,alpha=0.1)+theme_minimal()+ylab("Trait Value \n (n=5)")+xlab("")+xlim(0,50)+ylim(-5,5)
-n3p = ggplot(n_3,aes(x=time,y=x))+geom_point(size=0.25,stroke=0,alpha=0.1)+theme_minimal()+ylab("Trait Value \n (n=20)")+xlab("Time")+xlim(0,50)+ylim(-5,5)
-
-rescaled_plots = grid.arrange(n1p,n2p,n3p,nrow=3)
-
-ggsave("~/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/rescaled_plots.png",rescaled_plots,device="png",width=6,height=6,units="in",dpi=350)
-```
-
-```{r, message=FALSE, echo=F, fig.align='center', fig.cap='\\label{rescaled}Rescaled sample paths of a branching random walk under stabilizing selection and logistic growth.
-The top plot displays a sample path without scaling $(n=1)$, the middle plot shows a sample path rescaled by $n=5$ and the bottom plot shows a sample path rescaled by $n=20$.'}
-knitr::include_graphics('rescaled_plots.png')
-```
-
-Interactions that manifest in the spatial movement can be used to model mutation bias and those manisfesting in the reproductive law can model density-dependent growth rates and frequency-dependent selection.
-Perkins [-@Perkins1992; -@edwinarendperkins1995] developed a theory of stochastic integration with respect to the so-called _Brownian trees_ to characterize interacting superprocesses and establish properties of existence and uniqueness.
-Li [-@zeng1998absolute] built directly off of the construction of Méléard and Roelly [-@meleard1992interacting; -@Mlard1993] to study properties of interacting superprocesses and, by assuming individual spatial movement occurs independently of location $x$ and the entire population $X_t$, showed the evolution of associated density processes can be described by SPDE.
-
-Recall $\nu(x,t)$ denotes the density of a superprocess, given it exists.
-Assuming the interactions manifest only in the reproductive law and that spatial movement follows Brownian motion with diffusion parameter $\sqrt\mu$ independent of both $X_t$ and $x$ and the growth rate is bounded both above and below, Li's [-@zeng1998absolute] result implies the interacting superprocess on one dimensional trait space has a density $\nu(x,t)$ which is non-negative, integrable, continuous in time and space and satisfies the SPDE
-\begin{equation}\label{SPDE}
-\frac{\partial}{\partial t}\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)+\sqrt{V\nu(x,t)}\dot W(x,t).
-\end{equation}
-As mentioned above, we refer to SPDE (\ref{SPDE}) as the Stochastic Asexual Gaussian allelic model with Abundance dynamics (abbreviated SAGA).
-Comparing equation (\ref{SPDE}) to equation (3.5) of Li [-@zeng1998absolute], our $m$ and $V$ correspond to Li's $b$ and $c$ respectively.
-Using techniques developed by Evans and Perkins [-@Evans1994], Champagnat, Ferrière and Méléard [-@Champagnat2006] generalized this result to the case when $m(h,x)$ is bounded above and is linear or more than linear in $h$.
-In particular, SAGA corresponds to a special case of equation (4.6) in Champagnat, Ferrière and Méléard [-@Champagnat2006] and is valid for the assumptions on $m$ made in \S\ref{deterministic}.
-
-With solutions to SAGA well defined, we can calculate the total mass process $N(t)$ using the mild solution of (\ref{SPDE}) with $f(x)\equiv1\in C_b^2(\mathbb{R})$ (the symbol "$\equiv$" means equal to for every $x$).
-That is,
-
-\begin{multline}
-N(t)=N(0)+\int_0^t\int_\mathbb{R}\nu(x,s)\left(m(\nu,x)\cdot1+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}1\right)+1\sqrt{V\nu(x,s)}\dot W(x,s)dsdx \\
-=N(0)+\int_0^t\bar m(s)N(s)dt+\int_0^t\sqrt{VN(s)}d\hat{\mathbf{W}}_s(\sqrt{\nu(x,s)}),
-\end{multline}
-where
-\begin{equation}
-\bar m(t)=\frac{1}{N(t)}\int_\mathbb{R}m(\nu,x)\nu(x,t)dx,
-\end{equation}
-and
-\begin{equation}
-\int_0^td\hat{\mathbf{W}}_s(\sqrt{\nu(x,s)})=\int_0^t\int_\mathbb{R}\frac{\sqrt{\nu(x,s)}}{\sqrt{\int_\mathbb{R}\nu(x,s)dx}}\dot W(x,s)dxds.
-\end{equation}
-Setting $W_1(t)=\hat{\mathbf{W}}_t(\sqrt{\nu(x,t)})$, we can use traditional stochastic differential notation to write
-\begin{equation}
-dN=\bar mNdt+\sqrt{VN}dW_1.
-\end{equation}
-To find the associated SDE for $\bar x(t)$ and $\sigma^2(t)$, we want to repeat the same approach for $f(x)=x,x^2$ and apply It&ocirc;'s lemma.
-However, for these cases $f\notin C^2_b(\mathbb{R})$ since $f$ will not be bounded.
-But if we can show $\int_\mathbb{R}(|x|+x^2+x^4)\nu(x,t)dx<+\infty$ for all $t>0$ given this condition is satisfied by $\nu(x,0)$, then we can apply the mild solution of (\ref{SPDE}) to derive SDE for $\bar x(t)$ and $\sigma^2(t)$.
-To illustrate, let us suppose this is the case.
-Setting $\tilde x(t)=\int_\mathbb{R}x\nu(x,t)dx$, we have
-\begin{equation}
-\tilde x(t)=\tilde x(0)+\int_0^t\int_\mathbb{R}\nu(x,s)m(\nu,x)x+x\sqrt{V\nu(x,s)}\dot W(x,s)dxds.
-\end{equation}
-Similarly, setting $\tilde{\tilde\sigma}^2(t)=\int_\mathbb{R}x^2\nu(x,t)dx$, we have
-\begin{equation}
-\tilde{\tilde\sigma}^2(t)=\tilde{\tilde\sigma}^2(0)+\int_0^t\int_\mathbb{R}\nu(x,s)\left(m(\nu,x)x^2+\mu\right)+x^2\sqrt{V\nu(x,s)}\dot W(x,s)dxds.
-\end{equation}
-Since $\bar x(t)=\tilde x(t)/N(t)$ and $\sigma^2(t)=\tilde{\tilde\sigma}^2(t)/N(t)-\bar x^2(t)$, we can use It&ocirc;'s lemma to derive SDE for $\bar x(t)$ and $\sigma^2(t)$, which we perform in SM \S\ref{SDE_DERIV}.
-We make no attempt in finding sufficient conditions to ensure $\int_\mathbb{R}(|x|+x^2+x^4)\nu(x,t)dx<+\infty$ and hence make no general assertions about the existence or uniqueness of $\bar x(t)$ or $\sigma^2(t)$.
-Regardless, we will later assume $\nu(x,t)$ can be approximated by a Gaussian curve in $x$ for all $t\geq0$.
-This assumption implies $\int_\mathbb{R}|x|^n\nu(x,t)dx<+\infty$ for all $n\in\mathbb{N}$ and for all $t\geq0$.
-
-<!--
-
-__Superprocesses without densities__
-
-The superprocesses we consider here allow function-valued representations (i.e., densities such as $\nu(x,t)$) only in the case of spatial dimension $d=1$.
-However, if we are willing to let go of function-valued representations, we are free to study superprocesses for greater spatial dimensions.
-Dawson [-@dawson1993measure] provides an excellent resource for this subject.
-Futhermore, the literature on these more general processes includes models of ecological interactions including competition.
-We summarize a few results important to our model of diffuse coevolution here.
-
-Denoting $C^k(\mathbb{R}^d)$ the set of functions on $\mathbb{R}^d$ with continuous derivatives up to order $k$ and $\Delta=\sum_{i=1}^d\frac{\partial^2}{\partial x_i^2}$ the Laplacian operator on $C^2(\mathbb{R}^d)$, we define the operator $A:C^2(\mathbb{R}^d)\to C^0(\mathbb{R}^d)$ by
-\begin{equation}
-A=\frac{\mu}{2}\Delta+r-\frac{a}{2}(\theta-x)^2,
-\end{equation}
-for fixed $r,\theta\in\mathbb{R}$ and $\mu,a>0$.
-
--->
-
-## Equations of evolutionary and demographic dynamics \label{equations}
-
-In SM \S\ref{SDE_DERIV} we show SDE for $N(t)$, $\bar x(t)$ and $\sigma^2(t)$ can be expressed as
-\begin{subequations}\label{gen_eqns}
-\begin{equation}\label{N}
-dN(t)=\bar m(t)N(t)dt+\sqrt{V N(t)}dW_1(t),
-\end{equation}
-\begin{equation}\label{xbar_gen}
-d\bar x(t)=\mathrm{Cov}_t\Big(x,m(\nu,x)\Big)dt+\sqrt{V\frac{\sigma^2(t)}{N(t)}}dW_2(t),
-\end{equation}
-\begin{equation}\label{sig2_gen}
-d\sigma^2(t)=\mathrm{Cov}_t\Big((x-\bar x(t))^2,m(\nu,x)\Big)dt+\left(\mu-V\frac{\sigma^2(t)}{N(t)}\right)dt+\sqrt{V\frac{\overline{(x-\bar x(t))^4}-\sigma^4(t)}{N(t)}}dW_3(t),
-\end{equation}
-\end{subequations}
-where $W_1$, $W_2$ and $W_3$ are standard Brownian motions.
-We note that conditions on the growth rate $m$ to guarantee existence and uniqueness of solutions to (\ref{xbar_gen}) and (\ref{sig2_gen}) have yet to be investigated.
-However, our results on the deterministic PDE suggest that $m(h,x)$ bounded above and differentiable in both arguments is sufficient.
-Dividing by $dt$ one can interpret equations (\ref{gen_eqns}) as if they are ordinary differential equations, but this not technically rigorous since Brownian motion is nowhere differentiable with respect to time.
-In SM \S\ref{SDE_DERIV} we show that in general $W_1$ is independent of both $W_2$ and $W_3$, but $W_2$ and $W_3$ covary.
-
-There is quite a bit we can learn from expressions (\ref{gen_eqns}).
-Firstly, setting $V=0$ recovers the deterministic dynamics derived in \S\ref{deterministic}.
-Alternatively, one can take $N(t)\to\infty$ to recover the deterministic dynamics for $\bar x(t)$ and $\sigma^2(t)$.
-Characteristically, we note the effect of demographic stochasticity on abundance grows with $\sqrt{N(t)}$.
-Hence, dividing by $N(t)$, we find the effects of demographic stochasticity on the per-capita growth rate diminish with increased abundance.
-Relating the response to demographic stochasticity derived here to the effect of random genetic drift derived in classic quantitative genetic theory, if we set $\sigma^2(t)=\sigma^2$ and $N(t)=N$ constant with respect to time, then integrating the stochastic term in equation (\ref{xbar_gen}) over a single unit of time returns a normally distributed random variable with mean zero and variance equal to $V\sigma^2/N$.
-In particular, assuming perfect inheritance, when reproductive variance is unity ($V=1$) this random variable coincides with the effect of random genetic drift on the change in mean trait over a single generation derived using sampling arguments [@Lande1976].
-There is also an interesting connection with classical population genetics.
-A fundamental result from early population genetic theory is the expected reduction in diversity due to the chance loss of alleles in finite populations [@Fisher1923; @Wright97].
-This expected reduction in diversity due to random genetic drift is captured by the third term in the deterministic component of expression (\ref{sig2_gen}), particularly $-V\sigma^2(t)/N(t)$.
-The component of SDE (\ref{sig2_gen}) describing random fluctuations in $\sigma^2(t)$ is more complicated and is proportional to the root of the difference between the centralized fourth moment of $p(x,t)$ and $\sigma^4(t)$.
-
-These expressions can be used to investigate the dynamics of the mean and variance for general $\nu(x,t)$, assuming $\int_\mathbb{R}(|x|+x^2+x^4)\nu(x,t)dx<+\infty$ for all $t\geq0$.
-However, in the next subsection we simplify these expressions by approximating $\nu(x,t)$ with a Gaussian curve.
-By assuming $\nu(x,t)$ is Gaussian for $t\geq0$, we guarantee the existence of $\bar x(t)$ and $\sigma^2(t)$ for all $t\geq0$.
-Furthermore, in SM \S\ref{SDE_DERIV} we show that under the Gaussian case $W_1,W_2$ and $W_3$ are independent.
-
-### Particular results assuming a Gaussian phenotypic distribution \label{particular}
-
-By approximating $\nu(x,t)$ with a Gaussian curve in $x$ for each $t\geq0$, expressions (\ref{gen_eqns}) transform into efficient tools for deriving the dynamics of populations given a fitness function $m(\nu,x)$.
-Gaussian phenotypic distributions are often obtained through Gaussian, exponential or weak selection approximations together with a simplified model of mutation and asexual reproduction or random mating [@pmid17248993; @Turelli1984; @Turelli1986; @Turelli2017; @9780471986539].
-Alternatively, it has been shown that a Gaussian distribution can provide a reasonable approximation even when selection is strong and non-Gaussian [@pmid7851785].
-Hence, given appropriate assumptions on selection, mutation and reproduction, the abundance density $\nu(x,t)$ can be approximated as a Gaussian curve in $x$ when the ratio $V/N$ is small (i.e., when the variance in reproductive output is much smaller than the population size).
-As with any diffusion approximation, this requires a sufficiently large abundance to accurately reflect the dynamics of populations.
-Hence, models developed in this framework are not suitable for studies involving very small population sizes.
-Allowing for these restrictions, we assume
-\begin{equation}
-\nu(x,t)=\frac{N(t)}{\sqrt{2\pi\sigma^2(t)}}\exp\left(-\frac{\big(x-\bar x(t)\big)^2}{2\sigma^2(t)}\right).
-\end{equation}
-Under this assumption we find in SM \S\ref{cov2deriv} the results (suppressing the dependency on $t$)
-\begin{equation}\label{covxm}
-\mathrm{Cov}(x,m)=\sigma^2\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right),
-\end{equation}
-\begin{equation}
-\mathrm{Cov}\Big((x-\bar x)^2,m\Big)=2\sigma^4\left(\frac{\partial\bar m}{\partial\sigma^2}-\overline{\frac{\partial m}{\partial\sigma^2}}\right)
-\end{equation}
-and $\overline{(x-\bar x)^4}=3\sigma^4$. In particular, these results imply
-\begin{subequations}\label{no_inher}
-\begin{equation}\label{xbar}
-d\bar x=\sigma^2\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right)dt+\sqrt{V\frac{\sigma^2}{N}}dW_2,
-\end{equation}
-\begin{equation}\label{G}
-d\sigma^2=2\sigma^4\left(\frac{\partial\bar m}{\partial\sigma^2}-\overline{\frac{\partial m}{\partial\sigma^2}}\right)dt +\left(\mu-V\frac{\sigma^2}{N}\right)dt+\sigma^2\sqrt{\frac{2V}{N}}dW_3.
-\end{equation}
-\end{subequations}
-These equations allow us to derive the response in trait mean and variance by taking derivatives of fitness, a much more straightforward operation than calculating a covariance for general phenotypic distributions.
-Note that in the above expressions, the partial derivatives of $\bar m$ represent frequency independent selection and the averaged partial derivatives of $m$ represent frequency dependent selection.
-This relationship has already been pointed out by Lande [-@Lande1976] for the evolution of the mean trait in discrete time, but here we see an analogous relationship holds in continuous time and also for the evolution of trait variance.
-
-In the next subsection we generalize this result to the case when traits are imperfectly inherited.
-In this case, the phenotypic variance $\sigma^2$ is replaced by a genetic variance $G$.
-This genetic variance represents the component of the variance in expressed traits $\sigma^2$ explained by additive effects of different alleles among genetic loci encoding for the focal phenotype [@joanroughgarden1979; @bulmer1980; @michaellynch1998].
-It is therefore fitting that $G$ is referred to as the additive genetic variance.
-
-### The evolution of additive genetic variance \label{inheritance}
-
-To model imperfect heritability we consider the relationship between expressed phenotypes $x\in\mathbb{R}$ and associated genetic values $g\in\mathbb{R}$ known as _breeding values_.
-The breeding value of an individual is the sum of additive effects of the alleles carried by the individual on its expressed trait.
-Since our derivations of evolutionary equations are based on branching processes that assume asexually reproducing populations (\S\ref{stochastic}), the additive genetic variance $G$ is just the variance of breeding values in a population.
-For a detailed treatment of breeding values and additive genetic variances, see Bulmer [-@bulmer1980] and Lynch and Walsh [-@michaellynch1998].
-
-Our treatment of the relationship between breeding values and expressed traits follows classical quantitative genetic assumptions such as those used in the seminal paper by Lande [-@Lande1975] to investigate the maintenance of genetic variation.
-In particular, we assume that the expressed trait for any given individual is independent of environmental conditions and normally distributed around their breeding value with variance $\eta$.
-Hence, $\sigma^2=G+\eta$.
-In the case that all of the effects of alleles on an expressed trait are additive, $\eta$ is known as the _variance of environmental deviation_ [@Lande1975; @michaellynch1998].
-For a given breeding value, we denote the probability density of a randomly drawn expressed trait by $\psi(x,g)$.
-Hence,
-\begin{equation}
-\psi(x,g)=\frac{1}{\sqrt{2\pi\eta}}\exp\left(-\frac{(x-g)^2}{2\eta}\right).
-\end{equation}
-To include this relationship in our framework, we write $\rho(g,t)$ as the abundance density of breeding values at time $t$ so that $\int_{-\infty}^{+\infty}\rho(g,t)dg=\int_{-\infty}^{+\infty}\nu(x,t)dx=N(t)$.
-We switch our focus from directly modelling the evolution of $\nu(x,t)$ to modelling the evolution of $\rho(g,t)$.
-Once $\rho(g,t)$ is determined, we can compute $\nu(x,t)$ via
-\begin{equation}\label{nu_from_rho}
-\nu(x,t)=\int_{-\infty}^{+\infty}\rho(g,t)\psi(x,g)dg.
-\end{equation}
-However, since selection acts on expressed phenotypes, we use the assumed relationship between breeding values and expressed traits to calculate the fitness of breeding values.
-To motivate the approach taken, consider the problem of inferring the breeding value of an individual given its expressed trait $x$.
-Denote $\mathfrak{g}$ a random variable representing the unkown breeding value.
-Under this model of inheritance we know $x$ is a random sample from a normal distribution with mean $\mathfrak{g}$ and variance $\eta$.
-Maximizing likelihood suggests $x$ is our best guess for $\mathfrak{g}$, but the actual value of $\mathfrak{g}$ is normally distributed around $x$ with the variance $\eta$.
-Hence, for fixed $x$, we obtain $\psi(x,g)$ as the probability density of $\mathfrak{g}$.
-Thus, the mean fitness of a breeding value $g$ across all individuals carrying $g$ can be written as
-\begin{equation}\label{m_rel}
-m^*(\rho,g)=\int_{-\infty}^{+\infty}m(\nu,x)\psi(x,g)dx.
-\end{equation}
-This is similar to the approach taken by Kimura and Crow [-@Kimura1978] to calculate the overall effects of selection for expressed characters onto the changes in the distribution of alleles encoding those characters.
-However, instead of focusing on the frequencies of alleles at particular loci, our results focus on the densities of breeding values.
-With the relationship between $m(\nu,x)$ and $m^*(\rho,g)$ established, we define the evolution of $\rho(g,t)$ by the SPDE
-\begin{equation}\label{rho_SPDE}
-\dot\rho(g,t)=m^*(\rho,g)\rho(g,t)+\frac{\mu}{2}\frac{\partial^2}{\partial^2 g}\rho(g,t)+\sqrt{V\rho(g,t)}\dot W(g,t).
-\end{equation}
-Equation (\ref{rho_SPDE}) is a stochastic generalization of the deterministic PDE (\ref{eq1}) from \S\ref{deterministic}, but describes the evolution of the distribution of breeding values instead of expressed characters.
-However, whether modelling expressed characters or breeding values, we refer to SPDE of the form (\ref{rho_SPDE}) as Stochastic Asexual Gaussian allelic models with Abundance dynamics (abbreviated SAGA).
-Replacing $\psi(x,g)$ with different probability densities produces different developmental models.
-This includes densities that may depend on environmental variables and hence model phenotypic plasticy.
-To leave room for the naming of other models where $\psi(x,g)$ is Gaussian, we refer to the model of development we have chosen as the _centered_ Gaussian developmental model.
-In \S\ref{wnc} we review the origins of equation (\ref{rho_SPDE}) and provide some theory to help make sense of it, particularly the term $\dot W$.
-
-Assuming $\rho(g,t)$ is Gaussian implies its mode coincides with $\bar x$.
-Furthermore, since $\sigma^2=G+\eta$, we can use equation (\ref{m_rel}) and the chain rule from calculus (see SM \S\ref{fit2fit}) to find
-\begin{subequations}
-\begin{equation}
-\frac{\partial\bar m}{\partial G}=\frac{\partial\bar m}{\partial\sigma^2}\frac{\partial\sigma^2}{\partial G}=\frac{\partial\bar m}{\partial\sigma^2},
-\end{equation}
-\begin{equation}
-\overline{\frac{\partial m}{\partial G}}=\overline{\frac{\partial m}{\partial\sigma^2}\frac{\partial\sigma^2}{\partial G}}=\overline{\frac{\partial m}{\partial\sigma^2}}.
-\end{equation}
-\end{subequations}
-Thus, equations (\ref{no_inher}) become
-\begin{subequations}\label{inher}
-\begin{equation}\label{xbarfinal}
-d\bar x=G\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right)dt+\sqrt{V\frac{G}{N}}dW_2,
-\end{equation}
-\begin{equation}\label{Gfinal}
-dG=2G^2\left(\frac{\partial\bar m}{\partial G}-\overline{\frac{\partial m}{\partial G}}\right)dt+\left(\mu-V\frac{G}{N}\right)dt+ G\sqrt{\frac{2V}{N}}dW_3.
-\end{equation}
-\end{subequations}
-
-From expressions (\ref{inher}) we see that, under this model of inheritance, focusing on additive genetic variance $G$ instead the variance in expressed traits $\sigma^2$ makes no structural changes to the basic equations describing the dynamics of populations.
-In the next section, we make use of these expressions to develop a model of diffuse coevolution in a guild of $S$ species competing along a resource continuum.
-
-# A model of diffuse coevolution \label{coev}
-
-## Formulation \label{form_coev}
-
-In this section we demonstrate the use of our framework by developing a model of diffuse coevolution across a guild of $S$ species whose interactions are mediated by resource competition along a single niche axis.
-Because our approach treats abundance dynamics and evolutionary dynamics simultaneously, this model allows us to investigate the relationship between selection gradients and competition coefficients, which we carry out in what follows.
-
-The dynamics of phenotypic distributions and abundances have been derived above and so the only task remaining is the formulation of a fitness function.
-Our approach mirrors closely the theory developed by MacArthur and Levins [-@Macarthur1967], Levins [-@9780691080628] and MacArthur [-@Arthur1969; -@MacArthur1970; -@9780691023823].
-The most significant difference, aside from allowing evolution to occur, is the treatment of resource quality, which we replace with a model of abiotic stabilizing selection.
-A derivation is provided in SM \S\ref{diffuse}.
-
-For species $i$ we inherit the above notation for trait value, distribution, average, variance, abundance, et cetera, except with an $i$ in the subscript.
-Real world examples of niche axes include the size of seeds consumed competing finch species and the date of activity in a season for pollinators competing for floral resources.
-For mathematical convenience, we model the axis of resources by the real line $\mathbb{R}$.
-The value of a resouce along this axis is denoted by the symbol $\zeta$.
-For an individual in species $i$, we assume the resource utilization curve $u_i$ can be written as
-
-\begin{equation}
-u_i(\zeta,x_i)=\frac{U_i}{\sqrt{2\pi w_i}}\exp\left(-\frac{(x_i-\zeta)^2}{2w_i}\right).
-\end{equation}
-
-We further assume the niche center $x_i$ is normally distributed among individuals in species $i$, but the niche breadth $w_i$ and total niche utilization $U_i$ are constant across individuals in species $i$ and therefore cannot evolve.
-Suppose $\theta_i\in\mathbb R$ is the optimal location along the niche axis for species $i$ such that, in the absence of competition, individuals leave on average $Q_i$ offspring when concentrated at $\theta_i$.
-We assume the expected reproductive output of an individual in species $i$ decreases as niche location $\zeta$ leaves the optimum $\theta_i$ at the rate $A_i>0$.
-Hence, abiotic stabilizing selection along the resource axis can be modelled by the curve
-
-\begin{equation}
-e_i(\zeta)=Q_i\exp\left(-\frac{A_i}{2}(\theta_i-\zeta)^2\right).
-\end{equation}
-
-The effect of abiotic stabilizing selection on the fitness for an individual of species $i$ with niche location $x_i$ is then given by
-
-\begin{equation}
-\int_{-\infty}^{+\infty}e_i(\zeta)u_i(\zeta,x_i)d\zeta=\frac{Q_iU_i}{\sqrt{A_iw_i+1}}\exp\left(-\frac{A_i}{2(A_iw_i+1)}(\theta_i-x_i)^2\right).
-\end{equation}
-
-To determine the potential for competition between individuals with niche locations $x_i$ and $x_j$, belonging to species $i$ and $j$ respectively, we compute the niche overlap
-
-\begin{equation}
-\mathcal O_{ij}(x_i-x_j)=\int_{-\infty}^{+\infty}u_i(\zeta,x_i)u_j(\zeta,x_j)d\zeta=\frac{U_iU_j}{\sqrt{2\pi(w_i+w_j)}}\exp\left(-\frac{(x_i-x_j)^2}{2(w_i+w_j)}\right).
-\end{equation}
-
-A notable criticism of using niche overlap to measure the intensity of competition points to cases where populations competiting on multiple niche axes exhibit overlap on at least one of the axes, but no overall niche overlap [@Holt1987].
-Thus niche overlap on lower-dimensional projections of some multivariate niche space does not imply the populations compete.
-To illustrate with a simple example, consider two populations competing for space on the plane $\mathbb{R}^2$.
-If the spatial distributions of the two populations overlap, then they will overlap on both spatial axes.
-However, if the populations do not overlap on at least one of the spatial axes, they will have no overall spatial overlap.
-Furthermore, even if the species overlap on both spatial axes, they need not have any overall spatial overlap.
-This final result corresponds to the fact that components of niche space do not necessarily interact multiplicatively to determine the consequences for the intensity of competition.
-In another component of Holt's [-@Holt1987] critique, an argument is made for the potential of competition occuring without any overlap in niche space.
-However, this argument is based on the practical difficulty of identifying every resource axis populations are competing on and how these axes interact to determine fitness consequences.
-Our model avoids these caveats by assuming competition only occurs along a single dimensional resource gradient.
-
-To map the degree of niche overlap to fitness, we assume competition between individuals with niche locations $x_i$ and $x_j$ decreases the expected reproductive output for the individual in species $i$ at the rate $c_i\mathcal{O}_{ij}(x_i-x_j)$ for some $c_i>0$.
-We refer to $c_i$ as the sensitivity to competition for species $i$.
-The term $c_i\mathcal{O}_{ij}(x_i-x_j)$ coincides with a special case of a term used to capture competition in Dawson's geostochastic logistic model, an SPDE model developed to study the combined effects of demographic stochasticity, spatial dispersion and locally finite carrying capacity [@Dawson1978].
-In relation to the example fitness function discussed in \S\ref{deterministic}, consider $\kappa(x_i-x_j)=\mathcal{O}_{ij}(x_i-x_j)$.
-In SM \S\ref{diffuse} we combine our treatment of resource competition with equations (\ref{N}), (\ref{xbarfinal}) and (\ref{Gfinal}) to find
-\begin{subequations}\label{comm_dynamics}
-\begin{multline}\label{N_comm}
-dN_i = \Bigg\{R_i-\frac{a_i}{2}\Big((\bar x_i-\theta_i)^2+G_i+\eta_i\Big) - c_i\sum_{j=1}^SN_jU_iU_j\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\Bigg\}N_idt + \sqrt{V_iN_i}dW_1,
-\end{multline}
-\begin{multline}\label{x_comm}
-d\bar x_i = \Bigg\{a_iG_i(\theta_i-\bar x_i)-c_iG_i\bigg(\sum_{j=1}^SN_jU_iU_jb_{ij}(\bar x_j-\bar x_i)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\bigg)\Bigg\}dt+\sqrt{V_i\frac{G_i}{N_i}}dW_2,
-\end{multline}
-\begin{multline}\label{G_comm}
-dG_i =  \Bigg\{c_i{G_i}^2\bigg(N_iU_i^2b_{ii}\sqrt\frac{b_{ii}}{2\pi}+\sum_{j=1}^SN_jU_iU_jb_{ij}\left(1-b_{ij}(\bar x_i-\bar x_j)^2\right)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\bigg) \\
-+\mu_i-a_i{G_i}^2-V_i\frac{G_i}{N_i}\Bigg\}dt+G_i\sqrt{\frac{2V_i}{N_i}}dW_3,
-\end{multline}
-\end{subequations}
-where
-\begin{subequations}
-\begin{align}
-R_i = & \ \ln \left(\frac{Q_iU_i}{\sqrt{1+A_iw_i}}\right), \\
-a_i = & \ \frac{A_i}{1+A_iw_i}, \\
-b_{ij}(t) = b_{ji}(t) = & \ \big(w_i+w_j+\eta_i+\eta_j+G_i(t)+G_j(t)\big)^{-1}, \\
-c_i \geq & \ 0.
-\end{align}
-\end{subequations}
-Despite the convoluted appearance of system (\ref{comm_dynamics}), there are some nice features that reflect biological reasoning.
-For example, the dynamics of abundance are just a generalization of Lotka-Volterra dynamics.
-In particular, the effect of competition with species $j$ on the fitness of species $i$ grows linearly with $N_j$.
-However, as biotic selection pushes $\bar x_i$ away from $\bar x_j$, the effect of competition with species $j$ on the fitness of species $i$ rapidly diminishes, reflecting a reduction in niche overlap.
-The divergence of $\bar x_i$ and $\bar x_j$ due to competition is referred to in the community ecology literature as character displacement [@Brown1956].
-We also see that the fitness of species $i$ drops quadratically with the difference between $\bar x_i$ and the abiotic optimum $\theta_i$.
-Hence, abiotic selection acts to pull $\bar x_i$ towards $\theta_i$.
-The response in mean trait $\bar x_i$ to natural selection is proportional to the amount of heritable variation in the population, represented by the additive genetic variance $G_i$.
-However, we have that $G_i$ is itself a dynamic quantity.
-Under our model, abiotic stabilizing selection erodes away heritable variation at a rate that is independent of both $N_i$ and $\bar x_i$.
-The effect of competition on $G_i$ is a bit more complicated.
-When $b_{ij}(\bar x_i-\bar x_j)^2<1$, competition with species $j$ acts as diversifying selection which tends to increase the amount of heritable variation.
-However, when $b_{ij}(\bar x_i-\bar x_j)^2>1$, competition with species $j$ acts as directional selection and reduces $G_i$.
-In the following subsections we demonstrate the behavior of system (\ref{comm_dynamics}) by plotting numerical solutions and investigate implications for the relationship between the strength of ecological interactions and selection.
-
-## Community dynamics \label{dynamics}
-
-For the sake of illustration we numerically integrated system (\ref{comm_dynamics}) for a richness of $S=100$ species.
-We assumed homogeneous model parameters across species in the community as summarized by Table \ref{par_vals}.
-We repeated numerical integration under the two scenarios of weak and strong competition.
-For the first scenario of weak competition we set $c=1.0\times10^{-7}$ and for the second scenario of strong competition we set $c=5.0\times10^{-6}$.
-With these two sets of model parameters, we simulated our model for $1000.0$ units of time.
-For both scenarios, we initialized the trait means to $\bar x_i=0.0$, additive genetic variances to $G_i=10.0$ and abundances to $N_i=1000.0$ for each $i=1,\dots,S$.
-
-Temporal dynamics for each scenario are provided in Figure \ref{temporal}.
-This figure suggests weaker competition leads to smoother dynamics and a higher degree of organization within the community.
-Considering expression (\ref{N_comm}) we note that, all else equal, relaxed competition allows for larger growth rates which promote greater abundances.
-From (\ref{N_comm}) we also note that the per-capita effects on demographic stochasticity diminish with abundance.
-To see this, divide both sides by $N_i$.
-Inspecting expressions (\ref{x_comm}) and (\ref{G_comm}), we see that larger abundances also erode the effects of demographic stochasticity on the evolution of mean trait and additive genetic variance.
-These effects were already noted in \S\ref{equations}, and thus are not a consequence of our model of coevolution per-se, but we revisit them here since Figure \ref{temporal} demonstrates the importance of demographic stochasticity in structuring ecological communities even when populations are very large.
-Hence, contrary to the common assumption that stochastic effects can be ignored for large populations, we find that minute asymmetries generated by demographic stochasticity remain significant drivers of community structure.
-In particular, we initialized the species with identical state variables and model parameters, but found an enourmous amount of asymmetry and even some potential phase changes.
-In the following two paragraphs we describe the natural history of the community as illustrated in Figure \ref{temporal}.
-
-We begin by describing the weak competition scenario.
-After about $125.0$ units of time, the community appears to have shaken off the initial conditions and entered into a qualitatively distinct phase of dynamics.
-Aside from a few outliers, most of the species remain clustered together in their state variables.
-This lasts for approximately $375.0$ units of time until, at around time $500.0$, a drastic change occurs.
-At this moment the tightly packed cluster of species begins to fan out in all three state variables.
-Simultaneously, we observe large a shift in mean traits for higher values and in additive genetic variances for lower values.
-Upon inspecting our calculations, we diagnose the reason for this shift.
-The outlier species that were initially pushed away from the common abiotic optimum ($0.0$ in this case) evolved a significant reduction in the amount of heritable variation ($\approx60\%$) due to directional selection induced by competition.
-This reduction in heritable variation slowed adaptation, causing these species to linger on the outskirts of niche space, some longer than others.
-In the meantime the rest of the community, being tightly packed, experienced greater competition which led to diminished abundances for these species and caused some members of the core group to veer away from the abiotic optimum.
-The reduced abundances of the core group led to reduced competition overall.
-As a result, the outlier populations were given a slight increase in growth rate, enough to allow them to increase their abundances orders of magnitude higher than the species in the core group and giving them more weight in driving the evolution of other species.
-Many of these heavy-hitting outlier species had already been maintaining negative mean traits, but around time $500.0$ the high abundance species with positive mean traits began to experience enough intraspecific competition to override interspecific competition.
-This generated a net selection gradient and associated evolutionary response towards the abiotic optimum.
-The sudden imbalance of these high abundance species effectively induced a single large competitive exclusion event pushing the majority of the community far away from the abiotic optimum.
-After this shift the cluster began to slowly bloom in all three state variables as species took advantage of novel asymmetries in their competitive abilities mediated by a new distribution of mean trait values across the community.
-About $125.0$ units of time later, the community reached a qualitatively new phase of dynamics.
-If we kept running the numerical integrator, we would continue to see similar drama unfolding over and over again as minute stochastic changes contribute to asymmetries which slowly build into drastic shifts.
-
-The strong competition scenario is not quite as showy.
-Although the dynamics of trait means and variances tend to be far more stochastic than in the weak competition scenario, the community overall appears to quickly reach some statistical equilibrium and remain there.
-However, the abundances across all species in the community are very low due to competition sensitivies being an order of magnitude higher than in the weak competition case.
-Most of the species maintain abundances greater than $1000.0$, but we found one species that dropped to an abundance of about $50.0$.
-If we let the numerical integrator run long enough in this case, we will likely see many of the species go extinct.
-
-Finding ways to interpret simulated dynamics provides a useful arena to exercise biological reasoning.
-However, it does not fulfill our desire to quantify the patterns and processes present in competing communities.
-In the next subsection we take a step in this direction by using our model to derive formula for selection gradients and competition coefficients.
-To investigate their relationship, we calculate their covariances using simplifying assumptions on species abundances and intraspecific trait variances.
-We then investigate how these covariances change with the ratio of variance of interspecific mean traits to variance of intraspecific individual traits and use a numerical approach to investigate correlations between the strength of pairwise coevolution and competition coefficients.
-\begin{table}
-\centering\caption{Values of model parameters used for numerical integration.}\vspace{0.2cm}
-\begin{tabular}{l|l|l}
-Parameter     & Description & Value                                           \\ \hline
-$R$         & innate growth rate, see \S\ref{ecoevo}  & $1.0$      \\
-$\theta$    & abiotic optimum & $0.0$                              \\
-$a$         & strength of abiotic selection & $0.01$               \\
-$c$         & sensitivity to competition & $\{1.0\times10^{-7},5.0\times10^{-6}\}$ \\
-$w$         & niche breadth & $0.1$                               \\
-$U$         & total niche use & $1.0$                              \\
-$\eta$      & segregation variance & $1.0$                         \\
-$\mu$       & mutation rate & $1.0\times10^{-7}$                             \\
-$V$ & variance of reproductive output & $5.0$
-\end{tabular}\label{par_vals}
-\end{table}
-
-```{r, eval=F, echo=F}
-require("ggplot2")
-require("gridExtra")
-require("reshape")
-
-####################
-#                  #
-# weak competition #
-#                  #
-####################
-
-smpl_wc = read.csv("/home/bb/Gits/branching.brownian.motion.and.spde/sample_path_wc.csv")
-
-x_wc = ggplot(smpl_wc,aes(x=time,y=x,group=spp))+theme_minimal()+xlab("")+ylab("Mean Trait")+
-  ylim(-12,12)+geom_line(show.legend=F, lwd=0.1)+
-  ggtitle("Weak Competition")+theme(plot.title = element_text(hjust = 0.5))
-
-G_wc = ggplot(smpl_wc,aes(x=time,y=G,group=spp))+theme_minimal()+xlab("")+ylab("Additive Genetic \n Variance")+
-  ylim(0,45)+geom_line(show.legend=F, lwd=0.1)
-
-N_wc = ggplot(smpl_wc,aes(x=time,y=N,group=spp))+theme_minimal()+xlab("time")+ylab("Abundance")+
-  scale_y_log10(limits=c(50,1e8))+geom_line(show.legend=F, lwd=0.1)
-
-smpl_plt_wc = grid.arrange(x_wc,G_wc,N_wc,nrow=3)
-
-#
-# calculating strength of coevolution
-#
-
-tmax = max(smpl_wc$time)
-
-final_comm = subset(smpl_wc,time==tmax)
-
-x = final_comm$x
-G = final_comm$G
-N = final_comm$N
-
-w = 0.1
-c = 1e-7
-eta = 1.0
-R = 1.0
-a = 1e-2
-th = 0
-S = length(x)
-
-b = matrix(0,nc=S,nr=S)
-C = matrix(0,nc=S,nr=S)
-alpha = matrix(0,nc=S,nr=S)
-beta = matrix(0,nc=S,nr=S)
-gamma = matrix(0,nc=S,nr=S)
-for( i in 1:S){
-  for( j in 1:S ){
-
-    r = R - a*( (x[i]-th)^2 + G[i] + eta )/2
-
-    b[i,j] = 1/( G[i]+G[j]+2*(w+eta) )
-
-    alpha[i,j] = c*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2) / r
-
-    beta[i,j] = c*N[j]*b[i,j]*(x[i]-x[j])*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2)
-
-    gamma[i,j] = c*N[j]*b[i,j]*(1 - b[i,j]*(x[i]-x[j])^2)*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2)
-
-    C[i,j] = abs(beta[i,j]) + abs(gamma[i,j])
-
-  }
-}
-
-Csp = C # defines a matrix that ignores intraspecific interactions
-for( i in 1:S ){
-  Csp[i,i] = 0
-}
-
-Cvec = NULL
-for(i in 1:S){
-  for(j in i:S)
-    if( i!=j ) Cvec = c(Cvec,C[i,j])
-}
-
-#
-# distribution of coevolution
-#
-
-Cdf = data.frame(val=Cvec)
-
-WC_co <- ggplot(Cdf, aes(x=val)) + geom_histogram(aes(y=..count../sum(..count..)), fill='black', alpha=1.0, bins=50) + 
-  theme_minimal() + xlab("Pairwise Coevolution") + ylab("Relative Frequency") + 
-  scale_x_log10(limits=c(1e-7,1e-1))
-
-#
-# building a network
-#
-require(igraph)
-require(plotwidgets)
-require(network)
-require(GGally)
-
-asp = alpha # defines a matrix that ignores intraspecific interactions
-for( i in 1:S ){
-  asp[i,i] = 0
-}
-
-avec = NULL
-for(i in 1:S){
-  for(j in i:S)
-    if( i!=j ) avec = c(avec,alpha[i,j])
-}
-
-# bivariate plot
-Cadf_WC = data.frame(Coev=Cvec,Comp=avec, Which="WC")
-Ca_WC_pl = ggplot(Cadf_WC, aes(Coev,Comp)) + geom_bin2d(bins=20, show.legend = FALSE) + 
-  geom_density_2d(col='white') + scale_fill_gradient(low="black",high="grey",trans="log") +
-  scale_x_log10(limits=c(5e-5,1e-2)) + scale_y_log10(limits=c(2e-9,2e-8)) + 
-  theme_minimal() + xlab("Pairwise Coevolution") + ylab("Competition Coefficient") + 
-  ggtitle("Weak \n Competition") + theme(plot.title = element_text(hjust = 0.5))
-
-network <- graph_from_adjacency_matrix(ceiling(asp/max(asp)))
-
-CDF <- function(x,X){
-  L <- length(X)
-  k <- 0
-  for(i in 1:L){
-    if(x >= X[i]) k <- k+1
-  }
-  k/L
-}
-
-resc_wts <- NULL
-for(i in 1:length(avec)){
-  resc_wts <- c(resc_wts, CDF(avec[i],avec))
-}
-
-colors <- colorRamp(c("white","black"))
-
-edge_colors <- NULL
-for(i in 1:length(avec)){
-  edge_colors <- c(edge_colors,rgb2col(t(colors(resc_wts[i]))))
-}
-
-E(network)$width <- resc_wts
-E(network)$weights <- resc_wts
-E(network)$arrow.size <- 0.01
-E(network)$color <- edge_colors
-V(network)$size <- N/sum(N)
-V(network)$frame.color <- "white"
-V(network)$color <- "black"
-V(network)$label <- ""
-# l <- layout_in_circle(network)
-l <- layout_with_fr(network)
-
-WC_net <- network(asp, directed = F)
-
-netWC <- ggnet2(WC_net, mode = "circle", edge.alpha=.1, edge.color = edge_colors, edge.size = 2*min(1/avec)*avec, size = N/sum(N), max_size = 3, node.color = 'black')+
-  ggtitle("Weak \n Competition")+theme(plot.title = element_text(hjust = 0.5), legend.position = "none")
-
-######################
-#                    #
-# strong competition #
-#                    #
-######################
-
-smpl_mc = read.csv("/home/bb/Gits/branching.brownian.motion.and.spde/sample_path_mc.csv")
-
-x_mc = ggplot(smpl_mc,aes(x=time,y=x,group=spp))+theme_bw()+theme_minimal()+xlab("")+ylab("")+
-  ylim(-12,12)+geom_line(show.legend=F, lwd=0.1)+ggtitle("Strong Competition")+theme(plot.title = element_text(hjust = 0.5))
-
-G_mc = ggplot(smpl_mc,aes(x=time,y=G,group=spp))+theme_bw()+theme_minimal()+xlab("")+ylab("")+
-  ylim(0,45)+geom_line(show.legend=F, lwd=0.1)
-
-N_mc = ggplot(smpl_mc,aes(x=time,y=N,group=spp))+theme_bw()+theme_minimal()+xlab("time")+ylab("")+
-  scale_y_log10(limits=c(50,1e8))+geom_line(show.legend=F, lwd=0.1)
-
-smpl_plt_mc = grid.arrange(x_mc,G_mc,N_mc,nrow=3)
-
-community_plot = grid.arrange(smpl_plt_wc,smpl_plt_mc,nrow=1)
-
-ggsave("~/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/community_dynamics.png",community_plot,width=6.5,height=5,dpi=320)
-
-#
-# calculating strength of coevolution
-#
-
-tmax = max(smpl_mc$time)
-
-final_comm = subset(smpl_mc,time==tmax)
-
-x = final_comm$x
-G = final_comm$G
-N = final_comm$N
-
-w = 0.1
-c = 1e-7
-eta = 1.0
-R = 1.0
-a = 1e-2
-th = 0
-S = length(x)
-
-b = matrix(0,nc=S,nr=S)
-C = matrix(0,nc=S,nr=S)
-alpha = matrix(0,nc=S,nr=S)
-beta = matrix(0,nc=S,nr=S)
-gamma = matrix(0,nc=S,nr=S)
-for( i in 1:S){
-  for( j in 1:S ){
-
-    r = R - a*( (x[i]-th)^2 + G[i] + eta )/2
-
-    b[i,j] = 1/( G[i]+G[j]+2*(w+eta) )
-
-    alpha[i,j] = c*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2) / r
-
-    beta[i,j] = c*N[j]*b[i,j]*(x[i]-x[j])*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2)
-
-    gamma[i,j] = c*N[j]*b[i,j]*(1 - b[i,j]*(x[i]-x[j])^2)*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2)
-
-    C[i,j] = abs(beta[i,j]) + abs(gamma[i,j])
-
-  }
-}
-
-Csp = C # defines a matrix that ignores intraspecific interactions
-for( i in 1:S ){
-  Csp[i,i] = 0
-}
-
-Cvec = NULL
-for(i in 1:S){
-  for(j in i:S)
-    if( i!=j ) Cvec = c(Cvec,C[i,j])
-}
-
-#
-# distribution of coevolution
-#
-
-Cdf = data.frame(val=Cvec)
-
-MC_co <- ggplot(Cdf, aes(x=val)) + geom_histogram(aes(y=..count../sum(..count..)), fill='black', alpha=1.0, bins=50) + 
-  theme_minimal() + xlab("Pairwise Coevolution") + ylab("") + 
-  scale_x_log10(limits=c(1e-7,1e-1))
-
-#
-# building a network
-#
-require(igraph)
-require(plotwidgets)
-require(network)
-require(GGally)
-
-asp = alpha # defines a matrix that ignores intraspecific interactions
-for( i in 1:S ){
-  asp[i,i] = 0
-}
-
-avec = NULL
-for(i in 1:S){
-  for(j in i:S)
-    if( i!=j ) avec = c(avec,alpha[i,j])
-}
-
-# bivariate plot
-Cadf_SC = data.frame(Coev=Cvec,Comp=avec, Which="SC")
-Ca_SC_pl = ggplot(Cadf_SC, aes(Coev,Comp)) + geom_bin2d(bins=20, show.legend = FALSE) + 
-  geom_density_2d(col='white') + scale_fill_gradient(low="black",high="grey",trans="log") +
-  scale_x_log10(limits=c(5e-7,2e-4)) + scale_y_log10(limits=c(3e-9,1e-8)) + 
-  theme_minimal() + xlab("Pairwise Coevolution") + ylab("") + ggtitle("Strong \n Competition") +
-  theme(plot.title = element_text(hjust = 0.5))
-
-bivplot = grid.arrange(Ca_WC_pl,Ca_SC_pl,nrow=1)
-
-Cadf = rbind(Cadf_WC,Cadf_SC)
-head(Cadf)
-
-on_pl = ggplot(Cadf_WC, aes(Coev,Comp)) + geom_density_2d(col="black") +
-  geom_density_2d(data=Cadf_SC,col="red") +
-  scale_x_log10(limits=c(5e-7,1e-2)) + scale_y_log10(limits=c(2.5e-9,1.5e-8)) + 
-  theme_minimal() + xlab("Pairwise Coevolution") + ylab("Competition Coefficient")
-
-ggsave("~/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/on_pl.png",on_pl,device="png",width=3.5,height=3.5,units="in",dpi=350)
-
-network <- graph_from_adjacency_matrix(ceiling(asp/max(asp)))
-
-CDF <- function(x,X){
-  L <- length(X)
-  k <- 0
-  for(i in 1:L){
-    if(x >= X[i]) k <- k+1
-  }
-  k/L
-}
-
-resc_wts <- NULL
-for(i in 1:length(avec)){
-  resc_wts <- c(resc_wts, CDF(avec[i],avec))
-}
-
-colors <- colorRamp(c("white","black"))
-
-edge_colors <- NULL
-for(i in 1:length(avec)){
-  edge_colors <- c(edge_colors,rgb2col(t(colors(resc_wts[i]))))
-}
-
-E(network)$width <- resc_wts
-E(network)$weights <- resc_wts
-E(network)$arrow.size <- 0.01
-E(network)$color <- edge_colors
-V(network)$size <- N/sum(N)
-V(network)$frame.color <- "white"
-V(network)$color <- "black"
-V(network)$label <- ""
-# l <- layout_in_circle(network)
-l <- layout_with_fr(network)
-
-MC_net <- network(asp, directed = F)
-
-netMC <- ggnet2(MC_net, mode = "circle", edge.alpha=.1, edge.color = edge_colors, edge.size = 2*min(1/avec)*avec, size = N/sum(N), max_size = 3, node.color = 'black')+
-  ggtitle("Strong \n Competition")+theme(plot.title = element_text(hjust = 0.5), legend.position = "none")
-
-coevs <- grid.arrange(netWC, netMC,
-                      WC_co, MC_co, nrow=2)
-
-ggsave("~/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/interaction_networks.png",coevs,device="png",width=6,height=6,units="in",dpi=350)
-
-```
-
-```{r, message=FALSE, echo=F, fig.align='center', out.width='100%', fig.cap='\\label{temporal}Temporal dynamics of mean trait (top), additive genetic variance (middle) and abundance (bottom) for the scenario of weak competition (left) and strong competition (right).'}
-knitr::include_graphics('community_dynamics.jpg')
-```
-
-## The relation between the strength of ecological interactions and coevolution \label{ecoevo}
-
-Relating our treatment of resource competition to modern coexistence theory [@Chesson2000], the absolute competition coefficient $\alpha_{ij}$ becomes a dynamical quantity that can be written as
-\begin{equation}
-\alpha_{ij}(t)=\frac{c_i}{r_i(t)}\int_{-\infty}^{+\infty}\int_{-\infty}^{+\infty}p_i(x,t)p_j(y,t)\mathcal O_{ij}(x,y) dxdy =\frac{c_iU_iU_j}{r_i(t)}\sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right),
-\end{equation}
-where
-\begin{equation}
-r_i(t)=R_i-\frac{a_i}{2}\Big((\bar x_i(t)-\theta_i)^2+G_i(t)+\eta_i\Big).
-\end{equation}
-Hence, $dN_i(t)$ can be expressed as
-\begin{equation}
-dN_i(t)=r_i(t)\left(1-\sum_{j=1}^S\alpha_{ij}(t)N_j(t)\right)N_i(t)dt+\sqrt{V_iN_i(t)}dW_1(t).
-\end{equation}
-Note that although $r_i(t)$ is referred to in the coexistence literature as the intrinsic growth rate of the population, $R_i$ is a deeper intrinsic quantity.
-For now we refer to $R_i$ as the _innate_ growth rate.
-Previous work has shown the importance of demographic stochasticity [@Turelli1980; @Schreiber2017] and evolutionary dynamics [@Case1986; @Schreiber2018] in determining coexistence of competing species.
-However, these studies consider evolution and demographic stochasticity separately.
-To our knowledge, there are no rigorous theoretical investigations of coexistence that account for evolutionary dynamics of quantitative characters and demographic stochasticity simultaneously and, in particular, no studies combining these processes that also include the evolution of additive genetic variance.
-With the connection outlined in this section formally established, researchers may pursue a postmodern coexistence theory that naturally includes both the evolutionary dynamics of quantitative characters and the effects of demographic stochasticity in a simple synthetic framework.
-
-In SM \S\ref{diffuse} we show that the standardized directional selection gradient [sensu @Lande1983] induced by species $j$ on species $i$ can be computed as
-\begin{equation}
-\beta_{ij}(t)=c_iU_iU_jN_j(t)b_{ij}(t)\big(\bar x_i(t)-\bar x_j(t)\big) \sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right).
-\end{equation}
-Our notation differs from Lande and Arnold [-@Lande1983] in that subscripts here denote species instead of components of multivariate traits and we drop the prime that distinguishes between selection gradients and standardized selection gradients.
-
-__Metric of pairwise coevolution__
-
-Below we investigate the correspondence of interaction intensity and coevolutionary change.
-However, we can already identify one major discrepancy; $\alpha_{ij}$ is maximized when $\bar x_i=\bar x_j$, but $\beta_{ij}=0$ under the same condition.
-We therefore include in our metric of selection the standardized stabilizing selection gradient $\gamma$ which measures the effect of stabilizing or disruptive selection on phenotypic variance [@Lande1983].
-In SM \S\ref{diffuse} we show that the standardized stabilizing selection gradient induced by species $j$ on species $i$ can be computed as
-\begin{equation}
-\gamma_{ij}(t)=c_iU_iU_jN_j(t)b_{ij}(t)\left(1-b_{ij}(t)\big(\bar x_i(t)-\bar x_j(t)\big)^2\right) \sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right).
-\end{equation}
-To measure evolutionary change in species $i$ induced by species $j$, we form the metric $\Psi_{ij}=|\beta_{ij}|+|\gamma_{ij}|$ and define the strength of coevolution as $\mathfrak{C}_{ij}=\Psi_{ij}\Psi_{ji}$.
-Figure \ref{net} displays the joint distribution of $\mathfrak{C}_{ij}$ and the product of competition coefficients $\alpha_{ij}\alpha_{ji}$ for two simulated communities, both with richness $S=1000$.
-The solid contour represents the case of relatively high sensitivity to competition ($c=5\times10^{-7}$) and the dashed contour represents the case of relatively weak sensitivity to competition ($c=1\times10^{-8}$).
-In both cases we see that $\mathfrak{C}_{ij}$ and $\alpha_{ij}\alpha_{ji}$ are, essentially, unrelated.
-However, this depiction is only representative of a specific set of parameters.
-Next, we provide analytical approximations of the covariance between selection gradients and competition coefficients and a numerical estimate for the relationship between pairwise coevolution and competition coefficients for a range of parameters.
-
-```{r, eval=F, echo=F}
-require("ggplot2")
-require("gridExtra")
-require("reshape")
-
-####################
-#                  #
-# weak competition #
-#                  #
-####################
-
-smpl_wc1e3 = read.csv("/home/bb/Gits/branching.brownian.motion.and.spde/sample_path_wc1e3.csv")
-
-#
-# calculating strength of coevolution
-#
-
-x = smpl_wc1e3$x
-G = smpl_wc1e3$G
-N = smpl_wc1e3$N
-
-w = 0.1
-c = 1e-8
-eta = 1.0
-R = 1.0
-a = 1e-2
-th = 0
-S = length(x)
-
-b = matrix(0,nc=S,nr=S)
-C = matrix(0,nc=S,nr=S)
-alpha = matrix(0,nc=S,nr=S)
-beta = matrix(0,nc=S,nr=S)
-gamma = matrix(0,nc=S,nr=S)
-for( i in 1:S){
-  for( j in 1:S ){
-
-    r = R - a*( (x[i]-th)^2 + G[i] + eta )/2
-
-    b[i,j] = 1/( G[i]+G[j]+2*(w+eta) )
-
-    alpha[i,j] = c*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2) / r
-
-    beta[i,j] = c*N[j]*b[i,j]*(x[i]-x[j])*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2)
-
-    gamma[i,j] = c*N[j]*b[i,j]*(1 - b[i,j]*(x[i]-x[j])^2)*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2)
-
-    C[i,j] = abs(beta[i,j]) + abs(gamma[i,j])
-
-  }
-}
-
-Cvec = NULL
-for(i in 1:S){
-  for(j in i:S)
-    if( i!=j ) Cvec = c(Cvec,C[i,j])
-}
-
-avec = NULL
-for(i in 1:S){
-  for(j in i:S)
-    if( i!=j ) avec = c(avec,alpha[i,j])
-}
-
-Cadf_WC = data.frame(Coev=Cvec,Comp=avec, Which="WC")
-
-######################
-#                    #
-# strong competition #
-#                    #
-######################
-
-smpl_mc1e3 = read.csv("/home/bb/Gits/branching.brownian.motion.and.spde/sample_path_mc1e3.csv")
-
-#
-# calculating strength of coevolution
-#
-
-x = smpl_mc1e3$x
-G = smpl_mc1e3$G
-N = smpl_mc1e3$N
-
-w = 0.1
-c = 5e-7
-eta = 1.0
-R = 1.0
-a = 1e-2
-th = 0
-S = length(x)
-
-b = matrix(0,nc=S,nr=S)
-C = matrix(0,nc=S,nr=S)
-alpha = matrix(0,nc=S,nr=S)
-beta = matrix(0,nc=S,nr=S)
-gamma = matrix(0,nc=S,nr=S)
-for( i in 1:S){
-  for( j in 1:S ){
-
-    r = R - a*( (x[i]-th)^2 + G[i] + eta )/2
-
-    b[i,j] = 1/( G[i]+G[j]+2*(w+eta) )
-
-    alpha[i,j] = c*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2) / r
-
-    beta[i,j] = c*N[j]*b[i,j]*(x[i]-x[j])*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2)
-
-    gamma[i,j] = c*N[j]*b[i,j]*(1 - b[i,j]*(x[i]-x[j])^2)*sqrt(b[i,j]/(2*pi))*exp(-b[i,j]*(x[i]-x[j])^2/2)
-
-    C[i,j] = abs(beta[i,j]) + abs(gamma[i,j])
-
-  }
-}
-
-Cvec = NULL
-for(i in 1:S){
-  for(j in i:S)
-    if( i!=j ) Cvec = c(Cvec,C[i,j])
-}
-
-avec = NULL
-for(i in 1:S){
-  for(j in i:S)
-    if( i!=j ) avec = c(avec,alpha[i,j])
-}
-
-Cadf_SC = data.frame(Coev=Cvec,Comp=avec, Which="SC")
-
-# bivariate plot
-Cadf = rbind(Cadf_WC,Cadf_SC)
-
-on_pl = ggplot(Cadf_SC, aes(Coev,Comp)) + 
-  scale_x_log10(limits=c(1e-5,2e-4)) + 
-  geom_density_2d(col="black",n=200) +
-  geom_density_2d(data=Cadf_WC, linetype="dashed", col="black") +
-  theme_minimal() + xlab("Pairwise Coevolution") + ylab("")
-
-on_pl2 = on_pl + facet_grid(rows=vars(Which), scales="free_y") + 
-  ylab("Competition Coefficient") + theme(strip.background = element_blank(), strip.text.y = element_blank())
-
-ggsave("~/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/on_pl.png",on_pl2,device="png",width=3.5,height=3.5,units="in")
-```
-
-```{r, echo=F, fig.align='center', fig.cap='\\label{net}Bivariate distributions of competition coffecients (y-axis) and pairwise coevolution (x-axis) under the scenarios of weak competition (dashed line) and strong competition (solid line) after simulating for $t=1000.0$ units of time.
-Simulations were ran for $S=1000$ species.
-Parameters are the same as in Table 2, except we used $c=1\\times 10^{-8}$ for weak competition and $c=5\\times 10^{-7}$ for strong competition in order to account for increased species richness.'}
-knitr::include_graphics('on_pl.png')
-```
-
-__Covariance of selection and competition as a function of diversity__
-
-We now make use of the expressions derived for competition coeffecients and selection gradients to investigate their relationship.
-As a first pass, let us assume all model parameters are equivalent across species and that each species has the same abundance and trait variance.
-Let us further assume that richness $S$ is large and the distribution of mean trait values is normal with mean $\bar{\bar x}$, variance $V_{\bar X}$ and density $f_{\bar X}$.
-Such assumptions are typical when deriving analytical results in the field of theoretical coevolutionary community ecology [@Nuismer2012; @Nuismer2018].
-If $\bar{\bar x}$ is near $\theta$ and $V_{\bar X}$ is much smaller than $|2\frac{R}{a}-G-\eta|$, then we may approximate $r_i$ with
-\begin{equation}
-\bar r=\int_{-\infty}^{+\infty}\left(R-\frac{a}{2}\Big((\bar x-\theta)^2+G+\eta\Big)\right)f_{\bar X}(\bar x)d\bar x
-=R-\frac{a}{2}\Big((\bar{\bar x}-\theta)^2+V_{\bar X}+G+\eta\Big).
-\end{equation}
-In SM \S\ref{coef_grad_moms} we use these assumptions to calculate the first and second order moments describing the joint distribution of competition coefficients and selection gradients across the community.
-We find that the covariance between linear selection gradients and competition coeffecients are zero due to the symmetry implied by our assumptions.
-However, setting $\alpha(\bar x_i,\bar x_j)=\alpha_{ij}$, $\beta(\bar x_i,\bar x_j)=\beta_{ij}$ and $\gamma(\bar x_i,\bar x_j)=\gamma_{ij}$, the covariances between the magnitude of linear selection gradients and competition coefficients and between stabilizing selection gradients and competition coefficients can be written as
-\begin{subequations}\label{cov_eqns}
-\begin{equation}\label{cov_alpha_beta}
-\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)=\frac{2c^2b^2U^4N}{\pi\bar r}\sqrt\frac{V_{\bar X}}{2\pi}\left(\frac{1}{(1+8bV_{\bar X})^{3/4}}-\frac{1}{(1+4bV_{\bar X})^{3/4}(1+2bV_{\bar X})^{1/2}}\right),
-\end{equation}
-\begin{equation}\label{cov_alpha_gamma}
-\mathrm{Cov}_{f_{\bar X}}(\alpha,\gamma)=\frac{c^2b^2U^4N}{2\pi\bar r}(1-2bV_{\bar X})\left(\frac{1}{\sqrt{1+4bV_{\bar X}}}-\frac{1}{1+2bV_{\bar X}}\right).
-\end{equation}
-\end{subequations}
-```{r, eval=F, echo=F}
-c=1e-4
-b=1e-1
-r=0.1
-N=1e10
-Cov_ab <- function(x){
-  # x <- 2*b*V_{\bar X}
-  val = 2*c^2*b^2*N / (pi*r)
-  val = val * sqrt(x/(2*b*pi))
-  val2 = (1+8*x/2)^(-0.75)
-  val3 = (1+4*x/2)^(-0.75) / sqrt(1+2*x/2)
-  val = val * (val2 - val3)
-  return(val)  
-}
-Cov_ag <- function(x){
-  # x <- 2*b*V_{\bar X}
-  val = 2*c^2*b^2*N / (2*pi*r)
-  val = val * (1-2*x/2)
-  val2 = 1 / sqrt(1+4*x/2)
-  val3 = 1 / (1+2*x/2)
-  val = val * (val2 - val3)
-  return(val)  
-}
-require(ggplot2)
-require(ggthemes)
-require(gridExtra)
-dummy <- ggplot(data = data.frame(x = 0), mapping = aes(x = x))
-cov_ab_p <- dummy + stat_function(fun = Cov_ab) + ylim(-0.4,0.4) + xlim(0,2) + theme_minimal() + xlab("Approximate ratio of variance among \n species to variance within species") + ylab("Covariance of \u03B1 and |\u03B2|")
-cov_ag_p <- dummy + stat_function(fun = Cov_ag) + ylim(-0.4,0.4) + xlim(0,2) + theme_minimal() + xlab("Approximate ratio of variance among \n species to variance within species") + ylab("Covariance of \u03B1 and \u03B3")
-cov_plot <- grid.arrange(cov_ab_p,cov_ag_p,nrow=1) 
-ggsave("~/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/cov.png",cov_plot,width=6.5,height=2.5)
-```
-
-```{r, echo=F, fig.align='center', fig.cap='\\label{cov_fig}Curves representing the covariance between the magnitude of linear selection gradients and competition coefficients (left) and between stabilizing selection gradients and competition coefficients (right) as a function of $2bV_{\\bar X}$ which is approximately equal to the ratio of variance in mean traits among species to the intraspecific trait variance.
-In both plots we set $c=1.0\\times10^{-4}$, $b=0.1$, $\\bar r=0.1$ and $N=1.0\\times10^{10}$ and let $V_{\\bar X}$ vary.'}
-knitr::include_graphics('/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/cov.png')
-```
-We visualize these relationships in Figure \ref{cov_fig} for fixed $c,b,\bar r,U$ and $N$.
-In SM \S\ref{coef_grad_moms} we verify the qualitative nature of these relationships by repeating numerical integrations of system (\ref{comm_dynamics}) across a set of randomly drawn model parameters.
-We designed our numerical approach to inherit the assumptions of symmetric model parameters such as mutation rates $\mu_1=\dots=\mu_S$ and abiotic optima $\theta_1=\dots=\theta_S$ and to relax the key assumptions of normally distributed mean traits along with constant additive genetic variances and abundances across species.
-
-Our assumptions in this section imply $b^{-1}=2(\sigma^2+w)$.
-If we further assume $\sigma^2+w\approx\sigma^2$, then $2bV_{\bar X}\approx V_{\bar X}/\sigma^2$.
-That is, when populations are generalists and are comprised of specialist individuals, the value $2bV_{\bar X}$ is approximately equal to the ratio of interspecific mean trait variation to intraspecific individual trait variation.
-Glancing at Figure (\ref{cov_fig}) we see there is no relationship between selection gradients and competition coefficients when this ratio is zero.
-From equation (\ref{cov_alpha_beta}) we can use numerical optimization to find that when $V_{\bar X}/\sigma^2\approx1.25$ the relationship between the magnitudes of linear selection gradients and competition coefficients disappears, but when $V_{\bar X}/\sigma^2\lessapprox1.25$ ($\gtrapprox1.25$), this covariance becomes negative (positive).
-Equation (\ref{cov_alpha_gamma}) states that when $V_{\bar X}/\sigma^2$ is approximately equal to one (or slightly larger), there is no expected relationship between competition coefficients and quadratic selection gradients.
-However, when $V_{\bar X}/\sigma^2<1.0$ ($>1.0$), then we expect a positive (negative) relationship between $\alpha$ and $\gamma$.
-These results are true regardless of the chosen parameter values.
-In SM \S\ref{coef_grad_moms} we use simulations of system (\ref{comm_dynamics}) to show that these results do not qualitatively differ when allowing for heterogeneous population sizes and additive genetic variances across species.
-
-From a biological perspective, if the ratio $V_{\bar X}/\sigma^2$ is small, then species are packed tightly in phenotypic space.
-In our model this occurs when fitness is more sensitive to abiotic stabilizing selection than competition ($a\gg c$).
-This causes species to overlap more in niche space (i.e., large $\alpha$) and creates disruptive selection for greater intraspecific variance (i.e., positive $\gamma$), which explains the positive region of covariance between $\alpha$ and $\gamma$.
-However, as species begin to overlap in niche space, directional selection begins to vanish (i.e., small $|\beta|$), leading to a negative covariance between $\alpha$ and $|\beta|$.
-In the limiting case that two species have perfectly overlapping niches, they will exhibit zero directional selection since a shift in either direction will yield identical fitness advantages.
-
-In the opposite scenario where fitness is much more sensitive to competition than abiotic stabilizing selection ($c\gg a$), species will not evolve to be as tightly packed and instead their niche-centers will be spread out with little ovelap in their resouce utilization curves (i.e., small $\alpha$).
-In this case biotic directional selection will be strong (i.e., large $|\beta|$), particularly for species towards the outer regions of niche space due to asymmetric fitness advantages conferred by shifts in niche-centers.
-This leads to a positive covariance between $\alpha$ and $|\beta|$.
-However, as noted above, this directional selection will also erode away at standing heritable variation (i.e., negative $\gamma$), reducing the rate at which adaptation can occur and creating a negative covariance between $\alpha$ and $\gamma$.
-
-In summary, we see the relations between competition coefficients and selection gradients are non-trivial and depend on the relative magnitudes of different ecological processes shaping the community.
-However, this does not address the relation between competition coefficients and coevolution per se.
-In SM \S\ref{coef_grad_moms} we show that calculating a formula for the covariance between competition coefficients and the metric of coevolution $\mathfrak{C}$ introduced above provides a difficult analytical challenge.
-Instead of confronting this challenge we build on our numerical approach used to justify analytical approximations of $\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)$ and $\mathrm{Cov}_{f_{\bar X}}(\alpha,\gamma)$ to approximate the correlation of $\alpha$ and $\mathfrak{C}$.
-In Figure \ref{cor_coev_fig} we plot values of $\mathrm{Corr}_{f_{\bar X}}(\alpha,\mathfrak{C})$ against the exact ratio of interspecific trait variance to intraspecific trait variance.
-Using a cubic regression, we see the correlation of coevolutionary selection gradients and competition coefficeints is negative at variance ratios below $0.5$, zero at variance ratios between $0.5$ and $1.0$, and drops below zero again at variance ratios above $1.0$.
-
-```{r, eval=F, echo=F}
-require(ggplot2)
-require(gridExtra)
-df = read.csv("~/Research/Branching Brownian Motion/corrs.csv")
-Cαβp = ggplot(df,aes(x=df$V/df$σ,y=df$Cαβ))+geom_point(size=0.1)+stat_smooth(color="black", method = "lm", formula = y ~ x + I(x^2) + I(x^3))+theme_minimal() + ylab("Correlation of \u03B1 and \u03B2")+xlab("Ratio of variance among \n species to variance within species")
-Cαabsβp = ggplot(df,aes(x=df$V/df$σ,y=df$Cαabsβ))+geom_point(size=0.1)+stat_smooth(color="black", method = "lm", formula = y ~ x + I(x^2) + I(x^3))+theme_minimal()+ylab("Correlation of \u03B1 and |\u03B2|")+xlab("Ratio of variance among \n species to variance within species")
-Cαγp = ggplot(df,aes(x=df$V/df$σ,y=df$Cαγ))+geom_point(size=0.1)+stat_smooth(color="black", method = "lm", formula = y ~ x + I(x^2) + I(x^3))+theme_minimal()+ylab("Correlation of \u03B1 and \u03B3")+xlab("Ratio of variance among \n species to variance within species")
-Cαℭp = ggplot(df,aes(x=df$V/df$σ,y=df$Cαℭ))+geom_point(size=0.1)+stat_smooth(color="black", method = "lm", formula = y ~ x + I(x^2) + I(x^3))+theme_minimal()+ylab("Correlation of \u03B1 and ℭ")+xlab("Ratio of variance among \n species to variance within species")
-corrsp = grid.arrange(Cαabsβp,Cαγp,nrow=1)
-ggsave("/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/corrs.png",  corrsp,  width=6.5,height=2.5)
-ggsave("/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/Cαℭp.png",   Cαℭp,  width=4,height=2.5)
-```
-
-```{r, echo=F, fig.align='center', fig.cap='\\label{cor_coev_fig}Numerical estimates for the correlation of the strength of coevolution measured by $\\mathfrak{C}$ and competition coefficeints $\\alpha$ plotted against the variance of mean traits among species divided by the mean variance of traits within species.
-Each dot represents the result from a single simulation.
-The red line is a cubic regression.'}
-knitr::include_graphics('/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/Cαℭp.png')
-```
-
-# Conclusion \label{conclusion}
-
-We have introduced an approach to derive models of evolutionary ecology using the calculus of white noise and coined SAGA, a SPDE model of phenotypic evolution that accounts for demographic stochasticity.
-From SAGA we derived SDE that track the dynamics of abundance, mean trait and additive genetic variance.
-Observing the expressions of these SDE we find the effects of demographic stochasticity on the evolution of mean trait and additive genetic variance characterize the effects of random genetic drift.
-Although Lande [-@Lande1976] has previously characterized the effects of random genetic drift on mean trait evolution in quantitative genetic models, the approach taken assumed constant effective population size and discrete non-overlapping generations.
-In contrast, our approach shows random genetic drift is a result of demographic stochasticity for continuously reproducing populations with fluctuating abundances.
-To illustrate the relevance of our approach to studies of evolutionary ecology, we comined our SDE with classical competition theory to derive a model of diffuse coevolution and used this model to investigate the relationship between standardized selection gradients and competition coefficients.
-We found the absolute value of linear selection gradients are negatively related with competition coefficients when the ratio of interspecific phenotypic variation to intraspecific phenotypic variation is approximately less than $1.25$ and positive when this ratio is approximately greater than $1.25$.
-We found the opposite relationship between quadratic selection gradients and competition coefficients with a threshold diversity ratio equal to unity instead of $1.25$.
-This work demonstrates that connecting contemporary theoretical approaches to evolutionary ecology with some fundamental results in the theory of measure-valued stochastic processes allows for the development of a rigorous, yet flexible approach to synthesizing the dynamics of abundance and distributions of quantitative characters, in agreement with Champagnat, Ferrière and Méléard [-@Champagnat2006].
-Although this approach enjoys these important merits, there remains biological details and their associated technical challenges that need to be confronted for gaining a more thorough and accurate understanding of ecological communities.
-We touch on just four of them here and provide suggestions for future research to approach these challenges.
-For the first three, we recommend the use of individual-based simulations of measure-valued processes to confront these challenges.
-To summarize why, we state here that analytical results become difficult to derive for small populations and complicated models of mutation and selection.
-
-__Limitations of diffusion limits and generalizing coexistence theory__
-
-As noted by Feller [-@feller1951], diffusion limits provide reasonable approximations for large populations, but relatively small populations require discrete models.
-Hence, as a diffusion limit, SPDE (\ref{SPDE}) restricts parameter values to regions that maintain large population sizes.
-Although this suggests an important restriction on any model developed under this framework by implying populations are not at risk of extinction, the SDE describing abundance dynamics technically permits extinction.
-However, for small abundances, pathology emerges in the evolution of trait means and variances.
-In particular, stochastic components of the SDE describing the evolution of $\bar x$ and $\sigma^2$ diverge towards infinite values as $N\to0$.
-Hence, studies involving extinction and phenotypic evolution should be pursued through a different approach.
-A natural alternative can be developed utilizing the underlying measure-valued branching process that SAGA is a diffusion limit of (see section \ref{stochastic}).
-This process explicitly tracks individuals throughout their life-history and captures reproduction as branching events.
-Hence, branching processes model population size as a non-negative integer instead of a continuously varying number.
-In particular, the pathological behavior described for the SDE of $\bar x$ and $\sigma^2$ does not occur for measure-valued branching processes.
-This concern relates to our comment in \S\ref{ecoevo} on using our model of diffuse coevolution to extend modern coexistence theory [sensu @Chesson2000] in a way that accounts for phenotypic evolution and demographic stochasticity.
-Just as previous work has sought to understand the implications of processes for coexistence separately, it seems necessary to determine whether or not results pertaining to the coexistence of competing populations hold using models that account for phenotypic evolution and demographic stochasticity simultaneously.
-Because such investigations consider both extinction and phenotypic evolution we argue the models used should be either simulations of, or whenever possible, analytical results based on measure-valued branching processes.
-
-__Genetic architecture and intraspecific trait distributions__
-
-Our treatment of inheritance and our approach to model coevolution rests on the assumptions of normally distributed breeding values and expressed phenotypes along with asexual reproduction.
-However, real traits are not encoded by an infinite number of loci each contributing an additive infinitesimal effect (as required by the infinitesimal model), mutations are not inherited as normally distributed deviations from parental breeding values (as required by the Gaussian allelic model) and many populations of interest exhibit non-random sexual reproduction.
-Departures from this model of genetic architecture can produce more complex distributions of breeding values and expressed traits.
-Such deviations can be reinforced via strong non-Gaussian selection surfaces, including the surface $m(\nu,x)$ we have derived from niche theory, along with non-random mating in sexually reproducing populations.
-However, Gaussian approximations are convenient since they are defined by their mean and variance.
-Future work investigating the effects of non-normally distributed traits on the structure and dynamics of ecological communities will therefore need to confront higher moments such as skew and kurtosis, ideally integrating previously established mathematical approaches to derive the dynamics of these higher moments [@Dbarre2015EvolutionOQ].
-
-Another approach to breaking the assumption of normally distributed trait values is the development of explicit multilocus models.
-These models describe the contributions of alleles at particular loci in the genome to the development of quantitative traits.
-Tracking the fluctuations of allele frequencies then allows theoreticians to investigate the dynamics of phenotypic distributions that deviate from normality.
-This approach has a long history in theoretical quantitative genetics [@bulmer1980; @pmid7851785; @Kirkpatrick1727] and coevolutionary theory [@Nuismer2005; @Kopp2006; @Nuismer2007].
-However, work to investigate the role of complex genetic architectures in mediating feedbacks between the dynamics of population abundances and the distributions of traits mediating ecological interactions has apparently only just begun [@Patel2019].
-Since analytical results are difficult to obtain for general models of mutation, development and selection [c.f., @9780471986539], we again suggest the use of measure-valued branching processes and their individual-based simulations.
-For example, by assuming the mutational effects in the BBM model is Gaussian with variance an order of magnitude greater than the equilibrium phenotypic (or additive genetic) variance, we obtain the _House-of-cards_ model [@Turelli1984].
-However, perhaps considering heavy-tailed distributions such as the Laplace-distributions considered by Park et.
-al.
-[-@Park2017] will provide more realistic yet less analytically tractable models of mutational effects.
-Such mutation models have important implications in conservation genetics where small populations evolve primarily in response to mutation and random genetic drift [@Lande1995].
-
-__Ecological stoichiometry__
-
-Our treatment of both biotic and abiotic selection neglects important chemical constraints of biological reality.
-For instance, the resource we assume species are competing over is modelled as a static quantity.
-However, real resources can be dynamic quantities.
-Such theoretical quantities may reflect abiotic cycles of material and energy or whole trophic layers comprised of prey, hosts or mutualists.
-Although resource dynamics have been captured theoretically in consumer-resource models [@davidtilman1982], developing a more realistic model of resource competition must incorporate details on the ecophysiology of organisms [@michelloreau2010].
-Doing so will help clarify the feedback between the evolution of populations and the ecosystem processes they are a part of.
-
-Using plant-pollinator communities as an example, the role of nitrogen mediating interspecific interactions has been reviewed by David, Storkey and Stevens [-@David2019] and the evolutionary ecology of the nutritional content of nectar has been reviewed by Parachnowitsch, Manson and Sletvold [-@Parachnowitsch2018].
-These studies demonstrate the need for further research to understand how soil nutrient availability and organismal ecophysiology structures communities of plants and pollinators.
-Theoretical pursuits in this direction may profit from interfacing the framework we have outlined here with population-ecosystem models such as that developed by Fridley [-@Fridley2017].
-However, these population-ecosystem models will likely yield complicated fitness functions that are difficult to analyze using rigorous mathematical techniques.
-Hence, we again recommend the use of simulations of measure-valued branching processes that generalize BBM to understand when analytical approximations hold.
-
-__Accounting for macroevolutionary history__
-
-To understand patterns found in ecological communities considerations must push beyond microevolutionary and contemporary ecological processes and consider the macroevolutionary dynamics of ancestral lineages leading to extant populations.
-Using sub-alpine flower communities as an example, one can observe a very strict ordering of phenology across a broad geographic range.
-In particular, whether in the Colorado Rocky mountains (such as Gothic, Colorado) or on an outlier of the Idaho batholith (such as Kamiak butte near Palouse, Washington), one will almost surely observe a very conspicuous order of flowers emerging in early spring: at the very beginning of the season blooms _Claytonia lanceolata_ followed by _Erythronium grandiflorum_ which precedes _Delphinium nuttallianum_ (B.
-Week, personal observations).
-If contemporary phenological coevolution is rampant, why should this pattern be so well preserved across a thousand miles of rugged and diverse terrain?
-Although it would be exciting to find that these species repeatedly coevolved this pattern in each location, a more parsimoniuous hypothesis suggests the phenology and physiology of these species slowly evolved independently over macroevolutionary time scales to take advantage of the specific conditions available within each of these windows of the flowering season.
-However, this could not have carried out in the Rocky mountains since this terrain only became habitable just over ten thousand years ago as the glaciers of the Pleistocene began to retreat (Paul CaraDonna, personal communications).
-Hence, given these considerations, it appears that an understanding of early season phenology patterns should focus on how these communities are assembled as opposed to contemporary evolutionary dynamics.
-Indeed, recent work testing models of phylogeography ignores the potential for contemporary evolution and instead suggests alpine flower communities tend to follow neutral assembly where flowers merely compete for who can disperse to new habitat first, as opposed to a selective process where a regional species pool is filtered for those species adapted to the newly available habitat [@Marx2017].
-
-Of course microevolutionary and ecological dynamics are not completely irrelevant for understanding patterns in communities that are primarily structured by deep evolutionary processes.
-In particular, macroevolutionary trait evolution is simply the aggregation of microevolutionary change occuring over large spans of time.
-This suggests a road forward to connect the theory we have introduced to models of macroevolutionary trait evolution.
-
-Some approaches to modelling macroevolutionary trait change simply repurpose microevolutionary models by blindly rescaling time from the units of generations to millions of years [e.g., @Nuismer2014].
-Such an approach makes the implicit assumption that trait evolution is statistically self-similar [sensu @kennethfalconer2014] so that the stochastic evolution of traits on macroevolutionary time scales has the same properties of trait evolution on microevolutionary time scales.
-Although some stochastic processes, including Brownian motion, do exhibit self-similarity, others do not.
-For example, consider a modification of the Ornstein-Uhlenbeck process defined by the SDE
-
-\begin{equation}
-dX_t=a(\theta_t-X_t)dt+bdW_t
-\end{equation}
-
-where $a,b>0$, $W_t$ is a standard Brownian motion and $\theta_t$ is itself a Markov process that takes normally distributed jumps centered on its current location at exponentially distributed time intervals.
-If we assume the rate $\lambda$ at which jumps occur is much smaller than $a$ and the variance in jumping is much larger than $b^2$, then, even though the sample paths of $X_t$ are actually continuous (if we zoom in close enough, they look like Brownian motion), over long intervals of time sample paths of $X_t$ will begin to appear to have periods of continuity interrupted by an occasional discontinuous jump and thus approach a qualitatively distinct process.
-These emergent properties can be formally characterized by L&eacute;vy processes and have been successfully employed in comparative phylogenetics to fit phenotypic data from extant populations and the fossil record [@Landis2017].
-It would therefore be interesting to investigate whether an application of a separation of time scales argument for the rate of environmental change ($\lambda$) versus the rate of evolutionary and ecological change ($a$) to microevolutionary models derived via our framework can be used to obtain macroevolutionary models that include not only mean trait evolution, but also the evolution of trait variance and abundance.
-The resulting macroevolutionary models can give rise to novel comparative phylogenetic methods and provide initial conditions for microevolutionary models that capture contemporary dynamics.
-
-__Final remarks__
-
-Although top-down approaches to community ecology, such as the Maximum Entropy Theory of Ecology [@Harte2011], have enjoyed some success in describing community-level patterns [@Harte2014; @Xiao2015], a mechanistic understanding of why these patterns emerge and how they will change remains a formidable task.
-Such an understanding must take both bottom-up and top-down approaches integrating considerations from the ecophysiology of individual organisms that reveal the economics of interspecific interactions [@Sterner2008], to the phylogeographic history of taxa that sets the stage for contemporary dynamics [@Hickerson2010].
-Through connecting these dots we can increase the variance explained in observations of ecological communities by specific mechanisms and come closer to a predictive theory of evolutionary community ecology.
-Despite the long list of equations derived in this paper, this work takes just one small step towards capturing these many details.
-However, we hope the theoretical framework outlined here along with the demonstration of its use in modelling competitive communities provides some helpful tools to aid quantitative evolutionary ecologists in reaching such lofty goals.
-
-\newpage
-
-# Supplementary material (SM)
 
@@
 
-## Sufficient conditions for finite mean, variance and total abundance in the deterministic case \label{finite}
+# Sufficient conditions for finite mean, variance and total abundance in the deterministic case \label{finite}
 
@@
 
-## Equilibrium moments for a deterministic population experiencing logistic growth and stabilizing selection \label{equilib}
+# Equilibrium moments for a deterministic population experiencing logistic growth and stabilizing selection \label{equilib}
 
@@
 
-## The relation between diffusion and convolution with a Gaussian kernel \label{diffconvequiv}
+# The relation between diffusion and convolution with a Gaussian kernel \label{diffconvequiv}
 
@@
 
-## Deterministic dynamics of $\sigma^2(t)$ \label{var_deriv}
+# Deterministic dynamics of $\sigma^2(t)$ \label{var_deriv}
 
@@
 
-## Comparing our treatment of white noise to Da Prato and Zabczyk (2014) \label{DPZ14}
+# Comparing our treatment of white noise to Da Prato and Zabczyk (2014) \label{DPZ14}
 
@@
 
-## Simulating the rescaled process \label{numerical}
+# Simulating the rescaled process \label{numerical}
 
@@
 
-### Description of simulation \label{description}
+## Description of simulation \label{description}
 
@@
 
-__Rescaling__
+## Rescaling
 
@@
 
-## Derivation of SDE for $\bar x$ and $\sigma^2$ \label{SDE_DERIV}
+# Derivation of SDE for $\bar x$ and $\sigma^2$ \label{SDE_DERIV}
 
@@
 
-### Derivation for trait mean
+## Derivation for trait mean
 
@@
 
-### Derivation for trait variance
+## Derivation for trait variance
 
@@
 
-## Relating fitness of expressed traits to fitness of breeding values \label{fit2fit}
+# Relating fitness of expressed traits to fitness of breeding values \label{fit2fit}
 
@@
 
-## Derivation of diffuse coevolution model \label{diffuse}
+# Derivation of diffuse coevolution model \label{diffuse}
 
@@
 
-## The relation between competition coefficients and selection \label{coef_grad_moms}
-
-### Derivation of analytical approximations
+# The relation between competition coefficients and selection \label{coef_grad_moms}
+
+## Derivation of analytical approximations
 
@@
 
-#### Caclulating $\mathrm{Cov}_{f_{\bar X}}(\alpha,\gamma)$
+### Caclulating $\mathrm{Cov}_{f_{\bar X}}(\alpha,\gamma)$
 
@@
 
-#### Caclulating $\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)$
+### Caclulating $\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)$
 
@@
 
-#### Starting the caclulation of $\mathrm{Cov}_{f_{\bar X}}(\alpha,\mathfrak{C})$ \label{caC}
+### Starting the caclulation of $\mathrm{Cov}_{f_{\bar X}}(\alpha,\mathfrak{C})$ \label{caC}
 
@@
 
-### Numerical estimates for heterogeneous $N$ and $G$
+## Numerical estimates for heterogeneous $N$ and $G$
 

### 2020-04-18 .A White Noise Approach To Evolutionary Ecology/SM.tex
split main text and SM

--- 2020-04-18 .A White Noise Approach To Evolutionary Ecology/draft.tex
@@
 \hypersetup{
-            pdftitle={A White Noise Approach to Evolutionary Ecology},
-            pdfauthor={Bob Week, Steve Krone, Scott L. Nuismer, Luke J. Harmon},
+            pdftitle={Supplementary Material (SM)},
             pdfborder={0 0 0},
@@
 
+\usepackage{xr}
 \usepackage{mathtools}
@@
 \linenumbers
+\externaldocument{draft}
 \usepackage{float}
@@
 
-\title{A White Noise Approach to Evolutionary Ecology}
-\author{Bob Week, Steve Krone, Scott L. Nuismer, Luke J. Harmon}
-\date{3/1/2020}
+\title{Supplementary Material (SM)}
+\author{}
+\date{\vspace{-2.5em}}
 
@@
 \maketitle
-\begin{abstract}
-We derive the dynamics of the distribution of a quantitative character
-and the abundance of a biological population from a stochastic partial
-differential equation driven by space-time white noise. In the process
-we develop a useful set of heuristics to operationalize the powerful,
-but abstract theory of white noise and measure-valued stochastic
-processes. This approach allows us to compute the full implications of
-demographic stochasticity on phenotypic distributions and abundances of
-populations. We demonstrate the utility of our approach by deriving a
-quantitative genetic model of diffuse coevolution mediated by
-exploitative competition for a continuum of resources. Other than trait
-and abundance distributions, this model predicts interaction networks
-parameterized by rates of interactions, competition coefficients, and
-selection gradients. We briefly investigate the relationship between
-selection gradients and competition coefficients. This illustrative
-investigation suggests absolute values of linear selection gradients are
-negatively correlated with competition coefficients when the ratio of
-interspecific trait variation to intraspecific trait variation is
-approximately less than \(1.25\) and positively correlated otherwise.
-The inverse relationship holds for quadratic selection gradients except
-with threshold diversity ratio equal to unity instead of \(1.25\).
-Hence, this approach contributes to the development of a synthetic
-theory of evolutionary ecology by formalizing first principle
-derivations of stochastic models that underlie rigorous investigations
-of the relationship between feedbacks of biological processes and the
-patterns of diversity they produce.
-\end{abstract}
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-Our goal in this manuscript is to develop a rigorous, but accessible
-approach to synthesize the stochastic dynamics of abundance, mean trait
-and heritable variation in biological populations for the study of
-theoretical evolutionary ecology. A primary aim of theoretical
-evolutionary ecology is the development of mathematical approaches to
-describe the evolution of populations and their interactions with both
-the biotic and abiotic environments in which they are embedded. Given
-this consideration, a natural scope for such an approach centers on
-quantifying the abundance dynamics of populations and the evolution of
-traits mediating their interactions as functions of relevant abiotic
-factors. Although taking into account abundance, phenotype and
-environment provides the basis for a partial understanding of the
-complex nature of biological communities, a deeper understanding must
-account for the effects of dispersal and the phylogeographic history of
-interacting lineages (Kraft et al. 2007; Hickerson et al. 2010; Manceau,
-Lambert, and Morlon 2016; McPeek 2017) along with the genetic basis of
-ecologically relevant traits (Conner 2004; Fussman, Loreau, and Abrams
-2007) and feedbacks between populations and the biogeochemical cycles
-they ultimately depend on (Loreau 2010; Ågren and Andersson 2012). It is
-therefore ideal that the development of any such mathematical approach
-anticipates extensions to account for these important factors shaping
-ecological communities, especially as empirical and conceptual work in
-these directions continues to grow (Abdala-Roberts and Mooney 2014;
-Kölzsch et al. 2015; Crutsinger 2015; Fitzpatrick et al. 2015, 2017;
-Marx et al. 2017; Rudman et al. 2017; Skovmand et al. 2018; Nuland et
-al. 2019; Harmon et al. 2019). Furthermore, the approach would benefit
-from a stochastic component to capture the chance nature of biological
-reality (Lande, Engen, and SÆther 2003; Meester et al. 2018; Mubayi et
-al. 2019) and serve as a basis for the construction of statistical
-methods that measure evolutionary and ecological processes occuring in
-the wild. Such methods will tether theory to reality and allow for
-rigorous tests of hypotheses on the structure and behavior of ecological
-communities. In this paper we introduce a framework that establishes a
-formal connection between the continuous-time dynamics of abundance and
-quantitative traits in stochastically evolving populations. We then
-demonstrate the utility of our framework through the derivation and
-analysis of a model of diffuse coevolution and discuss how it can be
-extended to account for the details mentioned above.
-
-Current theoretical approaches to synthesize evolution and ecology have
-capitalized on the fact that biological fitness plays a key role in
-determining both sets of dynamics. While correlation of fitness and
-genotype is the basis of evolution by natural selection, the mean
-fitness across all individuals in a population determines the growth,
-stasis or decline of abundance. In section \ref{deterministic} we review
-the mathematical formalization of this connection, which has been
-established in the contexts of population genetics (Crow and Kimura
-1970; Roughgarden 1979), evolutionary game theory (Hofbauer and Sigmund
-1998; Nowak 2006; Lion 2018), quantitative genetics (Lande 1982; Doebeli
-1996; Lion 2018) and a unifying framework for these three distinct
-approaches to evolutionary theory (Champagnat, Ferrière, and Méléard
-2006) which includes our approach as a special case. We note here there
-are two primary differences between the results presented in this paper
-and those of Champagnat, Ferrière and Méléard (2006). First, instead of
-developing a unifying approach to evolutionary theory, we focus on
-developing a stochastic synthesis of population dynamics and
-quantitative genetics including specific expressions for the dynamics of
-abundance, mean trait and additive genetic variance. Second, although
-Champagnat, Ferrière and Méléard (2006) provide rigorous proofs of
-results relevant to our approach, they do not communicate the nessecary
-tools we make use of for deriving the dynamics of abundance, mean trait
-and additive genetic variance. We aim to communicate these tools to a
-wide audience by striking a balance between intuitive appeal and
-mathematical rigor.
-
-Although reviewing previously established theoretical results reveals a
-beautiful synthesis of evolution and population ecology, it also reveals
-a gap in theoretical approaches to incorporate the intrinsically random
-nature of populations. Specifically, in theoretical quantitative
-genetics the derivation of a population's response to random genetic
-drift is derived in discrete time under the assumption of constant
-effective population size using arguments based on properties of random
-samples (Lande 1976). Though this approach conveniently mimics the
-formalism provided by the Wright-Fisher model of population genetics,
-real population sizes fluctuate over time. Furthermore, since these
-fluctuations are themselves stochastic, it seems natural to derive
-expressions for the evolutionary response to demographic stochasticity
-and consider how the results relate to characterizations of random
-genetic drift. This can be done in continuous time for population
-genetic models without too much technical overhead, assuming a finite
-number of alleles (Lande, Engen, and Sæther 2009; Parsons, Quince, and
-Plotkin 2010; Gomulkiewicz, Krone, and Remien 2017). However, for
-populations with a continuum of types, such as a quantitative trait,
-finding a formal approach to derive the evolutionary response to
-demographic stochasticity has remained a vexing mathematical challenge.
-In this paper we close this gap by combining the calculus of white noise
-with results on rescaled limits of branching Brownian motion processes
-(BBM) and stochastic partial differential equations (SPDE), which are
-stochastic analogs of deterministic partial differential equations
-(PDE). Our goal has two components: 1) Establish a novel synthetic
-approach to theoretical evolutionary ecology that provides a formal
-connection between demographic stochasticity and random genetic drift in
-the context of quantitative traits. 2) Communicate some useful
-properties of space-time white noise, BBM and SPDE to as wide of
-audience as possible. With this goal in mind we will not provide a
-rigorous treatment of any of these deep subjects. Instead, we introduce
-a set of heuristics that only require the basic concepts of Reimann
-integration, partial differentiation and some exposure to Brownian
-motion and stochastic ordinary differential equations (SDE). For a
-concise introduction to SDE and Brownian motion, we recommend the primer
-by Evans (2014). Rigorous treatments of SPDE and rescaled limits of BBM
-can be found in Walsh (1986) and Dawson (1993) respectively.
-
-Through accomplishing these goals this work yields five main results.
-Firstly, we develop a concise, but accurate approach for understanding
-SPDE and provide heuristics to empower mathematical evolutionary
-ecologists to derive novel models that relax key assumptions commonly
-made in combining models of evolutionary and abundance dynamics. Second,
-we formally derive general SDE driven by demographic stochasticity that
-express the joint dynamics of abundance, mean trait and additive genetic
-variance for populations exhibiting a wide range of potential phenotypic
-distributions. Although these dynamics have been considered before while
-accounting for demographic stochasticity (Polechová and Barton 2015), to
-our knowledge this is the first time they have been rigorously derived
-from first principles. Third, we simplify these SDE by approximating the
-abundance density of phenotypes with a Gaussian curve and integrate the
-results with a model of imperfect inheritance to find formula that
-generalize classical quantitative genetic expressions. This includes an
-expression for the response of additive genetic variance to selection in
-terms of fitness gradients analogous to the expression for the response
-in mean phenotype found by Lande (1976). Fourth, following classical
-niche theory we provide a rigorous mechanistic derivation of fitness for
-populations experiencing resource competition and abiotic stabilizing
-selection using diffusion limits of measure-valued branching processes.
-This calculation provides a demonstration for performing such diffusion
-limits that can be applied in other contexts. Fifth, we explore
-implications of the resulting model of diffuse coevolution by
-investigating the relationship between selection gradients and
-competition coefficients, discovering the conditions under which these
-quantities positively or negatively covary.
-
-We now summarize the basic structure of the paper. To provide motivation
-for the stochastic equations developed later and background for our
-model of coevolution, we begin with \S\ref{deterministic} by briefly
-summarizing derivations of deterministic evolutionary and demographic
-dynamics. Starting with a PDE (which we refer to as the Deterministic
-Asexual Gaussian allelic model with Abundance dynamics, abbreviated
-DAGA), we arrive at a general set of ordinary differential equations
-modelling the dynamics of abundance, trait mean and trait variance. We
-discuss the relationship between DAGA and previous models of phenotypic
-evolution that have been employed in theoretical quantitative genetics.
-In \S\ref{wnc} we discuss how a diffusion limit of a spatially
-structured branching process (i.e., a measure-valued branching process)
-leads to a SPDE that provides a natural stochastic generalization of
-DAGA. We refer to this SPDE as the Stochastic Asexual Gaussian allelic
-model with Abundance dynamics (abbreviated SAGA). The diffusion limit in
-turn provides a rigorous method for constructing fitness functions used
-in models of evolutionary ecology. However, before we discuss
-measure-valued braching processes and their diffusion limits, we find it
-necessary to review some fundamental results on the calculus of white
-noise and introduce a set of mathematical tools based on these results.
-In SM \S\ref{SDE_DERIV} we employ these tools to derive a system of SDE
-generalizing our deterministic results to account for demographic
-stochasticity. We find the effect of demographic stochasticity on the
-evolution of mean trait and phenotypic variance characterizes the
-process of random genetic drift in continuous time. However, although
-biologically insightful, these equations remain difficult to analyze and
-implement numerically. In \S\ref{equations} we use an assumption of
-normally distributed trait values to simplify these expressions into
-tractable formula that express the evolutionary response to selection in
-terms of fitness gradients. We then account for the constraint of
-adaptive evolution on the amount of heritable variation available in a
-population by extending these results via a model of imperfect
-inheritance. The resulting equations coincide with classical results in
-quantitative genetics as a special case. In \S\ref{coev} we combine the
-derived equations of population dynamics with classical niche theory to
-formulate a model of coevolution across a guild of \(S\) species
-participating in exploitative competition along a common resource
-continuum. In SM \S\ref{diffuse} we apply a result on rescaled limits of
-BBM to provide a rigorous derivation. In \S\ref{dynamics} we illustrate
-our model of coevolution through numerical integration for \(S=100\)
-species, tracking the dynamics of traits and abundances under scenarios
-of weak and strong competition. We include an account of the natural
-history of the simulated system and discuss the significance of
-demographic stochasticity for structuring ecological communities. In
-\S\ref{ecoevo} we provide expressions for selection gradients and
-competition coefficients implied by our model and use these expressions
-to investigate the relationship between the degree of competition and
-strength of coevolution. Finally, \S\ref{conclusion} concludes with a
-summary of accomplishments, a few remarks on the limits of this approach
-and future directions to incorporate more explicitly the effects of
-small populations, the genetic architecture of quantitative traits,
-feedbacks with ecosystem processes and the macroevolutionary history of
-interacting lineages.
-
-\hypertarget{the-framework}{%
-\section{The framework}\label{the-framework}}
-
-At the core of our approach is a stochastic analog of the replicator
-equation with mutation in continuous time and phenotypic space (Taylor
-and Jonker 1978; Schuster and Sigmund 1983). From this stochastic
-replicator-mutator equation we derive a system of SDE for the dynamics
-of abundance, mean trait and additive genetic variance of a population.
-Hence, our approach develops a quantitative genetic theory of
-evolutionary ecology. A popular alternative to quantitative genetics is
-the theory of adaptive dynamics. As demonstrated by Page and Nowak
-(2002) and Champagnat, Ferrière and Méléard (2006), the canonical
-equation of adaptive dynamics can be derived from the replicator-mutator
-equation. Thus, one could start from the atomic roots of our approach
-and pursue a stochastic adaptive dynamic theory instead. We choose the
-former in anticipation of an extension of our approach that explicitly
-models the genetic details of populations.
-
-In this section we review the derivations of the replicator-mutator
-equation and trait dynamics from abundance dynamics and extend these
-formula along with related results to the stochastic case. The results
-established in this section provide the framework from which larger
-scale ecological stuctures, such as species abundance distributions and
-interaction networks, can be computed.
-
-\hypertarget{deterministic-dynamics}{%
-\subsection{\texorpdfstring{Deterministic dynamics
-\label{deterministic}}{Deterministic dynamics }}\label{deterministic-dynamics}}
-
-Our review begins by considering the dynamics of an asexually
-reproducing population in a homogeneous environment. For simplicity, we
-first assume individuals are haploid and carry one of \(K\) alleles each
-with a different fitness expressed as growth rate before introducing a
-model involving a quantitative trait. Under these assumptions, the
-derivation of the evolution of allele frequencies due to natural
-selection can be derived from expressions of exponential growth. This,
-and a few related approaches, have been provided by Crow and Kimura
-(1970, \S 5.3). Specifically, denoting \(\nu_i\) the abundance of
-individuals with allele \(i\) and \(m_i\) the growth rate of allele
-\(i\) (called the Malthusian parameter in Crow and Kimura 1970), we have
-\begin{equation}\label{first}
-\frac{d\nu_i}{dt}=m_i\nu_i.
-\end{equation} Starting from this model, we get the total abundance of
-the population as \(N=\sum_{i=1}^K\nu_i\), the frequency of allele \(i\)
-as \(p_i=\nu_i/N\) and the mean fitness of the population as
-\(\bar m=\sum_{i=1}^Kp_im_i\). Hence, we can employ some elementary
-calculus to derive the dynamics of abundance \(dN/dt\) and the dynamics
-of allele frequencies \(dp_1/dt,\dots,dp_K/dt\) as \begin{equation}
-\frac{dN}{dt}=\sum_{i=1}^K\nu_im_i=N\sum_{i=1}^Kp_im_i=\bar mN,
-\end{equation} \begin{equation}\label{popgen}
-\frac{dp_i}{dt}=\frac{d}{dt}\frac{\nu_i}{N}=\frac{1}{N^2}\left(N\frac{d\nu_i}{dt}-\frac{dN}{dt}\nu_i\right)
-=\frac{1}{N}\left(m_i\nu_i-\bar m Np_i\right)=(m_i-\bar m)p_i.
-\end{equation} Two important observations of these equations include (i)
-mean fitness \(\bar m\) determines the abundance dynamics of the entire
-population and (ii) allele \(i\) will increase (decrease) in frequency
-if \(m_i>\bar m\) (\(<\bar m\)). Equation (\ref{popgen}) is known in the
-field of evolutionary game theory as the replicator equation (Hofbauer
-and Sigmund 1998; Nowak 2006; Lion 2018; Taylor and Jonker 1978;
-Schuster and Sigmund 1983). Instead of being explicitly focused on
-alleles, the replicator equation describes the fluctuations of relative
-abundances of various \emph{types} in a population in terms of the vital
-rates of each type. Using a matrix of transition rates between differing
-types, it is straight-forward to extend the replicator equation to
-include mutation, which is known as the replicator-mutator equation
-(Nowak 2006).
-
-Inspired by equations (\ref{first})-(\ref{popgen}), we derive an analog
-of the replicator-mutator equation for a continuum of types (that is,
-for a quantitative trait). In particular, we model a continuously
-reproducing population with trait values \(x\in\mathbb{R}\) and an
-abundance density \(\nu(x,t)\) that represents the amount of individuals
-in the population with trait value \(x\) at time \(t\). Hence, the
-abundance density satisfies \(N(t)=\int_{-\infty}^{+\infty}\nu(x,t)dx\)
-and \(p(x,t)=\nu(x,t)/N(t)\) is the relative density of trait \(x\)
-which we also refer to as the phenotypic distribution.
-
-To stay within the realm of biological plausibility we require a set
-technical assumptions. First, we assume the initial abundance density is
-continuous, non-negative, integrable and has finite trait mean and
-variance. That is, we assume \(\nu(x,0)\) is continuous in \(x\),
-satisfies \(\nu(x,0)\geq0\) for all \(x\in\mathbb{R}\) and
-\begin{equation}
-N(0)=\int_{-\infty}^{+\infty}\nu(x,0)dx<+\infty,
-\end{equation} \begin{equation}
--\infty<\bar x(0)=\int_{-\infty}^{+\infty}xp(x,0)dx<+\infty,
-\end{equation} \begin{equation}
-\sigma^2(0)=\int_{-\infty}^{+\infty}(x-\bar x(0))^2p(x,0)dx<+\infty,
-\end{equation} where \(\bar x(t)\) and \(\sigma^2(t)\) are respectively
-the mean trait and phenotypic variance at time \(t\geq0\). Second, we
-assume selection is determined by the growth rate \(m(h,x)\) that is
-differentiable with respect to both arguments and is bounded above by
-some \(r\in\mathbb{R}\) for all \(x\in\mathbb{R}\) and \(h\geq0\). We
-also assume \((1+h)C\leq m(h,x)\) for some constant \(C>0\) so that
-\(m(h,x)\) is linear or more than linear in \(h\). We denote by \(K\) an
-operator that takes functions such as \(\nu(x,t)\) as its argument and
-returns other functions that may still depend on the spatial variable
-\(x\) and the temporal variable \(t\). For our application we will be
-focused on operators \(K\) that can be written as
-\((K\nu)(x,t)=\int_{-\infty}^{+\infty}\kappa(x-y)\nu(y,t)dy\) for some
-non-negative and bounded function \(\kappa\). We set \(h=(K\nu)(x,t)\)
-to account for nonlocal effects of the abundance density \(\nu(x,t)\) on
-the fitness of individuals with trait value \(x\) (Volpert 2014). As an
-example of a nonlocal effect, consider a population experiencing
-intraspecific competition for a common resource such that the
-competitive interactions are not mediated by any set of traits. In this
-case all individuals in the population effect the fitness of one another
-even though they may exhibit a diverse range of trait values. The term
-nonlocal comes from interpreting the trait value as spatial location.
-From here on we abbreviate \(m(h,x)=m\big((K\nu)(x,t),x\big)\) to just
-\(m(\nu,x)\). Third, we assume mutation is captured by diffusion with
-coefficient \(\frac{\mu}{2}\). With these technicalities addressed we
-model the demographic dynamics of a population and the dynamics of a
-quantitative character simultaneously by the PDE
-\begin{equation}\label{eq1}
-\frac{\partial}{\partial t}\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)
-\end{equation} with the initial condition \(\nu(x,0)\) described above.
-Equation (\ref{eq1}) is a semilinear parabolic PDE and is referred to as
-a scalar reaction-diffusion equation (Zheng 2004; Evans 2010). To ensure
-existence and uniqueness of solutions to (\ref{eq1}), we further assume
-that for each \(M>0\) there exists a constant \(L_M>0\) depending on
-\(M\) such that
-\(\int_\mathbb{R}\nu_1(x)dx,\int_\mathbb{R}\nu_2(x)dx\leq M\) implies
-\begin{equation}\label{local_lipschitz}
-\int_\mathbb{R}\left|m(\nu_1,x)\nu_1(x)-m(\nu_2,x)\nu_2(x)\right|dx\leq L_M\int_\mathbb{R}|\nu_1(x)-\nu_2(x)|dx.
-\end{equation} In SM \S\ref{finite} we combine the above assumptions on
-growth rate, mutation and initial condition to prove that solutions to
-(\ref{eq1}) satisfy \(N(t),|\bar x(t)|,\sigma^2(t)<+\infty\) for all
-\(t\geq0\).
-
-Interpreting the trait value \(x\) as spatial location and ignoring
-nonlocal effects, which can be satisfied by setting \(K\nu=\nu\), this
-model of spatially distributed population dynamics has been intensely
-studied (for a review, see Cantrell and Cosner 2004). When \(\mu=0\),
-equation (\ref{eq1}) can be seen as an analog of equation (\ref{first})
-for a continuum of types. By assuming mutation acts as diffusion, the
-effect of mutation causes \(\nu(x,t)\) to flatten out over time. In
-fact, if the growth rate is constant across \(x\), then this model of
-mutation will cause \(\nu(x,t)\) to converge to a flat line as
-\(t\to\infty\).
-
-Although clearly an idealized representation of biological reality, this
-model is sufficiently general to capture a large class of dynamics
-including density dependent growth and frequency dependent selection. As
-an example, logistic growth combined with quadratic stabilizing
-selection can be captured using the growth rate
-\begin{equation}\label{special_m}
-m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_{-\infty}^{+\infty}\nu(y,t)dy
-\end{equation} where \(r\in\mathbb{R}\) is the intrinsic rate of growth
-in the absence of abiotic stabilizing selection, \(a>0\) is the strength
-of abiotic stabilizing selection around the phenotypic optimum
-\(\theta\in\mathbb{R}\), \(c>0\) captures the sensitivity of fitness to
-intraspecific competition and \(\kappa(x-y)=1\) so that competitive
-interactions cause the same reduction in fitness regardless of trait
-value. This exemplary fitness function has a few convenient properties.
-First, the effect of competition induces a local carrying capacity on
-the population, leading to a finite equilibrium abundance over bounded
-subsets of \(\mathbb{R}\). Second, abiotic selection prevents the
-abundance density from diffusing too far from the abiotic optimum,
-leading to a finite equilibrial abundance across all of \(\mathbb{R}\),
-\(\lim_{t\to\infty}N(t)=\tfrac{1}{c}(r-\tfrac{1}{2}\sqrt{a\mu})\),
-finite equilibrial mean trait \(\lim_{t\to\infty}\bar x(t)=\theta\) and
-finite equilibrial phenotypic variance
-\(\lim_{t\to\infty}\sigma^2(t)=\sqrt{\frac{\mu}{a}}\), so long as
-\(\bar x(0)\in\mathbb{R}\), \(\sigma^2(0)\in[0,+\infty)\),
-\(N(0)\in(0,+\infty)\) and \(r>\frac{1}{2}\sqrt{a\mu}>0\). We prove
-these statements in SM \S\ref{equilib}. Note in particular the
-coincidence of equilibrial phenotypic variance predicted by this model
-and that predicted by the classic Gaussian allelic model of quantitative
-genetics (Lande 1975; Bürger 2000; Johnson and Barton 2005). In
-\S\ref{coev} and SM \S\ref{diffuse} we find this fitness function is
-closely related to a fitness function derived from niche theory that we
-use to model coevolution driven by resource competition.
-
-To derive a replicator-mutator equation from equation (\ref{eq1}), we
-employ the chain rule from calculus. Writing
-\(\bar m(t)=\int_{-\infty}^{+\infty}m(\nu,x)p(x,t)dx\) for the mean
-fitness, we have
-
-\begin{multline}\label{N_det}
-\frac{d}{dt}N(t)=\frac{d}{dt}\int_{-\infty}^{+\infty}\nu(x,t)dx=\int_{-\infty}^{+\infty}\frac{\partial}{\partial t}\nu(x,t)dx \\
-=\int_{-\infty}^{+\infty} m(\nu,x)\nu(x,t)dx+\int_{-\infty}^{+\infty}\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)dx \\
-=N(t)\int_{-\infty}^{+\infty} m(\nu,x)p(x,t)dx=\bar m(t) N(t).
-\end{multline} Using our assumptions on mutation and rate of growth, we
-show in SM \S\ref{finite} \(\nu(x,t)\) is twice differentiable with
-respect to \(x\) and \(\int_{-\infty}^{+\infty}\nu(x,t)dx<\infty\) for
-all \(t\geq0\). This implies that we are justified in swapping the order
-of differentiation and integration and the result
-\(\int_{-\infty}^{+\infty}\frac{\partial^2}{\partial x^2}\nu(x,t)dx=0\)
-can be derived from the fundamental theorem of calculus. Biological
-reasoning agrees with this latter result since mutation neither creates
-nor destroys individuals, but merely changes their type from their
-parental type. Taking the same approach, we derive the dynamics of the
-phenotypic distribution \(p(x,t)\) in response to selection and muation
-as
-
-\begin{multline}\label{p_PDE}
-\frac{\partial}{\partial t}p(x,t)=\frac{\partial}{\partial t}\frac{\nu(x,t)}{N(t)}=\frac{1}{N^2(t)}\left(N(t)\frac{\partial}{\partial t}\nu(x,t)-\nu(x,t)\frac{d}{dt}N(t)\right) \\
-=\frac{1}{N(t)}\Big(m(\nu,x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)-\bar m(t)\nu(x,t)\Big) \\
-=\Big(m(\nu,x)-\bar m(t)\Big) p(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t).
-\end{multline} This result closely resembles Kimura's
-continuum-of-alleles model (Kimura 1965; Bürger 1986). The primary
-difference being that our model utilizes diffusion instead of
-convolution with an arbitrary mutation kernel. Of course, our model of
-mutation can be derived as an approximation to Kimura's model, which,
-for sexually reproducing populations, has been referred to as the
-Gaussian allelic approximation in reference to the distribution of
-mutational effects at loci in a genome on the values of traits (Lande
-1975; but see also Bürger 2000; along with Johnson and Barton 2005), the
-infinitesimal model in reference to modelling continuous traits as being
-encoded by an infinite number of loci each having infinitesimal effect
-(Barton, Etheridge, and Véber 2017) and the Gaussian descendants
-approximation in reference to offspring trait values being normally
-distributed around their parental values (Turelli 2017). Alternatively,
-since diffusion is the continuous-time equivalent to convolution against
-a Gaussian kernel (SM \S\ref{diffconvequiv}), equation (\ref{p_PDE}) can
-also be seen as a special case of Kimura's continuum-of-alleles model.
-We return to these concepts in \S\ref{inheritance} where we review a
-simple model of inheritance that leads us to an expression for the
-stochastic evolution of additive genetic variance. To distinguish this
-model from previous models of phenotypic evolution we refer to PDE
-(\ref{eq1}) from which (\ref{p_PDE}) was derived as the Deterministic
-Asexual Gaussian allelic model with Abundance dynamics (abbreviated
-DAGA). Later, we will extend this model to include the effects of
-demographic stochasticity, which we refer to as the Stochastic Asexual
-Gaussian allelic model with Abundance dynamics (abbreviated SAGA).
-
-The covariance of fitness and phenotype across the population is defined
-as \begin{equation}
-\mathrm{Cov}_t\Big(m(\nu,x),x\Big)=\int_{-\infty}^{+\infty}\Big(m(\nu,x)-\bar m(t)\Big)\Big(x-\bar x(t)\Big)p(x,t)dx.
-\end{equation} Hence, the dynamics of the mean trait \(\bar x(t)\) can
-be derived as
-
-\begin{multline}\label{detmean}
-\frac{d}{dt}\bar x(t)=\frac{d}{dt}\int_{-\infty}^{+\infty} x p(x,t)dx=\int_{-\infty}^{+\infty} x\frac{\partial}{\partial t} p(x,t)dx \\
-=\int_{-\infty}^{+\infty} x\big(m(\nu,x)-\bar m(t)\big) p(x,t)+x\frac{\mu}{2}\frac{\partial^2}{\partial x^2} p(x,t)dx
-=\mathrm{Cov}_t\Big(m(\nu,x),x\Big).
-\end{multline}
-
-Equation (\ref{detmean}) is a continuous time analog of the well known
-Robertson-Price equation without transmission bias (Robertson 1966;
-Price 1970; Frank 2012; Queller 2017; Lion 2018). Whether or not the
-covariance of fitness and phenotype creates change in \(\bar x\) to
-maximize mean fitness \(\bar m\) depends on the degree to which
-selection is frequency dependent (Lande 1976). Since this change is
-driven by a covariance with respect to phenotypic diversity, the
-response in mean trait to selection is mediated by the phenotypic
-variance. In particular, when \(\sigma^2=0\), \(\bar x\) will not
-respond to selection. The result
-\(\int_{-\infty}^{+\infty} x\frac{\partial^2}{\partial x^2} p(x,t)dx=0\)
-can be found by applying integration by parts. Following the approach
-taken to calculate the evolution of \(\bar x\), we find the response of
-phenotypic variation to this model of selection and mutation is
-\begin{equation}\label{vardyn_det}
-\frac{d}{dt}\sigma^2(t)=\mathrm{Cov}_t\Big(m(\nu,x),(x-\bar x)^2\Big)+\mu.
-\end{equation} For the sake of space we relegate the derivation of
-\(d\sigma^2/dt\) to SM \S\ref{var_deriv}. In the absence of mutation
-equation (\ref{vardyn_det}) agrees with the result derived by Lion
-(2018) for discrete phenotypes. From a statistical perspective, if we
-think of \((x-\bar x)^2\) as a square error, then in analogy to the
-dynamics of the mean trait, we see that the response in \(\sigma^2\) to
-selection can be expressed as a covariance of fitness and square error,
-which is defined in analogy to \(\mathrm{Cov}_t(m(\nu,x),x)\). Just as
-for the evolution of \(\bar x(t)\), this covariance also creates change
-in \(\sigma^2\) that can either increase or decrease mean fitness
-\(\bar m\), depending on whether or not individual fitness depends on
-the distribution of traits across the entire population. The effect of
-selection on phenotypic variance can be positive or negative depending
-on whether selection is stabilizing or disruptive.
-
-In SM \S\ref{SDE_DERIV}, we extend these results to include the effects
-of demographic stochasticity. The key realization in this direction is
-that a rigorous derivation of a SPDE that naturally generalizes DAGA and
-accounts for ecological interactions has been provided by Li (1998) and
-slightly generalized by Champagnat, Ferriére and Méléard (2006). To
-understand how this SPDE was derived, we provide an informal discussion
-of measure-valued branching processes (which treat populations as sets
-of discrete individuals) and their diffusion limits, the so-called
-superprocesses (which treat population size as a continuously varying
-number). Similar to the approach taken by Champagnat, Ferriére and
-Méléard (2006), we begin with a BBM that models populations as discrete
-sets of reproducing individuals whose vital rates depend on their trait
-value as well as the state of the entire population. Taking a large
-population size limit and keeping our assumption of single dimensional
-traits, we employ a pair of classical results that show, under the
-appropriate rescaling in time, phenotypic space and population density,
-a sequence of rescaled BBM converges to a limiting process that can be
-characterized by a SPDE (Méléard and Roelly 1992, 1993; Li 1998;
-Champagnat, Ferrière, and Méléard 2006). The limiting processes of
-rescaled BBM have been referred to as measure-valued Markov processes
-(Dawson 1993) or superprocesses (Etheridge 2000). Under the simplifying
-assumptions inherited from our treatment of deterministic dynamics and
-the additional assumption that the magnitude of demographic
-stochasticity is independent of trait values, we obtain as a special
-case a relatively simple expression for an SPDE that generalizes DAGA.
-The simplicity of our special case allows us to use properties of
-space-time white noise processes to derive a set of SDE that generalize
-equations (\ref{N_det}), (\ref{detmean}) and (\ref{vardyn_det}) to
-include the effects of demographic stochasticity. Classical expressions
-for the effects of random genetic drift on the evolution of mean traits
-are obtained as a further special case.
-
-In the following section we provide the necessary mathematical tools
-needed to derive SDE from SPDE. Since our aim is to present this
-material to as wide of audience as possible, our approach deviates from
-standard definitions to remove the need for a detailed technical
-treatment. In addition to the notions of Reimann integration and partial
-differentiation already employed, the reader will only need some
-elementary probability and an intuitive understanding of SDE, including
-Brownian motion. Because space-time white noise, denoted by
-\(\dot W(x,t)\), appears in the SPDE characterizing diffusion limits of
-BBM, we begin by defining \(\dot W(x,t)\) and illustrating its relevant
-properties including a set of heuristics for performing calculations.
-Treating only the simplist of cases, we then provide a brief review of
-BBM, their diffusion limits and the SPDE that characterize them. For
-those not interested in the white noise calculus or superprocesses and
-would rather jump straight into more biologically relevant results, we
-recommend skipping to \S\ref{equations}.
-
-\hypertarget{white-noise-calculus-and-superprocesses}{%
-\subsection{\texorpdfstring{White noise calculus and superprocesses
-\label{wnc}}{White noise calculus and superprocesses }}\label{white-noise-calculus-and-superprocesses}}
-
-\hypertarget{introduction-to-white-noise}{%
-\subsubsection{Introduction to white
-noise}\label{introduction-to-white-noise}}
-
-One can think of white noise as the static seen on old television sets
-or infinitely detailed random dust spread across both time and space.
-From a more mathematical, yet still informal perspective, white noise
-can be thought of as a stochastic process. That is, we can picture white
-noise as a collection of random variables indexed by time and possibly
-space. In relation to Brownian motion, denoted by \(W\), white noise can
-be interpreted of as the derivative of Brownian motion with respect to
-time, denoted \(\dot W\). Since Brownian motion can be thought to take
-infinitesimally small Gaussian distributed jumps at each time point,
-this leads to the conceptualization of white noise as a collection of
-independent Gaussian distributed random variables. Figure \ref{wn}
-illustrates realizations of this conceptualized white noise in one
-(left) and two (right) dimensions.
-
-\begin{figure}
-
-{\centering \includegraphics{wn}
-
-}
-
-\caption{\label{wn}Approximations of sample paths of temporal white noise (left) and space-time white noise (right) with brightness scaled to value.}\label{fig:unnamed-chunk-2}
-\end{figure}
-
-However, it turns out that realizations of white noise do not exist as
-functions in the classical sense. Indeed, since Brownian motion is
-nowhere differentiable with respect to time, white noise cannot be
-formally understood as a time derivative. Thus our notation \(\dot W\)
-is only meant to aid intuition and not be taken as formal. A formal
-understanding is possible by considering white noise as a
-\emph{measure}-valued process (Dawson 1975; Walsh 1986) or as a
-\emph{generalized} process that acts on classically defined functions or
-stochastic processes to return either random variables or stochastic
-processes (Krylov and Rozovskii 1981; Da Prato and Zabczyk 2014). Since
-a measure-valued process can be defined from a generalized process and a
-generalized process can be defined from a measure-valued process, the
-distinction between the two is more or less a matter of perspective.
-However, we find the perspective of white-noise as a generalized process
-to be a more efficient route for developing heuristics to help with some
-routine calculations involved with deriving SDE from SPDE. Hence, the
-notion of a generalized process provides the general idea implemented
-here. Although the treatments of SPDE provided by Krylov and Rozovskii
-(1981) and Da Prato and Zabczyk (2014) extend the theory of SDE to
-formally treat SPDE in a general and elegant fashion, they require the
-navigation of an enourmous amount of technical definitions and detailed
-proofs. To extract some particularly useful results from this theory
-relevant to our goal of synthesizing the stochastic dynamics of
-biological populations, we provide a streamlined approach by
-capitolizing on the solid ground these authors have established. For
-instance, instead of rigorously proving properties of white noise, we
-simply define them to be so, taking solice in the fact that the
-technical details have been worked out elsewhere. In SM \S\ref{DPZ14} we
-show how our informal treatment is related to the rigorous treatment
-provided by Da Prato and Zabczyk (2014).
-
-Before diving in, we shed a bit of light on the idea of a generalized
-process. A generalized process is the stochastic analog of a generalized
-function, such as the Dirac delta function \(\delta\). Often one sees
-\(\delta\) defined as a function satisfying the properties
-\(\delta(x)=0\) for \(x\neq0\) and
-\(\int_{-\infty}^{+\infty}\delta(x)dx=1\). However, since there is no
-function that satisfies these properties, we refer to \(\delta\) as a
-generalized function. An alternative definition of \(\delta\) considers
-its action on classically defined functions \(f\). In particular,
-\(\delta(f)=f(0)\), which can be heuristically written as
-\(\int_{-\infty}^{+\infty}f(x)\delta(x)dx=f(0)\). Similarly, other
-generalized functions can be defined by their action on classically
-defined functions. Then, just as a generalized function operates on
-classical functions to return a value, a generalized process acts on a
-set of functions (or processes) to return a random variable (or a
-classically defined stochastic process). For a brief primer on the
-theory of generalized functions, see the addendum to chapter 3 of
-Kolmogorov and Fomin (1999).
-
-\hypertarget{definition-and-basic-properties-of-white-noise}{%
-\subsubsection{\texorpdfstring{Definition and basic properties of white
-noise
-\label{wnc_intro}}{Definition and basic properties of white noise }}\label{definition-and-basic-properties-of-white-noise}}
-
-Throughout this section, we minimize notation by writing
-\(\int_{\mathbb{R}}f(x)dx=\int_{-\infty}^{+\infty}f(x)dx\) and similarly
-\(\int_Df(x)dx\) for the integral of \(f\) over \(D\subset\mathbb{R}\).
-We define \(\mathscr{N}_2\) as the set of stochastic processes
-\(f(x,t)\) that are continuous in \(t\) and satisfy
-\(\mathbb{E}\left(\int_0^t\int_\mathbb{R}f^2(x,s)dxds\right)<+\infty\)
-for each \(t\geq0\). The operator \(\mathbb{E}\) denotes expectation
-with respect to the underlying probability space. For each \(t\geq0\) we
-set
-
-\begin{equation}
-\|f\|_t=\sqrt{\mathbb{E}\left(\int_0^t\int_\mathbb{R}f^2(x,s)dxds\right)},
-\end{equation}
-
-and make use of the convention \(f=g\) if \(\|f-g\|_t=0\) for all
-\(t\geq0\). Later, when we investigate stochastic dynamics of the
-abundance density \(\nu(x,t)\), we will find
-\(\sqrt\nu\in\mathscr{N}_2\). This enables us to utilize the heuristics
-developed in this section for the derivation of SDE describing the
-stochastic dynamics of \(N(t)\), \(\bar x(t)\) and \(\sigma^2(t)\).
-
-We define a generalized stochastic process \(\mathbf W\) that maps
-processes \(f\in\mathscr{N}_2\) to real-valued stochastic processes
-indexed by time \(t\geq0\), but not by space. To evaluate \(\mathbf W\)
-for a process \(f\in\mathscr{N}_2\) and some time \(t\geq0\) we write
-\(\mathbf W_t(f)\). Specifically, for any \(f,g\in\mathscr{N}_2\), we
-define \(\mathbf W(f)\) and \(\mathbf W(g)\) to be Gaussian processes
-satisfying, for any \(t,t_1,t_2\geq0\), \begin{equation}\label{exp_WN}
-\mathbb{E}\big(\mathbf W_t(f)\big)=\mathbb{E}\big(\mathbf W_t(g)\big)=0,
-\end{equation} \begin{equation}\label{cov_WN}
-\mathbb{C}\big(\mathbf W_{t_1}(f),\mathbf W_{t_2}(g)\big)=\mathbb{E}\left(\int_0^{t_1\wedge t_2}\int_\mathbb{R} f(x,s)g(x,s)dxds\right),
-\end{equation} where \(t_1\wedge t_2=\min(t_1,t_2)\) and \(\mathbb{C}\)
-denotes covariance with respect to the underlying probability space. In
-particular, denoting \(\mathbb{V}\) the variance operator with respect
-to the underlying probability space, we have
-\(\mathbb{V}\big(\mathbf W_t(f)\big)=\|f\|_t^2\) for all \(t\geq0\) and
-\(f\in\mathscr{N}_2\). The operators \(\mathbb{E}\) and \(\mathbb{C}\)
-are to be distinguished from \(\bar f(t)\) and \(\mathrm{Cov}_t(f,g)\)
-which denote expectation and covariance with respect to phenotypic
-diversity at time \(t\geq0\).
-
-Since Gaussian processes are characterized by their expectations and
-covariances and since we assume the \(\mathscr{N}_2\) processes are
-continuous in time, the processes \(\mathbf W(f)\) and \(\mathbf W(g)\)
-are well defined. As an example, if \(f\in\mathscr{N}_2\) is independent
-of time, then \(\mathbf W(f)\) is a Brownian motion with variance at
-time \(t\geq0\) equal to
-\(\|f\|_t^2=t \ \mathbb{E}(\int_\mathbb{R}f^2(x,0)dx)\). With the
-generalized process \(\mathbf W\) defined, we define the space-time
-white noise \(\dot W(x,t)\) implicitly via the stochastic integral
-\begin{equation}\label{informal}
-``\int_0^t\int_\mathbb{R}f(x,s)\dot W(x,s)dxds"=\mathbf W_t(f), \ \forall \ f\in\mathscr{N}_2, \ t\geq0.
-\end{equation} We place quotations in the above expression to emphasize
-its informal nature and that it should not be confused with classical
-Riemann integration. Following this definition of white noise, we
-compute its value by sampling it using \(\mathscr{N}_2\) processes. For
-example, integrating white noise over a region \(D\times[0,t]\), with
-\(t>0\) and \(D\) a bounded subset of \(\mathbb{R}\), is equivalent to
-evaluating \(\mathbf W_t(I_D)\) for the deterministic process
-\begin{equation}
-I_D(x,t)=\left\{\begin{matrix}
-0, & x\notin D \\
-1, & x\in D
-\end{matrix}\right..
-\end{equation} Since \begin{equation}
-\|I_D\|_t^2=\mathbb{E}\left(\int_0^t\int_\mathbb{R}I_D^2(x,s)dxds\right)=t\int_Ddx=t|D|<+\infty, 
-\end{equation} where \(|D|\) denotes the length of \(D\), we have
-\(I_D\in\mathscr{N}_2\). Thus, using equations (\ref{exp_WN}) and
-(\ref{cov_WN}) and adopting the informal notation introduced in equation
-(\ref{informal}), we can write the following \begin{equation}
-\mathbb{E}\left(\int_0^t\int_D\dot W(x,s)dxds\right)=0,
-\end{equation} \begin{equation}
-\mathbb{V}\left(\int_0^t\int_D\dot W(x,s)dxds\right)=t|D|.
-\end{equation} Using this informal notation, equations (\ref{exp_WN})
-and (\ref{cov_WN}) can be rewritten as \begin{equation}\label{exp_WN_xi}
-\mathbb{E}\left(\int_0^t\int_\mathbb{R}f(x,s)\dot W(x,s)dxds\right)=0,
-\end{equation} \begin{equation}\label{cov_WN_xi}
-\mathbb{C}\left(\int_0^{t_1}\int_\mathbb{R}f(x,s)\dot W(x,s)dxds,\int_0^{t_2}\int_\mathbb{R}g(x,s)\dot W(x,s)dxds\right)
-=\int_0^{t_1\wedge t_2}\int_\mathbb{R}f(x,s)g(x,s)dxds.
-\end{equation} To relate these formula to the common notation used for
-SDE, we write \begin{equation}
-\hat f(x,t)=\frac{f(x,t)}{\sqrt{\int_\mathbb{R}f^2(y,t)dy}} \ \text{ and } \ 
-d\hat{\mathbf W}_t(f)=\left(\int_\mathbb{R}\hat f(x,t)\dot W(x,t)dx\right)dt
-\end{equation} so that \begin{equation}
-\int_0^td\hat{\mathbf W}_s(f)=\int_0^t\int_\mathbb{R}\frac{f(x,s)}{\sqrt{\int_\mathbb{R}f^2(s,y)dy}}\dot W(x,s)dxds.
-\end{equation} This implies \begin{equation}
-\mathbb{E}\left(\int_0^td\hat{\mathbf W}_s(f)\right)=0, \ \mathbb{C}\left(\int_0^{t_1}d\hat{\mathbf W}_s(f),\int_0^{t_2}d\hat{\mathbf W}_s(f)\right)=t_1\wedge t_2
-\end{equation} and in particular, as a function of \(t\),
-\(\int_0^td\hat{\mathbf W}_s(f)\) is a standard Brownian motion for any
-\(f\in\mathscr{N}_2\). Hence, \(d\hat{\mathbf W}_t(f)\) is analogous to
-the traditional shorthand used to denote stochastic differentials. Thus,
-equation (\ref{cov_WN_xi}) effectively extends Itô's multiplication
-table to:
-
-\begin{table}[H]
-\centering\caption{An extension of It\^o's multiplication table.}\vspace{0.2cm}
-\begin{tabular}{l|lll}
-             & $d\hat{\mathbf W}_t(f)$               & $d\hat{\mathbf W}_t(g)$                & $dt$ \\ \hline
-             &                            &                            &      \\
-$d\hat{\mathbf W}_t(f)$ & $dt$                       & $\left(\int_\mathbb{R}\hat f(x,t)\hat g(x,t)dx\right)dt$ & $0$  \\
-             &                            &                            &      \\
-$d\hat{\mathbf W}_t(g)$ & $\left(\int_\mathbb{R}\hat f(x,t)\hat g(x,t)dx\right)dt$ & $dt$                       & $0$  \\
-             &                            &                            &      \\
-$dt$         & $0$                        & $0$                        & $0$
-\end{tabular}\label{mult}
-\end{table}
-
-The extension of Itô's multiplication table and properties of white
-noise outlined in this subsection provide a useful set of tools for
-working with SPDE. In SM \S\ref{SDE_DERIV} we employ these tools to
-derive SDE that track the dynamics of abundance, mean trait and
-phenotypic variance of a population from a particular SPDE. In the
-following subsection, we review how this particular SPDE naturally
-arises as the diffusion limit of a BBM.
-
-\hypertarget{from-branching-processes-to-spde}{%
-\subsubsection{\texorpdfstring{From branching processes to SPDE
-\label{stochastic}}{From branching processes to SPDE }}\label{from-branching-processes-to-spde}}
-
-In real populations individuals are born and potentially reproduce
-before they ultimately die. These three events provide the basic
-ingredients of a branching process. Mathematical investigations of such
-processes have a relatively deep history (Kendall 1966). The most simple
-branching process, known as the Galton-Watson process, describes the
-number of individuals alive at a given time \(t\geq0\) as a non-negative
-integer (Kimmel and Axelrod 2015). Feller (1951) introduced a formal
-method to approximate branching processes with diffusion processes which
-are continuous in state (i.e., population size is approximated as a
-continuous quantity). Since diffusion processes possess greater
-analytical tractability than branching processes, Feller's method, known
-as the diffusion limit, has acquired immense popularity particularly in
-the field of mathematical population genetics (Ewens 2004). For over the
-past half of a century a great deal of accomplishments have been
-achieved in formalizing the diffusion limits of branching processes that
-describe populations of individuals occuring in some continuous space
-(Watanabe 1968; Dawson 1975, 1978; Perkins 1992, 1995; Méléard and
-Roelly 1993; Li 1998; Bertoin and Le Gall 2003; Etheridge 2008; Barton
-and Etheridge 2019). This space can represent geographic space or,
-relevant to our context, phenotypic space. In the following subsection,
-we describe the BBM process, which is a particularly important branching
-process with spatial structure. This process has been very useful in the
-study of SPDE due to its simplifying assumption that individuals do not
-interact. However, this assumption imposes an unfortunate restriction by
-precluding the modelling of ecological interactions. We therefore follow
-our discussion of BBM with a review of a few important results on
-spatially structured branching processes that account for interactions.
-
-\textbf{Branching Brownian motion}
-
-A BBM tracks individuals navigating \(d\)-dimensional Euclidean space
-that reproduce and senesce between exponentially distributed intervals.
-Unlike other stochastic processes that take values in \(\mathbb{R}^d\),
-BBM takes values in the set of \emph{non-negative finite measures} over
-\(\mathbb{R}^d\). Intuitively, one can think of a finite measure as a
-function that maps subsets of \(\mathbb{R}^d\) to real numbers. To be
-formal, we only consider the Borel subsets of \(\mathbb{R}^d\)
-corresponding to the Euclidean metric, but understanding this
-technicality is not crucial to our discussion. In particular, denoting
-\(X_t\) a BBM, for a subset \(D\subset\mathbb{R}^d\), \(X_t(D)\) returns
-the (random) number of individuals alive within the region \(D\) at time
-\(t\geq0\). The BBM has three main components:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi})}
-\item
-  \textbf{Branching rate:} In our formulation of BBM we assume Lifetimes
-  of individuals are exponentially distributed with death rate
-  \(\lambda>0\) and reproduction occurs simultaneously with death.
-  Biologically, this implies individuals are semelparous An alternative
-  formulation treats birth and death events separately to model
-  iteroparity. However, under the appropriate rescaling, both approaches
-  converge to the same diffusion limit (Champagnat, Ferrière, and
-  Méléard 2006). We therefore choose the former approach for the sake of
-  simplicity.
-\item
-  \textbf{Reproductive law:} When a birth event occurs we assume a
-  random (possibly zero) number of offspring are produced. The
-  distribution of offspring left is called the reproductive law or
-  branching mechanism. We denote the mean and variance in reproductive
-  output by \(\mathscr{W}\) and \(V\) respectively. The case of
-  \(\mathscr{W}=1\) is referred to as the critical condition. Since
-  \(X_t(\mathbb{R}^d)\) is a Galton-Watson branching process, it is well
-  known the critical condition implies extinction in finite time with
-  probability one, given finite \(X_0(\mathbb{R}^d)\) (Athreya and Ney
-  1972).
-\item
-  \textbf{Spatial movement:} Each offspring is born at the current
-  location of their parent. Immediately after birth they move around
-  space according to \(d\)-dimensional Brownian motion with diffusion
-  parameter \(\sqrt\mu\). In our context we interpret spatial movement
-  as mutation so that the location of an individual at death represents
-  the value of its phenotype. Then an individual born at location
-  \(x\in\mathbb{R}^d\) that lives for \(\tau>0\) units of time will have
-  a normally distributed trait centered on \(x\) with covariance matrix
-  equal to \(\tau\mu\) times the \(d\times d\) identity matrix. Hence,
-  offspring inherit normally distributed traits centered on their
-  parental trait. This fact creates a vital link to the deterministic
-  dynamics reviewed above. Indeed, in the absence of selection, the
-  deterministic PDE (\ref{eq1}) reduces to the \(d=1\)-dimensional
-  Kolmogorov forward equation for a Brownian motion with diffusion
-  parameter \(\sqrt\mu\).
-\end{enumerate}
-
-To obtain a SPDE from a BBM we take a diffusion limit. There are several
-ways to do this, but a simple approach is to rescale the mass of
-individuals and time by \(1/n\), diffusion by \(\mu\to\mu/n\), branching
-rate by \(\lambda\to n\lambda\), fitness by
-\(\mathscr{W}\to\mathscr{W}^{1/n}\) and consider the limit as
-\(n\to\infty\). Denoting the rescaled process by \(X_t^{(n)}(D)\), the
-limiting process \(\mathscr{X}_t=\lim_{n\to\infty}X_t^{(n)}\) is called
-a super-Brownian motion and is also a non-negative finite measure-valued
-process (Watanabe 1968). However, instead of returning the number of
-individuals alive in a region of space, super-Brownian motion returns
-the \emph{mass} of the population concentrated in a region of space.
-Since we have rescaled individual mass by \(1/n\) and took the limit
-\(n\to\infty\), individuals are no longer discrete units. Instead, the
-particle view of the population gets replaced by a blob spread across
-\(d\)-dimensional space. In particular, the value of
-\(\mathscr{X}_t(D)\) is a continuously varying non-negative random
-variable for any \(t\geq0\) and \(D\subset\mathbb{R}^d\).
-
-Unfortunately, just as with cream cheese spread across too much toast,
-the blob perspective of the population may exhibit some complicated
-spatial discontinuities. However, for spatial dimension \(d=1\), it
-turns out that \(\mathscr{X}_t\) is absolutely continuous with respect
-to the Lebesgue measure for each \(t\geq0\) (Konno and Shiga 1988;
-Reimers 1989). This means that we can write
-\(\mathscr{X}_t(D)=\int_D \nu(x,t)dx\) for some density process
-\(\nu(x,t)\). Setting \(\lambda=1\) and \(m=\ln\mathscr{W}\) this
-density process satisfies the SPDE \begin{equation}\label{neutral_SPDE}
-\frac{\partial}{\partial t}\nu(x,t)=m\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)+\sqrt{V\nu(x,t)}\dot W(x,t).
-\end{equation} Since \(\nu(x,t)\) is not generally differentiable in
-\(x\) or \(t\), the derivatives in expression (\ref{neutral_SPDE}) are
-taken in the \emph{weak} sense (for a definition see Evans 2010). That
-is, to rigorously interpret SPDE (\ref{neutral_SPDE}), we integrate the
-solution \(\nu(x,t)\) against functions \(f\in C_b^2(\mathbb{R})\) where
-\(C_b^2(\mathbb{R})\) is the set of bounded and twice continuously
-differentiable functions on \(\mathbb{R}\). Hence, equation
-(\ref{neutral_SPDE}) is just an abbreviation for
-
-\begin{multline}\label{neutral_int}
-\int_\mathbb{R}\nu(x,t)f(x)dx-\int_\mathbb{R}\nu(x,0)f(x)dx=\int_0^t\int_\mathbb{R}\nu(x,s)\left(mf(x)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}f(x)\right)dsdx \\
-+\int_0^t\int_\mathbb{R}f(x)\sqrt{V\nu(x,s)}\dot W(x,s)dxds, \ \forall \ f\in C_b^2(\mathbb{R}).
-\end{multline} This expression is referred to as the \emph{mild}
-solution of (\ref{neutral_SPDE}). For more on the general theory of SPDE
-see Walsh (1986) and Da Prato and Zabczyk (2014). Note that since
-\(\nu(x,t)\) is the density of a finite measure, it is integrable for
-each \(t\geq0\). Thus, since for some \(M>0\), \(|f(x)|\leq M\) for
-every \(x\in\mathbb{R}\), setting \(\varphi(x,t)=f(x)\sqrt{V\nu(x,t)}\)
-implies \(\varphi\in\mathscr{N}_2\). Hence, the white noise integral on
-the right-hand side of equation (\ref{neutral_int}) can be understood
-using the heuristics introduced above. Evaluating equation
-(\ref{neutral_int}) in the particular case of \(f(x)\equiv1\) returns
-the total mass process, which we refer to as the total abundance
-\(N(t)\).
-
-A convergence theorem for the diffusion limit of a generalization of BBM
-was established by Watanabe (1968). Dawson (1975) suggested that, for
-spatial dimension \(d=1\), this diffusion limit should admit a density
-process that satisfies a SPDE. Konna and Shiga (1988) and Reimers (1989)
-independently proved Dawson's suggestion was indeed correct. The
-diffusion limit of this more general branching process (in arbitrary
-spatial dimension) is referred to as a Dawson-Watanabe superprocess
-(Etheridge 2000). Conditioning a Dawson-Watanabe superprocess to have
-constant mass returns a Fleming-Viot process (Etheridge and March 1991;
-Perkins 1991) which has been popular in studies of spatial population
-genetics. In particular, an extension of the Fleming-Viot process, known
-as the \(\Lambda\)-Fleming-Viot process, was introduced by Bertoin and
-Le Gall (2003) and coined by Etheridge (2008) where it was used to
-resolve some technical challenges in modelling isolation by distance
-(Felsenstein 1975; see also Barton, Etheridge, and Véber 2013; and
-Barton and Etheridge 2019). Although this provides an impressive list of
-accomplishments, the Dawson-Watanabe superprocess falls short of our
-needs. In particular this process assumes individuals do not interact
-and thus precludes its ability to model nonlocal effects on the fitness
-of individuals, such as those produced via competitive interactions.
-However, this concern has been addressed, leading to constuctions of
-superprocesses that account for interactions among individuals. In the
-next subsection we summarize the main results in this area and introduce
-the SPDE that provides the basis for our approach to theoretical
-evolutionary ecology.
-
-\textbf{Interacting superprocesses}
-
-In the above subsection we reviewed the diffusion limit of an especially
-tractable measure-valued branching process. However, we found the
-simplicity of this process restricts us from modelling nonlocal effects
-on the fitness of individuals. Here, we discuss superprocesses that
-account for such effects. The existence of diffusion limits for a class
-of measure-valued branching processes involving nonlocal effects has
-been treated by Méléard and Roelly (1992, 1993). The interactions can
-manifest as dependencies of the spatial movement or reproductive law of
-individuals on their position \(x\) and the state of the whole
-population described by \(X_t\). An important result of Méléard and
-Roelly (1992, 1993) is a theorem that provides sufficient conditions to
-construct a sequence of rescaled measure-valued branching processes that
-converge to a generalization of the Dawson-Watanabe superprocess that
-includes interactions. The rescaling is analogous to that described
-above for non-interacting Dawson-Watanabe superprocesses, but now the
-reproductive law described by \(\mathscr{W}(X_t,x)\) and \(V(X_t,x)\),
-branching rate \(\lambda(X_t,x)\) and diffusion parameter
-\(\sqrt{\mu(X_t,x)}\) are allowed to depend on the whole population
-\(X_t\) and individual location \(x\). In Figure \ref{rescaled} we
-demonstrate this rescaling in discrete time for a population
-experiencing stabilizing selection and logistic growth. Since time is
-discretized, the process we simulate is formally a branching random
-walk. For further details on our simulation see SM \S\ref{numerical}.
-
-\begin{figure}
-
-{\centering \includegraphics{rescaled_plots}
-
-}
-
-\caption{\label{rescaled}Rescaled sample paths of a branching random walk under stabilizing selection and logistic growth.
-The top plot displays a sample path without scaling $(n=1)$, the middle plot shows a sample path rescaled by $n=5$ and the bottom plot shows a sample path rescaled by $n=20$.}\label{fig:unnamed-chunk-4}
-\end{figure}
-
-Interactions that manifest in the spatial movement can be used to model
-mutation bias and those manisfesting in the reproductive law can model
-density-dependent growth rates and frequency-dependent selection.
-Perkins (1992, 1995) developed a theory of stochastic integration with
-respect to the so-called \emph{Brownian trees} to characterize
-interacting superprocesses and establish properties of existence and
-uniqueness. Li (1998) built directly off of the construction of Méléard
-and Roelly (1992, 1993) to study properties of interacting
-superprocesses and, by assuming individual spatial movement occurs
-independently of location \(x\) and the entire population \(X_t\),
-showed the evolution of associated density processes can be described by
-SPDE.
-
-Recall \(\nu(x,t)\) denotes the density of a superprocess, given it
-exists. Assuming the interactions manifest only in the reproductive law
-and that spatial movement follows Brownian motion with diffusion
-parameter \(\sqrt\mu\) independent of both \(X_t\) and \(x\) and the
-growth rate is bounded both above and below, Li's (1998) result implies
-the interacting superprocess on one dimensional trait space has a
-density \(\nu(x,t)\) which is non-negative, integrable, continuous in
-time and space and satisfies the SPDE \begin{equation}\label{SPDE}
-\frac{\partial}{\partial t}\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)+\sqrt{V\nu(x,t)}\dot W(x,t).
-\end{equation} As mentioned above, we refer to SPDE (\ref{SPDE}) as the
-Stochastic Asexual Gaussian allelic model with Abundance dynamics
-(abbreviated SAGA). Comparing equation (\ref{SPDE}) to equation (3.5) of
-Li (1998), our \(m\) and \(V\) correspond to Li's \(b\) and \(c\)
-respectively. Using techniques developed by Evans and Perkins (1994),
-Champagnat, Ferrière and Méléard (2006) generalized this result to the
-case when \(m(h,x)\) is bounded above and is linear or more than linear
-in \(h\). In particular, SAGA corresponds to a special case of equation
-(4.6) in Champagnat, Ferrière and Méléard (2006) and is valid for the
-assumptions on \(m\) made in \S\ref{deterministic}.
-
-With solutions to SAGA well defined, we can calculate the total mass
-process \(N(t)\) using the mild solution of (\ref{SPDE}) with
-\(f(x)\equiv1\in C_b^2(\mathbb{R})\) (the symbol ``\(\equiv\)'' means
-equal to for every \(x\)). That is,
-
-\begin{multline}
-N(t)=N(0)+\int_0^t\int_\mathbb{R}\nu(x,s)\left(m(\nu,x)\cdot1+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}1\right)+1\sqrt{V\nu(x,s)}\dot W(x,s)dsdx \\
-=N(0)+\int_0^t\bar m(s)N(s)dt+\int_0^t\sqrt{VN(s)}d\hat{\mathbf{W}}_s(\sqrt{\nu(x,s)}),
-\end{multline} where \begin{equation}
-\bar m(t)=\frac{1}{N(t)}\int_\mathbb{R}m(\nu,x)\nu(x,t)dx,
-\end{equation} and \begin{equation}
-\int_0^td\hat{\mathbf{W}}_s(\sqrt{\nu(x,s)})=\int_0^t\int_\mathbb{R}\frac{\sqrt{\nu(x,s)}}{\sqrt{\int_\mathbb{R}\nu(x,s)dx}}\dot W(x,s)dxds.
-\end{equation} Setting \(W_1(t)=\hat{\mathbf{W}}_t(\sqrt{\nu(x,t)})\),
-we can use traditional stochastic differential notation to write
-\begin{equation}
-dN=\bar mNdt+\sqrt{VN}dW_1.
-\end{equation} To find the associated SDE for \(\bar x(t)\) and
-\(\sigma^2(t)\), we want to repeat the same approach for \(f(x)=x,x^2\)
-and apply Itô's lemma. However, for these cases
-\(f\notin C^2_b(\mathbb{R})\) since \(f\) will not be bounded. But if we
-can show \(\int_\mathbb{R}(|x|+x^2+x^4)\nu(x,t)dx<+\infty\) for all
-\(t>0\) given this condition is satisfied by \(\nu(x,0)\), then we can
-apply the mild solution of (\ref{SPDE}) to derive SDE for \(\bar x(t)\)
-and \(\sigma^2(t)\). To illustrate, let us suppose this is the case.
-Setting \(\tilde x(t)=\int_\mathbb{R}x\nu(x,t)dx\), we have
-\begin{equation}
-\tilde x(t)=\tilde x(0)+\int_0^t\int_\mathbb{R}\nu(x,s)m(\nu,x)x+x\sqrt{V\nu(x,s)}\dot W(x,s)dxds.
-\end{equation} Similarly, setting
-\(\tilde{\tilde\sigma}^2(t)=\int_\mathbb{R}x^2\nu(x,t)dx\), we have
-\begin{equation}
-\tilde{\tilde\sigma}^2(t)=\tilde{\tilde\sigma}^2(0)+\int_0^t\int_\mathbb{R}\nu(x,s)\left(m(\nu,x)x^2+\mu\right)+x^2\sqrt{V\nu(x,s)}\dot W(x,s)dxds.
-\end{equation} Since \(\bar x(t)=\tilde x(t)/N(t)\) and
-\(\sigma^2(t)=\tilde{\tilde\sigma}^2(t)/N(t)-\bar x^2(t)\), we can use
-Itô's lemma to derive SDE for \(\bar x(t)\) and \(\sigma^2(t)\), which
-we perform in SM \S\ref{SDE_DERIV}. We make no attempt in finding
-sufficient conditions to ensure
-\(\int_\mathbb{R}(|x|+x^2+x^4)\nu(x,t)dx<+\infty\) and hence make no
-general assertions about the existence or uniqueness of \(\bar x(t)\) or
-\(\sigma^2(t)\). Regardless, we will later assume \(\nu(x,t)\) can be
-approximated by a Gaussian curve in \(x\) for all \(t\geq0\). This
-assumption implies \(\int_\mathbb{R}|x|^n\nu(x,t)dx<+\infty\) for all
-\(n\in\mathbb{N}\) and for all \(t\geq0\).
-
-\hypertarget{equations-of-evolutionary-and-demographic-dynamics}{%
-\subsection{\texorpdfstring{Equations of evolutionary and demographic
-dynamics
-\label{equations}}{Equations of evolutionary and demographic dynamics }}\label{equations-of-evolutionary-and-demographic-dynamics}}
-
-In SM \S\ref{SDE_DERIV} we show SDE for \(N(t)\), \(\bar x(t)\) and
-\(\sigma^2(t)\) can be expressed as
-
-\begin{subequations}\label{gen_eqns}
-\begin{equation}\label{N}
-dN(t)=\bar m(t)N(t)dt+\sqrt{V N(t)}dW_1(t),
-\end{equation}
-\begin{equation}\label{xbar_gen}
-d\bar x(t)=\mathrm{Cov}_t\Big(x,m(\nu,x)\Big)dt+\sqrt{V\frac{\sigma^2(t)}{N(t)}}dW_2(t),
-\end{equation}
-\begin{equation}\label{sig2_gen}
-d\sigma^2(t)=\mathrm{Cov}_t\Big((x-\bar x(t))^2,m(\nu,x)\Big)dt+\left(\mu-V\frac{\sigma^2(t)}{N(t)}\right)dt+\sqrt{V\frac{\overline{(x-\bar x(t))^4}-\sigma^4(t)}{N(t)}}dW_3(t),
-\end{equation}
-\end{subequations}
-
-where \(W_1\), \(W_2\) and \(W_3\) are standard Brownian motions. We
-note that conditions on the growth rate \(m\) to guarantee existence and
-uniqueness of solutions to (\ref{xbar_gen}) and (\ref{sig2_gen}) have
-yet to be investigated. However, our results on the deterministic PDE
-suggest that \(m(h,x)\) bounded above and differentiable in both
-arguments is sufficient. Dividing by \(dt\) one can interpret equations
-(\ref{gen_eqns}) as if they are ordinary differential equations, but
-this not technically rigorous since Brownian motion is nowhere
-differentiable with respect to time. In SM \S\ref{SDE_DERIV} we show
-that in general \(W_1\) is independent of both \(W_2\) and \(W_3\), but
-\(W_2\) and \(W_3\) covary.
-
-There is quite a bit we can learn from expressions (\ref{gen_eqns}).
-Firstly, setting \(V=0\) recovers the deterministic dynamics derived in
-\S\ref{deterministic}. Alternatively, one can take \(N(t)\to\infty\) to
-recover the deterministic dynamics for \(\bar x(t)\) and
-\(\sigma^2(t)\). Characteristically, we note the effect of demographic
-stochasticity on abundance grows with \(\sqrt{N(t)}\). Hence, dividing
-by \(N(t)\), we find the effects of demographic stochasticity on the
-per-capita growth rate diminish with increased abundance. Relating the
-response to demographic stochasticity derived here to the effect of
-random genetic drift derived in classic quantitative genetic theory, if
-we set \(\sigma^2(t)=\sigma^2\) and \(N(t)=N\) constant with respect to
-time, then integrating the stochastic term in equation (\ref{xbar_gen})
-over a single unit of time returns a normally distributed random
-variable with mean zero and variance equal to \(V\sigma^2/N\). In
-particular, assuming perfect inheritance, when reproductive variance is
-unity (\(V=1\)) this random variable coincides with the effect of random
-genetic drift on the change in mean trait over a single generation
-derived using sampling arguments (Lande 1976). There is also an
-interesting connection with classical population genetics. A fundamental
-result from early population genetic theory is the expected reduction in
-diversity due to the chance loss of alleles in finite populations
-(Fisher 1923; Wright 1931). This expected reduction in diversity due to
-random genetic drift is captured by the third term in the deterministic
-component of expression (\ref{sig2_gen}), particularly
-\(-V\sigma^2(t)/N(t)\). The component of SDE (\ref{sig2_gen}) describing
-random fluctuations in \(\sigma^2(t)\) is more complicated and is
-proportional to the root of the difference between the centralized
-fourth moment of \(p(x,t)\) and \(\sigma^4(t)\).
-
-These expressions can be used to investigate the dynamics of the mean
-and variance for general \(\nu(x,t)\), assuming
-\(\int_\mathbb{R}(|x|+x^2+x^4)\nu(x,t)dx<+\infty\) for all \(t\geq0\).
-However, in the next subsection we simplify these expressions by
-approximating \(\nu(x,t)\) with a Gaussian curve. By assuming
-\(\nu(x,t)\) is Gaussian for \(t\geq0\), we guarantee the existence of
-\(\bar x(t)\) and \(\sigma^2(t)\) for all \(t\geq0\). Furthermore, in SM
-\S\ref{SDE_DERIV} we show that under the Gaussian case \(W_1,W_2\) and
-\(W_3\) are independent.
-
-\hypertarget{particular-results-assuming-a-gaussian-phenotypic-distribution}{%
-\subsubsection{\texorpdfstring{Particular results assuming a Gaussian
-phenotypic distribution
-\label{particular}}{Particular results assuming a Gaussian phenotypic distribution }}\label{particular-results-assuming-a-gaussian-phenotypic-distribution}}
-
-By approximating \(\nu(x,t)\) with a Gaussian curve in \(x\) for each
-\(t\geq0\), expressions (\ref{gen_eqns}) transform into efficient tools
-for deriving the dynamics of populations given a fitness function
-\(m(\nu,x)\). Gaussian phenotypic distributions are often obtained
-through Gaussian, exponential or weak selection approximations together
-with a simplified model of mutation and asexual reproduction or random
-mating (Lande 1980; Turelli 1984, 1986, 2017; Bürger 2000).
-Alternatively, it has been shown that a Gaussian distribution can
-provide a reasonable approximation even when selection is strong and
-non-Gaussian (Turelli and Barton 1994). Hence, given appropriate
-assumptions on selection, mutation and reproduction, the abundance
-density \(\nu(x,t)\) can be approximated as a Gaussian curve in \(x\)
-when the ratio \(V/N\) is small (i.e., when the variance in reproductive
-output is much smaller than the population size). As with any diffusion
-approximation, this requires a sufficiently large abundance to
-accurately reflect the dynamics of populations. Hence, models developed
-in this framework are not suitable for studies involving very small
-population sizes. Allowing for these restrictions, we assume
-\begin{equation}
-\nu(x,t)=\frac{N(t)}{\sqrt{2\pi\sigma^2(t)}}\exp\left(-\frac{\big(x-\bar x(t)\big)^2}{2\sigma^2(t)}\right).
-\end{equation} Under this assumption we find in SM \S\ref{cov2deriv} the
-results (suppressing the dependency on \(t\))
-\begin{equation}\label{covxm}
-\mathrm{Cov}(x,m)=\sigma^2\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right),
-\end{equation} \begin{equation}
-\mathrm{Cov}\Big((x-\bar x)^2,m\Big)=2\sigma^4\left(\frac{\partial\bar m}{\partial\sigma^2}-\overline{\frac{\partial m}{\partial\sigma^2}}\right)
-\end{equation} and \(\overline{(x-\bar x)^4}=3\sigma^4\). In particular,
-these results imply
-
-\begin{subequations}\label{no_inher}
-\begin{equation}\label{xbar}
-d\bar x=\sigma^2\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right)dt+\sqrt{V\frac{\sigma^2}{N}}dW_2,
-\end{equation}
-\begin{equation}\label{G}
-d\sigma^2=2\sigma^4\left(\frac{\partial\bar m}{\partial\sigma^2}-\overline{\frac{\partial m}{\partial\sigma^2}}\right)dt +\left(\mu-V\frac{\sigma^2}{N}\right)dt+\sigma^2\sqrt{\frac{2V}{N}}dW_3.
-\end{equation}
-\end{subequations}
-
-These equations allow us to derive the response in trait mean and
-variance by taking derivatives of fitness, a much more straightforward
-operation than calculating a covariance for general phenotypic
-distributions. Note that in the above expressions, the partial
-derivatives of \(\bar m\) represent frequency independent selection and
-the averaged partial derivatives of \(m\) represent frequency dependent
-selection. This relationship has already been pointed out by Lande
-(1976) for the evolution of the mean trait in discrete time, but here we
-see an analogous relationship holds in continuous time and also for the
-evolution of trait variance.
-
-In the next subsection we generalize this result to the case when traits
-are imperfectly inherited. In this case, the phenotypic variance
-\(\sigma^2\) is replaced by a genetic variance \(G\). This genetic
-variance represents the component of the variance in expressed traits
-\(\sigma^2\) explained by additive effects of different alleles among
-genetic loci encoding for the focal phenotype (Roughgarden 1979; Bulmer
-1980; Lynch and Walsh 1998). It is therefore fitting that \(G\) is
-referred to as the additive genetic variance.
-
-\hypertarget{the-evolution-of-additive-genetic-variance}{%
-\subsubsection{\texorpdfstring{The evolution of additive genetic
-variance
-\label{inheritance}}{The evolution of additive genetic variance }}\label{the-evolution-of-additive-genetic-variance}}
-
-To model imperfect heritability we consider the relationship between
-expressed phenotypes \(x\in\mathbb{R}\) and associated genetic values
-\(g\in\mathbb{R}\) known as \emph{breeding values}. The breeding value
-of an individual is the sum of additive effects of the alleles carried
-by the individual on its expressed trait. Since our derivations of
-evolutionary equations are based on branching processes that assume
-asexually reproducing populations (\S\ref{stochastic}), the additive
-genetic variance \(G\) is just the variance of breeding values in a
-population. For a detailed treatment of breeding values and additive
-genetic variances, see Bulmer (1980) and Lynch and Walsh (1998).
-
-Our treatment of the relationship between breeding values and expressed
-traits follows classical quantitative genetic assumptions such as those
-used in the seminal paper by Lande (1975) to investigate the maintenance
-of genetic variation. In particular, we assume that the expressed trait
-for any given individual is independent of environmental conditions and
-normally distributed around their breeding value with variance \(\eta\).
-Hence, \(\sigma^2=G+\eta\). In the case that all of the effects of
-alleles on an expressed trait are additive, \(\eta\) is known as the
-\emph{variance of environmental deviation} (Lande 1975; Lynch and Walsh
-1998). For a given breeding value, we denote the probability density of
-a randomly drawn expressed trait by \(\psi(x,g)\). Hence,
-\begin{equation}
-\psi(x,g)=\frac{1}{\sqrt{2\pi\eta}}\exp\left(-\frac{(x-g)^2}{2\eta}\right).
-\end{equation} To include this relationship in our framework, we write
-\(\rho(g,t)\) as the abundance density of breeding values at time \(t\)
-so that
-\(\int_{-\infty}^{+\infty}\rho(g,t)dg=\int_{-\infty}^{+\infty}\nu(x,t)dx=N(t)\).
-We switch our focus from directly modelling the evolution of
-\(\nu(x,t)\) to modelling the evolution of \(\rho(g,t)\). Once
-\(\rho(g,t)\) is determined, we can compute \(\nu(x,t)\) via
-\begin{equation}\label{nu_from_rho}
-\nu(x,t)=\int_{-\infty}^{+\infty}\rho(g,t)\psi(x,g)dg.
-\end{equation} However, since selection acts on expressed phenotypes, we
-use the assumed relationship between breeding values and expressed
-traits to calculate the fitness of breeding values. To motivate the
-approach taken, consider the problem of inferring the breeding value of
-an individual given its expressed trait \(x\). Denote \(\mathfrak{g}\) a
-random variable representing the unkown breeding value. Under this model
-of inheritance we know \(x\) is a random sample from a normal
-distribution with mean \(\mathfrak{g}\) and variance \(\eta\).
-Maximizing likelihood suggests \(x\) is our best guess for
-\(\mathfrak{g}\), but the actual value of \(\mathfrak{g}\) is normally
-distributed around \(x\) with the variance \(\eta\). Hence, for fixed
-\(x\), we obtain \(\psi(x,g)\) as the probability density of
-\(\mathfrak{g}\). Thus, the mean fitness of a breeding value \(g\)
-across all individuals carrying \(g\) can be written as
-\begin{equation}\label{m_rel}
-m^*(\rho,g)=\int_{-\infty}^{+\infty}m(\nu,x)\psi(x,g)dx.
-\end{equation} This is similar to the approach taken by Kimura and Crow
-(1978) to calculate the overall effects of selection for expressed
-characters onto the changes in the distribution of alleles encoding
-those characters. However, instead of focusing on the frequencies of
-alleles at particular loci, our results focus on the densities of
-breeding values. With the relationship between \(m(\nu,x)\) and
-\(m^*(\rho,g)\) established, we define the evolution of \(\rho(g,t)\) by
-the SPDE \begin{equation}\label{rho_SPDE}
-\dot\rho(g,t)=m^*(\rho,g)\rho(g,t)+\frac{\mu}{2}\frac{\partial^2}{\partial^2 g}\rho(g,t)+\sqrt{V\rho(g,t)}\dot W(g,t).
-\end{equation} Equation (\ref{rho_SPDE}) is a stochastic generalization
-of the deterministic PDE (\ref{eq1}) from \S\ref{deterministic}, but
-describes the evolution of the distribution of breeding values instead
-of expressed characters. However, whether modelling expressed characters
-or breeding values, we refer to SPDE of the form (\ref{rho_SPDE}) as
-Stochastic Asexual Gaussian allelic models with Abundance dynamics
-(abbreviated SAGA). Replacing \(\psi(x,g)\) with different probability
-densities produces different developmental models. This includes
-densities that may depend on environmental variables and hence model
-phenotypic plasticy. To leave room for the naming of other models where
-\(\psi(x,g)\) is Gaussian, we refer to the model of development we have
-chosen as the \emph{centered} Gaussian developmental model. In
-\S\ref{wnc} we review the origins of equation (\ref{rho_SPDE}) and
-provide some theory to help make sense of it, particularly the term
-\(\dot W\).
-
-Assuming \(\rho(g,t)\) is Gaussian implies its mode coincides with
-\(\bar x\). Furthermore, since \(\sigma^2=G+\eta\), we can use equation
-(\ref{m_rel}) and the chain rule from calculus (see SM \S\ref{fit2fit})
-to find
-
-\begin{subequations}
-\begin{equation}
-\frac{\partial\bar m}{\partial G}=\frac{\partial\bar m}{\partial\sigma^2}\frac{\partial\sigma^2}{\partial G}=\frac{\partial\bar m}{\partial\sigma^2},
-\end{equation}
-\begin{equation}
-\overline{\frac{\partial m}{\partial G}}=\overline{\frac{\partial m}{\partial\sigma^2}\frac{\partial\sigma^2}{\partial G}}=\overline{\frac{\partial m}{\partial\sigma^2}}.
-\end{equation}
-\end{subequations}
-
-Thus, equations (\ref{no_inher}) become
-
-\begin{subequations}\label{inher}
-\begin{equation}\label{xbarfinal}
-d\bar x=G\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right)dt+\sqrt{V\frac{G}{N}}dW_2,
-\end{equation}
-\begin{equation}\label{Gfinal}
-dG=2G^2\left(\frac{\partial\bar m}{\partial G}-\overline{\frac{\partial m}{\partial G}}\right)dt+\left(\mu-V\frac{G}{N}\right)dt+ G\sqrt{\frac{2V}{N}}dW_3.
-\end{equation}
-\end{subequations}
-
-From expressions (\ref{inher}) we see that, under this model of
-inheritance, focusing on additive genetic variance \(G\) instead the
-variance in expressed traits \(\sigma^2\) makes no structural changes to
-the basic equations describing the dynamics of populations. In the next
-section, we make use of these expressions to develop a model of diffuse
-coevolution in a guild of \(S\) species competing along a resource
-continuum.
-
-\hypertarget{a-model-of-diffuse-coevolution}{%
-\section{\texorpdfstring{A model of diffuse coevolution
-\label{coev}}{A model of diffuse coevolution }}\label{a-model-of-diffuse-coevolution}}
-
-\hypertarget{formulation}{%
-\subsection{\texorpdfstring{Formulation
-\label{form_coev}}{Formulation }}\label{formulation}}
-
-In this section we demonstrate the use of our framework by developing a
-model of diffuse coevolution across a guild of \(S\) species whose
-interactions are mediated by resource competition along a single niche
-axis. Because our approach treats abundance dynamics and evolutionary
-dynamics simultaneously, this model allows us to investigate the
-relationship between selection gradients and competition coefficients,
-which we carry out in what follows.
-
-The dynamics of phenotypic distributions and abundances have been
-derived above and so the only task remaining is the formulation of a
-fitness function. Our approach mirrors closely the theory developed by
-MacArthur and Levins (1967), Levins (1968) and MacArthur (1969, 1970,
-1972). The most significant difference, aside from allowing evolution to
-occur, is the treatment of resource quality, which we replace with a
-model of abiotic stabilizing selection. A derivation is provided in SM
-\S\ref{diffuse}.
-
-For species \(i\) we inherit the above notation for trait value,
-distribution, average, variance, abundance, et cetera, except with an
-\(i\) in the subscript. Real world examples of niche axes include the
-size of seeds consumed competing finch species and the date of activity
-in a season for pollinators competing for floral resources. For
-mathematical convenience, we model the axis of resources by the real
-line \(\mathbb{R}\). The value of a resouce along this axis is denoted
-by the symbol \(\zeta\). For an individual in species \(i\), we assume
-the resource utilization curve \(u_i\) can be written as
-
-\begin{equation}
-u_i(\zeta,x_i)=\frac{U_i}{\sqrt{2\pi w_i}}\exp\left(-\frac{(x_i-\zeta)^2}{2w_i}\right).
-\end{equation}
-
-We further assume the niche center \(x_i\) is normally distributed among
-individuals in species \(i\), but the niche breadth \(w_i\) and total
-niche utilization \(U_i\) are constant across individuals in species
-\(i\) and therefore cannot evolve. Suppose \(\theta_i\in\mathbb R\) is
-the optimal location along the niche axis for species \(i\) such that,
-in the absence of competition, individuals leave on average \(Q_i\)
-offspring when concentrated at \(\theta_i\). We assume the expected
-reproductive output of an individual in species \(i\) decreases as niche
-location \(\zeta\) leaves the optimum \(\theta_i\) at the rate
-\(A_i>0\). Hence, abiotic stabilizing selection along the resource axis
-can be modelled by the curve
-
-\begin{equation}
-e_i(\zeta)=Q_i\exp\left(-\frac{A_i}{2}(\theta_i-\zeta)^2\right).
-\end{equation}
-
-The effect of abiotic stabilizing selection on the fitness for an
-individual of species \(i\) with niche location \(x_i\) is then given by
-
-\begin{equation}
-\int_{-\infty}^{+\infty}e_i(\zeta)u_i(\zeta,x_i)d\zeta=\frac{Q_iU_i}{\sqrt{A_iw_i+1}}\exp\left(-\frac{A_i}{2(A_iw_i+1)}(\theta_i-x_i)^2\right).
-\end{equation}
-
-To determine the potential for competition between individuals with
-niche locations \(x_i\) and \(x_j\), belonging to species \(i\) and
-\(j\) respectively, we compute the niche overlap
-
-\begin{equation}
-\mathcal O_{ij}(x_i-x_j)=\int_{-\infty}^{+\infty}u_i(\zeta,x_i)u_j(\zeta,x_j)d\zeta=\frac{U_iU_j}{\sqrt{2\pi(w_i+w_j)}}\exp\left(-\frac{(x_i-x_j)^2}{2(w_i+w_j)}\right).
-\end{equation}
-
-A notable criticism of using niche overlap to measure the intensity of
-competition points to cases where populations competiting on multiple
-niche axes exhibit overlap on at least one of the axes, but no overall
-niche overlap (Holt 1987). Thus niche overlap on lower-dimensional
-projections of some multivariate niche space does not imply the
-populations compete. To illustrate with a simple example, consider two
-populations competing for space on the plane \(\mathbb{R}^2\). If the
-spatial distributions of the two populations overlap, then they will
-overlap on both spatial axes. However, if the populations do not overlap
-on at least one of the spatial axes, they will have no overall spatial
-overlap. Furthermore, even if the species overlap on both spatial axes,
-they need not have any overall spatial overlap. This final result
-corresponds to the fact that components of niche space do not
-necessarily interact multiplicatively to determine the consequences for
-the intensity of competition. In another component of Holt's (1987)
-critique, an argument is made for the potential of competition occuring
-without any overlap in niche space. However, this argument is based on
-the practical difficulty of identifying every resource axis populations
-are competing on and how these axes interact to determine fitness
-consequences. Our model avoids these caveats by assuming competition
-only occurs along a single dimensional resource gradient.
-
-To map the degree of niche overlap to fitness, we assume competition
-between individuals with niche locations \(x_i\) and \(x_j\) decreases
-the expected reproductive output for the individual in species \(i\) at
-the rate \(c_i\mathcal{O}_{ij}(x_i-x_j)\) for some \(c_i>0\). We refer
-to \(c_i\) as the sensitivity to competition for species \(i\). The term
-\(c_i\mathcal{O}_{ij}(x_i-x_j)\) coincides with a special case of a term
-used to capture competition in Dawson's geostochastic logistic model, an
-SPDE model developed to study the combined effects of demographic
-stochasticity, spatial dispersion and locally finite carrying capacity
-(Dawson 1978). In relation to the example fitness function discussed in
-\S\ref{deterministic}, consider
-\(\kappa(x_i-x_j)=\mathcal{O}_{ij}(x_i-x_j)\). In SM \S\ref{diffuse} we
-combine our treatment of resource competition with equations (\ref{N}),
-(\ref{xbarfinal}) and (\ref{Gfinal}) to find
-
-\begin{subequations}\label{comm_dynamics}
-\begin{multline}\label{N_comm}
-dN_i = \Bigg\{R_i-\frac{a_i}{2}\Big((\bar x_i-\theta_i)^2+G_i+\eta_i\Big) - c_i\sum_{j=1}^SN_jU_iU_j\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\Bigg\}N_idt + \sqrt{V_iN_i}dW_1,
-\end{multline}
-\begin{multline}\label{x_comm}
-d\bar x_i = \Bigg\{a_iG_i(\theta_i-\bar x_i)-c_iG_i\bigg(\sum_{j=1}^SN_jU_iU_jb_{ij}(\bar x_j-\bar x_i)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\bigg)\Bigg\}dt+\sqrt{V_i\frac{G_i}{N_i}}dW_2,
-\end{multline}
-\begin{multline}\label{G_comm}
-dG_i =  \Bigg\{c_i{G_i}^2\bigg(N_iU_i^2b_{ii}\sqrt\frac{b_{ii}}{2\pi}+\sum_{j=1}^SN_jU_iU_jb_{ij}\left(1-b_{ij}(\bar x_i-\bar x_j)^2\right)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\bigg) \\
-+\mu_i-a_i{G_i}^2-V_i\frac{G_i}{N_i}\Bigg\}dt+G_i\sqrt{\frac{2V_i}{N_i}}dW_3,
-\end{multline}
-\end{subequations}
-
-where
-
-\begin{subequations}
-\begin{align}
-R_i = & \ \ln \left(\frac{Q_iU_i}{\sqrt{1+A_iw_i}}\right), \\
-a_i = & \ \frac{A_i}{1+A_iw_i}, \\
-b_{ij}(t) = b_{ji}(t) = & \ \big(w_i+w_j+\eta_i+\eta_j+G_i(t)+G_j(t)\big)^{-1}, \\
-c_i \geq & \ 0.
-\end{align}
-\end{subequations}
-
-Despite the convoluted appearance of system (\ref{comm_dynamics}), there
-are some nice features that reflect biological reasoning. For example,
-the dynamics of abundance are just a generalization of Lotka-Volterra
-dynamics. In particular, the effect of competition with species \(j\) on
-the fitness of species \(i\) grows linearly with \(N_j\). However, as
-biotic selection pushes \(\bar x_i\) away from \(\bar x_j\), the effect
-of competition with species \(j\) on the fitness of species \(i\)
-rapidly diminishes, reflecting a reduction in niche overlap. The
-divergence of \(\bar x_i\) and \(\bar x_j\) due to competition is
-referred to in the community ecology literature as character
-displacement (Brown and Wilson 1956). We also see that the fitness of
-species \(i\) drops quadratically with the difference between
-\(\bar x_i\) and the abiotic optimum \(\theta_i\). Hence, abiotic
-selection acts to pull \(\bar x_i\) towards \(\theta_i\). The response
-in mean trait \(\bar x_i\) to natural selection is proportional to the
-amount of heritable variation in the population, represented by the
-additive genetic variance \(G_i\). However, we have that \(G_i\) is
-itself a dynamic quantity. Under our model, abiotic stabilizing
-selection erodes away heritable variation at a rate that is independent
-of both \(N_i\) and \(\bar x_i\). The effect of competition on \(G_i\)
-is a bit more complicated. When \(b_{ij}(\bar x_i-\bar x_j)^2<1\),
-competition with species \(j\) acts as diversifying selection which
-tends to increase the amount of heritable variation. However, when
-\(b_{ij}(\bar x_i-\bar x_j)^2>1\), competition with species \(j\) acts
-as directional selection and reduces \(G_i\). In the following
-subsections we demonstrate the behavior of system (\ref{comm_dynamics})
-by plotting numerical solutions and investigate implications for the
-relationship between the strength of ecological interactions and
-selection.
-
-\hypertarget{community-dynamics}{%
-\subsection{\texorpdfstring{Community dynamics
-\label{dynamics}}{Community dynamics }}\label{community-dynamics}}
-
-For the sake of illustration we numerically integrated system
-(\ref{comm_dynamics}) for a richness of \(S=100\) species. We assumed
-homogeneous model parameters across species in the community as
-summarized by Table \ref{par_vals}. We repeated numerical integration
-under the two scenarios of weak and strong competition. For the first
-scenario of weak competition we set \(c=1.0\times10^{-7}\) and for the
-second scenario of strong competition we set \(c=5.0\times10^{-6}\).
-With these two sets of model parameters, we simulated our model for
-\(1000.0\) units of time. For both scenarios, we initialized the trait
-means to \(\bar x_i=0.0\), additive genetic variances to \(G_i=10.0\)
-and abundances to \(N_i=1000.0\) for each \(i=1,\dots,S\).
-
-Temporal dynamics for each scenario are provided in Figure
-\ref{temporal}. This figure suggests weaker competition leads to
-smoother dynamics and a higher degree of organization within the
-community. Considering expression (\ref{N_comm}) we note that, all else
-equal, relaxed competition allows for larger growth rates which promote
-greater abundances. From (\ref{N_comm}) we also note that the per-capita
-effects on demographic stochasticity diminish with abundance. To see
-this, divide both sides by \(N_i\). Inspecting expressions
-(\ref{x_comm}) and (\ref{G_comm}), we see that larger abundances also
-erode the effects of demographic stochasticity on the evolution of mean
-trait and additive genetic variance. These effects were already noted in
-\S\ref{equations}, and thus are not a consequence of our model of
-coevolution per-se, but we revisit them here since Figure \ref{temporal}
-demonstrates the importance of demographic stochasticity in structuring
-ecological communities even when populations are very large. Hence,
-contrary to the common assumption that stochastic effects can be ignored
-for large populations, we find that minute asymmetries generated by
-demographic stochasticity remain significant drivers of community
-structure. In particular, we initialized the species with identical
-state variables and model parameters, but found an enourmous amount of
-asymmetry and even some potential phase changes. In the following two
-paragraphs we describe the natural history of the community as
-illustrated in Figure \ref{temporal}.
-
-We begin by describing the weak competition scenario. After about
-\(125.0\) units of time, the community appears to have shaken off the
-initial conditions and entered into a qualitatively distinct phase of
-dynamics. Aside from a few outliers, most of the species remain
-clustered together in their state variables. This lasts for
-approximately \(375.0\) units of time until, at around time \(500.0\), a
-drastic change occurs. At this moment the tightly packed cluster of
-species begins to fan out in all three state variables. Simultaneously,
-we observe large a shift in mean traits for higher values and in
-additive genetic variances for lower values. Upon inspecting our
-calculations, we diagnose the reason for this shift. The outlier species
-that were initially pushed away from the common abiotic optimum (\(0.0\)
-in this case) evolved a significant reduction in the amount of heritable
-variation (\(\approx60\%\)) due to directional selection induced by
-competition. This reduction in heritable variation slowed adaptation,
-causing these species to linger on the outskirts of niche space, some
-longer than others. In the meantime the rest of the community, being
-tightly packed, experienced greater competition which led to diminished
-abundances for these species and caused some members of the core group
-to veer away from the abiotic optimum. The reduced abundances of the
-core group led to reduced competition overall. As a result, the outlier
-populations were given a slight increase in growth rate, enough to allow
-them to increase their abundances orders of magnitude higher than the
-species in the core group and giving them more weight in driving the
-evolution of other species. Many of these heavy-hitting outlier species
-had already been maintaining negative mean traits, but around time
-\(500.0\) the high abundance species with positive mean traits began to
-experience enough intraspecific competition to override interspecific
-competition. This generated a net selection gradient and associated
-evolutionary response towards the abiotic optimum. The sudden imbalance
-of these high abundance species effectively induced a single large
-competitive exclusion event pushing the majority of the community far
-away from the abiotic optimum. After this shift the cluster began to
-slowly bloom in all three state variables as species took advantage of
-novel asymmetries in their competitive abilities mediated by a new
-distribution of mean trait values across the community. About \(125.0\)
-units of time later, the community reached a qualitatively new phase of
-dynamics. If we kept running the numerical integrator, we would continue
-to see similar drama unfolding over and over again as minute stochastic
-changes contribute to asymmetries which slowly build into drastic
-shifts.
-
-The strong competition scenario is not quite as showy. Although the
-dynamics of trait means and variances tend to be far more stochastic
-than in the weak competition scenario, the community overall appears to
-quickly reach some statistical equilibrium and remain there. However,
-the abundances across all species in the community are very low due to
-competition sensitivies being an order of magnitude higher than in the
-weak competition case. Most of the species maintain abundances greater
-than \(1000.0\), but we found one species that dropped to an abundance
-of about \(50.0\). If we let the numerical integrator run long enough in
-this case, we will likely see many of the species go extinct.
-
-Finding ways to interpret simulated dynamics provides a useful arena to
-exercise biological reasoning. However, it does not fulfill our desire
-to quantify the patterns and processes present in competing communities.
-In the next subsection we take a step in this direction by using our
-model to derive formula for selection gradients and competition
-coefficients. To investigate their relationship, we calculate their
-covariances using simplifying assumptions on species abundances and
-intraspecific trait variances. We then investigate how these covariances
-change with the ratio of variance of interspecific mean traits to
-variance of intraspecific individual traits and use a numerical approach
-to investigate correlations between the strength of pairwise coevolution
-and competition coefficients.
-
-\begin{table}
-\centering\caption{Values of model parameters used for numerical integration.}\vspace{0.2cm}
-\begin{tabular}{l|l|l}
-Parameter     & Description & Value                                           \\ \hline
-$R$         & innate growth rate, see \S\ref{ecoevo}  & $1.0$      \\
-$\theta$    & abiotic optimum & $0.0$                              \\
-$a$         & strength of abiotic selection & $0.01$               \\
-$c$         & sensitivity to competition & $\{1.0\times10^{-7},5.0\times10^{-6}\}$ \\
-$w$         & niche breadth & $0.1$                               \\
-$U$         & total niche use & $1.0$                              \\
-$\eta$      & segregation variance & $1.0$                         \\
-$\mu$       & mutation rate & $1.0\times10^{-7}$                             \\
-$V$ & variance of reproductive output & $5.0$
-\end{tabular}\label{par_vals}
-\end{table}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{community_dynamics}
-
-}
-
-\caption{\label{temporal}Temporal dynamics of mean trait (top), additive genetic variance (middle) and abundance (bottom) for the scenario of weak competition (left) and strong competition (right).}\label{fig:unnamed-chunk-6}
-\end{figure}
-
-\hypertarget{the-relation-between-the-strength-of-ecological-interactions-and-coevolution}{%
-\subsection{\texorpdfstring{The relation between the strength of
-ecological interactions and coevolution
-\label{ecoevo}}{The relation between the strength of ecological interactions and coevolution }}\label{the-relation-between-the-strength-of-ecological-interactions-and-coevolution}}
-
-Relating our treatment of resource competition to modern coexistence
-theory (Chesson 2000), the absolute competition coefficient
-\(\alpha_{ij}\) becomes a dynamical quantity that can be written as
-\begin{equation}
-\alpha_{ij}(t)=\frac{c_i}{r_i(t)}\int_{-\infty}^{+\infty}\int_{-\infty}^{+\infty}p_i(x,t)p_j(y,t)\mathcal O_{ij}(x,y) dxdy =\frac{c_iU_iU_j}{r_i(t)}\sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right),
-\end{equation} where \begin{equation}
-r_i(t)=R_i-\frac{a_i}{2}\Big((\bar x_i(t)-\theta_i)^2+G_i(t)+\eta_i\Big).
-\end{equation} Hence, \(dN_i(t)\) can be expressed as \begin{equation}
-dN_i(t)=r_i(t)\left(1-\sum_{j=1}^S\alpha_{ij}(t)N_j(t)\right)N_i(t)dt+\sqrt{V_iN_i(t)}dW_1(t).
-\end{equation} Note that although \(r_i(t)\) is referred to in the
-coexistence literature as the intrinsic growth rate of the population,
-\(R_i\) is a deeper intrinsic quantity. For now we refer to \(R_i\) as
-the \emph{innate} growth rate. Previous work has shown the importance of
-demographic stochasticity (Turelli 1980; Schreiber 2017) and
-evolutionary dynamics (Case and Taper 1986; Schreiber, Patel, and
-terHorst 2018) in determining coexistence of competing species. However,
-these studies consider evolution and demographic stochasticity
-separately. To our knowledge, there are no rigorous theoretical
-investigations of coexistence that account for evolutionary dynamics of
-quantitative characters and demographic stochasticity simultaneously
-and, in particular, no studies combining these processes that also
-include the evolution of additive genetic variance. With the connection
-outlined in this section formally established, researchers may pursue a
-postmodern coexistence theory that naturally includes both the
-evolutionary dynamics of quantitative characters and the effects of
-demographic stochasticity in a simple synthetic framework.
-
-In SM \S\ref{diffuse} we show that the standardized directional
-selection gradient (sensu Lande and Arnold 1983) induced by species
-\(j\) on species \(i\) can be computed as \begin{equation}
-\beta_{ij}(t)=c_iU_iU_jN_j(t)b_{ij}(t)\big(\bar x_i(t)-\bar x_j(t)\big) \sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right).
-\end{equation} Our notation differs from Lande and Arnold (1983) in that
-subscripts here denote species instead of components of multivariate
-traits and we drop the prime that distinguishes between selection
-gradients and standardized selection gradients.
-
-\textbf{Metric of pairwise coevolution}
-
-Below we investigate the correspondence of interaction intensity and
-coevolutionary change. However, we can already identify one major
-discrepancy; \(\alpha_{ij}\) is maximized when \(\bar x_i=\bar x_j\),
-but \(\beta_{ij}=0\) under the same condition. We therefore include in
-our metric of selection the standardized stabilizing selection gradient
-\(\gamma\) which measures the effect of stabilizing or disruptive
-selection on phenotypic variance (Lande and Arnold 1983). In SM
-\S\ref{diffuse} we show that the standardized stabilizing selection
-gradient induced by species \(j\) on species \(i\) can be computed as
-\begin{equation}
-\gamma_{ij}(t)=c_iU_iU_jN_j(t)b_{ij}(t)\left(1-b_{ij}(t)\big(\bar x_i(t)-\bar x_j(t)\big)^2\right) \sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right).
-\end{equation} To measure evolutionary change in species \(i\) induced
-by species \(j\), we form the metric
-\(\Psi_{ij}=|\beta_{ij}|+|\gamma_{ij}|\) and define the strength of
-coevolution as \(\mathfrak{C}_{ij}=\Psi_{ij}\Psi_{ji}\). Figure
-\ref{net} displays the joint distribution of \(\mathfrak{C}_{ij}\) and
-the product of competition coefficients \(\alpha_{ij}\alpha_{ji}\) for
-two simulated communities, both with richness \(S=1000\). The solid
-contour represents the case of relatively high sensitivity to
-competition (\(c=5\times10^{-7}\)) and the dashed contour represents the
-case of relatively weak sensitivity to competition
-(\(c=1\times10^{-8}\)). In both cases we see that \(\mathfrak{C}_{ij}\)
-and \(\alpha_{ij}\alpha_{ji}\) are, essentially, unrelated. However,
-this depiction is only representative of a specific set of parameters.
-Next, we provide analytical approximations of the covariance between
-selection gradients and competition coefficients and a numerical
-estimate for the relationship between pairwise coevolution and
-competition coefficients for a range of parameters.
-
-\begin{figure}
-
-{\centering \includegraphics{on_pl}
-
-}
-
-\caption{\label{net}Bivariate distributions of competition coffecients (y-axis) and pairwise coevolution (x-axis) under the scenarios of weak competition (dashed line) and strong competition (solid line) after simulating for $t=1000.0$ units of time.
-Simulations were ran for $S=1000$ species.
-Parameters are the same as in Table 2, except we used $c=1\times 10^{-8}$ for weak competition and $c=5\times 10^{-7}$ for strong competition in order to account for increased species richness.}\label{fig:unnamed-chunk-8}
-\end{figure}
-
-\textbf{Covariance of selection and competition as a function of
-diversity}
-
-We now make use of the expressions derived for competition coeffecients
-and selection gradients to investigate their relationship. As a first
-pass, let us assume all model parameters are equivalent across species
-and that each species has the same abundance and trait variance. Let us
-further assume that richness \(S\) is large and the distribution of mean
-trait values is normal with mean \(\bar{\bar x}\), variance
-\(V_{\bar X}\) and density \(f_{\bar X}\). Such assumptions are typical
-when deriving analytical results in the field of theoretical
-coevolutionary community ecology (Nuismer, Jordano, and Bascompte 2012;
-Nuismer, Week, and Aizen 2018). If \(\bar{\bar x}\) is near \(\theta\)
-and \(V_{\bar X}\) is much smaller than \(|2\frac{R}{a}-G-\eta|\), then
-we may approximate \(r_i\) with \begin{equation}
-\bar r=\int_{-\infty}^{+\infty}\left(R-\frac{a}{2}\Big((\bar x-\theta)^2+G+\eta\Big)\right)f_{\bar X}(\bar x)d\bar x
-=R-\frac{a}{2}\Big((\bar{\bar x}-\theta)^2+V_{\bar X}+G+\eta\Big).
-\end{equation} In SM \S\ref{coef_grad_moms} we use these assumptions to
-calculate the first and second order moments describing the joint
-distribution of competition coefficients and selection gradients across
-the community. We find that the covariance between linear selection
-gradients and competition coeffecients are zero due to the symmetry
-implied by our assumptions. However, setting
-\(\alpha(\bar x_i,\bar x_j)=\alpha_{ij}\),
-\(\beta(\bar x_i,\bar x_j)=\beta_{ij}\) and
-\(\gamma(\bar x_i,\bar x_j)=\gamma_{ij}\), the covariances between the
-magnitude of linear selection gradients and competition coefficients and
-between stabilizing selection gradients and competition coefficients can
-be written as
-
-\begin{subequations}\label{cov_eqns}
-\begin{equation}\label{cov_alpha_beta}
-\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)=\frac{2c^2b^2U^4N}{\pi\bar r}\sqrt\frac{V_{\bar X}}{2\pi}\left(\frac{1}{(1+8bV_{\bar X})^{3/4}}-\frac{1}{(1+4bV_{\bar X})^{3/4}(1+2bV_{\bar X})^{1/2}}\right),
-\end{equation}
-\begin{equation}\label{cov_alpha_gamma}
-\mathrm{Cov}_{f_{\bar X}}(\alpha,\gamma)=\frac{c^2b^2U^4N}{2\pi\bar r}(1-2bV_{\bar X})\left(\frac{1}{\sqrt{1+4bV_{\bar X}}}-\frac{1}{1+2bV_{\bar X}}\right).
-\end{equation}
-\end{subequations}
-
-\begin{figure}
-
-{\centering \includegraphics{/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/cov}
-
-}
-
-\caption{\label{cov_fig}Curves representing the covariance between the magnitude of linear selection gradients and competition coefficients (left) and between stabilizing selection gradients and competition coefficients (right) as a function of $2bV_{\bar X}$ which is approximately equal to the ratio of variance in mean traits among species to the intraspecific trait variance.
-In both plots we set $c=1.0\times10^{-4}$, $b=0.1$, $\bar r=0.1$ and $N=1.0\times10^{10}$ and let $V_{\bar X}$ vary.}\label{fig:unnamed-chunk-10}
-\end{figure}
-
-We visualize these relationships in Figure \ref{cov_fig} for fixed
-\(c,b,\bar r,U\) and \(N\). In SM \S\ref{coef_grad_moms} we verify the
-qualitative nature of these relationships by repeating numerical
-integrations of system (\ref{comm_dynamics}) across a set of randomly
-drawn model parameters. We designed our numerical approach to inherit
-the assumptions of symmetric model parameters such as mutation rates
-\(\mu_1=\dots=\mu_S\) and abiotic optima \(\theta_1=\dots=\theta_S\) and
-to relax the key assumptions of normally distributed mean traits along
-with constant additive genetic variances and abundances across species.
-
-Our assumptions in this section imply \(b^{-1}=2(\sigma^2+w)\). If we
-further assume \(\sigma^2+w\approx\sigma^2\), then
-\(2bV_{\bar X}\approx V_{\bar X}/\sigma^2\). That is, when populations
-are generalists and are comprised of specialist individuals, the value
-\(2bV_{\bar X}\) is approximately equal to the ratio of interspecific
-mean trait variation to intraspecific individual trait variation.
-Glancing at Figure (\ref{cov_fig}) we see there is no relationship
-between selection gradients and competition coefficients when this ratio
-is zero. From equation (\ref{cov_alpha_beta}) we can use numerical
-optimization to find that when \(V_{\bar X}/\sigma^2\approx1.25\) the
-relationship between the magnitudes of linear selection gradients and
-competition coefficients disappears, but when
-\(V_{\bar X}/\sigma^2\lessapprox1.25\) (\(\gtrapprox1.25\)), this
-covariance becomes negative (positive). Equation (\ref{cov_alpha_gamma})
-states that when \(V_{\bar X}/\sigma^2\) is approximately equal to one
-(or slightly larger), there is no expected relationship between
-competition coefficients and quadratic selection gradients. However,
-when \(V_{\bar X}/\sigma^2<1.0\) (\(>1.0\)), then we expect a positive
-(negative) relationship between \(\alpha\) and \(\gamma\). These results
-are true regardless of the chosen parameter values. In SM
-\S\ref{coef_grad_moms} we use simulations of system
-(\ref{comm_dynamics}) to show that these results do not qualitatively
-differ when allowing for heterogeneous population sizes and additive
-genetic variances across species.
-
-From a biological perspective, if the ratio \(V_{\bar X}/\sigma^2\) is
-small, then species are packed tightly in phenotypic space. In our model
-this occurs when fitness is more sensitive to abiotic stabilizing
-selection than competition (\(a\gg c\)). This causes species to overlap
-more in niche space (i.e., large \(\alpha\)) and creates disruptive
-selection for greater intraspecific variance (i.e., positive
-\(\gamma\)), which explains the positive region of covariance between
-\(\alpha\) and \(\gamma\). However, as species begin to overlap in niche
-space, directional selection begins to vanish (i.e., small \(|\beta|\)),
-leading to a negative covariance between \(\alpha\) and \(|\beta|\). In
-the limiting case that two species have perfectly overlapping niches,
-they will exhibit zero directional selection since a shift in either
-direction will yield identical fitness advantages.
-
-In the opposite scenario where fitness is much more sensitive to
-competition than abiotic stabilizing selection (\(c\gg a\)), species
-will not evolve to be as tightly packed and instead their niche-centers
-will be spread out with little ovelap in their resouce utilization
-curves (i.e., small \(\alpha\)). In this case biotic directional
-selection will be strong (i.e., large \(|\beta|\)), particularly for
-species towards the outer regions of niche space due to asymmetric
-fitness advantages conferred by shifts in niche-centers. This leads to a
-positive covariance between \(\alpha\) and \(|\beta|\). However, as
-noted above, this directional selection will also erode away at standing
-heritable variation (i.e., negative \(\gamma\)), reducing the rate at
-which adaptation can occur and creating a negative covariance between
-\(\alpha\) and \(\gamma\).
-
-In summary, we see the relations between competition coefficients and
-selection gradients are non-trivial and depend on the relative
-magnitudes of different ecological processes shaping the community.
-However, this does not address the relation between competition
-coefficients and coevolution per se. In SM \S\ref{coef_grad_moms} we
-show that calculating a formula for the covariance between competition
-coefficients and the metric of coevolution \(\mathfrak{C}\) introduced
-above provides a difficult analytical challenge. Instead of confronting
-this challenge we build on our numerical approach used to justify
-analytical approximations of
-\(\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)\) and
-\(\mathrm{Cov}_{f_{\bar X}}(\alpha,\gamma)\) to approximate the
-correlation of \(\alpha\) and \(\mathfrak{C}\). In Figure
-\ref{cor_coev_fig} we plot values of
-\(\mathrm{Corr}_{f_{\bar X}}(\alpha,\mathfrak{C})\) against the exact
-ratio of interspecific trait variance to intraspecific trait variance.
-Using a cubic regression, we see the correlation of coevolutionary
-selection gradients and competition coefficeints is negative at variance
-ratios below \(0.5\), zero at variance ratios between \(0.5\) and
-\(1.0\), and drops below zero again at variance ratios above \(1.0\).
-
-\begin{figure}
-
-{\centering \includegraphics{/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/Cαℭp}
-
-}
-
-\caption{\label{cor_coev_fig}Numerical estimates for the correlation of the strength of coevolution measured by $\mathfrak{C}$ and competition coefficeints $\alpha$ plotted against the variance of mean traits among species divided by the mean variance of traits within species.
-Each dot represents the result from a single simulation.
-The red line is a cubic regression.}\label{fig:unnamed-chunk-12}
-\end{figure}
-
-\hypertarget{conclusion}{%
-\section{\texorpdfstring{Conclusion
-\label{conclusion}}{Conclusion }}\label{conclusion}}
-
-We have introduced an approach to derive models of evolutionary ecology
-using the calculus of white noise and coined SAGA, a SPDE model of
-phenotypic evolution that accounts for demographic stochasticity. From
-SAGA we derived SDE that track the dynamics of abundance, mean trait and
-additive genetic variance. Observing the expressions of these SDE we
-find the effects of demographic stochasticity on the evolution of mean
-trait and additive genetic variance characterize the effects of random
-genetic drift. Although Lande (1976) has previously characterized the
-effects of random genetic drift on mean trait evolution in quantitative
-genetic models, the approach taken assumed constant effective population
-size and discrete non-overlapping generations. In contrast, our approach
-shows random genetic drift is a result of demographic stochasticity for
-continuously reproducing populations with fluctuating abundances. To
-illustrate the relevance of our approach to studies of evolutionary
-ecology, we comined our SDE with classical competition theory to derive
-a model of diffuse coevolution and used this model to investigate the
-relationship between standardized selection gradients and competition
-coefficients. We found the absolute value of linear selection gradients
-are negatively related with competition coefficients when the ratio of
-interspecific phenotypic variation to intraspecific phenotypic variation
-is approximately less than \(1.25\) and positive when this ratio is
-approximately greater than \(1.25\). We found the opposite relationship
-between quadratic selection gradients and competition coefficients with
-a threshold diversity ratio equal to unity instead of \(1.25\). This
-work demonstrates that connecting contemporary theoretical approaches to
-evolutionary ecology with some fundamental results in the theory of
-measure-valued stochastic processes allows for the development of a
-rigorous, yet flexible approach to synthesizing the dynamics of
-abundance and distributions of quantitative characters, in agreement
-with Champagnat, Ferrière and Méléard (2006). Although this approach
-enjoys these important merits, there remains biological details and
-their associated technical challenges that need to be confronted for
-gaining a more thorough and accurate understanding of ecological
-communities. We touch on just four of them here and provide suggestions
-for future research to approach these challenges. For the first three,
-we recommend the use of individual-based simulations of measure-valued
-processes to confront these challenges. To summarize why, we state here
-that analytical results become difficult to derive for small populations
-and complicated models of mutation and selection.
-
-\textbf{Limitations of diffusion limits and generalizing coexistence
-theory}
-
-As noted by Feller (1951), diffusion limits provide reasonable
-approximations for large populations, but relatively small populations
-require discrete models. Hence, as a diffusion limit, SPDE (\ref{SPDE})
-restricts parameter values to regions that maintain large population
-sizes. Although this suggests an important restriction on any model
-developed under this framework by implying populations are not at risk
-of extinction, the SDE describing abundance dynamics technically permits
-extinction. However, for small abundances, pathology emerges in the
-evolution of trait means and variances. In particular, stochastic
-components of the SDE describing the evolution of \(\bar x\) and
-\(\sigma^2\) diverge towards infinite values as \(N\to0\). Hence,
-studies involving extinction and phenotypic evolution should be pursued
-through a different approach. A natural alternative can be developed
-utilizing the underlying measure-valued branching process that SAGA is a
-diffusion limit of (see section \ref{stochastic}). This process
-explicitly tracks individuals throughout their life-history and captures
-reproduction as branching events. Hence, branching processes model
-population size as a non-negative integer instead of a continuously
-varying number. In particular, the pathological behavior described for
-the SDE of \(\bar x\) and \(\sigma^2\) does not occur for measure-valued
-branching processes. This concern relates to our comment in
-\S\ref{ecoevo} on using our model of diffuse coevolution to extend
-modern coexistence theory (sensu Chesson 2000) in a way that accounts
-for phenotypic evolution and demographic stochasticity. Just as previous
-work has sought to understand the implications of processes for
-coexistence separately, it seems necessary to determine whether or not
-results pertaining to the coexistence of competing populations hold
-using models that account for phenotypic evolution and demographic
-stochasticity simultaneously. Because such investigations consider both
-extinction and phenotypic evolution we argue the models used should be
-either simulations of, or whenever possible, analytical results based on
-measure-valued branching processes.
-
-\textbf{Genetic architecture and intraspecific trait distributions}
-
-Our treatment of inheritance and our approach to model coevolution rests
-on the assumptions of normally distributed breeding values and expressed
-phenotypes along with asexual reproduction. However, real traits are not
-encoded by an infinite number of loci each contributing an additive
-infinitesimal effect (as required by the infinitesimal model), mutations
-are not inherited as normally distributed deviations from parental
-breeding values (as required by the Gaussian allelic model) and many
-populations of interest exhibit non-random sexual reproduction.
-Departures from this model of genetic architecture can produce more
-complex distributions of breeding values and expressed traits. Such
-deviations can be reinforced via strong non-Gaussian selection surfaces,
-including the surface \(m(\nu,x)\) we have derived from niche theory,
-along with non-random mating in sexually reproducing populations.
-However, Gaussian approximations are convenient since they are defined
-by their mean and variance. Future work investigating the effects of
-non-normally distributed traits on the structure and dynamics of
-ecological communities will therefore need to confront higher moments
-such as skew and kurtosis, ideally integrating previously established
-mathematical approaches to derive the dynamics of these higher moments
-(Débarre, Yeaman, and Guillaume 2015).
-
-Another approach to breaking the assumption of normally distributed
-trait values is the development of explicit multilocus models. These
-models describe the contributions of alleles at particular loci in the
-genome to the development of quantitative traits. Tracking the
-fluctuations of allele frequencies then allows theoreticians to
-investigate the dynamics of phenotypic distributions that deviate from
-normality. This approach has a long history in theoretical quantitative
-genetics (Bulmer 1980; Turelli and Barton 1994; Kirkpatrick, Johnson,
-and Barton 2002) and coevolutionary theory (Nuismer, Doebeli, and
-Browning 2005; Kopp and Gavrilets 2006; Nuismer, Ridenhour, and Oswald
-2007). However, work to investigate the role of complex genetic
-architectures in mediating feedbacks between the dynamics of population
-abundances and the distributions of traits mediating ecological
-interactions has apparently only just begun (Patel and Bürger 2019).
-Since analytical results are difficult to obtain for general models of
-mutation, development and selection (c.f., Bürger 2000), we again
-suggest the use of measure-valued branching processes and their
-individual-based simulations. For example, by assuming the mutational
-effects in the BBM model is Gaussian with variance an order of magnitude
-greater than the equilibrium phenotypic (or additive genetic) variance,
-we obtain the \emph{House-of-cards} model (Turelli 1984). However,
-perhaps considering heavy-tailed distributions such as the
-Laplace-distributions considered by Park et. al. (2017) will provide
-more realistic yet less analytically tractable models of mutational
-effects. Such mutation models have important implications in
-conservation genetics where small populations evolve primarily in
-response to mutation and random genetic drift (Lande 1995).
-
-\textbf{Ecological stoichiometry}
-
-Our treatment of both biotic and abiotic selection neglects important
-chemical constraints of biological reality. For instance, the resource
-we assume species are competing over is modelled as a static quantity.
-However, real resources can be dynamic quantities. Such theoretical
-quantities may reflect abiotic cycles of material and energy or whole
-trophic layers comprised of prey, hosts or mutualists. Although resource
-dynamics have been captured theoretically in consumer-resource models
-(Tilman 1982), developing a more realistic model of resource competition
-must incorporate details on the ecophysiology of organisms (Loreau
-2010). Doing so will help clarify the feedback between the evolution of
-populations and the ecosystem processes they are a part of.
-
-Using plant-pollinator communities as an example, the role of nitrogen
-mediating interspecific interactions has been reviewed by David, Storkey
-and Stevens (2019) and the evolutionary ecology of the nutritional
-content of nectar has been reviewed by Parachnowitsch, Manson and
-Sletvold (2018). These studies demonstrate the need for further research
-to understand how soil nutrient availability and organismal
-ecophysiology structures communities of plants and pollinators.
-Theoretical pursuits in this direction may profit from interfacing the
-framework we have outlined here with population-ecosystem models such as
-that developed by Fridley (2017). However, these population-ecosystem
-models will likely yield complicated fitness functions that are
-difficult to analyze using rigorous mathematical techniques. Hence, we
-again recommend the use of simulations of measure-valued branching
-processes that generalize BBM to understand when analytical
-approximations hold.
-
-\textbf{Accounting for macroevolutionary history}
-
-To understand patterns found in ecological communities considerations
-must push beyond microevolutionary and contemporary ecological processes
-and consider the macroevolutionary dynamics of ancestral lineages
-leading to extant populations. Using sub-alpine flower communities as an
-example, one can observe a very strict ordering of phenology across a
-broad geographic range. In particular, whether in the Colorado Rocky
-mountains (such as Gothic, Colorado) or on an outlier of the Idaho
-batholith (such as Kamiak butte near Palouse, Washington), one will
-almost surely observe a very conspicuous order of flowers emerging in
-early spring: at the very beginning of the season blooms \emph{Claytonia
-lanceolata} followed by \emph{Erythronium grandiflorum} which precedes
-\emph{Delphinium nuttallianum} (B. Week, personal observations). If
-contemporary phenological coevolution is rampant, why should this
-pattern be so well preserved across a thousand miles of rugged and
-diverse terrain? Although it would be exciting to find that these
-species repeatedly coevolved this pattern in each location, a more
-parsimoniuous hypothesis suggests the phenology and physiology of these
-species slowly evolved independently over macroevolutionary time scales
-to take advantage of the specific conditions available within each of
-these windows of the flowering season. However, this could not have
-carried out in the Rocky mountains since this terrain only became
-habitable just over ten thousand years ago as the glaciers of the
-Pleistocene began to retreat (Paul CaraDonna, personal communications).
-Hence, given these considerations, it appears that an understanding of
-early season phenology patterns should focus on how these communities
-are assembled as opposed to contemporary evolutionary dynamics. Indeed,
-recent work testing models of phylogeography ignores the potential for
-contemporary evolution and instead suggests alpine flower communities
-tend to follow neutral assembly where flowers merely compete for who can
-disperse to new habitat first, as opposed to a selective process where a
-regional species pool is filtered for those species adapted to the newly
-available habitat (Marx et al. 2017).
-
-Of course microevolutionary and ecological dynamics are not completely
-irrelevant for understanding patterns in communities that are primarily
-structured by deep evolutionary processes. In particular,
-macroevolutionary trait evolution is simply the aggregation of
-microevolutionary change occuring over large spans of time. This
-suggests a road forward to connect the theory we have introduced to
-models of macroevolutionary trait evolution.
-
-Some approaches to modelling macroevolutionary trait change simply
-repurpose microevolutionary models by blindly rescaling time from the
-units of generations to millions of years (e.g., Nuismer and Harmon
-2014). Such an approach makes the implicit assumption that trait
-evolution is statistically self-similar (sensu Falconer 2014) so that
-the stochastic evolution of traits on macroevolutionary time scales has
-the same properties of trait evolution on microevolutionary time scales.
-Although some stochastic processes, including Brownian motion, do
-exhibit self-similarity, others do not. For example, consider a
-modification of the Ornstein-Uhlenbeck process defined by the SDE
-
-\begin{equation}
-dX_t=a(\theta_t-X_t)dt+bdW_t
-\end{equation}
-
-where \(a,b>0\), \(W_t\) is a standard Brownian motion and \(\theta_t\)
-is itself a Markov process that takes normally distributed jumps
-centered on its current location at exponentially distributed time
-intervals. If we assume the rate \(\lambda\) at which jumps occur is
-much smaller than \(a\) and the variance in jumping is much larger than
-\(b^2\), then, even though the sample paths of \(X_t\) are actually
-continuous (if we zoom in close enough, they look like Brownian motion),
-over long intervals of time sample paths of \(X_t\) will begin to appear
-to have periods of continuity interrupted by an occasional discontinuous
-jump and thus approach a qualitatively distinct process. These emergent
-properties can be formally characterized by Lévy processes and have been
-successfully employed in comparative phylogenetics to fit phenotypic
-data from extant populations and the fossil record (Landis and Schraiber
-2017). It would therefore be interesting to investigate whether an
-application of a separation of time scales argument for the rate of
-environmental change (\(\lambda\)) versus the rate of evolutionary and
-ecological change (\(a\)) to microevolutionary models derived via our
-framework can be used to obtain macroevolutionary models that include
-not only mean trait evolution, but also the evolution of trait variance
-and abundance. The resulting macroevolutionary models can give rise to
-novel comparative phylogenetic methods and provide initial conditions
-for microevolutionary models that capture contemporary dynamics.
-
-\textbf{Final remarks}
-
-Although top-down approaches to community ecology, such as the Maximum
-Entropy Theory of Ecology (Harte 2011), have enjoyed some success in
-describing community-level patterns (Harte and Newman 2014; Xiao,
-McGlinn, and White 2015), a mechanistic understanding of why these
-patterns emerge and how they will change remains a formidable task. Such
-an understanding must take both bottom-up and top-down approaches
-integrating considerations from the ecophysiology of individual
-organisms that reveal the economics of interspecific interactions
-(Sterner and Elser 2008), to the phylogeographic history of taxa that
-sets the stage for contemporary dynamics (Hickerson et al. 2010).
-Through connecting these dots we can increase the variance explained in
-observations of ecological communities by specific mechanisms and come
-closer to a predictive theory of evolutionary community ecology. Despite
-the long list of equations derived in this paper, this work takes just
-one small step towards capturing these many details. However, we hope
-the theoretical framework outlined here along with the demonstration of
-its use in modelling competitive communities provides some helpful tools
-to aid quantitative evolutionary ecologists in reaching such lofty
-goals.
-
-\newpage
-
-\hypertarget{supplementary-material-sm}{%
-\section{Supplementary material (SM)}\label{supplementary-material-sm}}
 
@@
 \hypertarget{sufficient-conditions-for-finite-mean-variance-and-total-abundance-in-the-deterministic-case}{%
-\subsection{\texorpdfstring{Sufficient conditions for finite mean,
-variance and total abundance in the deterministic case
+\section{\texorpdfstring{Sufficient conditions for finite mean, variance
+and total abundance in the deterministic case
 \label{finite}}{Sufficient conditions for finite mean, variance and total abundance in the deterministic case }}\label{sufficient-conditions-for-finite-mean-variance-and-total-abundance-in-the-deterministic-case}}
@@
 \hypertarget{equilibrium-moments-for-a-deterministic-population-experiencing-logistic-growth-and-stabilizing-selection}{%
-\subsection{\texorpdfstring{Equilibrium moments for a deterministic
+\section{\texorpdfstring{Equilibrium moments for a deterministic
 population experiencing logistic growth and stabilizing selection
@@
 \hypertarget{the-relation-between-diffusion-and-convolution-with-a-gaussian-kernel}{%
-\subsection{\texorpdfstring{The relation between diffusion and
-convolution with a Gaussian kernel
+\section{\texorpdfstring{The relation between diffusion and convolution
+with a Gaussian kernel
 \label{diffconvequiv}}{The relation between diffusion and convolution with a Gaussian kernel }}\label{the-relation-between-diffusion-and-convolution-with-a-gaussian-kernel}}
@@
 \hypertarget{deterministic-dynamics-of-sigma2t}{%
-\subsection{\texorpdfstring{Deterministic dynamics of \(\sigma^2(t)\)
+\section{\texorpdfstring{Deterministic dynamics of \(\sigma^2(t)\)
 \label{var_deriv}}{Deterministic dynamics of \textbackslash{}sigma\^{}2(t) }}\label{deterministic-dynamics-of-sigma2t}}
@@
 \hypertarget{comparing-our-treatment-of-white-noise-to-da-prato-and-zabczyk-2014}{%
-\subsection{\texorpdfstring{Comparing our treatment of white noise to Da
+\section{\texorpdfstring{Comparing our treatment of white noise to Da
 Prato and Zabczyk (2014)
@@
 \hypertarget{simulating-the-rescaled-process}{%
-\subsection{\texorpdfstring{Simulating the rescaled process
+\section{\texorpdfstring{Simulating the rescaled process
 \label{numerical}}{Simulating the rescaled process }}\label{simulating-the-rescaled-process}}
@@
 \hypertarget{description-of-simulation}{%
-\subsubsection{\texorpdfstring{Description of simulation
+\subsection{\texorpdfstring{Description of simulation
 \label{description}}{Description of simulation }}\label{description-of-simulation}}
@@
 
-\textbf{Rescaling}
+\hypertarget{rescaling}{%
+\subsection{Rescaling}\label{rescaling}}
 
@@
 \hypertarget{derivation-of-sde-for-bar-x-and-sigma2}{%
-\subsection{\texorpdfstring{Derivation of SDE for \(\bar x\) and
+\section{\texorpdfstring{Derivation of SDE for \(\bar x\) and
 \(\sigma^2\)
@@
 \hypertarget{derivation-for-trait-mean}{%
-\subsubsection{Derivation for trait
-mean}\label{derivation-for-trait-mean}}
+\subsection{Derivation for trait mean}\label{derivation-for-trait-mean}}
 
@@
 \hypertarget{derivation-for-trait-variance}{%
-\subsubsection{Derivation for trait
+\subsection{Derivation for trait
 variance}\label{derivation-for-trait-variance}}
@@
 \hypertarget{relating-fitness-of-expressed-traits-to-fitness-of-breeding-values}{%
-\subsection{\texorpdfstring{Relating fitness of expressed traits to
-fitness of breeding values
+\section{\texorpdfstring{Relating fitness of expressed traits to fitness
+of breeding values
 \label{fit2fit}}{Relating fitness of expressed traits to fitness of breeding values }}\label{relating-fitness-of-expressed-traits-to-fitness-of-breeding-values}}
@@
 \hypertarget{derivation-of-diffuse-coevolution-model}{%
-\subsection{\texorpdfstring{Derivation of diffuse coevolution model
+\section{\texorpdfstring{Derivation of diffuse coevolution model
 \label{diffuse}}{Derivation of diffuse coevolution model }}\label{derivation-of-diffuse-coevolution-model}}
@@
 \hypertarget{the-relation-between-competition-coefficients-and-selection}{%
-\subsection{\texorpdfstring{The relation between competition
-coefficients and selection
+\section{\texorpdfstring{The relation between competition coefficients
+and selection
 \label{coef_grad_moms}}{The relation between competition coefficients and selection }}\label{the-relation-between-competition-coefficients-and-selection}}
@@
 \hypertarget{derivation-of-analytical-approximations}{%
-\subsubsection{Derivation of analytical
+\subsection{Derivation of analytical
 approximations}\label{derivation-of-analytical-approximations}}
@@
 \hypertarget{caclulating-mathrmcov_f_bar-xalphagamma}{%
-\paragraph{\texorpdfstring{Caclulating
+\subsubsection{\texorpdfstring{Caclulating
 \(\mathrm{Cov}_{f_{\bar X}}(\alpha,\gamma)\)}{Caclulating \textbackslash{}mathrm\{Cov\}\_\{f\_\{\textbackslash{}bar X\}\}(\textbackslash{}alpha,\textbackslash{}gamma)}}\label{caclulating-mathrmcov_f_bar-xalphagamma}}
@@
 \hypertarget{caclulating-mathrmcov_f_bar-xalphabeta}{%
-\paragraph{\texorpdfstring{Caclulating
+\subsubsection{\texorpdfstring{Caclulating
 \(\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)\)}{Caclulating \textbackslash{}mathrm\{Cov\}\_\{f\_\{\textbackslash{}bar X\}\}(\textbackslash{}alpha,\textbar{}\textbackslash{}beta\textbar{})}}\label{caclulating-mathrmcov_f_bar-xalphabeta}}
@@
 \hypertarget{starting-the-caclulation-of-mathrmcov_f_bar-xalphamathfrakc}{%
-\paragraph{\texorpdfstring{Starting the caclulation of
+\subsubsection{\texorpdfstring{Starting the caclulation of
 \(\mathrm{Cov}_{f_{\bar X}}(\alpha,\mathfrak{C})\)
@@
 \hypertarget{numerical-estimates-for-heterogeneous-n-and-g}{%
-\subsubsection{\texorpdfstring{Numerical estimates for heterogeneous
-\(N\) and
+\subsection{\texorpdfstring{Numerical estimates for heterogeneous \(N\)
+and
 \(G\)}{Numerical estimates for heterogeneous N and G}}\label{numerical-estimates-for-heterogeneous-n-and-g}}
@@
 
-\caption{\label{num_corr} Numerical estimate for the correlations of selection gradients and competition coefficients.}\label{fig:unnamed-chunk-13}
+\caption{\label{num_corr} Numerical estimate for the correlations of selection gradients and competition coefficients.}\label{fig:unnamed-chunk-1}
 \end{figure}
@@
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-AbdalaRoberts2014}{}%
-Abdala-Roberts, Luis, and Kailen A. Mooney. 2014. ``Ecological and
-Evolutionary Consequences of Plant Genotype Diversity in a Tri-Trophic
-System.'' \emph{Ecology} 95 (10). Wiley: 2879--93.
-
-\leavevmode\hypertarget{ref-Athreya1972}{}%
-Athreya, Krishna B., and Peter E. Ney. 1972. \emph{Branching Processes}.
-Springer Berlin Heidelberg.
-
-\leavevmode\hypertarget{ref-9781107648258}{}%
-Ågren, Göran I., and Folke O. Andersson. 2012. \emph{Terrestrial
-Ecosystem Ecology: Principles and Applications}. Cambridge University
-Press.
-
-\leavevmode\hypertarget{ref-Barton2017}{}%
-Barton, N.H., A.M. Etheridge, and A. Véber. 2017. ``The Infinitesimal
-Model: Definition, Derivation, and Implications.'' \emph{Theoretical
-Population Biology} 118 (December). Elsevier BV: 50--73.
-
-\leavevmode\hypertarget{ref-Barton2019}{}%
-Barton, Nick, and Alison Etheridge. 2019. ``Mathematical Models in
-Population Genetics.'' Wiley.
-
-\leavevmode\hypertarget{ref-Barton2013}{}%
-Barton, Nick, Alison Etheridge, and Amandine Véber. 2013. ``Modelling
-Evolution in a Spatial Continuum.'' \emph{Journal of Statistical
-Mechanics: Theory and Experiment} 2013 (01). IOP Publishing: P01002.
-
-\leavevmode\hypertarget{ref-Bertoin2003}{}%
-Bertoin, Jean, and Jean-François Le Gall. 2003. ``Stochastic Flows
-Associated to Coalescent Processes.'' \emph{Probability Theory and
-Related Fields} 126 (2). Springer Science; Business Media: 261--88.
-
-\leavevmode\hypertarget{ref-Brown1956}{}%
-Brown, W. L., and E. O. Wilson. 1956. ``Character Displacement.''
-\emph{Systematic Zoology} 5 (2). Oxford University Press (OUP): 49.
-
-\leavevmode\hypertarget{ref-bulmer1980}{}%
-Bulmer. 1980. \emph{The Mathematical Theory of Quantitative Genetics}.
-Oxford University Press.
-
-\leavevmode\hypertarget{ref-Brger1986}{}%
-Bürger, Reinhard. 1986. ``On the Maintenance of Genetic Variation:
-Global Analysis of Kimuras Continuum-of-Alleles Model.'' \emph{Journal
-of Mathematical Biology} 24 (3). Springer Nature: 341--51.
-
-\leavevmode\hypertarget{ref-9780471986539}{}%
----------. 2000. \emph{The Mathematical Theory of Selection,
-Recombination, and Mutation}. Wiley.
-
-\leavevmode\hypertarget{ref-Cantrell2004}{}%
-Cantrell, Robert Stephen, and Chris Cosner. 2004. \emph{Spatial Ecology
-via Reaction-Diffusion Equations}. Wiley.
-
-\leavevmode\hypertarget{ref-Case1986}{}%
-Case, Ted J., and Mark L. Taper. 1986. ``On the Coexistence and
-Coevolution of Asexual and Sexual Competitors.'' \emph{Evolution} 40
-(2). Wiley: 366--87.
-
 \leavevmode\hypertarget{ref-Champagnat2006}{}%
@@
 
-\leavevmode\hypertarget{ref-Chesson2000}{}%
-Chesson, Peter. 2000. ``Mechanisms of Maintenance of Species
-Diversity.'' \emph{Annual Review of Ecology and Systematics} 31 (1).
-Annual Reviews: 343--66.
-
-\leavevmode\hypertarget{ref-jeffreyconner2004}{}%
-Conner, Jeffrey K. 2004. \emph{A Primer of Ecological Genetics}. Sinauer
-Associates.
-
-\leavevmode\hypertarget{ref-9781932846126}{}%
-Crow, James F., and Motoo Kimura. 1970. \emph{An Introduction to
-Population Genetics Theory}. The Blackburn Press.
-
-\leavevmode\hypertarget{ref-Crutsinger2015}{}%
-Crutsinger, Gregory M. 2015. ``A Community Genetics Perspective:
-Opportunities for the Coming Decade.'' \emph{New Phytologist} 210 (1).
-Wiley: 65--70.
-
 \leavevmode\hypertarget{ref-DaPrato2014}{}%
@@
 
-\leavevmode\hypertarget{ref-David2019}{}%
-David, Thomas I., Jonathan Storkey, and Carly J. Stevens. 2019.
-``Understanding How Changing Soil Nitrogen Affects Plantpollinator
-Interactions.'' \emph{Arthropod-Plant Interactions} 13 (5). Springer
-Science; Business Media LLC: 671--84.
-
-\leavevmode\hypertarget{ref-Dawson1975}{}%
-Dawson, Donald A. 1975. ``Stochastic Evolution Equations and Related
-Measure Processes.'' \emph{Journal of Multivariate Analysis} 5 (1).
-Elsevier BV: 1--52.
-
-\leavevmode\hypertarget{ref-Dawson1978}{}%
----------. 1978. ``Geostochastic Calculus.'' \emph{Canadian Journal of
-Statistics} 6 (2). Wiley: 143--68.
-
-\leavevmode\hypertarget{ref-dawson1993measure}{}%
----------. 1993. ``Measure-Valued Markov Processes.'' In \emph{École
-d'été de Probabilités de Saint-Flour Xxi-1991}, 1--260. Springer.
-
-\leavevmode\hypertarget{ref-Dbarre2015EvolutionOQ}{}%
-Débarre, Florence, Sam Yeaman, and Frédéric Guillaume. 2015. ``Evolution
-of Quantitative Traits Under a Migration-Selection Balance: When Does
-Skew Matter?'' \emph{The American Naturalist} 186 Suppl 1: S37--47.
-
-\leavevmode\hypertarget{ref-Doebeli1996}{}%
-Doebeli, Michael. 1996. ``Quantitative Genetics and Population
-Dynamics.'' \emph{Evolution} 50 (2). Wiley: 532--46.
-
-\leavevmode\hypertarget{ref-alisonetheridge2000}{}%
-Etheridge, Alison M. 2000. \emph{An Introduction to Superprocesses}.
-American Mathematical Society.
-
-\leavevmode\hypertarget{ref-Etheridge2008}{}%
----------. 2008. ``Drift, Draft and Structure: Some Mathematical Models
-of Evolution.'' In \emph{Stochastic Models in Biological Sciences}.
-Institute of Mathematics Polish Academy of Sciences.
-
-\leavevmode\hypertarget{ref-Etheridge1991}{}%
-Etheridge, Alison, and Peter March. 1991. ``A Note on Superprocesses.''
-\emph{Probability Theory and Related Fields} 89 (2). Springer: 141--47.
-
 \leavevmode\hypertarget{ref-lawrenceevans2010}{}%
@@
 
-\leavevmode\hypertarget{ref-9781470410544}{}%
----------. 2014. \emph{An Introduction to Stochastic Differential
-Equations}. American Mathematical Society.
-
-\leavevmode\hypertarget{ref-Evans1994}{}%
-Evans, Steven N., and Edwin A. Perkins. 1994. ``Measure-Valued Branching
-Diffusions with Singular Interactions.'' \emph{Canadian Journal of
-Mathematics} 46 (1). Canadian Mathematical Society: 120--68.
-
-\leavevmode\hypertarget{ref-Ewens2004}{}%
-Ewens, Warren J. 2004. \emph{Mathematical Population Genetics}. Springer
-New York.
-
-\leavevmode\hypertarget{ref-kennethfalconer2014}{}%
-Falconer, Kenneth. 2014. \emph{Fractal Geometry: Mathematical
-Foundations and Applications}. Wiley.
-
 \leavevmode\hypertarget{ref-stanleyfarlow1993}{}%
@@
 Scientists and Engineers}. Dover.
-
-\leavevmode\hypertarget{ref-feller1951}{}%
-Feller, William. 1951. ``Diffusion Processes in Genetics.'' In
-\emph{Proceedings of the Second Berkeley Symposium on Mathematical
-Statistics and Probability}, 227--46. University of California Press.
-
-\leavevmode\hypertarget{ref-Felsenstein1975}{}%
-Felsenstein, Joseph. 1975. ``A Pain in the Torus: Some Difficulties with
-Models of Isolation by Distance.'' \emph{The American Naturalist} 109
-(967). University of Chicago Press: 359--68.
-
-\leavevmode\hypertarget{ref-Fisher1923}{}%
-Fisher, R. A. 1923. ``XXI.-on the Dominance Ratio.'' \emph{Proceedings
-of the Royal Society of Edinburgh} 42. Cambridge University Press:
-321--41.
-
-\leavevmode\hypertarget{ref-Fitzpatrick2015}{}%
-Fitzpatrick, Connor R., Anurag A. Agrawal, Nathan Basiliko, Amy P.
-Hastings, Marney E. Isaac, Michael Preston, and Marc T. J. Johnson.
-2015. ``The Importance of Plant Genotype and Contemporary Evolution for
-Terrestrial Ecosystem Processes.'' \emph{Ecology} 96 (10). Wiley:
-2632--42.
-
-\leavevmode\hypertarget{ref-Fitzpatrick2017}{}%
-Fitzpatrick, Connor R., Anna V. Mikhailitchenko, Daniel N. Anstett, and
-Marc T. J. Johnson. 2017. ``The Influence of Range-Wide Plant Genetic
-Variation on Soil Invertebrate Communities.'' \emph{Ecography} 41 (7).
-Wiley: 1135--46.
-
-\leavevmode\hypertarget{ref-FRANK2012}{}%
-Frank, S. A. 2012. ``Natural Selection. IV. The Price Equation.''
-\emph{Journal of Evolutionary Biology} 25 (6). Wiley: 1002--19.
-
-\leavevmode\hypertarget{ref-Fridley2017}{}%
-Fridley, Jason D. 2017. ``Plant Energetics and the Synthesis of
-Population and Ecosystem Ecology.'' Edited by David Gibson.
-\emph{Journal of Ecology} 105 (1). Wiley: 95--110.
-
-\leavevmode\hypertarget{ref-FUSSMANN2007}{}%
-Fussman, G. F., M. Loreau, and P. A. Abrams. 2007. ``Eco-Evolutionary
-Dynamics of Communities and Ecosystems.'' \emph{Functional Ecology} 21
-(3). Wiley: 465--77.
-
-\leavevmode\hypertarget{ref-Gomulkiewicz2017}{}%
-Gomulkiewicz, Richard, Stephen M. Krone, and Christopher H. Remien.
-2017. ``Evolution and the Duration of a Doomed Population.''
-\emph{Evolutionary Applications} 10 (5). Wiley: 471--84.
-
-\leavevmode\hypertarget{ref-Harmon2019}{}%
-Harmon, Luke J., Cecilia S. Andreazzi, Florence Débarre, Jonathan Drury,
-Emma E. Goldberg, Ayana B. Martins, Carlos J. Melián, et al. 2019.
-``Detecting the Macroevolutionary Signal of Species Interactions.''
-\emph{Journal of Evolutionary Biology} 32 (8). Wiley: 769--82.
-
-\leavevmode\hypertarget{ref-Harte2011}{}%
-Harte, John. 2011. \emph{Maximum Entropy and Ecology}. Oxford University
-Press.
-
-\leavevmode\hypertarget{ref-Harte2014}{}%
-Harte, John, and Erica A. Newman. 2014. ``Maximum Information Entropy: A
-Foundation for Ecological Theory.'' \emph{Trends in Ecology \&
-Evolution} 29 (7). Elsevier BV: 384--89.
-
-\leavevmode\hypertarget{ref-Hickerson2010}{}%
-Hickerson, M.J., B.C. Carstens, J. Cavender-Bares, K.A. Crandall, C.H.
-Graham, J.B. Johnson, L. Rissler, P.F. Victoriano, and A.D. Yoder. 2010.
-``Phylogeography's Past, Present, and Future: 10 Years After Avise,
-2000.'' \emph{Molecular Phylogenetics and Evolution} 54 (1). Elsevier
-BV: 291--301.
-
-\leavevmode\hypertarget{ref-Hofbauer1998}{}%
-Hofbauer, Josef, and Karl Sigmund. 1998. \emph{Evolutionary Games and
-Population Dynamics}. Cambridge University Press.
-
-\leavevmode\hypertarget{ref-Holt1987}{}%
-Holt, Robert D. 1987. ``On the Relation Between Niche Overlap and
-Competition: The Effect of Incommensurable Niche Dimensions.''
-\emph{Oikos} 48 (1). JSTOR: 110.
-
-\leavevmode\hypertarget{ref-Johnson2005}{}%
-Johnson, Toby, and Nick Barton. 2005. ``Theoretical Models of Selection
-and Mutation on Quantitative Traits.'' \emph{Philosophical Transactions
-of the Royal Society B: Biological Sciences} 360 (1459). The Royal
-Society: 1411--25.
-
-\leavevmode\hypertarget{ref-Kendall1966}{}%
-Kendall, David G. 1966. ``Branching Processes Since 1873.''
-\emph{Journal of the London Mathematical Society} s1-41 (1). Wiley:
-385--406.
-
-\leavevmode\hypertarget{ref-Kimmel2015}{}%
-Kimmel, Marek, and David E. Axelrod. 2015. \emph{Branching Processes in
-Biology}. Springer New York.
-
-\leavevmode\hypertarget{ref-Kimura1965}{}%
-Kimura, M. 1965. ``A Stochastic Model Concerning the Maintenance of
-Genetic Variability in Quantitative Characters.'' \emph{Proceedings of
-the National Academy of Sciences} 54 (3). Proceedings of the National
-Academy of Sciences: 731--36.
-
-\leavevmode\hypertarget{ref-Kimura1978}{}%
-Kimura, M., and J. F. Crow. 1978. ``Effect of Overall Phenotypic
-Selection on Genetic Change at Individual Loci.'' \emph{Proceedings of
-the National Academy of Sciences} 75 (12). Proceedings of the National
-Academy of Sciences: 6168--71.
-
-\leavevmode\hypertarget{ref-Kirkpatrick1727}{}%
-Kirkpatrick, Mark, Toby Johnson, and Nick Barton. 2002. ``General Models
-of Multilocus Evolution.'' \emph{Genetics} 161 (4). Genetics: 1727--50.
-
-\leavevmode\hypertarget{ref-kolmogorov1999elements}{}%
-Kolmogorov, A.N., and S.V. Fomin. 1999. \emph{Elements of the Theory of
-Functions and Functional Analysis}. v. 1. Dover.
-
-\leavevmode\hypertarget{ref-Konno1988}{}%
-Konno, N., and T. Shiga. 1988. ``Stochastic Partial Differential
-Equations for Some Measure-Valued Diffusions.'' \emph{Probability Theory
-and Related Fields} 79 (2). Springer Nature: 201--25.
-
-\leavevmode\hypertarget{ref-Kopp2006}{}%
-Kopp, Michael, and Sergey Gavrilets. 2006. ``Multilocus Gentics and the
-Coevolution of Quantitative Traits.'' \emph{Evolution} 60 (7). Wiley:
-1321--36.
-
-\leavevmode\hypertarget{ref-Klzsch2015}{}%
-Kölzsch, Andrea, Adriana Alzate, Frederic Bartumeus, Monique de Jager,
-Ellen J. Weerman, Geerten M. Hengeveld, Marc Naguib, Bart A. Nolet, and
-Johan van de Koppel. 2015. ``Experimental Evidence for Inherent Lévy
-Search Behaviour in Foraging Animals.'' \emph{Proceedings of the Royal
-Society B: Biological Sciences} 282 (1807). The Royal Society: 20150424.
-
-\leavevmode\hypertarget{ref-Kraft2007}{}%
-Kraft, Nathan J. B., William K. Cornwell, Campbell O. Webb, and David D.
-Ackerly. 2007. ``Trait Evolution, Community Assembly, and the
-Phylogenetic Structure of Ecological Communities.'' \emph{The American
-Naturalist} 170 (2). University of Chicago Press: 271--83.
-
-\leavevmode\hypertarget{ref-Krylov1981}{}%
-Krylov, N. V., and B. L. Rozovskii. 1981. ``Stochastic Evolution
-Equations.'' \emph{Journal of Soviet Mathematics} 16 (4). Springer
-Science; Business Media LLC: 1233--77.
-
-\leavevmode\hypertarget{ref-Lande1975}{}%
-Lande, Russell. 1975. ``The Maintenance of Genetic Variability by
-Mutation in a Polygenic Character with Linked Loci.'' \emph{Genetical
-Research} 26 (3). Cambridge University Press (CUP): 221--35.
-
-\leavevmode\hypertarget{ref-Lande1976}{}%
----------. 1976. ``Natural Selection and Random Genetic Drift in
-Phenotypic Evolution.'' \emph{Evolution} 30 (2). Wiley: 314--34.
-
-\leavevmode\hypertarget{ref-pmid17248993}{}%
----------. 1980. ``The Genetic Covariance between Characters Maintained
-by Pleiotropic Mutations.'' \emph{Genetics} 94 (1): 203--15.
-
-\leavevmode\hypertarget{ref-Lande1982}{}%
----------. 1982. ``A Quantitative Genetic Theory of Life History
-Evolution.'' \emph{Ecology} 63 (3). Wiley: 607--15.
-
-\leavevmode\hypertarget{ref-Lande1995}{}%
----------. 1995. ``Mutation and Conservation.'' \emph{Conservation
-Biology} 9 (4). Wiley: 782--91.
-
-\leavevmode\hypertarget{ref-Lande1983}{}%
-Lande, Russell, and Stevan J. Arnold. 1983. ``The Measurement of
-Selection on Correlated Characters.'' \emph{Evolution} 37 (6). JSTOR:
-1210.
-
-\leavevmode\hypertarget{ref-Lande2003}{}%
-Lande, Russell, Steinar Engen, and Bernt-Erik SÆther. 2003.
-``Demographic and Environmental Stochasticity.'' In \emph{Stochastic
-Population Dynamics in Ecology and Conservation}, 1--24. Oxford
-University Press.
-
-\leavevmode\hypertarget{ref-Lande2009}{}%
-Lande, Russell, Steinar Engen, and Bernt-Erik Sæther. 2009. ``An
-Evolutionary Maximum Principle for Density-Dependent Population Dynamics
-in a Fluctuating Environment.'' \emph{Philosophical Transactions of the
-Royal Society B: Biological Sciences} 364 (1523). The Royal Society:
-1511--8.
-
-\leavevmode\hypertarget{ref-Landis2017}{}%
-Landis, Michael J., and Joshua G. Schraiber. 2017. ``Pulsed Evolution
-Shaped Modern Vertebrate Body Sizes.'' \emph{Proceedings of the National
-Academy of Sciences} 114 (50). Proceedings of the National Academy of
-Sciences: 13224--9.
-
-\leavevmode\hypertarget{ref-9780691080628}{}%
-Levins, Richard. 1968. \emph{Evolution in Changing Environments: Some
-Theoretical Explorations. (MPB-2) (Monographs in Population Biology)}.
-Princeton University Press.
-
-\leavevmode\hypertarget{ref-zeng1998absolute}{}%
-Li, Zeng-Hu. 1998. ``Absolute Continuity of Measure Branching Processes
-with Interaction.'' \emph{Chinese Journal of Applied Probability and
-Statistics} 14. Citeseer: 231--42.
-
-\leavevmode\hypertarget{ref-Lion2018}{}%
-Lion, Sébastien. 2018. ``Theoretical Approaches in Evolutionary Ecology:
-Environmental Feedback as a Unifying Perspective.'' \emph{The American
-Naturalist} 191 (1). University of Chicago Press: 21--44.
-
-\leavevmode\hypertarget{ref-michelloreau2010}{}%
-Loreau, Michel. 2010. \emph{From Populations to Ecosystems: Theoretical
-Foundations for a New Ecological Synthesis}. Princeton University Press.
-
-\leavevmode\hypertarget{ref-michaellynch1998}{}%
-Lynch, Michael, and Bruce Walsh. 1998. \emph{Genetics and Analysis of
-Quantitative Traits}. Sinauer Associates is an imprint of Oxford
-University Press.
-
-\leavevmode\hypertarget{ref-Arthur1969}{}%
-MacArthur, Robert H. 1969. ``Species Packing, and what Competition
-Minimizes.'' \emph{Proceedings of the National Academy of Sciences} 64
-(4). Proceedings of the National Academy of Sciences: 1369--71.
-
-\leavevmode\hypertarget{ref-MacArthur1970}{}%
----------. 1970. ``Species Packing and Competitive Equilibrium for Many
-Species.'' \emph{Theoretical Population Biology} 1 (1). Elsevier BV:
-1--11.
-
-\leavevmode\hypertarget{ref-9780691023823}{}%
----------. 1972. \emph{Geographical Ecology}. Princeton University
-Press.
-
-\leavevmode\hypertarget{ref-Macarthur1967}{}%
-MacArthur, Robert H., and Richard Levins. 1967. ``The Limiting
-Similarity, Convergence, and Divergence of Coexisting Species.''
-\emph{The American Naturalist} 101 (921). University of Chicago Press:
-377--85.
-
-\leavevmode\hypertarget{ref-Manceau2016}{}%
-Manceau, Marc, Amaury Lambert, and Hélène Morlon. 2016. ``A Unifying
-Comparative Phylogenetic Framework Including Traits Coevolving Across
-Interacting Lineages.'' \emph{Systematic Biology}, December. Oxford
-University Press (OUP), syw115.
-
-\leavevmode\hypertarget{ref-Marx2017}{}%
-Marx, Hannah E., Cédric Dentant, Julien Renaud, Romain Delunel, David C.
-Tank, and Sébastien Lavergne. 2017. ``Riders in the Sky (Islands): Using
-a Mega-Phylogenetic Approach to Understand Plant Species Distribution
-and Coexistence at the Altitudinal Limits of Angiosperm Plant Life.''
-\emph{Journal of Biogeography} 44 (11). Wiley: 2618--30.
-
-\leavevmode\hypertarget{ref-markmcpeek2017}{}%
-McPeek, Mark A. 2017. \emph{Evolutionary Community Ecology}. Princeton
-University Press.
-
-\leavevmode\hypertarget{ref-DeMeester2018}{}%
-Meester, Luc De, Kristien I. Brans, Lynn Govaert, Caroline Souffreau,
-Shinjini Mukherjee, Héléne Vanvelk, Konrad Korzeniowski, et al. 2018.
-``Analyzing Eco-Evolutionary Dynamics - the Challenging Complexity of
-the Real World.'' \emph{Functional Ecology}. Wiley.
 
@@
 
-\leavevmode\hypertarget{ref-Mubayi2019}{}%
-Mubayi, Anuj, Christopher Kribs, Viswanathan Arunachalam, and Carlos
-Castillo-Chavez. 2019. ``Studying Complexity and Risk Through Stochastic
-Population Dynamics: Persistence, Resonance, and Extinction in
-Ecosystems.'' In \emph{Handbook of Statistics}, 157--93. Elsevier.
-
-\leavevmode\hypertarget{ref-9780674023383}{}%
-Nowak, Martin A. 2006. \emph{Evolutionary Dynamics: Exploring the
-Equations of Life}. Belknap Press.
-
-\leavevmode\hypertarget{ref-Nuismer2005}{}%
-Nuismer, Scott L., Michael Doebeli, and Danny Browning. 2005. ``The
-Coevolutionary Dynamics of Antagonistic Interactions Mediated by
-Quantitative Traits with Evolving Variances.'' \emph{Evolution} 59 (10).
-The Society for the Study of Evolution: 2073.
-
-\leavevmode\hypertarget{ref-Nuismer2014}{}%
-Nuismer, Scott L., and Luke J. Harmon. 2014. ``Predicting Rates of
-Interspecific Interaction from Phylogenetic Trees.'' Edited by Jerome
-Chave. \emph{Ecology Letters} 18 (1). Wiley: 17--27.
-
-\leavevmode\hypertarget{ref-Nuismer2012}{}%
-Nuismer, Scott L., Pedro Jordano, and Jordi Bascompte. 2012.
-``Coevolution and the Architecture of Mutualistic Networks.''
-\emph{Evolution} 67 (2). Wiley: 338--54.
-
-\leavevmode\hypertarget{ref-Nuismer2007}{}%
-Nuismer, Scott L., Benjamin J. Ridenhour, and Benjamin P. Oswald. 2007.
-``Antagonistic Coevolution Mediated by Phenotypic Differenes Between
-Quantitative Traits.'' \emph{Evolution} 61 (8). Wiley: 1823--34.
-
-\leavevmode\hypertarget{ref-Nuismer2018}{}%
-Nuismer, Scott L., Bob Week, and Marcelo A. Aizen. 2018. ``Coevolution
-Slows the Disassembly of Mutualistic Networks.'' \emph{The American
-Naturalist} 192 (4). University of Chicago Press: 490--502.
-
-\leavevmode\hypertarget{ref-VanNuland2019}{}%
-Nuland, Michael E. Van, Ian M. Ware, Joseph K. Bailey, and Jennifer A.
-Schweitzer. 2019. ``Ecosystem Feedbacks Contribute to Geographic
-Variation in Plant-Soil Eco-Evolutionary Dynamics Across a Fertility
-Gradient.'' Edited by Franziska Brunner. \emph{Functional Ecology} 33
-(1). Wiley: 95--106.
-
-\leavevmode\hypertarget{ref-PAGE2002}{}%
-Page, Karen M., and Martin A. Nowak. 2002. ``Unifying Evolutionary
-Dynamics.'' \emph{Journal of Theoretical Biology} 219 (1). Elsevier BV:
-93--98.
-
-\leavevmode\hypertarget{ref-Parachnowitsch2018}{}%
-Parachnowitsch, Amy L, Jessamyn S Manson, and Nina Sletvold. 2018.
-``Evolutionary Ecology of Nectar.'' \emph{Annals of Botany} 123 (2).
-Oxford University Press (OUP): 247--61.
-
-\leavevmode\hypertarget{ref-Park2017}{}%
-Park, Briton, Matthew T. Rutter, Charles B. Fenster, V. Vaughan Symonds,
-Mark C. Ungerer, and Jeffrey P. Townsend. 2017. ``Distributions of
-Mutational Effects and the Estimation of Directional Selection in
-Divergent Lineages ofArabidopsis Thaliana.'' \emph{Genetics} 206 (4).
-Genetics Society of America: 2105--17.
-
-\leavevmode\hypertarget{ref-Parsons2010}{}%
-Parsons, Todd L., Christopher Quince, and Joshua B. Plotkin. 2010.
-``Some Consequences of Demographic Stochasticity in Population
-Genetics.'' \emph{Genetics} 185 (4). Genetics Society of America:
-1345--54.
-
-\leavevmode\hypertarget{ref-Patel2019}{}%
-Patel, Swati, and Reinhard Bürger. 2019. ``Eco-Evolutionary Feedbacks
-Between Prey Densities and Linkage Disequilibrium in the Predator
-Maintain Diversity.'' \emph{Evolution} 73 (8). Wiley: 1533--48.
-
-\leavevmode\hypertarget{ref-Perkins1991}{}%
-Perkins, Edwin A. 1991. ``Conditional Dawson-Watanabe Processes and
-Fleming-Viot Processes.'' In \emph{Seminar on Stochastic Processes,
-1991}, 143--56. Birkhäuser Boston.
-
-\leavevmode\hypertarget{ref-Perkins1992}{}%
----------. 1992. ``Measure-Valued Branching Diffusions with Spatial
-Interactions.'' \emph{Probability Theory and Related Fields} 94 (2).
-Springer Science; Business Media LLC: 189--245.
-
-\leavevmode\hypertarget{ref-edwinarendperkins1995}{}%
----------. 1995. \emph{On the Martingale Problem for Interactive
-Measure-Valued Branching Diffusions}. Amer Mathematical Society.
-
-\leavevmode\hypertarget{ref-Polechov2015}{}%
-Polechová, Jitka, and Nicholas H. Barton. 2015. ``Limits to Adaptation
-Along Environmental Gradients.'' \emph{Proceedings of the National
-Academy of Sciences} 112 (20). Proceedings of the National Academy of
-Sciences: 6401--6.
-
-\leavevmode\hypertarget{ref-PRICE1970}{}%
-Price, George R. 1970. ``Selection and Covariance.'' \emph{Nature} 227
-(5257). Springer Nature: 520--21.
-
-\leavevmode\hypertarget{ref-Queller2017}{}%
-Queller, David C. 2017. ``Fundamental Theorems of Evolution.'' \emph{The
-American Naturalist} 189 (4). University of Chicago Press: 345--53.
-
-\leavevmode\hypertarget{ref-Reimers1989}{}%
-Reimers, Mark. 1989. ``One Dimensional Stochastic Partial Differential
-Equations and the Branching Measure Diffusion.'' \emph{Probability
-Theory and Related Fields} 81 (3). Springer Nature: 319--40.
-
-\leavevmode\hypertarget{ref-Robertson1966}{}%
-Robertson, Alan. 1966. ``A Mathematical Model of the Culling Process in
-Dairy Cattle.'' \emph{Animal Science} 8 (1). Cambridge University Press:
-95--108.
-
-\leavevmode\hypertarget{ref-joanroughgarden1979}{}%
-Roughgarden, Joan. 1979. \emph{Theory of Population Genetics and
-Evolutionary Ecology: An Introduction}. Macmillan.
-
-\leavevmode\hypertarget{ref-Rudman2017}{}%
-Rudman, Seth M., Matthew A. Barbour, Katalin Csilléry, Phillip Gienapp,
-Frederic Guillaume, Nelson G. Hairston Jr, Andrew P. Hendry, et al.
-2017. ``What Genomic Data Can Reveal About Eco-Evolutionary Dynamics.''
-\emph{Nature Ecology \& Evolution} 2 (1). Springer Nature: 9--15.
-
-\leavevmode\hypertarget{ref-Schreiber2017}{}%
-Schreiber, Sebastian J. 2017. ``Coexistence in the Face of
-Uncertainty.'' In \emph{Recent Progress and Modern Challenges in Applied
-Mathematics, Modeling and Computational Science}, 349--84. Springer New
-York.
-
-\leavevmode\hypertarget{ref-Schreiber2018}{}%
-Schreiber, Sebastian J., Swati Patel, and Casey terHorst. 2018.
-``Evolution as a Coexistence Mechanism: Does Genetic Architecture
-Matter?'' \emph{The American Naturalist} 191 (3). University of Chicago
-Press: 407--20.
-
-\leavevmode\hypertarget{ref-Schuster1983}{}%
-Schuster, Peter, and Karl Sigmund. 1983. ``Replicator Dynamics.''
-\emph{Journal of Theoretical Biology} 100 (3). Elsevier BV: 533--38.
-
-\leavevmode\hypertarget{ref-Skovmand2018}{}%
-Skovmand, Lotte H., Charles C.Y. Xu, Maria R. Servedio, Patrik Nosil,
-Rowan D.H. Barrett, and Andrew P. Hendry. 2018. ``Keystone Genes.''
-\emph{Trends in Ecology \& Evolution} 33 (9). Elsevier BV: 689--700.
-
-\leavevmode\hypertarget{ref-Sterner2008}{}%
-Sterner, R.W., and J.J. Elser. 2008. ``Ecological Stoichiometry:
-Overview.'' In \emph{Encyclopedia of Ecology}, 1101--16. Elsevier.
-
-\leavevmode\hypertarget{ref-Taylor1978}{}%
-Taylor, Peter D., and Leo B. Jonker. 1978. ``Evolutionary Stable
-Strategies and Game Dynamics.'' \emph{Mathematical Biosciences} 40
-(1-2). Elsevier BV: 145--56.
-
-\leavevmode\hypertarget{ref-davidtilman1982}{}%
-Tilman, David. 1982. \emph{Resource Competition and Community
-Structure}. Princeton University Press.
-
-\leavevmode\hypertarget{ref-Turelli1980}{}%
-Turelli, Michael. 1980. ``Niche Overlap and Invasion of Competitors in
-Random Environments II. The Effects of Demographic Stochasticity.'' In
-\emph{Biological Growth and Spread}, 119--29. Springer Berlin
-Heidelberg.
-
-\leavevmode\hypertarget{ref-Turelli1984}{}%
----------. 1984. ``Heritable Genetic Variation via Mutation-Selection
-Balance: Lerchs Zeta Meets the Abdominal Bristle.'' \emph{Theoretical
-Population Biology} 25 (2). Elsevier: 138--93.
-
-\leavevmode\hypertarget{ref-Turelli1986}{}%
----------. 1986. ``Gaussian Versus Non-Gaussian Genetic Analyses of
-Polygenic Mutation-Selection Balance.'' In \emph{Evolutionary Processes
-and Theory}, 607--28. Academic Press.
-
-\leavevmode\hypertarget{ref-Turelli2017}{}%
----------. 2017. ``Commentary: Fisher's Infinitesimal Model: A Story for
-the Ages.'' \emph{Theoretical Population Biology} 118 (December).
-Elsevier BV: 46--49.
-
-\leavevmode\hypertarget{ref-pmid7851785}{}%
-Turelli, Michael, and Nick Barton. 1994. ``Genetic and statistical
-analyses of strong selection on polygenic traits: what, me normal?''
-\emph{Genetics} 138 (3): 913--41.
-
-\leavevmode\hypertarget{ref-Volpert2014}{}%
-Volpert, Vitaly. 2014. \emph{Elliptic Partial Differential Equations:
-Volume 2: Reaction-Diffusion Equations}. Springer Basel.
-
-\leavevmode\hypertarget{ref-Walsh}{}%
-Walsh, John B. 1986. ``An Introduction to Stochastic Partial
-Differential Equations.'' In \emph{Lecture Notes in Mathematics},
-265--439. Springer Berlin Heidelberg.
-
-\leavevmode\hypertarget{ref-Watanabe1968}{}%
-Watanabe, Shinzo. 1968. ``A Limit Theorem of Branching Processes and
-Continuous State Branching Processes.'' \emph{Journal of Mathematics of
-Kyoto University} 8 (1). Duke University Press: 141--67.
-
-\leavevmode\hypertarget{ref-Wright97}{}%
-Wright, Sewall. 1931. ``Evolution in Mendelian Populations.''
-\emph{Genetics} 16 (2). Genetics: 97--159.
-
-\leavevmode\hypertarget{ref-Xiao2015}{}%
-Xiao, Xiao, Daniel J. McGlinn, and Ethan P. White. 2015. ``A Strong Test
-of the Maximum Entropy Theory of Ecology.'' \emph{The American
-Naturalist} 185 (3). University of Chicago Press: E70--E80.
-
 \leavevmode\hypertarget{ref-zheng2004nonlinear}{}%

### 2020-04-18 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
split main text and SM

@@
 header-includes:
+  - \usepackage{xr} # for cross-referencing between documents
   - \usepackage{mathtools}
@@
   - \linenumbers
+  - \externaldocument{SM} # externel document to cross-reference
 editor_options: 
@@
 ---
+
+<!--
+
+For cross-referencing to work we need the .aux files.
+Hence, we need to run:
+
+  tinytex::pdflatex("SM.tex", clean=FALSE)
+  tinytex::pdflatex("draft.tex", clean=FALSE)
+  tinytex::pdflatex("SM.tex", clean=FALSE)
+
+-->
 
@@
 
-# Supplementary material (SM)
-
-Throughout this supplement, we set use dot notation for time derivatives so that $\dot f(x,t)=\frac{\partial }{\partial t}f(x,t)$ and set $\Delta=\sum_{i=1}^d\frac{\partial^2}{\partial x_i^2}$ the Laplace operator on $\mathbb{R}^d$, except in \S\ref{caC} where $\Delta$ represents a random variable.
-
-<!--
-
-## Existence, uniqueness and regularity of solution to deterministic PDE (\ref{eq1}) of the main text \label{regular}
-
-Here we show that solutions to the PDE described in the main text exist, are unique and are twice continuously differentiable with respect to space.
-Our approach is an adaptation of that taken in chapter 9 \S2.2 of Volpert [-@Volpert2014].
-
-Recall that the PDE we consider is of the form
-\begin{equation}\label{PDE_supp}
-\begin{matrix}
-\dot\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\Delta\nu(x,t), \\
-\\
-\nu(x,0)=u(x),
-\end{matrix}
-\end{equation}
-where $m(\nu,x)\leq r\in\mathbb{R}$ is shorthand for $m((K\nu)(x,t),x)$ and $(K\nu)(x,t)=\int_\mathbb{R}\kappa(x-y)\nu(y,t)dy$ for some non-negative, bounded $\kappa$.
-Denoting $C(\mathbb{R})$ the set of continuous functions on $\mathbb{R}$ and $L^1(\mathbb{R})$ the Lebesgue space of integrable functions on $\mathbb{R}$ (ie, the set of functions $f:\mathbb{R}\to\mathbb{R}$ that satisfy $\|f\|_1=\int_\mathbb{R}|f(x)|dx<+\infty$ under the equivalence relation $f=g$ iff $\|f-g\|_1=0$), we assume a continuous and integrable initial condition $\nu(\cdot,0)=u\in L^1(\mathbb{R})\cap C(\mathbb{R})$.
-We now prove the existence and uniqueness of solutions to this Cauchy problem.
-
-Set 
-\begin{equation}
-\Phi(x,t)=\frac{\exp\left(-x^2/2\mu t\right)}{\sqrt{2\pi\mu t}}.
-\end{equation}
-Then the semigroup $\{S(t)\}_{t\geq0}$ given by 
-\begin{equation}
-S(t)v=\left\{\begin{matrix}
-\int_0^t\frac{\mu}{2}\Delta\nu(x,s)ds=\int_\mathbb{R}\Phi(x-y,t)v(y)dy, & t>0 \\
- & \\
-v, &t=0
-\end{matrix}\right.
-\end{equation}
-for all $v\in L^\infty(\mathbb{R})$ satisfies $\|S(t)v\|_1=\|v\|_1$ for all $t\geq0$ and $v\in L^1(\mathbb{R})$. We also define the operator
-\begin{equation}
-(Fv)(x)=m\left(\int_\mathbb{R}\kappa(x-y)v(y)dy,x\right)v(x), \ \ \forall \ \ v\in L^1(\mathbb{R}).
-\end{equation}
-Thus $\|Fv\|_1\leq |r| \ \|v\|_1$ for all $v\in L^1(\mathbb{R})$. Then, for any $v,w\in L^1(\mathbb{R})$,
-\begin{multline}
-Fv-Fw=F(v+w)-F(v+w)+Fv-Fw \\
-=m(v+w,x)(v+w)-m(w,x)w-m(v+w,x)(v+w)+m(v,x)v \\
-=\Big(m(v+w,x)-m(w,x)\Big)w-\Big(m(v+w,x)-m(v,x)\Big)v+m(v+w,x)(v-w).
-\end{multline}
-Hence $\|Fv-Fw\|_1\leq2|r|(\|w\|_1+\|v\|_2)+|r|\|v-w\|_1\leq\|r\|\|v-w\|_1$ which implies $F$ is Lipschitz continuous on $L^1(\mathbb{R})$.
-For $T>0$ set $L^{\infty,1}([0,T]\times\mathbb{R})$ as the set of functions $\nu:[0,T]\to L^1(\mathbb{R})$ that satisfy $\|\nu\|_{1,T}=\sup_{t\in[0,T]}\|\nu(t)\|_1<+\infty$ and consider the operator $G$ defined on $L^{\infty,1}([0,T]\times\mathbb{R})$ for fixed $u\in L^1(\mathbb{R})$ by
-\begin{equation}
-(Gv)(t)=S(t)u+\int_0^tS(t-s)(Fv(s))ds \ \ \forall \ \ t\in[0,T].
-\end{equation}
-Then for any $t\in[0,T]$ and $\nu_1,\nu_2\in L^{\infty,1}([0,T]\times\mathbb{R})$,
-
-\begin{multline}
-\|G\nu_1-G\nu_2\|_{1,T}=\sup_{t\in[0,T]}\left\|\int_0^tS(t-s)(F\nu_1(s)-F\nu_2(s))ds\right\|_1 \\ \leq \sup_{t\in[0,T]}t|r|\|\nu_1(t)-\nu_2(t)\|_1\leq T|r|\|\nu_1-\nu_2\|_{1,T}.
-\end{multline}
-Hence, when $T<\frac{1}{|r|}$ the operator $G$ is a contraction operator.
-In this case we apply Banach's fixed point theorem to demonstrate $G$ has a unique fixed point.
-That is, there exists a unique $\nu\in L^{\infty,1}([0,T]\times\mathbb{R})$ such that $G\nu=\nu$.
-Then, for $\nu=\nu(x,t)$ and considering $\nu(t)$ the function $x\to\nu(x,t)$, we have by definition $\nu$ is the mild solution of PDE (\ref{PDE_supp}) with initial condition $\nu(0)=u$.
-To extend this to arbitrarily large $T$ we take a similar approach as Evans [-@lawrenceevans2010] and choose $T_1<\frac{1}{|r|}$ and apply the above argument to the interval $[T_1,2T_1]$.
-Repeating this task a finite number of times shows $\nu$ is the mild solution of PDE (\ref{PDE_supp}) with initial condition $\nu(0)=u$ for arbitrarily large $T>0$.
-
-We shift our focus to showing mild solutions $\nu$ to PDE (\ref{PDE_supp}) are twice differentiable in $x$ and hence are classical solutions.
-Applying lemma 3.3.2 of Henry [-@Henry1981] we have that the mild solution $\nu$ is in fact a strong solution (i.e., differentiable with respect to $t\in(0,T)$) to the abstract Cauchy problem defined by PDE (\ref{PDE_supp}) with initial condition $\nu(0)=u$ for $t\in[0,T]$.
-In particular, this implies that $\nu(t)\in D(\Delta)$ for each $t\in(0,T)$ where $D(\Delta)=W^{2,1}(\mathbb{R})$ is the Sobolov space of integrable functions with integrable weak derivatives at least of order $2$.
-Since $\frac{1}{p}\geq\frac{1}{p+1}\geq\frac{1}{p}-2$ for all $p\geq1$, the Sobolov embedding theorem [@Henry1981] implies $D(\Delta)\subset L^p(\mathbb{R})$, the Lebesgue space of $p$-integrable functions on $\mathbb{R}$, for all $p\geq1$.
-Hence $D(\Delta)\subset W^{2,p}(\mathbb{R})$, the Sobolov space of $p$-integrable functions with weak derivatives at least of order $2$, for all $p\geq1$.
-
-If $v\in C^{2+\delta}(\mathbb{R})$ for some $\delta>0$, then 
-\begin{equation}
-(Kv)(\cdot)=\int_\mathbb{R}\kappa(\cdot-y)v(y)dy=\int_\mathbb{R}\kappa(y)v(\cdot-y)dy\in C^{2+\delta}(\mathbb{R}),
-\end{equation}
-and, since we assume $m(h,x)$ is continuous in both arguments, $Fv=m(Kv,\cdot)v\in C^\delta$.
-
-Why is $F\nu\in C^\delta(\mathbb{R})$ for some $\delta>0$?
-Why does this imply $\nu(\cdot,t)\in C^{2+\delta}(\mathbb{R})$ for all $t\in(0,T)$?
-This argument holds for all $T>0$.
-Hence, solutions to the Cauchy problem (\ref{PDE_supp}) are twice continuously differentiable in their spatial parameter for all $t>0$.
-
--->
-
-## Sufficient conditions for finite mean, variance and total abundance in the deterministic case \label{finite}
-
-Recall that $m(\nu,x)$ is shorthand for $m\big((K\nu)(x,t),x\big)$.
-That is, $m:[0,+\infty)\times\mathbb{R}\to\mathbb{R}$.
-Following our assumptions of the main text, we have that $m(h,x)$ is differentiable with respect to both $x$ and $h$ and there exists $r\in\mathbb{R}$ such that $m(h,x)\leq r$ across all $h\geq0$ and $x\in\mathbb{R}$.
-As in the main text, we also assume the initial condition $u(x)$ is continuous and integrable in $x$ and satisfies
-\begin{equation}
-0\leq\int_{\mathbb{R}}(|x|+x^2)u(x)dx<+\infty
-\end{equation}
-and consider the Cauchy problem
-\begin{equation}\label{PDE_SM}
-\left\{\begin{matrix}
-\dot\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\Delta\nu(x,t) & t>0\\
-\nu(x,0)=u(x) & t=0.
-\end{matrix}\right.
-\end{equation}
-
-According to Theorem 2.5.6 of Zheng [-@zheng2004nonlinear], if the operator $F$ defined by $\nu(x,t)\to m(\nu,x)\nu(x,t)$ is locally Lipschitz, corresponding to equation (\ref{local_lipschitz}) of the main text, then for some maximal $T>0$ problem (\ref{PDE_SM}) admits a unique local classical solution $\nu(x,t)$ for $t\in[0,T)$.
-Furthermore, either $T=+\infty$ or $T<+\infty$ and $\lim_{t\uparrow T}\int_\mathbb{R}|\nu(x,t)|dx=+\infty$.
-In this section we show that our assumption $m(h,x)\leq r$ for all $h\geq0$ and $x\in\mathbb{R}$ implies $T=+\infty$.
-Replacing $m$ with it's upper bound $r\in\mathbb{R}$, equation (\ref{PDE_SM}) reduces to a simple parabolic equation that can be solved using elementary techniques [@stanleyfarlow1993].
-In particular, when $m(\nu,x)\equiv0$ denote the solution to (\ref{PDE_SM}) by $\nu_0(x,t)$.
-Then, denoting
-\begin{equation}
-\Phi(x,t)=\frac{\exp\left(-x^2/2\mu t\right)}{\sqrt{2\pi\mu t}},
-\end{equation}
-we have
-\begin{equation}
-\nu_0(x,t)=\int_{\mathbb{R}}\Phi(x-y,t)\nu(y,0)dy.
-\end{equation}
-In the more general case, when $m(\nu,x)\equiv r\in\mathbb{R}$, equation (\ref{PDE_SM}) has the solution $\nu_r(x,t)=e^{rt}\nu_0(x,t)$.
-Hence, $\nu_r(x,t)\geq0$ for all $x\in\mathbb{R}$ and $\int_\mathbb{R}\nu_r(x,t)dx=e^{rt}N(0)<+\infty$ for all $t\geq0$.
-Furthermore, denoting
-\begin{equation}
-\begin{matrix}
-N_r(t)&=&\int_\mathbb{R}\nu_r(x,t)dx, \\
-p_r(x,t)&=&\nu_r(x,t)/N_r(t), \\
-\bar x_r(t)&=&\int_\mathbb{R}xp_r(x,t)dx, \\
-\sigma^2_r(t)&=&\int_\mathbb{R}(x-\bar x_r(t))^2p_r(x,t)dx, 
-\end{matrix}
-\end{equation}
-we have
-\begin{equation}
-\bar x_r(t)=\int_\mathbb{R}x\int_\mathbb{R}\Phi(x-y,t)p(y,0)dydx=\int_\mathbb{R}yp(y,0)dy=\bar x(0),
-\end{equation}
-\begin{equation}
-\sigma_r^2(t)=\int_\mathbb{R}(x-\bar x_r(t))^2\int_\mathbb{R}\Phi(x-y,t)p(y,0)dydx=\int_\mathbb{R}\Big((y-\bar x(0))^2+\mu t\Big)p(y,0)dy=\sigma^2(0)+\mu t.
-\end{equation}
-Hence, $|\bar x_r(t)|,\sigma_r^2(t)<+\infty$ for all $t\geq0$. For the sake of contradiction, suppose there exists $x\in\mathbb{R}$ and $t\geq0$ such that $\nu(x,t)>\nu_r(x,t)$. Then
-\begin{equation}
-\nu(x,t)-\nu(x,0)=\int_0^tm(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)ds>\int_0^tr\nu_r(x,s)+\frac{\mu}{2}\Delta\nu_r(x,s)ds=\nu_r(x,t)-\nu(x,0)
-\end{equation}
-which implies there exists $h\geq0$ and $x\in\mathbb{R}$ such that $m(h,x)>r$.
-But this contradicts our assumption $m(h,x)\leq r$ for all $h\geq0$ and $x\in\mathbb{R}$.
-So we have $\nu(x,t)\leq\nu_r(x,t)$ for each $x\in\mathbb{R}$ and $t\geq0$.
-This implies that $N(t)=\int_\mathbb{R}\nu(x,t)dx<+\infty$,
-\begin{equation}
-0\leq\int_\mathbb{R}x^2\nu(x,t)dx\leq\int_\mathbb{R}x^2\nu_r(x,t)dx<+\infty
-\end{equation}
-and in particular
-\begin{equation}
-0\leq\sigma^2(t)+\bar x^2(t)=\frac{1}{N(t)}\int_\mathbb{R}x^2\nu(x,t)dx<+\infty
-\end{equation}
-for each $t\geq0$.
-
-## Equilibrium moments for a deterministic population experiencing logistic growth and stabilizing selection \label{equilib}
-
-Here we set out to show, given the initial conditions $\nu(\cdot,0)\in L^1(\mathbb{R})$ such that $\bar x(0)\in\mathbb{R}$, $\sigma^2(0)\in[0,+\infty)$ and $N(0)\in(0,+\infty)$, and given the growth rate
-\begin{equation}\label{m_log_stab}
-m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_\mathbb{R}\nu(y,t)dy
-\end{equation}
-such that $\theta\in\mathbb{R}$, $a,c,\mu>0$ and $r>\tfrac{1}{2}\sqrt{\mu a}$, we have the following stable equilibrium $N_\infty=\lim_{t\to\infty}N(t)=\tfrac{1}{c}(r-\tfrac{1}{2}\sqrt{a\mu})$, $\bar x_\infty=\lim_{t\to\infty}\bar x(t)=\theta$ and $\sigma^2_\infty=\lim_{t\to\infty}\sigma^2(t)=\sqrt{\tfrac{\mu}{a}}$.
-
-The mean fitness can be calculated as
-\begin{equation}
-\bar m(t)=r-\frac{a}{2}((\theta-\bar x(t))^2+\sigma^2(t))-cN(t).
-\end{equation}
-Recalling the ODE derived for $N(t)$,
-\begin{equation}
-\frac{d}{dt}N(t)=\bar m(t)N(t)=(r-\frac{a}{2}((\theta-\bar x(t))^2+\sigma^2(t))-cN(t))N(t),
-\end{equation}
-solving for equilibrium total abundance $\hat N$ amounts to setting $\frac{d}{dt}N(t)=0$ and solving for $N(t)$.
-Ignoring the equilibrium $N(t)=0$, this reduces to solving $\bar m(t)=0$ for $N(t)$, which returns
-\begin{equation}
-\hat N=\frac{1}{c}\left(r-\frac{a}{2}\left((\theta-\hat{\bar x})^2-\hat\sigma^2\right)\right).
-\end{equation}
-Unfortunately, deriving ODE for $\bar x(t)$ and $\sigma^2(t)$ leads to expressions involving higher moments and finding ODE for these higher moments will lead to expressions involving yet even higher moments.
-To avoid this infinite regression, we find the equilibrium abundance density $\hat\nu(x)$ by solving $\frac{\partial}{\partial t}\nu(x,t)=0$ for $\nu(x,t)$.
-This implies the following ordinary differential equation
-\begin{equation}
-\frac{d^2}{dx^2}\hat\nu(x)=\left(\frac{2c}{\mu}\hat N+\frac{a}{\mu}(\theta-x)^2-\frac{2r}{\mu}\right)\hat\nu(x)
-\end{equation}
-which has the solution
-\begin{equation}
-\hat\nu(x)=\frac{\hat N}{\sqrt{2\pi}}\left(\frac{a}{\mu}\right)^{\frac{1}{4}}\exp\left(-\sqrt\frac{a}{\mu}\frac{(\theta-x)^2}{2}\right).
-\end{equation}
-From this expression we infer $\hat{\bar x}=\theta$ and $\hat\sigma^2=\sqrt{\frac{\mu}{a}}$.
-Hence $\hat N=\frac{1}{c}\left(r-\frac{1}{2}\sqrt{a\mu}\right)$.
-To show that $N_\infty=\hat N$, $\bar x_\infty=\hat{\bar x}$ and $\sigma^2_\infty=\hat\sigma^2$, we linearize ODE for $N(t)$, $\bar x(t)$ and $\sigma^2(t)$ using the equilibrium $\hat\nu(x)$.
-In particular, since $\hat\nu(x)$ is Gaussian, we do not run into the same issue with higher moments as above.
-Assuming the equilibrium $\hat\nu(x)$ does not change the ODE for $N(t)$, but the ODE for $\bar x(t)$ and $\sigma^2(t)$ can now be expressed as
-\begin{equation}
-\frac{d}{dt}\bar x(t)=\sigma^2(t)\left(\frac{\partial\bar m(t)}{\partial\bar x(t)}-\overline{\frac{\partial m(t)}{\partial\bar x(t)}}\right)=a\sigma^2(t)(\theta-\bar x(t)),
-\end{equation}
-\begin{equation}
-\frac{d}{dt}\sigma^2(t)=2\sigma^4(t)\left(\frac{\partial\bar m(t)}{\partial\sigma^2(t)}-\overline{\frac{\partial m(t)}{\partial\sigma^2(t)}}\right)+\mu=\mu-a\sigma^4(t),
-\end{equation}
-where $\hat p(x)=\hat\nu(x)/\hat N$. These expressions confirm our findings that $\hat{\bar x}=\theta$ and $\hat\sigma^2=\sqrt{\frac{\mu}{a}}$. Furthermore, calculating
-\begin{equation}
-\frac{\partial}{\partial\sigma^2(t)}\frac{d}{dt}\sigma^2(t)=-2a\sigma^2(t)
-\end{equation}
-and evaluating at $\sigma^2(t)=\hat\sigma^2$ demonstrates the equilibrium phenotypic variance is stable when $a,\mu>0$. Hence, calculating
-\begin{equation}
-\frac{\partial}{\partial\bar x(t)}\frac{d}{dt}\bar x(t)=-a\sigma^2(t)
-\end{equation}
-and evaluating at $\sigma^2(t)=\hat\sigma^2$ and $\bar x(t)=\hat{\bar x}$ demonstrates the equilibrium phenotypic mean is stable when $a,\mu>0$. Finally, calculating
-\begin{equation}
-\frac{\partial}{\partial N(t)}\frac{d}{dt}N(t)=r-\frac{a}{2}((\theta-\bar x(t))^2+\sigma^2(t))-2cN(t)
-\end{equation}
-and evaluating at $\sigma^2(t)=\hat\sigma^2$, $\bar x(t)=\hat{\bar x}$ and $N(t)=\hat N$ demonstrates the equilibrium total abundance is stable when $a,c,\mu>0$, and $r>\frac{1}{2}\sqrt{a\mu}$.
-
-## The relation between diffusion and convolution with a Gaussian kernel \label{diffconvequiv}
-
-For continuous $g:\mathbb{R}^d\to\mathbb{R}$, consider the deterministic Cauchy problem
-\begin{equation}\label{heateqn}
-\left\{\begin{matrix}
-\dot f(x,t)=&\Delta f(x,t), & (x,t)\in\mathbb{R}^d\times(0,\infty)\\
-f(x,t)=&g(x), & (x,t)\in\mathbb{R}^d\times\{0\}.
-\end{matrix}\right.
-\end{equation}
-According to Evans [-@lawrenceevans2010], the fundamental solution of (\ref{heateqn}) is
-\begin{equation}
-\Phi(x,t)=\frac{1}{(4\pi t)^{d/2}}\exp\left(-\frac{|x|^2}{4t}\right), \ (x,t)\in(0,\infty)\times\mathbb{R}^d,
-\end{equation}
-where $|x|=\sqrt{\sum_ix_i^2}$. The solution $f(x,t)$ of PDE (\ref{heateqn}) is then given by the convolution
-\begin{equation}
-f(x,t)=\int_{\mathbb{R}^d}\Phi(x-y,t)g(y)dy, \ (x,t)\in(0,\infty)\times\mathbb{R}^d.
-\end{equation}
-Hence, by the fundamental theorem of calculus,
-
-\begin{multline}
-f(x,t)+\int_t^{t+1}\dot f(x,s)ds=f(x,t+1) \\
-=\int_{\mathbb{R}^d}\Phi(x-y,t+1)g(y)dy=\int_{\mathbb{R}^d}\int_{\mathbb{R}^d}\Phi(x-y,1)\Phi(y-z,t)g(z)dzdy \\
-=\int_{\mathbb{R}^d}\Phi(x-y,1)f(t,y)dy.
-\end{multline}
-In particular,
-\begin{equation}
-f(x,t)+\int_t^{t+1}\Delta f(x,s)ds=\int_{\mathbb{R}^d}\Phi(1,x-y)f(y,t)dy.
-\end{equation}
-
-## Deterministic dynamics of $\sigma^2(t)$ \label{var_deriv}
-
-Picking up from the main text \S\ref{deterministic},
-
-\begin{multline}
-\frac{d}{dt}\sigma^2(t)=\frac{d}{dt}\int_\mathbb{R}(x-\bar x(t))^2p(x,t)dx=\int_\mathbb{R}2(x-\bar x(t))\frac{d}{dt}{\bar x}(t)+(x-\bar x(t))^2\frac{\partial}{\partial t} p(x,t)dx\\=\int_\mathbb{R}(x-\bar x(t))^2\left((m(\nu,x)-\bar m(t))p(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t)\right)dx\\=\int_\mathbb{R}\left((x-\bar x(t))^2-\sigma^2(t)+\sigma^2(t)\right)(m(\nu,x)-\bar m(t))p(x,t)+(x-\bar x(t))^2\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t)dx\\=\mathrm{Cov}_t\Big((x-\bar x(t))^2,m(\nu,x)\Big)+\frac{\mu}{2}\int_\mathbb{R}(x-\bar x(t))^2\frac{\partial^2}{\partial x^2}p(x,t)dx.
-\end{multline}
-Applying integration by parts twice yields
-\begin{equation}
-\int_{-\infty}^{+\infty}(x-\bar x(t))^2\frac{\partial^2}{\partial x^2}p(x,t)dx=2.
-\end{equation}
-
-## Simplifying covariances with fitness under the assumption of a Gaussian density \label{cov2deriv}
-
-By assuming
-
-\begin{equation}
-p(x,t)=\frac{\exp\left(-\frac{(x-\bar x(t))^2}{2\sigma^2(t)}\right)}{\sqrt{2\pi\sigma^2(t)}}
-\end{equation}
-
-we have
-
-\begin{multline}
-\sigma^2\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right)=\sigma^2\left(\frac{\partial}{\partial\bar x}\int_\mathbb{R}m(\nu,x)p(x,t)dx-\int_\mathbb{R}p(x,t)\frac{\partial}{\partial \bar x}m(\nu,x)dx\right)\\=\sigma^2\int_\mathbb{R}m(\nu,x)\frac{\partial}{\partial\bar x}p(x,t)dx=\sigma^2\int_\mathbb{R}\frac{x-\bar x(t)}{\sigma^2}m(\nu,x)p(x,t)dx\\=\int_\mathbb{R}(x-\bar x)(m(\nu,x)-\bar m)p(x,t)dx=\mathrm{Cov}_t(m,x),
-\end{multline}
-
-and
-
-\begin{multline}
-2\sigma^4\left(\frac{\partial\bar m}{\partial\sigma^2}-\overline{\frac{\partial m}{\partial\sigma^2}}\right)=2\sigma^4\left(\frac{\partial}{\partial\sigma^2}\int_\mathbb{R}m(\nu,x)p(x,t)dx-\int_\mathbb{R}p(x,t)\frac{\partial}{\partial\sigma^2}m(\nu,x)dx\right)\\=2\sigma^4\int_\mathbb{R}\frac{(x-\bar x)^2-\sigma^2}{2\sigma^4}m(\nu,x)p(x,t)dx=\int_\mathbb{R}\left((x-\bar x)^2-\sigma^2\right)\big(m(\nu,x)-\bar m\big)p(x,t)dx\\=\mathrm{Cov}_t\Big((x-\bar x)^2,m\Big).
-\end{multline}
-
-## Comparing our treatment of white noise to Da Prato and Zabczyk (2014) \label{DPZ14}
-
-Our approach in the main text is inspired by the treatment provided in \S4.2 of Da Prato and Zabczyk [-@DaPrato2014].
-Here the authors develop a stochastic integral of operator-valued processes.
-In particular, they consider processes indexed by time $t\geq0$ valued as Hilbert-Schmidt operators $\Phi(t)$ and define the norm
-\begin{equation}
-\|\Phi\|_t=\sqrt{\mathbb{E}\left(\int_0^t\mathrm{Tr}[\Phi(s)\Phi^*(s)]ds\right)}, \ t\geq0.
-\end{equation}
-In our case we only consider the so-called multiplication operators.
-That is, processes that consist of operators $\Phi(t)$ having the form $\Phi(t)g(x)=\varphi(x,t)g(x)$ such that $\varphi(\cdot,t)\in L^2(\mathbb{R})$ a.s.
-for each $t\geq0$.
-In this case $\Phi(t)=\Phi^*(t)$ and
-\begin{equation}
-\|\Phi\|_t=\|\varphi\|_t=\sqrt{\mathbb{E}\left(\int_0^t\int_\mathbb{R}\varphi^2(x,s)dxds\right)}, \ t\geq0.
-\end{equation}
-Da Prato and Zabczyk [-@DaPrato2014] form the space $\mathscr{N}^2_W(0,T)$ of Hilbert-Schmidt operator-valued predictable processes $\Phi(t)$ that satisfy $\|\Phi\|_T<+\infty$ for some $T>0$.
-This corresponds to our more specialized space $\mathscr{N}_2$ that consists of $L^2(\mathbb{R})$ valued processes $\varphi(x,t)$ such that $\|\varphi\|_t<+\infty$ for all $t\geq0$.
-In their treatment, $W(t)$ plays a similar role to our generalized process $\mathbf{W}_t$.
-For $\Phi\in\mathscr{N}^2_W(0,T)$, they denote the stochastic integral for $t\in[0,T]$ by $\Phi\cdot W(t)$.
-Hence, for $\Phi(t) g(x)=\varphi(x,t)g(x)$ as above, $\mathbf{W}_t(\varphi)=\Phi\cdot W(t)$.
-The authors then prove the following:
-
-  > __Proposition 4.28__ _Assume that $\Phi_1,\Phi_2\in\mathscr{N}^2_W(0,T)$. Then_
-  >
-  > $$\mathbb{E}(\Phi_i\cdot W(t))=0, \ \ \mathbb{E}(\|\Phi_i\cdot W(t)\|^2)<+\infty, \ \ \forall \ t\in[0,T].$$
-  >
-  > __Corollory 4.29__ _Under the same assumptions as Proposition 4.28,_
-  >
-  > $$\mathbb{C}(\Phi_1\cdot W(t),\Phi_2\cdot W(s))=\mathbb{E}\left(\int_0^{t\wedge s}\mathrm{Tr}[\Phi_2(r)\Phi_1^*(r)]dr\right), \ \ \forall \ t,s\in[0,T].$$
-
-Simplifying these expressions for the multiplication operators described above returns equations (\ref{exp_WN_xi}) and (\ref{cov_WN_xi}) of the main text.
-
-## Simulating the rescaled process and numerical evidence of approximate normality \label{numerical}
-
-Here we use a numerical argument to suggest, for
-
-\begin{equation}
-r-\frac{a}{2}(\theta-x)^2-c\int_\mathbb{R}\nu(x,t)dx \leq m(\nu,x)\leq r-\frac{a}{2}(\theta-x)^2, \ \forall (\nu,x)\in C_{1,c}^+(\mathbb{R})\times\mathbb{R},
-\end{equation}
-
-the density process $\nu(x,t)$ defined by SPDE (\ref{SPDE}) of the main text satisfies $\int_\mathbb{R}(|x|+x^2)\nu(x,t)dx<\infty$.
-
-and setting $c<r-(\eta a+\sqrt{\mu a})/2$ ensures $\hat N>1$, which is important for numerical simulations when $N$ is an integer.
-We use these results to help guide our choice of parameter values for simulations of the branching random walk.
-In the following section we provide a detailed description of the branching random walk and how we have chosen to rescale it.
-We then use the rescaled branching random walk to investigate finiteness of moments and normality.
-
-### Description of simulation \label{description}
-
-We begin by describing the branching random walk before introducing our scheme to rescale it.
-Our branching random walk follows closely the description of branching Brownian motion in the main text.
-However, we replace exponentially distributed lifetimes with deterministic unit time steps for easier implementation.
-Hence, we restrict time to $t=0,1,2,\dots$.
-Furthermore, we allow individual fitness to depend on both trait value and the state of the entire population.
-For time $t$ we write $\{\xi_1(t),\dots,\xi_{N(t)}(t)\}$ as the set of trait values across all $N(t)$ individuals alive in the population.
-Since our simulation treats individuals instead of continuous distributions of trait values, we can write
-
-\begin{equation}
-\nu(x,t)=\sum_{i=1}^{N(t)}\delta(x-\xi_i(t)),
-\end{equation}
-
-where $\delta(x-\xi_i(t))$ denotes the point mass located at $\xi_i(t)$.
-To allow for imperfect heritability, we also track the set of breeding values which, at time $t$, is denoted by $\{\gamma_1(t),\dots,\gamma_{N(t)}(t)\}$ and should not be confused with the quadratic selection gradients discussed in \S\ref{diffuse} of the main text.
-Then the distribution of breeding values can be written as
-
-\begin{equation}
-\rho(g,t)=\sum_{i=1}^{N(t)}\delta(g-\gamma_i(t)).
-\end{equation}
-
-Following our model of heritiability, the trait value $\xi_i(t)$ is drawn from a normal distribution centered on $\gamma_i(t)$ with variance $\eta$.
-At each iteration we draw, for each individual, a random number of offspring from a Negative-Binomial distribution.
-Recall the Negative-Binomial distribution models the number of failed Bernoulli trials that occur before a given number of successful trials.
-Denoting $q$ the probability of success for each trial and $s$ the number of successes, the mean and variance is given respectively by
-
-\begin{equation}
-\frac{s(1-q)}{q}, \ \frac{s(1-q)}{q^2}.
-\end{equation}
-
-Then if we require the $i$th individual to have mean number offspring $\mathscr W(\nu,\xi_i)$ and variance equal to $V$, the parameters of the associated Negative-Binomial distribution become
-
-\begin{equation}
-q(\nu,\xi_i)=\frac{\mathscr{W}(\nu,\xi_i)}{V}, \ s(\nu,\xi_i)=\frac{\mathscr{W}^2(\nu,\xi_i)}{V-\mathscr{W}(\nu,\xi_i)}.
-\end{equation}
-
-The imposes the restriction $V>\mathscr{W}(\nu,\xi_i)$.
-For each offpsring produced by the individual with breeding value $\gamma_i(t)$, we assign indepently drawn breeding values normally distributed around $\gamma_i(t)$ with variance $\mu$.
-After breeding values have been assigned, we randomly draw trait values for each offpsring as described above.
-For an overview of our model of inheritance, see \S\ref{inheritance} of the main text.
-This summarizes the basic structure of our simulation.
-To impose selection and density dependent growth rates, we set
-
-\begin{equation}
-\mathscr{W}(\nu,\xi_i)=\exp\left(r-\frac{a}{2}(\theta-\xi_i)^2-c\int_\mathbb{R}\nu(x,t)dx\right),
-\end{equation}
-
-where the above integral becomes $\int_\mathbb{R}\nu(x,t)dx=\sum_{i=1}^{N(t)}1=N(t)$.
-
-__Rescaling__
-
-To rescale the branching random walk by a positive integer $n$, we rescale segregation and mutational variance according to $\eta\to \eta$ and $\mu\to \mu/n$, time by $t\to t/n$ and the reproductive law by $V\to V$ and
-
-\begin{equation}
-\mathscr W(\nu,\xi_i)\to\mathscr W^{(n)}(\nu,\xi_i)=\exp\left(\frac{r}{n}-\frac{a}{2n}(\theta-\xi_i)^2-\frac{c}{n^2}N(t)\right)=\exp\left(\frac{r}{n}-\frac{a}{2n}(\theta-\xi_i)^2-\frac{c}{n}N^{(n)}(t)\right).
-\end{equation}
-
-We also replace individual mass with $\frac{1}{n}$ and write rescaled abundance as $N^{(n)}(t)=\frac{1}{n}N(nt)$.
-Under this rescaling the deterministic equilibrium of the raw numerical abundance becomes
-
-\begin{equation}
-\hat N=\frac{n^2}{c}\left(\frac{r}{n}-\frac{1}{2n}(\eta a+\sqrt{\mu a})\right)=\frac{n}{c}\left(r-\frac{1}{2}(\eta a+\sqrt{\mu a})\right).
-\end{equation}
-
-The deterministic equilibrium of the rescaled abundance is then
-
-\begin{equation}
-\hat N^{(n)}=\frac{1}{c}\left(r-\frac{1}{2}(\eta a+\sqrt{\mu a})\right).
-\end{equation}
-
-When it exists, we denote by $N^{(\infty)}(t)$ the limiting process of $N^{(n)}(t)$. Then
-
-\begin{equation}
-\lim_{n\to\infty}n\left(\mathscr W^{(n)}(\nu,\xi_i)-1\right)=r-\frac{a}{2}(\theta-\xi_i)^2-cN^{(\infty)}(t).
-\end{equation}
-
-Note that, following the notation of Theorem 1 in Méléard and Roelly [-@meleard1992interacting], setting $\lambda_n=n$, $m_n(\nu)=\mathscr W^{(n)}(\nu,\cdot)$ and $\varepsilon_n=1/n$ satisfies their hypotheses $(\mathscr H_0)$-$(\mathscr H_3)$ when $c=0$.
-We have implemented this simulation in the programming language Julia.
-A copy can be found at the url:
-
-\texttt{\url{https://github.com/bobweek/branching.brownian.motion.and.spde}}
-
-For the sake of illustration, we simulated the unscaled process ($n=1$) and the rescaled process with $n=5$ and $n=20$ for $50$ units of time.
-Results are shown in Figure \ref{rescaled}.
-In the following section we use a statistical test to show, for the lower bound on $m(\nu,x)$, the rescaled process converges to a Gaussian density as $n\to\infty$ and $V/N\to0$.
-
-### Evidence of normality \label{ev_normal}
-
-To demonstrate approximate normality of the phenotypic distribution when $V/N$ is small we utilized the one-sided Kolmogorov-Smirnov test.
-This test compares an empirical cumulative distribution function (i.e., a cumulative distribution function generated from simulated data) to a hypothetical cumulative distribution function by providing a distribution for the maximum distance between these curves.
-More precisely if $F_n(x)$ is the empricial distribution function for a sample of size $n$ and $F(x)$ is the hypothetical distribution function, Kolmogorov-Smirnov statistic is $D_n=\sup_x|F_n(x)-F(x)|$.
-
-## Derivation of SDE for $\bar x$ and $\sigma^2$ \label{SDE_DERIV}
-
-Picking up from \S\ref{stochastic} of the main text, we have
-\begin{equation}
-\tilde x(t)=\int_\mathbb{R}x\nu(x,t)dx, \ \ \tilde{\tilde\sigma}^2(t)=\int_\mathbb{R}x^2\nu(x,t)dx
-\end{equation}
-and
-\begin{equation}\label{xtilde}
-\tilde x(t)=\tilde x(0)+\int_0^t\int_\mathbb{R}\nu(x,s)m(\nu,x)x+x\sqrt{V\nu(x,s)}\dot W(x,s)dxds,
-\end{equation}
-\begin{equation}\label{stilde}
-\tilde{\tilde\sigma}^2(t)=\tilde{\tilde\sigma}^2(0)+\int_0^t\int_\mathbb{R}\nu(x,s)\left(m(\nu,x)x^2+\mu\right)+x^2\sqrt{V\nu(x,s)}\dot W(x,s)dxds.
-\end{equation}
-
-<!--
-
-For $\nu(x,t)$ defined in the main text, $h\in C(\mathbb{R})$ and $t\geq0$ we make the following assumptions:
-
-\begin{equation}\label{L1_nu}
-\mathbb{E}\left(\int_\mathbb{R}|h(x)|\nu(x,t)dx\right)<\infty,
-\end{equation}
-\begin{equation}\label{L2_nu}
-\mathbb{E}\left(\int_\mathbb{R}h^2(x)\nu(x,t)dx\right)<\infty,
-\end{equation}
-\begin{equation}\label{L1_deriv}
-\mathbb{E}\left(\int_0^t\int_\mathbb{R}\nu(x,s)\Big|h(x)\Big|\left|m(\nu,x)f(x)+\frac{\mu}{2}\Delta f(x)\right|dxds\right)<\infty, \ \forall f\in C^2_b(\mathbb{R}).
-\end{equation}
-
-Put $H(t)=\int_\mathbb{R}h(x)\nu(x,t)dx$. Then, for non-random and non-negative $\nu_1(x,t)$ that is continuous in both arguments and integrable in $x$,
-\begin{equation}
-\star:=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{E}[H(t+\varepsilon)-H(t)|\nu(x,t)=\nu_1(x,t)]
-\end{equation}
-\begin{equation}
-=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{E}\left[\int_\mathbb{R}h(x)\big(\nu(x,t+\varepsilon)-\nu(x,t)\big)dx\Big|\nu(x,t)=\nu_1(x,t)\right]
-\end{equation}
-\begin{equation}
-=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{E}\left[\int_\mathbb{R}\int_t^{t+\varepsilon}f(x)\left(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)
-+\sqrt{V\nu(x,s)}\dot W(x,s)\right)dsdx\Big|\nu(x,t)=\nu_1(x,t)\right].
-\end{equation}
-
-By assumption (\ref{L1_deriv}) we can use Fubini's theorem to write, with probability one,
-
-\begin{equation}
-\int_\mathbb{R}\int_t^{t+\varepsilon}h(x)\big(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)dsdx=\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\big(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)dxds.
-\end{equation}
-
-By assumption (\ref{L2_nu}) we have $\varphi(x,t)=h(x)\sqrt{V\nu(x,t)}$ implies $\varphi\in L^2_c(\mathbb{R}\times[0,\infty))$. Hence, the following is true by definition;
-
-\begin{equation}
-\int_\mathbb{R}\int_t^{t+\varepsilon}h(x)\sqrt{V\nu(x,s)}\dot W(x,s)dsdx = 
-\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\sqrt{V\nu(x,s)}\dot W(x,s)dxds.
-\end{equation}
-
-Hence,
-
-\begin{multline}
-\mathbb{E}\left[\int_\mathbb{R}\int_t^{t+\varepsilon}h(x)\sqrt{V\nu(x,s)}\dot W(x,s)dsdx\Big|\nu(x,t)=\nu_1(x,t)\right] \\ =\mathbb{E}\left[\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\sqrt{V\nu(x,s)}\dot W(x,s)dxds\Big|\nu(x,t)=\nu_1(x,t)\right]=0
-\end{multline}
-
-and
-
-\begin{equation}
-\star=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{E}\left[\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\left(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)\right)dxds\Big|\nu(x,t)=\nu_1(x,t)\right].
-\end{equation}
-
-By assumption (\ref{L1_deriv}) we know that there exists a $\delta>0$ such that for each positive $\varepsilon<\delta$ the following holds almost surely:
-
-\begin{multline}
-\left|\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\left(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)\right)dxds\right| \\
-\leq\int_t^{t+\delta}\int_\mathbb{R}\Big|h(x)\Big|\left|m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)\right|dxds<\infty.
-\end{multline}
-
-Thus, by Lebesgue's dominated convergence theorem, the drift component of the process $H(t)$ can be computed as
-
-\begin{multline}
-\star=\mathbb{E}\left[\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\left(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)\right)dxds\Big|\nu(x,t)=\nu_1(x,t)\right]\\
-=\int_\mathbb{R}h(x)\big(m(\nu_1,x)\nu_1(x,t)+\frac{\mu}{2}\Delta\nu_1(x,t)\big)dx.
-\end{multline}
-
-To find an expression for the diffusion component of $H(t)$ set
-
-\begin{multline}\label{star2}
-\star\star:=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{V}[H(t+\varepsilon)-H(t)|\nu(x,t)=\nu_1(x,t)] \\
-=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{V}\left[\int_\mathbb{R}h(x)\big(\nu(x,t+h)-\nu(x,t)\big)dx\Big|\nu(x,t)=\nu_1(x,t)\right].
-\end{multline}
-
-We can rewrite the integral inside expression (\ref{star2}) as
-
-\begin{equation}
-\int_\mathbb{R}\int_t^{t+\varepsilon}h(x)\big(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)+\sqrt{V\nu(x,s)}\dot W(x,s)\big)dsdx.
-\end{equation}
-
-We have already found
-
-\begin{multline}
-\mathbb{E}\left[\int_\mathbb{R}\int_t^{t+\varepsilon}h(x)\big(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)+\sqrt{V\nu(x,s)}\dot W(x,s)\big)dsdx\Big|\nu(x,t)=\nu_1(x,t)\right] \\
-=\mathbb{E}\left[\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\big(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)\big)dxds\Big|\nu(x,t)=\nu_1(x,t)\right].
-\end{multline}
-
-Then, since $h^2(x)\nu(x,t)$ is integrable by assumption (\ref{L2_nu}), we have
-
-\begin{multline}
-\star\star=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{E}\left[\left(\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\sqrt{V\nu(x,s)}\dot W(x,s)dsdx\right)^2\Big| \ \nu(x,t)=\nu_1(x,t)\right] \\
-=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{E}\left[\int_t^{t+\varepsilon}\int_\mathbb{R}V h^2(x)\nu(x,s)dxds\Big|\nu(x,t)=\nu_1(x,t)\right].
-\end{multline}
-
-Thus, for any $\delta\geq\varepsilon\geq0$, we have, with probability one,
-
-\begin{equation}
-\int_t^{t+\varepsilon}\int_\mathbb{R}V h^2(x)\nu(x,s)dxds\leq\int_t^{t+\delta}\int_\mathbb{R}V h^2(x)\nu(x,s)dxds.
-\end{equation}
-
-We can therefore use Lebesgue's dominated convergence theorem to justify
-
-\begin{equation}
-\star\star=\mathbb{E}\left[\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\int_t^{t+\varepsilon}\int_\mathbb{R}V h^2(x)\nu(x,s)dxds\Big|\nu(x,t)=\nu_1(x,t)\right]=\int_\mathbb{R}V h^2(x)\nu_1(x,t)dx.
-\end{equation}
-
-Then, using the notation of stochastic differentials, we have
-
-\begin{equation}\label{SDE_deriv}
-dH(t)=\left(\int_\mathbb{R}h(x)\left(m(\nu,x)\nu(x,t)+\frac{\mu}{2}\Delta\nu(x,t)\right)dx\right)dt+\sqrt{V\int_\mathbb{R}h^2(x)\nu(x,t)dx} \ dW(t)
-\end{equation}
-
-where $W$ is a standard Brownian motion. In the following subsections we employ this formula under the cases $h(x)=x,x^2$ to obtain SDE for the phenotypic mean and variance.
-
--->
-
-### Derivation for trait mean
-
-We make use of the notation
-
-\begin{equation}
-\left\{\begin{matrix}
-\|N\|_2 = \sqrt{V\int_\mathbb{R}\nu(x,t)dx}= \sqrt{VN} \\ \\
-\|\tilde x\|_2 = \sqrt{V\int_\mathbb{R}x^2\nu(x,t)dx} \\ \\
-\langle\tilde x,N\rangle = V\int_\mathbb{R}x\nu(x,t)dx=\bar x VN.
-\end{matrix}\right.
-\end{equation}
-
-Rewriting formula (\ref{xtilde}) as an SDE provides
-
-\begin{equation}
-d\tilde x=\left(\overline{xm}N+\frac{\mu}{2}\int_\mathbb{R}x\Delta\nu(x,t)dx\right)dt+\|\tilde x\|_2d\tilde W_2,
-\end{equation}
-
-where,
-
-\begin{equation}
-d\tilde{ W}_2=d\hat{\mathbf W}_t(\sqrt{V x^2\nu})=\frac{1}{\|\tilde x\|_2}\int_\mathbb{R}x\sqrt{V\nu(x,t)}\dot W(x,t)dxdt.
-\end{equation}
-
-Using It&ocirc;'s quotient rule on $\bar x=\tilde x/N$, we obtain
-
-\begin{equation}
-d\bar x = d\left(\frac{\tilde x}{N}\right)=\frac{\tilde x}{N}\left(\frac{d\tilde x}{\tilde x}-\frac{dN}{N}-\frac{d\tilde x}{\tilde x}\frac{dN}{N}+\Big(\frac{dN}{N}\Big)^2\right)
-= \frac{d\tilde x}{N}-\bar x\frac{dN}{N}-\frac{d\tilde x}{N}\frac{dN}{N}+\bar x\Big(\frac{dN}{N}\Big)^2.
-\end{equation}
-
-From Table \ref{mult} of the main text $d\tilde xdN=\langle\tilde x,N\rangle$ and $dN^2=\|N\|_2^2$. Hence,
-
-\begin{equation}
-d\bar x= \overline{xm}dt+\frac{\|\tilde x\|_2}{N}d\tilde W_2-\bar x\left(\bar m dt+\sqrt\frac{V}{N}dW_1\right)-\frac{\langle\tilde x,N\rangle}{N^2}dt+\bar x\frac{\|N\|_2^2}{N^2}dt
-\end{equation}
-
-\begin{equation*}
-= (\overline{xm}-\bar x\bar m)dt+\frac{\|\tilde x\|_2}{N}d\tilde W_2
--\bar x\sqrt{\frac{V}{N}}d W_1-V\frac{\bar x}{N}dt+V\frac{\bar x}{N}dt
-\end{equation*}
-
-\begin{equation*}
-=\mathrm{Cov}_t(x,m)
-+\frac{\|\tilde x\|_2}{N}d\tilde W_2-\bar x\sqrt{\frac{V}{N}}d W_1.
-\end{equation*}
-
-Note that
-
-\begin{equation}
-\frac{\|\tilde x\|_2}{N}d\tilde W_2-\bar x\sqrt{\frac{V}{N}}d W_1=\frac{1}{N}\int_\mathbb{R}x\sqrt{V\nu(x,t)}\dot W(x,t)dx-\frac{\bar x}{N}\int_\mathbb{R}\sqrt{V\nu(x,t)}\dot W(x,t)dx
-\end{equation}
-
-\begin{equation*}
-=\int_\mathbb{R}\frac{x-\bar x}{N}\sqrt{V\nu(x,t)}\dot W(x,t)dx
-\end{equation*}
-
-and
-
-\begin{equation}
-\mathbb{V}\left(\int_\mathbb{R}\frac{x-\bar x}{N}\sqrt{V\nu(x,t)}\dot W(x,t)dx\right)=\frac{V}{N}\int_\mathbb{R}(x-\bar x)^2p(x,t)dx=V\frac{\sigma^2}{N}.
-\end{equation}
-
-Hence, by setting
-
-\begin{equation}
-d W_2=\frac{\int_\mathbb{R}\frac{(x-\bar x)}{N}\sqrt{V\nu(x,t)}\dot W(x,t)dx}{\sqrt{V\sigma^2/N}}
-\end{equation}
-
-we can write
-
-\begin{equation}
-d\bar x=\mathrm{Cov}_t(x,m)dt
-+\sqrt{V\frac{\sigma^2}{N}}d W_2.
-\end{equation}
-
-### Derivation for trait variance
-
-We make use of the notation
-
-\begin{equation}
-\left\{\begin{matrix}
-\|\tilde{\tilde\sigma}^2\|_2 = \sqrt{V\int_\mathbb{R}x^4\nu(x,t)dx} \\ \\
-\langle\tilde{\tilde\sigma}^2,N\rangle=V\int_\mathbb{R}x^2\nu(x,t)dx=\overline{x^2}VN.
-\end{matrix}\right.
-\end{equation}
-
-Applying formula (\ref{stilde}) provides
-
-\begin{equation}
-d\tilde{\tilde\sigma}^2=\left(\overline{x^2m}N+\mu N\right)dt+\|\tilde{\tilde\sigma}^2\|_2d\tilde W_3
-\end{equation}
-
-where
-
-\begin{equation}
-d\tilde{W}_3=d\hat{\mathbf W}_t(\sqrt{V x^4\nu})
-=\frac{1}{\|\tilde{\tilde\sigma}^2\|_2}\int_\mathbb{R}x^2\sqrt{V\nu(x,t)}\dot W(x,t)dx.
-\end{equation}
-
-Using It&ocirc;'s quotient rule on $\overline{x^2}=\tilde{\tilde\sigma}^2/N$, we obtain
-
-\begin{equation}
-d\overline{x^2} = d\left(\frac{\tilde{\tilde\sigma}^2}{N}\right)=\frac{\tilde{\tilde\sigma}^2}{N}\left(\frac{d\tilde{\tilde\sigma}^2}{\tilde{\tilde\sigma}^2}-\frac{dN}{N}-\frac{d\tilde{\tilde\sigma}^2}{\tilde{\tilde\sigma}^2}\frac{dN}{N}+\Big(\frac{dN}{N}\Big)^2\right)
-=\frac{d\tilde{\tilde\sigma}^2}{N}-\overline{x^2}\frac{dN}{N}-\frac{d\tilde{\tilde\sigma}^2}{N}\frac{dN}{N}+\overline{x^2}\left(\frac{dN}{N}\right)^2.
-\end{equation}
-
-Table \ref{mult} of the main text implies $d\tilde W_3d W_1=\langle\tilde{\tilde\sigma}^2,N\rangle$ and hence
-
-\begin{multline}
-d\overline{x^2}=\left(\overline{x^2m} + \mu \right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3
--\overline{x^2}\left(\bar m dt+\sqrt\frac{V}{N}dW_1\right)-\frac{\langle\tilde{\tilde\sigma}^2,N\rangle}{N^2}dt+\overline{x^2}\frac{\|N\|_2^2}{N^2}dt \\ \\
-=\left(\overline{x^2m}-\overline{x^2}\bar mdt + \mu\right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3
--\overline{x^2}\sqrt{\frac{V}{N}}d W_1
--\overline{x^2}\frac{V}{N}dt+\overline{x^2}\frac{V}{N}dt \\ \\
-=\left(\mathrm{Cov}_t\Big(x^2,m\Big) + \mu \right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3-\overline{x^2}\sqrt\frac{V}{N}d W_1.
-\end{multline}
-
-Setting $F(y,z)=y-z^2$, use It&ocirc;'s formula on $\sigma^2=F(\overline{x^2},\bar x)=\overline{x^2}-\bar x^2$ to obtain:
-
-\begin{multline}
-d\sigma^2=d\overline{x^2}-2\bar xd\bar x-(d\bar x)^2
-=\left(\mathrm{Cov}_t\Big(x^2,m\Big) + \mu \right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3-\overline{x^2}\sqrt\frac{V}{N}d W_1 \\ \\
--2\bar x\left(\mathrm{Cov}_t(x,m)+\mu dt+\sqrt{\frac{V\sigma^2}{N}}d W_2\right)-\left(\mathrm{Cov}_t(x,m)dt+\mu dt+\sqrt{\frac{V\sigma^2}{N}}d W_2\right)^2 \\ \\
-=\left(\mathrm{Cov}_t\Big(x^2-2\bar xx,m\Big) + \mu \right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3-\overline{x^2}\sqrt\frac{V}{N}d W_1-2\bar x\sqrt{\frac{V\sigma^2}{N}}d W_2-\left(\frac{V\sigma^2}{N}\right)dt \\ \\
-=\left(\mathrm{Cov}_t\Big(x-\bar x)^2,m\Big) + \mu - \frac{V\sigma^2}{N}\right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3-\overline{x^2}\sqrt\frac{V}{N}d W_1-2\bar x\sqrt{\frac{V\sigma^2}{N}}d W_2.
-\end{multline}
-
-In light of
-
-\begin{multline}
-\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3-\overline{x^2}\sqrt\frac{V}{N}d W_1-2\bar x\sqrt{\frac{V\sigma^2}{N}}d W_2=\frac{1}{N}\int_\mathbb{R}\left(x^2-\tilde\sigma^2-2\bar x(x-\bar x)\right)\sqrt{V\nu(x,t)}\dot W(x,t)dx \\
-=\frac{1}{N}\int_\mathbb{R}\left((x-\bar x)^2-\sigma^2\right)\sqrt{V\nu(x,t)}\dot W(x,t)dx
-\end{multline}
-
-and
-
-\begin{multline}
-\frac{1}{N}\int_\mathbb{R}\Big(\left((x-\bar x)^2-\sigma^2\right)\sqrt{V\nu(x,s)}\Big)^2dx
-=\frac{V}{N}\left(\int_\mathbb{R}((x-\bar x)^4-2(x-\bar x)^2\sigma^2+\sigma^4)p(x,t)dx\right) \\
-=\frac{V}{N}\left(\overline{(x-\bar x)^4}-\sigma^4\right)
-\end{multline}
-
-we set
-
-\begin{equation}
-d W_3=\frac{\int_\mathbb{R}\left((x-\bar x)^2-\sigma^2\right)\sqrt{V\nu(x,t)}\dot W(x,t)dx}{V\left(\overline{(x-\bar x)^4}-\sigma^4\right)}
-\end{equation}
-
-so that
-
-\begin{equation}
-d\sigma^2=\mathrm{Cov}_t\Big((x-\bar x)^2,m\Big)dt+\left(\mu-V\frac{\sigma^2}{N}\right)dt+\sqrt{V\frac{\overline{(x-\bar x)^4}-\sigma^4}{N}}d W_3.
-\end{equation}
-
-Table \ref{mult} of the main text implies
-
-\begin{equation}
-d W_1d W_2=\frac{\int_\mathbb{R}(x-\bar x)\nu(x,t)dx}{\sqrt{N\sigma^2}}dt=0,
-\end{equation}
-
-\begin{equation}
-d W_1d W_3=\frac{\int_\mathbb{R}\big((x-\bar x)^2-\sigma^2\big)\nu(x,t)dx}{\sqrt{\overline{(x-\bar x)^4}-\sigma^4}}dt=0,
-\end{equation}
-
-\begin{equation}
-d W_2d W_3=\frac{\int_\mathbb{R}(x-\bar x)\big((x-\bar x)^2-\sigma^2\big)p(x,t)dx}{\sqrt{\sigma^2\big(\overline{(x-\bar x)^4}-\sigma^4\big)}}dt
-=\frac{N\overline{(x-\bar x)^3}}{\sqrt{\sigma^2\big(\overline{(x-\bar x)^4}-\sigma^4\big)}}dt.
-\end{equation}
-
-In particular, when $p$ is a Gaussian curve $d W_2d W_3=0$.
-
-## Relating fitness of expressed traits to fitness of breeding values \label{fit2fit}
-
-Following \S\ref{inheritance} of the main text, we have $\sigma^2=G+\eta$ and
-
-\begin{equation}
-m^*(\rho,g)=\int_\mathbb{R}m(\nu,x)\psi(x,g)dx.
-\end{equation}
-
-Hence,
-
-\begin{multline}
-\overline{\frac{\partial m^*}{\partial\bar x}}=\int_{\mathbb{R}}\frac{\rho(g,t)}{N(t)}\frac{\partial}{\partial\bar x}\int_\mathbb{R}m(\nu,x)\psi(x,g)dxdg= \\
-\int_{\mathbb{R}}\int_\mathbb{R}\frac{\rho(g,t)}{N(t)}\psi(x,g)dg\frac{\partial}{\partial\bar x}m(\nu,x)dx=\int_\mathbb{R}p(x,t)\frac{\partial}{\partial\bar x}m(\nu,x)dx=\overline{\frac{\partial m}{\partial\bar x}}
-\end{multline}
-and
-
-\begin{multline}
-\overline{\frac{\partial m^*}{\partial G}}=\int_{\mathbb{R}}\frac{\rho(g,t)}{N(t)}\frac{\partial}{\partial G}\int_\mathbb{R}m(\nu,x)\psi(x,g)dxdg=\int_{\mathbb{R}}\int_\mathbb{R}\frac{\rho(g,t)}{N(t)}\psi(x,g)dg\frac{\partial m}{\partial G}dx= \\
-\int_\mathbb{R}p(x,t)\frac{\partial m}{\partial\sigma^2}\frac{\partial\sigma^2}{\partial G}dx=\overline{\frac{\partial m}{\partial\sigma^2}}.
-\end{multline}
-
-## Derivation of diffuse coevolution model \label{diffuse}
-
-In this section we provide a derivation of our model of diffuse coevolution driven by competition.
-Since most of the work in this derivation has already been completed in Supplementary Material \S\ref{SDE_DERIV}, we focus here on deriving the Malthusian fitness $m$ as a function of trait value $x$.
-We begin with discrete populations of individuals.
-In particular, we begin by assuming population size $n_i$ is an integer for each species $i=1,\dots,S$ before passing to the large population size limit.
-
-The reduction in fitness for an individual of species $i$ caused by competition is captured multiplicatively by $0<C_i\leq1$.
-Biologically this assumes all competitors affect individuals of a given species equally by consuming the same amount of resources.
-This is a mean-field interaction since any individual that consumes resources has an effect on the fitness of all other individuals competing for the same resources.
-Denote by $x_{ij}$ the trait value of the $j$-th individual belonging to species $i$.
-The set of trait values across all individuals in the community at time $t\geq0$ is written $X=\{x_{ij}\}$.
-We denote by $\mathscr{B}_{ij}$ a function that maps $X$ to the cumulative effect of all competitive interactions on the fitness of the $j$-th individual in species $i$.
-Since individuals do not compete with themselves the net multiplicative effects on fitness of both interspecific and intraspecific competition on the $j$-th individual in species $i$ can be summarized by
-
-\begin{equation}
-\mathscr B_{ij}(X)=C_i^{\sum_{l\neq j}\mathcal{O}_{ii}(x_{ij}-x_{il})+\sum_{k\neq i}\sum_{l=1}^{n_k}\mathcal{O}_{ik}(x_{ij}-x_{kl})},
-\end{equation}
-
-where $\mathcal{O}_{ij}$, defined in the main text, measures the overlap in resource use between individuals of species $i$ and $j$ as a function of their niche-centers.
-Writing $\mathscr W_{ij}(X)$ as the average number of offspring left by the $j$-th individual of species $i$, we have
-
-\begin{equation}
-\mathscr W_{ij}(X)=\mathscr{A}_i(x_{ij})\mathscr{B}_{ij}(X),
-\end{equation}
-
-where $\mathscr{A}_i(x)=\int_\mathbb{R}e_i(\zeta)u_i(\zeta,x)d\zeta$ accounts for abiotic selection and $e_i$ has been defined in the main text.
-
-We now turn to a diffusion limit.
-Since we have more than one population, we take the diffusion limit for each population one at a time starting with population $1$.
-We write $\pmb n=(n_1,\dots,n_S)$.
-Following Méléard and Roelly [-@Mlard1993; -@meleard1992interacting] we rescale generation time and individual mass to $\frac{1}{n_1}$ and mean of the reproductive law to
-
-\begin{equation}
-\mathscr W_{1j}^{(\pmb n)}(X)=\mathscr{A}_1(x_{1j})^{1/n_1}\exp\left(\frac{\ln C_1}{n_1^2}\sum_{l\neq j}\mathcal{O}_{11}(x_{1j}-x_{1l})+\frac{\ln C_1}{n_1}\sum_{k\neq i}\frac{1}{n_k}\sum_{l=1}^{n_k}\mathcal{O}_{1k}(x_{1j}-x_{kl})\right).
-\end{equation}
-
-For large $n_1$, we have the approximation
-
-\begin{equation}
-\mathscr W_{1j}^{(\pmb n)}(X)\approx\mathscr{A}_1(x_{1j})^{1/n_1}\left(1+\frac{\ln C_1}{n_1^2}\sum_{l\neq j}\mathcal{O}_{11}(x_{1j}-x_{1l})+\frac{\ln C_1}{n_1}\sum_{k\neq 1}\frac{1}{n_k}\sum_{l=1}^{n_k}\mathcal{O}_{1k}(x_{1j}-x_{kl})\right).
-\end{equation}
-Hence
-\begin{equation}
-\lim_{n_1\to\infty}n_1\left(\mathscr W_{1j}^{(\pmb n)}(X)-1\right)=\ln\mathscr{A}_1(x_{1j})+\left(\int_\mathbb{R}\mathcal{O}_{11}(x_{1j}-y)\nu_1(y,t)dy+\sum_{k\neq 1}\frac{1}{n_k}\sum_{l=1}^{n_k}\mathcal{O}_{1k}(x_{1j}-x_{kl})\right)\ln C_1.
-\end{equation}
-
-We write $\lim_{\pmb n\to\pmb\infty}$ for the iterated limit $\lim_{n_S\to\infty}\cdots\lim_{n_1\to\infty}$ and, assuming  $\nu_i(\cdot,t)\in C_1^+(\mathbb{R})$ for $i=1,\dots,S$ and $t\in[0,\infty)$, we set $\pmb\nu=(\nu_1,\dots,\nu_S)$.
-Then, for any $\pmb\nu$, the the diffusion limits for the remaining populations provides the Malthusian parameter for individuals in species $i$ with trait value $x_{1j}$ as
-
-\begin{equation}
-m_1(\pmb\nu,x_{1j}):=\lim_{\pmb n\to\pmb\infty}n_1\left(\mathscr W_{1j}^{(\pmb n)}(X)-1\right)=\ln\mathscr{A}_1(x)+\left(\sum_{k=1}^S\int_\mathbb{R}\mathcal{O}_{1k}(x_{1j}-y)\nu_k(y,t)dy\right)\ln C_1.
-\end{equation}
-
-We compute the average niche overlap of an individual in species $i$ with nich location $x$ across all individuals in species $j$ as
-
-\begin{equation}
-\bar{\mathcal{O}}_{ij}(x,t)=\frac{\int_\mathbb{R}\mathcal{O}_{ij}(x-y)\nu_j(y,t)dy}{\int_\mathbb{R}\nu_j(y,t)dy}.
-\end{equation}
-
-We now assume the resource utilization curves $u_i(\zeta)$ and phenotypic densities $\nu_i(x,t)$ are Gaussian curves for $i=1,\dots,S$. In this case $\bar{\mathcal{O}}_{ij}(x,t)$ simplifies to
-
-\begin{equation}
-\bar{\mathcal{O}}_{ij}(x,t)=\frac{\int_\mathbb{R}\mathcal O_{ij}(x-y)\nu_j(y,t)dy}{\int_\mathbb{R}\nu_j(y,t)dy}
-=\frac{U_iU_j}{\sqrt{2\pi(w_i+w_j+{\sigma_j}^2(t))}}\exp\left(-\frac{(x-\bar x_j(t))^2}{2(w_i+w_j+{\sigma_j}^2(t))}\right).
-\end{equation}
-
-Setting
-
-\begin{subequations}
-\begin{align}
-\sigma_i^2(t) = & \ G_i(t)+\eta_i, \\
-R_i = & \ \ln \left(\frac{Q_iU_i}{\sqrt{1+A_iw_i}}\right), \\
-a_i = & \ \frac{A_i}{1+A_iw_i}, \\
-\tilde b_{ij}(t) = & \ \frac{1}{w_i+w_j+\sigma_j^2(t)}, \\
-c_i = & -\ln C_i,
-\end{align}
-\end{subequations}
-
-we get
-
-\begin{equation}\label{malth_fit}
-m_i(\pmb\nu,x)=R_i-\frac{a_i}{2}(x-\theta_i)^2-c_i\sum_{j=1}^SN_j(t)U_iU_j\sqrt{\frac{\tilde b_{ij}(t)}{2\pi}}e^{-\frac{\tilde b_{ij}(t)}{2}(x-\bar x_j(t))^2}.
-\end{equation}
-
-Hence, our fitness function is bounded above, as required in the main text.
-
-For the remainder of the derivation we suppress notation indicating dependency on $\pmb\nu$, $x$ and $t$. From (\ref{malth_fit}) we calculate
-
-\begin{equation}
-\frac{\partial m_i}{\partial\bar x_i}=c_iN_iU_i^2\tilde b_{ii}(x-\bar x_i)\sqrt{\frac{\tilde b_{ii}}{2\pi}}e^{-\frac{\tilde b_{ii}}{2}(x-\bar x_i)^2}
-\end{equation}
-
-\begin{multline}
-\frac{\partial m_i}{\partial G_i}=\frac{c_iN_iU_i^2}{2}\left(\frac{(x-\bar x_i)^2-G_i-\eta_i-2w_i}{(G_i+\eta_i+2w_i)^2}\right)\sqrt{\frac{\tilde b_{ii}}{2\pi}}e^{-\frac{\tilde b_{ii}}{2}(x-\bar x_i)^2} \\
-=\frac{c_iN_iU_i^2{\tilde b_{ii}}^2}{2}\left((x-\bar x_i)^2-{\sigma_i}^2-2w_i\right)\sqrt{\frac{\tilde b_{ii}}{2\pi}}e^{-\frac{\tilde b_{ii}}{2}(x-\bar x_i)^2}.
-\end{multline}
-
-Note that
-
-\begin{multline}
-\sqrt\frac{\tilde b_{ii}}{2\pi}\exp\left(-\frac{\tilde b_{ii}}{2}(x-\bar x_i)^2\right)\sqrt\frac{1}{2\pi{\sigma_i}^2}\exp\left(-\frac{(x-\bar x_i)^2}{2{\sigma_i}^2}\right) \\
-=\sqrt\frac{1}{2\pi({\sigma_i}^2+1/\tilde b_{ii})}\sqrt\frac{{\sigma_i}^2+1/\tilde b_{ii}}{2\pi{\sigma_i}^2/\tilde b_{ii}}\exp\left(-\frac{{\sigma_i}^2+1/\tilde b_{ii}}{2{\sigma_i}^2/\tilde b_{ii}}(x-\bar x_i)^2\right) \\
-=\sqrt{\frac{1}{4\pi({\sigma_i}^2+w_i)}}\sqrt\frac{2({\sigma_i}^2+w_i)}{2\pi{\sigma_i}^2({\sigma_i}^2+2w_i)}\exp\left(-\frac{{\sigma_i}^2({\sigma_i}^2+2w_i)}{4({\sigma_i}^2+w_i)}(x-\bar x_i)^2\right).
-\end{multline}
-
-Hence,
-
-\begin{equation}
-\overline{\frac{\partial m_i}{\partial\bar x_i}}=0,
-\end{equation}
-
-\begin{multline}
-\overline{\frac{\partial m_i}{\partial G_i}}=\frac{c_iN_iU_i^2}{2({\sigma_i}^2+2w_i)^2}\left(\frac{({\sigma_i}^2+2w_i){\sigma_i}^2}{2(w_i+{\sigma_i}^2)}-{\sigma_i}^2-2w_i\right)\sqrt{\frac{b_{ii}}{2\pi}} \\
-=\frac{c_iN_iU_i^2}{2({\sigma_i}^2+2w_i)}\left(\frac{{\sigma_i}^2}{2({\sigma_i}^2+w_i)}-1\right)\sqrt{\frac{b_{ii}}{2\pi}}=-\frac{c_iN_iU_i^2b_{ii}}{2}\sqrt{\frac{b_{ii}}{2\pi}},
-\end{multline}
-
-where
-
-\begin{equation}
-b_{ij}=\frac{1}{w_i+w_j+{\sigma_i}^2+{\sigma_j}^2}.
-\end{equation}
-
-The average fitness for species $i$ is
-
-\begin{equation}
-\bar m_i=R_i-\frac{a_i}{2}\Big((\bar x_i-\theta_i)^2+G_i+\eta_i\Big) - c_i\sum_{j=1}^SN_jU_iU_j\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}.
-\end{equation}
-
-Thus,
-
-\begin{equation}
-\frac{\partial\bar m_i}{\partial\bar x_i}=a_i(\theta_i-\bar x_i)-c_i\sum_jN_jU_iU_jb_{ij}(\bar x_j-\bar x_i)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2},
-\end{equation}
-
-\begin{equation}
-\frac{\partial\bar m_i}{\partial G_i}=-\frac{a_i}{2}+\frac{c_i}{2}\sum_{j=1}^SN_jU_iU_jb_{ij}\left(1-b_{ij}(\bar x_i-\bar x_j)^2\right)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}.
-\end{equation}
-
-In particular
-
-\begin{equation}
-\frac{\partial\bar m_i}{\partial G_i}-\overline{\frac{\partial m_i}{\partial G_i}}=-\frac{a_i}{2}+\frac{c_i}{2}\left(N_iU_i^2b_{ii}\sqrt\frac{b_{ii}}{2\pi}+\sum_{j=1}^SN_jU_iU_jb_{ij}\left(1-b_{ij}(\bar x_i-\bar x_j)^2\right)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\right).
-\end{equation}
-
-Applying equations (\ref{N}), (\ref{xbarfinal}) and (\ref{Gfinal}) of the main text recovers system (\ref{comm_dynamics}) of the main text.
-
-## The relation between competition coefficients and selection \label{coef_grad_moms}
-
-### Derivation of analytical approximations
-
-Just as with most caclulations in this work, the derivations are straightforward applications of Gaussian products. That is, if
-
-\begin{equation}
-f_1(x)=\frac{1}{\sqrt{2\pi\sigma_1^2}}\exp\left(-\frac{(\mu_1-x)^2}{2\sigma_1^2}\right), \ f_2(x)=\frac{1}{\sqrt{2\pi\sigma_2^2}}\exp\left(-\frac{(\mu_2-x)^2}{2\sigma_2^2}\right),
-\end{equation}
-
-then
-
-\begin{equation}
-f_1(x)f_2(x)=\frac{1}{\sqrt{2\pi(\sigma_1^2+\sigma_2^2)}}\exp\left(-\frac{(\mu_1-\mu_2)^2}{2(\sigma_1^2+\sigma_2^2)}\right)\frac{1}{\sqrt{2\pi\tilde\sigma^2}}\exp\left(-\frac{(\tilde\mu-x)^2}{2\tilde\sigma^2}\right),
-\end{equation}
-
-where
-
-\begin{equation}
-\tilde\mu=\frac{\sigma_2^2\mu_1+\sigma_1^2\mu_2}{\sigma_1^2+\sigma_2^2}, \ \tilde\sigma^2=\frac{\sigma_1^2\sigma_2^2}{\sigma_1^2+\sigma_2^2}.
-\end{equation}
-
-#### Caclulating $\mathrm{Cov}_{f_{\bar X}}(\alpha,\gamma)$
-
-Recalling
-
-\begin{equation}
-\alpha(\bar x_i,\bar x_j)=\frac{c}{\bar r}\sqrt\frac{b}{2\pi}\exp\left(-\frac{b}{2}\big(\bar x_i-\bar x_j\big)^2\right),
-\end{equation}
-
-\begin{equation}
-\gamma(\bar x_i,\bar x_j)=cNb\left(1-b\big(\bar x_i-\bar x_j\big)^2\right) \sqrt\frac{b}{2\pi}\exp\left(-\frac{b}{2}\big(\bar x_i-\bar x_j\big)^2\right),
-\end{equation}
-
-we have
-
-\begin{multline}
-\bar\alpha=
-\int_\mathbb{R}\int_\mathbb{R}\alpha(\bar x_i,\bar x_j)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
-=\frac{c}{\bar r}\int_\mathbb{R}\frac{1}{\sqrt{2\pi(b^{-1}+V_{\bar X})}}\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(b^{-1}+V_{\bar X})}\right)f_{\bar X}(\bar x_j)d\bar x_j
-=\frac{c/\bar r}{\sqrt{2\pi(b^{-1}+2V_{\bar X})}},
-\end{multline}
-
-\begin{multline}
-\bar\gamma=
-\int_\mathbb{R}\int_\mathbb{R}\gamma(\bar x_i,\bar x_j)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
-= cNb\int_\mathbb{R}\left\{1-\left[\left(\frac{\bar{\bar x}+bV_{\bar X}\bar x_j}{1+bV_{\bar X}}-\bar x_j\right)^2+\frac{V_{\bar X}}{1+bV_{\bar X}}\right]\right\}\frac{1}{\sqrt{2\pi(b^{-1}+V_{\bar X})}}\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(b^{-1}+V_{\bar X})}\right)f_{\bar X}(\bar x_j)d\bar x_j \\
-= cNb\int_\mathbb{R}\left\{1-\left[\left(\frac{\bar{\bar x}-\bar x_j}{1+bV_{\bar X}}\right)^2+\frac{V_{\bar X}}{1+bV_{\bar X}}\right]\right\}\frac{1}{\sqrt{2\pi(b^{-1}+V_{\bar X})}}\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(b^{-1}+V_{\bar X})}\right)f_{\bar X}(\bar x_j)d\bar x_j \\
-= cNb\left(1-\frac{(1+bV_{\bar X})V_{\bar X}}{1+2bV_{\bar X}}\frac{1}{(1+bV_{\bar X})^2}-\frac{V_{\bar X}}{1+bV_{\bar X}}\right)\frac{1}{\sqrt{2\pi(b^{-1}+2V_{\bar X})}} \\
-= cNb\left[1-\left(\frac{1}{1+2bV_{\bar X}}+1\right)\frac{V_{\bar X}}{1+bV_{\bar X}}\right]\frac{1}{\sqrt{2\pi(b^{-1}+2V_{\bar X})}} \\
-= cNb\left(1-\frac{2V_{\bar X}}{1+2bV_{\bar X}}\right)\sqrt\frac{b}{2\pi(1+2bV_{\bar X})},
-\end{multline}
-
-\begin{multline}
-\mathrm{Var}_{f_{\bar X}}(\alpha)=
-\int_\mathbb{R}\int_\mathbb{R}\big(\bar\alpha-\alpha(\bar x_i,\bar x_j)\big)^2f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
-=\frac{c^2}{\bar r^2}\left(\sqrt\frac{b}{4\pi}\int_\mathbb{R}\int_\mathbb{R}\sqrt\frac{b}{\pi}\exp\left(-b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j-\frac{1}{2\pi(b^{-1}+2V_{\bar X})}\right) \\
-=\frac{c^2}{\bar r^2}\left(\sqrt\frac{b}{4\pi}\int_\mathbb{R}\sqrt\frac{1}{2\pi(\frac{1}{2b}+V_{\bar X})}\exp\left(-b(\bar{\bar x}-\bar x_j)^2\right)f_{\bar X}(\bar x_j)d\bar x_j-\frac{1}{2\pi(b^{-1}+2V_{\bar X})}\right) \\
-=\frac{c^2}{\bar r^2}\left(\sqrt\frac{b}{4\pi}\sqrt\frac{1}{2\pi(\frac{1}{2b}+2V_{\bar X})}-\frac{1}{2\pi(b^{-1}+2V_{\bar X})}\right)=\frac{c^2b}{2\pi\bar r^2}\left(\frac{1}{\sqrt{1+4bV_{\bar X}}}-\frac{1}{1+2bV_{\bar X}}\right),
-\end{multline}
-
-\begin{multline}
-\mathrm{Cov}_{f_{\bar X}}(\alpha,\gamma)=
-\int_\mathbb{R}\int_\mathbb{R}\big(\bar\alpha-\alpha(\bar x_i,\bar x_j)\big)\big(\bar\gamma-\gamma(\bar x_i,\bar x_j)\big)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
-=\frac{c^2Nb}{2\bar r}\sqrt\frac{b}{\pi}\int_\mathbb{R}\int_\mathbb{R}\big(1-b(\bar x_i-\bar x_j)^2\big)\sqrt\frac{b}{\pi}\exp\left(-b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j-\bar\alpha\bar\gamma \\
-=\frac{c^2Nb}{2\bar r}\sqrt\frac{b}{\pi}\frac{1-2bV_{\bar X}}{\sqrt{2\pi((2b)^{-1}+2V_{\bar X})}}-\frac{c^2Nb}{\bar r}\frac{1-2bV_{\bar X}}{2\pi(b^{-1}+2V_{\bar X})} \\
-=\frac{c^2b^2N}{2\pi\bar r}(1-2bV_{\bar X})\left(\frac{1}{\sqrt{1+4bV_{\bar X}}}-\frac{1}{1+2bV_{\bar X}}\right).
-\end{multline}
-
-#### Caclulating $\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)$
-
-To calculate moments of $|\beta|$ we note that, as a random variable, $|\beta|$ takes a folded normal distribution.
-Setting $\Phi(x)$ equal to the cumulative density function of the standard normal distribution and using the properties of the folded normal distribution, we find
-
-\begin{equation}
-\overline{|\beta|}=\sqrt\frac{2\mathrm{Var}_{f_{\bar X}}(\beta)}{\pi}\exp\left(-\frac{\bar\beta^2}{2\mathrm{Var}_{f_{\bar X}}(\beta)}\right)-\bar\beta\left[1-2\Phi\left(\frac{\bar\beta}{\mathrm{Var}_{f_{\bar X}}(\beta)}\right)\right]
-\end{equation}
-
-\begin{equation}
-\mathrm{Var}_{f_{\bar X}}(|\beta|)=\bar\beta^2+\mathrm{Var}_{f_{\bar X}}(\beta)-\overline{|\beta|}^2.
-\end{equation}
-
-Recall that
-
-\begin{equation}
-\beta(\bar x_i,\bar x_j)=cNb\big(\bar x_i-\bar x_j\big)\sqrt\frac{b}{2\pi}\exp\left(-\frac{b}{2}\big(\bar x_i-\bar x_j\big)^2\right)
-\end{equation}
-
-and hence
-
-\begin{multline}
-\bar\beta=
-\int_\mathbb{R}\int_\mathbb{R}\beta(\bar x_i,\bar x_j)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
-= cNb\int_\mathbb{R}(\bar{\bar x}-\bar x_j)\frac{1}{\sqrt{2\pi(b^{-1}+V_{\bar X})}}\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(b^{-1}+V_{\bar X})}\right)f_{\bar X}(\bar x_j)d\bar x_j=0,
-\end{multline}
-
-\begin{multline}
-\mathrm{Var}_{f_{\bar X}}(\beta)=
-\int_\mathbb{R}\int_\mathbb{R}\big(\bar\beta-\beta(\bar x_i,\bar x_j)\big)^2f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
-=\int_\mathbb{R}\int_\mathbb{R}c^2N^2b^2(\bar x_i-\bar x_j)^2\frac{b}{2\pi}\exp\left(-b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
-=\sqrt{\frac{b}{4\pi}}c^2N^2b^2\int_\mathbb{R}\left[\left(\frac{\bar{\bar x}+2bV_{\bar X}\bar x_j}{1+2bV_{\bar X}}-\bar x_j\right)^2+\frac{V_{\bar X}}{1+2bV_{\bar X}}\right]\frac{\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(\frac{1}{2b}+V_{\bar X})}\right)}{\sqrt{2\pi(\frac{1}{2b}+V_{\bar X})}}f_{\bar X}(\bar x_j)d\bar x_j \\
-=\sqrt{\frac{b}{4\pi}}c^2N^2b^2\int_\mathbb{R}\left[\frac{(\bar{\bar x}-\bar x_j)^2}{(1+2bV_{\bar X})^2}+\frac{V_{\bar X}}{1+2bV_{\bar X}}\right]\frac{\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(\frac{1}{2b}+V_{\bar X})}\right)}{\sqrt{2\pi(\frac{1}{2b}+V_{\bar X})}}f_{\bar X}(\bar x_j)d\bar x_j \\
-=\sqrt{\frac{b}{4\pi}}c^2N^2b^2\left[\frac{(1+2bV_{\bar X})V_{\bar X}}{1+4bV_{\bar X}}\frac{1}{(1+2bV_{\bar X})^2}+\frac{V_{\bar X}}{1+2bV_{\bar X}}\right]\frac{1}{\sqrt{2\pi(\frac{1}{2b}+2V_{\bar X})}} \\
-=\frac{b}{\pi}\frac{c^2N^2b^2}{\sqrt{1+4bV_{\bar X}}}\frac{V_{\bar X}}{1+2bV_{\bar X}}\left(\frac{1}{1+4bV_{\bar X}}+1\right)=\frac{2c^2N^2b^3V_{\bar X}}{\pi(1+4bV_{\bar X})^{3/2}}.
-\end{multline}
-
-Thus, using properties of the folded normal distribution, we find
-
-\begin{equation}
-\overline{|\beta|}=\sqrt{\frac{2}{\pi}}\frac{cNb^{3/2}}{(1+4bV_{\bar X})^{3/4}}\sqrt\frac{2V_{\bar X}}{\pi}=\frac{2}{\pi}\frac{cNb^{3/2}}{(1+4bV_{\bar X})^{3/4}}\sqrt{V_{\bar X}},
-\end{equation}
-
-\begin{equation}
-\mathrm{Var}_{f_{\bar X}}(|\beta|)=\frac{c^2N^2b^3}{(1+4bV_{\bar X})^{3/2}}\frac{2V_{\bar X}}{\pi}\left(1-\frac{2}{\pi}\right).
-\end{equation}
-
-We also calculate
-
-\begin{multline}
-\mathrm{Cov}_{f_{\bar X}}(\alpha,\beta)=
-\int_\mathbb{R}\int_\mathbb{R}\big(\bar\alpha-\alpha(\bar x_i,\bar x_j)\big)\big(\bar\beta-\beta(\bar x_i,\bar x_j)\big)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
-=\frac{c^2Nb}{2\bar r}\sqrt\frac{b}{\pi}\int_\mathbb{R}\int_\mathbb{R}(\bar x_i-\bar x_j)\sqrt\frac{b}{\pi}\exp\left(-b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j=0.
-\end{multline}
-
-In attempt to calculate $\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)$ we find
-
-\begin{multline}
-\mathrm{Cov}f_{\bar X}(\alpha,|\beta|)=\int_\mathbb{R}\int_\mathbb{R}\alpha(\bar x_i,\bar x_j)|\beta(\bar x_i,\bar x_j)|f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j-\bar\alpha\overline{|\beta|} \\
-=\int_\mathbb{R}\int_\mathbb{R}\frac{c}{\bar r}\sqrt\frac{b}{2\pi}\exp\left(-\frac{b}{2}(\bar x_i-\bar x_j)^2\right)cNb|\bar x_i-\bar x_j|\sqrt\frac{b}{2\pi}\exp\left(-\frac{b}{2}(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j-\bar\alpha\overline{|\beta|} \\
-=\frac{c^2Nb}{\bar r}\sqrt\frac{b}{4\pi}\int_\mathbb{R}\int_\mathbb{R}|\bar x_i-\bar x_j|\sqrt\frac{b}{\pi}\exp\left(-b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j-\bar\alpha\overline{|\beta|}.
-\end{multline}
-
-Just as we used the folded normal to find $\overline{|\beta|}$ and $\mathrm{Var}_{f_{\bar X}}(|\beta|)$, we can calculate $\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)$ by considering
-
-\begin{equation}
-\int_\mathbb{R}\int_\mathbb{R}(\bar x_i-\bar x_j)\sqrt\frac{b}{\pi}\exp\left(-b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j=0
-\end{equation}
-
-and
-
-\begin{multline}
-\int_\mathbb{R}\int_\mathbb{R}(\bar x_i-\bar x_j)^2\frac{b}{\pi}\exp\left(-2b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
-=\sqrt\frac{2b}{\pi}\int_\mathbb{R}\int_\mathbb{R}(\bar x_i-\bar x_j)^2\frac{1}{\sqrt{2\pi\frac{1}{4b}}}\exp\left(-\frac{(\bar x_i-\bar x_j)^2}{2\frac{1}{4b}}\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
-=\sqrt\frac{2b}{\pi}\int_\mathbb{R}\left[\left(\frac{\bar{\bar x}+4bV_{\bar X}\bar x_j}{1+4bV_{\bar X}}-\bar x_j\right)^2+\frac{V_{\bar X}}{1+4bV_{\bar X}}\right]\frac{1}{\sqrt{2\pi(\frac{1}{4b}+V_{\bar X})}}\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(\frac{1}{4b}+V_{\bar X})}\right)f_{\bar X}(\bar x_j)d\bar x_j \\
-=\sqrt\frac{2b}{\pi}\int_\mathbb{R}\left[\left(\frac{\bar{\bar x}-\bar x_j}{1+4bV_{\bar X}}\right)^2+\frac{V_{\bar X}}{1+4bV_{\bar X}}\right]\frac{1}{\sqrt{2\pi(\frac{1}{4b}+V_{\bar X})}}\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(\frac{1}{4b}+V_{\bar X})}\right)f_{\bar X}(\bar x_j)d\bar x_j \\
-=\sqrt\frac{2b}{\pi}\left[\frac{(1+4bV_{\bar X})V_{\bar X}}{1+8bV_{\bar X}}\frac{1}{(1+4bV_{\bar X})^2}+\frac{V_{\bar X}}{1+4bV_{\bar X}}\right]\frac{1}{\sqrt{2\pi(\frac{1}{4b}+2V_{\bar X})}} \\
-=\sqrt\frac{2b}{\pi}\frac{2V_{\bar X}}{1+8bV_{\bar X}}\sqrt{\frac{4b}{2\pi(1+8bV_{\bar X})}}=\frac{b}{\pi}\frac{4V_{\bar X}}{(1+8bV_{\bar X})^{3/2}}.
-\end{multline}
-
-Hence
-
-\begin{equation}
-\int_\mathbb{R}\int_\mathbb{R}|\bar x_i-\bar x_j|\sqrt\frac{b}{\pi}\exp\left(-b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j=\sqrt\frac{2}{\pi}\sqrt{\frac{b}{\pi}\frac{4V_{\bar X}}{(1+8bV_{\bar X})^{3/2}}}=\frac{2}{\pi}\frac{\sqrt{2bV_{\bar X}}}{(1+8bV_{\bar X})^{3/4}}
-\end{equation}
-
-and
-
-\begin{multline}
-\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)=\frac{c^2Nb}{\bar r}\sqrt\frac{b}{4\pi}\frac{2}{\pi}\frac{\sqrt{2bV_{\bar X}}}{(1+8bV_{\bar X})^{3/4}}-\bar\alpha\overline{|\beta|} \\
-=\frac{2c^2Nb^2}{\pi\bar r(1+8bV_{\bar X})^{3/4}}\sqrt\frac{V_{\bar X}}{2\pi}-\frac{c}{\bar r}\sqrt\frac{b}{2\pi(1+2bV_{\bar X})}\frac{2}{\pi}\frac{cNb^{3/2}}{(1+4bV_{\bar X})^{3/4}}\sqrt{V_{\bar X}} \\
-=\frac{2c^2Nb^2}{\pi\bar r(1+8bV_{\bar X})^{3/4}}\sqrt\frac{V_{\bar X}}{2\pi}-\frac{2c^2Nb^2}{\pi\bar r(1+4bV_{\bar X})^{3/4}}\sqrt\frac{V_{\bar X}}{2\pi(1+2bV_{\bar X})} \\
-=\frac{2c^2Nb^2}{\pi\bar r}\sqrt\frac{V_{\bar X}}{2\pi}\left(\frac{1}{(1+8bV_{\bar X})^{3/4}}-\frac{1}{(1+4bV_{\bar X})^{3/4}(1+2bV_{\bar X})^{1/2}}\right).
-\end{multline}
-
-#### Starting the caclulation of $\mathrm{Cov}_{f_{\bar X}}(\alpha,\mathfrak{C})$ \label{caC}
-
-We have
-
-\begin{equation}
-\mathfrak{C}(\bar x_i,\bar x_j)=c^2N^2b^2\Big(|\bar x_i-\bar x_j|+\big|1-b(\bar x_i-\bar x_j)^2\big|\Big)^2\exp\left(-\frac{b}{2}(\bar x_i-\bar x_j)^2\right).
-\end{equation}
-
-Note that the random variable $\delta=\bar x_i-\bar x_j$ is a mean zero Gaussian random variable with variance $2V_{\bar X}$.
-We write the probability density function of $\delta$ as $f_{\Delta}(\delta)$.
-Substituting in $\delta$, we can write
-
-\begin{multline}
-\mathfrak{C}(\delta,0)=c^2N^2b^2\Big(|\delta|+\big|1-b\delta^2\big|\Big)^2\exp\left(-\frac{b}{2}\delta^2\right) \\
-=c^2N^2b^2\Big(\delta^2+2\big||\delta|-b|\delta|^3\big|+\big(1-b\delta^2\big)^2\Big)\exp\left(-\frac{b}{2}\delta^2\right).
-\end{multline}
-
-From this expression, we see properties of the folded normal distribution can be used to calculate several components of the intregal $\mathrm{Cov}_{f_{\bar X}}(\alpha,\mathfrak{C})$, but a major technical challenge lies in calculating
-
-\begin{equation}
-\int_{\mathbb{R}}\big||\delta|-b|\delta|^3\big|\exp\left(-\frac{b}{2}\delta^2\right)f_{\Delta}(\delta)d\delta.
-\end{equation}
-
-Instead of overcoming this challenge to find an analytical form of $\mathrm{Cov}_{f_{\bar X}}(\alpha,\mathfrak{C})$ we turn to a numerical approach outlined in the following section.
-
-### Numerical estimates for heterogeneous $N$ and $G$
-
-```{r, echo=F, fig.align='center', fig.cap='Numerical estimate for the correlations of selection gradients and competition coefficients.'}
-knitr::include_graphics('/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/corrs.png')
-```
-
-Details on simulations, table of parameters, distributions of $a$ and $c$.
-
-\newpage
-
 # References

### 2020-04-18 .A White Noise Approach To Evolutionary Ecology/draft.tex
split main text and SM

@@
 \linenumbers
+\externaldocument{SM}
 \usepackage{float}
@@
 
-\hypertarget{supplementary-material-sm}{%
-\section{Supplementary material (SM)}\label{supplementary-material-sm}}
-
-Throughout this supplement, we set use dot notation for time derivatives
-so that \(\dot f(x,t)=\frac{\partial }{\partial t}f(x,t)\) and set
-\(\Delta=\sum_{i=1}^d\frac{\partial^2}{\partial x_i^2}\) the Laplace
-operator on \(\mathbb{R}^d\), except in \S\ref{caC} where \(\Delta\)
-represents a random variable.
-
-\hypertarget{sufficient-conditions-for-finite-mean-variance-and-total-abundance-in-the-deterministic-case}{%
-\subsection{\texorpdfstring{Sufficient conditions for finite mean,
-variance and total abundance in the deterministic case
-\label{finite}}{Sufficient conditions for finite mean, variance and total abundance in the deterministic case }}\label{sufficient-conditions-for-finite-mean-variance-and-total-abundance-in-the-deterministic-case}}
-
-Recall that \(m(\nu,x)\) is shorthand for \(m\big((K\nu)(x,t),x\big)\).
-That is, \(m:[0,+\infty)\times\mathbb{R}\to\mathbb{R}\). Following our
-assumptions of the main text, we have that \(m(h,x)\) is differentiable
-with respect to both \(x\) and \(h\) and there exists \(r\in\mathbb{R}\)
-such that \(m(h,x)\leq r\) across all \(h\geq0\) and \(x\in\mathbb{R}\).
-As in the main text, we also assume the initial condition \(u(x)\) is
-continuous and integrable in \(x\) and satisfies \begin{equation}
-0\leq\int_{\mathbb{R}}(|x|+x^2)u(x)dx<+\infty
-\end{equation} and consider the Cauchy problem
-\begin{equation}\label{PDE_SM}
-\left\{\begin{matrix}
-\dot\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\Delta\nu(x,t) & t>0\\
-\nu(x,0)=u(x) & t=0.
-\end{matrix}\right.
-\end{equation}
-
-According to Theorem 2.5.6 of Zheng (2004), if the operator \(F\)
-defined by \(\nu(x,t)\to m(\nu,x)\nu(x,t)\) is locally Lipschitz,
-corresponding to equation (\ref{local_lipschitz}) of the main text, then
-for some maximal \(T>0\) problem (\ref{PDE_SM}) admits a unique local
-classical solution \(\nu(x,t)\) for \(t\in[0,T)\). Furthermore, either
-\(T=+\infty\) or \(T<+\infty\) and
-\(\lim_{t\uparrow T}\int_\mathbb{R}|\nu(x,t)|dx=+\infty\). In this
-section we show that our assumption \(m(h,x)\leq r\) for all \(h\geq0\)
-and \(x\in\mathbb{R}\) implies \(T=+\infty\). Replacing \(m\) with it's
-upper bound \(r\in\mathbb{R}\), equation (\ref{PDE_SM}) reduces to a
-simple parabolic equation that can be solved using elementary techniques
-(Farlow 1993). In particular, when \(m(\nu,x)\equiv0\) denote the
-solution to (\ref{PDE_SM}) by \(\nu_0(x,t)\). Then, denoting
-\begin{equation}
-\Phi(x,t)=\frac{\exp\left(-x^2/2\mu t\right)}{\sqrt{2\pi\mu t}},
-\end{equation} we have \begin{equation}
-\nu_0(x,t)=\int_{\mathbb{R}}\Phi(x-y,t)\nu(y,0)dy.
-\end{equation} In the more general case, when
-\(m(\nu,x)\equiv r\in\mathbb{R}\), equation (\ref{PDE_SM}) has the
-solution \(\nu_r(x,t)=e^{rt}\nu_0(x,t)\). Hence, \(\nu_r(x,t)\geq0\) for
-all \(x\in\mathbb{R}\) and
-\(\int_\mathbb{R}\nu_r(x,t)dx=e^{rt}N(0)<+\infty\) for all \(t\geq0\).
-Furthermore, denoting \begin{equation}
-\begin{matrix}
-N_r(t)&=&\int_\mathbb{R}\nu_r(x,t)dx, \\
-p_r(x,t)&=&\nu_r(x,t)/N_r(t), \\
-\bar x_r(t)&=&\int_\mathbb{R}xp_r(x,t)dx, \\
-\sigma^2_r(t)&=&\int_\mathbb{R}(x-\bar x_r(t))^2p_r(x,t)dx, 
-\end{matrix}
-\end{equation} we have \begin{equation}
-\bar x_r(t)=\int_\mathbb{R}x\int_\mathbb{R}\Phi(x-y,t)p(y,0)dydx=\int_\mathbb{R}yp(y,0)dy=\bar x(0),
-\end{equation} \begin{equation}
-\sigma_r^2(t)=\int_\mathbb{R}(x-\bar x_r(t))^2\int_\mathbb{R}\Phi(x-y,t)p(y,0)dydx=\int_\mathbb{R}\Big((y-\bar x(0))^2+\mu t\Big)p(y,0)dy=\sigma^2(0)+\mu t.
-\end{equation} Hence, \(|\bar x_r(t)|,\sigma_r^2(t)<+\infty\) for all
-\(t\geq0\). For the sake of contradiction, suppose there exists
-\(x\in\mathbb{R}\) and \(t\geq0\) such that \(\nu(x,t)>\nu_r(x,t)\).
-Then \begin{equation}
-\nu(x,t)-\nu(x,0)=\int_0^tm(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)ds>\int_0^tr\nu_r(x,s)+\frac{\mu}{2}\Delta\nu_r(x,s)ds=\nu_r(x,t)-\nu(x,0)
-\end{equation} which implies there exists \(h\geq0\) and
-\(x\in\mathbb{R}\) such that \(m(h,x)>r\). But this contradicts our
-assumption \(m(h,x)\leq r\) for all \(h\geq0\) and \(x\in\mathbb{R}\).
-So we have \(\nu(x,t)\leq\nu_r(x,t)\) for each \(x\in\mathbb{R}\) and
-\(t\geq0\). This implies that
-\(N(t)=\int_\mathbb{R}\nu(x,t)dx<+\infty\), \begin{equation}
-0\leq\int_\mathbb{R}x^2\nu(x,t)dx\leq\int_\mathbb{R}x^2\nu_r(x,t)dx<+\infty
-\end{equation} and in particular \begin{equation}
-0\leq\sigma^2(t)+\bar x^2(t)=\frac{1}{N(t)}\int_\mathbb{R}x^2\nu(x,t)dx<+\infty
-\end{equation} for each \(t\geq0\).
-
-\hypertarget{equilibrium-moments-for-a-deterministic-population-experiencing-logistic-growth-and-stabilizing-selection}{%
-\subsection{\texorpdfstring{Equilibrium moments for a deterministic
-population experiencing logistic growth and stabilizing selection
-\label{equilib}}{Equilibrium moments for a deterministic population experiencing logistic growth and stabilizing selection }}\label{equilibrium-moments-for-a-deterministic-population-experiencing-logistic-growth-and-stabilizing-selection}}
-
-Here we set out to show, given the initial conditions
-\(\nu(\cdot,0)\in L^1(\mathbb{R})\) such that
-\(\bar x(0)\in\mathbb{R}\), \(\sigma^2(0)\in[0,+\infty)\) and
-\(N(0)\in(0,+\infty)\), and given the growth rate
-\begin{equation}\label{m_log_stab}
-m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_\mathbb{R}\nu(y,t)dy
-\end{equation} such that \(\theta\in\mathbb{R}\), \(a,c,\mu>0\) and
-\(r>\tfrac{1}{2}\sqrt{\mu a}\), we have the following stable equilibrium
-\(N_\infty=\lim_{t\to\infty}N(t)=\tfrac{1}{c}(r-\tfrac{1}{2}\sqrt{a\mu})\),
-\(\bar x_\infty=\lim_{t\to\infty}\bar x(t)=\theta\) and
-\(\sigma^2_\infty=\lim_{t\to\infty}\sigma^2(t)=\sqrt{\tfrac{\mu}{a}}\).
-
-The mean fitness can be calculated as \begin{equation}
-\bar m(t)=r-\frac{a}{2}((\theta-\bar x(t))^2+\sigma^2(t))-cN(t).
-\end{equation} Recalling the ODE derived for \(N(t)\), \begin{equation}
-\frac{d}{dt}N(t)=\bar m(t)N(t)=(r-\frac{a}{2}((\theta-\bar x(t))^2+\sigma^2(t))-cN(t))N(t),
-\end{equation} solving for equilibrium total abundance \(\hat N\)
-amounts to setting \(\frac{d}{dt}N(t)=0\) and solving for \(N(t)\).
-Ignoring the equilibrium \(N(t)=0\), this reduces to solving
-\(\bar m(t)=0\) for \(N(t)\), which returns \begin{equation}
-\hat N=\frac{1}{c}\left(r-\frac{a}{2}\left((\theta-\hat{\bar x})^2-\hat\sigma^2\right)\right).
-\end{equation} Unfortunately, deriving ODE for \(\bar x(t)\) and
-\(\sigma^2(t)\) leads to expressions involving higher moments and
-finding ODE for these higher moments will lead to expressions involving
-yet even higher moments. To avoid this infinite regression, we find the
-equilibrium abundance density \(\hat\nu(x)\) by solving
-\(\frac{\partial}{\partial t}\nu(x,t)=0\) for \(\nu(x,t)\). This implies
-the following ordinary differential equation \begin{equation}
-\frac{d^2}{dx^2}\hat\nu(x)=\left(\frac{2c}{\mu}\hat N+\frac{a}{\mu}(\theta-x)^2-\frac{2r}{\mu}\right)\hat\nu(x)
-\end{equation} which has the solution \begin{equation}
-\hat\nu(x)=\frac{\hat N}{\sqrt{2\pi}}\left(\frac{a}{\mu}\right)^{\frac{1}{4}}\exp\left(-\sqrt\frac{a}{\mu}\frac{(\theta-x)^2}{2}\right).
-\end{equation} From this expression we infer \(\hat{\bar x}=\theta\) and
-\(\hat\sigma^2=\sqrt{\frac{\mu}{a}}\). Hence
-\(\hat N=\frac{1}{c}\left(r-\frac{1}{2}\sqrt{a\mu}\right)\). To show
-that \(N_\infty=\hat N\), \(\bar x_\infty=\hat{\bar x}\) and
-\(\sigma^2_\infty=\hat\sigma^2\), we linearize ODE for \(N(t)\),
-\(\bar x(t)\) and \(\sigma^2(t)\) using the equilibrium \(\hat\nu(x)\).
-In particular, since \(\hat\nu(x)\) is Gaussian, we do not run into the
-same issue with higher moments as above. Assuming the equilibrium
-\(\hat\nu(x)\) does not change the ODE for \(N(t)\), but the ODE for
-\(\bar x(t)\) and \(\sigma^2(t)\) can now be expressed as
-\begin{equation}
-\frac{d}{dt}\bar x(t)=\sigma^2(t)\left(\frac{\partial\bar m(t)}{\partial\bar x(t)}-\overline{\frac{\partial m(t)}{\partial\bar x(t)}}\right)=a\sigma^2(t)(\theta-\bar x(t)),
-\end{equation} \begin{equation}
-\frac{d}{dt}\sigma^2(t)=2\sigma^4(t)\left(\frac{\partial\bar m(t)}{\partial\sigma^2(t)}-\overline{\frac{\partial m(t)}{\partial\sigma^2(t)}}\right)+\mu=\mu-a\sigma^4(t),
-\end{equation} where \(\hat p(x)=\hat\nu(x)/\hat N\). These expressions
-confirm our findings that \(\hat{\bar x}=\theta\) and
-\(\hat\sigma^2=\sqrt{\frac{\mu}{a}}\). Furthermore, calculating
-\begin{equation}
-\frac{\partial}{\partial\sigma^2(t)}\frac{d}{dt}\sigma^2(t)=-2a\sigma^2(t)
-\end{equation} and evaluating at \(\sigma^2(t)=\hat\sigma^2\)
-demonstrates the equilibrium phenotypic variance is stable when
-\(a,\mu>0\). Hence, calculating \begin{equation}
-\frac{\partial}{\partial\bar x(t)}\frac{d}{dt}\bar x(t)=-a\sigma^2(t)
-\end{equation} and evaluating at \(\sigma^2(t)=\hat\sigma^2\) and
-\(\bar x(t)=\hat{\bar x}\) demonstrates the equilibrium phenotypic mean
-is stable when \(a,\mu>0\). Finally, calculating \begin{equation}
-\frac{\partial}{\partial N(t)}\frac{d}{dt}N(t)=r-\frac{a}{2}((\theta-\bar x(t))^2+\sigma^2(t))-2cN(t)
-\end{equation} and evaluating at \(\sigma^2(t)=\hat\sigma^2\),
-\(\bar x(t)=\hat{\bar x}\) and \(N(t)=\hat N\) demonstrates the
-equilibrium total abundance is stable when \(a,c,\mu>0\), and
-\(r>\frac{1}{2}\sqrt{a\mu}\).
-
-\hypertarget{the-relation-between-diffusion-and-convolution-with-a-gaussian-kernel}{%
-\subsection{\texorpdfstring{The relation between diffusion and
-convolution with a Gaussian kernel
-\label{diffconvequiv}}{The relation between diffusion and convolution with a Gaussian kernel }}\label{the-relation-between-diffusion-and-convolution-with-a-gaussian-kernel}}
-
-For continuous \(g:\mathbb{R}^d\to\mathbb{R}\), consider the
-deterministic Cauchy problem \begin{equation}\label{heateqn}
-\left\{\begin{matrix}
-\dot f(x,t)=&\Delta f(x,t), & (x,t)\in\mathbb{R}^d\times(0,\infty)\\
-f(x,t)=&g(x), & (x,t)\in\mathbb{R}^d\times\{0\}.
-\end{matrix}\right.
-\end{equation} According to Evans (2010), the fundamental solution of
-(\ref{heateqn}) is \begin{equation}
-\Phi(x,t)=\frac{1}{(4\pi t)^{d/2}}\exp\left(-\frac{|x|^2}{4t}\right), \ (x,t)\in(0,\infty)\times\mathbb{R}^d,
-\end{equation} where \(|x|=\sqrt{\sum_ix_i^2}\). The solution \(f(x,t)\)
-of PDE (\ref{heateqn}) is then given by the convolution \begin{equation}
-f(x,t)=\int_{\mathbb{R}^d}\Phi(x-y,t)g(y)dy, \ (x,t)\in(0,\infty)\times\mathbb{R}^d.
-\end{equation} Hence, by the fundamental theorem of calculus,
-
-\begin{multline}
-f(x,t)+\int_t^{t+1}\dot f(x,s)ds=f(x,t+1) \\
-=\int_{\mathbb{R}^d}\Phi(x-y,t+1)g(y)dy=\int_{\mathbb{R}^d}\int_{\mathbb{R}^d}\Phi(x-y,1)\Phi(y-z,t)g(z)dzdy \\
-=\int_{\mathbb{R}^d}\Phi(x-y,1)f(t,y)dy.
-\end{multline} In particular, \begin{equation}
-f(x,t)+\int_t^{t+1}\Delta f(x,s)ds=\int_{\mathbb{R}^d}\Phi(1,x-y)f(y,t)dy.
-\end{equation}
-
-\hypertarget{deterministic-dynamics-of-sigma2t}{%
-\subsection{\texorpdfstring{Deterministic dynamics of \(\sigma^2(t)\)
-\label{var_deriv}}{Deterministic dynamics of \textbackslash{}sigma\^{}2(t) }}\label{deterministic-dynamics-of-sigma2t}}
-
-Picking up from the main text \S\ref{deterministic},
-
-\begin{multline}
-\frac{d}{dt}\sigma^2(t)=\frac{d}{dt}\int_\mathbb{R}(x-\bar x(t))^2p(x,t)dx=\int_\mathbb{R}2(x-\bar x(t))\frac{d}{dt}{\bar x}(t)+(x-\bar x(t))^2\frac{\partial}{\partial t} p(x,t)dx\\=\int_\mathbb{R}(x-\bar x(t))^2\left((m(\nu,x)-\bar m(t))p(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t)\right)dx\\=\int_\mathbb{R}\left((x-\bar x(t))^2-\sigma^2(t)+\sigma^2(t)\right)(m(\nu,x)-\bar m(t))p(x,t)+(x-\bar x(t))^2\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t)dx\\=\mathrm{Cov}_t\Big((x-\bar x(t))^2,m(\nu,x)\Big)+\frac{\mu}{2}\int_\mathbb{R}(x-\bar x(t))^2\frac{\partial^2}{\partial x^2}p(x,t)dx.
-\end{multline} Applying integration by parts twice yields
-\begin{equation}
-\int_{-\infty}^{+\infty}(x-\bar x(t))^2\frac{\partial^2}{\partial x^2}p(x,t)dx=2.
-\end{equation}
-
-\hypertarget{simplifying-covariances-with-fitness-under-the-assumption-of-a-gaussian-density}{%
-\subsection{\texorpdfstring{Simplifying covariances with fitness under
-the assumption of a Gaussian density
-\label{cov2deriv}}{Simplifying covariances with fitness under the assumption of a Gaussian density }}\label{simplifying-covariances-with-fitness-under-the-assumption-of-a-gaussian-density}}
-
-By assuming
-
-\begin{equation}
-p(x,t)=\frac{\exp\left(-\frac{(x-\bar x(t))^2}{2\sigma^2(t)}\right)}{\sqrt{2\pi\sigma^2(t)}}
-\end{equation}
-
-we have
-
-\begin{multline}
-\sigma^2\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right)=\sigma^2\left(\frac{\partial}{\partial\bar x}\int_\mathbb{R}m(\nu,x)p(x,t)dx-\int_\mathbb{R}p(x,t)\frac{\partial}{\partial \bar x}m(\nu,x)dx\right)\\=\sigma^2\int_\mathbb{R}m(\nu,x)\frac{\partial}{\partial\bar x}p(x,t)dx=\sigma^2\int_\mathbb{R}\frac{x-\bar x(t)}{\sigma^2}m(\nu,x)p(x,t)dx\\=\int_\mathbb{R}(x-\bar x)(m(\nu,x)-\bar m)p(x,t)dx=\mathrm{Cov}_t(m,x),
-\end{multline}
-
-and
-
-\begin{multline}
-2\sigma^4\left(\frac{\partial\bar m}{\partial\sigma^2}-\overline{\frac{\partial m}{\partial\sigma^2}}\right)=2\sigma^4\left(\frac{\partial}{\partial\sigma^2}\int_\mathbb{R}m(\nu,x)p(x,t)dx-\int_\mathbb{R}p(x,t)\frac{\partial}{\partial\sigma^2}m(\nu,x)dx\right)\\=2\sigma^4\int_\mathbb{R}\frac{(x-\bar x)^2-\sigma^2}{2\sigma^4}m(\nu,x)p(x,t)dx=\int_\mathbb{R}\left((x-\bar x)^2-\sigma^2\right)\big(m(\nu,x)-\bar m\big)p(x,t)dx\\=\mathrm{Cov}_t\Big((x-\bar x)^2,m\Big).
-\end{multline}
-
-\hypertarget{comparing-our-treatment-of-white-noise-to-da-prato-and-zabczyk-2014}{%
-\subsection{\texorpdfstring{Comparing our treatment of white noise to Da
-Prato and Zabczyk (2014)
-\label{DPZ14}}{Comparing our treatment of white noise to Da Prato and Zabczyk (2014) }}\label{comparing-our-treatment-of-white-noise-to-da-prato-and-zabczyk-2014}}
-
-Our approach in the main text is inspired by the treatment provided in
-\S4.2 of Da Prato and Zabczyk (2014). Here the authors develop a
-stochastic integral of operator-valued processes. In particular, they
-consider processes indexed by time \(t\geq0\) valued as Hilbert-Schmidt
-operators \(\Phi(t)\) and define the norm \begin{equation}
-\|\Phi\|_t=\sqrt{\mathbb{E}\left(\int_0^t\mathrm{Tr}[\Phi(s)\Phi^*(s)]ds\right)}, \ t\geq0.
-\end{equation} In our case we only consider the so-called multiplication
-operators. That is, processes that consist of operators \(\Phi(t)\)
-having the form \(\Phi(t)g(x)=\varphi(x,t)g(x)\) such that
-\(\varphi(\cdot,t)\in L^2(\mathbb{R})\) a.s. for each \(t\geq0\). In
-this case \(\Phi(t)=\Phi^*(t)\) and \begin{equation}
-\|\Phi\|_t=\|\varphi\|_t=\sqrt{\mathbb{E}\left(\int_0^t\int_\mathbb{R}\varphi^2(x,s)dxds\right)}, \ t\geq0.
-\end{equation} Da Prato and Zabczyk (2014) form the space
-\(\mathscr{N}^2_W(0,T)\) of Hilbert-Schmidt operator-valued predictable
-processes \(\Phi(t)\) that satisfy \(\|\Phi\|_T<+\infty\) for some
-\(T>0\). This corresponds to our more specialized space
-\(\mathscr{N}_2\) that consists of \(L^2(\mathbb{R})\) valued processes
-\(\varphi(x,t)\) such that \(\|\varphi\|_t<+\infty\) for all \(t\geq0\).
-In their treatment, \(W(t)\) plays a similar role to our generalized
-process \(\mathbf{W}_t\). For \(\Phi\in\mathscr{N}^2_W(0,T)\), they
-denote the stochastic integral for \(t\in[0,T]\) by \(\Phi\cdot W(t)\).
-Hence, for \(\Phi(t) g(x)=\varphi(x,t)g(x)\) as above,
-\(\mathbf{W}_t(\varphi)=\Phi\cdot W(t)\). The authors then prove the
-following:
-
-\begin{quote}
-\textbf{Proposition 4.28} \emph{Assume that
-\(\Phi_1,\Phi_2\in\mathscr{N}^2_W(0,T)\). Then}
-
-\[\mathbb{E}(\Phi_i\cdot W(t))=0, \ \ \mathbb{E}(\|\Phi_i\cdot W(t)\|^2)<+\infty, \ \ \forall \ t\in[0,T].\]
-
-\textbf{Corollory 4.29} \emph{Under the same assumptions as Proposition
-4.28,}
-
-\[\mathbb{C}(\Phi_1\cdot W(t),\Phi_2\cdot W(s))=\mathbb{E}\left(\int_0^{t\wedge s}\mathrm{Tr}[\Phi_2(r)\Phi_1^*(r)]dr\right), \ \ \forall \ t,s\in[0,T].\]
-\end{quote}
-
-Simplifying these expressions for the multiplication operators described
-above returns equations (\ref{exp_WN_xi}) and (\ref{cov_WN_xi}) of the
-main text.
-
-\hypertarget{simulating-the-rescaled-process}{%
-\subsection{\texorpdfstring{Simulating the rescaled process
-\label{numerical}}{Simulating the rescaled process }}\label{simulating-the-rescaled-process}}
-
-Here we provide a detailed description of the branching random walk and
-how we have chosen to rescale it. We focus on the case \begin{equation}
-m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_\mathbb{R}\nu(x,t)dx.
-\end{equation}
-
-\hypertarget{description-of-simulation}{%
-\subsubsection{\texorpdfstring{Description of simulation
-\label{description}}{Description of simulation }}\label{description-of-simulation}}
-
-We begin by describing the branching random walk before introducing our
-scheme to rescale it. Our branching random walk follows closely the
-description of branching Brownian motion in the main text. However, we
-replace exponentially distributed lifetimes with deterministic unit time
-steps for easier implementation. Hence, we restrict time to
-\(t=0,1,2,\dots,\) and so on. Furthermore, we allow individual fitness
-to depend on both trait value and the state of the entire population.
-For time \(t\) we write \(\{\xi_1(t),\dots,\xi_{N(t)}(t)\}\) as the set
-of trait values across all \(N(t)\) individuals alive in the population.
-Since our simulation follows discrete individuals instead of continuous
-distributions of trait values, we can write \begin{equation}
-\nu(x,t)=\sum_{i=1}^{N(t)}\delta(x-\xi_i(t)),
-\end{equation} where \(\delta(x-\xi_i(t))\) denotes the point mass
-located at \(\xi_i(t)\). For simplicity we assume perfect heritability.
-At each iteration we draw, for each individual, a random number of
-offspring from a Negative-Binomial distribution. Recall the
-Negative-Binomial distribution models the number of failed Bernoulli
-trials that occur before a given number of successful trials. Denoting
-\(q\) the probability of success for each trial and \(s\) the number of
-successes, the mean and variance is given respectively by
-\begin{equation}
-\frac{s(1-q)}{q}, \ \frac{s(1-q)}{q^2}.
-\end{equation} Then if we require the \(i\)th individual to have mean
-number offspring \(\mathscr W(\nu,\xi_i)\) and variance equal to \(V\),
-the parameters of the associated Negative-Binomial distribution become
-\begin{equation}
-q(\nu,\xi_i)=\frac{\mathscr{W}(\nu,\xi_i)}{V}, \ s(\nu,\xi_i)=\frac{\mathscr{W}^2(\nu,\xi_i)}{V-\mathscr{W}(\nu,\xi_i)}.
-\end{equation} This imposes the restriction
-\(V>\mathscr{W}(\nu,\xi_i)\). For each offpsring produced by the
-individual with trait value \(\xi_i(t)\), we assign indepently drawn
-trait values normally distributed around \(\xi_i(t)\) with variance
-\(\mu\). This summarizes the basic structure of our simulation. To
-impose selection and density dependent growth rates, we set
-\begin{equation}
-\mathscr{W}(\nu,\xi_i)=\exp\left(r-\frac{a}{2}(\theta-\xi_i)^2-c\int_\mathbb{R}\nu(x,t)dx\right),
-\end{equation} where the above integral becomes
-\(\int_\mathbb{R}\nu(x,t)dx=\sum_{i=1}^{N(t)}1=N(t)\).
-
-\textbf{Rescaling}
-
-To rescale the branching random walk by a positive integer \(n\), we
-rescale mutational variance by \(\mu\to \mu/n\), time by \(t\to t/n\)
-and the reproductive law by \(V\to V\) and \begin{equation}
-\mathscr W(\nu,\xi_i)\to\mathscr W^{(n)}(\nu,\xi_i)=\exp\left(\frac{r}{n}-\frac{a}{2n}(\theta-\xi_i)^2-\frac{c}{n^2}N(t)\right)=\exp\left(\frac{r}{n}-\frac{a}{2n}(\theta-\xi_i)^2-\frac{c}{n}N^{(n)}(t)\right).
-\end{equation} We also replace individual mass with \(\frac{1}{n}\) and
-write rescaled abundance as \(N^{(n)}(t)=\frac{1}{n}N(nt)\). When it
-exists, we denote by \(N^{(\infty)}(t)\) the limiting process of
-\(N^{(n)}(t)\). Then \begin{equation}
-\lim_{n\to\infty}n\left(\mathscr W^{(n)}(\nu,\xi_i)-1\right)=r-\frac{a}{2}(\theta-\xi_i)^2-cN^{(\infty)}(t).
-\end{equation} Note that, since the limiting fitness function is bounded
-above and decreases linearly with respect to \(N^{(\infty)}(t)\), it
-satisfies the hypotheses of Champagnat, Ferrière and Méléard (2006). We
-have implemented this simulation in the programming language Julia. A
-copy can be found at the url:
-\[\texttt{\url{https://github.com/bobweek/branching.brownian.motion.and.spde}}\]
-For the sake of illustration, we simulated the unscaled process
-(\(n=1\)) and the rescaled process with \(n=5\) and \(n=20\) for \(50\)
-units of time. Results are shown in Figure \ref{rescaled} of the main
-text.
-
-\hypertarget{derivation-of-sde-for-bar-x-and-sigma2}{%
-\subsection{\texorpdfstring{Derivation of SDE for \(\bar x\) and
-\(\sigma^2\)
-\label{SDE_DERIV}}{Derivation of SDE for \textbackslash{}bar x and \textbackslash{}sigma\^{}2 }}\label{derivation-of-sde-for-bar-x-and-sigma2}}
-
-Picking up from \S\ref{stochastic} of the main text, we have
-\begin{equation}
-\tilde x(t)=\int_\mathbb{R}x\nu(x,t)dx, \ \ \tilde{\tilde\sigma}^2(t)=\int_\mathbb{R}x^2\nu(x,t)dx
-\end{equation} and \begin{equation}\label{xtilde}
-\tilde x(t)=\tilde x(0)+\int_0^t\int_\mathbb{R}\nu(x,s)m(\nu,x)x+x\sqrt{V\nu(x,s)}\dot W(x,s)dxds,
-\end{equation} \begin{equation}\label{stilde}
-\tilde{\tilde\sigma}^2(t)=\tilde{\tilde\sigma}^2(0)+\int_0^t\int_\mathbb{R}\nu(x,s)\left(m(\nu,x)x^2+\mu\right)+x^2\sqrt{V\nu(x,s)}\dot W(x,s)dxds.
-\end{equation}
-
-\hypertarget{derivation-for-trait-mean}{%
-\subsubsection{Derivation for trait
-mean}\label{derivation-for-trait-mean}}
-
-We make use of the notation
-
-\begin{equation}
-\left\{\begin{matrix}
-\|N\|_2 = \sqrt{V\int_\mathbb{R}\nu(x,t)dx}= \sqrt{VN} \\ \\
-\|\tilde x\|_2 = \sqrt{V\int_\mathbb{R}x^2\nu(x,t)dx} \\ \\
-\langle\tilde x,N\rangle = V\int_\mathbb{R}x\nu(x,t)dx=\bar x VN.
-\end{matrix}\right.
-\end{equation}
-
-Rewriting formula (\ref{xtilde}) as an SDE provides
-
-\begin{equation}
-d\tilde x=\left(\overline{xm}N+\frac{\mu}{2}\int_\mathbb{R}x\Delta\nu(x,t)dx\right)dt+\|\tilde x\|_2d\tilde W_2,
-\end{equation}
-
-where,
-
-\begin{equation}
-d\tilde{ W}_2=d\hat{\mathbf W}_t(\sqrt{V x^2\nu})=\frac{1}{\|\tilde x\|_2}\int_\mathbb{R}x\sqrt{V\nu(x,t)}\dot W(x,t)dxdt.
-\end{equation}
-
-Using Itô's quotient rule on \(\bar x=\tilde x/N\), we obtain
-
-\begin{equation}
-d\bar x = d\left(\frac{\tilde x}{N}\right)=\frac{\tilde x}{N}\left(\frac{d\tilde x}{\tilde x}-\frac{dN}{N}-\frac{d\tilde x}{\tilde x}\frac{dN}{N}+\Big(\frac{dN}{N}\Big)^2\right)
-= \frac{d\tilde x}{N}-\bar x\frac{dN}{N}-\frac{d\tilde x}{N}\frac{dN}{N}+\bar x\Big(\frac{dN}{N}\Big)^2.
-\end{equation}
-
-From Table \ref{mult} of the main text
-\(d\tilde xdN=\langle\tilde x,N\rangle\) and \(dN^2=\|N\|_2^2\). Hence,
-
-\begin{equation}
-d\bar x= \overline{xm}dt+\frac{\|\tilde x\|_2}{N}d\tilde W_2-\bar x\left(\bar m dt+\sqrt\frac{V}{N}dW_1\right)-\frac{\langle\tilde x,N\rangle}{N^2}dt+\bar x\frac{\|N\|_2^2}{N^2}dt
-\end{equation}
-
-\begin{equation*}
-= (\overline{xm}-\bar x\bar m)dt+\frac{\|\tilde x\|_2}{N}d\tilde W_2
--\bar x\sqrt{\frac{V}{N}}d W_1-V\frac{\bar x}{N}dt+V\frac{\bar x}{N}dt
-\end{equation*}
-
-\begin{equation*}
-=\mathrm{Cov}_t(x,m)
-+\frac{\|\tilde x\|_2}{N}d\tilde W_2-\bar x\sqrt{\frac{V}{N}}d W_1.
-\end{equation*}
-
-Note that
-
-\begin{equation}
-\frac{\|\tilde x\|_2}{N}d\tilde W_2-\bar x\sqrt{\frac{V}{N}}d W_1=\frac{1}{N}\int_\mathbb{R}x\sqrt{V\nu(x,t)}\dot W(x,t)dx-\frac{\bar x}{N}\int_\mathbb{R}\sqrt{V\nu(x,t)}\dot W(x,t)dx
-\end{equation}
-
-\begin{equation*}
-=\int_\mathbb{R}\frac{x-\bar x}{N}\sqrt{V\nu(x,t)}\dot W(x,t)dx
-\end{equation*}
-
-and
-
-\begin{equation}
-\mathbb{V}\left(\int_\mathbb{R}\frac{x-\bar x}{N}\sqrt{V\nu(x,t)}\dot W(x,t)dx\right)=\frac{V}{N}\int_\mathbb{R}(x-\bar x)^2p(x,t)dx=V\frac{\sigma^2}{N}.
-\end{equation}
-
-Hence, by setting
-
-\begin{equation}
-d W_2=\frac{\int_\mathbb{R}\frac{(x-\bar x)}{N}\sqrt{V\nu(x,t)}\dot W(x,t)dx}{\sqrt{V\sigma^2/N}}
-\end{equation}
-
-we can write
-
-\begin{equation}
-d\bar x=\mathrm{Cov}_t(x,m)dt
-+\sqrt{V\frac{\sigma^2}{N}}d W_2.
-\end{equation}
-
-\hypertarget{derivation-for-trait-variance}{%
-\subsubsection{Derivation for trait
-variance}\label{derivation-for-trait-variance}}
-
-We make use of the notation
-
-\begin{equation}
-\left\{\begin{matrix}
-\|\tilde{\tilde\sigma}^2\|_2 = \sqrt{V\int_\mathbb{R}x^4\nu(x,t)dx} \\ \\
-\langle\tilde{\tilde\sigma}^2,N\rangle=V\int_\mathbb{R}x^2\nu(x,t)dx=\overline{x^2}VN.
-\end{matrix}\right.
-\end{equation}
-
-Applying formula (\ref{stilde}) provides
-
-\begin{equation}
-d\tilde{\tilde\sigma}^2=\left(\overline{x^2m}N+\mu N\right)dt+\|\tilde{\tilde\sigma}^2\|_2d\tilde W_3
-\end{equation}
-
-where
-
-\begin{equation}
-d\tilde{W}_3=d\hat{\mathbf W}_t(\sqrt{V x^4\nu})
-=\frac{1}{\|\tilde{\tilde\sigma}^2\|_2}\int_\mathbb{R}x^2\sqrt{V\nu(x,t)}\dot W(x,t)dx.
-\end{equation}
-
-Using Itô's quotient rule on
-\(\overline{x^2}=\tilde{\tilde\sigma}^2/N\), we obtain
-
-\begin{equation}
-d\overline{x^2} = d\left(\frac{\tilde{\tilde\sigma}^2}{N}\right)=\frac{\tilde{\tilde\sigma}^2}{N}\left(\frac{d\tilde{\tilde\sigma}^2}{\tilde{\tilde\sigma}^2}-\frac{dN}{N}-\frac{d\tilde{\tilde\sigma}^2}{\tilde{\tilde\sigma}^2}\frac{dN}{N}+\Big(\frac{dN}{N}\Big)^2\right)
-=\frac{d\tilde{\tilde\sigma}^2}{N}-\overline{x^2}\frac{dN}{N}-\frac{d\tilde{\tilde\sigma}^2}{N}\frac{dN}{N}+\overline{x^2}\left(\frac{dN}{N}\right)^2.
-\end{equation}
-
-Table \ref{mult} of the main text implies
-\(d\tilde W_3d W_1=\langle\tilde{\tilde\sigma}^2,N\rangle\) and hence
-
-\begin{multline}
-d\overline{x^2}=\left(\overline{x^2m} + \mu \right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3
--\overline{x^2}\left(\bar m dt+\sqrt\frac{V}{N}dW_1\right)-\frac{\langle\tilde{\tilde\sigma}^2,N\rangle}{N^2}dt+\overline{x^2}\frac{\|N\|_2^2}{N^2}dt \\ \\
-=\left(\overline{x^2m}-\overline{x^2}\bar mdt + \mu\right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3
--\overline{x^2}\sqrt{\frac{V}{N}}d W_1
--\overline{x^2}\frac{V}{N}dt+\overline{x^2}\frac{V}{N}dt \\ \\
-=\left(\mathrm{Cov}_t\Big(x^2,m\Big) + \mu \right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3-\overline{x^2}\sqrt\frac{V}{N}d W_1.
-\end{multline}
-
-Setting \(F(y,z)=y-z^2\), use Itô's formula on
-\(\sigma^2=F(\overline{x^2},\bar x)=\overline{x^2}-\bar x^2\) to obtain:
-
-\begin{multline}
-d\sigma^2=d\overline{x^2}-2\bar xd\bar x-(d\bar x)^2
-=\left(\mathrm{Cov}_t\Big(x^2,m\Big) + \mu \right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3-\overline{x^2}\sqrt\frac{V}{N}d W_1 \\ \\
--2\bar x\left(\mathrm{Cov}_t(x,m)+\mu dt+\sqrt{\frac{V\sigma^2}{N}}d W_2\right)-\left(\mathrm{Cov}_t(x,m)dt+\mu dt+\sqrt{\frac{V\sigma^2}{N}}d W_2\right)^2 \\ \\
-=\left(\mathrm{Cov}_t\Big(x^2-2\bar xx,m\Big) + \mu \right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3-\overline{x^2}\sqrt\frac{V}{N}d W_1-2\bar x\sqrt{\frac{V\sigma^2}{N}}d W_2-\left(\frac{V\sigma^2}{N}\right)dt \\ \\
-=\left(\mathrm{Cov}_t\Big(x-\bar x)^2,m\Big) + \mu - \frac{V\sigma^2}{N}\right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3-\overline{x^2}\sqrt\frac{V}{N}d W_1-2\bar x\sqrt{\frac{V\sigma^2}{N}}d W_2.
-\end{multline}
-
-In light of
-
-\begin{multline}
-\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3-\overline{x^2}\sqrt\frac{V}{N}d W_1-2\bar x\sqrt{\frac{V\sigma^2}{N}}d W_2=\frac{1}{N}\int_\mathbb{R}\left(x^2-\tilde\sigma^2-2\bar x(x-\bar x)\right)\sqrt{V\nu(x,t)}\dot W(x,t)dx \\
-=\frac{1}{N}\int_\mathbb{R}\left((x-\bar x)^2-\sigma^2\right)\sqrt{V\nu(x,t)}\dot W(x,t)dx
-\end{multline}
-
-and
-
-\begin{multline}
-\frac{1}{N}\int_\mathbb{R}\Big(\left((x-\bar x)^2-\sigma^2\right)\sqrt{V\nu(x,s)}\Big)^2dx
-=\frac{V}{N}\left(\int_\mathbb{R}((x-\bar x)^4-2(x-\bar x)^2\sigma^2+\sigma^4)p(x,t)dx\right) \\
-=\frac{V}{N}\left(\overline{(x-\bar x)^4}-\sigma^4\right)
-\end{multline}
-
-we set
-
-\begin{equation}
-d W_3=\frac{\int_\mathbb{R}\left((x-\bar x)^2-\sigma^2\right)\sqrt{V\nu(x,t)}\dot W(x,t)dx}{V\left(\overline{(x-\bar x)^4}-\sigma^4\right)}
-\end{equation}
-
-so that
-
-\begin{equation}
-d\sigma^2=\mathrm{Cov}_t\Big((x-\bar x)^2,m\Big)dt+\left(\mu-V\frac{\sigma^2}{N}\right)dt+\sqrt{V\frac{\overline{(x-\bar x)^4}-\sigma^4}{N}}d W_3.
-\end{equation}
-
-Table \ref{mult} of the main text implies
-
-\begin{equation}
-d W_1d W_2=\frac{\int_\mathbb{R}(x-\bar x)\nu(x,t)dx}{\sqrt{N\sigma^2}}dt=0,
-\end{equation}
-
-\begin{equation}
-d W_1d W_3=\frac{\int_\mathbb{R}\big((x-\bar x)^2-\sigma^2\big)\nu(x,t)dx}{\sqrt{\overline{(x-\bar x)^4}-\sigma^4}}dt=0,
-\end{equation}
-
-\begin{equation}
-d W_2d W_3=\frac{\int_\mathbb{R}(x-\bar x)\big((x-\bar x)^2-\sigma^2\big)p(x,t)dx}{\sqrt{\sigma^2\big(\overline{(x-\bar x)^4}-\sigma^4\big)}}dt
-=\frac{N\overline{(x-\bar x)^3}}{\sqrt{\sigma^2\big(\overline{(x-\bar x)^4}-\sigma^4\big)}}dt.
-\end{equation}
-
-In particular, when \(p\) is a Gaussian curve \(d W_2d W_3=0\).
-
-\hypertarget{relating-fitness-of-expressed-traits-to-fitness-of-breeding-values}{%
-\subsection{\texorpdfstring{Relating fitness of expressed traits to
-fitness of breeding values
-\label{fit2fit}}{Relating fitness of expressed traits to fitness of breeding values }}\label{relating-fitness-of-expressed-traits-to-fitness-of-breeding-values}}
-
-Following \S\ref{inheritance} of the main text, we have
-\(\sigma^2=G+\eta\) and
-
-\begin{equation}
-m^*(\rho,g)=\int_\mathbb{R}m(\nu,x)\psi(x,g)dx.
-\end{equation}
-
-Hence,
-
-\begin{multline}
-\overline{\frac{\partial m^*}{\partial\bar x}}=\int_{\mathbb{R}}\frac{\rho(g,t)}{N(t)}\frac{\partial}{\partial\bar x}\int_\mathbb{R}m(\nu,x)\psi(x,g)dxdg= \\
-\int_{\mathbb{R}}\int_\mathbb{R}\frac{\rho(g,t)}{N(t)}\psi(x,g)dg\frac{\partial}{\partial\bar x}m(\nu,x)dx=\int_\mathbb{R}p(x,t)\frac{\partial}{\partial\bar x}m(\nu,x)dx=\overline{\frac{\partial m}{\partial\bar x}}
-\end{multline} and
-
-\begin{multline}
-\overline{\frac{\partial m^*}{\partial G}}=\int_{\mathbb{R}}\frac{\rho(g,t)}{N(t)}\frac{\partial}{\partial G}\int_\mathbb{R}m(\nu,x)\psi(x,g)dxdg=\int_{\mathbb{R}}\int_\mathbb{R}\frac{\rho(g,t)}{N(t)}\psi(x,g)dg\frac{\partial m}{\partial G}dx= \\
-\int_\mathbb{R}p(x,t)\frac{\partial m}{\partial\sigma^2}\frac{\partial\sigma^2}{\partial G}dx=\overline{\frac{\partial m}{\partial\sigma^2}}.
-\end{multline}
-
-\hypertarget{derivation-of-diffuse-coevolution-model}{%
-\subsection{\texorpdfstring{Derivation of diffuse coevolution model
-\label{diffuse}}{Derivation of diffuse coevolution model }}\label{derivation-of-diffuse-coevolution-model}}
-
-In this section we provide a derivation of our model of diffuse
-coevolution driven by competition. Since most of the work in this
-derivation has already been completed in Supplementary Material
-\S\ref{SDE_DERIV}, we focus here on deriving the Malthusian fitness
-\(m\) as a function of trait value \(x\). We begin with discrete
-populations of individuals. In particular, we begin by assuming
-population size \(n_i\) is an integer for each species \(i=1,\dots,S\)
-before passing to the large population size limit.
-
-The reduction in fitness for an individual of species \(i\) caused by
-competition is captured multiplicatively by \(0<C_i\leq1\). Biologically
-this assumes all competitors affect individuals of a given species
-equally by consuming the same amount of resources. This is a mean-field
-interaction since any individual that consumes resources has an effect
-on the fitness of all other individuals competing for the same
-resources. Denote by \(x_{ij}\) the trait value of the \(j\)-th
-individual belonging to species \(i\). The set of trait values across
-all individuals in the community at time \(t\geq0\) is written
-\(X=\{x_{ij}\}\). We denote by \(\mathscr{B}_{ij}\) a function that maps
-\(X\) to the cumulative effect of all competitive interactions on the
-fitness of the \(j\)-th individual in species \(i\). Since individuals
-do not compete with themselves the net multiplicative effects on fitness
-of both interspecific and intraspecific competition on the \(j\)-th
-individual in species \(i\) can be summarized by
-
-\begin{equation}
-\mathscr B_{ij}(X)=C_i^{\sum_{l\neq j}\mathcal{O}_{ii}(x_{ij}-x_{il})+\sum_{k\neq i}\sum_{l=1}^{n_k}\mathcal{O}_{ik}(x_{ij}-x_{kl})},
-\end{equation}
-
-where \(\mathcal{O}_{ij}\), defined in the main text, measures the
-overlap in resource use between individuals of species \(i\) and \(j\)
-as a function of their niche-centers. Writing \(\mathscr W_{ij}(X)\) as
-the average number of offspring left by the \(j\)-th individual of
-species \(i\), we have
-
-\begin{equation}
-\mathscr W_{ij}(X)=\mathscr{A}_i(x_{ij})\mathscr{B}_{ij}(X),
-\end{equation}
-
-where \(\mathscr{A}_i(x)=\int_\mathbb{R}e_i(\zeta)u_i(\zeta,x)d\zeta\)
-accounts for abiotic selection and \(e_i\) has been defined in the main
-text.
-
-We now turn to a diffusion limit. Since we have more than one
-population, we take the diffusion limit for each population one at a
-time starting with population \(1\). We write
-\(\pmb n=(n_1,\dots,n_S)\). Following Méléard and Roelly (1993, 1992) we
-rescale generation time and individual mass to \(\frac{1}{n_1}\) and
-mean of the reproductive law to
-
-\begin{equation}
-\mathscr W_{1j}^{(\pmb n)}(X)=\mathscr{A}_1(x_{1j})^{1/n_1}\exp\left(\frac{\ln C_1}{n_1^2}\sum_{l\neq j}\mathcal{O}_{11}(x_{1j}-x_{1l})+\frac{\ln C_1}{n_1}\sum_{k\neq i}\frac{1}{n_k}\sum_{l=1}^{n_k}\mathcal{O}_{1k}(x_{1j}-x_{kl})\right).
-\end{equation}
-
-For large \(n_1\), we have the approximation
-
-\begin{equation}
-\mathscr W_{1j}^{(\pmb n)}(X)\approx\mathscr{A}_1(x_{1j})^{1/n_1}\left(1+\frac{\ln C_1}{n_1^2}\sum_{l\neq j}\mathcal{O}_{11}(x_{1j}-x_{1l})+\frac{\ln C_1}{n_1}\sum_{k\neq 1}\frac{1}{n_k}\sum_{l=1}^{n_k}\mathcal{O}_{1k}(x_{1j}-x_{kl})\right).
-\end{equation} Hence \begin{equation}
-\lim_{n_1\to\infty}n_1\left(\mathscr W_{1j}^{(\pmb n)}(X)-1\right)=\ln\mathscr{A}_1(x_{1j})+\left(\int_\mathbb{R}\mathcal{O}_{11}(x_{1j}-y)\nu_1(y,t)dy+\sum_{k\neq 1}\frac{1}{n_k}\sum_{l=1}^{n_k}\mathcal{O}_{1k}(x_{1j}-x_{kl})\right)\ln C_1.
-\end{equation}
-
-We write \(\lim_{\pmb n\to\pmb\infty}\) for the iterated limit
-\(\lim_{n_S\to\infty}\cdots\lim_{n_1\to\infty}\) and, assuming
-\(\nu_i(\cdot,t)\in C_1^+(\mathbb{R})\) for \(i=1,\dots,S\) and
-\(t\in[0,\infty)\), we set \(\pmb\nu=(\nu_1,\dots,\nu_S)\). Then, for
-any \(\pmb\nu\), the the diffusion limits for the remaining populations
-provides the Malthusian parameter for individuals in species \(i\) with
-trait value \(x_{1j}\) as
-
-\begin{equation}
-m_1(\pmb\nu,x_{1j}):=\lim_{\pmb n\to\pmb\infty}n_1\left(\mathscr W_{1j}^{(\pmb n)}(X)-1\right)=\ln\mathscr{A}_1(x)+\left(\sum_{k=1}^S\int_\mathbb{R}\mathcal{O}_{1k}(x_{1j}-y)\nu_k(y,t)dy\right)\ln C_1.
-\end{equation}
-
-We compute the average niche overlap of an individual in species \(i\)
-with nich location \(x\) across all individuals in species \(j\) as
-
-\begin{equation}
-\bar{\mathcal{O}}_{ij}(x,t)=\frac{\int_\mathbb{R}\mathcal{O}_{ij}(x-y)\nu_j(y,t)dy}{\int_\mathbb{R}\nu_j(y,t)dy}.
-\end{equation}
-
-We now assume the resource utilization curves \(u_i(\zeta)\) and
-phenotypic densities \(\nu_i(x,t)\) are Gaussian curves for
-\(i=1,\dots,S\). In this case \(\bar{\mathcal{O}}_{ij}(x,t)\) simplifies
-to
-
-\begin{equation}
-\bar{\mathcal{O}}_{ij}(x,t)=\frac{\int_\mathbb{R}\mathcal O_{ij}(x-y)\nu_j(y,t)dy}{\int_\mathbb{R}\nu_j(y,t)dy}
-=\frac{U_iU_j}{\sqrt{2\pi(w_i+w_j+{\sigma_j}^2(t))}}\exp\left(-\frac{(x-\bar x_j(t))^2}{2(w_i+w_j+{\sigma_j}^2(t))}\right).
-\end{equation}
-
-Setting
-
-\begin{subequations}
-\begin{align}
-\sigma_i^2(t) = & \ G_i(t)+\eta_i, \\
-R_i = & \ \ln \left(\frac{Q_iU_i}{\sqrt{1+A_iw_i}}\right), \\
-a_i = & \ \frac{A_i}{1+A_iw_i}, \\
-\tilde b_{ij}(t) = & \ \frac{1}{w_i+w_j+\sigma_j^2(t)}, \\
-c_i = & -\ln C_i,
-\end{align}
-\end{subequations}
-
-we get
-
-\begin{equation}\label{malth_fit}
-m_i(\pmb\nu,x)=R_i-\frac{a_i}{2}(x-\theta_i)^2-c_i\sum_{j=1}^SN_j(t)U_iU_j\sqrt{\frac{\tilde b_{ij}(t)}{2\pi}}e^{-\frac{\tilde b_{ij}(t)}{2}(x-\bar x_j(t))^2}.
-\end{equation}
-
-Hence, our fitness function is bounded above and decreases linearly with
-total abundance, as required by the assumptions of Champagnat, Ferrière
-and Méléard (2006). For the remainder of the derivation we suppress
-notation indicating dependency on \(\pmb\nu\), \(x\) and \(t\). From
-(\ref{malth_fit}) we calculate
-
-\begin{equation}
-\frac{\partial m_i}{\partial\bar x_i}=c_iN_iU_i^2\tilde b_{ii}(x-\bar x_i)\sqrt{\frac{\tilde b_{ii}}{2\pi}}e^{-\frac{\tilde b_{ii}}{2}(x-\bar x_i)^2}
-\end{equation}
-
-\begin{multline}
-\frac{\partial m_i}{\partial G_i}=\frac{c_iN_iU_i^2}{2}\left(\frac{(x-\bar x_i)^2-G_i-\eta_i-2w_i}{(G_i+\eta_i+2w_i)^2}\right)\sqrt{\frac{\tilde b_{ii}}{2\pi}}e^{-\frac{\tilde b_{ii}}{2}(x-\bar x_i)^2} \\
-=\frac{c_iN_iU_i^2{\tilde b_{ii}}^2}{2}\left((x-\bar x_i)^2-{\sigma_i}^2-2w_i\right)\sqrt{\frac{\tilde b_{ii}}{2\pi}}e^{-\frac{\tilde b_{ii}}{2}(x-\bar x_i)^2}.
-\end{multline}
-
-Note that
-
-\begin{multline}
-\sqrt\frac{\tilde b_{ii}}{2\pi}\exp\left(-\frac{\tilde b_{ii}}{2}(x-\bar x_i)^2\right)\sqrt\frac{1}{2\pi{\sigma_i}^2}\exp\left(-\frac{(x-\bar x_i)^2}{2{\sigma_i}^2}\right) \\
-=\sqrt\frac{1}{2\pi({\sigma_i}^2+1/\tilde b_{ii})}\sqrt\frac{{\sigma_i}^2+1/\tilde b_{ii}}{2\pi{\sigma_i}^2/\tilde b_{ii}}\exp\left(-\frac{{\sigma_i}^2+1/\tilde b_{ii}}{2{\sigma_i}^2/\tilde b_{ii}}(x-\bar x_i)^2\right) \\
-=\sqrt{\frac{1}{4\pi({\sigma_i}^2+w_i)}}\sqrt\frac{2({\sigma_i}^2+w_i)}{2\pi{\sigma_i}^2({\sigma_i}^2+2w_i)}\exp\left(-\frac{{\sigma_i}^2({\sigma_i}^2+2w_i)}{4({\sigma_i}^2+w_i)}(x-\bar x_i)^2\right).
-\end{multline}
-
-Hence,
-
-\begin{equation}
-\overline{\frac{\partial m_i}{\partial\bar x_i}}=0,
-\end{equation}
-
-\begin{multline}
-\overline{\frac{\partial m_i}{\partial G_i}}=\frac{c_iN_iU_i^2}{2({\sigma_i}^2+2w_i)^2}\left(\frac{({\sigma_i}^2+2w_i){\sigma_i}^2}{2(w_i+{\sigma_i}^2)}-{\sigma_i}^2-2w_i\right)\sqrt{\frac{b_{ii}}{2\pi}} \\
-=\frac{c_iN_iU_i^2}{2({\sigma_i}^2+2w_i)}\left(\frac{{\sigma_i}^2}{2({\sigma_i}^2+w_i)}-1\right)\sqrt{\frac{b_{ii}}{2\pi}}=-\frac{c_iN_iU_i^2b_{ii}}{2}\sqrt{\frac{b_{ii}}{2\pi}},
-\end{multline}
-
-where
-
-\begin{equation}
-b_{ij}=\frac{1}{w_i+w_j+{\sigma_i}^2+{\sigma_j}^2}.
-\end{equation}
-
-The average fitness for species \(i\) is
-
-\begin{equation}
-\bar m_i=R_i-\frac{a_i}{2}\Big((\bar x_i-\theta_i)^2+G_i+\eta_i\Big) - c_i\sum_{j=1}^SN_jU_iU_j\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}.
-\end{equation}
-
-Thus,
-
-\begin{equation}
-\frac{\partial\bar m_i}{\partial\bar x_i}=a_i(\theta_i-\bar x_i)-c_i\sum_jN_jU_iU_jb_{ij}(\bar x_j-\bar x_i)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2},
-\end{equation}
-
-\begin{equation}
-\frac{\partial\bar m_i}{\partial G_i}=-\frac{a_i}{2}+\frac{c_i}{2}\sum_{j=1}^SN_jU_iU_jb_{ij}\left(1-b_{ij}(\bar x_i-\bar x_j)^2\right)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}.
-\end{equation}
-
-In particular
-
-\begin{equation}
-\frac{\partial\bar m_i}{\partial G_i}-\overline{\frac{\partial m_i}{\partial G_i}}=-\frac{a_i}{2}+\frac{c_i}{2}\left(N_iU_i^2b_{ii}\sqrt\frac{b_{ii}}{2\pi}+\sum_{j=1}^SN_jU_iU_jb_{ij}\left(1-b_{ij}(\bar x_i-\bar x_j)^2\right)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\right).
-\end{equation}
-
-Applying equations (\ref{N}), (\ref{xbarfinal}) and (\ref{Gfinal}) of
-the main text recovers system (\ref{comm_dynamics}) of the main text.
-
-\hypertarget{the-relation-between-competition-coefficients-and-selection}{%
-\subsection{\texorpdfstring{The relation between competition
-coefficients and selection
-\label{coef_grad_moms}}{The relation between competition coefficients and selection }}\label{the-relation-between-competition-coefficients-and-selection}}
-
-\hypertarget{derivation-of-analytical-approximations}{%
-\subsubsection{Derivation of analytical
-approximations}\label{derivation-of-analytical-approximations}}
-
-Just as with most caclulations in this work, the derivations are
-straightforward applications of Gaussian products. That is, if
-
-\begin{equation}
-f_1(x)=\frac{1}{\sqrt{2\pi\sigma_1^2}}\exp\left(-\frac{(\mu_1-x)^2}{2\sigma_1^2}\right), \ f_2(x)=\frac{1}{\sqrt{2\pi\sigma_2^2}}\exp\left(-\frac{(\mu_2-x)^2}{2\sigma_2^2}\right),
-\end{equation}
-
-then
-
-\begin{equation}
-f_1(x)f_2(x)=\frac{1}{\sqrt{2\pi(\sigma_1^2+\sigma_2^2)}}\exp\left(-\frac{(\mu_1-\mu_2)^2}{2(\sigma_1^2+\sigma_2^2)}\right)\frac{1}{\sqrt{2\pi\tilde\sigma^2}}\exp\left(-\frac{(\tilde\mu-x)^2}{2\tilde\sigma^2}\right),
-\end{equation}
-
-where
-
-\begin{equation}
-\tilde\mu=\frac{\sigma_2^2\mu_1+\sigma_1^2\mu_2}{\sigma_1^2+\sigma_2^2}, \ \tilde\sigma^2=\frac{\sigma_1^2\sigma_2^2}{\sigma_1^2+\sigma_2^2}.
-\end{equation}
-
-\hypertarget{caclulating-mathrmcov_f_bar-xalphagamma}{%
-\paragraph{\texorpdfstring{Caclulating
-\(\mathrm{Cov}_{f_{\bar X}}(\alpha,\gamma)\)}{Caclulating \textbackslash{}mathrm\{Cov\}\_\{f\_\{\textbackslash{}bar X\}\}(\textbackslash{}alpha,\textbackslash{}gamma)}}\label{caclulating-mathrmcov_f_bar-xalphagamma}}
-
-Recalling
-
-\begin{equation}
-\alpha(\bar x_i,\bar x_j)=\frac{c}{\bar r}\sqrt\frac{b}{2\pi}\exp\left(-\frac{b}{2}\big(\bar x_i-\bar x_j\big)^2\right),
-\end{equation}
-
-\begin{equation}
-\gamma(\bar x_i,\bar x_j)=cNb\left(1-b\big(\bar x_i-\bar x_j\big)^2\right) \sqrt\frac{b}{2\pi}\exp\left(-\frac{b}{2}\big(\bar x_i-\bar x_j\big)^2\right),
-\end{equation}
-
-we have
-
-\begin{multline}
-\bar\alpha=
-\int_\mathbb{R}\int_\mathbb{R}\alpha(\bar x_i,\bar x_j)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
-=\frac{c}{\bar r}\int_\mathbb{R}\frac{1}{\sqrt{2\pi(b^{-1}+V_{\bar X})}}\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(b^{-1}+V_{\bar X})}\right)f_{\bar X}(\bar x_j)d\bar x_j
-=\frac{c/\bar r}{\sqrt{2\pi(b^{-1}+2V_{\bar X})}},
-\end{multline}
-
-\begin{multline}
-\bar\gamma=
-\int_\mathbb{R}\int_\mathbb{R}\gamma(\bar x_i,\bar x_j)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
-= cNb\int_\mathbb{R}\left\{1-\left[\left(\frac{\bar{\bar x}+bV_{\bar X}\bar x_j}{1+bV_{\bar X}}-\bar x_j\right)^2+\frac{V_{\bar X}}{1+bV_{\bar X}}\right]\right\}\frac{1}{\sqrt{2\pi(b^{-1}+V_{\bar X})}}\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(b^{-1}+V_{\bar X})}\right)f_{\bar X}(\bar x_j)d\bar x_j \\
-= cNb\int_\mathbb{R}\left\{1-\left[\left(\frac{\bar{\bar x}-\bar x_j}{1+bV_{\bar X}}\right)^2+\frac{V_{\bar X}}{1+bV_{\bar X}}\right]\right\}\frac{1}{\sqrt{2\pi(b^{-1}+V_{\bar X})}}\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(b^{-1}+V_{\bar X})}\right)f_{\bar X}(\bar x_j)d\bar x_j \\
-= cNb\left(1-\frac{(1+bV_{\bar X})V_{\bar X}}{1+2bV_{\bar X}}\frac{1}{(1+bV_{\bar X})^2}-\frac{V_{\bar X}}{1+bV_{\bar X}}\right)\frac{1}{\sqrt{2\pi(b^{-1}+2V_{\bar X})}} \\
-= cNb\left[1-\left(\frac{1}{1+2bV_{\bar X}}+1\right)\frac{V_{\bar X}}{1+bV_{\bar X}}\right]\frac{1}{\sqrt{2\pi(b^{-1}+2V_{\bar X})}} \\
-= cNb\left(1-\frac{2V_{\bar X}}{1+2bV_{\bar X}}\right)\sqrt\frac{b}{2\pi(1+2bV_{\bar X})},
-\end{multline}
-
-\begin{multline}
-\mathrm{Var}_{f_{\bar X}}(\alpha)=
-\int_\mathbb{R}\int_\mathbb{R}\big(\bar\alpha-\alpha(\bar x_i,\bar x_j)\big)^2f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
-=\frac{c^2}{\bar r^2}\left(\sqrt\frac{b}{4\pi}\int_\mathbb{R}\int_\mathbb{R}\sqrt\frac{b}{\pi}\exp\left(-b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j-\frac{1}{2\pi(b^{-1}+2V_{\bar X})}\right) \\
-=\frac{c^2}{\bar r^2}\left(\sqrt\frac{b}{4\pi}\int_\mathbb{R}\sqrt\frac{1}{2\pi(\frac{1}{2b}+V_{\bar X})}\exp\left(-b(\bar{\bar x}-\bar x_j)^2\right)f_{\bar X}(\bar x_j)d\bar x_j-\frac{1}{2\pi(b^{-1}+2V_{\bar X})}\right) \\
-=\frac{c^2}{\bar r^2}\left(\sqrt\frac{b}{4\pi}\sqrt\frac{1}{2\pi(\frac{1}{2b}+2V_{\bar X})}-\frac{1}{2\pi(b^{-1}+2V_{\bar X})}\right)=\frac{c^2b}{2\pi\bar r^2}\left(\frac{1}{\sqrt{1+4bV_{\bar X}}}-\frac{1}{1+2bV_{\bar X}}\right),
-\end{multline}
-
-\begin{multline}
-\mathrm{Cov}_{f_{\bar X}}(\alpha,\gamma)=
-\int_\mathbb{R}\int_\mathbb{R}\big(\bar\alpha-\alpha(\bar x_i,\bar x_j)\big)\big(\bar\gamma-\gamma(\bar x_i,\bar x_j)\big)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
-=\frac{c^2Nb}{2\bar r}\sqrt\frac{b}{\pi}\int_\mathbb{R}\int_\mathbb{R}\big(1-b(\bar x_i-\bar x_j)^2\big)\sqrt\frac{b}{\pi}\exp\left(-b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j-\bar\alpha\bar\gamma \\
-=\frac{c^2Nb}{2\bar r}\sqrt\frac{b}{\pi}\frac{1-2bV_{\bar X}}{\sqrt{2\pi((2b)^{-1}+2V_{\bar X})}}-\frac{c^2Nb}{\bar r}\frac{1-2bV_{\bar X}}{2\pi(b^{-1}+2V_{\bar X})} \\
-=\frac{c^2b^2N}{2\pi\bar r}(1-2bV_{\bar X})\left(\frac{1}{\sqrt{1+4bV_{\bar X}}}-\frac{1}{1+2bV_{\bar X}}\right).
-\end{multline}
-
-\hypertarget{caclulating-mathrmcov_f_bar-xalphabeta}{%
-\paragraph{\texorpdfstring{Caclulating
-\(\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)\)}{Caclulating \textbackslash{}mathrm\{Cov\}\_\{f\_\{\textbackslash{}bar X\}\}(\textbackslash{}alpha,\textbar{}\textbackslash{}beta\textbar{})}}\label{caclulating-mathrmcov_f_bar-xalphabeta}}
-
-To calculate moments of \(|\beta|\) we note that, as a random variable,
-\(|\beta|\) takes a folded normal distribution. Setting \(\Phi(x)\)
-equal to the cumulative density function of the standard normal
-distribution and using the properties of the folded normal distribution,
-we find
-
-\begin{equation}
-\overline{|\beta|}=\sqrt\frac{2\mathrm{Var}_{f_{\bar X}}(\beta)}{\pi}\exp\left(-\frac{\bar\beta^2}{2\mathrm{Var}_{f_{\bar X}}(\beta)}\right)-\bar\beta\left[1-2\Phi\left(\frac{\bar\beta}{\mathrm{Var}_{f_{\bar X}}(\beta)}\right)\right]
-\end{equation}
-
-\begin{equation}
-\mathrm{Var}_{f_{\bar X}}(|\beta|)=\bar\beta^2+\mathrm{Var}_{f_{\bar X}}(\beta)-\overline{|\beta|}^2.
-\end{equation}
-
-Recall that
-
-\begin{equation}
-\beta(\bar x_i,\bar x_j)=cNb\big(\bar x_i-\bar x_j\big)\sqrt\frac{b}{2\pi}\exp\left(-\frac{b}{2}\big(\bar x_i-\bar x_j\big)^2\right)
-\end{equation}
-
-and hence
-
-\begin{multline}
-\bar\beta=
-\int_\mathbb{R}\int_\mathbb{R}\beta(\bar x_i,\bar x_j)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
-= cNb\int_\mathbb{R}(\bar{\bar x}-\bar x_j)\frac{1}{\sqrt{2\pi(b^{-1}+V_{\bar X})}}\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(b^{-1}+V_{\bar X})}\right)f_{\bar X}(\bar x_j)d\bar x_j=0,
-\end{multline}
-
-\begin{multline}
-\mathrm{Var}_{f_{\bar X}}(\beta)=
-\int_\mathbb{R}\int_\mathbb{R}\big(\bar\beta-\beta(\bar x_i,\bar x_j)\big)^2f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
-=\int_\mathbb{R}\int_\mathbb{R}c^2N^2b^2(\bar x_i-\bar x_j)^2\frac{b}{2\pi}\exp\left(-b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
-=\sqrt{\frac{b}{4\pi}}c^2N^2b^2\int_\mathbb{R}\left[\left(\frac{\bar{\bar x}+2bV_{\bar X}\bar x_j}{1+2bV_{\bar X}}-\bar x_j\right)^2+\frac{V_{\bar X}}{1+2bV_{\bar X}}\right]\frac{\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(\frac{1}{2b}+V_{\bar X})}\right)}{\sqrt{2\pi(\frac{1}{2b}+V_{\bar X})}}f_{\bar X}(\bar x_j)d\bar x_j \\
-=\sqrt{\frac{b}{4\pi}}c^2N^2b^2\int_\mathbb{R}\left[\frac{(\bar{\bar x}-\bar x_j)^2}{(1+2bV_{\bar X})^2}+\frac{V_{\bar X}}{1+2bV_{\bar X}}\right]\frac{\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(\frac{1}{2b}+V_{\bar X})}\right)}{\sqrt{2\pi(\frac{1}{2b}+V_{\bar X})}}f_{\bar X}(\bar x_j)d\bar x_j \\
-=\sqrt{\frac{b}{4\pi}}c^2N^2b^2\left[\frac{(1+2bV_{\bar X})V_{\bar X}}{1+4bV_{\bar X}}\frac{1}{(1+2bV_{\bar X})^2}+\frac{V_{\bar X}}{1+2bV_{\bar X}}\right]\frac{1}{\sqrt{2\pi(\frac{1}{2b}+2V_{\bar X})}} \\
-=\frac{b}{\pi}\frac{c^2N^2b^2}{\sqrt{1+4bV_{\bar X}}}\frac{V_{\bar X}}{1+2bV_{\bar X}}\left(\frac{1}{1+4bV_{\bar X}}+1\right)=\frac{2c^2N^2b^3V_{\bar X}}{\pi(1+4bV_{\bar X})^{3/2}}.
-\end{multline}
-
-Thus, using properties of the folded normal distribution, we find
-
-\begin{equation}
-\overline{|\beta|}=\sqrt{\frac{2}{\pi}}\frac{cNb^{3/2}}{(1+4bV_{\bar X})^{3/4}}\sqrt\frac{2V_{\bar X}}{\pi}=\frac{2}{\pi}\frac{cNb^{3/2}}{(1+4bV_{\bar X})^{3/4}}\sqrt{V_{\bar X}},
-\end{equation}
-
-\begin{equation}
-\mathrm{Var}_{f_{\bar X}}(|\beta|)=\frac{c^2N^2b^3}{(1+4bV_{\bar X})^{3/2}}\frac{2V_{\bar X}}{\pi}\left(1-\frac{2}{\pi}\right).
-\end{equation}
-
-We also calculate
-
-\begin{multline}
-\mathrm{Cov}_{f_{\bar X}}(\alpha,\beta)=
-\int_\mathbb{R}\int_\mathbb{R}\big(\bar\alpha-\alpha(\bar x_i,\bar x_j)\big)\big(\bar\beta-\beta(\bar x_i,\bar x_j)\big)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
-=\frac{c^2Nb}{2\bar r}\sqrt\frac{b}{\pi}\int_\mathbb{R}\int_\mathbb{R}(\bar x_i-\bar x_j)\sqrt\frac{b}{\pi}\exp\left(-b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j=0.
-\end{multline}
-
-In attempt to calculate \(\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)\) we
-find
-
-\begin{multline}
-\mathrm{Cov}f_{\bar X}(\alpha,|\beta|)=\int_\mathbb{R}\int_\mathbb{R}\alpha(\bar x_i,\bar x_j)|\beta(\bar x_i,\bar x_j)|f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j-\bar\alpha\overline{|\beta|} \\
-=\int_\mathbb{R}\int_\mathbb{R}\frac{c}{\bar r}\sqrt\frac{b}{2\pi}\exp\left(-\frac{b}{2}(\bar x_i-\bar x_j)^2\right)cNb|\bar x_i-\bar x_j|\sqrt\frac{b}{2\pi}\exp\left(-\frac{b}{2}(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j-\bar\alpha\overline{|\beta|} \\
-=\frac{c^2Nb}{\bar r}\sqrt\frac{b}{4\pi}\int_\mathbb{R}\int_\mathbb{R}|\bar x_i-\bar x_j|\sqrt\frac{b}{\pi}\exp\left(-b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j-\bar\alpha\overline{|\beta|}.
-\end{multline}
-
-Just as we used the folded normal to find \(\overline{|\beta|}\) and
-\(\mathrm{Var}_{f_{\bar X}}(|\beta|)\), we can calculate
-\(\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)\) by considering
-
-\begin{equation}
-\int_\mathbb{R}\int_\mathbb{R}(\bar x_i-\bar x_j)\sqrt\frac{b}{\pi}\exp\left(-b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j=0
-\end{equation}
-
-and
-
-\begin{multline}
-\int_\mathbb{R}\int_\mathbb{R}(\bar x_i-\bar x_j)^2\frac{b}{\pi}\exp\left(-2b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
-=\sqrt\frac{2b}{\pi}\int_\mathbb{R}\int_\mathbb{R}(\bar x_i-\bar x_j)^2\frac{1}{\sqrt{2\pi\frac{1}{4b}}}\exp\left(-\frac{(\bar x_i-\bar x_j)^2}{2\frac{1}{4b}}\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
-=\sqrt\frac{2b}{\pi}\int_\mathbb{R}\left[\left(\frac{\bar{\bar x}+4bV_{\bar X}\bar x_j}{1+4bV_{\bar X}}-\bar x_j\right)^2+\frac{V_{\bar X}}{1+4bV_{\bar X}}\right]\frac{1}{\sqrt{2\pi(\frac{1}{4b}+V_{\bar X})}}\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(\frac{1}{4b}+V_{\bar X})}\right)f_{\bar X}(\bar x_j)d\bar x_j \\
-=\sqrt\frac{2b}{\pi}\int_\mathbb{R}\left[\left(\frac{\bar{\bar x}-\bar x_j}{1+4bV_{\bar X}}\right)^2+\frac{V_{\bar X}}{1+4bV_{\bar X}}\right]\frac{1}{\sqrt{2\pi(\frac{1}{4b}+V_{\bar X})}}\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(\frac{1}{4b}+V_{\bar X})}\right)f_{\bar X}(\bar x_j)d\bar x_j \\
-=\sqrt\frac{2b}{\pi}\left[\frac{(1+4bV_{\bar X})V_{\bar X}}{1+8bV_{\bar X}}\frac{1}{(1+4bV_{\bar X})^2}+\frac{V_{\bar X}}{1+4bV_{\bar X}}\right]\frac{1}{\sqrt{2\pi(\frac{1}{4b}+2V_{\bar X})}} \\
-=\sqrt\frac{2b}{\pi}\frac{2V_{\bar X}}{1+8bV_{\bar X}}\sqrt{\frac{4b}{2\pi(1+8bV_{\bar X})}}=\frac{b}{\pi}\frac{4V_{\bar X}}{(1+8bV_{\bar X})^{3/2}}.
-\end{multline}
-
-Hence
-
-\begin{equation}
-\int_\mathbb{R}\int_\mathbb{R}|\bar x_i-\bar x_j|\sqrt\frac{b}{\pi}\exp\left(-b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j=\sqrt\frac{2}{\pi}\sqrt{\frac{b}{\pi}\frac{4V_{\bar X}}{(1+8bV_{\bar X})^{3/2}}}=\frac{2}{\pi}\frac{\sqrt{2bV_{\bar X}}}{(1+8bV_{\bar X})^{3/4}}
-\end{equation}
-
-and
-
-\begin{multline}
-\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)=\frac{c^2Nb}{\bar r}\sqrt\frac{b}{4\pi}\frac{2}{\pi}\frac{\sqrt{2bV_{\bar X}}}{(1+8bV_{\bar X})^{3/4}}-\bar\alpha\overline{|\beta|} \\
-=\frac{2c^2Nb^2}{\pi\bar r(1+8bV_{\bar X})^{3/4}}\sqrt\frac{V_{\bar X}}{2\pi}-\frac{c}{\bar r}\sqrt\frac{b}{2\pi(1+2bV_{\bar X})}\frac{2}{\pi}\frac{cNb^{3/2}}{(1+4bV_{\bar X})^{3/4}}\sqrt{V_{\bar X}} \\
-=\frac{2c^2Nb^2}{\pi\bar r(1+8bV_{\bar X})^{3/4}}\sqrt\frac{V_{\bar X}}{2\pi}-\frac{2c^2Nb^2}{\pi\bar r(1+4bV_{\bar X})^{3/4}}\sqrt\frac{V_{\bar X}}{2\pi(1+2bV_{\bar X})} \\
-=\frac{2c^2Nb^2}{\pi\bar r}\sqrt\frac{V_{\bar X}}{2\pi}\left(\frac{1}{(1+8bV_{\bar X})^{3/4}}-\frac{1}{(1+4bV_{\bar X})^{3/4}(1+2bV_{\bar X})^{1/2}}\right).
-\end{multline}
-
-\hypertarget{starting-the-caclulation-of-mathrmcov_f_bar-xalphamathfrakc}{%
-\paragraph{\texorpdfstring{Starting the caclulation of
-\(\mathrm{Cov}_{f_{\bar X}}(\alpha,\mathfrak{C})\)
-\label{caC}}{Starting the caclulation of \textbackslash{}mathrm\{Cov\}\_\{f\_\{\textbackslash{}bar X\}\}(\textbackslash{}alpha,\textbackslash{}mathfrak\{C\}) }}\label{starting-the-caclulation-of-mathrmcov_f_bar-xalphamathfrakc}}
-
-We have
-
-\begin{equation}
-\mathfrak{C}(\bar x_i,\bar x_j)=c^2N^2b^2\Big(|\bar x_i-\bar x_j|+\big|1-b(\bar x_i-\bar x_j)^2\big|\Big)^2\exp\left(-\frac{b}{2}(\bar x_i-\bar x_j)^2\right).
-\end{equation}
-
-Note that the random variable \(\delta=\bar x_i-\bar x_j\) is a mean
-zero Gaussian random variable with variance \(2V_{\bar X}\). We write
-the probability density function of \(\delta\) as
-\(f_{\Delta}(\delta)\). Substituting in \(\delta\), we can write
-
-\begin{multline}
-\mathfrak{C}(\delta,0)=c^2N^2b^2\Big(|\delta|+\big|1-b\delta^2\big|\Big)^2\exp\left(-\frac{b}{2}\delta^2\right) \\
-=c^2N^2b^2\Big(\delta^2+2\big||\delta|-b|\delta|^3\big|+\big(1-b\delta^2\big)^2\Big)\exp\left(-\frac{b}{2}\delta^2\right).
-\end{multline}
-
-From this expression, we see properties of the folded normal
-distribution can be used to calculate several components of the intregal
-\(\mathrm{Cov}_{f_{\bar X}}(\alpha,\mathfrak{C})\), but a major
-technical challenge lies in calculating
-
-\begin{equation}
-\int_{\mathbb{R}}\big||\delta|-b|\delta|^3\big|\exp\left(-\frac{b}{2}\delta^2\right)f_{\Delta}(\delta)d\delta.
-\end{equation}
-
-Instead of overcoming this challenge to find an analytical form of
-\(\mathrm{Cov}_{f_{\bar X}}(\alpha,\mathfrak{C})\) we turn to a
-numerical approach outlined in the following section.
-
-\hypertarget{numerical-estimates-for-heterogeneous-n-and-g}{%
-\subsubsection{\texorpdfstring{Numerical estimates for heterogeneous
-\(N\) and
-\(G\)}{Numerical estimates for heterogeneous N and G}}\label{numerical-estimates-for-heterogeneous-n-and-g}}
-
-To verify our analytical estimates for the covariance of selection
-gradients and competition coefficients, equations (\ref{cov_eqns}) of
-the main text, and for approximating the correlation of competition
-coefficients and a metric of pairwise coevolution, Figure \ref{cov_fig}
-of the main text, we developed a numerical approach by repeatedly
-simulating our system (\ref{comm_dynamics}) of the main text across a
-set of randomly drawn model parameters. In particular, we ran each
-numerical integration with \(S=100\) species for \(T_1=1000.0\) units of
-time. We then continued to integrate for \(T_2=1000.0\) units of time
-and calculated the covariances between competition coefficients and the
-quantities \(|\beta|\), \(\gamma\) and \(\mathfrak{C}\) for each of the
-last \(T_2\) time steps. We assume the temporal average of these
-covariances across the last \(T_2\) units of time approximates the
-expectation at equilibrium. We repeated this approach for randomly drawn
-\(a\) and \(c\) until our sample size reached \(1000\). We drew values
-for \(a\) and \(c\) from condition log-normal distributions. For \(a\),
-we set the log-normal parameters equal to \(\mu=-10\) and \(\sigma=6\)
-and conditioned on \(a<0.01\) to prevent extinction. For \(c\) we set
-the log-normal parameters equal to \(\mu=-4\) and \(\sigma=4\) and
-conditioned on \(c<1\times 10^{-4}\) to prevent very large interspecific
-phenotypic variation and extinction. When extinction occurs, we aborted
-the simulation and restarted with newly drawn parameters. Results for
-\(\mathrm{Corr}_{f_{\bar X}}(\alpha,|\beta|)\) and
-\(\mathrm{Corr}_{f_{\bar X}}(\alpha,\gamma)\) are displayed in SM Figure
-\ref{num_corr}.
-
-\begin{figure}
-
-{\centering \includegraphics{/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/corrs}
-
-}
-
-\caption{\label{num_corr} Numerical estimate for the correlations of selection gradients and competition coefficients.}\label{fig:unnamed-chunk-13}
-\end{figure}
-
-\newpage
-
 \hypertarget{references}{%
@@
 Foundations and Applications}. Wiley.
-
-\leavevmode\hypertarget{ref-stanleyfarlow1993}{}%
-Farlow, Stanley J. 1993. \emph{Partial Differential Equations for
-Scientists and Engineers}. Dover.
 

### 2020-04-18 .A White Noise Approach To Evolutionary Ecology/SM.Rmd
split main txt & SM; improved formatting

@@
 ```{r global_options, include=FALSE}
-#knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
+knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
 ```
@@
 
-```{r, echo=F, fig.align='center', fig.cap='\\label{num_corr} Numerical estimate for the correlations of selection gradients and competition coefficients.'}
+```{r, echo=F, fig.align='center', out.width='100%', fig.cap='\\label{num_corr} Numerical estimate for the correlations of selection gradients and competition coefficients.'}
 knitr::include_graphics('/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/corrs.png')
@@
 
-\newpage
-
 # References

### 2020-04-18 .A White Noise Approach To Evolutionary Ecology/SM.tex
split main txt & SM; improved formatting

@@
-\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
-\PassOptionsToPackage{hyphens}{url}
-%
 \documentclass[]{article}
@@
   \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provides euro and other symbols
 \else % if luatex or xelatex
-  \usepackage{unicode-math}
+  \ifxetex
+    \usepackage{mathspec}
+  \else
+    \usepackage{fontspec}
+  \fi
   \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
@@
 \IfFileExists{microtype.sty}{%
-\usepackage[]{microtype}
+\usepackage{microtype}
 \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
 }{}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
+\usepackage[margin=1in]{geometry}
 \usepackage{hyperref}
-\hypersetup{
+\hypersetup{unicode=true,
             pdftitle={Supplementary Material (SM)},
@@
 \urlstyle{same}  % don't use monospace font for urls
-\usepackage[margin=1in]{geometry}
 \usepackage{graphicx,grffile}
@@
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
 \setlength{\emergencystretch}{3em}  % prevent overfull lines
@@
 
-% set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
+%%% Use protect on footnotes to avoid problems with footnotes in titles
+\let\rmarkdownfootnote\footnote%
+\def\footnote{\protect\rmarkdownfootnote}
+
+%%% Change title format to be more compact
+\usepackage{titling}
+
+% Create subtitle command for use in maketitle
+\providecommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
+    }
+}
+
+\setlength{\droptitle}{-2em}
+
+  \title{Supplementary Material (SM)}
+    \pretitle{\vspace{\droptitle}\centering\huge}
+  \posttitle{\par}
+    \author{}
+    \preauthor{}\postauthor{}
+    \date{}
+    \predate{}\postdate{}
+
+\usepackage{float}
+\usepackage{flafter}
 
@@
 \externaldocument{draft}
-\usepackage{float}
-\usepackage{flafter}
-
-\title{Supplementary Material (SM)}
-\author{}
-\date{\vspace{-2.5em}}
 
@@
 
-\begin{figure}
-
-{\centering \includegraphics{/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/corrs}
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\linewidth,]{/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/corrs}
 
@@
 \end{figure}
-
-\newpage
 

### 2020-04-18 .A White Noise Approach To Evolutionary Ecology/draft.tex
split main txt & SM; improved formatting

@@
 \IfFileExists{microtype.sty}{%
-\usepackage[]{microtype}
+\usepackage{microtype}
 \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
 }{}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
+\usepackage[margin=1in]{geometry}
 \usepackage{hyperref}
-\hypersetup{
+\hypersetup{unicode=true,
             pdftitle={A White Noise Approach to Evolutionary Ecology},
@@
 
-% set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
+%%% Use protect on footnotes to avoid problems with footnotes in titles
+\let\rmarkdownfootnote\footnote%
+\def\footnote{\protect\rmarkdownfootnote}
+
+%%% Change title format to be more compact
+\usepackage{titling}
+
+% Create subtitle command for use in maketitle
+\providecommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
+    }
+}
+
+\setlength{\droptitle}{-2em}
+
+  \title{A White Noise Approach to Evolutionary Ecology}
+    \pretitle{\vspace{\droptitle}\centering\huge}
+  \posttitle{\par}
+    \author{Bob Week, Steve Krone, Scott L. Nuismer, Luke J. Harmon}
+    \preauthor{\centering\large\emph}
+  \postauthor{\par}
+      \predate{\centering\large\emph}
+  \postdate{\par}
+    \date{3/1/2020}
+
+\usepackage{float}
+\usepackage{flafter}
 
@@
 \externaldocument{SM}
-\usepackage{float}
-\usepackage{flafter}
-
-\title{A White Noise Approach to Evolutionary Ecology}
-\author{Bob Week, Steve Krone, Scott L. Nuismer, Luke J. Harmon}
-\date{3/1/2020}
 
@@
 
-{\centering \includegraphics{wn}
+{\centering \includegraphics[width=20.83in]{wn}
 
@@
 
-{\centering \includegraphics{rescaled_plots}
+{\centering \includegraphics[width=29.17in]{rescaled_plots}
 
@@
 
-{\centering \includegraphics{on_pl}
+{\centering \includegraphics[width=14.58in]{on_pl}
 
@@
 
-{\centering \includegraphics{/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/cov}
+{\centering \includegraphics[width=27.08in]{/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/cov}
 
@@
 
-{\centering \includegraphics{/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/Cαℭp}
+{\centering \includegraphics[width=16.67in]{/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/Cαℭp}
 

### 2020-04-28 Current Version/Untitled.Rmd
reorganize

--- 2020-04-18 .A White Noise Approach To Evolutionary Ecology/draft_04_18_20.Rmd
@@
 ---
-title: "A White Noise Approach to Evolutionary Ecology"
-author: "Bob Week, Steve Krone, Scott L. Nuismer, Luke J. Harmon"
-date: "3/1/2020"
-abstract: "We derive the dynamics of the distribution of a quantitative character and the abundance of a biological population from a stochastic partial differential equation driven by space-time white noise.
+title: A White Noise Approach to Evolutionary Ecology
+author:
+  - name: Bob Week
+    email: bobweek@gmail.com
+    affiliation: Bio
+    footnote: 1
+  - name: Steve Krone
+    affiliation: Stats
+  - name: Scott L. Nuismer
+    affiliation: Bio
+  - name: Luke J. Harmon
+    affiliation: Bio
+address:
+  - code: Bio
+    address: University of Idaho, Department of Biological Sciences, 875 Perimeter Drive MS 3051, Moscow, Idaho, 83844
+  - code: Stats
+    address: University of Idaho, Department of Statistical Science, 875 Perimeter Drive MS 1104, Moscow, Idaho, 83844
+footnote:
+  - code: 1
+    text: "Corresponding Author"
+abstract: |
+  We derive the dynamics of the distribution of a quantitative character and the abundance of a biological population from a stochastic partial differential equation driven by space-time white noise.
 In the process we develop a useful set of heuristics to operationalize the powerful, but abstract theory of white noise and measure-valued stochastic processes.
@@
 The inverse relationship holds for quadratic selection gradients except with threshold diversity ratio equal to unity instead of $1.25$.
-Hence, this approach contributes to the development of a synthetic theory of evolutionary ecology by formalizing first principle derivations of stochastic models that underlie rigorous investigations of the relationship between feedbacks of biological processes and the patterns of diversity they produce."
+Hence, this approach contributes to the development of a synthetic theory of evolutionary ecology by formalizing first principle derivations of stochastic models that underlie rigorous investigations of the relationship between feedbacks of biological processes and the patterns of diversity they produce.
+journal: "bioarxiv"
+date: "`r Sys.Date()`"
 bibliography: ref.bib
+linenumbers: true
+numbersections: true
+csl: elsevier-harvard.csl
 output: 
-  pdf_document: 
+  rticles::elsevier_article:
     extra_dependencies:
@@
   - \usepackage{hyperref}
-  - \usepackage[left]{lineno}
   - \usepackage{amsmath}
-  - \linenumbers
-editor_options: 
-  chunk_output_type: console
+  - \externaldocument{SM} # externel document to cross-reference
+  - \modulolinenumbers[2]
 ---
@@
 
-# Introduction
+Introduction
+==========================
 
@@
 
-# The framework
+The framework
+============
 
@@
 ## Deterministic dynamics \label{deterministic}
-
 Our review begins by considering the dynamics of an asexually reproducing population in a homogeneous environment.
@@
 Specifically, denoting $\nu_i$ the abundance of individuals with allele $i$ and $m_i$ the growth rate of allele $i$ [called the Malthusian parameter in @9781932846126], we have
+
 \begin{equation}\label{first}
@@
 \end{equation}
+
 Starting from this model, we get the total abundance of the population as $N=\sum_{i=1}^K\nu_i$, the frequency of allele $i$ as $p_i=\nu_i/N$ and the mean fitness of the population as $\bar m=\sum_{i=1}^Kp_im_i$.
 Hence, we can employ some elementary calculus to derive the dynamics of abundance $dN/dt$ and the dynamics of allele frequencies $dp_1/dt,\dots,dp_K/dt$ as
+
 \begin{equation}
@@
 \end{equation}
+
 Two important observations of these equations include (i) mean fitness $\bar m$ determines the abundance dynamics of the entire population and (ii) allele $i$ will increase (decrease) in frequency if $m_i>\bar m$ ($<\bar m$).
@@
 That is, we assume $\nu(x,0)$ is continuous in $x$, satisfies $\nu(x,0)\geq0$ for all $x\in\mathbb{R}$ and
+
 \begin{equation}
@@
 \end{equation}
+
 where $\bar x(t)$ and $\sigma^2(t)$ are respectively the mean trait and phenotypic variance at time $t\geq0$.
@@
 With these technicalities addressed we model the demographic dynamics of a population and the dynamics of a quantitative character simultaneously by the PDE
+
 \begin{equation}\label{eq1}
@@
 \end{equation}
+
 with the initial condition $\nu(x,0)$ described above.
@@
 To ensure existence and uniqueness of solutions to (\ref{eq1}), we further assume that for each $M>0$ there exists a constant $L_M>0$ depending on $M$ such that $\int_\mathbb{R}\nu_1(x)dx,\int_\mathbb{R}\nu_2(x)dx\leq M$ implies
+
 \begin{equation}\label{local_lipschitz}
@@
 \end{equation}
+
 In SM \S\ref{finite} we combine the above assumptions on growth rate, mutation and initial condition to prove that solutions to (\ref{eq1}) satisfy $N(t),|\bar x(t)|,\sigma^2(t)<+\infty$ for all $t\geq0$.
@@
 As an example, logistic growth combined with quadratic stabilizing selection can be captured using the growth rate
+
 \begin{equation}\label{special_m}
@@
 \end{equation}
+
 where $r\in\mathbb{R}$ is the intrinsic rate of growth in the absence of abiotic stabilizing selection, $a>0$ is the strength of abiotic stabilizing selection around the phenotypic optimum $\theta\in\mathbb{R}$, $c>0$ captures the sensitivity of fitness to intraspecific competition and $\kappa(x-y)=1$ so that competitive interactions cause the same reduction in fitness regardless of trait value.
@@
 \end{multline}
+
 Using our assumptions on mutation and rate of growth, we show in SM \S\ref{finite} $\nu(x,t)$ is twice differentiable with respect to $x$ and $\int_{-\infty}^{+\infty}\nu(x,t)dx<\infty$ for all $t\geq0$.
@@
 \end{multline}
+
 This result closely resembles Kimura's continuum-of-alleles model [@Kimura1965; @Brger1986].
@@
 The covariance of fitness and phenotype across the population is defined as
+
 \begin{equation}
@@
 \end{equation}
+
 Hence, the dynamics of the mean trait $\bar x(t)$ can be derived as
@@
 Following the approach taken to calculate the evolution of $\bar x$, we find the response of phenotypic variation to this model of selection and mutation is
+
 \begin{equation}\label{vardyn_det}
@@
 \end{equation}
+
 For the sake of space we relegate the derivation of $d\sigma^2/dt$ to SM \S\ref{var_deriv}.
@@
 
-### Introduction to white noise
-
 One can think of white noise as the static seen on old television sets or infinitely detailed random dust spread across both time and space.
@@
 
-```{r, echo=F, fig.align='center', fig.cap='\\label{wn}Approximations of sample paths of temporal white noise (left) and space-time white noise (right) with brightness scaled to value.'}
+```{r, echo=F, fig.align='center', out.width='80%', fig.cap='\\label{wn}Approximations of sample paths of temporal white noise (left) and space-time white noise (right) with brightness scaled to value.'}
 knitr::include_graphics('wn.png')
@@
 Specifically, for any $f,g\in\mathscr{N}_2$, we define $\mathbf W(f)$ and $\mathbf W(g)$ to be Gaussian processes satisfying, for any $t,t_1,t_2\geq0$,
+
 \begin{equation}\label{exp_WN}
@@
 \end{equation}
+
 where $t_1\wedge t_2=\min(t_1,t_2)$ and $\mathbb{C}$ denotes covariance with respect to the underlying probability space.
@@
 With the generalized process $\mathbf W$ defined, we define the space-time white noise $\dot W(x,t)$ implicitly via the stochastic integral
+
 \begin{equation}\label{informal}
@@
 \end{equation}
+
 We place quotations in the above expression to emphasize its informal nature and that it should not be confused with classical Riemann integration.
@@
 For example, integrating white noise over a region $D\times[0,t]$, with $t>0$ and $D$ a bounded subset of $\mathbb{R}$, is equivalent to evaluating $\mathbf W_t(I_D)$ for the deterministic process
+
 \begin{equation}
@@
 \end{equation}
+
 Since
+
 \begin{equation}
@@
 \end{equation}
+
 where $|D|$ denotes the length of $D$, we have $I_D\in\mathscr{N}_2$.
 Thus, using equations (\ref{exp_WN}) and (\ref{cov_WN}) and adopting the informal notation introduced in equation (\ref{informal}), we can write the following
+
 \begin{equation}
@@
 \end{equation}
+
 Using this informal notation, equations (\ref{exp_WN}) and (\ref{cov_WN}) can be rewritten as
+
 \begin{equation}\label{exp_WN_xi}
@@
 \end{equation}
-\begin{equation}\label{cov_WN_xi}
-\mathbb{C}\left(\int_0^{t_1}\int_\mathbb{R}f(x,s)\dot W(x,s)dxds,\int_0^{t_2}\int_\mathbb{R}g(x,s)\dot W(x,s)dxds\right)
+\begin{multline}\label{cov_WN_xi}
+\mathbb{C}\left(\int_0^{t_1}\int_\mathbb{R}f(x,s)\dot W(x,s)dxds,\int_0^{t_2}\int_\mathbb{R}g(x,s)\dot W(x,s)dxds\right) \\
 =\int_0^{t_1\wedge t_2}\int_\mathbb{R}f(x,s)g(x,s)dxds.
-\end{equation}
+\end{multline}
+
 To relate these formula to the common notation used for SDE, we write
+
 \begin{equation}
@@
 \end{equation}
+
 so that
+
 \begin{equation}
@@
 \end{equation}
+
 This implies
+
 \begin{equation}
@@
 \end{equation}
+
 and in particular, as a function of $t$, $\int_0^td\hat{\mathbf W}_s(f)$ is a standard Brownian motion for any $f\in\mathscr{N}_2$.
@@
 Thus, equation (\ref{cov_WN_xi}) effectively extends It&ocirc;'s multiplication table to:
+
 \begin{table}[H]
@@
 \end{table}
+
 The extension of It&ocirc;'s multiplication table and properties of white noise outlined in this subsection provide a useful set of tools for working with SPDE.
@@
 
-__Branching Brownian motion__
-
+_Branching Brownian motion (BBM):_
 A BBM tracks individuals navigating $d$-dimensional Euclidean space that reproduce and senesce between exponentially distributed intervals.
@@
 Setting $\lambda=1$ and $m=\ln\mathscr{W}$ this density process satisfies the SPDE
+
 \begin{equation}\label{neutral_SPDE}
@@
 \end{equation}
+
 Since $\nu(x,t)$ is not generally differentiable in $x$ or $t$, the derivatives in expression (\ref{neutral_SPDE}) are taken in the _weak_ sense [for a definition see @lawrenceevans2010].
@@
 \end{multline}
+
 This expression is referred to as the _mild_ solution of (\ref{neutral_SPDE}).
@@
 
-__Interacting superprocesses__
-
+_Interacting superprocesses:_
 In the above subsection we reviewed the diffusion limit of an especially tractable measure-valued branching process.
@@
 
-```{r, message=FALSE, echo=F, fig.align='center', fig.cap='\\label{rescaled}Rescaled sample paths of a branching random walk under stabilizing selection and logistic growth.
+```{r, message=FALSE, echo=F, fig.align='center', out.width='80%', fig.cap='\\label{rescaled}Rescaled sample paths of a branching random walk under stabilizing selection and logistic growth.
 The top plot displays a sample path without scaling $(n=1)$, the middle plot shows a sample path rescaled by $n=5$ and the bottom plot shows a sample path rescaled by $n=20$.'}
@@
 Assuming the interactions manifest only in the reproductive law and that spatial movement follows Brownian motion with diffusion parameter $\sqrt\mu$ independent of both $X_t$ and $x$ and the growth rate is bounded both above and below, Li's [-@zeng1998absolute] result implies the interacting superprocess on one dimensional trait space has a density $\nu(x,t)$ which is non-negative, integrable, continuous in time and space and satisfies the SPDE
+
 \begin{equation}\label{SPDE}
@@
 \end{equation}
+
 As mentioned above, we refer to SPDE (\ref{SPDE}) as the Stochastic Asexual Gaussian allelic model with Abundance dynamics (abbreviated SAGA).
@@
 \end{multline}
+
 where
+
 \begin{equation}
@@
 \end{equation}
+
 and
+
 \begin{equation}
@@
 \end{equation}
+
 Setting $W_1(t)=\hat{\mathbf{W}}_t(\sqrt{\nu(x,t)})$, we can use traditional stochastic differential notation to write
+
 \begin{equation}
@@
 \end{equation}
+
 To find the associated SDE for $\bar x(t)$ and $\sigma^2(t)$, we want to repeat the same approach for $f(x)=x,x^2$ and apply It&ocirc;'s lemma.
@@
 Setting $\tilde x(t)=\int_\mathbb{R}x\nu(x,t)dx$, we have
+
 \begin{equation}
@@
 \end{equation}
+
 Similarly, setting $\tilde{\tilde\sigma}^2(t)=\int_\mathbb{R}x^2\nu(x,t)dx$, we have
+
 \begin{equation}
@@
 \end{equation}
+
 Since $\bar x(t)=\tilde x(t)/N(t)$ and $\sigma^2(t)=\tilde{\tilde\sigma}^2(t)/N(t)-\bar x^2(t)$, we can use It&ocirc;'s lemma to derive SDE for $\bar x(t)$ and $\sigma^2(t)$, which we perform in SM \S\ref{SDE_DERIV}.
@@
 In SM \S\ref{SDE_DERIV} we show SDE for $N(t)$, $\bar x(t)$ and $\sigma^2(t)$ can be expressed as
+
 \begin{subequations}\label{gen_eqns}
@@
 \end{subequations}
+
 where $W_1$, $W_2$ and $W_3$ are standard Brownian motions.
@@
 Allowing for these restrictions, we assume
+
 \begin{equation}
@@
 \end{equation}
+
 Under this assumption we find in SM \S\ref{cov2deriv} the results (suppressing the dependency on $t$)
+
 \begin{equation}\label{covxm}
@@
 \end{equation}
+
 and $\overline{(x-\bar x)^4}=3\sigma^4$. In particular, these results imply
+
 \begin{subequations}\label{no_inher}
@@
 \end{subequations}
+
 These equations allow us to derive the response in trait mean and variance by taking derivatives of fitness, a much more straightforward operation than calculating a covariance for general phenotypic distributions.
@@
 Hence,
+
 \begin{equation}
@@
 \end{equation}
+
 To include this relationship in our framework, we write $\rho(g,t)$ as the abundance density of breeding values at time $t$ so that $\int_{-\infty}^{+\infty}\rho(g,t)dg=\int_{-\infty}^{+\infty}\nu(x,t)dx=N(t)$.
@@
 Once $\rho(g,t)$ is determined, we can compute $\nu(x,t)$ via
+
 \begin{equation}\label{nu_from_rho}
@@
 \end{equation}
+
 However, since selection acts on expressed phenotypes, we use the assumed relationship between breeding values and expressed traits to calculate the fitness of breeding values.
@@
 Thus, the mean fitness of a breeding value $g$ across all individuals carrying $g$ can be written as
+
 \begin{equation}\label{m_rel}
@@
 \end{equation}
+
 This is similar to the approach taken by Kimura and Crow [-@Kimura1978] to calculate the overall effects of selection for expressed characters onto the changes in the distribution of alleles encoding those characters.
@@
 With the relationship between $m(\nu,x)$ and $m^*(\rho,g)$ established, we define the evolution of $\rho(g,t)$ by the SPDE
+
 \begin{equation}\label{rho_SPDE}
@@
 \end{equation}
+
 Equation (\ref{rho_SPDE}) is a stochastic generalization of the deterministic PDE (\ref{eq1}) from \S\ref{deterministic}, but describes the evolution of the distribution of breeding values instead of expressed characters.
@@
 Furthermore, since $\sigma^2=G+\eta$, we can use equation (\ref{m_rel}) and the chain rule from calculus (see SM \S\ref{fit2fit}) to find
+
 \begin{subequations}
@@
 
-# A model of diffuse coevolution \label{coev}
+A model of diffuse coevolution \label{coev}
+============
 
@@
 In SM \S\ref{diffuse} we combine our treatment of resource competition with equations (\ref{N}), (\ref{xbarfinal}) and (\ref{Gfinal}) to find
+
 \begin{subequations}\label{comm_dynamics}
 \begin{multline}\label{N_comm}
-dN_i = \Bigg\{R_i-\frac{a_i}{2}\Big((\bar x_i-\theta_i)^2+G_i+\eta_i\Big) - c_i\sum_{j=1}^SN_jU_iU_j\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\Bigg\}N_idt + \sqrt{V_iN_i}dW_1,
+dN_i = \Bigg\{R_i-\frac{a_i}{2}\Big((\bar x_i-\theta_i)^2+G_i+\eta_i\Big) - c_i\sum_{j=1}^SN_jU_iU_j\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\Bigg\}N_idt \\ + \sqrt{V_iN_i}dW_1,
 \end{multline}
 \begin{multline}\label{x_comm}
-d\bar x_i = \Bigg\{a_iG_i(\theta_i-\bar x_i)-c_iG_i\bigg(\sum_{j=1}^SN_jU_iU_jb_{ij}(\bar x_j-\bar x_i)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\bigg)\Bigg\}dt+\sqrt{V_i\frac{G_i}{N_i}}dW_2,
+d\bar x_i = \Bigg\{a_iG_i(\theta_i-\bar x_i)-c_iG_i\bigg(\sum_{j=1}^SN_jU_iU_jb_{ij}(\bar x_j-\bar x_i)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\bigg)\Bigg\}dt \\ +\sqrt{V_i\frac{G_i}{N_i}}dW_2,
 \end{multline}
 \begin{multline}\label{G_comm}
-dG_i =  \Bigg\{c_i{G_i}^2\bigg(N_iU_i^2b_{ii}\sqrt\frac{b_{ii}}{2\pi}+\sum_{j=1}^SN_jU_iU_jb_{ij}\left(1-b_{ij}(\bar x_i-\bar x_j)^2\right)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\bigg) \\
-+\mu_i-a_i{G_i}^2-V_i\frac{G_i}{N_i}\Bigg\}dt+G_i\sqrt{\frac{2V_i}{N_i}}dW_3,
+dG_i =  \Bigg\{c_i{G_i}^2\bigg(\sum_{j=1}^SN_jU_iU_jb_{ij}\left(1-b_{ij}(\bar x_i-\bar x_j)^2\right)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2} \\
++N_iU_i^2b_{ii}\sqrt\frac{b_{ii}}{2\pi}+\mu_i-a_i{G_i}^2-V_i\frac{G_i}{N_i}\bigg)\Bigg\}dt+G_i\sqrt{\frac{2V_i}{N_i}}dW_3,
 \end{multline}
 \end{subequations}
+
 where
@@
 Relating our treatment of resource competition to modern coexistence theory [@Chesson2000], the absolute competition coefficient $\alpha_{ij}$ becomes a dynamical quantity that can be written as
-\begin{equation}
-\alpha_{ij}(t)=\frac{c_i}{r_i(t)}\int_{-\infty}^{+\infty}\int_{-\infty}^{+\infty}p_i(x,t)p_j(y,t)\mathcal O_{ij}(x,y) dxdy =\frac{c_iU_iU_j}{r_i(t)}\sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right),
-\end{equation}
+
+\begin{multline}
+\alpha_{ij}(t)=\frac{c_i}{r_i(t)}\int_{-\infty}^{+\infty}\int_{-\infty}^{+\infty}p_i(x,t)p_j(y,t)\mathcal O_{ij}(x,y) dxdy \\ =\frac{c_iU_iU_j}{r_i(t)}\sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right),
+\end{multline}
+
 where
+
 \begin{equation}
@@
 \end{equation}
+
 Hence, $dN_i(t)$ can be expressed as
+
 \begin{equation}
@@
 \end{equation}
+
 Note that although $r_i(t)$ is referred to in the coexistence literature as the intrinsic growth rate of the population, $R_i$ is a deeper intrinsic quantity.
@@
 In SM \S\ref{diffuse} we show that the standardized directional selection gradient [sensu @Lande1983] induced by species $j$ on species $i$ can be computed as
-\begin{equation}
-\beta_{ij}(t)=c_iU_iU_jN_j(t)b_{ij}(t)\big(\bar x_i(t)-\bar x_j(t)\big) \sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right).
-\end{equation}
+
+\begin{multline}
+\beta_{ij}(t)=c_iU_iU_jN_j(t)b_{ij}(t)\big(\bar x_i(t)-\bar x_j(t)\big) \\
+\times\sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right).
+\end{multline}
+
 Our notation differs from Lande and Arnold [-@Lande1983] in that subscripts here denote species instead of components of multivariate traits and we drop the prime that distinguishes between selection gradients and standardized selection gradients.
 
-__Metric of pairwise coevolution__
-
+_Metric of pairwise coevolution:_
 Below we investigate the correspondence of interaction intensity and coevolutionary change.
@@
 In SM \S\ref{diffuse} we show that the standardized stabilizing selection gradient induced by species $j$ on species $i$ can be computed as
-\begin{equation}
-\gamma_{ij}(t)=c_iU_iU_jN_j(t)b_{ij}(t)\left(1-b_{ij}(t)\big(\bar x_i(t)-\bar x_j(t)\big)^2\right) \sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right).
-\end{equation}
+
+\begin{multline}
+\gamma_{ij}(t)=c_iU_iU_jN_j(t)b_{ij}(t)\left(1-b_{ij}(t)\big(\bar x_i(t)-\bar x_j(t)\big)^2\right) \\
+\times\sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right).
+\end{multline}
+
 To measure evolutionary change in species $i$ induced by species $j$, we form the metric $\Psi_{ij}=|\beta_{ij}|+|\gamma_{ij}|$ and define the strength of coevolution as $\mathfrak{C}_{ij}=\Psi_{ij}\Psi_{ji}$.
@@
 
-```{r, echo=F, fig.align='center', fig.cap='\\label{net}Bivariate distributions of competition coffecients (y-axis) and pairwise coevolution (x-axis) under the scenarios of weak competition (dashed line) and strong competition (solid line) after simulating for $t=1000.0$ units of time.
+```{r, echo=F, fig.align='center', out.width='60%', fig.cap='\\label{net}Bivariate distributions of competition coffecients (y-axis) and pairwise coevolution (x-axis) under the scenarios of weak competition (dashed line) and strong competition (solid line) after simulating for $t=1000.0$ units of time.
 Simulations were ran for $S=1000$ species.
@@
 
-__Covariance of selection and competition as a function of diversity__
-
+_Covariance of selection and competition as a function of diversity:_
 We now make use of the expressions derived for competition coeffecients and selection gradients to investigate their relationship.
@@
 If $\bar{\bar x}$ is near $\theta$ and $V_{\bar X}$ is much smaller than $|2\frac{R}{a}-G-\eta|$, then we may approximate $r_i$ with
-\begin{equation}
-\bar r=\int_{-\infty}^{+\infty}\left(R-\frac{a}{2}\Big((\bar x-\theta)^2+G+\eta\Big)\right)f_{\bar X}(\bar x)d\bar x
+
+\begin{multline}
+\bar r=\int_{-\infty}^{+\infty}\left(R-\frac{a}{2}\Big((\bar x-\theta)^2+G+\eta\Big)\right)f_{\bar X}(\bar x)d\bar x \\
 =R-\frac{a}{2}\Big((\bar{\bar x}-\theta)^2+V_{\bar X}+G+\eta\Big).
-\end{equation}
+\end{multline}
+
 In SM \S\ref{coef_grad_moms} we use these assumptions to calculate the first and second order moments describing the joint distribution of competition coefficients and selection gradients across the community.
@@
 However, setting $\alpha(\bar x_i,\bar x_j)=\alpha_{ij}$, $\beta(\bar x_i,\bar x_j)=\beta_{ij}$ and $\gamma(\bar x_i,\bar x_j)=\gamma_{ij}$, the covariances between the magnitude of linear selection gradients and competition coefficients and between stabilizing selection gradients and competition coefficients can be written as
+
 \begin{subequations}\label{cov_eqns}
-\begin{equation}\label{cov_alpha_beta}
-\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)=\frac{2c^2b^2U^4N}{\pi\bar r}\sqrt\frac{V_{\bar X}}{2\pi}\left(\frac{1}{(1+8bV_{\bar X})^{3/4}}-\frac{1}{(1+4bV_{\bar X})^{3/4}(1+2bV_{\bar X})^{1/2}}\right),
-\end{equation}
+\begin{multline}\label{cov_alpha_beta}
+\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)=\frac{2c^2b^2U^4N}{\pi\bar r}\sqrt\frac{V_{\bar X}}{2\pi} \\
+\times\left(\frac{1}{(1+8bV_{\bar X})^{3/4}}-\frac{1}{(1+4bV_{\bar X})^{3/4}(1+2bV_{\bar X})^{1/2}}\right),
+\end{multline}
 \begin{equation}\label{cov_alpha_gamma}
@@
 
-```{r, echo=F, fig.align='center', fig.cap='\\label{cov_fig}Curves representing the covariance between the magnitude of linear selection gradients and competition coefficients (left) and between stabilizing selection gradients and competition coefficients (right) as a function of $2bV_{\\bar X}$ which is approximately equal to the ratio of variance in mean traits among species to the intraspecific trait variance.
+```{r, echo=F, fig.align='center', out.width='100%', fig.cap='\\label{cov_fig}Curves representing the covariance between the magnitude of linear selection gradients and competition coefficients (left) and between stabilizing selection gradients and competition coefficients (right) as a function of $2bV_{\\bar X}$ which is approximately equal to the ratio of variance in mean traits among species to the intraspecific trait variance.
 In both plots we set $c=1.0\\times10^{-4}$, $b=0.1$, $\\bar r=0.1$ and $N=1.0\\times10^{10}$ and let $V_{\\bar X}$ vary.'}
@@
 
-```{r, echo=F, fig.align='center', fig.cap='\\label{cor_coev_fig}Numerical estimates for the correlation of the strength of coevolution measured by $\\mathfrak{C}$ and competition coefficeints $\\alpha$ plotted against the variance of mean traits among species divided by the mean variance of traits within species.
+```{r, echo=F, fig.align='center', out.width='60%', fig.cap='\\label{cor_coev_fig}Numerical estimates for the correlation of the strength of coevolution measured by $\\mathfrak{C}$ and competition coefficeints $\\alpha$ plotted against the variance of mean traits among species divided by the mean variance of traits within species.
 Each dot represents the result from a single simulation.
@@
 
-# Conclusion \label{conclusion}
+Conclusion \label{conclusion}
+===================
 
@@
 
-__Limitations of diffusion limits and generalizing coexistence theory__
+## Limitations of diffusion limits and generalizing coexistence theory
 
@@
 
-__Genetic architecture and intraspecific trait distributions__
+## Genetic architecture and intraspecific trait distributions
 
@@
 
-__Ecological stoichiometry__
+## Ecological stoichiometry
 
@@
 
-__Accounting for macroevolutionary history__
+## Accounting for macroevolutionary history
 
@@
 
-__Final remarks__
+# Final remarks
 
@@
 
-\newpage
-
-# Supplementary material (SM)
-
-Throughout this supplement, we set use dot notation for time derivatives so that $\dot f(x,t)=\frac{\partial }{\partial t}f(x,t)$ and set $\Delta=\sum_{i=1}^d\frac{\partial^2}{\partial x_i^2}$ the Laplace operator on $\mathbb{R}^d$, except in \S\ref{caC} where $\Delta$ represents a random variable.
-
-<!--
-
-## Existence, uniqueness and regularity of solution to deterministic PDE (\ref{eq1}) of the main text \label{regular}
-
-Here we show that solutions to the PDE described in the main text exist, are unique and are twice continuously differentiable with respect to space.
-Our approach is an adaptation of that taken in chapter 9 \S2.2 of Volpert [-@Volpert2014].
-
-Recall that the PDE we consider is of the form
-\begin{equation}\label{PDE_supp}
-\begin{matrix}
-\dot\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\Delta\nu(x,t), \\
-\\
-\nu(x,0)=u(x),
-\end{matrix}
-\end{equation}
-where $m(\nu,x)\leq r\in\mathbb{R}$ is shorthand for $m((K\nu)(x,t),x)$ and $(K\nu)(x,t)=\int_\mathbb{R}\kappa(x-y)\nu(y,t)dy$ for some non-negative, bounded $\kappa$.
-Denoting $C(\mathbb{R})$ the set of continuous functions on $\mathbb{R}$ and $L^1(\mathbb{R})$ the Lebesgue space of integrable functions on $\mathbb{R}$ (ie, the set of functions $f:\mathbb{R}\to\mathbb{R}$ that satisfy $\|f\|_1=\int_\mathbb{R}|f(x)|dx<+\infty$ under the equivalence relation $f=g$ iff $\|f-g\|_1=0$), we assume a continuous and integrable initial condition $\nu(\cdot,0)=u\in L^1(\mathbb{R})\cap C(\mathbb{R})$.
-We now prove the existence and uniqueness of solutions to this Cauchy problem.
-
-Set 
-\begin{equation}
-\Phi(x,t)=\frac{\exp\left(-x^2/2\mu t\right)}{\sqrt{2\pi\mu t}}.
-\end{equation}
-Then the semigroup $\{S(t)\}_{t\geq0}$ given by 
-\begin{equation}
-S(t)v=\left\{\begin{matrix}
-\int_0^t\frac{\mu}{2}\Delta\nu(x,s)ds=\int_\mathbb{R}\Phi(x-y,t)v(y)dy, & t>0 \\
- & \\
-v, &t=0
-\end{matrix}\right.
-\end{equation}
-for all $v\in L^\infty(\mathbb{R})$ satisfies $\|S(t)v\|_1=\|v\|_1$ for all $t\geq0$ and $v\in L^1(\mathbb{R})$. We also define the operator
-\begin{equation}
-(Fv)(x)=m\left(\int_\mathbb{R}\kappa(x-y)v(y)dy,x\right)v(x), \ \ \forall \ \ v\in L^1(\mathbb{R}).
-\end{equation}
-Thus $\|Fv\|_1\leq |r| \ \|v\|_1$ for all $v\in L^1(\mathbb{R})$. Then, for any $v,w\in L^1(\mathbb{R})$,
-\begin{multline}
-Fv-Fw=F(v+w)-F(v+w)+Fv-Fw \\
-=m(v+w,x)(v+w)-m(w,x)w-m(v+w,x)(v+w)+m(v,x)v \\
-=\Big(m(v+w,x)-m(w,x)\Big)w-\Big(m(v+w,x)-m(v,x)\Big)v+m(v+w,x)(v-w).
-\end{multline}
-Hence $\|Fv-Fw\|_1\leq2|r|(\|w\|_1+\|v\|_2)+|r|\|v-w\|_1\leq\|r\|\|v-w\|_1$ which implies $F$ is Lipschitz continuous on $L^1(\mathbb{R})$.
-For $T>0$ set $L^{\infty,1}([0,T]\times\mathbb{R})$ as the set of functions $\nu:[0,T]\to L^1(\mathbb{R})$ that satisfy $\|\nu\|_{1,T}=\sup_{t\in[0,T]}\|\nu(t)\|_1<+\infty$ and consider the operator $G$ defined on $L^{\infty,1}([0,T]\times\mathbb{R})$ for fixed $u\in L^1(\mathbb{R})$ by
-\begin{equation}
-(Gv)(t)=S(t)u+\int_0^tS(t-s)(Fv(s))ds \ \ \forall \ \ t\in[0,T].
-\end{equation}
-Then for any $t\in[0,T]$ and $\nu_1,\nu_2\in L^{\infty,1}([0,T]\times\mathbb{R})$,
-
-\begin{multline}
-\|G\nu_1-G\nu_2\|_{1,T}=\sup_{t\in[0,T]}\left\|\int_0^tS(t-s)(F\nu_1(s)-F\nu_2(s))ds\right\|_1 \\ \leq \sup_{t\in[0,T]}t|r|\|\nu_1(t)-\nu_2(t)\|_1\leq T|r|\|\nu_1-\nu_2\|_{1,T}.
-\end{multline}
-Hence, when $T<\frac{1}{|r|}$ the operator $G$ is a contraction operator.
-In this case we apply Banach's fixed point theorem to demonstrate $G$ has a unique fixed point.
-That is, there exists a unique $\nu\in L^{\infty,1}([0,T]\times\mathbb{R})$ such that $G\nu=\nu$.
-Then, for $\nu=\nu(x,t)$ and considering $\nu(t)$ the function $x\to\nu(x,t)$, we have by definition $\nu$ is the mild solution of PDE (\ref{PDE_supp}) with initial condition $\nu(0)=u$.
-To extend this to arbitrarily large $T$ we take a similar approach as Evans [-@lawrenceevans2010] and choose $T_1<\frac{1}{|r|}$ and apply the above argument to the interval $[T_1,2T_1]$.
-Repeating this task a finite number of times shows $\nu$ is the mild solution of PDE (\ref{PDE_supp}) with initial condition $\nu(0)=u$ for arbitrarily large $T>0$.
-
-We shift our focus to showing mild solutions $\nu$ to PDE (\ref{PDE_supp}) are twice differentiable in $x$ and hence are classical solutions.
-Applying lemma 3.3.2 of Henry [-@Henry1981] we have that the mild solution $\nu$ is in fact a strong solution (i.e., differentiable with respect to $t\in(0,T)$) to the abstract Cauchy problem defined by PDE (\ref{PDE_supp}) with initial condition $\nu(0)=u$ for $t\in[0,T]$.
-In particular, this implies that $\nu(t)\in D(\Delta)$ for each $t\in(0,T)$ where $D(\Delta)=W^{2,1}(\mathbb{R})$ is the Sobolov space of integrable functions with integrable weak derivatives at least of order $2$.
-Since $\frac{1}{p}\geq\frac{1}{p+1}\geq\frac{1}{p}-2$ for all $p\geq1$, the Sobolov embedding theorem [@Henry1981] implies $D(\Delta)\subset L^p(\mathbb{R})$, the Lebesgue space of $p$-integrable functions on $\mathbb{R}$, for all $p\geq1$.
-Hence $D(\Delta)\subset W^{2,p}(\mathbb{R})$, the Sobolov space of $p$-integrable functions with weak derivatives at least of order $2$, for all $p\geq1$.
-
-If $v\in C^{2+\delta}(\mathbb{R})$ for some $\delta>0$, then 
-\begin{equation}
-(Kv)(\cdot)=\int_\mathbb{R}\kappa(\cdot-y)v(y)dy=\int_\mathbb{R}\kappa(y)v(\cdot-y)dy\in C^{2+\delta}(\mathbb{R}),
-\end{equation}
-and, since we assume $m(h,x)$ is continuous in both arguments, $Fv=m(Kv,\cdot)v\in C^\delta$.
-
-Why is $F\nu\in C^\delta(\mathbb{R})$ for some $\delta>0$?
-Why does this imply $\nu(\cdot,t)\in C^{2+\delta}(\mathbb{R})$ for all $t\in(0,T)$?
-This argument holds for all $T>0$.
-Hence, solutions to the Cauchy problem (\ref{PDE_supp}) are twice continuously differentiable in their spatial parameter for all $t>0$.
-
--->
-
-## Sufficient conditions for finite mean, variance and total abundance in the deterministic case \label{finite}
-
-Recall that $m(\nu,x)$ is shorthand for $m\big((K\nu)(x,t),x\big)$.
-That is, $m:[0,+\infty)\times\mathbb{R}\to\mathbb{R}$.
-Following our assumptions of the main text, we have that $m(h,x)$ is differentiable with respect to both $x$ and $h$ and there exists $r\in\mathbb{R}$ such that $m(h,x)\leq r$ across all $h\geq0$ and $x\in\mathbb{R}$.
-As in the main text, we also assume the initial condition $u(x)$ is continuous and integrable in $x$ and satisfies
-\begin{equation}
-0\leq\int_{\mathbb{R}}(|x|+x^2)u(x)dx<+\infty
-\end{equation}
-and consider the Cauchy problem
-\begin{equation}\label{PDE_SM}
-\left\{\begin{matrix}
-\dot\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\Delta\nu(x,t) & t>0\\
-\nu(x,0)=u(x) & t=0.
-\end{matrix}\right.
-\end{equation}
-
-According to Theorem 2.5.6 of Zheng [-@zheng2004nonlinear], if the operator $F$ defined by $\nu(x,t)\to m(\nu,x)\nu(x,t)$ is locally Lipschitz, corresponding to equation (\ref{local_lipschitz}) of the main text, then for some maximal $T>0$ problem (\ref{PDE_SM}) admits a unique local classical solution $\nu(x,t)$ for $t\in[0,T)$.
-Furthermore, either $T=+\infty$ or $T<+\infty$ and $\lim_{t\uparrow T}\int_\mathbb{R}|\nu(x,t)|dx=+\infty$.
-In this section we show that our assumption $m(h,x)\leq r$ for all $h\geq0$ and $x\in\mathbb{R}$ implies $T=+\infty$.
-Replacing $m$ with it's upper bound $r\in\mathbb{R}$, equation (\ref{PDE_SM}) reduces to a simple parabolic equation that can be solved using elementary techniques [@stanleyfarlow1993].
-In particular, when $m(\nu,x)\equiv0$ denote the solution to (\ref{PDE_SM}) by $\nu_0(x,t)$.
-Then, denoting
-\begin{equation}
-\Phi(x,t)=\frac{\exp\left(-x^2/2\mu t\right)}{\sqrt{2\pi\mu t}},
-\end{equation}
-we have
-\begin{equation}
-\nu_0(x,t)=\int_{\mathbb{R}}\Phi(x-y,t)\nu(y,0)dy.
-\end{equation}
-In the more general case, when $m(\nu,x)\equiv r\in\mathbb{R}$, equation (\ref{PDE_SM}) has the solution $\nu_r(x,t)=e^{rt}\nu_0(x,t)$.
-Hence, $\nu_r(x,t)\geq0$ for all $x\in\mathbb{R}$ and $\int_\mathbb{R}\nu_r(x,t)dx=e^{rt}N(0)<+\infty$ for all $t\geq0$.
-Furthermore, denoting
-\begin{equation}
-\begin{matrix}
-N_r(t)&=&\int_\mathbb{R}\nu_r(x,t)dx, \\
-p_r(x,t)&=&\nu_r(x,t)/N_r(t), \\
-\bar x_r(t)&=&\int_\mathbb{R}xp_r(x,t)dx, \\
-\sigma^2_r(t)&=&\int_\mathbb{R}(x-\bar x_r(t))^2p_r(x,t)dx, 
-\end{matrix}
-\end{equation}
-we have
-\begin{equation}
-\bar x_r(t)=\int_\mathbb{R}x\int_\mathbb{R}\Phi(x-y,t)p(y,0)dydx=\int_\mathbb{R}yp(y,0)dy=\bar x(0),
-\end{equation}
-\begin{equation}
-\sigma_r^2(t)=\int_\mathbb{R}(x-\bar x_r(t))^2\int_\mathbb{R}\Phi(x-y,t)p(y,0)dydx=\int_\mathbb{R}\Big((y-\bar x(0))^2+\mu t\Big)p(y,0)dy=\sigma^2(0)+\mu t.
-\end{equation}
-Hence, $|\bar x_r(t)|,\sigma_r^2(t)<+\infty$ for all $t\geq0$. For the sake of contradiction, suppose there exists $x\in\mathbb{R}$ and $t\geq0$ such that $\nu(x,t)>\nu_r(x,t)$. Then
-\begin{equation}
-\nu(x,t)-\nu(x,0)=\int_0^tm(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)ds>\int_0^tr\nu_r(x,s)+\frac{\mu}{2}\Delta\nu_r(x,s)ds=\nu_r(x,t)-\nu(x,0)
-\end{equation}
-which implies there exists $h\geq0$ and $x\in\mathbb{R}$ such that $m(h,x)>r$.
-But this contradicts our assumption $m(h,x)\leq r$ for all $h\geq0$ and $x\in\mathbb{R}$.
-So we have $\nu(x,t)\leq\nu_r(x,t)$ for each $x\in\mathbb{R}$ and $t\geq0$.
-This implies that $N(t)=\int_\mathbb{R}\nu(x,t)dx<+\infty$,
-\begin{equation}
-0\leq\int_\mathbb{R}x^2\nu(x,t)dx\leq\int_\mathbb{R}x^2\nu_r(x,t)dx<+\infty
-\end{equation}
-and in particular
-\begin{equation}
-0\leq\sigma^2(t)+\bar x^2(t)=\frac{1}{N(t)}\int_\mathbb{R}x^2\nu(x,t)dx<+\infty
-\end{equation}
-for each $t\geq0$.
-
-## Equilibrium moments for a deterministic population experiencing logistic growth and stabilizing selection \label{equilib}
-
-Here we set out to show, given the initial conditions $\nu(\cdot,0)\in L^1(\mathbb{R})$ such that $\bar x(0)\in\mathbb{R}$, $\sigma^2(0)\in[0,+\infty)$ and $N(0)\in(0,+\infty)$, and given the growth rate
-\begin{equation}\label{m_log_stab}
-m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_\mathbb{R}\nu(y,t)dy
-\end{equation}
-such that $\theta\in\mathbb{R}$, $a,c,\mu>0$ and $r>\tfrac{1}{2}\sqrt{\mu a}$, we have the following stable equilibrium $N_\infty=\lim_{t\to\infty}N(t)=\tfrac{1}{c}(r-\tfrac{1}{2}\sqrt{a\mu})$, $\bar x_\infty=\lim_{t\to\infty}\bar x(t)=\theta$ and $\sigma^2_\infty=\lim_{t\to\infty}\sigma^2(t)=\sqrt{\tfrac{\mu}{a}}$.
-
-The mean fitness can be calculated as
-\begin{equation}
-\bar m(t)=r-\frac{a}{2}((\theta-\bar x(t))^2+\sigma^2(t))-cN(t).
-\end{equation}
-Recalling the ODE derived for $N(t)$,
-\begin{equation}
-\frac{d}{dt}N(t)=\bar m(t)N(t)=(r-\frac{a}{2}((\theta-\bar x(t))^2+\sigma^2(t))-cN(t))N(t),
-\end{equation}
-solving for equilibrium total abundance $\hat N$ amounts to setting $\frac{d}{dt}N(t)=0$ and solving for $N(t)$.
-Ignoring the equilibrium $N(t)=0$, this reduces to solving $\bar m(t)=0$ for $N(t)$, which returns
-\begin{equation}
-\hat N=\frac{1}{c}\left(r-\frac{a}{2}\left((\theta-\hat{\bar x})^2-\hat\sigma^2\right)\right).
-\end{equation}
-Unfortunately, deriving ODE for $\bar x(t)$ and $\sigma^2(t)$ leads to expressions involving higher moments and finding ODE for these higher moments will lead to expressions involving yet even higher moments.
-To avoid this infinite regression, we find the equilibrium abundance density $\hat\nu(x)$ by solving $\frac{\partial}{\partial t}\nu(x,t)=0$ for $\nu(x,t)$.
-This implies the following ordinary differential equation
-\begin{equation}
-\frac{d^2}{dx^2}\hat\nu(x)=\left(\frac{2c}{\mu}\hat N+\frac{a}{\mu}(\theta-x)^2-\frac{2r}{\mu}\right)\hat\nu(x)
-\end{equation}
-which has the solution
-\begin{equation}
-\hat\nu(x)=\frac{\hat N}{\sqrt{2\pi}}\left(\frac{a}{\mu}\right)^{\frac{1}{4}}\exp\left(-\sqrt\frac{a}{\mu}\frac{(\theta-x)^2}{2}\right).
-\end{equation}
-From this expression we infer $\hat{\bar x}=\theta$ and $\hat\sigma^2=\sqrt{\frac{\mu}{a}}$.
-Hence $\hat N=\frac{1}{c}\left(r-\frac{1}{2}\sqrt{a\mu}\right)$.
-To show that $N_\infty=\hat N$, $\bar x_\infty=\hat{\bar x}$ and $\sigma^2_\infty=\hat\sigma^2$, we linearize ODE for $N(t)$, $\bar x(t)$ and $\sigma^2(t)$ using the equilibrium $\hat\nu(x)$.
-In particular, since $\hat\nu(x)$ is Gaussian, we do not run into the same issue with higher moments as above.
-Assuming the equilibrium $\hat\nu(x)$ does not change the ODE for $N(t)$, but the ODE for $\bar x(t)$ and $\sigma^2(t)$ can now be expressed as
-\begin{equation}
-\frac{d}{dt}\bar x(t)=\sigma^2(t)\left(\frac{\partial\bar m(t)}{\partial\bar x(t)}-\overline{\frac{\partial m(t)}{\partial\bar x(t)}}\right)=a\sigma^2(t)(\theta-\bar x(t)),
-\end{equation}
-\begin{equation}
-\frac{d}{dt}\sigma^2(t)=2\sigma^4(t)\left(\frac{\partial\bar m(t)}{\partial\sigma^2(t)}-\overline{\frac{\partial m(t)}{\partial\sigma^2(t)}}\right)+\mu=\mu-a\sigma^4(t),
-\end{equation}
-where $\hat p(x)=\hat\nu(x)/\hat N$. These expressions confirm our findings that $\hat{\bar x}=\theta$ and $\hat\sigma^2=\sqrt{\frac{\mu}{a}}$. Furthermore, calculating
-\begin{equation}
-\frac{\partial}{\partial\sigma^2(t)}\frac{d}{dt}\sigma^2(t)=-2a\sigma^2(t)
-\end{equation}
-and evaluating at $\sigma^2(t)=\hat\sigma^2$ demonstrates the equilibrium phenotypic variance is stable when $a,\mu>0$. Hence, calculating
-\begin{equation}
-\frac{\partial}{\partial\bar x(t)}\frac{d}{dt}\bar x(t)=-a\sigma^2(t)
-\end{equation}
-and evaluating at $\sigma^2(t)=\hat\sigma^2$ and $\bar x(t)=\hat{\bar x}$ demonstrates the equilibrium phenotypic mean is stable when $a,\mu>0$. Finally, calculating
-\begin{equation}
-\frac{\partial}{\partial N(t)}\frac{d}{dt}N(t)=r-\frac{a}{2}((\theta-\bar x(t))^2+\sigma^2(t))-2cN(t)
-\end{equation}
-and evaluating at $\sigma^2(t)=\hat\sigma^2$, $\bar x(t)=\hat{\bar x}$ and $N(t)=\hat N$ demonstrates the equilibrium total abundance is stable when $a,c,\mu>0$, and $r>\frac{1}{2}\sqrt{a\mu}$.
-
-## The relation between diffusion and convolution with a Gaussian kernel \label{diffconvequiv}
-
-For continuous $g:\mathbb{R}^d\to\mathbb{R}$, consider the deterministic Cauchy problem
-\begin{equation}\label{heateqn}
-\left\{\begin{matrix}
-\dot f(x,t)=&\Delta f(x,t), & (x,t)\in\mathbb{R}^d\times(0,\infty)\\
-f(x,t)=&g(x), & (x,t)\in\mathbb{R}^d\times\{0\}.
-\end{matrix}\right.
-\end{equation}
-According to Evans [-@lawrenceevans2010], the fundamental solution of (\ref{heateqn}) is
-\begin{equation}
-\Phi(x,t)=\frac{1}{(4\pi t)^{d/2}}\exp\left(-\frac{|x|^2}{4t}\right), \ (x,t)\in(0,\infty)\times\mathbb{R}^d,
-\end{equation}
-where $|x|=\sqrt{\sum_ix_i^2}$. The solution $f(x,t)$ of PDE (\ref{heateqn}) is then given by the convolution
-\begin{equation}
-f(x,t)=\int_{\mathbb{R}^d}\Phi(x-y,t)g(y)dy, \ (x,t)\in(0,\infty)\times\mathbb{R}^d.
-\end{equation}
-Hence, by the fundamental theorem of calculus,
-
-\begin{multline}
-f(x,t)+\int_t^{t+1}\dot f(x,s)ds=f(x,t+1) \\
-=\int_{\mathbb{R}^d}\Phi(x-y,t+1)g(y)dy=\int_{\mathbb{R}^d}\int_{\mathbb{R}^d}\Phi(x-y,1)\Phi(y-z,t)g(z)dzdy \\
-=\int_{\mathbb{R}^d}\Phi(x-y,1)f(t,y)dy.
-\end{multline}
-In particular,
-\begin{equation}
-f(x,t)+\int_t^{t+1}\Delta f(x,s)ds=\int_{\mathbb{R}^d}\Phi(1,x-y)f(y,t)dy.
-\end{equation}
-
-## Deterministic dynamics of $\sigma^2(t)$ \label{var_deriv}
-
-Picking up from the main text \S\ref{deterministic},
-
-\begin{multline}
-\frac{d}{dt}\sigma^2(t)=\frac{d}{dt}\int_\mathbb{R}(x-\bar x(t))^2p(x,t)dx=\int_\mathbb{R}2(x-\bar x(t))\frac{d}{dt}{\bar x}(t)+(x-\bar x(t))^2\frac{\partial}{\partial t} p(x,t)dx\\=\int_\mathbb{R}(x-\bar x(t))^2\left((m(\nu,x)-\bar m(t))p(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t)\right)dx\\=\int_\mathbb{R}\left((x-\bar x(t))^2-\sigma^2(t)+\sigma^2(t)\right)(m(\nu,x)-\bar m(t))p(x,t)+(x-\bar x(t))^2\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t)dx\\=\mathrm{Cov}_t\Big((x-\bar x(t))^2,m(\nu,x)\Big)+\frac{\mu}{2}\int_\mathbb{R}(x-\bar x(t))^2\frac{\partial^2}{\partial x^2}p(x,t)dx.
-\end{multline}
-Applying integration by parts twice yields
-\begin{equation}
-\int_{-\infty}^{+\infty}(x-\bar x(t))^2\frac{\partial^2}{\partial x^2}p(x,t)dx=2.
-\end{equation}
-
-## Simplifying covariances with fitness under the assumption of a Gaussian density \label{cov2deriv}
-
-By assuming
-
-\begin{equation}
-p(x,t)=\frac{\exp\left(-\frac{(x-\bar x(t))^2}{2\sigma^2(t)}\right)}{\sqrt{2\pi\sigma^2(t)}}
-\end{equation}
-
-we have
-
-\begin{multline}
-\sigma^2\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right)=\sigma^2\left(\frac{\partial}{\partial\bar x}\int_\mathbb{R}m(\nu,x)p(x,t)dx-\int_\mathbb{R}p(x,t)\frac{\partial}{\partial \bar x}m(\nu,x)dx\right)\\=\sigma^2\int_\mathbb{R}m(\nu,x)\frac{\partial}{\partial\bar x}p(x,t)dx=\sigma^2\int_\mathbb{R}\frac{x-\bar x(t)}{\sigma^2}m(\nu,x)p(x,t)dx\\=\int_\mathbb{R}(x-\bar x)(m(\nu,x)-\bar m)p(x,t)dx=\mathrm{Cov}_t(m,x),
-\end{multline}
-
-and
-
-\begin{multline}
-2\sigma^4\left(\frac{\partial\bar m}{\partial\sigma^2}-\overline{\frac{\partial m}{\partial\sigma^2}}\right)=2\sigma^4\left(\frac{\partial}{\partial\sigma^2}\int_\mathbb{R}m(\nu,x)p(x,t)dx-\int_\mathbb{R}p(x,t)\frac{\partial}{\partial\sigma^2}m(\nu,x)dx\right)\\=2\sigma^4\int_\mathbb{R}\frac{(x-\bar x)^2-\sigma^2}{2\sigma^4}m(\nu,x)p(x,t)dx=\int_\mathbb{R}\left((x-\bar x)^2-\sigma^2\right)\big(m(\nu,x)-\bar m\big)p(x,t)dx\\=\mathrm{Cov}_t\Big((x-\bar x)^2,m\Big).
-\end{multline}
-
-## Comparing our treatment of white noise to Da Prato and Zabczyk (2014) \label{DPZ14}
-
-Our approach in the main text is inspired by the treatment provided in \S4.2 of Da Prato and Zabczyk [-@DaPrato2014].
-Here the authors develop a stochastic integral of operator-valued processes.
-In particular, they consider processes indexed by time $t\geq0$ valued as Hilbert-Schmidt operators $\Phi(t)$ and define the norm
-\begin{equation}
-\|\Phi\|_t=\sqrt{\mathbb{E}\left(\int_0^t\mathrm{Tr}[\Phi(s)\Phi^*(s)]ds\right)}, \ t\geq0.
-\end{equation}
-In our case we only consider the so-called multiplication operators.
-That is, processes that consist of operators $\Phi(t)$ having the form $\Phi(t)g(x)=\varphi(x,t)g(x)$ such that $\varphi(\cdot,t)\in L^2(\mathbb{R})$ a.s.
-for each $t\geq0$.
-In this case $\Phi(t)=\Phi^*(t)$ and
-\begin{equation}
-\|\Phi\|_t=\|\varphi\|_t=\sqrt{\mathbb{E}\left(\int_0^t\int_\mathbb{R}\varphi^2(x,s)dxds\right)}, \ t\geq0.
-\end{equation}
-Da Prato and Zabczyk [-@DaPrato2014] form the space $\mathscr{N}^2_W(0,T)$ of Hilbert-Schmidt operator-valued predictable processes $\Phi(t)$ that satisfy $\|\Phi\|_T<+\infty$ for some $T>0$.
-This corresponds to our more specialized space $\mathscr{N}_2$ that consists of $L^2(\mathbb{R})$ valued processes $\varphi(x,t)$ such that $\|\varphi\|_t<+\infty$ for all $t\geq0$.
-In their treatment, $W(t)$ plays a similar role to our generalized process $\mathbf{W}_t$.
-For $\Phi\in\mathscr{N}^2_W(0,T)$, they denote the stochastic integral for $t\in[0,T]$ by $\Phi\cdot W(t)$.
-Hence, for $\Phi(t) g(x)=\varphi(x,t)g(x)$ as above, $\mathbf{W}_t(\varphi)=\Phi\cdot W(t)$.
-The authors then prove the following:
-
-  > __Proposition 4.28__ _Assume that $\Phi_1,\Phi_2\in\mathscr{N}^2_W(0,T)$. Then_
-  >
-  > $$\mathbb{E}(\Phi_i\cdot W(t))=0, \ \ \mathbb{E}(\|\Phi_i\cdot W(t)\|^2)<+\infty, \ \ \forall \ t\in[0,T].$$
-  >
-  > __Corollory 4.29__ _Under the same assumptions as Proposition 4.28,_
-  >
-  > $$\mathbb{C}(\Phi_1\cdot W(t),\Phi_2\cdot W(s))=\mathbb{E}\left(\int_0^{t\wedge s}\mathrm{Tr}[\Phi_2(r)\Phi_1^*(r)]dr\right), \ \ \forall \ t,s\in[0,T].$$
-
-Simplifying these expressions for the multiplication operators described above returns equations (\ref{exp_WN_xi}) and (\ref{cov_WN_xi}) of the main text.
-
-## Simulating the rescaled process \label{numerical}
-
-Here we provide a detailed description of the branching random walk and how we have chosen to rescale it. We focus on the case
-\begin{equation}
-m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_\mathbb{R}\nu(x,t)dx.
-\end{equation}
-
-### Description of simulation \label{description}
-
-We begin by describing the branching random walk before introducing our scheme to rescale it.
-Our branching random walk follows closely the description of branching Brownian motion in the main text.
-However, we replace exponentially distributed lifetimes with deterministic unit time steps for easier implementation.
-Hence, we restrict time to $t=0,1,2,\dots,$ and so on.
-Furthermore, we allow individual fitness to depend on both trait value and the state of the entire population.
-For time $t$ we write $\{\xi_1(t),\dots,\xi_{N(t)}(t)\}$ as the set of trait values across all $N(t)$ individuals alive in the population.
-Since our simulation follows discrete individuals instead of continuous distributions of trait values, we can write
-\begin{equation}
-\nu(x,t)=\sum_{i=1}^{N(t)}\delta(x-\xi_i(t)),
-\end{equation}
-where $\delta(x-\xi_i(t))$ denotes the point mass located at $\xi_i(t)$.
-For simplicity we assume perfect heritability.
-At each iteration we draw, for each individual, a random number of offspring from a Negative-Binomial distribution.
-Recall the Negative-Binomial distribution models the number of failed Bernoulli trials that occur before a given number of successful trials.
-Denoting $q$ the probability of success for each trial and $s$ the number of successes, the mean and variance is given respectively by
-\begin{equation}
-\frac{s(1-q)}{q}, \ \frac{s(1-q)}{q^2}.
-\end{equation}
-Then if we require the $i$th individual to have mean number offspring $\mathscr W(\nu,\xi_i)$ and variance equal to $V$, the parameters of the associated Negative-Binomial distribution become
-\begin{equation}
-q(\nu,\xi_i)=\frac{\mathscr{W}(\nu,\xi_i)}{V}, \ s(\nu,\xi_i)=\frac{\mathscr{W}^2(\nu,\xi_i)}{V-\mathscr{W}(\nu,\xi_i)}.
-\end{equation}
-This imposes the restriction $V>\mathscr{W}(\nu,\xi_i)$.
-For each offpsring produced by the individual with trait value $\xi_i(t)$, we assign indepently drawn trait values normally distributed around $\xi_i(t)$ with variance $\mu$.
-This summarizes the basic structure of our simulation.
-To impose selection and density dependent growth rates, we set
-\begin{equation}
-\mathscr{W}(\nu,\xi_i)=\exp\left(r-\frac{a}{2}(\theta-\xi_i)^2-c\int_\mathbb{R}\nu(x,t)dx\right),
-\end{equation}
-where the above integral becomes $\int_\mathbb{R}\nu(x,t)dx=\sum_{i=1}^{N(t)}1=N(t)$.
-
-__Rescaling__
-
-To rescale the branching random walk by a positive integer $n$, we rescale mutational variance by $\mu\to \mu/n$, time by $t\to t/n$ and the reproductive law by $V\to V$ and
-\begin{equation}
-\mathscr W(\nu,\xi_i)\to\mathscr W^{(n)}(\nu,\xi_i)=\exp\left(\frac{r}{n}-\frac{a}{2n}(\theta-\xi_i)^2-\frac{c}{n^2}N(t)\right)=\exp\left(\frac{r}{n}-\frac{a}{2n}(\theta-\xi_i)^2-\frac{c}{n}N^{(n)}(t)\right).
-\end{equation}
-We also replace individual mass with $\frac{1}{n}$ and write rescaled abundance as $N^{(n)}(t)=\frac{1}{n}N(nt)$.
-When it exists, we denote by $N^{(\infty)}(t)$ the limiting process of $N^{(n)}(t)$.
-Then
-\begin{equation}
-\lim_{n\to\infty}n\left(\mathscr W^{(n)}(\nu,\xi_i)-1\right)=r-\frac{a}{2}(\theta-\xi_i)^2-cN^{(\infty)}(t).
-\end{equation}
-Note that, since the limiting fitness function is bounded above and decreases linearly with respect to $N^{(\infty)}(t)$, it satisfies the hypotheses of Champagnat, Ferrière and Méléard [-@Champagnat2006].
-We have implemented this simulation in the programming language Julia.
-A copy can be found at the url:
-$$\texttt{\url{https://github.com/bobweek/branching.brownian.motion.and.spde}}$$
-For the sake of illustration, we simulated the unscaled process ($n=1$) and the rescaled process with $n=5$ and $n=20$ for $50$ units of time.
-Results are shown in Figure \ref{rescaled} of the main text.
-
-## Derivation of SDE for $\bar x$ and $\sigma^2$ \label{SDE_DERIV}
-
-Picking up from \S\ref{stochastic} of the main text, we have
-\begin{equation}
-\tilde x(t)=\int_\mathbb{R}x\nu(x,t)dx, \ \ \tilde{\tilde\sigma}^2(t)=\int_\mathbb{R}x^2\nu(x,t)dx
-\end{equation}
-and
-\begin{equation}\label{xtilde}
-\tilde x(t)=\tilde x(0)+\int_0^t\int_\mathbb{R}\nu(x,s)m(\nu,x)x+x\sqrt{V\nu(x,s)}\dot W(x,s)dxds,
-\end{equation}
-\begin{equation}\label{stilde}
-\tilde{\tilde\sigma}^2(t)=\tilde{\tilde\sigma}^2(0)+\int_0^t\int_\mathbb{R}\nu(x,s)\left(m(\nu,x)x^2+\mu\right)+x^2\sqrt{V\nu(x,s)}\dot W(x,s)dxds.
-\end{equation}
-
-<!--
-
-For $\nu(x,t)$ defined in the main text, $h\in C(\mathbb{R})$ and $t\geq0$ we make the following assumptions:
-
-\begin{equation}\label{L1_nu}
-\mathbb{E}\left(\int_\mathbb{R}|h(x)|\nu(x,t)dx\right)<\infty,
-\end{equation}
-\begin{equation}\label{L2_nu}
-\mathbb{E}\left(\int_\mathbb{R}h^2(x)\nu(x,t)dx\right)<\infty,
-\end{equation}
-\begin{equation}\label{L1_deriv}
-\mathbb{E}\left(\int_0^t\int_\mathbb{R}\nu(x,s)\Big|h(x)\Big|\left|m(\nu,x)f(x)+\frac{\mu}{2}\Delta f(x)\right|dxds\right)<\infty, \ \forall f\in C^2_b(\mathbb{R}).
-\end{equation}
-
-Put $H(t)=\int_\mathbb{R}h(x)\nu(x,t)dx$. Then, for non-random and non-negative $\nu_1(x,t)$ that is continuous in both arguments and integrable in $x$,
-\begin{equation}
-\star:=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{E}[H(t+\varepsilon)-H(t)|\nu(x,t)=\nu_1(x,t)]
-\end{equation}
-\begin{equation}
-=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{E}\left[\int_\mathbb{R}h(x)\big(\nu(x,t+\varepsilon)-\nu(x,t)\big)dx\Big|\nu(x,t)=\nu_1(x,t)\right]
-\end{equation}
-\begin{equation}
-=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{E}\left[\int_\mathbb{R}\int_t^{t+\varepsilon}f(x)\left(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)
-+\sqrt{V\nu(x,s)}\dot W(x,s)\right)dsdx\Big|\nu(x,t)=\nu_1(x,t)\right].
-\end{equation}
-
-By assumption (\ref{L1_deriv}) we can use Fubini's theorem to write, with probability one,
-
-\begin{equation}
-\int_\mathbb{R}\int_t^{t+\varepsilon}h(x)\big(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)dsdx=\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\big(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)dxds.
-\end{equation}
-
-By assumption (\ref{L2_nu}) we have $\varphi(x,t)=h(x)\sqrt{V\nu(x,t)}$ implies $\varphi\in L^2_c(\mathbb{R}\times[0,\infty))$. Hence, the following is true by definition;
-
-\begin{equation}
-\int_\mathbb{R}\int_t^{t+\varepsilon}h(x)\sqrt{V\nu(x,s)}\dot W(x,s)dsdx = 
-\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\sqrt{V\nu(x,s)}\dot W(x,s)dxds.
-\end{equation}
-
-Hence,
-
-\begin{multline}
-\mathbb{E}\left[\int_\mathbb{R}\int_t^{t+\varepsilon}h(x)\sqrt{V\nu(x,s)}\dot W(x,s)dsdx\Big|\nu(x,t)=\nu_1(x,t)\right] \\ =\mathbb{E}\left[\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\sqrt{V\nu(x,s)}\dot W(x,s)dxds\Big|\nu(x,t)=\nu_1(x,t)\right]=0
-\end{multline}
-
-and
-
-\begin{equation}
-\star=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{E}\left[\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\left(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)\right)dxds\Big|\nu(x,t)=\nu_1(x,t)\right].
-\end{equation}
-
-By assumption (\ref{L1_deriv}) we know that there exists a $\delta>0$ such that for each positive $\varepsilon<\delta$ the following holds almost surely:
-
-\begin{multline}
-\left|\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\left(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)\right)dxds\right| \\
-\leq\int_t^{t+\delta}\int_\mathbb{R}\Big|h(x)\Big|\left|m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)\right|dxds<\infty.
-\end{multline}
-
-Thus, by Lebesgue's dominated convergence theorem, the drift component of the process $H(t)$ can be computed as
-
-\begin{multline}
-\star=\mathbb{E}\left[\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\left(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)\right)dxds\Big|\nu(x,t)=\nu_1(x,t)\right]\\
-=\int_\mathbb{R}h(x)\big(m(\nu_1,x)\nu_1(x,t)+\frac{\mu}{2}\Delta\nu_1(x,t)\big)dx.
-\end{multline}
-
-To find an expression for the diffusion component of $H(t)$ set
-
-\begin{multline}\label{star2}
-\star\star:=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{V}[H(t+\varepsilon)-H(t)|\nu(x,t)=\nu_1(x,t)] \\
-=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{V}\left[\int_\mathbb{R}h(x)\big(\nu(x,t+h)-\nu(x,t)\big)dx\Big|\nu(x,t)=\nu_1(x,t)\right].
-\end{multline}
-
-We can rewrite the integral inside expression (\ref{star2}) as
-
-\begin{equation}
-\int_\mathbb{R}\int_t^{t+\varepsilon}h(x)\big(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)+\sqrt{V\nu(x,s)}\dot W(x,s)\big)dsdx.
-\end{equation}
-
-We have already found
-
-\begin{multline}
-\mathbb{E}\left[\int_\mathbb{R}\int_t^{t+\varepsilon}h(x)\big(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)+\sqrt{V\nu(x,s)}\dot W(x,s)\big)dsdx\Big|\nu(x,t)=\nu_1(x,t)\right] \\
-=\mathbb{E}\left[\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\big(m(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)\big)dxds\Big|\nu(x,t)=\nu_1(x,t)\right].
-\end{multline}
-
-Then, since $h^2(x)\nu(x,t)$ is integrable by assumption (\ref{L2_nu}), we have
-
-\begin{multline}
-\star\star=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{E}\left[\left(\int_t^{t+\varepsilon}\int_\mathbb{R}h(x)\sqrt{V\nu(x,s)}\dot W(x,s)dsdx\right)^2\Big| \ \nu(x,t)=\nu_1(x,t)\right] \\
-=\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\mathbb{E}\left[\int_t^{t+\varepsilon}\int_\mathbb{R}V h^2(x)\nu(x,s)dxds\Big|\nu(x,t)=\nu_1(x,t)\right].
-\end{multline}
-
-Thus, for any $\delta\geq\varepsilon\geq0$, we have, with probability one,
-
-\begin{equation}
-\int_t^{t+\varepsilon}\int_\mathbb{R}V h^2(x)\nu(x,s)dxds\leq\int_t^{t+\delta}\int_\mathbb{R}V h^2(x)\nu(x,s)dxds.
-\end{equation}
-
-We can therefore use Lebesgue's dominated convergence theorem to justify
-
-\begin{equation}
-\star\star=\mathbb{E}\left[\lim_{\varepsilon\to0}\frac{1}{\varepsilon}\int_t^{t+\varepsilon}\int_\mathbb{R}V h^2(x)\nu(x,s)dxds\Big|\nu(x,t)=\nu_1(x,t)\right]=\int_\mathbb{R}V h^2(x)\nu_1(x,t)dx.
-\end{equation}
-
-Then, using the notation of stochastic differentials, we have
-
-\begin{equation}\label{SDE_deriv}
-dH(t)=\left(\int_\mathbb{R}h(x)\left(m(\nu,x)\nu(x,t)+\frac{\mu}{2}\Delta\nu(x,t)\right)dx\right)dt+\sqrt{V\int_\mathbb{R}h^2(x)\nu(x,t)dx} \ dW(t)
-\end{equation}
-
-where $W$ is a standard Brownian motion. In the following subsections we employ this formula under the cases $h(x)=x,x^2$ to obtain SDE for the phenotypic mean and variance.
-
--->
-
-### Derivation for trait mean
-
-We make use of the notation
-
-\begin{equation}
-\left\{\begin{matrix}
-\|N\|_2 = \sqrt{V\int_\mathbb{R}\nu(x,t)dx}= \sqrt{VN} \\ \\
-\|\tilde x\|_2 = \sqrt{V\int_\mathbb{R}x^2\nu(x,t)dx} \\ \\
-\langle\tilde x,N\rangle = V\int_\mathbb{R}x\nu(x,t)dx=\bar x VN.
-\end{matrix}\right.
-\end{equation}
-
-Rewriting formula (\ref{xtilde}) as an SDE provides
-
-\begin{equation}
-d\tilde x=\left(\overline{xm}N+\frac{\mu}{2}\int_\mathbb{R}x\Delta\nu(x,t)dx\right)dt+\|\tilde x\|_2d\tilde W_2,
-\end{equation}
-
-where,
-
-\begin{equation}
-d\tilde{ W}_2=d\hat{\mathbf W}_t(\sqrt{V x^2\nu})=\frac{1}{\|\tilde x\|_2}\int_\mathbb{R}x\sqrt{V\nu(x,t)}\dot W(x,t)dxdt.
-\end{equation}
-
-Using It&ocirc;'s quotient rule on $\bar x=\tilde x/N$, we obtain
-
-\begin{equation}
-d\bar x = d\left(\frac{\tilde x}{N}\right)=\frac{\tilde x}{N}\left(\frac{d\tilde x}{\tilde x}-\frac{dN}{N}-\frac{d\tilde x}{\tilde x}\frac{dN}{N}+\Big(\frac{dN}{N}\Big)^2\right)
-= \frac{d\tilde x}{N}-\bar x\frac{dN}{N}-\frac{d\tilde x}{N}\frac{dN}{N}+\bar x\Big(\frac{dN}{N}\Big)^2.
-\end{equation}
-
-From Table \ref{mult} of the main text $d\tilde xdN=\langle\tilde x,N\rangle$ and $dN^2=\|N\|_2^2$. Hence,
-
-\begin{equation}
-d\bar x= \overline{xm}dt+\frac{\|\tilde x\|_2}{N}d\tilde W_2-\bar x\left(\bar m dt+\sqrt\frac{V}{N}dW_1\right)-\frac{\langle\tilde x,N\rangle}{N^2}dt+\bar x\frac{\|N\|_2^2}{N^2}dt
-\end{equation}
-
-\begin{equation*}
-= (\overline{xm}-\bar x\bar m)dt+\frac{\|\tilde x\|_2}{N}d\tilde W_2
--\bar x\sqrt{\frac{V}{N}}d W_1-V\frac{\bar x}{N}dt+V\frac{\bar x}{N}dt
-\end{equation*}
-
-\begin{equation*}
-=\mathrm{Cov}_t(x,m)
-+\frac{\|\tilde x\|_2}{N}d\tilde W_2-\bar x\sqrt{\frac{V}{N}}d W_1.
-\end{equation*}
-
-Note that
-
-\begin{equation}
-\frac{\|\tilde x\|_2}{N}d\tilde W_2-\bar x\sqrt{\frac{V}{N}}d W_1=\frac{1}{N}\int_\mathbb{R}x\sqrt{V\nu(x,t)}\dot W(x,t)dx-\frac{\bar x}{N}\int_\mathbb{R}\sqrt{V\nu(x,t)}\dot W(x,t)dx
-\end{equation}
-
-\begin{equation*}
-=\int_\mathbb{R}\frac{x-\bar x}{N}\sqrt{V\nu(x,t)}\dot W(x,t)dx
-\end{equation*}
-
-and
-
-\begin{equation}
-\mathbb{V}\left(\int_\mathbb{R}\frac{x-\bar x}{N}\sqrt{V\nu(x,t)}\dot W(x,t)dx\right)=\frac{V}{N}\int_\mathbb{R}(x-\bar x)^2p(x,t)dx=V\frac{\sigma^2}{N}.
-\end{equation}
-
-Hence, by setting
-
-\begin{equation}
-d W_2=\frac{\int_\mathbb{R}\frac{(x-\bar x)}{N}\sqrt{V\nu(x,t)}\dot W(x,t)dx}{\sqrt{V\sigma^2/N}}
-\end{equation}
-
-we can write
-
-\begin{equation}
-d\bar x=\mathrm{Cov}_t(x,m)dt
-+\sqrt{V\frac{\sigma^2}{N}}d W_2.
-\end{equation}
-
-### Derivation for trait variance
-
-We make use of the notation
-
-\begin{equation}
-\left\{\begin{matrix}
-\|\tilde{\tilde\sigma}^2\|_2 = \sqrt{V\int_\mathbb{R}x^4\nu(x,t)dx} \\ \\
-\langle\tilde{\tilde\sigma}^2,N\rangle=V\int_\mathbb{R}x^2\nu(x,t)dx=\overline{x^2}VN.
-\end{matrix}\right.
-\end{equation}
-
-Applying formula (\ref{stilde}) provides
-
-\begin{equation}
-d\tilde{\tilde\sigma}^2=\left(\overline{x^2m}N+\mu N\right)dt+\|\tilde{\tilde\sigma}^2\|_2d\tilde W_3
-\end{equation}
-
-where
-
-\begin{equation}
-d\tilde{W}_3=d\hat{\mathbf W}_t(\sqrt{V x^4\nu})
-=\frac{1}{\|\tilde{\tilde\sigma}^2\|_2}\int_\mathbb{R}x^2\sqrt{V\nu(x,t)}\dot W(x,t)dx.
-\end{equation}
-
-Using It&ocirc;'s quotient rule on $\overline{x^2}=\tilde{\tilde\sigma}^2/N$, we obtain
-
-\begin{equation}
-d\overline{x^2} = d\left(\frac{\tilde{\tilde\sigma}^2}{N}\right)=\frac{\tilde{\tilde\sigma}^2}{N}\left(\frac{d\tilde{\tilde\sigma}^2}{\tilde{\tilde\sigma}^2}-\frac{dN}{N}-\frac{d\tilde{\tilde\sigma}^2}{\tilde{\tilde\sigma}^2}\frac{dN}{N}+\Big(\frac{dN}{N}\Big)^2\right)
-=\frac{d\tilde{\tilde\sigma}^2}{N}-\overline{x^2}\frac{dN}{N}-\frac{d\tilde{\tilde\sigma}^2}{N}\frac{dN}{N}+\overline{x^2}\left(\frac{dN}{N}\right)^2.
-\end{equation}
-
-Table \ref{mult} of the main text implies $d\tilde W_3d W_1=\langle\tilde{\tilde\sigma}^2,N\rangle$ and hence
-
-\begin{multline}
-d\overline{x^2}=\left(\overline{x^2m} + \mu \right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3
--\overline{x^2}\left(\bar m dt+\sqrt\frac{V}{N}dW_1\right)-\frac{\langle\tilde{\tilde\sigma}^2,N\rangle}{N^2}dt+\overline{x^2}\frac{\|N\|_2^2}{N^2}dt \\ \\
-=\left(\overline{x^2m}-\overline{x^2}\bar mdt + \mu\right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3
--\overline{x^2}\sqrt{\frac{V}{N}}d W_1
--\overline{x^2}\frac{V}{N}dt+\overline{x^2}\frac{V}{N}dt \\ \\
-=\left(\mathrm{Cov}_t\Big(x^2,m\Big) + \mu \right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3-\overline{x^2}\sqrt\frac{V}{N}d W_1.
-\end{multline}
-
-Setting $F(y,z)=y-z^2$, use It&ocirc;'s formula on $\sigma^2=F(\overline{x^2},\bar x)=\overline{x^2}-\bar x^2$ to obtain:
-
-\begin{multline}
-d\sigma^2=d\overline{x^2}-2\bar xd\bar x-(d\bar x)^2
-=\left(\mathrm{Cov}_t\Big(x^2,m\Big) + \mu \right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3-\overline{x^2}\sqrt\frac{V}{N}d W_1 \\ \\
--2\bar x\left(\mathrm{Cov}_t(x,m)+\mu dt+\sqrt{\frac{V\sigma^2}{N}}d W_2\right)-\left(\mathrm{Cov}_t(x,m)dt+\mu dt+\sqrt{\frac{V\sigma^2}{N}}d W_2\right)^2 \\ \\
-=\left(\mathrm{Cov}_t\Big(x^2-2\bar xx,m\Big) + \mu \right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3-\overline{x^2}\sqrt\frac{V}{N}d W_1-2\bar x\sqrt{\frac{V\sigma^2}{N}}d W_2-\left(\frac{V\sigma^2}{N}\right)dt \\ \\
-=\left(\mathrm{Cov}_t\Big(x-\bar x)^2,m\Big) + \mu - \frac{V\sigma^2}{N}\right)dt+\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3-\overline{x^2}\sqrt\frac{V}{N}d W_1-2\bar x\sqrt{\frac{V\sigma^2}{N}}d W_2.
-\end{multline}
-
-In light of
-
-\begin{multline}
-\frac{\|\tilde{\tilde\sigma}^2\|_2}{N}d\tilde W_3-\overline{x^2}\sqrt\frac{V}{N}d W_1-2\bar x\sqrt{\frac{V\sigma^2}{N}}d W_2=\frac{1}{N}\int_\mathbb{R}\left(x^2-\tilde\sigma^2-2\bar x(x-\bar x)\right)\sqrt{V\nu(x,t)}\dot W(x,t)dx \\
-=\frac{1}{N}\int_\mathbb{R}\left((x-\bar x)^2-\sigma^2\right)\sqrt{V\nu(x,t)}\dot W(x,t)dx
-\end{multline}
-
-and
-
-\begin{multline}
-\frac{1}{N}\int_\mathbb{R}\Big(\left((x-\bar x)^2-\sigma^2\right)\sqrt{V\nu(x,s)}\Big)^2dx
-=\frac{V}{N}\left(\int_\mathbb{R}((x-\bar x)^4-2(x-\bar x)^2\sigma^2+\sigma^4)p(x,t)dx\right) \\
-=\frac{V}{N}\left(\overline{(x-\bar x)^4}-\sigma^4\right)
-\end{multline}
-
-we set
-
-\begin{equation}
-d W_3=\frac{\int_\mathbb{R}\left((x-\bar x)^2-\sigma^2\right)\sqrt{V\nu(x,t)}\dot W(x,t)dx}{V\left(\overline{(x-\bar x)^4}-\sigma^4\right)}
-\end{equation}
-
-so that
-
-\begin{equation}
-d\sigma^2=\mathrm{Cov}_t\Big((x-\bar x)^2,m\Big)dt+\left(\mu-V\frac{\sigma^2}{N}\right)dt+\sqrt{V\frac{\overline{(x-\bar x)^4}-\sigma^4}{N}}d W_3.
-\end{equation}
-
-Table \ref{mult} of the main text implies
-
-\begin{equation}
-d W_1d W_2=\frac{\int_\mathbb{R}(x-\bar x)\nu(x,t)dx}{\sqrt{N\sigma^2}}dt=0,
-\end{equation}
-
-\begin{equation}
-d W_1d W_3=\frac{\int_\mathbb{R}\big((x-\bar x)^2-\sigma^2\big)\nu(x,t)dx}{\sqrt{\overline{(x-\bar x)^4}-\sigma^4}}dt=0,
-\end{equation}
-
-\begin{equation}
-d W_2d W_3=\frac{\int_\mathbb{R}(x-\bar x)\big((x-\bar x)^2-\sigma^2\big)p(x,t)dx}{\sqrt{\sigma^2\big(\overline{(x-\bar x)^4}-\sigma^4\big)}}dt
-=\frac{N\overline{(x-\bar x)^3}}{\sqrt{\sigma^2\big(\overline{(x-\bar x)^4}-\sigma^4\big)}}dt.
-\end{equation}
-
-In particular, when $p$ is a Gaussian curve $d W_2d W_3=0$.
-
-## Relating fitness of expressed traits to fitness of breeding values \label{fit2fit}
-
-Following \S\ref{inheritance} of the main text, we have $\sigma^2=G+\eta$ and
-
-\begin{equation}
-m^*(\rho,g)=\int_\mathbb{R}m(\nu,x)\psi(x,g)dx.
-\end{equation}
-
-Hence,
-
-\begin{multline}
-\overline{\frac{\partial m^*}{\partial\bar x}}=\int_{\mathbb{R}}\frac{\rho(g,t)}{N(t)}\frac{\partial}{\partial\bar x}\int_\mathbb{R}m(\nu,x)\psi(x,g)dxdg= \\
-\int_{\mathbb{R}}\int_\mathbb{R}\frac{\rho(g,t)}{N(t)}\psi(x,g)dg\frac{\partial}{\partial\bar x}m(\nu,x)dx=\int_\mathbb{R}p(x,t)\frac{\partial}{\partial\bar x}m(\nu,x)dx=\overline{\frac{\partial m}{\partial\bar x}}
-\end{multline}
-and
-
-\begin{multline}
-\overline{\frac{\partial m^*}{\partial G}}=\int_{\mathbb{R}}\frac{\rho(g,t)}{N(t)}\frac{\partial}{\partial G}\int_\mathbb{R}m(\nu,x)\psi(x,g)dxdg=\int_{\mathbb{R}}\int_\mathbb{R}\frac{\rho(g,t)}{N(t)}\psi(x,g)dg\frac{\partial m}{\partial G}dx= \\
-\int_\mathbb{R}p(x,t)\frac{\partial m}{\partial\sigma^2}\frac{\partial\sigma^2}{\partial G}dx=\overline{\frac{\partial m}{\partial\sigma^2}}.
-\end{multline}
-
-## Derivation of diffuse coevolution model \label{diffuse}
-
-In this section we provide a derivation of our model of diffuse coevolution driven by competition.
-Since most of the work in this derivation has already been completed in Supplementary Material \S\ref{SDE_DERIV}, we focus here on deriving the Malthusian fitness $m$ as a function of trait value $x$.
-We begin with discrete populations of individuals.
-In particular, we begin by assuming population size $n_i$ is an integer for each species $i=1,\dots,S$ before passing to the large population size limit.
-
-The reduction in fitness for an individual of species $i$ caused by competition is captured multiplicatively by $0<C_i\leq1$.
-Biologically this assumes all competitors affect individuals of a given species equally by consuming the same amount of resources.
-This is a mean-field interaction since any individual that consumes resources has an effect on the fitness of all other individuals competing for the same resources.
-Denote by $x_{ij}$ the trait value of the $j$-th individual belonging to species $i$.
-The set of trait values across all individuals in the community at time $t\geq0$ is written $X=\{x_{ij}\}$.
-We denote by $\mathscr{B}_{ij}$ a function that maps $X$ to the cumulative effect of all competitive interactions on the fitness of the $j$-th individual in species $i$.
-Since individuals do not compete with themselves the net multiplicative effects on fitness of both interspecific and intraspecific competition on the $j$-th individual in species $i$ can be summarized by
-
-\begin{equation}
-\mathscr B_{ij}(X)=C_i^{\sum_{l\neq j}\mathcal{O}_{ii}(x_{ij}-x_{il})+\sum_{k\neq i}\sum_{l=1}^{n_k}\mathcal{O}_{ik}(x_{ij}-x_{kl})},
-\end{equation}
-
-where $\mathcal{O}_{ij}$, defined in the main text, measures the overlap in resource use between individuals of species $i$ and $j$ as a function of their niche-centers.
-Writing $\mathscr W_{ij}(X)$ as the average number of offspring left by the $j$-th individual of species $i$, we have
-
-\begin{equation}
-\mathscr W_{ij}(X)=\mathscr{A}_i(x_{ij})\mathscr{B}_{ij}(X),
-\end{equation}
-
-where $\mathscr{A}_i(x)=\int_\mathbb{R}e_i(\zeta)u_i(\zeta,x)d\zeta$ accounts for abiotic selection and $e_i$ has been defined in the main text.
-
-We now turn to a diffusion limit.
-Since we have more than one population, we take the diffusion limit for each population one at a time starting with population $1$.
-We write $\pmb n=(n_1,\dots,n_S)$.
-Following Méléard and Roelly [-@Mlard1993; -@meleard1992interacting] we rescale generation time and individual mass to $\frac{1}{n_1}$ and mean of the reproductive law to
-
-\begin{equation}
-\mathscr W_{1j}^{(\pmb n)}(X)=\mathscr{A}_1(x_{1j})^{1/n_1}\exp\left(\frac{\ln C_1}{n_1^2}\sum_{l\neq j}\mathcal{O}_{11}(x_{1j}-x_{1l})+\frac{\ln C_1}{n_1}\sum_{k\neq i}\frac{1}{n_k}\sum_{l=1}^{n_k}\mathcal{O}_{1k}(x_{1j}-x_{kl})\right).
-\end{equation}
-
-For large $n_1$, we have the approximation
-
-\begin{equation}
-\mathscr W_{1j}^{(\pmb n)}(X)\approx\mathscr{A}_1(x_{1j})^{1/n_1}\left(1+\frac{\ln C_1}{n_1^2}\sum_{l\neq j}\mathcal{O}_{11}(x_{1j}-x_{1l})+\frac{\ln C_1}{n_1}\sum_{k\neq 1}\frac{1}{n_k}\sum_{l=1}^{n_k}\mathcal{O}_{1k}(x_{1j}-x_{kl})\right).
-\end{equation}
-Hence
-\begin{equation}
-\lim_{n_1\to\infty}n_1\left(\mathscr W_{1j}^{(\pmb n)}(X)-1\right)=\ln\mathscr{A}_1(x_{1j})+\left(\int_\mathbb{R}\mathcal{O}_{11}(x_{1j}-y)\nu_1(y,t)dy+\sum_{k\neq 1}\frac{1}{n_k}\sum_{l=1}^{n_k}\mathcal{O}_{1k}(x_{1j}-x_{kl})\right)\ln C_1.
-\end{equation}
-
-We write $\lim_{\pmb n\to\pmb\infty}$ for the iterated limit $\lim_{n_S\to\infty}\cdots\lim_{n_1\to\infty}$ and, assuming  $\nu_i(\cdot,t)\in C_1^+(\mathbb{R})$ for $i=1,\dots,S$ and $t\in[0,\infty)$, we set $\pmb\nu=(\nu_1,\dots,\nu_S)$.
-Then, for any $\pmb\nu$, the the diffusion limits for the remaining populations provides the Malthusian parameter for individuals in species $i$ with trait value $x_{1j}$ as
-
-\begin{equation}
-m_1(\pmb\nu,x_{1j}):=\lim_{\pmb n\to\pmb\infty}n_1\left(\mathscr W_{1j}^{(\pmb n)}(X)-1\right)=\ln\mathscr{A}_1(x)+\left(\sum_{k=1}^S\int_\mathbb{R}\mathcal{O}_{1k}(x_{1j}-y)\nu_k(y,t)dy\right)\ln C_1.
-\end{equation}
-
-We compute the average niche overlap of an individual in species $i$ with nich location $x$ across all individuals in species $j$ as
-
-\begin{equation}
-\bar{\mathcal{O}}_{ij}(x,t)=\frac{\int_\mathbb{R}\mathcal{O}_{ij}(x-y)\nu_j(y,t)dy}{\int_\mathbb{R}\nu_j(y,t)dy}.
-\end{equation}
-
-We now assume the resource utilization curves $u_i(\zeta)$ and phenotypic densities $\nu_i(x,t)$ are Gaussian curves for $i=1,\dots,S$. In this case $\bar{\mathcal{O}}_{ij}(x,t)$ simplifies to
-
-\begin{equation}
-\bar{\mathcal{O}}_{ij}(x,t)=\frac{\int_\mathbb{R}\mathcal O_{ij}(x-y)\nu_j(y,t)dy}{\int_\mathbb{R}\nu_j(y,t)dy}
-=\frac{U_iU_j}{\sqrt{2\pi(w_i+w_j+{\sigma_j}^2(t))}}\exp\left(-\frac{(x-\bar x_j(t))^2}{2(w_i+w_j+{\sigma_j}^2(t))}\right).
-\end{equation}
-
-Setting
-
-\begin{subequations}
-\begin{align}
-\sigma_i^2(t) = & \ G_i(t)+\eta_i, \\
-R_i = & \ \ln \left(\frac{Q_iU_i}{\sqrt{1+A_iw_i}}\right), \\
-a_i = & \ \frac{A_i}{1+A_iw_i}, \\
-\tilde b_{ij}(t) = & \ \frac{1}{w_i+w_j+\sigma_j^2(t)}, \\
-c_i = & -\ln C_i,
-\end{align}
-\end{subequations}
-
-we get
-
-\begin{equation}\label{malth_fit}
-m_i(\pmb\nu,x)=R_i-\frac{a_i}{2}(x-\theta_i)^2-c_i\sum_{j=1}^SN_j(t)U_iU_j\sqrt{\frac{\tilde b_{ij}(t)}{2\pi}}e^{-\frac{\tilde b_{ij}(t)}{2}(x-\bar x_j(t))^2}.
-\end{equation}
-
-Hence, our fitness function is bounded above and decreases linearly with total abundance, as required by the assumptions of Champagnat, Ferrière and Méléard [-@Champagnat2006].
-For the remainder of the derivation we suppress notation indicating dependency on $\pmb\nu$, $x$ and $t$.
-From (\ref{malth_fit}) we calculate
-
-\begin{equation}
-\frac{\partial m_i}{\partial\bar x_i}=c_iN_iU_i^2\tilde b_{ii}(x-\bar x_i)\sqrt{\frac{\tilde b_{ii}}{2\pi}}e^{-\frac{\tilde b_{ii}}{2}(x-\bar x_i)^2}
-\end{equation}
-
-\begin{multline}
-\frac{\partial m_i}{\partial G_i}=\frac{c_iN_iU_i^2}{2}\left(\frac{(x-\bar x_i)^2-G_i-\eta_i-2w_i}{(G_i+\eta_i+2w_i)^2}\right)\sqrt{\frac{\tilde b_{ii}}{2\pi}}e^{-\frac{\tilde b_{ii}}{2}(x-\bar x_i)^2} \\
-=\frac{c_iN_iU_i^2{\tilde b_{ii}}^2}{2}\left((x-\bar x_i)^2-{\sigma_i}^2-2w_i\right)\sqrt{\frac{\tilde b_{ii}}{2\pi}}e^{-\frac{\tilde b_{ii}}{2}(x-\bar x_i)^2}.
-\end{multline}
-
-Note that
-
-\begin{multline}
-\sqrt\frac{\tilde b_{ii}}{2\pi}\exp\left(-\frac{\tilde b_{ii}}{2}(x-\bar x_i)^2\right)\sqrt\frac{1}{2\pi{\sigma_i}^2}\exp\left(-\frac{(x-\bar x_i)^2}{2{\sigma_i}^2}\right) \\
-=\sqrt\frac{1}{2\pi({\sigma_i}^2+1/\tilde b_{ii})}\sqrt\frac{{\sigma_i}^2+1/\tilde b_{ii}}{2\pi{\sigma_i}^2/\tilde b_{ii}}\exp\left(-\frac{{\sigma_i}^2+1/\tilde b_{ii}}{2{\sigma_i}^2/\tilde b_{ii}}(x-\bar x_i)^2\right) \\
-=\sqrt{\frac{1}{4\pi({\sigma_i}^2+w_i)}}\sqrt\frac{2({\sigma_i}^2+w_i)}{2\pi{\sigma_i}^2({\sigma_i}^2+2w_i)}\exp\left(-\frac{{\sigma_i}^2({\sigma_i}^2+2w_i)}{4({\sigma_i}^2+w_i)}(x-\bar x_i)^2\right).
-\end{multline}
-
-Hence,
-
-\begin{equation}
-\overline{\frac{\partial m_i}{\partial\bar x_i}}=0,
-\end{equation}
-
-\begin{multline}
-\overline{\frac{\partial m_i}{\partial G_i}}=\frac{c_iN_iU_i^2}{2({\sigma_i}^2+2w_i)^2}\left(\frac{({\sigma_i}^2+2w_i){\sigma_i}^2}{2(w_i+{\sigma_i}^2)}-{\sigma_i}^2-2w_i\right)\sqrt{\frac{b_{ii}}{2\pi}} \\
-=\frac{c_iN_iU_i^2}{2({\sigma_i}^2+2w_i)}\left(\frac{{\sigma_i}^2}{2({\sigma_i}^2+w_i)}-1\right)\sqrt{\frac{b_{ii}}{2\pi}}=-\frac{c_iN_iU_i^2b_{ii}}{2}\sqrt{\frac{b_{ii}}{2\pi}},
-\end{multline}
-
-where
-
-\begin{equation}
-b_{ij}=\frac{1}{w_i+w_j+{\sigma_i}^2+{\sigma_j}^2}.
-\end{equation}
-
-The average fitness for species $i$ is
-
-\begin{equation}
-\bar m_i=R_i-\frac{a_i}{2}\Big((\bar x_i-\theta_i)^2+G_i+\eta_i\Big) - c_i\sum_{j=1}^SN_jU_iU_j\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}.
-\end{equation}
-
-Thus,
-
-\begin{equation}
-\frac{\partial\bar m_i}{\partial\bar x_i}=a_i(\theta_i-\bar x_i)-c_i\sum_jN_jU_iU_jb_{ij}(\bar x_j-\bar x_i)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2},
-\end{equation}
-
-\begin{equation}
-\frac{\partial\bar m_i}{\partial G_i}=-\frac{a_i}{2}+\frac{c_i}{2}\sum_{j=1}^SN_jU_iU_jb_{ij}\left(1-b_{ij}(\bar x_i-\bar x_j)^2\right)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}.
-\end{equation}
-
-In particular
-
-\begin{equation}
-\frac{\partial\bar m_i}{\partial G_i}-\overline{\frac{\partial m_i}{\partial G_i}}=-\frac{a_i}{2}+\frac{c_i}{2}\left(N_iU_i^2b_{ii}\sqrt\frac{b_{ii}}{2\pi}+\sum_{j=1}^SN_jU_iU_jb_{ij}\left(1-b_{ij}(\bar x_i-\bar x_j)^2\right)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\right).
-\end{equation}
-
-Applying equations (\ref{N}), (\ref{xbarfinal}) and (\ref{Gfinal}) of the main text recovers system (\ref{comm_dynamics}) of the main text.
-
-## The relation between competition coefficients and selection \label{coef_grad_moms}
-
-### Derivation of analytical approximations
-
-Just as with most caclulations in this work, the derivations are straightforward applications of Gaussian products. That is, if
-
-\begin{equation}
-f_1(x)=\frac{1}{\sqrt{2\pi\sigma_1^2}}\exp\left(-\frac{(\mu_1-x)^2}{2\sigma_1^2}\right), \ f_2(x)=\frac{1}{\sqrt{2\pi\sigma_2^2}}\exp\left(-\frac{(\mu_2-x)^2}{2\sigma_2^2}\right),
-\end{equation}
-
-then
-
-\begin{equation}
-f_1(x)f_2(x)=\frac{1}{\sqrt{2\pi(\sigma_1^2+\sigma_2^2)}}\exp\left(-\frac{(\mu_1-\mu_2)^2}{2(\sigma_1^2+\sigma_2^2)}\right)\frac{1}{\sqrt{2\pi\tilde\sigma^2}}\exp\left(-\frac{(\tilde\mu-x)^2}{2\tilde\sigma^2}\right),
-\end{equation}
-
-where
-
-\begin{equation}
-\tilde\mu=\frac{\sigma_2^2\mu_1+\sigma_1^2\mu_2}{\sigma_1^2+\sigma_2^2}, \ \tilde\sigma^2=\frac{\sigma_1^2\sigma_2^2}{\sigma_1^2+\sigma_2^2}.
-\end{equation}
-
-#### Caclulating $\mathrm{Cov}_{f_{\bar X}}(\alpha,\gamma)$
-
-Recalling
-
-\begin{equation}
-\alpha(\bar x_i,\bar x_j)=\frac{c}{\bar r}\sqrt\frac{b}{2\pi}\exp\left(-\frac{b}{2}\big(\bar x_i-\bar x_j\big)^2\right),
-\end{equation}
-
-\begin{equation}
-\gamma(\bar x_i,\bar x_j)=cNb\left(1-b\big(\bar x_i-\bar x_j\big)^2\right) \sqrt\frac{b}{2\pi}\exp\left(-\frac{b}{2}\big(\bar x_i-\bar x_j\big)^2\right),
-\end{equation}
-
-we have
-
-\begin{multline}
-\bar\alpha=
-\int_\mathbb{R}\int_\mathbb{R}\alpha(\bar x_i,\bar x_j)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
-=\frac{c}{\bar r}\int_\mathbb{R}\frac{1}{\sqrt{2\pi(b^{-1}+V_{\bar X})}}\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(b^{-1}+V_{\bar X})}\right)f_{\bar X}(\bar x_j)d\bar x_j
-=\frac{c/\bar r}{\sqrt{2\pi(b^{-1}+2V_{\bar X})}},
-\end{multline}
-
-\begin{multline}
-\bar\gamma=
-\int_\mathbb{R}\int_\mathbb{R}\gamma(\bar x_i,\bar x_j)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
-= cNb\int_\mathbb{R}\left\{1-\left[\left(\frac{\bar{\bar x}+bV_{\bar X}\bar x_j}{1+bV_{\bar X}}-\bar x_j\right)^2+\frac{V_{\bar X}}{1+bV_{\bar X}}\right]\right\}\frac{1}{\sqrt{2\pi(b^{-1}+V_{\bar X})}}\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(b^{-1}+V_{\bar X})}\right)f_{\bar X}(\bar x_j)d\bar x_j \\
-= cNb\int_\mathbb{R}\left\{1-\left[\left(\frac{\bar{\bar x}-\bar x_j}{1+bV_{\bar X}}\right)^2+\frac{V_{\bar X}}{1+bV_{\bar X}}\right]\right\}\frac{1}{\sqrt{2\pi(b^{-1}+V_{\bar X})}}\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(b^{-1}+V_{\bar X})}\right)f_{\bar X}(\bar x_j)d\bar x_j \\
-= cNb\left(1-\frac{(1+bV_{\bar X})V_{\bar X}}{1+2bV_{\bar X}}\frac{1}{(1+bV_{\bar X})^2}-\frac{V_{\bar X}}{1+bV_{\bar X}}\right)\frac{1}{\sqrt{2\pi(b^{-1}+2V_{\bar X})}} \\
-= cNb\left[1-\left(\frac{1}{1+2bV_{\bar X}}+1\right)\frac{V_{\bar X}}{1+bV_{\bar X}}\right]\frac{1}{\sqrt{2\pi(b^{-1}+2V_{\bar X})}} \\
-= cNb\left(1-\frac{2V_{\bar X}}{1+2bV_{\bar X}}\right)\sqrt\frac{b}{2\pi(1+2bV_{\bar X})},
-\end{multline}
-
-\begin{multline}
-\mathrm{Var}_{f_{\bar X}}(\alpha)=
-\int_\mathbb{R}\int_\mathbb{R}\big(\bar\alpha-\alpha(\bar x_i,\bar x_j)\big)^2f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
-=\frac{c^2}{\bar r^2}\left(\sqrt\frac{b}{4\pi}\int_\mathbb{R}\int_\mathbb{R}\sqrt\frac{b}{\pi}\exp\left(-b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j-\frac{1}{2\pi(b^{-1}+2V_{\bar X})}\right) \\
-=\frac{c^2}{\bar r^2}\left(\sqrt\frac{b}{4\pi}\int_\mathbb{R}\sqrt\frac{1}{2\pi(\frac{1}{2b}+V_{\bar X})}\exp\left(-b(\bar{\bar x}-\bar x_j)^2\right)f_{\bar X}(\bar x_j)d\bar x_j-\frac{1}{2\pi(b^{-1}+2V_{\bar X})}\right) \\
-=\frac{c^2}{\bar r^2}\left(\sqrt\frac{b}{4\pi}\sqrt\frac{1}{2\pi(\frac{1}{2b}+2V_{\bar X})}-\frac{1}{2\pi(b^{-1}+2V_{\bar X})}\right)=\frac{c^2b}{2\pi\bar r^2}\left(\frac{1}{\sqrt{1+4bV_{\bar X}}}-\frac{1}{1+2bV_{\bar X}}\right),
-\end{multline}
-
-\begin{multline}
-\mathrm{Cov}_{f_{\bar X}}(\alpha,\gamma)=
-\int_\mathbb{R}\int_\mathbb{R}\big(\bar\alpha-\alpha(\bar x_i,\bar x_j)\big)\big(\bar\gamma-\gamma(\bar x_i,\bar x_j)\big)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
-=\frac{c^2Nb}{2\bar r}\sqrt\frac{b}{\pi}\int_\mathbb{R}\int_\mathbb{R}\big(1-b(\bar x_i-\bar x_j)^2\big)\sqrt\frac{b}{\pi}\exp\left(-b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j-\bar\alpha\bar\gamma \\
-=\frac{c^2Nb}{2\bar r}\sqrt\frac{b}{\pi}\frac{1-2bV_{\bar X}}{\sqrt{2\pi((2b)^{-1}+2V_{\bar X})}}-\frac{c^2Nb}{\bar r}\frac{1-2bV_{\bar X}}{2\pi(b^{-1}+2V_{\bar X})} \\
-=\frac{c^2b^2N}{2\pi\bar r}(1-2bV_{\bar X})\left(\frac{1}{\sqrt{1+4bV_{\bar X}}}-\frac{1}{1+2bV_{\bar X}}\right).
-\end{multline}
-
-#### Caclulating $\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)$
-
-To calculate moments of $|\beta|$ we note that, as a random variable, $|\beta|$ takes a folded normal distribution.
-Setting $\Phi(x)$ equal to the cumulative density function of the standard normal distribution and using the properties of the folded normal distribution, we find
-
-\begin{equation}
-\overline{|\beta|}=\sqrt\frac{2\mathrm{Var}_{f_{\bar X}}(\beta)}{\pi}\exp\left(-\frac{\bar\beta^2}{2\mathrm{Var}_{f_{\bar X}}(\beta)}\right)-\bar\beta\left[1-2\Phi\left(\frac{\bar\beta}{\mathrm{Var}_{f_{\bar X}}(\beta)}\right)\right]
-\end{equation}
-
-\begin{equation}
-\mathrm{Var}_{f_{\bar X}}(|\beta|)=\bar\beta^2+\mathrm{Var}_{f_{\bar X}}(\beta)-\overline{|\beta|}^2.
-\end{equation}
-
-Recall that
-
-\begin{equation}
-\beta(\bar x_i,\bar x_j)=cNb\big(\bar x_i-\bar x_j\big)\sqrt\frac{b}{2\pi}\exp\left(-\frac{b}{2}\big(\bar x_i-\bar x_j\big)^2\right)
-\end{equation}
-
-and hence
-
-\begin{multline}
-\bar\beta=
-\int_\mathbb{R}\int_\mathbb{R}\beta(\bar x_i,\bar x_j)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
-= cNb\int_\mathbb{R}(\bar{\bar x}-\bar x_j)\frac{1}{\sqrt{2\pi(b^{-1}+V_{\bar X})}}\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(b^{-1}+V_{\bar X})}\right)f_{\bar X}(\bar x_j)d\bar x_j=0,
-\end{multline}
-
-\begin{multline}
-\mathrm{Var}_{f_{\bar X}}(\beta)=
-\int_\mathbb{R}\int_\mathbb{R}\big(\bar\beta-\beta(\bar x_i,\bar x_j)\big)^2f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
-=\int_\mathbb{R}\int_\mathbb{R}c^2N^2b^2(\bar x_i-\bar x_j)^2\frac{b}{2\pi}\exp\left(-b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
-=\sqrt{\frac{b}{4\pi}}c^2N^2b^2\int_\mathbb{R}\left[\left(\frac{\bar{\bar x}+2bV_{\bar X}\bar x_j}{1+2bV_{\bar X}}-\bar x_j\right)^2+\frac{V_{\bar X}}{1+2bV_{\bar X}}\right]\frac{\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(\frac{1}{2b}+V_{\bar X})}\right)}{\sqrt{2\pi(\frac{1}{2b}+V_{\bar X})}}f_{\bar X}(\bar x_j)d\bar x_j \\
-=\sqrt{\frac{b}{4\pi}}c^2N^2b^2\int_\mathbb{R}\left[\frac{(\bar{\bar x}-\bar x_j)^2}{(1+2bV_{\bar X})^2}+\frac{V_{\bar X}}{1+2bV_{\bar X}}\right]\frac{\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(\frac{1}{2b}+V_{\bar X})}\right)}{\sqrt{2\pi(\frac{1}{2b}+V_{\bar X})}}f_{\bar X}(\bar x_j)d\bar x_j \\
-=\sqrt{\frac{b}{4\pi}}c^2N^2b^2\left[\frac{(1+2bV_{\bar X})V_{\bar X}}{1+4bV_{\bar X}}\frac{1}{(1+2bV_{\bar X})^2}+\frac{V_{\bar X}}{1+2bV_{\bar X}}\right]\frac{1}{\sqrt{2\pi(\frac{1}{2b}+2V_{\bar X})}} \\
-=\frac{b}{\pi}\frac{c^2N^2b^2}{\sqrt{1+4bV_{\bar X}}}\frac{V_{\bar X}}{1+2bV_{\bar X}}\left(\frac{1}{1+4bV_{\bar X}}+1\right)=\frac{2c^2N^2b^3V_{\bar X}}{\pi(1+4bV_{\bar X})^{3/2}}.
-\end{multline}
-
-Thus, using properties of the folded normal distribution, we find
-
-\begin{equation}
-\overline{|\beta|}=\sqrt{\frac{2}{\pi}}\frac{cNb^{3/2}}{(1+4bV_{\bar X})^{3/4}}\sqrt\frac{2V_{\bar X}}{\pi}=\frac{2}{\pi}\frac{cNb^{3/2}}{(1+4bV_{\bar X})^{3/4}}\sqrt{V_{\bar X}},
-\end{equation}
-
-\begin{equation}
-\mathrm{Var}_{f_{\bar X}}(|\beta|)=\frac{c^2N^2b^3}{(1+4bV_{\bar X})^{3/2}}\frac{2V_{\bar X}}{\pi}\left(1-\frac{2}{\pi}\right).
-\end{equation}
-
-We also calculate
-
-\begin{multline}
-\mathrm{Cov}_{f_{\bar X}}(\alpha,\beta)=
-\int_\mathbb{R}\int_\mathbb{R}\big(\bar\alpha-\alpha(\bar x_i,\bar x_j)\big)\big(\bar\beta-\beta(\bar x_i,\bar x_j)\big)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
-=\frac{c^2Nb}{2\bar r}\sqrt\frac{b}{\pi}\int_\mathbb{R}\int_\mathbb{R}(\bar x_i-\bar x_j)\sqrt\frac{b}{\pi}\exp\left(-b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j=0.
-\end{multline}
-
-In attempt to calculate $\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)$ we find
-
-\begin{multline}
-\mathrm{Cov}f_{\bar X}(\alpha,|\beta|)=\int_\mathbb{R}\int_\mathbb{R}\alpha(\bar x_i,\bar x_j)|\beta(\bar x_i,\bar x_j)|f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j-\bar\alpha\overline{|\beta|} \\
-=\int_\mathbb{R}\int_\mathbb{R}\frac{c}{\bar r}\sqrt\frac{b}{2\pi}\exp\left(-\frac{b}{2}(\bar x_i-\bar x_j)^2\right)cNb|\bar x_i-\bar x_j|\sqrt\frac{b}{2\pi}\exp\left(-\frac{b}{2}(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j-\bar\alpha\overline{|\beta|} \\
-=\frac{c^2Nb}{\bar r}\sqrt\frac{b}{4\pi}\int_\mathbb{R}\int_\mathbb{R}|\bar x_i-\bar x_j|\sqrt\frac{b}{\pi}\exp\left(-b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j-\bar\alpha\overline{|\beta|}.
-\end{multline}
-
-Just as we used the folded normal to find $\overline{|\beta|}$ and $\mathrm{Var}_{f_{\bar X}}(|\beta|)$, we can calculate $\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)$ by considering
-
-\begin{equation}
-\int_\mathbb{R}\int_\mathbb{R}(\bar x_i-\bar x_j)\sqrt\frac{b}{\pi}\exp\left(-b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j=0
-\end{equation}
-
-and
-
-\begin{multline}
-\int_\mathbb{R}\int_\mathbb{R}(\bar x_i-\bar x_j)^2\frac{b}{\pi}\exp\left(-2b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
-=\sqrt\frac{2b}{\pi}\int_\mathbb{R}\int_\mathbb{R}(\bar x_i-\bar x_j)^2\frac{1}{\sqrt{2\pi\frac{1}{4b}}}\exp\left(-\frac{(\bar x_i-\bar x_j)^2}{2\frac{1}{4b}}\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j \\
-=\sqrt\frac{2b}{\pi}\int_\mathbb{R}\left[\left(\frac{\bar{\bar x}+4bV_{\bar X}\bar x_j}{1+4bV_{\bar X}}-\bar x_j\right)^2+\frac{V_{\bar X}}{1+4bV_{\bar X}}\right]\frac{1}{\sqrt{2\pi(\frac{1}{4b}+V_{\bar X})}}\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(\frac{1}{4b}+V_{\bar X})}\right)f_{\bar X}(\bar x_j)d\bar x_j \\
-=\sqrt\frac{2b}{\pi}\int_\mathbb{R}\left[\left(\frac{\bar{\bar x}-\bar x_j}{1+4bV_{\bar X}}\right)^2+\frac{V_{\bar X}}{1+4bV_{\bar X}}\right]\frac{1}{\sqrt{2\pi(\frac{1}{4b}+V_{\bar X})}}\exp\left(-\frac{(\bar{\bar x}-\bar x_j)^2}{2(\frac{1}{4b}+V_{\bar X})}\right)f_{\bar X}(\bar x_j)d\bar x_j \\
-=\sqrt\frac{2b}{\pi}\left[\frac{(1+4bV_{\bar X})V_{\bar X}}{1+8bV_{\bar X}}\frac{1}{(1+4bV_{\bar X})^2}+\frac{V_{\bar X}}{1+4bV_{\bar X}}\right]\frac{1}{\sqrt{2\pi(\frac{1}{4b}+2V_{\bar X})}} \\
-=\sqrt\frac{2b}{\pi}\frac{2V_{\bar X}}{1+8bV_{\bar X}}\sqrt{\frac{4b}{2\pi(1+8bV_{\bar X})}}=\frac{b}{\pi}\frac{4V_{\bar X}}{(1+8bV_{\bar X})^{3/2}}.
-\end{multline}
-
-Hence
-
-\begin{equation}
-\int_\mathbb{R}\int_\mathbb{R}|\bar x_i-\bar x_j|\sqrt\frac{b}{\pi}\exp\left(-b(\bar x_i-\bar x_j)^2\right)f_{\bar X}(\bar x_i)f_{\bar X}(\bar x_j)d\bar x_id\bar x_j=\sqrt\frac{2}{\pi}\sqrt{\frac{b}{\pi}\frac{4V_{\bar X}}{(1+8bV_{\bar X})^{3/2}}}=\frac{2}{\pi}\frac{\sqrt{2bV_{\bar X}}}{(1+8bV_{\bar X})^{3/4}}
-\end{equation}
-
-and
-
-\begin{multline}
-\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)=\frac{c^2Nb}{\bar r}\sqrt\frac{b}{4\pi}\frac{2}{\pi}\frac{\sqrt{2bV_{\bar X}}}{(1+8bV_{\bar X})^{3/4}}-\bar\alpha\overline{|\beta|} \\
-=\frac{2c^2Nb^2}{\pi\bar r(1+8bV_{\bar X})^{3/4}}\sqrt\frac{V_{\bar X}}{2\pi}-\frac{c}{\bar r}\sqrt\frac{b}{2\pi(1+2bV_{\bar X})}\frac{2}{\pi}\frac{cNb^{3/2}}{(1+4bV_{\bar X})^{3/4}}\sqrt{V_{\bar X}} \\
-=\frac{2c^2Nb^2}{\pi\bar r(1+8bV_{\bar X})^{3/4}}\sqrt\frac{V_{\bar X}}{2\pi}-\frac{2c^2Nb^2}{\pi\bar r(1+4bV_{\bar X})^{3/4}}\sqrt\frac{V_{\bar X}}{2\pi(1+2bV_{\bar X})} \\
-=\frac{2c^2Nb^2}{\pi\bar r}\sqrt\frac{V_{\bar X}}{2\pi}\left(\frac{1}{(1+8bV_{\bar X})^{3/4}}-\frac{1}{(1+4bV_{\bar X})^{3/4}(1+2bV_{\bar X})^{1/2}}\right).
-\end{multline}
-
-#### Starting the caclulation of $\mathrm{Cov}_{f_{\bar X}}(\alpha,\mathfrak{C})$ \label{caC}
-
-We have
-
-\begin{equation}
-\mathfrak{C}(\bar x_i,\bar x_j)=c^2N^2b^2\Big(|\bar x_i-\bar x_j|+\big|1-b(\bar x_i-\bar x_j)^2\big|\Big)^2\exp\left(-\frac{b}{2}(\bar x_i-\bar x_j)^2\right).
-\end{equation}
-
-Note that the random variable $\delta=\bar x_i-\bar x_j$ is a mean zero Gaussian random variable with variance $2V_{\bar X}$.
-We write the probability density function of $\delta$ as $f_{\Delta}(\delta)$.
-Substituting in $\delta$, we can write
-
-\begin{multline}
-\mathfrak{C}(\delta,0)=c^2N^2b^2\Big(|\delta|+\big|1-b\delta^2\big|\Big)^2\exp\left(-\frac{b}{2}\delta^2\right) \\
-=c^2N^2b^2\Big(\delta^2+2\big||\delta|-b|\delta|^3\big|+\big(1-b\delta^2\big)^2\Big)\exp\left(-\frac{b}{2}\delta^2\right).
-\end{multline}
-
-From this expression, we see properties of the folded normal distribution can be used to calculate several components of the intregal $\mathrm{Cov}_{f_{\bar X}}(\alpha,\mathfrak{C})$, but a major technical challenge lies in calculating
-
-\begin{equation}
-\int_{\mathbb{R}}\big||\delta|-b|\delta|^3\big|\exp\left(-\frac{b}{2}\delta^2\right)f_{\Delta}(\delta)d\delta.
-\end{equation}
-
-Instead of overcoming this challenge to find an analytical form of $\mathrm{Cov}_{f_{\bar X}}(\alpha,\mathfrak{C})$ we turn to a numerical approach outlined in the following section.
-
-### Numerical estimates for heterogeneous $N$ and $G$
-
-To verify our analytical estimates for the covariance of selection gradients and competition coefficients, equations (\ref{cov_eqns}) of the main text, and for approximating the correlation of competition coefficients and a metric of pairwise coevolution, Figure \ref{cov_fig} of the main text, we developed a numerical approach by repeatedly simulating our system (\ref{comm_dynamics}) of the main text across a set of randomly drawn model parameters.
-In particular, we ran each numerical integration with $S=100$ species for $T_1=1000.0$ units of time.
-We then continued to integrate for $T_2=1000.0$ units of time and calculated the covariances between competition coefficients and the quantities $|\beta|$, $\gamma$ and $\mathfrak{C}$ for each of the last $T_2$ time steps.
-We assume the temporal average of these covariances across the last $T_2$ units of time approximates the expectation at equilibrium.
-We repeated this approach for randomly drawn $a$ and $c$ until our sample size reached $1000$.
-We drew values for $a$ and $c$ from condition log-normal distributions.
-For $a$, we set the log-normal parameters equal to $\mu=-10$ and $\sigma=6$ and conditioned on $a<0.01$ to prevent extinction.
-For $c$ we set the log-normal parameters equal to $\mu=-4$ and $\sigma=4$ and conditioned on $c<1\times 10^{-4}$ to prevent very large interspecific phenotypic variation and extinction.
-When extinction occurs, we aborted the simulation and restarted with newly drawn parameters.
-Results for $\mathrm{Corr}_{f_{\bar X}}(\alpha,|\beta|)$ and $\mathrm{Corr}_{f_{\bar X}}(\alpha,\gamma)$ are displayed in SM Figure \ref{num_corr}.
-
-```{r, echo=F, fig.align='center', fig.cap='\\label{num_corr} Numerical estimate for the correlations of selection gradients and competition coefficients.'}
-knitr::include_graphics('/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/corrs.png')
-```
-
-\newpage
-
-# References
+References {#references .unnumbered}
+==========

### 2020-04-28 Current Version/Untitled.tex
reorganize

--- 2020-04-18 .A White Noise Approach To Evolutionary Ecology/draft.tex
@@
-\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
-\PassOptionsToPackage{hyphens}{url}
-%
-\documentclass[]{article}
+\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
+%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
+\usepackage[hyphens]{url}
+
+  \journal{bioarxiv} % Sets Journal name
+
+\usepackage{lineno} % add
+  \linenumbers % turns line numbering on
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+\usepackage{graphicx}
+\usepackage{booktabs} % book-quality tables
+%%%%%%%%%%%%%%%% end my additions to header
+
+\usepackage[T1]{fontenc}
 \usepackage{lmodern}
@@
 \usepackage{fixltx2e} % provides \textsubscript
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provides euro and other symbols
-\else % if luatex or xelatex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
 % use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \usepackage{fontspec}
+  \ifxetex
+    \usepackage{xltxtra,xunicode}
+  \fi
+  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
+  \newcommand{\euro}{€}
+\fi
 % use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage[]{microtype}
-\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
-\usepackage{hyperref}
-\hypersetup{
+\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
+\bibliographystyle{elsarticle-harv}
+\ifxetex
+  \usepackage[setpagesize=false, % page size defined by xetex
+              unicode=false, % unicode breaks when used with xetex
+              xetex]{hyperref}
+\else
+  \usepackage[unicode=true]{hyperref}
+\fi
+\hypersetup{breaklinks=true,
+            bookmarks=true,
+            pdfauthor={},
             pdftitle={A White Noise Approach to Evolutionary Ecology},
-            pdfauthor={Bob Week, Steve Krone, Scott L. Nuismer, Luke J. Harmon},
-            pdfborder={0 0 0},
-            breaklinks=true}
+            colorlinks=false,
+            urlcolor=blue,
+            linkcolor=magenta,
+            pdfborder={0 0 0}}
 \urlstyle{same}  % don't use monospace font for urls
-\usepackage[margin=1in]{geometry}
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
 \setcounter{secnumdepth}{5}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-% set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-
+% Pandoc toggle for numbering sections (defaults to be off)
+
+% Pandoc header
 \usepackage{xr}
@@
 \usepackage{hyperref}
-\usepackage[left]{lineno}
 \usepackage{amsmath}
-\linenumbers
 \externaldocument{SM}
+\modulolinenumbers[2]
 \usepackage{float}
@@
 
+\begin{document}
+\begin{frontmatter}
+
   \title{A White Noise Approach to Evolutionary Ecology}
-\author{Bob Week, Steve Krone, Scott L. Nuismer, Luke J. Harmon}
-\date{3/1/2020}
-
-\begin{document}
-\maketitle
+    \author[Bio]{Bob Week\corref{1}}
+   \ead{bobweek@gmail.com} 
+    \author[Stats]{Steve Krone}
+
+    \author[Bio]{Scott L. Nuismer}
+
+    \author[Bio]{Luke J. Harmon}
+
+      \address[Bio]{University of Idaho, Department of Biological Sciences, 875 Perimeter
+Drive MS 3051, Moscow, Idaho, 83844}
+    \address[Stats]{University of Idaho, Department of Statistical Science, 875 Perimeter
+Drive MS 1104, Moscow, Idaho, 83844}
+      \cortext[1]{Corresponding Author}
+
   \begin{abstract}
@@
 
+ \end{frontmatter}
+
 \hypertarget{introduction}{%
@@
 account for the effects of dispersal and the phylogeographic history of
-interacting lineages (Kraft et al. 2007; Hickerson et al. 2010; Manceau,
-Lambert, and Morlon 2016; McPeek 2017) along with the genetic basis of
-ecologically relevant traits (Conner 2004; Fussman, Loreau, and Abrams
-2007) and feedbacks between populations and the biogeochemical cycles
-they ultimately depend on (Loreau 2010; Ågren and Andersson 2012). It is
+interacting lineages (Hickerson et al., 2010; Kraft et al., 2007;
+Manceau et al., 2016; McPeek, 2017) along with the genetic basis of
+ecologically relevant traits (Conner, 2004; Fussman et al., 2007) and
+feedbacks between populations and the biogeochemical cycles they
+ultimately depend on (Ågren and Andersson, 2012; Loreau, 2010). It is
 therefore ideal that the development of any such mathematical approach
@@
 ecological communities, especially as empirical and conceptual work in
-these directions continues to grow (Abdala-Roberts and Mooney 2014;
-Kölzsch et al. 2015; Crutsinger 2015; Fitzpatrick et al. 2015, 2017;
-Marx et al. 2017; Rudman et al. 2017; Skovmand et al. 2018; Nuland et
-al. 2019; Harmon et al. 2019). Furthermore, the approach would benefit
-from a stochastic component to capture the chance nature of biological
-reality (Lande, Engen, and SÆther 2003; Meester et al. 2018; Mubayi et
-al. 2019) and serve as a basis for the construction of statistical
+these directions continues to grow (Abdala-Roberts and Mooney, 2014;
+Crutsinger, 2015; Fitzpatrick et al., 2015, 2017; Harmon et al., 2019;
+Kölzsch et al., 2015; Marx et al., 2017; Nuland et al., 2019; Rudman et
+al., 2017; Skovmand et al., 2018). Furthermore, the approach would
+benefit from a stochastic component to capture the chance nature of
+biological reality (Lande et al., 2003; Meester et al., 2018; Mubayi et
+al., 2019) and serve as a basis for the construction of statistical
 methods that measure evolutionary and ecological processes occuring in
@@
 the mathematical formalization of this connection, which has been
-established in the contexts of population genetics (Crow and Kimura
-1970; Roughgarden 1979), evolutionary game theory (Hofbauer and Sigmund
-1998; Nowak 2006; Lion 2018), quantitative genetics (Lande 1982; Doebeli
-1996; Lion 2018) and a unifying framework for these three distinct
-approaches to evolutionary theory (Champagnat, Ferrière, and Méléard
-2006) which includes our approach as a special case. We note here there
-are two primary differences between the results presented in this paper
-and those of Champagnat, Ferrière and Méléard (2006). First, instead of
+established in the contexts of population genetics (Crow and Kimura,
+1970; Roughgarden, 1979), evolutionary game theory (Hofbauer and
+Sigmund, 1998; Lion, 2018; Nowak, 2006), quantitative genetics (Doebeli,
+1996; Lande, 1982; Lion, 2018) and a unifying framework for these three
+distinct approaches to evolutionary theory (Champagnat et al., 2006)
+which includes our approach as a special case. We note here there are
+two primary differences between the results presented in this paper and
+those of Champagnat, Ferrière and Méléard (2006). First, instead of
 developing a unifying approach to evolutionary theory, we focus on
@@
 effective population size using arguments based on properties of random
-samples (Lande 1976). Though this approach conveniently mimics the
+samples (Lande, 1976). Though this approach conveniently mimics the
 formalism provided by the Wright-Fisher model of population genetics,
@@
 genetic models without too much technical overhead, assuming a finite
-number of alleles (Lande, Engen, and Sæther 2009; Parsons, Quince, and
-Plotkin 2010; Gomulkiewicz, Krone, and Remien 2017). However, for
-populations with a continuum of types, such as a quantitative trait,
-finding a formal approach to derive the evolutionary response to
-demographic stochasticity has remained a vexing mathematical challenge.
-In this paper we close this gap by combining the calculus of white noise
-with results on rescaled limits of branching Brownian motion processes
-(BBM) and stochastic partial differential equations (SPDE), which are
-stochastic analogs of deterministic partial differential equations
-(PDE). Our goal has two components: 1) Establish a novel synthetic
-approach to theoretical evolutionary ecology that provides a formal
-connection between demographic stochasticity and random genetic drift in
-the context of quantitative traits. 2) Communicate some useful
-properties of space-time white noise, BBM and SPDE to as wide of
-audience as possible. With this goal in mind we will not provide a
-rigorous treatment of any of these deep subjects. Instead, we introduce
-a set of heuristics that only require the basic concepts of Reimann
-integration, partial differentiation and some exposure to Brownian
-motion and stochastic ordinary differential equations (SDE). For a
-concise introduction to SDE and Brownian motion, we recommend the primer
-by Evans (2014). Rigorous treatments of SPDE and rescaled limits of BBM
-can be found in Walsh (1986) and Dawson (1993) respectively.
+number of alleles (Gomulkiewicz et al., 2017; Lande et al., 2009;
+Parsons et al., 2010). However, for populations with a continuum of
+types, such as a quantitative trait, finding a formal approach to derive
+the evolutionary response to demographic stochasticity has remained a
+vexing mathematical challenge. In this paper we close this gap by
+combining the calculus of white noise with results on rescaled limits of
+branching Brownian motion processes (BBM) and stochastic partial
+differential equations (SPDE), which are stochastic analogs of
+deterministic partial differential equations (PDE). Our goal has two
+components: 1) Establish a novel synthetic approach to theoretical
+evolutionary ecology that provides a formal connection between
+demographic stochasticity and random genetic drift in the context of
+quantitative traits. 2) Communicate some useful properties of space-time
+white noise, BBM and SPDE to as wide of audience as possible. With this
+goal in mind we will not provide a rigorous treatment of any of these
+deep subjects. Instead, we introduce a set of heuristics that only
+require the basic concepts of Reimann integration, partial
+differentiation and some exposure to Brownian motion and stochastic
+ordinary differential equations (SDE). For a concise introduction to SDE
+and Brownian motion, we recommend the primer by Evans (2014). Rigorous
+treatments of SPDE and rescaled limits of BBM can be found in Walsh
+(1986) and Dawson (1993) respectively.
 
@@
 distributions. Although these dynamics have been considered before while
-accounting for demographic stochasticity (Polechová and Barton 2015), to
-our knowledge this is the first time they have been rigorously derived
-from first principles. Third, we simplify these SDE by approximating the
-abundance density of phenotypes with a Gaussian curve and integrate the
-results with a model of imperfect inheritance to find formula that
-generalize classical quantitative genetic expressions. This includes an
-expression for the response of additive genetic variance to selection in
-terms of fitness gradients analogous to the expression for the response
-in mean phenotype found by Lande (1976). Fourth, following classical
-niche theory we provide a rigorous mechanistic derivation of fitness for
-populations experiencing resource competition and abiotic stabilizing
-selection using diffusion limits of measure-valued branching processes.
-This calculation provides a demonstration for performing such diffusion
-limits that can be applied in other contexts. Fifth, we explore
-implications of the resulting model of diffuse coevolution by
+accounting for demographic stochasticity (Polechová and Barton, 2015),
+to our knowledge this is the first time they have been rigorously
+derived from first principles. Third, we simplify these SDE by
+approximating the abundance density of phenotypes with a Gaussian curve
+and integrate the results with a model of imperfect inheritance to find
+formula that generalize classical quantitative genetic expressions. This
+includes an expression for the response of additive genetic variance to
+selection in terms of fitness gradients analogous to the expression for
+the response in mean phenotype found by Lande (1976). Fourth, following
+classical niche theory we provide a rigorous mechanistic derivation of
+fitness for populations experiencing resource competition and abiotic
+stabilizing selection using diffusion limits of measure-valued branching
+processes. This calculation provides a demonstration for performing such
+diffusion limits that can be applied in other contexts. Fifth, we
+explore implications of the resulting model of diffuse coevolution by
 investigating the relationship between selection gradients and
@@
 At the core of our approach is a stochastic analog of the replicator
-equation with mutation in continuous time and phenotypic space (Taylor
-and Jonker 1978; Schuster and Sigmund 1983). From this stochastic
+equation with mutation in continuous time and phenotypic space (Schuster
+and Sigmund, 1983; Taylor and Jonker, 1978). From this stochastic
 replicator-mutator equation we derive a system of SDE for the dynamics
@@
 individuals with allele \(i\) and \(m_i\) the growth rate of allele
-\(i\) (called the Malthusian parameter in Crow and Kimura 1970), we have
+\(i\) (called the Malthusian parameter in Crow and Kimura, 1970), we
+have
+
 \begin{equation}\label{first}
 \frac{d\nu_i}{dt}=m_i\nu_i.
-\end{equation} Starting from this model, we get the total abundance of
-the population as \(N=\sum_{i=1}^K\nu_i\), the frequency of allele \(i\)
-as \(p_i=\nu_i/N\) and the mean fitness of the population as
+\end{equation}
+
+Starting from this model, we get the total abundance of the population
+as \(N=\sum_{i=1}^K\nu_i\), the frequency of allele \(i\) as
+\(p_i=\nu_i/N\) and the mean fitness of the population as
 \(\bar m=\sum_{i=1}^Kp_im_i\). Hence, we can employ some elementary
 calculus to derive the dynamics of abundance \(dN/dt\) and the dynamics
-of allele frequencies \(dp_1/dt,\dots,dp_K/dt\) as \begin{equation}
+of allele frequencies \(dp_1/dt,\dots,dp_K/dt\) as
+
+\begin{equation}
 \frac{dN}{dt}=\sum_{i=1}^K\nu_im_i=N\sum_{i=1}^Kp_im_i=\bar mN,
@@
 =\frac{1}{N}\left(m_i\nu_i-\bar m Np_i\right)=(m_i-\bar m)p_i.
-\end{equation} Two important observations of these equations include (i)
-mean fitness \(\bar m\) determines the abundance dynamics of the entire
-population and (ii) allele \(i\) will increase (decrease) in frequency
-if \(m_i>\bar m\) (\(<\bar m\)). Equation (\ref{popgen}) is known in the
+\end{equation}
+
+Two important observations of these equations include (i) mean fitness
+\(\bar m\) determines the abundance dynamics of the entire population
+and (ii) allele \(i\) will increase (decrease) in frequency if
+\(m_i>\bar m\) (\(<\bar m\)). Equation (\ref{popgen}) is known in the
 field of evolutionary game theory as the replicator equation (Hofbauer
-and Sigmund 1998; Nowak 2006; Lion 2018; Taylor and Jonker 1978;
-Schuster and Sigmund 1983). Instead of being explicitly focused on
+and Sigmund, 1998; Lion, 2018; Nowak, 2006; Schuster and Sigmund, 1983;
+Taylor and Jonker, 1978). Instead of being explicitly focused on
 alleles, the replicator equation describes the fluctuations of relative
@@
 include mutation, which is known as the replicator-mutator equation
-(Nowak 2006).
+(Nowak, 2006).
 
@@
 satisfies \(\nu(x,0)\geq0\) for all \(x\in\mathbb{R}\) and
+
 \begin{equation}
@@
 \sigma^2(0)=\int_{-\infty}^{+\infty}(x-\bar x(0))^2p(x,0)dx<+\infty,
-\end{equation} where \(\bar x(t)\) and \(\sigma^2(t)\) are respectively
-the mean trait and phenotypic variance at time \(t\geq0\). Second, we
-assume selection is determined by the growth rate \(m(h,x)\) that is
-differentiable with respect to both arguments and is bounded above by
-some \(r\in\mathbb{R}\) for all \(x\in\mathbb{R}\) and \(h\geq0\). We
-also assume \((1+h)C\leq m(h,x)\) for some constant \(C>0\) so that
+\end{equation}
+
+where \(\bar x(t)\) and \(\sigma^2(t)\) are respectively the mean trait
+and phenotypic variance at time \(t\geq0\). Second, we assume selection
+is determined by the growth rate \(m(h,x)\) that is differentiable with
+respect to both arguments and is bounded above by some
+\(r\in\mathbb{R}\) for all \(x\in\mathbb{R}\) and \(h\geq0\). We also
+assume \((1+h)C\leq m(h,x)\) for some constant \(C>0\) so that
 \(m(h,x)\) is linear or more than linear in \(h\). We denote by \(K\) an
@@
 to account for nonlocal effects of the abundance density \(\nu(x,t)\) on
-the fitness of individuals with trait value \(x\) (Volpert 2014). As an
+the fitness of individuals with trait value \(x\) (Volpert, 2014). As an
 example of a nonlocal effect, consider a population experiencing
@@
 quantitative character simultaneously by the PDE
+
 \begin{equation}\label{eq1}
 \frac{\partial}{\partial t}\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)
-\end{equation} with the initial condition \(\nu(x,0)\) described above.
-Equation (\ref{eq1}) is a semilinear parabolic PDE and is referred to as
-a scalar reaction-diffusion equation (Zheng 2004; Evans 2010). To ensure
+\end{equation}
+
+with the initial condition \(\nu(x,0)\) described above. Equation
+(\ref{eq1}) is a semilinear parabolic PDE and is referred to as a scalar
+reaction-diffusion equation (Evans, 2010; Zheng, 2004). To ensure
 existence and uniqueness of solutions to (\ref{eq1}), we further assume
@@
 \(\int_\mathbb{R}\nu_1(x)dx,\int_\mathbb{R}\nu_2(x)dx\leq M\) implies
+
 \begin{equation}\label{local_lipschitz}
 \int_\mathbb{R}\left|m(\nu_1,x)\nu_1(x)-m(\nu_2,x)\nu_2(x)\right|dx\leq L_M\int_\mathbb{R}|\nu_1(x)-\nu_2(x)|dx.
-\end{equation} In SM \S\ref{finite} we combine the above assumptions on
-growth rate, mutation and initial condition to prove that solutions to
-(\ref{eq1}) satisfy \(N(t),|\bar x(t)|,\sigma^2(t)<+\infty\) for all
-\(t\geq0\).
+\end{equation}
+
+In SM \S\ref{finite} we combine the above assumptions on growth rate,
+mutation and initial condition to prove that solutions to (\ref{eq1})
+satisfy \(N(t),|\bar x(t)|,\sigma^2(t)<+\infty\) for all \(t\geq0\).
 
@@
 model of spatially distributed population dynamics has been intensely
-studied (for a review, see Cantrell and Cosner 2004). When \(\mu=0\),
+studied (for a review, see Cantrell and Cosner, 2004). When \(\mu=0\),
 equation (\ref{eq1}) can be seen as an analog of equation (\ref{first})
@@
 selection can be captured using the growth rate
+
 \begin{equation}\label{special_m}
 m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_{-\infty}^{+\infty}\nu(y,t)dy
-\end{equation} where \(r\in\mathbb{R}\) is the intrinsic rate of growth
-in the absence of abiotic stabilizing selection, \(a>0\) is the strength
-of abiotic stabilizing selection around the phenotypic optimum
+\end{equation}
+
+where \(r\in\mathbb{R}\) is the intrinsic rate of growth in the absence
+of abiotic stabilizing selection, \(a>0\) is the strength of abiotic
+stabilizing selection around the phenotypic optimum
 \(\theta\in\mathbb{R}\), \(c>0\) captures the sensitivity of fitness to
@@
 and that predicted by the classic Gaussian allelic model of quantitative
-genetics (Lande 1975; Bürger 2000; Johnson and Barton 2005). In
+genetics (Bürger, 2000; Johnson and Barton, 2005; Lande, 1975). In
 \S\ref{coev} and SM \S\ref{diffuse} we find this fitness function is
@@
 =N(t)\int_{-\infty}^{+\infty} m(\nu,x)p(x,t)dx=\bar m(t) N(t).
-\end{multline} Using our assumptions on mutation and rate of growth, we
-show in SM \S\ref{finite} \(\nu(x,t)\) is twice differentiable with
-respect to \(x\) and \(\int_{-\infty}^{+\infty}\nu(x,t)dx<\infty\) for
-all \(t\geq0\). This implies that we are justified in swapping the order
-of differentiation and integration and the result
+\end{multline}
+
+Using our assumptions on mutation and rate of growth, we show in SM
+\S\ref{finite} \(\nu(x,t)\) is twice differentiable with respect to
+\(x\) and \(\int_{-\infty}^{+\infty}\nu(x,t)dx<\infty\) for all
+\(t\geq0\). This implies that we are justified in swapping the order of
+differentiation and integration and the result
 \(\int_{-\infty}^{+\infty}\frac{\partial^2}{\partial x^2}\nu(x,t)dx=0\)
@@
 =\Big(m(\nu,x)-\bar m(t)\Big) p(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t).
-\end{multline} This result closely resembles Kimura's
-continuum-of-alleles model (Kimura 1965; Bürger 1986). The primary
-difference being that our model utilizes diffusion instead of
-convolution with an arbitrary mutation kernel. Of course, our model of
-mutation can be derived as an approximation to Kimura's model, which,
-for sexually reproducing populations, has been referred to as the
-Gaussian allelic approximation in reference to the distribution of
-mutational effects at loci in a genome on the values of traits (Lande
-1975; but see also Bürger 2000; along with Johnson and Barton 2005), the
-infinitesimal model in reference to modelling continuous traits as being
-encoded by an infinite number of loci each having infinitesimal effect
-(Barton, Etheridge, and Véber 2017) and the Gaussian descendants
-approximation in reference to offspring trait values being normally
-distributed around their parental values (Turelli 2017). Alternatively,
-since diffusion is the continuous-time equivalent to convolution against
-a Gaussian kernel (SM \S\ref{diffconvequiv}), equation (\ref{p_PDE}) can
-also be seen as a special case of Kimura's continuum-of-alleles model.
-We return to these concepts in \S\ref{inheritance} where we review a
-simple model of inheritance that leads us to an expression for the
-stochastic evolution of additive genetic variance. To distinguish this
-model from previous models of phenotypic evolution we refer to PDE
-(\ref{eq1}) from which (\ref{p_PDE}) was derived as the Deterministic
-Asexual Gaussian allelic model with Abundance dynamics (abbreviated
-DAGA). Later, we will extend this model to include the effects of
-demographic stochasticity, which we refer to as the Stochastic Asexual
-Gaussian allelic model with Abundance dynamics (abbreviated SAGA).
+\end{multline}
+
+This result closely resembles Kimura's continuum-of-alleles model
+(Bürger, 1986; Kimura, 1965). The primary difference being that our
+model utilizes diffusion instead of convolution with an arbitrary
+mutation kernel. Of course, our model of mutation can be derived as an
+approximation to Kimura's model, which, for sexually reproducing
+populations, has been referred to as the Gaussian allelic approximation
+in reference to the distribution of mutational effects at loci in a
+genome on the values of traits (Lande, 1975; but see also Bürger, 2000;
+along with Johnson and Barton, 2005), the infinitesimal model in
+reference to modelling continuous traits as being encoded by an infinite
+number of loci each having infinitesimal effect (Barton et al., 2017)
+and the Gaussian descendants approximation in reference to offspring
+trait values being normally distributed around their parental values
+(Turelli, 2017). Alternatively, since diffusion is the continuous-time
+equivalent to convolution against a Gaussian kernel (SM
+\S\ref{diffconvequiv}), equation (\ref{p_PDE}) can also be seen as a
+special case of Kimura's continuum-of-alleles model. We return to these
+concepts in \S\ref{inheritance} where we review a simple model of
+inheritance that leads us to an expression for the stochastic evolution
+of additive genetic variance. To distinguish this model from previous
+models of phenotypic evolution we refer to PDE (\ref{eq1}) from which
+(\ref{p_PDE}) was derived as the Deterministic Asexual Gaussian allelic
+model with Abundance dynamics (abbreviated DAGA). Later, we will extend
+this model to include the effects of demographic stochasticity, which we
+refer to as the Stochastic Asexual Gaussian allelic model with Abundance
+dynamics (abbreviated SAGA).
 
 The covariance of fitness and phenotype across the population is defined
-as \begin{equation}
+as
+
+\begin{equation}
 \mathrm{Cov}_t\Big(m(\nu,x),x\Big)=\int_{-\infty}^{+\infty}\Big(m(\nu,x)-\bar m(t)\Big)\Big(x-\bar x(t)\Big)p(x,t)dx.
-\end{equation} Hence, the dynamics of the mean trait \(\bar x(t)\) can
-be derived as
+\end{equation}
+
+Hence, the dynamics of the mean trait \(\bar x(t)\) can be derived as
 
@@
 Equation (\ref{detmean}) is a continuous time analog of the well known
-Robertson-Price equation without transmission bias (Robertson 1966;
-Price 1970; Frank 2012; Queller 2017; Lion 2018). Whether or not the
+Robertson-Price equation without transmission bias (Frank, 2012; Lion,
+2018; Price, 1970; Queller, 2017; Robertson, 1966). Whether or not the
 covariance of fitness and phenotype creates change in \(\bar x\) to
 maximize mean fitness \(\bar m\) depends on the degree to which
-selection is frequency dependent (Lande 1976). Since this change is
+selection is frequency dependent (Lande, 1976). Since this change is
 driven by a covariance with respect to phenotypic diversity, the
@@
 phenotypic variation to this model of selection and mutation is
+
 \begin{equation}\label{vardyn_det}
 \frac{d}{dt}\sigma^2(t)=\mathrm{Cov}_t\Big(m(\nu,x),(x-\bar x)^2\Big)+\mu.
-\end{equation} For the sake of space we relegate the derivation of
-\(d\sigma^2/dt\) to SM \S\ref{var_deriv}. In the absence of mutation
-equation (\ref{vardyn_det}) agrees with the result derived by Lion
-(2018) for discrete phenotypes. From a statistical perspective, if we
-think of \((x-\bar x)^2\) as a square error, then in analogy to the
-dynamics of the mean trait, we see that the response in \(\sigma^2\) to
-selection can be expressed as a covariance of fitness and square error,
-which is defined in analogy to \(\mathrm{Cov}_t(m(\nu,x),x)\). Just as
-for the evolution of \(\bar x(t)\), this covariance also creates change
-in \(\sigma^2\) that can either increase or decrease mean fitness
+\end{equation}
+
+For the sake of space we relegate the derivation of \(d\sigma^2/dt\) to
+SM \S\ref{var_deriv}. In the absence of mutation equation
+(\ref{vardyn_det}) agrees with the result derived by Lion (2018) for
+discrete phenotypes. From a statistical perspective, if we think of
+\((x-\bar x)^2\) as a square error, then in analogy to the dynamics of
+the mean trait, we see that the response in \(\sigma^2\) to selection
+can be expressed as a covariance of fitness and square error, which is
+defined in analogy to \(\mathrm{Cov}_t(m(\nu,x),x)\). Just as for the
+evolution of \(\bar x(t)\), this covariance also creates change in
+\(\sigma^2\) that can either increase or decrease mean fitness
 \(\bar m\), depending on whether or not individual fitness depends on
@@
 a sequence of rescaled BBM converges to a limiting process that can be
-characterized by a SPDE (Méléard and Roelly 1992, 1993; Li 1998;
-Champagnat, Ferrière, and Méléard 2006). The limiting processes of
-rescaled BBM have been referred to as measure-valued Markov processes
-(Dawson 1993) or superprocesses (Etheridge 2000). Under the simplifying
-assumptions inherited from our treatment of deterministic dynamics and
-the additional assumption that the magnitude of demographic
-stochasticity is independent of trait values, we obtain as a special
-case a relatively simple expression for an SPDE that generalizes DAGA.
-The simplicity of our special case allows us to use properties of
-space-time white noise processes to derive a set of SDE that generalize
-equations (\ref{N_det}), (\ref{detmean}) and (\ref{vardyn_det}) to
-include the effects of demographic stochasticity. Classical expressions
-for the effects of random genetic drift on the evolution of mean traits
-are obtained as a further special case.
+characterized by a SPDE (Champagnat et al., 2006; Li, 1998; Méléard and
+Roelly, 1993, 1992). The limiting processes of rescaled BBM have been
+referred to as measure-valued Markov processes (Dawson, 1993) or
+superprocesses (Etheridge, 2000). Under the simplifying assumptions
+inherited from our treatment of deterministic dynamics and the
+additional assumption that the magnitude of demographic stochasticity is
+independent of trait values, we obtain as a special case a relatively
+simple expression for an SPDE that generalizes DAGA. The simplicity of
+our special case allows us to use properties of space-time white noise
+processes to derive a set of SDE that generalize equations
+(\ref{N_det}), (\ref{detmean}) and (\ref{vardyn_det}) to include the
+effects of demographic stochasticity. Classical expressions for the
+effects of random genetic drift on the evolution of mean traits are
+obtained as a further special case.
 
@@
 
-\hypertarget{introduction-to-white-noise}{%
-\subsubsection{Introduction to white
-noise}\label{introduction-to-white-noise}}
-
 One can think of white noise as the static seen on old television sets
@@
 
-{\centering \includegraphics{wn}
+{\centering \includegraphics[width=0.8\linewidth]{wn}
 
@@
 understanding is possible by considering white noise as a
-\emph{measure}-valued process (Dawson 1975; Walsh 1986) or as a
+\emph{measure}-valued process (Dawson, 1975; Walsh, 1986) or as a
 \emph{generalized} process that acts on classically defined functions or
 stochastic processes to return either random variables or stochastic
-processes (Krylov and Rozovskii 1981; Da Prato and Zabczyk 2014). Since
-a measure-valued process can be defined from a generalized process and a
-generalized process can be defined from a measure-valued process, the
-distinction between the two is more or less a matter of perspective.
+processes (Da Prato and Zabczyk, 2014; Krylov and Rozovskii, 1981).
+Since a measure-valued process can be defined from a generalized process
+and a generalized process can be defined from a measure-valued process,
+the distinction between the two is more or less a matter of perspective.
 However, we find the perspective of white-noise as a generalized process
@@
 define \(\mathbf W(f)\) and \(\mathbf W(g)\) to be Gaussian processes
-satisfying, for any \(t,t_1,t_2\geq0\), \begin{equation}\label{exp_WN}
+satisfying, for any \(t,t_1,t_2\geq0\),
+
+\begin{equation}\label{exp_WN}
 \mathbb{E}\big(\mathbf W_t(f)\big)=\mathbb{E}\big(\mathbf W_t(g)\big)=0,
@@
 \mathbb{C}\big(\mathbf W_{t_1}(f),\mathbf W_{t_2}(g)\big)=\mathbb{E}\left(\int_0^{t_1\wedge t_2}\int_\mathbb{R} f(x,s)g(x,s)dxds\right),
-\end{equation} where \(t_1\wedge t_2=\min(t_1,t_2)\) and \(\mathbb{C}\)
-denotes covariance with respect to the underlying probability space. In
+\end{equation}
+
+where \(t_1\wedge t_2=\min(t_1,t_2)\) and \(\mathbb{C}\) denotes
+covariance with respect to the underlying probability space. In
 particular, denoting \(\mathbb{V}\) the variance operator with respect
@@
 white noise \(\dot W(x,t)\) implicitly via the stochastic integral
+
 \begin{equation}\label{informal}
 ``\int_0^t\int_\mathbb{R}f(x,s)\dot W(x,s)dxds"=\mathbf W_t(f), \ \forall \ f\in\mathscr{N}_2, \ t\geq0.
-\end{equation} We place quotations in the above expression to emphasize
-its informal nature and that it should not be confused with classical
-Riemann integration. Following this definition of white noise, we
-compute its value by sampling it using \(\mathscr{N}_2\) processes. For
-example, integrating white noise over a region \(D\times[0,t]\), with
-\(t>0\) and \(D\) a bounded subset of \(\mathbb{R}\), is equivalent to
-evaluating \(\mathbf W_t(I_D)\) for the deterministic process
+\end{equation}
+
+We place quotations in the above expression to emphasize its informal
+nature and that it should not be confused with classical Riemann
+integration. Following this definition of white noise, we compute its
+value by sampling it using \(\mathscr{N}_2\) processes. For example,
+integrating white noise over a region \(D\times[0,t]\), with \(t>0\) and
+\(D\) a bounded subset of \(\mathbb{R}\), is equivalent to evaluating
+\(\mathbf W_t(I_D)\) for the deterministic process
+
 \begin{equation}
@@
 \end{matrix}\right..
-\end{equation} Since \begin{equation}
+\end{equation}
+
+Since
+
+\begin{equation}
 \|I_D\|_t^2=\mathbb{E}\left(\int_0^t\int_\mathbb{R}I_D^2(x,s)dxds\right)=t\int_Ddx=t|D|<+\infty, 
-\end{equation} where \(|D|\) denotes the length of \(D\), we have
+\end{equation}
+
+where \(|D|\) denotes the length of \(D\), we have
 \(I_D\in\mathscr{N}_2\). Thus, using equations (\ref{exp_WN}) and
 (\ref{cov_WN}) and adopting the informal notation introduced in equation
-(\ref{informal}), we can write the following \begin{equation}
+(\ref{informal}), we can write the following
+
+\begin{equation}
 \mathbb{E}\left(\int_0^t\int_D\dot W(x,s)dxds\right)=0,
@@
 \mathbb{V}\left(\int_0^t\int_D\dot W(x,s)dxds\right)=t|D|.
-\end{equation} Using this informal notation, equations (\ref{exp_WN})
-and (\ref{cov_WN}) can be rewritten as \begin{equation}\label{exp_WN_xi}
+\end{equation}
+
+Using this informal notation, equations (\ref{exp_WN}) and
+(\ref{cov_WN}) can be rewritten as
+
+\begin{equation}\label{exp_WN_xi}
 \mathbb{E}\left(\int_0^t\int_\mathbb{R}f(x,s)\dot W(x,s)dxds\right)=0,
-\end{equation} \begin{equation}\label{cov_WN_xi}
-\mathbb{C}\left(\int_0^{t_1}\int_\mathbb{R}f(x,s)\dot W(x,s)dxds,\int_0^{t_2}\int_\mathbb{R}g(x,s)\dot W(x,s)dxds\right)
+\end{equation} \begin{multline}\label{cov_WN_xi}
+\mathbb{C}\left(\int_0^{t_1}\int_\mathbb{R}f(x,s)\dot W(x,s)dxds,\int_0^{t_2}\int_\mathbb{R}g(x,s)\dot W(x,s)dxds\right) \\
 =\int_0^{t_1\wedge t_2}\int_\mathbb{R}f(x,s)g(x,s)dxds.
-\end{equation} To relate these formula to the common notation used for
-SDE, we write \begin{equation}
+\end{multline}
+
+To relate these formula to the common notation used for SDE, we write
+
+\begin{equation}
 \hat f(x,t)=\frac{f(x,t)}{\sqrt{\int_\mathbb{R}f^2(y,t)dy}} \ \text{ and } \ 
 d\hat{\mathbf W}_t(f)=\left(\int_\mathbb{R}\hat f(x,t)\dot W(x,t)dx\right)dt
-\end{equation} so that \begin{equation}
+\end{equation}
+
+so that
+
+\begin{equation}
 \int_0^td\hat{\mathbf W}_s(f)=\int_0^t\int_\mathbb{R}\frac{f(x,s)}{\sqrt{\int_\mathbb{R}f^2(s,y)dy}}\dot W(x,s)dxds.
-\end{equation} This implies \begin{equation}
+\end{equation}
+
+This implies
+
+\begin{equation}
 \mathbb{E}\left(\int_0^td\hat{\mathbf W}_s(f)\right)=0, \ \mathbb{C}\left(\int_0^{t_1}d\hat{\mathbf W}_s(f),\int_0^{t_2}d\hat{\mathbf W}_s(f)\right)=t_1\wedge t_2
-\end{equation} and in particular, as a function of \(t\),
+\end{equation}
+
+and in particular, as a function of \(t\),
 \(\int_0^td\hat{\mathbf W}_s(f)\) is a standard Brownian motion for any
@@
 ingredients of a branching process. Mathematical investigations of such
-processes have a relatively deep history (Kendall 1966). The most simple
-branching process, known as the Galton-Watson process, describes the
-number of individuals alive at a given time \(t\geq0\) as a non-negative
-integer (Kimmel and Axelrod 2015). Feller (1951) introduced a formal
-method to approximate branching processes with diffusion processes which
-are continuous in state (i.e., population size is approximated as a
-continuous quantity). Since diffusion processes possess greater
-analytical tractability than branching processes, Feller's method, known
-as the diffusion limit, has acquired immense popularity particularly in
-the field of mathematical population genetics (Ewens 2004). For over the
-past half of a century a great deal of accomplishments have been
-achieved in formalizing the diffusion limits of branching processes that
-describe populations of individuals occuring in some continuous space
-(Watanabe 1968; Dawson 1975, 1978; Perkins 1992, 1995; Méléard and
-Roelly 1993; Li 1998; Bertoin and Le Gall 2003; Etheridge 2008; Barton
-and Etheridge 2019). This space can represent geographic space or,
-relevant to our context, phenotypic space. In the following subsection,
-we describe the BBM process, which is a particularly important branching
-process with spatial structure. This process has been very useful in the
-study of SPDE due to its simplifying assumption that individuals do not
-interact. However, this assumption imposes an unfortunate restriction by
-precluding the modelling of ecological interactions. We therefore follow
-our discussion of BBM with a review of a few important results on
-spatially structured branching processes that account for interactions.
-
-\textbf{Branching Brownian motion}
-
-A BBM tracks individuals navigating \(d\)-dimensional Euclidean space
-that reproduce and senesce between exponentially distributed intervals.
-Unlike other stochastic processes that take values in \(\mathbb{R}^d\),
-BBM takes values in the set of \emph{non-negative finite measures} over
-\(\mathbb{R}^d\). Intuitively, one can think of a finite measure as a
-function that maps subsets of \(\mathbb{R}^d\) to real numbers. To be
-formal, we only consider the Borel subsets of \(\mathbb{R}^d\)
-corresponding to the Euclidean metric, but understanding this
-technicality is not crucial to our discussion. In particular, denoting
-\(X_t\) a BBM, for a subset \(D\subset\mathbb{R}^d\), \(X_t(D)\) returns
-the (random) number of individuals alive within the region \(D\) at time
-\(t\geq0\). The BBM has three main components:
+processes have a relatively deep history (Kendall, 1966). The most
+simple branching process, known as the Galton-Watson process, describes
+the number of individuals alive at a given time \(t\geq0\) as a
+non-negative integer (Kimmel and Axelrod, 2015). Feller (1951)
+introduced a formal method to approximate branching processes with
+diffusion processes which are continuous in state (i.e., population size
+is approximated as a continuous quantity). Since diffusion processes
+possess greater analytical tractability than branching processes,
+Feller's method, known as the diffusion limit, has acquired immense
+popularity particularly in the field of mathematical population genetics
+(Ewens, 2004). For over the past half of a century a great deal of
+accomplishments have been achieved in formalizing the diffusion limits
+of branching processes that describe populations of individuals occuring
+in some continuous space (Barton and Etheridge, 2019; Bertoin and Le
+Gall, 2003; Dawson, 1978, 1975; Etheridge, 2008; Li, 1998; Méléard and
+Roelly, 1993; Perkins, 1995, 1992; Watanabe, 1968). This space can
+represent geographic space or, relevant to our context, phenotypic
+space. In the following subsection, we describe the BBM process, which
+is a particularly important branching process with spatial structure.
+This process has been very useful in the study of SPDE due to its
+simplifying assumption that individuals do not interact. However, this
+assumption imposes an unfortunate restriction by precluding the
+modelling of ecological interactions. We therefore follow our discussion
+of BBM with a review of a few important results on spatially structured
+branching processes that account for interactions.
+
+\emph{Branching Brownian motion (BBM):} A BBM tracks individuals
+navigating \(d\)-dimensional Euclidean space that reproduce and senesce
+between exponentially distributed intervals. Unlike other stochastic
+processes that take values in \(\mathbb{R}^d\), BBM takes values in the
+set of \emph{non-negative finite measures} over \(\mathbb{R}^d\).
+Intuitively, one can think of a finite measure as a function that maps
+subsets of \(\mathbb{R}^d\) to real numbers. To be formal, we only
+consider the Borel subsets of \(\mathbb{R}^d\) corresponding to the
+Euclidean metric, but understanding this technicality is not crucial to
+our discussion. In particular, denoting \(X_t\) a BBM, for a subset
+\(D\subset\mathbb{R}^d\), \(X_t(D)\) returns the (random) number of
+individuals alive within the region \(D\) at time \(t\geq0\). The BBM
+has three main components:
 
@@
   iteroparity. However, under the appropriate rescaling, both approaches
-  converge to the same diffusion limit (Champagnat, Ferrière, and
-  Méléard 2006). We therefore choose the former approach for the sake of
-  simplicity.
+  converge to the same diffusion limit (Champagnat et al., 2006). We
+  therefore choose the former approach for the sake of simplicity.
 \item
@@
   known the critical condition implies extinction in finite time with
-  probability one, given finite \(X_0(\mathbb{R}^d)\) (Athreya and Ney
+  probability one, given finite \(X_0(\mathbb{R}^d)\) (Athreya and Ney,
   1972).
@@
 a super-Brownian motion and is also a non-negative finite measure-valued
-process (Watanabe 1968). However, instead of returning the number of
+process (Watanabe, 1968). However, instead of returning the number of
 individuals alive in a region of space, super-Brownian motion returns
@@
 turns out that \(\mathscr{X}_t\) is absolutely continuous with respect
-to the Lebesgue measure for each \(t\geq0\) (Konno and Shiga 1988;
-Reimers 1989). This means that we can write
+to the Lebesgue measure for each \(t\geq0\) (Konno and Shiga, 1988;
+Reimers, 1989). This means that we can write
 \(\mathscr{X}_t(D)=\int_D \nu(x,t)dx\) for some density process
 \(\nu(x,t)\). Setting \(\lambda=1\) and \(m=\ln\mathscr{W}\) this
-density process satisfies the SPDE \begin{equation}\label{neutral_SPDE}
+density process satisfies the SPDE
+
+\begin{equation}\label{neutral_SPDE}
 \frac{\partial}{\partial t}\nu(x,t)=m\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)+\sqrt{V\nu(x,t)}\dot W(x,t).
-\end{equation} Since \(\nu(x,t)\) is not generally differentiable in
-\(x\) or \(t\), the derivatives in expression (\ref{neutral_SPDE}) are
-taken in the \emph{weak} sense (for a definition see Evans 2010). That
-is, to rigorously interpret SPDE (\ref{neutral_SPDE}), we integrate the
+\end{equation}
+
+Since \(\nu(x,t)\) is not generally differentiable in \(x\) or \(t\),
+the derivatives in expression (\ref{neutral_SPDE}) are taken in the
+\emph{weak} sense (for a definition see Evans, 2010). That is, to
+rigorously interpret SPDE (\ref{neutral_SPDE}), we integrate the
 solution \(\nu(x,t)\) against functions \(f\in C_b^2(\mathbb{R})\) where
@@
 +\int_0^t\int_\mathbb{R}f(x)\sqrt{V\nu(x,s)}\dot W(x,s)dxds, \ \forall \ f\in C_b^2(\mathbb{R}).
-\end{multline} This expression is referred to as the \emph{mild}
-solution of (\ref{neutral_SPDE}). For more on the general theory of SPDE
-see Walsh (1986) and Da Prato and Zabczyk (2014). Note that since
-\(\nu(x,t)\) is the density of a finite measure, it is integrable for
-each \(t\geq0\). Thus, since for some \(M>0\), \(|f(x)|\leq M\) for
-every \(x\in\mathbb{R}\), setting \(\varphi(x,t)=f(x)\sqrt{V\nu(x,t)}\)
+\end{multline}
+
+This expression is referred to as the \emph{mild} solution of
+(\ref{neutral_SPDE}). For more on the general theory of SPDE see Walsh
+(1986) and Da Prato and Zabczyk (2014). Note that since \(\nu(x,t)\) is
+the density of a finite measure, it is integrable for each \(t\geq0\).
+Thus, since for some \(M>0\), \(|f(x)|\leq M\) for every
+\(x\in\mathbb{R}\), setting \(\varphi(x,t)=f(x)\sqrt{V\nu(x,t)}\)
 implies \(\varphi\in\mathscr{N}_2\). Hence, the white noise integral on
@@
 spatial dimension) is referred to as a Dawson-Watanabe superprocess
-(Etheridge 2000). Conditioning a Dawson-Watanabe superprocess to have
-constant mass returns a Fleming-Viot process (Etheridge and March 1991;
-Perkins 1991) which has been popular in studies of spatial population
+(Etheridge, 2000). Conditioning a Dawson-Watanabe superprocess to have
+constant mass returns a Fleming-Viot process (Etheridge and March, 1991;
+Perkins, 1991) which has been popular in studies of spatial population
 genetics. In particular, an extension of the Fleming-Viot process, known
@@
 resolve some technical challenges in modelling isolation by distance
-(Felsenstein 1975; see also Barton, Etheridge, and Véber 2013; and
-Barton and Etheridge 2019). Although this provides an impressive list of
+(Felsenstein, 1975; see also Barton et al., 2013; and Barton and
+Etheridge, 2019). Although this provides an impressive list of
 accomplishments, the Dawson-Watanabe superprocess falls short of our
@@
 
-\textbf{Interacting superprocesses}
-
-In the above subsection we reviewed the diffusion limit of an especially
-tractable measure-valued branching process. However, we found the
-simplicity of this process restricts us from modelling nonlocal effects
-on the fitness of individuals. Here, we discuss superprocesses that
-account for such effects. The existence of diffusion limits for a class
-of measure-valued branching processes involving nonlocal effects has
-been treated by Méléard and Roelly (1992, 1993). The interactions can
-manifest as dependencies of the spatial movement or reproductive law of
-individuals on their position \(x\) and the state of the whole
-population described by \(X_t\). An important result of Méléard and
-Roelly (1992, 1993) is a theorem that provides sufficient conditions to
-construct a sequence of rescaled measure-valued branching processes that
-converge to a generalization of the Dawson-Watanabe superprocess that
-includes interactions. The rescaling is analogous to that described
-above for non-interacting Dawson-Watanabe superprocesses, but now the
-reproductive law described by \(\mathscr{W}(X_t,x)\) and \(V(X_t,x)\),
-branching rate \(\lambda(X_t,x)\) and diffusion parameter
-\(\sqrt{\mu(X_t,x)}\) are allowed to depend on the whole population
-\(X_t\) and individual location \(x\). In Figure \ref{rescaled} we
-demonstrate this rescaling in discrete time for a population
-experiencing stabilizing selection and logistic growth. Since time is
-discretized, the process we simulate is formally a branching random
-walk. For further details on our simulation see SM \S\ref{numerical}.
+\emph{Interacting superprocesses:} In the above subsection we reviewed
+the diffusion limit of an especially tractable measure-valued branching
+process. However, we found the simplicity of this process restricts us
+from modelling nonlocal effects on the fitness of individuals. Here, we
+discuss superprocesses that account for such effects. The existence of
+diffusion limits for a class of measure-valued branching processes
+involving nonlocal effects has been treated by Méléard and Roelly (1993,
+1992). The interactions can manifest as dependencies of the spatial
+movement or reproductive law of individuals on their position \(x\) and
+the state of the whole population described by \(X_t\). An important
+result of Méléard and Roelly (1993, 1992) is a theorem that provides
+sufficient conditions to construct a sequence of rescaled measure-valued
+branching processes that converge to a generalization of the
+Dawson-Watanabe superprocess that includes interactions. The rescaling
+is analogous to that described above for non-interacting Dawson-Watanabe
+superprocesses, but now the reproductive law described by
+\(\mathscr{W}(X_t,x)\) and \(V(X_t,x)\), branching rate
+\(\lambda(X_t,x)\) and diffusion parameter \(\sqrt{\mu(X_t,x)}\) are
+allowed to depend on the whole population \(X_t\) and individual
+location \(x\). In Figure \ref{rescaled} we demonstrate this rescaling
+in discrete time for a population experiencing stabilizing selection and
+logistic growth. Since time is discretized, the process we simulate is
+formally a branching random walk. For further details on our simulation
+see SM \S\ref{numerical}.
 
@@
 
-{\centering \includegraphics{rescaled_plots}
+{\centering \includegraphics[width=0.8\linewidth]{rescaled_plots}
 
@@
 density-dependent growth rates and frequency-dependent selection.
-Perkins (1992, 1995) developed a theory of stochastic integration with
+Perkins (1995, 1992) developed a theory of stochastic integration with
 respect to the so-called \emph{Brownian trees} to characterize
@@
 uniqueness. Li (1998) built directly off of the construction of Méléard
-and Roelly (1992, 1993) to study properties of interacting
+and Roelly (1993, 1992) to study properties of interacting
 superprocesses and, by assuming individual spatial movement occurs
@@
 density \(\nu(x,t)\) which is non-negative, integrable, continuous in
-time and space and satisfies the SPDE \begin{equation}\label{SPDE}
+time and space and satisfies the SPDE
+
+\begin{equation}\label{SPDE}
 \frac{\partial}{\partial t}\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)+\sqrt{V\nu(x,t)}\dot W(x,t).
-\end{equation} As mentioned above, we refer to SPDE (\ref{SPDE}) as the
-Stochastic Asexual Gaussian allelic model with Abundance dynamics
-(abbreviated SAGA). Comparing equation (\ref{SPDE}) to equation (3.5) of
-Li (1998), our \(m\) and \(V\) correspond to Li's \(b\) and \(c\)
-respectively. Using techniques developed by Evans and Perkins (1994),
-Champagnat, Ferrière and Méléard (2006) generalized this result to the
-case when \(m(h,x)\) is bounded above and is linear or more than linear
-in \(h\). In particular, SAGA corresponds to a special case of equation
-(4.6) in Champagnat, Ferrière and Méléard (2006) and is valid for the
-assumptions on \(m\) made in \S\ref{deterministic}.
+\end{equation}
+
+As mentioned above, we refer to SPDE (\ref{SPDE}) as the Stochastic
+Asexual Gaussian allelic model with Abundance dynamics (abbreviated
+SAGA). Comparing equation (\ref{SPDE}) to equation (3.5) of Li (1998),
+our \(m\) and \(V\) correspond to Li's \(b\) and \(c\) respectively.
+Using techniques developed by Evans and Perkins (1994), Champagnat,
+Ferrière and Méléard (2006) generalized this result to the case when
+\(m(h,x)\) is bounded above and is linear or more than linear in \(h\).
+In particular, SAGA corresponds to a special case of equation (4.6) in
+Champagnat, Ferrière and Méléard (2006) and is valid for the assumptions
+on \(m\) made in \S\ref{deterministic}.
 
@@
 =N(0)+\int_0^t\bar m(s)N(s)dt+\int_0^t\sqrt{VN(s)}d\hat{\mathbf{W}}_s(\sqrt{\nu(x,s)}),
-\end{multline} where \begin{equation}
+\end{multline}
+
+where
+
+\begin{equation}
 \bar m(t)=\frac{1}{N(t)}\int_\mathbb{R}m(\nu,x)\nu(x,t)dx,
-\end{equation} and \begin{equation}
+\end{equation}
+
+and
+
+\begin{equation}
 \int_0^td\hat{\mathbf{W}}_s(\sqrt{\nu(x,s)})=\int_0^t\int_\mathbb{R}\frac{\sqrt{\nu(x,s)}}{\sqrt{\int_\mathbb{R}\nu(x,s)dx}}\dot W(x,s)dxds.
-\end{equation} Setting \(W_1(t)=\hat{\mathbf{W}}_t(\sqrt{\nu(x,t)})\),
-we can use traditional stochastic differential notation to write
+\end{equation}
+
+Setting \(W_1(t)=\hat{\mathbf{W}}_t(\sqrt{\nu(x,t)})\), we can use
+traditional stochastic differential notation to write
+
 \begin{equation}
 dN=\bar mNdt+\sqrt{VN}dW_1.
-\end{equation} To find the associated SDE for \(\bar x(t)\) and
-\(\sigma^2(t)\), we want to repeat the same approach for \(f(x)=x,x^2\)
-and apply Itô's lemma. However, for these cases
-\(f\notin C^2_b(\mathbb{R})\) since \(f\) will not be bounded. But if we
-can show \(\int_\mathbb{R}(|x|+x^2+x^4)\nu(x,t)dx<+\infty\) for all
-\(t>0\) given this condition is satisfied by \(\nu(x,0)\), then we can
-apply the mild solution of (\ref{SPDE}) to derive SDE for \(\bar x(t)\)
-and \(\sigma^2(t)\). To illustrate, let us suppose this is the case.
-Setting \(\tilde x(t)=\int_\mathbb{R}x\nu(x,t)dx\), we have
+\end{equation}
+
+To find the associated SDE for \(\bar x(t)\) and \(\sigma^2(t)\), we
+want to repeat the same approach for \(f(x)=x,x^2\) and apply Itô's
+lemma. However, for these cases \(f\notin C^2_b(\mathbb{R})\) since
+\(f\) will not be bounded. But if we can show
+\(\int_\mathbb{R}(|x|+x^2+x^4)\nu(x,t)dx<+\infty\) for all \(t>0\) given
+this condition is satisfied by \(\nu(x,0)\), then we can apply the mild
+solution of (\ref{SPDE}) to derive SDE for \(\bar x(t)\) and
+\(\sigma^2(t)\). To illustrate, let us suppose this is the case. Setting
+\(\tilde x(t)=\int_\mathbb{R}x\nu(x,t)dx\), we have
+
 \begin{equation}
 \tilde x(t)=\tilde x(0)+\int_0^t\int_\mathbb{R}\nu(x,s)m(\nu,x)x+x\sqrt{V\nu(x,s)}\dot W(x,s)dxds.
-\end{equation} Similarly, setting
+\end{equation}
+
+Similarly, setting
 \(\tilde{\tilde\sigma}^2(t)=\int_\mathbb{R}x^2\nu(x,t)dx\), we have
+
 \begin{equation}
 \tilde{\tilde\sigma}^2(t)=\tilde{\tilde\sigma}^2(0)+\int_0^t\int_\mathbb{R}\nu(x,s)\left(m(\nu,x)x^2+\mu\right)+x^2\sqrt{V\nu(x,s)}\dot W(x,s)dxds.
-\end{equation} Since \(\bar x(t)=\tilde x(t)/N(t)\) and
+\end{equation}
+
+Since \(\bar x(t)=\tilde x(t)/N(t)\) and
 \(\sigma^2(t)=\tilde{\tilde\sigma}^2(t)/N(t)-\bar x^2(t)\), we can use
@@
 genetic drift on the change in mean trait over a single generation
-derived using sampling arguments (Lande 1976). There is also an
+derived using sampling arguments (Lande, 1976). There is also an
 interesting connection with classical population genetics. A fundamental
@@
 diversity due to the chance loss of alleles in finite populations
-(Fisher 1923; Wright 1931). This expected reduction in diversity due to
-random genetic drift is captured by the third term in the deterministic
-component of expression (\ref{sig2_gen}), particularly
+(Fisher, 1923; Wright, 1931). This expected reduction in diversity due
+to random genetic drift is captured by the third term in the
+deterministic component of expression (\ref{sig2_gen}), particularly
 \(-V\sigma^2(t)/N(t)\). The component of SDE (\ref{sig2_gen}) describing
@@
 with a simplified model of mutation and asexual reproduction or random
-mating (Lande 1980; Turelli 1984, 1986, 2017; Bürger 2000).
+mating (Bürger, 2000; Lande, 1980; Turelli, 2017, 1986, 1984).
 Alternatively, it has been shown that a Gaussian distribution can
 provide a reasonable approximation even when selection is strong and
-non-Gaussian (Turelli and Barton 1994). Hence, given appropriate
+non-Gaussian (Turelli and Barton, 1994). Hence, given appropriate
 assumptions on selection, mutation and reproduction, the abundance
@@
 population sizes. Allowing for these restrictions, we assume
+
 \begin{equation}
 \nu(x,t)=\frac{N(t)}{\sqrt{2\pi\sigma^2(t)}}\exp\left(-\frac{\big(x-\bar x(t)\big)^2}{2\sigma^2(t)}\right).
-\end{equation} Under this assumption we find in SM \S\ref{cov2deriv} the
-results (suppressing the dependency on \(t\))
+\end{equation}
+
+Under this assumption we find in SM \S\ref{cov2deriv} the results
+(suppressing the dependency on \(t\))
+
 \begin{equation}\label{covxm}
@@
 \mathrm{Cov}\Big((x-\bar x)^2,m\Big)=2\sigma^4\left(\frac{\partial\bar m}{\partial\sigma^2}-\overline{\frac{\partial m}{\partial\sigma^2}}\right)
-\end{equation} and \(\overline{(x-\bar x)^4}=3\sigma^4\). In particular,
-these results imply
+\end{equation}
+
+and \(\overline{(x-\bar x)^4}=3\sigma^4\). In particular, these results
+imply
 
@@
 \(\sigma^2\) explained by additive effects of different alleles among
-genetic loci encoding for the focal phenotype (Roughgarden 1979; Bulmer
-1980; Lynch and Walsh 1998). It is therefore fitting that \(G\) is
+genetic loci encoding for the focal phenotype (Bulmer, 1980; Lynch and
+Walsh, 1998; Roughgarden, 1979). It is therefore fitting that \(G\) is
 referred to as the additive genetic variance.
@@
 alleles on an expressed trait are additive, \(\eta\) is known as the
-\emph{variance of environmental deviation} (Lande 1975; Lynch and Walsh
-1998). For a given breeding value, we denote the probability density of
-a randomly drawn expressed trait by \(\psi(x,g)\). Hence,
+\emph{variance of environmental deviation} (Lande, 1975; Lynch and
+Walsh, 1998). For a given breeding value, we denote the probability
+density of a randomly drawn expressed trait by \(\psi(x,g)\). Hence,
+
 \begin{equation}
 \psi(x,g)=\frac{1}{\sqrt{2\pi\eta}}\exp\left(-\frac{(x-g)^2}{2\eta}\right).
-\end{equation} To include this relationship in our framework, we write
-\(\rho(g,t)\) as the abundance density of breeding values at time \(t\)
-so that
+\end{equation}
+
+To include this relationship in our framework, we write \(\rho(g,t)\) as
+the abundance density of breeding values at time \(t\) so that
 \(\int_{-\infty}^{+\infty}\rho(g,t)dg=\int_{-\infty}^{+\infty}\nu(x,t)dx=N(t)\).
@@
 \(\rho(g,t)\) is determined, we can compute \(\nu(x,t)\) via
+
 \begin{equation}\label{nu_from_rho}
 \nu(x,t)=\int_{-\infty}^{+\infty}\rho(g,t)\psi(x,g)dg.
-\end{equation} However, since selection acts on expressed phenotypes, we
-use the assumed relationship between breeding values and expressed
-traits to calculate the fitness of breeding values. To motivate the
-approach taken, consider the problem of inferring the breeding value of
-an individual given its expressed trait \(x\). Denote \(\mathfrak{g}\) a
+\end{equation}
+
+However, since selection acts on expressed phenotypes, we use the
+assumed relationship between breeding values and expressed traits to
+calculate the fitness of breeding values. To motivate the approach
+taken, consider the problem of inferring the breeding value of an
+individual given its expressed trait \(x\). Denote \(\mathfrak{g}\) a
 random variable representing the unkown breeding value. Under this model
@@
 across all individuals carrying \(g\) can be written as
+
 \begin{equation}\label{m_rel}
 m^*(\rho,g)=\int_{-\infty}^{+\infty}m(\nu,x)\psi(x,g)dx.
-\end{equation} This is similar to the approach taken by Kimura and Crow
-(1978) to calculate the overall effects of selection for expressed
-characters onto the changes in the distribution of alleles encoding
-those characters. However, instead of focusing on the frequencies of
-alleles at particular loci, our results focus on the densities of
-breeding values. With the relationship between \(m(\nu,x)\) and
-\(m^*(\rho,g)\) established, we define the evolution of \(\rho(g,t)\) by
-the SPDE \begin{equation}\label{rho_SPDE}
+\end{equation}
+
+This is similar to the approach taken by Kimura and Crow (1978) to
+calculate the overall effects of selection for expressed characters onto
+the changes in the distribution of alleles encoding those characters.
+However, instead of focusing on the frequencies of alleles at particular
+loci, our results focus on the densities of breeding values. With the
+relationship between \(m(\nu,x)\) and \(m^*(\rho,g)\) established, we
+define the evolution of \(\rho(g,t)\) by the SPDE
+
+\begin{equation}\label{rho_SPDE}
 \dot\rho(g,t)=m^*(\rho,g)\rho(g,t)+\frac{\mu}{2}\frac{\partial^2}{\partial^2 g}\rho(g,t)+\sqrt{V\rho(g,t)}\dot W(g,t).
-\end{equation} Equation (\ref{rho_SPDE}) is a stochastic generalization
-of the deterministic PDE (\ref{eq1}) from \S\ref{deterministic}, but
-describes the evolution of the distribution of breeding values instead
-of expressed characters. However, whether modelling expressed characters
-or breeding values, we refer to SPDE of the form (\ref{rho_SPDE}) as
+\end{equation}
+
+Equation (\ref{rho_SPDE}) is a stochastic generalization of the
+deterministic PDE (\ref{eq1}) from \S\ref{deterministic}, but describes
+the evolution of the distribution of breeding values instead of
+expressed characters. However, whether modelling expressed characters or
+breeding values, we refer to SPDE of the form (\ref{rho_SPDE}) as
 Stochastic Asexual Gaussian allelic models with Abundance dynamics
@@
 fitness function. Our approach mirrors closely the theory developed by
-MacArthur and Levins (1967), Levins (1968) and MacArthur (1969, 1970,
-1972). The most significant difference, aside from allowing evolution to
+MacArthur and Levins (1967), Levins (1968) and MacArthur (1972, 1970,
+1969). The most significant difference, aside from allowing evolution to
 occur, is the treatment of resource quality, which we replace with a
@@
 niche axes exhibit overlap on at least one of the axes, but no overall
-niche overlap (Holt 1987). Thus niche overlap on lower-dimensional
+niche overlap (Holt, 1987). Thus niche overlap on lower-dimensional
 projections of some multivariate niche space does not imply the
@@
 stochasticity, spatial dispersion and locally finite carrying capacity
-(Dawson 1978). In relation to the example fitness function discussed in
+(Dawson, 1978). In relation to the example fitness function discussed in
 \S\ref{deterministic}, consider
@@
 \begin{multline}\label{N_comm}
-dN_i = \Bigg\{R_i-\frac{a_i}{2}\Big((\bar x_i-\theta_i)^2+G_i+\eta_i\Big) - c_i\sum_{j=1}^SN_jU_iU_j\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\Bigg\}N_idt + \sqrt{V_iN_i}dW_1,
+dN_i = \Bigg\{R_i-\frac{a_i}{2}\Big((\bar x_i-\theta_i)^2+G_i+\eta_i\Big) - c_i\sum_{j=1}^SN_jU_iU_j\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\Bigg\}N_idt \\ + \sqrt{V_iN_i}dW_1,
 \end{multline}
 \begin{multline}\label{x_comm}
-d\bar x_i = \Bigg\{a_iG_i(\theta_i-\bar x_i)-c_iG_i\bigg(\sum_{j=1}^SN_jU_iU_jb_{ij}(\bar x_j-\bar x_i)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\bigg)\Bigg\}dt+\sqrt{V_i\frac{G_i}{N_i}}dW_2,
+d\bar x_i = \Bigg\{a_iG_i(\theta_i-\bar x_i)-c_iG_i\bigg(\sum_{j=1}^SN_jU_iU_jb_{ij}(\bar x_j-\bar x_i)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\bigg)\Bigg\}dt \\ +\sqrt{V_i\frac{G_i}{N_i}}dW_2,
 \end{multline}
 \begin{multline}\label{G_comm}
-dG_i =  \Bigg\{c_i{G_i}^2\bigg(N_iU_i^2b_{ii}\sqrt\frac{b_{ii}}{2\pi}+\sum_{j=1}^SN_jU_iU_jb_{ij}\left(1-b_{ij}(\bar x_i-\bar x_j)^2\right)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2}\bigg) \\
-+\mu_i-a_i{G_i}^2-V_i\frac{G_i}{N_i}\Bigg\}dt+G_i\sqrt{\frac{2V_i}{N_i}}dW_3,
+dG_i =  \Bigg\{c_i{G_i}^2\bigg(\sum_{j=1}^SN_jU_iU_jb_{ij}\left(1-b_{ij}(\bar x_i-\bar x_j)^2\right)\sqrt{\frac{b_{ij}}{2\pi}}e^{-\frac{b_{ij}}{2}(\bar x_i-\bar x_j)^2} \\
++N_iU_i^2b_{ii}\sqrt\frac{b_{ii}}{2\pi}+\mu_i-a_i{G_i}^2-V_i\frac{G_i}{N_i}\bigg)\Bigg\}dt+G_i\sqrt{\frac{2V_i}{N_i}}dW_3,
 \end{multline}
@@
 referred to in the community ecology literature as character
-displacement (Brown and Wilson 1956). We also see that the fitness of
+displacement (Brown and Wilson, 1956). We also see that the fitness of
 species \(i\) drops quadratically with the difference between
@@
 Relating our treatment of resource competition to modern coexistence
-theory (Chesson 2000), the absolute competition coefficient
+theory (Chesson, 2000), the absolute competition coefficient
 \(\alpha_{ij}\) becomes a dynamical quantity that can be written as
+
+\begin{multline}
+\alpha_{ij}(t)=\frac{c_i}{r_i(t)}\int_{-\infty}^{+\infty}\int_{-\infty}^{+\infty}p_i(x,t)p_j(y,t)\mathcal O_{ij}(x,y) dxdy \\ =\frac{c_iU_iU_j}{r_i(t)}\sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right),
+\end{multline}
+
+where
+
 \begin{equation}
-\alpha_{ij}(t)=\frac{c_i}{r_i(t)}\int_{-\infty}^{+\infty}\int_{-\infty}^{+\infty}p_i(x,t)p_j(y,t)\mathcal O_{ij}(x,y) dxdy =\frac{c_iU_iU_j}{r_i(t)}\sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right),
-\end{equation} where \begin{equation}
 r_i(t)=R_i-\frac{a_i}{2}\Big((\bar x_i(t)-\theta_i)^2+G_i(t)+\eta_i\Big).
-\end{equation} Hence, \(dN_i(t)\) can be expressed as \begin{equation}
+\end{equation}
+
+Hence, \(dN_i(t)\) can be expressed as
+
+\begin{equation}
 dN_i(t)=r_i(t)\left(1-\sum_{j=1}^S\alpha_{ij}(t)N_j(t)\right)N_i(t)dt+\sqrt{V_iN_i(t)}dW_1(t).
-\end{equation} Note that although \(r_i(t)\) is referred to in the
-coexistence literature as the intrinsic growth rate of the population,
-\(R_i\) is a deeper intrinsic quantity. For now we refer to \(R_i\) as
-the \emph{innate} growth rate. Previous work has shown the importance of
-demographic stochasticity (Turelli 1980; Schreiber 2017) and
-evolutionary dynamics (Case and Taper 1986; Schreiber, Patel, and
-terHorst 2018) in determining coexistence of competing species. However,
-these studies consider evolution and demographic stochasticity
-separately. To our knowledge, there are no rigorous theoretical
-investigations of coexistence that account for evolutionary dynamics of
-quantitative characters and demographic stochasticity simultaneously
-and, in particular, no studies combining these processes that also
-include the evolution of additive genetic variance. With the connection
-outlined in this section formally established, researchers may pursue a
-postmodern coexistence theory that naturally includes both the
-evolutionary dynamics of quantitative characters and the effects of
-demographic stochasticity in a simple synthetic framework.
+\end{equation}
+
+Note that although \(r_i(t)\) is referred to in the coexistence
+literature as the intrinsic growth rate of the population, \(R_i\) is a
+deeper intrinsic quantity. For now we refer to \(R_i\) as the
+\emph{innate} growth rate. Previous work has shown the importance of
+demographic stochasticity (Schreiber, 2017; Turelli, 1980) and
+evolutionary dynamics (Case and Taper, 1986; Schreiber et al., 2018) in
+determining coexistence of competing species. However, these studies
+consider evolution and demographic stochasticity separately. To our
+knowledge, there are no rigorous theoretical investigations of
+coexistence that account for evolutionary dynamics of quantitative
+characters and demographic stochasticity simultaneously and, in
+particular, no studies combining these processes that also include the
+evolution of additive genetic variance. With the connection outlined in
+this section formally established, researchers may pursue a postmodern
+coexistence theory that naturally includes both the evolutionary
+dynamics of quantitative characters and the effects of demographic
+stochasticity in a simple synthetic framework.
 
 In SM \S\ref{diffuse} we show that the standardized directional
-selection gradient (sensu Lande and Arnold 1983) induced by species
-\(j\) on species \(i\) can be computed as \begin{equation}
-\beta_{ij}(t)=c_iU_iU_jN_j(t)b_{ij}(t)\big(\bar x_i(t)-\bar x_j(t)\big) \sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right).
-\end{equation} Our notation differs from Lande and Arnold (1983) in that
-subscripts here denote species instead of components of multivariate
-traits and we drop the prime that distinguishes between selection
-gradients and standardized selection gradients.
-
-\textbf{Metric of pairwise coevolution}
-
-Below we investigate the correspondence of interaction intensity and
-coevolutionary change. However, we can already identify one major
-discrepancy; \(\alpha_{ij}\) is maximized when \(\bar x_i=\bar x_j\),
-but \(\beta_{ij}=0\) under the same condition. We therefore include in
-our metric of selection the standardized stabilizing selection gradient
-\(\gamma\) which measures the effect of stabilizing or disruptive
-selection on phenotypic variance (Lande and Arnold 1983). In SM
-\S\ref{diffuse} we show that the standardized stabilizing selection
-gradient induced by species \(j\) on species \(i\) can be computed as
-\begin{equation}
-\gamma_{ij}(t)=c_iU_iU_jN_j(t)b_{ij}(t)\left(1-b_{ij}(t)\big(\bar x_i(t)-\bar x_j(t)\big)^2\right) \sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right).
-\end{equation} To measure evolutionary change in species \(i\) induced
-by species \(j\), we form the metric
-\(\Psi_{ij}=|\beta_{ij}|+|\gamma_{ij}|\) and define the strength of
-coevolution as \(\mathfrak{C}_{ij}=\Psi_{ij}\Psi_{ji}\). Figure
-\ref{net} displays the joint distribution of \(\mathfrak{C}_{ij}\) and
-the product of competition coefficients \(\alpha_{ij}\alpha_{ji}\) for
-two simulated communities, both with richness \(S=1000\). The solid
-contour represents the case of relatively high sensitivity to
-competition (\(c=5\times10^{-7}\)) and the dashed contour represents the
-case of relatively weak sensitivity to competition
-(\(c=1\times10^{-8}\)). In both cases we see that \(\mathfrak{C}_{ij}\)
-and \(\alpha_{ij}\alpha_{ji}\) are, essentially, unrelated. However,
-this depiction is only representative of a specific set of parameters.
-Next, we provide analytical approximations of the covariance between
-selection gradients and competition coefficients and a numerical
-estimate for the relationship between pairwise coevolution and
-competition coefficients for a range of parameters.
+selection gradient (sensu Lande and Arnold, 1983) induced by species
+\(j\) on species \(i\) can be computed as
+
+\begin{multline}
+\beta_{ij}(t)=c_iU_iU_jN_j(t)b_{ij}(t)\big(\bar x_i(t)-\bar x_j(t)\big) \\
+\times\sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right).
+\end{multline}
+
+Our notation differs from Lande and Arnold (1983) in that subscripts
+here denote species instead of components of multivariate traits and we
+drop the prime that distinguishes between selection gradients and
+standardized selection gradients.
+
+\emph{Metric of pairwise coevolution:} Below we investigate the
+correspondence of interaction intensity and coevolutionary change.
+However, we can already identify one major discrepancy; \(\alpha_{ij}\)
+is maximized when \(\bar x_i=\bar x_j\), but \(\beta_{ij}=0\) under the
+same condition. We therefore include in our metric of selection the
+standardized stabilizing selection gradient \(\gamma\) which measures
+the effect of stabilizing or disruptive selection on phenotypic variance
+(Lande and Arnold, 1983). In SM \S\ref{diffuse} we show that the
+standardized stabilizing selection gradient induced by species \(j\) on
+species \(i\) can be computed as
+
+\begin{multline}
+\gamma_{ij}(t)=c_iU_iU_jN_j(t)b_{ij}(t)\left(1-b_{ij}(t)\big(\bar x_i(t)-\bar x_j(t)\big)^2\right) \\
+\times\sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right).
+\end{multline}
+
+To measure evolutionary change in species \(i\) induced by species
+\(j\), we form the metric \(\Psi_{ij}=|\beta_{ij}|+|\gamma_{ij}|\) and
+define the strength of coevolution as
+\(\mathfrak{C}_{ij}=\Psi_{ij}\Psi_{ji}\). Figure \ref{net} displays the
+joint distribution of \(\mathfrak{C}_{ij}\) and the product of
+competition coefficients \(\alpha_{ij}\alpha_{ji}\) for two simulated
+communities, both with richness \(S=1000\). The solid contour represents
+the case of relatively high sensitivity to competition
+(\(c=5\times10^{-7}\)) and the dashed contour represents the case of
+relatively weak sensitivity to competition (\(c=1\times10^{-8}\)). In
+both cases we see that \(\mathfrak{C}_{ij}\) and
+\(\alpha_{ij}\alpha_{ji}\) are, essentially, unrelated. However, this
+depiction is only representative of a specific set of parameters. Next,
+we provide analytical approximations of the covariance between selection
+gradients and competition coefficients and a numerical estimate for the
+relationship between pairwise coevolution and competition coefficients
+for a range of parameters.
 
@@
 
-{\centering \includegraphics{on_pl}
+{\centering \includegraphics[width=0.6\linewidth]{on_pl}
 
@@
 
-\textbf{Covariance of selection and competition as a function of
-diversity}
-
-We now make use of the expressions derived for competition coeffecients
-and selection gradients to investigate their relationship. As a first
-pass, let us assume all model parameters are equivalent across species
-and that each species has the same abundance and trait variance. Let us
-further assume that richness \(S\) is large and the distribution of mean
-trait values is normal with mean \(\bar{\bar x}\), variance
-\(V_{\bar X}\) and density \(f_{\bar X}\). Such assumptions are typical
-when deriving analytical results in the field of theoretical
-coevolutionary community ecology (Nuismer, Jordano, and Bascompte 2012;
-Nuismer, Week, and Aizen 2018). If \(\bar{\bar x}\) is near \(\theta\)
-and \(V_{\bar X}\) is much smaller than \(|2\frac{R}{a}-G-\eta|\), then
-we may approximate \(r_i\) with \begin{equation}
-\bar r=\int_{-\infty}^{+\infty}\left(R-\frac{a}{2}\Big((\bar x-\theta)^2+G+\eta\Big)\right)f_{\bar X}(\bar x)d\bar x
+\emph{Covariance of selection and competition as a function of
+diversity:} We now make use of the expressions derived for competition
+coeffecients and selection gradients to investigate their relationship.
+As a first pass, let us assume all model parameters are equivalent
+across species and that each species has the same abundance and trait
+variance. Let us further assume that richness \(S\) is large and the
+distribution of mean trait values is normal with mean \(\bar{\bar x}\),
+variance \(V_{\bar X}\) and density \(f_{\bar X}\). Such assumptions are
+typical when deriving analytical results in the field of theoretical
+coevolutionary community ecology (Nuismer et al., 2012, 2018). If
+\(\bar{\bar x}\) is near \(\theta\) and \(V_{\bar X}\) is much smaller
+than \(|2\frac{R}{a}-G-\eta|\), then we may approximate \(r_i\) with
+
+\begin{multline}
+\bar r=\int_{-\infty}^{+\infty}\left(R-\frac{a}{2}\Big((\bar x-\theta)^2+G+\eta\Big)\right)f_{\bar X}(\bar x)d\bar x \\
 =R-\frac{a}{2}\Big((\bar{\bar x}-\theta)^2+V_{\bar X}+G+\eta\Big).
-\end{equation} In SM \S\ref{coef_grad_moms} we use these assumptions to
-calculate the first and second order moments describing the joint
-distribution of competition coefficients and selection gradients across
-the community. We find that the covariance between linear selection
-gradients and competition coeffecients are zero due to the symmetry
-implied by our assumptions. However, setting
-\(\alpha(\bar x_i,\bar x_j)=\alpha_{ij}\),
+\end{multline}
+
+In SM \S\ref{coef_grad_moms} we use these assumptions to calculate the
+first and second order moments describing the joint distribution of
+competition coefficients and selection gradients across the community.
+We find that the covariance between linear selection gradients and
+competition coeffecients are zero due to the symmetry implied by our
+assumptions. However, setting \(\alpha(\bar x_i,\bar x_j)=\alpha_{ij}\),
 \(\beta(\bar x_i,\bar x_j)=\beta_{ij}\) and
@@
 \begin{subequations}\label{cov_eqns}
-\begin{equation}\label{cov_alpha_beta}
-\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)=\frac{2c^2b^2U^4N}{\pi\bar r}\sqrt\frac{V_{\bar X}}{2\pi}\left(\frac{1}{(1+8bV_{\bar X})^{3/4}}-\frac{1}{(1+4bV_{\bar X})^{3/4}(1+2bV_{\bar X})^{1/2}}\right),
-\end{equation}
+\begin{multline}\label{cov_alpha_beta}
+\mathrm{Cov}_{f_{\bar X}}(\alpha,|\beta|)=\frac{2c^2b^2U^4N}{\pi\bar r}\sqrt\frac{V_{\bar X}}{2\pi} \\
+\times\left(\frac{1}{(1+8bV_{\bar X})^{3/4}}-\frac{1}{(1+4bV_{\bar X})^{3/4}(1+2bV_{\bar X})^{1/2}}\right),
+\end{multline}
 \begin{equation}\label{cov_alpha_gamma}
@@
 
-{\centering \includegraphics{/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/cov}
+{\centering \includegraphics[width=1\linewidth]{/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/cov}
 
@@
 
-{\centering \includegraphics{/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/Cαℭp}
+{\centering \includegraphics[width=0.6\linewidth]{/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/Cαℭp}
 
@@
 
-\textbf{Limitations of diffusion limits and generalizing coexistence
-theory}
+\hypertarget{limitations-of-diffusion-limits-and-generalizing-coexistence-theory}{%
+\subsection{Limitations of diffusion limits and generalizing coexistence
+theory}\label{limitations-of-diffusion-limits-and-generalizing-coexistence-theory}}
 
@@
 \S\ref{ecoevo} on using our model of diffuse coevolution to extend
-modern coexistence theory (sensu Chesson 2000) in a way that accounts
+modern coexistence theory (sensu Chesson, 2000) in a way that accounts
 for phenotypic evolution and demographic stochasticity. Just as previous
@@
 
-\textbf{Genetic architecture and intraspecific trait distributions}
+\hypertarget{genetic-architecture-and-intraspecific-trait-distributions}{%
+\subsection{Genetic architecture and intraspecific trait
+distributions}\label{genetic-architecture-and-intraspecific-trait-distributions}}
 
@@
 mathematical approaches to derive the dynamics of these higher moments
-(Débarre, Yeaman, and Guillaume 2015).
+(Débarre et al., 2015).
 
@@
 normality. This approach has a long history in theoretical quantitative
-genetics (Bulmer 1980; Turelli and Barton 1994; Kirkpatrick, Johnson,
-and Barton 2002) and coevolutionary theory (Nuismer, Doebeli, and
-Browning 2005; Kopp and Gavrilets 2006; Nuismer, Ridenhour, and Oswald
-2007). However, work to investigate the role of complex genetic
-architectures in mediating feedbacks between the dynamics of population
-abundances and the distributions of traits mediating ecological
-interactions has apparently only just begun (Patel and Bürger 2019).
-Since analytical results are difficult to obtain for general models of
-mutation, development and selection (c.f., Bürger 2000), we again
-suggest the use of measure-valued branching processes and their
-individual-based simulations. For example, by assuming the mutational
-effects in the BBM model is Gaussian with variance an order of magnitude
-greater than the equilibrium phenotypic (or additive genetic) variance,
-we obtain the \emph{House-of-cards} model (Turelli 1984). However,
-perhaps considering heavy-tailed distributions such as the
-Laplace-distributions considered by Park et. al. (2017) will provide
+genetics (Bulmer, 1980; Kirkpatrick et al., 2002; Turelli and Barton,
+1994) and coevolutionary theory (Kopp and Gavrilets, 2006; Nuismer et
+al., 2005, 2007). However, work to investigate the role of complex
+genetic architectures in mediating feedbacks between the dynamics of
+population abundances and the distributions of traits mediating
+ecological interactions has apparently only just begun (Patel and
+Bürger, 2019). Since analytical results are difficult to obtain for
+general models of mutation, development and selection (c.f., Bürger,
+2000), we again suggest the use of measure-valued branching processes
+and their individual-based simulations. For example, by assuming the
+mutational effects in the BBM model is Gaussian with variance an order
+of magnitude greater than the equilibrium phenotypic (or additive
+genetic) variance, we obtain the \emph{House-of-cards} model (Turelli,
+1984). However, perhaps considering heavy-tailed distributions such as
+the Laplace-distributions considered by Park et. al. (2017) will provide
 more realistic yet less analytically tractable models of mutational
@@
 conservation genetics where small populations evolve primarily in
-response to mutation and random genetic drift (Lande 1995).
-
-\textbf{Ecological stoichiometry}
+response to mutation and random genetic drift (Lande, 1995).
+
+\hypertarget{ecological-stoichiometry}{%
+\subsection{Ecological stoichiometry}\label{ecological-stoichiometry}}
 
@@
 dynamics have been captured theoretically in consumer-resource models
-(Tilman 1982), developing a more realistic model of resource competition
-must incorporate details on the ecophysiology of organisms (Loreau
-2010). Doing so will help clarify the feedback between the evolution of
-populations and the ecosystem processes they are a part of.
+(Tilman, 1982), developing a more realistic model of resource
+competition must incorporate details on the ecophysiology of organisms
+(Loreau, 2010). Doing so will help clarify the feedback between the
+evolution of populations and the ecosystem processes they are a part of.
 
@@
 
-\textbf{Accounting for macroevolutionary history}
+\hypertarget{accounting-for-macroevolutionary-history}{%
+\subsection{Accounting for macroevolutionary
+history}\label{accounting-for-macroevolutionary-history}}
 
@@
 regional species pool is filtered for those species adapted to the newly
-available habitat (Marx et al. 2017).
+available habitat (Marx et al., 2017).
 
@@
 repurpose microevolutionary models by blindly rescaling time from the
-units of generations to millions of years (e.g., Nuismer and Harmon
+units of generations to millions of years (e.g., Nuismer and Harmon,
 2014). Such an approach makes the implicit assumption that trait
-evolution is statistically self-similar (sensu Falconer 2014) so that
+evolution is statistically self-similar (sensu Falconer, 2014) so that
 the stochastic evolution of traits on macroevolutionary time scales has
@@
 successfully employed in comparative phylogenetics to fit phenotypic
-data from extant populations and the fossil record (Landis and Schraiber
-2017). It would therefore be interesting to investigate whether an
-application of a separation of time scales argument for the rate of
-environmental change (\(\lambda\)) versus the rate of evolutionary and
-ecological change (\(a\)) to microevolutionary models derived via our
-framework can be used to obtain macroevolutionary models that include
-not only mean trait evolution, but also the evolution of trait variance
-and abundance. The resulting macroevolutionary models can give rise to
-novel comparative phylogenetic methods and provide initial conditions
-for microevolutionary models that capture contemporary dynamics.
-
-\textbf{Final remarks}
+data from extant populations and the fossil record (Landis and
+Schraiber, 2017). It would therefore be interesting to investigate
+whether an application of a separation of time scales argument for the
+rate of environmental change (\(\lambda\)) versus the rate of
+evolutionary and ecological change (\(a\)) to microevolutionary models
+derived via our framework can be used to obtain macroevolutionary models
+that include not only mean trait evolution, but also the evolution of
+trait variance and abundance. The resulting macroevolutionary models can
+give rise to novel comparative phylogenetic methods and provide initial
+conditions for microevolutionary models that capture contemporary
+dynamics.
+
+\hypertarget{final-remarks}{%
+\section{Final remarks}\label{final-remarks}}
 
 Although top-down approaches to community ecology, such as the Maximum
-Entropy Theory of Ecology (Harte 2011), have enjoyed some success in
-describing community-level patterns (Harte and Newman 2014; Xiao,
-McGlinn, and White 2015), a mechanistic understanding of why these
-patterns emerge and how they will change remains a formidable task. Such
-an understanding must take both bottom-up and top-down approaches
-integrating considerations from the ecophysiology of individual
-organisms that reveal the economics of interspecific interactions
-(Sterner and Elser 2008), to the phylogeographic history of taxa that
-sets the stage for contemporary dynamics (Hickerson et al. 2010).
-Through connecting these dots we can increase the variance explained in
-observations of ecological communities by specific mechanisms and come
-closer to a predictive theory of evolutionary community ecology. Despite
-the long list of equations derived in this paper, this work takes just
-one small step towards capturing these many details. However, we hope
-the theoretical framework outlined here along with the demonstration of
-its use in modelling competitive communities provides some helpful tools
-to aid quantitative evolutionary ecologists in reaching such lofty
-goals.
-
-\newpage
+Entropy Theory of Ecology (Harte, 2011), have enjoyed some success in
+describing community-level patterns (Harte and Newman, 2014; Xiao et
+al., 2015), a mechanistic understanding of why these patterns emerge and
+how they will change remains a formidable task. Such an understanding
+must take both bottom-up and top-down approaches integrating
+considerations from the ecophysiology of individual organisms that
+reveal the economics of interspecific interactions (Sterner and Elser,
+2008), to the phylogeographic history of taxa that sets the stage for
+contemporary dynamics (Hickerson et al., 2010). Through connecting these
+dots we can increase the variance explained in observations of
+ecological communities by specific mechanisms and come closer to a
+predictive theory of evolutionary community ecology. Despite the long
+list of equations derived in this paper, this work takes just one small
+step towards capturing these many details. However, we hope the
+theoretical framework outlined here along with the demonstration of its
+use in modelling competitive communities provides some helpful tools to
+aid quantitative evolutionary ecologists in reaching such lofty goals.
 
@@
 \leavevmode\hypertarget{ref-AbdalaRoberts2014}{}%
-Abdala-Roberts, Luis, and Kailen A. Mooney. 2014. ``Ecological and
-Evolutionary Consequences of Plant Genotype Diversity in a Tri-Trophic
-System.'' \emph{Ecology} 95 (10). Wiley: 2879--93.
+Abdala-Roberts, L., Mooney, K.A., 2014. Ecological and evolutionary
+consequences of plant genotype diversity in a tri-trophic system.
+Ecology 95, 2879--2893.
 
 \leavevmode\hypertarget{ref-Athreya1972}{}%
-Athreya, Krishna B., and Peter E. Ney. 1972. \emph{Branching Processes}.
-Springer Berlin Heidelberg.
+Athreya, K.B., Ney, P.E., 1972. Branching processes. Springer Berlin
+Heidelberg.
 
 \leavevmode\hypertarget{ref-9781107648258}{}%
-Ågren, Göran I., and Folke O. Andersson. 2012. \emph{Terrestrial
-Ecosystem Ecology: Principles and Applications}. Cambridge University
+Ågren, G.I., Andersson, F.O., 2012. Terrestrial ecosystem ecology:
+Principles and applications. Cambridge University Press.
+
+\leavevmode\hypertarget{ref-Barton2019}{}%
+Barton, N., Etheridge, A., 2019. Mathematical models in population
+genetics.
+
+\leavevmode\hypertarget{ref-Barton2013}{}%
+Barton, N., Etheridge, A., Véber, A., 2013. Modelling evolution in a
+spatial continuum. Journal of Statistical Mechanics: Theory and
+Experiment 2013, P01002.
+
+\leavevmode\hypertarget{ref-Barton2017}{}%
+Barton, N., Etheridge, A., Véber, A., 2017. The infinitesimal model:
+Definition, derivation, and implications. Theoretical Population Biology
+118, 50--73.
+
+\leavevmode\hypertarget{ref-Bertoin2003}{}%
+Bertoin, J., Le Gall, J.-F., 2003. Stochastic flows associated to
+coalescent processes. Probability Theory and Related Fields 126,
+261--288.
+
+\leavevmode\hypertarget{ref-Brown1956}{}%
+Brown, W.L., Wilson, E.O., 1956. Character displacement. Systematic
+Zoology 5, 49.
+
+\leavevmode\hypertarget{ref-bulmer1980}{}%
+Bulmer, 1980. The mathematical theory of quantitative genetics. Oxford
+University Press.
+
+\leavevmode\hypertarget{ref-Brger1986}{}%
+Bürger, R., 1986. On the maintenance of genetic variation: Global
+analysis of kimuras continuum-of-alleles model. Journal of Mathematical
+Biology 24, 341--351.
+
+\leavevmode\hypertarget{ref-9780471986539}{}%
+Bürger, R., 2000. The mathematical theory of selection, recombination,
+and mutation. Wiley.
+
+\leavevmode\hypertarget{ref-Cantrell2004}{}%
+Cantrell, R.S., Cosner, C., 2004. Spatial ecology via reaction-diffusion
+equations. Wiley.
+
+\leavevmode\hypertarget{ref-Case1986}{}%
+Case, T.J., Taper, M.L., 1986. On the coexistence and coevolution of
+asexual and sexual competitors. Evolution 40, 366--387.
+
+\leavevmode\hypertarget{ref-Champagnat2006}{}%
+Champagnat, N., Ferrière, R., Méléard, S., 2006. Unifying evolutionary
+dynamics: From individual stochastic processes to macroscopic models.
+Theoretical Population Biology 69, 297--321.
+
+\leavevmode\hypertarget{ref-Chesson2000}{}%
+Chesson, P., 2000. Mechanisms of maintenance of species diversity.
+Annual Review of Ecology and Systematics 31, 343--366.
+
+\leavevmode\hypertarget{ref-jeffreyconner2004}{}%
+Conner, J.K., 2004. A primer of ecological genetics. Sinauer Associates.
+
+\leavevmode\hypertarget{ref-9781932846126}{}%
+Crow, J.F., Kimura, M., 1970. An introduction to population genetics
+theory. The Blackburn Press.
+
+\leavevmode\hypertarget{ref-Crutsinger2015}{}%
+Crutsinger, G.M., 2015. A community genetics perspective: Opportunities
+for the coming decade. New Phytologist 210, 65--70.
+
+\leavevmode\hypertarget{ref-DaPrato2014}{}%
+Da Prato, G., Zabczyk, J., 2014. Stochastic equations in infinite
+dimensions. Cambridge University Press.
+
+\leavevmode\hypertarget{ref-David2019}{}%
+David, T.I., Storkey, J., Stevens, C.J., 2019. Understanding how
+changing soil nitrogen affects plantpollinator interactions.
+Arthropod-Plant Interactions 13, 671--684.
+
+\leavevmode\hypertarget{ref-Dawson1975}{}%
+Dawson, D.A., 1975. Stochastic evolution equations and related measure
+processes. Journal of Multivariate Analysis 5, 1--52.
+
+\leavevmode\hypertarget{ref-Dawson1978}{}%
+Dawson, D.A., 1978. Geostochastic calculus. Canadian Journal of
+Statistics 6, 143--168.
+
+\leavevmode\hypertarget{ref-dawson1993measure}{}%
+Dawson, D.A., 1993. Measure-valued markov processes, in: École d'été de
+Probabilités de Saint-Flour Xxi-1991. Springer, pp. 1--260.
+
+\leavevmode\hypertarget{ref-Dbarre2015EvolutionOQ}{}%
+Débarre, F., Yeaman, S., Guillaume, F., 2015. Evolution of quantitative
+traits under a migration-selection balance: When does skew matter? The
+American naturalist 186 Suppl 1, S37--47.
+
+\leavevmode\hypertarget{ref-Doebeli1996}{}%
+Doebeli, M., 1996. Quantitative genetics and population dynamics.
+Evolution 50, 532--546.
+
+\leavevmode\hypertarget{ref-alisonetheridge2000}{}%
+Etheridge, A.M., 2000. An introduction to superprocesses. American
+Mathematical Society.
+
+\leavevmode\hypertarget{ref-Etheridge2008}{}%
+Etheridge, A.M., 2008. Drift, draft and structure: Some mathematical
+models of evolution, in: Stochastic Models in Biological Sciences.
+Institute of Mathematics Polish Academy of Sciences.
+
+\leavevmode\hypertarget{ref-Etheridge1991}{}%
+Etheridge, A., March, P., 1991. A note on superprocesses. Probability
+Theory and Related Fields 89, 141--147.
+
+\leavevmode\hypertarget{ref-lawrenceevans2010}{}%
+Evans, L.C., 2010. Partial differential equations: Second edition.
+American Mathematical Society.
+
+\leavevmode\hypertarget{ref-9781470410544}{}%
+Evans, L.C., 2014. An introduction to stochastic differential equations.
+American Mathematical Society.
+
+\leavevmode\hypertarget{ref-Evans1994}{}%
+Evans, S.N., Perkins, E.A., 1994. Measure-valued branching diffusions
+with singular interactions. Canadian Journal of Mathematics 46,
+120--168.
+
+\leavevmode\hypertarget{ref-Ewens2004}{}%
+Ewens, W.J., 2004. Mathematical population genetics. Springer New York.
+
+\leavevmode\hypertarget{ref-kennethfalconer2014}{}%
+Falconer, K., 2014. Fractal geometry: Mathematical foundations and
+applications. Wiley.
+
+\leavevmode\hypertarget{ref-feller1951}{}%
+Feller, W., 1951. Diffusion processes in genetics, in: Proceedings of
+the Second Berkeley Symposium on Mathematical Statistics and
+Probability. University of California Press, pp. 227--246.
+
+\leavevmode\hypertarget{ref-Felsenstein1975}{}%
+Felsenstein, J., 1975. A pain in the torus: Some difficulties with
+models of isolation by distance. The American Naturalist 109, 359--368.
+
+\leavevmode\hypertarget{ref-Fisher1923}{}%
+Fisher, R.A., 1923. XXI.-on the dominance ratio. Proceedings of the
+Royal Society of Edinburgh 42, 321--341.
+
+\leavevmode\hypertarget{ref-Fitzpatrick2015}{}%
+Fitzpatrick, C.R., Agrawal, A.A., Basiliko, N., Hastings, A.P., Isaac,
+M.E., Preston, M., Johnson, M.T.J., 2015. The importance of plant
+genotype and contemporary evolution for terrestrial ecosystem processes.
+Ecology 96, 2632--2642.
+
+\leavevmode\hypertarget{ref-Fitzpatrick2017}{}%
+Fitzpatrick, C.R., Mikhailitchenko, A.V., Anstett, D.N., Johnson,
+M.T.J., 2017. The influence of range-wide plant genetic variation on
+soil invertebrate communities. Ecography 41, 1135--1146.
+
+\leavevmode\hypertarget{ref-FRANK2012}{}%
+Frank, S.A., 2012. Natural selection. IV. The price equation. Journal of
+Evolutionary Biology 25, 1002--1019.
+
+\leavevmode\hypertarget{ref-Fridley2017}{}%
+Fridley, J.D., 2017. Plant energetics and the synthesis of population
+and ecosystem ecology. Journal of Ecology 105, 95--110.
+
+\leavevmode\hypertarget{ref-FUSSMANN2007}{}%
+Fussman, G.F., Loreau, M., Abrams, P.A., 2007. Eco-evolutionary dynamics
+of communities and ecosystems. Functional Ecology 21, 465--477.
+
+\leavevmode\hypertarget{ref-Gomulkiewicz2017}{}%
+Gomulkiewicz, R., Krone, S.M., Remien, C.H., 2017. Evolution and the
+duration of a doomed population. Evolutionary Applications 10, 471--484.
+
+\leavevmode\hypertarget{ref-Harmon2019}{}%
+Harmon, L.J., Andreazzi, C.S., Débarre, F., Drury, J., Goldberg, E.E.,
+Martins, A.B., Melián, C.J., Narwani, A., Nuismer, S.L., Pennell, M.W.,
+Rudman, S.M., Seehausen, O., Silvestro, D., Weber, M., Matthews, B.,
+2019. Detecting the macroevolutionary signal of species interactions.
+Journal of Evolutionary Biology 32, 769--782.
+
+\leavevmode\hypertarget{ref-Harte2011}{}%
+Harte, J., 2011. Maximum entropy and ecology. Oxford University Press.
+
+\leavevmode\hypertarget{ref-Harte2014}{}%
+Harte, J., Newman, E.A., 2014. Maximum information entropy: A foundation
+for ecological theory. Trends in Ecology \& Evolution 29, 384--389.
+
+\leavevmode\hypertarget{ref-Hickerson2010}{}%
+Hickerson, M., Carstens, B., Cavender-Bares, J., Crandall, K., Graham,
+C., Johnson, J., Rissler, L., Victoriano, P., Yoder, A., 2010.
+Phylogeography's past, present, and future: 10 years after avise, 2000.
+Molecular Phylogenetics and Evolution 54, 291--301.
+
+\leavevmode\hypertarget{ref-Hofbauer1998}{}%
+Hofbauer, J., Sigmund, K., 1998. Evolutionary games and population
+dynamics. Cambridge University Press.
+
+\leavevmode\hypertarget{ref-Holt1987}{}%
+Holt, R.D., 1987. On the relation between niche overlap and competition:
+The effect of incommensurable niche dimensions. Oikos 48, 110.
+
+\leavevmode\hypertarget{ref-Johnson2005}{}%
+Johnson, T., Barton, N., 2005. Theoretical models of selection and
+mutation on quantitative traits. Philosophical Transactions of the Royal
+Society B: Biological Sciences 360, 1411--1425.
+
+\leavevmode\hypertarget{ref-Kendall1966}{}%
+Kendall, D.G., 1966. Branching processes since 1873. Journal of the
+London Mathematical Society s1-41, 385--406.
+
+\leavevmode\hypertarget{ref-Kimmel2015}{}%
+Kimmel, M., Axelrod, D.E., 2015. Branching processes in biology.
+Springer New York.
+
+\leavevmode\hypertarget{ref-Kimura1965}{}%
+Kimura, M., 1965. A stochastic model concerning the maintenance of
+genetic variability in quantitative characters. Proceedings of the
+National Academy of Sciences 54, 731--736.
+
+\leavevmode\hypertarget{ref-Kimura1978}{}%
+Kimura, M., Crow, J.F., 1978. Effect of overall phenotypic selection on
+genetic change at individual loci. Proceedings of the National Academy
+of Sciences 75, 6168--6171.
+
+\leavevmode\hypertarget{ref-Kirkpatrick1727}{}%
+Kirkpatrick, M., Johnson, T., Barton, N., 2002. General models of
+multilocus evolution. Genetics 161, 1727--1750.
+
+\leavevmode\hypertarget{ref-kolmogorov1999elements}{}%
+Kolmogorov, A., Fomin, S., 1999. Elements of the theory of functions and
+functional analysis. Dover.
+
+\leavevmode\hypertarget{ref-Konno1988}{}%
+Konno, N., Shiga, T., 1988. Stochastic partial differential equations
+for some measure-valued diffusions. Probability Theory and Related
+Fields 79, 201--225.
+
+\leavevmode\hypertarget{ref-Kopp2006}{}%
+Kopp, M., Gavrilets, S., 2006. Multilocus gentics and the coevolution of
+quantitative traits. Evolution 60, 1321--1336.
+
+\leavevmode\hypertarget{ref-Klzsch2015}{}%
+Kölzsch, A., Alzate, A., Bartumeus, F., Jager, M. de, Weerman, E.J.,
+Hengeveld, G.M., Naguib, M., Nolet, B.A., Koppel, J. van de, 2015.
+Experimental evidence for inherent lévy search behaviour in foraging
+animals. Proceedings of the Royal Society B: Biological Sciences 282,
+20150424.
+
+\leavevmode\hypertarget{ref-Kraft2007}{}%
+Kraft, N.J.B., Cornwell, W.K., Webb, C.O., Ackerly, D.D., 2007. Trait
+evolution, community assembly, and the phylogenetic structure of
+ecological communities. The American Naturalist 170, 271--283.
+
+\leavevmode\hypertarget{ref-Krylov1981}{}%
+Krylov, N.V., Rozovskii, B.L., 1981. Stochastic evolution equations.
+Journal of Soviet Mathematics 16, 1233--1277.
+
+\leavevmode\hypertarget{ref-Lande1975}{}%
+Lande, R., 1975. The maintenance of genetic variability by mutation in a
+polygenic character with linked loci. Genetical Research 26, 221--235.
+
+\leavevmode\hypertarget{ref-Lande1976}{}%
+Lande, R., 1976. Natural selection and random genetic drift in
+phenotypic evolution. Evolution 30, 314--334.
+
+\leavevmode\hypertarget{ref-pmid17248993}{}%
+Lande, R., 1980. The Genetic Covariance between Characters Maintained by
+Pleiotropic Mutations. Genetics 94, 203--215.
+
+\leavevmode\hypertarget{ref-Lande1982}{}%
+Lande, R., 1982. A quantitative genetic theory of life history
+evolution. Ecology 63, 607--615.
+
+\leavevmode\hypertarget{ref-Lande1995}{}%
+Lande, R., 1995. Mutation and conservation. Conservation Biology 9,
+782--791.
+
+\leavevmode\hypertarget{ref-Lande1983}{}%
+Lande, R., Arnold, S.J., 1983. The measurement of selection on
+correlated characters. Evolution 37, 1210.
+
+\leavevmode\hypertarget{ref-Lande2003}{}%
+Lande, R., Engen, S., SÆther, B.-E., 2003. Demographic and environmental
+stochasticity, in: Stochastic Population Dynamics in Ecology and
+Conservation. Oxford University Press, pp. 1--24.
+
+\leavevmode\hypertarget{ref-Lande2009}{}%
+Lande, R., Engen, S., Sæther, B.-E., 2009. An evolutionary maximum
+principle for density-dependent population dynamics in a fluctuating
+environment. Philosophical Transactions of the Royal Society B:
+Biological Sciences 364, 1511--1518.
+
+\leavevmode\hypertarget{ref-Landis2017}{}%
+Landis, M.J., Schraiber, J.G., 2017. Pulsed evolution shaped modern
+vertebrate body sizes. Proceedings of the National Academy of Sciences
+114, 13224--13229.
+
+\leavevmode\hypertarget{ref-9780691080628}{}%
+Levins, R., 1968. Evolution in changing environments: Some theoretical
+explorations. (MPB-2) (monographs in population biology). Princeton
+University Press.
+
+\leavevmode\hypertarget{ref-zeng1998absolute}{}%
+Li, Z.-H., 1998. Absolute continuity of measure branching processes with
+interaction. Chinese Journal of Applied Probability and Statistics 14,
+231--242.
+
+\leavevmode\hypertarget{ref-Lion2018}{}%
+Lion, S., 2018. Theoretical approaches in evolutionary ecology:
+Environmental feedback as a unifying perspective. The American
+Naturalist 191, 21--44.
+
+\leavevmode\hypertarget{ref-michelloreau2010}{}%
+Loreau, M., 2010. From populations to ecosystems: Theoretical
+foundations for a new ecological synthesis. Princeton University Press.
+
+\leavevmode\hypertarget{ref-michaellynch1998}{}%
+Lynch, M., Walsh, B., 1998. Genetics and analysis of quantitative
+traits. Sinauer Associates is an imprint of Oxford University Press.
+
+\leavevmode\hypertarget{ref-Arthur1969}{}%
+MacArthur, R.H., 1969. Species Packing, and what Competition Minimizes.
+Proceedings of the National Academy of Sciences 64, 1369--1371.
+
+\leavevmode\hypertarget{ref-MacArthur1970}{}%
+MacArthur, R.H., 1970. Species packing and competitive equilibrium for
+many species. Theoretical Population Biology 1, 1--11.
+
+\leavevmode\hypertarget{ref-9780691023823}{}%
+MacArthur, R.H., 1972. Geographical ecology. Princeton University Press.
+
+\leavevmode\hypertarget{ref-Macarthur1967}{}%
+MacArthur, R.H., Levins, R., 1967. The limiting similarity, convergence,
+and divergence of coexisting species. The American Naturalist 101,
+377--385.
+
+\leavevmode\hypertarget{ref-Manceau2016}{}%
+Manceau, M., Lambert, A., Morlon, H., 2016. A unifying comparative
+phylogenetic framework including traits coevolving across interacting
+lineages. Systematic Biology syw115.
+
+\leavevmode\hypertarget{ref-Marx2017}{}%
+Marx, H.E., Dentant, C., Renaud, J., Delunel, R., Tank, D.C., Lavergne,
+S., 2017. Riders in the sky (islands): Using a mega-phylogenetic
+approach to understand plant species distribution and coexistence at the
+altitudinal limits of angiosperm plant life. Journal of Biogeography 44,
+2618--2630.
+
+\leavevmode\hypertarget{ref-markmcpeek2017}{}%
+McPeek, M.A., 2017. Evolutionary community ecology. Princeton University
 Press.
 
-\leavevmode\hypertarget{ref-Barton2017}{}%
-Barton, N.H., A.M. Etheridge, and A. Véber. 2017. ``The Infinitesimal
-Model: Definition, Derivation, and Implications.'' \emph{Theoretical
-Population Biology} 118 (December). Elsevier BV: 50--73.
-
-\leavevmode\hypertarget{ref-Barton2019}{}%
-Barton, Nick, and Alison Etheridge. 2019. ``Mathematical Models in
-Population Genetics.'' Wiley.
-
-\leavevmode\hypertarget{ref-Barton2013}{}%
-Barton, Nick, Alison Etheridge, and Amandine Véber. 2013. ``Modelling
-Evolution in a Spatial Continuum.'' \emph{Journal of Statistical
-Mechanics: Theory and Experiment} 2013 (01). IOP Publishing: P01002.
-
-\leavevmode\hypertarget{ref-Bertoin2003}{}%
-Bertoin, Jean, and Jean-François Le Gall. 2003. ``Stochastic Flows
-Associated to Coalescent Processes.'' \emph{Probability Theory and
-Related Fields} 126 (2). Springer Science; Business Media: 261--88.
-
-\leavevmode\hypertarget{ref-Brown1956}{}%
-Brown, W. L., and E. O. Wilson. 1956. ``Character Displacement.''
-\emph{Systematic Zoology} 5 (2). Oxford University Press (OUP): 49.
-
-\leavevmode\hypertarget{ref-bulmer1980}{}%
-Bulmer. 1980. \emph{The Mathematical Theory of Quantitative Genetics}.
-Oxford University Press.
-
-\leavevmode\hypertarget{ref-Brger1986}{}%
-Bürger, Reinhard. 1986. ``On the Maintenance of Genetic Variation:
-Global Analysis of Kimuras Continuum-of-Alleles Model.'' \emph{Journal
-of Mathematical Biology} 24 (3). Springer Nature: 341--51.
-
-\leavevmode\hypertarget{ref-9780471986539}{}%
----------. 2000. \emph{The Mathematical Theory of Selection,
-Recombination, and Mutation}. Wiley.
-
-\leavevmode\hypertarget{ref-Cantrell2004}{}%
-Cantrell, Robert Stephen, and Chris Cosner. 2004. \emph{Spatial Ecology
-via Reaction-Diffusion Equations}. Wiley.
-
-\leavevmode\hypertarget{ref-Case1986}{}%
-Case, Ted J., and Mark L. Taper. 1986. ``On the Coexistence and
-Coevolution of Asexual and Sexual Competitors.'' \emph{Evolution} 40
-(2). Wiley: 366--87.
-
-\leavevmode\hypertarget{ref-Champagnat2006}{}%
-Champagnat, Nicolas, Régis Ferrière, and Sylvie Méléard. 2006.
-``Unifying Evolutionary Dynamics: From Individual Stochastic Processes
-to Macroscopic Models.'' \emph{Theoretical Population Biology} 69 (3).
-Elsevier BV: 297--321.
-
-\leavevmode\hypertarget{ref-Chesson2000}{}%
-Chesson, Peter. 2000. ``Mechanisms of Maintenance of Species
-Diversity.'' \emph{Annual Review of Ecology and Systematics} 31 (1).
-Annual Reviews: 343--66.
-
-\leavevmode\hypertarget{ref-jeffreyconner2004}{}%
-Conner, Jeffrey K. 2004. \emph{A Primer of Ecological Genetics}. Sinauer
-Associates.
-
-\leavevmode\hypertarget{ref-9781932846126}{}%
-Crow, James F., and Motoo Kimura. 1970. \emph{An Introduction to
-Population Genetics Theory}. The Blackburn Press.
-
-\leavevmode\hypertarget{ref-Crutsinger2015}{}%
-Crutsinger, Gregory M. 2015. ``A Community Genetics Perspective:
-Opportunities for the Coming Decade.'' \emph{New Phytologist} 210 (1).
-Wiley: 65--70.
-
-\leavevmode\hypertarget{ref-DaPrato2014}{}%
-Da Prato, Giuseppe, and Jerzy Zabczyk. 2014. \emph{Stochastic Equations
-in Infinite Dimensions}. Cambridge University Press.
-
-\leavevmode\hypertarget{ref-David2019}{}%
-David, Thomas I., Jonathan Storkey, and Carly J. Stevens. 2019.
-``Understanding How Changing Soil Nitrogen Affects Plantpollinator
-Interactions.'' \emph{Arthropod-Plant Interactions} 13 (5). Springer
-Science; Business Media LLC: 671--84.
-
-\leavevmode\hypertarget{ref-Dawson1975}{}%
-Dawson, Donald A. 1975. ``Stochastic Evolution Equations and Related
-Measure Processes.'' \emph{Journal of Multivariate Analysis} 5 (1).
-Elsevier BV: 1--52.
-
-\leavevmode\hypertarget{ref-Dawson1978}{}%
----------. 1978. ``Geostochastic Calculus.'' \emph{Canadian Journal of
-Statistics} 6 (2). Wiley: 143--68.
-
-\leavevmode\hypertarget{ref-dawson1993measure}{}%
----------. 1993. ``Measure-Valued Markov Processes.'' In \emph{École
-d'été de Probabilités de Saint-Flour Xxi-1991}, 1--260. Springer.
-
-\leavevmode\hypertarget{ref-Dbarre2015EvolutionOQ}{}%
-Débarre, Florence, Sam Yeaman, and Frédéric Guillaume. 2015. ``Evolution
-of Quantitative Traits Under a Migration-Selection Balance: When Does
-Skew Matter?'' \emph{The American Naturalist} 186 Suppl 1: S37--47.
-
-\leavevmode\hypertarget{ref-Doebeli1996}{}%
-Doebeli, Michael. 1996. ``Quantitative Genetics and Population
-Dynamics.'' \emph{Evolution} 50 (2). Wiley: 532--46.
-
-\leavevmode\hypertarget{ref-alisonetheridge2000}{}%
-Etheridge, Alison M. 2000. \emph{An Introduction to Superprocesses}.
-American Mathematical Society.
-
-\leavevmode\hypertarget{ref-Etheridge2008}{}%
----------. 2008. ``Drift, Draft and Structure: Some Mathematical Models
-of Evolution.'' In \emph{Stochastic Models in Biological Sciences}.
-Institute of Mathematics Polish Academy of Sciences.
-
-\leavevmode\hypertarget{ref-Etheridge1991}{}%
-Etheridge, Alison, and Peter March. 1991. ``A Note on Superprocesses.''
-\emph{Probability Theory and Related Fields} 89 (2). Springer: 141--47.
-
-\leavevmode\hypertarget{ref-lawrenceevans2010}{}%
-Evans, Lawrence C. 2010. \emph{Partial Differential Equations: Second
-Edition}. American Mathematical Society.
-
-\leavevmode\hypertarget{ref-9781470410544}{}%
----------. 2014. \emph{An Introduction to Stochastic Differential
-Equations}. American Mathematical Society.
-
-\leavevmode\hypertarget{ref-Evans1994}{}%
-Evans, Steven N., and Edwin A. Perkins. 1994. ``Measure-Valued Branching
-Diffusions with Singular Interactions.'' \emph{Canadian Journal of
-Mathematics} 46 (1). Canadian Mathematical Society: 120--68.
-
-\leavevmode\hypertarget{ref-Ewens2004}{}%
-Ewens, Warren J. 2004. \emph{Mathematical Population Genetics}. Springer
-New York.
-
-\leavevmode\hypertarget{ref-kennethfalconer2014}{}%
-Falconer, Kenneth. 2014. \emph{Fractal Geometry: Mathematical
-Foundations and Applications}. Wiley.
-
-\leavevmode\hypertarget{ref-feller1951}{}%
-Feller, William. 1951. ``Diffusion Processes in Genetics.'' In
-\emph{Proceedings of the Second Berkeley Symposium on Mathematical
-Statistics and Probability}, 227--46. University of California Press.
-
-\leavevmode\hypertarget{ref-Felsenstein1975}{}%
-Felsenstein, Joseph. 1975. ``A Pain in the Torus: Some Difficulties with
-Models of Isolation by Distance.'' \emph{The American Naturalist} 109
-(967). University of Chicago Press: 359--68.
-
-\leavevmode\hypertarget{ref-Fisher1923}{}%
-Fisher, R. A. 1923. ``XXI.-on the Dominance Ratio.'' \emph{Proceedings
-of the Royal Society of Edinburgh} 42. Cambridge University Press:
-321--41.
-
-\leavevmode\hypertarget{ref-Fitzpatrick2015}{}%
-Fitzpatrick, Connor R., Anurag A. Agrawal, Nathan Basiliko, Amy P.
-Hastings, Marney E. Isaac, Michael Preston, and Marc T. J. Johnson.
-2015. ``The Importance of Plant Genotype and Contemporary Evolution for
-Terrestrial Ecosystem Processes.'' \emph{Ecology} 96 (10). Wiley:
-2632--42.
-
-\leavevmode\hypertarget{ref-Fitzpatrick2017}{}%
-Fitzpatrick, Connor R., Anna V. Mikhailitchenko, Daniel N. Anstett, and
-Marc T. J. Johnson. 2017. ``The Influence of Range-Wide Plant Genetic
-Variation on Soil Invertebrate Communities.'' \emph{Ecography} 41 (7).
-Wiley: 1135--46.
-
-\leavevmode\hypertarget{ref-FRANK2012}{}%
-Frank, S. A. 2012. ``Natural Selection. IV. The Price Equation.''
-\emph{Journal of Evolutionary Biology} 25 (6). Wiley: 1002--19.
-
-\leavevmode\hypertarget{ref-Fridley2017}{}%
-Fridley, Jason D. 2017. ``Plant Energetics and the Synthesis of
-Population and Ecosystem Ecology.'' Edited by David Gibson.
-\emph{Journal of Ecology} 105 (1). Wiley: 95--110.
-
-\leavevmode\hypertarget{ref-FUSSMANN2007}{}%
-Fussman, G. F., M. Loreau, and P. A. Abrams. 2007. ``Eco-Evolutionary
-Dynamics of Communities and Ecosystems.'' \emph{Functional Ecology} 21
-(3). Wiley: 465--77.
-
-\leavevmode\hypertarget{ref-Gomulkiewicz2017}{}%
-Gomulkiewicz, Richard, Stephen M. Krone, and Christopher H. Remien.
-2017. ``Evolution and the Duration of a Doomed Population.''
-\emph{Evolutionary Applications} 10 (5). Wiley: 471--84.
-
-\leavevmode\hypertarget{ref-Harmon2019}{}%
-Harmon, Luke J., Cecilia S. Andreazzi, Florence Débarre, Jonathan Drury,
-Emma E. Goldberg, Ayana B. Martins, Carlos J. Melián, et al. 2019.
-``Detecting the Macroevolutionary Signal of Species Interactions.''
-\emph{Journal of Evolutionary Biology} 32 (8). Wiley: 769--82.
-
-\leavevmode\hypertarget{ref-Harte2011}{}%
-Harte, John. 2011. \emph{Maximum Entropy and Ecology}. Oxford University
-Press.
-
-\leavevmode\hypertarget{ref-Harte2014}{}%
-Harte, John, and Erica A. Newman. 2014. ``Maximum Information Entropy: A
-Foundation for Ecological Theory.'' \emph{Trends in Ecology \&
-Evolution} 29 (7). Elsevier BV: 384--89.
-
-\leavevmode\hypertarget{ref-Hickerson2010}{}%
-Hickerson, M.J., B.C. Carstens, J. Cavender-Bares, K.A. Crandall, C.H.
-Graham, J.B. Johnson, L. Rissler, P.F. Victoriano, and A.D. Yoder. 2010.
-``Phylogeography's Past, Present, and Future: 10 Years After Avise,
-2000.'' \emph{Molecular Phylogenetics and Evolution} 54 (1). Elsevier
-BV: 291--301.
-
-\leavevmode\hypertarget{ref-Hofbauer1998}{}%
-Hofbauer, Josef, and Karl Sigmund. 1998. \emph{Evolutionary Games and
-Population Dynamics}. Cambridge University Press.
-
-\leavevmode\hypertarget{ref-Holt1987}{}%
-Holt, Robert D. 1987. ``On the Relation Between Niche Overlap and
-Competition: The Effect of Incommensurable Niche Dimensions.''
-\emph{Oikos} 48 (1). JSTOR: 110.
-
-\leavevmode\hypertarget{ref-Johnson2005}{}%
-Johnson, Toby, and Nick Barton. 2005. ``Theoretical Models of Selection
-and Mutation on Quantitative Traits.'' \emph{Philosophical Transactions
-of the Royal Society B: Biological Sciences} 360 (1459). The Royal
-Society: 1411--25.
-
-\leavevmode\hypertarget{ref-Kendall1966}{}%
-Kendall, David G. 1966. ``Branching Processes Since 1873.''
-\emph{Journal of the London Mathematical Society} s1-41 (1). Wiley:
-385--406.
-
-\leavevmode\hypertarget{ref-Kimmel2015}{}%
-Kimmel, Marek, and David E. Axelrod. 2015. \emph{Branching Processes in
-Biology}. Springer New York.
-
-\leavevmode\hypertarget{ref-Kimura1965}{}%
-Kimura, M. 1965. ``A Stochastic Model Concerning the Maintenance of
-Genetic Variability in Quantitative Characters.'' \emph{Proceedings of
-the National Academy of Sciences} 54 (3). Proceedings of the National
-Academy of Sciences: 731--36.
-
-\leavevmode\hypertarget{ref-Kimura1978}{}%
-Kimura, M., and J. F. Crow. 1978. ``Effect of Overall Phenotypic
-Selection on Genetic Change at Individual Loci.'' \emph{Proceedings of
-the National Academy of Sciences} 75 (12). Proceedings of the National
-Academy of Sciences: 6168--71.
-
-\leavevmode\hypertarget{ref-Kirkpatrick1727}{}%
-Kirkpatrick, Mark, Toby Johnson, and Nick Barton. 2002. ``General Models
-of Multilocus Evolution.'' \emph{Genetics} 161 (4). Genetics: 1727--50.
-
-\leavevmode\hypertarget{ref-kolmogorov1999elements}{}%
-Kolmogorov, A.N., and S.V. Fomin. 1999. \emph{Elements of the Theory of
-Functions and Functional Analysis}. v. 1. Dover.
-
-\leavevmode\hypertarget{ref-Konno1988}{}%
-Konno, N., and T. Shiga. 1988. ``Stochastic Partial Differential
-Equations for Some Measure-Valued Diffusions.'' \emph{Probability Theory
-and Related Fields} 79 (2). Springer Nature: 201--25.
-
-\leavevmode\hypertarget{ref-Kopp2006}{}%
-Kopp, Michael, and Sergey Gavrilets. 2006. ``Multilocus Gentics and the
-Coevolution of Quantitative Traits.'' \emph{Evolution} 60 (7). Wiley:
-1321--36.
-
-\leavevmode\hypertarget{ref-Klzsch2015}{}%
-Kölzsch, Andrea, Adriana Alzate, Frederic Bartumeus, Monique de Jager,
-Ellen J. Weerman, Geerten M. Hengeveld, Marc Naguib, Bart A. Nolet, and
-Johan van de Koppel. 2015. ``Experimental Evidence for Inherent Lévy
-Search Behaviour in Foraging Animals.'' \emph{Proceedings of the Royal
-Society B: Biological Sciences} 282 (1807). The Royal Society: 20150424.
-
-\leavevmode\hypertarget{ref-Kraft2007}{}%
-Kraft, Nathan J. B., William K. Cornwell, Campbell O. Webb, and David D.
-Ackerly. 2007. ``Trait Evolution, Community Assembly, and the
-Phylogenetic Structure of Ecological Communities.'' \emph{The American
-Naturalist} 170 (2). University of Chicago Press: 271--83.
-
-\leavevmode\hypertarget{ref-Krylov1981}{}%
-Krylov, N. V., and B. L. Rozovskii. 1981. ``Stochastic Evolution
-Equations.'' \emph{Journal of Soviet Mathematics} 16 (4). Springer
-Science; Business Media LLC: 1233--77.
-
-\leavevmode\hypertarget{ref-Lande1975}{}%
-Lande, Russell. 1975. ``The Maintenance of Genetic Variability by
-Mutation in a Polygenic Character with Linked Loci.'' \emph{Genetical
-Research} 26 (3). Cambridge University Press (CUP): 221--35.
-
-\leavevmode\hypertarget{ref-Lande1976}{}%
----------. 1976. ``Natural Selection and Random Genetic Drift in
-Phenotypic Evolution.'' \emph{Evolution} 30 (2). Wiley: 314--34.
-
-\leavevmode\hypertarget{ref-pmid17248993}{}%
----------. 1980. ``The Genetic Covariance between Characters Maintained
-by Pleiotropic Mutations.'' \emph{Genetics} 94 (1): 203--15.
-
-\leavevmode\hypertarget{ref-Lande1982}{}%
----------. 1982. ``A Quantitative Genetic Theory of Life History
-Evolution.'' \emph{Ecology} 63 (3). Wiley: 607--15.
-
-\leavevmode\hypertarget{ref-Lande1995}{}%
----------. 1995. ``Mutation and Conservation.'' \emph{Conservation
-Biology} 9 (4). Wiley: 782--91.
-
-\leavevmode\hypertarget{ref-Lande1983}{}%
-Lande, Russell, and Stevan J. Arnold. 1983. ``The Measurement of
-Selection on Correlated Characters.'' \emph{Evolution} 37 (6). JSTOR:
-1210.
-
-\leavevmode\hypertarget{ref-Lande2003}{}%
-Lande, Russell, Steinar Engen, and Bernt-Erik SÆther. 2003.
-``Demographic and Environmental Stochasticity.'' In \emph{Stochastic
-Population Dynamics in Ecology and Conservation}, 1--24. Oxford
-University Press.
-
-\leavevmode\hypertarget{ref-Lande2009}{}%
-Lande, Russell, Steinar Engen, and Bernt-Erik Sæther. 2009. ``An
-Evolutionary Maximum Principle for Density-Dependent Population Dynamics
-in a Fluctuating Environment.'' \emph{Philosophical Transactions of the
-Royal Society B: Biological Sciences} 364 (1523). The Royal Society:
-1511--8.
-
-\leavevmode\hypertarget{ref-Landis2017}{}%
-Landis, Michael J., and Joshua G. Schraiber. 2017. ``Pulsed Evolution
-Shaped Modern Vertebrate Body Sizes.'' \emph{Proceedings of the National
-Academy of Sciences} 114 (50). Proceedings of the National Academy of
-Sciences: 13224--9.
-
-\leavevmode\hypertarget{ref-9780691080628}{}%
-Levins, Richard. 1968. \emph{Evolution in Changing Environments: Some
-Theoretical Explorations. (MPB-2) (Monographs in Population Biology)}.
+\leavevmode\hypertarget{ref-DeMeester2018}{}%
+Meester, L.D., Brans, K.I., Govaert, L., Souffreau, C., Mukherjee, S.,
+Vanvelk, H., Korzeniowski, K., Kilsdonk, L., Decaestecker, E., Stoks,
+R., Urban, M.C., 2018. Analyzing eco-evolutionary dynamics - the
+challenging complexity of the real world. Functional Ecology.
+
+\leavevmode\hypertarget{ref-meleard1992interacting}{}%
+Méléard, M., Roelly, S., 1992. Interacting branching measure processes.
+Stochastic Partial Differential Equations and Applications (G. Da Prato
+and L. Tubaro, eds.) 246--256.
+
+\leavevmode\hypertarget{ref-Mlard1993}{}%
+Méléard, M., Roelly, S., 1993. Interacting measure branching processes.
+Some bounds for the support. Stochastics and Stochastic Reports 44,
+103--121.
+
+\leavevmode\hypertarget{ref-Mubayi2019}{}%
+Mubayi, A., Kribs, C., Arunachalam, V., Castillo-Chavez, C., 2019.
+Studying complexity and risk through stochastic population dynamics:
+Persistence, resonance, and extinction in ecosystems, in: Handbook of
+Statistics. Elsevier, pp. 157--193.
+
+\leavevmode\hypertarget{ref-9780674023383}{}%
+Nowak, M.A., 2006. Evolutionary dynamics: Exploring the equations of
+life. Belknap Press.
+
+\leavevmode\hypertarget{ref-Nuismer2005}{}%
+Nuismer, S.L., Doebeli, M., Browning, D., 2005. The coevolutionary
+dynamics of antagonistic interactions mediated by quantitative traits
+with evolving variances. Evolution 59, 2073.
+
+\leavevmode\hypertarget{ref-Nuismer2014}{}%
+Nuismer, S.L., Harmon, L.J., 2014. Predicting rates of interspecific
+interaction from phylogenetic trees. Ecology Letters 18, 17--27.
+
+\leavevmode\hypertarget{ref-Nuismer2012}{}%
+Nuismer, S.L., Jordano, P., Bascompte, J., 2012. Coevolution and the
+architecture of mutualistic networks. Evolution 67, 338--354.
+
+\leavevmode\hypertarget{ref-Nuismer2007}{}%
+Nuismer, S.L., Ridenhour, B.J., Oswald, B.P., 2007. Antagonistic
+coevolution mediated by phenotypic differenes between quantitative
+traits. Evolution 61, 1823--1834.
+
+\leavevmode\hypertarget{ref-Nuismer2018}{}%
+Nuismer, S.L., Week, B., Aizen, M.A., 2018. Coevolution slows the
+disassembly of mutualistic networks. The American Naturalist 192,
+490--502.
+
+\leavevmode\hypertarget{ref-VanNuland2019}{}%
+Nuland, M.E.V., Ware, I.M., Bailey, J.K., Schweitzer, J.A., 2019.
+Ecosystem feedbacks contribute to geographic variation in plant-soil
+eco-evolutionary dynamics across a fertility gradient. Functional
+Ecology 33, 95--106.
+
+\leavevmode\hypertarget{ref-PAGE2002}{}%
+Page, K.M., Nowak, M.A., 2002. Unifying evolutionary dynamics. Journal
+of Theoretical Biology 219, 93--98.
+
+\leavevmode\hypertarget{ref-Parachnowitsch2018}{}%
+Parachnowitsch, A.L., Manson, J.S., Sletvold, N., 2018. Evolutionary
+ecology of nectar. Annals of Botany 123, 247--261.
+
+\leavevmode\hypertarget{ref-Park2017}{}%
+Park, B., Rutter, M.T., Fenster, C.B., Symonds, V.V., Ungerer, M.C.,
+Townsend, J.P., 2017. Distributions of mutational effects and the
+estimation of directional selection in divergent lineages ofArabidopsis
+thaliana. Genetics 206, 2105--2117.
+
+\leavevmode\hypertarget{ref-Parsons2010}{}%
+Parsons, T.L., Quince, C., Plotkin, J.B., 2010. Some consequences of
+demographic stochasticity in population genetics. Genetics 185,
+1345--1354.
+
+\leavevmode\hypertarget{ref-Patel2019}{}%
+Patel, S., Bürger, R., 2019. Eco-evolutionary feedbacks between prey
+densities and linkage disequilibrium in the predator maintain diversity.
+Evolution 73, 1533--1548.
+
+\leavevmode\hypertarget{ref-Perkins1991}{}%
+Perkins, E.A., 1991. Conditional dawson-watanabe processes and
+fleming-viot processes, in: Seminar on Stochastic Processes, 1991.
+Birkhäuser Boston, pp. 143--156.
+
+\leavevmode\hypertarget{ref-Perkins1992}{}%
+Perkins, E.A., 1992. Measure-valued branching diffusions with spatial
+interactions. Probability Theory and Related Fields 94, 189--245.
+
+\leavevmode\hypertarget{ref-edwinarendperkins1995}{}%
+Perkins, E.A., 1995. On the martingale problem for interactive
+measure-valued branching diffusions. Amer Mathematical Society.
+
+\leavevmode\hypertarget{ref-Polechov2015}{}%
+Polechová, J., Barton, N.H., 2015. Limits to adaptation along
+environmental gradients. Proceedings of the National Academy of Sciences
+112, 6401--6406.
+
+\leavevmode\hypertarget{ref-PRICE1970}{}%
+Price, G.R., 1970. Selection and covariance. Nature 227, 520--521.
+
+\leavevmode\hypertarget{ref-Queller2017}{}%
+Queller, D.C., 2017. Fundamental theorems of evolution. The American
+Naturalist 189, 345--353.
+
+\leavevmode\hypertarget{ref-Reimers1989}{}%
+Reimers, M., 1989. One dimensional stochastic partial differential
+equations and the branching measure diffusion. Probability Theory and
+Related Fields 81, 319--340.
+
+\leavevmode\hypertarget{ref-Robertson1966}{}%
+Robertson, A., 1966. A mathematical model of the culling process in
+dairy cattle. Animal Science 8, 95--108.
+
+\leavevmode\hypertarget{ref-joanroughgarden1979}{}%
+Roughgarden, J., 1979. Theory of population genetics and evolutionary
+ecology: An introduction. Macmillan.
+
+\leavevmode\hypertarget{ref-Rudman2017}{}%
+Rudman, S.M., Barbour, M.A., Csilléry, K., Gienapp, P., Guillaume, F.,
+Jr, N.G.H., Hendry, A.P., Lasky, J.R., Rafajlović, M., Räsänen, K.,
+Schmidt, P.S., Seehausen, O., Therkildsen, N.O., Turcotte, M.M., Levine,
+J.M., 2017. What genomic data can reveal about eco-evolutionary
+dynamics. Nature Ecology \& Evolution 2, 9--15.
+
+\leavevmode\hypertarget{ref-Schreiber2017}{}%
+Schreiber, S.J., 2017. Coexistence in the face of uncertainty, in:
+Recent Progress and Modern Challenges in Applied Mathematics, Modeling
+and Computational Science. Springer New York, pp. 349--384.
+
+\leavevmode\hypertarget{ref-Schreiber2018}{}%
+Schreiber, S.J., Patel, S., terHorst, C., 2018. Evolution as a
+coexistence mechanism: Does genetic architecture matter? The American
+Naturalist 191, 407--420.
+
+\leavevmode\hypertarget{ref-Schuster1983}{}%
+Schuster, P., Sigmund, K., 1983. Replicator dynamics. Journal of
+Theoretical Biology 100, 533--538.
+
+\leavevmode\hypertarget{ref-Skovmand2018}{}%
+Skovmand, L.H., Xu, C.C., Servedio, M.R., Nosil, P., Barrett, R.D.,
+Hendry, A.P., 2018. Keystone genes. Trends in Ecology \& Evolution 33,
+689--700.
+
+\leavevmode\hypertarget{ref-Sterner2008}{}%
+Sterner, R., Elser, J., 2008. Ecological stoichiometry: Overview, in:
+Encyclopedia of Ecology. Elsevier, pp. 1101--1116.
+
+\leavevmode\hypertarget{ref-Taylor1978}{}%
+Taylor, P.D., Jonker, L.B., 1978. Evolutionary stable strategies and
+game dynamics. Mathematical Biosciences 40, 145--156.
+
+\leavevmode\hypertarget{ref-davidtilman1982}{}%
+Tilman, D., 1982. Resource competition and community structure.
 Princeton University Press.
 
-\leavevmode\hypertarget{ref-zeng1998absolute}{}%
-Li, Zeng-Hu. 1998. ``Absolute Continuity of Measure Branching Processes
-with Interaction.'' \emph{Chinese Journal of Applied Probability and
-Statistics} 14. Citeseer: 231--42.
-
-\leavevmode\hypertarget{ref-Lion2018}{}%
-Lion, Sébastien. 2018. ``Theoretical Approaches in Evolutionary Ecology:
-Environmental Feedback as a Unifying Perspective.'' \emph{The American
-Naturalist} 191 (1). University of Chicago Press: 21--44.
-
-\leavevmode\hypertarget{ref-michelloreau2010}{}%
-Loreau, Michel. 2010. \emph{From Populations to Ecosystems: Theoretical
-Foundations for a New Ecological Synthesis}. Princeton University Press.
-
-\leavevmode\hypertarget{ref-michaellynch1998}{}%
-Lynch, Michael, and Bruce Walsh. 1998. \emph{Genetics and Analysis of
-Quantitative Traits}. Sinauer Associates is an imprint of Oxford
-University Press.
-
-\leavevmode\hypertarget{ref-Arthur1969}{}%
-MacArthur, Robert H. 1969. ``Species Packing, and what Competition
-Minimizes.'' \emph{Proceedings of the National Academy of Sciences} 64
-(4). Proceedings of the National Academy of Sciences: 1369--71.
-
-\leavevmode\hypertarget{ref-MacArthur1970}{}%
----------. 1970. ``Species Packing and Competitive Equilibrium for Many
-Species.'' \emph{Theoretical Population Biology} 1 (1). Elsevier BV:
-1--11.
-
-\leavevmode\hypertarget{ref-9780691023823}{}%
----------. 1972. \emph{Geographical Ecology}. Princeton University
-Press.
-
-\leavevmode\hypertarget{ref-Macarthur1967}{}%
-MacArthur, Robert H., and Richard Levins. 1967. ``The Limiting
-Similarity, Convergence, and Divergence of Coexisting Species.''
-\emph{The American Naturalist} 101 (921). University of Chicago Press:
-377--85.
-
-\leavevmode\hypertarget{ref-Manceau2016}{}%
-Manceau, Marc, Amaury Lambert, and Hélène Morlon. 2016. ``A Unifying
-Comparative Phylogenetic Framework Including Traits Coevolving Across
-Interacting Lineages.'' \emph{Systematic Biology}, December. Oxford
-University Press (OUP), syw115.
-
-\leavevmode\hypertarget{ref-Marx2017}{}%
-Marx, Hannah E., Cédric Dentant, Julien Renaud, Romain Delunel, David C.
-Tank, and Sébastien Lavergne. 2017. ``Riders in the Sky (Islands): Using
-a Mega-Phylogenetic Approach to Understand Plant Species Distribution
-and Coexistence at the Altitudinal Limits of Angiosperm Plant Life.''
-\emph{Journal of Biogeography} 44 (11). Wiley: 2618--30.
-
-\leavevmode\hypertarget{ref-markmcpeek2017}{}%
-McPeek, Mark A. 2017. \emph{Evolutionary Community Ecology}. Princeton
-University Press.
-
-\leavevmode\hypertarget{ref-DeMeester2018}{}%
-Meester, Luc De, Kristien I. Brans, Lynn Govaert, Caroline Souffreau,
-Shinjini Mukherjee, Héléne Vanvelk, Konrad Korzeniowski, et al. 2018.
-``Analyzing Eco-Evolutionary Dynamics - the Challenging Complexity of
-the Real World.'' \emph{Functional Ecology}. Wiley.
-
-\leavevmode\hypertarget{ref-meleard1992interacting}{}%
-Méléard, M, and S Roelly. 1992. ``Interacting Branching Measure
-Processes.'' \emph{Stochastic Partial Differential Equations and
-Applications (G. Da Prato and L. Tubaro, Eds.)}, 246--56.
-
-\leavevmode\hypertarget{ref-Mlard1993}{}%
----------. 1993. ``Interacting Measure Branching Processes. Some Bounds
-for the Support.'' \emph{Stochastics and Stochastic Reports} 44 (1-2).
-Informa UK Limited: 103--21.
-
-\leavevmode\hypertarget{ref-Mubayi2019}{}%
-Mubayi, Anuj, Christopher Kribs, Viswanathan Arunachalam, and Carlos
-Castillo-Chavez. 2019. ``Studying Complexity and Risk Through Stochastic
-Population Dynamics: Persistence, Resonance, and Extinction in
-Ecosystems.'' In \emph{Handbook of Statistics}, 157--93. Elsevier.
-
-\leavevmode\hypertarget{ref-9780674023383}{}%
-Nowak, Martin A. 2006. \emph{Evolutionary Dynamics: Exploring the
-Equations of Life}. Belknap Press.
-
-\leavevmode\hypertarget{ref-Nuismer2005}{}%
-Nuismer, Scott L., Michael Doebeli, and Danny Browning. 2005. ``The
-Coevolutionary Dynamics of Antagonistic Interactions Mediated by
-Quantitative Traits with Evolving Variances.'' \emph{Evolution} 59 (10).
-The Society for the Study of Evolution: 2073.
-
-\leavevmode\hypertarget{ref-Nuismer2014}{}%
-Nuismer, Scott L., and Luke J. Harmon. 2014. ``Predicting Rates of
-Interspecific Interaction from Phylogenetic Trees.'' Edited by Jerome
-Chave. \emph{Ecology Letters} 18 (1). Wiley: 17--27.
-
-\leavevmode\hypertarget{ref-Nuismer2012}{}%
-Nuismer, Scott L., Pedro Jordano, and Jordi Bascompte. 2012.
-``Coevolution and the Architecture of Mutualistic Networks.''
-\emph{Evolution} 67 (2). Wiley: 338--54.
-
-\leavevmode\hypertarget{ref-Nuismer2007}{}%
-Nuismer, Scott L., Benjamin J. Ridenhour, and Benjamin P. Oswald. 2007.
-``Antagonistic Coevolution Mediated by Phenotypic Differenes Between
-Quantitative Traits.'' \emph{Evolution} 61 (8). Wiley: 1823--34.
-
-\leavevmode\hypertarget{ref-Nuismer2018}{}%
-Nuismer, Scott L., Bob Week, and Marcelo A. Aizen. 2018. ``Coevolution
-Slows the Disassembly of Mutualistic Networks.'' \emph{The American
-Naturalist} 192 (4). University of Chicago Press: 490--502.
-
-\leavevmode\hypertarget{ref-VanNuland2019}{}%
-Nuland, Michael E. Van, Ian M. Ware, Joseph K. Bailey, and Jennifer A.
-Schweitzer. 2019. ``Ecosystem Feedbacks Contribute to Geographic
-Variation in Plant-Soil Eco-Evolutionary Dynamics Across a Fertility
-Gradient.'' Edited by Franziska Brunner. \emph{Functional Ecology} 33
-(1). Wiley: 95--106.
-
-\leavevmode\hypertarget{ref-PAGE2002}{}%
-Page, Karen M., and Martin A. Nowak. 2002. ``Unifying Evolutionary
-Dynamics.'' \emph{Journal of Theoretical Biology} 219 (1). Elsevier BV:
-93--98.
-
-\leavevmode\hypertarget{ref-Parachnowitsch2018}{}%
-Parachnowitsch, Amy L, Jessamyn S Manson, and Nina Sletvold. 2018.
-``Evolutionary Ecology of Nectar.'' \emph{Annals of Botany} 123 (2).
-Oxford University Press (OUP): 247--61.
-
-\leavevmode\hypertarget{ref-Park2017}{}%
-Park, Briton, Matthew T. Rutter, Charles B. Fenster, V. Vaughan Symonds,
-Mark C. Ungerer, and Jeffrey P. Townsend. 2017. ``Distributions of
-Mutational Effects and the Estimation of Directional Selection in
-Divergent Lineages ofArabidopsis Thaliana.'' \emph{Genetics} 206 (4).
-Genetics Society of America: 2105--17.
-
-\leavevmode\hypertarget{ref-Parsons2010}{}%
-Parsons, Todd L., Christopher Quince, and Joshua B. Plotkin. 2010.
-``Some Consequences of Demographic Stochasticity in Population
-Genetics.'' \emph{Genetics} 185 (4). Genetics Society of America:
-1345--54.
-
-\leavevmode\hypertarget{ref-Patel2019}{}%
-Patel, Swati, and Reinhard Bürger. 2019. ``Eco-Evolutionary Feedbacks
-Between Prey Densities and Linkage Disequilibrium in the Predator
-Maintain Diversity.'' \emph{Evolution} 73 (8). Wiley: 1533--48.
-
-\leavevmode\hypertarget{ref-Perkins1991}{}%
-Perkins, Edwin A. 1991. ``Conditional Dawson-Watanabe Processes and
-Fleming-Viot Processes.'' In \emph{Seminar on Stochastic Processes,
-1991}, 143--56. Birkhäuser Boston.
-
-\leavevmode\hypertarget{ref-Perkins1992}{}%
----------. 1992. ``Measure-Valued Branching Diffusions with Spatial
-Interactions.'' \emph{Probability Theory and Related Fields} 94 (2).
-Springer Science; Business Media LLC: 189--245.
-
-\leavevmode\hypertarget{ref-edwinarendperkins1995}{}%
----------. 1995. \emph{On the Martingale Problem for Interactive
-Measure-Valued Branching Diffusions}. Amer Mathematical Society.
-
-\leavevmode\hypertarget{ref-Polechov2015}{}%
-Polechová, Jitka, and Nicholas H. Barton. 2015. ``Limits to Adaptation
-Along Environmental Gradients.'' \emph{Proceedings of the National
-Academy of Sciences} 112 (20). Proceedings of the National Academy of
-Sciences: 6401--6.
-
-\leavevmode\hypertarget{ref-PRICE1970}{}%
-Price, George R. 1970. ``Selection and Covariance.'' \emph{Nature} 227
-(5257). Springer Nature: 520--21.
-
-\leavevmode\hypertarget{ref-Queller2017}{}%
-Queller, David C. 2017. ``Fundamental Theorems of Evolution.'' \emph{The
-American Naturalist} 189 (4). University of Chicago Press: 345--53.
-
-\leavevmode\hypertarget{ref-Reimers1989}{}%
-Reimers, Mark. 1989. ``One Dimensional Stochastic Partial Differential
-Equations and the Branching Measure Diffusion.'' \emph{Probability
-Theory and Related Fields} 81 (3). Springer Nature: 319--40.
-
-\leavevmode\hypertarget{ref-Robertson1966}{}%
-Robertson, Alan. 1966. ``A Mathematical Model of the Culling Process in
-Dairy Cattle.'' \emph{Animal Science} 8 (1). Cambridge University Press:
-95--108.
-
-\leavevmode\hypertarget{ref-joanroughgarden1979}{}%
-Roughgarden, Joan. 1979. \emph{Theory of Population Genetics and
-Evolutionary Ecology: An Introduction}. Macmillan.
-
-\leavevmode\hypertarget{ref-Rudman2017}{}%
-Rudman, Seth M., Matthew A. Barbour, Katalin Csilléry, Phillip Gienapp,
-Frederic Guillaume, Nelson G. Hairston Jr, Andrew P. Hendry, et al.
-2017. ``What Genomic Data Can Reveal About Eco-Evolutionary Dynamics.''
-\emph{Nature Ecology \& Evolution} 2 (1). Springer Nature: 9--15.
-
-\leavevmode\hypertarget{ref-Schreiber2017}{}%
-Schreiber, Sebastian J. 2017. ``Coexistence in the Face of
-Uncertainty.'' In \emph{Recent Progress and Modern Challenges in Applied
-Mathematics, Modeling and Computational Science}, 349--84. Springer New
-York.
-
-\leavevmode\hypertarget{ref-Schreiber2018}{}%
-Schreiber, Sebastian J., Swati Patel, and Casey terHorst. 2018.
-``Evolution as a Coexistence Mechanism: Does Genetic Architecture
-Matter?'' \emph{The American Naturalist} 191 (3). University of Chicago
-Press: 407--20.
-
-\leavevmode\hypertarget{ref-Schuster1983}{}%
-Schuster, Peter, and Karl Sigmund. 1983. ``Replicator Dynamics.''
-\emph{Journal of Theoretical Biology} 100 (3). Elsevier BV: 533--38.
-
-\leavevmode\hypertarget{ref-Skovmand2018}{}%
-Skovmand, Lotte H., Charles C.Y. Xu, Maria R. Servedio, Patrik Nosil,
-Rowan D.H. Barrett, and Andrew P. Hendry. 2018. ``Keystone Genes.''
-\emph{Trends in Ecology \& Evolution} 33 (9). Elsevier BV: 689--700.
-
-\leavevmode\hypertarget{ref-Sterner2008}{}%
-Sterner, R.W., and J.J. Elser. 2008. ``Ecological Stoichiometry:
-Overview.'' In \emph{Encyclopedia of Ecology}, 1101--16. Elsevier.
-
-\leavevmode\hypertarget{ref-Taylor1978}{}%
-Taylor, Peter D., and Leo B. Jonker. 1978. ``Evolutionary Stable
-Strategies and Game Dynamics.'' \emph{Mathematical Biosciences} 40
-(1-2). Elsevier BV: 145--56.
-
-\leavevmode\hypertarget{ref-davidtilman1982}{}%
-Tilman, David. 1982. \emph{Resource Competition and Community
-Structure}. Princeton University Press.
-
 \leavevmode\hypertarget{ref-Turelli1980}{}%
-Turelli, Michael. 1980. ``Niche Overlap and Invasion of Competitors in
-Random Environments II. The Effects of Demographic Stochasticity.'' In
-\emph{Biological Growth and Spread}, 119--29. Springer Berlin
-Heidelberg.
+Turelli, M., 1980. Niche overlap and invasion of competitors in random
+environments II. The effects of demographic stochasticity, in:
+Biological Growth and Spread. Springer Berlin Heidelberg, pp. 119--129.
 
 \leavevmode\hypertarget{ref-Turelli1984}{}%
----------. 1984. ``Heritable Genetic Variation via Mutation-Selection
-Balance: Lerchs Zeta Meets the Abdominal Bristle.'' \emph{Theoretical
-Population Biology} 25 (2). Elsevier: 138--93.
+Turelli, M., 1984. Heritable genetic variation via mutation-selection
+balance: Lerchs zeta meets the abdominal bristle. Theoretical Population
+Biology 25, 138--193.
 
 \leavevmode\hypertarget{ref-Turelli1986}{}%
----------. 1986. ``Gaussian Versus Non-Gaussian Genetic Analyses of
-Polygenic Mutation-Selection Balance.'' In \emph{Evolutionary Processes
-and Theory}, 607--28. Academic Press.
+Turelli, M., 1986. Gaussian versus non-gaussian genetic analyses of
+polygenic mutation-selection balance, in: Evolutionary Processes and
+Theory. Academic Press, pp. 607--628.
 
 \leavevmode\hypertarget{ref-Turelli2017}{}%
----------. 2017. ``Commentary: Fisher's Infinitesimal Model: A Story for
-the Ages.'' \emph{Theoretical Population Biology} 118 (December).
-Elsevier BV: 46--49.
+Turelli, M., 2017. Commentary: Fisher's infinitesimal model: A story for
+the ages. Theoretical Population Biology 118, 46--49.
 
 \leavevmode\hypertarget{ref-pmid7851785}{}%
-Turelli, Michael, and Nick Barton. 1994. ``Genetic and statistical
-analyses of strong selection on polygenic traits: what, me normal?''
-\emph{Genetics} 138 (3): 913--41.
+Turelli, M., Barton, N., 1994. Genetic and statistical analyses of
+strong selection on polygenic traits: what, me normal? Genetics 138,
+913--941.
 
 \leavevmode\hypertarget{ref-Volpert2014}{}%
-Volpert, Vitaly. 2014. \emph{Elliptic Partial Differential Equations:
-Volume 2: Reaction-Diffusion Equations}. Springer Basel.
+Volpert, V., 2014. Elliptic partial differential equations: Volume 2:
+Reaction-diffusion equations. Springer Basel.
 
 \leavevmode\hypertarget{ref-Walsh}{}%
-Walsh, John B. 1986. ``An Introduction to Stochastic Partial
-Differential Equations.'' In \emph{Lecture Notes in Mathematics},
-265--439. Springer Berlin Heidelberg.
+Walsh, J.B., 1986. An introduction to stochastic partial differential
+equations, in: Lecture Notes in Mathematics. Springer Berlin Heidelberg,
+pp. 265--439.
 
 \leavevmode\hypertarget{ref-Watanabe1968}{}%
-Watanabe, Shinzo. 1968. ``A Limit Theorem of Branching Processes and
-Continuous State Branching Processes.'' \emph{Journal of Mathematics of
-Kyoto University} 8 (1). Duke University Press: 141--67.
+Watanabe, S., 1968. A limit theorem of branching processes and
+continuous state branching processes. Journal of Mathematics of Kyoto
+University 8, 141--167.
 
 \leavevmode\hypertarget{ref-Wright97}{}%
-Wright, Sewall. 1931. ``Evolution in Mendelian Populations.''
-\emph{Genetics} 16 (2). Genetics: 97--159.
+Wright, S., 1931. Evolution in mendelian populations. Genetics 16,
+97--159.
 
 \leavevmode\hypertarget{ref-Xiao2015}{}%
-Xiao, Xiao, Daniel J. McGlinn, and Ethan P. White. 2015. ``A Strong Test
-of the Maximum Entropy Theory of Ecology.'' \emph{The American
-Naturalist} 185 (3). University of Chicago Press: E70--E80.
+Xiao, X., McGlinn, D.J., White, E.P., 2015. A strong test of the maximum
+entropy theory of ecology. The American Naturalist 185, E70--E80.
 
 \leavevmode\hypertarget{ref-zheng2004nonlinear}{}%
-Zheng, Songmu. 2004. \emph{Nonlinear Evolution Equations}. Boca Raton,
-Fla: Chapman \& Hall/CRC Press.
+Zheng, S., 2004. Nonlinear evolution equations. Chapman \& Hall/CRC
+Press, Boca Raton, Fla.
 

### 2020-04-28 Current Version/draft.Rmd
reorganize

--- 2020-04-18 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
@@
 ---
-title: "A White Noise Approach to Evolutionary Ecology"
-author: "Bob Week, Steve Krone, Scott L. Nuismer, Luke J. Harmon"
-date: "3/1/2020"
-abstract: "We derive the dynamics of the distribution of a quantitative character and the abundance of a biological population from a stochastic partial differential equation driven by space-time white noise.
+title: A White Noise Approach to Evolutionary Ecology
+author:
+  - name: Bob Week
+    email: bobweek@gmail.com
+    affiliation: University of Idaho
+    footnote: 1
+  - name: Steve Krone
+    affiliation: University of Idaho
+  - name: Scott L. Nuismer
+    affiliation: University of Idaho
+  - name: Luke J. Harmon
+    affiliation: University of Idaho
+address:
+  - code: University of Idaho
+    address: Department of Biological Sciences, 875 Perimeter Drive MS 3051, Moscow, Idaho, 83843
+footnote:
+  - code: 1
+    text: "Corresponding Author"
+  - code: 2
+    text: "Equal contribution"
+date: "`r Sys.Date()`"
+abstract: |
+  We derive the dynamics of the distribution of a quantitative character and the abundance of a biological population from a stochastic partial differential equation driven by space-time white noise.
 In the process we develop a useful set of heuristics to operationalize the powerful, but abstract theory of white noise and measure-valued stochastic processes.
@@
 The inverse relationship holds for quadratic selection gradients except with threshold diversity ratio equal to unity instead of $1.25$.
-Hence, this approach contributes to the development of a synthetic theory of evolutionary ecology by formalizing first principle derivations of stochastic models that underlie rigorous investigations of the relationship between feedbacks of biological processes and the patterns of diversity they produce."
+Hence, this approach contributes to the development of a synthetic theory of evolutionary ecology by formalizing first principle derivations of stochastic models that underlie rigorous investigations of the relationship between feedbacks of biological processes and the patterns of diversity they produce.
 bibliography: ref.bib
@@
   - \usepackage{amsmath}
+  - \usepackage[authoryear,sectionbib,sort]{natbib}  
+  - \usepackage{fullpage}
   - \linenumbers
@@
 ## White noise calculus and superprocesses \label{wnc}
-
-### Introduction to white noise
 

### 2020-04-28 Current Version/draft.tex
reorganize

--- 2020-04-18 .A White Noise Approach To Evolutionary Ecology/draft.tex
@@
             pdftitle={A White Noise Approach to Evolutionary Ecology},
-            pdfauthor={Bob Week, Steve Krone, Scott L. Nuismer, Luke J. Harmon},
             pdfborder={0 0 0},
@@
 \usepackage{amsmath}
+\usepackage[authoryear,sectionbib,sort]{natbib}
+\usepackage{fullpage}
 \linenumbers
@@
 \title{A White Noise Approach to Evolutionary Ecology}
-\author{Bob Week, Steve Krone, Scott L. Nuismer, Luke J. Harmon}
-\date{3/1/2020}
+\author{true \and true \and true \and true}
+\date{2020-04-21}
 

### 2020-04-28 Current Version/mapping.Rmd
reorganize

---
title: "Mapping the notation of Méléard and Roelly (1992) to Week et al"
output:
  pdf_document:
    toc: false
    number_sections: true
    fig_width: 4
    fig_height: 3
    fig_caption: true
header-includes:
  - \usepackage{float}
  - \usepackage{mathrsfs}
  - \usepackage{hyperref}
  - \usepackage[left]{lineno}
  - \usepackage{amsmath}
  - \usepackage{makecell}
editor_options: 
  chunk_output_type: console
---

\begin{table}[H]
\begin{tabular}{l|l|l}
Description           & Méléard and Roelly 92 & Week et al    \\ \hline
&& \\
Arbitrary finite measure & $\mu\in M_F(\mathbb{R}^d)$ & NA \\
&& \\
Bounded function & $f\in C_b(\mathbb{R}^d)$ & $f\in C_b(\mathbb{R})$ \\
&& \\
\makecell{Non-negative continuous \\ integrable function} & NA & $\nu(\cdot,t)\in C_1^+(\mathbb{R})$ \\
&& \\
"Spatial" location    & $x\in\mathbb{R}^d$ & $x\in\mathbb{R}\dots$ \makecell{interpreted as \\ trait value} \\
&& \\
\makecell{Infinitesimal generator \\ of spatial movement} & \makecell{$A(\mu)\dots$ \\ satisfies ($\mathrm{H}_1$) and ($\mathrm{H}_2$) below} & \makecell{$\frac{\mu}{2}\Delta\dots$ Laplace operator \\ with diffusion coefficient $\mu/2$} \\
&& \\
Superprocess          & $X_t\in M_F(\mathbb{R}^d)$ for $t\in[0,\infty)$ & NA                      \\
&& \\
Density process       & NA               & $\nu(x,t)$              \\
&& \\
Duality bracket       & $\begin{matrix}\langle X_t,f\rangle\dots \\ X_t\in M_F(\mathbb{R}^d), \ f\in C_b(\mathbb{R}^d)\end{matrix}$ & $\begin{matrix}\int_\mathbb{R}\nu(x,t)f(x)dx\dots \\ \nu(\cdot,t)\in C_1^+(\mathbb{R}), \ f\in C_b(\mathbb{R})\end{matrix}$ \\
&& \\
Malthusian parameter  & \makecell{$b(\mu)(x)\dots \ b(\mu)\in C_b(\mathbb{R}^d)$ \\ for $\mu\in M_F(\mathbb{R}^d)$} & $m\big(\nu(x,t),x\big)\dots$ \makecell{bounded above, \\ but not below} \\
&& \\
Stochastic parameter  & \makecell{$c(\mu)(x)\dots \ c(\mu)\in C_b(\mathbb{R}^d)$ \\ for $\mu\in M_F(\mathbb{R}^d)$} & $V>0\dots$ constant \\
&& \\
Reproductive law      & $p_k(x,\mu)$ & NA \\
&& \\
Mean of reproductive law & $\begin{matrix}m(x,\mu)=\sum_{k\in\mathbb{N}}kp_k(x,\mu)\dots \\ \sup_{x,\mu}m(x,\mu)<\infty\end{matrix}$ & $\begin{matrix}\mathscr{W}(\nu,x)\dots\\\sup_{\nu,x}\mathscr{W}(\nu,x)<\infty\end{matrix}$ \makecell{introduced in \\ appendix} \\
&& \\
Unnamed                  & $v^2(x,\mu)=\sum_{k\in\mathbb{N}}(k-1)^2p_k(x,\mu)$ & NA \\
&& \\
Death rate               & $\lambda(x,\mu)$ & we assume constant "lifetimes"
\end{tabular}\label{mult}
\end{table}

- ($\mathrm{H}_1$): For all $\mu\in M_F(\mathbb{R}^d)$ and all $f$ in the domain of $A(\mu)$, there is a constant $K>0$ such that $\|A(\mu)f\|_\infty\leq K\langle\mu,1\rangle$.

- ($\mathrm{H}_2$): For all $f$ in the domain of $A(\mu)$, the map $\mu\to\langle\mu,A(\mu)f\rangle$ is continuous.

- In Week et al, spatial movement is interpreted as mutation. Hence, durations of spatial movement are not interpreted as lifetimes. Instead, we assume discrete non-overlapping generations for the branching diffusion process.

### 2020-04-28 Julia Scripts/bbm.jl

@@ -1,0 +1,165 @@
+using Parameters, Statistics, Random, LinearAlgebra, Distributions,
+	StatsBase, StatsPlots, Plots, DataFrames, CSV, Optim, KernelDensity,
+	QuadGK, HypothesisTests
+
+include("/home/bob/Research/Branching Brownian Motion/bbm_functions_structs.jl")
+
+#######################################################
+#                                                     #
+# a branching Brownian motion for a single population #
+#                                                     #
+#######################################################
+
+# parameter values
+w = 0.1  # niche breadths
+U = 1.0  # total niche use
+c = 2e-3 # strengths of competition
+η = 1e-5 # segregation variances
+μ = 1e-3 # mutation rates
+V = 2.0  # magnitudes of drift
+r = 1.0  # innate rate of growth
+a = 1e-2 # strengths of abiotic selection
+θ = 0.0  # phenotypic optima
+n = 1.0 # scaling parameter
+
+##
+## VERY IMPORTANT REQUIREMENT   -->  V >= exp(r)
+##
+## this inequality must be satisfied to use negative binomial sampling
+##
+
+# equilibrium abundance an the absence of interspecific interactions
+# we use this as the initial abundance
+
+N₀ = Int64(floor( n*( r -0.5*( η*a + √(μ*a) ) )/c ) )
+
+# initial breeding values
+g₀ = rand(Normal(0.0,1.0),N₀)
+
+# initial trait values
+ηₘ = √η*Matrix(I, N₀, N₀)
+x₀ = vec(rand(MvNormal(g₀,ηₘ),1))
+
+# set up initial population
+X = community(S=1, x=fill(x₀,1), g=fill(g₀,1), N=fill(N₀,1), n=fill(n,1),
+	x̄=mean.(fill(x₀,1)), σ²=var.(fill(x₀,1)), G=var.(fill(g₀,1)),
+	R=fill(r,1), a=fill(a,1), θ=fill(θ,1), c=fill(c,1), w=fill(w,1),
+	U=fill(U,1), η=fill(η,1), μ=fill(μ,1), V=fill(V,1) )
+
[…]

### 2020-04-28 Julia Scripts/bbm_comm.jl

@@ -1,0 +1,127 @@
+using Parameters, Statistics, Random, LinearAlgebra, Distributions,
+	StatsBase, StatsPlots, Plots, DataFrames, CSV, Optim
+
+include("/home/bob/Research/Branching Brownian Motion/bbm_functions_structs.jl")
+
+########################################################
+#                                                      #
+# a branching Brownian motion for the entire community #
+#                                                      #
+########################################################
+
+# parameter values
+S = 1
+w = fill(0.1, S)  # niche breadths
+U = fill(1.0, S)  # total niche use
+c = fill(0.002,S) # strengths of competition
+Ω = sum(U)        # niche use scaling
+η = fill(1e-3, S)  # segregation variances
+μ = fill(1e-5, S) # mutation rates
+V = fill(2.0, S)  # magnitudes of drift
+R = fill(0.1, S)  # innate rate of growth
+a = fill(1e-2,S)  # strengths of abiotic selection
+θ = fill(0.0, S)  # phenotypic optima
+n = fill(10.0, S) # scaling parameter
+
+# equilibrium abundance an the absence of interspecific interactions
+# we use this as the initial abundance
+
+C = c.*(U.^2)./.√(4 .*π.*w)
+N₀ = Int64.(floor.( (n.^2).*( (R./n).-0.5*( η.*a .+ .√(μ.*a) ) )./C ) )
+
+# initial breeding values
+g₀ = rand.(Normal(0.0,1.0),N₀)
+
+# initial trait values
+x₀ = fill(zeros(0),S)
+for i in 1:S
+	ηₘ = √η[i]*Matrix(I, N₀[i], N₀[i])
+	x₀[i] = vec(rand(MvNormal(g₀[i],ηₘ),1))
+end
+
+##
+## VERY IMPORTANT REQUIREMENT   -->  V >= exp(r)
+##
+## this inequality must be satisfied to use negative binomial sampling
+##
+##
+## TWO MORE IMPORTANT REQUIREMENTS --> 2*r > √(μ*a) && c > r - √(μ*a)/2
+##
[…]

### 2020-04-28 Julia Scripts/sde_comp_coev.jl

@@ -1,0 +1,187 @@
+#
+# using numerical solutions to estimate
+# Cov(α,|β|), Cov(α,γ) Cov(α,ℭ) and
+# Cor(α,|β|), Cor(α,γ) Cor(α,ℭ)
+#
+
+
+using Parameters,
+      Statistics,
+      Random,
+      LinearAlgebra,
+      Distributions,
+      DifferentialEquations,
+      StatsBase,
+      StatsPlots,
+      Plots,
+      DataFrames,
+      CSV,
+      Optim
+
+include("/home/bob/Research/Branching Brownian Motion/sde_functions.jl")
+
+# background parameters
+S = 100
+w = fill(0.1, S)  # niche breadths
+U = fill(1.0, S)  # total niche use
+Ω = sum(U) # niche use scaling
+η = fill(1.0, S)  # segregation variances
+μ = fill(1e-5, S) # mutation rates
+V = fill(1.0, S)  # magnitudes of drift
+R = fill(1.0, S) # innate rate of growth
+θ = fill(0.0, S)  # phenotypic optima
+
+# set the timespan
+# T₁ corresponds to the 'burn-in'
+# T₂ corresponds the region we estimate the covariances over
+T₁ = 5e2
+T₂ = 5e2
+tspan = (0.0, T₁ + T₂)
+
+# containers for covariances and correlations
+Cov_αβ_mean = zeros(0)
+Cov_αabsβ_mean = zeros(0)
+Cov_αγ_mean = zeros(0)
+Cov_αℭ_mean = zeros(0)
+Cor_αβ_mean = zeros(0)
+Cor_αabsβ_mean = zeros(0)
+Cor_αγ_mean = zeros(0)
+Cor_αℭ_mean = zeros(0)
+
+# containers for variance in mean traits among species
+# and mean of the variance within species
+Vₓ_mean = zeros(0)
+σ²_mean = zeros(0)
+
+# counts number of points accepted
+num = 0
+
+# accumulate 1000 points
+while num < 1000
+
+    # random draws for a and c
+    adraw = exp.(rand(Normal(-10,6),1))[1]
+    cdraw = exp.(rand(Normal(-20,4),1))[1]
+
+    # continue to draw a until it is less than 1e-2
+    # this prevents small N
+    while maximum(adraw)>1e-2
[…]

### 2020-04-28 Julia Scripts/sde_functions.jl

@@ -1,0 +1,429 @@
+###############################
+#                             #
+# data and parameter structs  #
+#                             #
+###############################
+
+
+# data type that holds data on the community
+@with_kw mutable struct CommData
+	x̄::Vector{Float64}  # mean traits
+	G::Vector{Float64}  # additive genetic variances
+	N::Vector{Float64}  # abundances
+end
+
+# data type that holds model parameters
+@with_kw mutable struct ModelParameters
+	S::Int64            # species richness
+	w::Vector{Float64}  # niche breadths, always zero for now...
+	U::Vector{Float64}  # total niche use
+	η::Vector{Float64}  # segregation variances
+	c::Vector{Float64}	# strengths of competition
+	a::Vector{Float64}	# strengths of abiotic selection
+	μ::Vector{Float64}	# mutation rates
+	V::Vector{Float64}	# variance of reproductive output
+	R::Vector{Float64}	# intrinsic rates of growth
+	θ::Vector{Float64}	# phenotypic optima
+	Ω::Float64        	# total niche-use scaling
+end
+
+# data type that holds known parameters
+@with_kw mutable struct FixedParameters
+	S::Int64            # species richness
+	w::Vector{Float64}  # niche breadths, always zero for now...
+	U::Vector{Float64}  # total niche use
+	x̃::Vector{Float64}  # initial condition (≈ deterministic equilibrium)
+	ε::Float64          # convergence threshold
+	T::Float64          # duration of time to integrate over
+	dT::Float64					# duration of time to find deterministic equilibrium
+	ΔT::Float64					# length of time to check convergence over
[…]

### 2020-04-28 Julia Scripts/sde_num_sol.jl

@@ -1,0 +1,95 @@
+using Parameters, Statistics, Random, LinearAlgebra, Distributions,
+	DifferentialEquations, StatsBase, StatsPlots, Plots, DataFrames, CSV, Optim
+
+include("/home/bb/Gits/branching.brownian.motion.and.spde/sde_functions.jl")
+
+j=1
+k=1
+# strengths of competition
+cc = [1e-7,5e-6]
+
+# for S = 1000
+cc = [1e-8,5e-7]
+
+
+# durations
+TT = [1e3,5e3]
+
+tspan = (0.0,TT[j])
+
+# background parameters
+S = 1000
+w = fill(0.1, S)  # niche breadths
+U = fill(1.0, S)  # total niche use
+c = fill(cc[k],S)      # strengths of competition
+Ω = sum(U) # niche use scaling
+η = fill(1.0, S)  # segregation variances
+μ = fill(1e-7, S) # mutation rates
+V = fill(5.0, S)  # magnitudes of drift
+R = fill(1.0, S) # innate rate of growth
+a = fill(1e-2,S)       # strengths of abiotic selection
+θ = fill(0.0, S)  # phenotypic optima
+
+pars = ModelParameters(S=S, w=w, U=U, η=η, c=c, a=a, μ=μ, V=V, R=R, θ=θ, Ω=Ω)
+
+#
+# find deterministic equilibrium
+#
+
+# approximation for equilibrium abundance
+C = c.*.√( U.^2 ./ .√(4*π.*w) )
+N₀ = Int64.(floor.( (R.-0.5*.√(μ.*a))./C ) )
+
+# initial condition
+u₀ = cat(θ,fill(10.0, S),fill(1000.0, S),dims=1)
+
+# numerically solve SDE
+prob = SDEProblem(f,g,u₀,tspan,pars)
+sol = solve(prob)
+
+# extract state variables
+x̄_sol = copy(.-sol[(0*S+1):(1*S),:])
+G_sol = copy(sol[(1*S+1):(2*S),:])
+N_sol = copy(sol[(2*S+1):(3*S),:])
+
+# build dataframes
+spp = string("Species", 1)
+df = DataFrame(spp = spp, x = x̄_sol[1,:], G = G_sol[1,:], N = N_sol[1,:], time = sol.t)
+for i in 2:S
+	spp = string("Species", i)
+	df_dummy = DataFrame(spp = spp, x = x̄_sol[i,:], G = G_sol[i,:],
[…]

### 2020-04-28 bbm.jl

@@ -1,165 +1,0 @@
-using Parameters, Statistics, Random, LinearAlgebra, Distributions,
-	StatsBase, StatsPlots, Plots, DataFrames, CSV, Optim, KernelDensity,
-	QuadGK, HypothesisTests
-
-include("/home/bob/Research/Branching Brownian Motion/bbm_functions_structs.jl")
-
-#######################################################
-#                                                     #
-# a branching Brownian motion for a single population #
-#                                                     #
-#######################################################
-
-# parameter values
-w = 0.1  # niche breadths
-U = 1.0  # total niche use
-c = 2e-3 # strengths of competition
-η = 1e-5 # segregation variances
-μ = 1e-3 # mutation rates
-V = 2.0  # magnitudes of drift
-r = 1.0  # innate rate of growth
-a = 1e-2 # strengths of abiotic selection
-θ = 0.0  # phenotypic optima
-n = 1.0 # scaling parameter
-
-##
-## VERY IMPORTANT REQUIREMENT   -->  V >= exp(r)
-##
-## this inequality must be satisfied to use negative binomial sampling
-##
-
-# equilibrium abundance an the absence of interspecific interactions
-# we use this as the initial abundance
-
-N₀ = Int64(floor( n*( r -0.5*( η*a + √(μ*a) ) )/c ) )
-
-# initial breeding values
-g₀ = rand(Normal(0.0,1.0),N₀)
-
-# initial trait values
-ηₘ = √η*Matrix(I, N₀, N₀)
-x₀ = vec(rand(MvNormal(g₀,ηₘ),1))
-
-# set up initial population
-X = community(S=1, x=fill(x₀,1), g=fill(g₀,1), N=fill(N₀,1), n=fill(n,1),
-	x̄=mean.(fill(x₀,1)), σ²=var.(fill(x₀,1)), G=var.(fill(g₀,1)),
-	R=fill(r,1), a=fill(a,1), θ=fill(θ,1), c=fill(c,1), w=fill(w,1),
-	U=fill(U,1), η=fill(η,1), μ=fill(μ,1), V=fill(V,1) )
-
[…]

### 2020-04-28 bbm_functions_structs.jl

@@ -1,365 +1,0 @@
-# data type that holds population and parameters
-
-@with_kw mutable struct population
-    x::Vector{Float64}  # trait values
-    N::Int64            # poulation size
-    x̄::Float64          # mean trait
-    σ²::Float64         # phenotypic variance
-    R::Float64          # innate rate of growth
-    a::Float64          # strength of abiotic selection
-    θ::Float64          # abiotic optimum
-    c::Float64          # strength of competition
-    μ::Float64          # rate of diffusion (mutation)
-    V::Float64          # variance in reproductive output
-end
-
-@with_kw mutable struct community
-    S::Int64                   # number of species
-    x::Vector{Vector{Float64}} # trait values
-    g::Vector{Vector{Float64}} # breeding values
-    N::Vector{Int64}           # population sizes
-    n::Vector{Int64}           # index of rescaled sequence
-    x̄::Vector{Float64}         # mean traits
-    σ²::Vector{Float64}        # phenotypic variances
-    G::Vector{Float64}         # additive genetic variances
-    R::Vector{Float64}         # innate rates of growth
-    a::Vector{Float64}         # strengths of abiotic selection
-    θ::Vector{Float64}         # abiotic optima
-    c::Vector{Float64}         # strengths of competition
-    w::Vector{Float64}         # individual niche widths
-    U::Vector{Float64}         # total niche uses
-    η::Vector{Float64}         # segregation variance
-    μ::Vector{Float64}         # rates of diffusion (mutation)
-    V::Vector{Float64}         # variances in reproductive output
-end
-
-# update for community
-function comm_update(X)
-
-    @unpack S, x, N, x̄, σ², R, a, θ, c, w, U, μ, V = X
-
-    # creates array of offspring trait values
-    # first index is species
-    # second index is individual
-    xₚ = fill(zeros(0),S)
-
-    for i in 1:S
-
-        W = fill(0,N[i])
-
-        for j in 1:N[i]
-
-            #
-            # mean fitness of individual j in species i
-            # this follows exactly from SM §5.6
-            #
-
-            # container for aggregating effects of competition
[…]

### 2020-04-28 bbm_gaussian.jl

@@ -1,165 +1,0 @@
-using Parameters, Statistics, Random, LinearAlgebra, Distributions,
-	StatsBase, StatsPlots, Plots, DataFrames, CSV, Optim, KernelDensity,
-	QuadGK, HypothesisTests
-
-include("/home/bob/Research/Branching Brownian Motion/bbm_functions_structs.jl")
-
-#######################################################
-#                                                     #
-# a branching Brownian motion for a single population #
-#                                                     #
-#######################################################
-
-# parameter values
-w = 0.1  # niche breadths
-U = 1.0  # total niche use
-c = 2e-3 # strengths of competition
-η = 1e-5 # segregation variances
-μ = 1e-2 # mutation rates
-V = 2.0  # magnitudes of drift
-r = 1.0  # innate rate of growth
-a = 1e-2 # strengths of abiotic selection
-θ = 0.0  # phenotypic optima
-n = 20.0 # scaling parameter
-
-##
-## VERY IMPORTANT REQUIREMENT   -->  V >= exp(r)
-##
-## this inequality must be satisfied to use negative binomial sampling
-##
-
-# equilibrium abundance an the absence of interspecific interactions
-# we use this as the initial abundance
-
-N₀ = Int64(floor( n*( r -0.5*( η*a + √(μ*a) ) )/c ) )
-
-# initial breeding values
-g₀ = rand(Normal(θ,√(μ/a)),N₀)
-
-# initial trait values
-ηₘ = √η*Matrix(I, N₀, N₀)
-x₀ = vec(rand(MvNormal(g₀,ηₘ),1))
-
-# set up initial population
-X = community(S=1, x=fill(x₀,1), g=fill(g₀,1), N=fill(N₀,1), n=fill(n,1),
-	x̄=mean.(fill(x₀,1)), σ²=var.(fill(x₀,1)), G=var.(fill(g₀,1)),
-	R=fill(r,1), a=fill(a,1), θ=fill(θ,1), c=fill(c,1), w=fill(w,1),
[…]

### 2020-04-28 plot_sde_sample_path.R

@@ -1,24 +1,0 @@
-require("ggplot2")
-require("gridExtra")
-
-smpl_wc = read.csv("/home/bob/Research/WNEE/sample_path_wc.csv")
-
-x_wc = ggplot(smpl_wc,aes(x=time,y=x,group=spp,color=spp))+theme_bw()+theme_minimal()+xlab("")+ylab("x")+
-  geom_line(show.legend=F, lwd=0.1)+scale_fill_viridis_c()+ggtitle("Weak Competition")
-G_wc = ggplot(smpl_wc,aes(x=time,y=G,group=spp,color=spp))+theme_bw()+theme_minimal()+xlab("")+ylab("G")+
-  geom_line(show.legend=F, lwd=0.1)+scale_fill_viridis_c()
-N_wc = ggplot(smpl_wc,aes(x=time,y=N,group=spp,color=spp))+theme_bw()+theme_minimal()+xlab("time")+ylab("N")+
-  geom_line(show.legend=F, lwd=0.1)+scale_fill_viridis_c()
-smpl_plt_wc = grid.arrange(x_wc,G_wc,N_wc,nrow=3)
-ggsave("~/Research/WNEE/smpl_plt_wc.png",smpl_plt_wc,width=5,height=2.5)
-
-smpl_mc = read.csv("/home/bob/Research/WNEE/sample_path_mc.csv")
-
-x_mc = ggplot(smpl_mc,aes(x=time,y=x,group=spp,color=spp))+theme_bw()+theme_minimal()+xlab("")+ylab("x")+
-  geom_line(show.legend=F, lwd=0.1)+scale_fill_viridis_c()+ggtitle("Moderate Competition")
-G_mc = ggplot(smpl_mc,aes(x=time,y=G,group=spp,color=spp))+theme_bw()+theme_minimal()+xlab("")+ylab("G")+
-  geom_line(show.legend=F, lwd=0.1)+scale_fill_viridis_c()
-N_mc = ggplot(smpl_mc,aes(x=time,y=N,group=spp,color=spp))+theme_bw()+theme_minimal()+xlab("time")+ylab("N")+
-  geom_line(show.legend=F, lwd=0.1)+scale_fill_viridis_c()
-smpl_plt_mc = grid.arrange(x_mc,G_mc,N_mc,nrow=3)
-ggsave("~/Research/WNEE/smpl_plt_mc.png",smpl_plt_mc,width=5,height=2.5)

### 2020-04-28 sde_functions.jl

@@ -1,429 +1,0 @@
-###############################
-#                             #
-# data and parameter structs  #
-#                             #
-###############################
-
-
-# data type that holds data on the community
-@with_kw mutable struct CommData
-	x̄::Vector{Float64}  # mean traits
-	G::Vector{Float64}  # additive genetic variances
-	N::Vector{Float64}  # abundances
-end
-
-# data type that holds model parameters
-@with_kw mutable struct ModelParameters
-	S::Int64            # species richness
-	w::Vector{Float64}  # niche breadths, always zero for now...
-	U::Vector{Float64}  # total niche use
-	η::Vector{Float64}  # segregation variances
-	c::Vector{Float64}	# strengths of competition
-	a::Vector{Float64}	# strengths of abiotic selection
-	μ::Vector{Float64}	# mutation rates
-	V::Vector{Float64}	# variance of reproductive output
-	R::Vector{Float64}	# intrinsic rates of growth
-	θ::Vector{Float64}	# phenotypic optima
-	Ω::Float64        	# total niche-use scaling
-end
-
-# data type that holds known parameters
-@with_kw mutable struct FixedParameters
-	S::Int64            # species richness
-	w::Vector{Float64}  # niche breadths, always zero for now...
-	U::Vector{Float64}  # total niche use
-	x̃::Vector{Float64}  # initial condition (≈ deterministic equilibrium)
-	ε::Float64          # convergence threshold
-	T::Float64          # duration of time to integrate over
-	dT::Float64					# duration of time to find deterministic equilibrium
-	ΔT::Float64					# length of time to check convergence over
[…]

### 2020-04-28 sde_num_sol.jl

@@ -1,95 +1,0 @@
-using Parameters, Statistics, Random, LinearAlgebra, Distributions,
-	DifferentialEquations, StatsBase, StatsPlots, Plots, DataFrames, CSV, Optim
-
-include("/home/bb/Gits/branching.brownian.motion.and.spde/sde_functions.jl")
-
-j=1
-k=1
-# strengths of competition
-cc = [1e-7,5e-6]
-
-# for S = 1000
-cc = [1e-8,5e-7]
-
-
-# durations
-TT = [1e3,5e3]
-
-tspan = (0.0,TT[j])
-
-# background parameters
-S = 1000
-w = fill(0.1, S)  # niche breadths
-U = fill(1.0, S)  # total niche use
-c = fill(cc[k],S)      # strengths of competition
-Ω = sum(U) # niche use scaling
-η = fill(1.0, S)  # segregation variances
-μ = fill(1e-7, S) # mutation rates
-V = fill(5.0, S)  # magnitudes of drift
-R = fill(1.0, S) # innate rate of growth
-a = fill(1e-2,S)       # strengths of abiotic selection
-θ = fill(0.0, S)  # phenotypic optima
-
-pars = ModelParameters(S=S, w=w, U=U, η=η, c=c, a=a, μ=μ, V=V, R=R, θ=θ, Ω=Ω)
-
-#
-# find deterministic equilibrium
-#
-
-# approximation for equilibrium abundance
-C = c.*.√( U.^2 ./ .√(4*π.*w) )
-N₀ = Int64.(floor.( (R.-0.5*.√(μ.*a))./C ) )
-
-# initial condition
-u₀ = cat(θ,fill(10.0, S),fill(1000.0, S),dims=1)
-
-# numerically solve SDE
-prob = SDEProblem(f,g,u₀,tspan,pars)
-sol = solve(prob)
-
-# extract state variables
-x̄_sol = copy(.-sol[(0*S+1):(1*S),:])
-G_sol = copy(sol[(1*S+1):(2*S),:])
-N_sol = copy(sol[(2*S+1):(3*S),:])
-
-# build dataframes
-spp = string("Species", 1)
-df = DataFrame(spp = spp, x = x̄_sol[1,:], G = G_sol[1,:], N = N_sol[1,:], time = sol.t)
-for i in 2:S
-	spp = string("Species", i)
-	df_dummy = DataFrame(spp = spp, x = x̄_sol[i,:], G = G_sol[i,:],
[…]

### 2020-04-28 spde.jl

@@ -1,52 +1,0 @@
-"""
-Solve the heat equation ∂ₜu = Δu on [0,1] using OrdinaryDiffEq for time stepping.
-Boundary conditions are such that an analytic solution is given by:
-    sin(2π*x) * exp(-(2*π)^2*t)
-This example demonstrates how to solve time dependent PDE's combining `fenics` and `DiffEq`.
-"""
-#module HeatEquationExample
-
-using Test
-using OrdinaryDiffEq
-using FEniCS
-
-mesh = UnitIntervalMesh(50)
-V = FunctionSpace(mesh,"P",1)
-u_str = "sin(2*pi*x[0]) * exp(-t*pow(2*pi, 2))"
-
-tspan = (0.0, 0.1)
-u = interpolate(Expression(u_str, degree=2, pi=Float64(pi), t=tspan[1]), V)
-
-bc = DirichletBC(V, 0.0, "on_boundary")
-dudt = TrialFunction(V)
-v = TestFunction(V)
-Dudt = FeFunction(V)
-F = dot(dudt, v)*dx + dot(grad(u), grad(v))*dx
-a = lhs(F)
-L = rhs(F)
-p = (u=u, Dudt=Dudt, bc=bc, a=a, L=L)
-
-function f!(dudt_vec, u_vec, p, t)
-    assign(p.u, u_vec)
-    lvsolve(p.a, p.L, p.Dudt, p.bc)
-    copy!(dudt_vec, get_array(p.Dudt))
-end
-
-u0_vec = get_array(u)
-t_final = 0.1
-prob = ODEProblem(f!, u0_vec, tspan, p)
-u_true = interpolate(Expression(u_str, degree=2, pi=Float64(pi), t=tspan[2]), V)
-
-# some algorithms to try:
-# alg = ImplicitEuler(autodiff=false)
-# alg = Rosenbrock23(autodiff=false)
-# alg = Rodas5(autodiff=false)
-
-alg = KenCarp4(autodiff=false)
-sol = OrdinaryDiffEq.solve(prob, alg, reltol=1e-6, abstol=1e-6)
-
-vec_sol = get_array(p.u)
-vec_true = get_array(u_true)
-@test vec_sol ≈ vec_true rtol=1e-2
-
-#end#module

### 2020-04-28 Current Version/Current-Version.Rmd
reorganize

--- 2020-04-28 Current Version/Untitled.Rmd
@@
 ```{r global_options, include=FALSE}
+options(tinytex.clean = FALSE)
 #knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
@@
 \end{equation}
-
 where $\bar x(t)$ and $\sigma^2(t)$ are respectively the mean trait and phenotypic variance at time $t\geq0$.
@@
 \end{equation}
-
 with the initial condition $\nu(x,0)$ described above.
@@
 \end{equation}
-
 where $r\in\mathbb{R}$ is the intrinsic rate of growth in the absence of abiotic stabilizing selection, $a>0$ is the strength of abiotic stabilizing selection around the phenotypic optimum $\theta\in\mathbb{R}$, $c>0$ captures the sensitivity of fitness to intraspecific competition and $\kappa(x-y)=1$ so that competitive interactions cause the same reduction in fitness regardless of trait value.
@@
 \frac{d}{dt}\bar x(t)=\frac{d}{dt}\int_{-\infty}^{+\infty} x p(x,t)dx=\int_{-\infty}^{+\infty} x\frac{\partial}{\partial t} p(x,t)dx \\
-=\int_{-\infty}^{+\infty} x\big(m(\nu,x)-\bar m(t)\big) p(x,t)+x\frac{\mu}{2}\frac{\partial^2}{\partial x^2} p(x,t)dx
+=\int_{-\infty}^{+\infty} x\big(m(\nu,x)-\bar m(t)\big) p(x,t)+x\frac{\mu}{2}\frac{\partial^2}{\partial x^2} p(x,t)dx \\
 =\mathrm{Cov}_t\Big(m(\nu,x),x\Big).
@@
 
-ggsave("~/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/wn.png",wn,width=5,height=2.5)
+ggsave("~/Gits/branching.brownian.motion.and.spde/Current Version/wn.png",wn,width=5,height=2.5)
 ```
@@
 \end{equation}
-
 and make use of the convention $f=g$ if $\|f-g\|_t=0$ for all $t\geq0$.
@@
 \end{equation}
-
 where $t_1\wedge t_2=\min(t_1,t_2)$ and $\mathbb{C}$ denotes covariance with respect to the underlying probability space.
@@
 \end{equation}
-
 where $|D|$ denotes the length of $D$, we have $I_D\in\mathscr{N}_2$.
@@
 \end{equation}
-
 so that
@@
 \end{equation}
-
 and in particular, as a function of $t$, $\int_0^td\hat{\mathbf W}_s(f)$ is a standard Brownian motion for any $f\in\mathscr{N}_2$.
@@
 
-_Branching Brownian motion (BBM):_
+\paragraph{Branching Brownian motion (BBM)}
 A BBM tracks individuals navigating $d$-dimensional Euclidean space that reproduce and senesce between exponentially distributed intervals.
@@
 \begin{multline}\label{neutral_int}
-\int_\mathbb{R}\nu(x,t)f(x)dx-\int_\mathbb{R}\nu(x,0)f(x)dx=\int_0^t\int_\mathbb{R}\nu(x,s)\left(mf(x)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}f(x)\right)dsdx \\
+\int_\mathbb{R}\nu(x,t)f(x)dx-\int_\mathbb{R}\nu(x,0)f(x)dx\\=\int_0^t\int_\mathbb{R}\nu(x,s)\left(mf(x)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}f(x)\right)dsdx \\
 +\int_0^t\int_\mathbb{R}f(x)\sqrt{V\nu(x,s)}\dot W(x,s)dxds, \ \forall \ f\in C_b^2(\mathbb{R}).
@@
 
-_Interacting superprocesses:_
+\paragraph{Interacting superprocesses}
 In the above subsection we reviewed the diffusion limit of an especially tractable measure-valued branching process.
@@
 
-ggsave("~/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/rescaled_plots.png",rescaled_plots,device="png",width=6,height=6,units="in",dpi=350)
+ggsave("~/Gits/branching.brownian.motion.and.spde/Current Version/rescaled_plots.png",rescaled_plots,device="png",width=6,height=6,units="in",dpi=350)
 ```
@@
 \begin{multline}
-N(t)=N(0)+\int_0^t\int_\mathbb{R}\nu(x,s)\left(m(\nu,x)\cdot1+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}1\right)+1\sqrt{V\nu(x,s)}\dot W(x,s)dsdx \\
-=N(0)+\int_0^t\bar m(s)N(s)dt+\int_0^t\sqrt{VN(s)}d\hat{\mathbf{W}}_s(\sqrt{\nu(x,s)}),
+N(t)-N(0)=\\
+\int_0^t\int_\mathbb{R}\nu(x,s)\left(m(\nu,x)\cdot1+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}1\right)+1\sqrt{V\nu(x,s)}\dot W(x,s)dsdx \\
+=\int_0^t\bar m(s)N(s)dt+\int_0^t\sqrt{VN(s)}d\hat{\mathbf{W}}_s(\sqrt{\nu(x,s)}),
 \end{multline}
-
 where
@@
 \end{equation}
-
 and
@@
 \end{equation}
-\begin{equation}\label{sig2_gen}
-d\sigma^2(t)=\mathrm{Cov}_t\Big((x-\bar x(t))^2,m(\nu,x)\Big)dt+\left(\mu-V\frac{\sigma^2(t)}{N(t)}\right)dt+\sqrt{V\frac{\overline{(x-\bar x(t))^4}-\sigma^4(t)}{N(t)}}dW_3(t),
-\end{equation}
+\begin{multline}\label{sig2_gen}
+d\sigma^2(t)=\mathrm{Cov}_t\Big((x-\bar x(t))^2,m(\nu,x)\Big)dt+\left(\mu-V\frac{\sigma^2(t)}{N(t)}\right)dt\\+\sqrt{V\frac{\overline{(x-\bar x(t))^4}-\sigma^4(t)}{N(t)}}dW_3(t),
+\end{multline}
 \end{subequations}
-
 where $W_1$, $W_2$ and $W_3$ are standard Brownian motions.
@@
 \end{equation}
-
 and $\overline{(x-\bar x)^4}=3\sigma^4$. In particular, these results imply
@@
 
+\paragraph{Description of abiotic selection and competition}
 For species $i$ we inherit the above notation for trait value, distribution, average, variance, abundance, et cetera, except with an $i$ in the subscript.
@@
 
-\begin{equation}
-\mathcal O_{ij}(x_i-x_j)=\int_{-\infty}^{+\infty}u_i(\zeta,x_i)u_j(\zeta,x_j)d\zeta=\frac{U_iU_j}{\sqrt{2\pi(w_i+w_j)}}\exp\left(-\frac{(x_i-x_j)^2}{2(w_i+w_j)}\right).
-\end{equation}
-
+\begin{multline}
+\mathcal O_{ij}(x_i-x_j)=\int_{-\infty}^{+\infty}u_i(\zeta,x_i)u_j(\zeta,x_j)d\zeta\\=\frac{U_iU_j}{\sqrt{2\pi(w_i+w_j)}}\exp\left(-\frac{(x_i-x_j)^2}{2(w_i+w_j)}\right).
+\end{multline}
+
+\paragraph{Avoiding caveats of niche theory}
 A notable criticism of using niche overlap to measure the intensity of competition points to cases where populations competiting on multiple niche axes exhibit overlap on at least one of the axes, but no overall niche overlap [@Holt1987].
@@
 
+\paragraph{The fitness function}
 To map the degree of niche overlap to fitness, we assume competition between individuals with niche locations $x_i$ and $x_j$ decreases the expected reproductive output for the individual in species $i$ at the rate $c_i\mathcal{O}_{ij}(x_i-x_j)$ for some $c_i>0$.
@@
 In relation to the example fitness function discussed in \S\ref{deterministic}, consider $\kappa(x_i-x_j)=\mathcal{O}_{ij}(x_i-x_j)$.
+
+\paragraph{The model}
 In SM \S\ref{diffuse} we combine our treatment of resource competition with equations (\ref{N}), (\ref{xbarfinal}) and (\ref{Gfinal}) to find
@@
 \end{subequations}
+
+\paragraph{Description of model behavior}
 Despite the convoluted appearance of system (\ref{comm_dynamics}), there are some nice features that reflect biological reasoning.
@@
 
+\paragraph{Natural history of an in-silico community}
 We begin by describing the weak competition scenario.
@@
 
-ggsave("~/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/community_dynamics.png",community_plot,width=6.5,height=5,dpi=320)
+ggsave("~/Gits/branching.brownian.motion.and.spde/Current Version/community_dynamics.png",community_plot,width=6.5,height=5,dpi=320)
 
@@
 
-ggsave("~/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/on_pl.png",on_pl,device="png",width=3.5,height=3.5,units="in",dpi=350)
+ggsave("~/Gits/branching.brownian.motion.and.spde/Current Version/on_pl.png",on_pl,device="png",width=3.5,height=3.5,units="in",dpi=350)
 
@@
 
-ggsave("~/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/interaction_networks.png",coevs,device="png",width=6,height=6,units="in",dpi=350)
+ggsave("~/Gits/branching.brownian.motion.and.spde/Current Version/interaction_networks.png",coevs,device="png",width=6,height=6,units="in",dpi=350)
 
@@
 
-_Metric of pairwise coevolution:_
+\paragraph{Metric of pairwise coevolution}
 Below we investigate the correspondence of interaction intensity and coevolutionary change.
@@
 
-ggsave("~/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/on_pl.png",on_pl2,device="png",width=3.5,height=3.5,units="in")
+ggsave("~/Gits/branching.brownian.motion.and.spde/Current Version/on_pl.png",on_pl2,device="png",width=3.5,height=3.5,units="in")
 ```
@@
 
-_Covariance of selection and competition as a function of diversity:_
+\paragraph{Covariance of selection and competition as a function of diversity}
 We now make use of the expressions derived for competition coeffecients and selection gradients to investigate their relationship.
@@
 cov_plot <- grid.arrange(cov_ab_p,cov_ag_p,nrow=1) 
-ggsave("~/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/cov.png",cov_plot,width=6.5,height=2.5)
+ggsave("~/Gits/branching.brownian.motion.and.spde/Current Version/cov.png",cov_plot,width=6.5,height=2.5)
 ```
@@
 In both plots we set $c=1.0\\times10^{-4}$, $b=0.1$, $\\bar r=0.1$ and $N=1.0\\times10^{10}$ and let $V_{\\bar X}$ vary.'}
-knitr::include_graphics('/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/cov.png')
+knitr::include_graphics('/home/bb/Gits/branching.brownian.motion.and.spde/Current Version/cov.png')
 ```
@@
 corrsp = grid.arrange(Cαabsβp,Cαγp,nrow=1)
-ggsave("/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/corrs.png",  corrsp,  width=6.5,height=2.5)
-ggsave("/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/Cαℭp.png",   Cαℭp,  width=4,height=2.5)
+ggsave("/home/bb/Gits/branching.brownian.motion.and.spde/Current Version/corrs.png",  corrsp,  width=6.5,height=2.5)
+ggsave("/home/bb/Gits/branching.brownian.motion.and.spde/Current Version/Cαℭp.png",   Cαℭp,  width=4,height=2.5)
 ```
@@
 The red line is a cubic regression.'}
-knitr::include_graphics('/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/Cαℭp.png')
+knitr::include_graphics('/home/bb/Gits/branching.brownian.motion.and.spde/Current Version/Cαℭp.png')
 ```
@@
 If contemporary phenological coevolution is rampant, why should this pattern be so well preserved across a thousand miles of rugged and diverse terrain?
-Although it would be exciting to find that these species repeatedly coevolved this pattern in each location, a more parsimoniuous hypothesis suggests the phenology and physiology of these species slowly evolved independently over macroevolutionary time scales to take advantage of the specific conditions available within each of these windows of the flowering season.
-However, this could not have carried out in the Rocky mountains since this terrain only became habitable just over ten thousand years ago as the glaciers of the Pleistocene began to retreat (Paul CaraDonna, personal communications).
-Hence, given these considerations, it appears that an understanding of early season phenology patterns should focus on how these communities are assembled as opposed to contemporary evolutionary dynamics.
-Indeed, recent work testing models of phylogeography ignores the potential for contemporary evolution and instead suggests alpine flower communities tend to follow neutral assembly where flowers merely compete for who can disperse to new habitat first, as opposed to a selective process where a regional species pool is filtered for those species adapted to the newly available habitat [@Marx2017].
+Although it would be exciting to find that these species repeatedly coevolved this pattern in each location, a more parsimonious hypothesis suggests the phenology and physiology of these species slowly evolved independently over macroevolutionary time scales to take advantage of the specific conditions available within each of these windows of the flowering season.
+However, considering genera of these species have many early-season representatives occurring further east, such as in the Appalachian Mountains, these adaptations potentially arose before the glaciers of the Pleistocene began to retreat (Paul CaraDonna & Dave Tank, personal communications).
+Hence, given these considerations, it appears that an understanding of early season phenology patterns of sub-alpine flower communities should focus on how these communities are assembled as opposed to contemporary evolutionary dynamics.
+Indeed, recent work testing phylogeographic models suggests alpine flower communities tend to follow neutral assembly where flowers merely compete for who can disperse to new habitat first [@Marx2017].
+Due to spatial and altitudinal proximity, a similar result is likely to hold for sub-alpine flower communities.
 
@@
 In particular, macroevolutionary trait evolution is simply the aggregation of microevolutionary change occuring over large spans of time.
-This suggests a road forward to connect the theory we have introduced to models of macroevolutionary trait evolution.
+Hence, it would be ideal to formally derive macroevolutionary models from microevolutionary models.
+Furthermore, since tests based on phylogeographic models do not take into account rapid coevolutionary dynamics, it has not been rigorously shown that rapid coevolution is not a driver of alpine flower community phenology patterns.
+Hence, there is a need for models that simultaneously account for both deep macroevolutionary history of taxa and recent microevolutionary change.
+Here we suggest a road forward to accomplish both of these tasks.
 
@@
 \end{equation}
-
 where $a,b>0$, $W_t$ is a standard Brownian motion and $\theta_t$ is itself a Markov process that takes normally distributed jumps centered on its current location at exponentially distributed time intervals.
@@
 
+\newpage
+
 References {#references .unnumbered}

### 2020-04-28 Current Version/Current-Version.tex
reorganize

--- 2020-04-28 Current Version/Untitled.tex
@@
 \sigma^2(0)=\int_{-\infty}^{+\infty}(x-\bar x(0))^2p(x,0)dx<+\infty,
-\end{equation}
-
-where \(\bar x(t)\) and \(\sigma^2(t)\) are respectively the mean trait
-and phenotypic variance at time \(t\geq0\). Second, we assume selection
-is determined by the growth rate \(m(h,x)\) that is differentiable with
-respect to both arguments and is bounded above by some
-\(r\in\mathbb{R}\) for all \(x\in\mathbb{R}\) and \(h\geq0\). We also
-assume \((1+h)C\leq m(h,x)\) for some constant \(C>0\) so that
+\end{equation} where \(\bar x(t)\) and \(\sigma^2(t)\) are respectively
+the mean trait and phenotypic variance at time \(t\geq0\). Second, we
+assume selection is determined by the growth rate \(m(h,x)\) that is
+differentiable with respect to both arguments and is bounded above by
+some \(r\in\mathbb{R}\) for all \(x\in\mathbb{R}\) and \(h\geq0\). We
+also assume \((1+h)C\leq m(h,x)\) for some constant \(C>0\) so that
 \(m(h,x)\) is linear or more than linear in \(h\). We denote by \(K\) an
@@
 \frac{\partial}{\partial t}\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)
-\end{equation}
-
-with the initial condition \(\nu(x,0)\) described above. Equation
-(\ref{eq1}) is a semilinear parabolic PDE and is referred to as a scalar
-reaction-diffusion equation (Evans, 2010; Zheng, 2004). To ensure
-existence and uniqueness of solutions to (\ref{eq1}), we further assume
-that for each \(M>0\) there exists a constant \(L_M>0\) depending on
-\(M\) such that
+\end{equation} with the initial condition \(\nu(x,0)\) described above.
+Equation (\ref{eq1}) is a semilinear parabolic PDE and is referred to as
+a scalar reaction-diffusion equation (Evans, 2010; Zheng, 2004). To
+ensure existence and uniqueness of solutions to (\ref{eq1}), we further
+assume that for each \(M>0\) there exists a constant \(L_M>0\) depending
+on \(M\) such that
 \(\int_\mathbb{R}\nu_1(x)dx,\int_\mathbb{R}\nu_2(x)dx\leq M\) implies
@@
 m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_{-\infty}^{+\infty}\nu(y,t)dy
-\end{equation}
-
-where \(r\in\mathbb{R}\) is the intrinsic rate of growth in the absence
-of abiotic stabilizing selection, \(a>0\) is the strength of abiotic
-stabilizing selection around the phenotypic optimum
+\end{equation} where \(r\in\mathbb{R}\) is the intrinsic rate of growth
+in the absence of abiotic stabilizing selection, \(a>0\) is the strength
+of abiotic stabilizing selection around the phenotypic optimum
 \(\theta\in\mathbb{R}\), \(c>0\) captures the sensitivity of fitness to
@@
 \|f\|_t=\sqrt{\mathbb{E}\left(\int_0^t\int_\mathbb{R}f^2(x,s)dxds\right)},
-\end{equation}
-
-and make use of the convention \(f=g\) if \(\|f-g\|_t=0\) for all
-\(t\geq0\). Later, when we investigate stochastic dynamics of the
-abundance density \(\nu(x,t)\), we will find
+\end{equation} and make use of the convention \(f=g\) if \(\|f-g\|_t=0\)
+for all \(t\geq0\). Later, when we investigate stochastic dynamics of
+the abundance density \(\nu(x,t)\), we will find
 \(\sqrt\nu\in\mathscr{N}_2\). This enables us to utilize the heuristics
@@
 \mathbb{C}\big(\mathbf W_{t_1}(f),\mathbf W_{t_2}(g)\big)=\mathbb{E}\left(\int_0^{t_1\wedge t_2}\int_\mathbb{R} f(x,s)g(x,s)dxds\right),
-\end{equation}
-
-where \(t_1\wedge t_2=\min(t_1,t_2)\) and \(\mathbb{C}\) denotes
-covariance with respect to the underlying probability space. In
+\end{equation} where \(t_1\wedge t_2=\min(t_1,t_2)\) and \(\mathbb{C}\)
+denotes covariance with respect to the underlying probability space. In
 particular, denoting \(\mathbb{V}\) the variance operator with respect
@@
 \|I_D\|_t^2=\mathbb{E}\left(\int_0^t\int_\mathbb{R}I_D^2(x,s)dxds\right)=t\int_Ddx=t|D|<+\infty, 
-\end{equation}
-
-where \(|D|\) denotes the length of \(D\), we have
+\end{equation} where \(|D|\) denotes the length of \(D\), we have
 \(I_D\in\mathscr{N}_2\). Thus, using equations (\ref{exp_WN}) and
@@
 d\hat{\mathbf W}_t(f)=\left(\int_\mathbb{R}\hat f(x,t)\dot W(x,t)dx\right)dt
-\end{equation}
-
-so that
+\end{equation} so that
 
@@
 \mathbb{E}\left(\int_0^td\hat{\mathbf W}_s(f)\right)=0, \ \mathbb{C}\left(\int_0^{t_1}d\hat{\mathbf W}_s(f),\int_0^{t_2}d\hat{\mathbf W}_s(f)\right)=t_1\wedge t_2
-\end{equation}
-
-and in particular, as a function of \(t\),
+\end{equation} and in particular, as a function of \(t\),
 \(\int_0^td\hat{\mathbf W}_s(f)\) is a standard Brownian motion for any
@@
 
-\emph{Branching Brownian motion (BBM):} A BBM tracks individuals
-navigating \(d\)-dimensional Euclidean space that reproduce and senesce
-between exponentially distributed intervals. Unlike other stochastic
-processes that take values in \(\mathbb{R}^d\), BBM takes values in the
-set of \emph{non-negative finite measures} over \(\mathbb{R}^d\).
-Intuitively, one can think of a finite measure as a function that maps
-subsets of \(\mathbb{R}^d\) to real numbers. To be formal, we only
-consider the Borel subsets of \(\mathbb{R}^d\) corresponding to the
-Euclidean metric, but understanding this technicality is not crucial to
-our discussion. In particular, denoting \(X_t\) a BBM, for a subset
-\(D\subset\mathbb{R}^d\), \(X_t(D)\) returns the (random) number of
-individuals alive within the region \(D\) at time \(t\geq0\). The BBM
-has three main components:
+\paragraph{Branching Brownian motion (BBM)}
+
+A BBM tracks individuals navigating \(d\)-dimensional Euclidean space
+that reproduce and senesce between exponentially distributed intervals.
+Unlike other stochastic processes that take values in \(\mathbb{R}^d\),
+BBM takes values in the set of \emph{non-negative finite measures} over
+\(\mathbb{R}^d\). Intuitively, one can think of a finite measure as a
+function that maps subsets of \(\mathbb{R}^d\) to real numbers. To be
+formal, we only consider the Borel subsets of \(\mathbb{R}^d\)
+corresponding to the Euclidean metric, but understanding this
+technicality is not crucial to our discussion. In particular, denoting
+\(X_t\) a BBM, for a subset \(D\subset\mathbb{R}^d\), \(X_t(D)\) returns
+the (random) number of individuals alive within the region \(D\) at time
+\(t\geq0\). The BBM has three main components:
 
@@
 
-\emph{Interacting superprocesses:} In the above subsection we reviewed
-the diffusion limit of an especially tractable measure-valued branching
-process. However, we found the simplicity of this process restricts us
-from modelling nonlocal effects on the fitness of individuals. Here, we
-discuss superprocesses that account for such effects. The existence of
-diffusion limits for a class of measure-valued branching processes
-involving nonlocal effects has been treated by Méléard and Roelly (1993,
-1992). The interactions can manifest as dependencies of the spatial
-movement or reproductive law of individuals on their position \(x\) and
-the state of the whole population described by \(X_t\). An important
-result of Méléard and Roelly (1993, 1992) is a theorem that provides
-sufficient conditions to construct a sequence of rescaled measure-valued
-branching processes that converge to a generalization of the
-Dawson-Watanabe superprocess that includes interactions. The rescaling
-is analogous to that described above for non-interacting Dawson-Watanabe
-superprocesses, but now the reproductive law described by
-\(\mathscr{W}(X_t,x)\) and \(V(X_t,x)\), branching rate
-\(\lambda(X_t,x)\) and diffusion parameter \(\sqrt{\mu(X_t,x)}\) are
-allowed to depend on the whole population \(X_t\) and individual
-location \(x\). In Figure \ref{rescaled} we demonstrate this rescaling
-in discrete time for a population experiencing stabilizing selection and
-logistic growth. Since time is discretized, the process we simulate is
-formally a branching random walk. For further details on our simulation
-see SM \S\ref{numerical}.
+\paragraph{Interacting superprocesses}
+
+In the above subsection we reviewed the diffusion limit of an especially
+tractable measure-valued branching process. However, we found the
+simplicity of this process restricts us from modelling nonlocal effects
+on the fitness of individuals. Here, we discuss superprocesses that
+account for such effects. The existence of diffusion limits for a class
+of measure-valued branching processes involving nonlocal effects has
+been treated by Méléard and Roelly (1993, 1992). The interactions can
+manifest as dependencies of the spatial movement or reproductive law of
+individuals on their position \(x\) and the state of the whole
+population described by \(X_t\). An important result of Méléard and
+Roelly (1993, 1992) is a theorem that provides sufficient conditions to
+construct a sequence of rescaled measure-valued branching processes that
+converge to a generalization of the Dawson-Watanabe superprocess that
+includes interactions. The rescaling is analogous to that described
+above for non-interacting Dawson-Watanabe superprocesses, but now the
+reproductive law described by \(\mathscr{W}(X_t,x)\) and \(V(X_t,x)\),
+branching rate \(\lambda(X_t,x)\) and diffusion parameter
+\(\sqrt{\mu(X_t,x)}\) are allowed to depend on the whole population
+\(X_t\) and individual location \(x\). In Figure \ref{rescaled} we
+demonstrate this rescaling in discrete time for a population
+experiencing stabilizing selection and logistic growth. Since time is
+discretized, the process we simulate is formally a branching random
+walk. For further details on our simulation see SM \S\ref{numerical}.
 
@@
 \begin{multline}
-N(t)=N(0)+\int_0^t\int_\mathbb{R}\nu(x,s)\left(m(\nu,x)\cdot1+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}1\right)+1\sqrt{V\nu(x,s)}\dot W(x,s)dsdx \\
-=N(0)+\int_0^t\bar m(s)N(s)dt+\int_0^t\sqrt{VN(s)}d\hat{\mathbf{W}}_s(\sqrt{\nu(x,s)}),
-\end{multline}
-
-where
+N(t)-N(0)=\\
+\int_0^t\int_\mathbb{R}\nu(x,s)\left(m(\nu,x)\cdot1+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}1\right)+1\sqrt{V\nu(x,s)}\dot W(x,s)dsdx \\
+=\int_0^t\bar m(s)N(s)dt+\int_0^t\sqrt{VN(s)}d\hat{\mathbf{W}}_s(\sqrt{\nu(x,s)}),
+\end{multline} where
 
@@
 \bar m(t)=\frac{1}{N(t)}\int_\mathbb{R}m(\nu,x)\nu(x,t)dx,
-\end{equation}
-
-and
+\end{equation} and
 
@@
 \end{equation}
-\begin{equation}\label{sig2_gen}
-d\sigma^2(t)=\mathrm{Cov}_t\Big((x-\bar x(t))^2,m(\nu,x)\Big)dt+\left(\mu-V\frac{\sigma^2(t)}{N(t)}\right)dt+\sqrt{V\frac{\overline{(x-\bar x(t))^4}-\sigma^4(t)}{N(t)}}dW_3(t),
-\end{equation}
+\begin{multline}\label{sig2_gen}
+d\sigma^2(t)=\mathrm{Cov}_t\Big((x-\bar x(t))^2,m(\nu,x)\Big)dt+\left(\mu-V\frac{\sigma^2(t)}{N(t)}\right)dt\\+\sqrt{V\frac{\overline{(x-\bar x(t))^4}-\sigma^4(t)}{N(t)}}dW_3(t),
+\end{multline}
 \end{subequations}
@@
 \mathrm{Cov}\Big((x-\bar x)^2,m\Big)=2\sigma^4\left(\frac{\partial\bar m}{\partial\sigma^2}-\overline{\frac{\partial m}{\partial\sigma^2}}\right)
-\end{equation}
-
-and \(\overline{(x-\bar x)^4}=3\sigma^4\). In particular, these results
-imply
+\end{equation} and \(\overline{(x-\bar x)^4}=3\sigma^4\). In particular,
+these results imply
 
@@
 
+\paragraph{Description of abiotic selection and competition}
+
 For species \(i\) we inherit the above notation for trait value,
@@
 
-\begin{equation}
-\mathcal O_{ij}(x_i-x_j)=\int_{-\infty}^{+\infty}u_i(\zeta,x_i)u_j(\zeta,x_j)d\zeta=\frac{U_iU_j}{\sqrt{2\pi(w_i+w_j)}}\exp\left(-\frac{(x_i-x_j)^2}{2(w_i+w_j)}\right).
-\end{equation}
+\begin{multline}
+\mathcal O_{ij}(x_i-x_j)=\int_{-\infty}^{+\infty}u_i(\zeta,x_i)u_j(\zeta,x_j)d\zeta\\=\frac{U_iU_j}{\sqrt{2\pi(w_i+w_j)}}\exp\left(-\frac{(x_i-x_j)^2}{2(w_i+w_j)}\right).
+\end{multline}
+
+\paragraph{Avoiding caveats of niche theory}
 
@@
 
+\paragraph{The fitness function}
+
 To map the degree of niche overlap to fitness, we assume competition
@@
 \S\ref{deterministic}, consider
-\(\kappa(x_i-x_j)=\mathcal{O}_{ij}(x_i-x_j)\). In SM \S\ref{diffuse} we
-combine our treatment of resource competition with equations (\ref{N}),
-(\ref{xbarfinal}) and (\ref{Gfinal}) to find
+\(\kappa(x_i-x_j)=\mathcal{O}_{ij}(x_i-x_j)\).
+
+\paragraph{The model}
+
+In SM \S\ref{diffuse} we combine our treatment of resource competition
+with equations (\ref{N}), (\ref{xbarfinal}) and (\ref{Gfinal}) to find
 
@@
 \end{subequations}
+
+\paragraph{Description of model behavior}
 
@@
 illustrated in Figure \ref{temporal}.
+
+\paragraph{Natural history of an in-silico community}
 
@@
 \alpha_{ij}(t)=\frac{c_i}{r_i(t)}\int_{-\infty}^{+\infty}\int_{-\infty}^{+\infty}p_i(x,t)p_j(y,t)\mathcal O_{ij}(x,y) dxdy \\ =\frac{c_iU_iU_j}{r_i(t)}\sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right),
-\end{multline}
-
-where
+\end{multline} where
 
@@
 
-\emph{Metric of pairwise coevolution:} Below we investigate the
-correspondence of interaction intensity and coevolutionary change.
-However, we can already identify one major discrepancy; \(\alpha_{ij}\)
-is maximized when \(\bar x_i=\bar x_j\), but \(\beta_{ij}=0\) under the
-same condition. We therefore include in our metric of selection the
-standardized stabilizing selection gradient \(\gamma\) which measures
-the effect of stabilizing or disruptive selection on phenotypic variance
-(Lande and Arnold, 1983). In SM \S\ref{diffuse} we show that the
-standardized stabilizing selection gradient induced by species \(j\) on
-species \(i\) can be computed as
+\paragraph{Metric of pairwise coevolution}
+
+Below we investigate the correspondence of interaction intensity and
+coevolutionary change. However, we can already identify one major
+discrepancy; \(\alpha_{ij}\) is maximized when \(\bar x_i=\bar x_j\),
+but \(\beta_{ij}=0\) under the same condition. We therefore include in
+our metric of selection the standardized stabilizing selection gradient
+\(\gamma\) which measures the effect of stabilizing or disruptive
+selection on phenotypic variance (Lande and Arnold, 1983). In SM
+\S\ref{diffuse} we show that the standardized stabilizing selection
+gradient induced by species \(j\) on species \(i\) can be computed as
 
@@
 
-\emph{Covariance of selection and competition as a function of
-diversity:} We now make use of the expressions derived for competition
-coeffecients and selection gradients to investigate their relationship.
-As a first pass, let us assume all model parameters are equivalent
-across species and that each species has the same abundance and trait
-variance. Let us further assume that richness \(S\) is large and the
-distribution of mean trait values is normal with mean \(\bar{\bar x}\),
-variance \(V_{\bar X}\) and density \(f_{\bar X}\). Such assumptions are
-typical when deriving analytical results in the field of theoretical
+\paragraph{Covariance of selection and competition as a function of diversity}
+
+We now make use of the expressions derived for competition coeffecients
+and selection gradients to investigate their relationship. As a first
+pass, let us assume all model parameters are equivalent across species
+and that each species has the same abundance and trait variance. Let us
+further assume that richness \(S\) is large and the distribution of mean
+trait values is normal with mean \(\bar{\bar x}\), variance
+\(V_{\bar X}\) and density \(f_{\bar X}\). Such assumptions are typical
+when deriving analytical results in the field of theoretical
 coevolutionary community ecology (Nuismer et al., 2012, 2018). If
@@
 
-{\centering \includegraphics[width=1\linewidth]{/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/cov}
+{\centering \includegraphics[width=1\linewidth]{/home/bb/Gits/branching.brownian.motion.and.spde/Current Version/cov}
 
@@
 
-{\centering \includegraphics[width=0.6\linewidth]{/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/Cαℭp}
+{\centering \includegraphics[width=0.6\linewidth]{/home/bb/Gits/branching.brownian.motion.and.spde/Current Version/Cαℭp}
 
@@
 species repeatedly coevolved this pattern in each location, a more
-parsimoniuous hypothesis suggests the phenology and physiology of these
+parsimonious hypothesis suggests the phenology and physiology of these
 species slowly evolved independently over macroevolutionary time scales
 to take advantage of the specific conditions available within each of
-these windows of the flowering season. However, this could not have
-carried out in the Rocky mountains since this terrain only became
-habitable just over ten thousand years ago as the glaciers of the
-Pleistocene began to retreat (Paul CaraDonna, personal communications).
-Hence, given these considerations, it appears that an understanding of
-early season phenology patterns should focus on how these communities
-are assembled as opposed to contemporary evolutionary dynamics. Indeed,
-recent work testing models of phylogeography ignores the potential for
-contemporary evolution and instead suggests alpine flower communities
-tend to follow neutral assembly where flowers merely compete for who can
-disperse to new habitat first, as opposed to a selective process where a
-regional species pool is filtered for those species adapted to the newly
-available habitat (Marx et al., 2017).
+these windows of the flowering season. However, considering genera of
+these species have many early-season representatives occurring further
+east, such as in the Appalachian Mountains, these adaptations
+potentially arose before the glaciers of the Pleistocene began to
+retreat (Paul CaraDonna \& Dave Tank, personal communications). Hence,
+given these considerations, it appears that an understanding of early
+season phenology patterns of sub-alpine flower communities should focus
+on how these communities are assembled as opposed to contemporary
+evolutionary dynamics. Indeed, recent work testing phylogeographic
+models suggests alpine flower communities tend to follow neutral
+assembly where flowers merely compete for who can disperse to new
+habitat first (Marx et al., 2017). Due to spatial and altitudinal
+proximity, a similar result is likely to hold for sub-alpine flower
+communities.
 
@@
 macroevolutionary trait evolution is simply the aggregation of
-microevolutionary change occuring over large spans of time. This
-suggests a road forward to connect the theory we have introduced to
-models of macroevolutionary trait evolution.
+microevolutionary change occuring over large spans of time. Hence, it
+would be ideal to formally derive macroevolutionary models from
+microevolutionary models. Furthermore, since tests based on
+phylogeographic models do not take into account rapid coevolutionary
+dynamics, it has not been rigorously shown that rapid coevolution is not
+a driver of alpine flower community phenology patterns. Hence, there is
+a need for models that simultaneously account for both deep
+macroevolutionary history of taxa and recent microevolutionary change.
+Here we suggest a road forward to accomplish both of these tasks.
 
@@
 dX_t=a(\theta_t-X_t)dt+bdW_t
-\end{equation}
-
-where \(a,b>0\), \(W_t\) is a standard Brownian motion and \(\theta_t\)
-is itself a Markov process that takes normally distributed jumps
-centered on its current location at exponentially distributed time
-intervals. If we assume the rate \(\lambda\) at which jumps occur is
-much smaller than \(a\) and the variance in jumping is much larger than
-\(b^2\), then, even though the sample paths of \(X_t\) are actually
-continuous (if we zoom in close enough, they look like Brownian motion),
-over long intervals of time sample paths of \(X_t\) will begin to appear
-to have periods of continuity interrupted by an occasional discontinuous
-jump and thus approach a qualitatively distinct process. These emergent
-properties can be formally characterized by Lévy processes and have been
-successfully employed in comparative phylogenetics to fit phenotypic
-data from extant populations and the fossil record (Landis and
-Schraiber, 2017). It would therefore be interesting to investigate
-whether an application of a separation of time scales argument for the
-rate of environmental change (\(\lambda\)) versus the rate of
-evolutionary and ecological change (\(a\)) to microevolutionary models
-derived via our framework can be used to obtain macroevolutionary models
-that include not only mean trait evolution, but also the evolution of
-trait variance and abundance. The resulting macroevolutionary models can
-give rise to novel comparative phylogenetic methods and provide initial
-conditions for microevolutionary models that capture contemporary
-dynamics.
+\end{equation} where \(a,b>0\), \(W_t\) is a standard Brownian motion
+and \(\theta_t\) is itself a Markov process that takes normally
+distributed jumps centered on its current location at exponentially
+distributed time intervals. If we assume the rate \(\lambda\) at which
+jumps occur is much smaller than \(a\) and the variance in jumping is
+much larger than \(b^2\), then, even though the sample paths of \(X_t\)
+are actually continuous (if we zoom in close enough, they look like
+Brownian motion), over long intervals of time sample paths of \(X_t\)
+will begin to appear to have periods of continuity interrupted by an
+occasional discontinuous jump and thus approach a qualitatively distinct
+process. These emergent properties can be formally characterized by Lévy
+processes and have been successfully employed in comparative
+phylogenetics to fit phenotypic data from extant populations and the
+fossil record (Landis and Schraiber, 2017). It would therefore be
+interesting to investigate whether an application of a separation of
+time scales argument for the rate of environmental change (\(\lambda\))
+versus the rate of evolutionary and ecological change (\(a\)) to
+microevolutionary models derived via our framework can be used to obtain
+macroevolutionary models that include not only mean trait evolution, but
+also the evolution of trait variance and abundance. The resulting
+macroevolutionary models can give rise to novel comparative phylogenetic
+methods and provide initial conditions for microevolutionary models that
+capture contemporary dynamics.
 
@@
 
+\newpage
+
 \hypertarget{references}{%

### 2020-04-28 Current Version/SM.Rmd
reorganize

--- 2020-04-18 .A White Noise Approach To Evolutionary Ecology/SM.Rmd
@@
   - \linenumbers
-  - \externaldocument{draft}  
+  - \externaldocument{Current-Version}  
 editor_options: 
@@
 ```{r global_options, include=FALSE}
-#knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
+options(tinytex.clean = FALSE)
+knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
 ```
@@
 
-```{r, echo=F, fig.align='center', fig.cap='\\label{num_corr} Numerical estimate for the correlations of selection gradients and competition coefficients.'}
-knitr::include_graphics('/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/corrs.png')
+```{r, echo=F, fig.align='center', out.width='100%', fig.cap='\\label{num_corr} Numerical estimate for the correlations of selection gradients and competition coefficients.'}
+knitr::include_graphics('/home/bb/Gits/branching.brownian.motion.and.spde/Current Version/corrs.png')
 ```

### 2020-04-28 Current-Version.Rmd
organize for overleaf

--- 2020-04-28 Current Version/Current-Version.Rmd
@@
 
-ggsave("~/Gits/branching.brownian.motion.and.spde/Current Version/wn.png",wn,width=5,height=2.5)
+ggsave("~/Gits/branching.brownian.motion.and.spde/wn.png",wn,width=5,height=2.5)
 ```
@@
 
-ggsave("~/Gits/branching.brownian.motion.and.spde/Current Version/rescaled_plots.png",rescaled_plots,device="png",width=6,height=6,units="in",dpi=350)
+ggsave("~/Gits/branching.brownian.motion.and.spde/rescaled_plots.png",rescaled_plots,device="png",width=6,height=6,units="in",dpi=350)
 ```
@@
 
-ggsave("~/Gits/branching.brownian.motion.and.spde/Current Version/community_dynamics.png",community_plot,width=6.5,height=5,dpi=320)
+ggsave("~/Gits/branching.brownian.motion.and.spde/community_dynamics.png",community_plot,width=6.5,height=5,dpi=320)
 
@@
 
-ggsave("~/Gits/branching.brownian.motion.and.spde/Current Version/on_pl.png",on_pl,device="png",width=3.5,height=3.5,units="in",dpi=350)
+ggsave("~/Gits/branching.brownian.motion.and.spde/on_pl.png",on_pl,device="png",width=3.5,height=3.5,units="in",dpi=350)
 
@@
 
-ggsave("~/Gits/branching.brownian.motion.and.spde/Current Version/interaction_networks.png",coevs,device="png",width=6,height=6,units="in",dpi=350)
+ggsave("~/Gits/branching.brownian.motion.and.spde/interaction_networks.png",coevs,device="png",width=6,height=6,units="in",dpi=350)
 
@@
 
-ggsave("~/Gits/branching.brownian.motion.and.spde/Current Version/on_pl.png",on_pl2,device="png",width=3.5,height=3.5,units="in")
+ggsave("~/Gits/branching.brownian.motion.and.spde/on_pl.png",on_pl2,device="png",width=3.5,height=3.5,units="in")
 ```
@@
 cov_plot <- grid.arrange(cov_ab_p,cov_ag_p,nrow=1) 
-ggsave("~/Gits/branching.brownian.motion.and.spde/Current Version/cov.png",cov_plot,width=6.5,height=2.5)
+ggsave("~/Gits/branching.brownian.motion.and.spde/cov.png",cov_plot,width=6.5,height=2.5)
 ```
@@
 In both plots we set $c=1.0\\times10^{-4}$, $b=0.1$, $\\bar r=0.1$ and $N=1.0\\times10^{10}$ and let $V_{\\bar X}$ vary.'}
-knitr::include_graphics('/home/bb/Gits/branching.brownian.motion.and.spde/Current Version/cov.png')
+knitr::include_graphics('/home/bb/Gits/branching.brownian.motion.and.spde/cov.png')
 ```
@@
 corrsp = grid.arrange(Cαabsβp,Cαγp,nrow=1)
-ggsave("/home/bb/Gits/branching.brownian.motion.and.spde/Current Version/corrs.png",  corrsp,  width=6.5,height=2.5)
-ggsave("/home/bb/Gits/branching.brownian.motion.and.spde/Current Version/Cαℭp.png",   Cαℭp,  width=4,height=2.5)
+ggsave("/home/bb/Gits/branching.brownian.motion.and.spde/corrs.png",  corrsp,  width=6.5,height=2.5)
+ggsave("/home/bb/Gits/branching.brownian.motion.and.spde/Cαℭp.png",   Cαℭp,  width=4,height=2.5)
 ```
@@
 The red line is a cubic regression.'}
-knitr::include_graphics('/home/bb/Gits/branching.brownian.motion.and.spde/Current Version/Cαℭp.png')
+knitr::include_graphics('/home/bb/Gits/branching.brownian.motion.and.spde/Cαℭp.png')
 ```

### 2020-04-28 Current-Version.tex
organize for overleaf

--- 2020-04-28 Current Version/Current-Version.tex
@@
 
-{\centering \includegraphics[width=1\linewidth]{/home/bb/Gits/branching.brownian.motion.and.spde/Current Version/cov}
+{\centering \includegraphics[width=1\linewidth]{/home/bb/Gits/branching.brownian.motion.and.spde/cov}
 
@@
 
-{\centering \includegraphics[width=0.6\linewidth]{/home/bb/Gits/branching.brownian.motion.and.spde/Current Version/Cαℭp}
+{\centering \includegraphics[width=0.6\linewidth]{/home/bb/Gits/branching.brownian.motion.and.spde/Cαℭp}
 

### 2020-04-28 SM.Rmd
organize for overleaf

--- 2020-04-28 Current Version/SM.Rmd
@@
 ```{r, echo=F, fig.align='center', out.width='100%', fig.cap='\\label{num_corr} Numerical estimate for the correlations of selection gradients and competition coefficients.'}
-knitr::include_graphics('/home/bb/Gits/branching.brownian.motion.and.spde/Current Version/corrs.png')
+knitr::include_graphics('/home/bb/Gits/branching.brownian.motion.and.spde/corrs.png')
 ```

### 2020-04-28 mapping.Rmd
organize for overleaf

---
title: "Mapping Méléard and Roelly (1992) to Week et. al."
output:
  pdf_document:
    toc: false
    number_sections: true
    fig_width: 4
    fig_height: 3
    fig_caption: true
header-includes:
  - \usepackage{float}
  - \usepackage{mathrsfs}
  - \usepackage{hyperref}
  - \usepackage[left]{lineno}
  - \usepackage{amsmath}
  - \usepackage{makecell}
  - \usepackage{arydshln}
editor_options: 
  chunk_output_type: console
---

\begin{table}[H]
\begin{tabular}{l|l|l}
Description           & Méléard and Roelly 92 & Week et al    \\ \hline
&& \\
Arbitrary finite measure & $\mu\in M_F(\mathbb{R}^d)$ & NA \\ \hdashline
&& \\
Bounded function & $f\in C_b(\mathbb{R}^d)$ & $f\in C_b(\mathbb{R})$ \\ \hdashline
&& \\
\makecell{Non-negative continuous \\ integrable function} & NA & $\nu(\cdot,t)\in C_1^+(\mathbb{R})$ \\ \hdashline
&& \\
"Spatial" location    & $x\in\mathbb{R}^d$ & $x\in\mathbb{R}\dots$ \makecell{interpreted as \\ trait value} \\ \hdashline
&& \\
\makecell{Infinitesimal generator \\ of spatial movement} & \makecell{$A(\mu)\dots$ \\ satisfies ($\mathrm{H}_1$) and ($\mathrm{H}_2$) below} & \makecell{$\frac{\mu}{2}\Delta\dots$ Laplace operator \\ with diffusion coefficient $\mu/2$} \\ \hdashline
&& \\
Superprocess          & $X_t\in M_F(\mathbb{R}^d)$ for $t\in[0,\infty)$ & NA                      \\ \hdashline
&& \\
Density process       & NA               & $\nu(x,t)$              \\ \hdashline
&& \\
Duality bracket       & $\begin{matrix}\langle X_t,f\rangle\dots \\ X_t\in M_F(\mathbb{R}^d), \ f\in C_b(\mathbb{R}^d)\end{matrix}$ & $\begin{matrix}\int_\mathbb{R}\nu(x,t)f(x)dx\dots \\ \nu(\cdot,t)\in C_1^+(\mathbb{R}), \ f\in C^2_b(\mathbb{R})\end{matrix}$ \\ \hdashline
&& \\
Malthusian parameter  & \makecell{$b(\mu)(x)\dots \ b(\mu)\in C_b(\mathbb{R}^d)$ \\ for $\mu\in M_F(\mathbb{R}^d)$} & $m\big(\nu(x,t),x\big)\dots$ \makecell{bounded above, \\ but not below} \\ \hdashline
&& \\
Stochastic parameter  & \makecell{$c(\mu)(x)\dots \ c(\mu)\in C_b(\mathbb{R}^d)$ \\ for $\mu\in M_F(\mathbb{R}^d)$} & $V>0\dots$ constant \\ \hdashline
&& \\
Reproductive law      & $p_k(x,\mu)$ & NA \\ \hdashline
&& \\
Mean of reproductive law & $\begin{matrix}m(x,\mu)=\sum_{k\in\mathbb{N}}kp_k(x,\mu)\dots \\ \sup_{x,\mu}|m(x,\mu)|<\infty\end{matrix}$ & $\begin{matrix}\mathscr{W}(\nu,x)\dots\\\sup_{\nu,x}\mathscr{W}(\nu,x)<\infty\end{matrix}$ \makecell{introduced in \\ appendix} \\ \hdashline
&& \\
Unnamed                  & $v^2(x,\mu)=\sum_{k\in\mathbb{N}}(k-1)^2p_k(x,\mu)$ & NA \\ \hdashline
&& \\
Death rate               & $\lambda(x,\mu)$ & we assume constant "lifetimes"
\end{tabular}\label{mult}
\end{table}
\newpage
\section*{Notes}

- ($\mathrm{H}_1$): For all $\mu\in M_F(\mathbb{R}^d)$ and all $f$ in the domain of $A(\mu)$, there is a constant $K>0$ such that $\|A(\mu)f\|_\infty\leq K\langle\mu,1\rangle$.

- ($\mathrm{H}_2$): For all $f$ in the domain of $A(\mu)$, the map $\mu\to\langle\mu,A(\mu)f\rangle$ is continuous.

- In Week et al, spatial movement is interpreted as mutation. Hence, durations of spatial movement are not interpreted as lifetimes. Instead, we assume discrete non-overlapping generations for the branching diffusion process.

- In Champagnat, Ferrière and Méléard (2006) the condition $\sup_{x,\mu}|m(x,\mu)|<\infty$ has been weakened to $\sup_{x,\mu}m(x,\mu)<\infty$ and $\ln m(x,\mu)\geq C(1+\mu(\mathbb{R}^d))$ for some constant $K>0$ which they refer to as being linear or "more than linear".
